### 2020-03-23 VegOzone.tex
Initialize Copernicus package 5.8.

%% Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
%% ---------------------------------
%% This template should be used for copernicus.cls
%% The class file and some style files are bundled in the Copernicus Latex Package, which can be downloaded from the different journal webpages.
%% For further assistance please contact Copernicus Publications at: production@copernicus.org
%% https://publications.copernicus.org/for_authors/manuscript_preparation.html

%% Please use the following documentclass and journal abbreviations for discussion papers and final revised papers.

%% 2-column papers and discussion papers
\documentclass[journal abbreviation, manuscript]{copernicus}

%% Journal abbreviations (please use the same for discussion papers and final revised papers)

% Advances in Geosciences (adgeo)
% Advances in Radio Science (ars)
% Advances in Science and Research (asr)
% Advances in Statistical Climatology, Meteorology and Oceanography (ascmo)
% Annales Geophysicae (angeo)
% Archives Animal Breeding (aab)
% ASTRA Proceedings (ap)
% Atmospheric Chemistry and Physics (acp)
% Atmospheric Measurement Techniques (amt)
% Biogeosciences (bg)
% Climate of the Past (cp)
% DEUQUA Special Publications (deuquasp)
% Drinking Water Engineering and Science (dwes)
% Earth Surface Dynamics (esurf)
% Earth System Dynamics (esd)
% Earth System Science Data (essd)
% E&G Quaternary Science Journal (egqsj)
% European Journal of Mineralogy (ejm)
% Fossil Record (fr)
% Geochronology (gchron)
% Geographica Helvetica (gh)
% Geoscience Communication (gc)
% Geoscientific Instrumentation, Methods and Data Systems (gi)
% Geoscientific Model Development (gmd)
% History of Geo- and Space Sciences (hgss)
% Hydrology and Earth System Sciences (hess)
% Journal of Micropalaeontology (jm)
% Journal of Sensors and Sensor Systems (jsss)
% Magnetic Resonance (mr)
% Mechanical Sciences (ms)
% Natural Hazards and Earth System Sciences (nhess)
% Nonlinear Processes in Geophysics (npg)
% Ocean Science (os)
% Primate Biology (pb)
% Proceedings of the International Association of Hydrological Sciences (piahs)
% Scientific Drilling (sd)
% SOIL (soil)
% Solid Earth (se)
% The Cryosphere (tc)
% Weather and Climate Dynamics (wcd)
% Web Ecology (we)
% Wind Energy Science (wes)

%% \usepackage commands included in the copernicus.cls:
%\usepackage[german, english]{babel}
%\usepackage{tabularx}
%\usepackage{cancel}
%\usepackage{multirow}
%\usepackage{supertabular}
%\usepackage{algorithmic}
%\usepackage{algorithm}
%\usepackage{amsthm}
%\usepackage{float}
%\usepackage{subfig}
%\usepackage{rotating}

\begin{document}

\title{TEXT}

% \Author[affil]{given_name}{surname}

\Author[]{}{}
\Author[]{}{}
\Author[]{}{}

\affil[]{ADDRESS}
\affil[]{ADDRESS}

%% The [] brackets identify the author with the corresponding affiliation. 1, 2, 3, etc. should be inserted.

%% If an author is deceased, please mark the respective author name(s) with a dagger, e.g. "\Author[2,$\dag$]{Anton}{Aman}", and add a further "\affil[$\dag$]{deceased, 1 July 2019}".

%% If authors contributed equally, please mark the respective author names with an asterisk, e.g. "\Author[2,*]{Anton}{Aman}" and "\Author[3,*]{Bradley}{Bman}" and add a further affiliation: "\affil[*]{These authors contributed equally to this work.}".

\correspondence{NAME (EMAIL)}

\runningtitle{TEXT}

\runningauthor{TEXT}

\received{}
\pubdiscuss{} %% only important for two-stage journals
\revised{}
\accepted{}
\published{}

%% These dates will be inserted by Copernicus Publications during the typesetting process.

\firstpage{1}

\maketitle

\begin{abstract}
TEXT
\end{abstract}

\copyrightstatement{TEXT}

\introduction  %% \introduction[modified heading if necessary]
TEXT

\section{HEADING}
TEXT

\subsection{HEADING}
TEXT

\subsubsection{HEADING}
TEXT

\conclusions  %% \conclusions[modified heading if necessary]
TEXT

%% The following commands are for the statements about the availability of data sets and/or software code corresponding to the manuscript.
%% It is strongly recommended to make use of these sections in case data sets and/or software code have been part of your research the article is based on.

\codeavailability{TEXT} %% use this section when having only software code available

\dataavailability{TEXT} %% use this section when having only data sets available

\codedataavailability{TEXT} %% use this section when having data sets and software code available

\sampleavailability{TEXT} %% use this section when having geoscientific samples available

\videosupplement{TEXT} %% use this section when having video supplements available

\appendix
\section{}    %% Appendix A

\subsection{}     %% Appendix A1, A2, etc.

\noappendix       %% use this to mark the end of the appendix section. Otherwise the figures might be numbered incorrectly (e.g. 10 instead of 1).

%% Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:

%% Option 1: If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
%% They will be correctly named automatically.

%% Option 2: If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.
%% To rename them correctly to A1, A2, etc., please add the following commands in front of them:

\appendixfigures  %% needs to be added in front of appendix figures

\appendixtables   %% needs to be added in front of appendix tables

%% Please add \clearpage between each table and/or figure. Further guidelines on figures and tables can be found below.

\authorcontribution{TEXT} %% this section is mandatory

\competinginterests{TEXT} %% this section is mandatory even if you declare that no competing interests are present

\disclaimer{TEXT} %% optional section

\begin{acknowledgements}
TEXT
\end{acknowledgements}

%% REFERENCES

%% The reference list is compiled as follows:

\begin{thebibliography}{}

\bibitem[AUTHOR(YEAR)]{LABEL1}
REFERENCE 1

\bibitem[AUTHOR(YEAR)]{LABEL2}
REFERENCE 2

\end{thebibliography}

%% Since the Copernicus LaTeX package includes the BibTeX style file copernicus.bst,
%% authors experienced with BibTeX only have to include the following two lines:
%%
%% \bibliographystyle{copernicus}
%% \bibliography{example.bib}
%%
%% URLs and DOIs can be entered in your BibTeX file as:
%%
%% URL = {http://www.xyz.org/~jones/idx_g.htm}
%% DOI = {10.5194/xyz}

%% LITERATURE CITATIONS
%%
%% command                        & example result
%% \citet{jones90}|               & Jones et al. (1990)
%% \citep{jones90}|               & (Jones et al., 1990)
%% \citep{jones90,jones93}|       & (Jones et al., 1990, 1993)
%% \citep[p.~32]{jones90}|        & (Jones et al., 1990, p.~32)
%% \citep[e.g.,][]{jones90}|      & (e.g., Jones et al., 1990)
%% \citep[e.g.,][p.~32]{jones90}| & (e.g., Jones et al., 1990, p.~32)
%% \citeauthor{jones90}|          & Jones et al.
%% \citeyear{jones90}|            & 1990

%% FIGURES

%% When figures and tables are placed at the end of the MS (article in one-column style), please add \clearpage
%% between bibliography and first table and/or figure as well as between each table and/or figure.

% The figure files should be labelled correctly with Arabic numerals (e.g. fig01.jpg, fig02.png).

%% ONE-COLUMN FIGURES

%%f
%\begin{figure}[t]
%\includegraphics[width=8.3cm]{FILE NAME}
%\caption{TEXT}
%\end{figure}
%
%%% TWO-COLUMN FIGURES
%
%%f
%\begin{figure*}[t]
%\includegraphics[width=12cm]{FILE NAME}
%\caption{TEXT}
%\end{figure*}
%
%
%%% TABLES
%%%
%%% The different columns must be seperated with a & command and should
%%% end with \\ to identify the column brake.
%
%%% ONE-COLUMN TABLE
%
%%t
%\begin{table}[t]
%\caption{TEXT}
%\begin{tabular}{column = lcr}
%\tophline
%
%\middlehline
%
%\bottomhline
%\end{tabular}
%\belowtable{} % Table Footnotes
%\end{table}
%
%%% TWO-COLUMN TABLE
%
%%t
%\begin{table*}[t]
%\caption{TEXT}
%\begin{tabular}{column = lcr}
%\tophline
%
%\middlehline
%
%\bottomhline
%\end{tabular}
%\belowtable{} % Table Footnotes
%\end{table*}
%
%%% LANDSCAPE TABLE
%
%%t
%\begin{sidewaystable*}[t]
%\caption{TEXT}
%\begin{tabular}{column = lcr}
%\tophline
%
%\middlehline
%
%\bottomhline
%\end{tabular}
%\belowtable{} % Table Footnotes
%\end{sidewaystable*}
%
%
%%% MATHEMATICAL EXPRESSIONS
%
%%% All papers typeset by Copernicus Publications follow the math typesetting regulations
%%% given by the IUPAC Green Book (IUPAC: Quantities, Units and Symbols in Physical Chemistry,
%%% 2nd Edn., Blackwell Science, available at: http://old.iupac.org/publications/books/gbook/green_book_2ed.pdf, 1993).
%%%
%%% Physical quantities/variables are typeset in italic font (t for time, T for Temperature)
%%% Indices which are not defined are typeset in italic font (x, y, z, a, b, c)
%%% Items/objects which are defined are typeset in roman font (Car A, Car B)
%%% Descriptions/specifications which are defined by itself are typeset in roman font (abs, rel, ref, tot, net, ice)
%%% Abbreviations from 2 letters are typeset in roman font (RH, LAI)
%%% Vectors are identified in bold italic font using \vec{x}
%%% Matrices are identified in bold roman font
%%% Multiplication signs are typeset using the LaTeX commands \times (for vector products, grids, and exponential notations) or \cdot
%%% The character * should not be applied as mutliplication sign
%
%
%%% EQUATIONS
%
%%% Single-row equation
%
%\begin{equation}
%
%\end{equation}
%
%%% Multiline equation
%
%\begin{align}
%& 3 + 5 = 8\\
%& 3 + 5 = 8\\
%& 3 + 5 = 8
%\end{align}
%
%
%%% MATRICES
%
%\begin{matrix}
%x & y & z\\
%x & y & z\\
%x & y & z\\
%\end{matrix}
%
%
%%% ALGORITHM
%
%\begin{algorithm}
%\caption{...}
%\label{a1}
%\begin{algorithmic}
%...
%\end{algorithmic}
%\end{algorithm}
%
%
%%% CHEMICAL FORMULAS AND REACTIONS
%
%%% For formulas embedded in the text, please use \chem{}
%
%%% The reaction environment creates labels including the letter R, i.e. (R1), (R2), etc.
%
%\begin{reaction}
%%% \rightarrow should be used for normal (one-way) chemical reactions
%%% \rightleftharpoons should be used for equilibria
%%% \leftrightarrow should be used for resonance structures
%\end{reaction}
%
%
%%% PHYSICAL UNITS
%%%
%%% Please use \unit{} and apply the exponential notation

\end{document}

### 2020-03-23 VegOzone.tex
Outline.

@@
 %% 2-column papers and discussion papers
-\documentclass[journal abbreviation, manuscript]{copernicus}
+\documentclass[bg, manuscript]{copernicus}
 
@@
 
-\title{TEXT}
+\title{The 2018 heatwave and its implications on ozone induced damage on vegetation in a subartic climate}
 
@@
 
-\Author[]{}{}
-\Author[]{}{}
-\Author[]{}{}
+\Author[1]{Stefanie}{Falk}
+\Author[2]{Ane Victoria}{Vollsnes}
+\author[3]{Lisa}{Emberson}
+\Author[1]{Frode}{Stordal}
 
-\affil[]{ADDRESS}
-\affil[]{ADDRESS}
-
+\affil[1]{Department of Geosciences, University of Oslo, Oslo, Norway}
+\affil[2]{Department of Biosciences, University of Oslo, Oslo, Norway}
+\affil[3]{Department of Environment and Geography, University of York, UK}
 %% The [] brackets identify the author with the corresponding affiliation. 1, 2, 3, etc. should be inserted.
@@
 
-\correspondence{NAME (EMAIL)}
+\correspondence{Stefanie Falk (stefanie.falk@geo.uio.no)}
 
-\runningtitle{TEXT}
+\runningtitle{2018 heatwave and implication on ozone induced damage on vegetation in subarctic climate}
 
-\runningauthor{TEXT}
+\runningauthor{Falk et al.}
 
@@
 \begin{abstract}
-TEXT
+  \begin{itemize}
+  \item ozone damage
+  \item present subarctic climate and conditions
+    \begin{itemize}
+    \item[$\rightarrow$] midnight sun and 23\,\unit{h} open stomata (biological details will be published in a compaignend paper)
+    \item[$\rightarrow$] ozone climatology and spring peak
+    \end{itemize}
+  \item future climate and conditions
+    \begin{itemize}
+    \item[$\rightarrow$] increase in heatwaves and dry periodes
+    \item[$\rightarrow$] shift in start of growing season (into vincity of ozone spring peak)
+    \item[$\rightarrow$] increase in tropospheric background ozone
+    \end{itemize}
+  \item present day test case 2018 vs 2019 ozone induced damage at NIBIO research station Svanhovd
+  \item findings and results 
+  \end{itemize}
 \end{abstract}
@@
 \introduction  %% \introduction[modified heading if necessary]
+\label{sec:intro}
+\begin{itemize}
+\item ozone in general
+\item ozone induced damage in vegetation
+\item present subarctic climate and conditions
+  \begin{itemize}
+  \item[$\rightarrow$] midnight sun and 23\,\unit{h} open stomata (biological details will be published in a compaignend paper)
+  \item[$\rightarrow$] ozone climatology and spring peak
+  \end{itemize}
+\item future climate and conditions
+  \begin{itemize}
+  \item[$\rightarrow$] increase in heatwaves and dry periodes
+  \item[$\rightarrow$] shift in start of growing season (into vincity of ozone spring peak)
+  \item[$\rightarrow$] increase in tropospheric background ozone
+  \end{itemize}
+\end{itemize}
+
+\section{Data}
 TEXT
 
-\section{HEADING}
+\subsection{NIBIO research station Svanhovd}
 TEXT
 
-\subsection{HEADING}
+\subsection{Other observational data}
 TEXT
 
-\subsubsection{HEADING}
+\subsection{Model and reanalysis data}
 TEXT
+
+\section{Statistical analysis}
+TEXT
+
+\subsection{Derived climatologies for Svanhovd}
+\begin{itemize}
+\item temperature
+\item precipitation
+\item ozone
+\end{itemize}
+
+\subsection{2018/19 anomalies}
+\begin{itemize}
+\item temperature
+\item precipitation
+\item ozone
+\end{itemize}
+
+\section{DO3SE modeling}
+Model results for natural and semi-natural vegetation (birch, pine, clover)
+\begin{itemize}
+\item climatological ozone impact
+\item 2018
+\item 2019
+\end{itemize}
 
@@
 
-\codeavailability{TEXT} %% use this section when having only software code available
+%\codeavailability{TEXT} %% use this section when having only software code available
 
@@
 
-\codedataavailability{TEXT} %% use this section when having data sets and software code available
+%\codedataavailability{TEXT} %% use this section when having data sets and software code available
 
@@
 
-\videosupplement{TEXT} %% use this section when having video supplements available
+%\videosupplement{TEXT} %% use this section when having video supplements available
 
@@
 
-\begin{thebibliography}{}
+%\begin{thebibliography}{}
 
-\bibitem[AUTHOR(YEAR)]{LABEL1}
-REFERENCE 1
+%\bibitem[AUTHOR(YEAR)]{LABEL1}
+%REFERENCE 1
 
-\bibitem[AUTHOR(YEAR)]{LABEL2}
-REFERENCE 2
+%\bibitem[AUTHOR(YEAR)]{LABEL2}
+%REFERENCE 2
 
-\end{thebibliography}
+%\end{thebibliography}
 
@@
 %%
-%% \bibliographystyle{copernicus}
-%% \bibliography{example.bib}
+\bibliographystyle{copernicus}
+\bibliography{VegOzone.bib}
 %%

### 2020-03-23 VegOzone.tex
Include the site overviews.
Some plots included. Added text to distribute work load.

@@
 %\usepackage{rotating}
+\graphicspath{{pictures/}}
 
@@
 \item ozone in general
-\item ozone induced damage in vegetation
+\item ozone induced damage in vegetation \emph{Contributor: Ane - maybe also ask Aud?}
 \item present subarctic climate and conditions
@@
 TEXT
+\begin{figure}[t]
+\includegraphics[width=8.3cm]{station_map_fennoscandia}
+\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
+\end{figure}
 
 \subsection{NIBIO research station Svanhovd}
-TEXT
 
-\subsection{Other observational data}
-TEXT
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{svanhovd_researchstation}
+
+  \caption{NIBIO research station Svanhovd. Atmospheric monitoring site and ozone garden have been marked in the picture.}
+\end{figure}
+TEXT: Ozone monitoring and ozone garden. \emph{Invite co-authors from NILU and NIBIO?}
+
+\subsection{Observational data}
+Table: EBAS - Station list
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{ozone_timeseries_fenoscandic_obs.png}
+
+  \caption{Timeseries of ozone observations in northern Fennoscadia.
+Data taken from EBAS and ozone monitoring at Svanhovd in the summer seasons 2018/19.
+The hatched area indicates time periods with unsufficient quality control - cite: NILU ozone report 200x.}
+\end{figure*}
 
@@
 TEXT
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{ozone_climatologies}
+
+  \caption{Comparison of observational data with various ozone reanlysis products.}
+\end{figure}
 
@@
 
-\subsection{Derived climatologies for Svanhovd}
+\subsection{Derived climatologies}
 \begin{itemize}

### 2020-03-24 VegOzone.tex
Included plots.

@@
 \Author[2]{Ane Victoria}{Vollsnes}
-\author[3]{Lisa}{Emberson}
+\Author[3]{Lisa}{Emberson}
 \Author[1]{Frode}{Stordal}
@@
   \includegraphics[width=8.3cm]{station_map_fennoscandia}
+  \label{fig:station_map_fennoscandia}
 \caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
@@
   \includegraphics[width=12cm]{ozone_timeseries_fenoscandic_obs.png}
-
+  \label{fig:ozone_timesseries_fenoscandic_obs}
   \caption{Timeseries of ozone observations in northern Fennoscadia.
@@
 \subsection{Model and reanalysis data}
-TEXT
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_climatologies}
-
-  \caption{Comparison of observational data with various ozone reanlysis products.}
-\end{figure}
+Ozone: MACC-rea, CAMS-rea, Copernicus European regional model rea
+Precipitation/Temperature: CRU biascorrected rea
+Ozone uptake / POD: DO3SE
 
@@
 
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs}
+  \label{fig:ozone_climatology_fenoscandic_obs}
+  \caption{Comparison of observational data with various ozone reanlysis products.}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{ozone_climatologies}
+  \label{fig:ozone_climatologies}
+  \caption{Comparison of observational data with various ozone reanlysis products.}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_spline}
+  \label{fig:ozone_climatology_fenoscandic_obs_spline}
+\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_residuals}
+  \label{fig:ozone_climatology_fenoscandic_obs_residuals}
+\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_residuals-Svanvik}
+  \label{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}
+\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_test}
+  \label{fig:ozone_climatology_fenoscandic_obs_test}
+\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_test-Svanvik}
+  \label{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}
+\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{plot_temperature_anomalies_svanvik}
+  \label{fig:plot_temperature_anomalies_svanvik}
+\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{plot_precipitation_anomalies_svanvik}
+  \label{fig:plot_precipitation_anomalies_svanvik}
+\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
+\end{figure}
+
 \subsection{2018/19 anomalies}
@@
 \section{DO3SE modeling}
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{ozone_obs_2018_07}
+  \label{fig:ozone_obs_2018_07}
+\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
+\end{figure}
 Model results for natural and semi-natural vegetation (birch, pine, clover)
@@
 \appendixfigures  %% needs to be added in front of appendix figures
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{greening_season_change_Svanvik}
+  \label{fig:greening_season_change_Svanvik}
+\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{test_fennoscandia}
+  \label{fig:test_fennoscandia}
+\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{density_distribution}
+  \label{fig:density_distribution}
+\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{ozone_obs_2018_07}
+  \label{fig:ozone_obs_2018_rrea}
+\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
+\end{figure}
 

### 2020-03-24 VegOzone.tex
Started to add captions.

@@
   \includegraphics[width=8.3cm]{station_map_fennoscandia}
+  \caption{Cap of the North. Location of ozone observation sites in northern Fennoscandia used within this study.}
   \label{fig:station_map_fennoscandia}
-\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
 \end{figure}
@@
   \includegraphics[width=8.3cm]{svanhovd_researchstation}
-
-  \caption{NIBIO research station Svanhovd. Atmospheric monitoring site and ozone garden have been marked in the picture.}
+  \caption{NIBIO research station Svanhovd close by the settlement of Svanvik, Norway. Atmospheric monitoring site and ozone garden have been marked in the picture.}
+  \label{fig:svanhovd_research_station}
 \end{figure}
@@
 \subsection{Observational data}
-Table: EBAS - Station list
+Table: EBAS - Station list: ID - name - location - country - operational
 \begin{figure*}[t]
@@
   \caption{Timeseries of ozone observations in northern Fennoscadia.
-Data taken from EBAS and ozone monitoring at Svanhovd in the summer seasons 2018/19.
-The hatched area indicates time periods with unsufficient quality control - cite: NILU ozone report 200x.}
+Data taken from EBAS as well as from dedicated ozone monitoring at Svanhovd in the summer seasons 2018/19.
+The hatched areas indicate time periods with unsufficient quality control - \bf{TODO: citation NILU ozone report 200x}.}
 \end{figure*}
@@
   \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs}
+  \caption{Multi-annual mean of daily mean ozone at observation sites in Fennoscandia.
+For comparison also data from Prestbakke (southern Norway) is shown.
+The large spread in case of Svanvik and Janiskoski is due to the lower statistics.
+All northern Fennoscandic climatologies peak in spring, with peak values in April.
+In southern Norway, peak values appear only in May.
+The annual average ozone concentration \chem{\left<[O_3]\right>} at Svanvik is $6.6\,\unit{ppb}$ lower than at the sites at Jergul/Karasjok (NOR), Esrange (SWE), and Pallas (FIN).}
   \label{fig:ozone_climatology_fenoscandic_obs}
-  \caption{Comparison of observational data with various ozone reanlysis products.}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_spline}
+  \caption{Derivation of ozone climatologies.
+Density distributions are shown together with multi-annual mean of daily ozone concentrations \chem{[O_3]}.
+Splines have been fitted though daily mean and max \chem{[O_3]}.
+The climatology of daily mean ozone is shown together with $1\,\sigma$ standard deviation and standard error of the mean.
+Since \chem{[O_3]} are strongly correlated for sites at Jergul/Karasjok (NOR), Esrange (SWE), and Pallas (FIN) (see Fig.~\ref{fig:density_distribution}), data of these have been averaged together to derive a climatology for northern Fennoscandia.
+(a) Northern Fennoscandia; (b) Svanvik}
+  \label{fig:ozone_climatology_fenoscandic_obs_spline}
 \end{figure}
@@
   \includegraphics[width=8.3cm]{ozone_climatologies}
+  \caption{Comparison of climatologies (northern Fennoscandia -- black and southern Fennoscadia -- grey) derived from observational data (EBAS, see Fig.~\ref{fig:ozone_climatology_fenoscandic_obs}) with various ozone reanlysis products (ECMWF: MACC/CAMS reanalysis, Copernicus: Regional model reanalysis ensemble mean).
+As representative for the sites, the nearest neighboring grid point has been chosen.
+The ECMWF products do not reproduce observed surface ozone well.
+The MACC reanalysis is in general too low and displays a erronous seasonal cycle.
+The CAMS reanalysis matches quite well in southern Fennoscandia but does not reproduce the seasonality in northern Fennoscandia.
+The ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in both northern and southern Fennoscadia, though the ozone concentrations in the north are slightly to low.}
   \label{fig:ozone_climatologies}
-  \caption{Comparison of observational data with various ozone reanlysis products.}
 \end{figure}
 
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_spline}
-  \label{fig:ozone_climatology_fenoscandic_obs_spline}
-\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
-\end{figure}
+\subsection{2018/19 anomalies}
+\begin{itemize}
+\item temperature
+\item precipitation
+\item ozone
+\end{itemize}
 
@@
   \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_residuals-Svanvik}
+  \caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
    \label{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}
-\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
 \end{figure}
@@
 \end{figure}
-
-\subsection{2018/19 anomalies}
-\begin{itemize}
-\item temperature
-\item precipitation
-\item ozone
-\end{itemize}
 
@@
   \includegraphics[width=8.3cm]{test_fennoscandia}
-  \label{fig:test_fennoscandia}
+  \label{fig:sunlight_fennoscandia}
 \caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}

### 2020-03-24 VegOzone.tex
Finished captions and spell check.

@@
 
-\title{The 2018 heatwave and its implications on ozone induced damage on vegetation in a subartic climate}
+\title{The 2018 heatwave and its implications on ozone induced damage on vegetation in a subarctic climate}
 
@@
     \begin{itemize}
-    \item[$\rightarrow$] midnight sun and 23\,\unit{h} open stomata (biological details will be published in a compaignend paper)
+    \item[$\rightarrow$] midnight sun and 23\,\unit{h} open stomata (biological details will be published in a component paper)
     \item[$\rightarrow$] ozone climatology and spring peak
@@
     \begin{itemize}
-    \item[$\rightarrow$] increase in heatwaves and dry periodes
-    \item[$\rightarrow$] shift in start of growing season (into vincity of ozone spring peak)
+    \item[$\rightarrow$] increase in heatwaves and dry periods
+    \item[$\rightarrow$] shift in start of growing season (into vicinity of ozone spring peak)
     \item[$\rightarrow$] increase in tropospheric background ozone
@@
   \begin{itemize}
-  \item[$\rightarrow$] midnight sun and 23\,\unit{h} open stomata (biological details will be published in a compaignend paper)
+  \item[$\rightarrow$] midnight sun and 23\,\unit{h} open stomata (biological details will be published in a component paper)
   \item[$\rightarrow$] ozone climatology and spring peak
@@
   \begin{itemize}
-  \item[$\rightarrow$] increase in heatwaves and dry periodes
-  \item[$\rightarrow$] shift in start of growing season (into vincity of ozone spring peak)
+  \item[$\rightarrow$] increase in heatwaves and dry periods
+  \item[$\rightarrow$] shift in start of growing season (into vicinity of ozone spring peak)
   \item[$\rightarrow$] increase in tropospheric background ozone
@@
   \label{fig:ozone_timesseries_fenoscandic_obs}
-  \caption{Timeseries of ozone observations in northern Fennoscadia.
+  \caption{Time series of ozone observations in northern Fennoscandia.
 Data taken from EBAS as well as from dedicated ozone monitoring at Svanhovd in the summer seasons 2018/19.
-The hatched areas indicate time periods with unsufficient quality control - \bf{TODO: citation NILU ozone report 200x}.}
+The hatched areas indicate time periods with insufficient quality control - \bf{TODO: citation NILU ozone report 200x}.}
 \end{figure*}
@@
 Ozone: MACC-rea, CAMS-rea, Copernicus European regional model rea
-Precipitation/Temperature: CRU biascorrected rea
+Precipitation/Temperature: CRU bias corrected rea
 Ozone uptake / POD: DO3SE
@@
   \caption{Multi-annual mean of daily mean ozone at observation sites in Fennoscandia.
-For comparison also data from Prestbakke (southern Norway) is shown.
+For comparison also data from Prestebakke (southern Norway) is shown.
 The large spread in case of Svanvik and Janiskoski is due to the lower statistics.
@@
   \includegraphics[width=8.3cm]{ozone_climatologies}
-  \caption{Comparison of climatologies (northern Fennoscandia -- black and southern Fennoscadia -- grey) derived from observational data (EBAS, see Fig.~\ref{fig:ozone_climatology_fenoscandic_obs}) with various ozone reanlysis products (ECMWF: MACC/CAMS reanalysis, Copernicus: Regional model reanalysis ensemble mean).
+  \caption{Comparison of climatologies (northern Fennoscandia -- black and southern Fennoscandia -- grey) derived from observational data (EBAS, see Fig.~\ref{fig:ozone_climatology_fenoscandic_obs}) with various ozone reanalysis products (ECMWF: MACC/CAMS reanalysis, Copernicus: Regional model reanalysis ensemble mean).
 As representative for the sites, the nearest neighboring grid point has been chosen.
 The ECMWF products do not reproduce observed surface ozone well.
-The MACC reanalysis is in general too low and displays a erronous seasonal cycle.
+The MACC reanalysis is in general too low and displays a erroneous seasonal cycle.
 The CAMS reanalysis matches quite well in southern Fennoscandia but does not reproduce the seasonality in northern Fennoscandia.
-The ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in both northern and southern Fennoscadia, though the ozone concentrations in the north are slightly to low.}
+The ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in both northern and southern Fennoscandia, though the ozone concentrations in the north are slightly to low.}
   \label{fig:ozone_climatologies}
@@
   \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_residuals}
+  \caption{Probability density functions of ozone concentration residuals. 2018 with respect to respective climatology for different sites in Fennoscandia. (a) Esrange; (b) Pallas; (c) Prestebakke.}
   \label{fig:ozone_climatology_fenoscandic_obs_residuals}
-\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
 \end{figure}
@@
   \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_residuals-Svanvik}
-  \caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
+  \caption{Probability density functions of ozone concentration residuals. 2018/19 observations at Svanhovd with respect to derived climatologies for (a) Northern Fennoscandia; (b) Svanvik.}
   \label{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}
@@
   \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_test}
+  \caption{Student's t-test assuming same sample uncertainty in both, climatology and 2018 observations for different sites in Fennoscandia.
+\chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals} are significantly different from zero-hypothesis on the $1\,\sigma$ level.
+(a) Esrange; (b) Pallas; (c) Prestebakke.}
   \label{fig:ozone_climatology_fenoscandic_obs_test}
-\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
 \end{figure}
@@
   \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_test-Svanvik}
+  \caption{Student's t-test assuming same sample uncertainty in both, climatology and 2018/19 observations at Svanhovd with respect to derived climatologies for (a) Northern Fennoscandia; (b) Svanvik.
+\chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}a) are significantly different from zero-hypothesis on the $2\,\sigma$,  $1\,\sigma$ level, respectively.
+\chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}b) are not significantly different from zero-hypothesis.}
   \label{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}
-\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
 \end{figure}
@@
   \includegraphics[width=8.3cm]{plot_temperature_anomalies_svanvik}
+  \caption{Time series over the past 6 years of (a) temperature observed at Svanhovd and climatology derived from CRU bias-corrected reanalysis; (b) Student's t-test of temperature.
+The summer of 2018 has been significantly warmer on the $2-4\,\sigma$ level.
+The summer of 2019 was fairly average but on the colder side.}
   \label{fig:plot_temperature_anomalies_svanvik}
-\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
 \end{figure}
@@
   \includegraphics[width=8.3cm]{plot_precipitation_anomalies_svanvik}
+  \caption{Time series over the past 6 years of (a) precipitation observed at Svanhovd and climatology derived from CRU bias-corrected reanalysis; (b) Student's t-test of precipitation.}
   \label{fig:plot_precipitation_anomalies_svanvik}
-\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
 \end{figure}
@@
 \section{DO3SE modeling}
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_obs_2018_07}
-  \label{fig:ozone_obs_2018_07}
-\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
-\end{figure}
+
 Model results for natural and semi-natural vegetation (birch, pine, clover)
@@
   \includegraphics[width=8.3cm]{greening_season_change_Svanvik}
+  \caption{Estimated shift and prolongation of greening season at Svanhovd over the past two decades based on the method and data described in \citet{GMD:Falk2019}.}
   \label{fig:greening_season_change_Svanvik}
-\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
 \end{figure}
@@
   \includegraphics[width=8.3cm]{test_fennoscandia}
+  \caption{Maximum daily sun shine duration based on geometrical calculations at different sites in Fennoscandia.
+Midnight sun conditions at Jergul/Karasjok and Svanvik prevail from the end of May until the end of July, while at Esrange and Pallas they only prevail from the beginning of June until mid of July.
+For comparison with southern Fennoscandia, Prestebakke is shown.}
   \label{fig:sunlight_fennoscandia}
-\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
 \end{figure}
@@
   \includegraphics[width=8.3cm]{density_distribution}
+  \caption{Probability densities and correlation coefficient for \chem{[O_3]} between different sites in northern Fennoscandia.
+Jergul/Karasjok is well correlated with Esrange, and Pallas.
+Svanvik displays highest correlation with Pallas.
+(a) Jergul/Karasjok--Esrange; (b) Jergul/Karasjok--Pallas; (c) Svanvik--Esrange; (d) Svanvik--Pallas.}
   \label{fig:density_distribution}
-\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
 \end{figure}
@@
   \includegraphics[width=8.3cm]{ozone_obs_2018_07}
+  \caption{Estimate of missing observation \chem{[O_3]} data at Svanhovd in July 2018 based on derived hourly climatologies of northern Fennoscandia, Svanvik, and Copernicus regional model reanalysis ensemble mean.
+(a) Time series; (b) Student's t-test.}
   \label{fig:ozone_obs_2018_rrea}
-\caption{Cap of the North. Reasearch stations with given ozone time series in northern Fennoscandia.}
 \end{figure}

### 2020-03-25 VegOzone.tex
Some text and table added.

@@
   \end{itemize}
+\item meteorological situation 2018
 \end{itemize}
@@
 \section{Data}
+\label{sec:data}
 TEXT
@@
   \includegraphics[width=8.3cm]{station_map_fennoscandia}
-  \caption{Cap of the North. Location of ozone observation sites in northern Fennoscandia used within this study.}
+  \caption{Cap of the North. Location of past and present ozone observation sites in northern Fennoscandia used within this study.}
   \label{fig:station_map_fennoscandia}
@@
 \subsection{NIBIO research station Svanhovd}
+\label{subsec:svanhovd}
+"Svanhovd is located beside the Pasvik river, in the middle of the wedge of Norwegian land separating Russia and Finland in the north.
+It is where the eastern Siberian taiga meets the western Boreal forest.
+[...] Svanhovd was originally established as a State Demonstration and Experimental Farm in the 1930's, in order to expand agriculture in Pasvik valley [...]" - \href{https://www.nibio.no/en/about-eng/addresses/northern-norway/svanhovd}{NIBIO}\\
+\emph{Comment: Invite co-authors from NILU and NIBIO?}
 
@@
   \includegraphics[width=8.3cm]{svanhovd_researchstation}
-  \caption{NIBIO research station Svanhovd close by the settlement of Svanvik, Norway. Atmospheric monitoring site and ozone garden have been marked in the picture.}
+  \caption{NIBIO research station Svanhovd close by the settlement of Svanvik, Norway. Atmospheric monitoring site and ozone garden have been marked. Aerial photography (c) Norges Kartverk}
   \label{fig:svanhovd_research_station}
 \end{figure}
-TEXT: Ozone monitoring and ozone garden. \emph{Invite co-authors from NILU and NIBIO?}
 
-\subsection{Observational data}
-Table: EBAS - Station list: ID - name - location - country - operational
+\subsubsection{Ozone monitoring}
+\label{subsubsec}
+{\bf TODO: (Technical) details about ozone monitoring at Svanhovd - historical periode. Project related monitoring (summer seasons 2018/19) $\rightarrow$ NILU -- Sverre?}
+
+\subsubsection{Ozone garden}
+\label{subsubsec:ozone_garden}
+{\bf TODO: Details about the ozone garden.} \emph{Author: Ane/Aud}
+
+\subsection{Other observational data}
+\label{subsec:obs_data}
+
+\subsubsection{EBAS database}
+\label{subsubsec:ebas}
+
+\begin{table}[t]
+  \caption{Ozone observations in northern Fennoscandia available from EBAS.}
+  \label{tab:ebas_obs}
+\begin{tabular}{lllllrc}
+\tophline
+Name       & Country & ID      & \multicolumn{3}{c}{Location} & Operational\\
+           &         &         & lat             & lon               & alt            &\\
+           &         &         & (\unit{^\circ N}) & (\unit{^\circ E})  & (\unit{m})     &\\
+\middlehline
+Esrange    & SWE     & SE0013R & 67.83           & 21.07             & 475            & 1991 -- present\\
+Janiskoski & RUS     & RU0001R & 68.93           & 28.85             & 118            & 1995 -- 1997\\
+Jergul     & NOR     & NO0030R & 69.45           & 24.60             & 255            & 1997 -- 2011\\
+Karasjok   & NOR     & NO0055R & 69.467          & 25.217            & 333            & 1988 -- 1997\\
+Pallas     & FIN     & FI0096G & 69.97           & 24.12             & 565            & 1995 -- present\\
+Svanvik    & NOR     & NO0047R & 69.45           & 30.03             & 30             & 1986 -- 1996 $^\dagger$\\
+\bottomhline
+\end{tabular}
+\belowtable{$^\dagger$ Exclusive data taking summer seasons 2018/19.} % Table Footnotes
+\end{table}
+
 \begin{figure*}[t]
@@
 
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_climatologies}
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{ozone_climatologies}
   \caption{Comparison of climatologies (northern Fennoscandia -- black and southern Fennoscandia -- grey) derived from observational data (EBAS, see Fig.~\ref{fig:ozone_climatology_fenoscandic_obs}) with various ozone reanalysis products (ECMWF: MACC/CAMS reanalysis, Copernicus: Regional model reanalysis ensemble mean).
@@
   \label{fig:ozone_climatologies}
-\end{figure}
+\end{figure*}
 

### 2020-03-25 VegOzone.tex
Added tables for data sources and some text.

@@
 \label{sec:data}
-TEXT
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{station_map_fennoscandia}
-  \caption{Cap of the North. Location of past and present ozone observation sites in northern Fennoscandia used within this study.}
-  \label{fig:station_map_fennoscandia}
-\end{figure}
+
+In our assessment on ozone damage on vegetation is the subarctic in relation to the 2018 heatwave, we focus on the NIBIO operated local site Svanhovd which lies close by the settlement of Svanvik in the Pasvik valley.
+For the purpose of this study, we evaluate data taken at Svanhovd with observational data (Section~\ref{subsec:svanhovd}) with different monitoring sites in northern Fennoscandia (Section~\ref{subsec:obs_data}).
+Reanalysis data is used for complementation were observational data was not available or temporally sparse (Section~\ref{subsec:model_data}).
 
@@
 It is where the eastern Siberian taiga meets the western Boreal forest.
-[...] Svanhovd was originally established as a State Demonstration and Experimental Farm in the 1930's, in order to expand agriculture in Pasvik valley [...]" - \href{https://www.nibio.no/en/about-eng/addresses/northern-norway/svanhovd}{NIBIO}\\
+[...] Svanhovd was originally established as a State Demonstration and Experimental Farm in the 1930's, in order to expand agriculture in Pasvik valley [...]" - \href{https://www.nibio.no/en/about-eng/addresses/northern-norway/svanhovd}{NIBIO}\\ The research station and locations of the atmospheric monitoring site and the ozone garden are shown in the aerial photography (Fig.~\ref{fig:station_map_fennoscandia}).
+Svanhovd comprises atmospheric monitoring (ozone monitoring exclusively for the summer seasons 2018/19) and observed ozone damage in plants.
 \emph{Comment: Invite co-authors from NILU and NIBIO?}
@@
   \includegraphics[width=8.3cm]{svanhovd_researchstation}
-  \caption{NIBIO research station Svanhovd close by the settlement of Svanvik, Norway. Atmospheric monitoring site and ozone garden have been marked. Aerial photography (c) Norges Kartverk}
+  \caption{NIBIO research station Svanhovd close by the settlement of Svanvik, Norway. Atmospheric monitoring site and ozone garden have been marked. Aerial photography \copyright Norges Kartverk.}
   \label{fig:svanhovd_research_station}
@@
 \label{subsubsec}
-{\bf TODO: (Technical) details about ozone monitoring at Svanhovd - historical periode. Project related monitoring (summer seasons 2018/19) $\rightarrow$ NILU -- Sverre?}
+{\bf TODO: (Technical) details about ozone monitoring at Svanhovd - historical periode. Project related monitoring (summer seasons 2018/19.) $\rightarrow$ NILU -- Sverre?}
 
@@
 \label{subsec:obs_data}
+For the purpose of this study, we are interested in past and present observations of surface ozone concentrations in northern Fennoscandia.
+This limits the available observation data to a handfull of atmospheric monitoring sites with data available from the NILU EBAS database.
+An overview over the geographical locations is given in Fig.~\ref{fig:station_map_fennoscandia}.
 
-\subsubsection{EBAS database}
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{station_map_fennoscandia}
+  \caption{Cap of the North. Location of past and present ozone observation sites in northern Fennoscandia used within this study.}
+  \label{fig:station_map_fennoscandia}
+\end{figure}
+
+\subsubsection{Ozone monitoring}
 \label{subsubsec:ebas}
@@
 \begin{table}[t]
-  \caption{Ozone observations in northern Fennoscandia available from EBAS.}
+  \caption{Past and present ozone observation sites in northern Fennoscandia. Data available on EBAS.}
   \label{tab:ebas_obs}
-\begin{tabular}{lllllrc}
+\begin{tabular}{lllllrl}
 \tophline
@@
 Pallas     & FIN     & FI0096G & 69.97           & 24.12             & 565            & 1995 -- present\\
-Svanvik    & NOR     & NO0047R & 69.45           & 30.03             & 30             & 1986 -- 1996 $^\dagger$\\
+Svanvik    & NOR     & NO0047R & 69.45           & 30.03             & 30             & 1986 -- 1996$^\dagger$\\
 \bottomhline
 \end{tabular}
-\belowtable{$^\dagger$ Exclusive data taking summer seasons 2018/19.} % Table Footnotes
+\belowtable{$^\dagger$ Exclusive ozone monitoring summer seasons 2018/19 operated by NILU.} % Table Footnotes
 \end{table}
@@
   \caption{Time series of ozone observations in northern Fennoscandia.
-Data taken from EBAS as well as from dedicated ozone monitoring at Svanhovd in the summer seasons 2018/19.
+Data taken from EBAS as well as from dedicated ozone monitoring at Svanhovd in the summer seasons 2018/19 (Tab.~\ref{tab_ebas_obs}).
 The hatched areas indicate time periods with insufficient quality control - \bf{TODO: citation NILU ozone report 200x}.}
@@
 
-\subsection{Model and reanalysis data}
-Ozone: MACC-rea, CAMS-rea, Copernicus European regional model rea
+\subsubsection{Atmospheric monitoring}
+Monitoring data from Svanvik/Pasvik.
+Ozone/Temperature (2018/2019): Luftkvalitet.no (NILU), LandbruksMeteorologiskTjeneste (NIBIO)
+
+\subsection{Reanalysis data}
+\label{subsec:model_data}
+Range of reanlysis data.
+\subsubsection{Ozone}
+
+\begin{table}[t]
+\caption{Ozone related ranalysis products used in this study: \href{https://www.ecmwf.int/en/newsletter/158/meteorology/new-cams-global-reanalysis-atmospheric-composition}{Global reanalysis (ECMWF)}; \href{https://www.regional.atmosphere.copernicus.eu/}{ensemble mean of regional reanalysis for Europe (Copernicus)}.}
+\begin{tabular}{llccl}
+\tophline
+Name & Organization & \multicolumn{2}{c}{Resolution} & Time period \\
+&              & spatial & temporal & \\
+\middlehline
+MACC & ECMWF & $0.75^\circ \times 0.75^\circ$ & 3-hourly & 2003 -- 2012\\
+CAMSRA & ECMWF & $0.75^\circ \times 0.75^\circ$ & 3-hourly & 2003 -- 2012$^\dagger$\\
+CAMS Regional Air Quality & Copernicus & $0.1^\circ \times 0.1^\circ$ & 1-hourly & 2014 -- 2018$^\dagger$\\
+\bottomhline
+\end{tabular}
+\belowtable{$^\dagger$ Subset of the data used in this study. The full dataset extends into near real time.} % Table Footnotes
+\end{table}
+
+Ozone: MACC-rea, CAMS-rea, Copernicus European regional model rea ensemble mean
+
+\subsubsection{Temperature and precipitation}
 Precipitation/Temperature: CRU bias corrected rea
-Ozone uptake / POD: DO3SE
 
 \section{Statistical analysis}
-TEXT
+Statistical evaluation of Svanhovd (Svanvik/Pasvik) ozone data in a regional and climatological context.
 

### 2020-03-25 VegOzone.tex
Added some more text. Spell-check.

@@
 \begin{abstract}
-  \begin{itemize}
-  \item ozone damage
-  \item present subarctic climate and conditions
-    \begin{itemize}
-    \item[$\rightarrow$] midnight sun and 23\,\unit{h} open stomata (biological details will be published in a component paper)
-    \item[$\rightarrow$] ozone climatology and spring peak
-    \end{itemize}
-  \item future climate and conditions
-    \begin{itemize}
-    \item[$\rightarrow$] increase in heatwaves and dry periods
-    \item[$\rightarrow$] shift in start of growing season (into vicinity of ozone spring peak)
-    \item[$\rightarrow$] increase in tropospheric background ozone
-    \end{itemize}
-  \item present day test case 2018 vs 2019 ozone induced damage at NIBIO research station Svanhovd
-  \item findings and results 
-  \end{itemize}
+  The summer 2018 and summer 2019 did not differ significantly in the level of ozone.
+On the other hand, the temperatures (precipitation and cloud cover/PAR?) differed quite a lot.
+We also observed ozone-induced visible injuries on clovers in the ozone garden at Svanhovd in 2018, but not in 2019.
+This may be due to differences in uptake of ozone (POD) between the two years.
+Since DO3SE can be run on birch/beech and conifers, we would like to investigate whether the critical levels for these tree species were crossed during the two summer seasons.
+Ozone concentrations for the two missing weeks in July 2018 have been estimated based on the measurements from the Swedish and Finnish ozone measurement stations in the region.
 \end{abstract}
@@
   \end{itemize}
-\item meteorological situation 2018
+\item meteorological situation 2018 vs 2019
 \end{itemize}
@@
 In our assessment on ozone damage on vegetation is the subarctic in relation to the 2018 heatwave, we focus on the NIBIO operated local site Svanhovd which lies close by the settlement of Svanvik in the Pasvik valley.
-For the purpose of this study, we evaluate data taken at Svanhovd with observational data (Section~\ref{subsec:svanhovd}) with different monitoring sites in northern Fennoscandia (Section~\ref{subsec:obs_data}).
-Reanalysis data is used for complementation were observational data was not available or temporally sparse (Section~\ref{subsec:model_data}).
+Throughout the paper, we will use Svanvik synonymously for the atmospheric monitoring site at Svanhovd.
+For the purpose of this study, we evaluate data taken at Svanhovd (Section~\ref{subsec:svanhovd}) with observational data from other monitoring sites in northern Fennoscandia (Section~\ref{subsec:obs_data}).
+Reanalysis data is used as complement where observational data was not available or temporally sparse (Section~\ref{subsec:model_data}).
 
@@
 \label{subsubsec}
-{\bf TODO: (Technical) details about ozone monitoring at Svanhovd - historical periode. Project related monitoring (summer seasons 2018/19.) $\rightarrow$ NILU -- Sverre?}
+{\bf TODO: (Technical) details about ozone monitoring at Svanhovd - historical period. Project related monitoring (summer seasons 2018/19.) $\rightarrow$ NILU -- Sverre?}
 
@@
 \label{subsec:obs_data}
-For the purpose of this study, we are interested in past and present observations of surface ozone concentrations in northern Fennoscandia.
-This limits the available observation data to a handfull of atmospheric monitoring sites with data available from the NILU EBAS database.
-An overview over the geographical locations is given in Fig.~\ref{fig:station_map_fennoscandia}.
+For the purpose of this study, we are interested in past and present observations of surface ozone concentrations and meteorological conditions in northern Fennoscandia.
+This limits the available observational data to a handful of atmospheric monitoring sites.
+An overview over their geographical locations is given in Fig.~\ref{fig:station_map_fennoscandia}.
 
@@
   \includegraphics[width=8.3cm]{station_map_fennoscandia}
-  \caption{Cap of the North. Location of past and present ozone observation sites in northern Fennoscandia used within this study.}
+  \caption{Cap of the North. Locations of past and present ozone observation sites in northern Fennoscandia used within this study. For more details see Tab.~\ref{tab:ebas_obs}.}
   \label{fig:station_map_fennoscandia}
@@
 
-\subsubsection{Ozone monitoring}
+\subsubsection{Surface ozone observations}
 \label{subsubsec:ebas}
+All surface ozone observations used in this study are available from NILU EBAS database except for the data for the 2018/19 summer season at Svanhovd.
+In Table~\ref{tab:ebas_obs}, we give an overview over the exact locations and time range of operational data taking.
 
 \begin{table}[t]
-  \caption{Past and present ozone observation sites in northern Fennoscandia. Data available on EBAS.}
+  \caption{Past and present ozone observation sites in northern Fennoscandia. Data available on \href{http://ebas.nilu.no/}{EBAS}.}
   \label{tab:ebas_obs}
@@
   \caption{Time series of ozone observations in northern Fennoscandia.
-Data taken from EBAS as well as from dedicated ozone monitoring at Svanhovd in the summer seasons 2018/19 (Tab.~\ref{tab_ebas_obs}).
+Data taken from EBAS as well as from dedicated ozone monitoring at Svanhovd in the summer seasons 2018/19 (Tab.~\ref{tab:ebas_obs}).
 The hatched areas indicate time periods with insufficient quality control - \bf{TODO: citation NILU ozone report 200x}.}
@@
 Ozone/Temperature (2018/2019): Luftkvalitet.no (NILU), LandbruksMeteorologiskTjeneste (NIBIO)
+We refrain from doing a full cross-evaluation for these meteorological fields, since this has been done elsewhere {\bf TODO: It's somewhat clear that this is a job of the operators in the monitoring network.
+Find a citation for this.}.
+\begin{itemize}
+\item Temperature
+\item Precipitation
+\item Incoming radiation \emph{Comment to Ane: Same plot that you've included in the Nordic ozone report?}
+\item ...
+\end{itemize}
 
@@
 \label{subsec:model_data}
-Range of reanlysis data.
+Different reanalysis products have been used to evaluate and complement the observational data.
+
 \subsubsection{Ozone}
+\label{subsubsec:ozone_rea}
+There are two global reanalysis products of ECMWF available that include also ozone.
+In addition, higher resolution surface ozone reanalysis for Europe is available from Copernicus service for regional air quality control.
+The details are given in Tab.~\ref{tab:ozone_rea}.
+The covered time periods differ but are sufficient for our purpose of comparison of derived ozone climatologies.
 
 \begin{table}[t]
-\caption{Ozone related ranalysis products used in this study: \href{https://www.ecmwf.int/en/newsletter/158/meteorology/new-cams-global-reanalysis-atmospheric-composition}{Global reanalysis (ECMWF)}; \href{https://www.regional.atmosphere.copernicus.eu/}{ensemble mean of regional reanalysis for Europe (Copernicus)}.}
+  \caption{Ozone related reanalysis products used in this study: \href{https://www.ecmwf.int/en/newsletter/158/meteorology/new-cams-global-reanalysis-atmospheric-composition}{Global reanalysis (ECMWF)}; \href{https://www.regional.atmosphere.copernicus.eu/}{ensemble mean of regional reanalysis for Europe (Copernicus)}.}
+  \label{tab:ozone_rea}
 \begin{tabular}{llccl}
@@
 \end{tabular}
-\belowtable{$^\dagger$ Subset of the data used in this study. The full dataset extends into near real time.} % Table Footnotes
+\belowtable{$^\dagger$ Subset of the data used in this study. The full data set extends into near real time.} % Table Footnotes
 \end{table}
-
-Ozone: MACC-rea, CAMS-rea, Copernicus European regional model rea ensemble mean
 

### 2020-03-26 VegOzone.tex
Some more text in the data section and a new plot dedicaded to monitoring data from Svanvik in 2018/19.

@@
 
-In our assessment on ozone damage on vegetation is the subarctic in relation to the 2018 heatwave, we focus on the NIBIO operated local site Svanhovd which lies close by the settlement of Svanvik in the Pasvik valley.
+In our assessment on ozone damage on vegetation in the subarctic in relation to the 2018 heatwave, we focus on the NIBIO operated local site Svanhovd which lies close by the settlement of Svanvik in the Pasvik valley (Section~\ref{subsec:svanhovd}).
 Throughout the paper, we will use Svanvik synonymously for the atmospheric monitoring site at Svanhovd.
-For the purpose of this study, we evaluate data taken at Svanhovd (Section~\ref{subsec:svanhovd}) with observational data from other monitoring sites in northern Fennoscandia (Section~\ref{subsec:obs_data}).
-Reanalysis data is used as complement where observational data was not available or temporally sparse (Section~\ref{subsec:model_data}).
+The available agrameteorological monitoring data are described in Section~\ref{subsec:atmo_svanvik}.
+In Section~\ref{subsubsec:ozone_garden}, the ozone garden -- a plant-based ozone monitoring -- will be described in more detail.
+For evaluation of the ozone data taken at Svanhovd, we compare with observational data from other monitoring sites in northern Fennoscandia (Section~\ref{subsec:obs_data}).
+Reanalysis data is used as complement where observational data was not available or temporally sparse.
+The chosen data sets are described in Section~\ref{subsec:model_data}.
 
@@
 It is where the eastern Siberian taiga meets the western Boreal forest.
-[...] Svanhovd was originally established as a State Demonstration and Experimental Farm in the 1930's, in order to expand agriculture in Pasvik valley [...]" - \href{https://www.nibio.no/en/about-eng/addresses/northern-norway/svanhovd}{NIBIO}\\ The research station and locations of the atmospheric monitoring site and the ozone garden are shown in the aerial photography (Fig.~\ref{fig:station_map_fennoscandia}).
-Svanhovd comprises atmospheric monitoring (ozone monitoring exclusively for the summer seasons 2018/19) and observed ozone damage in plants.
+[...] Svanhovd was originally established as a State Demonstration and Experimental Farm in the 1930's, in order to expand agriculture in Pasvik valley [...]" - \href{https://www.nibio.no/en/about-eng/addresses/northern-norway/svanhovd}{NIBIO}.\\
+The research station at Svanhovd comprises long-term observations of various agrameteorological variables, i.a.
+$2\,\unit{m}$ temperature, precipitation, global radiation, and soil moisture and temperature at different depths.
+For the purpose of relating observed damage on ozone sensitive plants (e.g.
+clover and tobacco) an ozone monitor had been installed during the growing seasons 2018/19).
+The relevant data are described in more detail Section~\ref{subsubsec:atmo_svanvik}.
+Note that from 1986--1996, ozone monitoring has been conducted operationally at Svanhovd. Details about these historical ozone data will be given in Section~\ref{subsubsec:ebas}.\\
+The locations of the atmospheric monitoring site as well as the ozone garden with respect to the main buildings of the NIBIO research station are marked in the aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}.
 \emph{Comment: Invite co-authors from NILU and NIBIO?}
@@
 
-\subsubsection{Ozone monitoring}
-\label{subsubsec}
-{\bf TODO: (Technical) details about ozone monitoring at Svanhovd - historical period. Project related monitoring (summer seasons 2018/19.) $\rightarrow$ NILU -- Sverre?}
+\subsubsection{Atmospheric monitoring}
+\label{subsubsec:atmo_svanvik}
+Monitoring data of key variables in the growing seasons 2018/19 have been downloaded from \href{luftkvalitet.no} operated by NILU and are shown in Fig.~\ref{fig:data_svanvik_2018_2019}.
+The hatched areas mark times when no ozone data has been taken.
+Note, while the downtime during winter was planed, missing data in two weeks of July 2018 were due to problems in data acquisition.
+For longer time series: LandbruksMeteorologiskTjeneste (NIBIO)
+
+      {\bf TODO: (Technical) details about ozone monitoring at Svanhovd - historical period. Project related monitoring (growing seasons 2018/19.) $\rightarrow$ NILU -- Sverre?}
+
+\emph{Comment: We refrain from doing a full cross-evaluation for these meteorological fields, since this has been done elsewhere.
+It's somewhat clear that this is a job of the operators in the monitoring network.} {\bf TODO: Find a citation for this.}.
+\begin{itemize}
+\item Incoming radiation \emph{Comment to Ane: Same plot that you've included in the Nordic ozone report?}
+\item ...
+\end{itemize}
+
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{data_svanvik_2018_2019}
+  \caption{Observational data from atmospheric monitoring at Svanhovd in 2018/19.
+The hatched areas indicate missing ozone monitoring data.
+(a) Ozone; (b) temperature; (c) daily accumulated precipitation.}
+  \label{fig:data_svanvik_2018_2019}
+\end{figure*}
 
@@
 \label{subsubsec:ebas}
-All surface ozone observations used in this study are available from NILU EBAS database except for the data for the 2018/19 summer season at Svanhovd.
+All surface ozone observations used in this study are available from NILU EBAS database except for the data for the 2018/19 growing season at Svanhovd.
 In Table~\ref{tab:ebas_obs}, we give an overview over the exact locations and time range of operational data taking.
@@
   \end{tabular}
-  \belowtable{$^\dagger$ Exclusive ozone monitoring summer seasons 2018/19 operated by NILU.} % Table Footnotes
+  \belowtable{$^\dagger$ Exclusive ozone monitoring growing seasons 2018/19 operated by NILU.} % Table Footnotes
 \end{table}
@@
   \caption{Time series of ozone observations in northern Fennoscandia.
-Data taken from EBAS as well as from dedicated ozone monitoring at Svanhovd in the summer seasons 2018/19 (Tab.~\ref{tab:ebas_obs}).
+Data taken from EBAS as well as from dedicated ozone monitoring at Svanhovd in the growing seasons 2018/19 (Tab.~\ref{tab:ebas_obs}).
 The hatched areas indicate time periods with insufficient quality control - \bf{TODO: citation NILU ozone report 200x}.}
 \end{figure*}
-
-\subsubsection{Atmospheric monitoring}
-Monitoring data from Svanvik/Pasvik.
-Ozone/Temperature (2018/2019): Luftkvalitet.no (NILU), LandbruksMeteorologiskTjeneste (NIBIO)
-We refrain from doing a full cross-evaluation for these meteorological fields, since this has been done elsewhere {\bf TODO: It's somewhat clear that this is a job of the operators in the monitoring network.
-Find a citation for this.}.
-\begin{itemize}
-\item Temperature
-\item Precipitation
-\item Incoming radiation \emph{Comment to Ane: Same plot that you've included in the Nordic ozone report?}
-\item ...
-\end{itemize}
 

### 2020-03-27 VegOzone.tex
Fixed broken reference.
Ane - added text about stomata and ozone.

@@
 \item ozone induced damage in vegetation \emph{Contributor: Ane - maybe also ask Aud?}
+  Ozone acts as an oxidative stress to plants.
+Its main action is imposed through reactions occuring in the cell walls and cell membranes of mesophyll cells inside the leaves.Ozone enters the leaves through stomata, that are kept open to allow for gas exchange between the leaf air spaces and the exterior air.
+When the stomata are closed, ozone does not enter.
 \item present subarctic climate and conditions
@@
 Throughout the paper, we will use Svanvik synonymously for the atmospheric monitoring site at Svanhovd.
-The available agrameteorological monitoring data are described in Section~\ref{subsec:atmo_svanvik}.
+The available agrameteorological monitoring data are described in Section~\ref{subsubsec:atmo_svanvik}.
 In Section~\ref{subsubsec:ozone_garden}, the ozone garden -- a plant-based ozone monitoring -- will be described in more detail.

### 2020-03-27 VegOzone.tex
Added suggestions by Frode.

@@
 It is where the eastern Siberian taiga meets the western Boreal forest.
-[...] Svanhovd was originally established as a State Demonstration and Experimental Farm in the 1930's, in order to expand agriculture in Pasvik valley [...]" - \href{https://www.nibio.no/en/about-eng/addresses/northern-norway/svanhovd}{NIBIO}.\\
+[...] Svanhovd was originally established as a State Demonstration and Experimental Farm in the 1930's, in order to expand agriculture in Pasvik valley [...]" - \href{https://www.nibio.no/en/about-eng/addresses/northern-norway/svanhovd}{NIBIO webpage, accessed March 2020}.\\
 The research station at Svanhovd comprises long-term observations of various agrameteorological variables, i.a.
@@
   \includegraphics[width=8.3cm]{station_map_fennoscandia}
-  \caption{Cap of the North. Locations of past and present ozone observation sites in northern Fennoscandia used within this study. For more details see Tab.~\ref{tab:ebas_obs}.}
+  \caption{Cap of the North.
+Locations of past and present ozone observation sites in northern Fennoscandia used within this study.
+For more details see Tab.~\ref{tab:ebas_obs}.
+The same color coding for each station as displayed herein will be used in all following figures.}
   \label{fig:station_map_fennoscandia}
@@
 \label{subsubsec:ebas}
-All surface ozone observations used in this study are available from NILU EBAS database except for the data for the 2018/19 growing season at Svanhovd.
+All surface ozone observations used in this study are available from NILU EBAS database.
+Data for the 2018/19 growing season at Svanhovd have not yet been made available \emph{alternative: shall be made available at a later point}.
 In Table~\ref{tab:ebas_obs}, we give an overview over the exact locations and time range of operational data taking.

### 2020-03-30 VegOzone.tex
Removed Prestebakke.

@@
   \caption{Multi-annual mean of daily mean ozone at observation sites in Fennoscandia.
-For comparison also data from Prestebakke (southern Norway) is shown.
 The large spread in case of Svanvik and Janiskoski is due to the lower statistics.
@@
   \includegraphics[width=12cm]{ozone_climatologies}
-  \caption{Comparison of climatologies (northern Fennoscandia -- black and southern Fennoscandia -- grey) derived from observational data (EBAS, see Fig.~\ref{fig:ozone_climatology_fenoscandic_obs}) with various ozone reanalysis products (ECMWF: MACC/CAMS reanalysis, Copernicus: Regional model reanalysis ensemble mean).
+  \caption{Comparison of climatologies (northern Fennoscandia -- black and southern Fennoscandia -- grey) derived from observational data (EBAS, see Fig.~\ref{fig:ozone_climatology_fenoscandic_obs}) with various ozone reanalysis products: (a) ECMWF: MACC reanalysis; (b) ECMWF: CAMS reanalysis; (c) Copernicus: Regional model reanalysis ensemble mean.
 As representative for the sites, the nearest neighboring grid point has been chosen.
@@
   \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_residuals}
-  \caption{Probability density functions of ozone concentration residuals. 2018 with respect to respective climatology for different sites in Fennoscandia. (a) Esrange; (b) Pallas; (c) Prestebakke.}
+  \caption{Probability density functions of ozone concentration residuals. 2018 with respect to respective climatology for different sites in Fennoscandia. (a) Esrange; (b) Pallas.}
   \label{fig:ozone_climatology_fenoscandic_obs_residuals}
@@
 \chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals} are significantly different from zero-hypothesis on the $1\,\sigma$ level.
-(a) Esrange; (b) Pallas; (c) Prestebakke.}
+(a) Esrange; (b) Pallas.}
   \label{fig:ozone_climatology_fenoscandic_obs_test}
@@
   \caption{Maximum daily sun shine duration based on geometrical calculations at different sites in Fennoscandia.
-Midnight sun conditions at Jergul/Karasjok and Svanvik prevail from the end of May until the end of July, while at Esrange and Pallas they only prevail from the beginning of June until mid of July.
-For comparison with southern Fennoscandia, Prestebakke is shown.}
+Midnight sun conditions at Jergul/Karasjok and Svanvik prevail from the end of May until the end of July, while at Esrange and Pallas they only prevail from the beginning of June until mid of July.}
   \label{fig:sunlight_fennoscandia}

### 2020-03-31 VegOzone.tex
Updated text. Included new plots.

@@
 The hatched areas indicate missing ozone monitoring data.
-(a) Ozone; (b) temperature; (c) daily accumulated precipitation.}
+(a) Ozone; (b) temperature; (c) daily accumulated precipitation; (d) hourly global radiation.}
   \label{fig:data_svanvik_2018_2019}
@@
 \label{subsubsec:ebas}
-All surface ozone observations used in this study are available from NILU EBAS database.
-Data for the 2018/19 growing season at Svanhovd have not yet been made available \emph{alternative: shall be made available at a later point}.
+All surface ozone observations used in this study are available from the EBAS database.
 In Table~\ref{tab:ebas_obs}, we give an overview over the exact locations and time range of operational data taking.
@@
 
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{global_rad_signific}
+  \caption{Global radiation at Svanvik.
+Percentage of days above/below $\pm 1\,\sigma$ from climatological mean displayed for each month.
+(a) 2018; (b) 2019.
+Negative deviations from climatology are shown as negative percentaged.
+The annual positive/negative deviation are indicated in the respective corners (right upper/ right lower).}
+  \label{fig:global_rad_signific}
+\end{figure}
+
 \section{DO3SE modeling}
@@
   \label{fig:ozone_obs_2018_rrea}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{global_rad_clim}
+  \caption{Observed global radition at Svanvik.
+The grey lines indicate the $1\,\sigma$ level deducted from the climatology.
+(a) Climatology (hourly maxima, mean, and minima); deviation from climatology (b) 2018; (c) 2019.}
+  \label{fig:global_rad_clim}
 \end{figure}

### 2020-04-01 VegOzone.tex
First draft of large parts of Section 2.

@@
   Ozone acts as an oxidative stress to plants.
-Its main action is imposed through reactions occuring in the cell walls and cell membranes of mesophyll cells inside the leaves.Ozone enters the leaves through stomata, that are kept open to allow for gas exchange between the leaf air spaces and the exterior air.
+Its main action is imposed through reactions occurring in the cell walls and cell membranes of mesophyll cells inside the leaves.Ozone enters the leaves through stomata, that are kept open to allow for gas exchange between the leaf air spaces and the exterior air.
 When the stomata are closed, ozone does not enter.
@@
 Throughout the paper, we will use Svanvik synonymously for the atmospheric monitoring site at Svanhovd.
-The available agrameteorological monitoring data are described in Section~\ref{subsubsec:atmo_svanvik}.
+The available meteorological monitoring data are described in Section~\ref{subsubsec:atmo_svanvik}.
 In Section~\ref{subsubsec:ozone_garden}, the ozone garden -- a plant-based ozone monitoring -- will be described in more detail.
-For evaluation of the ozone data taken at Svanhovd, we compare with observational data from other monitoring sites in northern Fennoscandia (Section~\ref{subsec:obs_data}).
-Reanalysis data is used as complement where observational data was not available or temporally sparse.
-The chosen data sets are described in Section~\ref{subsec:model_data}.
+For evaluation of the ozone data taken at Svanhovd, we compare with observational data from other monitoring sites in northern Fennoscandia and use as well as reanalysis products, i.a., as complement where observational data was not available or spatially sparse, (Section~\ref{subsec:obs_data}).
 
@@
 The research station at Svanhovd comprises long-term observations of various agrameteorological variables, i.a.
-$2\,\unit{m}$ temperature, precipitation, global radiation, and soil moisture and temperature at different depths.
+$2\,\unit{m}$ temperature, precipitation, global irradiance, and soil moisture and temperature at different depths.
+
 For the purpose of relating observed damage on ozone sensitive plants (e.g.
-clover and tobacco) an ozone monitor had been installed during the growing seasons 2018/19).
+clover and tobacco) an ozone monitor had been installed during the growing seasons 2018/19.
 The relevant data are described in more detail Section~\ref{subsubsec:atmo_svanvik}.
-Note that from 1986--1996, ozone monitoring has been conducted operationally at Svanhovd. Details about these historical ozone data will be given in Section~\ref{subsubsec:ebas}.\\
+
 The locations of the atmospheric monitoring site as well as the ozone garden with respect to the main buildings of the NIBIO research station are marked in the aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}.
 \emph{Comment: Invite co-authors from NILU and NIBIO?}
+{\bf TODO: Some more general information about the research station? E.g., what NIBIO us currently researching there, that it also hosts a conference site, and national park center.}
 
@@
 \label{subsubsec:atmo_svanvik}
-Monitoring data of key variables in the growing seasons 2018/19 have been downloaded from \href{luftkvalitet.no} operated by NILU and are shown in Fig.~\ref{fig:data_svanvik_2018_2019}.
-The hatched areas mark times when no ozone data has been taken.
-Note, while the downtime during winter was planed, missing data in two weeks of July 2018 were due to problems in data acquisition.
-For longer time series: LandbruksMeteorologiskTjeneste (NIBIO)
+For the growing seasons 2018/19, monitoring data of key variables (temperature, precipitation, and ozone) have been downloaded from \href{luftkvalitet.no}{luftkvalitet.no} which is operated by NILU.
+
+All agrameterological variables including temperature and precipitation are available from approximately 1992 to present day from LandbruksMeteorologisk Tjeneste provided by NIBIO (\href{https://lmt.nibio.no/}{lmt.nibio.no}, note the station name here is Pasvik).
 
@@
 It's somewhat clear that this is a job of the operators in the monitoring network.} {\bf TODO: Find a citation for this.}.
-\begin{itemize}
-\item Incoming radiation \emph{Comment to Ane: Same plot that you've included in the Nordic ozone report?}
-\item ...
-\end{itemize}
+
+Relevant data for the growing season 2018/19 are shown in Fig.~\ref{fig:data_svanvik_2018_2019}.
+The hatched areas mark times when no ozone data has been taken.
+Note, while the downtime during winter was planed, missing data in two weeks of July 2018 were due to problems in data acquisition.
+
+Ozone concentrations \chem{[O_3]} measured in $2\,\unit{m}$ height above ground are averaged hourly.
+As can be seen in Fig.~\ref{fig:data_svanvik_2018_2019}a), \chem{[O_3]} peaks in spring (April/May) and reaches its minimum in late summer (July/August).
+The spring peak has not been captured in 2019, for data acquisition started later than in 2018.
+In 2018, high summer ozone concentrations ($\chem{[O_3]}> 40\,\unit{ppb}$) were observed in July.
+A part of these peak values may, however, coincident with the time period of missing data.
+In the summer of 2019, ozone concentrations stayed below the threshold of $40\,\unit{ppb}$.
+
+Hourly averaged $2\,\unit{m}$ temperatures below $-20\,\unit{^\circ C}$ were observed in winter (January/February), while high temperatures (above $20\,\unit{^\circ C}$) occurred in summer (July), though more regularly in 2018 than in 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}b).
+In 2018, temperature regularly rose above freezing only in May, while in 2019 this occurred already early in March/April.
+2019 saw a very cold November with temperatures regularly below $-10\unit{^\circ C}$.
+
+More rain events with accumulated daily precipitation $\sum_d \mathrm{Precip}$ above $10\,\unit{mm}$ occurred in the summer of 2018 compared to 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}c).
+$\sum_d \mathrm{Precip} \ge 20\,\unit{mm}$ was only breached twice in 2018 and once in 2019.
+
+In 2018, global irradiance $Q_0$ (Fig.~\ref{fig:data_svanvik_2018_2019}d) displays more sequentially high values in spring (May) and summer (July) compared to 2019, while June 2019 saw more consecutive high irradiance than 2018.
 
@@
 The hatched areas indicate missing ozone monitoring data.
-(a) Ozone; (b) temperature; (c) daily accumulated precipitation; (d) hourly global radiation.}
+(a) Hourly averaged ozone concentration; (b) hourly averaged temperature; (c) daily accumulated precipitation; (d) hourly averaged global irradiance.}
   \label{fig:data_svanvik_2018_2019}
@@
 \label{subsubsec:ozone_garden}
-      {\bf TODO: Details about the ozone garden.} \emph{Author: Ane/Aud}
+      {\bf TODO: Details about the ozone garden.} \emph{Author: Ane/Aud -- pictures are welcome, just upload them on overleaf or send them.}
 
-\subsection{Other observational data}
+\subsection{Other data}
 \label{subsec:obs_data}
-For the purpose of this study, we are interested in past and present observations of surface ozone concentrations and meteorological conditions in northern Fennoscandia.
-This limits the available observational data to a handful of atmospheric monitoring sites.
+To determine the significance of the conditions in 2018, we are interested not only in present (2018/19) surface ozone concentrations at Svanvik but also in long term observations and a climatology derived thereof.
+While data for the period 1986--1996 exists for Svanvik, no ozone measurements have been conducted at Svanhovd for the past $14\,\unit{years}$.
+Hence, we also have to have a look at ozone observations at other sites in northern Fennoscandia.
+In regard to ozone, available observational data are limited to a handful of atmospheric monitoring sites.
 An overview over their geographical locations is given in Fig.~\ref{fig:station_map_fennoscandia}.
+The data is descibed in Section~\ref{subsubsec:ebas}.
+As additional complement to observational data which was not available at all times or spatially sparse, we also look at verious ozone reanalysis products (Section~\ref{subsubsec:ozone_rea}).
 
@@
 \label{subsubsec:ebas}
-All surface ozone observations used in this study are available from the EBAS database.
+All surface ozone observations used in this study are available from the EBAS database operated by NILU (\href{http://ebas.nilu.no/}{ebas.nilu.no}).
 In Table~\ref{tab:ebas_obs}, we give an overview over the exact locations and time range of operational data taking.
+Note that from 1986--1996, ozone monitoring has been conducted operationally by NILU at Svanhovd.
 
@@
 
-\subsection{Reanalysis data}
-\label{subsec:model_data}
+\subsubsection{Ozone reanalysis}
+\label{subsubsec:ozone_rea}
+
 Different reanalysis products have been used to evaluate and complement the observational data.
 
-\subsubsection{Ozone}
-\label{subsubsec:ozone_rea}
 There are two global reanalysis products of ECMWF available that include also ozone.
@@
 
-\subsubsection{Temperature and precipitation}
-Precipitation/Temperature: CRU bias corrected rea
-
 \section{Statistical analysis}
@@
 \subsection{Derived climatologies}
-\begin{itemize}
-\item temperature
-\item precipitation
-\item ozone
-\end{itemize}
+\label{subsec:climatologies}
+
+\subsubsection{Temperature and precipitation}
+\label{subsubsec:clim_temp_prec}
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{climatediagram}
+  \caption{Climate diagram following Walter and Lieth.
+The diagram is based on climatological data for Svanvik/Pasvik (1992--2017) downloaded from LandbruksMeteorologiske Tjeneste.
+Monthly average temperatures (red line) are displayed with standard error of mean as error band.
+Average monthly accumulated precipitation (blue bars) is shown with standard deviation of mean as error bars.
+Months with average temperatures below freezing are denoted with a star.}
+  \label{fig:climatediagram}
+\end{figure}
+
+\subsubsection{Ozone}
+\label{subsubsec:clim_ozone}
 
@@
 Since \chem{[O_3]} are strongly correlated for sites at Jergul/Karasjok (NOR), Esrange (SWE), and Pallas (FIN) (see Fig.~\ref{fig:density_distribution}), data of these have been averaged together to derive a climatology for northern Fennoscandia.
-(a) Northern Fennoscandia; (b) Svanvik}
+(a) Northern Fennoscandia; (b) Svanvik.}
   \label{fig:ozone_climatology_fenoscandic_obs_spline}
@@
   \includegraphics[width=8.3cm]{plot_temperature_anomalies_svanvik}
-  \caption{Time series over the past 6 years of (a) temperature observed at Svanhovd and climatology derived from CRU bias-corrected reanalysis; (b) Student's t-test of temperature.
+  \caption{{\bf TODO: Replace by analysis based on Svanvik observed climatology.} Time series over the past 6 years of (a) temperature observed at Svanhovd and climatology derived from CRU bias-corrected reanalysis; (b) Student's t-test of temperature.
 The summer of 2018 has been significantly warmer on the $2-4\,\sigma$ level.
@@
   \includegraphics[width=8.3cm]{plot_precipitation_anomalies_svanvik}
-  \caption{Time series over the past 6 years of (a) precipitation observed at Svanhovd and climatology derived from CRU bias-corrected reanalysis; (b) Student's t-test of precipitation.}
+  \caption{{\bf TODO: Replace by analysis based on Svanvik observed climatology.} Time series over the past 6 years of (a) precipitation observed at Svanhovd and climatology derived from CRU bias-corrected reanalysis; (b) Student's t-test of precipitation.}
   \label{fig:plot_precipitation_anomalies_svanvik}
@@
   \includegraphics[width=8.3cm]{global_rad_signific}
-  \caption{Global radiation at Svanvik.
+  \caption{Global irradiance at Svanvik.
 Percentage of days above/below $\pm 1\,\sigma$ from climatological mean displayed for each month.
 (a) 2018; (b) 2019.
-Negative deviations from climatology are shown as negative percentaged.
+Negative deviations from climatology are shown as negative percentage.
 The annual positive/negative deviation are indicated in the respective corners (right upper/ right lower).}
@@
   \includegraphics[width=8.3cm]{global_rad_clim}
-  \caption{Observed global radition at Svanvik.
+  \caption{Observed global irradiance at Svanvik.
 The grey lines indicate the $1\,\sigma$ level deducted from the climatology.

### 2020-04-02 VegOzone.tex
Updated plots and text in accordance.

@@
 
-\subsubsection{Temperature and precipitation}
+\subsubsection{Temperature, precipitation, and global irradiance}
 \label{subsubsec:clim_temp_prec}
@@
 \subsection{2018/19 anomalies}
-\begin{itemize}
-\item temperature
-\item precipitation
-\item ozone
-\end{itemize}
+\label{subsec:anomalies}
+
+\subsubsection{Temperature, precipitation, and global irradiance}
+\label{subsubsec:anomal_tpq}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{temperature_signific}
+  \caption{Temperature at Svanvik.
+Percentage of days warmer/colder than $\pm 1\,\sigma$ from climatological mean displayed for each month.
+(a) 2018; (b) 2019.
+Negative deviations from climatology are shown as negative percentage.
+The annual positive/negative deviations are indicated in the respective corners (right upper/ right lower).
+The summer of 2018 was significantly warmer than average, especially May and July.
+The summer of 2019 was fairly average but significantly warmer than average in April and colder in July/August.}
+  \label{fig:plot_temperature_anomalies_svanvik}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{precipitation_signific}
+  \caption{Precipitation at Svanvik.
+Percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ from climatological mean displayed for each month.
+(a) 2018; (b) 2019.
+Negative deviations from climatology are shown as negative percentage.
+The annual positive/negative deviations are indicated in the respective corners (right upper/ right lower).
+The summer of 2018 was significantly drier than in 2019.}
+  \label{fig:plot_precipitation_anomalies_svanvik}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{global_rad_signific}
+  \caption{Global irradiance at Svanvik.
+Percentage of days above/below $\pm 1\,\sigma$ from climatological mean displayed for each month.
+(a) 2018; (b) 2019.
+Negative deviations from climatology are shown as negative percentage.
+The annual positive/negative deviations are indicated in the respective corners (right upper/ right lower).
+In summer 2018 (May/July), gobal irradiance was significantly higher than on average, while 2019 was generally an average year.}
+  \label{fig:global_rad_signific}
+\end{figure}
+
+\subsubsection{Ozone}
+\label{subsubsec:anomal_ozone}
 
@@
   \label{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}
-\end{figure}
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{plot_temperature_anomalies_svanvik}
-  \caption{{\bf TODO: Replace by analysis based on Svanvik observed climatology.} Time series over the past 6 years of (a) temperature observed at Svanhovd and climatology derived from CRU bias-corrected reanalysis; (b) Student's t-test of temperature.
-The summer of 2018 has been significantly warmer on the $2-4\,\sigma$ level.
-The summer of 2019 was fairly average but on the colder side.}
-  \label{fig:plot_temperature_anomalies_svanvik}
-\end{figure}
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{plot_precipitation_anomalies_svanvik}
-  \caption{{\bf TODO: Replace by analysis based on Svanvik observed climatology.} Time series over the past 6 years of (a) precipitation observed at Svanhovd and climatology derived from CRU bias-corrected reanalysis; (b) Student's t-test of precipitation.}
-  \label{fig:plot_precipitation_anomalies_svanvik}
-\end{figure}
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{global_rad_signific}
-  \caption{Global irradiance at Svanvik.
-Percentage of days above/below $\pm 1\,\sigma$ from climatological mean displayed for each month.
-(a) 2018; (b) 2019.
-Negative deviations from climatology are shown as negative percentage.
-The annual positive/negative deviation are indicated in the respective corners (right upper/ right lower).}
-  \label{fig:global_rad_signific}
 \end{figure}

### 2020-04-03 VegOzone.tex
Some more text in Section 2. Started with section 3.

@@
 In Section~\ref{subsubsec:ozone_garden}, the ozone garden -- a plant-based ozone monitoring -- will be described in more detail.
-For evaluation of the ozone data taken at Svanhovd, we compare with observational data from other monitoring sites in northern Fennoscandia and use as well as reanalysis products, i.a., as complement where observational data was not available or spatially sparse, (Section~\ref{subsec:obs_data}).
+Other data used for evaluation of the ozone data taken at Svanhovd is described in Section~\ref{subsec:other_data}.
+We compare with observational data from other monitoring sites in northern Fennoscandia (Section~\ref{subsubsec:ebas}) and use as well reanalysis products (Section~\ref{subsubsec:ozone_rea}) as complement when observational data was not available.
 
@@
 The research station at Svanhovd comprises long-term observations of various agrameteorological variables, i.a.
-$2\,\unit{m}$ temperature, precipitation, global irradiance, and soil moisture and temperature at different depths.
+$2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths.
 
@@
 clover and tobacco) an ozone monitor had been installed during the growing seasons 2018/19.
-The relevant data are described in more detail Section~\ref{subsubsec:atmo_svanvik}.
+The relevant data are described in more detail in Section~\ref{subsubsec:atmo_svanvik}.
 
@@
 
-All agrameterological variables including temperature and precipitation are available from approximately 1992 to present day from LandbruksMeteorologisk Tjeneste provided by NIBIO (\href{https://lmt.nibio.no/}{lmt.nibio.no}, note the station name here is Pasvik).
+All agrameterological variables including temperature and precipitation are available from September 1992 to present day (LandbruksMeteorologisk Tjeneste provided by NIBIO -- \href{https://lmt.nibio.no/}{lmt.nibio.no}, note the station name here is Pasvik).
 
-{\bf TODO: (Technical) details about ozone monitoring at Svanhovd - historical period. Project related monitoring (growing seasons 2018/19.) $\rightarrow$ NILU -- Sverre?}
+{\bf TODO: Some more (technical) details about ozone and atmospheric monitoring at Svanhovd. $\rightarrow$ Sverre?}
 
@@
 \label{subsubsec:ozone_garden}
-      {\bf TODO: Details about the ozone garden.} \emph{Author: Ane/Aud -- pictures are welcome, just upload them on overleaf or send them.}
+      {\bf TODO: Details about the ozone garden and observed damage on ozone sensitive plants.
+We will need to come back to this in the discussion section after the Do3se modelling part.} \emph{Author: Ane/Aud -- pictures are welcome, just upload them on overleaf or send them to me.}
 
 \subsection{Other data}
-\label{subsec:obs_data}
-To determine the significance of the conditions in 2018, we are interested not only in present (2018/19) surface ozone concentrations at Svanvik but also in long term observations and a climatology derived thereof.
-While data for the period 1986--1996 exists for Svanvik, no ozone measurements have been conducted at Svanhovd for the past $14\,\unit{years}$.
-Hence, we also have to have a look at ozone observations at other sites in northern Fennoscandia.
-In regard to ozone, available observational data are limited to a handful of atmospheric monitoring sites.
-An overview over their geographical locations is given in Fig.~\ref{fig:station_map_fennoscandia}.
+\label{subsec:other_data}
+To determine the significance of the ozone conditions in 2018, we have to derive a surface ozone concentration climatology at Svanvik.
+From 1986--1996 ozone monitoring has been operated by NILU at Svanhovd, but no ozone measurements have been conducted for the past $14\,\unit{years}$.
+Due to probabel changes in tropospheric background ozone and a reduction of summertime peak values related to precursor that fell under airquality regulations, we also have to have a look at ozone observations at other sites in northern Fennoscandia.
+In regard to ozone, available observational data are sparse in this region.
+An overview over the geographical locations of conducted atmospheric monitoring in past and present is given in Fig.~\ref{fig:station_map_fennoscandia}.
 The data is descibed in Section~\ref{subsubsec:ebas}.
-As additional complement to observational data which was not available at all times or spatially sparse, we also look at verious ozone reanalysis products (Section~\ref{subsubsec:ozone_rea}).
+As additional complement to observational data which is not available at all times and spatially sparse, we also look at verious ozone reanalysis products (Section~\ref{subsubsec:ozone_rea}).
 
@@
 In Table~\ref{tab:ebas_obs}, we give an overview over the exact locations and time range of operational data taking.
-Note that from 1986--1996, ozone monitoring has been conducted operationally by NILU at Svanhovd.
+All stations with available long term observation are locate on higher ground than Svanhovd which will result in higher surface ozone concentrations compared to data taken at Svanvik as ozone abundance decreases with altitude in the absence of local precursors.
+{\bf TODO: The following has to be confirmed and ellaborated on!} The vegetation at all sites is similar and consists mainly of pine forests, birch, and heath.
+
+Data from before the 1990s is available solely from Svanvik and Karasjok.
+These data, however, do not follow the high quality standard procedures implemented nowadays and have to be treated with care {\bf TODO: citation NILU ozone report 200x}.
+Especially, ozone monitors did not undergo a regular re-calibration.
+This leads likely to drifts in the observed data and may impose a false periodicity.
+In 1997, the ozone monitoring at Jergul had been relocated downstream the river Karasjokka to a site close by Karasjok.
+Although the same equipment has been used, this might have intruduced a shift in the data series denotes as Jergul/Karasjok in the following.
+Only few data have been taken upstream the Pasvik river at Janiskovski, hence these will not be taken into further consideration.
+As can be seen in Fig.~\ref{fig:ozone_timesseries_fenoscandic_obs}, there is an overlapp of about $4\,\unit{years}$ between data taken at Esrange, Jergul/Karasjok, and Svanvik.
+The hatched areas herin indicate the time periods with insufficient quality control as mentioned above.
 
@@
   \includegraphics[width=12cm]{ozone_timeseries_fenoscandic_obs.png}
+  \caption{Time series of ozone observations in northern Fennoscandia (Tab.~\ref{tab:ebas_obs}).
+Data taken from EBAS.
+The hatched areas indicate time periods with insufficient quality control.
+\bf{TODO: citation NILU ozone report 200x}.}
   \label{fig:ozone_timesseries_fenoscandic_obs}
-  \caption{Time series of ozone observations in northern Fennoscandia.
-Data taken from EBAS as well as from dedicated ozone monitoring at Svanhovd in the growing seasons 2018/19 (Tab.~\ref{tab:ebas_obs}).
-The hatched areas indicate time periods with insufficient quality control - \bf{TODO: citation NILU ozone report 200x}.}
 \end{figure*}
@@
 
-Different reanalysis products have been used to evaluate and complement the observational data.
-
-There are two global reanalysis products of ECMWF available that include also ozone.
-In addition, higher resolution surface ozone reanalysis for Europe is available from Copernicus service for regional air quality control.
-The details are given in Tab.~\ref{tab:ozone_rea}.
-The covered time periods differ but are sufficient for our purpose of comparison of derived ozone climatologies.
+Details of renalysis data sets refered to in this work are listed in Table~\ref{tab:ozone_rea}.
+There are two global reanalysis products available at the European Centre for Medium-Range Weather Forecasts (ECMWF) that include atmospheric tracers, in particular ozone (MACC and CAMSRA).
+The temporal as well as spatial resolution of these reanalysis products is rather corse: 3-hourly and $0.75^\circ\times 0.75^\circ$ or roughly $(78.1\times 83.4)\,\unit{km}$ at the location of Svanvik, respectively.
+In addition, higher resolution surface ozone reanalysis for whole Europe is available from Copernicus service for regional air quality control.
+The regional model reanalysis ensemble mean is at fairly high spatial and temporal resolution compared to the data from ECMWF: $0.1^\circ\times 0.1^\circ$ or roughly $(10.4\times 11.1)\,\unit{km}$ and 1-hourly, respectively.
+The covered time periods differ and no data are available for before the turn of the millenium.
+The CAMSRA and CAMS Regional Air Quality ensemble mean are available in near realtime, but only former covers a period of sufficient length for climate analysis (2003 -- present).
+Their quality with respect to observations will be discussed in Section~\ref{subsubsec:clim_ozone}.
+Without too much foreshadowing, the results prevented us from fetching the whole CAMSRA data.
 
@@
 
-\section{Statistical analysis}
-Statistical evaluation of Svanhovd (Svanvik/Pasvik) ozone data in a regional and climatological context.
+\section{Statistical analysis of atmospheric key variables}
+In this section, we will evaluate atmospheric conditions in the growing seasons of 2018/19 by statistical means.
+We will derive climatologies for both, temperature, precipitation, global irradiance, and ozone for Svanvik (Section~\ref{subsec:climatologies}).
+The ozone measurements will be put into a climatological as well as regional context.
+We will look at anomalies of the atmospheric key variable mentioned above (Section~\ref{subsec:anomalies}).
 
@@
 Percentage of days warmer/colder than $\pm 1\,\sigma$ from climatological mean displayed for each month.
-(a) 2018; (b) 2019.
 Negative deviations from climatology are shown as negative percentage.
-The annual positive/negative deviations are indicated in the respective corners (right upper/ right lower).
+The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
+The dashed lines indicate the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
+Year displayed in panel (a) 2018; (b) 2019.
 The summer of 2018 was significantly warmer than average, especially May and July.
@@
 Percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ from climatological mean displayed for each month.
-(a) 2018; (b) 2019.
 Negative deviations from climatology are shown as negative percentage.
-The annual positive/negative deviations are indicated in the respective corners (right upper/ right lower).
+The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
+The dashed lines indicate the expected percentage of values falling above $\pm\frac{1}{2}\,\sigma$ if a normal distribution is assumed ($30.8\,\unit{\%}$).
+Year displayed in panel (a) 2018; (b) 2019.
 The summer of 2018 was significantly drier than in 2019.}
@@
 Percentage of days above/below $\pm 1\,\sigma$ from climatological mean displayed for each month.
-(a) 2018; (b) 2019.
 Negative deviations from climatology are shown as negative percentage.
-The annual positive/negative deviations are indicated in the respective corners (right upper/ right lower).
+The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
+The dashed lines indicate the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
+Year displayed in panel (a) 2018; (b) 2019.
 In summer 2018 (May/July), gobal irradiance was significantly higher than on average, while 2019 was generally an average year.}

### 2020-04-06 VegOzone.tex
Finished large parts of section 3.1. Missing ozone climatology and spell check.

@@
 When the stomata are closed, ozone does not enter.
+\item metrices of ozone: aot40 and pod
 \item present subarctic climate and conditions
@@
 
-In our assessment on ozone damage on vegetation in the subarctic in relation to the 2018 heatwave, we focus on the NIBIO operated local site Svanhovd which lies close by the settlement of Svanvik in the Pasvik valley (Section~\ref{subsec:svanhovd}).
+In our assessment of ozone damage on vegetation in the subarctic in relation to the 2018 heatwave, we focus on the NIBIO operated local site Svanhovd which lies close by the settlement of Svanvik in the Pasvik valley (Section~\ref{subsec:svanhovd}).
 Throughout the paper, we will use Svanvik synonymously for the atmospheric monitoring site at Svanhovd.
@@
 \emph{Comment: Invite co-authors from NILU and NIBIO?}
-{\bf TODO: Some more general information about the research station? E.g., what NIBIO us currently researching there, that it also hosts a conference site, and national park center.}
+     {\bf TODO: Some more general information about the research station? E.g., what NIBIO us currently researching there, that it also hosts a conference center and national park center.}
+     "The close proximity to Russia was a dominant factor when the new centre was built in 1992 and established as Svanhovd Environment Centre.
+The centre building has later been extended and modernised.
+It contains offices, a conference centre, a national park centre, a brown bear exhibition and a modern laboratory which is also accredited for DNA analysis of brown bear.
+The centre is surrounded by a botanical garden, one of three such gardens open to the public north of the Polar Circle.
+
+     NIBIO Svanhovd in Pasvik valley is strongly committed to environmental research cooperation in the Barents region.
+The station disseminates knowledge and performs R\&D based upon the natural environment of this region.
+It also runs an up-to date DNA-laboratory, a national park centre, a botanical garden and a conference centre.
+Svanhovd has extensive cooperation with all countries in the Barents region."  - \href{https://www.nibio.no/en/about-eng/addresses/northern-norway/svanhovd}{NIBIO webpage, accessed April 2020}.\\
 
@@
 As can be seen in Fig.~\ref{fig:data_svanvik_2018_2019}a), \chem{[O_3]} peaks in spring (April/May) and reaches its minimum in late summer (July/August).
-The spring peak has not been captured in 2019, for data acquisition started later than in 2018.
+The spring peak has not been captured completely in 2019, for data acquisition started later than in 2018.
 In 2018, high summer ozone concentrations ($\chem{[O_3]}> 40\,\unit{ppb}$) were observed in July.
 A part of these peak values may, however, coincident with the time period of missing data.
-In the summer of 2019, ozone concentrations stayed below the threshold of $40\,\unit{ppb}$.
+In the summer of 2019, ozone concentrations stayed below the threshold of $40\,\unit{ppb}$ most of the times.
 
@@
 \label{subsec:other_data}
-To determine the significance of the ozone conditions in 2018, we have to derive a surface ozone concentration climatology at Svanvik.
+To determine the significance of the ozone conditions in 2018, we have to derive a surface ozone concentration climatology for Svanvik.
 From 1986--1996 ozone monitoring has been operated by NILU at Svanhovd, but no ozone measurements have been conducted for the past $14\,\unit{years}$.
-Due to probabel changes in tropospheric background ozone and a reduction of summertime peak values related to precursor that fell under airquality regulations, we also have to have a look at ozone observations at other sites in northern Fennoscandia.
+Due to probabel changes in northern hemispheric tropospheric background ozone and a reduction of summertime peak values related to ozone precursors that fell under airquality regulations {\bf TODO: Citations!}, we also have to have a look at ozone observations at other sites in northern Fennoscandia.
 In regard to ozone, available observational data are sparse in this region.
-An overview over the geographical locations of conducted atmospheric monitoring in past and present is given in Fig.~\ref{fig:station_map_fennoscandia}.
+An overview over the geographical locations of conducted atmospheric monitoring in the past and present is given in Fig.~\ref{fig:station_map_fennoscandia}.
 The data is descibed in Section~\ref{subsubsec:ebas}.
@@
 In Table~\ref{tab:ebas_obs}, we give an overview over the exact locations and time range of operational data taking.
-All stations with available long term observation are locate on higher ground than Svanhovd which will result in higher surface ozone concentrations compared to data taken at Svanvik as ozone abundance decreases with altitude in the absence of local precursors.
-{\bf TODO: The following has to be confirmed and ellaborated on!} The vegetation at all sites is similar and consists mainly of pine forests, birch, and heath.
+All stations with available long term observation are locate on higher ground than Svanhovd which will result in higher surface ozone concentrations compared to data taken at Svanvik as ozone abundance decreases with decreasing altitude in the absence of local precursors.
+{\bf TODO: Citation!} {\bf TODO: The following has to be confirmed and ellaborated on!} The vegetation at all sites is similar and consists mainly of pine forests, birch, and heath.
 
@@
 This leads likely to drifts in the observed data and may impose a false periodicity.
-In 1997, the ozone monitoring at Jergul had been relocated downstream the river Karasjokka to a site close by Karasjok.
+In 1997, the ozone monitoring at Jergul was relocated downstream the river Karasjokka to a site close by Karasjok.
 Although the same equipment has been used, this might have intruduced a shift in the data series denotes as Jergul/Karasjok in the following.
@@
 There are two global reanalysis products available at the European Centre for Medium-Range Weather Forecasts (ECMWF) that include atmospheric tracers, in particular ozone (MACC and CAMSRA).
-The temporal as well as spatial resolution of these reanalysis products is rather corse: 3-hourly and $0.75^\circ\times 0.75^\circ$ or roughly $(78.1\times 83.4)\,\unit{km}$ at the location of Svanvik, respectively.
+The temporal as well as spatial resolution of these reanalysis products is rather coarse: 3-hourly and $0.75^\circ\times 0.75^\circ$ or roughly $78.1\,\unit{km}\times 83.4\,\unit{km}$ at the location of Svanvik, respectively.
 In addition, higher resolution surface ozone reanalysis for whole Europe is available from Copernicus service for regional air quality control.
-The regional model reanalysis ensemble mean is at fairly high spatial and temporal resolution compared to the data from ECMWF: $0.1^\circ\times 0.1^\circ$ or roughly $(10.4\times 11.1)\,\unit{km}$ and 1-hourly, respectively.
+The regional model reanalysis ensemble mean is at fairly high spatial and temporal resolution compared to the data from ECMWF: $0.1^\circ\times 0.1^\circ$ or roughly $10.4\,\unit{km}\times 11.1\,\unit{km}$ and 1-hourly, respectively.
 The covered time periods differ and no data are available for before the turn of the millenium.
@@
 \label{subsubsec:clim_temp_prec}
+The location of Svanvik at $69.45\,\unit{^\circ N}$ suggests a subarctic climate.
+The derived climate diagram following Walter and Lieth shown in Fig.~\ref{fig:climatediagram} supports this very well.
+The diagram is based on climatological data from Svanvik (1992--2012) downloaded from LandbruksMeteorologiske Tjeneste.
+Monthly averaged temperatures (red line) are displayed with standard error of mean as error band.
+Averaged monthly accumulated precipitation (blue bars) is shown with standard deviation of mean as error bars.
+Months with average temperatures below freezing are denoted with a star.
+As can be seen, temperatures stay below freezing for 5 consequative months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for K\"{o}ppens climate classifcation of a regular subarctic climate (Dfc).
+The highest monthly average temperature is $(13.1\pm 1.1)\,\unit{^\circ C}$ in July and the lowest $(-12.8\pm 2.0)\,\unit{^\circ C}$ in February.
+In the time period 1992--2012, the coldest measured temperature was $-45.2\,\unit{^\circ C}$ observed January 27, 1999, while the highest temperature on July 16 the same year was $29.4\,\unit{^\circ C}$.
+
+Winter and spring (November--April, except for March) are relatively dry ($\sum_m \mathrm{Precip} < 20\,\unit{mm}$).
+The driest month is January with $\sum_m \mathrm{Precip} = (16.7\pm 3.0)\,\unit{mm}$.
+The most precipitation occurres in the summer months, with a highest monthly averaged accumulated precipitation $\sum_m \mathrm{Precip} = (58.5\pm 9.2)\,\unit{mm}$ in August.
+The average annual accumulated precipitation given with standard error of mean is $(383\pm 86)\,\unit{mm}$.
+Precipitation shows a large veriablity throughout the years.
+
+A climatology derived for global irradiance $Q_0$ is shown in Fig.~\ref{fig:global_rad_clim}a) for both, climatological average of hourly maximum $\left<Q_0^\mathrm{max}\right>$, minimum $\left<Q_0^\mathrm{min}\right>$, and mean $\left<Q_0^\mathrm{mean}\right>$ irradiance observed at Svanvik.
+In accordance to the geometrically derived sunshine duration shown in Fig.~\ref{fig:sunlight_fennoscandia}, $\left<Q_0^\mathrm{max}\right>$ approaches zero in December/January commonly known as polar night and reaches its maximum in June/July commonly referred to as midnight sun conditions.
+The highest observed $\left<Q_0^\mathrm{max}\right>$ is about $800\,\unit{W\,m^{-2}}$, while on average $\left<Q_0^\mathrm{mean}\right>$ it is about $450\,\unit{W\,m^{-2}}$ (regularly reached in May--July).
+The maximum of the climatological minimum $\left<Q_0^\mathrm{min}\right>$ lies at about $200\,\unit{W\,m^{-2}}$.
+
 \begin{figure}[t]
@@
   \caption{Climate diagram following Walter and Lieth.
-The diagram is based on climatological data for Svanvik/Pasvik (1992--2017) downloaded from LandbruksMeteorologiske Tjeneste.
-Monthly average temperatures (red line) are displayed with standard error of mean as error band.
-Average monthly accumulated precipitation (blue bars) is shown with standard deviation of mean as error bars.
+The diagram is based on climatological data for Svanvik/Pasvik (1992--2012) downloaded from LandbruksMeteorologiske Tjeneste.
+Monthly averaged temperatures (red line) are displayed with standard error of mean as error band.
+Averaged monthly accumulated precipitation (blue bars) is shown with standard deviation of mean as error bars.
 Months with average temperatures below freezing are denoted with a star.}
@@
 \label{subsubsec:clim_ozone}
+The multi-annual means of daily mean ozone, for simplicity here without standard deviations, at observation sites in northern Fennoscandia (refer to Table~\ref{tab:ebas_obs} for details) are shown in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs}.
+All climatologies peak in spring, with peak values of about $48\,\unit{ppb}$ in late April.
+The annual average ozone concentration \chem{\left<[O_3]\right>} declines throughout the growing season until it reaches its minimum in late summer/fall ($24-30\,\unit{ppb}$).
+The turning points lie approximately in June and November.
+The ozone climatology laggs about 3 months behind the temperature climatology.
+At Svanvik, \chem{\left<[O_3]\right>} is $6.6\,\unit{ppb}$ lower than on sites at Jergul/Karasjok (NOR), Esrange (SWE), and Pallas (FIN).
+The large spread in case of Svanvik and Janiskoski can be explained due to the lower overall statistics.
 
@@
 The large spread in case of Svanvik and Janiskoski is due to the lower statistics.
-All northern Fennoscandic climatologies peak in spring, with peak values in April.
-In southern Norway, peak values appear only in May.
+All climatologies peak in spring, with peak values in April.
 The annual average ozone concentration \chem{\left<[O_3]\right>} at Svanvik is $6.6\,\unit{ppb}$ lower than at the sites at Jergul/Karasjok (NOR), Esrange (SWE), and Pallas (FIN).}
@@
 \end{figure}
+
+The correlation of \chem{[O_3]} between Esrange, Pallas, and Jergul/Karasjok is faily high (Fig.~\ref{fig:density_distribution}a, b), hence we can put those data series together to derive a general ozone climatology for northern Fennoscandia at similar altitudes.
+In Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}, the 2 dimensional density distributions for both, the thus defined northern Fennoscandia (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}a) and Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}) are shown.
+Darker colors indicate higher prohablity to observe these values.
+On top of the density distributions, 10 days averages of daily mean, daily maximum, and daily minimum \chem{[O_3]} are displayed together with the $1 \sigma$ uncertainties.
+For the daily mean the standard error of mean is shown in addition (albeit hardly visible).
+From the northern Fennoscandic climatology, one can deduce several features.
+The maximum of the 10 days average minimum ozone concentration peaks about $20\,\unit{days}$ earlier than the average mean and average maximum concentrations.
+From minimum concentration to maximum concentration it takes about $8.5\,\unit{months}$ while from maximum to minimum it only takes about $4.5\,\unit{months}$.
+The maximum of the average maximum ozone concentration is at $52\,\unit{ppb}$ (late April) and the minimum of the average minimum ocurres at $15\,\unit{ppb}$ (beginning of September).
+The steep rise of \chem{[O_3]} in spring coincident with increasing temperatures as well as increase in global irradiance.
 
@@
   \includegraphics[width=12cm]{ozone_climatologies}
-  \caption{Comparison of climatologies (northern Fennoscandia -- black and southern Fennoscandia -- grey) derived from observational data (EBAS, see Fig.~\ref{fig:ozone_climatology_fenoscandic_obs}) with various ozone reanalysis products: (a) ECMWF: MACC reanalysis; (b) ECMWF: CAMS reanalysis; (c) Copernicus: Regional model reanalysis ensemble mean.
+  \caption{Comparison of climatologies derived from observational data (EBAS, see Fig.~\ref{fig:ozone_climatology_fenoscandic_obs}) with various ozone reanalysis products: (a) ECMWF: MACC reanalysis; (b) ECMWF: CAMS reanalysis; (c) Copernicus: Regional model reanalysis ensemble mean.
 As representative for the sites, the nearest neighboring grid point has been chosen.
@@
 The MACC reanalysis is in general too low and displays a erroneous seasonal cycle.
-The CAMS reanalysis matches quite well in southern Fennoscandia but does not reproduce the seasonality in northern Fennoscandia.
-The ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in both northern and southern Fennoscandia, though the ozone concentrations in the north are slightly to low.}
+The CAMS reanalysis matches better but does not reproduce the seasonality in northern Fennoscandia.
+The ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in northern Fennoscandia, though the modelled ozone concentrations in the north are slightly to low.}
   \label{fig:ozone_climatologies}
@@
 \appendixfigures  %% needs to be added in front of appendix figures
+
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{greening_season_change_Svanvik}
-  \caption{Estimated shift and prolongation of greening season at Svanhovd over the past two decades based on the method and data described in \citet{GMD:Falk2019}.}
-  \label{fig:greening_season_change_Svanvik}
+  \includegraphics[width=8.3cm]{global_rad_clim}
+  \caption{Observed global irradiance at Svanvik.
+The cyan lines indicate the $1\,\sigma$ level deducted from the climatology.
+(a) Climatology (hourly maxima, mean, and minima); deviation from climatology (b) 2018; (c) 2019.}
+  \label{fig:global_rad_clim}
 \end{figure}
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{global_rad_clim}
-  \caption{Observed global irradiance at Svanvik.
-The grey lines indicate the $1\,\sigma$ level deducted from the climatology.
-(a) Climatology (hourly maxima, mean, and minima); deviation from climatology (b) 2018; (c) 2019.}
-  \label{fig:global_rad_clim}
+  \includegraphics[width=8.3cm]{greening_season_change_Svanvik}
+  \caption{Estimated shift and prolongation of greening season at Svanhovd over the past two decades based on the method and data described in \citet{GMD:Falk2019}.}
+  \label{fig:greening_season_change_Svanvik}
 \end{figure}

### 2020-04-06 VegOzone.tex
Finished first draft of Sec. 3.1. No spell-check done.

@@
 \label{subsubsec:clim_ozone}
+In the following, we will look at derived ozone climatologies from different perspectives.
 The multi-annual means of daily mean ozone, for simplicity here without standard deviations, at observation sites in northern Fennoscandia (refer to Table~\ref{tab:ebas_obs} for details) are shown in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs}.
@@
 
-The correlation of \chem{[O_3]} between Esrange, Pallas, and Jergul/Karasjok is faily high (Fig.~\ref{fig:density_distribution}a, b), hence we can put those data series together to derive a general ozone climatology for northern Fennoscandia at similar altitudes.
-In Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}, the 2 dimensional density distributions for both, the thus defined northern Fennoscandia (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}a) and Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}) are shown.
+The correlation of \chem{[O_3]} between Esrange, Pallas, and Jergul/Karasjok is fairly high ($r^2\approx 0.8$, Fig.~\ref{fig:density_distribution}a, b), hence we can put those data series together to derive a general ozone climatology for northern Fennoscandia.
+In Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}, the 2 dimensional density distributions for the thus defined northern Fennoscandia (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}a) and Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) are shown.
 Darker colors indicate higher prohablity to observe these values.
-On top of the density distributions, 10 days averages of daily mean, daily maximum, and daily minimum \chem{[O_3]} are displayed together with the $1 \sigma$ uncertainties.
-For the daily mean the standard error of mean is shown in addition (albeit hardly visible).
+On top of the density distributions, $10\,\unit{days}$ averages of daily mean \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, daily maximum \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}}, and daily minimum \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, min}} are displayed together with the $1 \sigma$ uncertainties, respectively.
+In addition, \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} is also shown with standard error of mean (albeit small and hardly visible).
+Splines have been fitted through the data for easier intercomparison with ranalyis data.
+
 From the northern Fennoscandic climatology, one can deduce several features.
-The maximum of the 10 days average minimum ozone concentration peaks about $20\,\unit{days}$ earlier than the average mean and average maximum concentrations.
-From minimum concentration to maximum concentration it takes about $8.5\,\unit{months}$ while from maximum to minimum it only takes about $4.5\,\unit{months}$.
-The maximum of the average maximum ozone concentration is at $52\,\unit{ppb}$ (late April) and the minimum of the average minimum ocurres at $15\,\unit{ppb}$ (beginning of September).
-The steep rise of \chem{[O_3]} in spring coincident with increasing temperatures as well as increase in global irradiance.
+The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, min}} peaks about $20\,\unit{days}$ earlier than the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, and \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}}.
+With respect of the periodicity of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, it takes about $8.5\,\unit{months}$ from minimum concentration to maximum concentration, while from maximum to minimum it only takes about $4.5\,\unit{months}$.
+The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}} is at $52\,\unit{ppb}$ in late April and the minimum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, min}} ocurres at $15\,\unit{ppb}$ in the beginning of September.
+The steep rise of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} in spring coincident with increasing temperatures as well as increase in global irradiance.
+The decrease in \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} coincides with the estimated onset of the greening season based on the agricultural rule of thumb that the daily average temperature has to be above $5\,\unit{^\circ C}$ for $5\,\unit{days}$ conseguatively (day of the year 110--150, Fig~\ref{fig:greening_season_change_Svanvik}).
+
+As mentioned above, the average ozone concentration at Svanvik is in general lower than at the other sites which is in league with a decrease in tropospheric ozone with decreasing altitude.
+The densitiy distribution for Svanvik and the respective \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, and \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}} (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) display a similar pattern as the northern Fennoscandic climatology.
+Only does the spring peak apparently occure slightly earlier (in the beginning of April).
 
@@
 \end{figure}
+
+Finally, we shall look at the comparison of climatologies derived from observational data ozone and the various reanalysis products (Fig.~\ref{fig:ozone_climatologies}).
+As representative for the sites, the nearest neighboring grid point has been chosen.
+The ECMWF products do not reproduce observed surface ozone well.
+The MACC reanalysis is in general too low and displays a erroneous seasonal cycle.
+The CAMS reanalysis matches better but does not reproduce the seasonality in northern Fennoscandia.
+The ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in northern Fennoscandia, though the modelled ozone concentrations in the north are slightly to low.
 

### 2020-04-07 VegOzone.tex
Ysterdays text spell-check.

@@
 When the stomata are closed, ozone does not enter.
-\item metrices of ozone: aot40 and pod
+\item metrics of ozone: aot40 and pod
 \item present subarctic climate and conditions
@@
 It is where the eastern Siberian taiga meets the western Boreal forest.
-[...] Svanhovd was originally established as a State Demonstration and Experimental Farm in the 1930's, in order to expand agriculture in Pasvik valley [...]" - \href{https://www.nibio.no/en/about-eng/addresses/northern-norway/svanhovd}{NIBIO webpage, accessed March 2020}.\\
+[...] Svanhovd was originally established as a State Demonstration and Experimental Farm in the 1930's, in order to expand agriculture in Pasvik valley [...]" - \href{https://www.nibio.no/en/about-eng/addresses/northern-norway/svanhovd}{NIBIO web page, accessed March 2020}.\\
 The research station at Svanhovd comprises long-term observations of various agrameteorological variables, i.a.
@@
 It also runs an up-to date DNA-laboratory, a national park centre, a botanical garden and a conference centre.
-Svanhovd has extensive cooperation with all countries in the Barents region."  - \href{https://www.nibio.no/en/about-eng/addresses/northern-norway/svanhovd}{NIBIO webpage, accessed April 2020}.\\
+Svanhovd has extensive cooperation with all countries in the Barents region."  - \href{https://www.nibio.no/en/about-eng/addresses/northern-norway/svanhovd}{NIBIO web page, accessed April 2020}.\\
 
@@
 
-All agrameterological variables including temperature and precipitation are available from September 1992 to present day (LandbruksMeteorologisk Tjeneste provided by NIBIO -- \href{https://lmt.nibio.no/}{lmt.nibio.no}, note the station name here is Pasvik).
+All agrameteorological variables including temperature and precipitation are available from September 1992 to present day (LandbruksMeteorologisk Tjeneste provided by NIBIO -- \href{https://lmt.nibio.no/}{lmt.nibio.no}, note the station name here is Pasvik).
 
@@
       {\bf TODO: Details about the ozone garden and observed damage on ozone sensitive plants.
-We will need to come back to this in the discussion section after the Do3se modelling part.} \emph{Author: Ane/Aud -- pictures are welcome, just upload them on overleaf or send them to me.}
+We will need to come back to this in the discussion section after the Do3se modeling part.} \emph{Author: Ane/Aud -- pictures are welcome, just upload them on overleaf or send them to me.}
 
@@
 From 1986--1996 ozone monitoring has been operated by NILU at Svanhovd, but no ozone measurements have been conducted for the past $14\,\unit{years}$.
-Due to probabel changes in northern hemispheric tropospheric background ozone and a reduction of summertime peak values related to ozone precursors that fell under airquality regulations {\bf TODO: Citations!}, we also have to have a look at ozone observations at other sites in northern Fennoscandia.
+Due to probable changes in northern hemispheric tropospheric background ozone and a reduction of summertime peak values related to ozone precursors that fell under air quality regulations {\bf TODO: Citations!}, we also have to have a look at ozone observations at other sites in northern Fennoscandia.
 In regard to ozone, available observational data are sparse in this region.
 An overview over the geographical locations of conducted atmospheric monitoring in the past and present is given in Fig.~\ref{fig:station_map_fennoscandia}.
-The data is descibed in Section~\ref{subsubsec:ebas}.
-As additional complement to observational data which is not available at all times and spatially sparse, we also look at verious ozone reanalysis products (Section~\ref{subsubsec:ozone_rea}).
+The data is described in Section~\ref{subsubsec:ebas}.
+As additional complement to observational data which is not available at all times and spatially sparse, we also look at various ozone reanalysis products (Section~\ref{subsubsec:ozone_rea}).
 
@@
 All stations with available long term observation are locate on higher ground than Svanhovd which will result in higher surface ozone concentrations compared to data taken at Svanvik as ozone abundance decreases with decreasing altitude in the absence of local precursors.
-{\bf TODO: Citation!} {\bf TODO: The following has to be confirmed and ellaborated on!} The vegetation at all sites is similar and consists mainly of pine forests, birch, and heath.
+{\bf TODO: Citation!} {\bf TODO: The following has to be confirmed and elaborated on!} The vegetation at all sites is similar and consists mainly of pine forests, birch, and heath.
 
@@
 In 1997, the ozone monitoring at Jergul was relocated downstream the river Karasjokka to a site close by Karasjok.
-Although the same equipment has been used, this might have intruduced a shift in the data series denotes as Jergul/Karasjok in the following.
+Although the same equipment has been used, this might have introduced a shift in the data series denotes as Jergul/Karasjok in the following.
 Only few data have been taken upstream the Pasvik river at Janiskovski, hence these will not be taken into further consideration.
-As can be seen in Fig.~\ref{fig:ozone_timesseries_fenoscandic_obs}, there is an overlapp of about $4\,\unit{years}$ between data taken at Esrange, Jergul/Karasjok, and Svanvik.
-The hatched areas herin indicate the time periods with insufficient quality control as mentioned above.
+As can be seen in Fig.~\ref{fig:ozone_timesseries_fenoscandic_obs}, there is an overlap of about $4\,\unit{years}$ between data taken at Esrange, Jergul/Karasjok, and Svanvik.
+The hatched areas herein indicate the time periods with insufficient quality control as mentioned above.
 
@@
 
-Details of renalysis data sets refered to in this work are listed in Table~\ref{tab:ozone_rea}.
+Details of reanalysis data sets referred to in this work are listed in Table~\ref{tab:ozone_rea}.
 There are two global reanalysis products available at the European Centre for Medium-Range Weather Forecasts (ECMWF) that include atmospheric tracers, in particular ozone (MACC and CAMSRA).
@@
 The regional model reanalysis ensemble mean is at fairly high spatial and temporal resolution compared to the data from ECMWF: $0.1^\circ\times 0.1^\circ$ or roughly $10.4\,\unit{km}\times 11.1\,\unit{km}$ and 1-hourly, respectively.
-The covered time periods differ and no data are available for before the turn of the millenium.
-The CAMSRA and CAMS Regional Air Quality ensemble mean are available in near realtime, but only former covers a period of sufficient length for climate analysis (2003 -- present).
+The covered time periods differ and no data are available for before the turn of the millennium.
+The CAMSRA and CAMS Regional Air Quality ensemble mean are available in near real time, but only former covers a period of sufficient length for climate analysis (2003 -- present).
 Their quality with respect to observations will be discussed in Section~\ref{subsubsec:clim_ozone}.
@@
 Months with average temperatures below freezing are denoted with a star.
-As can be seen, temperatures stay below freezing for 5 consequative months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for K\"{o}ppens climate classifcation of a regular subarctic climate (Dfc).
+As can be seen, temperatures stay below freezing for 5 consecutive months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for K\"{o}ppens climate classification of a regular subarctic climate (Dfc).
 The highest monthly average temperature is $(13.1\pm 1.1)\,\unit{^\circ C}$ in July and the lowest $(-12.8\pm 2.0)\,\unit{^\circ C}$ in February.
@@
 The driest month is January with $\sum_m \mathrm{Precip} = (16.7\pm 3.0)\,\unit{mm}$.
-The most precipitation occurres in the summer months, with a highest monthly averaged accumulated precipitation $\sum_m \mathrm{Precip} = (58.5\pm 9.2)\,\unit{mm}$ in August.
+The most precipitation occurs in the summer months, with a highest monthly averaged accumulated precipitation $\sum_m \mathrm{Precip} = (58.5\pm 9.2)\,\unit{mm}$ in August.
 The average annual accumulated precipitation given with standard error of mean is $(383\pm 86)\,\unit{mm}$.
-Precipitation shows a large veriablity throughout the years.
+Precipitation shows a large variability throughout the years.
 
@@
 The turning points lie approximately in June and November.
-The ozone climatology laggs about 3 months behind the temperature climatology.
+The ozone climatology lags about 3 months behind the temperature climatology.
 At Svanvik, \chem{\left<[O_3]\right>} is $6.6\,\unit{ppb}$ lower than on sites at Jergul/Karasjok (NOR), Esrange (SWE), and Pallas (FIN).
@@
 In Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}, the 2 dimensional density distributions for the thus defined northern Fennoscandia (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}a) and Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) are shown.
-Darker colors indicate higher prohablity to observe these values.
+Darker colors indicate higher probability to observe these values.
 On top of the density distributions, $10\,\unit{days}$ averages of daily mean \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, daily maximum \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}}, and daily minimum \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, min}} are displayed together with the $1 \sigma$ uncertainties, respectively.
 In addition, \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} is also shown with standard error of mean (albeit small and hardly visible).
-Splines have been fitted through the data for easier intercomparison with ranalyis data.
+Splines have been fitted through the data for easier inter-comparison with reanalysis data.
 
@@
 With respect of the periodicity of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, it takes about $8.5\,\unit{months}$ from minimum concentration to maximum concentration, while from maximum to minimum it only takes about $4.5\,\unit{months}$.
-The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}} is at $52\,\unit{ppb}$ in late April and the minimum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, min}} ocurres at $15\,\unit{ppb}$ in the beginning of September.
+The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}} is at $52\,\unit{ppb}$ in late April and the minimum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, min}} occurs at $15\,\unit{ppb}$ in the beginning of September.
 The steep rise of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} in spring coincident with increasing temperatures as well as increase in global irradiance.
-The decrease in \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} coincides with the estimated onset of the greening season based on the agricultural rule of thumb that the daily average temperature has to be above $5\,\unit{^\circ C}$ for $5\,\unit{days}$ conseguatively (day of the year 110--150, Fig~\ref{fig:greening_season_change_Svanvik}).
+The decrease in \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} coincides with the estimated onset of the greening season based on the agricultural rule of thumb that the daily average temperature has to be above $5\,\unit{^\circ C}$ for $5\,\unit{days}$ consecutively (day of the year 110--150, Fig~\ref{fig:greening_season_change_Svanvik}).
 
 As mentioned above, the average ozone concentration at Svanvik is in general lower than at the other sites which is in league with a decrease in tropospheric ozone with decreasing altitude.
-The densitiy distribution for Svanvik and the respective \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, and \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}} (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) display a similar pattern as the northern Fennoscandic climatology.
-Only does the spring peak apparently occure slightly earlier (in the beginning of April).
+The density distribution for Svanvik and the respective \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, and \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}} (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) display a similar pattern as the northern Fennoscandic climatology.
+Only does the spring peak apparently occur slightly earlier (in the beginning of April).
 
@@
 The CAMS reanalysis matches better but does not reproduce the seasonality in northern Fennoscandia.
-The ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in northern Fennoscandia, though the modelled ozone concentrations in the north are slightly to low.
+The ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in northern Fennoscandia, though the modeled ozone concentrations in the north are slightly to low.
 
@@
 The CAMS reanalysis matches better but does not reproduce the seasonality in northern Fennoscandia.
-The ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in northern Fennoscandia, though the modelled ozone concentrations in the north are slightly to low.}
+The ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in northern Fennoscandia, though the modeled ozone concentrations in the north are slightly to low.}
   \label{fig:ozone_climatologies}
@@
 Year displayed in panel (a) 2018; (b) 2019.
-In summer 2018 (May/July), gobal irradiance was significantly higher than on average, while 2019 was generally an average year.}
+In summer 2018 (May/July), global irradiance was significantly higher than on average, while 2019 was generally an average year.}
   \label{fig:global_rad_signific}

### 2020-04-07 VegOzone.tex
Finished most of Sec 3, only ozone anomalies missing; may need to do the same plot as for the other variables and place the ditailed histograms in an appendix or supplement. Spell-check done.

@@
   \item[$\rightarrow$] midnight sun and 23\,\unit{h} open stomata (biological details will be published in a component paper)
-  \item[$\rightarrow$] ozone climatology and spring peak
+  \item[$\rightarrow$] ozone climatology and spring peak {\bf TODO: Cite review article - find newer articles.}
   \end{itemize}
@@
 
-In our assessment of ozone damage on vegetation in the subarctic in relation to the 2018 heatwave, we focus on the NIBIO operated local site Svanhovd which lies close by the settlement of Svanvik in the Pasvik valley (Section~\ref{subsec:svanhovd}).
+In our assessment of ozone damage on vegetation in the subarctic in relation to the 2018 heatwave, we focus on the NIBIO operated Environment Centre Svanhovd which lies close by the settlement of Svanvik in the Pasvik valley (Section~\ref{subsec:svanhovd}).
 Throughout the paper, we will use Svanvik synonymously for the atmospheric monitoring site at Svanhovd.
@@
 
-\subsection{NIBIO research station Svanhovd}
+\subsection{NIBIO Environment Centre Svanhovd}
 \label{subsec:svanhovd}
@@
 It is where the eastern Siberian taiga meets the western Boreal forest.
-[...] Svanhovd was originally established as a State Demonstration and Experimental Farm in the 1930's, in order to expand agriculture in Pasvik valley [...]" - \href{https://www.nibio.no/en/about-eng/addresses/northern-norway/svanhovd}{NIBIO web page, accessed March 2020}.\\
-The research station at Svanhovd comprises long-term observations of various agrameteorological variables, i.a.
+[...] Svanhovd was originally established as a State Demonstration and Experimental Farm in the 1930's, in order to expand agriculture in Pasvik valley [...]" - \href{https://www.nibio.no/en/about-eng/addresses/northern-norway/svanhovd}{NIBIO web page, accessed March 2020}.
+
+The locations of the atmospheric monitoring site as well as the ozone garden with respect to the main buildings of the Svanhovd Environment Centre are marked in the aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}.
+The Centre is located south of the settlement of Svanvik, where as the border to Russia, the Pasvik river, lies only a few hundred meters to the east.
+The original farm buildings, protected by law, are located in the lower center, while the Svanhovd Environment Centre follows to the north-east.
+The Centre hosts offices, a conference center, a national park center, a brown bear exhibition as well as a modern laboratory for DNA analysis.
+Left of the Centre and north of the original farm buildings a botanical garden can be found.
+It is one amongst three such gardens north of the Polar Circle.
+The Svanhovd Environment Centre focuses on research of the special environment of the region, but also comprises long-term observations of various agrameteorological variables, i.a.
 $2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths.
@@
 For the purpose of relating observed damage on ozone sensitive plants (e.g.
-clover and tobacco) an ozone monitor had been installed during the growing seasons 2018/19.
+clover and tobacco) to local ozone concentrations, an ozone monitor had been installed during the growing seasons 2018/19.
 The relevant data are described in more detail in Section~\ref{subsubsec:atmo_svanvik}.
 
-The locations of the atmospheric monitoring site as well as the ozone garden with respect to the main buildings of the NIBIO research station are marked in the aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}.
 \emph{Comment: Invite co-authors from NILU and NIBIO?}
-     {\bf TODO: Some more general information about the research station? E.g., what NIBIO us currently researching there, that it also hosts a conference center and national park center.}
-     "The close proximity to Russia was a dominant factor when the new centre was built in 1992 and established as Svanhovd Environment Centre.
-The centre building has later been extended and modernised.
-It contains offices, a conference centre, a national park centre, a brown bear exhibition and a modern laboratory which is also accredited for DNA analysis of brown bear.
-The centre is surrounded by a botanical garden, one of three such gardens open to the public north of the Polar Circle.
-
-     NIBIO Svanhovd in Pasvik valley is strongly committed to environmental research cooperation in the Barents region.
-The station disseminates knowledge and performs R\&D based upon the natural environment of this region.
-It also runs an up-to date DNA-laboratory, a national park centre, a botanical garden and a conference centre.
-Svanhovd has extensive cooperation with all countries in the Barents region."  - \href{https://www.nibio.no/en/about-eng/addresses/northern-norway/svanhovd}{NIBIO web page, accessed April 2020}.\\
 
@@
   \includegraphics[width=8.3cm]{svanhovd_researchstation}
-  \caption{NIBIO research station Svanhovd close by the settlement of Svanvik, Norway. Atmospheric monitoring site and ozone garden have been marked. Aerial photography \copyright Norges Kartverk.}
+  \caption{NIBIO Environment Centre Svanhovd close by the settlement of Svanvik, Norway. Atmospheric monitoring site and ozone garden have been marked. Aerial photography \copyright Norges Kartverk.}
   \label{fig:svanhovd_research_station}
@@
 In 2018, global irradiance $Q_0$ (Fig.~\ref{fig:data_svanvik_2018_2019}d) displays more sequentially high values in spring (May) and summer (July) compared to 2019, while June 2019 saw more consecutive high irradiance than 2018.
+In both years, the maximum recorded $Q_0$ was roughly $750\,\unit{W\,m^{-2}}$ reached only in June.
 
@@
 In Table~\ref{tab:ebas_obs}, we give an overview over the exact locations and time range of operational data taking.
-All stations with available long term observation are locate on higher ground than Svanhovd which will result in higher surface ozone concentrations compared to data taken at Svanvik as ozone abundance decreases with decreasing altitude in the absence of local precursors.
-{\bf TODO: Citation!} {\bf TODO: The following has to be confirmed and elaborated on!} The vegetation at all sites is similar and consists mainly of pine forests, birch, and heath.
+All stations with available long term observations are locate on higher ground than Svanhovd which will result in higher surface ozone concentrations in comparison as ozone abundance, in general, decreases with decreasing altitude in the absence of local precursors.
+{\bf TODO: Citation!} {\bf TODO: The following has to be confirmed and elaborated on!
+Ane?} The vegetation at all sites is similar and consists mainly of pine forests, birch, and heath.
 
@@
 In 1997, the ozone monitoring at Jergul was relocated downstream the river Karasjokka to a site close by Karasjok.
-Although the same equipment has been used, this might have introduced a shift in the data series denotes as Jergul/Karasjok in the following.
-Only few data have been taken upstream the Pasvik river at Janiskovski, hence these will not be taken into further consideration.
+Although the same equipment has been used, this might have introduced a breaking point in the combined data series denoted as Jergul/Karasjok in the following which would be of a problem if this was the sole data to compare with.
+Only few data have been taken upstream the Pasvik river at Janiskovski, hence these will not be taken into further consideration, but are shown for completion.
 As can be seen in Fig.~\ref{fig:ozone_timesseries_fenoscandic_obs}, there is an overlap of about $4\,\unit{years}$ between data taken at Esrange, Jergul/Karasjok, and Svanvik.
-The hatched areas herein indicate the time periods with insufficient quality control as mentioned above.
+The hatched areas indicate time periods with insufficient quality control as mentioned above.
 
@@
 A climatology derived for global irradiance $Q_0$ is shown in Fig.~\ref{fig:global_rad_clim}a) for both, climatological average of hourly maximum $\left<Q_0^\mathrm{max}\right>$, minimum $\left<Q_0^\mathrm{min}\right>$, and mean $\left<Q_0^\mathrm{mean}\right>$ irradiance observed at Svanvik.
-In accordance to the geometrically derived sunshine duration shown in Fig.~\ref{fig:sunlight_fennoscandia}, $\left<Q_0^\mathrm{max}\right>$ approaches zero in December/January commonly known as polar night and reaches its maximum in June/July commonly referred to as midnight sun conditions.
+In agreement with the geometrically derived sunshine duration shown in Fig.~\ref{fig:sunlight_fennoscandia}, $\left<Q_0^\mathrm{max}\right>$ approaches zero in December/January commonly known as polar night and reaches its maximum in June/July commonly referred to as midnight sun conditions.
 The highest observed $\left<Q_0^\mathrm{max}\right>$ is about $800\,\unit{W\,m^{-2}}$, while on average $\left<Q_0^\mathrm{mean}\right>$ it is about $450\,\unit{W\,m^{-2}}$ (regularly reached in May--July).
@@
 The correlation of \chem{[O_3]} between Esrange, Pallas, and Jergul/Karasjok is fairly high ($r^2\approx 0.8$, Fig.~\ref{fig:density_distribution}a, b), hence we can put those data series together to derive a general ozone climatology for northern Fennoscandia.
+The correlation between Svanvik and Esrange and Pallas is rather low, 0.4 and 0.6, respectively.
+Thus, Svanvik cannot directly be represented by any measurements taken at either Esrange or Pallas.
+
 In Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}, the 2 dimensional density distributions for the thus defined northern Fennoscandia (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}a) and Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) are shown.
@@
 The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, min}} peaks about $20\,\unit{days}$ earlier than the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, and \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}}.
-With respect of the periodicity of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, it takes about $8.5\,\unit{months}$ from minimum concentration to maximum concentration, while from maximum to minimum it only takes about $4.5\,\unit{months}$.
+With respect to the periodicity of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, it takes about $8.5\,\unit{months}$ from minimum concentration to maximum concentration, while from maximum to minimum it only takes about $4.5\,\unit{months}$.
 The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}} is at $52\,\unit{ppb}$ in late April and the minimum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, min}} occurs at $15\,\unit{ppb}$ in the beginning of September.
 The steep rise of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} in spring coincident with increasing temperatures as well as increase in global irradiance.
-The decrease in \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} coincides with the estimated onset of the greening season based on the agricultural rule of thumb that the daily average temperature has to be above $5\,\unit{^\circ C}$ for $5\,\unit{days}$ consecutively (day of the year 110--150, Fig~\ref{fig:greening_season_change_Svanvik}).
+The decrease in \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} coincides with the estimated onset of the greening season based on the agricultural rule of thumb that the daily average temperature has to be above $5\,\unit{^\circ C}$ for $5\,\unit{days}$ consecutively (day of the year 110--150 at Svanvik, Fig~\ref{fig:greening_season_change_Svanvik}).
 
-As mentioned above, the average ozone concentration at Svanvik is in general lower than at the other sites which is in league with a decrease in tropospheric ozone with decreasing altitude.
 The density distribution for Svanvik and the respective \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, and \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}} (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) display a similar pattern as the northern Fennoscandic climatology.
-Only does the spring peak apparently occur slightly earlier (in the beginning of April).
+The spring peak apparently occurs slightly earlier (in the beginning of April).
+As Svanvik is located lower than the other sites, the earlier ozone peak may point to a strong interplay between ozone uptake by vegetation after the onset of greening, for at higher altitude plants sprout later.
+The average ozone concentration at Svanvik is lower than at the other sites, as already mentioned, which is in league with a decrease in tropospheric ozone with decreasing altitude.
+In July--September, ozone is occasionally almost completely depleted during night time.
 
@@
 
-Finally, we shall look at the comparison of climatologies derived from observational data ozone and the various reanalysis products (Fig.~\ref{fig:ozone_climatologies}).
+Finally, we shall look at the comparison of climatologies derived from observational ozone data and the various reanalysis products (Fig.~\ref{fig:ozone_climatologies}).
 As representative for the sites, the nearest neighboring grid point has been chosen.
 The ECMWF products do not reproduce observed surface ozone well.
-The MACC reanalysis is in general too low and displays a erroneous seasonal cycle.
-The CAMS reanalysis matches better but does not reproduce the seasonality in northern Fennoscandia.
-The ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in northern Fennoscandia, though the modeled ozone concentrations in the north are slightly to low.
+The MACC reanalysis (Fig.~\ref{fig:ozone_climatologies}a) is in general too low ($(9\pm 7)\,\unit{ppb}$) and displays a erroneous seasonal cycle.
+MACC reanalysis ozone concentrations peak already in March followed by a second peak in July.
+Also CAMSRA matches the observed ozone climatology hardly (Fig.~\ref{fig:ozone_climatologies}b).
+Despite displaying only one ozone peak, it still does not reproduce the seasonality in northern Fennoscandia at all.
+The CAMSRA spring peak appears to lag behind by $1\,\unit{month}$ and is $5\,\unit{ppb}$ lower, whereas the minimum occurs in January compared to August/September.
+The annual amplitude ($(26\pm 1)\,\unit{ppb}$) is larger than in the climatology derived from observations ($19\,\unit{ppb}$).
+Modeled \chem{[O_3]} at Svanvik are even slightly higher than at Pallas.
+In contrast, the ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in northern Fennoscandia quite well (Fig.~\ref{fig:ozone_climatologies}c), but is on average still slightly too low ($(2.8\pm 0.5)\,\unit{ppb}$).
 
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{ozone_climatologies}
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{ozone_climatologies_portrait}
   \caption{Comparison of climatologies derived from observational data (EBAS, see Fig.~\ref{fig:ozone_climatology_fenoscandic_obs}) with various ozone reanalysis products: (a) ECMWF: MACC reanalysis; (b) ECMWF: CAMS reanalysis; (c) Copernicus: Regional model reanalysis ensemble mean.
@@
 The CAMS reanalysis matches better but does not reproduce the seasonality in northern Fennoscandia.
-The ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in northern Fennoscandia, though the modeled ozone concentrations in the north are slightly to low.}
+The ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in northern Fennoscandia, though the modeled ozone concentrations are slightly too low.}
   \label{fig:ozone_climatologies}
-\end{figure*}
+\end{figure}
 
@@
 \label{subsec:anomalies}
+Using the established climatologies, we look at the anomalies in 2018/19 and the differences between the two years.
 
@@
 \label{subsubsec:anomal_tpq}
+In Fig.~\ref{fig:plot_temperature_anomalies_svanvik}, temperature anomalies at Svanvik are shown as percentage of days warmer/colder than $\pm 1\,\sigma$ from the climatological mean for each month.
+Negative deviations from climatology are displayed as negative percentage.
+The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
+The dashed lines indicate the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
+The summer of 2018 was significantly warmer than average.
+Especially in May and July more than $40\,\unit{\%}$ of days were significantly warmer than normal temperature.
+Temperatures continued to be significantly warmer throughout fall into early winter (August--November) with the exception of October.
+March 2018 was unusually cold.
+The summer of 2019 was fairly average, however significantly colder in July, while August was both significantly colder and warmer.
+Temperatures in April were significantly warmer than average.
 
@@
   \includegraphics[width=8.3cm]{temperature_signific}
-  \caption{Temperature at Svanvik.
+  \caption{Temperature anomalies at Svanvik.
 Percentage of days warmer/colder than $\pm 1\,\sigma$ from climatological mean displayed for each month.
@@
 
+In a similar way, the precipitation anomalies at Svanvik are displayed in Fig.~\ref{fig:plot_precipitation_anomalies_svanvik}.
+However, the percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ from the climatological mean is shown for each month, due to the larger variability in observed precipitation too few days showed a significance of $1 \sigma$ or higher.
+As above, negative deviations from the climatology are shown as negative percentage and annual positive/negative deviations are indicated in the respective corners.
+The dashed lines indicate the expected percentage of values falling above $\pm\frac{1}{2}\,\sigma$ if a normal distribution is assumed ($30.8\,\unit{\%}$).
+Unlike for the temperature anomalies, the picture is not as clear.
+While March 2018 was unusually wet, summer and fall (May/July, September/October) were significantly drier than average.
+2019 was rather average throughout the year, with drier than usual fall (September/October) and a wet December.
+
 \begin{figure}[t]
   \includegraphics[width=8.3cm]{precipitation_signific}
-  \caption{Precipitation at Svanvik.
+  \caption{Precipitation anomalies at Svanvik.
 Percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ from climatological mean displayed for each month.
@@
 
+Similarly, the global irradiance anomalies at Svanvik are presented in Fig.~\ref{fig:global_rad_signific}.
+The percentage of days above/below $\pm 1\,\sigma$ from the climatological mean is displayed for each month.
+The dashed lines indicate the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed.
+In summer 2018 (May/July), global irradiance was significantly higher than on average, while 2019 was generally an average year with a lightly darker spring (April/May).
+
 \begin{figure}[t]
   \includegraphics[width=8.3cm]{global_rad_signific}
-  \caption{Global irradiance at Svanvik.
+  \caption{Global irradiance anomalies at Svanvik.
 Percentage of days above/below $\pm 1\,\sigma$ from climatological mean displayed for each month.
@@
 
+In summary, 2018 was significantly warmer, drier, and lighter than the climatological average, while 2019 was a rather average year.
+
 \subsubsection{Ozone}
 \label{subsubsec:anomal_ozone}
+As we have found above, meteorological conditions in 2018 were significantly above average. How does this translate into observed ozone anomalies?
 

### 2020-04-08 VegOzone.tex
Some text about ozone residuals. No spell-check. Monthly analysis to be done.

@@
 \label{subsec:anomalies}
-Using the established climatologies, we look at the anomalies in 2018/19 and the differences between the two years.
+Using these established climatologies, we look at the anomalies in 2018/19 and discuss the differences between the two years.
 
@@
 The dashed lines indicate the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed.
-In summer 2018 (May/July), global irradiance was significantly higher than on average, while 2019 was generally an average year with a lightly darker spring (April/May).
+In summer 2018 (May/July), global irradiance was significantly higher than on average, while 2019 was generally an average year with a slightly darker spring (April/May).
 
@@
 \label{subsubsec:anomal_ozone}
-As we have found above, meteorological conditions in 2018 were significantly above average. How does this translate into observed ozone anomalies?
+As we discussed in the preceding Section~\ref{subsubsec:anomal_tpq}, meteorological conditions in 2018 were significantly different from the climatological average.
+How does this translate into observed ozone anomalies \chem{\Delta[O_3]}?
+If the data collection at Svanvik had been complete in 2018 and the climatology had recent years included, it would had been straight forward to deduct the significance of ozone anomalies in the same way as for temperature, precipitation, and global irradiance.
+In Section~\ref{subsubsec:clim_ozone}, we described the construction of climatologies for northern Fennoscandia based on long term observations at Esrange, Jergul/Karasjok, and Pallas.
+For Svanvik this assessment is based on rather uncertain observations from 1986--1996.
+We also discussed the general patterns, similarities and differences.
+We shall now investigate \chem{\Delta[O_3]} in 2018/19 with respect to these climatologies in more detail.
+
+In the following, we will use the interchangable terms anomalies and residuals to distinquish between a climatological (anomalies) and a statistical (residuals) point of view.
+All the above factors make it neccessary to look not only at anomalies but statistically assess ozone concentration residuals at Svanvik as well as at Esrange and Pallas.
+Since the climatology of latter is more reliable, we will start with surface ozone concentration residuals \chem{\Delta[O_3]} at Esrange and Pallas in 2018.
+As data for 2019 has not yet been available from EBAS (last access, February 2020), we refrain from looking into these.
+
+In Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}, the probability density functions of daily mean ozone concentration residuals for Esrange and Pallas are shown.
+Both stations display a clear deviation from the climatological mean, however, not significant.
+On average, \chem{[O_3]} had been elevated by $(1.9\pm 5.7)\,\unit{ppb}$ and $(1.9\pm 5.4)\,\unit{ppb}$, respectively.
+Both distributions are slightly skrewed towards positive deviations with $\frac{1}{4}$ of days (3.
+quantile, q3) falling above $5.8\,\unit{ppb}$ and $5.6\,\unit{ppb}$, respectively.
+
+Figure~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik} displays the probability density functions for Svanvik based on data taken in 2018/19 as well as relative to the climatology derived for northern Fennoscandia (denoted with NF in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}a) and the climatology derived for Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b).
+As previously mentioned, \chem{\left<[O_3]\right>} at Svanvik are lower compared to the northern Fennoscandic climatology ($(6.6\pm 7.1)\,\unit{ppb}$).
+For 2018 and 2019, we find $\Delta\chem{\left<[O_3]\right>} = (-3.8\pm 5.6)\,\unit{ppb}$ and $\Delta\chem{\left<[O_3]\right>} = (-5.2\pm 5.0)\,\unit{ppb}$, respectively.
+While the probablility density of the residuals between the two climatologies is skewed towards negative values, the skew of 2018/19 is positive.
+With respect to the climatology for Svanvik, the probability density functions average for 2018/19 lies at $\Delta\chem{\left<[O_3]\right>} = (2.6\pm 5.8)\,\unit{ppb}$ and $\Delta\chem{\left<[O_3]\right>} = (1.2\pm 5.0)\,\unit{ppb}$, respectively.
+Neither of the \chem{\Delta\left<[O_3]\right>} is significant by statistical means. But daily average ozone concentrations are slightly elevated in 2018 compared to 2019.
 
@@
 \end{figure}
+
+To further investigate this, we have a look at a Student's~t-test ($t = \frac{\chem{\Delta[O_3]}}{\sigma}$), assuming the same sample uncertainty $\sigma$ in both, climatology and the year in question.
+Results for Esrange and Pallas are shown in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test}.
+This t-test clearly shows, that at both sites for at least $50\,\unit{\%}$ of days the deviation from climatological mean is significant on the $1\sigma$ level.
+In Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}, the test statistics for Svanvik are displayed.
+From Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}a), we can deduce, that the deviation from the northern Fennoscandic climatology is significant on the $2\sigma$ level for at least half of the days in the Svanvik climatology, while the deviation is less certain in 2018 and 2019.
+In Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}b) the same is shown with respect to the climatology from Svanvik.
+We find that in less than $50\,\unit{\%}$ of days the ozone concentration deviation is significant on the $1\sigma$ level.
 
@@
 
-Model results for natural and semi-natural vegetation (birch, pine, clover)
+Model results for natural and semi-natural vegetation (birch, pine, clover, tobacco(?))
 \begin{itemize}

### 2020-04-08 VegOzone.tex
Finish Section 3 - spell-check and monthly ozone analysis remain to be done.

@@
 How does this translate into observed ozone anomalies \chem{\Delta[O_3]}?
-If the data collection at Svanvik had been complete in 2018 and the climatology had recent years included, it would had been straight forward to deduct the significance of ozone anomalies in the same way as for temperature, precipitation, and global irradiance.
+
+If the data collected at Svanvik had been complete in 2018 and the climatology had extended into the resent past, we could have assessed the significance of ozone anomalies in the same way as for temperature, precipitation, and global irradiance.
 In Section~\ref{subsubsec:clim_ozone}, we described the construction of climatologies for northern Fennoscandia based on long term observations at Esrange, Jergul/Karasjok, and Pallas.
@@
 In the following, we will use the interchangable terms anomalies and residuals to distinquish between a climatological (anomalies) and a statistical (residuals) point of view.
-All the above factors make it neccessary to look not only at anomalies but statistically assess ozone concentration residuals at Svanvik as well as at Esrange and Pallas.
+All the above factors make it neccessary to look not only at anomalies but statistically assess ozone concentration residuals at Svanvik as well as at Esrange and Pallas in comparison.
 Since the climatology of latter is more reliable, we will start with surface ozone concentration residuals \chem{\Delta[O_3]} at Esrange and Pallas in 2018.
@@
 
-Figure~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik} displays the probability density functions for Svanvik based on data taken in 2018/19 as well as relative to the climatology derived for northern Fennoscandia (denoted with NF in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}a) and the climatology derived for Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b).
-As previously mentioned, \chem{\left<[O_3]\right>} at Svanvik are lower compared to the northern Fennoscandic climatology ($(6.6\pm 7.1)\,\unit{ppb}$).
+Figure~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik} displays the probability density functions of daily mean ozone concentration residuals for Svanvik based on data taken in 2018/19 with respect to the climatology derived for northern Fennoscandia (denoted NF in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}a) and the climatology derived for Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b).
+As previously mentioned, \chem{\left<[O_3]\right>} at Svanvik are $(6.6\pm 7.1)\,\unit{ppb}$ lower than the northern Fennoscandic climatology.
 For 2018 and 2019, we find $\Delta\chem{\left<[O_3]\right>} = (-3.8\pm 5.6)\,\unit{ppb}$ and $\Delta\chem{\left<[O_3]\right>} = (-5.2\pm 5.0)\,\unit{ppb}$, respectively.
 While the probablility density of the residuals between the two climatologies is skewed towards negative values, the skew of 2018/19 is positive.
-With respect to the climatology for Svanvik, the probability density functions average for 2018/19 lies at $\Delta\chem{\left<[O_3]\right>} = (2.6\pm 5.8)\,\unit{ppb}$ and $\Delta\chem{\left<[O_3]\right>} = (1.2\pm 5.0)\,\unit{ppb}$, respectively.
-Neither of the \chem{\Delta\left<[O_3]\right>} is significant by statistical means. But daily average ozone concentrations are slightly elevated in 2018 compared to 2019.
+With respect to the climatology for Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b), the probability density functions average for 2018/19 lies at $\Delta\chem{\left<[O_3]\right>} = (2.6\pm 5.8)\,\unit{ppb}$ and $\Delta\chem{\left<[O_3]\right>} = (1.2\pm 5.0)\,\unit{ppb}$, respectively.
+Neither of the \chem{\Delta\left<[O_3]\right>} is significant by statistical means, but daily average ozone concentrations are slightly elevated in 2018 compared to 2019.
+
+The \chem{\left<\Delta\left<[O_3]\right>\right>} derived for Svanvik in 2018 is slightly larger than for Esrange and Pallas.
+This could mean either that 2018 had been more exceptional at Svanvik or the Svanvik climatology may need a correction for increased tropospheric background ozone concentration.
+Looking at daily \chem{\Delta_{present-past}\left<[O_3]\right>}, wherein present is defined as the average of 2010--2018 and past as the respective beginning of data taking until 1997, does not reveal a clear trend: We find for Esrange $(1.5\pm 3.1)\,\unit{ppb}$ and for Pallas $(-2.5\pm 3.7)\,\unit{ppb}$.
 
@@
 
-To further investigate this, we have a look at a Student's~t-test ($t = \frac{\chem{\Delta[O_3]}}{\sigma}$), assuming the same sample uncertainty $\sigma$ in both, climatology and the year in question.
+To further investigate the ozone residuals, we conduct a Student's~t-test ($t = \frac{\chem{\Delta[O_3]}}{\sigma}$), assuming the same sample uncertainty $\sigma$ in both, climatology and the year in question.
 Results for Esrange and Pallas are shown in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test}.
-This t-test clearly shows, that at both sites for at least $50\,\unit{\%}$ of days the deviation from climatological mean is significant on the $1\sigma$ level.
+This t-test clearly shows, that at both sites for at least $50\,\unit{\%}$ of days the residuals are significant on the $1\sigma$ level.
 In Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}, the test statistics for Svanvik are displayed.
-From Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}a), we can deduce, that the deviation from the northern Fennoscandic climatology is significant on the $2\sigma$ level for at least half of the days in the Svanvik climatology, while the deviation is less certain in 2018 and 2019.
-In Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}b) the same is shown with respect to the climatology from Svanvik.
-We find that in less than $50\,\unit{\%}$ of days the ozone concentration deviation is significant on the $1\sigma$ level.
+From Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}a), we can deduce, that the residuals of the Svanvik climatology from the northern Fennoscandic climatology are significant on the $2\sigma$ level for at least half of the days, while the residuals are less certain in 2018 and 2019.
+In Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}b) the same is shown for the years 2018/19 with respect to the climatology from Svanvik.
+We find that in less than $50\,\unit{\%}$ of days the ozone concentration residuals is significant on the $1\sigma$ level.
 

### 2020-04-16 VegOzone.tex
Spell-check.

@@
 
-In the following, we will use the interchangable terms anomalies and residuals to distinquish between a climatological (anomalies) and a statistical (residuals) point of view.
-All the above factors make it neccessary to look not only at anomalies but statistically assess ozone concentration residuals at Svanvik as well as at Esrange and Pallas in comparison.
+In the following, we will use the interchangeable terms anomalies and residuals to distinguish between a climatological (anomalies) and a statistical (residuals) point of view.
+All the above factors make it necessary to look not only at anomalies but statistically assess ozone concentration residuals at Svanvik as well as at Esrange and Pallas in comparison.
 Since the climatology of latter is more reliable, we will start with surface ozone concentration residuals \chem{\Delta[O_3]} at Esrange and Pallas in 2018.
@@
 On average, \chem{[O_3]} had been elevated by $(1.9\pm 5.7)\,\unit{ppb}$ and $(1.9\pm 5.4)\,\unit{ppb}$, respectively.
-Both distributions are slightly skrewed towards positive deviations with $\frac{1}{4}$ of days (3.
+Both distributions are slightly skewed towards positive deviations with $\frac{1}{4}$ of days (3.
 quantile, q3) falling above $5.8\,\unit{ppb}$ and $5.6\,\unit{ppb}$, respectively.
@@
 For 2018 and 2019, we find $\Delta\chem{\left<[O_3]\right>} = (-3.8\pm 5.6)\,\unit{ppb}$ and $\Delta\chem{\left<[O_3]\right>} = (-5.2\pm 5.0)\,\unit{ppb}$, respectively.
-While the probablility density of the residuals between the two climatologies is skewed towards negative values, the skew of 2018/19 is positive.
+While the probability density of the residuals between the two climatologies is skewed towards negative values, the skew of 2018/19 is positive.
 With respect to the climatology for Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b), the probability density functions average for 2018/19 lies at $\Delta\chem{\left<[O_3]\right>} = (2.6\pm 5.8)\,\unit{ppb}$ and $\Delta\chem{\left<[O_3]\right>} = (1.2\pm 5.0)\,\unit{ppb}$, respectively.

### 2020-04-16 VegOzone.tex
Some rearangement of figures and added a new figure.

@@
 This could mean either that 2018 had been more exceptional at Svanvik or the Svanvik climatology may need a correction for increased tropospheric background ozone concentration.
-Looking at daily \chem{\Delta_{present-past}\left<[O_3]\right>}, wherein present is defined as the average of 2010--2018 and past as the respective beginning of data taking until 1997, does not reveal a clear trend: We find for Esrange $(1.5\pm 3.1)\,\unit{ppb}$ and for Pallas $(-2.5\pm 3.7)\,\unit{ppb}$.
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_residuals}
-  \caption{Probability density functions of ozone concentration residuals. 2018 with respect to respective climatology for different sites in Fennoscandia. (a) Esrange; (b) Pallas.}
-  \label{fig:ozone_climatology_fenoscandic_obs_residuals}
-\end{figure}
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_residuals-Svanvik}
-  \caption{Probability density functions of ozone concentration residuals. 2018/19 observations at Svanhovd with respect to derived climatologies for (a) Northern Fennoscandia; (b) Svanvik.}
-  \label{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}
-\end{figure}
+Based on the meteorological conditions, we may assume 2019 as an average year which would lead to a deduction of a constant bias of about $1.2\,\unit{ppb}$ in the Svanvik climatology.
+Daily \chem{\Delta_{present-past}\left<[O_3]\right>} from Esrange and Pallas, wherein present is defined as the average of 2010--2018 and past as the respective beginning of data taking until 1997, do not reveal a clear trend: We find for Esrange $(1.5\pm 3.1)\,\unit{ppb}$ and for Pallas $(-2.5\pm 3.7)\,\unit{ppb}$.
+Former is compatible with our assumed bias for the Svanvik climatology.
 
@@
 In Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}b) the same is shown for the years 2018/19 with respect to the climatology from Svanvik.
-We find that in less than $50\,\unit{\%}$ of days the ozone concentration residuals is significant on the $1\sigma$ level.
+We find that in less than $50\,\unit{\%}$ of days the ozone concentration residuals are significant on the $1\sigma$ level.
 
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_test}
-  \caption{Student's t-test assuming same sample uncertainty in both, climatology and 2018 observations for different sites in Fennoscandia.
-\chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals} are significantly different from zero-hypothesis on the $1\,\sigma$ level.
-(a) Esrange; (b) Pallas.}
-  \label{fig:ozone_climatology_fenoscandic_obs_test}
-\end{figure}
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_test-Svanvik}
-  \caption{Student's t-test assuming same sample uncertainty in both, climatology and 2018/19 observations at Svanhovd with respect to derived climatologies for (a) Northern Fennoscandia; (b) Svanvik.
-\chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}a) are significantly different from zero-hypothesis on the $2\,\sigma$,  $1\,\sigma$ level, respectively.
-\chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}b) are not significantly different from zero-hypothesis.}
-  \label{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}
+  \includegraphics[width=8.3cm]{ozone_signific}
+  \caption{Ozone concentration anomalies at Svanvik.
+Percentage of days above/below $\pm 1\,\sigma$ from climatological mean displayed for each month.
+Negative deviations from climatology are shown as negative percentage.
+The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
+The dashed lines indicate the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
+Year displayed in panel (a) 2018; (b) 2019.
+In April/May 2018, ozone was significantly higher than on average, while 2019 was generally an average year.
+In both years ozone concentrations were significantly below average in November.}
+  \label{fig:ozone_signific}
 \end{figure}
@@
 \appendix
-\section{}    %% Appendix A
+\section{Derived climatologies}    %% Appendix A
 
-\subsection{}     %% Appendix A1, A2, etc.
+\subsection{Temperature, precipitation, and global irradiance}     %% Appendix A1, A2, etc.
+
+\appendixfigures  %% needs to be added in front of appendix figures
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{global_rad_clim}
+  \caption{Observed global irradiance at Svanvik.
+The cyan lines indicate the $1\,\sigma$ level deducted from the climatology.
+(a) Climatology (hourly maxima, mean, and minima); deviation from climatology (b) 2018; (c) 2019.}
+  \label{fig:global_rad_clim}
+\end{figure}
+\clearpage
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{test_fennoscandia}
+  \caption{Maximum daily sun shine duration based on geometrical calculations at different sites in Fennoscandia.
+Midnight sun conditions at Jergul/Karasjok and Svanvik prevail from the end of May until the end of July, while at Esrange and Pallas they only prevail from the beginning of June until mid of July.}
+  \label{fig:sunlight_fennoscandia}
+\end{figure}
+\clearpage
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{greening_season_change_Svanvik}
+  \caption{Estimated shift and prolongation of greening season at Svanhovd over the past two decades based on the method and data described in \citet{GMD:Falk2019}.}
+  \label{fig:greening_season_change_Svanvik}
+\end{figure}
+\clearpage
+\subsection{Ozone}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{density_distribution}
+  \caption{Probability densities and correlation coefficient for \chem{[O_3]} between different sites in northern Fennoscandia.
+Jergul/Karasjok is well correlated with Esrange, and Pallas.
+Svanvik displays highest correlation with Pallas.
+(a) Jergul/Karasjok--Esrange; (b) Jergul/Karasjok--Pallas; (c) Svanvik--Esrange; (d) Svanvik--Pallas.}
+  \label{fig:density_distribution}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_residuals}
+  \caption{Probability density functions of ozone concentration residuals. 2018 with respect to respective climatology for different sites in Fennoscandia. (a) Esrange; (b) Pallas.}
+  \label{fig:ozone_climatology_fenoscandic_obs_residuals}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_residuals-Svanvik}
+  \caption{Probability density functions of ozone concentration residuals. 2018/19 observations at Svanhovd with respect to derived climatologies for (a) Northern Fennoscandia; (b) Svanvik.}
+  \label{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_test}
+  \caption{Student's t-test assuming same sample uncertainty in both, climatology and 2018 observations for different sites in Fennoscandia.
+\chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals} are significantly different from zero-hypothesis on the $1\,\sigma$ level.
+(a) Esrange; (b) Pallas.}
+  \label{fig:ozone_climatology_fenoscandic_obs_test}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_test-Svanvik}
+  \caption{Student's t-test assuming same sample uncertainty in both, climatology and 2018/19 observations at Svanhovd with respect to derived climatologies for (a) Northern Fennoscandia; (b) Svanvik.
+\chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}a) are significantly different from zero-hypothesis on the $2\,\sigma$,  $1\,\sigma$ level, respectively.
+\chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}b) are not significantly different from zero-hypothesis.}
+  \label{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}
+\end{figure}
 
@@
 
-\appendixfigures  %% needs to be added in front of appendix figures
+%\appendixfigures  %% needs to be added in front of appendix figures
 
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{global_rad_clim}
-  \caption{Observed global irradiance at Svanvik.
-The cyan lines indicate the $1\,\sigma$ level deducted from the climatology.
-(a) Climatology (hourly maxima, mean, and minima); deviation from climatology (b) 2018; (c) 2019.}
-  \label{fig:global_rad_clim}
-\end{figure}
+%\begin{figure}[t]
+%  \includegraphics[width=8.3cm]{ozone_obs_2018_07}
+%  \caption{Estimate of missing observation \chem{[O_3]} data at Svanhovd in July 2018 based on derived hourly climatologies of northern Fennoscandia, Svanvik, and Copernicus regional model reanalysis ensemble mean.
+(a) Time series; (b) Student's t-test.}
+%  \label{fig:ozone_obs_2018_rrea}
+%\end{figure}
 
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{test_fennoscandia}
-  \caption{Maximum daily sun shine duration based on geometrical calculations at different sites in Fennoscandia.
-Midnight sun conditions at Jergul/Karasjok and Svanvik prevail from the end of May until the end of July, while at Esrange and Pallas they only prevail from the beginning of June until mid of July.}
-  \label{fig:sunlight_fennoscandia}
-\end{figure}
+%\noappendix
 
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{density_distribution}
-  \caption{Probability densities and correlation coefficient for \chem{[O_3]} between different sites in northern Fennoscandia.
-Jergul/Karasjok is well correlated with Esrange, and Pallas.
-Svanvik displays highest correlation with Pallas.
-(a) Jergul/Karasjok--Esrange; (b) Jergul/Karasjok--Pallas; (c) Svanvik--Esrange; (d) Svanvik--Pallas.}
-  \label{fig:density_distribution}
-\end{figure}
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_obs_2018_07}
-  \caption{Estimate of missing observation \chem{[O_3]} data at Svanhovd in July 2018 based on derived hourly climatologies of northern Fennoscandia, Svanvik, and Copernicus regional model reanalysis ensemble mean.
-(a) Time series; (b) Student's t-test.}
-  \label{fig:ozone_obs_2018_rrea}
-\end{figure}
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{greening_season_change_Svanvik}
-  \caption{Estimated shift and prolongation of greening season at Svanhovd over the past two decades based on the method and data described in \citet{GMD:Falk2019}.}
-  \label{fig:greening_season_change_Svanvik}
-\end{figure}
-
-\appendixtables   %% needs to be added in front of appendix tables
+%\appendixtables   %% needs to be added in front of appendix tables
 

### 2020-04-17 VegOzone.tex
Finished ozone section - finished Section 3. Wrote some introduction to Section 4.

@@
 
+In accordance to the above findings, we adjust the derived ozone climatology for Svanvik by adding a constant bias correction factor $\Delta\chem{[O_3]_\mathrm{bias-corr}} = 1.2\,\unit{ppb}$ and estimate the ozone concentration anomalies for 2018/19.
+In Figure~\ref{fig:ozone_signific}, the resulting percentage of days above/below $\pm 1\,\sigma$ from the bias-corrected climatological mean is displayed for each month.
+Negative deviations from climatology are shown as negative percentage.
+Note that missing bars for certain months are both due to data availability (see Section~\ref{subsubsec:atmo_svanvik} for details about the time range of data taking at Svanhovd) and anomalies not meeting the criteria of $\pm 1\sigma$ deviation from climatological mean.
+In April/May 2018, ozone was significantly enhanced, while 2019 was generally an average year.
+In both years ozone concentrations were significantly below average in November.
+
 \begin{figure}[t]
@@
   \caption{Ozone concentration anomalies at Svanvik.
+Ozone climatology has been corrected for an constant negative bias of $-1.2\,\unit{ppb}$.
 Percentage of days above/below $\pm 1\,\sigma$ from climatological mean displayed for each month.
@@
 
+To conclude, ozone concentrations in 2018 have been elevated compared to the respective climatologies, however, on average not significantly by statistical means.
+At Esrange and Pallas at least $25\,\unit{\%}$ (3.~quantile) of anomalies are significant on a $3\sigma$ level or more, while at Svanvik the 3.~quantile lies at about $2\sigma$.
+This means that in 2018 the stations at Esrange and Pallas saw more days with elevated ozone than was monitored at Svanvik.
+
 \section{DO3SE modeling}
+To investigate which of the evironmental factors lead to the observed ozone damage on clover and tobacco {\bf TODO: Ane confirm!} at Svanhovd, we have modeled the ozone uptake using the DO3SE model.
 
-Model results for natural and semi-natural vegetation (birch, pine, clover, tobacco(?))
+{\bf TODO: Model results for natural and semi-natural vegetation (birch, pine, clover, tobacco(?))}
 \begin{itemize}
@@
 \end{figure}
-\clearpage
+
 \begin{figure}[t]
@@
 \clearpage
+
 \subsection{Ozone}
@@
 \end{figure}
+\clearpage
 

### 2020-04-17 VegOzone.tex
Some more text for indroduction.

@@
 \label{sec:intro}
+
+{\bf OBS: The following is just a copy of the 2018 dry deposition paper. It has to be substantially changed to meet the no-copycat-criteria. TODO: Rewrite in context of this article.}
+
+Ozone is an important trace gas for all lifeforms on Earth.
+Depending on the place of its occurrence it has either a positive or negative connotation.
+In the stratosphere, ozone absorbs most of the ultraviolet (UV)-light from the sun within the range of 100--315\,\unit{nm}, thus shielding the Earth's surface from the most harmful UV-radiation.
+In addition, ozone is a potent greenhouse gas in both, stratosphere and troposphere.
+With a radiative forcing of $0.40 \pm 0.20\,\unit{W\,m^{-2}}$, it is placed third, only surpassed by \chem{CO_2} and \chem{CH_4} \citep[Chapter 8]{IPCC2013}.\\
+In the troposphere and in particular in ambient air, ozone is considered as a highly toxic pollutant.
+Since the industrial revolution, tropospheric background ozone concentrations have been increasing in the northern hemisphere \citep[Chapter 2]{IPCC2013}.
+In recent years, the number of episodes of peak concentrations has been, in general, decreasing in North America and Europe due to the implementation of air quality regulations \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
+At the same time, fast developing countries, like e.g., China or India, saw a significant increase in ozone related air pollution.
+Continuously high concentrations of ambient air ozone are hazardous to the whole ecosystem.
+It is estimated that ozone is cause to an increase in pre-mature deaths \citep{WHO2008}, an average global loss of yield in the four major crops (wheat, rice, maize, and soybean) of about $3-15\,\unit{\%}$ \citep{PJ:Ainsworth2017} as well as $7\,\unit{\%}$ loss in primary production in forestry \citep{GCB:Wittig2009,EP:Matyssek2012}.
+The impact of ozone damage on vegetation and agricultural plants may affect food security in the future especially in Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018} and might be an important additional feedback to climate change \citep{Nat:Sitch2007}.\\
+Elevated ozone levels at a site may originate from both, the local production of ozone from its precursors, which are transported, and from advection of ozone itself.
+Long-range ozone transport occurs regularly and might be most important in regions that otherwise lack precursors.
+Tropospheric ozone is produced in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) or volatile organic substances (VOCs -- also known as hydrocarbons) in the presents of nitrogen oxides (\chem{NO_x}).
+A typical reaction mechanism for \chem{CO} is sketched in the following.
+In a sequence of rapid reactions a peroxyl radical \chem{HO_2^\bullet} is formed through an initial reaction of \chem{CO} with a hydroxyl radical \chem{^\bullet OH}.
+Via a reaction between \chem{HO_2^\bullet} and \chem{NO}, \chem{NO_2} is formed which is then photolyzed.
+The resulting atomic oxygen reacts then with \chem{O_2} (and also under the presence of available co-reactants) to form an ozone molecule.
+
+Since ozone is highly reactive, its global mean life-time in the troposphere is roughly $22$ days but ranges between a few days in the tropical boundary layer to up to $1$ year in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}.
+The abundance of tropospheric ozone therefore varies, e.g., with time of the day, season, altitude, location \citep{ACP:Schnell2015}, or weather conditions in general \citep{ACP:Otero2018}.
+Typical concentrations of surface ozone range from $10\,\unit{ppb}$ over the tropical Pacific to $100\,\unit{ppb}$ in the downwind areas of highly emitting sources \citep[Chapter 8]{IPCC2013}.
+This variability poses a challenge on both, trend analysis from observations as well as validation and inter-comparison of models.
+From the observational side, the number of long-term observations (started before the 1950s) is limited and restricted to mainly European sites.
+Most of these indicate a doubling of tropospheric ozone since the 1950s \citep[Chapter 2]{IPCC2013}.
+
 \begin{itemize}
-\item ozone in general
 \item ozone induced damage in vegetation \emph{Contributor: Ane - maybe also ask Aud?}
@@
 When the stomata are closed, ozone does not enter.
-\item metrics of ozone: aot40 and pod
+\item metrics of ozone
+  \begin{equation}
+    \mathrm{AOT_x} = \int{\mathrm{max(\chem{[O_3]}-x)}\cdot \mathrm{dt}},
+  \end{equation}
+  with the ambient ozone concentration $\chem{[O_3]}$ and the concentration threshold $\mathrm{x}$. A typical value for $\mathrm{x}=40\,\unit{ppb}$.
+  \begin{equation}
+    \mathrm{POD_y} = \int{\mathrm{max(\Phi_\chem{O_3}-y)}\cdot \mathrm{dt}},
+  \end{equation}
+   with the ambient ozone flux through the stomata $\Phi_\chem{[O_3]}$ and the flux threshold $\mathrm{y}$.
 \item present subarctic climate and conditions
@@
 
-A climatology derived for global irradiance $Q_0$ is shown in Fig.~\ref{fig:global_rad_clim}a) for both, climatological average of hourly maximum $\left<Q_0^\mathrm{max}\right>$, minimum $\left<Q_0^\mathrm{min}\right>$, and mean $\left<Q_0^\mathrm{mean}\right>$ irradiance observed at Svanvik.
-In agreement with the geometrically derived sunshine duration shown in Fig.~\ref{fig:sunlight_fennoscandia}, $\left<Q_0^\mathrm{max}\right>$ approaches zero in December/January commonly known as polar night and reaches its maximum in June/July commonly referred to as midnight sun conditions.
+A climatology derived for global irradiance $Q_0$ is shown in Appendix Fig.~\ref{fig:global_rad_clim}a) for both, climatological average of hourly maximum $\left<Q_0^\mathrm{max}\right>$, minimum $\left<Q_0^\mathrm{min}\right>$, and mean $\left<Q_0^\mathrm{mean}\right>$ irradiance observed at Svanvik.
+In agreement with the geometrically derived sunshine duration shown in Appendix Fig.~\ref{fig:sunlight_fennoscandia}, $\left<Q_0^\mathrm{max}\right>$ approaches zero in December/January commonly known as polar night and reaches its maximum in June/July commonly referred to as midnight sun conditions.
 The highest observed $\left<Q_0^\mathrm{max}\right>$ is about $800\,\unit{W\,m^{-2}}$, while on average $\left<Q_0^\mathrm{mean}\right>$ it is about $450\,\unit{W\,m^{-2}}$ (regularly reached in May--July).
@@
 
-The correlation of \chem{[O_3]} between Esrange, Pallas, and Jergul/Karasjok is fairly high ($r^2\approx 0.8$, Fig.~\ref{fig:density_distribution}a, b), hence we can put those data series together to derive a general ozone climatology for northern Fennoscandia.
+The correlation of \chem{[O_3]} between Esrange, Pallas, and Jergul/Karasjok is fairly high ($r^2\approx 0.8$, Appendix Fig.~\ref{fig:density_distribution}a, b), hence we can put those data series together to derive a general ozone climatology for northern Fennoscandia.
 The correlation between Svanvik and Esrange and Pallas is rather low, 0.4 and 0.6, respectively.
@@
 The steep rise of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} in spring coincident with increasing temperatures as well as increase in global irradiance.
-The decrease in \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} coincides with the estimated onset of the greening season based on the agricultural rule of thumb that the daily average temperature has to be above $5\,\unit{^\circ C}$ for $5\,\unit{days}$ consecutively (day of the year 110--150 at Svanvik, Fig~\ref{fig:greening_season_change_Svanvik}).
+The decrease in \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} coincides with the estimated onset of the greening season based on the agricultural rule of thumb that the daily average temperature has to be above $5\,\unit{^\circ C}$ for $5\,\unit{days}$ consecutively (day of the year 110--150 at Svanvik, Appendix Fig~\ref{fig:greening_season_change_Svanvik}).
 
@@
 The spring peak apparently occurs slightly earlier (in the beginning of April).
-As Svanvik is located lower than the other sites, the earlier ozone peak may point to a strong interplay between ozone uptake by vegetation after the onset of greening, for at higher altitude plants sprout later.
-The average ozone concentration at Svanvik is lower than at the other sites, as already mentioned, which is in league with a decrease in tropospheric ozone with decreasing altitude.
+As Svanvik is located at lower altitude than the other sites, the earlier ozone peak may point to a strong interplay between ozone uptake by vegetation after the onset of greening, for at higher altitude plants sprout later.
+{\bf TODO: Is there a way to confirm this?
+Actual data for the three sites?
+Please correct if wrong!
+Ane?
+Aud?} The average ozone concentration at Svanvik is lower than at the other sites, as already mentioned, which is in league with a decrease in tropospheric ozone with decreasing altitude.
 In July--September, ozone is occasionally almost completely depleted during night time.
@@
 MACC reanalysis ozone concentrations peak already in March followed by a second peak in July.
-Also CAMSRA matches the observed ozone climatology hardly (Fig.~\ref{fig:ozone_climatologies}b).
+Also CAMSRA matches the observed ozone climatology poorly (Fig.~\ref{fig:ozone_climatologies}b).
 Despite displaying only one ozone peak, it still does not reproduce the seasonality in northern Fennoscandia at all.
@@
 
-In Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}, the probability density functions of daily mean ozone concentration residuals for Esrange and Pallas are shown.
+In Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}, the probability density functions of daily mean ozone concentration residuals for Esrange and Pallas are shown.
 Both stations display a clear deviation from the climatological mean, however, not significant.
@@
 
-Figure~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik} displays the probability density functions of daily mean ozone concentration residuals for Svanvik based on data taken in 2018/19 with respect to the climatology derived for northern Fennoscandia (denoted NF in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}a) and the climatology derived for Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b).
+Appendix Figure~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik} displays the probability density functions of daily mean ozone concentration residuals for Svanvik based on data taken in 2018/19 with respect to the climatology derived for northern Fennoscandia (denoted NF in Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}a) and the climatology derived for Svanvik (Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b).
 As previously mentioned, \chem{\left<[O_3]\right>} at Svanvik are $(6.6\pm 7.1)\,\unit{ppb}$ lower than the northern Fennoscandic climatology.
@@
 While the probability density of the residuals between the two climatologies is skewed towards negative values, the skew of 2018/19 is positive.
-With respect to the climatology for Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b), the probability density functions average for 2018/19 lies at $\Delta\chem{\left<[O_3]\right>} = (2.6\pm 5.8)\,\unit{ppb}$ and $\Delta\chem{\left<[O_3]\right>} = (1.2\pm 5.0)\,\unit{ppb}$, respectively.
+With respect to the climatology for Svanvik (Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b), the probability density functions average for 2018/19 lies at $\Delta\chem{\left<[O_3]\right>} = (2.6\pm 5.8)\,\unit{ppb}$ and $\Delta\chem{\left<[O_3]\right>} = (1.2\pm 5.0)\,\unit{ppb}$, respectively.
 Neither of the \chem{\Delta\left<[O_3]\right>} is significant by statistical means, but daily average ozone concentrations are slightly elevated in 2018 compared to 2019.
 
-The \chem{\left<\Delta\left<[O_3]\right>\right>} derived for Svanvik in 2018 is slightly larger than for Esrange and Pallas.
-This could mean either that 2018 had been more exceptional at Svanvik or the Svanvik climatology may need a correction for increased tropospheric background ozone concentration.
-Based on the meteorological conditions, we may assume 2019 as an average year which would lead to a deduction of a constant bias of about $1.2\,\unit{ppb}$ in the Svanvik climatology.
+The \chem{\left<\Delta\left<[O_3]\right>\right>} for Svanvik in 2018 is slightly larger than for Esrange and Pallas.
+This could mean either that 2018 had been more exceptional at Svanvik or the derived climatology may need a correction for increased tropospheric background ozone concentration.
+Based on the meteorological conditions, we assume 2019 as average year which leads to a constant bias of about $1.2\,\unit{ppb}$.
 Daily \chem{\Delta_{present-past}\left<[O_3]\right>} from Esrange and Pallas, wherein present is defined as the average of 2010--2018 and past as the respective beginning of data taking until 1997, do not reveal a clear trend: We find for Esrange $(1.5\pm 3.1)\,\unit{ppb}$ and for Pallas $(-2.5\pm 3.7)\,\unit{ppb}$.
-Former is compatible with our assumed bias for the Svanvik climatology.
+Former is compatible with our assumed bias for the climatology derived for Svanvik.
 
 To further investigate the ozone residuals, we conduct a Student's~t-test ($t = \frac{\chem{\Delta[O_3]}}{\sigma}$), assuming the same sample uncertainty $\sigma$ in both, climatology and the year in question.
-Results for Esrange and Pallas are shown in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test}.
-This t-test clearly shows, that at both sites for at least $50\,\unit{\%}$ of days the residuals are significant on the $1\sigma$ level.
-In Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}, the test statistics for Svanvik are displayed.
-From Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}a), we can deduce, that the residuals of the Svanvik climatology from the northern Fennoscandic climatology are significant on the $2\sigma$ level for at least half of the days, while the residuals are less certain in 2018 and 2019.
-In Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}b) the same is shown for the years 2018/19 with respect to the climatology from Svanvik.
-We find that in less than $50\,\unit{\%}$ of days the ozone concentration residuals are significant on the $1\sigma$ level.
+Results for Esrange and Pallas are shown in Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test}.
+This t-test clearly shows, that at both sites the residuals are significant on the $1\sigma$ level for at least $50\,\unit{\%}$.
+In Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}, the test statistics for Svanvik are displayed.
+From Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}a), we can deduce, that the residuals between the Svanvik climatology and the northern Fennoscandic climatology are significant on the $2\sigma$ level for at least half of the data, while the residuals are less certain in 2018 and 2019.
+In Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}b) the same is shown for the years 2018/19 with respect to the climatology from Svanvik.
+We find that less than $50\,\unit{\%}$ of ozone concentration residuals are significant on the $1\sigma$ level.
 
 In accordance to the above findings, we adjust the derived ozone climatology for Svanvik by adding a constant bias correction factor $\Delta\chem{[O_3]_\mathrm{bias-corr}} = 1.2\,\unit{ppb}$ and estimate the ozone concentration anomalies for 2018/19.
-In Figure~\ref{fig:ozone_signific}, the resulting percentage of days above/below $\pm 1\,\sigma$ from the bias-corrected climatological mean is displayed for each month.
+In Fig.~\ref{fig:ozone_signific}, the resulting percentage of days above/below $\pm 1\,\sigma$ from the bias-corrected climatological mean is displayed for each month.
 Negative deviations from climatology are shown as negative percentage.
@@
 In April/May 2018, ozone was significantly enhanced, while 2019 was generally an average year.
-In both years ozone concentrations were significantly below average in November.
+In both years, ozone concentrations were significantly below average in November.
 
@@
 \section{DO3SE modeling}
-To investigate which of the evironmental factors lead to the observed ozone damage on clover and tobacco {\bf TODO: Ane confirm!} at Svanhovd, we have modeled the ozone uptake using the DO3SE model.
+To investigate which of the environmental factors lead to the observed ozone damage on clover and tobacco {\bf TODO: Ane confirm!} at Svanhovd, we have modeled the ozone uptake using the DO3SE model.
 

### 2020-04-22 VegOzone.tex
Fix some mistakes. Add more info to reanalysis - some more info to come.

@@
 There are two global reanalysis products available at the European Centre for Medium-Range Weather Forecasts (ECMWF) that include atmospheric tracers, in particular ozone (MACC and CAMSRA).
-The temporal as well as spatial resolution of these reanalysis products is rather coarse: 3-hourly and $0.75^\circ\times 0.75^\circ$ or roughly $78.1\,\unit{km}\times 83.4\,\unit{km}$ at the location of Svanvik, respectively.
+The temporal as well as spatial resolution of these reanalysis products is rather coarse: 3-hourly and $0.75^\circ\times 0.75^\circ$ or roughly $10.2\,\unit{km}\times 83.4\,\unit{km}$ at the location of Svanvik, respectively.
 In addition, higher resolution surface ozone reanalysis for whole Europe is available from Copernicus service for regional air quality control.
-The regional model reanalysis ensemble mean is at fairly high spatial and temporal resolution compared to the data from ECMWF: $0.1^\circ\times 0.1^\circ$ or roughly $10.4\,\unit{km}\times 11.1\,\unit{km}$ and 1-hourly, respectively.
+The regional air quality model reanalysis ensemble (CAMS regional Ensemble) is based on nine European state-of-the-art numerical air quality models (CHIMERE, EMEP, EURAD-IM, LOTUS-EUROS, MATCH, MOCAGE, SILAM, DEHM, GEM-AQ) {\bf TODO: References for the models?
+https://atmosphere.copernicus.eu/documentation-regional-systems}.
+The ensemble mean is at fairly high spatial and temporal resolution compared to the global reanalysis from ECMWF: $0.1^\circ\times 0.1^\circ$ or roughly $1.4\,\unit{km}\times 11.1\,\unit{km}$ and 1-hourly, respectively.
 The covered time periods differ and no data are available for before the turn of the millennium.
@@
 CAMSRA & ECMWF & $0.75^\circ \times 0.75^\circ$ & 3-hourly & 2003 -- 2012$^\dagger$\\
-CAMS Regional Air Quality & Copernicus & $0.1^\circ \times 0.1^\circ$ & 1-hourly & 2014 -- 2018$^\dagger$\\
+CAMS Regional Ensemble & Copernicus & $0.1^\circ \times 0.1^\circ$ & 1-hourly & 2014 -- 2018$^\dagger$\\
 \bottomhline
@@
 The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, min}} peaks about $20\,\unit{days}$ earlier than the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, and \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}}.
-With respect to the periodicity of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, it takes about $8.5\,\unit{months}$ from minimum concentration to maximum concentration, while from maximum to minimum it only takes about $4.5\,\unit{months}$.
+With respect to the periodicity of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, it takes about $7.5\,\unit{months}$ from minimum concentration to maximum concentration, while from maximum to minimum it only takes about $4.5\,\unit{months}$.
 The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}} is at $52\,\unit{ppb}$ in late April and the minimum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, min}} occurs at $15\,\unit{ppb}$ in the beginning of September.
@@
 Also CAMSRA matches the observed ozone climatology poorly (Fig.~\ref{fig:ozone_climatologies}b).
-Despite displaying only one ozone peak, it still does not reproduce the seasonality in northern Fennoscandia at all.
+Despite displaying only one ozone peak, it still does not reproduce the seasonality in northern Fennoscandia well.
 The CAMSRA spring peak appears to lag behind by $1\,\unit{month}$ and is $5\,\unit{ppb}$ lower, whereas the minimum occurs in January compared to August/September.

### 2020-04-23 VegOzone.tex
Update of reanalyses sections. Took out the copied introduction. Included some motivation. Spell-checked.

@@
 \Author[3]{Lisa}{Emberson}
+\Author[2]{Aud Else}{Berglen Eriksen}
+\Author[1]{Bjørg}{Rognerud}
 \Author[1]{Frode}{Stordal}
+\Author[1]{Terje}{Koren Berntsen}
 
@@
 
-{\bf OBS: The following is just a copy of the 2018 dry deposition paper. It has to be substantially changed to meet the no-copycat-criteria. TODO: Rewrite in context of this article.}
+Following the report by the Norwegian Meteorological institute, the summer of 2018 was the warmest and driest ever recorded in Eastern, Western, and Southern Norway with 1906 registered forest fires of different sizes.
+In the north (including Finnmark) and northwest of Norway it was amongst the warmest and driest on record.
+{\bf TODO: How to cite report of Met NOR?} In Sweden, about 500 fires had been reported (five times more than in a usual summer) and an estimated total of $25000\,\unit{hectare}$ of forest burned down in the course of extensive forest fires especially in central(?) Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n).
+({\bf TODO: How to cite governmental report of Sweden?
+https://www.regeringen.se/rattsliga-dokument/statens-offentliga-utredningar/2019/02/sou-20197/}) The heat and drought stress on vegetation was exceptional not only in large parts of Fennoscandia but also in the rest of Europe.
 
-Ozone is an important trace gas for all lifeforms on Earth.
+In the scope of our project to assess probable effects of midnight sun conditions on ozone uptake by plants and resultant enhanced susceptibility to ozone inflicted damage, we conducted ambient air ozone concentration measurements at Svanhovd in the growing seasons of 2018/19 in collaboration with NILU.
+The observed average ozone concentration in these years did not differ significantly by statistical means.
+
+In 2018, we have observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden at NIBIO Environment Centre Svanhovd.
+In contrast, no such damage has been seen in 2019.
+
+Ozone damage is rather driven by the actual uptake of ozone (PODy) than the pure ambient air concentration (AOTx).
+The environmental drivers for the PODy calculation include ambient air temperature, water availability, and humidity which influence the actual capability of ozone uptake through a corresponding change in stomatal conductance.
+
+We are going to estimate the potential main drivers for the observed ozone damage in 2018 by computing PODy values for natural and semi-natural vegetation, based on derived climatological averages of environmental drivers for Svanhovd as well as for the growing seasons 2018/19 serving as case studies.
+
+%{\bf OBS: The following is just a copy of the 2018 dry deposition paper. It has to be substantially changed to meet the no-copycat-criteria. TODO: Rewrite in context of this article.}
+
+%Ozone is an important trace gas for all lifeforms on Earth.
 Depending on the place of its occurrence it has either a positive or negative connotation.
@@
 With a radiative forcing of $0.40 \pm 0.20\,\unit{W\,m^{-2}}$, it is placed third, only surpassed by \chem{CO_2} and \chem{CH_4} \citep[Chapter 8]{IPCC2013}.\\
-In the troposphere and in particular in ambient air, ozone is considered as a highly toxic pollutant.
+%In the troposphere and in particular in ambient air, ozone is considered as a highly toxic pollutant.
 Since the industrial revolution, tropospheric background ozone concentrations have been increasing in the northern hemisphere \citep[Chapter 2]{IPCC2013}.
@@
 The impact of ozone damage on vegetation and agricultural plants may affect food security in the future especially in Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018} and might be an important additional feedback to climate change \citep{Nat:Sitch2007}.\\
-Elevated ozone levels at a site may originate from both, the local production of ozone from its precursors, which are transported, and from advection of ozone itself.
+%Elevated ozone levels at a site may originate from both, the local production of ozone from its precursors, which are transported, and from advection of ozone itself.
 Long-range ozone transport occurs regularly and might be most important in regions that otherwise lack precursors.
@@
 In a sequence of rapid reactions a peroxyl radical \chem{HO_2^\bullet} is formed through an initial reaction of \chem{CO} with a hydroxyl radical \chem{^\bullet OH}.
-Via a reaction between \chem{HO_2^\bullet} and \chem{NO}, \chem{NO_2} is formed which is then photolyzed.
+%Via a reaction between \chem{HO_2^\bullet} and \chem{NO}, \chem{NO_2} is formed which is then photolyzed.
 The resulting atomic oxygen reacts then with \chem{O_2} (and also under the presence of available co-reactants) to form an ozone molecule.
 
-Since ozone is highly reactive, its global mean life-time in the troposphere is roughly $22$ days but ranges between a few days in the tropical boundary layer to up to $1$ year in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}.
+%Since ozone is highly reactive, its global mean life-time in the troposphere is roughly $22$ days but ranges between a few days in the tropical boundary layer to up to $1$ year in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}.
 The abundance of tropospheric ozone therefore varies, e.g., with time of the day, season, altitude, location \citep{ACP:Schnell2015}, or weather conditions in general \citep{ACP:Otero2018}.
@@
 \begin{itemize}
+\item general about ozone - focus on surface ozone
 \item ozone induced damage in vegetation \emph{Contributor: Ane - maybe also ask Aud?}
@@
 There are two global reanalysis products available at the European Centre for Medium-Range Weather Forecasts (ECMWF) that include atmospheric tracers, in particular ozone (MACC and CAMSRA).
-The temporal as well as spatial resolution of these reanalysis products is rather coarse: 3-hourly and $0.75^\circ\times 0.75^\circ$ or roughly $10.2\,\unit{km}\times 83.4\,\unit{km}$ at the location of Svanvik, respectively.
-In addition, higher resolution surface ozone reanalysis for whole Europe is available from Copernicus service for regional air quality control.
-The regional air quality model reanalysis ensemble (CAMS regional Ensemble) is based on nine European state-of-the-art numerical air quality models (CHIMERE, EMEP, EURAD-IM, LOTUS-EUROS, MATCH, MOCAGE, SILAM, DEHM, GEM-AQ) {\bf TODO: References for the models?
-https://atmosphere.copernicus.eu/documentation-regional-systems}.
-The ensemble mean is at fairly high spatial and temporal resolution compared to the global reanalysis from ECMWF: $0.1^\circ\times 0.1^\circ$ or roughly $1.4\,\unit{km}\times 11.1\,\unit{km}$ and 1-hourly, respectively.
+The temporal as well as spatial resolution of these reanalysis products is rather coarse: 3-hourly and $0.75^\circ\times 0.75^\circ$ or roughly $29.3\,\unit{km}\times 83.4\,\unit{km}$ at the location of Svanvik, respectively.
+In addition, higher resolution surface ozone reanalysis for whole Europe is available from Copernicus service for regional air quality control ({\bf TODO: From May 31st on through the CAMS Atmosphere Data Store (ADS)}).
+The regional air quality model reanalysis ensemble (CAMS regional Ensemble) is based on nine European state-of-the-art numerical air quality models ({\bf TODO: Reference -  https://atmosphere.copernicus.eu/documentation-regional-systems, last accessed April 2020}).
+The ensemble mean is at fairly high spatial and temporal resolution compared to the global reanalysis from ECMWF: $0.1^\circ\times 0.1^\circ$ or roughly $3.9\,\unit{km}\times 11.1\,\unit{km}$ and 1-hourly, respectively.
 The covered time periods differ and no data are available for before the turn of the millennium.
 The CAMSRA and CAMS Regional Air Quality ensemble mean are available in near real time, but only former covers a period of sufficient length for climate analysis (2003 -- present).
-Their quality with respect to observations will be discussed in Section~\ref{subsubsec:clim_ozone}.
+All reanalyses use different but, at that time, latest versions of the operational weather forecast system (OpenIFS) from ECMWF.
+They differ substantially in the assimilated observational ozone data.
+The MACC reanalysis assimilates only satellite derived tropospheric column ozone (TC), while CAMSRA also includes ozone profiles from satellite retrievals.
+Most importantly for our study, in situ observations from ozone surface station networks are only assimilated in the CAMS Regional Ensemble.
+The quality with respect to observations in northern Fennoscandia will be discussed in Section~\ref{subsubsec:clim_ozone}.
 Without too much foreshadowing, the results prevented us from fetching the whole CAMSRA data.
@@
   \label{tab:ozone_rea}
-\begin{tabular}{llccl}
+\begin{tabular}{llcclcc}
 \tophline
-Name & Organization & \multicolumn{2}{c}{Resolution} & Time period \\
-&              & spatial & temporal & \\
+Name & Provider & \multicolumn{2}{c}{Resolution} & Time period & Meteorological forcing & \chem{O_3} assimilation\\
+&              & spatial & temporal & & \\
 \middlehline
-MACC & ECMWF & $0.75^\circ \times 0.75^\circ$ & 3-hourly & 2003 -- 2012\\
-CAMSRA & ECMWF & $0.75^\circ \times 0.75^\circ$ & 3-hourly & 2003 -- 2012$^\dagger$\\
-CAMS Regional Ensemble & Copernicus & $0.1^\circ \times 0.1^\circ$ & 1-hourly & 2014 -- 2018$^\dagger$\\
+MACC & ECMWF & $0.75^\circ \times 0.75^\circ$ & 3-hourly & 2003 -- 2012 & OPS & satellite $^\triangledown$\\
+CAMSRA & ECMWF & $0.75^\circ \times 0.75^\circ$ & 3-hourly & 2003 -- 2012 $^\dagger$ & ERA5 / OPS $^\ddagger$ & satellite $^\blacktriangledown$\\
+CAMS Regional Ensemble & Copernicus & $0.1^\circ \times 0.1^\circ$ & 1-hourly & 2014 -- 2018 $^\dagger$ & OPS $^\star$ & in situ $^\vartriangle$\\
 \bottomhline
 \end{tabular}
-\belowtable{$^\dagger$ Subset of the data used in this study. The full data set extends into near real time.} % Table Footnotes
+\belowtable{$^\dagger$ Subset of the data used in this study.
+The full data set extends into near real time; $^\ddagger$ ERA5 (2003--2016), OPS (later); $^\star$ EURAD uses WRF for downscaling of operational IFS; $^\triangledown$ MLS, OMI - tropospheric column; $^\blacktriangledown$ SCIAMACHY, MIPAS, MLS, OMI, GOME2, SBUV2 - tropospheric column + profile; $^\vartriangle$ METEO France NRT.} % Table Footnotes
 \end{table}
@@
 The MACC reanalysis (Fig.~\ref{fig:ozone_climatologies}a) is in general too low ($(9\pm 7)\,\unit{ppb}$) and displays a erroneous seasonal cycle.
-MACC reanalysis ozone concentrations peak already in March followed by a second peak in July.
+Ozone concentrations peak already in March followed by a second peak in July.
 Also CAMSRA matches the observed ozone climatology poorly (Fig.~\ref{fig:ozone_climatologies}b).
@@
 \end{figure}
+
+There are many probable explanations for the discovered differences in the climatologies derived from the three reanalyses.
+The effective changes from the MACC reanalysis to CAMSRA have been reported and discussed by {\bf TODO: Find paper and citation.} on a global level.
+Amongst others, an assimilation of ozone profiles from satellite retrieval and an upgraded ozone chemistry scheme have lead to the enhanced performance of latter.
+The higher spatial resolution of the regional air quality models used in the CAMS Regional Ensemble compared to the global reanalyses does capture more localized depletion and peak episodes in \chem{[O_3]} better and enhances the capability of the models with respect to daily and seasonal cycles in ozone {\bf TODO: Any citable source for this assumption?}.
+Latter is, of course, also affected by upgrades in the OpenIFS used as meteorological forcing.
+However, we can assume that the main driver for the strikingly different results are the assimilated observational ozone data.
+Satellite retrievals resolve \chem{[O_3]} at the surface rather poorly and hence do not constrain the global models well enough {\bf TODO: Any citable source for this?}.
 

### 2020-04-24 VegOzone.tex
Updated text with respect to forest fires - probably needs some more ellaboration. Wrote large protions of introduction.

@@
 \Author[2]{Aud Else}{Berglen Eriksen}
-\Author[1]{Bjørg}{Rognerud}
+\Author[1]{Bj{\o}rg}{Rognerud}
 \Author[1]{Frode}{Stordal}
@@
 We also observed ozone-induced visible injuries on clovers in the ozone garden at Svanhovd in 2018, but not in 2019.
-This may be due to differences in uptake of ozone (POD) between the two years.
+This may be due to differences in uptake of ozone ($\mathrm{POD_y}$) between the two years.
 Since DO3SE can be run on birch/beech and conifers, we would like to investigate whether the critical levels for these tree species were crossed during the two summer seasons.
@@
 \label{sec:intro}
-
-Following the report by the Norwegian Meteorological institute, the summer of 2018 was the warmest and driest ever recorded in Eastern, Western, and Southern Norway with 1906 registered forest fires of different sizes.
-In the north (including Finnmark) and northwest of Norway it was amongst the warmest and driest on record.
-{\bf TODO: How to cite report of Met NOR?} In Sweden, about 500 fires had been reported (five times more than in a usual summer) and an estimated total of $25000\,\unit{hectare}$ of forest burned down in the course of extensive forest fires especially in central(?) Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n).
-({\bf TODO: How to cite governmental report of Sweden?
-https://www.regeringen.se/rattsliga-dokument/statens-offentliga-utredningar/2019/02/sou-20197/}) The heat and drought stress on vegetation was exceptional not only in large parts of Fennoscandia but also in the rest of Europe.
-
-In the scope of our project to assess probable effects of midnight sun conditions on ozone uptake by plants and resultant enhanced susceptibility to ozone inflicted damage, we conducted ambient air ozone concentration measurements at Svanhovd in the growing seasons of 2018/19 in collaboration with NILU.
-The observed average ozone concentration in these years did not differ significantly by statistical means.
-
-In 2018, we have observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden at NIBIO Environment Centre Svanhovd.
-In contrast, no such damage has been seen in 2019.
-
-Ozone damage is rather driven by the actual uptake of ozone (PODy) than the pure ambient air concentration (AOTx).
-The environmental drivers for the PODy calculation include ambient air temperature, water availability, and humidity which influence the actual capability of ozone uptake through a corresponding change in stomatal conductance.
-
-We are going to estimate the potential main drivers for the observed ozone damage in 2018 by computing PODy values for natural and semi-natural vegetation, based on derived climatological averages of environmental drivers for Svanhovd as well as for the growing seasons 2018/19 serving as case studies.
 
@@
 
+{\bf TODO:
 \begin{itemize}
-\item general about ozone - focus on surface ozone
-\item ozone induced damage in vegetation \emph{Contributor: Ane - maybe also ask Aud?}
+\item general about ozone $\rightarrow$ focus on surface ozone and in particular precursors with respect to the forest fires (\chem{CO}) and heat stress (\chem{VOCs}).
+\item ozone induced damage in vegetation \emph{Contributor: Ane - maybe also ask Aud?}\\
   Ozone acts as an oxidative stress to plants.
-Its main action is imposed through reactions occurring in the cell walls and cell membranes of mesophyll cells inside the leaves.Ozone enters the leaves through stomata, that are kept open to allow for gas exchange between the leaf air spaces and the exterior air.
+Its main action is imposed through reactions occurring in the cell walls and cell membranes of mesophyll cells inside the leaves.
+Ozone enters the leaves through stomata, that are kept open to allow for gas exchange between the leaf air spaces and the exterior air.
 When the stomata are closed, ozone does not enter.
+\item[$\rightarrow$] Some info on stomatal conductance and its empirical calculation:
+  \begin{equation}
+  g^k_\text{sto, m} = g^k_\text{max, m} \cdot f^k_\text{phen} \cdot f^k_\text{light} \cdot \max{\left\{f^k_\text{min}, f^k_T \cdot f^k_D \cdot f^k_\text{SW}\right\}}.
+  \label{eq:stomatal}
+\end{equation}
+%
+The factors herein are normalized and vary within the range $0-1$.
+They account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_T$), water vapor pressure deficit ($f_D$), and soil water content ($f_\text{SW}$).
+All factors differ with plant functional type $k$.
 \item metrics of ozone
+  \begin{itemize}
+  \item[$\rightarrow$] Concentration based
     \begin{equation}
@@
     with the ambient ozone concentration $\chem{[O_3]}$ and the concentration threshold $\mathrm{x}$. A typical value for $\mathrm{x}=40\,\unit{ppb}$.
+  \item[$\rightarrow$] Flux based
     \begin{equation}
-    \mathrm{POD_y} = \int{\mathrm{max(\Phi_\chem{O_3}-y)}\cdot \mathrm{dt}},
+      \mathrm{POD_y} = \int{\mathrm{max(\Phi_\chem{[O_3]}-y)}\cdot \mathrm{dt}},
     \end{equation}
-   with the ambient ozone flux through the stomata $\Phi_\chem{[O_3]}$ and the flux threshold $\mathrm{y}$.
+    with the ambient ozone flux through the stomata $\Phi_\chem{[O_3]} = g_\mathrm{sto} \cdot \chem{[O_3]}$ and the flux threshold $\mathrm{y}$.
+  \end{itemize}
 \item present subarctic climate and conditions
@@
   \end{itemize}
-\item future climate and conditions
+\end{itemize}
+}
+
+{\bf TODO: Frode/Terje, you may help elaborate on the following paragraph on weather systems.
+I'm not an expert on synoptic scale meteorology and basically translated from Norwegian/Swedish into English.}
+
+Following a report by the Norwegian Meteorological institute ({\bf TODO: How to cite report of Met NOR?}), the summer of 2018 was the warmest and driest ever recorded in eastern, western, and southern Norway.
+In the north (including Finnmark) and northwest of Norway, it was amongst the warmest and driest on record - favorable conditions for boundary layer ozone formation.
+
+An unusually weak and northward shifted jet stream allowed for a persistent high pressure system above Fennoscandia which blocked the east-bound low pressure systems for several consecutive months.
+While in May and July whole Fennoscandia saw exceptionally warm and dry weather, the high pressure system's influence subsided in northern Fennoscandia in June allowing for lower temperatures than in the south.
+
+The such induced heat and drought stress on vegetation was exceptional not only in large parts of Fennoscandia but also in the rest of Europe, where the influence of the high pressure system extended even over a period of five months (April/May, July--September).
+These conditions gave rise to massive forest fires in different parts of Europe and thus an increase in ozone precursors (\chem{CO}, \chem{VOCs}).
+
+A total of 2079 forest fires of different size were registered in Norway in 2018, twice as many as in the preceding years 2016/17 ({\bf TODO: Cite https://www.dsb.no/reportasjearkiv/brannstatistikk-2018/}, last accessed April 2020).
+In Sweden, about 500 fires had been reported (five times more than in a usual summer) and an estimated total of $25000\,\unit{hectare}$ burned down in the course of the most extensive fires in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) in younger history ({\bf TODO: How to cite governmental report of Sweden?
+https://www.regeringen.se/rattsliga-dokument/statens-offentliga-utredningar/2019/02/sou-20197/ and report therein}).\\
+
+{\bf TODO: Future climate and conditions
   \begin{itemize}
@@
   \end{itemize}
-\item meteorological situation 2018 vs 2019
-\end{itemize}
+}
+
+Within the scope of our project to assess probable effects of midnight sun conditions on ozone uptake by plants and resultant enhanced susceptibility to ozone inflicted damage, we conducted ambient air ozone concentration measurements at Svanhovd in the growing seasons of 2018/19 in collaboration with NILU.
+In 2018, we have observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden at NIBIO Environment Centre Svanhovd.
+In contrast, no such damage has been reported in 2019.
+The respective observed average ozone concentration in 2018/19, however, did not differ significantly by statistical means.
+
+As previously mentioned, ozone damage is rather driven by the actual uptake of ozone ($\mathrm{POD_y}$) than by the accumulated ambient air concentration above a certain threshold ($\mathrm{AOT_x}$).
+Through the stomatal conductance, $\mathrm{POD_y}$ depends on environmental drivers such as ambient air temperature, photoactive radiation, soil water availability, and water vapor pressure deficit.
+
+In this study, we investigate potential main drivers for the observed ozone damage in the growing season 2018 at NIBIO Environment Centre Svanhovd in contrast to 2019.
+We compare $\mathrm{POD_y}$ values for natural and semi-natural vegetation for derived climatological averages of environmental conditions at Svanhovd as well as for the growing seasons 2018/19 which shall serve as case studies.
+
+The observation site at NIBIO Environment Centre Svanhovd and all related data are presented in Section~\ref{sec:data}.
+In Section~\ref{sec:stats}, we derive climatologies for surface ozone concentration at Svanhovd and northern Fennoscandia based on in situ observations and also compute climatologies of surface temperature, precipitation, and global radiation for Svanhovd.
+Based on these, we evaluate anomalies of environmental key variables in 2018/19 by statistical means.
+In addition, the derived surface ozone concentration climatology for northern Fennoscandia will be compared with available ozone reanalysis products.
+We use the DO3SE model to estimate the ozone uptake by natural and semi-natural vegetation to investigate the main environmental driver of the observed ozone damage in 2018 (Section~\ref{sec:do3se}).
+In Section~\ref{sec:conc} finally, we will summarize our results, draw conclusions, and give an outlook on further work.
 
@@
 \emph{Comment: We refrain from doing a full cross-evaluation for these meteorological fields, since this has been done elsewhere.
-It's somewhat clear that this is a job of the operators in the monitoring network.} {\bf TODO: Find a citation for this.}.
+It's somewhat clear that this is a job of the operators in the monitoring network.} {\bf TODO: Find a citation for this -> Met NOR report!.}.
 
 Relevant data for the growing season 2018/19 are shown in Fig.~\ref{fig:data_svanvik_2018_2019}.
-The hatched areas mark times when no ozone data has been taken.
-Note, while the downtime during winter was planed, missing data in two weeks of July 2018 were due to problems in data acquisition.
+The hatched areas mark times when no ozone data was recorded.
+Note, while the downtime during winter was planed, missing data in two weeks of July 2018 (July 9--23) were due to problems in data acquisition.
 
@@
 The spring peak has not been captured completely in 2019, for data acquisition started later than in 2018.
-In 2018, high summer ozone concentrations ($\chem{[O_3]}> 40\,\unit{ppb}$) were observed in July.
-A part of these peak values may, however, coincident with the time period of missing data.
-In the summer of 2019, ozone concentrations stayed below the threshold of $40\,\unit{ppb}$ most of the times.
+In summer 2018 (June--August), high ozone concentrations ($\chem{[O_3]} > 40\,\unit{ppb}$) were recorded 50 times.
+The highest summer ozone concentration ($\chem{[O_3]} = 50.2\,\unit{ppb}$) was measured on July 25.
+This coincides with the period of the most extensive forest fires in central Swedish (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) which occurred from July 12--29 and destroyed about $18000\,\unit{hectare}$ of forest{\bf TODO: Cite chapter 11 Swedish government report.}.
+{\bf TODO: Do we need a corresponding \chem{CO_2} emission here?} However, due to the above mentioned data acquisition problems, we miss most of the corresponding ozone data for this event and most likely also the peak ozone concentration.
+In contrast, ozone concentration only rose 18 times above the threshold of $40\,\unit{ppb}$ during the summer of 2019.
 
@@
 
-\section{Statistical analysis of atmospheric key variables}
+\section{Statistical analysis of environmental key variables}
+\label{sec:stats}
 In this section, we will evaluate atmospheric conditions in the growing seasons of 2018/19 by statistical means.
@@
 The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}} is at $52\,\unit{ppb}$ in late April and the minimum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, min}} occurs at $15\,\unit{ppb}$ in the beginning of September.
-The steep rise of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} in spring coincident with increasing temperatures as well as increase in global irradiance.
+The steep rise of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} in spring is in coincidence with increasing temperatures as well as increase in global irradiance.
 The decrease in \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} coincides with the estimated onset of the greening season based on the agricultural rule of thumb that the daily average temperature has to be above $5\,\unit{^\circ C}$ for $5\,\unit{days}$ consecutively (day of the year 110--150 at Svanvik, Appendix Fig~\ref{fig:greening_season_change_Svanvik}).
@@
 \section{DO3SE modeling}
+\label{sec:do3se}
 To investigate which of the environmental factors lead to the observed ozone damage on clover and tobacco {\bf TODO: Ane confirm!} at Svanhovd, we have modeled the ozone uptake using the DO3SE model.
@@
 \conclusions  %% \conclusions[modified heading if necessary]
+\label{sec:conc}
 TEXT

### 2020-04-24 VegOzone.tex
Cosmetix.

@@
 This coincides with the period of the most extensive forest fires in central Swedish (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) which occurred from July 12--29 and destroyed about $18000\,\unit{hectare}$ of forest{\bf TODO: Cite chapter 11 Swedish government report.}.
-{\bf TODO: Do we need a corresponding \chem{CO_2} emission here?} However, due to the above mentioned data acquisition problems, we miss most of the corresponding ozone data for this event and most likely also the peak ozone concentration.
+{\bf TODO: Do we need a corresponding \chem{CO_2} emission here?} However, due to the above mentioned data acquisition problems, we missed most of the corresponding ozone data for this event and most likely also the peak ozone concentration.
 In contrast, ozone concentration only rose 18 times above the threshold of $40\,\unit{ppb}$ during the summer of 2019.

### 2020-04-27 VegOzone.tex
Updates Ane.

@@
 In 2018, we have observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden at NIBIO Environment Centre Svanhovd.
-In contrast, no such damage has been reported in 2019.
+In contrast, no such damage was found on the clovers in 2019, although it was found on the sensitive tobacco cultivar.
+This indicates that the vegetation may have been more affected by ozone in 2018 than in 2019.
 The respective observed average ozone concentration in 2018/19, however, did not differ significantly by statistical means.
@@
 The hatched areas mark times when no ozone data was recorded.
-Note, while the downtime during winter was planed, missing data in two weeks of July 2018 (July 9--23) were due to problems in data acquisition.
+Note, while the downtime during winter was planned, missing data in two weeks of July 2018 (July 9--23) were due to problems in data acquisition.
 
@@
 In Table~\ref{tab:ebas_obs}, we give an overview over the exact locations and time range of operational data taking.
-All stations with available long term observations are locate on higher ground than Svanhovd which will result in higher surface ozone concentrations in comparison as ozone abundance, in general, decreases with decreasing altitude in the absence of local precursors.
+All stations with available long term observations are located on higher ground than Svanhovd which will result in higher surface ozone concentrations in comparison as ozone abundance, in general, decreases with decreasing altitude in the absence of local precursors.
 {\bf TODO: Citation!} {\bf TODO: The following has to be confirmed and elaborated on!
@@
 Especially, ozone monitors did not undergo a regular re-calibration.
-This leads likely to drifts in the observed data and may impose a false periodicity.
+This probably leads to drifts in the observed data and may impose a false periodicity.
 In 1997, the ozone monitoring at Jergul was relocated downstream the river Karasjokka to a site close by Karasjok.
-Although the same equipment has been used, this might have introduced a breaking point in the combined data series denoted as Jergul/Karasjok in the following which would be of a problem if this was the sole data to compare with.
+Although the same equipment has been used, this might have introduced a breaking point in the combined data series, denoted as Jergul/Karasjok in the following, which would be of a problem if this was the sole data to compare with.
 Only few data have been taken upstream the Pasvik river at Janiskovski, hence these will not be taken into further consideration, but are shown for completion.

### 2020-04-27 VegOzone.tex
Corection of wording.

@@
 The original farm buildings, protected by law, are located in the lower center, while the Svanhovd Environment Centre follows to the north-east.
-The Centre hosts offices, a conference center, a national park center, a brown bear exhibition as well as a modern laboratory for DNA analysis.
+The Centre hosts offices, a conference center, a national park center, and a modern laboratory for DNA analysis.
 Left of the Centre and north of the original farm buildings a botanical garden can be found.
 It is one amongst three such gardens north of the Polar Circle.
-The Svanhovd Environment Centre focuses on research of the special environment of the region, but also comprises long-term observations of various agrameteorological variables, i.a.
+The Svanhovd Environment Centre focuses on research of the special environment of the region, but also comprises long-term observations of various agrometeorological variables, i.a.
 $2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths.
@@
 
-All agrameteorological variables including temperature and precipitation are available from September 1992 to present day (LandbruksMeteorologisk Tjeneste provided by NIBIO -- \href{https://lmt.nibio.no/}{lmt.nibio.no}, note the station name here is Pasvik).
+All agrometeorological variables including temperature and precipitation are available from September 1992 to present day (LandbruksMeteorologisk Tjeneste provided by NIBIO -- \href{https://lmt.nibio.no/}{lmt.nibio.no}, note the station name here is Pasvik).
 

### 2020-04-28 VegOzone.tex
Suggested changes by Frode.

@@
 
-{\bf TODO: Frode/Terje, you may help elaborate on the following paragraph on weather systems.
-I'm not an expert on synoptic scale meteorology and basically translated from Norwegian/Swedish into English.}
-
-Following a report by the Norwegian Meteorological institute ({\bf TODO: How to cite report of Met NOR?}), the summer of 2018 was the warmest and driest ever recorded in eastern, western, and southern Norway.
-In the north (including Finnmark) and northwest of Norway, it was amongst the warmest and driest on record - favorable conditions for boundary layer ozone formation.
-
-An unusually weak and northward shifted jet stream allowed for a persistent high pressure system above Fennoscandia which blocked the east-bound low pressure systems for several consecutive months.
-While in May and July whole Fennoscandia saw exceptionally warm and dry weather, the high pressure system's influence subsided in northern Fennoscandia in June allowing for lower temperatures than in the south.
-
-The such induced heat and drought stress on vegetation was exceptional not only in large parts of Fennoscandia but also in the rest of Europe, where the influence of the high pressure system extended even over a period of five months (April/May, July--September).
+Following a report by the Norwegian Meteorological institute ({\bf TODO: Citation https://www.met.no/publikasjoner/met-info/met-info-2019}), the summer of 2018 was the warmest and driest ever recorded in eastern, western, and southern Norway.
+In the north (including Finnmark), it was amongst the warmest on record - favorable conditions for boundary layer ozone formation.
+
+An unusually weak and northward shifted jet stream allowed for a persistent high pressure system above northern Europe, including Fennoscandia which blocked the pressure systems for several consecutive months.
+In the periode May -- July, southern Norway had temperatures $4\,\unit{^\circ C}$ above normal.
+Whgile southern Norway had only about $60\,\unit{\%}$ of normal precipitation, northern Norway as a whle had close to normalprecipitation, but with local variations.
+
+Heat and drought stress on vegetation was exceptional not only in large parts of Fennoscandia but also in much of Europe, where the influence of the high pressure system extended even over a period of five months (April/May, July--September).
 These conditions gave rise to massive forest fires in different parts of Europe and thus an increase in ozone precursors (\chem{CO}, \chem{VOCs}).
@@
 In Sweden, about 500 fires had been reported (five times more than in a usual summer) and an estimated total of $25000\,\unit{hectare}$ burned down in the course of the most extensive fires in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) in younger history ({\bf TODO: How to cite governmental report of Sweden?
-https://www.regeringen.se/rattsliga-dokument/statens-offentliga-utredningar/2019/02/sou-20197/ and report therein}).\\
+https://www.regeringen.se/rattsliga-dokument/statens-offentliga-utredningar/2019/02/sou-20197/ and report therein}).
+
+Elevated ozone in Europe was liekly due to several other factors as well, like in the 2003 drought period, such as biogenic VOC emissions, reduced dry deposition, reduced clouds and increased solar radiation, photochemical reactions, and stagnant weather conditions \citep{JGR:Solberg2018}\\
 

### 2020-04-28 VegOzone.tex
Comments by Ane.

@@
 In our assessment of ozone damage on vegetation in the subarctic in relation to the 2018 heatwave, we focus on the NIBIO operated Environment Centre Svanhovd which lies close by the settlement of Svanvik in the Pasvik valley (Section~\ref{subsec:svanhovd}).
-Throughout the paper, we will use Svanvik synonymously for the atmospheric monitoring site at Svanhovd.
+Throughout the paper, we will use Svanhovd synonymously for the atmospheric monitoring site at Svanvik.
 The available meteorological monitoring data are described in Section~\ref{subsubsec:atmo_svanvik}.
@@
 The original farm buildings, protected by law, are located in the lower center, while the Svanhovd Environment Centre follows to the north-east.
-The Centre hosts offices, a conference center, a national park center, and a modern laboratory for DNA analysis.
 Left of the Centre and north of the original farm buildings a botanical garden can be found.

### 2020-04-28 VegOzone.tex
Updated text with respect to data from SeNorge.

@@
 The steep rise of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} in spring is in coincidence with increasing temperatures as well as increase in global irradiance.
-The decrease in \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} coincides with the estimated onset of the greening season based on the agricultural rule of thumb that the daily average temperature has to be above $5\,\unit{^\circ C}$ for $5\,\unit{days}$ consecutively (day of the year 110--150 at Svanvik, Appendix Fig~\ref{fig:greening_season_change_Svanvik}).
+The decrease in \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} begins about one month ahead of the estimated onset of the greening season.
+Based on data from SeNorge.no and the agricultural rule of thumb that the daily average temperature has to be above $5\,\unit{^\circ C}$ for $5\,\unit{days}$ consecutively, we find $\left<G_\mathrm{begin}\right> = (150 \pm 14)\,\unit{days of the year}$ at Svanvik (Appendix Fig~\ref{fig:greening_season_change_Svanvik}), which translates into the end of May.
 
@@
 The spring peak apparently occurs slightly earlier (in the beginning of April).
-As Svanvik is located at lower altitude than the other sites, the earlier ozone peak may point to a strong interplay between ozone uptake by vegetation after the onset of greening, for at higher altitude plants sprout later.
-{\bf TODO: Is there a way to confirm this?
+As Svanvik is located at lower altitude than the other sites, the earlier ozone peak may point to an interplay with ozone uptake by vegetation.
+{\bf TODO: The data from SeNorge seems to point to a later onset of greening, than what I interfered from OpenIFS data.
+This means that the beginning of the growing season is not a good measure for the determination of the ozone spring peak.
+This does not exclude the early start of pine forest photosynthesis as a driver.
+Is there a way to confirm this?
 Actual data for the three sites?
@@
   \includegraphics[width=8.3cm]{greening_season_change_Svanvik}
-  \caption{Estimated shift and prolongation of greening season at Svanhovd over the past two decades based on the method and data described in \citet{GMD:Falk2019}.}
+  \caption{Estimated shift and prolongation of greening season at Svanhovd over the past 6 decades based on data from SeNorge.no {\bf TODO: citation http://www.senorge.no}.}
   \label{fig:greening_season_change_Svanvik}

### 2020-04-29 VegOzone.tex
Include ozone reconstruction plot again.
Added plots and analyse for ozone reconstruction ion 2018.

@@
 An unusually weak and northward shifted jet stream allowed for a persistent high pressure system above northern Europe, including Fennoscandia which blocked the pressure systems for several consecutive months.
-In the periode May -- July, southern Norway had temperatures $4\,\unit{^\circ C}$ above normal.
-Whgile southern Norway had only about $60\,\unit{\%}$ of normal precipitation, northern Norway as a whle had close to normalprecipitation, but with local variations.
+In the period May -- July, southern Norway had temperatures $4\,\unit{^\circ C}$ above normal.
+While southern Norway had only about $60\,\unit{\%}$ of normal precipitation, northern Norway as a whole had close to normal precipitation, but with local variations.
 
@@
 
-Elevated ozone in Europe was liekly due to several other factors as well, like in the 2003 drought period, such as biogenic VOC emissions, reduced dry deposition, reduced clouds and increased solar radiation, photochemical reactions, and stagnant weather conditions \citep{JGR:Solberg2018}\\
+Elevated ozone in Europe was likely due to several other factors as well, like in the 2003 drought period, such as biogenic VOC emissions, reduced dry deposition, reduced clouds and increased solar radiation, photochemical reactions, and stagnant weather conditions \citep{JGR:Solberg2018}.\\
 
@@
 {\bf TODO: The data from SeNorge seems to point to a later onset of greening, than what I interfered from OpenIFS data.
-This means that the beginning of the growing season is not a good measure for the determination of the ozone spring peak.
+This means that the beginning of the growing season is not a good measure for the termination of the ozone spring peak.
 This does not exclude the early start of pine forest photosynthesis as a driver.
+Coniferous forest starts to photosynthesis earlier than what the growing season begin would mark.
+I remember Håkan or Per Erik(?) talking about the \chem{CO_2} uptake and phenology of coniferous trees - but still have not found a related publication?
 Is there a way to confirm this?
@@
 \label{sec:do3se}
-To investigate which of the environmental factors lead to the observed ozone damage on clover and tobacco {\bf TODO: Ane confirm!} at Svanhovd, we have modeled the ozone uptake using the DO3SE model.
+To investigate which of the environmental factors lead to the observed ozone damage on clover and tobacco in 2018 in contrast to 2019, when no such damage was found on the clovers but only on the sensitive tobacco cultivar, we model the ozone uptake using the DO3SE model.
+
+\subsection{Reconstruction of missing ozone values in 2018}
+\label{subsec:ozone_reco}
+As we have shown, the available reanalysis products have their limitations.
+Although the quality of the Copernicus Regional Model Reanalysis is fairly good, we additionally deem it necessary to reconstruct the missing data in 2018 for the purpose of DO3SE modeling.
+In Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}, the procedure for a probable reconstruction is shown.
+First, we derive hourly climatologies for northern Fennoscandia \chem{\left<[O_3]\right>_{hourly}} and Svanvik \chem{\left<[O_3]\right>_{hourly}^{Svanvik}} (Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}a).
+Then we compute temporal correlations between Svanvik and all other stations with respect to a temporal shift (Appendix Fig.~\ref{fig:time_lag_correlation}) to estimate a probable time lag between the stations (e.g., $3\,\unit{h}$ for Esrange and Pallas).
+Because of the high correlation with Svanvik observations, we have chosen Pallas as reference station for the reconstruction.
+We correct the Fennoscandic climatology for the found time lag\chem{\left<[O_3]\right>_{hourly,\,t-3}} and use it to derive a mapping function:
+\begin{equation}
+  X = \frac{\chem{\left<[O_3]\right>_{hourly}^{Svanvik}}}{\chem{\left<[O_3]\right>_{hourly,\,t-3}}}.
+  \label{eq:mapping}
+\end{equation}
+Performing a time lag correction basically means to shift the respective time series back in time by $3\,\unit{h}$.
+As an intermediate step, hourly anomalies are computed
+\begin{equation}
+  \chem{\Delta[O_3]_{hourly}^{Pallas}} = \chem{[O_3]_{hourly}^{Pallas}}-\chem{\left<[O_3]\right>_{hourly}}.
+\end{equation}
+In Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}b), hourly ozone concentration anomalies are shown for Pallas as well as for Esrange and Svanvik.
+We also correct ozone anomalies for Pallas for the time lag \chem{\Delta[O_3]_{hourly,\,t-3}^{Pallas}} and use the mapping function in Eq.~(\ref{eq:mapping}) to derive reconstructed anomalies for the missing values:
+\begin{equation}
+    \chem{\Delta [O_3]_{hourly}^{Svanvik,\,reco}} = \chem{\Delta[O_3]_{hourly,\,t-3}^{Pallas}} \cdot X
+\end{equation}
+Finally, we add the bias corrected Svanvik ozone climatology and derive a reconstructed ozone concentration time series \chem{[O_3]_{hourly}^{Svanvik,\,reco}}.
+In Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}c), the result is shown together with the observed data and the Copernicus Regional Model Reanalysis at the nearest grid point for comparison.
+
+\subsection{Model description and input data}
+
+{\bf TODO: Some general words about DO3SE model?}
+
+\subsection{POD and ozone impact}
 
@@
 \end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{ozone_reconstruction_2018_07}
+  \caption{Reconstruction of missing \chem{[O_3]} observation data at Svanvik in July 2018 based on derived hourly climatologies of northern Fennoscandia, Svanvik, and data Pallas.
+(a) Time series; (b) anomalies; (c) reconstruction.}
+  \label{fig:ozone_reconstruction_2018_07}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{ozone_observation_timelag_svanvik}
+  \caption{Temporal correlation of \chem{[O_3]} data at Svanvik with other ozone monitoring stations in northern Fennoscandia.
+A negative lag means that Svanvik lags behind, while a positive lag mean the other station lags behind.
+The highest correlation with Pallas/Esrange is found at a time lag of $3\,\unit{h}$, for Jergul/Karasjok at $1\,\unit{h}$, and for Janiskoski at $-20\,\unit{h}$.}
+  \label{fig:time_lag_correlation}
+\end{figure}
+
 \clearpage
@@
 %\appendixfigures  %% needs to be added in front of appendix figures
-
-%\begin{figure}[t]
-%  \includegraphics[width=8.3cm]{ozone_obs_2018_07}
-%  \caption{Estimate of missing observation \chem{[O_3]} data at Svanhovd in July 2018 based on derived hourly climatologies of northern Fennoscandia, Svanvik, and Copernicus regional model reanalysis ensemble mean.
-(a) Time series; (b) Student's t-test.}
-%  \label{fig:ozone_obs_2018_rrea}
-%\end{figure}
 

### 2020-05-04 VegOzone.tex
Included some updated citations.

@@
   \item[$\rightarrow$] midnight sun and 23\,\unit{h} open stomata (biological details will be published in a component paper)
-  \item[$\rightarrow$] ozone climatology and spring peak {\bf TODO: Cite review article - find newer articles.}
+  \item[$\rightarrow$] ozone climatology and spring peak \citep{AE:Monks2000, AB:Klingberg2009, AE:Rummukainen1996} {\bf TODO: Find newer articles.}
   \end{itemize}
@@
   \item[$\rightarrow$] increase in heatwaves and dry periods
-  \item[$\rightarrow$] shift in start of growing season (into vicinity of ozone spring peak)
+  \item[$\rightarrow$] shift in start of growing season (into vicinity of ozone spring peak) \citep{GCB:Menzel2006,RS:Hogda2013}
   \item[$\rightarrow$] increase in tropospheric background ozone

### 2020-05-05 VegOzone.tex
Added the corresponding citeations for the reports from Norway and Sweden.

@@
 
-Following a report by the Norwegian Meteorological institute ({\bf TODO: Citation https://www.met.no/publikasjoner/met-info/met-info-2019}), the summer of 2018 was the warmest and driest ever recorded in eastern, western, and southern Norway.
+Following a report by the Norwegian Meteorological institute \citep{MetNOR2019}, the summer of 2018 was the warmest and driest ever recorded in eastern, western, and southern Norway.
 In the north (including Finnmark), it was amongst the warmest on record - favorable conditions for boundary layer ozone formation.
@@
 
-A total of 2079 forest fires of different size were registered in Norway in 2018, twice as many as in the preceding years 2016/17 ({\bf TODO: Cite https://www.dsb.no/reportasjearkiv/brannstatistikk-2018/}, last accessed April 2020).
-In Sweden, about 500 fires had been reported (five times more than in a usual summer) and an estimated total of $25000\,\unit{hectare}$ burned down in the course of the most extensive fires in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) in younger history ({\bf TODO: How to cite governmental report of Sweden?
-https://www.regeringen.se/rattsliga-dokument/statens-offentliga-utredningar/2019/02/sou-20197/ and report therein}).
+A total of 2079 forest fires of different size were registered in Norway in 2018, twice as many as in the preceding years 2016/17 \citep{DSB2019}, last accessed April 2020).
+In Sweden, about 500 fires had been reported (five times more than in a usual summer) and an estimated total of $25000\,\unit{hectare}$ burned down in the course of the most extensive fires in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) in younger history \citep{SOU2019}.
 
@@
 The respective observed average ozone concentration in 2018/19, however, did not differ significantly by statistical means.
+{\bf TODO: Citation NILU ozone report 2019 - high AOT40 in 2018 \citep{NILU2019}.}
 
@@
 
-{\bf TODO: Some more (technical) details about ozone and atmospheric monitoring at Svanhovd. $\rightarrow$ Sverre?}
+{\bf TODO: Some more (technical) details about ozone and atmospheric monitoring at Svanhovd? $\rightarrow$ Sverre?}
 
 \emph{Comment: We refrain from doing a full cross-evaluation for these meteorological fields, since this has been done elsewhere.
-It's somewhat clear that this is a job of the operators in the monitoring network.} {\bf TODO: Find a citation for this -> Met NOR report!.}.
+It's somewhat clear that this is a job of the operators in the monitoring network \citep{MetNOR2019}.}
 
@@
 The highest summer ozone concentration ($\chem{[O_3]} = 50.2\,\unit{ppb}$) was measured on July 25.
-This coincides with the period of the most extensive forest fires in central Swedish (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) which occurred from July 12--29 and destroyed about $18000\,\unit{hectare}$ of forest{\bf TODO: Cite chapter 11 Swedish government report.}.
+This coincides with the period of the most extensive forest fires in central Swedish (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) which occurred from July 12--29 and destroyed about $18000\,\unit{hectare}$ of forest \citep{SOU2019}.
 {\bf TODO: Do we need a corresponding \chem{CO_2} emission here?} However, due to the above mentioned data acquisition problems, we missed most of the corresponding ozone data for this event and most likely also the peak ozone concentration.
@@
 All stations with available long term observations are located on higher ground than Svanhovd which will result in higher surface ozone concentrations in comparison as ozone abundance, in general, decreases with decreasing altitude in the absence of local precursors.
-{\bf TODO: Citation!} {\bf TODO: The following has to be confirmed and elaborated on!
+{\bf TODO: Citation see references in \citet{AB:Klingberg2009}!} {\bf TODO: The following has to be confirmed and elaborated on!
 Ane?} The vegetation at all sites is similar and consists mainly of pine forests, birch, and heath.
@@
 Data taken from EBAS.
-The hatched areas indicate time periods with insufficient quality control.
-\bf{TODO: citation NILU ozone report 200x}.}
+The hatched areas indicate time periods with insufficient quality control \citep{NILU2003}.}
   \label{fig:ozone_timesseries_fenoscandic_obs}
@@
 Details of reanalysis data sets referred to in this work are listed in Table~\ref{tab:ozone_rea}.
-There are two global reanalysis products available at the European Centre for Medium-Range Weather Forecasts (ECMWF) that include atmospheric tracers, in particular ozone (MACC and CAMSRA).
+There are two global reanalysis products available at the European Centre for Medium-Range Weather Forecasts (ECMWF) that include atmospheric tracers, in particular ozone (MACC and CAMSRA) \citep{ACP:Inness2013, ACP:Inness2019}.
 The temporal as well as spatial resolution of these reanalysis products is rather coarse: 3-hourly and $0.75^\circ\times 0.75^\circ$ or roughly $29.3\,\unit{km}\times 83.4\,\unit{km}$ at the location of Svanvik, respectively.
@@
 Coniferous forest starts to photosynthesis earlier than what the growing season begin would mark.
-I remember Håkan or Per Erik(?) talking about the \chem{CO_2} uptake and phenology of coniferous trees - but still have not found a related publication?
-Is there a way to confirm this?
-Actual data for the three sites?
-Please correct if wrong!
-Ane?
-Aud?} The average ozone concentration at Svanvik is lower than at the other sites, as already mentioned, which is in league with a decrease in tropospheric ozone with decreasing altitude.
+\chem{CO_2} uptake and phenology of coniferous trees \citep{TB:Kolari2007} starts as early as doy 100, acclimates to leaf temperature history - photosynthesis drops at around doy 200--250.} The average ozone concentration at Svanvik is lower than at the other sites, as already mentioned, which is in league with a decrease in tropospheric ozone with decreasing altitude.
 In July--September, ozone is occasionally almost completely depleted during night time.
@@
 There are many probable explanations for the discovered differences in the climatologies derived from the three reanalyses.
-The effective changes from the MACC reanalysis to CAMSRA have been reported and discussed by {\bf TODO: Find paper and citation.} on a global level.
+The effective changes from the MACC reanalysis to CAMSRA have been reported and discussed by \citet{ACP:Inness2019} on a global level.
 Amongst others, an assimilation of ozone profiles from satellite retrieval and an upgraded ozone chemistry scheme have lead to the enhanced performance of latter.

### 2020-05-07 VegOzone.tex
An AOT40 plot added which shal be moved to supplements later on - some citations added.

@@
 
+In terms of AOT40, the threshold for protection of vegetation of $3000\,\unit{ppb}$ given by the EU directive has been well exceeded at many ozone monitoring sites in Southern Norway in the 2018 growing season \citep{NILU2019}.
+{\bf TODO: come back to this later on in the discussions(?) - it has not been breached at Svanvik in northern Fennoscandia.
+Later supplement figure \ref{fig:fennoscandic_aot40}}
+
 Within the scope of our project to assess probable effects of midnight sun conditions on ozone uptake by plants and resultant enhanced susceptibility to ozone inflicted damage, we conducted ambient air ozone concentration measurements at Svanhovd in the growing seasons of 2018/19 in collaboration with NILU.
@@
 The respective observed average ozone concentration in 2018/19, however, did not differ significantly by statistical means.
-{\bf TODO: Citation NILU ozone report 2019 - high AOT40 in 2018 \citep{NILU2019}.}
 
@@
 Data from before the 1990s is available solely from Svanvik and Karasjok.
-These data, however, do not follow the high quality standard procedures implemented nowadays and have to be treated with care {\bf TODO: citation NILU ozone report 200x}.
+These data, however, do not follow the high quality standard procedures implemented nowadays and have to be treated with care \citep{NILU2003}.
 Especially, ozone monitors did not undergo a regular re-calibration.
@@
 Thus, Svanvik cannot directly be represented by any measurements taken at either Esrange or Pallas.
+We shall come back to this in Section~\ref{subsec:ozone_reco}.
 
@@
 \appendixfigures  %% needs to be added in front of appendix figures
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{global_rad_clim}
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{global_rad_clim}
   \caption{Observed global irradiance at Svanvik.
@@
   \label{fig:global_rad_clim}
-\end{figure}
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{test_fennoscandia}
+\end{figure*}
+
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{test_fennoscandia}
   \caption{Maximum daily sun shine duration based on geometrical calculations at different sites in Fennoscandia.
@@
   \label{fig:sunlight_fennoscandia}
-\end{figure}
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{greening_season_change_Svanvik}
+\end{figure*}
+
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{greening_season_change_Svanvik}
   \caption{Estimated shift and prolongation of greening season at Svanhovd over the past 6 decades based on data from SeNorge.no {\bf TODO: citation http://www.senorge.no}.}
   \label{fig:greening_season_change_Svanvik}
-\end{figure}
+\end{figure*}
 \clearpage
@@
 
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{density_distribution}
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{density_distribution}
   \caption{Probability densities and correlation coefficient for \chem{[O_3]} between different sites in northern Fennoscandia.
@@
   \label{fig:density_distribution}
-\end{figure}
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_residuals}
+\end{figure*}
+
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{ozone_climatology_fenoscandic_obs_residuals}
   \caption{Probability density functions of ozone concentration residuals. 2018 with respect to respective climatology for different sites in Fennoscandia. (a) Esrange; (b) Pallas.}
   \label{fig:ozone_climatology_fenoscandic_obs_residuals}
-\end{figure}
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_residuals-Svanvik}
+\end{figure*}
+
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{ozone_climatology_fenoscandic_obs_residuals-Svanvik}
   \caption{Probability density functions of ozone concentration residuals. 2018/19 observations at Svanhovd with respect to derived climatologies for (a) Northern Fennoscandia; (b) Svanvik.}
   \label{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}
-\end{figure}
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_test}
+\end{figure*}
+
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{ozone_climatology_fenoscandic_obs_test}
   \caption{Student's t-test assuming same sample uncertainty in both, climatology and 2018 observations for different sites in Fennoscandia.
@@
   \label{fig:ozone_climatology_fenoscandic_obs_test}
-\end{figure}
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_test-Svanvik}
+\end{figure*}
+
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{ozone_climatology_fenoscandic_obs_test-Svanvik}
   \caption{Student's t-test assuming same sample uncertainty in both, climatology and 2018/19 observations at Svanhovd with respect to derived climatologies for (a) Northern Fennoscandia; (b) Svanvik.
@@
   \label{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}
-\end{figure}
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_reconstruction_2018_07}
+\end{figure*}
+
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{ozone_reconstruction_2018_07}
   \caption{Reconstruction of missing \chem{[O_3]} observation data at Svanvik in July 2018 based on derived hourly climatologies of northern Fennoscandia, Svanvik, and data Pallas.
@@
   \label{fig:ozone_reconstruction_2018_07}
-\end{figure}
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_observation_timelag_svanvik}
+\end{figure*}
+
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{ozone_observation_timelag_svanvik}
   \caption{Temporal correlation of \chem{[O_3]} data at Svanvik with other ozone monitoring stations in northern Fennoscandia.
@@
   \label{fig:time_lag_correlation}
-\end{figure}
+\end{figure*}
+
+% Shall go to supplement!
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{ozone_fennoscandic_obs_aot40}
+  \caption{Integrated AOT40 for various sites in northern Fennoscandia over the course of 33 years.
+For simplicity, we integrated hourly ozone between $1\,\unit{am}-11\,\unit{pm}$ in any case, although a night time light intensity $> 50\,\unit{W\,m^{-2}}$ is only meassured during midnight sun conditions.
+Since night time \chem{[O_3]} is mostly below $40\,\unit{ppb}$, the such induced high bias in spring should be tolerable.
+The dashed red lines indicate the respective threshold given by the EU directive.
+(a) May 1 -- July 31 (b) April 1 -- September 30.}
+  \label{fig:time_lag_correlation}
+\end{figure*}
 

### 2020-05-08 VegOzone.tex
Update captions accordingly.

@@
 Negative deviations from climatology are displayed as negative percentage.
-The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The dashed lines indicate the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
+The full annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
+The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
 The summer of 2018 was significantly warmer than average.
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{temperature_signific}
+  \includegraphics[width=8.3cm]{temperature_signific_1}
   \caption{Temperature anomalies at Svanvik.
@@
 The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The dashed lines indicate the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
-Year displayed in panel (a) 2018; (b) 2019.
+The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
 The summer of 2018 was significantly warmer than average, especially May and July.
@@
 As above, negative deviations from the climatology are shown as negative percentage and annual positive/negative deviations are indicated in the respective corners.
-The dashed lines indicate the expected percentage of values falling above $\pm\frac{1}{2}\,\sigma$ if a normal distribution is assumed ($30.8\,\unit{\%}$).
+The hatched area between the dashed lines indicates the expected percentage of values falling above $\pm\frac{1}{2}\,\sigma$ if a normal distribution is assumed ($30.8\,\unit{\%}$).
 Unlike for the temperature anomalies, the picture is not as clear.
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{precipitation_signific}
+  \includegraphics[width=8.3cm]{precipitation_signific_1}
   \caption{Precipitation anomalies at Svanvik.
@@
 The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The dashed lines indicate the expected percentage of values falling above $\pm\frac{1}{2}\,\sigma$ if a normal distribution is assumed ($30.8\,\unit{\%}$).
-Year displayed in panel (a) 2018; (b) 2019.
-The summer of 2018 was significantly drier than in 2019.}
+The hatched area between the dashed lines indicates the expected percentage of values falling above $\pm\frac{1}{2}\,\sigma$ if a normal distribution is assumed ($30.8\,\unit{\%}$).
+Summer was significantly dry only in 2018, while fall was significantly dry in both years.}
   \label{fig:plot_precipitation_anomalies_svanvik}
@@
 The percentage of days above/below $\pm 1\,\sigma$ from the climatological mean is displayed for each month.
-The dashed lines indicate the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed.
+The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed.
 In summer 2018 (May/July), global irradiance was significantly higher than on average, while 2019 was generally an average year with a slightly darker spring (April/May).
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{global_rad_signific}
+  \includegraphics[width=8.3cm]{global_rad_signific_1}
   \caption{Global irradiance anomalies at Svanvik.
@@
 The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The dashed lines indicate the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
-Year displayed in panel (a) 2018; (b) 2019.
-In summer 2018 (May/July), global irradiance was significantly higher than on average, while 2019 was generally an average year.}
+The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
+In summer 2018 (May/July), global irradiance was significantly higher than on average, while 2019 had significantly less than normal irradiance in April/May and more in June.}
   \label{fig:global_rad_signific}
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_signific}
+  \includegraphics[width=8.3cm]{ozone_signific_1}
   \caption{Ozone concentration anomalies at Svanvik.
@@
 The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The dashed lines indicate the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
-Year displayed in panel (a) 2018; (b) 2019.
+The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
+July 2018 has not been corrected for the missing data due to data acquisition problems.
+The corrected value (refer to Section~\ref{subsec:ozone_reco} for details) is indicated by the star.
 In April/May 2018, ozone was significantly higher than on average, while 2019 was generally an average year.
@@
   \caption{Integrated AOT40 for various sites in northern Fennoscandia over the course of 33 years.
-For simplicity, we integrated hourly ozone between $1\,\unit{am}-11\,\unit{pm}$ in any case, although a night time light intensity $> 50\,\unit{W\,m^{-2}}$ is only meassured during midnight sun conditions.
+For simplicity, we integrated hourly ozone between $1\,\unit{am}-11\,\unit{pm}$ in any case, although a night time light intensity $> 50\,\unit{W\,m^{-2}}$ is only measured during midnight sun conditions.
 Since night time \chem{[O_3]} is mostly below $40\,\unit{ppb}$, the such induced high bias in spring should be tolerable.
@@
 (a) May 1 -- July 31 (b) April 1 -- September 30.}
-  \label{fig:time_lag_correlation}
+  \label{fig:fennoscandic_aot40}
 \end{figure*}

### 2020-05-08 VegOzone.tex
Updated captions and text in accordance to new plots - AOT40 plot needs to be integrated.

@@
 \Author[3]{Lisa}{Emberson}
+\Author[3]{Connie}{O'Neill}
 \Author[2]{Aud Else}{Berglen Eriksen}
@@
 In both years, ozone concentrations were significantly below average in November.
+July 2018 has not been corrected for the missing data due to data acquisition problems.
+The corrected value is, however, indicated by the star (refer to Section~\ref{subsec:ozone_reco} for details of the used method for gap filling).
 
@@
 July 2018 has not been corrected for the missing data due to data acquisition problems.
-The corrected value (refer to Section~\ref{subsec:ozone_reco} for details) is indicated by the star.
+The corrected value is indicated by the star (refer to Section~\ref{subsec:ozone_reco} for details).
 In April/May 2018, ozone was significantly higher than on average, while 2019 was generally an average year.

### 2020-05-15 VegOzone.tex
Added new plots and citations.

@@
 Based on the meteorological conditions, we assume 2019 as average year which leads to a constant bias of about $1.2\,\unit{ppb}$.
-Daily \chem{\Delta_{present-past}\left<[O_3]\right>} from Esrange and Pallas, wherein present is defined as the average of 2010--2018 and past as the respective beginning of data taking until 1997, do not reveal a clear trend: We find for Esrange $(1.5\pm 3.1)\,\unit{ppb}$ and for Pallas $(-2.5\pm 3.7)\,\unit{ppb}$.
+Daily \chem{\Delta_{present-past}\left<[O_3]\right>} from Esrange and Pallas, wherein present is defined as the average of 2010--2018 and past as the respective beginning of data taking until 1997, do not reveal a clear trend: We find for Esrange $(1.5\pm 3.1)\,\unit{ppb}$ and for Pallas $(-2.5\pm 3.7)\,\unit{ppb}$ \citep{NILU2003, BER:Ruoho-Airola2015}.
 Former is compatible with our assumed bias for the climatology derived for Svanvik.

### 2020-06-05 VegOzone.tex
Included some of the citations which have been added.

@@
   \item[$\rightarrow$] increase in heatwaves and dry periods
-  \item[$\rightarrow$] shift in start of growing season (into vicinity of ozone spring peak) \citep{GCB:Menzel2006,RS:Hogda2013}
-  \item[$\rightarrow$] increase in tropospheric background ozone
+  \item[$\rightarrow$] shift in start of growing season (into vicinity of ozone spring peak) \citep{GCB:Menzel2006,RS:Hogda2013,IJB:Karlsen2007}
+  \item[$\rightarrow$] increase in tropospheric background ozone cite WMO?
   \end{itemize}
@@
 All stations with available long term observations are located on higher ground than Svanhovd which will result in higher surface ozone concentrations in comparison as ozone abundance, in general, decreases with decreasing altitude in the absence of local precursors.
-{\bf TODO: Citation see references in \citet{AB:Klingberg2009}!} {\bf TODO: The following has to be confirmed and elaborated on!
-Ane?} The vegetation at all sites is similar and consists mainly of pine forests, birch, and heath.
+{\bf TODO: Citation see references in \citet{AB:Klingberg2009}!} The periphery vegetation at all sites is similar and consists mainly of pine forests, birch, and heath.
+However, at close vicinity it varies substantially.
+Ozone measurements at Pallas are taken on top of mount Sammaltunturi $100\,\unit{m}$ above tree line (pine forest) and mainly consists of low vascular plants, mosses, and lichen \citep{BER:Hatakka2003}.
+At Svanhovd, the ozone measurements are taken amidst a meadow surrounded by agriculturally managed herbaceous plants and grass (c.c.
+Fig.\ref{fig:svanhovd_research_station}).
 
@@
 Based on data from SeNorge.no and the agricultural rule of thumb that the daily average temperature has to be above $5\,\unit{^\circ C}$ for $5\,\unit{days}$ consecutively, we find $\left<G_\mathrm{begin}\right> = (150 \pm 14)\,\unit{days of the year}$ at Svanvik (Appendix Fig~\ref{fig:greening_season_change_Svanvik}), which translates into the end of May.
+This finding is in line with more extensive studies \citep[][e.g.,]{IJB:Karlsen2007,AFM:Linderholm}.
+{\bf TODO: Rewrite how growing season has been calculated.}
 

### 2020-07-03 VegOzone.tex
A few more lines for the abstract.

@@
 \begin{abstract}
+The past two decades saw a steady reduction in episodic high peak summer concentrations of ozone in Europe and North America thanks to the implementation of air quality regultions.
+At the same time, tropospheric background ozone levels have been increasing in general.
 The summer 2018 and summer 2019 did not differ significantly in the level of ozone.

### 2020-07-13 VegOzone.tex
Addition to abstract - not spell checked.

--- 2020-06-05 VegOzone.tex
@@
 \begin{abstract}
-  The summer 2018 and summer 2019 did not differ significantly in the level of ozone.
-On the other hand, the temperatures (precipitation and cloud cover/PAR?) differed quite a lot.
-We also observed ozone-induced visible injuries on clovers in the ozone garden at Svanhovd in 2018, but not in 2019.
+  The past two decades saw a steady reduction in episodic high peak concentrations of ozone over Europe and North America -- a sucessful result of air quality regulations.
+At the same time, tropospheric background ozone levels kept increasing in the northern hemisphere until recently.
+An increased occurance of persitant heatwaves, as one possible future scenario, generates favourable conditions for the formation of ambient ozone.
+Vegetation, highly specialized to sub-artic climate, is vulnerable to rapid environmental changes due to global warming and might become more susceptible to ozone induced damage.
+The summer of 2018 can, hence, be regarded as a good test case for future scenarios.
+In both 2018 and 2019, we have monitored ambient ozone concentrations at the Svanhovd research station in Northern Norway.
+Ambient ozone concentrations did not differ significantly between the two years.
+On the other hand, temperatures, precipitation, and irradiance differed significantly to some extend.
+At the same time, we have observed ozone-induced visible injuries on clovers in the ozone garden at Svanhovd in 2018, but not in 2019.
+
   This may be due to differences in uptake of ozone ($\mathrm{POD_y}$) between the two years.

### 2020-07-15 VegOzone.tex
Spell checked. Added more to abstracted. Elaborated on introduction.

--- 2020-06-05 VegOzone.tex
@@
 \begin{abstract}
-  The summer 2018 and summer 2019 did not differ significantly in the level of ozone.
-On the other hand, the temperatures (precipitation and cloud cover/PAR?) differed quite a lot.
-We also observed ozone-induced visible injuries on clovers in the ozone garden at Svanhovd in 2018, but not in 2019.
-This may be due to differences in uptake of ozone ($\mathrm{POD_y}$) between the two years.
+  As a result of successful air quality regulations, the past two decades saw a steady reduction in episodic high summer peak concentrations of ozone over Europe and North America.
+At the same time, tropospheric background ozone levels kept increasing in the northern hemisphere until recently.
+An increased occurrence of persistent heatwaves, as one possible future scenario, generates favorable conditions for the formation of ambient air ozone.
+Vegetation, highly specialized to sub-arctic climate, is vulnerable to rapid environmental changes inflicted by global warming and might become more susceptible to ozone in the future.
+Over large parts of Europe the summer 2018 had been extraordinarily hot and dry and caused large wildfires in northern Sweden in particular.
+It can be regarded as a test case for future scenarios.
+In both 2018 and 2019, we have monitored ambient air ozone concentrations at NIBIO Environment Centre Svanhovd in Northern Norway.
+Ozone concentrations did not differ significantly between the two years.
+While temperatures and global irradiance differed significantly, precipitation varied only to some extend.
+Coincidentally, we have observed ozone-induced visible injuries on clovers in the ozone garden at Svanhovd in 2018, but not in 2019.
+
+  \emph{This may be due to differences in uptake of ozone ($\mathrm{POD_y}$) between the two years.
 Since DO3SE can be run on birch/beech and conifers, we would like to investigate whether the critical levels for these tree species were crossed during the two summer seasons.
-Ozone concentrations for the two missing weeks in July 2018 have been estimated based on the measurements from the Swedish and Finnish ozone measurement stations in the region.
+Ozone concentrations for the two missing weeks in July 2018 have been estimated based on the measurements from the Swedish and Finnish ozone measurement stations in the region.}
 \end{abstract}
@@
 
-%{\bf OBS: The following is just a copy of the 2018 dry deposition paper. It has to be substantially changed to meet the no-copycat-criteria. TODO: Rewrite in context of this article.}
-
-%Ozone is an important trace gas for all lifeforms on Earth.
-Depending on the place of its occurrence it has either a positive or negative connotation.
-In the stratosphere, ozone absorbs most of the ultraviolet (UV)-light from the sun within the range of 100--315\,\unit{nm}, thus shielding the Earth's surface from the most harmful UV-radiation.
-In addition, ozone is a potent greenhouse gas in both, stratosphere and troposphere.
-With a radiative forcing of $0.40 \pm 0.20\,\unit{W\,m^{-2}}$, it is placed third, only surpassed by \chem{CO_2} and \chem{CH_4} \citep[Chapter 8]{IPCC2013}.\\
-%In the troposphere and in particular in ambient air, ozone is considered as a highly toxic pollutant.
-Since the industrial revolution, tropospheric background ozone concentrations have been increasing in the northern hemisphere \citep[Chapter 2]{IPCC2013}.
-In recent years, the number of episodes of peak concentrations has been, in general, decreasing in North America and Europe due to the implementation of air quality regulations \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
-At the same time, fast developing countries, like e.g., China or India, saw a significant increase in ozone related air pollution.
-Continuously high concentrations of ambient air ozone are hazardous to the whole ecosystem.
-It is estimated that ozone is cause to an increase in pre-mature deaths \citep{WHO2008}, an average global loss of yield in the four major crops (wheat, rice, maize, and soybean) of about $3-15\,\unit{\%}$ \citep{PJ:Ainsworth2017} as well as $7\,\unit{\%}$ loss in primary production in forestry \citep{GCB:Wittig2009,EP:Matyssek2012}.
-The impact of ozone damage on vegetation and agricultural plants may affect food security in the future especially in Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018} and might be an important additional feedback to climate change \citep{Nat:Sitch2007}.\\
-%Elevated ozone levels at a site may originate from both, the local production of ozone from its precursors, which are transported, and from advection of ozone itself.
-Long-range ozone transport occurs regularly and might be most important in regions that otherwise lack precursors.
-Tropospheric ozone is produced in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) or volatile organic substances (VOCs -- also known as hydrocarbons) in the presents of nitrogen oxides (\chem{NO_x}).
-A typical reaction mechanism for \chem{CO} is sketched in the following.
-In a sequence of rapid reactions a peroxyl radical \chem{HO_2^\bullet} is formed through an initial reaction of \chem{CO} with a hydroxyl radical \chem{^\bullet OH}.
-%Via a reaction between \chem{HO_2^\bullet} and \chem{NO}, \chem{NO_2} is formed which is then photolyzed.
-The resulting atomic oxygen reacts then with \chem{O_2} (and also under the presence of available co-reactants) to form an ozone molecule.
-
-%Since ozone is highly reactive, its global mean life-time in the troposphere is roughly $22$ days but ranges between a few days in the tropical boundary layer to up to $1$ year in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}.
-The abundance of tropospheric ozone therefore varies, e.g., with time of the day, season, altitude, location \citep{ACP:Schnell2015}, or weather conditions in general \citep{ACP:Otero2018}.
-Typical concentrations of surface ozone range from $10\,\unit{ppb}$ over the tropical Pacific to $100\,\unit{ppb}$ in the downwind areas of highly emitting sources \citep[Chapter 8]{IPCC2013}.
-This variability poses a challenge on both, trend analysis from observations as well as validation and inter-comparison of models.
-From the observational side, the number of long-term observations (started before the 1950s) is limited and restricted to mainly European sites.
-Most of these indicate a doubling of tropospheric ozone since the 1950s \citep[Chapter 2]{IPCC2013}.
+Tropospheric and in particular ambient air ozone is a highly toxic pollutant to all ecosystems.
+In continuously high concentration, ozone negatively affects life expectancy of humans \citep{WHO2008} and is cause to early senescence as well as programmed cell death in plants {\bf TODO citations}.
+On average, a global loss of yield in four major crops (wheat, rice, maize, and soybean) of $\approx 3-15\,\unit{\%}$ \citep{PJ:Ainsworth2017} and a loss in primary production in forestry of $7\,\unit{\%}$ \citep{GCB:Wittig2009,EP:Matyssek2012} has been attributed to ozone.
+Former can proof a vital threat to food security in fast developing countries, e.g., in East and South-East Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018}, where air pollution related to ozone has increased in the past years.
+At the same time, a successful implementation of air quality regulations reduced the number of episodes of peak concentrations especially in summertime over Europe and North America \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
+
+With a radiative forcing of $0.40 \pm 0.20\,\unit{W\,m^{-2}}$, ozone itself is a potent climate forcer, ranked third behind \chem{CO_2} and \chem{CH_4} \citep[Chapter 8]{IPCC2013}, but in combination with its damaging effects on global vegetation it might provide an important additional feedback to climate change \citep{Nat:Sitch2007}.
+
+Tropospheric ozone is formed in complex photochemical cycles in the presents of nitrogen oxides (\chem{NO_x}) involving precursor gases such as carbon monoxide (\chem{CO}) and hydrocarbons known as (biogenic) volatile organic substances (VOCs and BVOCs).
+In the absence of local pollution sources, episodes of high ozone concentrations can be attributed to long-range transport of precursors.
+Despite a relatively short tropospheric life-time of roughly $22\,\unit{days}$ (a few days in the tropical boundary layer to up to $1\,\unit{year}$ in the upper troposphere) \citep{JGR:Stevenson2005,ACP:Young2013}, ozone itself is subject to advection over long distances.
+In northern Fennoscandia episodes of especially high concentrations of ambient air ozone in 2003 and 2006 have been traced back to emissions related to forest fires in southern Europe and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}.
+Boreal wildfires emit ozone precursors such as \chem{CO} ($6-13\,\unit{\%}$ relative to \chem{CO_2^{tot}}) and \chem{VOCs} ($0.5-1.5\,\unit{\%}$ relative to \chem{CO_2^{tot}}) \citep{AE:Cofer1990}.
+
+Ozone displays an annual cycle with a distinct maximum typically in spring \citep{AE:Monks2000}.
+This spring peak is reached earlier in northern Fennoscandia than in more southern locations but is highly variable throughout the years \citep{AB:Klingberg2009, BER:Klingberg2019}.
+The cause for this might be a complex interplay between atmospheric dynamics and chemistry \citep{AE:Laurila1996,BER:Hatakka2003} as well as anthropogenic and biogenic activity \citep{AE:Rummukainen1996,AE:Simpson2002,QJRMS:Galbally2007,NGS:Schnell2009}.
+
+\emph{Contributor: Ane}\\
+As mentioned above, ozone acts as an oxidative stress to plants.
+Its main action is imposed through reactions occurring in the cell walls and cell membranes of mesophyll cells inside the leaves.
+Ozone enters the leaves through stomata, that are kept open to allow for gas exchange between the leaf air spaces and the exterior air.
+When the stomata are closed, ozone does not enter.
 
@@
 \item general about ozone $\rightarrow$ focus on surface ozone and in particular precursors with respect to the forest fires (\chem{CO}) and heat stress (\chem{VOCs}).
-\item ozone induced damage in vegetation \emph{Contributor: Ane - maybe also ask Aud?}\\
-  Ozone acts as an oxidative stress to plants.
-Its main action is imposed through reactions occurring in the cell walls and cell membranes of mesophyll cells inside the leaves.
-Ozone enters the leaves through stomata, that are kept open to allow for gas exchange between the leaf air spaces and the exterior air.
-When the stomata are closed, ozone does not enter.
+\item ozone induced damage in vegetation 
 \item[$\rightarrow$] Some info on stomatal conductance and its empirical calculation:
@@
 
-Following a report by the Norwegian Meteorological institute \citep{MetNOR2019}, the summer of 2018 was the warmest and driest ever recorded in eastern, western, and southern Norway.
+According to a report by the Norwegian Meteorological institute \citep{MetNOR2019}, the summer of 2018 was the warmest and driest ever recorded in eastern, western, and southern Norway.
 In the north (including Finnmark), it was amongst the warmest on record - favorable conditions for boundary layer ozone formation.

### 2020-07-16 VegOzone.tex
Spell checked. Completed Introduction draft. Update abstract.

@@
 Over large parts of Europe the summer 2018 had been extraordinarily hot and dry and caused large wildfires in northern Sweden in particular.
-It can be regarded as a test case for future scenarios.
+This can be regarded as a test case for a future scenario.
 In both 2018 and 2019, we have monitored ambient air ozone concentrations at NIBIO Environment Centre Svanhovd in Northern Norway.
-Ozone concentrations did not differ significantly between the two years.
-While temperatures and global irradiance differed significantly, precipitation varied only to some extend.
+Over all, ozone concentrations did not differ significantly between the two years.
+While temperatures and global irradiance diverged significantly from multi annual mean, precipitation varied only to some extend.
 Coincidentally, we have observed ozone-induced visible injuries on clovers in the ozone garden at Svanhovd in 2018, but not in 2019.
-
-  \emph{This may be due to differences in uptake of ozone ($\mathrm{POD_y}$) between the two years.
-Since DO3SE can be run on birch/beech and conifers, we would like to investigate whether the critical levels for these tree species were crossed during the two summer seasons.
-Ozone concentrations for the two missing weeks in July 2018 have been estimated based on the measurements from the Swedish and Finnish ozone measurement stations in the region.}
+We investigate the difference in uptake of ozone using the DO3SE model, with respect to the typical vegetation (beech/birch and conifers) at the location.
+We assess whether critical levels for these species were breached during the summer seasons.
+Due to data acquisition problem, ozone concentrations for two weeks in July 2018 were missing from our record and have been reconstructed based on the measurements from the Swedish and Finnish ozone measurement stations in the region.
 \end{abstract}
@@
 In continuously high concentration, ozone negatively affects life expectancy of humans \citep{WHO2008} and is cause to early senescence as well as programmed cell death in plants {\bf TODO citations}.
-On average, a global loss of yield in four major crops (wheat, rice, maize, and soybean) of $\approx 3-15\,\unit{\%}$ \citep{PJ:Ainsworth2017} and a loss in primary production in forestry of $7\,\unit{\%}$ \citep{GCB:Wittig2009,EP:Matyssek2012} has been attributed to ozone.
-Former can proof a vital threat to food security in fast developing countries, e.g., in East and South-East Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018}, where air pollution related to ozone has increased in the past years.
+On average, a global loss of yield in four major crops (wheat, rice, maize, and soybean) of $3-15\,\unit{\%}$ \citep{PJ:Ainsworth2017} and a loss in primary production in forestry of $7\,\unit{\%}$ \citep{GCB:Wittig2009,EP:Matyssek2012} has been attributed to ozone.
+Former can proof a vital threat to food security in fast developing countries, e.g., in East and South-East Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018}, where ozone related to air pollution has increased in the past years.
 At the same time, a successful implementation of air quality regulations reduced the number of episodes of peak concentrations especially in summertime over Europe and North America \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
 
-With a radiative forcing of $0.40 \pm 0.20\,\unit{W\,m^{-2}}$, ozone itself is a potent climate forcer, ranked third behind \chem{CO_2} and \chem{CH_4} \citep[Chapter 8]{IPCC2013}, but in combination with its damaging effects on global vegetation it might provide an important additional feedback to climate change \citep{Nat:Sitch2007}.
-
-Tropospheric ozone is formed in complex photochemical cycles in the presents of nitrogen oxides (\chem{NO_x}) involving precursor gases such as carbon monoxide (\chem{CO}) and hydrocarbons known as (biogenic) volatile organic substances (VOCs and BVOCs).
-In the absence of local pollution sources, episodes of high ozone concentrations can be attributed to long-range transport of precursors.
-Despite a relatively short tropospheric life-time of roughly $22\,\unit{days}$ (a few days in the tropical boundary layer to up to $1\,\unit{year}$ in the upper troposphere) \citep{JGR:Stevenson2005,ACP:Young2013}, ozone itself is subject to advection over long distances.
-In northern Fennoscandia episodes of especially high concentrations of ambient air ozone in 2003 and 2006 have been traced back to emissions related to forest fires in southern Europe and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}.
-Boreal wildfires emit ozone precursors such as \chem{CO} ($6-13\,\unit{\%}$ relative to \chem{CO_2^{tot}}) and \chem{VOCs} ($0.5-1.5\,\unit{\%}$ relative to \chem{CO_2^{tot}}) \citep{AE:Cofer1990}.
-
-Ozone displays an annual cycle with a distinct maximum typically in spring \citep{AE:Monks2000}.
+With a radiative forcing of $0.40 \pm 0.20\,\unit{W\,m^{-2}}$, ozone in itself is a potent climate forcer, ranked third behind \chem{CO_2} and \chem{CH_4} \citep[Chapter 8]{IPCC2013}.
+Its damaging effects on global vegetation might provide an important additional feedback to climate change \citep{Nat:Sitch2007}.
+
+In the presents of nitrogen oxides (\chem{NO_x}), tropospheric ozone is formed in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) and hydrocarbons known as volatile organic substances (VOCs).
+In the absence of local pollution sources, episodes of high ozone concentrations are often attributed to long-range transport of precursors.
+In northern Fennoscandia episodes in 2003 and 2006 have been traced back to emissions related to forest fires in southern Europe and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}.
+But also living vegetation emits hydrocarbons in form of terpenes and monoterpenes, so called biogenic volatile substances (BVOCs).
+Those emissions are associated with, e.g., thermal stress, defense against herbivores, or attraction of pollinators \citep{TPS:Penuelas2003}.
+
+Despite a relatively short tropospheric life-time of roughly $22\,\unit{days}$ ranging from a few days in the tropical boundary layer to up to $1\,\unit{year}$ in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}, ozone itself is subject to advection over long distances.
+In the annual cycle a distinct maximum typically occurs in spring followed by lower concentrations throughout the summer probably due to a higher uptake on land vegetation than on bare ground or snow and ice \citep{AE:Monks2000}.
 This spring peak is reached earlier in northern Fennoscandia than in more southern locations but is highly variable throughout the years \citep{AB:Klingberg2009, BER:Klingberg2019}.
-The cause for this might be a complex interplay between atmospheric dynamics and chemistry \citep{AE:Laurila1996,BER:Hatakka2003} as well as anthropogenic and biogenic activity \citep{AE:Rummukainen1996,AE:Simpson2002,QJRMS:Galbally2007,NGS:Schnell2009}.
-
-\emph{Contributor: Ane}\\
-As mentioned above, ozone acts as an oxidative stress to plants.
+A cause for this is a complex interplay between atmospheric dynamics and chemistry \citep{AE:Laurila1996,BER:Hatakka2003} as well as anthropogenic and biogenic activity \citep{AE:Rummukainen1996,AE:Simpson2002,QJRMS:Galbally2007,NGS:Schnell2009}.
+\citet{ACP:Andersson2017} showed in a modeling study focused on Fennoscandia that the variability in winter can be attributed mainly to changes in atmospheric background (transport) of ozone, while summertime abundance is mostly affected by emissions of precursors in the rest of Europe.
+Meteorology plays an important role throughout the year.
+
+At the same time, the start of growing season in northern Fennoscandia is shifting to earlier days of the year in response to the general warming trend \citep[e.g.,]{GCB:Menzel2006,RS:Hogda2013,IJB:Karlsen2007} and thereby into close vicinity of the ozone spring peak.
+This may promote a higher susceptibility of deciduous species to ozone.
+
+Since the industrial revolution, tropospheric background concentrations of ozone in the northern hemisphere have doubled and kept increasing until recently.
+But long-term observations started before 1950 are scarce \citep[Chapter 2]{IPCC2013}.
+Recent trend analyses indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric ozone started to level-off or even decline \citep{ACP:Wespes2018,ESA:Gaudel2018}.\\
+
+\emph{Contributor: Ane} {\bf TODO: More detailed ozone damage description?}\\
+Ozone acts as an oxidative stress to plants.
 Its main action is imposed through reactions occurring in the cell walls and cell membranes of mesophyll cells inside the leaves.
@@
 
-{\bf TODO:
-\begin{itemize}
-\item general about ozone $\rightarrow$ focus on surface ozone and in particular precursors with respect to the forest fires (\chem{CO}) and heat stress (\chem{VOCs}).
-\item ozone induced damage in vegetation 
-\item[$\rightarrow$] Some info on stomatal conductance and its empirical calculation:
+The opening of stomata, described through their conductance to gases $g_\mathrm{sto}$, is governed by different environmental factors and is described by a multiplicative ansatz \citep{PTRS:Jarvis1976,ICP:MappingManual2017}:
 \begin{equation}
@@
 %
-The factors herein are normalized and vary within the range $0-1$.
-They account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_T$), water vapor pressure deficit ($f_D$), and soil water content ($f_\text{SW}$).
-All factors differ with plant functional type $k$.
-\item metrics of ozone
-  \begin{itemize}
-  \item[$\rightarrow$] Concentration based
+The factors herein are normalized and vary within a range $0-1$.
+They are empirically determined and account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_T$), water vapor pressure deficit ($f_D$), and soil water content ($f_\text{SW}$).
+All factors differ with plant functional type (PFT) $k$.
+Stomata are typically considered closed below a light intensity of $50\,\unit{W\,m^{-2}}$.
+
+To measure the risk of damage inflicted to vegetation through ozone, many different metrics have been defined.
+Two main types can be differentiated, concentration-based (e.g., $\mathrm{AOT_{40}}$) and flux-based (e.g.,$\mathrm{POD_{0.1}}$) metrics.
+Due to their simplicity, $\mathrm{AOT_{x}}$ are still widely used, despite the actual integrated flux of ozone entering through the stomata is more relevant for an estimate of the actual toxic dose.
+The Accumulated Ozone Exposure over a threshold concentration of $x\,\unit{ppb}$ is defined as:
 \begin{equation}
   \mathrm{AOT_x} = \int{\mathrm{max(\chem{[O_3]}-x)}\cdot \mathrm{dt}},
+  \label{eq:aot}
 \end{equation}
-    with the ambient ozone concentration $\chem{[O_3]}$ and the concentration threshold $\mathrm{x}$. A typical value for $\mathrm{x}=40\,\unit{ppb}$.
-  \item[$\rightarrow$] Flux based
+with the ambient ozone concentration $\chem{[O_3]}$ and the concentration threshold $\mathrm{x}$. The integration time varies, but is usually defined from $8\,\unit{am}-8\,\unit{pm}$.
+The Phototoxic Ozone Dose (POD) over a threshold y is defined as:
 \begin{equation}
   \mathrm{POD_y} = \int{\mathrm{max(\Phi_\chem{[O_3]}-y)}\cdot \mathrm{dt}},
+  \label{eq:pod}
 \end{equation}
 with the ambient ozone flux through the stomata $\Phi_\chem{[O_3]} = g_\mathrm{sto} \cdot \chem{[O_3]}$ and the flux threshold $\mathrm{y}$.
-  \end{itemize}
-\item present subarctic climate and conditions
-  \begin{itemize}
-  \item[$\rightarrow$] midnight sun and 23\,\unit{h} open stomata (biological details will be published in a component paper)
-  \item[$\rightarrow$] ozone climatology and spring peak \citep{AE:Monks2000, AB:Klingberg2009, AE:Rummukainen1996} {\bf TODO: Find newer articles.}
-  \end{itemize}
-\end{itemize}
-}
+The stomatal conductance is either measured directly or computed in accordance to Eq.~(\ref{eq:stomatal}).\\
 
@@
 In Sweden, about 500 fires had been reported (five times more than in a usual summer) and an estimated total of $25000\,\unit{hectare}$ burned down in the course of the most extensive fires in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) in younger history \citep{SOU2019}.
+Boreal wildfires emit in addition to \chem{CO_2} also \chem{CO} ($\chem{[CO]}/\chem{[CO_2]} \propto 6-13\,\unit{\%}$) and \chem{VOCs} ($\chem{[VOCs]}/\chem{[CO_2]} \propto 0.5-1.5\,\unit{\%}$) \citep{AE:Cofer1990}.
 
@@
 
-{\bf TODO: Future climate and conditions
-  \begin{itemize}
-  \item[$\rightarrow$] increase in heatwaves and dry periods
-  \item[$\rightarrow$] shift in start of growing season (into vicinity of ozone spring peak) \citep{GCB:Menzel2006,RS:Hogda2013,IJB:Karlsen2007}
-  \item[$\rightarrow$] increase in tropospheric background ozone cite WMO?
-  \end{itemize}
-}
-
-In terms of AOT40, the threshold for protection of vegetation of $3000\,\unit{ppb}$ given by the EU directive has been well exceeded at many ozone monitoring sites in Southern Norway in the 2018 growing season \citep{NILU2019}.
+In terms of AOT40, the threshold for protection of vegetation of $3000\,\unit{ppb}$ (integrated over the growing season) given by the EU directive has been well exceeded at many ozone monitoring sites in Southern Norway in the 2018 growing season \citep{NILU2019}.
 {\bf TODO: come back to this later on in the discussions(?) - it has not been breached at Svanvik in northern Fennoscandia.
@@
 This coincides with the period of the most extensive forest fires in central Swedish (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) which occurred from July 12--29 and destroyed about $18000\,\unit{hectare}$ of forest \citep{SOU2019}.
-{\bf TODO: Do we need a corresponding \chem{CO_2} emission here?} However, due to the above mentioned data acquisition problems, we missed most of the corresponding ozone data for this event and most likely also the peak ozone concentration.
+However, due to the above mentioned data acquisition problems, we missed most of the corresponding ozone data for this event and most likely also the peak ozone concentration.
 In contrast, ozone concentration only rose 18 times above the threshold of $40\,\unit{ppb}$ during the summer of 2019.
@@
 This finding is in line with more extensive studies \citep[][e.g.,]{IJB:Karlsen2007,AFM:Linderholm}.
-{\bf TODO: Rewrite how growing season has been calculated.}
 
@@
 Amongst others, an assimilation of ozone profiles from satellite retrieval and an upgraded ozone chemistry scheme have lead to the enhanced performance of latter.
-The higher spatial resolution of the regional air quality models used in the CAMS Regional Ensemble compared to the global reanalyses does capture more localized depletion and peak episodes in \chem{[O_3]} better and enhances the capability of the models with respect to daily and seasonal cycles in ozone {\bf TODO: Any citable source for this assumption?}.
+The higher spatial resolution of the regional air quality models used in the CAMS Regional Ensemble compared to the global reanalyses does capture more localized depletion and peak episodes in \chem{[O_3]} better and enhances the capability of the models with respect to daily and seasonal cycles in ozone.
 Latter is, of course, also affected by upgrades in the OpenIFS used as meteorological forcing.
 However, we can assume that the main driver for the strikingly different results are the assimilated observational ozone data.
-Satellite retrievals resolve \chem{[O_3]} at the surface rather poorly and hence do not constrain the global models well enough {\bf TODO: Any citable source for this?}.
+Satellite retrievals resolve \chem{[O_3]} at the surface rather poorly and hence do not constrain the global models well enough \citet{ACP:Andersson2017}.
 

### 2020-07-20 VegOzone.tex
Revision of introduction.

@@
 Tropospheric and in particular ambient air ozone is a highly toxic pollutant to all ecosystems.
-In continuously high concentration, ozone negatively affects life expectancy of humans \citep{WHO2008} and is cause to early senescence as well as programmed cell death in plants {\bf TODO citations}.
+In continuously high concentration, ozone negatively affects life expectancy of humans \citep{WHO2008} and is cause to early senescence as well as programmed cell death in plants \citep{PCE:Kangasjarvi2005}.
 On average, a global loss of yield in four major crops (wheat, rice, maize, and soybean) of $3-15\,\unit{\%}$ \citep{PJ:Ainsworth2017} and a loss in primary production in forestry of $7\,\unit{\%}$ \citep{GCB:Wittig2009,EP:Matyssek2012} has been attributed to ozone.
@@
 In the absence of local pollution sources, episodes of high ozone concentrations are often attributed to long-range transport of precursors.
-In northern Fennoscandia episodes in 2003 and 2006 have been traced back to emissions related to forest fires in southern Europe and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}.
-But also living vegetation emits hydrocarbons in form of terpenes and monoterpenes, so called biogenic volatile substances (BVOCs).
+In northern Fennoscandia, episodes in 2003 and 2006 have been traced back to emissions related to forest fires in southern and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}.
+Hydrocarbons are also emitted by living vegetation in form of terpenes and monoterpenes, so called biogenic volatile substances (BVOCs).
 Those emissions are associated with, e.g., thermal stress, defense against herbivores, or attraction of pollinators \citep{TPS:Penuelas2003}.
@@
 
-The opening of stomata, described through their conductance to gases $g_\mathrm{sto}$, is governed by different environmental factors and is described by a multiplicative ansatz \citep{PTRS:Jarvis1976,ICP:MappingManual2017}:
+The opening of stomata, typically quantified through a conductance to water vapor $g_\mathrm{sto}$, is governed by different environmental factors and is described by a multiplicative ansatz \citep{PTRS:Jarvis1976,ICP:MappingManual2017}:
 \begin{equation}
@@
 Two main types can be differentiated, concentration-based (e.g., $\mathrm{AOT_{40}}$) and flux-based (e.g.,$\mathrm{POD_{0.1}}$) metrics.
-Due to their simplicity, $\mathrm{AOT_{x}}$ are still widely used, despite the actual integrated flux of ozone entering through the stomata is more relevant for an estimate of the actual toxic dose.
-The Accumulated Ozone Exposure over a threshold concentration of $x\,\unit{ppb}$ is defined as:
+Due to their simplicity, $\mathrm{AOT_{x}}$ are still widely used, despite the actual integrated flux of ozone entering through the stomata is more relevant for an estimate of the actual toxic dose \citep{ICP:MappingManual2017}.
+The Accumulated Ozone Exposure over a threshold concentration of $\mathrm{x}$ in units of $\unit{ppb}$ is defined as:
 \begin{equation}
@@
 \end{equation}
-with the ambient ozone concentration $\chem{[O_3]}$ and the concentration threshold $\mathrm{x}$. The integration time varies, but is usually defined from $8\,\unit{am}-8\,\unit{pm}$.
+with the ambient ozone concentration $\chem{[O_3]}$. The integration time varies, but is usually defined from $8\,\unit{am}-8\,\unit{pm}$.
 The Phototoxic Ozone Dose (POD) over a threshold y is defined as:
@@
 Heat and drought stress on vegetation was exceptional not only in large parts of Fennoscandia but also in much of Europe, where the influence of the high pressure system extended even over a period of five months (April/May, July--September).
-These conditions gave rise to massive forest fires in different parts of Europe and thus an increase in ozone precursors (\chem{CO}, \chem{VOCs}).
+These conditions gave rise to massive forest fires in different parts of Europe and thus an increase in ozone precursors.
+Boreal wildfires emit in addition to \chem{CO_2} also \chem{CO} ($\chem{[CO]}/\chem{[CO_2]} \propto 6-13\,\unit{\%}$) and \chem{VOCs} ($\chem{[VOCs]}/\chem{[CO_2]} \propto 0.5-1.5\,\unit{\%}$) \citep{AE:Cofer1990}.
 
@@
 In Sweden, about 500 fires had been reported (five times more than in a usual summer) and an estimated total of $25000\,\unit{hectare}$ burned down in the course of the most extensive fires in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) in younger history \citep{SOU2019}.
-Boreal wildfires emit in addition to \chem{CO_2} also \chem{CO} ($\chem{[CO]}/\chem{[CO_2]} \propto 6-13\,\unit{\%}$) and \chem{VOCs} ($\chem{[VOCs]}/\chem{[CO_2]} \propto 0.5-1.5\,\unit{\%}$) \citep{AE:Cofer1990}.
 
@@
 
-In terms of AOT40, the threshold for protection of vegetation of $3000\,\unit{ppb}$ (integrated over the growing season) given by the EU directive has been well exceeded at many ozone monitoring sites in Southern Norway in the 2018 growing season \citep{NILU2019}.
+In terms of $\mathrm{AOT_{40}}$, the threshold for protection of vegetation of $3000\,\unit{ppb}$ (integrated over the growing season) given by the EU directive has been well exceeded at many ozone monitoring sites in Southern Norway in the 2018 growing season \citep{NILU2019}.
 {\bf TODO: come back to this later on in the discussions(?) - it has not been breached at Svanvik in northern Fennoscandia.

### 2020-07-28 VegOzone.tex
Update to Copernicus latex package 6.0.
Added acknowledgment and author contributions.

@@
 
-%% Please use the following documentclass and journal abbreviations for discussion papers and final revised papers.
-
-%% 2-column papers and discussion papers
+%% Please use the following documentclass and journal abbreviations for preprints and final revised papers.
+
+%% 2-column papers and preprints
 \documentclass[bg, manuscript]{copernicus}
 
-%% Journal abbreviations (please use the same for discussion papers and final revised papers)
+%% Journal abbreviations (please use the same for preprints and final revised papers)
 
@@
 % Hydrology and Earth System Sciences (hess)
+% Journal of Bone and Joint Infection (jbji)
 % Journal of Micropalaeontology (jm)
@@
 %\usepackage{rotating}
+
 \graphicspath{{pictures/}}
@@
 \Author[2]{Aud Else}{Berglen Eriksen}
-\Author[1]{Bj{\o}rg}{Rognerud}
+%\Author[1]{Bj{\o}rg}{Rognerud}
 \Author[1]{Frode}{Stordal}
@@
 \affil[3]{Department of Environment and Geography, University of York, UK}
+
 %% The [] brackets identify the author with the corresponding affiliation. 1, 2, 3, etc. should be inserted.
@@
 
-\authorcontribution{TEXT} %% this section is mandatory
-
-\competinginterests{TEXT} %% this section is mandatory even if you declare that no competing interests are present
-
-\disclaimer{TEXT} %% optional section
+\authorcontribution{SF has written the manuscripted, collected and processed ozone and environmental data, and performed all statistical analyses.
+AVV has performed on-site observation of vegetation demage induced by ozone and contributed to theory of ozone interaction with vegetation within the introduction.
+LE...
+CO...
+ALEB...
+FS...
+TB...
+All authors contributed to discussion and conclusion of the results.} %% this section is mandatory
+
+\competinginterests{The authors declare that they have no conflict of interest.} %% this section is mandatory even if you declare that no competing interests are present
+
+%\disclaimer{TEXT} %% optional section
 
 \begin{acknowledgements}
-TEXT
+  Bj{\o}rg Rognerud for processing SeNorge.no data with respect for begin of growing season.
+  Tore Flatlandsmo Berglen for hourly pressure data from Svanhovd.
 \end{acknowledgements}
@@
 %% DOI = {10.5194/xyz}
+
+%% LITERATURE CITATIONS
+%%
+%% command                        & example result
+%% \citet{jones90}|               & Jones et al. (1990)
+%% \citep{jones90}|               & (Jones et al., 1990)
+%% \citep{jones90,jones93}|       & (Jones et al., 1990, 1993)
+%% \citep[p.~32]{jones90}|        & (Jones et al., 1990, p.~32)
+%% \citep[e.g.,][]{jones90}|      & (e.g., Jones et al., 1990)
+%% \citep[e.g.,][p.~32]{jones90}| & (e.g., Jones et al., 1990, p.~32)
+%% \citeauthor{jones90}|          & Jones et al.
+%% \citeyear{jones90}|            & 1990
 

### 2020-07-28 VegOzone.tex
Edit intro.

@@
 \begin{abstract}
-  As a result of successful air quality regulations, the past two decades saw a steady reduction in episodic high summer peak concentrations of ozone over Europe and North America.
-At the same time, tropospheric background ozone levels kept increasing in the northern hemisphere until recently.
   An increased occurrence of persistent heatwaves, as one possible future scenario, generates favorable conditions for the formation of ambient air ozone.
-Vegetation, highly specialized to sub-arctic climate, is vulnerable to rapid environmental changes inflicted by global warming and might become more susceptible to ozone in the future.
+Vegetation highly specialized to sub-arctic climate is vulnerable to rapid environmental changes inflicted by global warming and might become more susceptible to ozone in the future.
 Over large parts of Europe the summer 2018 had been extraordinarily hot and dry and caused large wildfires in northern Sweden in particular.
-This can be regarded as a test case for a future scenario.
-In both 2018 and 2019, we have monitored ambient air ozone concentrations at NIBIO Environment Centre Svanhovd in Northern Norway.
+This can be regarded as a test case for such a future scenario.
+In both 2018 and 2019, we have monitored ambient air ozone concentrations at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd in Northern Norway.
+Due to data acquisition problems, ozone concentrations for two weeks in July 2018 were missing from our record.
+We present a reconstruction based on probability density function with respect to the Swedish and Finnish atmospheric monitoring sites in the region.
 Over all, ozone concentrations did not differ significantly between the two years.
@@
 We assess whether critical levels for these species were breached during the summer seasons.
-Due to data acquisition problem, ozone concentrations for two weeks in July 2018 were missing from our record and have been reconstructed based on the measurements from the Swedish and Finnish ozone measurement stations in the region.
+{\bf TODO: Summary of results.}
 \end{abstract}
@@
 
-With a radiative forcing of $0.40 \pm 0.20\,\unit{W\,m^{-2}}$, ozone in itself is a potent climate forcer, ranked third behind \chem{CO_2} and \chem{CH_4} \citep[Chapter 8]{IPCC2013}.
+With a radiative forcing of $0.40 \pm 0.20\,\unit{W\,m^{-2}}$, ozone is a potent climate forcer, ranked third behind \chem{CO_2} and \chem{CH_4} \citep[Chapter 8]{IPCC2013}.
 Its damaging effects on global vegetation might provide an important additional feedback to climate change \citep{Nat:Sitch2007}.
@@
 
-Despite a relatively short tropospheric life-time of roughly $22\,\unit{days}$ ranging from a few days in the tropical boundary layer to up to $1\,\unit{year}$ in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}, ozone itself is subject to advection over long distances.
-In the annual cycle a distinct maximum typically occurs in spring followed by lower concentrations throughout the summer probably due to a higher uptake on land vegetation than on bare ground or snow and ice \citep{AE:Monks2000}.
+Despite a relatively short tropospheric life-time of roughly $22\,\unit{days}$, ranging from a few days in the tropical boundary layer to up to $1\,\unit{year}$ in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}, ozone is subject to advection over long distances.
+In the annual cycle, a distinct maximum typically occurs in spring followed by lower concentrations throughout the summer.
+This is most likely due to stronger dry deposition on land vegetation than on bare ground or snow and ice \citep{AE:Monks2000}.
 This spring peak is reached earlier in northern Fennoscandia than in more southern locations but is highly variable throughout the years \citep{AB:Klingberg2009, BER:Klingberg2019}.
@@
 
-The opening of stomata, typically quantified through a conductance to water vapor $g_\mathrm{sto}$, is governed by different environmental factors and is described by a multiplicative ansatz \citep{PTRS:Jarvis1976,ICP:MappingManual2017}:
+The opening of stomata, quantified through their conductance to water vapor $g_\mathrm{sto}$, is governed by different environmental factors and can be described by a multiplicative ansatz \citep{PTRS:Jarvis1976,ICP:MappingManual2017}:
 \begin{equation}
@@
 Two main types can be differentiated, concentration-based (e.g., $\mathrm{AOT_{40}}$) and flux-based (e.g.,$\mathrm{POD_{0.1}}$) metrics.
-Due to their simplicity, $\mathrm{AOT_{x}}$ are still widely used, despite the actual integrated flux of ozone entering through the stomata is more relevant for an estimate of the actual toxic dose \citep{ICP:MappingManual2017}.
+Due to their simplicity, $\mathrm{AOT_{x}}$ are still widely used, despite the actual integrated flux of ozone entering through the stomata is more relevant for an estimation of the actual toxic dose \citep{ICP:MappingManual2017}.
 The Accumulated Ozone Exposure over a threshold concentration of $\mathrm{x}$ in units of $\unit{ppb}$ is defined as:
@@
 According to a report by the Norwegian Meteorological institute \citep{MetNOR2019}, the summer of 2018 was the warmest and driest ever recorded in eastern, western, and southern Norway.
-In the north (including Finnmark), it was amongst the warmest on record - favorable conditions for boundary layer ozone formation.
+In the north (including Finnmark), it was amongst the warmest on record -- favorable conditions for boundary layer ozone formation.
 
@@
 
-Within the scope of our project to assess probable effects of midnight sun conditions on ozone uptake by plants and resultant enhanced susceptibility to ozone inflicted damage, we conducted ambient air ozone concentration measurements at Svanhovd in the growing seasons of 2018/19 in collaboration with NILU.
-In 2018, we have observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden at NIBIO Environment Centre Svanhovd.
+Within the scope of our project to assess probable effects of midnight sun conditions on ozone uptake by plants and resultant enhanced susceptibility to ozone inflicted damage, we conducted ambient air ozone concentration measurements at Svanhovd in the growing seasons of 2018/19 in collaboration with the Norwegian Institute for Air Research (NILU).
+In 2018, we have observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd.
 In contrast, no such damage was found on the clovers in 2019, although it was found on the sensitive tobacco cultivar.

### 2020-09-15 VegOzone.tex
Conclusion and plots.
Remove Aud from list of authors.
Fixed location of Pallas which had been wrong.

@@
 \Author[3]{Connie}{O'Neill}
-\Author[2]{Aud Else}{Berglen Eriksen}
+%\Author[2]{Aud Else}{Berglen Eriksen}
 %\Author[1]{Bj{\o}rg}{Rognerud}
@@
 
-\copyrightstatement{TEXT}
+%\copyrightstatement{TEXT}
 
@@
     Karasjok   & NOR     & NO0055R & 69.467          & 25.217            & 333            & 1988 -- 1997\\
-    Pallas     & FIN     & FI0096G & 69.97           & 24.12             & 565            & 1995 -- present\\
+    Pallas     & FIN     & FI0096G & 67.97           & 24.12             & 565            & 1995 -- present\\
     Svanvik    & NOR     & NO0047R & 69.45           & 30.03             & 30             & 1986 -- 1996$^\dagger$\\

### 2020-10-09 VegOzone.tex
Fix gramma.

@@
 The ECMWF products do not reproduce observed surface ozone well.
-The MACC reanalysis (Fig.~\ref{fig:ozone_climatologies}a) is in general too low ($(9\pm 7)\,\unit{ppb}$) and displays a erroneous seasonal cycle.
+The MACC reanalysis (Fig.~\ref{fig:ozone_climatologies}a) is in general too low ($(9\pm 7)\,\unit{ppb}$) and displays an erroneous seasonal cycle.
 Ozone concentrations peak already in March followed by a second peak in July.
@@
 The ECMWF products do not reproduce observed surface ozone well.
-The MACC reanalysis is in general too low and displays a erroneous seasonal cycle.
+The MACC reanalysis is in general too low and displays an erroneous seasonal cycle.
 The CAMS reanalysis matches better but does not reproduce the seasonality in northern Fennoscandia.

### 2020-10-09 VegOzone.tex
Added appendix plot for BGS/EGS estiate for coniferous trees from MODIS data.

@@
 
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{modis_Psn}
+  \caption{Estimated begin (BGS) and end (EGS) of growing season for coniferous trees from MODIS Aqua/Terra net photosynthesis (PSN) product.
+A $1\times 1\,\unit{km}$ area around Svanhovd was selected.
+Daily averaged data for both 2018 and 2019 has been fitted with a quadratic polynomial function.
+The numerically computed root yields: BGS 122/106 day of year and EGS 261/274 day of year for 2018/2019, respectively.}
+  \label{fig:modis_Psn}
+\end{figure*}
+
 \clearpage
@@
 
-\authorcontribution{SF has written the manuscripted, collected and processed ozone and environmental data, and performed all statistical analyses.
-AVV has performed on-site observation of vegetation demage induced by ozone and contributed to theory of ozone interaction with vegetation within the introduction.
+\authorcontribution{SF has written the manuscript, collected and processed ozone and environmental data, and performed all statistical analyses.
+AVV has performed on-site observation of vegetation damage induced by ozone and contributed to theory of ozone interaction with vegetation within the introduction.
 LE...

### 2020-10-09 VegOzone.tex
Add note about DO3SE parameter list which shall go to supplement.

@@
 
-{\bf TODO: Some general words about DO3SE model?}
+{\bf TODO: Some general words about DO3SE model}
+
+\subsection{Model parameters}
+A comprehensive list of model parameters used in this study is found in Supplement~X.
 
@@
 
-{\bf TODO: Model results for natural and semi-natural vegetation (birch, pine, clover, tobacco(?))}
+{\bf TODO: Model results for natural and semi-natural vegetation (birch, pine, grassland)}
 \begin{itemize}

### 2020-11-13 VegOzone.tex
Update description of DO3SE model input and ozone reconstruction. Add parameter description and first pody results.

@@
 \label{sec:do3se}
-To investigate which of the environmental factors lead to the observed ozone damage on clover and tobacco in 2018 in contrast to 2019, when no such damage was found on the clovers but only on the sensitive tobacco cultivar, we model the ozone uptake using the DO3SE model.
-
-\subsection{Reconstruction of missing ozone values in 2018}
+To investigate which of the environmental factors lead to the observed ozone damage on clover and tobacco in 2018 in contrast to 2019, when no such damage was found on the clovers but only on the sensitive tobacco cultivar, we model the ozone uptake using the DO3SE model and give an estimate on biomass reduction due to ozone uptake.
+
+\subsection{Model description}
+
+{\bf TODO: Some general words about DO3SE model}
+
+\subsection{Input data and gapfilling methology}
 \label{subsec:ozone_reco}
-As we have shown, the available reanalysis products have their limitations.
-Although the quality of the Copernicus Regional Model Reanalysis is fairly good, we additionally deem it necessary to reconstruct the missing data in 2018 for the purpose of DO3SE modeling.
-In Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}, the procedure for a probable reconstruction is shown.
-First, we derive hourly climatologies for northern Fennoscandia \chem{\left<[O_3]\right>_{hourly}} and Svanvik \chem{\left<[O_3]\right>_{hourly}^{Svanvik}} (Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}a).
-Then we compute temporal correlations between Svanvik and all other stations with respect to a temporal shift (Appendix Fig.~\ref{fig:time_lag_correlation}) to estimate a probable time lag between the stations (e.g., $3\,\unit{h}$ for Esrange and Pallas).
-Because of the high correlation with Svanvik observations, we have chosen Pallas as reference station for the reconstruction.
-We correct the Fennoscandic climatology for the found time lag\chem{\left<[O_3]\right>_{hourly,\,t-3}} and use it to derive a mapping function:
+The DO3SE model requires hourly, continuous meteorological observations.
+In addition to variables covered in Section~\ref{sec:data}, $2\,\unit{m}$ wind $u_\text{2m}$ and vapor pressure deficit VPD are needed.
+VPD has been calculated from observed $T_\text{2m}$ and relative humidity
+\begin{equation}
+  \text{VPD} = P_s(T) \cdot \left(1-\frac{\text{relHum}}{100}\right),
+\end{equation}
+with satuation vapor pressure $P_s(T)$ in \unit{hPa} derived from Arden Buck equation~\citep{JAP:Buck1981, Buck2012}.
+
+The following gap filling methodology was devised for the meterological input data:
+\begin{itemize}
+\item Single hours of missing data were filled by taking the average of the hourly values coming before, and after, the missing value.
+\item Several consecutive hours of missing data (23 or less) were filled by taking the average of the corresponding hour the day before, and the day after; and repeating this for each missing hour of data.
+If data was unavailable from that hour of the previous day, then only the value from the day after was used and vice versa.
+Interpolated values were not used in calculating averages.
+\item Data gaps longer than $24\,\unit{h}$ were filled using weekly diurnal averages.
+i.e.
+an average was calculated using the corresponding hour throughout the week before and after.
+Interpolated values were not used in calculating averages.
+\end{itemize}
+
+Because the available reanalysis products for surface ozone have their limitations, we deem it necessary to reconstruct the missing data in 2018 for the purpose of DO3SE modeling by applying a Reynold's decomposition of the form:
+\begin{equation}
+  \chem{[O_3]} = \chem{\left<[O_3]\right>} + \chem{\Delta[O_3]},
+  \label{eq:reynolds_decomp}
+\end{equation}
+with climatology \chem{\left<[O_3]\right>} and annomalies \chem{\Delta[O_3]}.
+The exact procedure will be described in the following.
+We derive hourly climatologies at all operational stations for July and a combined northern Fennoscandic climatology based on both data from Esrange and Pallas (Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}a).
+We compute temporal correlations between Svanvik and the other stations in northern Fennoscandia to estimate a time lag (Appendix Fig.~\ref{fig:time_lag_correlation}.
+We find $3\,\unit{h}$ for Esrange and Pallas with respect to Svanvik.
+Due to the high correlation with Svanvik (Appendix~Fig.~\ref{fig:density_distribution}), we decide on Pallas as reference station for the reconstruction.
+We correct the Fennoscand climatology for the found time lag \chem{\left<[O_3]\right>_{hourly,\,t-3}} by shifting the respective time series back in time by $3\,\unit{h}$ and use it to derive a projection for Svanvik:
 \begin{equation}
@@
 \end{equation}
-Performing a time lag correction basically means to shift the respective time series back in time by $3\,\unit{h}$.
-As an intermediate step, hourly anomalies are computed
+As an intermediate step, we apply Eq.~(\ref{eq:reynolds_decomp}) to derive hourly anomalies for each station $i$
 \begin{equation}
-  \chem{\Delta[O_3]_{hourly}^{Pallas}} = \chem{[O_3]_{hourly}^{Pallas}}-\chem{\left<[O_3]\right>_{hourly}}.
+  \chem{\Delta[O_3]_{hourly}}^i = \chem{[O_3]_{hourly}}^i-\chem{\left<[O_3]\right>_{hourly}},
 \end{equation}
-In Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}b), hourly ozone concentration anomalies are shown for Pallas as well as for Esrange and Svanvik.
-We also correct ozone anomalies for Pallas for the time lag \chem{\Delta[O_3]_{hourly,\,t-3}^{Pallas}} and use the mapping function in Eq.~(\ref{eq:mapping}) to derive reconstructed anomalies for the missing values:
+with $i \in \{\text{Esrange, Pallas, Svanvik\}}$.
+In Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}b), these hourly ozone concentration anomalies are shown.
+We also correct the ozone anomalies for Pallas for the time lag \chem{\Delta[O_3]_{hourly,\,t-3}^{Pallas}} and use the mapping function in Eq.~(\ref{eq:mapping}) to derive reconstructed anomalies for the missing values at Svanvik:
 \begin{equation}
@@
 
-\subsection{Model description and input data}
-
-{\bf TODO: Some general words about DO3SE model}
-
 \subsection{Model parameters}
-A comprehensive list of model parameters used in this study is found in Supplement~X.
+A comprehensive list of model parameters used in this study is found in Supplement~Table~S1.
+Start ($A_\text{start}$) and end ($A_\text{end}$) of growing season for each species (Norway spruce, birch, perennial grassland) are estimated as follows.
+Evergreen needleleaf trees are active already at low temperatures \citep{TB:Kolari2007}.
+This leads to an increase in observable net photosynthesis.
+We use a MODIS (Aqua/Terra) retreived net photosynthesis product for a $1\times 1\,\unit{km}$ patch around Svanvik and calculate the roots from a quadratic fit through the data (Fig.~\ref{fig:modis_Psn}).
+We derive $A_\text{start}$ day of the year 122 for 2018 and 106 in 2019 and $A_\text{end}$ 261 and 274, respectively.
+This means the growing season in 2019 had been one month longer then in 2019.
+For birch tree $A_\text{start}$, we use the usual 5 consequative days above 5 degrees agricultural rule of thumb based on grided temperature data from seNorge.no.
+We find 129 and 130, respectively.
+For perennial grassland, we assume a dormancy of about $1\,\unit{month}$ after snowmelt.
+Data for first snow-free day is taken from the weather station at Øvre Neiden in Sør-Varanger.
+Grassland lags about $1\,\unit{month}$ behind birches in both years.
+As leafs of northern Fennoscadic birches are smaller than southern speciemen, we have measured their dimension for the purpose of this study.
+%\href{https://www.nina.no/archive/nina/PppBasePdf/oppdragsmelding/829.pdf​}{NINA report} (p.
+52)
+indicates an average tree height of $13.5\,\unit{m}$ in the area.
+To derive the leaf area index (LAI), we estimated land use for the $1\times 1\,\unit{km}$ patch around Svanhovd from areal pictures at kartverket.no: $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban.
+With these, we rescale the average maximum LAI from MODIS ($2.7$).
+We assumed LAI ratios between the species: $f_\text{con}=5/4\cdot f_\text{dec}$, $f_\text{grass}=7/8\cdot f_\text{dec}$.
+Another important factor is the soil texture. As Svanhovd is located in the bed of the Pasvik river, we assume a sandy loam texture.
 
@@
 {\bf TODO: Model results for natural and semi-natural vegetation (birch, pine, grassland)}
-\begin{itemize}
-\item climatological ozone impact
-\item 2018
-\item 2019
-\end{itemize}
+
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{DO3SE_results_pody_gsto_o3_Birch}
+  \caption{.}
+  \label{fig:pody_birch}
+\end{figure*}
+
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{DO3SE_results_pody_gsto_o3_Norway_spruce}
+  \caption{.}
+  \label{fig:pody_norway_spruce}
+\end{figure*}
+
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{DO3SE_results_pody_gsto_o3_Perennial_grass}
+  \caption{.}
+  \label{fig:pody_grass}
+\end{figure*}
+
+%\begin{table}[t]
+%\caption{TEXT}
+%\begin{tabular}{column = lcr}
+%\tophline
+%
+%\middlehline
+%
+%\bottomhline
+%\end{tabular}
+%\belowtable{} % Table Footnotes
+%\end{table}
 

### 2020-11-16 VegOzone.tex
Prove read and spell checked.

@@
 
-\subsection{Input data and gapfilling methology}
+\subsection{Input data and gap filling methodology}
 \label{subsec:ozone_reco}
@@
 \end{equation}
-with satuation vapor pressure $P_s(T)$ in \unit{hPa} derived from Arden Buck equation~\citep{JAP:Buck1981, Buck2012}.
-
-The following gap filling methodology was devised for the meterological input data:
+with saturation vapor pressure $P_s(T)$ in \unit{hPa} derived from Arden Buck equation~\citep{JAP:Buck1981, Buck2012}.
+
+The following gap filling methodology was devised for the meteorological input data:
 \begin{itemize}
@@
 
-Because the available reanalysis products for surface ozone have their limitations, we deem it necessary to reconstruct the missing data in 2018 for the purpose of DO3SE modeling by applying a Reynold's decomposition of the form:
+Because the available reanalysis products for surface ozone have their limitations, we deem it necessary to reconstruct the missing data in 2018 for the purpose of DO3SE modeling by applying a Reynolds's decomposition of the form:
 \begin{equation}
@@
 \end{equation}
-with climatology \chem{\left<[O_3]\right>} and annomalies \chem{\Delta[O_3]}.
+with climatology \chem{\left<[O_3]\right>} and anomalies \chem{\Delta[O_3]}.
 The exact procedure will be described in the following.
-We derive hourly climatologies at all operational stations for July and a combined northern Fennoscandic climatology based on both data from Esrange and Pallas (Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}a).
+We derive hourly climatologies at individual stations and for a combined Esrange/Pallas northern Fennoscandia for July (Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}a).
 We compute temporal correlations between Svanvik and the other stations in northern Fennoscandia to estimate a time lag (Appendix Fig.~\ref{fig:time_lag_correlation}.
@@
 Due to the high correlation with Svanvik (Appendix~Fig.~\ref{fig:density_distribution}), we decide on Pallas as reference station for the reconstruction.
-We correct the Fennoscand climatology for the found time lag \chem{\left<[O_3]\right>_{hourly,\,t-3}} by shifting the respective time series back in time by $3\,\unit{h}$ and use it to derive a projection for Svanvik:
+We correct the Fennoscandic climatology for the found time lag \chem{\left<[O_3]\right>_{hourly,\,t-3}} by shifting the respective time series back in time by $3\,\unit{h}$ and use it to derive a projection for Svanvik:
 \begin{equation}
@@
 \subsection{Model parameters}
-A comprehensive list of model parameters used in this study is found in Supplement~Table~S1.
-Start ($A_\text{start}$) and end ($A_\text{end}$) of growing season for each species (Norway spruce, birch, perennial grassland) are estimated as follows.
-Evergreen needleleaf trees are active already at low temperatures \citep{TB:Kolari2007}.
-This leads to an increase in observable net photosynthesis.
-We use a MODIS (Aqua/Terra) retreived net photosynthesis product for a $1\times 1\,\unit{km}$ patch around Svanvik and calculate the roots from a quadratic fit through the data (Fig.~\ref{fig:modis_Psn}).
-We derive $A_\text{start}$ day of the year 122 for 2018 and 106 in 2019 and $A_\text{end}$ 261 and 274, respectively.
-This means the growing season in 2019 had been one month longer then in 2019.
-For birch tree $A_\text{start}$, we use the usual 5 consequative days above 5 degrees agricultural rule of thumb based on grided temperature data from seNorge.no.
+A comprehensive list of model parameters used in this study can be found in Supplement~Table~S1.
+Start ($A_\text{start}$) and end ($A_\text{end}$) of growing season for each PFT (Norway spruce, birch, perennial grassland) are estimated as follows.
+Evergreen needle leaf trees are photosynthetic active already at low temperatures \citep{TB:Kolari2007}.
+This leads necessarily to an increase in observable net photosynthesis.
+Hence, we use a MODIS (Aqua/Terra) retrieved net photosynthesis product for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine onset and end of growing season.
+Net photosynthesis data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}).
+By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{day~of~year}$ for 2018 and day 106 in 2019.
+$A_\text{end}$ amounts to day 261 and 274, respectively, and shall be used for all vegetation alike.
+The resulting growing season for Norway spruce in 2019 was one month longer then in 2018.
+For estimating $A_\text{start}$ of birch tree, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Fig.~\ref{fig:greening_season_change_Svanvik}).
 We find 129 and 130, respectively.
-For perennial grassland, we assume a dormancy of about $1\,\unit{month}$ after snowmelt.
+For perennial grassland, we assume a latency period of about $1\,\unit{month}$ after snow melt.
 Data for first snow-free day is taken from the weather station at Øvre Neiden in Sør-Varanger.
-Grassland lags about $1\,\unit{month}$ behind birches in both years.
-As leafs of northern Fennoscadic birches are smaller than southern speciemen, we have measured their dimension for the purpose of this study.
-%\href{https://www.nina.no/archive/nina/PppBasePdf/oppdragsmelding/829.pdf​}{NINA report} (p.
-52)
-indicates an average tree height of $13.5\,\unit{m}$ in the area.
-To derive the leaf area index (LAI), we estimated land use for the $1\times 1\,\unit{km}$ patch around Svanhovd from areal pictures at kartverket.no: $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban.
-With these, we rescale the average maximum LAI from MODIS ($2.7$).
-We assumed LAI ratios between the species: $f_\text{con}=5/4\cdot f_\text{dec}$, $f_\text{grass}=7/8\cdot f_\text{dec}$.
-Another important factor is the soil texture. As Svanhovd is located in the bed of the Pasvik river, we assume a sandy loam texture.
+In both years, the first snow-free day incidents with the estimated date of bud burst.
+As leafs of birches in northern Fennoscandia {\bf (TODO: Ane, species name :-))} are smaller than southern specimen, we have measured their dimension for the purpose of this study and find $(3.0\pm 0.5)\,\unit{cm}$.
+\citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the arial.
+To localize leaf area index (LAI), we roughly characterize land use for the $1\times 1\,\unit{km}$ patch centered at Svanhovd from areal pictures at kartverket.no: $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban.
+We normalize LAI to deciduous trees based on standard DO3SE parameterization: $f_\text{con}=5/4\cdot f_\text{dec}$, $f_\text{grass}=7/8\cdot f_\text{dec}$.
+With these, we re-scale the average maximum LAI from MODIS ($2.7$).
+This results in 3.75, 3, and 2.625 for Norway spruce, birch, and grassland, respectively.
+
+As pointed out by \citet{ACP:Bueker2012}, a very important factor for deriving $\text{POD}_y$ is the soil texture or how much water the soil can hold.
+As Svanhovd is located in the bed of the Pasvik river, we assume a sandy loam texture at $60\,\unit{cm}$ depth.
 
@@
 
-\conclusions  %% \conclusions[modified heading if necessary]
+\conclusions[Discussion and conclusions]
 \label{sec:conc}
@@
 % Shall go to supplement!
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{ozone_fennoscandic_obs_aot40}
-  \caption{Integrated AOT40 for various sites in northern Fennoscandia over the course of 33 years.
+%\begin{figure*}[t]
+%  \includegraphics[width=12cm]{ozone_fennoscandic_obs_aot40}
+%  \caption{Integrated AOT40 for various sites in northern Fennoscandia over the course of 33 years.
 For simplicity, we integrated hourly ozone between $1\,\unit{am}-11\,\unit{pm}$ in any case, although a night time light intensity $> 50\,\unit{W\,m^{-2}}$ is only measured during midnight sun conditions.
@@
 (a) May 1 -- July 31 (b) April 1 -- September 30.}
-  \label{fig:fennoscandic_aot40}
-\end{figure*}
+%  \label{fig:fennoscandic_aot40}
+%\end{figure*}
 

### 2020-11-16 VegOzone.tex
First words about results.

@@
 In terms of $\mathrm{AOT_{40}}$, the threshold for protection of vegetation of $3000\,\unit{ppb}$ (integrated over the growing season) given by the EU directive has been well exceeded at many ozone monitoring sites in Southern Norway in the 2018 growing season \citep{NILU2019}.
-{\bf TODO: come back to this later on in the discussions(?) - it has not been breached at Svanvik in northern Fennoscandia.
-Later supplement figure \ref{fig:fennoscandic_aot40}}
+{\bf TODO: come back to this later on in the discussions(?) - it has not been breached at Svanvik in northern Fennoscandia.}
 
@@
 
-As pointed out by \citet{ACP:Bueker2012}, a very important factor for deriving $\text{POD}_y$ is the soil texture or how much water the soil can hold.
+As pointed out by \citet{ACP:Bueker2012}, a very important factor in deriving a proper $\text{POD}_y$ is the soil texture or how much water the soil can hold.
 As Svanhovd is located in the bed of the Pasvik river, we assume a sandy loam texture at $60\,\unit{cm}$ depth.
 
+The estimated total biomass reduction in \unit{\%} due to \chem{O_3} following \citet{ICP:MappingManual2017} is shown in Table~\ref{tab:biomass_reduction}.
+
 \subsection{POD and ozone impact}
 
-{\bf TODO: Model results for natural and semi-natural vegetation (birch, pine, grassland)}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{DO3SE_results_pody_gsto_o3_Birch}
-  \caption{.}
-  \label{fig:pody_birch}
-\end{figure*}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{DO3SE_results_pody_gsto_o3_Norway_spruce}
-  \caption{.}
-  \label{fig:pody_norway_spruce}
-\end{figure*}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{DO3SE_results_pody_gsto_o3_Perennial_grass}
-  \caption{.}
-  \label{fig:pody_grass}
-\end{figure*}
-
-%\begin{table}[t]
-%\caption{TEXT}
-%\begin{tabular}{column = lcr}
-%\tophline
-%
-%\middlehline
-%
-%\bottomhline
-%\end{tabular}
-%\belowtable{} % Table Footnotes
-%\end{table}
+We have modeled $\text{POD}_1$ for three natural and semi-natural vegetation PFTs, birch, Norway spruce, and perennial grassland.
+The results are shown comprehensively in Fig.~\ref{fig:pody} as a function of day of year (doy) for March--October.
+\chem{[O_3]} are plotted on the same axis and scales as $G_\text{sto}^\text{leaf}$ but in units of $\unit{ppb}$.
+Due to expected non linearities in stomatal conductance, the climatological year represents rather an optimal year then a true climatology in this case.
+To establish a true climatology in $\text{POD}_y$, it would therefore be necessary to run the DO3SE model for 10--20 years with true start and end of growing season for each PFT and the respective meteorological and ozone forcing.
+
+As expected $\text{POD}_y$ was higher in 2018 than in 2019 for all PFTs.
+
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{DO3SE_results_pody_gsto_o3}
+  \caption{$\text{POD}_y$ modeling results from DO3SE model for three different PFTs over day of year (doy), March--October.
+A flux threashold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) has been chosen.
+\chem{[O_3]} are plotted on the same axis and scales as $G_\text{sto}^\text{leaf}$ but in units of $\unit{ppb}$.
+Due to non linearities in stomatal conductance, the climatological year represents rather an optimal year then a true climatology.
+(a)--(c) Birch; (d)--(f) Norway spruce; (g)--(i) perennial grassland.
+From left to right: 2018, 2019, and climatological year.}
+  \label{fig:pody}
+\end{figure*}
+
+\begin{table}[t]
+  \caption{Estimated total biomass reduction in \unit{\%} due to \chem{O_3} following \citet{ICP:MappingManual2017}.}
+  \label{tab:biomass_reduction}
+\begin{tabular}{lccrl}
+\tophline
+Year & \multicolumn{4}{c}{PFT}\\
+& Birch & Norway spruce & \multicolumn{2}{c}{Perennial grassland}\\
+\middlehline
+2018 & 11.99 & 2.34 & 8.06 & (10.51)$^*$\\
+2019 & 10.07 & 2.26 & 6.43 & (7.90)$^*$\\
+clim. & 13.81 & 2.58 & 6.67 & (8.29)$^*$\\
+\bottomhline
+\end{tabular}
+\belowtable{$^*$ above ground bioamass reduction} % Table Footnotes
+\end{table}
 
@@
 Daily averaged data for both 2018 and 2019 has been fitted with a quadratic polynomial function.
-The numerically computed root yields: BGS 122/106 day of year and EGS 261/274 day of year for 2018/2019, respectively.}
+The numerically computed root yields: BGS 122/106 day of year (doy) and EGS 261/274 day of year for 2018/2019, respectively.}
   \label{fig:modis_Psn}

### 2020-11-23 VegOzone.tex
Update plots and write some more about results. Added first thoughts on conclusions and discussion.

@@
 The exact procedure will be described in the following.
-We derive hourly climatologies at individual stations and for a combined Esrange/Pallas northern Fennoscandia for July (Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}a).
+We derive hourly climatologies for individual stations and combine Esrange/Pallas to derive a climatology for northern Fennoscandia for July (Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}a).
 We compute temporal correlations between Svanvik and the other stations in northern Fennoscandia to estimate a time lag (Appendix Fig.~\ref{fig:time_lag_correlation}.
@@
 As leafs of birches in northern Fennoscandia {\bf (TODO: Ane, species name :-))} are smaller than southern specimen, we have measured their dimension for the purpose of this study and find $(3.0\pm 0.5)\,\unit{cm}$.
-\citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the arial.
+\citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area.
 To localize leaf area index (LAI), we roughly characterize land use for the $1\times 1\,\unit{km}$ patch centered at Svanhovd from areal pictures at kartverket.no: $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban.
 We normalize LAI to deciduous trees based on standard DO3SE parameterization: $f_\text{con}=5/4\cdot f_\text{dec}$, $f_\text{grass}=7/8\cdot f_\text{dec}$.
-With these, we re-scale the average maximum LAI from MODIS ($2.7$).
+With these, we re-scale the average maximum LAI from MODIS ($\left<f\right> = 2.7$).
 This results in 3.75, 3, and 2.625 for Norway spruce, birch, and grassland, respectively.
 
-As pointed out by \citet{ACP:Bueker2012}, a very important factor in deriving a proper $\text{POD}_y$ is the soil texture or how much water the soil can hold.
+As pointed out by \citet{ACP:Bueker2012}, a very important factor in deriving a proper $\mathrm{POD_y}$ is the soil texture or how much water the soil can hold.
 As Svanhovd is located in the bed of the Pasvik river, we assume a sandy loam texture at $60\,\unit{cm}$ depth.
 
-The estimated total biomass reduction in \unit{\%} due to \chem{O_3} following \citet{ICP:MappingManual2017} is shown in Table~\ref{tab:biomass_reduction}.
-
 \subsection{POD and ozone impact}
 
-We have modeled $\text{POD}_1$ for three natural and semi-natural vegetation PFTs, birch, Norway spruce, and perennial grassland.
+We have modeled $\mathrm{POD_y}$ for three natural/semi-natural vegetation PFTs, birch, Norway spruce, and perennial grassland.
 The results are shown comprehensively in Fig.~\ref{fig:pody} as a function of day of year (doy) for March--October.
+A flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) has been chosen.
 \chem{[O_3]} are plotted on the same axis and scales as $G_\text{sto}^\text{leaf}$ but in units of $\unit{ppb}$.
-Due to expected non linearities in stomatal conductance, the climatological year represents rather an optimal year then a true climatology in this case.
-To establish a true climatology in $\text{POD}_y$, it would therefore be necessary to run the DO3SE model for 10--20 years with true start and end of growing season for each PFT and the respective meteorological and ozone forcing.
-
-As expected $\text{POD}_y$ was higher in 2018 than in 2019 for all PFTs.
+
+As expected, $\mathrm{POD_y}$ is higher in 2018 than in 2019 for all PFTs.
+In the end of the 2018 season, Birch took up $13.1\,\unit{mmol\,m^{-2}\,PLA}$, while perennial grassland had an uptake of $4.5\,\unit{mmol\,m^{-2}\,PLA}$.
+The largest interannual variation is also found in Birch ($2.1\,\unit{mmol\,m^{-2}\,PLA}$) and perennial grassland ($2.6\,\unit{mmol\,m^{-2}\,PLA}$).
+Whereas, there is only a small difference for Norway spruce ($0.3\,\unit{mmol\,m^{-2}\,PLA}$) between the two years.
+
+Except for perennial grassland, model integrations with climatological forcings show higher $\mathrm{POD_y}$ than 2018.
+This can be explained by means of non linearity in modeled stomatal conductance (Eq.~(\ref{eq:stomatal})) in Eq.~(\ref{eq:pod}).
+The climatology in $\mathrm{POD_y}$ therefore represents rather an ideal year for growth then a true climatology.
+To establish a true climatology, it would be necessary to run the DO3SE model for 10--20 years with respective observed phenology as well as meteorological and ozone forcing.
+
+Based on this interpretation of climatology in $\mathrm{POD_y}$, we can learn more about potential ozone damage.
+The time derivative $\frac{\mathrm{\Delta POD_y}}{\Delta t}$ displays very different features for each PFT and meteorological forcing.
+Although Norway spruce shows a low interannual variation, $\frac{\mathrm{\Delta POD_y}}{\Delta t}$ (Fig.~\ref{fig:pody_rel}) indicates an extended period of higher than average ozone uptake in spring 2018 (steep slope), while ozone uptake in June/July was anomalously low in 2019 (shallow slope) but high in fall.
+This is most prominently visible in Fig.~\ref{fig:pody}h) for perennial grassland, where consecutive periods of low $G_\text{sto}^\text{leaf}$ manifest in three plateaus in $\mathrm{POD_y}$ throughout the season (1: $184-198\,\unit{doy}$, 2: $209-228\,\unit{doy}$, 3: $235-240\,\unit{doy}$).
+An episode with anomalous high ozone in mid June ($194\,\unit{doy}$) can be identified but does not amount to unusually high ozone uptake on an annual basis.
+
+The estimated biomass reduction in \unit{\%} due to \chem{O_3} following \citet{ICP:MappingManual2017} is shown in Table~\ref{tab:biomass_reduction}.
+{\bf TODO: Lisa?
+Add paragraph about biomass reduction to introduction?}
 
@@
   \includegraphics[width=12cm]{DO3SE_results_pody_gsto_o3}
-  \caption{$\text{POD}_y$ modeling results from DO3SE model for three different PFTs over day of year (doy), March--October.
-A flux threashold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) has been chosen.
+  \caption{$\mathrm{POD_y}$ modeling results from DO3SE model for three different PFTs over day of year (doy), March--October.
+A flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) has been chosen.
 \chem{[O_3]} are plotted on the same axis and scales as $G_\text{sto}^\text{leaf}$ but in units of $\unit{ppb}$.
-Due to non linearities in stomatal conductance, the climatological year represents rather an optimal year then a true climatology.
+Due to non linearity in stomatal conductance, the climatological year represents rather an optimal year then a true climatology.
 (a)--(c) Birch; (d)--(f) Norway spruce; (g)--(i) perennial grassland.
@@
   \label{tab:biomass_reduction}
-\begin{tabular}{lccrl}
+\begin{tabular}{lccrr}
 \tophline
@@
 \middlehline
-2018 & 11.99 & 2.34 & 8.06 & (10.51)$^*$\\
-2019 & 10.07 & 2.26 & 6.43 & (7.90)$^*$\\
-clim. & 13.81 & 2.58 & 6.67 & (8.29)$^*$\\
+2018 & 11.99 & 2.34 & 8.06 & 10.51$^*$\\
+2019 & 10.07 & 2.26 & 6.43 & 7.90$^*$\\
+clim. & 13.81 & 2.58 & 6.67 & 8.29$^*$\\
 \bottomhline
 \end{tabular}
-\belowtable{$^*$ above ground bioamass reduction} % Table Footnotes
+\belowtable{$^*$ above ground biomass} % Table Footnotes
 \end{table}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{DO3SE_results_rel}
+  \caption{Daily averaged $\mathrm{\Delta POD_y}/\Delta t$ from DO3SE model for three different PFTs over day of year (doy), March--October.
+Error bars/-bands indicate $1 \sigma$ standard deviation.
+(a) Birch; (b) Norway spruce; (c) perennial grassland.}
+  \label{fig:pody_rel}
+\end{figure}
 
@@
 \label{sec:conc} 
-TEXT
+\begin{itemize}
+  \item What did we learn from the 2018/2019 seasons in terms of ozone damage in the far north?
+  \item What are the main uncertainties in the tools we've used?
+  \item What implications could this have for future scenarios in which 2018 conditions may occur more often?
+\end{itemize}
+PODY is no good measure without a damage function with higher temporal resolution than annual.
+This has been already pointed out by \citet{AE:Musselman2006}!
+Higher than normal spring uptake (weather conditions) does not amount to critical uptake on an annual basis.
+Weather and climatic conditioning of plants is more important for PODy than episodes of higher than usual ozone concentrations.
+Therefor high $\Delta \mathrm{POD_y} / \Delta t$ is not accounted for properly.
+It is important to know the plant response functions to temperature, light, humidity/precipitation as best as possible if one wants to assess risk of damage.
+
+Regarding the shift (earlier) and prolongation (later) of the growing season in the far north no general assessment can be made.
+Data show that in both 2018 and 2019 ozone uptake in early season was anomalous high in all studied PFTs.
+The extended growing season in 2019 might have countered higher episodic ozone uptake in 2018.
+In both years, low mid-season uptake due to weather conditions washed out any effect of these.
 

### 2021-01-28 VegOzone.tex
First updates in accordance to meeting today.

@@
 We assess whether critical levels for these species were breached during the summer seasons.
-{\bf TODO: Summary of results.}
+We find that an unadjusted transfer of currently used standard parameters and methodes on ozone damage assessment (IPC Mapping Manual) to vegetation in the subarctics will result in an missinterpretation of POD1 values.
+{\bf TODO: Results.
+Difference in POD1 compared to mapping manual parameterization.}
 \end{abstract}
@@
 All factors differ with plant functional type (PFT) $k$.
-Stomata are typically considered closed below a light intensity of $50\,\unit{W\,m^{-2}}$.
+{\bf TODO: Put in f-functions.}
 
@@
 \end{equation}
-with the ambient ozone flux through the stomata $\Phi_\chem{[O_3]} = g_\mathrm{sto} \cdot \chem{[O_3]}$ and the flux threshold $\mathrm{y}$.
+with the ambient ozone flux through the stomata $\Phi_\chem{[O_3]} = g_\mathrm{sto} \cdot \chem{[O_3]}$ and a flux threshold $\mathrm{y}$ representing the detoxification potential of the plant.
 The stomatal conductance is either measured directly or computed in accordance to Eq.~(\ref{eq:stomatal}).\\
@@
 
+{\bf TODO: Write about boreal and cold parameter sets and their justification.
+Write about suggested method for optimizing PFTs in a simple but physiological way based on optimizing maximum climatological mean $g_\mathrm{sto}$ and morning responds.
+If plants have a short growing season up north optimizing \chem{CO_2} uptake around noon (highest light intensity) and speed up opening of stomata at lower intensities (morning), additional contraint lower climatological variance?}
+
 \subsection{POD and ozone impact}
@@
 
+Bigger picture:
+\begin{itemize}
+\item Ozone damage assessment using mapping manual "compromise" parameterization.
+\item Land-atmosphere modelling: Problems in productivity of species especially in the arctic regions due to PFTs which are not suited for the climate.
+Automating optimization of climate adaption of general PFTs.
+\end{itemize}
+
 %% The following commands are for the statements about the availability of data sets and/or software code corresponding to the manuscript.

### 2021-01-28 VegOzone.tex
Included f-functions for stomatal conductance computation.

@@
 \Author[2]{Ane Victoria}{Vollsnes}
+\Author[2]{Aud Else}{Berglen Erikson}
 \Author[3]{Lisa}{Emberson}
@@
 All factors differ with plant functional type (PFT) $k$.
-{\bf TODO: Put in f-functions.}
+
+The temperature adjustment function
+%
+\begin{equation}
+  f_T = \frac{T_\text{2m}-T_\text{min}}{T_\text{opt}-T_\text{min}} \cdot \left(\frac{T_{\text{max}}-T_\text{2m}}{T_{\text{max}}-T_\text{opt}}\right)^\beta,
+  \label{eq:f_temp}
+\end{equation}
+with $\beta = \frac{T_\text{max}-T_\text{opt}}{T_\text{opt}-T_\text{min}}$ is a mixed reciprocal polynomial function of temperature.
+The shape parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various plant functional types.
+%
+the water vapor pressure deficit penalty factor
+%
+\begin{equation}
+  f_D = f_\text{min}+(1-f_\text{min}) \cdot \frac{D_\text{min} - \text{VPD}}{D_\text{min}-D_\text{max}}.
+  \label{eq:f_vpd}
+\end{equation}
+%
+with $f_\text{min}$, $D_\text{min}$, $D_\text{max}$ describes the stomata reaction to imminent heat stress.
+
+The wavelength band $400-700\,\unit{nm}$ plant chlorophyll responds to is called photosynthetic active radiation (PAR).
+Its integral is the photosynthetic photon flux density (PPFD).
+The correction factor $f_\mathrm{light}$ in dependency of PPFD is:
+\begin{equation}
+  f_\text{light} = 1-\text{exp}(-\alpha_\text{light}\cdot \text{PPFD}).
+  \label{eq:flight}
+\end{equation}
+
+\citet{ACP:Bueker2012} discusses several formulations available to describe the reduction in stomata conductance due to water content in the soil.
 

### 2021-02-05 VegOzone.tex
First words about parameter update

@@
 
-\title{The 2018 heatwave and its implications on ozone induced damage on vegetation in a subarctic climate}
+%\title{The 2018 heatwave and its implications on ozone induced damage on vegetation in a subarctic climate}
+\title{Ozone induced damage on vegetation in a subarctic climate: implications of Plant Functional Types on Phototoxic Ozone Dose modeling}
 
@@
 \Author[3]{Connie}{O'Neill}
-%\Author[2]{Aud Else}{Berglen Eriksen}
-%\Author[1]{Bj{\o}rg}{Rognerud}
 \Author[1]{Frode}{Stordal}
@@
 
-\runningtitle{2018 heatwave and implication on ozone induced damage on vegetation in subarctic climate}
+\runningtitle{Ozone induced damage on vegetation in subarctic climate}
 
@@
 \begin{abstract}
+  In the future, persitent heatwaves may occur more often. This
   An increased occurrence of persistent heatwaves, as one possible future scenario, generates favorable conditions for the formation of ambient air ozone.
@@
 \subsection{Model parameters}
-A comprehensive list of model parameters used in this study can be found in Supplement~Table~S1.
+A comprehensive list of model parameters used as starting point in this study can be found in Supplement~Table~S1.
+These default PFTs are derived from \citet{ICP:MappingManual2017} and will be referred to as \emph{mapping manual} in the following.
 Start ($A_\text{start}$) and end ($A_\text{end}$) of growing season for each PFT (Norway spruce, birch, perennial grassland) are estimated as follows.
@@
 This results in 3.75, 3, and 2.625 for Norway spruce, birch, and grassland, respectively.
-
 As pointed out by \citet{ACP:Bueker2012}, a very important factor in deriving a proper $\mathrm{POD_y}$ is the soil texture or how much water the soil can hold.
-As Svanhovd is located in the bed of the Pasvik river, we assume a sandy loam texture at $60\,\unit{cm}$ depth.
+As Svanhovd is located in the bed of the Pasvik river, we assume a sandy loam texture at $60\,\unit{cm}$ depth.\\
+
+In addition to the mapping manual PFTs, we derived hypothetical PFTs optimized for the climatic conditions in the target region.
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{javis_funcs_temp_hist_grassland}\\
+  \includegraphics[width=8.3cm]{javis_funcs_f_light_grassland}
+\caption{TEXT}
+\label{fig:f_temp_grassland}
+\end{figure}
 

### 2021-02-08 VegOzone.tex
Change figure
Change title and rewrite abstract.

@@
 %\title{The 2018 heatwave and its implications on ozone induced damage on vegetation in a subarctic climate}
-\title{Ozone induced damage on vegetation in a subarctic climate: implications of Plant Functional Types on Phototoxic Ozone Dose modeling}
+\title{Characterizing subarctic biomes for land surface modeling of pollution and climate risk.}
 
@@
 \begin{abstract}
-  In the future, persitent heatwaves may occur more often. This
-  An increased occurrence of persistent heatwaves, as one possible future scenario, generates favorable conditions for the formation of ambient air ozone.
-Vegetation highly specialized to sub-arctic climate is vulnerable to rapid environmental changes inflicted by global warming and might become more susceptible to ozone in the future.
-Over large parts of Europe the summer 2018 had been extraordinarily hot and dry and caused large wildfires in northern Sweden in particular.
-This can be regarded as a test case for such a future scenario.
-In both 2018 and 2019, we have monitored ambient air ozone concentrations at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd in Northern Norway.
+Here we assess the importance of characterization of key meteorological variables and air pollution profiles on the modeled susceptibility of subarctic vegetation to ozone.
+Changes in climate (e.g.
+temperature, water availability) will manifest in both direct and indirect effects influencing the native species sensitivity to air pollution.
+A likely increase in frequency of persistent heatwaves in the future, will inflict heat related stress while generating favorable conditions for the formation of ambient air ozone, e.g.
+due to wildfires.
+We have monitored ambient air ozone concentrations at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd in Northern Norway in the vegetation period 2018/19.
 Due to data acquisition problems, ozone concentrations for two weeks in July 2018 were missing from our record.
-We present a reconstruction based on probability density function with respect to the Swedish and Finnish atmospheric monitoring sites in the region.
+We present a reconstruction based on probability density function with respect to the Swedish and Finnish atmospheric monitoring sites and compare with available surface ozone products for the region.
 Over all, ozone concentrations did not differ significantly between the two years.
-While temperatures and global irradiance diverged significantly from multi annual mean, precipitation varied only to some extend.
-Coincidentally, we have observed ozone-induced visible injuries on clovers in the ozone garden at Svanhovd in 2018, but not in 2019.
-We investigate the difference in uptake of ozone using the DO3SE model, with respect to the typical vegetation (beech/birch and conifers) at the location.
-We assess whether critical levels for these species were breached during the summer seasons.
-We find that an unadjusted transfer of currently used standard parameters and methodes on ozone damage assessment (IPC Mapping Manual) to vegetation in the subarctics will result in an missinterpretation of POD1 values.
-{\bf TODO: Results.
-Difference in POD1 compared to mapping manual parameterization.}
+While temperatures and global irradiance diverged substantially from the multi annual mean precipitation varied only to some extend.
+We investigate the difference in uptake of ozone using the DO3SE model, for typical vegetation at the location.
+We find that bespoke parameterization for subarctic vegetation is essential for an accurate estimation of Phototoxic Ozone Dose (POD) and an assessment of damage risk.
+Accurate estimates of the start and length of growing season are also crucial.
+We find that temperature and its derivative vapor pressure deficit are likely the most important climate variables while the variability of ozone concentration within climatologically reasonable limits plays a minor role.
+The use of subarctic parameterization suggests an improvement in risk assessment as POD values are increased in 2018; corroborated by evidence of visible injury in the Svanhovd ozone garden.
 \end{abstract}
@@
   \includegraphics[width=8.3cm]{javis_funcs_temp_hist_grassland}\\
-  \includegraphics[width=8.3cm]{javis_funcs_f_light_grassland}
+  \includegraphics[width=8.3cm]{javis_funcs_rad_hist_grassland}
 \caption{TEXT}

### 2021-02-08 VegOzone.tex
First suggested changes by Lisa. List with todos for introduction.

@@
 Changes in climate (e.g.
-temperature, water availability) will manifest in both direct and indirect effects influencing the native species sensitivity to air pollution.
+temperature, water availability) will manifest in both direct and indirect effects influencing the species' sensitivity to air pollution.
 A likely increase in frequency of persistent heatwaves in the future, will inflict heat related stress while generating favorable conditions for the formation of ambient air ozone, e.g.
@@
 
-Tropospheric and in particular ambient air ozone is a highly toxic pollutant to all ecosystems.
-In continuously high concentration, ozone negatively affects life expectancy of humans \citep{WHO2008} and is cause to early senescence as well as programmed cell death in plants \citep{PCE:Kangasjarvi2005}.
-On average, a global loss of yield in four major crops (wheat, rice, maize, and soybean) of $3-15\,\unit{\%}$ \citep{PJ:Ainsworth2017} and a loss in primary production in forestry of $7\,\unit{\%}$ \citep{GCB:Wittig2009,EP:Matyssek2012} has been attributed to ozone.
-Former can proof a vital threat to food security in fast developing countries, e.g., in East and South-East Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018}, where ozone related to air pollution has increased in the past years.
-At the same time, a successful implementation of air quality regulations reduced the number of episodes of peak concentrations especially in summertime over Europe and North America \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
-
-With a radiative forcing of $0.40 \pm 0.20\,\unit{W\,m^{-2}}$, ozone is a potent climate forcer, ranked third behind \chem{CO_2} and \chem{CH_4} \citep[Chapter 8]{IPCC2013}.
-Its damaging effects on global vegetation might provide an important additional feedback to climate change \citep{Nat:Sitch2007}.
-
-In the presents of nitrogen oxides (\chem{NO_x}), tropospheric ozone is formed in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) and hydrocarbons known as volatile organic substances (VOCs).
+\begin{itemize}
+\item Wealth of evidence demonstrating ozone damage to vegetation (perhaps focus on key evidence from northern climates and key species (e.g.
+birch, Norway srpuce etc...
+- this might help to make it clear why we focus on these species with the PODy modelling))
+\item Although O3 trends stagnant or decreasing in Europe there is unique photochemcistry over the arctic (causing spring time peaks in ozone) that may mean ozone damage continues to be an issue
+\item The earlier timing of growing seasons with climate change and 24 hour sunlight may also enhance the exposure of plants during those times when they are physiologically active to higher [O3]
+\item Research also clearly shows how ecosystem sensitivity to vegetation is influenced by climate variables, both because this can influence ozone uptake (dose) as well as the impact of altered ozone doses (can add in some relevant detail here on ozone damage mechanisms - again perhaps with a focus on what we know for colder climates)
+\item Risk assessment methods have been developed that allow for ozone uptake (i.e. PODy)
+\item Comparison of these methods with previously used concentration based methods (e.g.
+AOT40) show that PODy is better able to represent the risk to vegetation from ozone, especially in the more climate extreme parts of Europe (see Mills et al 2011 paper attached - that also includes some work for clover that we could use)
+\item These methods rely on accurate representation of plant physiological response to climate variables (i.e.
+the species or plant funtional type specific responses used to parameterise risk assessment models)
+\item These models are most often parameterised using data for less extreme climates (i.e.
+temperate or continental climates; and to a lesser extent, boreal and Mediterranean climates.
+Parameterisation of these models for more extreme bio-geographical regions such as sub arctic climates has not been perfromed and relies on assuming that boreal parameterisations will be applicable.
+\item It is important to understand the implications of using these more generic parameterisations for more extreme climate regions, here we explore how risk estimates might be over or underestimated using generic vs bespoke sub-arctic parameterisations
+\item we also explore the role [O3] profile and changing climate variables to understand how this risk might alter in the future with the advent of climate change 
+\item This allows both an understanding of how the threat from ozone may change over time but also which are the key aspects of plant physiology that might determine risk (and hence which parameters might require special attention to get right) both for pollution impact assessment, but also more generally for describing gas exchange response to changing climates that will influence bio-geochemcial cycling (could we add something in here about the paucity of PFTs representing subarctic conditions in biogeochemcial models?).
+\end{itemize}
+
+Ground level ozone is a highly toxic pollutant known to cause damage to a variety of ecosystems around the World \citep{PT:Emberson2020}.
+At continuously high concentration, ozone negatively affects life expectancy of humans \citep{WHO2008} and is a cause of visible injury, phtosynthetic damage, early senescence as well as programmed cell death \citep{PCE:Kangasjarvi2005}.
+Annual global yield losses of four major crops (wheat, rice, maize, and soybean) of $3-15\,\unit{\%}$ \citep{PJ:Ainsworth2017} and a suggested loss in primary production in forestry of $7\,\unit{\%}$ \citep{GCB:Wittig2009,EP:Matyssek2012} has been attributed to ozone.
+The crop yield loss studies indicate a threat to food security in rapidly developing countries, e.g., in East and South-East Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018}, where ozone related to enhanced pollutant emissions from the transport, residential, power generation and industrial sectors have increased in the past few decades.
+At the same time, a successful implementation of air quality regulations coupled with economic restructuring which saw a move from coal to gas, reduced the number of episodes of peak concentrations especially in summertime over Europe and North America \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
+
+With a radiative forcing value of $0.40 \pm 0.20\,\unit{W\,m^{-2}}$, ozone is a potent climate forcer, ranked third behind \chem{CO_2} and \chem{CH_4} \citep[Chapter 8]{IPCC2013}.
+Its damaging effects on the carbon sequestered by global vegetation might provide an important additional feedback to climate change \citep{Nat:Sitch2007}.
+
+In the presence of nitrogen oxides (\chem{NO_x}), tropospheric ozone is formed in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) and hydrocarbons known as volatile organic substances (VOCs).
 In the absence of local pollution sources, episodes of high ozone concentrations are often attributed to long-range transport of precursors.
-In northern Fennoscandia, episodes in 2003 and 2006 have been traced back to emissions related to forest fires in southern and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}.
+In northern Fennoscandia, ozone episodes in 2003 and 2006 have been traced back to \chem{NO_x} and \chem{VOC} emissions related to forest fires in southern and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}.
 Hydrocarbons are also emitted by living vegetation in form of terpenes and monoterpenes, so called biogenic volatile substances (BVOCs).
-Those emissions are associated with, e.g., thermal stress, defense against herbivores, or attraction of pollinators \citep{TPS:Penuelas2003}.
-
-Despite a relatively short tropospheric life-time of roughly $22\,\unit{days}$, ranging from a few days in the tropical boundary layer to up to $1\,\unit{year}$ in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}, ozone is subject to advection over long distances.
-In the annual cycle, a distinct maximum typically occurs in spring followed by lower concentrations throughout the summer.
-This is most likely due to stronger dry deposition on land vegetation than on bare ground or snow and ice \citep{AE:Monks2000}.
+These emissions are associated with, e.g., thermal stress, defense against herbivores, or attraction of pollinators \citep{TPS:Penuelas2003}.
+
+Despite a relatively short tropospheric life-time of approximately $22\,\unit{days}$, ranging from a few days in the tropical boundary layer to up to $1\,\unit{year}$ in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}, ozone is subject to advection over long distances.
+In the annual cycle in boreal and temperate climates, a distinct maximum typically occurs in spring followed by lower concentrations throughout the summer.
+This is most likely due to stronger dry deposition on land vegetation than on bare ground or snow and ice \citep{AE:Monks2000} causing a build off of ozone and its precursors in the course of polar night.
+Come spring they are photochemical activated again.
 This spring peak is reached earlier in northern Fennoscandia than in more southern locations but is highly variable throughout the years \citep{AB:Klingberg2009, BER:Klingberg2019}.
@@
 
-At the same time, the start of growing season in northern Fennoscandia is shifting to earlier days of the year in response to the general warming trend \citep[e.g.,]{GCB:Menzel2006,RS:Hogda2013,IJB:Karlsen2007} and thereby into close vicinity of the ozone spring peak.
-This may promote a higher susceptibility of deciduous species to ozone.
-
-Since the industrial revolution, tropospheric background concentrations of ozone in the northern hemisphere have doubled and kept increasing until recently.
-But long-term observations started before 1950 are scarce \citep[Chapter 2]{IPCC2013}.
-Recent trend analyses indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric ozone started to level-off or even decline \citep{ACP:Wespes2018,ESA:Gaudel2018}.\\
-
-\emph{Contributor: Ane} {\bf TODO: More detailed ozone damage description?}\\
+At the same time, the start of growing season in northern Fennoscandia is shifting to earlier days of the year in response to the general warming trend \citep[e.g.,]{GCB:Menzel2006,RS:Hogda2013,IJB:Karlsen2007} and thereby converging with the period of the ozone spring peak.
+This may promote a higher susceptibility of deciduous species to ozone and could also affect coniferous species which will have flushes of new needles in the spring time which are likely, through their immaturity, to be more susceptible to ozone damage.
+
+Even though long-term (pre 1950) observations of surface ozone concentrations are scarce; evidence from trend analysis of those few sites that have monitored concentrations suggests that since the industrial revolution, background concentrations in the northern Hemisphere have at least doubled and continued to increase until relatively recently \citep[Chapter 2]{IPCC2013}.
+Recent trend analyses indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric ozone started to level-off or even decline \citep{ESA:Cooper2014, ACP:Wespes2018,ESA:Gaudel2018}.\\
+
 Ozone acts as an oxidative stress to plants.
 Its main action is imposed through reactions occurring in the cell walls and cell membranes of mesophyll cells inside the leaves.
-Ozone enters the leaves through stomata, that are kept open to allow for gas exchange between the leaf air spaces and the exterior air.
-When the stomata are closed, ozone does not enter.
-
+Ozone enters the leaves through stomata, that are kept open to allow for gas exchange allowing the entry of CO2 for photosynthesis and loss of H2O vapour via the plant transpiration stream.
+Stomatal aperture, and hence stomatal conductance to these gases (including ozone) will vary over the course of the day and growing season primarily to balance CO2 uptake against H2O vapour  loss.
+The higher the stomatal conductance, the higher the potential for ozone uptake.\\
+
+{\bf TODO: Move to Section~\ref{sec:do3se}.}
 The opening of stomata, quantified through their conductance to water vapor $g_\mathrm{sto}$, is governed by different environmental factors and can be described by a multiplicative ansatz \citep{PTRS:Jarvis1976,ICP:MappingManual2017}:

### 2021-02-08 VegOzone.tex
Note about moving parts of section.

@@
 {\bf TODO: Move to Section~\ref{sec:do3se}.}
-The opening of stomata, quantified through their conductance to water vapor $g_\mathrm{sto}$, is governed by different environmental factors and can be described by a multiplicative ansatz \citep{PTRS:Jarvis1976,ICP:MappingManual2017}:
+The opening of stomata is governed by different environmental factors and can be described by a multiplicative model \citep{PTRS:Jarvis1976,ICP:MappingManual2017}:
 \begin{equation}

### 2021-02-09 VegOzone.tex
Included more comprehensive describtion of the DO3SE model. Still has to be moved to Section 4.

@@
 {\bf TODO: Move to Section~\ref{sec:do3se}.}
-The opening of stomata is governed by different environmental factors and can be described by a multiplicative model \citep{PTRS:Jarvis1976,ICP:MappingManual2017}:
+The opening of stomata is governed by different species-specific and environmental factors and can be described by a multiplicative model \citep{PTRS:Jarvis1976,ICP:MappingManual2017}:
 \begin{equation}
@@
 %
-The factors herein are normalized and vary within a range $0-1$.
+Where gmax is the specific-specific maximum stomatal conductance value which is then modified by seasonal and environmental factors that vary within a range $0-1$.
 They are empirically determined and account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_T$), water vapor pressure deficit ($f_D$), and soil water content ($f_\text{SW}$).
-All factors differ with plant functional type (PFT) $k$.
-
-The temperature adjustment function
-%
-\begin{equation}
-  f_T = \frac{T_\text{2m}-T_\text{min}}{T_\text{opt}-T_\text{min}} \cdot \left(\frac{T_{\text{max}}-T_\text{2m}}{T_{\text{max}}-T_\text{opt}}\right)^\beta,
+All factors differ with species or plant functional type (PFT) denoted with $k$.
+
+The temperature adjustment function is defined as
+%
+\begin{equation}
+  f_T = \frac{T_\text{air}-T_\text{min}}{T_\text{opt}-T_\text{min}} \cdot \left(\frac{T_{\text{max}}-T_\text{air}}{T_{\text{max}}-T_\text{opt}}\right)^\beta,
   \label{eq:f_temp}
 \end{equation}
-with $\beta = \frac{T_\text{max}-T_\text{opt}}{T_\text{opt}-T_\text{min}}$ is a mixed reciprocal polynomial function of temperature.
-The shape parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various plant functional types.
-%
-the water vapor pressure deficit penalty factor
-%
-\begin{equation}
-  f_D = f_\text{min}+(1-f_\text{min}) \cdot \frac{D_\text{min} - \text{VPD}}{D_\text{min}-D_\text{max}}.
+with $\beta = \frac{T_\text{max}-T_\text{opt}}{T_\text{opt}-T_\text{min}}$ a mixed reciprocal polynomial function of temperature.
+The shape parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various species or PFT and $T_\mathrm{air}$ is the surface air temperature.
+All temperatures are defined in units of \unit{^\circ C}.
+
+The water vapor pressure deficit function is
+%
+\begin{equation}
+  f_D = f_\text{min}+(1-f_\text{min}) \cdot \frac{D_\text{min} - \text{VPD}}{D_\text{min}-D_\text{max}}
   \label{eq:f_vpd}
@@
 %
-with $f_\text{min}$, $D_\text{min}$, $D_\text{max}$ describes the stomata reaction to imminent heat stress.
+where VPD is the leaf to air vapor pressure deficit in \unit{kPa} with $f_\text{min}$, $D_\text{min}$, $D_\text{max}$ describing the relative stomatal conductance to changes in vapor pressure deficit.
 
@@
 Its integral is the photosynthetic photon flux density (PPFD).
-The correction factor $f_\mathrm{light}$ in dependency of PPFD is:
-\begin{equation}
-  f_\text{light} = 1-\text{exp}(-\alpha_\text{light}\cdot \text{PPFD}).
+The relationship between relative gsto and PPFD ($f_\mathrm{light}$) is given by
+%
+\begin{equation}
+  f_\text{light} = 1-\text{exp}(-\alpha_\text{light}\cdot \text{PPFD})
   \label{eq:flight}
 \end{equation}
+%
+where $\alpha_\mathrm{light}$ is a slope parameter describing how fast stomata open or close at low light intensities.
 
 \citet{ACP:Bueker2012} discusses several formulations available to describe the reduction in stomata conductance due to water content in the soil.
+Here we use the relationship between reative $g_\mathrm{sto}$ and soil water potential (SWP) given by:
+%
+\begin{equation}
+  f_\mathrm{SW} = \text{min}\left\{1, \text{max}\left\{f_\mathrm{min}, \frac{(1 - f_\mathrm{min})\cdot(\mathrm{SWP_{min}} - \mathrm{SWP})}{\mathrm{SWP_{min}} - \mathrm{SWP_{max}}} + f_\mathrm{min} \right\} \right\}
+    \label{eq:fsw}
+\end{equation}
+%
+where SWP is the soil water potential across the root zone and $\mathrm{SWP_{min}}$ and $\mathrm{SWP_{max}}$ are the paramters describing the $f_\mathrm{SW}$ relationship.
+The DO3SE model as described in \citet{ACP:Bueker2012} is used to simulate SWP across a species or PFT specific root depth according to the Penman--Monteith energy balance method that drives water cycling through the soil--plant--atmosphere system.
 
 To measure the risk of damage inflicted to vegetation through ozone, many different metrics have been defined.
-Two main types can be differentiated, concentration-based (e.g., $\mathrm{AOT_{40}}$) and flux-based (e.g.,$\mathrm{POD_{0.1}}$) metrics.
-Due to their simplicity, $\mathrm{AOT_{x}}$ are still widely used, despite the actual integrated flux of ozone entering through the stomata is more relevant for an estimation of the actual toxic dose \citep{ICP:MappingManual2017}.
-The Accumulated Ozone Exposure over a threshold concentration of $\mathrm{x}$ in units of $\unit{ppb}$ is defined as:
+Two main types can be differentiated, concentration-based or flux-based metrics, e.g., Accumulated Over a Threshold $\mathrm{x}$ ($\mathrm{AOT_x}$) and Phyto-toxic Ozone Dose over a threshold $\mathrm{y}$ ($\mathrm{POD_y}$), respectively.
+Due to their simplicity, $\mathrm{AOT_x}$ are still widely used, despite the actual integrated flux of ozone entering through the stomata being more relevant for an estimation of the actual phyto-toxic ozone dose \citep{ICP:MappingManual2017}.
+The $\mathrm{AOT_x}$ in units of $\unit{ppb}$ is defined as:
 \begin{equation}
@@
 \end{equation}
-with the ambient ozone concentration $\chem{[O_3]}$. The integration time varies, but is usually defined from $8\,\unit{am}-8\,\unit{pm}$.
-The Phototoxic Ozone Dose (POD) over a threshold y is defined as:
-\begin{equation}
-  \mathrm{POD_y} = \int{\mathrm{max(\Phi_\chem{[O_3]}-y)}\cdot \mathrm{dt}},
+with the ambient ozone concentration $\chem{[O_3]}$. The  time over which the integration is calculated varies, but is usually defined from $8\,\unit{am}-8\,\unit{pm}$.
+The $\mathrm{POD_y}$ is calculated by first estimating the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf.
+The terms $r_c$ (leaf surface resistance) and $r_b$ (quasi-laminar resistance) allow for the deposition of \chem{O_3} to the leaf and the fraction that is taken up via the stomata.
+%
+\begin{equation}
+  \Phi_\mathrm{sto} = \chem{[O_3]} \cdot \frac{u(z_1) \cdot g_\mathrm{sto} \cdot r_c}{r_b + r_c}.
+  \label{eq:flux_stomata}
+\end{equation}
+%
+The quasi-laminar boundary layer resistance is calculated by
+%
+\begin{equation}
+  r_b = 1.3 \cdot 150 \cdot \sqrt{\frac{L}{u(z_1)}}.
+  \label{eq:quasi_laminar_resist}
+\end{equation}
+%
+Where $L$ is the cross-wind leaf dimension, $u(z_1)$ the wind speed at height $z_1$, and the factor $1.3$ accounts for the diffusivities between heat and \chem{O_3}.
+$\mathrm{POD_y}$ is then calculated according to:
+%
+\begin{equation}
+  \mathrm{POD_y} = \int{(\Phi_\mathrm{sto}-y)\cdot \mathrm{dt}},
   \label{eq:pod}
 \end{equation}
-with the ambient ozone flux through the stomata $\Phi_\chem{[O_3]} = g_\mathrm{sto} \cdot \chem{[O_3]}$ and a flux threshold $\mathrm{y}$ representing the detoxification potential of the plant.
-The stomatal conductance is either measured directly or computed in accordance to Eq.~(\ref{eq:stomatal}).\\
+%
+with the hourly averaged ambient air ozone flux through the stomata Eq.~(\ref{eq:flux_stomata}) and a flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}\,s^{-1}}.
+The flux threshold $\mathrm{y}$ is representing the detoxification potential of the plant and is typically only exceeded during daylight hours (i.e.
+when global radiation is above $50\,\unit{W\,m^{-2}}$).
+
+Finally, exceedance of the flux-based critical level (CL) is calculated by:
+%
+\begin{equation}
+  \mathrm{CL_{exeed}} = \mathrm{POD_y} - \mathrm{CL}.
+\end{equation}
+%
 

### 2021-02-09 VegOzone.tex
Finished initial comments of Lisa in Section 1. Started to remove TODOs from Section 2 and added picture of clover in ozone garden.

@@
 In the absence of local pollution sources, episodes of high ozone concentrations are often attributed to long-range transport of precursors.
-In northern Fennoscandia, ozone episodes in 2003 and 2006 have been traced back to \chem{NO_x} and \chem{VOC} emissions related to forest fires in southern and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}.
+In northern Fennoscandia, ozone episodes in 2003 and 2006 have been traced back to ozone precursor emissions related to forest fires in southern and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}.
 Hydrocarbons are also emitted by living vegetation in form of terpenes and monoterpenes, so called biogenic volatile substances (BVOCs).
@@
 According to a report by the Norwegian Meteorological institute \citep{MetNOR2019}, the summer of 2018 was the warmest and driest ever recorded in eastern, western, and southern Norway.
-In the north (including Finnmark), it was amongst the warmest on record -- favorable conditions for boundary layer ozone formation.
+In the north (including Finnmark), it was amongst the warmest on record -- favorable conditions for planetary boundary layer ozone formation.
 
@@
 In Sweden, about 500 fires had been reported (five times more than in a usual summer) and an estimated total of $25000\,\unit{hectare}$ burned down in the course of the most extensive fires in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) in younger history \citep{SOU2019}.
-
-Elevated ozone in Europe was likely due to several other factors as well, like in the 2003 drought period, such as biogenic VOC emissions, reduced dry deposition, reduced clouds and increased solar radiation, photochemical reactions, and stagnant weather conditions \citep{JGR:Solberg2018}.\\
+Coincident peak \chem{[O_3]} are found in ozone monitoring data from Svanhovd but also over whole northern Fennoscandia in July (see Fig.~\ref{fig:data_svanvik_2018_2019} or Fig.~\ref{fig:ozone_reconstruction_2018_07}).
+
+Like in the 2003 drought period, elevated ozone in Europe was likely promoted by a combination of various other factors as well, such as biogenic VOC emissions, reduced cloud and hence increased solar radiation, faster photochemical reactions, and reduced dry deposition and turbulent mixing due to stagnant weather conditions \citep{JGR:Solberg2018}.\\
 
 In terms of $\mathrm{AOT_{40}}$, the threshold for protection of vegetation of $3000\,\unit{ppb}$ (integrated over the growing season) given by the EU directive has been well exceeded at many ozone monitoring sites in Southern Norway in the 2018 growing season \citep{NILU2019}.
-{\bf TODO: come back to this later on in the discussions(?) - it has not been breached at Svanvik in northern Fennoscandia.}
+{\bf TODO: might be interesting to look at Mills et al 2003 and the different spatial extents of risk estimated when using AOT40 vs PODy.
+That could be useful to explain why we focus on PODy (rather then AOT40).- Lisa Come back to this later on in the discussions(?) - it has not been breached at Svanvik in northern Fennoscandia.}
 
@@
 In addition, the derived surface ozone concentration climatology for northern Fennoscandia will be compared with available ozone reanalysis products.
-We use the DO3SE model to estimate the ozone uptake by natural and semi-natural vegetation to investigate the main environmental driver of the observed ozone damage in 2018 (Section~\ref{sec:do3se}).
+We use the DO3SE model to estimate the ozone uptake by natural and semi-natural vegetation to investigate main environmental drivers for the observed ozone damage in 2018 (Section~\ref{sec:do3se}).
+In this context, we biefely introduce the main DO3SE model equations and derive bespoke parameterizations for subarctic species.
 In Section~\ref{sec:conc} finally, we will summarize our results, draw conclusions, and give an outlook on further work.
@@
 It is where the eastern Siberian taiga meets the western Boreal forest.
-[...] Svanhovd was originally established as a State Demonstration and Experimental Farm in the 1930's, in order to expand agriculture in Pasvik valley [...]" - \href{https://www.nibio.no/en/about-eng/addresses/northern-norway/svanhovd}{NIBIO web page, accessed March 2020}.
+[...] Svanhovd was originally established as a State Demonstration and Experimental Farm in the 1930's, in order to expand agriculture in Pasvik valley [...]" -- \href{https://www.nibio.no/en/about-eng/addresses/northern-norway/svanhovd}{NIBIO web page, last accessed March 2020}.
 
@@
 
-\emph{Comment: Invite co-authors from NILU and NIBIO?}
-
 \begin{figure}[t]
@@
 \label{subsubsec:atmo_svanvik}
-For the growing seasons 2018/19, monitoring data of key variables (temperature, precipitation, and ozone) have been downloaded from \href{luftkvalitet.no}{luftkvalitet.no} which is operated by NILU.
+For the growing seasons 2018/19, monitoring data of key variables (temperature, pressure, precipitation, and ozone) have been obtained from \href{luftkvalitet.no}{luftkvalitet.no} which is operated by NILU.
 
 All agrometeorological variables including temperature and precipitation are available from September 1992 to present day (LandbruksMeteorologisk Tjeneste provided by NIBIO -- \href{https://lmt.nibio.no/}{lmt.nibio.no}, note the station name here is Pasvik).
-
-{\bf TODO: Some more (technical) details about ozone and atmospheric monitoring at Svanhovd? $\rightarrow$ Sverre?}
-
-\emph{Comment: We refrain from doing a full cross-evaluation for these meteorological fields, since this has been done elsewhere.
-It's somewhat clear that this is a job of the operators in the monitoring network \citep{MetNOR2019}.}
 
@@
 The highest summer ozone concentration ($\chem{[O_3]} = 50.2\,\unit{ppb}$) was measured on July 25.
-This coincides with the period of the most extensive forest fires in central Swedish (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) which occurred from July 12--29 and destroyed about $18000\,\unit{hectare}$ of forest \citep{SOU2019}.
+This coincides with the period of the most extensive forest fires in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) which occurred from July 12--29 and destroyed about $18000\,\unit{hectare}$ of forest \citep{SOU2019}.
 However, due to the above mentioned data acquisition problems, we missed most of the corresponding ozone data for this event and most likely also the peak ozone concentration.
@@
 We will need to come back to this in the discussion section after the Do3se modeling part.} \emph{Author: Ane/Aud -- pictures are welcome, just upload them on overleaf or send them to me.}
+\begin{figure}
+  \includegraphics[width=8.3cm]{IMG_7739.JPG}
+  \caption{Clover in the Svanhovd ozone garden. The plants had to be secured against herbivores with a wire-mesh fence. The plants shown are approximately $6-12\,\unit{cm}$ in height.}
+\end{figure}
 
@@
 From 1986--1996 ozone monitoring has been operated by NILU at Svanhovd, but no ozone measurements have been conducted for the past $14\,\unit{years}$.
-Due to probable changes in northern hemispheric tropospheric background ozone and a reduction of summertime peak values related to ozone precursors that fell under air quality regulations {\bf TODO: Citations!}, we also have to have a look at ozone observations at other sites in northern Fennoscandia.
+Due to probable changes in northern hemispheric tropospheric background ozone and a reduction of summertime peak values related to ozone precursors that fell under air quality regulations ein Europe, we also have to have a look at ozone observations at other sites in northern Fennoscandia.
 In regard to ozone, available observational data are sparse in this region.
@@
 In Table~\ref{tab:ebas_obs}, we give an overview over the exact locations and time range of operational data taking.
-All stations with available long term observations are located on higher ground than Svanhovd which will result in higher surface ozone concentrations in comparison as ozone abundance, in general, decreases with decreasing altitude in the absence of local precursors.
-{\bf TODO: Citation see references in \citet{AB:Klingberg2009}!} The periphery vegetation at all sites is similar and consists mainly of pine forests, birch, and heath.
+All stations with available long term observations are located on higher ground than Svanhovd which will result in higher surface ozone concentrations in comparison as ozone abundance, in general, decreases with decreasing altitude in the absence of local precursors \citep[e.g.,][]{AB:Klingberg2009}.
+The periphery vegetation at all sites is similar and consists mainly of pine forests, birch, and heath.
 However, at close vicinity it varies substantially.

### 2021-02-09 VegOzone.tex
Moved in text web addresses to bibliography

@@
 It is where the eastern Siberian taiga meets the western Boreal forest.
-[...] Svanhovd was originally established as a State Demonstration and Experimental Farm in the 1930's, in order to expand agriculture in Pasvik valley [...]" -- \href{https://www.nibio.no/en/about-eng/addresses/northern-norway/svanhovd}{NIBIO web page, last accessed March 2020}.
+[...] Svanhovd was originally established as a State Demonstration and Experimental Farm in the 1930's, in order to expand agriculture in Pasvik valley [...]" \citep[Svanhovd][]{NIBIO_Svanhovd}.
 
@@
 \label{subsubsec:atmo_svanvik}
-For the growing seasons 2018/19, monitoring data of key variables (temperature, pressure, precipitation, and ozone) have been obtained from \href{luftkvalitet.no}{luftkvalitet.no} which is operated by NILU.
-
-All agrometeorological variables including temperature and precipitation are available from September 1992 to present day (LandbruksMeteorologisk Tjeneste provided by NIBIO -- \href{https://lmt.nibio.no/}{lmt.nibio.no}, note the station name here is Pasvik).
+For the growing seasons 2018/19, monitoring data of key variables (temperature, pressure, precipitation, and ozone) have been obtained from luftkvalitet.no which is operated by \citet{NILU_AIRQ}.
+
+All agrometeorological variables including temperature and precipitation are available from September 1992 to present day \citep[LandbruksMeteorologiske Tjeneste][note the station name here is Pasvik]{LMT_NIBIO}.
 
@@
 \label{subsubsec:ozone_garden}
-{\bf TODO: Details about the ozone garden and observed damage on ozone sensitive plants.
-We will need to come back to this in the discussion section after the Do3se modeling part.} \emph{Author: Ane/Aud -- pictures are welcome, just upload them on overleaf or send them to me.}
+{\bf TODO: Details about the ozone garden and observed damage on ozone sensitive plants. We will need to come back to this in the discussion section after the Do3se modeling part.}
 \begin{figure}
@@
 \label{subsubsec:ebas}
-All surface ozone observations used in this study are available from the EBAS database operated by NILU (\href{http://ebas.nilu.no/}{ebas.nilu.no}).
+All surface ozone observations used in this study are available from the EBAS database operated by \citet{NILU_EBAS}.
 In Table~\ref{tab:ebas_obs}, we give an overview over the exact locations and time range of operational data taking.
@@
 The temporal as well as spatial resolution of these reanalysis products is rather coarse: 3-hourly and $0.75^\circ\times 0.75^\circ$ or roughly $29.3\,\unit{km}\times 83.4\,\unit{km}$ at the location of Svanvik, respectively.
-In addition, higher resolution surface ozone reanalysis for whole Europe is available from Copernicus service for regional air quality control ({\bf TODO: From May 31st on through the CAMS Atmosphere Data Store (ADS)}).
-The regional air quality model reanalysis ensemble (CAMS regional Ensemble) is based on nine European state-of-the-art numerical air quality models ({\bf TODO: Reference -  https://atmosphere.copernicus.eu/documentation-regional-systems, last accessed April 2020}).
+In addition, higher resolution surface ozone reanalysis for whole Europe is available from Copernicus service for regional air quality control (note: since May 31st 2020 through the CAMS Atmosphere Data Store (ADS)).
+The regional air quality model reanalysis ensemble (CAMS regional Ensemble) is based on nine European state-of-the-art numerical air quality models \citep{Copernicus_Regio_AIRQ}.
 The ensemble mean is at fairly high spatial and temporal resolution compared to the global reanalysis from ECMWF: $0.1^\circ\times 0.1^\circ$ or roughly $3.9\,\unit{km}\times 11.1\,\unit{km}$ and 1-hourly, respectively.

### 2021-02-10 VegOzone.tex
Moved model describtion, added citation, removed wrong LAI estimate. Added some text about ozone garden.

@@
 %\title{The 2018 heatwave and its implications on ozone induced damage on vegetation in a subarctic climate}
-\title{Characterizing subarctic biomes for land surface modeling of pollution and climate risk.}
+\title{Characterizing subarctic biomes for land surface modeling of pollution and climate risk}
 
@@
 
-\runningtitle{Ozone induced damage on vegetation in subarctic climate}
+\runningtitle{Subarctic biomes for land surface modelling}
 
@@
 While temperatures and global irradiance diverged substantially from the multi annual mean precipitation varied only to some extend.
-We investigate the difference in uptake of ozone using the DO3SE model, for typical vegetation at the location.
+We investigate the difference in uptake of ozone using the $\mathrm{DO_3SE}$ model, for typical vegetation at the location.
 We find that bespoke parameterization for subarctic vegetation is essential for an accurate estimation of Phototoxic Ozone Dose (POD) and an assessment of damage risk.
@@
 
-{\bf TODO: Move to Section~\ref{sec:do3se}.}
-The opening of stomata is governed by different species-specific and environmental factors and can be described by a multiplicative model \citep{PTRS:Jarvis1976,ICP:MappingManual2017}:
-\begin{equation}
-  g^k_\text{sto, m} = g^k_\text{max, m} \cdot f^k_\text{phen} \cdot f^k_\text{light} \cdot \max{\left\{f^k_\text{min}, f^k_T \cdot f^k_D \cdot f^k_\text{SW}\right\}}.
-  \label{eq:stomatal}
-\end{equation}
-%
-Where gmax is the specific-specific maximum stomatal conductance value which is then modified by seasonal and environmental factors that vary within a range $0-1$.
-They are empirically determined and account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_T$), water vapor pressure deficit ($f_D$), and soil water content ($f_\text{SW}$).
-All factors differ with species or plant functional type (PFT) denoted with $k$.
-
-The temperature adjustment function is defined as
-%
-\begin{equation}
-  f_T = \frac{T_\text{air}-T_\text{min}}{T_\text{opt}-T_\text{min}} \cdot \left(\frac{T_{\text{max}}-T_\text{air}}{T_{\text{max}}-T_\text{opt}}\right)^\beta,
-  \label{eq:f_temp}
-\end{equation}
-with $\beta = \frac{T_\text{max}-T_\text{opt}}{T_\text{opt}-T_\text{min}}$ a mixed reciprocal polynomial function of temperature.
-The shape parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various species or PFT and $T_\mathrm{air}$ is the surface air temperature.
-All temperatures are defined in units of \unit{^\circ C}.
-
-The water vapor pressure deficit function is
-%
-\begin{equation}
-  f_D = f_\text{min}+(1-f_\text{min}) \cdot \frac{D_\text{min} - \text{VPD}}{D_\text{min}-D_\text{max}}
-  \label{eq:f_vpd}
-\end{equation}
-%
-where VPD is the leaf to air vapor pressure deficit in \unit{kPa} with $f_\text{min}$, $D_\text{min}$, $D_\text{max}$ describing the relative stomatal conductance to changes in vapor pressure deficit.
-
-The wavelength band $400-700\,\unit{nm}$ plant chlorophyll responds to is called photosynthetic active radiation (PAR).
-Its integral is the photosynthetic photon flux density (PPFD).
-The relationship between relative gsto and PPFD ($f_\mathrm{light}$) is given by
-%
-\begin{equation}
-  f_\text{light} = 1-\text{exp}(-\alpha_\text{light}\cdot \text{PPFD})
-  \label{eq:flight}
-\end{equation}
-%
-where $\alpha_\mathrm{light}$ is a slope parameter describing how fast stomata open or close at low light intensities.
-
-\citet{ACP:Bueker2012} discusses several formulations available to describe the reduction in stomata conductance due to water content in the soil.
-Here we use the relationship between reative $g_\mathrm{sto}$ and soil water potential (SWP) given by:
-%
-\begin{equation}
-  f_\mathrm{SW} = \text{min}\left\{1, \text{max}\left\{f_\mathrm{min}, \frac{(1 - f_\mathrm{min})\cdot(\mathrm{SWP_{min}} - \mathrm{SWP})}{\mathrm{SWP_{min}} - \mathrm{SWP_{max}}} + f_\mathrm{min} \right\} \right\}
-    \label{eq:fsw}
-\end{equation}
-%
-where SWP is the soil water potential across the root zone and $\mathrm{SWP_{min}}$ and $\mathrm{SWP_{max}}$ are the paramters describing the $f_\mathrm{SW}$ relationship.
-The DO3SE model as described in \citet{ACP:Bueker2012} is used to simulate SWP across a species or PFT specific root depth according to the Penman--Monteith energy balance method that drives water cycling through the soil--plant--atmosphere system.
+According to a report by the Norwegian Meteorological institute \citep{MetNOR2019}, the summer of 2018 was the warmest and driest ever recorded in eastern, western, and southern Norway.
+In the north (including Finnmark), it was amongst the warmest on record -- favorable conditions for planetary boundary layer ozone formation.
+
+An unusually weak and northward shifted jet stream allowed for a persistent high pressure system above northern Europe, including Fennoscandia which blocked the pressure systems for several consecutive months.
+In the period May -- July, southern Norway had temperatures $4\,\unit{^\circ C}$ above normal.
+While southern Norway had only about $60\,\unit{\%}$ of normal precipitation, northern Norway as a whole had close to normal precipitation, but with local variations.
+
+Heat and drought stress on vegetation was exceptional not only in large parts of Fennoscandia but also in much of Europe, where the influence of the high pressure system extended even over a period of five months (April/May, July--September).
+These conditions gave rise to massive forest fires in different parts of Europe and thus an increase in ozone precursors.
+Boreal wildfires emit in addition to \chem{CO_2} also \chem{CO} ($\chem{[CO]}/\chem{[CO_2]} \propto 6-13\,\unit{\%}$) and \chem{VOCs} ($\chem{[VOCs]}/\chem{[CO_2]} \propto 0.5-1.5\,\unit{\%}$) \citep{AE:Cofer1990}.
+
+A total of 2079 forest fires of different size were registered in Norway in 2018, twice as many as in the preceding years 2016/17 \citep{DSB2019}, last accessed April 2020).
+In Sweden, about 500 fires had been reported (five times more than in a usual summer) and an estimated total of $25000\,\unit{hectare}$ burned down in the course of the most extensive fires in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) in younger history \citep{SOU2019}.
+Coincident peak \chem{[O_3]} are found in ozone monitoring data from Svanhovd but also over whole northern Fennoscandia in July (see Fig.~\ref{fig:data_svanvik_2018_2019} or Fig.~\ref{fig:ozone_reconstruction_2018_07}).
+
+Like in the 2003 drought period, elevated ozone in Europe was likely promoted by a combination of various other factors as well, such as biogenic VOC emissions, reduced cloud and hence increased solar radiation, faster photochemical reactions, and reduced dry deposition and turbulent mixing due to stagnant weather conditions \citep{JGR:Solberg2018}.\\
 
@@
 Two main types can be differentiated, concentration-based or flux-based metrics, e.g., Accumulated Over a Threshold $\mathrm{x}$ ($\mathrm{AOT_x}$) and Phyto-toxic Ozone Dose over a threshold $\mathrm{y}$ ($\mathrm{POD_y}$), respectively.
-Due to their simplicity, $\mathrm{AOT_x}$ are still widely used, despite the actual integrated flux of ozone entering through the stomata being more relevant for an estimation of the actual phyto-toxic ozone dose \citep{ICP:MappingManual2017}.
+Due to their simplicity, concentration-based metrics are still widely used, despite the actual integrated flux of ozone entering through the stomata being more relevant for an estimation of the phyto-toxic ozone dose \citep{GCB:Mills2010,ICP:MappingManual2017}.
+
 The $\mathrm{AOT_x}$ in units of $\unit{ppb}$ is defined as:
@@
 %
-
-According to a report by the Norwegian Meteorological institute \citep{MetNOR2019}, the summer of 2018 was the warmest and driest ever recorded in eastern, western, and southern Norway.
-In the north (including Finnmark), it was amongst the warmest on record -- favorable conditions for planetary boundary layer ozone formation.
-
-An unusually weak and northward shifted jet stream allowed for a persistent high pressure system above northern Europe, including Fennoscandia which blocked the pressure systems for several consecutive months.
-In the period May -- July, southern Norway had temperatures $4\,\unit{^\circ C}$ above normal.
-While southern Norway had only about $60\,\unit{\%}$ of normal precipitation, northern Norway as a whole had close to normal precipitation, but with local variations.
-
-Heat and drought stress on vegetation was exceptional not only in large parts of Fennoscandia but also in much of Europe, where the influence of the high pressure system extended even over a period of five months (April/May, July--September).
-These conditions gave rise to massive forest fires in different parts of Europe and thus an increase in ozone precursors.
-Boreal wildfires emit in addition to \chem{CO_2} also \chem{CO} ($\chem{[CO]}/\chem{[CO_2]} \propto 6-13\,\unit{\%}$) and \chem{VOCs} ($\chem{[VOCs]}/\chem{[CO_2]} \propto 0.5-1.5\,\unit{\%}$) \citep{AE:Cofer1990}.
-
-A total of 2079 forest fires of different size were registered in Norway in 2018, twice as many as in the preceding years 2016/17 \citep{DSB2019}, last accessed April 2020).
-In Sweden, about 500 fires had been reported (five times more than in a usual summer) and an estimated total of $25000\,\unit{hectare}$ burned down in the course of the most extensive fires in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) in younger history \citep{SOU2019}.
-Coincident peak \chem{[O_3]} are found in ozone monitoring data from Svanhovd but also over whole northern Fennoscandia in July (see Fig.~\ref{fig:data_svanvik_2018_2019} or Fig.~\ref{fig:ozone_reconstruction_2018_07}).
-
-Like in the 2003 drought period, elevated ozone in Europe was likely promoted by a combination of various other factors as well, such as biogenic VOC emissions, reduced cloud and hence increased solar radiation, faster photochemical reactions, and reduced dry deposition and turbulent mixing due to stagnant weather conditions \citep{JGR:Solberg2018}.\\
 
@@
 This indicates that the vegetation may have been more affected by ozone in 2018 than in 2019.
-The respective observed average ozone concentration in 2018/19, however, did not differ significantly by statistical means.
 
@@
 In addition, the derived surface ozone concentration climatology for northern Fennoscandia will be compared with available ozone reanalysis products.
-We use the DO3SE model to estimate the ozone uptake by natural and semi-natural vegetation to investigate main environmental drivers for the observed ozone damage in 2018 (Section~\ref{sec:do3se}).
-In this context, we biefely introduce the main DO3SE model equations and derive bespoke parameterizations for subarctic species.
+We use the $\mathrm{DO_3SE}$ model to estimate the ozone uptake by natural and semi-natural vegetation to investigate main environmental drivers for the observed ozone damage in 2018 (Section~\ref{sec:do3se}).
+In this context, we biefely introduce the main $\mathrm{DO_3SE}$ model equations and derive bespoke parameterizations for subarctic species.
 In Section~\ref{sec:conc} finally, we will summarize our results, draw conclusions, and give an outlook on further work.
@@
 \label{subsubsec:ozone_garden}
-{\bf TODO: Details about the ozone garden and observed damage on ozone sensitive plants. We will need to come back to this in the discussion section after the Do3se modeling part.}
+An ozone garden consists of carefully selected plant species of which one cultivar is more sensitive to ozone than the other.
+At the NIBIO Environment Centre Svanhovd, the likely northern-most ozone garden to date, an ozone garden was installed in the course of our project.
+Cultivated species were, e.g.
+clover, tobacco, and beans.
+As shown in Fig.~\ref{fig:ozone_garden}, the plants had to be protected from herbivores with a wire-mesh fence.
+In 2018, we have observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden.
+In contrast, no such damage was found on the clovers in 2019, although it was found on the sensitive tobacco cultivar.
+This indicates that the vegetation may have been more affected by ozone in 2018 than in 2019.
+
 \begin{figure}
@@
   \caption{Clover in the Svanhovd ozone garden. The plants had to be secured against herbivores with a wire-mesh fence. The plants shown are approximately $6-12\,\unit{cm}$ in height.}
+  \label{fig:ozone_garden}
 \end{figure}
@@
 From 1986--1996 ozone monitoring has been operated by NILU at Svanhovd, but no ozone measurements have been conducted for the past $14\,\unit{years}$.
-Due to probable changes in northern hemispheric tropospheric background ozone and a reduction of summertime peak values related to ozone precursors that fell under air quality regulations ein Europe, we also have to have a look at ozone observations at other sites in northern Fennoscandia.
+Due to probable changes in northern hemispheric tropospheric background ozone and a reduction of summertime peak values related to ozone precursors which fell under air quality regulations in Europe, we also have to have a look at ozone observations at other sites in northern Fennoscandia.
 In regard to ozone, available observational data are sparse in this region.
@@
 \label{sec:do3se}
-To investigate which of the environmental factors lead to the observed ozone damage on clover and tobacco in 2018 in contrast to 2019, when no such damage was found on the clovers but only on the sensitive tobacco cultivar, we model the ozone uptake using the DO3SE model and give an estimate on biomass reduction due to ozone uptake.
+To investigate which of the environmental factors lead to the observed ozone damage on clover and tobacco in 2018 in contrast to 2019, when no such damage was found on the clovers but only on the sensitive tobacco cultivar, we model the ozone uptake using the $\mathrm{DO_3SE}$ model and give an estimate on biomass reduction due to ozone uptake.
 
 \subsection{Model description}
-
-{\bf TODO: Some general words about DO3SE model}
+\label{subsec:model_describtion}
+
+The opening of stomata is governed by different species-specific and environmental factors and can be described by a multiplicative model \citep{PTRS:Jarvis1976,ICP:MappingManual2017}:
+\begin{equation}
+  g^k_\text{sto} = g^k_\text{max} \cdot f^k_\text{phen} \cdot f^k_\text{light} \cdot \max{\left\{f^k_\text{min}, f^k_\mathrm{T} \cdot f^k_\mathrm{VPD} \cdot f^k_\text{SW}\right\}}.
+  \label{eq:stomatal}
+\end{equation}
+%
+Where $g_\mathrm{max}$ is the specific-specific maximum stomatal conductance value which is then modified by seasonal and environmental factors that vary within a range $0-1$.
+They are empirically determined and account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_\mathrm{T}$), water vapor pressure deficit ($f_\mathrm{VPD}$), and soil water content ($f_\text{SW}$).
+All factors differ with species or plant functional type (PFT) denoted with $k$.
+
+The temperature adjustment function is defined as
+%
+\begin{equation}
+  f_\mathrm{T} = \frac{T_\text{air}-T_\text{min}}{T_\text{opt}-T_\text{min}} \cdot \left(\frac{T_{\text{max}}-T_\text{air}}{T_{\text{max}}-T_\text{opt}}\right)^\beta,
+  \label{eq:f_temp}
+\end{equation}
+with $\beta = \frac{T_\text{max}-T_\text{opt}}{T_\text{opt}-T_\text{min}}$ a mixed reciprocal polynomial function of temperature.
+The shape parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various species or PFT and $T_\mathrm{air}$ is the surface air temperature.
+All temperatures are defined in units of \unit{^\circ C}.
+
+The water vapor pressure deficit function is
+%
+\begin{equation}
+  f_\mathrm{VPD} = f_\text{min}+(1-f_\text{min}) \cdot \frac{D_\text{min} - \text{VPD}}{D_\text{min}-D_\text{max}}
+  \label{eq:f_vpd}
+\end{equation}
+%
+where VPD is the leaf to air vapor pressure deficit in \unit{kPa} with $f_\text{min}$, $D_\text{min}$, $D_\text{max}$ describing the relative stomatal conductance to changes in vapor pressure deficit.
+
+The wavelength band $400-700\,\unit{nm}$ plant chlorophyll responds to is called photosynthetic active radiation (PAR).
+Its integral is the photosynthetic photon flux density (PPFD).
+The relationship between relative $g_\mathrm{sto}$ and PPFD is given by
+%
+\begin{equation}
+  f_\text{light} = 1-\text{exp}(-\alpha_\text{light}\cdot \text{PPFD})
+  \label{eq:flight}
+\end{equation}
+%
+where $\alpha_\mathrm{light}$ is a slope parameter describing how fast stomata open or close at low light intensities.
+
+The $\mathrm{DO_3SE}$ model as described in \citet{ACP:Bueker2012} is used to simulate SWP across a species or PFT specific root depth according to the Penman--Monteith energy balance method that drives water cycling through the soil--plant--atmosphere system.
+\citet{ACP:Bueker2012} discuss several formulations available to parameterize the reduction in stomata conductance due to water content in the soil.
+Here we use the relationship between reative $g_\mathrm{sto}$ and soil water potential (SWP) given by:
+%
+\begin{equation}
+  f_\mathrm{SW} = \text{min}\left\{1, \text{max}\left\{f_\mathrm{min}, \frac{(1 - f_\mathrm{min})\cdot(\mathrm{SWP_{min}} - \mathrm{SWP})}{\mathrm{SWP_{min}} - \mathrm{SWP_{max}}} + f_\mathrm{min} \right\} \right\}
+    \label{eq:fsw}
+\end{equation}
+%
+where SWP is the soil water potential across the root zone and $\mathrm{SWP_{min}}$ and $\mathrm{SWP_{max}}$ are the paramters describing the $f_\mathrm{SW}$ relationship.
 
@@
 \label{subsec:ozone_reco}
-The DO3SE model requires hourly, continuous meteorological observations.
+The $\mathrm{DO_3SE}$ model requires hourly, continuous meteorological observations.
 In addition to variables covered in Section~\ref{sec:data}, $2\,\unit{m}$ wind $u_\text{2m}$ and vapor pressure deficit VPD are needed.
@@
 
-Because the available reanalysis products for surface ozone have their limitations, we deem it necessary to reconstruct the missing data in 2018 for the purpose of DO3SE modeling by applying a Reynolds's decomposition of the form:
+Because the available reanalysis products for surface ozone have their limitations, we deem it necessary to reconstruct the missing data in 2018 for the purpose of $\mathrm{DO_3SE}$ modeling by applying a Reynolds's decomposition of the form:
 \begin{equation}
@@
 \subsection{Model parameters}
+Here, we will derive bespoke parameterizations for the main vegetation types (Norway spruce, birch, perennial grassland) surrounding the NIBIO Environment Centre Svanhovd.
+We roughly estimated land use types for a $1\times 1\,\unit{km}$ patch centered at Svanhovd from areal pictures at kartverket.no: $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban.
+
 A comprehensive list of model parameters used as starting point in this study can be found in Supplement~Table~S1.
-These default PFTs are derived from \citet{ICP:MappingManual2017} and will be referred to as \emph{mapping manual} in the following.
-Start ($A_\text{start}$) and end ($A_\text{end}$) of growing season for each PFT (Norway spruce, birch, perennial grassland) are estimated as follows.
+These default PFTs are derived from \citet{ICP:MappingManual2017} and will be referred to as \emph{mapping manual}.
+Start ($A_\text{start}$) and end ($A_\text{end}$) of growing season for each PFT are estimated as follows.
 Evergreen needle leaf trees are photosynthetic active already at low temperatures \citep{TB:Kolari2007}.
@@
 $A_\text{end}$ amounts to day 261 and 274, respectively, and shall be used for all vegetation alike.
-The resulting growing season for Norway spruce in 2019 was one month longer then in 2018.
+{\bf TODO: Compare with latitudinal model?} The resulting growing season for Norway spruce in 2019 was one month longer then in 2018.
 For estimating $A_\text{start}$ of birch tree, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Fig.~\ref{fig:greening_season_change_Svanvik}).
-We find 129 and 130, respectively.
+We find 129 and 130, respectively {\bf Accuracy?}.
 For perennial grassland, we assume a latency period of about $1\,\unit{month}$ after snow melt.
@@
 In both years, the first snow-free day incidents with the estimated date of bud burst.
-As leafs of birches in northern Fennoscandia {\bf (TODO: Ane, species name :-))} are smaller than southern specimen, we have measured their dimension for the purpose of this study and find $(3.0\pm 0.5)\,\unit{cm}$.
+As leafs of birches in northern Fennoscandia \emph{Betula nana} {\bf (TODO: Right name?)} are smaller than southern specimen, we have measured their dimension for the purpose of this study and find $(3.0\pm 0.5)\,\unit{cm}$.
 \citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area.
-To localize leaf area index (LAI), we roughly characterize land use for the $1\times 1\,\unit{km}$ patch centered at Svanhovd from areal pictures at kartverket.no: $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban.
-We normalize LAI to deciduous trees based on standard DO3SE parameterization: $f_\text{con}=5/4\cdot f_\text{dec}$, $f_\text{grass}=7/8\cdot f_\text{dec}$.
+%To localize leaf area index (LAI), we roughly characterize land use for the $1\times 1\,\unit{km}$ patch centered at Svanhovd from areal pictures at kartverket.no: $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban.
+We normalize LAI to deciduous trees based on standard $\mathrm{DO_3SE}$ parameterization: $f_\text{con}=5/4\cdot f_\text{dec}$, $f_\text{grass}=7/8\cdot f_\text{dec}$.
 With these, we re-scale the average maximum LAI from MODIS ($\left<f\right> = 2.7$).
 This results in 3.75, 3, and 2.625 for Norway spruce, birch, and grassland, respectively.
+
 As pointed out by \citet{ACP:Bueker2012}, a very important factor in deriving a proper $\mathrm{POD_y}$ is the soil texture or how much water the soil can hold.
@@
 The climatology in $\mathrm{POD_y}$ therefore represents rather an ideal year for growth then a true climatology.
-To establish a true climatology, it would be necessary to run the DO3SE model for 10--20 years with respective observed phenology as well as meteorological and ozone forcing.
+To establish a true climatology, it would be necessary to run the $\mathrm{DO_3SE}$ model for 10--20 years with respective observed phenology as well as meteorological and ozone forcing.
 
@@
   \includegraphics[width=12cm]{DO3SE_results_pody_gsto_o3}
-  \caption{$\mathrm{POD_y}$ modeling results from DO3SE model for three different PFTs over day of year (doy), March--October.
+  \caption{$\mathrm{POD_y}$ modeling results from $\mathrm{DO_3SE}$ model for three different PFTs over day of year (doy), March--October.
 A flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) has been chosen.
@@
   \includegraphics[width=8.3cm]{DO3SE_results_rel}
-  \caption{Daily averaged $\mathrm{\Delta POD_y}/\Delta t$ from DO3SE model for three different PFTs over day of year (doy), March--October.
+  \caption{Daily averaged $\mathrm{\Delta POD_y}/\Delta t$ from $\mathrm{DO_3SE}$ model for three different PFTs over day of year (doy), March--October.
 Error bars/-bands indicate $1 \sigma$ standard deviation.

### 2021-02-11 VegOzone.tex
Start to write about latest results, remove obsolete text.

@@
 $A_\text{end}$ amounts to day 261 and 274, respectively, and shall be used for all vegetation alike.
-{\bf TODO: Compare with latitudinal model?} The resulting growing season for Norway spruce in 2019 was one month longer then in 2018.
+{\bf TODO: Compare with laitude model?
+Birch: 100--307} The resulting growing season for Norway spruce in 2019 was one month longer then in 2018.
 For estimating $A_\text{start}$ of birch tree, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Fig.~\ref{fig:greening_season_change_Svanvik}).
@@
 
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{DO3SE_results_pody_gsto_o3}
-  \caption{$\mathrm{POD_y}$ modeling results from $\mathrm{DO_3SE}$ model for three different PFTs over day of year (doy), March--October.
+%\begin{figure*}[t]
+%  \includegraphics[width=12cm]{DO3SE_results_pody_gsto_o3}
+%  \caption{$\mathrm{POD_y}$ modeling results from $\mathrm{DO_3SE}$ model for three different PFTs over day of year (doy), March--October.
 A flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) has been chosen.
@@
 From left to right: 2018, 2019, and climatological year.}
-  \label{fig:pody}
-\end{figure*}
+%  \label{fig:pody}
+%\end{figure*}
 

### 2021-02-11 VegOzone.tex
Confirm name of bertula species. Add caption to figure. Remove more obsolete text.

--- 2021-02-10 VegOzone.tex
@@
 \subsection{Model parameters}
-Here, we will derive bespoke parameterizations for the main vegetation types (Norway spruce, birch, perennial grassland) surrounding the NIBIO Environment Centre Svanhovd.
-We roughly estimated land use types for a $1\times 1\,\unit{km}$ patch centered at Svanhovd from areal pictures at kartverket.no: $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban.
+We roughly estimated land use types for an approximately $1\times 1\,\unit{km}$ patch centered at the NIBIO Environment Centre Svanhovd from Fig.~\ref{fig:svanhovd_research_station}: $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban.
+In the following, we derive bespoke parameterizations for the main vegetation types, which we refer to as Norway spruce, birch, perennial grassland.
 
@@
 $A_\text{end}$ amounts to day 261 and 274, respectively, and shall be used for all vegetation alike.
-{\bf TODO: Compare with latitudinal model?} The resulting growing season for Norway spruce in 2019 was one month longer then in 2018.
+The resulting growing season for Norway spruce in 2019 was one month longer then in 2018.
 For estimating $A_\text{start}$ of birch tree, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Fig.~\ref{fig:greening_season_change_Svanvik}).
-We find 129 and 130, respectively {\bf Accuracy?}.
+We find 129 and 130, respectively.
+From the latitude model typically used for localizing $\mathrm{DO_3SE}$, we get $A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$ indicating an about $2\,\unit{months}$ longer growing season.
 For perennial grassland, we assume a latency period of about $1\,\unit{month}$ after snow melt.
@@
 In both years, the first snow-free day incidents with the estimated date of bud burst.
-As leafs of birches in northern Fennoscandia \emph{Betula nana} {\bf (TODO: Right name?)} are smaller than southern specimen, we have measured their dimension for the purpose of this study and find $(3.0\pm 0.5)\,\unit{cm}$.
+As leafs of birches in northern Fennoscandia \emph{Betula pubescens} are smaller than southern specimen, we have measured their dimension for the purpose of this study and find $(3.0\pm 0.5)\,\unit{cm}$.
 \citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area.
-%To localize leaf area index (LAI), we roughly characterize land use for the $1\times 1\,\unit{km}$ patch centered at Svanhovd from areal pictures at kartverket.no: $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban.
-We normalize LAI to deciduous trees based on standard $\mathrm{DO_3SE}$ parameterization: $f_\text{con}=5/4\cdot f_\text{dec}$, $f_\text{grass}=7/8\cdot f_\text{dec}$.
-With these, we re-scale the average maximum LAI from MODIS ($\left<f\right> = 2.7$).
-This results in 3.75, 3, and 2.625 for Norway spruce, birch, and grassland, respectively.
 
@@
   \includegraphics[width=8.3cm]{javis_funcs_rad_hist_grassland}
-\caption{TEXT}
+\caption{Construction of bespoke response functions for grassland.
+(a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown toghether with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability densitiy function - PDF), respectively.
+Original mapping manual parameterization is shown in comparison as solid line.
+Note that $Q_0$ has been truncated to $0.006$.
+PPFD0.8 and PPFD1.2 refer to $\alpha$ values increasing/decreasing PPFD at $f_\mathrm{light}=0.5$ by $\pm 20\,\%$, respectively.}
 \label{fig:f_temp_grassland}
@@
 
-Except for perennial grassland, model integrations with climatological forcings show higher $\mathrm{POD_y}$ than 2018.
+%Except for perennial grassland, model integrations with climatological forcings show higher $\mathrm{POD_y}$ than 2018.
 This can be explained by means of non linearity in modeled stomatal conductance (Eq.~(\ref{eq:stomatal})) in Eq.~(\ref{eq:pod}).
@@
 
-Based on this interpretation of climatology in $\mathrm{POD_y}$, we can learn more about potential ozone damage.
+%Based on this interpretation of climatology in $\mathrm{POD_y}$, we can learn more about potential ozone damage.
 The time derivative $\frac{\mathrm{\Delta POD_y}}{\Delta t}$ displays very different features for each PFT and meteorological forcing.
@@
 
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{DO3SE_results_rel}
-  \caption{Daily averaged $\mathrm{\Delta POD_y}/\Delta t$ from $\mathrm{DO_3SE}$ model for three different PFTs over day of year (doy), March--October.
+%\begin{figure}[t]
+%  \includegraphics[width=8.3cm]{DO3SE_results_rel}
+%  \caption{Daily averaged $\mathrm{\Delta POD_y}/\Delta t$ from $\mathrm{DO_3SE}$ model for three different PFTs over day of year (doy), March--October.
 Error bars/-bands indicate $1 \sigma$ standard deviation.
 (a) Birch; (b) Norway spruce; (c) perennial grassland.}
-  \label{fig:pody_rel}
-\end{figure}
+%  \label{fig:pody_rel}
+%\end{figure}
 

### 2021-02-11 VegOzone.tex
Include bespoke parameterization figures and write some initial text about the method.

@@
 We roughly estimated land use types for an approximately $1\times 1\,\unit{km}$ patch centered at the NIBIO Environment Centre Svanhovd from Fig.~\ref{fig:svanhovd_research_station}: $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban.
-In the following, we derive bespoke parameterizations for the main vegetation types, which we refer to as Norway spruce, birch, perennial grassland.
-
+In the following, we derive bespoke parameterizations for the main vegetation types Norway spruce, birch, and perennial grassland.
 A comprehensive list of model parameters used as starting point in this study can be found in Supplement~Table~S1.
-These default PFTs are derived from \citet{ICP:MappingManual2017} and will be referred to as \emph{mapping manual}.
-Start ($A_\text{start}$) and end ($A_\text{end}$) of growing season for each PFT are estimated as follows.
+Default PFTs are derived from \citet{ICP:MappingManual2017} and will be referred to as \emph{mapping manual}.
+
+These suggested, hypothetical PFTs or species parameterizations ought to be better adapted to the climatic conditions in the target region, but have to be varified by experiements.
+We hypothize that plants are likely to optimize their growth throughout the short growing season and hence the uptake of \chem{CO_2}.
+For simplicity, we assume a first order proptionality between photosynthesis and $g_\mathrm{sto}$.
+Plants are therefor likely to optimize their stomatal conductance with respect to climatological drivers.
+In Fig.~\ref{fig:f_temp_grassland}, we display the construction of bespoke PFTs or species temperature and light response function $f_\mathrm{T}$ and $f_\mathrm{light}$ for grassland (for Norway spruce and birch see Appendix~\ref{fig:f_temp_spruce_birch}).
+The response functions are shown together with climatology probability density functions (PDFs) of temperature and global radiation, respectively.
+In both cases, a parameterization is considered more optimized the larger the intersection area between the response function and PDF is.
+With respect to temperature (Fig.~\ref{fig:f_temp_grassland}a)), we assume two different types: \emph{boreal} and \emph{cold}.
+In both cases, we assume that the species are more tolerant to cold conditions than in the default mapping manual parameterization and shift $f_\mathrm{T}$ towards colder temperature.
+Herein, \emph{cold} represents a species that is more tolerant to warmer temperatures, but slightly less tolerant to cold temperatures than \emph{boreal}.
+For Norway spruce, $T_\mathrm{opt}$ are estimated from a longer timeseries of observed \chem{CO_2} uptake over temperature at two observations sites in Finland presented in \citet{TB:Kolari2007}.
+For the the senisitivity regarding light, we assume that the plants may react either faster or slower to lower light intensities.
+We qualitatively quantify this by a $\pm 20\,\%$ shift in respective halfway PPFD($f_\mathrm{light}=0.5$).
+The resulting functions are shown in Fig.~\ref{fig:f_temp_grassland}b).
+All derived parameters are given in Tables~\ref{tab:sensitivity_tests_temp}--\ref{tab:sensitivity_tests_light}.
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{javis_funcs_temp_hist_grassland}\\
+  \includegraphics[width=8.3cm]{javis_funcs_rad_hist_grassland}
+\caption{Construction of bespoke response functions for grassland.
+(a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown toghether with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability densitiy function - PDF), respectively.
+Original mapping manual parameterization is shown in comparison as solid line.
+Note that $Q_0$ has been truncated to $0.006$.
+PPFD0.8 and PPFD1.2 refer to $\alpha$ values increasing/decreasing PPFD at $f_\mathrm{light}=0.5$ by $\pm 20\,\%$, respectively.}
+\label{fig:f_temp_grassland}
+\end{figure}
+
+\begin{table}[t]
+  \caption{Bespoke temperature prameterizations. MM refers to mapping manual \citep{GCB:Mills2010,ICP:MappingManual2017}.}
+  \label{tab:sensitivity_tests_temp}
+  \begin{tabular}{llrrr}
+    \tophline
+    Species & type & $T_\mathrm{min}$ & $T_\mathrm{opt}$ & $T_\mathrm{max}$ \\
+    \middlehline
+    \multirow{3}{*}{Spruce} & MM & 0 & 20 & \multirow{3}{*}{100}\\
+    & Boreal & 0 & 10 & \\
+    & Cold & 0 & 15 & \\
+    \middlehline
+    \multirow{3}{*}{Birch} & MM & 5 & 20 & \multirow{3}{*}{100}\\
+    & Boreal & 0 & 10\\
+    & Cold & 5 & 15\\
+    \middlehline
+    \multirow{3}{*}{Grassland} & MM & 10 & 24 & 36\\
+    & Boreal & 0 & 10 & 24 \\
+    & Cold & 5 & 15 & 36\\
+    \bottomhline
+    \end{tabular}
+\end{table}
+
+\begin{table}[t]
+  \caption{Bespoke light prameterizations. MM refers to mapping manual \citep{GCB:Mills2010,ICP:MappingManual2017}.}
+  \label{tab:sensitivity_tests_light}
+  \begin{tabular}{llr}
+    \tophline
+    Species & type & $\alpha$\\
+    \middlehline
+    \multirow{3}{*}{Spruce} & MM & 0.006 \\
+    & PPFD0.8 & 0.008  \\
+    & PPFD1.2 & 0.005  \\
+    \middlehline
+    \multirow{3}{*}{Birch} & MM & 0.004 \\
+    & PPFD0.8 & 0.005\\
+    & PPFD1.2 & 0.004\\
+    \middlehline
+    \multirow{3}{*}{Grassland} & MM & 0.011\\
+    & PPFD0.8 & 0.014 \\
+    & PPFD1.2 & 0.009 \\
+    \bottomhline
+    \end{tabular}
+\end{table}
+
+{\bf TODO: Write about boreal and cold parameter sets and their justification.
+Write about suggested method for optimizing PFTs in a simple but physiological way based on optimizing maximum climatological mean $g_\mathrm{sto}$ and morning responds.
+If plants have a short growing season up north optimizing \chem{CO_2} uptake around noon (highest light intensity) and speed up opening of stomata at lower intensities (morning), additional contraint lower climatological variance?}
+An idea to justify our choice based on the fullset of climatological data for Svanvik is shown in Fig.~\ref{fig:javis_func_opt_mean}.
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{javis_func_opt_mean.pdf}
+  \caption{Proposed metric to test bespoke response functions.}
+  \label{fig:javis_func_opt_mean}
+\end{figure}
+
+Start ($A_\text{start}$) and end ($A_\text{end}$) of growing season for each PFT are given in units of day of year (doy) and estimated as follows.
 Evergreen needle leaf trees are photosynthetic active already at low temperatures \citep{TB:Kolari2007}.
@@
 Net photosynthesis data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}).
-By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{day~of~year}$ for 2018 and day 106 in 2019.
+By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{doy}$ for 2018 and day 106 in 2019.
 $A_\text{end}$ amounts to day 261 and 274, respectively, and shall be used for all vegetation alike.
@@
 
-In addition to the mapping manual PFTs, we derived hypothetical PFTs optimized for the climatic conditions in the target region.
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{javis_funcs_temp_hist_grassland}\\
-  \includegraphics[width=8.3cm]{javis_funcs_rad_hist_grassland}
-\caption{Construction of bespoke response functions for grassland.
-(a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown toghether with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability densitiy function - PDF), respectively.
-Original mapping manual parameterization is shown in comparison as solid line.
-Note that $Q_0$ has been truncated to $0.006$.
-PPFD0.8 and PPFD1.2 refer to $\alpha$ values increasing/decreasing PPFD at $f_\mathrm{light}=0.5$ by $\pm 20\,\%$, respectively.}
-\label{fig:f_temp_grassland}
-\end{figure}
-
-{\bf TODO: Write about boreal and cold parameter sets and their justification.
-Write about suggested method for optimizing PFTs in a simple but physiological way based on optimizing maximum climatological mean $g_\mathrm{sto}$ and morning responds.
-If plants have a short growing season up north optimizing \chem{CO_2} uptake around noon (highest light intensity) and speed up opening of stomata at lower intensities (morning), additional contraint lower climatological variance?}
-
 \subsection{POD and ozone impact}
@@
 We have modeled $\mathrm{POD_y}$ for three natural/semi-natural vegetation PFTs, birch, Norway spruce, and perennial grassland.
-The results are shown comprehensively in Fig.~\ref{fig:pody} as a function of day of year (doy) for March--October.
+The results are shown comprehensively in Fig.~\ref{fig:pody} as a function of doy for March--October.
 A flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) has been chosen.
@@
 \chem{[O_3]} are plotted on the same axis and scales as $G_\text{sto}^\text{leaf}$ but in units of $\unit{ppb}$.
-Due to non linearity in stomatal conductance, the climatological year represents rather an optimal year then a true climatology.
-(a)--(c) Birch; (d)--(f) Norway spruce; (g)--(i) perennial grassland.
-From left to right: 2018, 2019, and climatological year.}
+(a, b) Birch; (c, d) Norway spruce; (e, f) perennial grassland.
+From left to right: 2018, 2019.
+{\bf TODO: substitute with figure for boreal PFTs?}}
   \label{fig:pody}

### 2021-02-12 VegOzone.tex
Updated plot and wrote draft about bespoke parameters.

@@
 
+\newpage
 \section{Data}
@@
 We roughly estimated land use types for an approximately $1\times 1\,\unit{km}$ patch centered at the NIBIO Environment Centre Svanhovd from Fig.~\ref{fig:svanhovd_research_station}: $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban.
-In the following, we derive bespoke parameterizations for the main vegetation types Norway spruce, birch, and perennial grassland.
-A comprehensive list of model parameters used as starting point in this study can be found in Supplement~Table~S1.
-Default PFTs are derived from \citet{ICP:MappingManual2017} and will be referred to as \emph{mapping manual}.
-
-These suggested, hypothetical PFTs or species parameterizations ought to be better adapted to the climatic conditions in the target region, but have to be varified by experiements.
-We hypothize that plants are likely to optimize their growth throughout the short growing season and hence the uptake of \chem{CO_2}.
-For simplicity, we assume a first order proptionality between photosynthesis and $g_\mathrm{sto}$.
-Plants are therefor likely to optimize their stomatal conductance with respect to climatological drivers.
-In Fig.~\ref{fig:f_temp_grassland}, we display the construction of bespoke PFTs or species temperature and light response function $f_\mathrm{T}$ and $f_\mathrm{light}$ for grassland (for Norway spruce and birch see Appendix~\ref{fig:f_temp_spruce_birch}).
-The response functions are shown together with climatology probability density functions (PDFs) of temperature and global radiation, respectively.
-In both cases, a parameterization is considered more optimized the larger the intersection area between the response function and PDF is.
-With respect to temperature (Fig.~\ref{fig:f_temp_grassland}a)), we assume two different types: \emph{boreal} and \emph{cold}.
-In both cases, we assume that the species are more tolerant to cold conditions than in the default mapping manual parameterization and shift $f_\mathrm{T}$ towards colder temperature.
-Herein, \emph{cold} represents a species that is more tolerant to warmer temperatures, but slightly less tolerant to cold temperatures than \emph{boreal}.
-For Norway spruce, $T_\mathrm{opt}$ are estimated from a longer timeseries of observed \chem{CO_2} uptake over temperature at two observations sites in Finland presented in \citet{TB:Kolari2007}.
-For the the senisitivity regarding light, we assume that the plants may react either faster or slower to lower light intensities.
-We qualitatively quantify this by a $\pm 20\,\%$ shift in respective halfway PPFD($f_\mathrm{light}=0.5$).
+
+Default PFTs are derived from \citet{ICP:MappingManual2017} and will be referred to as \emph{mapping manual} abbrieved MM.
+A comprehensive list of these model parameters can be found in Supplement~Table~S1.
+Intitial test simulations with the mapping manual derived PFTs showed unrealistically low $g_\mathrm{sto}$ in 2019 (Supplememt~Figs.SX--SX+2) particularly for grassland.
+Only during an extended periode in late July did we find substential stomatal conductance.
+In ecological terms, this would mean that there was almost no growth of grass in the summer of 2019 -- a prediction which is easily falsified by reality.
+We identified $f_\mathrm{T}$ as abnormally low, being the limiting factor of stomatal conductance.
+We propose bespoke parameterizations for the main vegetation types Norway spruce, birch, and perennial grassland better adapted to the climatic conditions in the target region.
+Note that these parameterizations represent hypothetical PFTs or species that have to be varified by experiements.
+
+We hypothize that perannial vegetation will likely prioritize their growth or fast adolesence throughout the short, sub-boreal growing season and hence their carbon sequestration.
+Since the $\mathrm{DO_3SE}$ in its current state does not simulate $A_\mathrm{n}$, we assume for out purpose a first order proptionality between carbon uptake (photosynthesis) and $g_\mathrm{sto}$ and tune the species temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) seperately for higher stomatal conductance.
+For this purpose, we graphically intersected propability density functions (PDFs) of climatological data for temperature and global irradiace with the respective response function (Fig.~\ref{fig:f_temp_grassland} for grassland; for Norway spruce and birch see Appendix Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}).
+Formally, a parameterization is considered more optimial the larger the intersection area between the response function and PDF is.
+
+For the climatization of the temperature response functions, we assume two different types: \emph{boreal} and \emph{cold}.
+In either cases, we further assume that the hypothetical species are more tolerant to cold conditions than the mapping manual default suggests and shift $f_\mathrm{T}$ towards colder temperatures.
+We construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to the mapping manual default.
+In the same way, \emph{boreal} refers to a hypothetical species which is very tolerant to cold but is sensitive to high temperatures, and most efficient at cool temperatures (close to the climatological mean temperature of 1992--2000).
+All three types for grassland are depicted together with the temperature PDF in Fig.~\ref{fig:f_temp_grassland}a).
+
+For Norway spruce, our hypothetical PFTs $T_\mathrm{opt}$ are motivated by a longer timeseries of both observed \chem{CO_2} uptake and temperature at two observation sites in Southern (\emph{cold}) and Northern (\emph{boreal}) Finland presented in \citet{TB:Kolari2007}.
+We assume a similar adaption in birch.
+The parameters are comprehensibly listed in Table~\ref{tab:sensitivity_tests_temp}.
+
+For the the senisitivity regarding light, we assume that the plants may react either faster or slower to a change at lower light intensities.
+We analytically derive the inverse function $f_{\mathrm{light},k}^{-1}$ for each PFT or species $k$
+\begin{equation}
+  \gamma_k(f_\mathrm{light}) := f_{\mathrm{light}, k}^{-1} = -\frac{\ln(1-f_\mathrm{light})}{\alpha_k}.
+  \label{eq:inverse_function}
+\end{equation}
+We calculate it for the mapping manual $\alpha_\mathrm{MM}$ at its halfway value $0.5$ 
+\begin{equation}
+  \gamma(0.5) = -\frac{\ln(0.5)}{\alpha_\mathrm{MM}},
+  \label{eq:inverse_function_halfway}
+\end{equation}
+and define a variation $\gamma' = \gamma \cdot \epsilon$ with $\epsilon \in \{0.8, 1.2\}$.
+We resolve Eq.~(\ref{eq:inverse_function}) for $\alpha$ and derive $\alpha'(\gamma')$.
 The resulting functions are shown in Fig.~\ref{fig:f_temp_grassland}b).
-All derived parameters are given in Tables~\ref{tab:sensitivity_tests_temp}--\ref{tab:sensitivity_tests_light}.
+All derived parameters are given in Table~\ref{tab:sensitivity_tests_light}.
 
@@
 
+\section{DO3SE parameterization}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{javis_funcs_temp_hist_evergreen}\\
+  \includegraphics[width=8.3cm]{javis_funcs_rad_hist_evergreen}
+\caption{Construction of bespoke response functions for Norway spruce.
+(a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown toghether with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability densitiy function - PDF), respectively.
+Original mapping manual parameterization is shown in comparison as solid line.
+Note that $Q_0$ has been truncated to $0.006$.
+PPFD0.8 and PPFD1.2 refer to $\alpha$ values increasing/decreasing PPFD at $f_\mathrm{light}=0.5$ by $\pm 20\,\%$, respectively.}
+\label{fig:f_temp_spruce}
+\end{figure}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{javis_funcs_temp_hist_birch}\\
+  \includegraphics[width=8.3cm]{javis_funcs_rad_hist_birch}
+\caption{Construction of bespoke response functions for Norway birch.
+(a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown toghether with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability densitiy function - PDF), respectively.
+Original mapping manual parameterization is shown in comparison as solid line.
+Note that $Q_0$ has been truncated to $0.006$.
+PPFD0.8 and PPFD1.2 refer to $\alpha$ values increasing/decreasing PPFD at $f_\mathrm{light}=0.5$ by $\pm 20\,\%$, respectively.}
+\label{fig:f_temp_birch}
+\end{figure}
+
 \noappendix       %% use this to mark the end of the appendix section. Otherwise the figures might be numbered incorrectly (e.g. 10 instead of 1).

### 2021-03-01 VegOzone.tex
Restructure introduction in accordance to Lisa's comments - not spell checked.

@@
 \Author[2]{Ane Victoria}{Vollsnes}
-\Author[2]{Aud Else}{Berglen Erikson}
+\Author[2]{Aud E. B.}{Eriksen}
 \Author[3]{Lisa}{Emberson}
@@
 
-\begin{itemize}
-\item Wealth of evidence demonstrating ozone damage to vegetation (perhaps focus on key evidence from northern climates and key species (e.g.
+Ground level ozone is a highly toxic pollutant known to cause damage to a variety of ecosystems around the World \citep{PT:Emberson2020}.
+At continuously high concentration levels, ozone negatively affects life expectancy of humans \citep{WHO2008} and is a cause of visible injury, photosynthetic damage, early senescence as well as programmed cell death of plants \citep{PCE:Kangasjarvi2005}.
+Annual global yield losses of four major crops (wheat, rice, maize, and soybean) of $3-15\,\unit{\%}$ \citep{PJ:Ainsworth2017} and a suggested loss in primary production in forestry of $7\,\unit{\%}$ \citep{GCB:Wittig2009,EP:Matyssek2012} has been attributed to ozone.
+{\bf TODO: Wealth of evidence demonstrating ozone damage to vegetation (perhaps focus on key evidence from northern climates and key species (e.g.
 birch, Norway srpuce etc...
-- this might help to make it clear why we focus on these species with the PODy modelling))
-\item Although O3 trends stagnant or decreasing in Europe there is unique photochemcistry over the arctic (causing spring time peaks in ozone) that may mean ozone damage continues to be an issue
-\item The earlier timing of growing seasons with climate change and 24 hour sunlight may also enhance the exposure of plants during those times when they are physiologically active to higher [O3]
-\item Research also clearly shows how ecosystem sensitivity to vegetation is influenced by climate variables, both because this can influence ozone uptake (dose) as well as the impact of altered ozone doses (can add in some relevant detail here on ozone damage mechanisms - again perhaps with a focus on what we know for colder climates)
-\item Risk assessment methods have been developed that allow for ozone uptake (i.e. PODy)
-\item Comparison of these methods with previously used concentration based methods (e.g.
-AOT40) show that PODy is better able to represent the risk to vegetation from ozone, especially in the more climate extreme parts of Europe (see Mills et al 2011 paper attached - that also includes some work for clover that we could use)
-\item These methods rely on accurate representation of plant physiological response to climate variables (i.e.
-the species or plant funtional type specific responses used to parameterise risk assessment models)
-\item These models are most often parameterised using data for less extreme climates (i.e.
-temperate or continental climates; and to a lesser extent, boreal and Mediterranean climates.
-Parameterisation of these models for more extreme bio-geographical regions such as sub arctic climates has not been perfromed and relies on assuming that boreal parameterisations will be applicable.
-\item It is important to understand the implications of using these more generic parameterisations for more extreme climate regions, here we explore how risk estimates might be over or underestimated using generic vs bespoke sub-arctic parameterisations
-\item we also explore the role [O3] profile and changing climate variables to understand how this risk might alter in the future with the advent of climate change 
-\item This allows both an understanding of how the threat from ozone may change over time but also which are the key aspects of plant physiology that might determine risk (and hence which parameters might require special attention to get right) both for pollution impact assessment, but also more generally for describing gas exchange response to changing climates that will influence bio-geochemcial cycling (could we add something in here about the paucity of PFTs representing subarctic conditions in biogeochemcial models?).
-\end{itemize}
-
-Ground level ozone is a highly toxic pollutant known to cause damage to a variety of ecosystems around the World \citep{PT:Emberson2020}.
-At continuously high concentration, ozone negatively affects life expectancy of humans \citep{WHO2008} and is a cause of visible injury, phtosynthetic damage, early senescence as well as programmed cell death \citep{PCE:Kangasjarvi2005}.
-Annual global yield losses of four major crops (wheat, rice, maize, and soybean) of $3-15\,\unit{\%}$ \citep{PJ:Ainsworth2017} and a suggested loss in primary production in forestry of $7\,\unit{\%}$ \citep{GCB:Wittig2009,EP:Matyssek2012} has been attributed to ozone.
-The crop yield loss studies indicate a threat to food security in rapidly developing countries, e.g., in East and South-East Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018}, where ozone related to enhanced pollutant emissions from the transport, residential, power generation and industrial sectors have increased in the past few decades.
-At the same time, a successful implementation of air quality regulations coupled with economic restructuring which saw a move from coal to gas, reduced the number of episodes of peak concentrations especially in summertime over Europe and North America \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
-
-With a radiative forcing value of $0.40 \pm 0.20\,\unit{W\,m^{-2}}$, ozone is a potent climate forcer, ranked third behind \chem{CO_2} and \chem{CH_4} \citep[Chapter 8]{IPCC2013}.
-Its damaging effects on the carbon sequestered by global vegetation might provide an important additional feedback to climate change \citep{Nat:Sitch2007}.
-
-In the presence of nitrogen oxides (\chem{NO_x}), tropospheric ozone is formed in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) and hydrocarbons known as volatile organic substances (VOCs).
-In the absence of local pollution sources, episodes of high ozone concentrations are often attributed to long-range transport of precursors.
-In northern Fennoscandia, ozone episodes in 2003 and 2006 have been traced back to ozone precursor emissions related to forest fires in southern and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}.
-Hydrocarbons are also emitted by living vegetation in form of terpenes and monoterpenes, so called biogenic volatile substances (BVOCs).
-These emissions are associated with, e.g., thermal stress, defense against herbivores, or attraction of pollinators \citep{TPS:Penuelas2003}.
-
+- this might help to make it clear why we focus on these species with the PODy modelling))}
+The crop yield loss studies indicate a threat to food security in rapidly developing countries, e.g., in East and South-East Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018}, where ozone concentrations related to enhanced pollutant emissions from the transport, residential, power generation and industrial sectors have increased in the past few decades.
+Even though long-term (pre 1950) observations of surface ozone concentrations are scarce; evidence from trend analysis of those few sites suggests that since the industrial revolution, background concentrations in the northern Hemisphere have at least doubled and continued to increase until relatively recently \citep[Chapter 2]{IPCC2013}.
+Recent trend analyses indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric ozone started to level-off or even decline \citep{ESA:Cooper2014, ACP:Wespes2018,ESA:Gaudel2018}.
+This can be attributed to a successful implementation of air quality regulations coupled with economic restructuring which saw a move from coal to gas, reduced the number of episodes of peak concentrations especially in summertime over Europe and North America \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
+However, there are indications that an increasing frequency of thermal stress accompanying climate change will reduce vegetations' capacity for ozone uptake and therefor scotch these achievments \citep{NCC:Lin2020}.\\
+
+Ground level ozone is formed in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) and hydrocarbons known as volatile organic substances (VOCs) in the presence of nitrogen oxides (\chem{NO_x}).
 Despite a relatively short tropospheric life-time of approximately $22\,\unit{days}$, ranging from a few days in the tropical boundary layer to up to $1\,\unit{year}$ in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}, ozone is subject to advection over long distances.
-In the annual cycle in boreal and temperate climates, a distinct maximum typically occurs in spring followed by lower concentrations throughout the summer.
-This is most likely due to stronger dry deposition on land vegetation than on bare ground or snow and ice \citep{AE:Monks2000} causing a build off of ozone and its precursors in the course of polar night.
-Come spring they are photochemical activated again.
-This spring peak is reached earlier in northern Fennoscandia than in more southern locations but is highly variable throughout the years \citep{AB:Klingberg2009, BER:Klingberg2019}.
-A cause for this is a complex interplay between atmospheric dynamics and chemistry \citep{AE:Laurila1996,BER:Hatakka2003} as well as anthropogenic and biogenic activity \citep{AE:Rummukainen1996,AE:Simpson2002,QJRMS:Galbally2007,NGS:Schnell2009}.
-\citet{ACP:Andersson2017} showed in a modeling study focused on Fennoscandia that the variability in winter can be attributed mainly to changes in atmospheric background (transport) of ozone, while summertime abundance is mostly affected by emissions of precursors in the rest of Europe.
+Episodes of high ozone concentrations in the absence of local pollution sources can therefor often be attributed to long-range transport of both ozone and its precursors.
+In northern Fennoscandia, episodes of enhanced ozone concentrations in 2003 and 2006 have been traced back to ozone precursor emissions related to forest fires in southern and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}.
+In addition, hydrocarbons are also emitted by vegetation itself in form of terpenes and monoterpenes, so called biogenic volatile substances (BVOCs).
+These emissions are associated with, e.g., thermal stress, defense against herbivores, or even attraction of pollinators \citep{TPS:Penuelas2003}.
+
+In the annual cycle in boreal and temperate climates, a distinct maximum of ground level ozone concentrations typically occurs in spring followed by lower concentrations throughout the summer.
+Several conditions unique to the Actic are stimulating the formation of this spring peak.
+Major pathways of ozone and its precursors' removal from the boundary layer include dry deposition on vegetated surfaces and photochemistry \citep{RG:Clifton2020}.
+As dry deposition to bare ground or snow and ice covered surfaces was found to be low \citep{ACP:Helmig2007}, both pathways of removal are substantially suppressed during polar night \citep{AE:Monks2000}.
+This leads to a build up of ozone and its precursors in the course of winter.
+Come spring precursors are photochemical activated again, while snow cover typically prevails for a longer period of time continuing to suppress ozone dry deposition.
+On the other hand, very high ozone concentrations occuring over shorter periodes of time can often be attributed to deep tropopause folding events over the Actic during winter.
+These events facilitate an intrusion of stratospheric air masses which are enriched in ozone \citep{JGR:Skerlak2015}.
+The spring peak is more pronounced in northern Fennoscandia than in more southern locations but shows a  high interannual variability \citep{AB:Klingberg2009, BER:Klingberg2019}.
+A cause for this is the complex interplay between atmospheric dynamics and chemistry \citep{AE:Laurila1996,BER:Hatakka2003} as described above as well as anthropogenic and biogenic activity \citep{AE:Rummukainen1996,AE:Simpson2002,QJRMS:Galbally2007,NGS:Schnell2009}.
+\citet{ACP:Andersson2017} showed in a modeling study focused on Fennoscandia that the variability of ozone concentrations in winter can be attributed mainly to changes in atmospheric background (transport) of ozone, while summertime abundance is mostly affected by emissions of precursors in the rest of Europe.
 Meteorology plays an important role throughout the year.
 
-At the same time, the start of growing season in northern Fennoscandia is shifting to earlier days of the year in response to the general warming trend \citep[e.g.,]{GCB:Menzel2006,RS:Hogda2013,IJB:Karlsen2007} and thereby converging with the period of the ozone spring peak.
-This may promote a higher susceptibility of deciduous species to ozone and could also affect coniferous species which will have flushes of new needles in the spring time which are likely, through their immaturity, to be more susceptible to ozone damage.
-
-Even though long-term (pre 1950) observations of surface ozone concentrations are scarce; evidence from trend analysis of those few sites that have monitored concentrations suggests that since the industrial revolution, background concentrations in the northern Hemisphere have at least doubled and continued to increase until relatively recently \citep[Chapter 2]{IPCC2013}.
-Recent trend analyses indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric ozone started to level-off or even decline \citep{ESA:Cooper2014, ACP:Wespes2018,ESA:Gaudel2018}.\\
+The start of growing season in northern Fennoscandia has been shifting to earlier dates in response to the general warming trend \citep[e.g.,]{GCB:Menzel2006,RS:Hogda2013,IJB:Karlsen2007} and thereby converging with the period of the ozone spring peak.
+At the same time the growing season is also becoming longer.
+In conjunction with midnight sun conditions (almost $24\,\unit{h}$ photosynthetic activity), a longer growing season, is prolonging the time in which vegetation can accumulate ozone.
+These factors may promote a higher potential susceptibility of northern vegetation to ozone in the future.
 
@@
 Its main action is imposed through reactions occurring in the cell walls and cell membranes of mesophyll cells inside the leaves.
-Ozone enters the leaves through stomata, that are kept open to allow for gas exchange allowing the entry of CO2 for photosynthesis and loss of H2O vapour via the plant transpiration stream.
-Stomatal aperture, and hence stomatal conductance to these gases (including ozone) will vary over the course of the day and growing season primarily to balance CO2 uptake against H2O vapour  loss.
-The higher the stomatal conductance, the higher the potential for ozone uptake.\\
+Ozone enters the leaves through stomata, that are kept open to allow for gas exchange allowing the entry of \chem{CO_2} for photosynthesis and loss of \chem{H_2O} vapour via the plant transpiration stream.
+Stomatal aperture, and hence stomatal conductance to these gases (including ozone) will vary over the course of the day and growing season primarily to balance \chem{CO_2} uptake against \chem{H_2O} vapour loss.
+The higher the stomatal conductance, the higher the potential for ozone uptake.
+Stomatal conductance has been emperically linked to environmental factors such as air temperature, photosynthetic active radiation, vapor pressure deficit, and soil water potential as well as photosynthesis itself.\\
+
+To measure the risk of damage inflicted to vegetation through ozone, different metrics have been defined.
+Two main types can be differentiated, concentration-based or flux-based metrics, e.g., Accumulated Over a Threshold $\mathrm{x}$ ($\mathrm{AOT_x}$) and Phyto-toxic Ozone Dose over a threshold $\mathrm{y}$ ($\mathrm{POD_y}$), respectively.
+Due to their simplicity, concentration-based metrics are still widely used.
+Despite comparisons of exposure-based methods with previously used concentration based methods (e.g.
+$\mathrm{AOT_{40}}$) show that $\mathrm{POD_y}$ is better suited to represent the risk to vegetation from ozone, especially in the more climate extreme parts of Europe \citep{GCB:Mills2010,ICP:MappingManual2017}.
+
+The concentration-based metric $\mathrm{AOT_x}$ (in units of $\unit{ppb}$) is defined as:
+\begin{equation}
+  \mathrm{AOT_x} = \int{\mathrm{max(\chem{[O_3]}-x)}\cdot \mathrm{dt}},
+  \label{eq:aot}
+\end{equation}
+with the ambient ozone concentration $\chem{[O_3]}$. The  time over which the integration is calculated varies, but is usually defined from $8\,\unit{am}-8\,\unit{pm}$.
+
+For computing the $\mathrm{POD_y}$, an estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf.
+Details specific to the $\mathrm{DO_3SE}$ model are given in Section~\ref{sec_do3se}.
+$\mathrm{POD_y}$ is then calculated according to:
+%
+\begin{equation}
+  \mathrm{POD_y} = \int{(\Phi_\mathrm{sto}-y)\cdot \mathrm{dt}},
+  \label{eq:pod}
+\end{equation}
+%
+with the hourly averaged ambient air ozone flux through the stomata Eq.~(\ref{eq:flux_stomata}) and a flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}\,s^{-1}}.
+The flux threshold $\mathrm{y}$ is representing the detoxification potential of the plant and is typically only exceeded during daylight hours (i.e.
+when global radiation is above $50\,\unit{W\,m^{-2}}$).
+
+Finally, exceedance of the flux-based critical level (CL) is calculated by:
+%
+\begin{equation}
+  \mathrm{CL_{exeed}} = \mathrm{POD_y} - \mathrm{CL}.
+\end{equation}
+%
 
@@
 In the north (including Finnmark), it was amongst the warmest on record -- favorable conditions for planetary boundary layer ozone formation.
-
 An unusually weak and northward shifted jet stream allowed for a persistent high pressure system above northern Europe, including Fennoscandia which blocked the pressure systems for several consecutive months.
@@
 While southern Norway had only about $60\,\unit{\%}$ of normal precipitation, northern Norway as a whole had close to normal precipitation, but with local variations.
-
-Heat and drought stress on vegetation was exceptional not only in large parts of Fennoscandia but also in much of Europe, where the influence of the high pressure system extended even over a period of five months (April/May, July--September).
+Thermal stress on vegetation was exceptional not only in large parts of Fennoscandia but also in much of Europe, where the influence of the high pressure system extended even over a period of five months (April/May, July--September).
 These conditions gave rise to massive forest fires in different parts of Europe and thus an increase in ozone precursors.
 Boreal wildfires emit in addition to \chem{CO_2} also \chem{CO} ($\chem{[CO]}/\chem{[CO_2]} \propto 6-13\,\unit{\%}$) and \chem{VOCs} ($\chem{[VOCs]}/\chem{[CO_2]} \propto 0.5-1.5\,\unit{\%}$) \citep{AE:Cofer1990}.
-
 A total of 2079 forest fires of different size were registered in Norway in 2018, twice as many as in the preceding years 2016/17 \citep{DSB2019}, last accessed April 2020).
@@
 Coincident peak \chem{[O_3]} are found in ozone monitoring data from Svanhovd but also over whole northern Fennoscandia in July (see Fig.~\ref{fig:data_svanvik_2018_2019} or Fig.~\ref{fig:ozone_reconstruction_2018_07}).
-
 Like in the 2003 drought period, elevated ozone in Europe was likely promoted by a combination of various other factors as well, such as biogenic VOC emissions, reduced cloud and hence increased solar radiation, faster photochemical reactions, and reduced dry deposition and turbulent mixing due to stagnant weather conditions \citep{JGR:Solberg2018}.\\
 
-To measure the risk of damage inflicted to vegetation through ozone, many different metrics have been defined.
-Two main types can be differentiated, concentration-based or flux-based metrics, e.g., Accumulated Over a Threshold $\mathrm{x}$ ($\mathrm{AOT_x}$) and Phyto-toxic Ozone Dose over a threshold $\mathrm{y}$ ($\mathrm{POD_y}$), respectively.
-Due to their simplicity, concentration-based metrics are still widely used, despite the actual integrated flux of ozone entering through the stomata being more relevant for an estimation of the phyto-toxic ozone dose \citep{GCB:Mills2010,ICP:MappingManual2017}.
-
-The $\mathrm{AOT_x}$ in units of $\unit{ppb}$ is defined as:
-\begin{equation}
-  \mathrm{AOT_x} = \int{\mathrm{max(\chem{[O_3]}-x)}\cdot \mathrm{dt}},
-  \label{eq:aot}
-\end{equation}
-with the ambient ozone concentration $\chem{[O_3]}$. The  time over which the integration is calculated varies, but is usually defined from $8\,\unit{am}-8\,\unit{pm}$.
-The $\mathrm{POD_y}$ is calculated by first estimating the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf.
-The terms $r_c$ (leaf surface resistance) and $r_b$ (quasi-laminar resistance) allow for the deposition of \chem{O_3} to the leaf and the fraction that is taken up via the stomata.
-%
-\begin{equation}
-  \Phi_\mathrm{sto} = \chem{[O_3]} \cdot \frac{u(z_1) \cdot g_\mathrm{sto} \cdot r_c}{r_b + r_c}.
-  \label{eq:flux_stomata}
-\end{equation}
-%
-The quasi-laminar boundary layer resistance is calculated by
-%
-\begin{equation}
-  r_b = 1.3 \cdot 150 \cdot \sqrt{\frac{L}{u(z_1)}}.
-  \label{eq:quasi_laminar_resist}
-\end{equation}
-%
-Where $L$ is the cross-wind leaf dimension, $u(z_1)$ the wind speed at height $z_1$, and the factor $1.3$ accounts for the diffusivities between heat and \chem{O_3}.
-$\mathrm{POD_y}$ is then calculated according to:
-%
-\begin{equation}
-  \mathrm{POD_y} = \int{(\Phi_\mathrm{sto}-y)\cdot \mathrm{dt}},
-  \label{eq:pod}
-\end{equation}
-%
-with the hourly averaged ambient air ozone flux through the stomata Eq.~(\ref{eq:flux_stomata}) and a flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}\,s^{-1}}.
-The flux threshold $\mathrm{y}$ is representing the detoxification potential of the plant and is typically only exceeded during daylight hours (i.e.
-when global radiation is above $50\,\unit{W\,m^{-2}}$).
-
-Finally, exceedance of the flux-based critical level (CL) is calculated by:
-%
-\begin{equation}
-  \mathrm{CL_{exeed}} = \mathrm{POD_y} - \mathrm{CL}.
-\end{equation}
-%
-
-In terms of $\mathrm{AOT_{40}}$, the threshold for protection of vegetation of $3000\,\unit{ppb}$ (integrated over the growing season) given by the EU directive has been well exceeded at many ozone monitoring sites in Southern Norway in the 2018 growing season \citep{NILU2019}.
-{\bf TODO: might be interesting to look at Mills et al 2003 and the different spatial extents of risk estimated when using AOT40 vs PODy.
-That could be useful to explain why we focus on PODy (rather then AOT40).- Lisa Come back to this later on in the discussions(?) - it has not been breached at Svanvik in northern Fennoscandia.}
+These methods described above rely on accurate representation of plant physiological response to climate variables (i.e.
+the species or plant funtional type specific responses used to parameterise risk assessment models) and are most often parameterized using data for less extreme climates, i.e., temperate or continental climates; and to a lesser extent, boreal and Mediterranean climates.
+Parameterization of these models for more extreme bio-geographical regions such as sub-arctic climates has not been perfromed and relies on assuming that less extreme boreal parameterizations will be applicable.
+It is important to understand the implications of using these more generic parameterizations for more extreme climate regions, here we explore how risk estimates might be over or underestimated using generic vs bespoke sub-arctic parameterisations.
+we also explore the role [O3] profile and changing climate variables to understand how this risk might alter in the future with the advent of climate change.
+This allows both an understanding of how the threat from ozone may change over time but also which are the key aspects of plant physiology that might determine risk (and hence which parameters might require special attention to get right) both for pollution impact assessment, but also more generally for describing gas exchange response to changing climates that will influence bio-geochemcial cycling.
+This might be one key to solve long-standing issues of coupled earthsystem models (ESM) which generally lack PFTs representing sub-arctic conditions in their landsurface model.
 
@@
 This indicates that the vegetation may have been more affected by ozone in 2018 than in 2019.
-
-As previously mentioned, ozone damage is rather driven by the actual uptake of ozone ($\mathrm{POD_y}$) than by the accumulated ambient air concentration above a certain threshold ($\mathrm{AOT_x}$).
-Through the stomatal conductance, $\mathrm{POD_y}$ depends on environmental drivers such as ambient air temperature, photoactive radiation, soil water availability, and water vapor pressure deficit.
+In terms of $\mathrm{AOT_{40}}$, the threshold for protection of vegetation of $3000\,\unit{ppb}$ (integrated over the growing season) given by the EU directive has been well exceeded at many ozone monitoring sites in Southern Norway in the 2018 growing season \citep{NILU2019} but not at Svanhovd.
 
@@
 We use the $\mathrm{DO_3SE}$ model to estimate the ozone uptake by natural and semi-natural vegetation to investigate main environmental drivers for the observed ozone damage in 2018 (Section~\ref{sec:do3se}).
-In this context, we biefely introduce the main $\mathrm{DO_3SE}$ model equations and derive bespoke parameterizations for subarctic species.
+In this context, we introduce the$\mathrm{DO_3SE}$ model equations and derive bespoke parameterizations for sub-arctic species to assess systematic uncertainties arising from the use of default parameterizations.
 In Section~\ref{sec:conc} finally, we will summarize our results, draw conclusions, and give an outlook on further work.
 
-\newpage
 \section{Data}
@@
 where SWP is the soil water potential across the root zone and $\mathrm{SWP_{min}}$ and $\mathrm{SWP_{max}}$ are the paramters describing the $f_\mathrm{SW}$ relationship.
+
+To compute $\mathrm{POD_y}$, we estimate the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf.
+The terms $r_c$ (leaf surface resistance) and $r_b$ (quasi-laminar resistance) allow for the deposition of \chem{O_3} to the leaf and the fraction that is taken up via the stomata.
+%
+\begin{equation}
+  \Phi_\mathrm{sto} = \chem{[O_3]} \cdot \frac{u(z_1) \cdot g_\mathrm{sto} \cdot r_c}{r_b + r_c}.
+  \label{eq:flux_stomata}
+\end{equation}
+%
+The quasi-laminar boundary layer resistance is calculated by
+%
+\begin{equation}
+  r_b = 1.3 \cdot 150 \cdot \sqrt{\frac{L}{u(z_1)}}.
+  \label{eq:quasi_laminar_resist}
+\end{equation}
+%
+Where $L$ is the cross-wind leaf dimension, $u(z_1)$ the wind speed at height $z_1$, and the factor $1.3$ accounts for the diffusivities between heat and \chem{O_3}.
 
@@
 
+{\bf TODO: might be interesting to look at Mills et al 2003 and the different spatial extents of risk estimated when using AOT40 vs PODy.
+That could be useful to explain why we focus on PODy (rather then AOT40).- Lisa Come back to this later on in the discussions(?) - it has not been breached at Svanvik in northern Fennoscandia.}
+
 Bigger picture:

### 2021-03-02 VegOzone_supplement.tex
Moved obsolete figure to supplement

--- 2021-02-11 VegOzone.tex
@@
-%% Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
-%% ---------------------------------
-%% This template should be used for copernicus.cls
-%% The class file and some style files are bundled in the Copernicus Latex Package, which can be downloaded from the different journal webpages.
-%% For further assistance please contact Copernicus Publications at: production@copernicus.org
-%% https://publications.copernicus.org/for_authors/manuscript_preparation.html
-
-%% Please use the following documentclass and journal abbreviations for preprints and final revised papers.
-
-%% 2-column papers and preprints
-\documentclass[bg, manuscript]{copernicus}
-
-%% Journal abbreviations (please use the same for preprints and final revised papers)
-
-% Advances in Geosciences (adgeo)
-% Advances in Radio Science (ars)
-% Advances in Science and Research (asr)
-% Advances in Statistical Climatology, Meteorology and Oceanography (ascmo)
-% Annales Geophysicae (angeo)
-% Archives Animal Breeding (aab)
-% ASTRA Proceedings (ap)
-% Atmospheric Chemistry and Physics (acp)
-% Atmospheric Measurement Techniques (amt)
-% Biogeosciences (bg)
-% Climate of the Past (cp)
-% DEUQUA Special Publications (deuquasp)
-% Drinking Water Engineering and Science (dwes)
-% Earth Surface Dynamics (esurf)
-% Earth System Dynamics (esd)
-% Earth System Science Data (essd)
-% E&G Quaternary Science Journal (egqsj)
-% European Journal of Mineralogy (ejm)
-% Fossil Record (fr)
-% Geochronology (gchron)
-% Geographica Helvetica (gh)
-% Geoscience Communication (gc)
-% Geoscientific Instrumentation, Methods and Data Systems (gi)
-% Geoscientific Model Development (gmd)
-% History of Geo- and Space Sciences (hgss)
-% Hydrology and Earth System Sciences (hess)
-% Journal of Bone and Joint Infection (jbji)
-% Journal of Micropalaeontology (jm)
-% Journal of Sensors and Sensor Systems (jsss)
-% Magnetic Resonance (mr)
-% Mechanical Sciences (ms)
-% Natural Hazards and Earth System Sciences (nhess)
-% Nonlinear Processes in Geophysics (npg)
-% Ocean Science (os)
-% Primate Biology (pb)
-% Proceedings of the International Association of Hydrological Sciences (piahs)
-% Scientific Drilling (sd)
-% SOIL (soil)
-% Solid Earth (se)
-% The Cryosphere (tc)
-% Weather and Climate Dynamics (wcd)
-% Web Ecology (we)
-% Wind Energy Science (wes)
-
-%% \usepackage commands included in the copernicus.cls:
-%\usepackage[german, english]{babel}
-%\usepackage{tabularx}
-%\usepackage{cancel}
-%\usepackage{multirow}
-%\usepackage{supertabular}
-%\usepackage{algorithmic}
-%\usepackage{algorithm}
-%\usepackage{amsthm}
-%\usepackage{float}
-%\usepackage{subfig}
-%\usepackage{rotating}
-
-\graphicspath{{pictures/}}
-
-\begin{document}
-
-%\title{The 2018 heatwave and its implications on ozone induced damage on vegetation in a subarctic climate}
-\title{Characterizing subarctic biomes for land surface modeling of pollution and climate risk}
-
-% \Author[affil]{given_name}{surname}
-
-\Author[1]{Stefanie}{Falk}
-\Author[2]{Ane Victoria}{Vollsnes}
-\Author[2]{Aud Else}{Berglen Erikson}
-\Author[3]{Lisa}{Emberson}
-\Author[3]{Connie}{O'Neill}
-\Author[1]{Frode}{Stordal}
-\Author[1]{Terje}{Koren Berntsen}
-
-\affil[1]{Department of Geosciences, University of Oslo, Oslo, Norway}
-\affil[2]{Department of Biosciences, University of Oslo, Oslo, Norway}
-\affil[3]{Department of Environment and Geography, University of York, UK}
-
-%% The [] brackets identify the author with the corresponding affiliation. 1, 2, 3, etc. should be inserted.
-
-%% If an author is deceased, please mark the respective author name(s) with a dagger, e.g. "\Author[2,$\dag$]{Anton}{Aman}", and add a further "\affil[$\dag$]{deceased, 1 July 2019}".
-
-%% If authors contributed equally, please mark the respective author names with an asterisk, e.g.
-"\Author[2,*]{Anton}{Aman}" and "\Author[3,*]{Bradley}{Bman}" and add a further affiliation: "\affil[*]{These authors contributed equally to this work.}".
-
-\correspondence{Stefanie Falk (stefanie.falk@geo.uio.no)}
-
-\runningtitle{Subarctic biomes for land surface modelling}
-
-\runningauthor{Falk et al.}
-
-\received{}
-\pubdiscuss{} %% only important for two-stage journals
-\revised{}
-\accepted{}
-\published{}
-
-%% These dates will be inserted by Copernicus Publications during the typesetting process.
-
-\firstpage{1}
-
-\maketitle
-
-\begin{abstract}
-Here we assess the importance of characterization of key meteorological variables and air pollution profiles on the modeled susceptibility of subarctic vegetation to ozone.
-Changes in climate (e.g.
-temperature, water availability) will manifest in both direct and indirect effects influencing the species' sensitivity to air pollution.
-A likely increase in frequency of persistent heatwaves in the future, will inflict heat related stress while generating favorable conditions for the formation of ambient air ozone, e.g.
-due to wildfires.
-We have monitored ambient air ozone concentrations at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd in Northern Norway in the vegetation period 2018/19.
-Due to data acquisition problems, ozone concentrations for two weeks in July 2018 were missing from our record.
-We present a reconstruction based on probability density function with respect to the Swedish and Finnish atmospheric monitoring sites and compare with available surface ozone products for the region.
-Over all, ozone concentrations did not differ significantly between the two years.
-While temperatures and global irradiance diverged substantially from the multi annual mean precipitation varied only to some extend.
-We investigate the difference in uptake of ozone using the $\mathrm{DO_3SE}$ model, for typical vegetation at the location.
-We find that bespoke parameterization for subarctic vegetation is essential for an accurate estimation of Phototoxic Ozone Dose (POD) and an assessment of damage risk.
-Accurate estimates of the start and length of growing season are also crucial.
-We find that temperature and its derivative vapor pressure deficit are likely the most important climate variables while the variability of ozone concentration within climatologically reasonable limits plays a minor role.
-The use of subarctic parameterization suggests an improvement in risk assessment as POD values are increased in 2018; corroborated by evidence of visible injury in the Svanhovd ozone garden.
-\end{abstract}
-
-%\copyrightstatement{TEXT}
-
-\introduction  %% \introduction[modified heading if necessary]
-\label{sec:intro}
-
-\begin{itemize}
-\item Wealth of evidence demonstrating ozone damage to vegetation (perhaps focus on key evidence from northern climates and key species (e.g.
-birch, Norway srpuce etc...
-- this might help to make it clear why we focus on these species with the PODy modelling))
-\item Although O3 trends stagnant or decreasing in Europe there is unique photochemcistry over the arctic (causing spring time peaks in ozone) that may mean ozone damage continues to be an issue
-\item The earlier timing of growing seasons with climate change and 24 hour sunlight may also enhance the exposure of plants during those times when they are physiologically active to higher [O3]
-\item Research also clearly shows how ecosystem sensitivity to vegetation is influenced by climate variables, both because this can influence ozone uptake (dose) as well as the impact of altered ozone doses (can add in some relevant detail here on ozone damage mechanisms - again perhaps with a focus on what we know for colder climates)
-\item Risk assessment methods have been developed that allow for ozone uptake (i.e. PODy)
-\item Comparison of these methods with previously used concentration based methods (e.g.
-AOT40) show that PODy is better able to represent the risk to vegetation from ozone, especially in the more climate extreme parts of Europe (see Mills et al 2011 paper attached - that also includes some work for clover that we could use)
-\item These methods rely on accurate representation of plant physiological response to climate variables (i.e.
-the species or plant funtional type specific responses used to parameterise risk assessment models)
-\item These models are most often parameterised using data for less extreme climates (i.e.
-temperate or continental climates; and to a lesser extent, boreal and Mediterranean climates.
-Parameterisation of these models for more extreme bio-geographical regions such as sub arctic climates has not been perfromed and relies on assuming that boreal parameterisations will be applicable.
-\item It is important to understand the implications of using these more generic parameterisations for more extreme climate regions, here we explore how risk estimates might be over or underestimated using generic vs bespoke sub-arctic parameterisations
-\item we also explore the role [O3] profile and changing climate variables to understand how this risk might alter in the future with the advent of climate change 
-\item This allows both an understanding of how the threat from ozone may change over time but also which are the key aspects of plant physiology that might determine risk (and hence which parameters might require special attention to get right) both for pollution impact assessment, but also more generally for describing gas exchange response to changing climates that will influence bio-geochemcial cycling (could we add something in here about the paucity of PFTs representing subarctic conditions in biogeochemcial models?).
-\end{itemize}
-
-Ground level ozone is a highly toxic pollutant known to cause damage to a variety of ecosystems around the World \citep{PT:Emberson2020}.
-At continuously high concentration, ozone negatively affects life expectancy of humans \citep{WHO2008} and is a cause of visible injury, phtosynthetic damage, early senescence as well as programmed cell death \citep{PCE:Kangasjarvi2005}.
-Annual global yield losses of four major crops (wheat, rice, maize, and soybean) of $3-15\,\unit{\%}$ \citep{PJ:Ainsworth2017} and a suggested loss in primary production in forestry of $7\,\unit{\%}$ \citep{GCB:Wittig2009,EP:Matyssek2012} has been attributed to ozone.
-The crop yield loss studies indicate a threat to food security in rapidly developing countries, e.g., in East and South-East Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018}, where ozone related to enhanced pollutant emissions from the transport, residential, power generation and industrial sectors have increased in the past few decades.
-At the same time, a successful implementation of air quality regulations coupled with economic restructuring which saw a move from coal to gas, reduced the number of episodes of peak concentrations especially in summertime over Europe and North America \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
-
-With a radiative forcing value of $0.40 \pm 0.20\,\unit{W\,m^{-2}}$, ozone is a potent climate forcer, ranked third behind \chem{CO_2} and \chem{CH_4} \citep[Chapter 8]{IPCC2013}.
-Its damaging effects on the carbon sequestered by global vegetation might provide an important additional feedback to climate change \citep{Nat:Sitch2007}.
-
-In the presence of nitrogen oxides (\chem{NO_x}), tropospheric ozone is formed in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) and hydrocarbons known as volatile organic substances (VOCs).
-In the absence of local pollution sources, episodes of high ozone concentrations are often attributed to long-range transport of precursors.
-In northern Fennoscandia, ozone episodes in 2003 and 2006 have been traced back to ozone precursor emissions related to forest fires in southern and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}.
-Hydrocarbons are also emitted by living vegetation in form of terpenes and monoterpenes, so called biogenic volatile substances (BVOCs).
-These emissions are associated with, e.g., thermal stress, defense against herbivores, or attraction of pollinators \citep{TPS:Penuelas2003}.
-
-Despite a relatively short tropospheric life-time of approximately $22\,\unit{days}$, ranging from a few days in the tropical boundary layer to up to $1\,\unit{year}$ in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}, ozone is subject to advection over long distances.
-In the annual cycle in boreal and temperate climates, a distinct maximum typically occurs in spring followed by lower concentrations throughout the summer.
-This is most likely due to stronger dry deposition on land vegetation than on bare ground or snow and ice \citep{AE:Monks2000} causing a build off of ozone and its precursors in the course of polar night.
-Come spring they are photochemical activated again.
-This spring peak is reached earlier in northern Fennoscandia than in more southern locations but is highly variable throughout the years \citep{AB:Klingberg2009, BER:Klingberg2019}.
-A cause for this is a complex interplay between atmospheric dynamics and chemistry \citep{AE:Laurila1996,BER:Hatakka2003} as well as anthropogenic and biogenic activity \citep{AE:Rummukainen1996,AE:Simpson2002,QJRMS:Galbally2007,NGS:Schnell2009}.
-\citet{ACP:Andersson2017} showed in a modeling study focused on Fennoscandia that the variability in winter can be attributed mainly to changes in atmospheric background (transport) of ozone, while summertime abundance is mostly affected by emissions of precursors in the rest of Europe.
-Meteorology plays an important role throughout the year.
-
-At the same time, the start of growing season in northern Fennoscandia is shifting to earlier days of the year in response to the general warming trend \citep[e.g.,]{GCB:Menzel2006,RS:Hogda2013,IJB:Karlsen2007} and thereby converging with the period of the ozone spring peak.
-This may promote a higher susceptibility of deciduous species to ozone and could also affect coniferous species which will have flushes of new needles in the spring time which are likely, through their immaturity, to be more susceptible to ozone damage.
-
-Even though long-term (pre 1950) observations of surface ozone concentrations are scarce; evidence from trend analysis of those few sites that have monitored concentrations suggests that since the industrial revolution, background concentrations in the northern Hemisphere have at least doubled and continued to increase until relatively recently \citep[Chapter 2]{IPCC2013}.
-Recent trend analyses indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric ozone started to level-off or even decline \citep{ESA:Cooper2014, ACP:Wespes2018,ESA:Gaudel2018}.\\
-
-Ozone acts as an oxidative stress to plants.
-Its main action is imposed through reactions occurring in the cell walls and cell membranes of mesophyll cells inside the leaves.
-Ozone enters the leaves through stomata, that are kept open to allow for gas exchange allowing the entry of CO2 for photosynthesis and loss of H2O vapour via the plant transpiration stream.
-Stomatal aperture, and hence stomatal conductance to these gases (including ozone) will vary over the course of the day and growing season primarily to balance CO2 uptake against H2O vapour  loss.
-The higher the stomatal conductance, the higher the potential for ozone uptake.\\
-
-According to a report by the Norwegian Meteorological institute \citep{MetNOR2019}, the summer of 2018 was the warmest and driest ever recorded in eastern, western, and southern Norway.
-In the north (including Finnmark), it was amongst the warmest on record -- favorable conditions for planetary boundary layer ozone formation.
-
-An unusually weak and northward shifted jet stream allowed for a persistent high pressure system above northern Europe, including Fennoscandia which blocked the pressure systems for several consecutive months.
-In the period May -- July, southern Norway had temperatures $4\,\unit{^\circ C}$ above normal.
-While southern Norway had only about $60\,\unit{\%}$ of normal precipitation, northern Norway as a whole had close to normal precipitation, but with local variations.
-
-Heat and drought stress on vegetation was exceptional not only in large parts of Fennoscandia but also in much of Europe, where the influence of the high pressure system extended even over a period of five months (April/May, July--September).
-These conditions gave rise to massive forest fires in different parts of Europe and thus an increase in ozone precursors.
-Boreal wildfires emit in addition to \chem{CO_2} also \chem{CO} ($\chem{[CO]}/\chem{[CO_2]} \propto 6-13\,\unit{\%}$) and \chem{VOCs} ($\chem{[VOCs]}/\chem{[CO_2]} \propto 0.5-1.5\,\unit{\%}$) \citep{AE:Cofer1990}.
-
-A total of 2079 forest fires of different size were registered in Norway in 2018, twice as many as in the preceding years 2016/17 \citep{DSB2019}, last accessed April 2020).
-In Sweden, about 500 fires had been reported (five times more than in a usual summer) and an estimated total of $25000\,\unit{hectare}$ burned down in the course of the most extensive fires in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) in younger history \citep{SOU2019}.
-Coincident peak \chem{[O_3]} are found in ozone monitoring data from Svanhovd but also over whole northern Fennoscandia in July (see Fig.~\ref{fig:data_svanvik_2018_2019} or Fig.~\ref{fig:ozone_reconstruction_2018_07}).
-
-Like in the 2003 drought period, elevated ozone in Europe was likely promoted by a combination of various other factors as well, such as biogenic VOC emissions, reduced cloud and hence increased solar radiation, faster photochemical reactions, and reduced dry deposition and turbulent mixing due to stagnant weather conditions \citep{JGR:Solberg2018}.\\
-
-To measure the risk of damage inflicted to vegetation through ozone, many different metrics have been defined.
-Two main types can be differentiated, concentration-based or flux-based metrics, e.g., Accumulated Over a Threshold $\mathrm{x}$ ($\mathrm{AOT_x}$) and Phyto-toxic Ozone Dose over a threshold $\mathrm{y}$ ($\mathrm{POD_y}$), respectively.
-Due to their simplicity, concentration-based metrics are still widely used, despite the actual integrated flux of ozone entering through the stomata being more relevant for an estimation of the phyto-toxic ozone dose \citep{GCB:Mills2010,ICP:MappingManual2017}.
-
-The $\mathrm{AOT_x}$ in units of $\unit{ppb}$ is defined as:
-\begin{equation}
-  \mathrm{AOT_x} = \int{\mathrm{max(\chem{[O_3]}-x)}\cdot \mathrm{dt}},
-  \label{eq:aot}
-\end{equation}
-with the ambient ozone concentration $\chem{[O_3]}$. The  time over which the integration is calculated varies, but is usually defined from $8\,\unit{am}-8\,\unit{pm}$.
-The $\mathrm{POD_y}$ is calculated by first estimating the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf.
-The terms $r_c$ (leaf surface resistance) and $r_b$ (quasi-laminar resistance) allow for the deposition of \chem{O_3} to the leaf and the fraction that is taken up via the stomata.
-%
-\begin{equation}
-  \Phi_\mathrm{sto} = \chem{[O_3]} \cdot \frac{u(z_1) \cdot g_\mathrm{sto} \cdot r_c}{r_b + r_c}.
-  \label{eq:flux_stomata}
-\end{equation}
-%
-The quasi-laminar boundary layer resistance is calculated by
-%
-\begin{equation}
-  r_b = 1.3 \cdot 150 \cdot \sqrt{\frac{L}{u(z_1)}}.
-  \label{eq:quasi_laminar_resist}
-\end{equation}
-%
-Where $L$ is the cross-wind leaf dimension, $u(z_1)$ the wind speed at height $z_1$, and the factor $1.3$ accounts for the diffusivities between heat and \chem{O_3}.
-$\mathrm{POD_y}$ is then calculated according to:
-%
-\begin{equation}
-  \mathrm{POD_y} = \int{(\Phi_\mathrm{sto}-y)\cdot \mathrm{dt}},
-  \label{eq:pod}
-\end{equation}
-%
-with the hourly averaged ambient air ozone flux through the stomata Eq.~(\ref{eq:flux_stomata}) and a flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}\,s^{-1}}.
-The flux threshold $\mathrm{y}$ is representing the detoxification potential of the plant and is typically only exceeded during daylight hours (i.e.
-when global radiation is above $50\,\unit{W\,m^{-2}}$).
-
-Finally, exceedance of the flux-based critical level (CL) is calculated by:
-%
-\begin{equation}
-  \mathrm{CL_{exeed}} = \mathrm{POD_y} - \mathrm{CL}.
-\end{equation}
-%
-
-In terms of $\mathrm{AOT_{40}}$, the threshold for protection of vegetation of $3000\,\unit{ppb}$ (integrated over the growing season) given by the EU directive has been well exceeded at many ozone monitoring sites in Southern Norway in the 2018 growing season \citep{NILU2019}.
-{\bf TODO: might be interesting to look at Mills et al 2003 and the different spatial extents of risk estimated when using AOT40 vs PODy.
-That could be useful to explain why we focus on PODy (rather then AOT40).- Lisa Come back to this later on in the discussions(?) - it has not been breached at Svanvik in northern Fennoscandia.}
-
-Within the scope of our project to assess probable effects of midnight sun conditions on ozone uptake by plants and resultant enhanced susceptibility to ozone inflicted damage, we conducted ambient air ozone concentration measurements at Svanhovd in the growing seasons of 2018/19 in collaboration with the Norwegian Institute for Air Research (NILU).
-In 2018, we have observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd.
-In contrast, no such damage was found on the clovers in 2019, although it was found on the sensitive tobacco cultivar.
-This indicates that the vegetation may have been more affected by ozone in 2018 than in 2019.
-
-As previously mentioned, ozone damage is rather driven by the actual uptake of ozone ($\mathrm{POD_y}$) than by the accumulated ambient air concentration above a certain threshold ($\mathrm{AOT_x}$).
-Through the stomatal conductance, $\mathrm{POD_y}$ depends on environmental drivers such as ambient air temperature, photoactive radiation, soil water availability, and water vapor pressure deficit.
-
-In this study, we investigate potential main drivers for the observed ozone damage in the growing season 2018 at NIBIO Environment Centre Svanhovd in contrast to 2019.
-We compare $\mathrm{POD_y}$ values for natural and semi-natural vegetation for derived climatological averages of environmental conditions at Svanhovd as well as for the growing seasons 2018/19 which shall serve as case studies.
-
-The observation site at NIBIO Environment Centre Svanhovd and all related data are presented in Section~\ref{sec:data}.
-In Section~\ref{sec:stats}, we derive climatologies for surface ozone concentration at Svanhovd and northern Fennoscandia based on in situ observations and also compute climatologies of surface temperature, precipitation, and global radiation for Svanhovd.
-Based on these, we evaluate anomalies of environmental key variables in 2018/19 by statistical means.
-In addition, the derived surface ozone concentration climatology for northern Fennoscandia will be compared with available ozone reanalysis products.
-We use the $\mathrm{DO_3SE}$ model to estimate the ozone uptake by natural and semi-natural vegetation to investigate main environmental drivers for the observed ozone damage in 2018 (Section~\ref{sec:do3se}).
-In this context, we biefely introduce the main $\mathrm{DO_3SE}$ model equations and derive bespoke parameterizations for subarctic species.
-In Section~\ref{sec:conc} finally, we will summarize our results, draw conclusions, and give an outlook on further work.
-
-\section{Data}
-\label{sec:data}
-
-In our assessment of ozone damage on vegetation in the subarctic in relation to the 2018 heatwave, we focus on the NIBIO operated Environment Centre Svanhovd which lies close by the settlement of Svanvik in the Pasvik valley (Section~\ref{subsec:svanhovd}).
-Throughout the paper, we will use Svanhovd synonymously for the atmospheric monitoring site at Svanvik.
-The available meteorological monitoring data are described in Section~\ref{subsubsec:atmo_svanvik}.
-In Section~\ref{subsubsec:ozone_garden}, the ozone garden -- a plant-based ozone monitoring -- will be described in more detail.
-Other data used for evaluation of the ozone data taken at Svanhovd is described in Section~\ref{subsec:other_data}.
-We compare with observational data from other monitoring sites in northern Fennoscandia (Section~\ref{subsubsec:ebas}) and use as well reanalysis products (Section~\ref{subsubsec:ozone_rea}) as complement when observational data was not available.
-
-\subsection{NIBIO Environment Centre Svanhovd}
-\label{subsec:svanhovd}
-"Svanhovd is located beside the Pasvik river, in the middle of the wedge of Norwegian land separating Russia and Finland in the north.
-It is where the eastern Siberian taiga meets the western Boreal forest.
-[...] Svanhovd was originally established as a State Demonstration and Experimental Farm in the 1930's, in order to expand agriculture in Pasvik valley [...]" \citep[Svanhovd][]{NIBIO_Svanhovd}.
-
-The locations of the atmospheric monitoring site as well as the ozone garden with respect to the main buildings of the Svanhovd Environment Centre are marked in the aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}.
-The Centre is located south of the settlement of Svanvik, where as the border to Russia, the Pasvik river, lies only a few hundred meters to the east.
-The original farm buildings, protected by law, are located in the lower center, while the Svanhovd Environment Centre follows to the north-east.
-Left of the Centre and north of the original farm buildings a botanical garden can be found.
-It is one amongst three such gardens north of the Polar Circle.
-The Svanhovd Environment Centre focuses on research of the special environment of the region, but also comprises long-term observations of various agrometeorological variables, i.a.
-$2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths.
-
-For the purpose of relating observed damage on ozone sensitive plants (e.g.
-clover and tobacco) to local ozone concentrations, an ozone monitor had been installed during the growing seasons 2018/19.
-The relevant data are described in more detail in Section~\ref{subsubsec:atmo_svanvik}.
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{svanhovd_researchstation}
-  \caption{NIBIO Environment Centre Svanhovd close by the settlement of Svanvik, Norway. Atmospheric monitoring site and ozone garden have been marked. Aerial photography \copyright Norges Kartverk.}
-  \label{fig:svanhovd_research_station}
-\end{figure}
-
-\subsubsection{Atmospheric monitoring}
-\label{subsubsec:atmo_svanvik}
-For the growing seasons 2018/19, monitoring data of key variables (temperature, pressure, precipitation, and ozone) have been obtained from luftkvalitet.no which is operated by \citet{NILU_AIRQ}.
-
-All agrometeorological variables including temperature and precipitation are available from September 1992 to present day \citep[LandbruksMeteorologiske Tjeneste][note the station name here is Pasvik]{LMT_NIBIO}.
-
-Relevant data for the growing season 2018/19 are shown in Fig.~\ref{fig:data_svanvik_2018_2019}.
-The hatched areas mark times when no ozone data was recorded.
-Note, while the downtime during winter was planned, missing data in two weeks of July 2018 (July 9--23) were due to problems in data acquisition.
-
-Ozone concentrations \chem{[O_3]} measured in $2\,\unit{m}$ height above ground are averaged hourly.
-As can be seen in Fig.~\ref{fig:data_svanvik_2018_2019}a), \chem{[O_3]} peaks in spring (April/May) and reaches its minimum in late summer (July/August).
-The spring peak has not been captured completely in 2019, for data acquisition started later than in 2018.
-In summer 2018 (June--August), high ozone concentrations ($\chem{[O_3]} > 40\,\unit{ppb}$) were recorded 50 times.
-The highest summer ozone concentration ($\chem{[O_3]} = 50.2\,\unit{ppb}$) was measured on July 25.
-This coincides with the period of the most extensive forest fires in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) which occurred from July 12--29 and destroyed about $18000\,\unit{hectare}$ of forest \citep{SOU2019}.
-However, due to the above mentioned data acquisition problems, we missed most of the corresponding ozone data for this event and most likely also the peak ozone concentration.
-In contrast, ozone concentration only rose 18 times above the threshold of $40\,\unit{ppb}$ during the summer of 2019.
-
-Hourly averaged $2\,\unit{m}$ temperatures below $-20\,\unit{^\circ C}$ were observed in winter (January/February), while high temperatures (above $20\,\unit{^\circ C}$) occurred in summer (July), though more regularly in 2018 than in 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}b).
-In 2018, temperature regularly rose above freezing only in May, while in 2019 this occurred already early in March/April.
-2019 saw a very cold November with temperatures regularly below $-10\unit{^\circ C}$.
-
-More rain events with accumulated daily precipitation $\sum_d \mathrm{Precip}$ above $10\,\unit{mm}$ occurred in the summer of 2018 compared to 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}c).
-$\sum_d \mathrm{Precip} \ge 20\,\unit{mm}$ was only breached twice in 2018 and once in 2019.
-
-In 2018, global irradiance $Q_0$ (Fig.~\ref{fig:data_svanvik_2018_2019}d) displays more sequentially high values in spring (May) and summer (July) compared to 2019, while June 2019 saw more consecutive high irradiance than 2018.
-In both years, the maximum recorded $Q_0$ was roughly $750\,\unit{W\,m^{-2}}$ reached only in June.
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{data_svanvik_2018_2019}
-  \caption{Observational data from atmospheric monitoring at Svanhovd in 2018/19.
-The hatched areas indicate missing ozone monitoring data.
-(a) Hourly averaged ozone concentration; (b) hourly averaged temperature; (c) daily accumulated precipitation; (d) hourly averaged global irradiance.}
-  \label{fig:data_svanvik_2018_2019}
-\end{figure*}
-
-\subsubsection{Ozone garden}
-\label{subsubsec:ozone_garden}
-An ozone garden consists of carefully selected plant species of which one cultivar is more sensitive to ozone than the other.
-At the NIBIO Environment Centre Svanhovd, the likely northern-most ozone garden to date, an ozone garden was installed in the course of our project.
-Cultivated species were, e.g.
-clover, tobacco, and beans.
-As shown in Fig.~\ref{fig:ozone_garden}, the plants had to be protected from herbivores with a wire-mesh fence.
-In 2018, we have observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden.
-In contrast, no such damage was found on the clovers in 2019, although it was found on the sensitive tobacco cultivar.
-This indicates that the vegetation may have been more affected by ozone in 2018 than in 2019.
-
-\begin{figure}
-  \includegraphics[width=8.3cm]{IMG_7739.JPG}
-  \caption{Clover in the Svanhovd ozone garden. The plants had to be secured against herbivores with a wire-mesh fence. The plants shown are approximately $6-12\,\unit{cm}$ in height.}
-  \label{fig:ozone_garden}
-\end{figure}
-
-\subsection{Other data}
-\label{subsec:other_data}
-To determine the significance of the ozone conditions in 2018, we have to derive a surface ozone concentration climatology for Svanvik.
-From 1986--1996 ozone monitoring has been operated by NILU at Svanhovd, but no ozone measurements have been conducted for the past $14\,\unit{years}$.
-Due to probable changes in northern hemispheric tropospheric background ozone and a reduction of summertime peak values related to ozone precursors which fell under air quality regulations in Europe, we also have to have a look at ozone observations at other sites in northern Fennoscandia.
-In regard to ozone, available observational data are sparse in this region.
-An overview over the geographical locations of conducted atmospheric monitoring in the past and present is given in Fig.~\ref{fig:station_map_fennoscandia}.
-The data is described in Section~\ref{subsubsec:ebas}.
-As additional complement to observational data which is not available at all times and spatially sparse, we also look at various ozone reanalysis products (Section~\ref{subsubsec:ozone_rea}).
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{station_map_fennoscandia}
-  \caption{Cap of the North.
-Locations of past and present ozone observation sites in northern Fennoscandia used within this study.
-For more details see Tab.~\ref{tab:ebas_obs}.
-The same color coding for each station as displayed herein will be used in all following figures.}
-  \label{fig:station_map_fennoscandia}
-\end{figure}
-
-\subsubsection{Surface ozone observations}
-\label{subsubsec:ebas}
-All surface ozone observations used in this study are available from the EBAS database operated by \citet{NILU_EBAS}.
-In Table~\ref{tab:ebas_obs}, we give an overview over the exact locations and time range of operational data taking.
-All stations with available long term observations are located on higher ground than Svanhovd which will result in higher surface ozone concentrations in comparison as ozone abundance, in general, decreases with decreasing altitude in the absence of local precursors \citep[e.g.,][]{AB:Klingberg2009}.
-The periphery vegetation at all sites is similar and consists mainly of pine forests, birch, and heath.
-However, at close vicinity it varies substantially.
-Ozone measurements at Pallas are taken on top of mount Sammaltunturi $100\,\unit{m}$ above tree line (pine forest) and mainly consists of low vascular plants, mosses, and lichen \citep{BER:Hatakka2003}.
-At Svanhovd, the ozone measurements are taken amidst a meadow surrounded by agriculturally managed herbaceous plants and grass (c.c.
-Fig.\ref{fig:svanhovd_research_station}).
-
-Data from before the 1990s is available solely from Svanvik and Karasjok.
-These data, however, do not follow the high quality standard procedures implemented nowadays and have to be treated with care \citep{NILU2003}.
-Especially, ozone monitors did not undergo a regular re-calibration.
-This probably leads to drifts in the observed data and may impose a false periodicity.
-In 1997, the ozone monitoring at Jergul was relocated downstream the river Karasjokka to a site close by Karasjok.
-Although the same equipment has been used, this might have introduced a breaking point in the combined data series, denoted as Jergul/Karasjok in the following, which would be of a problem if this was the sole data to compare with.
-Only few data have been taken upstream the Pasvik river at Janiskovski, hence these will not be taken into further consideration, but are shown for completion.
-As can be seen in Fig.~\ref{fig:ozone_timesseries_fenoscandic_obs}, there is an overlap of about $4\,\unit{years}$ between data taken at Esrange, Jergul/Karasjok, and Svanvik.
-The hatched areas indicate time periods with insufficient quality control as mentioned above.
-
-\begin{table}[t]
-  \caption{Past and present ozone observation sites in northern Fennoscandia. Data available on \href{http://ebas.nilu.no/}{EBAS}.}
-  \label{tab:ebas_obs}
-  \begin{tabular}{lllllrl}
-    \tophline
-    Name       & Country & ID      & \multicolumn{3}{c}{Location} & Operational\\
-    &         &         & lat             & lon               & alt            &\\
-    &         &         & (\unit{^\circ N}) & (\unit{^\circ E})  & (\unit{m})     &\\
-    \middlehline
-    Esrange    & SWE     & SE0013R & 67.83           & 21.07             & 475            & 1991 -- present\\
-    Janiskoski & RUS     & RU0001R & 68.93           & 28.85             & 118            & 1995 -- 1997\\
-    Jergul     & NOR     & NO0030R & 69.45           & 24.60             & 255            & 1997 -- 2011\\
-    Karasjok   & NOR     & NO0055R & 69.467          & 25.217            & 333            & 1988 -- 1997\\
-    Pallas     & FIN     & FI0096G & 67.97           & 24.12             & 565            & 1995 -- present\\
-    Svanvik    & NOR     & NO0047R & 69.45           & 30.03             & 30             & 1986 -- 1996$^\dagger$\\
-    \bottomhline
-  \end{tabular}
-  \belowtable{$^\dagger$ Exclusive ozone monitoring growing seasons 2018/19 operated by NILU.} % Table Footnotes
-\end{table}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{ozone_timeseries_fenoscandic_obs.png}
-  \caption{Time series of ozone observations in northern Fennoscandia (Tab.~\ref{tab:ebas_obs}).
-Data taken from EBAS.
-The hatched areas indicate time periods with insufficient quality control \citep{NILU2003}.}
-  \label{fig:ozone_timesseries_fenoscandic_obs}
-\end{figure*}
-
-\subsubsection{Ozone reanalysis}
-\label{subsubsec:ozone_rea}
-
-Details of reanalysis data sets referred to in this work are listed in Table~\ref{tab:ozone_rea}.
-There are two global reanalysis products available at the European Centre for Medium-Range Weather Forecasts (ECMWF) that include atmospheric tracers, in particular ozone (MACC and CAMSRA) \citep{ACP:Inness2013, ACP:Inness2019}.
-The temporal as well as spatial resolution of these reanalysis products is rather coarse: 3-hourly and $0.75^\circ\times 0.75^\circ$ or roughly $29.3\,\unit{km}\times 83.4\,\unit{km}$ at the location of Svanvik, respectively.
-In addition, higher resolution surface ozone reanalysis for whole Europe is available from Copernicus service for regional air quality control (note: since May 31st 2020 through the CAMS Atmosphere Data Store (ADS)).
-The regional air quality model reanalysis ensemble (CAMS regional Ensemble) is based on nine European state-of-the-art numerical air quality models \citep{Copernicus_Regio_AIRQ}.
-The ensemble mean is at fairly high spatial and temporal resolution compared to the global reanalysis from ECMWF: $0.1^\circ\times 0.1^\circ$ or roughly $3.9\,\unit{km}\times 11.1\,\unit{km}$ and 1-hourly, respectively.
-The covered time periods differ and no data are available for before the turn of the millennium.
-The CAMSRA and CAMS Regional Air Quality ensemble mean are available in near real time, but only former covers a period of sufficient length for climate analysis (2003 -- present).
-All reanalyses use different but, at that time, latest versions of the operational weather forecast system (OpenIFS) from ECMWF.
-They differ substantially in the assimilated observational ozone data.
-The MACC reanalysis assimilates only satellite derived tropospheric column ozone (TC), while CAMSRA also includes ozone profiles from satellite retrievals.
-Most importantly for our study, in situ observations from ozone surface station networks are only assimilated in the CAMS Regional Ensemble.
-The quality with respect to observations in northern Fennoscandia will be discussed in Section~\ref{subsubsec:clim_ozone}.
-Without too much foreshadowing, the results prevented us from fetching the whole CAMSRA data.
-
-\begin{table}[t]
-  \caption{Ozone related reanalysis products used in this study: \href{https://www.ecmwf.int/en/newsletter/158/meteorology/new-cams-global-reanalysis-atmospheric-composition}{Global reanalysis (ECMWF)}; \href{https://www.regional.atmosphere.copernicus.eu/}{ensemble mean of regional reanalysis for Europe (Copernicus)}.}
-  \label{tab:ozone_rea}
-\begin{tabular}{llcclcc}
-\tophline
-Name & Provider & \multicolumn{2}{c}{Resolution} & Time period & Meteorological forcing & \chem{O_3} assimilation\\
-&              & spatial & temporal & & \\
-\middlehline
-MACC & ECMWF & $0.75^\circ \times 0.75^\circ$ & 3-hourly & 2003 -- 2012 & OPS & satellite $^\triangledown$\\
-CAMSRA & ECMWF & $0.75^\circ \times 0.75^\circ$ & 3-hourly & 2003 -- 2012 $^\dagger$ & ERA5 / OPS $^\ddagger$ & satellite $^\blacktriangledown$\\
-CAMS Regional Ensemble & Copernicus & $0.1^\circ \times 0.1^\circ$ & 1-hourly & 2014 -- 2018 $^\dagger$ & OPS $^\star$ & in situ $^\vartriangle$\\
-\bottomhline
-\end{tabular}
-\belowtable{$^\dagger$ Subset of the data used in this study.
-The full data set extends into near real time; $^\ddagger$ ERA5 (2003--2016), OPS (later); $^\star$ EURAD uses WRF for downscaling of operational IFS; $^\triangledown$ MLS, OMI - tropospheric column; $^\blacktriangledown$ SCIAMACHY, MIPAS, MLS, OMI, GOME2, SBUV2 - tropospheric column + profile; $^\vartriangle$ METEO France NRT.} % Table Footnotes
-\end{table}
-
-\section{Statistical analysis of environmental key variables}
-\label{sec:stats}
-In this section, we will evaluate atmospheric conditions in the growing seasons of 2018/19 by statistical means.
-We will derive climatologies for both, temperature, precipitation, global irradiance, and ozone for Svanvik (Section~\ref{subsec:climatologies}).
-The ozone measurements will be put into a climatological as well as regional context.
-We will look at anomalies of the atmospheric key variable mentioned above (Section~\ref{subsec:anomalies}).
-
-\subsection{Derived climatologies}
-\label{subsec:climatologies}
-
-\subsubsection{Temperature, precipitation, and global irradiance}
-\label{subsubsec:clim_temp_prec}
-The location of Svanvik at $69.45\,\unit{^\circ N}$ suggests a subarctic climate.
-The derived climate diagram following Walter and Lieth shown in Fig.~\ref{fig:climatediagram} supports this very well.
-The diagram is based on climatological data from Svanvik (1992--2012) downloaded from LandbruksMeteorologiske Tjeneste.
-Monthly averaged temperatures (red line) are displayed with standard error of mean as error band.
-Averaged monthly accumulated precipitation (blue bars) is shown with standard deviation of mean as error bars.
-Months with average temperatures below freezing are denoted with a star.
-As can be seen, temperatures stay below freezing for 5 consecutive months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for K\"{o}ppens climate classification of a regular subarctic climate (Dfc).
-The highest monthly average temperature is $(13.1\pm 1.1)\,\unit{^\circ C}$ in July and the lowest $(-12.8\pm 2.0)\,\unit{^\circ C}$ in February.
-In the time period 1992--2012, the coldest measured temperature was $-45.2\,\unit{^\circ C}$ observed January 27, 1999, while the highest temperature on July 16 the same year was $29.4\,\unit{^\circ C}$.
-
-Winter and spring (November--April, except for March) are relatively dry ($\sum_m \mathrm{Precip} < 20\,\unit{mm}$).
-The driest month is January with $\sum_m \mathrm{Precip} = (16.7\pm 3.0)\,\unit{mm}$.
-The most precipitation occurs in the summer months, with a highest monthly averaged accumulated precipitation $\sum_m \mathrm{Precip} = (58.5\pm 9.2)\,\unit{mm}$ in August.
-The average annual accumulated precipitation given with standard error of mean is $(383\pm 86)\,\unit{mm}$.
-Precipitation shows a large variability throughout the years.
-
-A climatology derived for global irradiance $Q_0$ is shown in Appendix Fig.~\ref{fig:global_rad_clim}a) for both, climatological average of hourly maximum $\left<Q_0^\mathrm{max}\right>$, minimum $\left<Q_0^\mathrm{min}\right>$, and mean $\left<Q_0^\mathrm{mean}\right>$ irradiance observed at Svanvik.
-In agreement with the geometrically derived sunshine duration shown in Appendix Fig.~\ref{fig:sunlight_fennoscandia}, $\left<Q_0^\mathrm{max}\right>$ approaches zero in December/January commonly known as polar night and reaches its maximum in June/July commonly referred to as midnight sun conditions.
-The highest observed $\left<Q_0^\mathrm{max}\right>$ is about $800\,\unit{W\,m^{-2}}$, while on average $\left<Q_0^\mathrm{mean}\right>$ it is about $450\,\unit{W\,m^{-2}}$ (regularly reached in May--July).
-The maximum of the climatological minimum $\left<Q_0^\mathrm{min}\right>$ lies at about $200\,\unit{W\,m^{-2}}$.
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{climatediagram}
-  \caption{Climate diagram following Walter and Lieth.
-The diagram is based on climatological data for Svanvik/Pasvik (1992--2012) downloaded from LandbruksMeteorologiske Tjeneste.
-Monthly averaged temperatures (red line) are displayed with standard error of mean as error band.
-Averaged monthly accumulated precipitation (blue bars) is shown with standard deviation of mean as error bars.
-Months with average temperatures below freezing are denoted with a star.}
-  \label{fig:climatediagram}
-\end{figure}
-
-\subsubsection{Ozone}
-\label{subsubsec:clim_ozone}
-In the following, we will look at derived ozone climatologies from different perspectives.
-The multi-annual means of daily mean ozone, for simplicity here without standard deviations, at observation sites in northern Fennoscandia (refer to Table~\ref{tab:ebas_obs} for details) are shown in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs}.
-All climatologies peak in spring, with peak values of about $48\,\unit{ppb}$ in late April.
-The annual average ozone concentration \chem{\left<[O_3]\right>} declines throughout the growing season until it reaches its minimum in late summer/fall ($24-30\,\unit{ppb}$).
-The turning points lie approximately in June and November.
-The ozone climatology lags about 3 months behind the temperature climatology.
-At Svanvik, \chem{\left<[O_3]\right>} is $6.6\,\unit{ppb}$ lower than on sites at Jergul/Karasjok (NOR), Esrange (SWE), and Pallas (FIN).
-The large spread in case of Svanvik and Janiskoski can be explained due to the lower overall statistics.
-
 \begin{figure}[t]
   \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs}
-  \caption{Multi-annual mean of daily mean ozone at observation sites in Fennoscandia.
+  \caption{S1 Multi-annual mean of daily mean ozone at observation sites in Fennoscandia.
 The large spread in case of Svanvik and Janiskoski is due to the lower statistics.
@@
 \end{figure}
-
-The correlation of \chem{[O_3]} between Esrange, Pallas, and Jergul/Karasjok is fairly high ($r^2\approx 0.8$, Appendix Fig.~\ref{fig:density_distribution}a, b), hence we can put those data series together to derive a general ozone climatology for northern Fennoscandia.
-The correlation between Svanvik and Esrange and Pallas is rather low, 0.4 and 0.6, respectively.
-Thus, Svanvik cannot directly be represented by any measurements taken at either Esrange or Pallas.
-We shall come back to this in Section~\ref{subsec:ozone_reco}.
-
-In Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}, the 2 dimensional density distributions for the thus defined northern Fennoscandia (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}a) and Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) are shown.
-Darker colors indicate higher probability to observe these values.
-On top of the density distributions, $10\,\unit{days}$ averages of daily mean \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, daily maximum \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}}, and daily minimum \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, min}} are displayed together with the $1 \sigma$ uncertainties, respectively.
-In addition, \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} is also shown with standard error of mean (albeit small and hardly visible).
-Splines have been fitted through the data for easier inter-comparison with reanalysis data.
-
-From the northern Fennoscandic climatology, one can deduce several features.
-The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, min}} peaks about $20\,\unit{days}$ earlier than the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, and \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}}.
-With respect to the periodicity of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, it takes about $7.5\,\unit{months}$ from minimum concentration to maximum concentration, while from maximum to minimum it only takes about $4.5\,\unit{months}$.
-The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}} is at $52\,\unit{ppb}$ in late April and the minimum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, min}} occurs at $15\,\unit{ppb}$ in the beginning of September.
-The steep rise of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} in spring is in coincidence with increasing temperatures as well as increase in global irradiance.
-The decrease in \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} begins about one month ahead of the estimated onset of the greening season.
-Based on data from SeNorge.no and the agricultural rule of thumb that the daily average temperature has to be above $5\,\unit{^\circ C}$ for $5\,\unit{days}$ consecutively, we find $\left<G_\mathrm{begin}\right> = (150 \pm 14)\,\unit{days of the year}$ at Svanvik (Appendix Fig~\ref{fig:greening_season_change_Svanvik}), which translates into the end of May.
-This finding is in line with more extensive studies \citep[][e.g.,]{IJB:Karlsen2007,AFM:Linderholm}.
-
-The density distribution for Svanvik and the respective \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, and \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}} (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) display a similar pattern as the northern Fennoscandic climatology.
-The spring peak apparently occurs slightly earlier (in the beginning of April).
-As Svanvik is located at lower altitude than the other sites, the earlier ozone peak may point to an interplay with ozone uptake by vegetation.
-{\bf TODO: The data from SeNorge seems to point to a later onset of greening, than what I interfered from OpenIFS data.
-This means that the beginning of the growing season is not a good measure for the termination of the ozone spring peak.
-This does not exclude the early start of pine forest photosynthesis as a driver.
-Coniferous forest starts to photosynthesis earlier than what the growing season begin would mark.
-\chem{CO_2} uptake and phenology of coniferous trees \citep{TB:Kolari2007} starts as early as doy 100, acclimates to leaf temperature history - photosynthesis drops at around doy 200--250.} The average ozone concentration at Svanvik is lower than at the other sites, as already mentioned, which is in league with a decrease in tropospheric ozone with decreasing altitude.
-In July--September, ozone is occasionally almost completely depleted during night time.
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_spline}
-  \caption{Derivation of ozone climatologies.
-Density distributions are shown together with multi-annual mean of daily ozone concentrations \chem{[O_3]}.
-Splines have been fitted though daily mean and max \chem{[O_3]}.
-The climatology of daily mean ozone is shown together with $1\,\sigma$ standard deviation and standard error of the mean.
-Since \chem{[O_3]} are strongly correlated for sites at Jergul/Karasjok (NOR), Esrange (SWE), and Pallas (FIN) (see Fig.~\ref{fig:density_distribution}), data of these have been averaged together to derive a climatology for northern Fennoscandia.
-(a) Northern Fennoscandia; (b) Svanvik.}
-  \label{fig:ozone_climatology_fenoscandic_obs_spline}
-\end{figure}
-
-Finally, we shall look at the comparison of climatologies derived from observational ozone data and the various reanalysis products (Fig.~\ref{fig:ozone_climatologies}).
-As representative for the sites, the nearest neighboring grid point has been chosen.
-The ECMWF products do not reproduce observed surface ozone well.
-The MACC reanalysis (Fig.~\ref{fig:ozone_climatologies}a) is in general too low ($(9\pm 7)\,\unit{ppb}$) and displays an erroneous seasonal cycle.
-Ozone concentrations peak already in March followed by a second peak in July.
-Also CAMSRA matches the observed ozone climatology poorly (Fig.~\ref{fig:ozone_climatologies}b).
-Despite displaying only one ozone peak, it still does not reproduce the seasonality in northern Fennoscandia well.
-The CAMSRA spring peak appears to lag behind by $1\,\unit{month}$ and is $5\,\unit{ppb}$ lower, whereas the minimum occurs in January compared to August/September.
-The annual amplitude ($(26\pm 1)\,\unit{ppb}$) is larger than in the climatology derived from observations ($19\,\unit{ppb}$).
-Modeled \chem{[O_3]} at Svanvik are even slightly higher than at Pallas.
-In contrast, the ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in northern Fennoscandia quite well (Fig.~\ref{fig:ozone_climatologies}c), but is on average still slightly too low ($(2.8\pm 0.5)\,\unit{ppb}$).
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_climatologies_portrait}
-  \caption{Comparison of climatologies derived from observational data (EBAS, see Fig.~\ref{fig:ozone_climatology_fenoscandic_obs}) with various ozone reanalysis products: (a) ECMWF: MACC reanalysis; (b) ECMWF: CAMS reanalysis; (c) Copernicus: Regional model reanalysis ensemble mean.
-As representative for the sites, the nearest neighboring grid point has been chosen.
-The ECMWF products do not reproduce observed surface ozone well.
-The MACC reanalysis is in general too low and displays an erroneous seasonal cycle.
-The CAMS reanalysis matches better but does not reproduce the seasonality in northern Fennoscandia.
-The ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in northern Fennoscandia, though the modeled ozone concentrations are slightly too low.}
-  \label{fig:ozone_climatologies}
-\end{figure}
-
-There are many probable explanations for the discovered differences in the climatologies derived from the three reanalyses.
-The effective changes from the MACC reanalysis to CAMSRA have been reported and discussed by \citet{ACP:Inness2019} on a global level.
-Amongst others, an assimilation of ozone profiles from satellite retrieval and an upgraded ozone chemistry scheme have lead to the enhanced performance of latter.
-The higher spatial resolution of the regional air quality models used in the CAMS Regional Ensemble compared to the global reanalyses does capture more localized depletion and peak episodes in \chem{[O_3]} better and enhances the capability of the models with respect to daily and seasonal cycles in ozone.
-Latter is, of course, also affected by upgrades in the OpenIFS used as meteorological forcing.
-However, we can assume that the main driver for the strikingly different results are the assimilated observational ozone data.
-Satellite retrievals resolve \chem{[O_3]} at the surface rather poorly and hence do not constrain the global models well enough \citet{ACP:Andersson2017}.
-
-\subsection{2018/19 anomalies}
-\label{subsec:anomalies}
-Using these established climatologies, we look at the anomalies in 2018/19 and discuss the differences between the two years.
-
-\subsubsection{Temperature, precipitation, and global irradiance}
-\label{subsubsec:anomal_tpq}
-In Fig.~\ref{fig:plot_temperature_anomalies_svanvik}, temperature anomalies at Svanvik are shown as percentage of days warmer/colder than $\pm 1\,\sigma$ from the climatological mean for each month.
-Negative deviations from climatology are displayed as negative percentage.
-The full annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
-The summer of 2018 was significantly warmer than average.
-Especially in May and July more than $40\,\unit{\%}$ of days were significantly warmer than normal temperature.
-Temperatures continued to be significantly warmer throughout fall into early winter (August--November) with the exception of October.
-March 2018 was unusually cold.
-The summer of 2019 was fairly average, however significantly colder in July, while August was both significantly colder and warmer.
-Temperatures in April were significantly warmer than average.
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{temperature_signific_1}
-  \caption{Temperature anomalies at Svanvik.
-Percentage of days warmer/colder than $\pm 1\,\sigma$ from climatological mean displayed for each month.
-Negative deviations from climatology are shown as negative percentage.
-The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
-The summer of 2018 was significantly warmer than average, especially May and July.
-The summer of 2019 was fairly average but significantly warmer than average in April and colder in July/August.}
-  \label{fig:plot_temperature_anomalies_svanvik}
-\end{figure}
-
-In a similar way, the precipitation anomalies at Svanvik are displayed in Fig.~\ref{fig:plot_precipitation_anomalies_svanvik}.
-However, the percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ from the climatological mean is shown for each month, due to the larger variability in observed precipitation too few days showed a significance of $1 \sigma$ or higher.
-As above, negative deviations from the climatology are shown as negative percentage and annual positive/negative deviations are indicated in the respective corners.
-The hatched area between the dashed lines indicates the expected percentage of values falling above $\pm\frac{1}{2}\,\sigma$ if a normal distribution is assumed ($30.8\,\unit{\%}$).
-Unlike for the temperature anomalies, the picture is not as clear.
-While March 2018 was unusually wet, summer and fall (May/July, September/October) were significantly drier than average.
-2019 was rather average throughout the year, with drier than usual fall (September/October) and a wet December.
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{precipitation_signific_1}
-  \caption{Precipitation anomalies at Svanvik.
-Percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ from climatological mean displayed for each month.
-Negative deviations from climatology are shown as negative percentage.
-The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The hatched area between the dashed lines indicates the expected percentage of values falling above $\pm\frac{1}{2}\,\sigma$ if a normal distribution is assumed ($30.8\,\unit{\%}$).
-Summer was significantly dry only in 2018, while fall was significantly dry in both years.}
-  \label{fig:plot_precipitation_anomalies_svanvik}
-\end{figure}
-
-Similarly, the global irradiance anomalies at Svanvik are presented in Fig.~\ref{fig:global_rad_signific}.
-The percentage of days above/below $\pm 1\,\sigma$ from the climatological mean is displayed for each month.
-The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed.
-In summer 2018 (May/July), global irradiance was significantly higher than on average, while 2019 was generally an average year with a slightly darker spring (April/May).
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{global_rad_signific_1}
-  \caption{Global irradiance anomalies at Svanvik.
-Percentage of days above/below $\pm 1\,\sigma$ from climatological mean displayed for each month.
-Negative deviations from climatology are shown as negative percentage.
-The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
-In summer 2018 (May/July), global irradiance was significantly higher than on average, while 2019 had significantly less than normal irradiance in April/May and more in June.}
-  \label{fig:global_rad_signific}
-\end{figure}
-
-In summary, 2018 was significantly warmer, drier, and lighter than the climatological average, while 2019 was a rather average year.
-
-\subsubsection{Ozone}
-\label{subsubsec:anomal_ozone}
-As we discussed in the preceding Section~\ref{subsubsec:anomal_tpq}, meteorological conditions in 2018 were significantly different from the climatological average.
-How does this translate into observed ozone anomalies \chem{\Delta[O_3]}?
-
-If the data collected at Svanvik had been complete in 2018 and the climatology had extended into the resent past, we could have assessed the significance of ozone anomalies in the same way as for temperature, precipitation, and global irradiance.
-In Section~\ref{subsubsec:clim_ozone}, we described the construction of climatologies for northern Fennoscandia based on long term observations at Esrange, Jergul/Karasjok, and Pallas.
-For Svanvik this assessment is based on rather uncertain observations from 1986--1996.
-We also discussed the general patterns, similarities and differences.
-We shall now investigate \chem{\Delta[O_3]} in 2018/19 with respect to these climatologies in more detail.
-
-In the following, we will use the interchangeable terms anomalies and residuals to distinguish between a climatological (anomalies) and a statistical (residuals) point of view.
-All the above factors make it necessary to look not only at anomalies but statistically assess ozone concentration residuals at Svanvik as well as at Esrange and Pallas in comparison.
-Since the climatology of latter is more reliable, we will start with surface ozone concentration residuals \chem{\Delta[O_3]} at Esrange and Pallas in 2018.
-As data for 2019 has not yet been available from EBAS (last access, February 2020), we refrain from looking into these.
-
-In Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}, the probability density functions of daily mean ozone concentration residuals for Esrange and Pallas are shown.
-Both stations display a clear deviation from the climatological mean, however, not significant.
-On average, \chem{[O_3]} had been elevated by $(1.9\pm 5.7)\,\unit{ppb}$ and $(1.9\pm 5.4)\,\unit{ppb}$, respectively.
-Both distributions are slightly skewed towards positive deviations with $\frac{1}{4}$ of days (3.
-quantile, q3) falling above $5.8\,\unit{ppb}$ and $5.6\,\unit{ppb}$, respectively.
-
-Appendix Figure~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik} displays the probability density functions of daily mean ozone concentration residuals for Svanvik based on data taken in 2018/19 with respect to the climatology derived for northern Fennoscandia (denoted NF in Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}a) and the climatology derived for Svanvik (Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b).
-As previously mentioned, \chem{\left<[O_3]\right>} at Svanvik are $(6.6\pm 7.1)\,\unit{ppb}$ lower than the northern Fennoscandic climatology.
-For 2018 and 2019, we find $\Delta\chem{\left<[O_3]\right>} = (-3.8\pm 5.6)\,\unit{ppb}$ and $\Delta\chem{\left<[O_3]\right>} = (-5.2\pm 5.0)\,\unit{ppb}$, respectively.
-While the probability density of the residuals between the two climatologies is skewed towards negative values, the skew of 2018/19 is positive.
-With respect to the climatology for Svanvik (Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b), the probability density functions average for 2018/19 lies at $\Delta\chem{\left<[O_3]\right>} = (2.6\pm 5.8)\,\unit{ppb}$ and $\Delta\chem{\left<[O_3]\right>} = (1.2\pm 5.0)\,\unit{ppb}$, respectively.
-Neither of the \chem{\Delta\left<[O_3]\right>} is significant by statistical means, but daily average ozone concentrations are slightly elevated in 2018 compared to 2019.
-
-The \chem{\left<\Delta\left<[O_3]\right>\right>} for Svanvik in 2018 is slightly larger than for Esrange and Pallas.
-This could mean either that 2018 had been more exceptional at Svanvik or the derived climatology may need a correction for increased tropospheric background ozone concentration.
-Based on the meteorological conditions, we assume 2019 as average year which leads to a constant bias of about $1.2\,\unit{ppb}$.
-Daily \chem{\Delta_{present-past}\left<[O_3]\right>} from Esrange and Pallas, wherein present is defined as the average of 2010--2018 and past as the respective beginning of data taking until 1997, do not reveal a clear trend: We find for Esrange $(1.5\pm 3.1)\,\unit{ppb}$ and for Pallas $(-2.5\pm 3.7)\,\unit{ppb}$ \citep{NILU2003, BER:Ruoho-Airola2015}.
-Former is compatible with our assumed bias for the climatology derived for Svanvik.
-
-To further investigate the ozone residuals, we conduct a Student's~t-test ($t = \frac{\chem{\Delta[O_3]}}{\sigma}$), assuming the same sample uncertainty $\sigma$ in both, climatology and the year in question.
-Results for Esrange and Pallas are shown in Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test}.
-This t-test clearly shows, that at both sites the residuals are significant on the $1\sigma$ level for at least $50\,\unit{\%}$.
-In Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}, the test statistics for Svanvik are displayed.
-From Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}a), we can deduce, that the residuals between the Svanvik climatology and the northern Fennoscandic climatology are significant on the $2\sigma$ level for at least half of the data, while the residuals are less certain in 2018 and 2019.
-In Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}b) the same is shown for the years 2018/19 with respect to the climatology from Svanvik.
-We find that less than $50\,\unit{\%}$ of ozone concentration residuals are significant on the $1\sigma$ level.
-
-In accordance to the above findings, we adjust the derived ozone climatology for Svanvik by adding a constant bias correction factor $\Delta\chem{[O_3]_\mathrm{bias-corr}} = 1.2\,\unit{ppb}$ and estimate the ozone concentration anomalies for 2018/19.
-In Fig.~\ref{fig:ozone_signific}, the resulting percentage of days above/below $\pm 1\,\sigma$ from the bias-corrected climatological mean is displayed for each month.
-Negative deviations from climatology are shown as negative percentage.
-Note that missing bars for certain months are both due to data availability (see Section~\ref{subsubsec:atmo_svanvik} for details about the time range of data taking at Svanhovd) and anomalies not meeting the criteria of $\pm 1\sigma$ deviation from climatological mean.
-In April/May 2018, ozone was significantly enhanced, while 2019 was generally an average year.
-In both years, ozone concentrations were significantly below average in November.
-July 2018 has not been corrected for the missing data due to data acquisition problems.
-The corrected value is, however, indicated by the star (refer to Section~\ref{subsec:ozone_reco} for details of the used method for gap filling).
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_signific_1}
-  \caption{Ozone concentration anomalies at Svanvik.
-Ozone climatology has been corrected for an constant negative bias of $-1.2\,\unit{ppb}$.
-Percentage of days above/below $\pm 1\,\sigma$ from climatological mean displayed for each month.
-Negative deviations from climatology are shown as negative percentage.
-The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
-July 2018 has not been corrected for the missing data due to data acquisition problems.
-The corrected value is indicated by the star (refer to Section~\ref{subsec:ozone_reco} for details).
-In April/May 2018, ozone was significantly higher than on average, while 2019 was generally an average year.
-In both years ozone concentrations were significantly below average in November.}
-  \label{fig:ozone_signific}
-\end{figure}
-
-To conclude, ozone concentrations in 2018 have been elevated compared to the respective climatologies, however, on average not significantly by statistical means.
-At Esrange and Pallas at least $25\,\unit{\%}$ (3.~quantile) of anomalies are significant on a $3\sigma$ level or more, while at Svanvik the 3.~quantile lies at about $2\sigma$.
-This means that in 2018 the stations at Esrange and Pallas saw more days with elevated ozone than was monitored at Svanvik.
-
-\section{DO3SE modeling}
-\label{sec:do3se}
-To investigate which of the environmental factors lead to the observed ozone damage on clover and tobacco in 2018 in contrast to 2019, when no such damage was found on the clovers but only on the sensitive tobacco cultivar, we model the ozone uptake using the $\mathrm{DO_3SE}$ model and give an estimate on biomass reduction due to ozone uptake.
-
-\subsection{Model description}
-\label{subsec:model_describtion}
-
-The opening of stomata is governed by different species-specific and environmental factors and can be described by a multiplicative model \citep{PTRS:Jarvis1976,ICP:MappingManual2017}:
-\begin{equation}
-  g^k_\text{sto} = g^k_\text{max} \cdot f^k_\text{phen} \cdot f^k_\text{light} \cdot \max{\left\{f^k_\text{min}, f^k_\mathrm{T} \cdot f^k_\mathrm{VPD} \cdot f^k_\text{SW}\right\}}.
-  \label{eq:stomatal}
-\end{equation}
-%
-Where $g_\mathrm{max}$ is the specific-specific maximum stomatal conductance value which is then modified by seasonal and environmental factors that vary within a range $0-1$.
-They are empirically determined and account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_\mathrm{T}$), water vapor pressure deficit ($f_\mathrm{VPD}$), and soil water content ($f_\text{SW}$).
-All factors differ with species or plant functional type (PFT) denoted with $k$.
-
-The temperature adjustment function is defined as
-%
-\begin{equation}
-  f_\mathrm{T} = \frac{T_\text{air}-T_\text{min}}{T_\text{opt}-T_\text{min}} \cdot \left(\frac{T_{\text{max}}-T_\text{air}}{T_{\text{max}}-T_\text{opt}}\right)^\beta,
-  \label{eq:f_temp}
-\end{equation}
-with $\beta = \frac{T_\text{max}-T_\text{opt}}{T_\text{opt}-T_\text{min}}$ a mixed reciprocal polynomial function of temperature.
-The shape parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various species or PFT and $T_\mathrm{air}$ is the surface air temperature.
-All temperatures are defined in units of \unit{^\circ C}.
-
-The water vapor pressure deficit function is
-%
-\begin{equation}
-  f_\mathrm{VPD} = f_\text{min}+(1-f_\text{min}) \cdot \frac{D_\text{min} - \text{VPD}}{D_\text{min}-D_\text{max}}
-  \label{eq:f_vpd}
-\end{equation}
-%
-where VPD is the leaf to air vapor pressure deficit in \unit{kPa} with $f_\text{min}$, $D_\text{min}$, $D_\text{max}$ describing the relative stomatal conductance to changes in vapor pressure deficit.
-
-The wavelength band $400-700\,\unit{nm}$ plant chlorophyll responds to is called photosynthetic active radiation (PAR).
-Its integral is the photosynthetic photon flux density (PPFD).
-The relationship between relative $g_\mathrm{sto}$ and PPFD is given by
-%
-\begin{equation}
-  f_\text{light} = 1-\text{exp}(-\alpha_\text{light}\cdot \text{PPFD})
-  \label{eq:flight}
-\end{equation}
-%
-where $\alpha_\mathrm{light}$ is a slope parameter describing how fast stomata open or close at low light intensities.
-
-The $\mathrm{DO_3SE}$ model as described in \citet{ACP:Bueker2012} is used to simulate SWP across a species or PFT specific root depth according to the Penman--Monteith energy balance method that drives water cycling through the soil--plant--atmosphere system.
-\citet{ACP:Bueker2012} discuss several formulations available to parameterize the reduction in stomata conductance due to water content in the soil.
-Here we use the relationship between reative $g_\mathrm{sto}$ and soil water potential (SWP) given by:
-%
-\begin{equation}
-  f_\mathrm{SW} = \text{min}\left\{1, \text{max}\left\{f_\mathrm{min}, \frac{(1 - f_\mathrm{min})\cdot(\mathrm{SWP_{min}} - \mathrm{SWP})}{\mathrm{SWP_{min}} - \mathrm{SWP_{max}}} + f_\mathrm{min} \right\} \right\}
-    \label{eq:fsw}
-\end{equation}
-%
-where SWP is the soil water potential across the root zone and $\mathrm{SWP_{min}}$ and $\mathrm{SWP_{max}}$ are the paramters describing the $f_\mathrm{SW}$ relationship.
-
-\subsection{Input data and gap filling methodology}
-\label{subsec:ozone_reco}
-The $\mathrm{DO_3SE}$ model requires hourly, continuous meteorological observations.
-In addition to variables covered in Section~\ref{sec:data}, $2\,\unit{m}$ wind $u_\text{2m}$ and vapor pressure deficit VPD are needed.
-VPD has been calculated from observed $T_\text{2m}$ and relative humidity
-\begin{equation}
-  \text{VPD} = P_s(T) \cdot \left(1-\frac{\text{relHum}}{100}\right),
-\end{equation}
-with saturation vapor pressure $P_s(T)$ in \unit{hPa} derived from Arden Buck equation~\citep{JAP:Buck1981, Buck2012}.
-
-The following gap filling methodology was devised for the meteorological input data:
-\begin{itemize}
-\item Single hours of missing data were filled by taking the average of the hourly values coming before, and after, the missing value.
-\item Several consecutive hours of missing data (23 or less) were filled by taking the average of the corresponding hour the day before, and the day after; and repeating this for each missing hour of data.
-If data was unavailable from that hour of the previous day, then only the value from the day after was used and vice versa.
-Interpolated values were not used in calculating averages.
-\item Data gaps longer than $24\,\unit{h}$ were filled using weekly diurnal averages.
-i.e.
-an average was calculated using the corresponding hour throughout the week before and after.
-Interpolated values were not used in calculating averages.
-\end{itemize}
-
-Because the available reanalysis products for surface ozone have their limitations, we deem it necessary to reconstruct the missing data in 2018 for the purpose of $\mathrm{DO_3SE}$ modeling by applying a Reynolds's decomposition of the form:
-\begin{equation}
-  \chem{[O_3]} = \chem{\left<[O_3]\right>} + \chem{\Delta[O_3]},
-  \label{eq:reynolds_decomp}
-\end{equation}
-with climatology \chem{\left<[O_3]\right>} and anomalies \chem{\Delta[O_3]}.
-The exact procedure will be described in the following.
-We derive hourly climatologies for individual stations and combine Esrange/Pallas to derive a climatology for northern Fennoscandia for July (Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}a).
-We compute temporal correlations between Svanvik and the other stations in northern Fennoscandia to estimate a time lag (Appendix Fig.~\ref{fig:time_lag_correlation}.
-We find $3\,\unit{h}$ for Esrange and Pallas with respect to Svanvik.
-Due to the high correlation with Svanvik (Appendix~Fig.~\ref{fig:density_distribution}), we decide on Pallas as reference station for the reconstruction.
-We correct the Fennoscandic climatology for the found time lag \chem{\left<[O_3]\right>_{hourly,\,t-3}} by shifting the respective time series back in time by $3\,\unit{h}$ and use it to derive a projection for Svanvik:
-\begin{equation}
-  X = \frac{\chem{\left<[O_3]\right>_{hourly}^{Svanvik}}}{\chem{\left<[O_3]\right>_{hourly,\,t-3}}}.
-  \label{eq:mapping}
-\end{equation}
-As an intermediate step, we apply Eq.~(\ref{eq:reynolds_decomp}) to derive hourly anomalies for each station $i$
-\begin{equation}
-  \chem{\Delta[O_3]_{hourly}}^i = \chem{[O_3]_{hourly}}^i-\chem{\left<[O_3]\right>_{hourly}},
-\end{equation}
-with $i \in \{\text{Esrange, Pallas, Svanvik\}}$.
-In Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}b), these hourly ozone concentration anomalies are shown.
-We also correct the ozone anomalies for Pallas for the time lag \chem{\Delta[O_3]_{hourly,\,t-3}^{Pallas}} and use the mapping function in Eq.~(\ref{eq:mapping}) to derive reconstructed anomalies for the missing values at Svanvik:
-\begin{equation}
-    \chem{\Delta [O_3]_{hourly}^{Svanvik,\,reco}} = \chem{\Delta[O_3]_{hourly,\,t-3}^{Pallas}} \cdot X
-\end{equation}
-Finally, we add the bias corrected Svanvik ozone climatology and derive a reconstructed ozone concentration time series \chem{[O_3]_{hourly}^{Svanvik,\,reco}}.
-In Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}c), the result is shown together with the observed data and the Copernicus Regional Model Reanalysis at the nearest grid point for comparison.
-
-\subsection{Model parameters}
-We roughly estimated land use types for an approximately $1\times 1\,\unit{km}$ patch centered at the NIBIO Environment Centre Svanhovd from Fig.~\ref{fig:svanhovd_research_station}: $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban.
-In the following, we derive bespoke parameterizations for the main vegetation types Norway spruce, birch, and perennial grassland.
-A comprehensive list of model parameters used as starting point in this study can be found in Supplement~Table~S1.
-Default PFTs are derived from \citet{ICP:MappingManual2017} and will be referred to as \emph{mapping manual}.
-
-These suggested, hypothetical PFTs or species parameterizations ought to be better adapted to the climatic conditions in the target region, but have to be varified by experiements.
-We hypothize that plants are likely to optimize their growth throughout the short growing season and hence the uptake of \chem{CO_2}.
-For simplicity, we assume a first order proptionality between photosynthesis and $g_\mathrm{sto}$.
-Plants are therefor likely to optimize their stomatal conductance with respect to climatological drivers.
-In Fig.~\ref{fig:f_temp_grassland}, we display the construction of bespoke PFTs or species temperature and light response function $f_\mathrm{T}$ and $f_\mathrm{light}$ for grassland (for Norway spruce and birch see Appendix~\ref{fig:f_temp_spruce_birch}).
-The response functions are shown together with climatology probability density functions (PDFs) of temperature and global radiation, respectively.
-In both cases, a parameterization is considered more optimized the larger the intersection area between the response function and PDF is.
-With respect to temperature (Fig.~\ref{fig:f_temp_grassland}a)), we assume two different types: \emph{boreal} and \emph{cold}.
-In both cases, we assume that the species are more tolerant to cold conditions than in the default mapping manual parameterization and shift $f_\mathrm{T}$ towards colder temperature.
-Herein, \emph{cold} represents a species that is more tolerant to warmer temperatures, but slightly less tolerant to cold temperatures than \emph{boreal}.
-For Norway spruce, $T_\mathrm{opt}$ are estimated from a longer timeseries of observed \chem{CO_2} uptake over temperature at two observations sites in Finland presented in \citet{TB:Kolari2007}.
-For the the senisitivity regarding light, we assume that the plants may react either faster or slower to lower light intensities.
-We qualitatively quantify this by a $\pm 20\,\%$ shift in respective halfway PPFD($f_\mathrm{light}=0.5$).
-The resulting functions are shown in Fig.~\ref{fig:f_temp_grassland}b).
-All derived parameters are given in Tables~\ref{tab:sensitivity_tests_temp}--\ref{tab:sensitivity_tests_light}.
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{javis_funcs_temp_hist_grassland}\\
-  \includegraphics[width=8.3cm]{javis_funcs_rad_hist_grassland}
-\caption{Construction of bespoke response functions for grassland.
-(a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown toghether with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability densitiy function - PDF), respectively.
-Original mapping manual parameterization is shown in comparison as solid line.
-Note that $Q_0$ has been truncated to $0.006$.
-PPFD0.8 and PPFD1.2 refer to $\alpha$ values increasing/decreasing PPFD at $f_\mathrm{light}=0.5$ by $\pm 20\,\%$, respectively.}
-\label{fig:f_temp_grassland}
-\end{figure}
-
-\begin{table}[t]
-  \caption{Bespoke temperature prameterizations. MM refers to mapping manual \citep{GCB:Mills2010,ICP:MappingManual2017}.}
-  \label{tab:sensitivity_tests_temp}
-  \begin{tabular}{llrrr}
-    \tophline
-    Species & type & $T_\mathrm{min}$ & $T_\mathrm{opt}$ & $T_\mathrm{max}$ \\
-    \middlehline
-    \multirow{3}{*}{Spruce} & MM & 0 & 20 & \multirow{3}{*}{100}\\
-    & Boreal & 0 & 10 & \\
-    & Cold & 0 & 15 & \\
-    \middlehline
-    \multirow{3}{*}{Birch} & MM & 5 & 20 & \multirow{3}{*}{100}\\
-    & Boreal & 0 & 10\\
-    & Cold & 5 & 15\\
-    \middlehline
-    \multirow{3}{*}{Grassland} & MM & 10 & 24 & 36\\
-    & Boreal & 0 & 10 & 24 \\
-    & Cold & 5 & 15 & 36\\
-    \bottomhline
-    \end{tabular}
-\end{table}
-
-\begin{table}[t]
-  \caption{Bespoke light prameterizations. MM refers to mapping manual \citep{GCB:Mills2010,ICP:MappingManual2017}.}
-  \label{tab:sensitivity_tests_light}
-  \begin{tabular}{llr}
-    \tophline
-    Species & type & $\alpha$\\
-    \middlehline
-    \multirow{3}{*}{Spruce} & MM & 0.006 \\
-    & PPFD0.8 & 0.008  \\
-    & PPFD1.2 & 0.005  \\
-    \middlehline
-    \multirow{3}{*}{Birch} & MM & 0.004 \\
-    & PPFD0.8 & 0.005\\
-    & PPFD1.2 & 0.004\\
-    \middlehline
-    \multirow{3}{*}{Grassland} & MM & 0.011\\
-    & PPFD0.8 & 0.014 \\
-    & PPFD1.2 & 0.009 \\
-    \bottomhline
-    \end{tabular}
-\end{table}
-
-{\bf TODO: Write about boreal and cold parameter sets and their justification.
-Write about suggested method for optimizing PFTs in a simple but physiological way based on optimizing maximum climatological mean $g_\mathrm{sto}$ and morning responds.
-If plants have a short growing season up north optimizing \chem{CO_2} uptake around noon (highest light intensity) and speed up opening of stomata at lower intensities (morning), additional contraint lower climatological variance?}
-An idea to justify our choice based on the fullset of climatological data for Svanvik is shown in Fig.~\ref{fig:javis_func_opt_mean}.
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{javis_func_opt_mean.pdf}
-  \caption{Proposed metric to test bespoke response functions.}
-  \label{fig:javis_func_opt_mean}
-\end{figure}
-
-Start ($A_\text{start}$) and end ($A_\text{end}$) of growing season for each PFT are given in units of day of year (doy) and estimated as follows.
-Evergreen needle leaf trees are photosynthetic active already at low temperatures \citep{TB:Kolari2007}.
-This leads necessarily to an increase in observable net photosynthesis.
-Hence, we use a MODIS (Aqua/Terra) retrieved net photosynthesis product for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine onset and end of growing season.
-Net photosynthesis data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}).
-By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{doy}$ for 2018 and day 106 in 2019.
-$A_\text{end}$ amounts to day 261 and 274, respectively, and shall be used for all vegetation alike.
-The resulting growing season for Norway spruce in 2019 was one month longer then in 2018.
-For estimating $A_\text{start}$ of birch tree, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Fig.~\ref{fig:greening_season_change_Svanvik}).
-We find 129 and 130, respectively.
-From the latitude model typically used for localizing $\mathrm{DO_3SE}$, we get $A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$ indicating an about $2\,\unit{months}$ longer growing season.
-For perennial grassland, we assume a latency period of about $1\,\unit{month}$ after snow melt.
-Data for first snow-free day is taken from the weather station at Øvre Neiden in Sør-Varanger.
-In both years, the first snow-free day incidents with the estimated date of bud burst.
-As leafs of birches in northern Fennoscandia \emph{Betula pubescens} are smaller than southern specimen, we have measured their dimension for the purpose of this study and find $(3.0\pm 0.5)\,\unit{cm}$.
-\citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area.
-
-As pointed out by \citet{ACP:Bueker2012}, a very important factor in deriving a proper $\mathrm{POD_y}$ is the soil texture or how much water the soil can hold.
-As Svanhovd is located in the bed of the Pasvik river, we assume a sandy loam texture at $60\,\unit{cm}$ depth.\\
-
-\subsection{POD and ozone impact}
-
-We have modeled $\mathrm{POD_y}$ for three natural/semi-natural vegetation PFTs, birch, Norway spruce, and perennial grassland.
-The results are shown comprehensively in Fig.~\ref{fig:pody} as a function of doy for March--October.
-A flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) has been chosen.
-\chem{[O_3]} are plotted on the same axis and scales as $G_\text{sto}^\text{leaf}$ but in units of $\unit{ppb}$.
-
-As expected, $\mathrm{POD_y}$ is higher in 2018 than in 2019 for all PFTs.
-In the end of the 2018 season, Birch took up $13.1\,\unit{mmol\,m^{-2}\,PLA}$, while perennial grassland had an uptake of $4.5\,\unit{mmol\,m^{-2}\,PLA}$.
-The largest interannual variation is also found in Birch ($2.1\,\unit{mmol\,m^{-2}\,PLA}$) and perennial grassland ($2.6\,\unit{mmol\,m^{-2}\,PLA}$).
-Whereas, there is only a small difference for Norway spruce ($0.3\,\unit{mmol\,m^{-2}\,PLA}$) between the two years.
-
-%Except for perennial grassland, model integrations with climatological forcings show higher $\mathrm{POD_y}$ than 2018.
-This can be explained by means of non linearity in modeled stomatal conductance (Eq.~(\ref{eq:stomatal})) in Eq.~(\ref{eq:pod}).
-The climatology in $\mathrm{POD_y}$ therefore represents rather an ideal year for growth then a true climatology.
-To establish a true climatology, it would be necessary to run the $\mathrm{DO_3SE}$ model for 10--20 years with respective observed phenology as well as meteorological and ozone forcing.
-
-%Based on this interpretation of climatology in $\mathrm{POD_y}$, we can learn more about potential ozone damage.
-The time derivative $\frac{\mathrm{\Delta POD_y}}{\Delta t}$ displays very different features for each PFT and meteorological forcing.
-Although Norway spruce shows a low interannual variation, $\frac{\mathrm{\Delta POD_y}}{\Delta t}$ (Fig.~\ref{fig:pody_rel}) indicates an extended period of higher than average ozone uptake in spring 2018 (steep slope), while ozone uptake in June/July was anomalously low in 2019 (shallow slope) but high in fall.
-This is most prominently visible in Fig.~\ref{fig:pody}h) for perennial grassland, where consecutive periods of low $G_\text{sto}^\text{leaf}$ manifest in three plateaus in $\mathrm{POD_y}$ throughout the season (1: $184-198\,\unit{doy}$, 2: $209-228\,\unit{doy}$, 3: $235-240\,\unit{doy}$).
-An episode with anomalous high ozone in mid June ($194\,\unit{doy}$) can be identified but does not amount to unusually high ozone uptake on an annual basis.
-
-The estimated biomass reduction in \unit{\%} due to \chem{O_3} following \citet{ICP:MappingManual2017} is shown in Table~\ref{tab:biomass_reduction}.
-{\bf TODO: Lisa?
-Add paragraph about biomass reduction to introduction?}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{DO3SE_results_pody_gsto_o3}
-  \caption{$\mathrm{POD_y}$ modeling results from $\mathrm{DO_3SE}$ model for three different PFTs over day of year (doy), March--October.
-A flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) has been chosen.
-\chem{[O_3]} are plotted on the same axis and scales as $G_\text{sto}^\text{leaf}$ but in units of $\unit{ppb}$.
-(a, b) Birch; (c, d) Norway spruce; (e, f) perennial grassland.
-From left to right: 2018, 2019.
-{\bf TODO: substitute with figure for boreal PFTs?}}
-  \label{fig:pody}
-\end{figure*}
-
-\begin{table}[t]
-  \caption{Estimated total biomass reduction in \unit{\%} due to \chem{O_3} following \citet{ICP:MappingManual2017}.}
-  \label{tab:biomass_reduction}
-\begin{tabular}{lccrr}
-\tophline
-Year & \multicolumn{4}{c}{PFT}\\
-& Birch & Norway spruce & \multicolumn{2}{c}{Perennial grassland}\\
-\middlehline
-2018 & 11.99 & 2.34 & 8.06 & 10.51$^*$\\
-2019 & 10.07 & 2.26 & 6.43 & 7.90$^*$\\
-clim. & 13.81 & 2.58 & 6.67 & 8.29$^*$\\
-\bottomhline
-\end{tabular}
-\belowtable{$^*$ above ground biomass} % Table Footnotes
-\end{table}
-
-%\begin{figure}[t]
-%  \includegraphics[width=8.3cm]{DO3SE_results_rel}
-%  \caption{Daily averaged $\mathrm{\Delta POD_y}/\Delta t$ from $\mathrm{DO_3SE}$ model for three different PFTs over day of year (doy), March--October.
-Error bars/-bands indicate $1 \sigma$ standard deviation.
-(a) Birch; (b) Norway spruce; (c) perennial grassland.}
-%  \label{fig:pody_rel}
-%\end{figure}
-
-\conclusions[Discussion and conclusions]
-\label{sec:conc} 
-\begin{itemize}
-  \item What did we learn from the 2018/2019 seasons in terms of ozone damage in the far north?
-  \item What are the main uncertainties in the tools we've used?
-  \item What implications could this have for future scenarios in which 2018 conditions may occur more often?
-\end{itemize}
-PODY is no good measure without a damage function with higher temporal resolution than annual.
-This has been already pointed out by \citet{AE:Musselman2006}!
-Higher than normal spring uptake (weather conditions) does not amount to critical uptake on an annual basis.
-Weather and climatic conditioning of plants is more important for PODy than episodes of higher than usual ozone concentrations.
-Therefor high $\Delta \mathrm{POD_y} / \Delta t$ is not accounted for properly.
-It is important to know the plant response functions to temperature, light, humidity/precipitation as best as possible if one wants to assess risk of damage.
-
-Regarding the shift (earlier) and prolongation (later) of the growing season in the far north no general assessment can be made.
-Data show that in both 2018 and 2019 ozone uptake in early season was anomalous high in all studied PFTs.
-The extended growing season in 2019 might have countered higher episodic ozone uptake in 2018.
-In both years, low mid-season uptake due to weather conditions washed out any effect of these.
-
-Bigger picture:
-\begin{itemize}
-\item Ozone damage assessment using mapping manual "compromise" parameterization.
-\item Land-atmosphere modelling: Problems in productivity of species especially in the arctic regions due to PFTs which are not suited for the climate.
-Automating optimization of climate adaption of general PFTs.
-\end{itemize}
-
-%% The following commands are for the statements about the availability of data sets and/or software code corresponding to the manuscript.
-%% It is strongly recommended to make use of these sections in case data sets and/or software code have been part of your research the article is based on.
-
-%\codeavailability{TEXT} %% use this section when having only software code available
-
-\dataavailability{TEXT} %% use this section when having only data sets available
-
-%\codedataavailability{TEXT} %% use this section when having data sets and software code available
-
-\sampleavailability{TEXT} %% use this section when having geoscientific samples available
-
-%\videosupplement{TEXT} %% use this section when having video supplements available
-
-\appendix
-\section{Derived climatologies}    %% Appendix A
-
-\subsection{Temperature, precipitation, and global irradiance}     %% Appendix A1, A2, etc.
-
-\appendixfigures  %% needs to be added in front of appendix figures
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{global_rad_clim}
-  \caption{Observed global irradiance at Svanvik.
-The cyan lines indicate the $1\,\sigma$ level deducted from the climatology.
-(a) Climatology (hourly maxima, mean, and minima); deviation from climatology (b) 2018; (c) 2019.}
-  \label{fig:global_rad_clim}
-\end{figure*}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{test_fennoscandia}
-  \caption{Maximum daily sun shine duration based on geometrical calculations at different sites in Fennoscandia.
-Midnight sun conditions at Jergul/Karasjok and Svanvik prevail from the end of May until the end of July, while at Esrange and Pallas they only prevail from the beginning of June until mid of July.}
-  \label{fig:sunlight_fennoscandia}
-\end{figure*}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{greening_season_change_Svanvik}
-  \caption{Estimated shift and prolongation of greening season at Svanhovd over the past 6 decades based on data from SeNorge.no {\bf TODO: citation http://www.senorge.no}.}
-  \label{fig:greening_season_change_Svanvik}
-\end{figure*}
-\clearpage
-
-\subsection{Ozone}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{density_distribution}
-  \caption{Probability densities and correlation coefficient for \chem{[O_3]} between different sites in northern Fennoscandia.
-Jergul/Karasjok is well correlated with Esrange, and Pallas.
-Svanvik displays highest correlation with Pallas.
-(a) Jergul/Karasjok--Esrange; (b) Jergul/Karasjok--Pallas; (c) Svanvik--Esrange; (d) Svanvik--Pallas.}
-  \label{fig:density_distribution}
-\end{figure*}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{ozone_climatology_fenoscandic_obs_residuals}
-  \caption{Probability density functions of ozone concentration residuals. 2018 with respect to respective climatology for different sites in Fennoscandia. (a) Esrange; (b) Pallas.}
-  \label{fig:ozone_climatology_fenoscandic_obs_residuals}
-\end{figure*}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{ozone_climatology_fenoscandic_obs_residuals-Svanvik}
-  \caption{Probability density functions of ozone concentration residuals. 2018/19 observations at Svanhovd with respect to derived climatologies for (a) Northern Fennoscandia; (b) Svanvik.}
-  \label{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}
-\end{figure*}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{ozone_climatology_fenoscandic_obs_test}
-  \caption{Student's t-test assuming same sample uncertainty in both, climatology and 2018 observations for different sites in Fennoscandia.
-\chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals} are significantly different from zero-hypothesis on the $1\,\sigma$ level.
-(a) Esrange; (b) Pallas.}
-  \label{fig:ozone_climatology_fenoscandic_obs_test}
-\end{figure*}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{ozone_climatology_fenoscandic_obs_test-Svanvik}
-  \caption{Student's t-test assuming same sample uncertainty in both, climatology and 2018/19 observations at Svanhovd with respect to derived climatologies for (a) Northern Fennoscandia; (b) Svanvik.
-\chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}a) are significantly different from zero-hypothesis on the $2\,\sigma$,  $1\,\sigma$ level, respectively.
-\chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}b) are not significantly different from zero-hypothesis.}
-  \label{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}
-\end{figure*}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{ozone_reconstruction_2018_07}
-  \caption{Reconstruction of missing \chem{[O_3]} observation data at Svanvik in July 2018 based on derived hourly climatologies of northern Fennoscandia, Svanvik, and data Pallas.
-(a) Time series; (b) anomalies; (c) reconstruction.}
-  \label{fig:ozone_reconstruction_2018_07}
-\end{figure*}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{ozone_observation_timelag_svanvik}
-  \caption{Temporal correlation of \chem{[O_3]} data at Svanvik with other ozone monitoring stations in northern Fennoscandia.
-A negative lag means that Svanvik lags behind, while a positive lag mean the other station lags behind.
-The highest correlation with Pallas/Esrange is found at a time lag of $3\,\unit{h}$, for Jergul/Karasjok at $1\,\unit{h}$, and for Janiskoski at $-20\,\unit{h}$.}
-  \label{fig:time_lag_correlation}
-\end{figure*}
-
-% Shall go to supplement!
-%\begin{figure*}[t]
-%  \includegraphics[width=12cm]{ozone_fennoscandic_obs_aot40}
-%  \caption{Integrated AOT40 for various sites in northern Fennoscandia over the course of 33 years.
-For simplicity, we integrated hourly ozone between $1\,\unit{am}-11\,\unit{pm}$ in any case, although a night time light intensity $> 50\,\unit{W\,m^{-2}}$ is only measured during midnight sun conditions.
-Since night time \chem{[O_3]} is mostly below $40\,\unit{ppb}$, the such induced high bias in spring should be tolerable.
-The dashed red lines indicate the respective threshold given by the EU directive.
-(a) May 1 -- July 31 (b) April 1 -- September 30.}
-%  \label{fig:fennoscandic_aot40}
-%\end{figure*}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{modis_Psn}
-  \caption{Estimated begin (BGS) and end (EGS) of growing season for coniferous trees from MODIS Aqua/Terra net photosynthesis (PSN) product.
-A $1\times 1\,\unit{km}$ area around Svanhovd was selected.
-Daily averaged data for both 2018 and 2019 has been fitted with a quadratic polynomial function.
-The numerically computed root yields: BGS 122/106 day of year (doy) and EGS 261/274 day of year for 2018/2019, respectively.}
-  \label{fig:modis_Psn}
-\end{figure*}
-
-\clearpage
-
-\noappendix       %% use this to mark the end of the appendix section. Otherwise the figures might be numbered incorrectly (e.g. 10 instead of 1).
-
-%% Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
-
-%% Option 1: If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
-%% They will be correctly named automatically.
-
-%% Option 2: If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.
-%% To rename them correctly to A1, A2, etc., please add the following commands in front of them:
-
-%\appendixfigures  %% needs to be added in front of appendix figures
-
-%\noappendix
-
-%\appendixtables   %% needs to be added in front of appendix tables
-
-%% Please add \clearpage between each table and/or figure. Further guidelines on figures and tables can be found below.
-
-\authorcontribution{SF has written the manuscript, collected and processed ozone and environmental data, and performed all statistical analyses.
-AVV has performed on-site observation of vegetation damage induced by ozone and contributed to theory of ozone interaction with vegetation within the introduction.
-LE...
-CO...
-ALEB...
-FS...
-TB...
-All authors contributed to discussion and conclusion of the results.} %% this section is mandatory
-
-\competinginterests{The authors declare that they have no conflict of interest.} %% this section is mandatory even if you declare that no competing interests are present
-
-%\disclaimer{TEXT} %% optional section
-
-\begin{acknowledgements}
-  Bj{\o}rg Rognerud for processing SeNorge.no data with respect for begin of growing season.
-  Tore Flatlandsmo Berglen for hourly pressure data from Svanhovd.
-\end{acknowledgements}
-
-%% REFERENCES
-
-%% The reference list is compiled as follows:
-
-%\begin{thebibliography}{}
-
-%\bibitem[AUTHOR(YEAR)]{LABEL1}
-%REFERENCE 1
-
-%\bibitem[AUTHOR(YEAR)]{LABEL2}
-%REFERENCE 2
-
-%\end{thebibliography}
-
-%% Since the Copernicus LaTeX package includes the BibTeX style file copernicus.bst,
-%% authors experienced with BibTeX only have to include the following two lines:
-%%
-\bibliographystyle{copernicus}
-\bibliography{VegOzone.bib}
-%%
-%% URLs and DOIs can be entered in your BibTeX file as:
-%%
-%% URL = {http://www.xyz.org/~jones/idx_g.htm}
-%% DOI = {10.5194/xyz}
-
-%% LITERATURE CITATIONS
-%%
-%% command                        & example result
-%% \citet{jones90}|               & Jones et al. (1990)
-%% \citep{jones90}|               & (Jones et al., 1990)
-%% \citep{jones90,jones93}|       & (Jones et al., 1990, 1993)
-%% \citep[p.~32]{jones90}|        & (Jones et al., 1990, p.~32)
-%% \citep[e.g.,][]{jones90}|      & (e.g., Jones et al., 1990)
-%% \citep[e.g.,][p.~32]{jones90}| & (e.g., Jones et al., 1990, p.~32)
-%% \citeauthor{jones90}|          & Jones et al.
-%% \citeyear{jones90}|            & 1990
-
-%% LITERATURE CITATIONS
-%%
-%% command                        & example result
-%% \citet{jones90}|               & Jones et al. (1990)
-%% \citep{jones90}|               & (Jones et al., 1990)
-%% \citep{jones90,jones93}|       & (Jones et al., 1990, 1993)
-%% \citep[p.~32]{jones90}|        & (Jones et al., 1990, p.~32)
-%% \citep[e.g.,][]{jones90}|      & (e.g., Jones et al., 1990)
-%% \citep[e.g.,][p.~32]{jones90}| & (e.g., Jones et al., 1990, p.~32)
-%% \citeauthor{jones90}|          & Jones et al.
-%% \citeyear{jones90}|            & 1990
-
-%% FIGURES
-
-%% When figures and tables are placed at the end of the MS (article in one-column style), please add \clearpage
-%% between bibliography and first table and/or figure as well as between each table and/or figure.
-
-% The figure files should be labelled correctly with Arabic numerals (e.g. fig01.jpg, fig02.png).
-
-%% ONE-COLUMN FIGURES
-
-%%f
-%\begin{figure}[t]
-%\includegraphics[width=8.3cm]{FILE NAME}
-%\caption{TEXT}
-%\end{figure}
-%
-%%% TWO-COLUMN FIGURES
-%
-%%f
-%\begin{figure*}[t]
-%\includegraphics[width=12cm]{FILE NAME}
-%\caption{TEXT}
-%\end{figure*}
-%
-%
-%%% TABLES
-%%%
-%%% The different columns must be seperated with a & command and should
-%%% end with \\ to identify the column brake.
-%
-%%% ONE-COLUMN TABLE
-%
-%%t
-%\begin{table}[t]
-%\caption{TEXT}
-%\begin{tabular}{column = lcr}
-%\tophline
-%
-%\middlehline
-%
-%\bottomhline
-%\end{tabular}
-%\belowtable{} % Table Footnotes
-%\end{table}
-%
-%%% TWO-COLUMN TABLE
-%
-%%t
-%\begin{table*}[t]
-%\caption{TEXT}
-%\begin{tabular}{column = lcr}
-%\tophline
-%
-%\middlehline
-%
-%\bottomhline
-%\end{tabular}
-%\belowtable{} % Table Footnotes
-%\end{table*}
-%
-%%% LANDSCAPE TABLE
-%
-%%t
-%\begin{sidewaystable*}[t]
-%\caption{TEXT}
-%\begin{tabular}{column = lcr}
-%\tophline
-%
-%\middlehline
-%
-%\bottomhline
-%\end{tabular}
-%\belowtable{} % Table Footnotes
-%\end{sidewaystable*}
-%
-%
-%%% MATHEMATICAL EXPRESSIONS
-%
-%%% All papers typeset by Copernicus Publications follow the math typesetting regulations
-%%% given by the IUPAC Green Book (IUPAC: Quantities, Units and Symbols in Physical Chemistry,
-%%% 2nd Edn., Blackwell Science, available at: http://old.iupac.org/publications/books/gbook/green_book_2ed.pdf, 1993).
-%%%
-%%% Physical quantities/variables are typeset in italic font (t for time, T for Temperature)
-%%% Indices which are not defined are typeset in italic font (x, y, z, a, b, c)
-%%% Items/objects which are defined are typeset in roman font (Car A, Car B)
-%%% Descriptions/specifications which are defined by itself are typeset in roman font (abs, rel, ref, tot, net, ice)
-%%% Abbreviations from 2 letters are typeset in roman font (RH, LAI)
-%%% Vectors are identified in bold italic font using \vec{x}
-%%% Matrices are identified in bold roman font
-%%% Multiplication signs are typeset using the LaTeX commands \times (for vector products, grids, and exponential notations) or \cdot
-%%% The character * should not be applied as mutliplication sign
-%
-%
-%%% EQUATIONS
-%
-%%% Single-row equation
-%
-%\begin{equation}
-%
-%\end{equation}
-%
-%%% Multiline equation
-%
-%\begin{align}
-%& 3 + 5 = 8\\
-%& 3 + 5 = 8\\
-%& 3 + 5 = 8
-%\end{align}
-%
-%
-%%% MATRICES
-%
-%\begin{matrix}
-%x & y & z\\
-%x & y & z\\
-%x & y & z\\
-%\end{matrix}
-%
-%
-%%% ALGORITHM
-%
-%\begin{algorithm}
-%\caption{...}
-%\label{a1}
-%\begin{algorithmic}
-%...
-%\end{algorithmic}
-%\end{algorithm}
-%
-%
-%%% CHEMICAL FORMULAS AND REACTIONS
-%
-%%% For formulas embedded in the text, please use \chem{}
-%
-%%% The reaction environment creates labels including the letter R, i.e. (R1), (R2), etc.
-%
-%\begin{reaction}
-%%% \rightarrow should be used for normal (one-way) chemical reactions
-%%% \rightleftharpoons should be used for equilibria
-%%% \leftrightarrow should be used for resonance structures
-%\end{reaction}
-%
-%
-%%% PHYSICAL UNITS
-%%%
-%%% Please use \unit{} and apply the exponential notation
-
-\end{document}

### 2021-03-02 VegOzone.tex
Fix hyperref
Restructured Section 2 and spell-ckeck.

@@
 
-\runningtitle{Subarctic biomes for land surface modelling}
+\runningtitle{Subarctic biomes for land surface modeling}
 
@@
 Annual global yield losses of four major crops (wheat, rice, maize, and soybean) of $3-15\,\unit{\%}$ \citep{PJ:Ainsworth2017} and a suggested loss in primary production in forestry of $7\,\unit{\%}$ \citep{GCB:Wittig2009,EP:Matyssek2012} has been attributed to ozone.
-{\bf TODO: Wealth of evidence demonstrating ozone damage to vegetation (perhaps focus on key evidence from northern climates and key species (e.g.
-birch, Norway srpuce etc...
-- this might help to make it clear why we focus on these species with the PODy modelling))}
 The crop yield loss studies indicate a threat to food security in rapidly developing countries, e.g., in East and South-East Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018}, where ozone concentrations related to enhanced pollutant emissions from the transport, residential, power generation and industrial sectors have increased in the past few decades.
-Even though long-term (pre 1950) observations of surface ozone concentrations are scarce; evidence from trend analysis of those few sites suggests that since the industrial revolution, background concentrations in the northern Hemisphere have at least doubled and continued to increase until relatively recently \citep[Chapter 2]{IPCC2013}.
+Even though long-term (pre 1950) observations of surface ozone concentrations are scarce; evidence from trend analysis of those few sites suggests that since the industrial revolution, background concentrations in the northern hemisphere have at least doubled and continued to increase until relatively recently \citep[Chapter 2]{IPCC2013}.
 Recent trend analyses indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric ozone started to level-off or even decline \citep{ESA:Cooper2014, ACP:Wespes2018,ESA:Gaudel2018}.
-This can be attributed to a successful implementation of air quality regulations coupled with economic restructuring which saw a move from coal to gas, reduced the number of episodes of peak concentrations especially in summertime over Europe and North America \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
-However, there are indications that an increasing frequency of thermal stress accompanying climate change will reduce vegetations' capacity for ozone uptake and therefor scotch these achievments \citep{NCC:Lin2020}.\\
+This can be attributed to a successful implementation of air quality regulations coupled with economic restructuring which saw a move from coal to gas, reducing the number of episodes of peak concentrations especially in summertime over Europe and North America \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
+However, there are indications that an increasing frequency of thermal stress accompanying climate change will reduce vegetation's' capacity for ozone uptake and therefor scotch these achievements \citep{NCC:Lin2020}.\\
 
@@
 
-In the annual cycle in boreal and temperate climates, a distinct maximum of ground level ozone concentrations typically occurs in spring followed by lower concentrations throughout the summer.
-Several conditions unique to the Actic are stimulating the formation of this spring peak.
+In the annual cycle in Boreal and temperate climates, a distinct maximum of ground level ozone concentrations typically occurs in spring followed by lower concentrations throughout the summer.
+Several conditions unique to the Arctic are stimulating the formation of this spring peak.
 Major pathways of ozone and its precursors' removal from the boundary layer include dry deposition on vegetated surfaces and photochemistry \citep{RG:Clifton2020}.
@@
 This leads to a build up of ozone and its precursors in the course of winter.
-Come spring precursors are photochemical activated again, while snow cover typically prevails for a longer period of time continuing to suppress ozone dry deposition.
-On the other hand, very high ozone concentrations occuring over shorter periodes of time can often be attributed to deep tropopause folding events over the Actic during winter.
+Come spring precursors are photochemical activated again, while snow cover typically continues to prevail and suppress ozone dry deposition.
+On the other hand, very high ozone concentrations occurring over shorter periods of time can often be attributed to deep tropopause folding events over the Arctic during winter.
 These events facilitate an intrusion of stratospheric air masses which are enriched in ozone \citep{JGR:Skerlak2015}.
@@
 
-The start of growing season in northern Fennoscandia has been shifting to earlier dates in response to the general warming trend \citep[e.g.,]{GCB:Menzel2006,RS:Hogda2013,IJB:Karlsen2007} and thereby converging with the period of the ozone spring peak.
+The start of growing season (GS) in northern Fennoscandia has been shifting to earlier dates in response to the general warming trend \citep[e.g.,]{GCB:Menzel2006,RS:Hogda2013,IJB:Karlsen2007} and thereby converging with the period of the ozone spring peak.
 At the same time the growing season is also becoming longer.
 In conjunction with midnight sun conditions (almost $24\,\unit{h}$ photosynthetic activity), a longer growing season, is prolonging the time in which vegetation can accumulate ozone.
-These factors may promote a higher potential susceptibility of northern vegetation to ozone in the future.
+For a natural and semi-natural vegetation already subject to rapid climatic changes, these factors could promote a higher potential risk for northern vegetation to ozone induced damage in the not so distant future.
 
@@
 Its main action is imposed through reactions occurring in the cell walls and cell membranes of mesophyll cells inside the leaves.
-Ozone enters the leaves through stomata, that are kept open to allow for gas exchange allowing the entry of \chem{CO_2} for photosynthesis and loss of \chem{H_2O} vapour via the plant transpiration stream.
-Stomatal aperture, and hence stomatal conductance to these gases (including ozone) will vary over the course of the day and growing season primarily to balance \chem{CO_2} uptake against \chem{H_2O} vapour loss.
+Ozone enters the leaves through stomata, that are kept open to allow for gas exchange allowing the entry of \chem{CO_2} for photosynthesis and loss of \chem{H_2O} vapor via the plant transpiration stream.
+Stomatal aperture, and hence stomatal conductance to these gases (including ozone) will vary over the course of the day and growing season primarily to balance \chem{CO_2} uptake against \chem{H_2O} vapor loss.
 The higher the stomatal conductance, the higher the potential for ozone uptake.
-Stomatal conductance has been emperically linked to environmental factors such as air temperature, photosynthetic active radiation, vapor pressure deficit, and soil water potential as well as photosynthesis itself.\\
+Stomatal conductance has been empirically linked to environmental factors such as air temperature $T$, photosynthetic photon flux density (PPFD), vapor pressure deficit (VPD), and soil water potential (SWP) as well as photosynthesis itself.
+Results from open top chamber (OTC) experiments on downy birch (\emph{Betula pubescens}) and mountain birch (\emph{Betula pubescens toruosa}), native to subarctic regions, indicated reductions in both biomass, in root:shoot ratio, and visible leaf damage under elevated ozone treatment ($\left<\chem{[O_3]}\right> = 36-54\,\unit{ppb}$) \citep{Amb:Manninen2009}.
+Although considered to be more ozone tolerant due to an absence of visible injuries \citep{Amb:Girgzdiene2009}, \citet{Amb:Manninen2009} also found chlorophyll:carotenoid ratio and polyamines reductions in Scots pine (\emph{Pinus sylvestris}).\\
 
 To measure the risk of damage inflicted to vegetation through ozone, different metrics have been defined.
-Two main types can be differentiated, concentration-based or flux-based metrics, e.g., Accumulated Over a Threshold $\mathrm{x}$ ($\mathrm{AOT_x}$) and Phyto-toxic Ozone Dose over a threshold $\mathrm{y}$ ($\mathrm{POD_y}$), respectively.
-Due to their simplicity, concentration-based metrics are still widely used.
-Despite comparisons of exposure-based methods with previously used concentration based methods (e.g.
-$\mathrm{AOT_{40}}$) show that $\mathrm{POD_y}$ is better suited to represent the risk to vegetation from ozone, especially in the more climate extreme parts of Europe \citep{GCB:Mills2010,ICP:MappingManual2017}.
+Two main types can be differentiated, concentration-based and exposure-based metrics, e.g., Accumulated Over a Threshold $\mathrm{x}$ ($\mathrm{AOT_x}$) and Phyto-toxic Ozone Dose over a threshold $\mathrm{y}$ ($\mathrm{POD_y}$).
+Due to their simplicity, concentration-based metrics are still widely used, despite comparisons between the two methods (e.g.
+$\mathrm{AOT_{40}}$ with $\mathrm{POD_3}$) show that exposure-based metrics are better suited to represent the actual risk to vegetation from ozone, especially in the more climatically extreme parts of Europe \citep{GCB:Mills2010,ICP:MappingManual2017}.
+Most of these previous studies focused on the Mediterranean region.
 
@@
 For computing the $\mathrm{POD_y}$, an estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf.
-Details specific to the $\mathrm{DO_3SE}$ model are given in Section~\ref{sec_do3se}.
+Details specific to the $\mathrm{DO_3SE}$ model are given in Section~\ref{sec:do3se}.
 $\mathrm{POD_y}$ is then calculated according to:
@@
 %
+
+Within the scope of this study to assess probable effects of conditions unique to northern Fennoscandia on ozone uptake by plants and resultant enhanced susceptibility to ozone inflicted damage, we conducted ambient air ozone concentration measurements at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd in the growing seasons of 2018/19 in collaboration with the Norwegian Institute for Air Research (NILU).
+In 2018, we have observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden at Svanhovd.
+In contrast, no such damage was found on the clovers in 2019, although it was found on the sensitive tobacco cultivar.
+This indicates that the vegetation may have been more affected by ozone in 2018 than in 2019.
+In terms of $\mathrm{AOT_{40}}$, the threshold for protection of vegetation of $3000\,\unit{ppb}$ (integrated over the growing season) given by the EU directive has been well exceeded at many ozone monitoring sites in Southern Norway in the 2018 growing season \citep{NILU2019} but not at Svanhovd.
 
@@
 
-These methods described above rely on accurate representation of plant physiological response to climate variables (i.e.
-the species or plant funtional type specific responses used to parameterise risk assessment models) and are most often parameterized using data for less extreme climates, i.e., temperate or continental climates; and to a lesser extent, boreal and Mediterranean climates.
-Parameterization of these models for more extreme bio-geographical regions such as sub-arctic climates has not been perfromed and relies on assuming that less extreme boreal parameterizations will be applicable.
-It is important to understand the implications of using these more generic parameterizations for more extreme climate regions, here we explore how risk estimates might be over or underestimated using generic vs bespoke sub-arctic parameterisations.
-we also explore the role [O3] profile and changing climate variables to understand how this risk might alter in the future with the advent of climate change.
+The methods described above rely on accurate representation of plant physiological response to climate variables (i.e.
+the species or plant functional type specific responses used to parameterize risk assessment models) and are most often parameterized using data for less extreme climates, i.e., temperate or continental climates; and to a lesser extent, Boreal and Mediterranean climates.
+Parameterization of these models for more extreme bio-geographical regions such as subarctic climates has not been performed and relies on assuming that less extreme Boreal parameterization will be applicable.
+In this study, conditions at Svanhovd for the growing seasons 2018/19 shall serve as reference.
+It is important to understand the implications of using these more generic parameterizations for more extreme climate regions, here we explore how risk estimates might be over or underestimated using generic vs bespoke subarctic parameterization.
+we also explore the role \chem{[O_3]} variability and changing climate variables to understand how this risk might alter in the future with the advent of climate change.
 This allows both an understanding of how the threat from ozone may change over time but also which are the key aspects of plant physiology that might determine risk (and hence which parameters might require special attention to get right) both for pollution impact assessment, but also more generally for describing gas exchange response to changing climates that will influence bio-geochemcial cycling.
-This might be one key to solve long-standing issues of coupled earthsystem models (ESM) which generally lack PFTs representing sub-arctic conditions in their landsurface model.
-
-Within the scope of our project to assess probable effects of midnight sun conditions on ozone uptake by plants and resultant enhanced susceptibility to ozone inflicted damage, we conducted ambient air ozone concentration measurements at Svanhovd in the growing seasons of 2018/19 in collaboration with the Norwegian Institute for Air Research (NILU).
-In 2018, we have observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd.
-In contrast, no such damage was found on the clovers in 2019, although it was found on the sensitive tobacco cultivar.
-This indicates that the vegetation may have been more affected by ozone in 2018 than in 2019.
-In terms of $\mathrm{AOT_{40}}$, the threshold for protection of vegetation of $3000\,\unit{ppb}$ (integrated over the growing season) given by the EU directive has been well exceeded at many ozone monitoring sites in Southern Norway in the 2018 growing season \citep{NILU2019} but not at Svanhovd.
-
-In this study, we investigate potential main drivers for the observed ozone damage in the growing season 2018 at NIBIO Environment Centre Svanhovd in contrast to 2019.
-We compare $\mathrm{POD_y}$ values for natural and semi-natural vegetation for derived climatological averages of environmental conditions at Svanhovd as well as for the growing seasons 2018/19 which shall serve as case studies.
+This might be one key to solve long-standing issues of coupled earth system models (ESM) which generally lack PFTs representing subarctic conditions in their land surface model.\\
 
@@
 We use the $\mathrm{DO_3SE}$ model to estimate the ozone uptake by natural and semi-natural vegetation to investigate main environmental drivers for the observed ozone damage in 2018 (Section~\ref{sec:do3se}).
-In this context, we introduce the$\mathrm{DO_3SE}$ model equations and derive bespoke parameterizations for sub-arctic species to assess systematic uncertainties arising from the use of default parameterizations.
+In this context, we introduce the$\mathrm{DO_3SE}$ model equations and derive bespoke parameterizations for subarctic species to assess systematic uncertainties arising from the use of default parameterizations.
 In Section~\ref{sec:conc} finally, we will summarize our results, draw conclusions, and give an outlook on further work.
@@
 
-In our assessment of ozone damage on vegetation in the subarctic in relation to the 2018 heatwave, we focus on the NIBIO operated Environment Centre Svanhovd which lies close by the settlement of Svanvik in the Pasvik valley (Section~\ref{subsec:svanhovd}).
-Throughout the paper, we will use Svanhovd synonymously for the atmospheric monitoring site at Svanvik.
-The available meteorological monitoring data are described in Section~\ref{subsubsec:atmo_svanvik}.
-In Section~\ref{subsubsec:ozone_garden}, the ozone garden -- a plant-based ozone monitoring -- will be described in more detail.
-Other data used for evaluation of the ozone data taken at Svanhovd is described in Section~\ref{subsec:other_data}.
-We compare with observational data from other monitoring sites in northern Fennoscandia (Section~\ref{subsubsec:ebas}) and use as well reanalysis products (Section~\ref{subsubsec:ozone_rea}) as complement when observational data was not available.
-
-\subsection{NIBIO Environment Centre Svanhovd}
-\label{subsec:svanhovd}
-"Svanhovd is located beside the Pasvik river, in the middle of the wedge of Norwegian land separating Russia and Finland in the north.
-It is where the eastern Siberian taiga meets the western Boreal forest.
-[...] Svanhovd was originally established as a State Demonstration and Experimental Farm in the 1930's, in order to expand agriculture in Pasvik valley [...]" \citep[Svanhovd][]{NIBIO_Svanhovd}.
-
-The locations of the atmospheric monitoring site as well as the ozone garden with respect to the main buildings of the Svanhovd Environment Centre are marked in the aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}.
-The Centre is located south of the settlement of Svanvik, where as the border to Russia, the Pasvik river, lies only a few hundred meters to the east.
-The original farm buildings, protected by law, are located in the lower center, while the Svanhovd Environment Centre follows to the north-east.
-Left of the Centre and north of the original farm buildings a botanical garden can be found.
-It is one amongst three such gardens north of the Polar Circle.
-The Svanhovd Environment Centre focuses on research of the special environment of the region, but also comprises long-term observations of various agrometeorological variables, i.a.
+In our assessment of primary drivers of ozone damage risk on vegetation in the subarctic, we focus on the 2018 growing season which was subject to an extended heatwave and its subsequent growing season of 2019.
+The NIBIO operated Environment Centre Svanhovd which lies close by the settlement of Svanvik in the Pasvik valley serves as main reference.
+Throughout the paper, we will use Svanhovd synonymously for the atmospheric monitoring site at Svanvik/Pasvik.
+The available meteorological monitoring data are described in Section~\ref{subsec:atmo_svanvik}.
+Other data used for cross-evaluation of the ozone monitoring data taken at Svanhovd is described in Section~\ref{subsec:other_data}.
+We compare with observational data from other monitoring sites in northern Fennoscandia (Section~\ref{subsubsec:ebas}) and use these to evaluate reanalysis products for ground level ozone (Section~\ref{subsubsec:ozone_rea}).
+
+\subsection{NIBIO Environment Centre Svanhovd and atmospheric monitoring}
+\label{subsec:atmo_svanvik}
+The foundation of the NIBIO Environment Centre Svanhovd dates back to the 1930s \citep{NIBIO_Svanhovd}.
+The center focuses on research of the special environment of the region, but also comprises long-term observations of various agrometeorological variables, i.a.
 $2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths.
-
-For the purpose of relating observed damage on ozone sensitive plants (e.g.
-clover and tobacco) to local ozone concentrations, an ozone monitor had been installed during the growing seasons 2018/19.
-The relevant data are described in more detail in Section~\ref{subsubsec:atmo_svanvik}.
+The locations of the atmospheric monitoring site and the ozone garden are marked in an aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}a).
+Svanhovd is located south of the settlement of Svanvik, where as the border to Russia, the Pasvik river, lies only a few hundred meters to the east.\\
 
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{svanhovd_researchstation}
-  \caption{NIBIO Environment Centre Svanhovd close by the settlement of Svanvik, Norway. Atmospheric monitoring site and ozone garden have been marked. Aerial photography \copyright Norges Kartverk.}
+  \includegraphics[width=8.3cm]{svanhovd_researchstation}\\
+  \includegraphics[width=8.3cm]{IMG_7739.JPG}  
+  \caption{NIBIO Environment Centre Svanhovd close by the settlement of Svanvik, Norway.
+(a) Atmospheric monitoring site and ozone garden have been marked.
+Aerial photography \copyright Norges Kartverk; (b) Clover in the Svanhovd ozone garden.
+The plants had to be secured against herbivores with a wire-mesh fence.
+The plants shown are approximately $6-12\,\unit{cm}$ in height.}
   \label{fig:svanhovd_research_station}
@@
 
-\subsubsection{Atmospheric monitoring}
-\label{subsubsec:atmo_svanvik}
-For the growing seasons 2018/19, monitoring data of key variables (temperature, pressure, precipitation, and ozone) have been obtained from luftkvalitet.no which is operated by \citet{NILU_AIRQ}.
-
-All agrometeorological variables including temperature and precipitation are available from September 1992 to present day \citep[LandbruksMeteorologiske Tjeneste][note the station name here is Pasvik]{LMT_NIBIO}.
-
-Relevant data for the growing season 2018/19 are shown in Fig.~\ref{fig:data_svanvik_2018_2019}.
+An ozone garden was installed in the course of our project at the NIBIO Environment Centre Svanhovd and was likely the northern-most ozone garden to date.
+An ozone garden consists of selected plant species of which one cultivar is more sensitive to ozone than the other.
+Cultivated species were, e.g.
+clover, tobacco, and beans.
+As shown in Fig.~\ref{fig:svanhovd_research_station}b), the plants had to be protected from herbivores with a wire-mesh fence.
+In 2018, we have observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden.
+In contrast, no such damage was found on the clovers in 2019, although it was found on the sensitive tobacco cultivar.
+This indicates that the vegetation may have been more affected by ozone in 2018 than in 2019.
+
+For the purpose of relating observed damage on ozone sensitive plants to local ozone concentrations, an ozone monitor had been installed during the growing seasons 2018/19 in addition to existing monitors of atmospheric key variables (temperature, pressure, precipitation).
+All variables have been obtained through \href{luftkvalitet.no}{luftkvalitet.no} operated by \citet{NILU_AIRQ}.
+Ozone monitoring data has also been added to EBAS database operated by \citet{NILU_EBAS}.
+
+Long-term data sets of agrometeorological variables including temperature and precipitation are available from September 1992 to present day \citep[LandbruksMeteorologiske Tjeneste][note the station name here is Pasvik]{LMT_NIBIO}.
+
+All relevant data for the growing season 2018/19 are shown in Fig.~\ref{fig:data_svanvik_2018_2019}.
+Ozone concentrations \chem{[O_3]} measured in $2\,\unit{m}$ height above ground are averaged hourly.
 The hatched areas mark times when no ozone data was recorded.
@@
 
-Ozone concentrations \chem{[O_3]} measured in $2\,\unit{m}$ height above ground are averaged hourly.
  As can be seen in Fig.~\ref{fig:data_svanvik_2018_2019}a), \chem{[O_3]} peaks in spring (April/May) and reaches its minimum in late summer (July/August).
@@
 
-\subsubsection{Ozone garden}
-\label{subsubsec:ozone_garden}
-An ozone garden consists of carefully selected plant species of which one cultivar is more sensitive to ozone than the other.
-At the NIBIO Environment Centre Svanhovd, the likely northern-most ozone garden to date, an ozone garden was installed in the course of our project.
-Cultivated species were, e.g.
-clover, tobacco, and beans.
-As shown in Fig.~\ref{fig:ozone_garden}, the plants had to be protected from herbivores with a wire-mesh fence.
-In 2018, we have observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden.
-In contrast, no such damage was found on the clovers in 2019, although it was found on the sensitive tobacco cultivar.
-This indicates that the vegetation may have been more affected by ozone in 2018 than in 2019.
-
-\begin{figure}
-  \includegraphics[width=8.3cm]{IMG_7739.JPG}
-  \caption{Clover in the Svanhovd ozone garden. The plants had to be secured against herbivores with a wire-mesh fence. The plants shown are approximately $6-12\,\unit{cm}$ in height.}
-  \label{fig:ozone_garden}
-\end{figure}
-
 \subsection{Other data}
 \label{subsec:other_data}
-To determine the significance of the ozone conditions in 2018, we have to derive a surface ozone concentration climatology for Svanvik.
+To determine the significance of the ozone conditions in 2018, we derive a surface ozone concentration climatology for Svanvik.
 From 1986--1996 ozone monitoring has been operated by NILU at Svanhovd, but no ozone measurements have been conducted for the past $14\,\unit{years}$.
-Due to probable changes in northern hemispheric tropospheric background ozone and a reduction of summertime peak values related to ozone precursors which fell under air quality regulations in Europe, we also have to have a look at ozone observations at other sites in northern Fennoscandia.
+Due to probable changes in northern hemispheric tropospheric background ozone and a reduction of summertime peak values related to ozone precursors which fell under air quality regulations in Europe, we will cross-calibrate our data with long-term ozone observations at other sites in northern Fennoscandia.
 In regard to ozone, available observational data are sparse in this region.
 An overview over the geographical locations of conducted atmospheric monitoring in the past and present is given in Fig.~\ref{fig:station_map_fennoscandia}.
-The data is described in Section~\ref{subsubsec:ebas}.
-As additional complement to observational data which is not available at all times and spatially sparse, we also look at various ozone reanalysis products (Section~\ref{subsubsec:ozone_rea}).
+The data are described in Section~\ref{subsubsec:ebas}.
+As additional complement to observational data which are not available at all times and spatially sparse, we also evaluate various ozone reanalysis products (Section~\ref{subsubsec:ozone_rea}).
 
@@
 In Table~\ref{tab:ebas_obs}, we give an overview over the exact locations and time range of operational data taking.
-All stations with available long term observations are located on higher ground than Svanhovd which will result in higher surface ozone concentrations in comparison as ozone abundance, in general, decreases with decreasing altitude in the absence of local precursors \citep[e.g.,][]{AB:Klingberg2009}.
+All stations with available long-term observations are located on higher ground than Svanhovd which will result in higher surface ozone concentrations in comparison as ozone abundance, in general, decreases with decreasing altitude in the absence of local precursors \citep[e.g.,][]{AB:Klingberg2009}.
 The periphery vegetation at all sites is similar and consists mainly of pine forests, birch, and heath.
@@
 
-Data from before the 1990s is available solely from Svanvik and Karasjok.
+Data from before the 1990s are available solely from Svanvik and Karasjok.
 These data, however, do not follow the high quality standard procedures implemented nowadays and have to be treated with care \citep{NILU2003}.
@@
 In the following, we will look at derived ozone climatologies from different perspectives.
-The multi-annual means of daily mean ozone, for simplicity here without standard deviations, at observation sites in northern Fennoscandia (refer to Table~\ref{tab:ebas_obs} for details) are shown in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs}.
+The multi-annual means of daily mean ozone, for simplicity here without standard deviations, at observation sites in northern Fennoscandia (refer to Table~\ref{tab:ebas_obs} for details) are shown in Supplement Fig.~S1.
 All climatologies peak in spring, with peak values of about $48\,\unit{ppb}$ in late April.
@@
 The large spread in case of Svanvik and Janiskoski can be explained due to the lower overall statistics.
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs}
-  \caption{Multi-annual mean of daily mean ozone at observation sites in Fennoscandia.
-The large spread in case of Svanvik and Janiskoski is due to the lower statistics.
-All climatologies peak in spring, with peak values in April.
-The annual average ozone concentration \chem{\left<[O_3]\right>} at Svanvik is $6.6\,\unit{ppb}$ lower than at the sites at Jergul/Karasjok (NOR), Esrange (SWE), and Pallas (FIN).}
-  \label{fig:ozone_climatology_fenoscandic_obs}
-\end{figure}
 
@@
 The spring peak apparently occurs slightly earlier (in the beginning of April).
-As Svanvik is located at lower altitude than the other sites, the earlier ozone peak may point to an interplay with ozone uptake by vegetation.
-{\bf TODO: The data from SeNorge seems to point to a later onset of greening, than what I interfered from OpenIFS data.
-This means that the beginning of the growing season is not a good measure for the termination of the ozone spring peak.
-This does not exclude the early start of pine forest photosynthesis as a driver.
-Coniferous forest starts to photosynthesis earlier than what the growing season begin would mark.
-\chem{CO_2} uptake and phenology of coniferous trees \citep{TB:Kolari2007} starts as early as doy 100, acclimates to leaf temperature history - photosynthesis drops at around doy 200--250.} The average ozone concentration at Svanvik is lower than at the other sites, as already mentioned, which is in league with a decrease in tropospheric ozone with decreasing altitude.
+As Svanvik is located at lower altitude than the other sites, the earlier decline of the ozone peak may point to a stronger ozone uptake by vegetation.
+\chem{CO_2} uptake coniferous trees can occur as early as doy 100 \citep{TB:Kolari2007}.
+The average ozone concentration at Svanvik is lower than at the other sites, as already mentioned, which is in league with a decrease in tropospheric ozone with decreasing altitude.
 In July--September, ozone is occasionally almost completely depleted during night time.
@@
 Density distributions are shown together with multi-annual mean of daily ozone concentrations \chem{[O_3]}.
-Splines have been fitted though daily mean and max \chem{[O_3]}.
+Splines have been fitted though daily mean \chem{[O_3]}.
 The climatology of daily mean ozone is shown together with $1\,\sigma$ standard deviation and standard error of the mean.
@@
   \includegraphics[width=8.3cm]{ozone_climatologies_portrait}
-  \caption{Comparison of climatologies derived from observational data (EBAS, see Fig.~\ref{fig:ozone_climatology_fenoscandic_obs}) with various ozone reanalysis products: (a) ECMWF: MACC reanalysis; (b) ECMWF: CAMS reanalysis; (c) Copernicus: Regional model reanalysis ensemble mean.
+  \caption{Comparison of climatologies derived from observational data (EBAS) with various ozone reanalysis products: (a) ECMWF: MACC reanalysis; (b) ECMWF: CAMS reanalysis; (c) Copernicus: Regional model reanalysis ensemble mean.
 As representative for the sites, the nearest neighboring grid point has been chosen.
@@
 If the data collected at Svanvik had been complete in 2018 and the climatology had extended into the resent past, we could have assessed the significance of ozone anomalies in the same way as for temperature, precipitation, and global irradiance.
-In Section~\ref{subsubsec:clim_ozone}, we described the construction of climatologies for northern Fennoscandia based on long term observations at Esrange, Jergul/Karasjok, and Pallas.
+In Section~\ref{subsubsec:clim_ozone}, we described the construction of climatologies for northern Fennoscandia based on long-term observations at Esrange, Jergul/Karasjok, and Pallas.
 For Svanvik this assessment is based on rather uncertain observations from 1986--1996.
@@
 \citet{ACP:Bueker2012} discuss several formulations available to parameterize the reduction in stomata conductance due to water content in the soil.
-Here we use the relationship between reative $g_\mathrm{sto}$ and soil water potential (SWP) given by:
+Here we use the relationship between relative $g_\mathrm{sto}$ and soil water potential (SWP) given by:
 %
@@
 %
-where SWP is the soil water potential across the root zone and $\mathrm{SWP_{min}}$ and $\mathrm{SWP_{max}}$ are the paramters describing the $f_\mathrm{SW}$ relationship.
+where SWP is the soil water potential across the root zone and $\mathrm{SWP_{min}}$ and $\mathrm{SWP_{max}}$ are the parameters describing the $f_\mathrm{SW}$ relationship.
 
@@
 %
-Where $L$ is the cross-wind leaf dimension, $u(z_1)$ the wind speed at height $z_1$, and the factor $1.3$ accounts for the diffusivities between heat and \chem{O_3}.
+Where $L$ is the cross-wind leaf dimension, $u(z_1)$ the wind speed at height $z_1$, and the factor $1.3$ accounts for the diffusivity between heat and \chem{O_3}.
 
@@
 
-Default PFTs are derived from \citet{ICP:MappingManual2017} and will be referred to as \emph{mapping manual} abbrieved MM.
+Default PFTs are derived from \citet{ICP:MappingManual2017} and will be referred to as \emph{mapping manual} abbreviate MM.
 A comprehensive list of these model parameters can be found in Supplement~Table~S1.
-Intitial test simulations with the mapping manual derived PFTs showed unrealistically low $g_\mathrm{sto}$ in 2019 (Supplememt~Figs.SX--SX+2) particularly for grassland.
-Only during an extended periode in late July did we find substential stomatal conductance.
+Initial test simulations with the mapping manual derived PFTs showed unrealistically low $g_\mathrm{sto}$ in 2019 (Supplement~Figs.SX--SX+2) particularly for grassland.
+Only during an extended period in late July did we find substantial stomatal conductance.
 In ecological terms, this would mean that there was almost no growth of grass in the summer of 2019 -- a prediction which is easily falsified by reality.
@@
 We propose bespoke parameterizations for the main vegetation types Norway spruce, birch, and perennial grassland better adapted to the climatic conditions in the target region.
-Note that these parameterizations represent hypothetical PFTs or species that have to be varified by experiements.
-
-We hypothize that perannial vegetation will likely prioritize their growth or fast adolesence throughout the short, sub-boreal growing season and hence their carbon sequestration.
-Since the $\mathrm{DO_3SE}$ in its current state does not simulate $A_\mathrm{n}$, we assume for out purpose a first order proptionality between carbon uptake (photosynthesis) and $g_\mathrm{sto}$ and tune the species temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) seperately for higher stomatal conductance.
-For this purpose, we graphically intersected propability density functions (PDFs) of climatological data for temperature and global irradiace with the respective response function (Fig.~\ref{fig:f_temp_grassland} for grassland; for Norway spruce and birch see Appendix Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}).
-Formally, a parameterization is considered more optimial the larger the intersection area between the response function and PDF is.
-
-For the climatization of the temperature response functions, we assume two different types: \emph{boreal} and \emph{cold}.
+Note that these parameterizations represent hypothetical PFTs or species that have to be verified by experiments.
+
+We hypothesize that perennial vegetation will likely prioritize their growth or fast adolescence throughout the short, sub-Boreal growing season and hence their carbon sequestration.
+Since the $\mathrm{DO_3SE}$ in its current state does not simulate $A_\mathrm{n}$, we assume for out purpose a first order proportionality between carbon uptake (photosynthesis) and $g_\mathrm{sto}$ and tune the species temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) separately for higher stomatal conductance.
+For this purpose, we graphically intersected probability density functions (PDFs) of climatological data for temperature and global irradiance with the respective response function (Fig.~\ref{fig:f_temp_grassland} for grassland; for Norway spruce and birch see Appendix Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}).
+Formally, a parameterization is considered more optimal the larger the intersection area between the response function and PDF is.
+
+For the acclimatization of the temperature response functions, we assume two different types: \emph{Boreal} and \emph{cold}.
 In either cases, we further assume that the hypothetical species are more tolerant to cold conditions than the mapping manual default suggests and shift $f_\mathrm{T}$ towards colder temperatures.
 We construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to the mapping manual default.
-In the same way, \emph{boreal} refers to a hypothetical species which is very tolerant to cold but is sensitive to high temperatures, and most efficient at cool temperatures (close to the climatological mean temperature of 1992--2000).
+In the same way, \emph{Boreal} refers to a hypothetical species which is very tolerant to cold but is sensitive to high temperatures, and most efficient at cool temperatures (close to the climatological mean temperature of 1992--2000).
 All three types for grassland are depicted together with the temperature PDF in Fig.~\ref{fig:f_temp_grassland}a).
 
-For Norway spruce, our hypothetical PFTs $T_\mathrm{opt}$ are motivated by a longer timeseries of both observed \chem{CO_2} uptake and temperature at two observation sites in Southern (\emph{cold}) and Northern (\emph{boreal}) Finland presented in \citet{TB:Kolari2007}.
+For Norway spruce, our hypothetical PFTs $T_\mathrm{opt}$ are motivated by a longer time series of both observed \chem{CO_2} uptake and temperature at two observation sites in Southern (\emph{cold}) and Northern (\emph{Boreal}) Finland presented in \citet{TB:Kolari2007}.
 We assume a similar adaption in birch.
@@
 
-For the the senisitivity regarding light, we assume that the plants may react either faster or slower to a change at lower light intensities.
+For the the sensitivity regarding light, we assume that the plants may react either faster or slower to a change at lower light intensities.
 We analytically derive the inverse function $f_{\mathrm{light},k}^{-1}$ for each PFT or species $k$
@@
 \caption{Construction of bespoke response functions for grassland.
-(a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown toghether with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability densitiy function - PDF), respectively.
+(a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively.
 Original mapping manual parameterization is shown in comparison as solid line.
@@
 \begin{table}[t]
-  \caption{Bespoke temperature prameterizations. MM refers to mapping manual \citep{GCB:Mills2010,ICP:MappingManual2017}.}
+  \caption{Bespoke temperature parameterizations. MM refers to mapping manual \citep{GCB:Mills2010,ICP:MappingManual2017}.}
   \label{tab:sensitivity_tests_temp}
@@
 \begin{table}[t]
-  \caption{Bespoke light prameterizations. MM refers to mapping manual \citep{GCB:Mills2010,ICP:MappingManual2017}.}
+  \caption{Bespoke light parameterizations. MM refers to mapping manual \citep{GCB:Mills2010,ICP:MappingManual2017}.}
   \label{tab:sensitivity_tests_light}
@@
 
-{\bf TODO: Write about boreal and cold parameter sets and their justification.
+{\bf TODO: Write about Boreal and cold parameter sets and their justification.
 Write about suggested method for optimizing PFTs in a simple but physiological way based on optimizing maximum climatological mean $g_\mathrm{sto}$ and morning responds.
-If plants have a short growing season up north optimizing \chem{CO_2} uptake around noon (highest light intensity) and speed up opening of stomata at lower intensities (morning), additional contraint lower climatological variance?}
-An idea to justify our choice based on the fullset of climatological data for Svanvik is shown in Fig.~\ref{fig:javis_func_opt_mean}.
+If plants have a short growing season up north optimizing \chem{CO_2} uptake around noon (highest light intensity) and speed up opening of stomata at lower intensities (morning), additional contained lower climatological variance?}
+An idea to justify our choice based on the full set of climatological data for Svanvik is shown in Fig.~\ref{fig:javis_func_opt_mean}.
 
@@
 From left to right: 2018, 2019.
-{\bf TODO: substitute with figure for boreal PFTs?}}
+{\bf TODO: substitute with figure for Boreal PFTs?}}
   \label{fig:pody}
@@
 
+Further stomatal conductance limit induced through ozone damage itself - varying strength and even sign (sluggishness) can shift the uptake into both directions
+
 Bigger picture:
@@
 \item Ozone damage assessment using mapping manual "compromise" parameterization.
-\item Land-atmosphere modelling: Problems in productivity of species especially in the arctic regions due to PFTs which are not suited for the climate.
+\item Land-atmosphere modeling: Problems in productivity of species especially in the arctic regions due to PFTs which are not suited for the climate.
 Automating optimization of climate adaption of general PFTs.
 \end{itemize}
+
+GS:
+The timing and trend in the SGS may differ substantially between individual species as compared to the thermal SGS though.
+Species-specific methods to estimate the SGS (Karlsson et al., 2018b) resulted in later SGS for European silver birch (Betula pendula) and earlier for Norway spruce (Picea abies), compared to the thermal SGS (Figure 4).
+The SGS of birch and spruce has also shifted to earlier in the year over time (significantly negative trend since 1994, -0.96 days year-1 and -0.98 days year-1 respectively) in northern Sweden.
+Wallin et al (2013) measured gas-exchange of 40-year-old Norway spruce trees with branch cuvettes in stands at Flakaliden in northern Sweden.
+Measurements of light-saturated photosynthesis were made during spring recovery from dormancy during three years 2002–2004.
+The recovery of gas-exchange from winter dormancy was non-linear over time and the first recovery, up to 60\% of summer maximum gas-exchange, occurred already in March, more than a month before the start of growing season calculated as when the 24-h mean temperatures exceeded 5°C for five consecutive days.
+
+Manninen et al.
+(2009a) summarized the results from open-top chamber (OTC) experiments performed in northern Finland where Scots pine and northern birch provenances were exposed to 36 – 54 ppb O3 as mean daylight concentrations.
+Biomass reductions of both downy birch (Betula pubescens) and mountain birch seedlings were demonstrated after one-season exposure.
+Moreover, the root:shoot ratio was reduced and the leaves of the mountain birch in the high O3 treatments showed necrotic stipples already in June.
+Mountain birch has a very rapid phenological development during spring-early summer that makes it difficult for this species to compensate for damaged leaves.
+It is also noteworthy that N fertilization increased the total biomass production of mountain birch markedly only under the ambient O3 both under the current climate (open-field plots) and in the about 1°C warmer climate (non-filtered air OTCs) but not under elevated O3 (Manninen et al.
+2009b).
+In contrast to mountain birches, northern Scots pine of local origin only showed invisible needle-level O3 impacts, such as increased respiration, decreased level of polyamines and reduced chlorophyll/carotenoid ratio and increased amount of epicuticular waxes after three years of exposure (Manninen et al.
+2009a).
+There is no field data comparing the O3 sensitivity of birches and Scots pine from northern Europe, but both Scots pine and Norway spruce have been ranked as more O3-tolerant species than deciduous birches and Alnus incana based on crown condition in a Lithuanian study (Girgzdiene et al.
+2009).
+Despite the relatively high O3 tolerance of Scots pine, invisible needle injuries may not be excluded in early spring conditions with high O3 concentrations though (Augustaitis and Bytnerowicz 2008, Huttunen and Manninen 2013).
 
@@
 \caption{Construction of bespoke response functions for Norway spruce.
-(a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown toghether with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability densitiy function - PDF), respectively.
+(a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively.
 Original mapping manual parameterization is shown in comparison as solid line.
@@
 \caption{Construction of bespoke response functions for Norway birch.
-(a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown toghether with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability densitiy function - PDF), respectively.
+(a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively.
 Original mapping manual parameterization is shown in comparison as solid line.

### 2021-03-02 VegOzone.tex
Start to revise Section 4.

@@
 \label{sec:do3se}
-To investigate which of the environmental factors lead to the observed ozone damage on clover and tobacco in 2018 in contrast to 2019, when no such damage was found on the clovers but only on the sensitive tobacco cultivar, we model the ozone uptake using the $\mathrm{DO_3SE}$ model and give an estimate on biomass reduction due to ozone uptake.
+To investigate which of the environmental factors lead to the observed ozone damage on clover and tobacco in 2018 in contrast to 2019, when no such damage was found on the clovers but only on the sensitive tobacco cultivar, we model ozone uptake of different natural and semi-natural vegetation using the $\mathrm{DO_3SE}$ model and develop bespoke subarctic prameterizations for common plant species.
+In the following, we give an account of the $\mathrm{DO_3SE}$ model and the used method for deriving the bespoke parameters.
 
@@
 \label{subsec:model_describtion}
-
 The opening of stomata is governed by different species-specific and environmental factors and can be described by a multiplicative model \citep{PTRS:Jarvis1976,ICP:MappingManual2017}:
@@
 
-We hypothesize that perennial vegetation will likely prioritize their growth or fast adolescence throughout the short, sub-Boreal growing season and hence their carbon sequestration.
-Since the $\mathrm{DO_3SE}$ in its current state does not simulate $A_\mathrm{n}$, we assume for out purpose a first order proportionality between carbon uptake (photosynthesis) and $g_\mathrm{sto}$ and tune the species temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) separately for higher stomatal conductance.
-For this purpose, we graphically intersected probability density functions (PDFs) of climatological data for temperature and global irradiance with the respective response function (Fig.~\ref{fig:f_temp_grassland} for grassland; for Norway spruce and birch see Appendix Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}).
+We hypothesize that perennial vegetation will likely prioritize their growth or fast adolescence throughout the short, subarctic growing season and hence their carbon sequestration.
+Since the $\mathrm{DO_3SE}$ in its current state does not simulate $A_\mathrm{n}$, we assume for our purpose a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011}.
+We then tune the species temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) separately for higher stomatal conductance.
+We illustrate this procedure graphically by intersecting probability density functions (PDFs) of climatological data for temperature and global irradiance with the respective response function (Fig.~\ref{fig:f_temp_grassland} for grassland; for Norway spruce and birch see Appendix Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}).
 Formally, a parameterization is considered more optimal the larger the intersection area between the response function and PDF is.

### 2021-03-03 supplements/VegOzone_supplement.tex
Add supplement material

--- 2021-03-01 VegOzone.tex
@@
-%% Copernicus Publications Manuscript Preparation Template for LaTeX Submissions
-%% ---------------------------------
-%% This template should be used for copernicus.cls
-%% The class file and some style files are bundled in the Copernicus Latex Package, which can be downloaded from the different journal webpages.
-%% For further assistance please contact Copernicus Publications at: production@copernicus.org
-%% https://publications.copernicus.org/for_authors/manuscript_preparation.html
-
-%% Please use the following documentclass and journal abbreviations for preprints and final revised papers.
-
-%% 2-column papers and preprints
-\documentclass[bg, manuscript]{copernicus}
-
-%% Journal abbreviations (please use the same for preprints and final revised papers)
-
-% Advances in Geosciences (adgeo)
-% Advances in Radio Science (ars)
-% Advances in Science and Research (asr)
-% Advances in Statistical Climatology, Meteorology and Oceanography (ascmo)
-% Annales Geophysicae (angeo)
-% Archives Animal Breeding (aab)
-% ASTRA Proceedings (ap)
-% Atmospheric Chemistry and Physics (acp)
-% Atmospheric Measurement Techniques (amt)
-% Biogeosciences (bg)
-% Climate of the Past (cp)
-% DEUQUA Special Publications (deuquasp)
-% Drinking Water Engineering and Science (dwes)
-% Earth Surface Dynamics (esurf)
-% Earth System Dynamics (esd)
-% Earth System Science Data (essd)
-% E&G Quaternary Science Journal (egqsj)
-% European Journal of Mineralogy (ejm)
-% Fossil Record (fr)
-% Geochronology (gchron)
-% Geographica Helvetica (gh)
-% Geoscience Communication (gc)
-% Geoscientific Instrumentation, Methods and Data Systems (gi)
-% Geoscientific Model Development (gmd)
-% History of Geo- and Space Sciences (hgss)
-% Hydrology and Earth System Sciences (hess)
-% Journal of Bone and Joint Infection (jbji)
-% Journal of Micropalaeontology (jm)
-% Journal of Sensors and Sensor Systems (jsss)
-% Magnetic Resonance (mr)
-% Mechanical Sciences (ms)
-% Natural Hazards and Earth System Sciences (nhess)
-% Nonlinear Processes in Geophysics (npg)
-% Ocean Science (os)
-% Primate Biology (pb)
-% Proceedings of the International Association of Hydrological Sciences (piahs)
-% Scientific Drilling (sd)
-% SOIL (soil)
-% Solid Earth (se)
-% The Cryosphere (tc)
-% Weather and Climate Dynamics (wcd)
-% Web Ecology (we)
-% Wind Energy Science (wes)
-
-%% \usepackage commands included in the copernicus.cls:
-%\usepackage[german, english]{babel}
-%\usepackage{tabularx}
-%\usepackage{cancel}
-%\usepackage{multirow}
-%\usepackage{supertabular}
-%\usepackage{algorithmic}
-%\usepackage{algorithm}
-%\usepackage{amsthm}
-%\usepackage{float}
-%\usepackage{subfig}
-%\usepackage{rotating}
-
-\graphicspath{{pictures/}}
-
-\begin{document}
-
-%\title{The 2018 heatwave and its implications on ozone induced damage on vegetation in a subarctic climate}
-\title{Characterizing subarctic biomes for land surface modeling of pollution and climate risk}
-
-% \Author[affil]{given_name}{surname}
-
-\Author[1]{Stefanie}{Falk}
-\Author[2]{Ane Victoria}{Vollsnes}
-\Author[2]{Aud E. B.}{Eriksen}
-\Author[3]{Lisa}{Emberson}
-\Author[3]{Connie}{O'Neill}
-\Author[1]{Frode}{Stordal}
-\Author[1]{Terje}{Koren Berntsen}
-
-\affil[1]{Department of Geosciences, University of Oslo, Oslo, Norway}
-\affil[2]{Department of Biosciences, University of Oslo, Oslo, Norway}
-\affil[3]{Department of Environment and Geography, University of York, UK}
-
-%% The [] brackets identify the author with the corresponding affiliation. 1, 2, 3, etc. should be inserted.
-
-%% If an author is deceased, please mark the respective author name(s) with a dagger, e.g. "\Author[2,$\dag$]{Anton}{Aman}", and add a further "\affil[$\dag$]{deceased, 1 July 2019}".
-
-%% If authors contributed equally, please mark the respective author names with an asterisk, e.g.
-"\Author[2,*]{Anton}{Aman}" and "\Author[3,*]{Bradley}{Bman}" and add a further affiliation: "\affil[*]{These authors contributed equally to this work.}".
-
-\correspondence{Stefanie Falk (stefanie.falk@geo.uio.no)}
-
-\runningtitle{Subarctic biomes for land surface modelling}
-
-\runningauthor{Falk et al.}
-
-\received{}
-\pubdiscuss{} %% only important for two-stage journals
-\revised{}
-\accepted{}
-\published{}
-
-%% These dates will be inserted by Copernicus Publications during the typesetting process.
-
-\firstpage{1}
-
-\maketitle
-
-\begin{abstract}
-Here we assess the importance of characterization of key meteorological variables and air pollution profiles on the modeled susceptibility of subarctic vegetation to ozone.
-Changes in climate (e.g.
-temperature, water availability) will manifest in both direct and indirect effects influencing the species' sensitivity to air pollution.
-A likely increase in frequency of persistent heatwaves in the future, will inflict heat related stress while generating favorable conditions for the formation of ambient air ozone, e.g.
-due to wildfires.
-We have monitored ambient air ozone concentrations at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd in Northern Norway in the vegetation period 2018/19.
-Due to data acquisition problems, ozone concentrations for two weeks in July 2018 were missing from our record.
-We present a reconstruction based on probability density function with respect to the Swedish and Finnish atmospheric monitoring sites and compare with available surface ozone products for the region.
-Over all, ozone concentrations did not differ significantly between the two years.
-While temperatures and global irradiance diverged substantially from the multi annual mean precipitation varied only to some extend.
-We investigate the difference in uptake of ozone using the $\mathrm{DO_3SE}$ model, for typical vegetation at the location.
-We find that bespoke parameterization for subarctic vegetation is essential for an accurate estimation of Phototoxic Ozone Dose (POD) and an assessment of damage risk.
-Accurate estimates of the start and length of growing season are also crucial.
-We find that temperature and its derivative vapor pressure deficit are likely the most important climate variables while the variability of ozone concentration within climatologically reasonable limits plays a minor role.
-The use of subarctic parameterization suggests an improvement in risk assessment as POD values are increased in 2018; corroborated by evidence of visible injury in the Svanhovd ozone garden.
-\end{abstract}
-
-%\copyrightstatement{TEXT}
-
-\introduction  %% \introduction[modified heading if necessary]
-\label{sec:intro}
-
-Ground level ozone is a highly toxic pollutant known to cause damage to a variety of ecosystems around the World \citep{PT:Emberson2020}.
-At continuously high concentration levels, ozone negatively affects life expectancy of humans \citep{WHO2008} and is a cause of visible injury, photosynthetic damage, early senescence as well as programmed cell death of plants \citep{PCE:Kangasjarvi2005}.
-Annual global yield losses of four major crops (wheat, rice, maize, and soybean) of $3-15\,\unit{\%}$ \citep{PJ:Ainsworth2017} and a suggested loss in primary production in forestry of $7\,\unit{\%}$ \citep{GCB:Wittig2009,EP:Matyssek2012} has been attributed to ozone.
-{\bf TODO: Wealth of evidence demonstrating ozone damage to vegetation (perhaps focus on key evidence from northern climates and key species (e.g.
-birch, Norway srpuce etc...
-- this might help to make it clear why we focus on these species with the PODy modelling))}
-The crop yield loss studies indicate a threat to food security in rapidly developing countries, e.g., in East and South-East Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018}, where ozone concentrations related to enhanced pollutant emissions from the transport, residential, power generation and industrial sectors have increased in the past few decades.
-Even though long-term (pre 1950) observations of surface ozone concentrations are scarce; evidence from trend analysis of those few sites suggests that since the industrial revolution, background concentrations in the northern Hemisphere have at least doubled and continued to increase until relatively recently \citep[Chapter 2]{IPCC2013}.
-Recent trend analyses indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric ozone started to level-off or even decline \citep{ESA:Cooper2014, ACP:Wespes2018,ESA:Gaudel2018}.
-This can be attributed to a successful implementation of air quality regulations coupled with economic restructuring which saw a move from coal to gas, reduced the number of episodes of peak concentrations especially in summertime over Europe and North America \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
-However, there are indications that an increasing frequency of thermal stress accompanying climate change will reduce vegetations' capacity for ozone uptake and therefor scotch these achievments \citep{NCC:Lin2020}.\\
-
-Ground level ozone is formed in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) and hydrocarbons known as volatile organic substances (VOCs) in the presence of nitrogen oxides (\chem{NO_x}).
-Despite a relatively short tropospheric life-time of approximately $22\,\unit{days}$, ranging from a few days in the tropical boundary layer to up to $1\,\unit{year}$ in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}, ozone is subject to advection over long distances.
-Episodes of high ozone concentrations in the absence of local pollution sources can therefor often be attributed to long-range transport of both ozone and its precursors.
-In northern Fennoscandia, episodes of enhanced ozone concentrations in 2003 and 2006 have been traced back to ozone precursor emissions related to forest fires in southern and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}.
-In addition, hydrocarbons are also emitted by vegetation itself in form of terpenes and monoterpenes, so called biogenic volatile substances (BVOCs).
-These emissions are associated with, e.g., thermal stress, defense against herbivores, or even attraction of pollinators \citep{TPS:Penuelas2003}.
-
-In the annual cycle in boreal and temperate climates, a distinct maximum of ground level ozone concentrations typically occurs in spring followed by lower concentrations throughout the summer.
-Several conditions unique to the Actic are stimulating the formation of this spring peak.
-Major pathways of ozone and its precursors' removal from the boundary layer include dry deposition on vegetated surfaces and photochemistry \citep{RG:Clifton2020}.
-As dry deposition to bare ground or snow and ice covered surfaces was found to be low \citep{ACP:Helmig2007}, both pathways of removal are substantially suppressed during polar night \citep{AE:Monks2000}.
-This leads to a build up of ozone and its precursors in the course of winter.
-Come spring precursors are photochemical activated again, while snow cover typically prevails for a longer period of time continuing to suppress ozone dry deposition.
-On the other hand, very high ozone concentrations occuring over shorter periodes of time can often be attributed to deep tropopause folding events over the Actic during winter.
-These events facilitate an intrusion of stratospheric air masses which are enriched in ozone \citep{JGR:Skerlak2015}.
-The spring peak is more pronounced in northern Fennoscandia than in more southern locations but shows a  high interannual variability \citep{AB:Klingberg2009, BER:Klingberg2019}.
-A cause for this is the complex interplay between atmospheric dynamics and chemistry \citep{AE:Laurila1996,BER:Hatakka2003} as described above as well as anthropogenic and biogenic activity \citep{AE:Rummukainen1996,AE:Simpson2002,QJRMS:Galbally2007,NGS:Schnell2009}.
-\citet{ACP:Andersson2017} showed in a modeling study focused on Fennoscandia that the variability of ozone concentrations in winter can be attributed mainly to changes in atmospheric background (transport) of ozone, while summertime abundance is mostly affected by emissions of precursors in the rest of Europe.
-Meteorology plays an important role throughout the year.
-
-The start of growing season in northern Fennoscandia has been shifting to earlier dates in response to the general warming trend \citep[e.g.,]{GCB:Menzel2006,RS:Hogda2013,IJB:Karlsen2007} and thereby converging with the period of the ozone spring peak.
-At the same time the growing season is also becoming longer.
-In conjunction with midnight sun conditions (almost $24\,\unit{h}$ photosynthetic activity), a longer growing season, is prolonging the time in which vegetation can accumulate ozone.
-These factors may promote a higher potential susceptibility of northern vegetation to ozone in the future.
-
-Ozone acts as an oxidative stress to plants.
-Its main action is imposed through reactions occurring in the cell walls and cell membranes of mesophyll cells inside the leaves.
-Ozone enters the leaves through stomata, that are kept open to allow for gas exchange allowing the entry of \chem{CO_2} for photosynthesis and loss of \chem{H_2O} vapour via the plant transpiration stream.
-Stomatal aperture, and hence stomatal conductance to these gases (including ozone) will vary over the course of the day and growing season primarily to balance \chem{CO_2} uptake against \chem{H_2O} vapour loss.
-The higher the stomatal conductance, the higher the potential for ozone uptake.
-Stomatal conductance has been emperically linked to environmental factors such as air temperature, photosynthetic active radiation, vapor pressure deficit, and soil water potential as well as photosynthesis itself.\\
-
-To measure the risk of damage inflicted to vegetation through ozone, different metrics have been defined.
-Two main types can be differentiated, concentration-based or flux-based metrics, e.g., Accumulated Over a Threshold $\mathrm{x}$ ($\mathrm{AOT_x}$) and Phyto-toxic Ozone Dose over a threshold $\mathrm{y}$ ($\mathrm{POD_y}$), respectively.
-Due to their simplicity, concentration-based metrics are still widely used.
-Despite comparisons of exposure-based methods with previously used concentration based methods (e.g.
-$\mathrm{AOT_{40}}$) show that $\mathrm{POD_y}$ is better suited to represent the risk to vegetation from ozone, especially in the more climate extreme parts of Europe \citep{GCB:Mills2010,ICP:MappingManual2017}.
-
-The concentration-based metric $\mathrm{AOT_x}$ (in units of $\unit{ppb}$) is defined as:
-\begin{equation}
-  \mathrm{AOT_x} = \int{\mathrm{max(\chem{[O_3]}-x)}\cdot \mathrm{dt}},
-  \label{eq:aot}
-\end{equation}
-with the ambient ozone concentration $\chem{[O_3]}$. The  time over which the integration is calculated varies, but is usually defined from $8\,\unit{am}-8\,\unit{pm}$.
-
-For computing the $\mathrm{POD_y}$, an estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf.
-Details specific to the $\mathrm{DO_3SE}$ model are given in Section~\ref{sec_do3se}.
-$\mathrm{POD_y}$ is then calculated according to:
-%
-\begin{equation}
-  \mathrm{POD_y} = \int{(\Phi_\mathrm{sto}-y)\cdot \mathrm{dt}},
-  \label{eq:pod}
-\end{equation}
-%
-with the hourly averaged ambient air ozone flux through the stomata Eq.~(\ref{eq:flux_stomata}) and a flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}\,s^{-1}}.
-The flux threshold $\mathrm{y}$ is representing the detoxification potential of the plant and is typically only exceeded during daylight hours (i.e.
-when global radiation is above $50\,\unit{W\,m^{-2}}$).
-
-Finally, exceedance of the flux-based critical level (CL) is calculated by:
-%
-\begin{equation}
-  \mathrm{CL_{exeed}} = \mathrm{POD_y} - \mathrm{CL}.
-\end{equation}
-%
-
-According to a report by the Norwegian Meteorological institute \citep{MetNOR2019}, the summer of 2018 was the warmest and driest ever recorded in eastern, western, and southern Norway.
-In the north (including Finnmark), it was amongst the warmest on record -- favorable conditions for planetary boundary layer ozone formation.
-An unusually weak and northward shifted jet stream allowed for a persistent high pressure system above northern Europe, including Fennoscandia which blocked the pressure systems for several consecutive months.
-In the period May -- July, southern Norway had temperatures $4\,\unit{^\circ C}$ above normal.
-While southern Norway had only about $60\,\unit{\%}$ of normal precipitation, northern Norway as a whole had close to normal precipitation, but with local variations.
-Thermal stress on vegetation was exceptional not only in large parts of Fennoscandia but also in much of Europe, where the influence of the high pressure system extended even over a period of five months (April/May, July--September).
-These conditions gave rise to massive forest fires in different parts of Europe and thus an increase in ozone precursors.
-Boreal wildfires emit in addition to \chem{CO_2} also \chem{CO} ($\chem{[CO]}/\chem{[CO_2]} \propto 6-13\,\unit{\%}$) and \chem{VOCs} ($\chem{[VOCs]}/\chem{[CO_2]} \propto 0.5-1.5\,\unit{\%}$) \citep{AE:Cofer1990}.
-A total of 2079 forest fires of different size were registered in Norway in 2018, twice as many as in the preceding years 2016/17 \citep{DSB2019}, last accessed April 2020).
-In Sweden, about 500 fires had been reported (five times more than in a usual summer) and an estimated total of $25000\,\unit{hectare}$ burned down in the course of the most extensive fires in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) in younger history \citep{SOU2019}.
-Coincident peak \chem{[O_3]} are found in ozone monitoring data from Svanhovd but also over whole northern Fennoscandia in July (see Fig.~\ref{fig:data_svanvik_2018_2019} or Fig.~\ref{fig:ozone_reconstruction_2018_07}).
-Like in the 2003 drought period, elevated ozone in Europe was likely promoted by a combination of various other factors as well, such as biogenic VOC emissions, reduced cloud and hence increased solar radiation, faster photochemical reactions, and reduced dry deposition and turbulent mixing due to stagnant weather conditions \citep{JGR:Solberg2018}.\\
-
-These methods described above rely on accurate representation of plant physiological response to climate variables (i.e.
-the species or plant funtional type specific responses used to parameterise risk assessment models) and are most often parameterized using data for less extreme climates, i.e., temperate or continental climates; and to a lesser extent, boreal and Mediterranean climates.
-Parameterization of these models for more extreme bio-geographical regions such as sub-arctic climates has not been perfromed and relies on assuming that less extreme boreal parameterizations will be applicable.
-It is important to understand the implications of using these more generic parameterizations for more extreme climate regions, here we explore how risk estimates might be over or underestimated using generic vs bespoke sub-arctic parameterisations.
-we also explore the role [O3] profile and changing climate variables to understand how this risk might alter in the future with the advent of climate change.
-This allows both an understanding of how the threat from ozone may change over time but also which are the key aspects of plant physiology that might determine risk (and hence which parameters might require special attention to get right) both for pollution impact assessment, but also more generally for describing gas exchange response to changing climates that will influence bio-geochemcial cycling.
-This might be one key to solve long-standing issues of coupled earthsystem models (ESM) which generally lack PFTs representing sub-arctic conditions in their landsurface model.
-
-Within the scope of our project to assess probable effects of midnight sun conditions on ozone uptake by plants and resultant enhanced susceptibility to ozone inflicted damage, we conducted ambient air ozone concentration measurements at Svanhovd in the growing seasons of 2018/19 in collaboration with the Norwegian Institute for Air Research (NILU).
-In 2018, we have observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd.
-In contrast, no such damage was found on the clovers in 2019, although it was found on the sensitive tobacco cultivar.
-This indicates that the vegetation may have been more affected by ozone in 2018 than in 2019.
-In terms of $\mathrm{AOT_{40}}$, the threshold for protection of vegetation of $3000\,\unit{ppb}$ (integrated over the growing season) given by the EU directive has been well exceeded at many ozone monitoring sites in Southern Norway in the 2018 growing season \citep{NILU2019} but not at Svanhovd.
-
-In this study, we investigate potential main drivers for the observed ozone damage in the growing season 2018 at NIBIO Environment Centre Svanhovd in contrast to 2019.
-We compare $\mathrm{POD_y}$ values for natural and semi-natural vegetation for derived climatological averages of environmental conditions at Svanhovd as well as for the growing seasons 2018/19 which shall serve as case studies.
-
-The observation site at NIBIO Environment Centre Svanhovd and all related data are presented in Section~\ref{sec:data}.
-In Section~\ref{sec:stats}, we derive climatologies for surface ozone concentration at Svanhovd and northern Fennoscandia based on in situ observations and also compute climatologies of surface temperature, precipitation, and global radiation for Svanhovd.
-Based on these, we evaluate anomalies of environmental key variables in 2018/19 by statistical means.
-In addition, the derived surface ozone concentration climatology for northern Fennoscandia will be compared with available ozone reanalysis products.
-We use the $\mathrm{DO_3SE}$ model to estimate the ozone uptake by natural and semi-natural vegetation to investigate main environmental drivers for the observed ozone damage in 2018 (Section~\ref{sec:do3se}).
-In this context, we introduce the$\mathrm{DO_3SE}$ model equations and derive bespoke parameterizations for sub-arctic species to assess systematic uncertainties arising from the use of default parameterizations.
-In Section~\ref{sec:conc} finally, we will summarize our results, draw conclusions, and give an outlook on further work.
-
-\section{Data}
-\label{sec:data}
-
-In our assessment of ozone damage on vegetation in the subarctic in relation to the 2018 heatwave, we focus on the NIBIO operated Environment Centre Svanhovd which lies close by the settlement of Svanvik in the Pasvik valley (Section~\ref{subsec:svanhovd}).
-Throughout the paper, we will use Svanhovd synonymously for the atmospheric monitoring site at Svanvik.
-The available meteorological monitoring data are described in Section~\ref{subsubsec:atmo_svanvik}.
-In Section~\ref{subsubsec:ozone_garden}, the ozone garden -- a plant-based ozone monitoring -- will be described in more detail.
-Other data used for evaluation of the ozone data taken at Svanhovd is described in Section~\ref{subsec:other_data}.
-We compare with observational data from other monitoring sites in northern Fennoscandia (Section~\ref{subsubsec:ebas}) and use as well reanalysis products (Section~\ref{subsubsec:ozone_rea}) as complement when observational data was not available.
-
-\subsection{NIBIO Environment Centre Svanhovd}
-\label{subsec:svanhovd}
-"Svanhovd is located beside the Pasvik river, in the middle of the wedge of Norwegian land separating Russia and Finland in the north.
-It is where the eastern Siberian taiga meets the western Boreal forest.
-[...] Svanhovd was originally established as a State Demonstration and Experimental Farm in the 1930's, in order to expand agriculture in Pasvik valley [...]" \citep[Svanhovd][]{NIBIO_Svanhovd}.
-
-The locations of the atmospheric monitoring site as well as the ozone garden with respect to the main buildings of the Svanhovd Environment Centre are marked in the aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}.
-The Centre is located south of the settlement of Svanvik, where as the border to Russia, the Pasvik river, lies only a few hundred meters to the east.
-The original farm buildings, protected by law, are located in the lower center, while the Svanhovd Environment Centre follows to the north-east.
-Left of the Centre and north of the original farm buildings a botanical garden can be found.
-It is one amongst three such gardens north of the Polar Circle.
-The Svanhovd Environment Centre focuses on research of the special environment of the region, but also comprises long-term observations of various agrometeorological variables, i.a.
-$2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths.
-
-For the purpose of relating observed damage on ozone sensitive plants (e.g.
-clover and tobacco) to local ozone concentrations, an ozone monitor had been installed during the growing seasons 2018/19.
-The relevant data are described in more detail in Section~\ref{subsubsec:atmo_svanvik}.
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{svanhovd_researchstation}
-  \caption{NIBIO Environment Centre Svanhovd close by the settlement of Svanvik, Norway. Atmospheric monitoring site and ozone garden have been marked. Aerial photography \copyright Norges Kartverk.}
-  \label{fig:svanhovd_research_station}
-\end{figure}
-
-\subsubsection{Atmospheric monitoring}
-\label{subsubsec:atmo_svanvik}
-For the growing seasons 2018/19, monitoring data of key variables (temperature, pressure, precipitation, and ozone) have been obtained from luftkvalitet.no which is operated by \citet{NILU_AIRQ}.
-
-All agrometeorological variables including temperature and precipitation are available from September 1992 to present day \citep[LandbruksMeteorologiske Tjeneste][note the station name here is Pasvik]{LMT_NIBIO}.
-
-Relevant data for the growing season 2018/19 are shown in Fig.~\ref{fig:data_svanvik_2018_2019}.
-The hatched areas mark times when no ozone data was recorded.
-Note, while the downtime during winter was planned, missing data in two weeks of July 2018 (July 9--23) were due to problems in data acquisition.
-
-Ozone concentrations \chem{[O_3]} measured in $2\,\unit{m}$ height above ground are averaged hourly.
-As can be seen in Fig.~\ref{fig:data_svanvik_2018_2019}a), \chem{[O_3]} peaks in spring (April/May) and reaches its minimum in late summer (July/August).
-The spring peak has not been captured completely in 2019, for data acquisition started later than in 2018.
-In summer 2018 (June--August), high ozone concentrations ($\chem{[O_3]} > 40\,\unit{ppb}$) were recorded 50 times.
-The highest summer ozone concentration ($\chem{[O_3]} = 50.2\,\unit{ppb}$) was measured on July 25.
-This coincides with the period of the most extensive forest fires in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) which occurred from July 12--29 and destroyed about $18000\,\unit{hectare}$ of forest \citep{SOU2019}.
-However, due to the above mentioned data acquisition problems, we missed most of the corresponding ozone data for this event and most likely also the peak ozone concentration.
-In contrast, ozone concentration only rose 18 times above the threshold of $40\,\unit{ppb}$ during the summer of 2019.
-
-Hourly averaged $2\,\unit{m}$ temperatures below $-20\,\unit{^\circ C}$ were observed in winter (January/February), while high temperatures (above $20\,\unit{^\circ C}$) occurred in summer (July), though more regularly in 2018 than in 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}b).
-In 2018, temperature regularly rose above freezing only in May, while in 2019 this occurred already early in March/April.
-2019 saw a very cold November with temperatures regularly below $-10\unit{^\circ C}$.
-
-More rain events with accumulated daily precipitation $\sum_d \mathrm{Precip}$ above $10\,\unit{mm}$ occurred in the summer of 2018 compared to 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}c).
-$\sum_d \mathrm{Precip} \ge 20\,\unit{mm}$ was only breached twice in 2018 and once in 2019.
-
-In 2018, global irradiance $Q_0$ (Fig.~\ref{fig:data_svanvik_2018_2019}d) displays more sequentially high values in spring (May) and summer (July) compared to 2019, while June 2019 saw more consecutive high irradiance than 2018.
-In both years, the maximum recorded $Q_0$ was roughly $750\,\unit{W\,m^{-2}}$ reached only in June.
-
 \begin{figure*}[t]
-  \includegraphics[width=12cm]{data_svanvik_2018_2019}
-  \caption{Observational data from atmospheric monitoring at Svanhovd in 2018/19.
-The hatched areas indicate missing ozone monitoring data.
-(a) Hourly averaged ozone concentration; (b) hourly averaged temperature; (c) daily accumulated precipitation; (d) hourly averaged global irradiance.}
-  \label{fig:data_svanvik_2018_2019}
-\end{figure*}
-
-\subsubsection{Ozone garden}
-\label{subsubsec:ozone_garden}
-An ozone garden consists of carefully selected plant species of which one cultivar is more sensitive to ozone than the other.
-At the NIBIO Environment Centre Svanhovd, the likely northern-most ozone garden to date, an ozone garden was installed in the course of our project.
-Cultivated species were, e.g.
-clover, tobacco, and beans.
-As shown in Fig.~\ref{fig:ozone_garden}, the plants had to be protected from herbivores with a wire-mesh fence.
-In 2018, we have observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden.
-In contrast, no such damage was found on the clovers in 2019, although it was found on the sensitive tobacco cultivar.
-This indicates that the vegetation may have been more affected by ozone in 2018 than in 2019.
-
-\begin{figure}
-  \includegraphics[width=8.3cm]{IMG_7739.JPG}
-  \caption{Clover in the Svanhovd ozone garden. The plants had to be secured against herbivores with a wire-mesh fence. The plants shown are approximately $6-12\,\unit{cm}$ in height.}
-  \label{fig:ozone_garden}
-\end{figure}
-
-\subsection{Other data}
-\label{subsec:other_data}
-To determine the significance of the ozone conditions in 2018, we have to derive a surface ozone concentration climatology for Svanvik.
-From 1986--1996 ozone monitoring has been operated by NILU at Svanhovd, but no ozone measurements have been conducted for the past $14\,\unit{years}$.
-Due to probable changes in northern hemispheric tropospheric background ozone and a reduction of summertime peak values related to ozone precursors which fell under air quality regulations in Europe, we also have to have a look at ozone observations at other sites in northern Fennoscandia.
-In regard to ozone, available observational data are sparse in this region.
-An overview over the geographical locations of conducted atmospheric monitoring in the past and present is given in Fig.~\ref{fig:station_map_fennoscandia}.
-The data is described in Section~\ref{subsubsec:ebas}.
-As additional complement to observational data which is not available at all times and spatially sparse, we also look at various ozone reanalysis products (Section~\ref{subsubsec:ozone_rea}).
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{station_map_fennoscandia}
-  \caption{Cap of the North.
-Locations of past and present ozone observation sites in northern Fennoscandia used within this study.
-For more details see Tab.~\ref{tab:ebas_obs}.
-The same color coding for each station as displayed herein will be used in all following figures.}
-  \label{fig:station_map_fennoscandia}
-\end{figure}
-
-\subsubsection{Surface ozone observations}
-\label{subsubsec:ebas}
-All surface ozone observations used in this study are available from the EBAS database operated by \citet{NILU_EBAS}.
-In Table~\ref{tab:ebas_obs}, we give an overview over the exact locations and time range of operational data taking.
-All stations with available long term observations are located on higher ground than Svanhovd which will result in higher surface ozone concentrations in comparison as ozone abundance, in general, decreases with decreasing altitude in the absence of local precursors \citep[e.g.,][]{AB:Klingberg2009}.
-The periphery vegetation at all sites is similar and consists mainly of pine forests, birch, and heath.
-However, at close vicinity it varies substantially.
-Ozone measurements at Pallas are taken on top of mount Sammaltunturi $100\,\unit{m}$ above tree line (pine forest) and mainly consists of low vascular plants, mosses, and lichen \citep{BER:Hatakka2003}.
-At Svanhovd, the ozone measurements are taken amidst a meadow surrounded by agriculturally managed herbaceous plants and grass (c.c.
-Fig.\ref{fig:svanhovd_research_station}).
-
-Data from before the 1990s is available solely from Svanvik and Karasjok.
-These data, however, do not follow the high quality standard procedures implemented nowadays and have to be treated with care \citep{NILU2003}.
-Especially, ozone monitors did not undergo a regular re-calibration.
-This probably leads to drifts in the observed data and may impose a false periodicity.
-In 1997, the ozone monitoring at Jergul was relocated downstream the river Karasjokka to a site close by Karasjok.
-Although the same equipment has been used, this might have introduced a breaking point in the combined data series, denoted as Jergul/Karasjok in the following, which would be of a problem if this was the sole data to compare with.
-Only few data have been taken upstream the Pasvik river at Janiskovski, hence these will not be taken into further consideration, but are shown for completion.
-As can be seen in Fig.~\ref{fig:ozone_timesseries_fenoscandic_obs}, there is an overlap of about $4\,\unit{years}$ between data taken at Esrange, Jergul/Karasjok, and Svanvik.
-The hatched areas indicate time periods with insufficient quality control as mentioned above.
-
-\begin{table}[t]
-  \caption{Past and present ozone observation sites in northern Fennoscandia. Data available on \href{http://ebas.nilu.no/}{EBAS}.}
-  \label{tab:ebas_obs}
-  \begin{tabular}{lllllrl}
-    \tophline
-    Name       & Country & ID      & \multicolumn{3}{c}{Location} & Operational\\
-    &         &         & lat             & lon               & alt            &\\
-    &         &         & (\unit{^\circ N}) & (\unit{^\circ E})  & (\unit{m})     &\\
-    \middlehline
-    Esrange    & SWE     & SE0013R & 67.83           & 21.07             & 475            & 1991 -- present\\
-    Janiskoski & RUS     & RU0001R & 68.93           & 28.85             & 118            & 1995 -- 1997\\
-    Jergul     & NOR     & NO0030R & 69.45           & 24.60             & 255            & 1997 -- 2011\\
-    Karasjok   & NOR     & NO0055R & 69.467          & 25.217            & 333            & 1988 -- 1997\\
-    Pallas     & FIN     & FI0096G & 67.97           & 24.12             & 565            & 1995 -- present\\
-    Svanvik    & NOR     & NO0047R & 69.45           & 30.03             & 30             & 1986 -- 1996$^\dagger$\\
-    \bottomhline
-  \end{tabular}
-  \belowtable{$^\dagger$ Exclusive ozone monitoring growing seasons 2018/19 operated by NILU.} % Table Footnotes
-\end{table}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{ozone_timeseries_fenoscandic_obs.png}
-  \caption{Time series of ozone observations in northern Fennoscandia (Tab.~\ref{tab:ebas_obs}).
-Data taken from EBAS.
-The hatched areas indicate time periods with insufficient quality control \citep{NILU2003}.}
-  \label{fig:ozone_timesseries_fenoscandic_obs}
-\end{figure*}
-
-\subsubsection{Ozone reanalysis}
-\label{subsubsec:ozone_rea}
-
-Details of reanalysis data sets referred to in this work are listed in Table~\ref{tab:ozone_rea}.
-There are two global reanalysis products available at the European Centre for Medium-Range Weather Forecasts (ECMWF) that include atmospheric tracers, in particular ozone (MACC and CAMSRA) \citep{ACP:Inness2013, ACP:Inness2019}.
-The temporal as well as spatial resolution of these reanalysis products is rather coarse: 3-hourly and $0.75^\circ\times 0.75^\circ$ or roughly $29.3\,\unit{km}\times 83.4\,\unit{km}$ at the location of Svanvik, respectively.
-In addition, higher resolution surface ozone reanalysis for whole Europe is available from Copernicus service for regional air quality control (note: since May 31st 2020 through the CAMS Atmosphere Data Store (ADS)).
-The regional air quality model reanalysis ensemble (CAMS regional Ensemble) is based on nine European state-of-the-art numerical air quality models \citep{Copernicus_Regio_AIRQ}.
-The ensemble mean is at fairly high spatial and temporal resolution compared to the global reanalysis from ECMWF: $0.1^\circ\times 0.1^\circ$ or roughly $3.9\,\unit{km}\times 11.1\,\unit{km}$ and 1-hourly, respectively.
-The covered time periods differ and no data are available for before the turn of the millennium.
-The CAMSRA and CAMS Regional Air Quality ensemble mean are available in near real time, but only former covers a period of sufficient length for climate analysis (2003 -- present).
-All reanalyses use different but, at that time, latest versions of the operational weather forecast system (OpenIFS) from ECMWF.
-They differ substantially in the assimilated observational ozone data.
-The MACC reanalysis assimilates only satellite derived tropospheric column ozone (TC), while CAMSRA also includes ozone profiles from satellite retrievals.
-Most importantly for our study, in situ observations from ozone surface station networks are only assimilated in the CAMS Regional Ensemble.
-The quality with respect to observations in northern Fennoscandia will be discussed in Section~\ref{subsubsec:clim_ozone}.
-Without too much foreshadowing, the results prevented us from fetching the whole CAMSRA data.
-
-\begin{table}[t]
-  \caption{Ozone related reanalysis products used in this study: \href{https://www.ecmwf.int/en/newsletter/158/meteorology/new-cams-global-reanalysis-atmospheric-composition}{Global reanalysis (ECMWF)}; \href{https://www.regional.atmosphere.copernicus.eu/}{ensemble mean of regional reanalysis for Europe (Copernicus)}.}
-  \label{tab:ozone_rea}
-\begin{tabular}{llcclcc}
-\tophline
-Name & Provider & \multicolumn{2}{c}{Resolution} & Time period & Meteorological forcing & \chem{O_3} assimilation\\
-&              & spatial & temporal & & \\
-\middlehline
-MACC & ECMWF & $0.75^\circ \times 0.75^\circ$ & 3-hourly & 2003 -- 2012 & OPS & satellite $^\triangledown$\\
-CAMSRA & ECMWF & $0.75^\circ \times 0.75^\circ$ & 3-hourly & 2003 -- 2012 $^\dagger$ & ERA5 / OPS $^\ddagger$ & satellite $^\blacktriangledown$\\
-CAMS Regional Ensemble & Copernicus & $0.1^\circ \times 0.1^\circ$ & 1-hourly & 2014 -- 2018 $^\dagger$ & OPS $^\star$ & in situ $^\vartriangle$\\
-\bottomhline
-\end{tabular}
-\belowtable{$^\dagger$ Subset of the data used in this study.
-The full data set extends into near real time; $^\ddagger$ ERA5 (2003--2016), OPS (later); $^\star$ EURAD uses WRF for downscaling of operational IFS; $^\triangledown$ MLS, OMI - tropospheric column; $^\blacktriangledown$ SCIAMACHY, MIPAS, MLS, OMI, GOME2, SBUV2 - tropospheric column + profile; $^\vartriangle$ METEO France NRT.} % Table Footnotes
-\end{table}
-
-\section{Statistical analysis of environmental key variables}
-\label{sec:stats}
-In this section, we will evaluate atmospheric conditions in the growing seasons of 2018/19 by statistical means.
-We will derive climatologies for both, temperature, precipitation, global irradiance, and ozone for Svanvik (Section~\ref{subsec:climatologies}).
-The ozone measurements will be put into a climatological as well as regional context.
-We will look at anomalies of the atmospheric key variable mentioned above (Section~\ref{subsec:anomalies}).
-
-\subsection{Derived climatologies}
-\label{subsec:climatologies}
-
-\subsubsection{Temperature, precipitation, and global irradiance}
-\label{subsubsec:clim_temp_prec}
-The location of Svanvik at $69.45\,\unit{^\circ N}$ suggests a subarctic climate.
-The derived climate diagram following Walter and Lieth shown in Fig.~\ref{fig:climatediagram} supports this very well.
-The diagram is based on climatological data from Svanvik (1992--2012) downloaded from LandbruksMeteorologiske Tjeneste.
-Monthly averaged temperatures (red line) are displayed with standard error of mean as error band.
-Averaged monthly accumulated precipitation (blue bars) is shown with standard deviation of mean as error bars.
-Months with average temperatures below freezing are denoted with a star.
-As can be seen, temperatures stay below freezing for 5 consecutive months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for K\"{o}ppens climate classification of a regular subarctic climate (Dfc).
-The highest monthly average temperature is $(13.1\pm 1.1)\,\unit{^\circ C}$ in July and the lowest $(-12.8\pm 2.0)\,\unit{^\circ C}$ in February.
-In the time period 1992--2012, the coldest measured temperature was $-45.2\,\unit{^\circ C}$ observed January 27, 1999, while the highest temperature on July 16 the same year was $29.4\,\unit{^\circ C}$.
-
-Winter and spring (November--April, except for March) are relatively dry ($\sum_m \mathrm{Precip} < 20\,\unit{mm}$).
-The driest month is January with $\sum_m \mathrm{Precip} = (16.7\pm 3.0)\,\unit{mm}$.
-The most precipitation occurs in the summer months, with a highest monthly averaged accumulated precipitation $\sum_m \mathrm{Precip} = (58.5\pm 9.2)\,\unit{mm}$ in August.
-The average annual accumulated precipitation given with standard error of mean is $(383\pm 86)\,\unit{mm}$.
-Precipitation shows a large variability throughout the years.
-
-A climatology derived for global irradiance $Q_0$ is shown in Appendix Fig.~\ref{fig:global_rad_clim}a) for both, climatological average of hourly maximum $\left<Q_0^\mathrm{max}\right>$, minimum $\left<Q_0^\mathrm{min}\right>$, and mean $\left<Q_0^\mathrm{mean}\right>$ irradiance observed at Svanvik.
-In agreement with the geometrically derived sunshine duration shown in Appendix Fig.~\ref{fig:sunlight_fennoscandia}, $\left<Q_0^\mathrm{max}\right>$ approaches zero in December/January commonly known as polar night and reaches its maximum in June/July commonly referred to as midnight sun conditions.
-The highest observed $\left<Q_0^\mathrm{max}\right>$ is about $800\,\unit{W\,m^{-2}}$, while on average $\left<Q_0^\mathrm{mean}\right>$ it is about $450\,\unit{W\,m^{-2}}$ (regularly reached in May--July).
-The maximum of the climatological minimum $\left<Q_0^\mathrm{min}\right>$ lies at about $200\,\unit{W\,m^{-2}}$.
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{climatediagram}
-  \caption{Climate diagram following Walter and Lieth.
-The diagram is based on climatological data for Svanvik/Pasvik (1992--2012) downloaded from LandbruksMeteorologiske Tjeneste.
-Monthly averaged temperatures (red line) are displayed with standard error of mean as error band.
-Averaged monthly accumulated precipitation (blue bars) is shown with standard deviation of mean as error bars.
-Months with average temperatures below freezing are denoted with a star.}
-  \label{fig:climatediagram}
-\end{figure}
-
-\subsubsection{Ozone}
-\label{subsubsec:clim_ozone}
-In the following, we will look at derived ozone climatologies from different perspectives.
-The multi-annual means of daily mean ozone, for simplicity here without standard deviations, at observation sites in northern Fennoscandia (refer to Table~\ref{tab:ebas_obs} for details) are shown in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs}.
-All climatologies peak in spring, with peak values of about $48\,\unit{ppb}$ in late April.
-The annual average ozone concentration \chem{\left<[O_3]\right>} declines throughout the growing season until it reaches its minimum in late summer/fall ($24-30\,\unit{ppb}$).
-The turning points lie approximately in June and November.
-The ozone climatology lags about 3 months behind the temperature climatology.
-At Svanvik, \chem{\left<[O_3]\right>} is $6.6\,\unit{ppb}$ lower than on sites at Jergul/Karasjok (NOR), Esrange (SWE), and Pallas (FIN).
-The large spread in case of Svanvik and Janiskoski can be explained due to the lower overall statistics.
-
-\begin{figure}[t]
   \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs}
-  \caption{Multi-annual mean of daily mean ozone at observation sites in Fennoscandia.
+  \caption{S1 Multi-annual mean of daily mean ozone at observation sites in Fennoscandia.
 The large spread in case of Svanvik and Janiskoski is due to the lower statistics.
@@
   \label{fig:ozone_climatology_fenoscandic_obs}
-\end{figure}
-
-The correlation of \chem{[O_3]} between Esrange, Pallas, and Jergul/Karasjok is fairly high ($r^2\approx 0.8$, Appendix Fig.~\ref{fig:density_distribution}a, b), hence we can put those data series together to derive a general ozone climatology for northern Fennoscandia.
-The correlation between Svanvik and Esrange and Pallas is rather low, 0.4 and 0.6, respectively.
-Thus, Svanvik cannot directly be represented by any measurements taken at either Esrange or Pallas.
-We shall come back to this in Section~\ref{subsec:ozone_reco}.
-
-In Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}, the 2 dimensional density distributions for the thus defined northern Fennoscandia (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}a) and Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) are shown.
-Darker colors indicate higher probability to observe these values.
-On top of the density distributions, $10\,\unit{days}$ averages of daily mean \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, daily maximum \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}}, and daily minimum \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, min}} are displayed together with the $1 \sigma$ uncertainties, respectively.
-In addition, \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} is also shown with standard error of mean (albeit small and hardly visible).
-Splines have been fitted through the data for easier inter-comparison with reanalysis data.
-
-From the northern Fennoscandic climatology, one can deduce several features.
-The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, min}} peaks about $20\,\unit{days}$ earlier than the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, and \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}}.
-With respect to the periodicity of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, it takes about $7.5\,\unit{months}$ from minimum concentration to maximum concentration, while from maximum to minimum it only takes about $4.5\,\unit{months}$.
-The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}} is at $52\,\unit{ppb}$ in late April and the minimum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, min}} occurs at $15\,\unit{ppb}$ in the beginning of September.
-The steep rise of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} in spring is in coincidence with increasing temperatures as well as increase in global irradiance.
-The decrease in \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} begins about one month ahead of the estimated onset of the greening season.
-Based on data from SeNorge.no and the agricultural rule of thumb that the daily average temperature has to be above $5\,\unit{^\circ C}$ for $5\,\unit{days}$ consecutively, we find $\left<G_\mathrm{begin}\right> = (150 \pm 14)\,\unit{days of the year}$ at Svanvik (Appendix Fig~\ref{fig:greening_season_change_Svanvik}), which translates into the end of May.
-This finding is in line with more extensive studies \citep[][e.g.,]{IJB:Karlsen2007,AFM:Linderholm}.
-
-The density distribution for Svanvik and the respective \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, and \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}} (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) display a similar pattern as the northern Fennoscandic climatology.
-The spring peak apparently occurs slightly earlier (in the beginning of April).
-As Svanvik is located at lower altitude than the other sites, the earlier ozone peak may point to an interplay with ozone uptake by vegetation.
-{\bf TODO: The data from SeNorge seems to point to a later onset of greening, than what I interfered from OpenIFS data.
-This means that the beginning of the growing season is not a good measure for the termination of the ozone spring peak.
-This does not exclude the early start of pine forest photosynthesis as a driver.
-Coniferous forest starts to photosynthesis earlier than what the growing season begin would mark.
-\chem{CO_2} uptake and phenology of coniferous trees \citep{TB:Kolari2007} starts as early as doy 100, acclimates to leaf temperature history - photosynthesis drops at around doy 200--250.} The average ozone concentration at Svanvik is lower than at the other sites, as already mentioned, which is in league with a decrease in tropospheric ozone with decreasing altitude.
-In July--September, ozone is occasionally almost completely depleted during night time.
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_spline}
-  \caption{Derivation of ozone climatologies.
-Density distributions are shown together with multi-annual mean of daily ozone concentrations \chem{[O_3]}.
-Splines have been fitted though daily mean and max \chem{[O_3]}.
-The climatology of daily mean ozone is shown together with $1\,\sigma$ standard deviation and standard error of the mean.
-Since \chem{[O_3]} are strongly correlated for sites at Jergul/Karasjok (NOR), Esrange (SWE), and Pallas (FIN) (see Fig.~\ref{fig:density_distribution}), data of these have been averaged together to derive a climatology for northern Fennoscandia.
-(a) Northern Fennoscandia; (b) Svanvik.}
-  \label{fig:ozone_climatology_fenoscandic_obs_spline}
-\end{figure}
-
-Finally, we shall look at the comparison of climatologies derived from observational ozone data and the various reanalysis products (Fig.~\ref{fig:ozone_climatologies}).
-As representative for the sites, the nearest neighboring grid point has been chosen.
-The ECMWF products do not reproduce observed surface ozone well.
-The MACC reanalysis (Fig.~\ref{fig:ozone_climatologies}a) is in general too low ($(9\pm 7)\,\unit{ppb}$) and displays an erroneous seasonal cycle.
-Ozone concentrations peak already in March followed by a second peak in July.
-Also CAMSRA matches the observed ozone climatology poorly (Fig.~\ref{fig:ozone_climatologies}b).
-Despite displaying only one ozone peak, it still does not reproduce the seasonality in northern Fennoscandia well.
-The CAMSRA spring peak appears to lag behind by $1\,\unit{month}$ and is $5\,\unit{ppb}$ lower, whereas the minimum occurs in January compared to August/September.
-The annual amplitude ($(26\pm 1)\,\unit{ppb}$) is larger than in the climatology derived from observations ($19\,\unit{ppb}$).
-Modeled \chem{[O_3]} at Svanvik are even slightly higher than at Pallas.
-In contrast, the ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in northern Fennoscandia quite well (Fig.~\ref{fig:ozone_climatologies}c), but is on average still slightly too low ($(2.8\pm 0.5)\,\unit{ppb}$).
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_climatologies_portrait}
-  \caption{Comparison of climatologies derived from observational data (EBAS, see Fig.~\ref{fig:ozone_climatology_fenoscandic_obs}) with various ozone reanalysis products: (a) ECMWF: MACC reanalysis; (b) ECMWF: CAMS reanalysis; (c) Copernicus: Regional model reanalysis ensemble mean.
-As representative for the sites, the nearest neighboring grid point has been chosen.
-The ECMWF products do not reproduce observed surface ozone well.
-The MACC reanalysis is in general too low and displays an erroneous seasonal cycle.
-The CAMS reanalysis matches better but does not reproduce the seasonality in northern Fennoscandia.
-The ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in northern Fennoscandia, though the modeled ozone concentrations are slightly too low.}
-  \label{fig:ozone_climatologies}
-\end{figure}
-
-There are many probable explanations for the discovered differences in the climatologies derived from the three reanalyses.
-The effective changes from the MACC reanalysis to CAMSRA have been reported and discussed by \citet{ACP:Inness2019} on a global level.
-Amongst others, an assimilation of ozone profiles from satellite retrieval and an upgraded ozone chemistry scheme have lead to the enhanced performance of latter.
-The higher spatial resolution of the regional air quality models used in the CAMS Regional Ensemble compared to the global reanalyses does capture more localized depletion and peak episodes in \chem{[O_3]} better and enhances the capability of the models with respect to daily and seasonal cycles in ozone.
-Latter is, of course, also affected by upgrades in the OpenIFS used as meteorological forcing.
-However, we can assume that the main driver for the strikingly different results are the assimilated observational ozone data.
-Satellite retrievals resolve \chem{[O_3]} at the surface rather poorly and hence do not constrain the global models well enough \citet{ACP:Andersson2017}.
-
-\subsection{2018/19 anomalies}
-\label{subsec:anomalies}
-Using these established climatologies, we look at the anomalies in 2018/19 and discuss the differences between the two years.
-
-\subsubsection{Temperature, precipitation, and global irradiance}
-\label{subsubsec:anomal_tpq}
-In Fig.~\ref{fig:plot_temperature_anomalies_svanvik}, temperature anomalies at Svanvik are shown as percentage of days warmer/colder than $\pm 1\,\sigma$ from the climatological mean for each month.
-Negative deviations from climatology are displayed as negative percentage.
-The full annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
-The summer of 2018 was significantly warmer than average.
-Especially in May and July more than $40\,\unit{\%}$ of days were significantly warmer than normal temperature.
-Temperatures continued to be significantly warmer throughout fall into early winter (August--November) with the exception of October.
-March 2018 was unusually cold.
-The summer of 2019 was fairly average, however significantly colder in July, while August was both significantly colder and warmer.
-Temperatures in April were significantly warmer than average.
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{temperature_signific_1}
-  \caption{Temperature anomalies at Svanvik.
-Percentage of days warmer/colder than $\pm 1\,\sigma$ from climatological mean displayed for each month.
-Negative deviations from climatology are shown as negative percentage.
-The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
-The summer of 2018 was significantly warmer than average, especially May and July.
-The summer of 2019 was fairly average but significantly warmer than average in April and colder in July/August.}
-  \label{fig:plot_temperature_anomalies_svanvik}
-\end{figure}
-
-In a similar way, the precipitation anomalies at Svanvik are displayed in Fig.~\ref{fig:plot_precipitation_anomalies_svanvik}.
-However, the percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ from the climatological mean is shown for each month, due to the larger variability in observed precipitation too few days showed a significance of $1 \sigma$ or higher.
-As above, negative deviations from the climatology are shown as negative percentage and annual positive/negative deviations are indicated in the respective corners.
-The hatched area between the dashed lines indicates the expected percentage of values falling above $\pm\frac{1}{2}\,\sigma$ if a normal distribution is assumed ($30.8\,\unit{\%}$).
-Unlike for the temperature anomalies, the picture is not as clear.
-While March 2018 was unusually wet, summer and fall (May/July, September/October) were significantly drier than average.
-2019 was rather average throughout the year, with drier than usual fall (September/October) and a wet December.
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{precipitation_signific_1}
-  \caption{Precipitation anomalies at Svanvik.
-Percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ from climatological mean displayed for each month.
-Negative deviations from climatology are shown as negative percentage.
-The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The hatched area between the dashed lines indicates the expected percentage of values falling above $\pm\frac{1}{2}\,\sigma$ if a normal distribution is assumed ($30.8\,\unit{\%}$).
-Summer was significantly dry only in 2018, while fall was significantly dry in both years.}
-  \label{fig:plot_precipitation_anomalies_svanvik}
-\end{figure}
-
-Similarly, the global irradiance anomalies at Svanvik are presented in Fig.~\ref{fig:global_rad_signific}.
-The percentage of days above/below $\pm 1\,\sigma$ from the climatological mean is displayed for each month.
-The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed.
-In summer 2018 (May/July), global irradiance was significantly higher than on average, while 2019 was generally an average year with a slightly darker spring (April/May).
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{global_rad_signific_1}
-  \caption{Global irradiance anomalies at Svanvik.
-Percentage of days above/below $\pm 1\,\sigma$ from climatological mean displayed for each month.
-Negative deviations from climatology are shown as negative percentage.
-The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
-In summer 2018 (May/July), global irradiance was significantly higher than on average, while 2019 had significantly less than normal irradiance in April/May and more in June.}
-  \label{fig:global_rad_signific}
-\end{figure}
-
-In summary, 2018 was significantly warmer, drier, and lighter than the climatological average, while 2019 was a rather average year.
-
-\subsubsection{Ozone}
-\label{subsubsec:anomal_ozone}
-As we discussed in the preceding Section~\ref{subsubsec:anomal_tpq}, meteorological conditions in 2018 were significantly different from the climatological average.
-How does this translate into observed ozone anomalies \chem{\Delta[O_3]}?
-
-If the data collected at Svanvik had been complete in 2018 and the climatology had extended into the resent past, we could have assessed the significance of ozone anomalies in the same way as for temperature, precipitation, and global irradiance.
-In Section~\ref{subsubsec:clim_ozone}, we described the construction of climatologies for northern Fennoscandia based on long term observations at Esrange, Jergul/Karasjok, and Pallas.
-For Svanvik this assessment is based on rather uncertain observations from 1986--1996.
-We also discussed the general patterns, similarities and differences.
-We shall now investigate \chem{\Delta[O_3]} in 2018/19 with respect to these climatologies in more detail.
-
-In the following, we will use the interchangeable terms anomalies and residuals to distinguish between a climatological (anomalies) and a statistical (residuals) point of view.
-All the above factors make it necessary to look not only at anomalies but statistically assess ozone concentration residuals at Svanvik as well as at Esrange and Pallas in comparison.
-Since the climatology of latter is more reliable, we will start with surface ozone concentration residuals \chem{\Delta[O_3]} at Esrange and Pallas in 2018.
-As data for 2019 has not yet been available from EBAS (last access, February 2020), we refrain from looking into these.
-
-In Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}, the probability density functions of daily mean ozone concentration residuals for Esrange and Pallas are shown.
-Both stations display a clear deviation from the climatological mean, however, not significant.
-On average, \chem{[O_3]} had been elevated by $(1.9\pm 5.7)\,\unit{ppb}$ and $(1.9\pm 5.4)\,\unit{ppb}$, respectively.
-Both distributions are slightly skewed towards positive deviations with $\frac{1}{4}$ of days (3.
-quantile, q3) falling above $5.8\,\unit{ppb}$ and $5.6\,\unit{ppb}$, respectively.
-
-Appendix Figure~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik} displays the probability density functions of daily mean ozone concentration residuals for Svanvik based on data taken in 2018/19 with respect to the climatology derived for northern Fennoscandia (denoted NF in Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}a) and the climatology derived for Svanvik (Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b).
-As previously mentioned, \chem{\left<[O_3]\right>} at Svanvik are $(6.6\pm 7.1)\,\unit{ppb}$ lower than the northern Fennoscandic climatology.
-For 2018 and 2019, we find $\Delta\chem{\left<[O_3]\right>} = (-3.8\pm 5.6)\,\unit{ppb}$ and $\Delta\chem{\left<[O_3]\right>} = (-5.2\pm 5.0)\,\unit{ppb}$, respectively.
-While the probability density of the residuals between the two climatologies is skewed towards negative values, the skew of 2018/19 is positive.
-With respect to the climatology for Svanvik (Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b), the probability density functions average for 2018/19 lies at $\Delta\chem{\left<[O_3]\right>} = (2.6\pm 5.8)\,\unit{ppb}$ and $\Delta\chem{\left<[O_3]\right>} = (1.2\pm 5.0)\,\unit{ppb}$, respectively.
-Neither of the \chem{\Delta\left<[O_3]\right>} is significant by statistical means, but daily average ozone concentrations are slightly elevated in 2018 compared to 2019.
-
-The \chem{\left<\Delta\left<[O_3]\right>\right>} for Svanvik in 2018 is slightly larger than for Esrange and Pallas.
-This could mean either that 2018 had been more exceptional at Svanvik or the derived climatology may need a correction for increased tropospheric background ozone concentration.
-Based on the meteorological conditions, we assume 2019 as average year which leads to a constant bias of about $1.2\,\unit{ppb}$.
-Daily \chem{\Delta_{present-past}\left<[O_3]\right>} from Esrange and Pallas, wherein present is defined as the average of 2010--2018 and past as the respective beginning of data taking until 1997, do not reveal a clear trend: We find for Esrange $(1.5\pm 3.1)\,\unit{ppb}$ and for Pallas $(-2.5\pm 3.7)\,\unit{ppb}$ \citep{NILU2003, BER:Ruoho-Airola2015}.
-Former is compatible with our assumed bias for the climatology derived for Svanvik.
-
-To further investigate the ozone residuals, we conduct a Student's~t-test ($t = \frac{\chem{\Delta[O_3]}}{\sigma}$), assuming the same sample uncertainty $\sigma$ in both, climatology and the year in question.
-Results for Esrange and Pallas are shown in Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test}.
-This t-test clearly shows, that at both sites the residuals are significant on the $1\sigma$ level for at least $50\,\unit{\%}$.
-In Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}, the test statistics for Svanvik are displayed.
-From Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}a), we can deduce, that the residuals between the Svanvik climatology and the northern Fennoscandic climatology are significant on the $2\sigma$ level for at least half of the data, while the residuals are less certain in 2018 and 2019.
-In Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}b) the same is shown for the years 2018/19 with respect to the climatology from Svanvik.
-We find that less than $50\,\unit{\%}$ of ozone concentration residuals are significant on the $1\sigma$ level.
-
-In accordance to the above findings, we adjust the derived ozone climatology for Svanvik by adding a constant bias correction factor $\Delta\chem{[O_3]_\mathrm{bias-corr}} = 1.2\,\unit{ppb}$ and estimate the ozone concentration anomalies for 2018/19.
-In Fig.~\ref{fig:ozone_signific}, the resulting percentage of days above/below $\pm 1\,\sigma$ from the bias-corrected climatological mean is displayed for each month.
-Negative deviations from climatology are shown as negative percentage.
-Note that missing bars for certain months are both due to data availability (see Section~\ref{subsubsec:atmo_svanvik} for details about the time range of data taking at Svanhovd) and anomalies not meeting the criteria of $\pm 1\sigma$ deviation from climatological mean.
-In April/May 2018, ozone was significantly enhanced, while 2019 was generally an average year.
-In both years, ozone concentrations were significantly below average in November.
-July 2018 has not been corrected for the missing data due to data acquisition problems.
-The corrected value is, however, indicated by the star (refer to Section~\ref{subsec:ozone_reco} for details of the used method for gap filling).
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_signific_1}
-  \caption{Ozone concentration anomalies at Svanvik.
-Ozone climatology has been corrected for an constant negative bias of $-1.2\,\unit{ppb}$.
-Percentage of days above/below $\pm 1\,\sigma$ from climatological mean displayed for each month.
-Negative deviations from climatology are shown as negative percentage.
-The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
-July 2018 has not been corrected for the missing data due to data acquisition problems.
-The corrected value is indicated by the star (refer to Section~\ref{subsec:ozone_reco} for details).
-In April/May 2018, ozone was significantly higher than on average, while 2019 was generally an average year.
-In both years ozone concentrations were significantly below average in November.}
-  \label{fig:ozone_signific}
-\end{figure}
-
-To conclude, ozone concentrations in 2018 have been elevated compared to the respective climatologies, however, on average not significantly by statistical means.
-At Esrange and Pallas at least $25\,\unit{\%}$ (3.~quantile) of anomalies are significant on a $3\sigma$ level or more, while at Svanvik the 3.~quantile lies at about $2\sigma$.
-This means that in 2018 the stations at Esrange and Pallas saw more days with elevated ozone than was monitored at Svanvik.
-
-\section{DO3SE modeling}
-\label{sec:do3se}
-To investigate which of the environmental factors lead to the observed ozone damage on clover and tobacco in 2018 in contrast to 2019, when no such damage was found on the clovers but only on the sensitive tobacco cultivar, we model the ozone uptake using the $\mathrm{DO_3SE}$ model and give an estimate on biomass reduction due to ozone uptake.
-
-\subsection{Model description}
-\label{subsec:model_describtion}
-
-The opening of stomata is governed by different species-specific and environmental factors and can be described by a multiplicative model \citep{PTRS:Jarvis1976,ICP:MappingManual2017}:
-\begin{equation}
-  g^k_\text{sto} = g^k_\text{max} \cdot f^k_\text{phen} \cdot f^k_\text{light} \cdot \max{\left\{f^k_\text{min}, f^k_\mathrm{T} \cdot f^k_\mathrm{VPD} \cdot f^k_\text{SW}\right\}}.
-  \label{eq:stomatal}
-\end{equation}
-%
-Where $g_\mathrm{max}$ is the specific-specific maximum stomatal conductance value which is then modified by seasonal and environmental factors that vary within a range $0-1$.
-They are empirically determined and account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_\mathrm{T}$), water vapor pressure deficit ($f_\mathrm{VPD}$), and soil water content ($f_\text{SW}$).
-All factors differ with species or plant functional type (PFT) denoted with $k$.
-
-The temperature adjustment function is defined as
-%
-\begin{equation}
-  f_\mathrm{T} = \frac{T_\text{air}-T_\text{min}}{T_\text{opt}-T_\text{min}} \cdot \left(\frac{T_{\text{max}}-T_\text{air}}{T_{\text{max}}-T_\text{opt}}\right)^\beta,
-  \label{eq:f_temp}
-\end{equation}
-with $\beta = \frac{T_\text{max}-T_\text{opt}}{T_\text{opt}-T_\text{min}}$ a mixed reciprocal polynomial function of temperature.
-The shape parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various species or PFT and $T_\mathrm{air}$ is the surface air temperature.
-All temperatures are defined in units of \unit{^\circ C}.
-
-The water vapor pressure deficit function is
-%
-\begin{equation}
-  f_\mathrm{VPD} = f_\text{min}+(1-f_\text{min}) \cdot \frac{D_\text{min} - \text{VPD}}{D_\text{min}-D_\text{max}}
-  \label{eq:f_vpd}
-\end{equation}
-%
-where VPD is the leaf to air vapor pressure deficit in \unit{kPa} with $f_\text{min}$, $D_\text{min}$, $D_\text{max}$ describing the relative stomatal conductance to changes in vapor pressure deficit.
-
-The wavelength band $400-700\,\unit{nm}$ plant chlorophyll responds to is called photosynthetic active radiation (PAR).
-Its integral is the photosynthetic photon flux density (PPFD).
-The relationship between relative $g_\mathrm{sto}$ and PPFD is given by
-%
-\begin{equation}
-  f_\text{light} = 1-\text{exp}(-\alpha_\text{light}\cdot \text{PPFD})
-  \label{eq:flight}
-\end{equation}
-%
-where $\alpha_\mathrm{light}$ is a slope parameter describing how fast stomata open or close at low light intensities.
-
-The $\mathrm{DO_3SE}$ model as described in \citet{ACP:Bueker2012} is used to simulate SWP across a species or PFT specific root depth according to the Penman--Monteith energy balance method that drives water cycling through the soil--plant--atmosphere system.
-\citet{ACP:Bueker2012} discuss several formulations available to parameterize the reduction in stomata conductance due to water content in the soil.
-Here we use the relationship between reative $g_\mathrm{sto}$ and soil water potential (SWP) given by:
-%
-\begin{equation}
-  f_\mathrm{SW} = \text{min}\left\{1, \text{max}\left\{f_\mathrm{min}, \frac{(1 - f_\mathrm{min})\cdot(\mathrm{SWP_{min}} - \mathrm{SWP})}{\mathrm{SWP_{min}} - \mathrm{SWP_{max}}} + f_\mathrm{min} \right\} \right\}
-    \label{eq:fsw}
-\end{equation}
-%
-where SWP is the soil water potential across the root zone and $\mathrm{SWP_{min}}$ and $\mathrm{SWP_{max}}$ are the paramters describing the $f_\mathrm{SW}$ relationship.
-
-To compute $\mathrm{POD_y}$, we estimate the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf.
-The terms $r_c$ (leaf surface resistance) and $r_b$ (quasi-laminar resistance) allow for the deposition of \chem{O_3} to the leaf and the fraction that is taken up via the stomata.
-%
-\begin{equation}
-  \Phi_\mathrm{sto} = \chem{[O_3]} \cdot \frac{u(z_1) \cdot g_\mathrm{sto} \cdot r_c}{r_b + r_c}.
-  \label{eq:flux_stomata}
-\end{equation}
-%
-The quasi-laminar boundary layer resistance is calculated by
-%
-\begin{equation}
-  r_b = 1.3 \cdot 150 \cdot \sqrt{\frac{L}{u(z_1)}}.
-  \label{eq:quasi_laminar_resist}
-\end{equation}
-%
-Where $L$ is the cross-wind leaf dimension, $u(z_1)$ the wind speed at height $z_1$, and the factor $1.3$ accounts for the diffusivities between heat and \chem{O_3}.
-
-\subsection{Input data and gap filling methodology}
-\label{subsec:ozone_reco}
-The $\mathrm{DO_3SE}$ model requires hourly, continuous meteorological observations.
-In addition to variables covered in Section~\ref{sec:data}, $2\,\unit{m}$ wind $u_\text{2m}$ and vapor pressure deficit VPD are needed.
-VPD has been calculated from observed $T_\text{2m}$ and relative humidity
-\begin{equation}
-  \text{VPD} = P_s(T) \cdot \left(1-\frac{\text{relHum}}{100}\right),
-\end{equation}
-with saturation vapor pressure $P_s(T)$ in \unit{hPa} derived from Arden Buck equation~\citep{JAP:Buck1981, Buck2012}.
-
-The following gap filling methodology was devised for the meteorological input data:
-\begin{itemize}
-\item Single hours of missing data were filled by taking the average of the hourly values coming before, and after, the missing value.
-\item Several consecutive hours of missing data (23 or less) were filled by taking the average of the corresponding hour the day before, and the day after; and repeating this for each missing hour of data.
-If data was unavailable from that hour of the previous day, then only the value from the day after was used and vice versa.
-Interpolated values were not used in calculating averages.
-\item Data gaps longer than $24\,\unit{h}$ were filled using weekly diurnal averages.
-i.e.
-an average was calculated using the corresponding hour throughout the week before and after.
-Interpolated values were not used in calculating averages.
-\end{itemize}
-
-Because the available reanalysis products for surface ozone have their limitations, we deem it necessary to reconstruct the missing data in 2018 for the purpose of $\mathrm{DO_3SE}$ modeling by applying a Reynolds's decomposition of the form:
-\begin{equation}
-  \chem{[O_3]} = \chem{\left<[O_3]\right>} + \chem{\Delta[O_3]},
-  \label{eq:reynolds_decomp}
-\end{equation}
-with climatology \chem{\left<[O_3]\right>} and anomalies \chem{\Delta[O_3]}.
-The exact procedure will be described in the following.
-We derive hourly climatologies for individual stations and combine Esrange/Pallas to derive a climatology for northern Fennoscandia for July (Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}a).
-We compute temporal correlations between Svanvik and the other stations in northern Fennoscandia to estimate a time lag (Appendix Fig.~\ref{fig:time_lag_correlation}.
-We find $3\,\unit{h}$ for Esrange and Pallas with respect to Svanvik.
-Due to the high correlation with Svanvik (Appendix~Fig.~\ref{fig:density_distribution}), we decide on Pallas as reference station for the reconstruction.
-We correct the Fennoscandic climatology for the found time lag \chem{\left<[O_3]\right>_{hourly,\,t-3}} by shifting the respective time series back in time by $3\,\unit{h}$ and use it to derive a projection for Svanvik:
-\begin{equation}
-  X = \frac{\chem{\left<[O_3]\right>_{hourly}^{Svanvik}}}{\chem{\left<[O_3]\right>_{hourly,\,t-3}}}.
-  \label{eq:mapping}
-\end{equation}
-As an intermediate step, we apply Eq.~(\ref{eq:reynolds_decomp}) to derive hourly anomalies for each station $i$
-\begin{equation}
-  \chem{\Delta[O_3]_{hourly}}^i = \chem{[O_3]_{hourly}}^i-\chem{\left<[O_3]\right>_{hourly}},
-\end{equation}
-with $i \in \{\text{Esrange, Pallas, Svanvik\}}$.
-In Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}b), these hourly ozone concentration anomalies are shown.
-We also correct the ozone anomalies for Pallas for the time lag \chem{\Delta[O_3]_{hourly,\,t-3}^{Pallas}} and use the mapping function in Eq.~(\ref{eq:mapping}) to derive reconstructed anomalies for the missing values at Svanvik:
-\begin{equation}
-    \chem{\Delta [O_3]_{hourly}^{Svanvik,\,reco}} = \chem{\Delta[O_3]_{hourly,\,t-3}^{Pallas}} \cdot X
-\end{equation}
-Finally, we add the bias corrected Svanvik ozone climatology and derive a reconstructed ozone concentration time series \chem{[O_3]_{hourly}^{Svanvik,\,reco}}.
-In Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}c), the result is shown together with the observed data and the Copernicus Regional Model Reanalysis at the nearest grid point for comparison.
-
-\subsection{Model parameters}
-We roughly estimated land use types for an approximately $1\times 1\,\unit{km}$ patch centered at the NIBIO Environment Centre Svanhovd from Fig.~\ref{fig:svanhovd_research_station}: $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban.
-
-Default PFTs are derived from \citet{ICP:MappingManual2017} and will be referred to as \emph{mapping manual} abbrieved MM.
-A comprehensive list of these model parameters can be found in Supplement~Table~S1.
-Intitial test simulations with the mapping manual derived PFTs showed unrealistically low $g_\mathrm{sto}$ in 2019 (Supplememt~Figs.SX--SX+2) particularly for grassland.
-Only during an extended periode in late July did we find substential stomatal conductance.
-In ecological terms, this would mean that there was almost no growth of grass in the summer of 2019 -- a prediction which is easily falsified by reality.
-We identified $f_\mathrm{T}$ as abnormally low, being the limiting factor of stomatal conductance.
-We propose bespoke parameterizations for the main vegetation types Norway spruce, birch, and perennial grassland better adapted to the climatic conditions in the target region.
-Note that these parameterizations represent hypothetical PFTs or species that have to be varified by experiements.
-
-We hypothize that perannial vegetation will likely prioritize their growth or fast adolesence throughout the short, sub-boreal growing season and hence their carbon sequestration.
-Since the $\mathrm{DO_3SE}$ in its current state does not simulate $A_\mathrm{n}$, we assume for out purpose a first order proptionality between carbon uptake (photosynthesis) and $g_\mathrm{sto}$ and tune the species temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) seperately for higher stomatal conductance.
-For this purpose, we graphically intersected propability density functions (PDFs) of climatological data for temperature and global irradiace with the respective response function (Fig.~\ref{fig:f_temp_grassland} for grassland; for Norway spruce and birch see Appendix Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}).
-Formally, a parameterization is considered more optimial the larger the intersection area between the response function and PDF is.
-
-For the climatization of the temperature response functions, we assume two different types: \emph{boreal} and \emph{cold}.
-In either cases, we further assume that the hypothetical species are more tolerant to cold conditions than the mapping manual default suggests and shift $f_\mathrm{T}$ towards colder temperatures.
-We construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to the mapping manual default.
-In the same way, \emph{boreal} refers to a hypothetical species which is very tolerant to cold but is sensitive to high temperatures, and most efficient at cool temperatures (close to the climatological mean temperature of 1992--2000).
-All three types for grassland are depicted together with the temperature PDF in Fig.~\ref{fig:f_temp_grassland}a).
-
-For Norway spruce, our hypothetical PFTs $T_\mathrm{opt}$ are motivated by a longer timeseries of both observed \chem{CO_2} uptake and temperature at two observation sites in Southern (\emph{cold}) and Northern (\emph{boreal}) Finland presented in \citet{TB:Kolari2007}.
-We assume a similar adaption in birch.
-The parameters are comprehensibly listed in Table~\ref{tab:sensitivity_tests_temp}.
-
-For the the senisitivity regarding light, we assume that the plants may react either faster or slower to a change at lower light intensities.
-We analytically derive the inverse function $f_{\mathrm{light},k}^{-1}$ for each PFT or species $k$
-\begin{equation}
-  \gamma_k(f_\mathrm{light}) := f_{\mathrm{light}, k}^{-1} = -\frac{\ln(1-f_\mathrm{light})}{\alpha_k}.
-  \label{eq:inverse_function}
-\end{equation}
-We calculate it for the mapping manual $\alpha_\mathrm{MM}$ at its halfway value $0.5$ 
-\begin{equation}
-  \gamma(0.5) = -\frac{\ln(0.5)}{\alpha_\mathrm{MM}},
-  \label{eq:inverse_function_halfway}
-\end{equation}
-and define a variation $\gamma' = \gamma \cdot \epsilon$ with $\epsilon \in \{0.8, 1.2\}$.
-We resolve Eq.~(\ref{eq:inverse_function}) for $\alpha$ and derive $\alpha'(\gamma')$.
-The resulting functions are shown in Fig.~\ref{fig:f_temp_grassland}b).
-All derived parameters are given in Table~\ref{tab:sensitivity_tests_light}.
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{javis_funcs_temp_hist_grassland}\\
-  \includegraphics[width=8.3cm]{javis_funcs_rad_hist_grassland}
-\caption{Construction of bespoke response functions for grassland.
-(a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown toghether with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability densitiy function - PDF), respectively.
-Original mapping manual parameterization is shown in comparison as solid line.
-Note that $Q_0$ has been truncated to $0.006$.
-PPFD0.8 and PPFD1.2 refer to $\alpha$ values increasing/decreasing PPFD at $f_\mathrm{light}=0.5$ by $\pm 20\,\%$, respectively.}
-\label{fig:f_temp_grassland}
-\end{figure}
-
-\begin{table}[t]
-  \caption{Bespoke temperature prameterizations. MM refers to mapping manual \citep{GCB:Mills2010,ICP:MappingManual2017}.}
-  \label{tab:sensitivity_tests_temp}
-  \begin{tabular}{llrrr}
-    \tophline
-    Species & type & $T_\mathrm{min}$ & $T_\mathrm{opt}$ & $T_\mathrm{max}$ \\
-    \middlehline
-    \multirow{3}{*}{Spruce} & MM & 0 & 20 & \multirow{3}{*}{100}\\
-    & Boreal & 0 & 10 & \\
-    & Cold & 0 & 15 & \\
-    \middlehline
-    \multirow{3}{*}{Birch} & MM & 5 & 20 & \multirow{3}{*}{100}\\
-    & Boreal & 0 & 10\\
-    & Cold & 5 & 15\\
-    \middlehline
-    \multirow{3}{*}{Grassland} & MM & 10 & 24 & 36\\
-    & Boreal & 0 & 10 & 24 \\
-    & Cold & 5 & 15 & 36\\
-    \bottomhline
-    \end{tabular}
-\end{table}
-
-\begin{table}[t]
-  \caption{Bespoke light prameterizations. MM refers to mapping manual \citep{GCB:Mills2010,ICP:MappingManual2017}.}
-  \label{tab:sensitivity_tests_light}
-  \begin{tabular}{llr}
-    \tophline
-    Species & type & $\alpha$\\
-    \middlehline
-    \multirow{3}{*}{Spruce} & MM & 0.006 \\
-    & PPFD0.8 & 0.008  \\
-    & PPFD1.2 & 0.005  \\
-    \middlehline
-    \multirow{3}{*}{Birch} & MM & 0.004 \\
-    & PPFD0.8 & 0.005\\
-    & PPFD1.2 & 0.004\\
-    \middlehline
-    \multirow{3}{*}{Grassland} & MM & 0.011\\
-    & PPFD0.8 & 0.014 \\
-    & PPFD1.2 & 0.009 \\
-    \bottomhline
-    \end{tabular}
-\end{table}
-
-{\bf TODO: Write about boreal and cold parameter sets and their justification.
-Write about suggested method for optimizing PFTs in a simple but physiological way based on optimizing maximum climatological mean $g_\mathrm{sto}$ and morning responds.
-If plants have a short growing season up north optimizing \chem{CO_2} uptake around noon (highest light intensity) and speed up opening of stomata at lower intensities (morning), additional contraint lower climatological variance?}
-An idea to justify our choice based on the fullset of climatological data for Svanvik is shown in Fig.~\ref{fig:javis_func_opt_mean}.
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{javis_func_opt_mean.pdf}
-  \caption{Proposed metric to test bespoke response functions.}
-  \label{fig:javis_func_opt_mean}
-\end{figure}
-
-Start ($A_\text{start}$) and end ($A_\text{end}$) of growing season for each PFT are given in units of day of year (doy) and estimated as follows.
-Evergreen needle leaf trees are photosynthetic active already at low temperatures \citep{TB:Kolari2007}.
-This leads necessarily to an increase in observable net photosynthesis.
-Hence, we use a MODIS (Aqua/Terra) retrieved net photosynthesis product for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine onset and end of growing season.
-Net photosynthesis data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}).
-By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{doy}$ for 2018 and day 106 in 2019.
-$A_\text{end}$ amounts to day 261 and 274, respectively, and shall be used for all vegetation alike.
-The resulting growing season for Norway spruce in 2019 was one month longer then in 2018.
-For estimating $A_\text{start}$ of birch tree, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Fig.~\ref{fig:greening_season_change_Svanvik}).
-We find 129 and 130, respectively.
-From the latitude model typically used for localizing $\mathrm{DO_3SE}$, we get $A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$ indicating an about $2\,\unit{months}$ longer growing season.
-For perennial grassland, we assume a latency period of about $1\,\unit{month}$ after snow melt.
-Data for first snow-free day is taken from the weather station at Øvre Neiden in Sør-Varanger.
-In both years, the first snow-free day incidents with the estimated date of bud burst.
-As leafs of birches in northern Fennoscandia \emph{Betula pubescens} are smaller than southern specimen, we have measured their dimension for the purpose of this study and find $(3.0\pm 0.5)\,\unit{cm}$.
-\citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area.
-
-As pointed out by \citet{ACP:Bueker2012}, a very important factor in deriving a proper $\mathrm{POD_y}$ is the soil texture or how much water the soil can hold.
-As Svanhovd is located in the bed of the Pasvik river, we assume a sandy loam texture at $60\,\unit{cm}$ depth.\\
-
-\subsection{POD and ozone impact}
-
-We have modeled $\mathrm{POD_y}$ for three natural/semi-natural vegetation PFTs, birch, Norway spruce, and perennial grassland.
-The results are shown comprehensively in Fig.~\ref{fig:pody} as a function of doy for March--October.
-A flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) has been chosen.
-\chem{[O_3]} are plotted on the same axis and scales as $G_\text{sto}^\text{leaf}$ but in units of $\unit{ppb}$.
-
-As expected, $\mathrm{POD_y}$ is higher in 2018 than in 2019 for all PFTs.
-In the end of the 2018 season, Birch took up $13.1\,\unit{mmol\,m^{-2}\,PLA}$, while perennial grassland had an uptake of $4.5\,\unit{mmol\,m^{-2}\,PLA}$.
-The largest interannual variation is also found in Birch ($2.1\,\unit{mmol\,m^{-2}\,PLA}$) and perennial grassland ($2.6\,\unit{mmol\,m^{-2}\,PLA}$).
-Whereas, there is only a small difference for Norway spruce ($0.3\,\unit{mmol\,m^{-2}\,PLA}$) between the two years.
-
-%Except for perennial grassland, model integrations with climatological forcings show higher $\mathrm{POD_y}$ than 2018.
-This can be explained by means of non linearity in modeled stomatal conductance (Eq.~(\ref{eq:stomatal})) in Eq.~(\ref{eq:pod}).
-The climatology in $\mathrm{POD_y}$ therefore represents rather an ideal year for growth then a true climatology.
-To establish a true climatology, it would be necessary to run the $\mathrm{DO_3SE}$ model for 10--20 years with respective observed phenology as well as meteorological and ozone forcing.
-
-%Based on this interpretation of climatology in $\mathrm{POD_y}$, we can learn more about potential ozone damage.
-The time derivative $\frac{\mathrm{\Delta POD_y}}{\Delta t}$ displays very different features for each PFT and meteorological forcing.
-Although Norway spruce shows a low interannual variation, $\frac{\mathrm{\Delta POD_y}}{\Delta t}$ (Fig.~\ref{fig:pody_rel}) indicates an extended period of higher than average ozone uptake in spring 2018 (steep slope), while ozone uptake in June/July was anomalously low in 2019 (shallow slope) but high in fall.
-This is most prominently visible in Fig.~\ref{fig:pody}h) for perennial grassland, where consecutive periods of low $G_\text{sto}^\text{leaf}$ manifest in three plateaus in $\mathrm{POD_y}$ throughout the season (1: $184-198\,\unit{doy}$, 2: $209-228\,\unit{doy}$, 3: $235-240\,\unit{doy}$).
-An episode with anomalous high ozone in mid June ($194\,\unit{doy}$) can be identified but does not amount to unusually high ozone uptake on an annual basis.
-
-The estimated biomass reduction in \unit{\%} due to \chem{O_3} following \citet{ICP:MappingManual2017} is shown in Table~\ref{tab:biomass_reduction}.
-{\bf TODO: Lisa?
-Add paragraph about biomass reduction to introduction?}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{DO3SE_results_pody_gsto_o3}
-  \caption{$\mathrm{POD_y}$ modeling results from $\mathrm{DO_3SE}$ model for three different PFTs over day of year (doy), March--October.
-A flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) has been chosen.
-\chem{[O_3]} are plotted on the same axis and scales as $G_\text{sto}^\text{leaf}$ but in units of $\unit{ppb}$.
-(a, b) Birch; (c, d) Norway spruce; (e, f) perennial grassland.
-From left to right: 2018, 2019.
-{\bf TODO: substitute with figure for boreal PFTs?}}
-  \label{fig:pody}
-\end{figure*}
-
-\begin{table}[t]
-  \caption{Estimated total biomass reduction in \unit{\%} due to \chem{O_3} following \citet{ICP:MappingManual2017}.}
-  \label{tab:biomass_reduction}
-\begin{tabular}{lccrr}
-\tophline
-Year & \multicolumn{4}{c}{PFT}\\
-& Birch & Norway spruce & \multicolumn{2}{c}{Perennial grassland}\\
-\middlehline
-2018 & 11.99 & 2.34 & 8.06 & 10.51$^*$\\
-2019 & 10.07 & 2.26 & 6.43 & 7.90$^*$\\
-clim. & 13.81 & 2.58 & 6.67 & 8.29$^*$\\
-\bottomhline
-\end{tabular}
-\belowtable{$^*$ above ground biomass} % Table Footnotes
-\end{table}
-
-%\begin{figure}[t]
-%  \includegraphics[width=8.3cm]{DO3SE_results_rel}
-%  \caption{Daily averaged $\mathrm{\Delta POD_y}/\Delta t$ from $\mathrm{DO_3SE}$ model for three different PFTs over day of year (doy), March--October.
-Error bars/-bands indicate $1 \sigma$ standard deviation.
-(a) Birch; (b) Norway spruce; (c) perennial grassland.}
-%  \label{fig:pody_rel}
-%\end{figure}
-
-\conclusions[Discussion and conclusions]
-\label{sec:conc} 
-\begin{itemize}
-  \item What did we learn from the 2018/2019 seasons in terms of ozone damage in the far north?
-  \item What are the main uncertainties in the tools we've used?
-  \item What implications could this have for future scenarios in which 2018 conditions may occur more often?
-\end{itemize}
-PODY is no good measure without a damage function with higher temporal resolution than annual.
-This has been already pointed out by \citet{AE:Musselman2006}!
-Higher than normal spring uptake (weather conditions) does not amount to critical uptake on an annual basis.
-Weather and climatic conditioning of plants is more important for PODy than episodes of higher than usual ozone concentrations.
-Therefor high $\Delta \mathrm{POD_y} / \Delta t$ is not accounted for properly.
-It is important to know the plant response functions to temperature, light, humidity/precipitation as best as possible if one wants to assess risk of damage.
-
-Regarding the shift (earlier) and prolongation (later) of the growing season in the far north no general assessment can be made.
-Data show that in both 2018 and 2019 ozone uptake in early season was anomalous high in all studied PFTs.
-The extended growing season in 2019 might have countered higher episodic ozone uptake in 2018.
-In both years, low mid-season uptake due to weather conditions washed out any effect of these.
-
-{\bf TODO: might be interesting to look at Mills et al 2003 and the different spatial extents of risk estimated when using AOT40 vs PODy.
-That could be useful to explain why we focus on PODy (rather then AOT40).- Lisa Come back to this later on in the discussions(?) - it has not been breached at Svanvik in northern Fennoscandia.}
-
-Bigger picture:
-\begin{itemize}
-\item Ozone damage assessment using mapping manual "compromise" parameterization.
-\item Land-atmosphere modelling: Problems in productivity of species especially in the arctic regions due to PFTs which are not suited for the climate.
-Automating optimization of climate adaption of general PFTs.
-\end{itemize}
-
-%% The following commands are for the statements about the availability of data sets and/or software code corresponding to the manuscript.
-%% It is strongly recommended to make use of these sections in case data sets and/or software code have been part of your research the article is based on.
-
-%\codeavailability{TEXT} %% use this section when having only software code available
-
-\dataavailability{TEXT} %% use this section when having only data sets available
-
-%\codedataavailability{TEXT} %% use this section when having data sets and software code available
-
-\sampleavailability{TEXT} %% use this section when having geoscientific samples available
-
-%\videosupplement{TEXT} %% use this section when having video supplements available
-
-\appendix
-\section{Derived climatologies}    %% Appendix A
-
-\subsection{Temperature, precipitation, and global irradiance}     %% Appendix A1, A2, etc.
-
-\appendixfigures  %% needs to be added in front of appendix figures
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{global_rad_clim}
-  \caption{Observed global irradiance at Svanvik.
-The cyan lines indicate the $1\,\sigma$ level deducted from the climatology.
-(a) Climatology (hourly maxima, mean, and minima); deviation from climatology (b) 2018; (c) 2019.}
-  \label{fig:global_rad_clim}
 \end{figure*}
@@
   \label{fig:greening_season_change_Svanvik}
-\end{figure*}
-\clearpage
-
-\subsection{Ozone}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{density_distribution}
-  \caption{Probability densities and correlation coefficient for \chem{[O_3]} between different sites in northern Fennoscandia.
-Jergul/Karasjok is well correlated with Esrange, and Pallas.
-Svanvik displays highest correlation with Pallas.
-(a) Jergul/Karasjok--Esrange; (b) Jergul/Karasjok--Pallas; (c) Svanvik--Esrange; (d) Svanvik--Pallas.}
-  \label{fig:density_distribution}
 \end{figure*}
@@
 \begin{figure*}[t]
-  \includegraphics[width=12cm]{ozone_reconstruction_2018_07}
-  \caption{Reconstruction of missing \chem{[O_3]} observation data at Svanvik in July 2018 based on derived hourly climatologies of northern Fennoscandia, Svanvik, and data Pallas.
-(a) Time series; (b) anomalies; (c) reconstruction.}
-  \label{fig:ozone_reconstruction_2018_07}
-\end{figure*}
-
-\begin{figure*}[t]
   \includegraphics[width=12cm]{ozone_observation_timelag_svanvik}
@@
 
-% Shall go to supplement!
-%\begin{figure*}[t]
-%  \includegraphics[width=12cm]{ozone_fennoscandic_obs_aot40}
-%  \caption{Integrated AOT40 for various sites in northern Fennoscandia over the course of 33 years.
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{ozone_fennoscandic_obs_aot40}
+  \caption{Integrated AOT40 for various sites in northern Fennoscandia over the course of 33 years.
 For simplicity, we integrated hourly ozone between $1\,\unit{am}-11\,\unit{pm}$ in any case, although a night time light intensity $> 50\,\unit{W\,m^{-2}}$ is only measured during midnight sun conditions.
@@
 (a) May 1 -- July 31 (b) April 1 -- September 30.}
-%  \label{fig:fennoscandic_aot40}
-%\end{figure*}
+  \label{fig:fennoscandic_aot40}
+\end{figure*}
 
 \begin{figure*}[t]
-  \includegraphics[width=12cm]{modis_Psn}
-  \caption{Estimated begin (BGS) and end (EGS) of growing season for coniferous trees from MODIS Aqua/Terra net photosynthesis (PSN) product.
-A $1\times 1\,\unit{km}$ area around Svanhovd was selected.
-Daily averaged data for both 2018 and 2019 has been fitted with a quadratic polynomial function.
-The numerically computed root yields: BGS 122/106 day of year (doy) and EGS 261/274 day of year for 2018/2019, respectively.}
-  \label{fig:modis_Psn}
+\include{supplement_do3se-parameterization_v2.pdf}
+\label{tab:do3se_mm_param}
 \end{figure*}
-
-\clearpage
-
-\section{DO3SE parameterization}
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{javis_funcs_temp_hist_evergreen}\\
-  \includegraphics[width=8.3cm]{javis_funcs_rad_hist_evergreen}
-\caption{Construction of bespoke response functions for Norway spruce.
-(a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown toghether with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability densitiy function - PDF), respectively.
-Original mapping manual parameterization is shown in comparison as solid line.
-Note that $Q_0$ has been truncated to $0.006$.
-PPFD0.8 and PPFD1.2 refer to $\alpha$ values increasing/decreasing PPFD at $f_\mathrm{light}=0.5$ by $\pm 20\,\%$, respectively.}
-\label{fig:f_temp_spruce}
-\end{figure}
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{javis_funcs_temp_hist_birch}\\
-  \includegraphics[width=8.3cm]{javis_funcs_rad_hist_birch}
-\caption{Construction of bespoke response functions for Norway birch.
-(a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown toghether with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability densitiy function - PDF), respectively.
-Original mapping manual parameterization is shown in comparison as solid line.
-Note that $Q_0$ has been truncated to $0.006$.
-PPFD0.8 and PPFD1.2 refer to $\alpha$ values increasing/decreasing PPFD at $f_\mathrm{light}=0.5$ by $\pm 20\,\%$, respectively.}
-\label{fig:f_temp_birch}
-\end{figure}
-
-\noappendix       %% use this to mark the end of the appendix section. Otherwise the figures might be numbered incorrectly (e.g. 10 instead of 1).
-
-%% Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
-
-%% Option 1: If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
-%% They will be correctly named automatically.
-
-%% Option 2: If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.
-%% To rename them correctly to A1, A2, etc., please add the following commands in front of them:
-
-%\appendixfigures  %% needs to be added in front of appendix figures
-
-%\noappendix
-
-%\appendixtables   %% needs to be added in front of appendix tables
-
-%% Please add \clearpage between each table and/or figure. Further guidelines on figures and tables can be found below.
-
-\authorcontribution{SF has written the manuscript, collected and processed ozone and environmental data, and performed all statistical analyses.
-AVV has performed on-site observation of vegetation damage induced by ozone and contributed to theory of ozone interaction with vegetation within the introduction.
-LE...
-CO...
-ALEB...
-FS...
-TB...
-All authors contributed to discussion and conclusion of the results.} %% this section is mandatory
-
-\competinginterests{The authors declare that they have no conflict of interest.} %% this section is mandatory even if you declare that no competing interests are present
-
-%\disclaimer{TEXT} %% optional section
-
-\begin{acknowledgements}
-  Bj{\o}rg Rognerud for processing SeNorge.no data with respect for begin of growing season.
-  Tore Flatlandsmo Berglen for hourly pressure data from Svanhovd.
-\end{acknowledgements}
-
-%% REFERENCES
-
-%% The reference list is compiled as follows:
-
-%\begin{thebibliography}{}
-
-%\bibitem[AUTHOR(YEAR)]{LABEL1}
-%REFERENCE 1
-
-%\bibitem[AUTHOR(YEAR)]{LABEL2}
-%REFERENCE 2
-
-%\end{thebibliography}
-
-%% Since the Copernicus LaTeX package includes the BibTeX style file copernicus.bst,
-%% authors experienced with BibTeX only have to include the following two lines:
-%%
-\bibliographystyle{copernicus}
-\bibliography{VegOzone.bib}
-%%
-%% URLs and DOIs can be entered in your BibTeX file as:
-%%
-%% URL = {http://www.xyz.org/~jones/idx_g.htm}
-%% DOI = {10.5194/xyz}
-
-%% LITERATURE CITATIONS
-%%
-%% command                        & example result
-%% \citet{jones90}|               & Jones et al. (1990)
-%% \citep{jones90}|               & (Jones et al., 1990)
-%% \citep{jones90,jones93}|       & (Jones et al., 1990, 1993)
-%% \citep[p.~32]{jones90}|        & (Jones et al., 1990, p.~32)
-%% \citep[e.g.,][]{jones90}|      & (e.g., Jones et al., 1990)
-%% \citep[e.g.,][p.~32]{jones90}| & (e.g., Jones et al., 1990, p.~32)
-%% \citeauthor{jones90}|          & Jones et al.
-%% \citeyear{jones90}|            & 1990
-
-%% LITERATURE CITATIONS
-%%
-%% command                        & example result
-%% \citet{jones90}|               & Jones et al. (1990)
-%% \citep{jones90}|               & (Jones et al., 1990)
-%% \citep{jones90,jones93}|       & (Jones et al., 1990, 1993)
-%% \citep[p.~32]{jones90}|        & (Jones et al., 1990, p.~32)
-%% \citep[e.g.,][]{jones90}|      & (e.g., Jones et al., 1990)
-%% \citep[e.g.,][p.~32]{jones90}| & (e.g., Jones et al., 1990, p.~32)
-%% \citeauthor{jones90}|          & Jones et al.
-%% \citeyear{jones90}|            & 1990
-
-%% FIGURES
-
-%% When figures and tables are placed at the end of the MS (article in one-column style), please add \clearpage
-%% between bibliography and first table and/or figure as well as between each table and/or figure.
-
-% The figure files should be labelled correctly with Arabic numerals (e.g. fig01.jpg, fig02.png).
-
-%% ONE-COLUMN FIGURES
-
-%%f
-%\begin{figure}[t]
-%\includegraphics[width=8.3cm]{FILE NAME}
-%\caption{TEXT}
-%\end{figure}
-%
-%%% TWO-COLUMN FIGURES
-%
-%%f
-%\begin{figure*}[t]
-%\includegraphics[width=12cm]{FILE NAME}
-%\caption{TEXT}
-%\end{figure*}
-%
-%
-%%% TABLES
-%%%
-%%% The different columns must be seperated with a & command and should
-%%% end with \\ to identify the column brake.
-%
-%%% ONE-COLUMN TABLE
-%
-%%t
-%\begin{table}[t]
-%\caption{TEXT}
-%\begin{tabular}{column = lcr}
-%\tophline
-%
-%\middlehline
-%
-%\bottomhline
-%\end{tabular}
-%\belowtable{} % Table Footnotes
-%\end{table}
-%
-%%% TWO-COLUMN TABLE
-%
-%%t
-%\begin{table*}[t]
-%\caption{TEXT}
-%\begin{tabular}{column = lcr}
-%\tophline
-%
-%\middlehline
-%
-%\bottomhline
-%\end{tabular}
-%\belowtable{} % Table Footnotes
-%\end{table*}
-%
-%%% LANDSCAPE TABLE
-%
-%%t
-%\begin{sidewaystable*}[t]
-%\caption{TEXT}
-%\begin{tabular}{column = lcr}
-%\tophline
-%
-%\middlehline
-%
-%\bottomhline
-%\end{tabular}
-%\belowtable{} % Table Footnotes
-%\end{sidewaystable*}
-%
-%
-%%% MATHEMATICAL EXPRESSIONS
-%
-%%% All papers typeset by Copernicus Publications follow the math typesetting regulations
-%%% given by the IUPAC Green Book (IUPAC: Quantities, Units and Symbols in Physical Chemistry,
-%%% 2nd Edn., Blackwell Science, available at: http://old.iupac.org/publications/books/gbook/green_book_2ed.pdf, 1993).
-%%%
-%%% Physical quantities/variables are typeset in italic font (t for time, T for Temperature)
-%%% Indices which are not defined are typeset in italic font (x, y, z, a, b, c)
-%%% Items/objects which are defined are typeset in roman font (Car A, Car B)
-%%% Descriptions/specifications which are defined by itself are typeset in roman font (abs, rel, ref, tot, net, ice)
-%%% Abbreviations from 2 letters are typeset in roman font (RH, LAI)
-%%% Vectors are identified in bold italic font using \vec{x}
-%%% Matrices are identified in bold roman font
-%%% Multiplication signs are typeset using the LaTeX commands \times (for vector products, grids, and exponential notations) or \cdot
-%%% The character * should not be applied as mutliplication sign
-%
-%
-%%% EQUATIONS
-%
-%%% Single-row equation
-%
-%\begin{equation}
-%
-%\end{equation}
-%
-%%% Multiline equation
-%
-%\begin{align}
-%& 3 + 5 = 8\\
-%& 3 + 5 = 8\\
-%& 3 + 5 = 8
-%\end{align}
-%
-%
-%%% MATRICES
-%
-%\begin{matrix}
-%x & y & z\\
-%x & y & z\\
-%x & y & z\\
-%\end{matrix}
-%
-%
-%%% ALGORITHM
-%
-%\begin{algorithm}
-%\caption{...}
-%\label{a1}
-%\begin{algorithmic}
-%...
-%\end{algorithmic}
-%\end{algorithm}
-%
-%
-%%% CHEMICAL FORMULAS AND REACTIONS
-%
-%%% For formulas embedded in the text, please use \chem{}
-%
-%%% The reaction environment creates labels including the letter R, i.e. (R1), (R2), etc.
-%
-%\begin{reaction}
-%%% \rightarrow should be used for normal (one-way) chemical reactions
-%%% \rightleftharpoons should be used for equilibria
-%%% \leftrightarrow should be used for resonance structures
-%\end{reaction}
-%
-%
-%%% PHYSICAL UNITS
-%%%
-%%% Please use \unit{} and apply the exponential notation
-
-\end{document}

### 2021-03-03 VegOzone.tex
Moved figures to supplementary material

@@
 
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{DO3SE_results_pody_gsto_o3}
-  \caption{$\mathrm{POD_y}$ modeling results from $\mathrm{DO_3SE}$ model for three different PFTs over day of year (doy), March--October.
-A flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) has been chosen.
-\chem{[O_3]} are plotted on the same axis and scales as $G_\text{sto}^\text{leaf}$ but in units of $\unit{ppb}$.
-(a, b) Birch; (c, d) Norway spruce; (e, f) perennial grassland.
-From left to right: 2018, 2019.
-{\bf TODO: substitute with figure for Boreal PFTs?}}
-  \label{fig:pody}
-\end{figure*}
-
 \begin{table}[t]
@@
 
-%\begin{figure}[t]
-%  \includegraphics[width=8.3cm]{DO3SE_results_rel}
-%  \caption{Daily averaged $\mathrm{\Delta POD_y}/\Delta t$ from $\mathrm{DO_3SE}$ model for three different PFTs over day of year (doy), March--October.
-Error bars/-bands indicate $1 \sigma$ standard deviation.
-(a) Birch; (b) Norway spruce; (c) perennial grassland.}
-%  \label{fig:pody_rel}
-%\end{figure}
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{do3se_results_pod_v3_Birch}\\
+  \includegraphics[width=8.3cm]{do3se_results_pod_v3_Spruce}\\
+  \includegraphics[width=8.3cm]{do3se_results_pod_v3_Grassland}
+  \caption{Comprehensive comparison of $\mathrm{DO_3SE}$ modeling results with default parameterization from mapping manual (MM)and bespoke parameters for Svanhovd.
+$\mathrm{POD_1}$ is shown for the different temperature (MM, boreal, cold) and light threshold (MM, PPFD0.8, PPFD1.2) parameterizations at the end of the MM GS (lighter colors and dotted lines) and a derived bespoke GS.
+(a) Birch; (b) Norway spruce; (c) Perennial grassland.}
+  \label{fig:pody_rel}
+\end{figure}
 
@@
 
-\subsection{Temperature, precipitation, and global irradiance}     %% Appendix A1, A2, etc.
+%\subsection{Temperature, precipitation, and global irradiance}     %% Appendix A1, A2, etc.
 
@@
   \caption{Observed global irradiance at Svanvik.
-The cyan lines indicate the $1\,\sigma$ level deducted from the climatology.
-(a) Climatology (hourly maxima, mean, and minima); deviation from climatology (b) 2018; (c) 2019.}
+(a) Climatology (hourly maxima, mean, and minima); deviation from climatology (b) 2018; (c) 2019.
+The cyan lines indicate the $1\,\sigma$ level derived from climatology.}
   \label{fig:global_rad_clim}
 \end{figure*}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{test_fennoscandia}
-  \caption{Maximum daily sun shine duration based on geometrical calculations at different sites in Fennoscandia.
-Midnight sun conditions at Jergul/Karasjok and Svanvik prevail from the end of May until the end of July, while at Esrange and Pallas they only prevail from the beginning of June until mid of July.}
-  \label{fig:sunlight_fennoscandia}
-\end{figure*}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{greening_season_change_Svanvik}
-  \caption{Estimated shift and prolongation of greening season at Svanhovd over the past 6 decades based on data from SeNorge.no {\bf TODO: citation http://www.senorge.no}.}
-  \label{fig:greening_season_change_Svanvik}
-\end{figure*}
-\clearpage
-
-\subsection{Ozone}
 
@@
 \begin{figure*}[t]
-  \includegraphics[width=12cm]{ozone_climatology_fenoscandic_obs_residuals}
-  \caption{Probability density functions of ozone concentration residuals. 2018 with respect to respective climatology for different sites in Fennoscandia. (a) Esrange; (b) Pallas.}
-  \label{fig:ozone_climatology_fenoscandic_obs_residuals}
-\end{figure*}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{ozone_climatology_fenoscandic_obs_residuals-Svanvik}
-  \caption{Probability density functions of ozone concentration residuals. 2018/19 observations at Svanhovd with respect to derived climatologies for (a) Northern Fennoscandia; (b) Svanvik.}
-  \label{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}
-\end{figure*}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{ozone_climatology_fenoscandic_obs_test}
-  \caption{Student's t-test assuming same sample uncertainty in both, climatology and 2018 observations for different sites in Fennoscandia.
-\chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals} are significantly different from zero-hypothesis on the $1\,\sigma$ level.
-(a) Esrange; (b) Pallas.}
-  \label{fig:ozone_climatology_fenoscandic_obs_test}
-\end{figure*}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{ozone_climatology_fenoscandic_obs_test-Svanvik}
-  \caption{Student's t-test assuming same sample uncertainty in both, climatology and 2018/19 observations at Svanhovd with respect to derived climatologies for (a) Northern Fennoscandia; (b) Svanvik.
-\chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}a) are significantly different from zero-hypothesis on the $2\,\sigma$,  $1\,\sigma$ level, respectively.
-\chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}b) are not significantly different from zero-hypothesis.}
-  \label{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}
-\end{figure*}
-
-\begin{figure*}[t]
   \includegraphics[width=12cm]{ozone_reconstruction_2018_07}
@@
 
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{ozone_observation_timelag_svanvik}
-  \caption{Temporal correlation of \chem{[O_3]} data at Svanvik with other ozone monitoring stations in northern Fennoscandia.
-A negative lag means that Svanvik lags behind, while a positive lag mean the other station lags behind.
-The highest correlation with Pallas/Esrange is found at a time lag of $3\,\unit{h}$, for Jergul/Karasjok at $1\,\unit{h}$, and for Janiskoski at $-20\,\unit{h}$.}
-  \label{fig:time_lag_correlation}
-\end{figure*}
-
-% Shall go to supplement!
-%\begin{figure*}[t]
-%  \includegraphics[width=12cm]{ozone_fennoscandic_obs_aot40}
-%  \caption{Integrated AOT40 for various sites in northern Fennoscandia over the course of 33 years.
-For simplicity, we integrated hourly ozone between $1\,\unit{am}-11\,\unit{pm}$ in any case, although a night time light intensity $> 50\,\unit{W\,m^{-2}}$ is only measured during midnight sun conditions.
-Since night time \chem{[O_3]} is mostly below $40\,\unit{ppb}$, the such induced high bias in spring should be tolerable.
-The dashed red lines indicate the respective threshold given by the EU directive.
-(a) May 1 -- July 31 (b) April 1 -- September 30.}
-%  \label{fig:fennoscandic_aot40}
-%\end{figure*}
+\clearpage
+
+\section{DO3SE parameterization}
 
@@
 \end{figure*}
-
-\clearpage
-
-\section{DO3SE parameterization}
 
@@
 \end{figure}
+
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{DO3SE_results_pody_gsto_o3}
+  \caption{$\mathrm{POD_y}$ modeling results from $\mathrm{DO_3SE}$ model for three different PFTs over day of year (doy), March--October.
+A flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) has been chosen.
+\chem{[O_3]} are plotted on the same axis and scales as $G_\text{sto}^\text{leaf}$ but in units of $\unit{ppb}$.
+(a, b) Birch; (c, d) Norway spruce; (e, f) perennial grassland.
+From left to right: 2018, 2019.
+{\bf TODO: substitute with figure for Boreal PFTs?}}
+  \label{fig:pody}
+\end{figure*}
 

### 2021-03-03 VegOzone.tex
Use final figures

@@
 
-\graphicspath{{pictures/}}
+%\graphicspath{{pictures/}}
+\graphicspath{{figures/}}
 
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{svanhovd_researchstation}\\
-  \includegraphics[width=8.3cm]{IMG_7739.JPG}  
+  %\includegraphics[width=8.3cm]{svanhovd_researchstation}\\
+  %\includegraphics[width=8.3cm]{IMG_7739.JPG}
+  \includegraphics[width=8.3cm]{fig01}
   \caption{NIBIO Environment Centre Svanhovd close by the settlement of Svanvik, Norway.
@@
 \begin{figure*}[t]
-  \includegraphics[width=12cm]{data_svanvik_2018_2019}
+  %\includegraphics[width=12cm]{data_svanvik_2018_2019}
+  \includegraphics[width=12cm]{fig02}
   \caption{Observational data from atmospheric monitoring at Svanhovd in 2018/19.
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{station_map_fennoscandia}
+  %\includegraphics[width=8.3cm]{station_map_fennoscandia}
+  \includegraphics[width=8.3cm]{fig03}
   \caption{Cap of the North.
@@
 \begin{figure*}[t]
-  \includegraphics[width=12cm]{ozone_timeseries_fenoscandic_obs.png}
+  %\includegraphics[width=12cm]{ozone_timeseries_fenoscandic_obs.png}
+  \includegraphics[width=12cm]{fig04}
   \caption{Time series of ozone observations in northern Fennoscandia (Tab.~\ref{tab:ebas_obs}).
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{climatediagram}
+  %\includegraphics[width=8.3cm]{climatediagram}
+  \includegraphics[width=8.3cm]{fig05}
   \caption{Climate diagram following Walter and Lieth.
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_spline}
+  %\includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_spline}
+  \includegraphics[width=8.3cm]{fig06}
   \caption{Derivation of ozone climatologies.
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_climatologies_portrait}
+  %\includegraphics[width=8.3cm]{ozone_climatologies_portrait}
+  \includegraphics[width=8.3cm]{fig07}
   \caption{Comparison of climatologies derived from observational data (EBAS) with various ozone reanalysis products: (a) ECMWF: MACC reanalysis; (b) ECMWF: CAMS reanalysis; (c) Copernicus: Regional model reanalysis ensemble mean.
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{temperature_signific_1}
+  %\includegraphics[width=8.3cm]{temperature_signific_1}
+  \includegraphics[width=8.3cm]{fig08}
   \caption{Temperature anomalies at Svanvik.
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{precipitation_signific_1}
+  %\includegraphics[width=8.3cm]{precipitation_signific_1}
+  \includegraphics[width=8.3cm]{fig09}
   \caption{Precipitation anomalies at Svanvik.
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{global_rad_signific_1}
+  %\includegraphics[width=8.3cm]{global_rad_signific_1}
+  \includegraphics[width=8.3cm]{fig10}
   \caption{Global irradiance anomalies at Svanvik.
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_signific_1}
+  %\includegraphics[width=8.3cm]{ozone_signific_1}
+  \includegraphics[width=8.3cm]{fig11}
   \caption{Ozone concentration anomalies at Svanvik.
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{javis_funcs_temp_hist_grassland}\\
-  \includegraphics[width=8.3cm]{javis_funcs_rad_hist_grassland}
+  %\includegraphics[width=8.3cm]{javis_funcs_temp_hist_grassland}\\
+  %\includegraphics[width=8.3cm]{javis_funcs_rad_hist_grassland}
+  \includegraphics[width=8.3cm]{fig12}
 \caption{Construction of bespoke response functions for grassland.
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{javis_func_opt_mean.pdf}
+  %\includegraphics[width=8.3cm]{javis_func_opt_mean}
+  \includegraphics[width=8.3cm]{fig13}
   \caption{Proposed metric to test bespoke response functions.}
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{do3se_results_pod_v3_Birch}\\
-  \includegraphics[width=8.3cm]{do3se_results_pod_v3_Spruce}\\
-  \includegraphics[width=8.3cm]{do3se_results_pod_v3_Grassland}
+  %\includegraphics[width=8.3cm]{do3se_results_pod_v3_Birch}\\
+  %\includegraphics[width=8.3cm]{do3se_results_pod_v3_Spruce}\\
+  %\includegraphics[width=8.3cm]{do3se_results_pod_v3_Grassland}
+  \includegraphics[width=8.3cm]{fig14}
   \caption{Comprehensive comparison of $\mathrm{DO_3SE}$ modeling results with default parameterization from mapping manual (MM)and bespoke parameters for Svanhovd.
@@
 %\videosupplement{TEXT} %% use this section when having video supplements available
+\clearpage
 
@@
 \begin{figure*}[t]
-  \includegraphics[width=12cm]{global_rad_clim}
+  \includegraphics[width=12cm]{figA1}
   \caption{Observed global irradiance at Svanvik.
@@
 \begin{figure*}[t]
-  \includegraphics[width=12cm]{density_distribution}
+  \includegraphics[width=12cm]{figA2}
   \caption{Probability densities and correlation coefficient for \chem{[O_3]} between different sites in northern Fennoscandia.
@@
 \begin{figure*}[t]
-  \includegraphics[width=12cm]{ozone_reconstruction_2018_07}
+  \includegraphics[width=12cm]{figA3}
   \caption{Reconstruction of missing \chem{[O_3]} observation data at Svanvik in July 2018 based on derived hourly climatologies of northern Fennoscandia, Svanvik, and data Pallas.
@@
 \begin{figure*}[t]
-  \includegraphics[width=12cm]{modis_Psn}
+  \includegraphics[width=12cm]{figB1}
   \caption{Estimated begin (BGS) and end (EGS) of growing season for coniferous trees from MODIS Aqua/Terra net photosynthesis (PSN) product.
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{javis_funcs_temp_hist_evergreen}\\
-  \includegraphics[width=8.3cm]{javis_funcs_rad_hist_evergreen}
+  %\includegraphics[width=8.3cm]{javis_funcs_temp_hist_evergreen}\\
+  %\includegraphics[width=8.3cm]{javis_funcs_rad_hist_evergreen}
+  \includegraphics[width=12cm]{figB2}
 \caption{Construction of bespoke response functions for Norway spruce.
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{javis_funcs_temp_hist_birch}\\
-  \includegraphics[width=8.3cm]{javis_funcs_rad_hist_birch}
+  %\includegraphics[width=8.3cm]{javis_funcs_temp_hist_birch}\\
+  %\includegraphics[width=8.3cm]{javis_funcs_rad_hist_birch}
+  \includegraphics[width=12cm]{figB3}
 \caption{Construction of bespoke response functions for Norway birch.
@@
 \begin{figure*}[t]
-  \includegraphics[width=12cm]{DO3SE_results_pody_gsto_o3}
+  \includegraphics[width=12cm]{figB4}
   \caption{$\mathrm{POD_y}$ modeling results from $\mathrm{DO_3SE}$ model for three different PFTs over day of year (doy), March--October.

### 2021-03-03 VegOzone.tex
Include combined figure and revise its caption and referencing.

@@
 \label{subsubsec:anomal_tpq}
-In Fig.~\ref{fig:plot_temperature_anomalies_svanvik}, temperature anomalies at Svanvik are shown as percentage of days warmer/colder than $\pm 1\,\sigma$ from the climatological mean for each month.
+In Fig.~\ref{fig:anomalies_svanvik}a), temperature anomalies at Svanvik are shown as percentage of days warmer/colder than $\pm 1\,\sigma$ from the climatological mean for each month.
 Negative deviations from climatology are displayed as negative percentage.
@@
 \begin{figure}[t]
+  \includegraphics[width=12cm]{fig08-11}
+  \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as percentages of days deviating from climatological mean by means of standart deviation for each month.
+Negative deviations from climatology are shown as negative percentage.
+The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
+The hatched area between the dashed lines indicates statistical significance under the assumption of a normal distribution ($15.9\,\unit{\%}$, $30.8\,\unit{\%}$ for precip.).
+(a) Temperature, days warmer/colder than $\pm 1\,\sigma$; (b) precipitation, days wetter/drier than $\pm \frac{1}{2}\,\sigma$; (c) global irradiance, days lighter/darker than $\pm 1\,\sigma$; (d) ozone, days higher/lower than $\pm 1\,\sigma$.
+The July 2018 ozone data corrected for data acquisition problems is indicated with the star (refer Section~\ref{subsec:ozone_reco} for details).
+The growing season of 2018 was significantly warmer, brighter, and slightly drier than average, especially in May and July.
+Ozone was significantly elevated in May.
+The growing season of 2019 was fairly average but significantly warmer than average in April and colder in July/August.
+No significant elevation in ozone was found.}
+  \label{fig:anomalies_svanvik}
+\end{figure}
+
+%\begin{figure}[t]
   %\includegraphics[width=8.3cm]{temperature_signific_1}
-  \includegraphics[width=8.3cm]{fig08}
-  \caption{Temperature anomalies at Svanvik.
+  %\includegraphics[width=8.3cm]{fig08}
+  %\caption{Temperature anomalies at Svanvik.
 Percentage of days warmer/colder than $\pm 1\,\sigma$ from climatological mean displayed for each month.
@@
 The summer of 2019 was fairly average but significantly warmer than average in April and colder in July/August.}
-  \label{fig:plot_temperature_anomalies_svanvik}
-\end{figure}
-
-In a similar way, the precipitation anomalies at Svanvik are displayed in Fig.~\ref{fig:plot_precipitation_anomalies_svanvik}.
+  %\label{fig:plot_temperature_anomalies_svanvik}
+%\end{figure}
+
+In a similar way, the precipitation anomalies at Svanvik are displayed in Fig.~\ref{fig:anomalies_svanvik}b).
 However, the percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ from the climatological mean is shown for each month, due to the larger variability in observed precipitation too few days showed a significance of $1 \sigma$ or higher.
@@
 
-\begin{figure}[t]
+%\begin{figure}[t]
   %\includegraphics[width=8.3cm]{precipitation_signific_1}
-  \includegraphics[width=8.3cm]{fig09}
-  \caption{Precipitation anomalies at Svanvik.
+  %\includegraphics[width=8.3cm]{fig09}
+  %\caption{Precipitation anomalies at Svanvik.
 Percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ from climatological mean displayed for each month.
@@
 Summer was significantly dry only in 2018, while fall was significantly dry in both years.}
-  \label{fig:plot_precipitation_anomalies_svanvik}
-\end{figure}
-
-Similarly, the global irradiance anomalies at Svanvik are presented in Fig.~\ref{fig:global_rad_signific}.
+  %\label{fig:plot_precipitation_anomalies_svanvik}
+%\end{figure}
+
+Similarly, the global irradiance anomalies at Svanvik are presented in Fig.~\ref{fig:anomalies_svanvik}c).
 The percentage of days above/below $\pm 1\,\sigma$ from the climatological mean is displayed for each month.
@@
 
-\begin{figure}[t]
+%\begin{figure}[t]
   %\includegraphics[width=8.3cm]{global_rad_signific_1}
-  \includegraphics[width=8.3cm]{fig10}
-  \caption{Global irradiance anomalies at Svanvik.
+  %\includegraphics[width=8.3cm]{fig10}
+  %\caption{Global irradiance anomalies at Svanvik.
 Percentage of days above/below $\pm 1\,\sigma$ from climatological mean displayed for each month.
@@
 In summer 2018 (May/July), global irradiance was significantly higher than on average, while 2019 had significantly less than normal irradiance in April/May and more in June.}
-  \label{fig:global_rad_signific}
-\end{figure}
+  %\label{fig:global_rad_signific}
+%\end{figure}
 
@@
 In accordance to the above findings, we adjust the derived ozone climatology for Svanvik by adding a constant bias correction factor $\Delta\chem{[O_3]_\mathrm{bias-corr}} = 1.2\,\unit{ppb}$ and estimate the ozone concentration anomalies for 2018/19.
-In Fig.~\ref{fig:ozone_signific}, the resulting percentage of days above/below $\pm 1\,\sigma$ from the bias-corrected climatological mean is displayed for each month.
+In Fig.~\ref{fig:anomalies_svanvik}d), the resulting percentage of days above/below $\pm 1\,\sigma$ from the bias-corrected climatological mean is displayed for each month.
 Negative deviations from climatology are shown as negative percentage.
@@
 
-\begin{figure}[t]
+%\begin{figure}[t]
   %\includegraphics[width=8.3cm]{ozone_signific_1}
-  \includegraphics[width=8.3cm]{fig11}
-  \caption{Ozone concentration anomalies at Svanvik.
+  %\includegraphics[width=8.3cm]{fig11}
+  %\caption{Ozone concentration anomalies at Svanvik.
 Ozone climatology has been corrected for an constant negative bias of $-1.2\,\unit{ppb}$.
@@
 In both years ozone concentrations were significantly below average in November.}
-  \label{fig:ozone_signific}
-\end{figure}
+  %\label{fig:ozone_signific}
+%\end{figure}
 

### 2021-03-03 VegOzone.tex
Input of supplementary material - final pdf needs to be seperated

@@
 A climatology derived for global irradiance $Q_0$ is shown in Appendix Fig.~\ref{fig:global_rad_clim}a) for both, climatological average of hourly maximum $\left<Q_0^\mathrm{max}\right>$, minimum $\left<Q_0^\mathrm{min}\right>$, and mean $\left<Q_0^\mathrm{mean}\right>$ irradiance observed at Svanvik.
-In agreement with the geometrically derived sunshine duration shown in Appendix Fig.~\ref{fig:sunlight_fennoscandia}, $\left<Q_0^\mathrm{max}\right>$ approaches zero in December/January commonly known as polar night and reaches its maximum in June/July commonly referred to as midnight sun conditions.
+In agreement with the geometrically derived sunshine duration shown in Supplement Fig.~\ref{fig:sunlight_fennoscandia}, $\left<Q_0^\mathrm{max}\right>$ approaches zero in December/January commonly known as polar night and reaches its maximum in June/July commonly referred to as midnight sun conditions.
 The highest observed $\left<Q_0^\mathrm{max}\right>$ is about $800\,\unit{W\,m^{-2}}$, while on average $\left<Q_0^\mathrm{mean}\right>$ it is about $450\,\unit{W\,m^{-2}}$ (regularly reached in May--July).
@@
 
+\input{supplements/VegOzone_supplement}
+
 \end{document}

### 2021-03-03 supplements/VegOzone_supplement.tex
Input of supplementary material - final pdf needs to be seperated

@@
-\begin{figure*}[t]
-  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs}
-  \caption{S1 Multi-annual mean of daily mean ozone at observation sites in Fennoscandia.
+\graphicspath{{pictures/}}          % graphics
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
+
+\clearpage
+\setcounter{page}{1}
+\setcounter{figure}{0}
+
+%\begin{figure*}[t]
+%  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs}
+%  \caption{Multi-annual mean of daily mean ozone at observation sites in Fennoscandia.
 The large spread in case of Svanvik and Janiskoski is due to the lower statistics.
@@
 The annual average ozone concentration \chem{\left<[O_3]\right>} at Svanvik is $6.6\,\unit{ppb}$ lower than at the sites at Jergul/Karasjok (NOR), Esrange (SWE), and Pallas (FIN).}
-  \label{fig:ozone_climatology_fenoscandic_obs}
-\end{figure*}
+%  \label{fig:ozone_climatology_fenoscandic_obs}
+%\end{figure*}
 
-\begin{figure*}[t]
+\begin{figure}[t]
   \includegraphics[width=12cm]{test_fennoscandia}
@@
   \label{fig:sunlight_fennoscandia}
-\end{figure*}
+\end{figure}
 
@@
 \begin{figure*}[t]
-\include{supplement_do3se-parameterization_v2.pdf}
+  \centering
+  \includegraphics[width=1\textwidth]{supplements/supplement_do3se-parameterization_v2}
 \label{tab:do3se_mm_param}

### 2021-03-03 VegOzone.tex
Change references accordingly to renamed files

@@
 In the following, we will look at derived ozone climatologies from different perspectives.
-The multi-annual means of daily mean ozone, for simplicity here without standard deviations, at observation sites in northern Fennoscandia (refer to Table~\ref{tab:ebas_obs} for details) are shown in Supplement Fig.~S1.
+The multi-annual means of daily mean ozone, for simplicity here without standard deviations, at observation sites in northern Fennoscandia (refer to Table~\ref{tab:ebas_obs} for details) are shown in Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs}.
 All climatologies peak in spring, with peak values of about $48\,\unit{ppb}$ in late April.
@@
 The decrease in \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} begins about one month ahead of the estimated onset of the greening season.
-Based on data from SeNorge.no and the agricultural rule of thumb that the daily average temperature has to be above $5\,\unit{^\circ C}$ for $5\,\unit{days}$ consecutively, we find $\left<G_\mathrm{begin}\right> = (150 \pm 14)\,\unit{days of the year}$ at Svanvik (Appendix Fig~\ref{fig:greening_season_change_Svanvik}), which translates into the end of May.
+Based on data from SeNorge.no and the agricultural rule of thumb that the daily average temperature has to be above $5\,\unit{^\circ C}$ for $5\,\unit{days}$ consecutively, we find $\left<G_\mathrm{begin}\right> = (150 \pm 14)\,\unit{days of the year}$ at Svanvik (Supplement Fig~\ref{fig:greening_season_change_Svanvik}), which translates into the end of May.
 This finding is in line with more extensive studies \citep[][e.g.,]{IJB:Karlsen2007,AFM:Linderholm}.
@@
 \begin{figure}[t]
-  \includegraphics[width=12cm]{fig08-11}
+  \includegraphics[width=12cm]{fig08}
   \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as percentages of days deviating from climatological mean by means of standart deviation for each month.
@@
 
-In Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}, the probability density functions of daily mean ozone concentration residuals for Esrange and Pallas are shown.
+In Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}, the probability density functions of daily mean ozone concentration residuals for Esrange and Pallas are shown.
 Both stations display a clear deviation from the climatological mean, however, not significant.
@@
 
-Appendix Figure~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik} displays the probability density functions of daily mean ozone concentration residuals for Svanvik based on data taken in 2018/19 with respect to the climatology derived for northern Fennoscandia (denoted NF in Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}a) and the climatology derived for Svanvik (Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b).
+Supplement Figure~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik} displays the probability density functions of daily mean ozone concentration residuals for Svanvik based on data taken in 2018/19 with respect to the climatology derived for northern Fennoscandia (denoted NF in Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}a) and the climatology derived for Svanvik (Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b).
 As previously mentioned, \chem{\left<[O_3]\right>} at Svanvik are $(6.6\pm 7.1)\,\unit{ppb}$ lower than the northern Fennoscandic climatology.
@@
 While the probability density of the residuals between the two climatologies is skewed towards negative values, the skew of 2018/19 is positive.
-With respect to the climatology for Svanvik (Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b), the probability density functions average for 2018/19 lies at $\Delta\chem{\left<[O_3]\right>} = (2.6\pm 5.8)\,\unit{ppb}$ and $\Delta\chem{\left<[O_3]\right>} = (1.2\pm 5.0)\,\unit{ppb}$, respectively.
+With respect to the climatology for Svanvik (Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b), the probability density functions average for 2018/19 lies at $\Delta\chem{\left<[O_3]\right>} = (2.6\pm 5.8)\,\unit{ppb}$ and $\Delta\chem{\left<[O_3]\right>} = (1.2\pm 5.0)\,\unit{ppb}$, respectively.
 Neither of the \chem{\Delta\left<[O_3]\right>} is significant by statistical means, but daily average ozone concentrations are slightly elevated in 2018 compared to 2019.
@@
 To further investigate the ozone residuals, we conduct a Student's~t-test ($t = \frac{\chem{\Delta[O_3]}}{\sigma}$), assuming the same sample uncertainty $\sigma$ in both, climatology and the year in question.
-Results for Esrange and Pallas are shown in Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test}.
+Results for Esrange and Pallas are shown in Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test}.
 This t-test clearly shows, that at both sites the residuals are significant on the $1\sigma$ level for at least $50\,\unit{\%}$.
-In Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}, the test statistics for Svanvik are displayed.
-From Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}a), we can deduce, that the residuals between the Svanvik climatology and the northern Fennoscandic climatology are significant on the $2\sigma$ level for at least half of the data, while the residuals are less certain in 2018 and 2019.
-In Appendix Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}b) the same is shown for the years 2018/19 with respect to the climatology from Svanvik.
+In Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}, the test statistics for Svanvik are displayed.
+From Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}a), we can deduce, that the residuals between the Svanvik climatology and the northern Fennoscandic climatology are significant on the $2\sigma$ level for at least half of the data, while the residuals are less certain in 2018 and 2019.
+In Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}b) the same is shown for the years 2018/19 with respect to the climatology from Svanvik.
 We find that less than $50\,\unit{\%}$ of ozone concentration residuals are significant on the $1\sigma$ level.
@@
 We derive hourly climatologies for individual stations and combine Esrange/Pallas to derive a climatology for northern Fennoscandia for July (Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}a).
-We compute temporal correlations between Svanvik and the other stations in northern Fennoscandia to estimate a time lag (Appendix Fig.~\ref{fig:time_lag_correlation}.
+We compute temporal correlations between Svanvik and the other stations in northern Fennoscandia to estimate a time lag (Supplement Fig.~\ref{fig:time_lag_correlation}.
 We find $3\,\unit{h}$ for Esrange and Pallas with respect to Svanvik.
@@
 We construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to the mapping manual default.
-In the same way, \emph{Boreal} refers to a hypothetical species which is very tolerant to cold but is sensitive to high temperatures, and most efficient at cool temperatures (close to the climatological mean temperature of 1992--2000).
+In the same way, \emph{boreal} refers to a hypothetical species which is very tolerant to cold but is sensitive to high temperatures, and most efficient at cool temperatures ($T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000).
 All three types for grassland are depicted together with the temperature PDF in Fig.~\ref{fig:f_temp_grassland}a).
@@
   %\includegraphics[width=8.3cm]{javis_funcs_rad_hist_grassland}
-  \includegraphics[width=8.3cm]{fig12}
+  \includegraphics[width=8.3cm]{fig09}
 \caption{Construction of bespoke response functions for grassland.
@@
   %\includegraphics[width=8.3cm]{javis_func_opt_mean}
-  \includegraphics[width=8.3cm]{fig13}
+  \includegraphics[width=8.3cm]{fig10}
   \caption{Proposed metric to test bespoke response functions.}
@@
 Start ($A_\text{start}$) and end ($A_\text{end}$) of growing season for each PFT are given in units of day of year (doy) and estimated as follows.
-Evergreen needle leaf trees are photosynthetic active already at low temperatures \citep{TB:Kolari2007}.
-This leads necessarily to an increase in observable net photosynthesis.
-Hence, we use a MODIS (Aqua/Terra) retrieved net photosynthesis product for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine onset and end of growing season.
+Evergreen needle leaf trees show photosynthetic activity already at low temperatures \citep{TP:Wallin2013, TB:Kolari2007}.
+We use a MODIS (Aqua/Terra) retrieved net photosynthesis product for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine onset and end of growing season.
 Net photosynthesis data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}).
@@
   %\includegraphics[width=8.3cm]{do3se_results_pod_v3_Grassland}
-  \includegraphics[width=8.3cm]{fig14}
+  \includegraphics[width=8.3cm]{fig11}
   \caption{Comprehensive comparison of $\mathrm{DO_3SE}$ modeling results with default parameterization from mapping manual (MM)and bespoke parameters for Svanhovd.
-$\mathrm{POD_1}$ is shown for the different temperature (MM, boreal, cold) and light threshold (MM, PPFD0.8, PPFD1.2) parameterizations at the end of the MM GS (lighter colors and dotted lines) and a derived bespoke GS.
+$\mathrm{POD_1}$ is shown for the different temperature response functions (MM, boreal, cold) and light thresholds (MM, PPFD0.8, PPFD1.2) at the end of a MM GS (lighter hue and dotted lines) and a derived bespoke GS.
 (a) Birch; (b) Norway spruce; (c) Perennial grassland.}

### 2021-03-03 VegOzone.tex
Start to work on bespoke metric plot and text.

@@
 To investigate which of the environmental factors lead to the observed ozone damage on clover and tobacco in 2018 in contrast to 2019, when no such damage was found on the clovers but only on the sensitive tobacco cultivar, we model ozone uptake of different natural and semi-natural vegetation using the $\mathrm{DO_3SE}$ model and develop bespoke subarctic prameterizations for common plant species.
-In the following, we give an account of the $\mathrm{DO_3SE}$ model and the used method for deriving the bespoke parameters.
+We first give an account of the $\mathrm{DO_3SE}$ model, input data, and the used method for deriving the bespoke parameters.
+Then we will present and discuss the modeling results.
 
@@
 \subsection{Model parameters}
-We roughly estimated land use types for an approximately $1\times 1\,\unit{km}$ patch centered at the NIBIO Environment Centre Svanhovd from Fig.~\ref{fig:svanhovd_research_station}: $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban.
-
-Default PFTs are derived from \citet{ICP:MappingManual2017} and will be referred to as \emph{mapping manual} abbreviate MM.
+From Fig.~\ref{fig:svanhovd_research_station}, we roughly estimated land use types for an approximately $1\times 1\,\unit{km}$ patch centered at the NIBIO Environment Centre Svanhovd : $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban.
+
+Default PFTs are derived from \citet{ICP:MappingManual2017} and will be referred to as \emph{mapping manual} abbreviated as MM.
 A comprehensive list of these model parameters can be found in Supplement~Table~S1.
-Initial test simulations with the mapping manual derived PFTs showed unrealistically low $g_\mathrm{sto}$ in 2019 (Supplement~Figs.SX--SX+2) particularly for grassland.
+Initial test simulations with the mapping manual derived PFTs showed unrealistically low $g_\mathrm{sto}$ in 2019 particularly for grassland (Appendix~Fig.~\ref{fig:pody_mm_composit}).
 Only during an extended period in late July did we find substantial stomatal conductance.
@@
 We identified $f_\mathrm{T}$ as abnormally low, being the limiting factor of stomatal conductance.
-We propose bespoke parameterizations for the main vegetation types Norway spruce, birch, and perennial grassland better adapted to the climatic conditions in the target region.
-Note that these parameterizations represent hypothetical PFTs or species that have to be verified by experiments.
+For these reasons, we propose bespoke parameterizations for the main vegetation types Norway spruce, birch, and perennial grassland which are better adapted to the climatic conditions in the target region.
+
+Note that these parameterizations represent hypothetical PFTs or species that have yet to be verified by experiments.
 
@@
 We then tune the species temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) separately for higher stomatal conductance.
-We illustrate this procedure graphically by intersecting probability density functions (PDFs) of climatological data for temperature and global irradiance with the respective response function (Fig.~\ref{fig:f_temp_grassland} for grassland; for Norway spruce and birch see Appendix Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}).
-Formally, a parameterization is considered more optimal the larger the intersection area between the response function and PDF is.
-
-For the acclimatization of the temperature response functions, we assume two different types: \emph{Boreal} and \emph{cold}.
-In either cases, we further assume that the hypothetical species are more tolerant to cold conditions than the mapping manual default suggests and shift $f_\mathrm{T}$ towards colder temperatures.
-We construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to the mapping manual default.
+We illustrate this procedure graphically in Fig.~\ref{fig:f_temp_grassland} for grassland.
+We aim to increase the enclosed area between probability density function (PDF) and the respective response function.
+For comparison, we define two climatological references 1990s and 2000s.
+For Norway spruce and birch see Appendix Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}.
+
+We assume two different types of hypothetical temperature acclimations: \emph{Boreal} and \emph{cold}.
+For either, we further assume that the hypothetical species are more tolerant to cold conditions than MM suggests and shift $f_\mathrm{T}$ towards colder temperatures.
+We construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to MM.
 In the same way, \emph{boreal} refers to a hypothetical species which is very tolerant to cold but is sensitive to high temperatures, and most efficient at cool temperatures ($T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000).
-All three types for grassland are depicted together with the temperature PDF in Fig.~\ref{fig:f_temp_grassland}a).
-
-For Norway spruce, our hypothetical PFTs $T_\mathrm{opt}$ are motivated by a longer time series of both observed \chem{CO_2} uptake and temperature at two observation sites in Southern (\emph{cold}) and Northern (\emph{Boreal}) Finland presented in \citet{TB:Kolari2007}.
-We assume a similar adaption in birch.
-The parameters are comprehensibly listed in Table~\ref{tab:sensitivity_tests_temp}.
-
-For the the sensitivity regarding light, we assume that the plants may react either faster or slower to a change at lower light intensities.
+For perennial grassland these are depicted together with the temperature PDF in Fig.~\ref{fig:f_temp_grassland}a).
+
+Norway spruce was experimentally found to be active already at rather low air temperatures and could reach $60\,\unit{\%}$ photosynthetic activity as early as doy~100 \citet{TB:Wallin2013, TB:Kolari2007}.
+We derived temperature acclimation, based on time series of observed \chem{CO_2} uptake and temperature at two observation sites in Southern ($\rightarrow$\emph{cold}) and Northern ($\rightarrow$\emph{boreal}) Finland \citet{TB:Kolari2007}.
+We assume a similar acclimation for birch.
+All bespoke temperature parameters are comprehensibly listed in Table~\ref{tab:sensitivity_tests_temp}.\\
+
+For an acclimation with respect to light conditions, we hypothize that the light threshold for a $50\,\unit{\%}$ opening of plants' stomata may differ in subarctic species.
 We analytically derive the inverse function $f_{\mathrm{light},k}^{-1}$ for each PFT or species $k$
@@
 \end{equation}
-We calculate it for the mapping manual $\alpha_\mathrm{MM}$ at its halfway value $0.5$ 
+We solve Eq.~(\ref{eq:inverse_function}) for the mapping manual default value for each PFT or species $\alpha_\mathrm{MM}$ at its halfway value $0.5$ 
 \begin{equation}
@@
 and define a variation $\gamma' = \gamma \cdot \epsilon$ with $\epsilon \in \{0.8, 1.2\}$.
-We resolve Eq.~(\ref{eq:inverse_function}) for $\alpha$ and derive $\alpha'(\gamma')$.
+We resolve Eq.~(\ref{eq:inverse_function}) again for $\alpha$ and find $\alpha'(\gamma')$.
 The resulting functions are shown in Fig.~\ref{fig:f_temp_grassland}b).
-All derived parameters are given in Table~\ref{tab:sensitivity_tests_light}.
+All derived parameters are tabulated in Table~\ref{tab:sensitivity_tests_light}.
 
@@
 
+To quantify the resulting bespoke parameterizations, we propose the following metric.
 {\bf TODO: Write about Boreal and cold parameter sets and their justification.
 Write about suggested method for optimizing PFTs in a simple but physiological way based on optimizing maximum climatological mean $g_\mathrm{sto}$ and morning responds.
-If plants have a short growing season up north optimizing \chem{CO_2} uptake around noon (highest light intensity) and speed up opening of stomata at lower intensities (morning), additional contained lower climatological variance?}
-An idea to justify our choice based on the full set of climatological data for Svanvik is shown in Fig.~\ref{fig:javis_func_opt_mean}.
+If plants have a short growing season up north optimizing \chem{CO_2} uptake around noon (highest light intensity) and speed up opening of stomata at lower intensities (morning), additional contained lower climatological variance?
+An idea to justify our choice based on the full set of climatological data for Svanvik is shown in Fig.~\ref{fig:javis_func_opt_mean}.}
 
@@
   \includegraphics[width=8.3cm]{fig10}
-  \caption{Proposed metric to test bespoke response functions.}
+  \caption{Proposed metric to test bespoke response functions.
+GS (May--August) climatological averages and standart deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time) are shown for (a) Birch; (b) Norway spruce; (c) perennial grassland.
+For simplicity, $f_\mathrm{SW}$ is neglected in this calculation.
+}
   \label{fig:javis_func_opt_mean}
@@
 {\bf TODO: substitute with figure for Boreal PFTs?}}
-  \label{fig:pody}
+  \label{fig:pody_mm_composit}
 \end{figure*}

### 2021-03-03 VegOzone.tex
Rearange labels

@@
   \caption{Proposed metric to test bespoke response functions.
-GS (May--August) climatological averages and standart deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time) are shown for (a) Birch; (b) Norway spruce; (c) perennial grassland.
+GS (May--August) climatological averages and standart deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time) are shown for (a) birch; (b) Norway spruce; (c) perennial grassland.
 For simplicity, $f_\mathrm{SW}$ is neglected in this calculation.

### 2021-03-04 VegOzone.tex
First draft of results.

@@
 As Svanvik is located at lower altitude than the other sites, the earlier decline of the ozone peak may point to a stronger ozone uptake by vegetation.
-\chem{CO_2} uptake coniferous trees can occur as early as doy 100 \citep{TB:Kolari2007}.
+\chem{CO_2} uptake coniferous trees can occur as early as day of the year (\unit{doy}) 100 \citep{TB:Kolari2007, TP:Wallin2013}.
 The average ozone concentration at Svanvik is lower than at the other sites, as already mentioned, which is in league with a decrease in tropospheric ozone with decreasing altitude.
@@
 \subsection{Model parameters}
+\label{subsec:do3se_parameters}
+
 From Fig.~\ref{fig:svanhovd_research_station}, we roughly estimated land use types for an approximately $1\times 1\,\unit{km}$ patch centered at the NIBIO Environment Centre Svanhovd : $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban.
 
-Default PFTs are derived from \citet{ICP:MappingManual2017} and will be referred to as \emph{mapping manual} abbreviated as MM.
-A comprehensive list of these model parameters can be found in Supplement~Table~S1.
-Initial test simulations with the mapping manual derived PFTs showed unrealistically low $g_\mathrm{sto}$ in 2019 particularly for grassland (Appendix~Fig.~\ref{fig:pody_mm_composit}).
-Only during an extended period in late July did we find substantial stomatal conductance.
-In ecological terms, this would mean that there was almost no growth of grass in the summer of 2019 -- a prediction which is easily falsified by reality.
+Default parameters for these species or PFTs are derived from \citet{ICP:MappingManual2017} and will be referred to as \emph{mapping manual} abbreviated as MM in the following.
+A comprehensive list of model parameters can be found in Supplement~Table~S1.
+Initial test simulations with the MM PFTs showed unrealistically low $g_\mathrm{sto}$ in 2019 particularly for perennial grassland (Appendix~Fig.~\ref{fig:pody_mm_composit}).
+Only during an extended warm period in late July did we find substantial stomatal conductance.
+In ecological terms this would mean that there was almost no growth of grass in the summer of 2019 -- a prediction which is easily falsified by reality.
 We identified $f_\mathrm{T}$ as abnormally low, being the limiting factor of stomatal conductance.
-For these reasons, we propose bespoke parameterizations for the main vegetation types Norway spruce, birch, and perennial grassland which are better adapted to the climatic conditions in the target region.
-
-Note that these parameterizations represent hypothetical PFTs or species that have yet to be verified by experiments.
-
-We hypothesize that perennial vegetation will likely prioritize their growth or fast adolescence throughout the short, subarctic growing season and hence their carbon sequestration.
-Since the $\mathrm{DO_3SE}$ in its current state does not simulate $A_\mathrm{n}$, we assume for our purpose a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011}.
-We then tune the species temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) separately for higher stomatal conductance.
-We illustrate this procedure graphically in Fig.~\ref{fig:f_temp_grassland} for grassland.
+Therefor, we propose bespoke parameterizations which may describe an actual acclimation of these species to the climatic conditions in the target region.
+
+Note, however, that these parameterizations represent hypothetical PFTs or species that have yet to be verified by experiments.
+
+We hypothesize that perennial vegetation will likely prioritize their growth or fast adolescence during a short, subarctic growing season and hence their carbon sequestration.
+Because the $\mathrm{DO_3SE}$ model currently does not simulate $A_\mathrm{n}$, we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$.
+The procedure is graphically illustrated in Fig.~\ref{fig:f_temp_grassland} for grassland.
 We aim to increase the enclosed area between probability density function (PDF) and the respective response function.
-For comparison, we define two climatological references 1990s and 2000s.
-For Norway spruce and birch see Appendix Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}.
-
-We assume two different types of hypothetical temperature acclimations: \emph{Boreal} and \emph{cold}.
-For either, we further assume that the hypothetical species are more tolerant to cold conditions than MM suggests and shift $f_\mathrm{T}$ towards colder temperatures.
+We show two climatological reference periods 1990s and 2000s to indicate that the vegation is subject to ongoing climate change and acclimations may be changing.
+Resulting reponse functions for Norway spruce and downy birch are displayed in Appendix~Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}.
+
+We define two different hypothetical temperature acclimations: \emph{Boreal} and \emph{cold}.
+For either, we assume that the hypothetical species are more tolerant to cold conditions and shift $f_\mathrm{T}$ towards colder temperatures accordingly.
 We construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to MM.
 In the same way, \emph{boreal} refers to a hypothetical species which is very tolerant to cold but is sensitive to high temperatures, and most efficient at cool temperatures ($T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000).
-For perennial grassland these are depicted together with the temperature PDF in Fig.~\ref{fig:f_temp_grassland}a).
-
-Norway spruce was experimentally found to be active already at rather low air temperatures and could reach $60\,\unit{\%}$ photosynthetic activity as early as doy~100 \citet{TB:Wallin2013, TB:Kolari2007}.
-We derived temperature acclimation, based on time series of observed \chem{CO_2} uptake and temperature at two observation sites in Southern ($\rightarrow$\emph{cold}) and Northern ($\rightarrow$\emph{boreal}) Finland \citet{TB:Kolari2007}.
-We assume a similar acclimation for birch.
-All bespoke temperature parameters are comprehensibly listed in Table~\ref{tab:sensitivity_tests_temp}.\\
+For perennial grassland the resulting response functions are depicted together with the PDF of temperatur in Fig.~\ref{fig:f_temp_grassland}a).
+
+Norway spruce was experimentally found to be active already at rather low air temperatures and could reach $60\,\unit{\%}$ photosynthetic activity as early as \unit{doy}~100 \citep{TB:Kolari2007, TP:Wallin2013}.
+Based on the time series of observed \chem{CO_2} uptake and temperature at two observation sites in Southern ($\rightarrow$ \emph{cold}) and Northern ($\rightarrow$ \emph{boreal}) Finland \citep{TB:Kolari2007}, we estimate $10\,\unit{^\circ C}\le T_\mathrm{opt} \le 15\,\unit{^\circ C}$.
+We assume a similar acclimation for downy birch.
+All bespoke temperature parameters are listed in Table~\ref{tab:sensitivity_tests_temp}.\\
 
@@
 \end{equation}
-We solve Eq.~(\ref{eq:inverse_function}) for the mapping manual default value for each PFT or species $\alpha_\mathrm{MM}$ at its halfway value $0.5$ 
-\begin{equation}
-  \gamma(0.5) = -\frac{\ln(0.5)}{\alpha_\mathrm{MM}},
+First, we solve Eq.~(\ref{eq:inverse_function}) for the MM default value $\alpha_\mathrm{MM}^k$ for each PFT or species $k$ at $0.5$ 
+\begin{equation}
+  \gamma^k(0.5) = -\frac{\ln(0.5)}{\alpha_\mathrm{MM}^k},
   \label{eq:inverse_function_halfway}
 \end{equation}
-and define a variation $\gamma' = \gamma \cdot \epsilon$ with $\epsilon \in \{0.8, 1.2\}$.
-We resolve Eq.~(\ref{eq:inverse_function}) again for $\alpha$ and find $\alpha'(\gamma')$.
-The resulting functions are shown in Fig.~\ref{fig:f_temp_grassland}b).
+and define a variation $\gamma' = \gamma \cdot \eta$ with $\eta \in \{0.8, 1.2\}$.
+We invert Eq.~(\ref{eq:inverse_function}) for $\alpha$ and find $\alpha'(\gamma')$.
+The resulting functions for perennial grassland are shown in Fig.~\ref{fig:f_temp_grassland}b) (for downy birch and Norway spruce refer Appendix~Figs.~\ref{fig:f_temp_spruce}b)--\ref{fig:f_temp_birch}b)).
 All derived parameters are tabulated in Table~\ref{tab:sensitivity_tests_light}.
@@
     \middlehline
-    \multirow{3}{*}{Spruce} & MM & 0 & 20 & \multirow{3}{*}{100}\\
+    \multirow{3}{*}{Downy birch} & MM & 5 & 20 & \multirow{3}{*}{100}\\
+    & Boreal & 0 & 10\\
+    & Cold & 5 & 15\\
+    \middlehline
+    \multirow{3}{*}{Norway spruce} & MM & 0 & 20 & \multirow{3}{*}{100}\\
     & Boreal & 0 & 10 & \\
@@
     \middlehline
-    \multirow{3}{*}{Birch} & MM & 5 & 20 & \multirow{3}{*}{100}\\
-    & Boreal & 0 & 10\\
-    & Cold & 5 & 15\\
-    \middlehline
-    \multirow{3}{*}{Grassland} & MM & 10 & 24 & 36\\
+    \multirow{3}{*}{Perennial grassland} & MM & 10 & 24 & 36\\
     & Boreal & 0 & 10 & 24 \\
@@
     \middlehline
-    \multirow{3}{*}{Spruce} & MM & 0.006 \\
+    \multirow{3}{*}{Downy birch} & MM & 0.004 \\
+    & PPFD0.8 & 0.005\\
+    & PPFD1.2 & 0.004\\
+    \middlehline
+    \multirow{3}{*}{Norway spruce} & MM & 0.006 \\
     & PPFD0.8 & 0.008  \\
@@
     \middlehline
-    \multirow{3}{*}{Birch} & MM & 0.004 \\
-    & PPFD0.8 & 0.005\\
-    & PPFD1.2 & 0.004\\
-    \middlehline
-    \multirow{3}{*}{Grassland} & MM & 0.011\\
+    \multirow{3}{*}{Perennial grassland} & MM & 0.011\\
     & PPFD0.8 & 0.014 \\
@@
 
+As indicated above, a better acclimation will result in a higher average $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$.
 To quantify the resulting bespoke parameterizations, we propose the following metric.
-{\bf TODO: Write about Boreal and cold parameter sets and their justification.
-Write about suggested method for optimizing PFTs in a simple but physiological way based on optimizing maximum climatological mean $g_\mathrm{sto}$ and morning responds.
-If plants have a short growing season up north optimizing \chem{CO_2} uptake around noon (highest light intensity) and speed up opening of stomata at lower intensities (morning), additional contained lower climatological variance?
-An idea to justify our choice based on the full set of climatological data for Svanvik is shown in Fig.~\ref{fig:javis_func_opt_mean}.}
+For a generic GS (May--August), we compute climatological average and standart deviation of $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal})) at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time).
+We presume that relative stomatal conductance of around noon (highest light intensity) and in the early morning (lower intensities) are good proxies for \chem{CO_2} uptake efficiency.
+Our arbitrary optimization target is $60\,\unit{\%}$.
+Additional constrains are a small deviation between noon and morning and a low standard deviation indicating a higher robustness to interannual variability of growing conditions.
+For simplicity, we neglect the dependency on SWP in this regard.
+From the results shown in Fig.~\ref{fig:javis_func_opt_mean}, we identify boreal, PPFD0.8 as the best acclimation for all hypothetical species or PFTs alike.
+Overall, Norway spruce displays the least deviation between the different types of parameterizations, while the divergance for perannial grassland is substantial.
 
@@
   \caption{Proposed metric to test bespoke response functions.
-GS (May--August) climatological averages and standart deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time) are shown for (a) birch; (b) Norway spruce; (c) perennial grassland.
+GS (May--August) climatological averages and standart deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time) are shown for (a) Downy birch; (b) Norway spruce; (c) perennial grassland.
 For simplicity, $f_\mathrm{SW}$ is neglected in this calculation.
@@
 
-Start ($A_\text{start}$) and end ($A_\text{end}$) of growing season for each PFT are given in units of day of year (doy) and estimated as follows.
+Another important factor for ozone exposure metrices is the timing and length of the GS.
+Start ($A_\text{start}$) and end ($A_\text{end}$) of GS for each PFT are given in units of \unit{doy} (Tab.~\ref{tab:sensitivity_tests_gs}) and estimated as follows.
 Evergreen needle leaf trees show photosynthetic activity already at low temperatures \citep{TP:Wallin2013, TB:Kolari2007}.
-We use a MODIS (Aqua/Terra) retrieved net photosynthesis product for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine onset and end of growing season.
+We use a MODIS (Aqua/Terra) retrieved net photosynthesis product for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine $A_\text{start}$ and $A_\text{end}$.
 Net photosynthesis data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}).
-By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{doy}$ for 2018 and day 106 in 2019.
-$A_\text{end}$ amounts to day 261 and 274, respectively, and shall be used for all vegetation alike.
+By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{doy}$ for 2018 and \unit{doy} 106 in 2019.
+$A_\text{end}$ amounts to \unit{doy} 261 and 274, respectively, and shall be used for all vegetation alike.
 The resulting growing season for Norway spruce in 2019 was one month longer then in 2018.
-For estimating $A_\text{start}$ of birch tree, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Fig.~\ref{fig:greening_season_change_Svanvik}).
+For estimating $A_\text{start}$ of downy birch, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Fig.~\ref{fig:greening_season_change_Svanvik}).
 We find 129 and 130, respectively.
+For both years, these dates coincident with the first snow-free day at the meterological weather station at Øvre Neiden (Sør-Varanger, NOR) which is relatively close-by.
 From the latitude model typically used for localizing $\mathrm{DO_3SE}$, we get $A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$ indicating an about $2\,\unit{months}$ longer growing season.
-For perennial grassland, we assume a latency period of about $1\,\unit{month}$ after snow melt.
-Data for first snow-free day is taken from the weather station at Øvre Neiden in Sør-Varanger.
-In both years, the first snow-free day incidents with the estimated date of bud burst.
-As leafs of birches in northern Fennoscandia \emph{Betula pubescens} are smaller than southern specimen, we have measured their dimension for the purpose of this study and find $(3.0\pm 0.5)\,\unit{cm}$.
+Due to a lack of quantitative field observation, we assume a latency period of about $1\,\unit{month}$ after snow-melt for perennial grassland.
+
+\begin{table}[t]
+  \caption{Bespoke start and end of growing season in units of \unit{doy}. MM refers to mapping manual \citep{GCB:Mills2010,ICP:MappingManual2017}.}
+  \label{tab:sensitivity_tests_gs}
+  \begin{tabular}{llcc}
+    \tophline
+    Species & Year & $A_\mathrm{start}$ & $A_\mathrm{end}$\\
+    \middlehline
+    \multirow{2}{*}{Downy birch} & 2018 & 129$^*$ & 261$^\dagger$ \\
+    & 2019 & 130$^*$ & 274$^\dagger$ \\
+    \multirow{2}{*}{Norway spruce} & 2018 & 122$^\dagger$ & 261$^\dagger$ \\
+    & 2019 & 106$^\dagger$ & 274$^\dagger$ \\
+    \multirow{2}{*}{Perennial grassland} & 2018 & 159$^\ddagger$ & 261$^\dagger$\\
+    & 2019 & 161$^\ddagger$ & 274$^\dagger$ \\
+    \bottomhline
+  \end{tabular}
+  \belowtable{$^\dagger$ MODIS (Aqua/Terra) net photosynthesis product;\\
+    $^*$ $5\,\unit{days}-5\,\unit{^\circ C}$-rule, $T_\mathrm{air}$ from \href{seNorge.no}{seNorge.no};\\
+    $^\ddagger$ One month after snow melt; reference station Øvre Neiden (Sør-Varanger, NOR)} % Table Footnotes
+\end{table}
+
+As pointed out by \citet{ACP:Bueker2012}, a very important factor in deriving a proper $\mathrm{POD_y}$ is the soil texture or how much water the soil can hold.
+As Svanhovd is located in the bed of the Pasvik river, we assume a sandy loam texture at $60\,\unit{cm}$ depth.
+Further, leaves of downy birch in northern Fennoscandia are smaller than southern specimens'.
+Here, we have measured their dimension and found $(3.0\pm 0.5)\,\unit{cm}$.
 \citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area.
 
-As pointed out by \citet{ACP:Bueker2012}, a very important factor in deriving a proper $\mathrm{POD_y}$ is the soil texture or how much water the soil can hold.
-As Svanhovd is located in the bed of the Pasvik river, we assume a sandy loam texture at $60\,\unit{cm}$ depth.\\
-
-\subsection{POD and ozone impact}
-
-We have modeled $\mathrm{POD_y}$ for three natural/semi-natural vegetation PFTs, birch, Norway spruce, and perennial grassland.
-The results are shown comprehensively in Fig.~\ref{fig:pody} as a function of doy for March--October.
-A flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) has been chosen.
-\chem{[O_3]} are plotted on the same axis and scales as $G_\text{sto}^\text{leaf}$ but in units of $\unit{ppb}$.
-
-As expected, $\mathrm{POD_y}$ is higher in 2018 than in 2019 for all PFTs.
-In the end of the 2018 season, Birch took up $13.1\,\unit{mmol\,m^{-2}\,PLA}$, while perennial grassland had an uptake of $4.5\,\unit{mmol\,m^{-2}\,PLA}$.
-The largest interannual variation is also found in Birch ($2.1\,\unit{mmol\,m^{-2}\,PLA}$) and perennial grassland ($2.6\,\unit{mmol\,m^{-2}\,PLA}$).
-Whereas, there is only a small difference for Norway spruce ($0.3\,\unit{mmol\,m^{-2}\,PLA}$) between the two years.
-
-%Except for perennial grassland, model integrations with climatological forcings show higher $\mathrm{POD_y}$ than 2018.
-This can be explained by means of non linearity in modeled stomatal conductance (Eq.~(\ref{eq:stomatal})) in Eq.~(\ref{eq:pod}).
-The climatology in $\mathrm{POD_y}$ therefore represents rather an ideal year for growth then a true climatology.
-To establish a true climatology, it would be necessary to run the $\mathrm{DO_3SE}$ model for 10--20 years with respective observed phenology as well as meteorological and ozone forcing.
-
-%Based on this interpretation of climatology in $\mathrm{POD_y}$, we can learn more about potential ozone damage.
-The time derivative $\frac{\mathrm{\Delta POD_y}}{\Delta t}$ displays very different features for each PFT and meteorological forcing.
-Although Norway spruce shows a low interannual variation, $\frac{\mathrm{\Delta POD_y}}{\Delta t}$ (Fig.~\ref{fig:pody_rel}) indicates an extended period of higher than average ozone uptake in spring 2018 (steep slope), while ozone uptake in June/July was anomalously low in 2019 (shallow slope) but high in fall.
-This is most prominently visible in Fig.~\ref{fig:pody}h) for perennial grassland, where consecutive periods of low $G_\text{sto}^\text{leaf}$ manifest in three plateaus in $\mathrm{POD_y}$ throughout the season (1: $184-198\,\unit{doy}$, 2: $209-228\,\unit{doy}$, 3: $235-240\,\unit{doy}$).
-An episode with anomalous high ozone in mid June ($194\,\unit{doy}$) can be identified but does not amount to unusually high ozone uptake on an annual basis.
+\subsection{$\mathrm{POD_1}$ results and implications on ozone impact}
+\label{subsec:do3se_results}
+
+We have modeled $\mathrm{POD_y}$ (flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA)) for three natural/semi-natural vegetation types, downy birch, Norway spruce, and perennial grassland using both MM parameterizations and a variaty of bespoke parameterizations in $f_\mathrm{T}$, $f_\mathrm{light}$, and GS (described in detail in Section~\ref{subsec:do3se_parameters}).
+The results are shown comprehensively in Fig.~\ref{fig:pody_rel}.
+For comparison, critical levels (CLs) for a $4\,\unit{\%}$ ($2\,\unit{\%}$ for Norway spruce) reduction in biomass in accordance to \citet{ICP:MappingManual2017} are shown as horizontal dashed lines.
+Different years are color coded.
+Lighter hue indicates simulations with MM GS and MM canopy and leaf parameters.
+Annotated $\pm$ refers to bespoke light response functions, while bespoke temperature response functions are indicated by categories on the x-axis.
+Closed/open symbols represent SWP sensitivity switched off/on in the model.
+
+For downy birch (Fig.~\ref{fig:pody_rel}a), all simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and display the largest spread among the species ($16\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 for boreal type).
+As expected, SWP reduces ozone uptake in 2018 mainly.
+The uncertainty induced by the light sensitivity threshold ranges between $(1.5..2.5)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $(-2.2..-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
+With MM GS, 2018 displays increasingly larger ozone uptake than 2019 with more acclimated paramters.
+However accounting for the shorter and later GS in 2018, we find a reversal of this relationship in the boreal type.
+We deduce an overall systematic uncertainty of $xx$.
+
+For Norway spruce (Fig.~\ref{fig:pody_rel}b), most simulations exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
+Norway spruce displays an overall lower spread ($10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 for boreal type) than downy birch, but appears more sensitive to GS and interannual veriability in conditions (boreal GS MM compared to GS bespoke).
+The sensitivity to the light threshold is also smaller compared to downy birch.
+Accounting for the shorter and later GS in 2018, we find again a reversal between predicted ozone damage in 2018 and 2019 for the boreal type.
+We deduce an overall systematic uncertainty of $xx$.
+
+For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations with bespoke GS stay below the CL.
+CL is only exceeded for boreal types and cold type in 2018.
+We find the same reveral between 2018 and 2019 as for the other species in the boreal type.
+Perannial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant for the MM GS.
+The spread in 2018 for boreal type is of the order of $10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
+We deduce an overall systematic uncertainty of $xx$.
+
+These results allow for an assessment of a large variety of systematic uncertainties in the exposure-based risk metric for ozone induced damage in subarctic vegetation.
+In summary, we identify the following main effects on the model predicted ozone uptake:
+\begin{itemize}
+\item Optimal acclimation of species or PFT $\rightarrow$ increased uptake in any year,
+\item Identification of proper timing of the growing season $\rightarrow$ reduced uptake in any year,
+\item Drought conditions $\rightarrow$ reduced uptake,
+\item Early onset of growing season $\rightarrow$ increased uptake.
+\end{itemize}
+The exact order of magnitude of these effects varies with species or PFT and from year to year.
+The systematic uncertainty induced by our proposed temperature acclimation is, in general, larger than for a $\pm 20\,\unit{\%}$ variation in light response threshold and can likely exceed $xx\,\unit{\%}$ for some years and species.
+This would lead to a huge underestimation of induced ozone damage.
+At the same time, the magnitude of light response uncertainty is increasing with acclimation to colder temperature.
+The impact of the timing of the GS can be at the same oder of magnitude as the acclimation of response function but points into the oposite direction.
+Especially, the timing of $A_\mathrm{start}$ with respect to the ozone spring peak (compare 2018 GS MM and GS bespoke) induces large uncertainties which are increasing if an acclimation to colder climate is anticipated.
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{fig11}
+  \caption{$\mathrm{DO_3SE}$ modeling results for $\mathrm{POD_1}$ with default parameterization (MM) and bespoke parameters for Svanhovd.
+Bespoke parameterizations include temperature response functions for boreal, cold and light thresholds (PPFD0.8, PPFD1.2) at the end of a MM GS (lighter hue and dotted lines) and the estimated GS for Svanhoved (2018, 2019).
+(a) Downy birch; (b) Norway spruce; (c) Perennial grassland.
+The dashed horizontal lines indicate the critical levels defined in \citet{ICP:MappingManual2017} for each species, respectively.}
+  \label{fig:pody_rel}
+\end{figure}
 
@@
 Year & \multicolumn{4}{c}{PFT}\\
-& Birch & Norway spruce & \multicolumn{2}{c}{Perennial grassland}\\
+& Downy birch & Norway spruce & \multicolumn{2}{c}{Perennial grassland}\\
 \middlehline
@@
 \end{table}
-
-\begin{figure}[t]
-  %\includegraphics[width=8.3cm]{do3se_results_pod_v3_Birch}\\
-  %\includegraphics[width=8.3cm]{do3se_results_pod_v3_Spruce}\\
-  %\includegraphics[width=8.3cm]{do3se_results_pod_v3_Grassland}
-  \includegraphics[width=8.3cm]{fig11}
-  \caption{Comprehensive comparison of $\mathrm{DO_3SE}$ modeling results with default parameterization from mapping manual (MM)and bespoke parameters for Svanhovd.
-$\mathrm{POD_1}$ is shown for the different temperature response functions (MM, boreal, cold) and light thresholds (MM, PPFD0.8, PPFD1.2) at the end of a MM GS (lighter hue and dotted lines) and a derived bespoke GS.
-(a) Birch; (b) Norway spruce; (c) Perennial grassland.}
-  \label{fig:pody_rel}
-\end{figure}
 
@@
 Daily averaged data for both 2018 and 2019 has been fitted with a quadratic polynomial function.
-The numerically computed root yields: BGS 122/106 day of year (doy) and EGS 261/274 day of year for 2018/2019, respectively.}
+The numerically computed root yields: BGS 122/106 day of year (\unit{doy}) and EGS 261/274 day of year for 2018/2019, respectively.}
   \label{fig:modis_Psn}
@@
   \includegraphics[width=12cm]{figB4}
-  \caption{$\mathrm{POD_y}$ modeling results from $\mathrm{DO_3SE}$ model for three different PFTs over day of year (doy), March--October.
+  \caption{$\mathrm{POD_y}$ modeling results from $\mathrm{DO_3SE}$ model for three different PFTs over day of year (\unit{doy}), March--October.
 A flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) has been chosen.

### 2021-03-17 supplements/VegOzone_supplement.tex
Update
Add reference to senorge~

--- 2021-03-03 supplements/VegOzone_supplement.tex
@@
-\begin{figure*}[t]
+\graphicspath{{pictures/}}          % graphics
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
+
+\clearpage
+\setcounter{page}{1}
+\setcounter{figure}{0}
+
+\begin{figure}[t]
+  \includegraphics[width=12cm]{test_fennoscandia}
+  \caption{Maximum daily sun shine duration based on geometrical calculations at different sites in Fennoscandia.
+Midnight sun conditions at Jergul/Karasjok and Svanvik prevail from the end of May until the end of July, while at Esrange and Pallas they only prevail from the beginning of June until mid of July.}
+  \label{fig:sunlight_fennoscandia}
+\end{figure}
+
+\begin{figure}[t]
   \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs}
-  \caption{S1 Multi-annual mean of daily mean ozone at observation sites in Fennoscandia.
+  \caption{Multi-annual mean of daily mean ozone at observation sites in Fennoscandia.
 The large spread in case of Svanvik and Janiskoski is due to the lower statistics.
@@
   \label{fig:ozone_climatology_fenoscandic_obs}
-\end{figure*}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{test_fennoscandia}
-  \caption{Maximum daily sun shine duration based on geometrical calculations at different sites in Fennoscandia.
-Midnight sun conditions at Jergul/Karasjok and Svanvik prevail from the end of May until the end of July, while at Esrange and Pallas they only prevail from the beginning of June until mid of July.}
-  \label{fig:sunlight_fennoscandia}
-\end{figure*}
+\end{figure}
 
@@
   \includegraphics[width=12cm]{greening_season_change_Svanvik}
-  \caption{Estimated shift and prolongation of greening season at Svanhovd over the past 6 decades based on data from SeNorge.no {\bf TODO: citation http://www.senorge.no}.}
+  \caption{Estimated shift and prolongation of greening season at Svanhovd over the past 6 decades based on data from \citet{SeNorge}.}
   \label{fig:greening_season_change_Svanvik}

### 2021-03-17 VegOzone.tex
Change reference to bibtex file
Write about results

@@
 
-We have modeled $\mathrm{POD_y}$ (flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA)) for three natural/semi-natural vegetation types, downy birch, Norway spruce, and perennial grassland using both MM parameterizations and a variaty of bespoke parameterizations in $f_\mathrm{T}$, $f_\mathrm{light}$, and GS (described in detail in Section~\ref{subsec:do3se_parameters}).
+We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for three natural/semi-natural vegetation types, downy birch, Norway spruce, and perennial grassland using both MM parameterizations and a variaty of bespoke response function in $f_\mathrm{T}$ and $f_\mathrm{light}$, and localized GS (described in detail in Section~\ref{subsec:do3se_parameters}).
 The results are shown comprehensively in Fig.~\ref{fig:pody_rel}.
-For comparison, critical levels (CLs) for a $4\,\unit{\%}$ ($2\,\unit{\%}$ for Norway spruce) reduction in biomass in accordance to \citet{ICP:MappingManual2017} are shown as horizontal dashed lines.
-Different years are color coded.
-Lighter hue indicates simulations with MM GS and MM canopy and leaf parameters.
-Annotated $\pm$ refers to bespoke light response functions, while bespoke temperature response functions are indicated by categories on the x-axis.
+For comparison, critical levels for a $4\,\unit{\%}$ (silver birch), $2\,\unit{\%}$ (Norway spruce), and $10\,\unit{\%}$ (perennial grassland, above ground) reduction in biomass \citep{ICP:MappingManual2017,ESPR:Hayes2021} are shown as horizontal dashed lines.
+Different years are color coded as previously.
+Simulations with MM GS, canopy, and leaf parameters are indicated with squared markers, bespoke GS with circles.
+The annotation $\pm$ refers to bespoke light response function threshold ($+$ increased, $-$ decreased), while bespoke temperature response functions are indicated by categories on the x-axis.
 Closed/open symbols represent SWP sensitivity switched off/on in the model.
 
-For downy birch (Fig.~\ref{fig:pody_rel}a), all simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and display the largest spread among the species ($16\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 for boreal type).
-As expected, SWP reduces ozone uptake in 2018 mainly.
-The uncertainty induced by the light sensitivity threshold ranges between $(1.5..2.5)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $(-2.2..-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
-With MM GS, 2018 displays increasingly larger ozone uptake than 2019 with more acclimated paramters.
-However accounting for the shorter and later GS in 2018, we find a reversal of this relationship in the boreal type.
-We deduce an overall systematic uncertainty of $xx$.
-
-For Norway spruce (Fig.~\ref{fig:pody_rel}b), most simulations exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
-Norway spruce displays an overall lower spread ($10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 for boreal type) than downy birch, but appears more sensitive to GS and interannual veriability in conditions (boreal GS MM compared to GS bespoke).
-The sensitivity to the light threshold is also smaller compared to downy birch.
-Accounting for the shorter and later GS in 2018, we find again a reversal between predicted ozone damage in 2018 and 2019 for the boreal type.
-We deduce an overall systematic uncertainty of $xx$.
+The results depicted in Fig.~\ref{fig:pody_rel} allow for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric for ozone induced damage in subarctic vegetation.
+In summary, we identify the following main effects on the model predicted ozone uptake with respect to MM default parameterization:
+\begin{itemize}
+  \itemsep0pt
+\item Increasing temperature acclimation $\rightarrow$ increased uptake regardless of individual year,
+\item usage of proper growing season $\rightarrow$ reduced uptake with respect to the above,
+\item an earlier assumed onset of growing season $\rightarrow$ increased uptake,
+\item drought conditions (SWP) $\rightarrow$ reduced uptake,
+\item lower/higher light threshold $\rightarrow$ incresed/reduced uptake,
+\item bespoke light threshold $\rightarrow$ asymmetric response in POD,
+\item temperature acclimation $\rightarrow$ amplified effect of light threshold and drought.
+\end{itemize}
+The exact order of magnitude of these effects varies for individual species or PFT and between years, but the thus predicted ozone uptake is in any case larger than for MM default parameterizations.
+This indicates a considerable underestimation of ozone risk for subarctic species.
+Drought effects (SWP) do only matter if a bespoke GS is not taken into account.
+
+In more detail for downy birch (Fig.~\ref{fig:pody_rel}a) all simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
+Downy birch displays the largest range of ozone uptake among all species or PFTs ($10-31\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
+In particular, the sensitivity to the light threshold is especially pronounced, e.g.
+$\Delta\mathrm{POD_1} = ^{+2.5}_{-3.6}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of boreal type without bespoke GS in 2019.
+Without bespoke GS 2018 stands out as a higher ozone damage risk year compared to 2019.
+However, taking the later onset and overall shorter GS in 2018 into account, this is reversed for the boreal acclimation type.
+SWP induced reduction in ozone uptake is only effectiv in 2018 and without bespoke GS.
+The overall bias deduced from bespoke GS for boreal and cold with respect to MM ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to light acclimation of $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
+
+For Norway spruce (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
+Norway spruce is the least sensitive to acclimation in temperature and therefor the least biased.
+Norway spruce displayes a medium sensitivity to the light threshold, e.g.
+$\Delta\mathrm{POD_1} = ^{+1.7}_{-1.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of boreal type without bespoke GS in 2018.
+Taking the bespoke GS into account, we find a higher ozone uptake in 2019 for all bespoke temperature response functions.
+The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to light acclimation of $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
 
 For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations with bespoke GS stay below the CL.
-CL is only exceeded for boreal types and cold type in 2018.
-We find the same reveral between 2018 and 2019 as for the other species in the boreal type.
-Perannial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant for the MM GS.
-The spread in 2018 for boreal type is of the order of $10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
-We deduce an overall systematic uncertainty of $xx$.
-
-These results allow for an assessment of a large variety of systematic uncertainties in the exposure-based risk metric for ozone induced damage in subarctic vegetation.
-In summary, we identify the following main effects on the model predicted ozone uptake:
-\begin{itemize}
-\item Optimal acclimation of species or PFT $\rightarrow$ increased uptake in any year,
-\item Identification of proper timing of the growing season $\rightarrow$ reduced uptake in any year,
-\item Drought conditions $\rightarrow$ reduced uptake,
-\item Early onset of growing season $\rightarrow$ increased uptake.
-\end{itemize}
-The exact order of magnitude of these effects varies with species or PFT and from year to year.
-The systematic uncertainty induced by our proposed temperature acclimation is, in general, larger than for a $\pm 20\,\unit{\%}$ variation in light response threshold and can likely exceed $xx\,\unit{\%}$ for some years and species.
-This would lead to a huge underestimation of induced ozone damage.
-At the same time, the magnitude of light response uncertainty is increasing with acclimation to colder temperature.
-The impact of the timing of the GS can be at the same oder of magnitude as the acclimation of response function but points into the oposite direction.
-Especially, the timing of $A_\mathrm{start}$ with respect to the ozone spring peak (compare 2018 GS MM and GS bespoke) induces large uncertainties which are increasing if an acclimation to colder climate is anticipated.
+Perennial grassland shows the smallest ozone uptake, but a similarily large bias as downy birch.
+We also find a reversal for predicted ozone damage risk in $\mathrm{POD_1}$ between 2018 and 2019 for the boreal type.
+Perannial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant without bespoke GS.
+The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to light acclimation of $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
 
@@
 (a) Downy birch; (b) Norway spruce; (c) Perennial grassland.
-The dashed horizontal lines indicate the critical levels defined in \citet{ICP:MappingManual2017} for each species, respectively.}
+The dashed horizontal lines indicate critical levels defined in \citet{ICP:MappingManual2017,ESPR:Hayes2021} for each species, respectively.}
   \label{fig:pody_rel}
@@
 
+---
+The results depicted in Fig.~\ref{fig:pody_rel} allow for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric for ozone induced damage in subarctic vegetation.
+In summary, we identify the following main effects on the model predicted ozone uptake with respect to MM default parameterization:
+\begin{itemize}
+\item Increasing temperature acclimation $\rightarrow$ increased uptake regardless of individual year,
+\item Usage of proper growing season $\rightarrow$ reduced uptake with respect to the above,
+\item An earlier assumed onset of growing season $\rightarrow$ increased uptake,
+\item Drought conditions (SWP) $\rightarrow$ reduced uptake,
+\item Lower/higher light threshold $\rightarrow$ incresed/reduced uptake.
+\end{itemize}
+The exact order of magnitude of these effects varies for individual species or PFT and between years, but the thus predicted ozone uptake is in any case larger than for MM default parameterizations.
+This indicates a considerable underestimation of ozone risk for subarctic species.
+ The impact of the timing of the GS can be at the same oder of magnitude as the acclimation of response function but points into the oposite direction.
+Especially, the timing of $A_\mathrm{start}$ with respect to the ozone spring peak induces large uncertainties in anticipation of an acclimation to colder climates.
+
 Bigger picture:
@@
 Manninen et al.
-(2009a) summarized the results from open-top chamber (OTC) experiments performed in northern Finland where Scots pine and northern birch provenances were exposed to 36 – 54 ppb O3 as mean daylight concentrations.
+(2009a) summarized the results from open-top chamber (OTC) experiments performed in northern Finland where Scots pine and downy birch provenances were exposed to 36 – 54 ppb O3 as mean daylight concentrations.
 Biomass reductions of both downy birch (Betula pubescens) and mountain birch seedlings were demonstrated after one-season exposure.
@@
 Svanvik displays highest correlation with Pallas.
+Note: The stripe at $26\,\unit{ppb}$ is an artifact from plotting.
 (a) Jergul/Karasjok--Esrange; (b) Jergul/Karasjok--Pallas; (c) Svanvik--Esrange; (d) Svanvik--Pallas.}
@@
   \includegraphics[width=12cm]{figB3}
-\caption{Construction of bespoke response functions for Norway birch.
+\caption{Construction of bespoke response functions for downy birch.
 (a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively.
@@
   \includegraphics[width=12cm]{figB4}
-  \caption{$\mathrm{POD_y}$ modeling results from $\mathrm{DO_3SE}$ model for three different PFTs over day of year (\unit{doy}), March--October.
+  \caption{$\mathrm{DO_3SE}$ modeling results for mapping manual default parameterazation.
+$\mathrm{POD_y}$ is shown over day of year (\unit{doy}), March--October.
 A flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) has been chosen.
 \chem{[O_3]} are plotted on the same axis and scales as $G_\text{sto}^\text{leaf}$ but in units of $\unit{ppb}$.
-(a, b) Birch; (c, d) Norway spruce; (e, f) perennial grassland.
-From left to right: 2018, 2019.
-{\bf TODO: substitute with figure for Boreal PFTs?}}
+(a, b) Downy birch; (c, d) Norway spruce; (e, f) perennial grassland.
+From left to right: 2018, 2019.}
   \label{fig:pody_mm_composit}
@@
 \bibliographystyle{copernicus}
-\bibliography{VegOzone.bib}
+\bibliography{atmo.bib}
 %%

### 2021-03-17 VegOzone.tex
Added biomass reduction

@@
 
-The estimated biomass reduction in \unit{\%} due to \chem{O_3} following \citet{ICP:MappingManual2017} is shown in Table~\ref{tab:biomass_reduction}.
+The estimated biomass reduction in \unit{\%} due to \chem{O_3} for bespoke type boreal in accordance to \citet{ICP:MappingManual2017, ESPR:Hayes2021} is shown in Table~\ref{tab:biomass_reduction}.
+The estimated uncertainies correspond to a $\pm 20\,\unit{\%}$ variation in light threshold and SWP=on.
 {\bf TODO: Lisa?
@@
 \begin{table}[t]
-  \caption{Estimated total biomass reduction in \unit{\%} due to \chem{O_3} following \citet{ICP:MappingManual2017}.}
+  \caption{Estimated total biomass reduction in \unit{\%} due to \chem{O_3} for bespoke type boreal. Uncertainies correspond to a $\pm 20\,\unit{\%}$ variation in light threshold and SWP=on.}
   \label{tab:biomass_reduction}
-\begin{tabular}{lccrr}
+\begin{tabular}{lllll}
 \tophline
@@
 \middlehline
-2018 & 11.99 & 2.34 & 8.06 & 10.51$^*$\\
-2019 & 10.07 & 2.26 & 6.43 & 7.90$^*$\\
-clim. & 13.81 & 2.58 & 6.67 & 8.29$^*$\\
+2018 & $15.5^{+(1.9...2.1)}_{-(0.8...1.5)}$ & $2.5^{+(0.2)}_{-(0.2)}$ & $9.7^{+(0.2)}_{-(0.2)}$ & $^*13.1^{+(0.3)}_{-(0.3)}$\\
+2019 & $17.4^{+(2.5)}_{-(1.8)}$ & $3.0^{+(0.2)}_{-(0.3)}$ & $10.5^{+(0.1...0.2)}_{-(0.0...0.1)}$ & $^*14.5^{+(0.2...0.3)}_{-(0.1...0.3)}$\\
 \bottomhline

### 2021-03-17 VegOzone.tex
First draft of dos3e results

--- 2021-03-17 VegOzone.tex
@@
 In this study, conditions at Svanhovd for the growing seasons 2018/19 shall serve as reference.
-It is important to understand the implications of using these more generic parameterizations for more extreme climate regions, here we explore how risk estimates might be over or underestimated using generic vs bespoke subarctic parameterization.
+It is important to understand the implications of using these more generic parameterization for more extreme climate regions, here we explore how risk estimates might be over or underestimated using generic vs bespoke subarctic parameterization.
 we also explore the role \chem{[O_3]} variability and changing climate variables to understand how this risk might alter in the future with the advent of climate change.
-This allows both an understanding of how the threat from ozone may change over time but also which are the key aspects of plant physiology that might determine risk (and hence which parameters might require special attention to get right) both for pollution impact assessment, but also more generally for describing gas exchange response to changing climates that will influence bio-geochemcial cycling.
+This allows both an understanding of how the threat from ozone may change over time but also which are the key aspects of plant physiology that might determine risk (and hence which parameters might require special attention to get right) both for pollution impact assessment, but also more generally for describing gas exchange response to changing climates that will influence bio-chemical cycling.
 This might be one key to solve long-standing issues of coupled earth system models (ESM) which generally lack PFTs representing subarctic conditions in their land surface model.\\
@@
   \includegraphics[width=12cm]{fig08}
-  \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as percentages of days deviating from climatological mean by means of standart deviation for each month.
+  \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as percentages of days deviating from climatological mean by means of standard deviation for each month.
 Negative deviations from climatology are shown as negative percentage.
 The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The hatched area between the dashed lines indicates statistical significance under the assumption of a normal distribution ($15.9\,\unit{\%}$, $30.8\,\unit{\%}$ for precip.).
+The hatched area between the dashed lines indicates statistical significance under the assumption of a normal distribution ($15.9\,\unit{\%}$, $30.8\,\unit{\%}$ for Precip.).
 (a) Temperature, days warmer/colder than $\pm 1\,\sigma$; (b) precipitation, days wetter/drier than $\pm \frac{1}{2}\,\sigma$; (c) global irradiance, days lighter/darker than $\pm 1\,\sigma$; (d) ozone, days higher/lower than $\pm 1\,\sigma$.
@@
 \label{sec:do3se}
-To investigate which of the environmental factors lead to the observed ozone damage on clover and tobacco in 2018 in contrast to 2019, when no such damage was found on the clovers but only on the sensitive tobacco cultivar, we model ozone uptake of different natural and semi-natural vegetation using the $\mathrm{DO_3SE}$ model and develop bespoke subarctic prameterizations for common plant species.
+To investigate which of the environmental factors lead to the observed ozone damage on clover and tobacco in 2018 in contrast to 2019, when no such damage was found on the clovers but only on the sensitive tobacco cultivar, we model ozone uptake of different natural and semi-natural vegetation using the $\mathrm{DO_3SE}$ model and develop bespoke subarctic parameterizations for common plant species.
 We first give an account of the $\mathrm{DO_3SE}$ model, input data, and the used method for deriving the bespoke parameters.
@@
 We aim to increase the enclosed area between probability density function (PDF) and the respective response function.
-We show two climatological reference periods 1990s and 2000s to indicate that the vegation is subject to ongoing climate change and acclimations may be changing.
-Resulting reponse functions for Norway spruce and downy birch are displayed in Appendix~Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}.
-
-We define two different hypothetical temperature acclimations: \emph{Boreal} and \emph{cold}.
+We show two climatological reference periods 1990s and 2000s to indicate that the vegetation is subject to ongoing climate change and acclimation may be changing.
+Resulting response functions for Norway spruce and downy birch are displayed in Appendix~Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}.
+
+We define two different hypothetical temperature acclimation: \emph{Boreal} and \emph{cold}.
 For either, we assume that the hypothetical species are more tolerant to cold conditions and shift $f_\mathrm{T}$ towards colder temperatures accordingly.
@@
 In the same way, \emph{boreal} refers to a hypothetical species which is very tolerant to cold but is sensitive to high temperatures, and most efficient at cool temperatures ($T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000).
-For perennial grassland the resulting response functions are depicted together with the PDF of temperatur in Fig.~\ref{fig:f_temp_grassland}a).
+For perennial grassland the resulting response functions are depicted together with the PDF of temperature in Fig.~\ref{fig:f_temp_grassland}a).
 
@@
 
-For an acclimation with respect to light conditions, we hypothize that the light threshold for a $50\,\unit{\%}$ opening of plants' stomata may differ in subarctic species.
+For an acclimation with respect to light conditions, we hypothesize that the light threshold for a $50\,\unit{\%}$ opening of plants' stomata may differ in subarctic species.
 We analytically derive the inverse function $f_{\mathrm{light},k}^{-1}$ for each PFT or species $k$
@@
 To quantify the resulting bespoke parameterizations, we propose the following metric.
-For a generic GS (May--August), we compute climatological average and standart deviation of $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal})) at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time).
+For a generic GS (May--August), we compute climatological average and standard deviation of $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal})) at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time).
 We presume that relative stomatal conductance of around noon (highest light intensity) and in the early morning (lower intensities) are good proxies for \chem{CO_2} uptake efficiency.
@@
 From the results shown in Fig.~\ref{fig:javis_func_opt_mean}, we identify boreal, PPFD0.8 as the best acclimation for all hypothetical species or PFTs alike.
-Overall, Norway spruce displays the least deviation between the different types of parameterizations, while the divergance for perannial grassland is substantial.
+Overall, Norway spruce displays the least deviation between the different types of parameterizations, while the divergence for perennial grassland is substantial.
 
@@
   \caption{Proposed metric to test bespoke response functions.
-GS (May--August) climatological averages and standart deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time) are shown for (a) Downy birch; (b) Norway spruce; (c) perennial grassland.
+GS (May--August) climatological averages and standard deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time) are shown for (a) Downy birch; (b) Norway spruce; (c) perennial grassland.
 For simplicity, $f_\mathrm{SW}$ is neglected in this calculation.
@@
 
-Another important factor for ozone exposure metrices is the timing and length of the GS.
+Another important factor for ozone exposure metrics is the timing and length of the GS.
 Start ($A_\text{start}$) and end ($A_\text{end}$) of GS for each PFT are given in units of \unit{doy} (Tab.~\ref{tab:sensitivity_tests_gs}) and estimated as follows.
@@
 We find 129 and 130, respectively.
-For both years, these dates coincident with the first snow-free day at the meterological weather station at Øvre Neiden (Sør-Varanger, NOR) which is relatively close-by.
+For both years, these dates coincident with the first snow-free day at the meteorological weather station at Øvre Neiden (Sør-Varanger, NOR) which is relatively close-by.
 From the latitude model typically used for localizing $\mathrm{DO_3SE}$, we get $A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$ indicating an about $2\,\unit{months}$ longer growing season.
@@
 
-We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for three natural/semi-natural vegetation types, downy birch, Norway spruce, and perennial grassland using both MM parameterizations and a variaty of bespoke response function in $f_\mathrm{T}$ and $f_\mathrm{light}$, and localized GS (described in detail in Section~\ref{subsec:do3se_parameters}).
+We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for three natural/semi-natural vegetation types, downy birch, Norway spruce, and perennial grassland using both MM parameterizations and a variety of bespoke response function in $f_\mathrm{T}$ and $f_\mathrm{light}$, and localized GS (described in detail in Section~\ref{subsec:do3se_parameters}).
 The results are shown comprehensively in Fig.~\ref{fig:pody_rel}.
@@
   \itemsep0pt
-\item Increasing temperature acclimation $\rightarrow$ increased uptake regardless of individual year,
-\item usage of proper growing season $\rightarrow$ reduced uptake with respect to the above,
-\item an earlier assumed onset of growing season $\rightarrow$ increased uptake,
-\item drought conditions (SWP) $\rightarrow$ reduced uptake,
-\item lower/higher light threshold $\rightarrow$ incresed/reduced uptake,
-\item bespoke light threshold $\rightarrow$ asymmetric response in POD,
-\item temperature acclimation $\rightarrow$ amplified effect of light threshold and drought.
+\item Increasing temperature acclimation $\Rightarrow$ increased uptake regardless of individual year,
+\item usage of proper growing season $\Rightarrow$ reduced uptake with respect to the above,
+\item an earlier assumed onset of growing season $\Rightarrow$ increased uptake,
+\item drought conditions (SWP) $\Rightarrow$ reduced uptake,
+\item lower/higher light threshold $\Rightarrow$ increased/reduced uptake,
+\item bespoke light threshold $\Rightarrow$ asymmetric response in POD,
+\item temperature acclimation $\Rightarrow$ amplified effect of light threshold and drought.
 \end{itemize}
@@
 However, taking the later onset and overall shorter GS in 2018 into account, this is reversed for the boreal acclimation type.
-SWP induced reduction in ozone uptake is only effectiv in 2018 and without bespoke GS.
+SWP induced reduction in ozone uptake is only effective in 2018 and without bespoke GS.
 The overall bias deduced from bespoke GS for boreal and cold with respect to MM ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to light acclimation of $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
@@
 Norway spruce is the least sensitive to acclimation in temperature and therefor the least biased.
-Norway spruce displayes a medium sensitivity to the light threshold, e.g.
+Norway spruce displays a medium sensitivity to the light threshold, e.g.
 $\Delta\mathrm{POD_1} = ^{+1.7}_{-1.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of boreal type without bespoke GS in 2018.
@@
 For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations with bespoke GS stay below the CL.
-Perennial grassland shows the smallest ozone uptake, but a similarily large bias as downy birch.
+Perennial grassland shows the smallest ozone uptake, but a similarly large bias as downy birch.
 We also find a reversal for predicted ozone damage risk in $\mathrm{POD_1}$ between 2018 and 2019 for the boreal type.
-Perannial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant without bespoke GS.
+Perennial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant without bespoke GS.
 The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to light acclimation of $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
@@
   \caption{$\mathrm{DO_3SE}$ modeling results for $\mathrm{POD_1}$ with default parameterization (MM) and bespoke parameters for Svanhovd.
-Bespoke parameterizations include temperature response functions for boreal, cold and light thresholds (PPFD0.8, PPFD1.2) at the end of a MM GS (lighter hue and dotted lines) and the estimated GS for Svanhoved (2018, 2019).
+Bespoke parameterizations include temperature response functions for boreal, cold and light thresholds (PPFD0.8, PPFD1.2) at the end of a MM GS (lighter hue and dotted lines) and the estimated GS for Svanhovd (2018, 2019).
 (a) Downy birch; (b) Norway spruce; (c) Perennial grassland.
@@
 
-The estimated biomass reduction in \unit{\%} due to \chem{O_3} following \citet{ICP:MappingManual2017} is shown in Table~\ref{tab:biomass_reduction}.
-{\bf TODO: Lisa?
-Add paragraph about biomass reduction to introduction?}
+The estimated biomass reduction in \unit{\%} for the bespoke temperature response function of type boreal in accordance to \citet{ICP:MappingManual2017, ESPR:Hayes2021} is shown in Table~\ref{tab:biomass_reduction}.
+The estimated uncertainty range corresponds to the deviation given by the $\pm 20\,\unit{\%}$ variations in light threshold and SWP=on.
+For downy birch and perennial grassland, substantial reductions in total biomass are predicted for any of the two years.
+Norway spruce is least affected.
+The systematic uncertainties for downy birch are one order of magnitude larger than for Norway spruce and perennial grassland.
+For all species or PFTs the interannual variability is larger than the systematic uncertainties estimated.
+In comparison with average results for the default MM parameterization (with bespoke GS), we find a non negligible (above systematic uncertainties) low-bias for both downy birch and perennial grassland.
+For Norway spruce, the default MM parameterization average lies within the given systematic uncertainty range.
 
 \begin{table}[t]
-  \caption{Estimated total biomass reduction in \unit{\%} due to \chem{O_3} following \citet{ICP:MappingManual2017}.}
+  \caption{Estimated total biomass reduction in \unit{\%} for bespoke temperature response function of type boreal with bespoke GS and SWP=off.
+Uncertainty ranges reported here correspond to deviations by the $\pm 20\,\unit{\%}$ variation in light threshold and SWP=on, respectively.
+For comparison results for MM parameterization are shown with standard deviations over both years and SWP treatments.}
   \label{tab:biomass_reduction}
-\begin{tabular}{lccrr}
+\begin{tabular}{lllll}
 \tophline
@@
 \middlehline
-2018 & 11.99 & 2.34 & 8.06 & 10.51$^*$\\
-2019 & 10.07 & 2.26 & 6.43 & 7.90$^*$\\
-clim. & 13.81 & 2.58 & 6.67 & 8.29$^*$\\
+2018 & $15.5^{+(1.9...2.1)}_{-(0.8...1.5)}$ & $2.5^{+0.2}_{-0.2}$ & $9.7^{+0.2}_{-0.2}$ & $^*12.4^{+0.2}_{-0.2}$\\
+2019 & $17.4^{+2.5}_{-1.8}$ & $3.0^{+0.2}_{-0.3}$ & $10.5^{+(0.1...0.2)}_{-(0.0...0.1)}$ & $^*13.7^{+(0.1...0.3)}_{-(0.1...0.3)}$\\
+\middlehline
+$\left<\mathrm{MM}\right>$ & $11.2\pm 1.1$ & $2.31\pm 0.04$ & $7.5\pm 0.9$ & $^*9.6\pm 1.4$\\
 \bottomhline
@@
 The results depicted in Fig.~\ref{fig:pody_rel} allow for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric for ozone induced damage in subarctic vegetation.
-In summary, we identify the following main effects on the model predicted ozone uptake with respect to MM default parameterization:
+In summary, we identify the following main effects on the model predicted ozone uptake with respect to the MM default parameterization:
 \begin{itemize}
-\item Increasing temperature acclimation $\rightarrow$ increased uptake regardless of individual year,
-\item Usage of proper growing season $\rightarrow$ reduced uptake with respect to the above,
-\item An earlier assumed onset of growing season $\rightarrow$ increased uptake,
-\item Drought conditions (SWP) $\rightarrow$ reduced uptake,
-\item Lower/higher light threshold $\rightarrow$ incresed/reduced uptake.
+\item Increasing temperature acclimation $\Rightarrow$ increased uptake regardless of individual year,
+\item usage of proper growing season $\Rightarrow$ reduced uptake with respect to the above,
+\item an earlier assumed onset of growing season $\Rightarrow$ increased uptake,
+\item drought conditions (SWP) $\Rightarrow$ reduced uptake,
+\item lower/higher light threshold $\Rightarrow$ increased/reduced uptake,
+\item bespoke light threshold $\Rightarrow$ asymmetric response in POD,
+\item temperature acclimation $\Rightarrow$ amplified effect of light threshold and drought.
 \end{itemize}
@@
 This indicates a considerable underestimation of ozone risk for subarctic species.
- The impact of the timing of the GS can be at the same oder of magnitude as the acclimation of response function but points into the oposite direction.
+ The impact of the timing of the GS can be at the same order of magnitude as the acclimation of response function but points into the opposite direction.
 Especially, the timing of $A_\mathrm{start}$ with respect to the ozone spring peak induces large uncertainties in anticipation of an acclimation to colder climates.
@@
   \includegraphics[width=12cm]{figB4}
-  \caption{$\mathrm{DO_3SE}$ modeling results for mapping manual default parameterazation.
+  \caption{$\mathrm{DO_3SE}$ modeling results for mapping manual default parameterization.
 $\mathrm{POD_y}$ is shown over day of year (\unit{doy}), March--October.

### 2021-03-17 VegOzone.tex
Second draft of results

@@
 The results depicted in Fig.~\ref{fig:pody_rel} allow for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric for ozone induced damage in subarctic vegetation.
-In summary, we identify the following main effects on the model predicted ozone uptake with respect to MM default parameterization:
+In summary, we identify the following main effects on the model predicted ozone uptake with respect to the MM default parameterization:
 \begin{itemize}
   \itemsep0pt
-\item Increasing temperature acclimation $\Rightarrow$ increased uptake regardless of individual year,
+\item Increasing cold tolerance/heat intolerance $\Rightarrow$ increased uptake regardless of individual year,
 \item usage of proper growing season $\Rightarrow$ reduced uptake with respect to the above,
@@
 \item bespoke light threshold $\Rightarrow$ asymmetric response in POD,
-\item temperature acclimation $\Rightarrow$ amplified effect of light threshold and drought.
+\item temperature acclimation $\Rightarrow$ amplifying effect on light threshold and drought.
 \end{itemize}
-The exact order of magnitude of these effects varies for individual species or PFT and between years, but the thus predicted ozone uptake is in any case larger than for MM default parameterizations.
+The exact order of magnitude of these effects varies for individual species or PFT and between years, but the predicted ozone uptake for the bespoke parameterization is in any case larger than for the MM default parameterizations.
 This indicates a considerable underestimation of ozone risk for subarctic species.
 Drought effects (SWP) do only matter if a bespoke GS is not taken into account.
-
-In more detail for downy birch (Fig.~\ref{fig:pody_rel}a) all simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
-Downy birch displays the largest range of ozone uptake among all species or PFTs ($10-31\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
+We will look at the specific species or PFTs in more detail in the following.
+
+For downy birch (Fig.~\ref{fig:pody_rel}a) all simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
+Downy birch displays the largest variance of ozone uptake among all species or PFTs ($10-31\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
 In particular, the sensitivity to the light threshold is especially pronounced, e.g.
-$\Delta\mathrm{POD_1} = ^{+2.5}_{-3.6}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of boreal type without bespoke GS in 2019.
-Without bespoke GS 2018 stands out as a higher ozone damage risk year compared to 2019.
+$\Delta\mathrm{POD_1} = ^{+2.5}_{-3.6}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of the boreal type without bespoke GS in 2019.
+Without bespoke GS, 2018 stands out as the year with the higher ozone damage risk than 2019.
 However, taking the later onset and overall shorter GS in 2018 into account, this is reversed for the boreal acclimation type.
 SWP induced reduction in ozone uptake is only effective in 2018 and without bespoke GS.
-The overall bias deduced from bespoke GS for boreal and cold with respect to MM ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to light acclimation of $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
+The overall bias deduced from bespoke GS for boreal and cold with respect to MM ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to the variation of the light threshold of $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
 
@@
 Taking the bespoke GS into account, we find a higher ozone uptake in 2019 for all bespoke temperature response functions.
-The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to light acclimation of $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
+The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to the variation of the light threshold of $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
 
@@
 Perennial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant without bespoke GS.
-The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to light acclimation of $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
+The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to the variation of the light threshold of $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
 
@@
 The estimated biomass reduction in \unit{\%} for the bespoke temperature response function of type boreal in accordance to \citet{ICP:MappingManual2017, ESPR:Hayes2021} is shown in Table~\ref{tab:biomass_reduction}.
-The estimated uncertainty range corresponds to the deviation given by the $\pm 20\,\unit{\%}$ variations in light threshold and SWP=on.
+The estimated uncertainty range corresponds to the deviation given by the $\pm 20\,\unit{\%}$ variations of the light threshold and SWP=on.
 For downy birch and perennial grassland, substantial reductions in total biomass are predicted for any of the two years.
@@
 In comparison with average results for the default MM parameterization (with bespoke GS), we find a non negligible (above systematic uncertainties) low-bias for both downy birch and perennial grassland.
-For Norway spruce, the default MM parameterization average lies within the given systematic uncertainty range.
+For Norway spruce, the default MM parameterization average lies within the given systematic uncertainty range, this may reflect the more comprehensive research on this species or PFT.
 
@@
 \begin{itemize}
-\item Increasing temperature acclimation $\Rightarrow$ increased uptake regardless of individual year,
+  \itemsep0pt
+\item Increasing cold tolerance/heat intolerance $\Rightarrow$ increased uptake regardless of individual year,
 \item usage of proper growing season $\Rightarrow$ reduced uptake with respect to the above,
@@
 \item bespoke light threshold $\Rightarrow$ asymmetric response in POD,
-\item temperature acclimation $\Rightarrow$ amplified effect of light threshold and drought.
+\item temperature acclimation $\Rightarrow$ amplifying effect on light threshold and drought.
 \end{itemize}
-The exact order of magnitude of these effects varies for individual species or PFT and between years, but the thus predicted ozone uptake is in any case larger than for MM default parameterizations.
+The exact order of magnitude of these effects varies for individual species or PFTs and between years.
+Thus predicted ozone uptake is XSlarger than for MM default parameterizations.
 This indicates a considerable underestimation of ozone risk for subarctic species.

### 2021-03-19 VegOzone.tex
Draft for conclusions

@@
 We show two climatological reference periods 1990s and 2000s to indicate that the vegetation is subject to ongoing climate change and acclimation may be changing.
-Resulting response functions for Norway spruce and downy birch are displayed in Appendix~Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}.
+Resulting response functions for Norway spruce (Picea abies) and downy birch are displayed in Appendix~Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}.
 
@@
 Simulations with MM GS, canopy, and leaf parameters are indicated with squared markers, bespoke GS with circles.
-The annotation $\pm$ refers to bespoke light response function threshold ($+$ increased, $-$ decreased), while bespoke temperature response functions are indicated by categories on the x-axis.
+The annotation $\pm$ refers to bespoke light response function threshold ($+\rightarrow\mathrm{PPFD1.2}$, $-\rightarrow\mathrm{PPFD0.8}$), while bespoke temperature response functions are indicated by categories on the x-axis.
 Closed/open symbols represent SWP sensitivity switched off/on in the model.
@@
 \item usage of proper growing season $\Rightarrow$ reduced uptake with respect to the above,
-\item an earlier assumed onset of growing season $\Rightarrow$ increased uptake,
+\item an earlier assumed onset of GS $\Rightarrow$ increased uptake,
 \item drought conditions (SWP) $\Rightarrow$ reduced uptake,
@@
 \item bespoke light threshold $\Rightarrow$ asymmetric response in POD,
-\item temperature acclimation $\Rightarrow$ amplifying effect on light threshold and drought.
+\item temperature acclimation $\Rightarrow$ amplifying effect on light threshold and drought effects.
 \end{itemize}
-The exact order of magnitude of these effects varies for individual species or PFT and between years, but the predicted ozone uptake for the bespoke parameterization is in any case larger than for the MM default parameterizations.
-This indicates a considerable underestimation of ozone risk for subarctic species.
+The exact order of magnitude of these effects varies for individual species or PFTs and between years, but the predicted ozone uptake for the bespoke parameterization is in any case larger than for the MM default parameterizations and of the same order of magnitude as the variability between the years studyed here.
+Although our temperatue acclimations are hypothetical, this indicates a possibly considerable underestimation of ozone risk for subarctic species.
+Especially, an earlier GS due to climate change, that is overlapping with the ozone spring peak, has the potential to increase the ozone damage risk for vegetation with an acclimation to cold climates.
 Drought effects (SWP) do only matter if a bespoke GS is not taken into account.
@@
 $\Delta\mathrm{POD_1} = ^{+2.5}_{-3.6}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of the boreal type without bespoke GS in 2019.
-Without bespoke GS, 2018 stands out as the year with the higher ozone damage risk than 2019.
+Without bespoke GS, 2018 displayes a higher ozone damage risk than 2019.
 However, taking the later onset and overall shorter GS in 2018 into account, this is reversed for the boreal acclimation type.
+This can be partly explained by very high light response factors (over $f_\mathrm{light}>0.8$) in the beginning of the GS in these simulations.
+Whether this ia a bug or a feature of the $\mathrm{DO_3SE}$ model is not yet clear.
 SWP induced reduction in ozone uptake is only effective in 2018 and without bespoke GS.
@@
 For Norway spruce (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
-Norway spruce is the least sensitive to acclimation in temperature and therefor the least biased.
+Norway spruce is the least sensitive to an acclimation in temperature and therefor the least biased.
 Norway spruce displays a medium sensitivity to the light threshold, e.g.
@@
 The systematic uncertainties for downy birch are one order of magnitude larger than for Norway spruce and perennial grassland.
-For all species or PFTs the interannual variability is larger than the systematic uncertainties estimated.
-In comparison with average results for the default MM parameterization (with bespoke GS), we find a non negligible (above systematic uncertainties) low-bias for both downy birch and perennial grassland.
+For all species or PFTs the interannual variability is larger than the systematic uncertainties estimated for a specific acclimation and is of the same order of magnitude as the divergance between the hypothetical acclimations.
+For comparison, we averaged the corresponding biomass reduction for the MM temperature and light response functions with both SWP on/off but with bespoke GS over both years, we find a non negligible, which means above systematic uncertainties, low-bias for both downy birch and perennial grassland with respect to the MM default parameteraziation.
 For Norway spruce, the default MM parameterization average lies within the given systematic uncertainty range, this may reflect the more comprehensive research on this species or PFT.
@@
   \caption{Estimated total biomass reduction in \unit{\%} for bespoke temperature response function of type boreal with bespoke GS and SWP=off.
-Uncertainty ranges reported here correspond to deviations by the $\pm 20\,\unit{\%}$ variation in light threshold and SWP=on, respectively.
-For comparison results for MM parameterization are shown with standard deviations over both years and SWP treatments.}
+The uncertainty ranges reported here correspond to maximum a divergance by varying the light threshold and turning SWP sensitivity on in $\mathrm{DO_3SE}$.
+For comparison, the corresponding biomass reduction for the MM temperature and light response functions, SWP on/off, and bespoke GS averaged over both years is shown with its standard deviation.}
   \label{tab:biomass_reduction}
@@
 \label{sec:conc}
-\begin{itemize}
-  \item What did we learn from the 2018/2019 seasons in terms of ozone damage in the far north?
-  \item What are the main uncertainties in the tools we've used?
-  \item What implications could this have for future scenarios in which 2018 conditions may occur more often?
-\end{itemize}
-PODY is no good measure without a damage function with higher temporal resolution than annual.
-This has been already pointed out by \citet{AE:Musselman2006}!
-Higher than normal spring uptake (weather conditions) does not amount to critical uptake on an annual basis.
-Weather and climatic conditioning of plants is more important for PODy than episodes of higher than usual ozone concentrations.
-Therefor high $\Delta \mathrm{POD_y} / \Delta t$ is not accounted for properly.
-It is important to know the plant response functions to temperature, light, humidity/precipitation as best as possible if one wants to assess risk of damage.
-
-Regarding the shift (earlier) and prolongation (later) of the growing season in the far north no general assessment can be made.
-Data show that in both 2018 and 2019 ozone uptake in early season was anomalous high in all studied PFTs.
-The extended growing season in 2019 might have countered higher episodic ozone uptake in 2018.
-In both years, low mid-season uptake due to weather conditions washed out any effect of these.
-
+We have examplarily studied the effect of acclimations of biomes to a subarctic climate on phyto-toxic ozone uptake.
+We also assessed implications on potential ozone damage risks arising at the advant of climate change.
+We found that conditions for ozone formation were more favourable in the 2018 growing season than in 2019, with 2018 being significantly warmer and less cloudy in spring and early summer.
+Accordingly, peak ozone concentrations occured more frequently and at higher concentration levels in 2018.
+This was in particular caused by the extended heatwave in spring and early summer over whole Fennoscandia and associated, extensive wildfires esecially in Central Sweden (mid--end July).
+
+The timing and trend in the onset of growing season may differ substantially between individual species in comparison to the thermal or latitudinal estimated growing season.
+Species-specific methods to estimate the onset \citep{IVL:Karlsson2018} show a later onset for European silver birch and earlier for Norway spruce, compared to the thermal growing season estimate.
+Based on MODIS data, we estimated the growing season in 2018 to be at least one month shorter than in 2019 for Norway spruce, and half a month shorter for downy birch and perennial grassland, respectively.
+
+We have proposed a routine for gap filling of ozone data following a Reynolds's decomposition into annomalies and climatology (mulit-annual mean).
+We also assessed the quality of available global (ECWMF MACC and CAM reanalysis) and regional (Copernicus) near-surface ozone concentrations products and conclude that only the Copernicus product can sufficiently reproduce observed ozone concentrations in northern Fennoscandia.
+
+Although high and more frequent ozone peak concentrations were observed in 2018, ozone concentrations were not significantly elevated by statistical means.
+It therefor occurs that weather conditions and constantly high ozone concentration are more important in risk assessment using flux-based metrices like $\mathrm{POD_y}$ than episodes of peak concentrations over shorter periodes of time.
+Consequently, accute cellular damage caused by a very high fluxes over a short periode of time is not accounted for properly in $\mathrm{POD_y}$.
+As pointed out by \citet{AE:Musselman2006}, a higher temporal resolution, e.g.
+higher susceptibility to ozone depending on leaf age, in the associated damage functions might give a more differenciated picture.
+
+Besides a currently missing coupling to photosynthesis or nutrient cycle, ozone inflicted damage on stomatal conductance is also not accounted for in the $\mathrm{DO_3SE}$ model.
+Cumulative ozone uptake has been reported to reduce stomatal conductance significantly \citep[e.g.]{BGS:Lombardozzi2012} and subsequently reduces uptake of ozone.
+On the other hand, an ozone induced sluggishness \citep{SR:Hoshika2015} renders stomata inapt to fully close and would lead to increased ozone uptake.
+The presented $\mathrm{POD_y}$ may therefor be regarded as an upper limit.
 {\bf TODO: might be interesting to look at Mills et al 2003 and the different spatial extents of risk estimated when using AOT40 vs PODy.
@@
 
-Further stomatal conductance limit induced through ozone damage itself - varying strength and even sign (sluggishness) can shift the uptake into both directions
-
----
-The results depicted in Fig.~\ref{fig:pody_rel} allow for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric for ozone induced damage in subarctic vegetation.
-In summary, we identify the following main effects on the model predicted ozone uptake with respect to the MM default parameterization:
+We defined hypothetical temperature response functions representing different levels of species assumed acclimation to cold climates.
+We assume the acclimation to increase the enclosed area between probability density function and the respective response function.
+These hypothetical species or PFTs still need varification by experiemental data.
+Our results, however, allowed for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric for ozone induced damage in subarctic vegetation.
+In particular, we identified the following main effects on the model predicted ozone uptake with respect to the default parameterization used in risk assessment:
 \begin{itemize}
@@
 The exact order of magnitude of these effects varies for individual species or PFTs and between years.
-Thus predicted ozone uptake is XSlarger than for MM default parameterizations.
-This indicates a considerable underestimation of ozone risk for subarctic species.
- The impact of the timing of the GS can be at the same order of magnitude as the acclimation of response function but points into the opposite direction.
-Especially, the timing of $A_\mathrm{start}$ with respect to the ozone spring peak induces large uncertainties in anticipation of an acclimation to colder climates.
-
-Bigger picture:
-\begin{itemize}
-\item Ozone damage assessment using mapping manual "compromise" parameterization.
-\item Land-atmosphere modeling: Problems in productivity of species especially in the arctic regions due to PFTs which are not suited for the climate.
-Automating optimization of climate adaption of general PFTs.
-\end{itemize}
-
-GS:
-The timing and trend in the SGS may differ substantially between individual species as compared to the thermal SGS though.
-Species-specific methods to estimate the SGS (Karlsson et al., 2018b) resulted in later SGS for European silver birch (Betula pendula) and earlier for Norway spruce (Picea abies), compared to the thermal SGS (Figure 4).
-The SGS of birch and spruce has also shifted to earlier in the year over time (significantly negative trend since 1994, -0.96 days year-1 and -0.98 days year-1 respectively) in northern Sweden.
-Wallin et al (2013) measured gas-exchange of 40-year-old Norway spruce trees with branch cuvettes in stands at Flakaliden in northern Sweden.
-Measurements of light-saturated photosynthesis were made during spring recovery from dormancy during three years 2002–2004.
-The recovery of gas-exchange from winter dormancy was non-linear over time and the first recovery, up to 60\% of summer maximum gas-exchange, occurred already in March, more than a month before the start of growing season calculated as when the 24-h mean temperatures exceeded 5°C for five consecutive days.
-
-Manninen et al.
-(2009a) summarized the results from open-top chamber (OTC) experiments performed in northern Finland where Scots pine and downy birch provenances were exposed to 36 – 54 ppb O3 as mean daylight concentrations.
-Biomass reductions of both downy birch (Betula pubescens) and mountain birch seedlings were demonstrated after one-season exposure.
-Moreover, the root:shoot ratio was reduced and the leaves of the mountain birch in the high O3 treatments showed necrotic stipples already in June.
-Mountain birch has a very rapid phenological development during spring-early summer that makes it difficult for this species to compensate for damaged leaves.
-It is also noteworthy that N fertilization increased the total biomass production of mountain birch markedly only under the ambient O3 both under the current climate (open-field plots) and in the about 1°C warmer climate (non-filtered air OTCs) but not under elevated O3 (Manninen et al.
-2009b).
-In contrast to mountain birches, northern Scots pine of local origin only showed invisible needle-level O3 impacts, such as increased respiration, decreased level of polyamines and reduced chlorophyll/carotenoid ratio and increased amount of epicuticular waxes after three years of exposure (Manninen et al.
-2009a).
-There is no field data comparing the O3 sensitivity of birches and Scots pine from northern Europe, but both Scots pine and Norway spruce have been ranked as more O3-tolerant species than deciduous birches and Alnus incana based on crown condition in a Lithuanian study (Girgzdiene et al.
-2009).
-Despite the relatively high O3 tolerance of Scots pine, invisible needle injuries may not be excluded in early spring conditions with high O3 concentrations though (Augustaitis and Bytnerowicz 2008, Huttunen and Manninen 2013).
+Predicted ozone uptake with bespoke temperature and light response function and length and timing of growing season is larger than for default parameterizations.
+This indicates a considerable underestimation of ozone risk for subarctic species, especially for downy birch and perennial grassland which have been studied less intensively than Norway spruce which showes a lower susceptibility to ozone \citep{Amb:Girgzdiene2009}.
+
+Open-top chamber (OTC) experiments performed in northern Finland where Scots pine and downy birch were exposed elevated ozone concentrations showed biomass reductions of both downy birch and mountain birch seedlings after one-season exposure \citep{Amb:Manninen2009}.
+The leaves of the mountain birch in the high ozone treatments showed necrotic stipples as early as June.
+As mountain birch very rapidly develops during spring and early summer, compensation for damaged leaves is difficult for this species.
+In contrast to mountain birches, northern Scots pine of local origin only showed invisible needle-level \chem{O_3} impacts after three years of exposure.
+There is no field data comparing the ozone sensitivity of birches and Scots pine from northern Europe.
+
+The impact of the timing of the growing season can be of the same order of magnitude as the acclimation of response functions but points into the opposite direction.
+Especially, the trend towards an earlier onset of the growing season in northern Fennoscandia with respect to the ozone spring peak induces large uncertainties for future risk assessment and includes potential higher ozone damage risk.
+However, the decline of ozone concentrations in arctic spring is partly caused by the removal of ozone through active vegetation at the start ot the growing season.
+It is therefor not clear if the peak concentrations of ozone experienced by vegetation in the future with an earlier onset of the growing season, will be reduced compared to present day.
+Hence, a simulated earlier onset of growing season with present day ozone concentration might overestimate ozone uptake and hence ozone induced risk of damage.
+
+Beyond the risk assessment of ozone induced damage, out hypothetical bespoke temperature reponse function may have important indications for landsurface modeling, where problems in productivity of species especially in the arctic regions occure due to PFTs which are not suitable for the climate.
+An automatition of the here proposed PDF-based acclimation using machine learning techniques could overcome these issues in the future.
 
@@
 
-\dataavailability{TEXT} %% use this section when having only data sets available
+\dataavailability{All data is available from public databases or through institutional access.
+$\mathrm{DO_3SE}$ modeling results can be made available upon request {\bf TODO: Upload to an archieve together with processed, gapfilled input data?}.} %% use this section when having only data sets available
 
@@
 \authorcontribution{SF has written the manuscript, collected and processed ozone and environmental data, and performed all statistical analyses.
-AVV has performed on-site observation of vegetation damage induced by ozone and contributed to theory of ozone interaction with vegetation within the introduction.
-LE...
-CO...
-ALEB...
-FS...
-TB...
-All authors contributed to discussion and conclusion of the results.} %% this section is mandatory
+AVV helped in framing the research articel, has performed on-site observation of vegetation damage induced by ozone, advised with regard to plant physiological processes and contributed text to the introduction.
+LE has contributed to discussions, setup and evaluation of $\mathrm{DO_3SE}$ modeling results, and development of the PDF-based temperature acclimation.
+CO has collected species or PFT parameters from literature, performed all $\mathrm{DO_3SE}$ simulations and initial graphical checks of the results.
+ALEB contributed with her experience regarding subartic vegetation in Finnmark.
+FS contributed with a summary of the 2018 meteorological conditions.
+TB gave guidance in a broader research sense.
+All authors contributed to discussions, writing, and conclusions.} %% this section is mandatory
 
@@
   Tore Flatlandsmo Berglen for hourly pressure data from Svanhovd.
+  This work was financed by the Research Council of Norway (NRC) grand No. xxxxx.
 \end{acknowledgements}

### 2021-03-22 VegOzone.tex
Second version of conclusions - not spell ckecked

@@
 We investigate the difference in uptake of ozone using the $\mathrm{DO_3SE}$ model, for typical vegetation at the location.
-We find that bespoke parameterization for subarctic vegetation is essential for an accurate estimation of Phototoxic Ozone Dose (POD) and an assessment of damage risk.
+We find that bespoke parameterization for subarctic vegetation is essential for an accurate estimation of Phyto-toxic Ozone Dose (POD) and an assessment of damage risk.
 Accurate estimates of the start and length of growing season are also crucial.
@@
 Due to their simplicity, concentration-based metrics are still widely used, despite comparisons between the two methods (e.g.
-$\mathrm{AOT_{40}}$ with $\mathrm{POD_3}$) show that exposure-based metrics are better suited to represent the actual risk to vegetation from ozone, especially in the more climatically extreme parts of Europe \citep{GCB:Mills2010,ICP:MappingManual2017}.
+$\mathrm{AOT_{40}}$ with $\mathrm{POD_3}$) show that exposure-based metrics are better suited to represent the actual risk to vegetation from ozone, especially in the more climatically extreme parts of Europe \citep{GCB:Mills2010,ICP:MappingManual2017,iFJuran2018}.
 Most of these previous studies focused on the Mediterranean region.
@@
 Based on data from SeNorge.no and the agricultural rule of thumb that the daily average temperature has to be above $5\,\unit{^\circ C}$ for $5\,\unit{days}$ consecutively, we find $\left<G_\mathrm{begin}\right> = (150 \pm 14)\,\unit{days of the year}$ at Svanvik (Supplement Fig~\ref{fig:greening_season_change_Svanvik}), which translates into the end of May.
-This finding is in line with more extensive studies \citep[][e.g.,]{IJB:Karlsen2007,AFM:Linderholm}.
+This finding is in line with more extensive studies \citep[][e.g.,]{IJB:Karlsen2007,AFM:Linderholm2006}.
 
@@
 We found that conditions for ozone formation were more favourable in the 2018 growing season than in 2019, with 2018 being significantly warmer and less cloudy in spring and early summer.
-Accordingly, peak ozone concentrations occured more frequently and at higher concentration levels in 2018.
-This was in particular caused by the extended heatwave in spring and early summer over whole Fennoscandia and associated, extensive wildfires esecially in Central Sweden (mid--end July).
-
-The timing and trend in the onset of growing season may differ substantially between individual species in comparison to the thermal or latitudinal estimated growing season.
-Species-specific methods to estimate the onset \citep{IVL:Karlsson2018} show a later onset for European silver birch and earlier for Norway spruce, compared to the thermal growing season estimate.
-Based on MODIS data, we estimated the growing season in 2018 to be at least one month shorter than in 2019 for Norway spruce, and half a month shorter for downy birch and perennial grassland, respectively.
-
-We have proposed a routine for gap filling of ozone data following a Reynolds's decomposition into annomalies and climatology (mulit-annual mean).
-We also assessed the quality of available global (ECWMF MACC and CAM reanalysis) and regional (Copernicus) near-surface ozone concentrations products and conclude that only the Copernicus product can sufficiently reproduce observed ozone concentrations in northern Fennoscandia.
+Accordingly, peak ozone concentrations occured more frequently and at higher levels in 2018.
+This was in particular caused by the extended heatwave in spring and early summer and associated, extensive wildfires in Central Sweden (in particular in mid--end July).
+
+We found that the length and timing of the growing season is crucial for risk assessment based on accumulated ozone flux metrices ($\mathrm{POD_y}$).
+Based on a MODIS photosynthesis product, we estimated the growing season for Norway spruce in 2018 to be at least one month shorter than in 2019.
+The found onsets ($106/122\,\unit{doy}$) are within given ranges by empirical observations \citep{TB:Kolari2007,IVL:Karlsson2018}.
+Based on gridded observational temperature and snow depth, we found a half a month shorter growing season for downy birch and perennial grassland in 2018 compared to 2019, respectively.
+The timing and emerging trend in the onset of growing season may differ substantially between local observation and common thermal or latitudinal estimations.
+Species-specific methods to estimators show a later onset for European silver birch and an earlier for Norway spruce, compared to the thermal growing season \citep{IVL:Karlsson2018}.
+With respect to ongoing climate change, a clear positive trend in length ($5.2\,\unit{d\,decade^{-1}}$) almost equally distributed between earlier start ($2.9\,\unit{days\,decade^{-1}}$) and later end ($2.3\,\unit{d\,decade^{-1}}$) could be deducted from the thermal growing season at Svanhovd.
+
+We found representative ozone climatologies (multi-annual means) for northern Fennoscandia and Svanhovd based on surface ozone monitoring.
+To account for missing data from our 2018 record, we proposed a routine for gap filling of ozone observations following a Reynolds's decomposition into annomalies and climatology.
+We also assessed the quality of available global reanalysis (ECWMF MACC and CAM) and regional model reanalysis (Copernicus) near-surface ozone concentrations products and conclude that only the Copernicus product can sufficiently reproduce observed ozone concentrations in northern Fennoscandia.
 
@@
 It therefor occurs that weather conditions and constantly high ozone concentration are more important in risk assessment using flux-based metrices like $\mathrm{POD_y}$ than episodes of peak concentrations over shorter periodes of time.
-Consequently, accute cellular damage caused by a very high fluxes over a short periode of time is not accounted for properly in $\mathrm{POD_y}$.
+Consequently, accute cellular damage caused by very high fluxes over a short periode of time is not accounted for properly in $\mathrm{POD_y}$.
 As pointed out by \citet{AE:Musselman2006}, a higher temporal resolution, e.g.
-higher susceptibility to ozone depending on leaf age, in the associated damage functions might give a more differenciated picture.
-
-Besides a currently missing coupling to photosynthesis or nutrient cycle, ozone inflicted damage on stomatal conductance is also not accounted for in the $\mathrm{DO_3SE}$ model.
-Cumulative ozone uptake has been reported to reduce stomatal conductance significantly \citep[e.g.]{BGS:Lombardozzi2012} and subsequently reduces uptake of ozone.
-On the other hand, an ozone induced sluggishness \citep{SR:Hoshika2015} renders stomata inapt to fully close and would lead to increased ozone uptake.
-The presented $\mathrm{POD_y}$ may therefor be regarded as an upper limit.
-{\bf TODO: might be interesting to look at Mills et al 2003 and the different spatial extents of risk estimated when using AOT40 vs PODy.
-That could be useful to explain why we focus on PODy (rather then AOT40).- Lisa Come back to this later on in the discussions(?) - it has not been breached at Svanvik in northern Fennoscandia.}
-
-We defined hypothetical temperature response functions representing different levels of species assumed acclimation to cold climates.
-We assume the acclimation to increase the enclosed area between probability density function and the respective response function.
+susceptibility to ozone depending on leaf age, in the associated damage functions might give a more differenciated picture.
+For herbal and grassland species (similar for crops) a strong dependency on phenological state, e.g.
+high susceptibility to ozone in the reproductive state, were found \citep{EP:Bassin2004}.
+
+To estimate potential biases in assessment of ozone induced risk on Fennoscandic vegetation, we defined hypothetical temperature response functions representing different levels of assumed acclimation to cold climates.
+We assume northern Fennoscandic vegetation to be more tolerant towards cold temperatures and less tolerant to heat.
+Based on probability density functions (PDF) of observed light and temperature, we assume that an acclimation to increase the enclosed area between the respective response functions and the PDF.
 These hypothetical species or PFTs still need varification by experiemental data.
-Our results, however, allowed for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric for ozone induced damage in subarctic vegetation.
-In particular, we identified the following main effects on the model predicted ozone uptake with respect to the default parameterization used in risk assessment:
+Our results, however, allowed for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric.
+In particular, we identified the following main effects on the model predicted ozone uptake with respect to the default parameterization:
 \begin{itemize}
@@
 The exact order of magnitude of these effects varies for individual species or PFTs and between years.
-Predicted ozone uptake with bespoke temperature and light response function and length and timing of growing season is larger than for default parameterizations.
-This indicates a considerable underestimation of ozone risk for subarctic species, especially for downy birch and perennial grassland which have been studied less intensively than Norway spruce which showes a lower susceptibility to ozone \citep{Amb:Girgzdiene2009}.
-
-Open-top chamber (OTC) experiments performed in northern Finland where Scots pine and downy birch were exposed elevated ozone concentrations showed biomass reductions of both downy birch and mountain birch seedlings after one-season exposure \citep{Amb:Manninen2009}.
-The leaves of the mountain birch in the high ozone treatments showed necrotic stipples as early as June.
-As mountain birch very rapidly develops during spring and early summer, compensation for damaged leaves is difficult for this species.
-In contrast to mountain birches, northern Scots pine of local origin only showed invisible needle-level \chem{O_3} impacts after three years of exposure.
-There is no field data comparing the ozone sensitivity of birches and Scots pine from northern Europe.
+Only accounting for acclimation, the predicted ozone uptake is typically larger than for default parameterizations.
+This indicates a considerable underestimation of ozone risk for subarctic species, especially for downy birch and perennial grassland.
+However, taking bespoke length and timing of growing season into account the opposite might be the case under certain conditions, e.g.
+a later start of the growing season with respect to the ozone spring peak for downy birch annd perennial grassland.
+Simulations of Norway spruce with bespoke (shorter than default) growing season show only little sensitivity to acclimations and inter-annual variation of meteorological conditions.
+We have not assessed the effect of VPD response function here, but our climatological data show that conditions at Svanhovd are shifting towards a more VPD limited regime.
 
@@
 
-Beyond the risk assessment of ozone induced damage, out hypothetical bespoke temperature reponse function may have important indications for landsurface modeling, where problems in productivity of species especially in the arctic regions occure due to PFTs which are not suitable for the climate.
-An automatition of the here proposed PDF-based acclimation using machine learning techniques could overcome these issues in the future.
+Based on the methodes described in \citet{ICP:MappingManual2017}, we estimated biomass reductions due to ozone uptake for both 2018 and 2019 and find significant reductions for downy birch ($15.5-17.4\,\unit{\%}$) and perennial grassland (above ground, $12.4-13.7\,\unit{\%}$) with respect to estimates based on $\mathrm{POD_1}$ from default parameterizations.
+For Norway spruce, the results are compatible with the default parameterization estimates and are contained within the range of systematic uncertainties deduced from our bespoke response function ($2.5-3\,\unit{\%}$).
+These results have to be treated with care, because the biomass reduction functions have been evaluated based on central European species or PFTs and might as well be subject to biases.
+
+At the same time, intra-specied variability even at the same location is typically non neglicable \citep{EP:Bassin2004,Amb:Girgzdiene2009}.
+This may explain contradictory results regarding the ozone sensitivity of natural vegetation in northern Europe.
+While \citet{FS:Subramanian2014} reported a modeled reduction of net primary production in Norway spruce ($4.3-15.5\,\unit{\%}$) and birch ($1.4-4.3\,\unit{\%}$) under elevated ozone in Sweden, Lithuanian forest decidous trees were observed to be more susceptible to visible damage by ozone than Norway spruce \citep{Amb:Girgzdiene2009}.
+Open-top chamber (OTC) experiments performed in northern Finland where Scots pine and downy birch were exposed to elevated ozone concentrations showed biomass reductions of both downy birch and mountain birch seedlings after one-season exposure \citep{Amb:Manninen2009}.
+The leaves of the mountain birch in the high ozone treatments showed necrotic stipples as early as June.
+As mountain birch very rapidly develops during spring and early summer, compensation for damaged leaves is difficult for this species.
+In contrast to mountain birches, northern Scots pine of local origin only showed invisible needle-level \chem{O_3} impacts after three years of exposure.
+There is no field data comparing the ozone sensitivity of birches and Scots pine from northern Europe.
+In our study, we do not explicitly account for this natural distribution of individual plant traits, but the chosen hypothetical, bespoke response functions might reflect this diversity within a given species or PFT.
+The resiliance of northern Fennoscandic plant communities to climate change and predictions in this regard strongly depends on both the existing range of acclimations as well as the actual type of acclimation.
+
+Cumulative ozone uptake has been reported to reduce stomatal conductance significatly and independently from photosynthesis \citep[e.g.]{BGS:Lombardozzi2012}, with photosynthesis being more strongly affacted.
+This might be due competing effects, as an ozone induced sluggishness \citep{SR:Hoshika2015} renders stomata inapt to fully close and leads to higher stomatal conductance than expected and thus higher ozone uptake as well as transpiration.
+At present, the $\mathrm{DO_3SE}$ model does not account for any ozone inflicted reduction of stomatal conductance throughout the growing season and is not coupled to photosynthesis.
+Hence, the presented $\mathrm{POD_y}$ may be regarded as an estimated upper limit.
+
+Beyond the risk assessment of ozone induced damage, our hypothetical bespoke temperature reponse function may have important indications for landsurface modeling, where problems in productivity of species especially in the Arctic regions occure due to PFTs which are not suitable for the climate.
+We pointed out with respect to default perennial grassland parameterization that that did not show any substential stomatal conductance in 2019 which we characterized as a normal year.
+In termas of GPP this would mean unrealistically low carbon sequestation.
+An automation of the here proposed PDF-based acclimation using machine learning techniques could overcome these issues in the future.
 

### 2021-03-22 VegOzone.tex
Spell-checked

@@
 \end{itemize}
-The exact order of magnitude of these effects varies for individual species or PFTs and between years, but the predicted ozone uptake for the bespoke parameterization is in any case larger than for the MM default parameterizations and of the same order of magnitude as the variability between the years studyed here.
-Although our temperatue acclimations are hypothetical, this indicates a possibly considerable underestimation of ozone risk for subarctic species.
+The exact order of magnitude of these effects varies for individual species or PFTs and between years, but the predicted ozone uptake for the bespoke parameterization is in any case larger than for the MM default parameterizations and of the same order of magnitude as the variability between the years studied here.
+Although our temperature acclimations are hypothetical, this indicates a possibly considerable underestimation of ozone risk for subarctic species.
 Especially, an earlier GS due to climate change, that is overlapping with the ozone spring peak, has the potential to increase the ozone damage risk for vegetation with an acclimation to cold climates.
@@
 $\Delta\mathrm{POD_1} = ^{+2.5}_{-3.6}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of the boreal type without bespoke GS in 2019.
-Without bespoke GS, 2018 displayes a higher ozone damage risk than 2019.
+Without bespoke GS, 2018 displays a higher ozone damage risk than 2019.
 However, taking the later onset and overall shorter GS in 2018 into account, this is reversed for the boreal acclimation type.
 This can be partly explained by very high light response factors (over $f_\mathrm{light}>0.8$) in the beginning of the GS in these simulations.
-Whether this ia a bug or a feature of the $\mathrm{DO_3SE}$ model is not yet clear.
+Whether this is a bug or a feature of the $\mathrm{DO_3SE}$ model is not yet clear.
 SWP induced reduction in ozone uptake is only effective in 2018 and without bespoke GS.
@@
 The systematic uncertainties for downy birch are one order of magnitude larger than for Norway spruce and perennial grassland.
-For all species or PFTs the interannual variability is larger than the systematic uncertainties estimated for a specific acclimation and is of the same order of magnitude as the divergance between the hypothetical acclimations.
-For comparison, we averaged the corresponding biomass reduction for the MM temperature and light response functions with both SWP on/off but with bespoke GS over both years, we find a non negligible, which means above systematic uncertainties, low-bias for both downy birch and perennial grassland with respect to the MM default parameteraziation.
+For all species or PFTs the interannual variability is larger than the systematic uncertainties estimated for a specific acclimation and is of the same order of magnitude as the divergence between the hypothetical acclimations.
+For comparison, we averaged the corresponding biomass reduction for the MM temperature and light response functions with both SWP on/off but with bespoke GS over both years, we find a non negligible, which means above systematic uncertainties, low-bias for both downy birch and perennial grassland with respect to the MM default parameterization.
 For Norway spruce, the default MM parameterization average lies within the given systematic uncertainty range, this may reflect the more comprehensive research on this species or PFT.
@@
   \caption{Estimated total biomass reduction in \unit{\%} for bespoke temperature response function of type boreal with bespoke GS and SWP=off.
-The uncertainty ranges reported here correspond to maximum a divergance by varying the light threshold and turning SWP sensitivity on in $\mathrm{DO_3SE}$.
+The uncertainty ranges reported here correspond to maximum a divergence by varying the light threshold and turning SWP sensitivity on in $\mathrm{DO_3SE}$.
 For comparison, the corresponding biomass reduction for the MM temperature and light response functions, SWP on/off, and bespoke GS averaged over both years is shown with its standard deviation.}
@@
 \label{sec:conc}
-We have examplarily studied the effect of acclimations of biomes to a subarctic climate on phyto-toxic ozone uptake.
-We also assessed implications on potential ozone damage risks arising at the advant of climate change.
-We found that conditions for ozone formation were more favourable in the 2018 growing season than in 2019, with 2018 being significantly warmer and less cloudy in spring and early summer.
-Accordingly, peak ozone concentrations occured more frequently and at higher levels in 2018.
+We have exemplarily studied the effect of acclimations of biomes to a subarctic climate on phyto-toxic ozone uptake.
+We also assessed implications on potential ozone damage risks arising at the advent of climate change.
+We found that conditions for ozone formation were more favorable in the 2018 growing season than in 2019, with 2018 being significantly warmer and less cloudy in spring and early summer.
+Accordingly, peak ozone concentrations occurred more frequently and at higher levels in 2018.
 This was in particular caused by the extended heatwave in spring and early summer and associated, extensive wildfires in Central Sweden (in particular in mid--end July).
 
-We found that the length and timing of the growing season is crucial for risk assessment based on accumulated ozone flux metrices ($\mathrm{POD_y}$).
+We found that the length and timing of the growing season is crucial for risk assessment based on accumulated ozone flux metrics ($\mathrm{POD_y}$).
 Based on a MODIS photosynthesis product, we estimated the growing season for Norway spruce in 2018 to be at least one month shorter than in 2019.
@@
 We found representative ozone climatologies (multi-annual means) for northern Fennoscandia and Svanhovd based on surface ozone monitoring.
-To account for missing data from our 2018 record, we proposed a routine for gap filling of ozone observations following a Reynolds's decomposition into annomalies and climatology.
-We also assessed the quality of available global reanalysis (ECWMF MACC and CAM) and regional model reanalysis (Copernicus) near-surface ozone concentrations products and conclude that only the Copernicus product can sufficiently reproduce observed ozone concentrations in northern Fennoscandia.
+To account for missing data from our 2018 record, we proposed a routine for gap filling of ozone observations following a Reynolds's decomposition into anomalies and climatology.
+We also assessed the quality of available global reanalysis (ECMWF MACC and CAM) and regional model reanalysis (Copernicus) near-surface ozone concentrations products and conclude that only the Copernicus product can sufficiently reproduce observed ozone concentrations in northern Fennoscandia.
 
 Although high and more frequent ozone peak concentrations were observed in 2018, ozone concentrations were not significantly elevated by statistical means.
-It therefor occurs that weather conditions and constantly high ozone concentration are more important in risk assessment using flux-based metrices like $\mathrm{POD_y}$ than episodes of peak concentrations over shorter periodes of time.
-Consequently, accute cellular damage caused by very high fluxes over a short periode of time is not accounted for properly in $\mathrm{POD_y}$.
+It therefor occurs that weather conditions and constantly high ozone concentration are more important in risk assessment using flux-based metrics like $\mathrm{POD_y}$ than episodes of peak concentrations over shorter periods of time.
+Consequently, acute cellular damage caused by very high fluxes over a short period of time is not accounted for properly in $\mathrm{POD_y}$.
 As pointed out by \citet{AE:Musselman2006}, a higher temporal resolution, e.g.
-susceptibility to ozone depending on leaf age, in the associated damage functions might give a more differenciated picture.
+susceptibility to ozone depending on leaf age, in the associated damage functions might give a more differentiate picture.
 For herbal and grassland species (similar for crops) a strong dependency on phenological state, e.g.
@@
 Based on probability density functions (PDF) of observed light and temperature, we assume that an acclimation to increase the enclosed area between the respective response functions and the PDF.
-These hypothetical species or PFTs still need varification by experiemental data.
+These hypothetical species or PFTs still need verification by experimental data.
 Our results, however, allowed for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric.
@@
 However, taking bespoke length and timing of growing season into account the opposite might be the case under certain conditions, e.g.
-a later start of the growing season with respect to the ozone spring peak for downy birch annd perennial grassland.
+a later start of the growing season with respect to the ozone spring peak for downy birch and perennial grassland.
 Simulations of Norway spruce with bespoke (shorter than default) growing season show only little sensitivity to acclimations and inter-annual variation of meteorological conditions.
@@
 Especially, the trend towards an earlier onset of the growing season in northern Fennoscandia with respect to the ozone spring peak induces large uncertainties for future risk assessment and includes potential higher ozone damage risk.
-However, the decline of ozone concentrations in arctic spring is partly caused by the removal of ozone through active vegetation at the start ot the growing season.
+However, the decline of ozone concentrations in arctic spring is partly caused by the removal of ozone through active vegetation at the start of the growing season.
 It is therefor not clear if the peak concentrations of ozone experienced by vegetation in the future with an earlier onset of the growing season, will be reduced compared to present day.
@@
 
-Based on the methodes described in \citet{ICP:MappingManual2017}, we estimated biomass reductions due to ozone uptake for both 2018 and 2019 and find significant reductions for downy birch ($15.5-17.4\,\unit{\%}$) and perennial grassland (above ground, $12.4-13.7\,\unit{\%}$) with respect to estimates based on $\mathrm{POD_1}$ from default parameterizations.
+Based on the methods described in \citet{ICP:MappingManual2017}, we estimated biomass reductions due to ozone uptake for both 2018 and 2019 and find significant reductions for downy birch ($15.5-17.4\,\unit{\%}$) and perennial grassland (above ground, $12.4-13.7\,\unit{\%}$) with respect to estimates based on $\mathrm{POD_1}$ from default parameterizations.
 For Norway spruce, the results are compatible with the default parameterization estimates and are contained within the range of systematic uncertainties deduced from our bespoke response function ($2.5-3\,\unit{\%}$).
@@
 
-At the same time, intra-specied variability even at the same location is typically non neglicable \citep{EP:Bassin2004,Amb:Girgzdiene2009}.
+At the same time, intra-species variability even at the same location is typically non negligible \citep{EP:Bassin2004,Amb:Girgzdiene2009}.
 This may explain contradictory results regarding the ozone sensitivity of natural vegetation in northern Europe.
@@
 In our study, we do not explicitly account for this natural distribution of individual plant traits, but the chosen hypothetical, bespoke response functions might reflect this diversity within a given species or PFT.
-The resiliance of northern Fennoscandic plant communities to climate change and predictions in this regard strongly depends on both the existing range of acclimations as well as the actual type of acclimation.
-
-Cumulative ozone uptake has been reported to reduce stomatal conductance significatly and independently from photosynthesis \citep[e.g.]{BGS:Lombardozzi2012}, with photosynthesis being more strongly affacted.
+The resilience of northern Fennoscandic plant communities to climate change and predictions in this regard strongly depends on both the existing range of acclimations as well as the actual type of acclimation.
+
+Cumulative ozone uptake has been reported to reduce stomatal conductance significantly and independently from photosynthesis \citep[e.g.]{BGS:Lombardozzi2012}, with photosynthesis being more strongly affected.
 This might be due competing effects, as an ozone induced sluggishness \citep{SR:Hoshika2015} renders stomata inapt to fully close and leads to higher stomatal conductance than expected and thus higher ozone uptake as well as transpiration.
@@
 
-Beyond the risk assessment of ozone induced damage, our hypothetical bespoke temperature reponse function may have important indications for landsurface modeling, where problems in productivity of species especially in the Arctic regions occure due to PFTs which are not suitable for the climate.
-We pointed out with respect to default perennial grassland parameterization that that did not show any substential stomatal conductance in 2019 which we characterized as a normal year.
-In termas of GPP this would mean unrealistically low carbon sequestation.
+Beyond the risk assessment of ozone induced damage, our hypothetical bespoke temperature response function may have important indications for land-surface modeling, where problems in productivity of species especially in the Arctic regions occur due to PFTs which are not suitable for the climate.
+We pointed out with respect to default perennial grassland parameterization that that did not show any substantial stomatal conductance in 2019 which we characterized as a normal year.
+In terms of GPP this would mean unrealistically low carbon sequestration.
 An automation of the here proposed PDF-based acclimation using machine learning techniques could overcome these issues in the future.
@@
 \dataavailability{All data is available from public databases or through institutional access.
-$\mathrm{DO_3SE}$ modeling results can be made available upon request {\bf TODO: Upload to an archieve together with processed, gapfilled input data?}.} %% use this section when having only data sets available
+$\mathrm{DO_3SE}$ modeling results can be made available upon request {\bf TODO: Upload to an achieve together with processed, gap filled input data?}.} %% use this section when having only data sets available
 
@@
 \authorcontribution{SF has written the manuscript, collected and processed ozone and environmental data, and performed all statistical analyses.
-AVV helped in framing the research articel, has performed on-site observation of vegetation damage induced by ozone, advised with regard to plant physiological processes and contributed text to the introduction.
+AVV helped in framing the research article, has performed on-site observation of vegetation damage induced by ozone, advised with regard to plant physiological processes and contributed text to the introduction.
 LE has contributed to discussions, setup and evaluation of $\mathrm{DO_3SE}$ modeling results, and development of the PDF-based temperature acclimation.
 CO has collected species or PFT parameters from literature, performed all $\mathrm{DO_3SE}$ simulations and initial graphical checks of the results.
-ALEB contributed with her experience regarding subartic vegetation in Finnmark.
+ALEB contributed with her experience regarding subarctic vegetation in Finnmark.
 FS contributed with a summary of the 2018 meteorological conditions.

### 2021-03-22 VegOzone.tex
Edited conclusions

@@
 The steep rise of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} in spring is in coincidence with increasing temperatures as well as increase in global irradiance.
-The decrease in \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} begins about one month ahead of the estimated onset of the greening season.
+The decrease in \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} begins about one month ahead of the estimated start of the greening season.
 Based on data from SeNorge.no and the agricultural rule of thumb that the daily average temperature has to be above $5\,\unit{^\circ C}$ for $5\,\unit{days}$ consecutively, we find $\left<G_\mathrm{begin}\right> = (150 \pm 14)\,\unit{days of the year}$ at Svanvik (Supplement Fig~\ref{fig:greening_season_change_Svanvik}), which translates into the end of May.
@@
 \item usage of proper growing season $\Rightarrow$ reduced uptake with respect to the above,
-\item an earlier assumed onset of GS $\Rightarrow$ increased uptake,
+\item an earlier assumed start of GS $\Rightarrow$ increased uptake,
 \item drought conditions (SWP) $\Rightarrow$ reduced uptake,
@@
 Without bespoke GS, 2018 displays a higher ozone damage risk than 2019.
-However, taking the later onset and overall shorter GS in 2018 into account, this is reversed for the boreal acclimation type.
+However, taking the later start and overall shorter GS in 2018 into account, this is reversed for the boreal acclimation type.
 This can be partly explained by very high light response factors (over $f_\mathrm{light}>0.8$) in the beginning of the GS in these simulations.
@@
 SWP induced reduction in ozone uptake is only effective in 2018 and without bespoke GS.
-The overall bias deduced from bespoke GS for boreal and cold with respect to MM ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to the variation of the light threshold of $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
+The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to the variation of the light threshold of $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
 
@@
 For all species or PFTs the interannual variability is larger than the systematic uncertainties estimated for a specific acclimation and is of the same order of magnitude as the divergence between the hypothetical acclimations.
-For comparison, we averaged the corresponding biomass reduction for the MM temperature and light response functions with both SWP on/off but with bespoke GS over both years, we find a non negligible, which means above systematic uncertainties, low-bias for both downy birch and perennial grassland with respect to the MM default parameterization.
+For comparison, we averaged the corresponding biomass reduction for the MM temperature and light response functions with both SWP on/off but with bespoke GS over both years and find a non negligible, which means above systematic uncertainties, low-bias for both downy birch and perennial grassland with respect to the MM default parameterization.
 For Norway spruce, the default MM parameterization average lies within the given systematic uncertainty range, this may reflect the more comprehensive research on this species or PFT.
@@
 Based on a MODIS photosynthesis product, we estimated the growing season for Norway spruce in 2018 to be at least one month shorter than in 2019.
-The found onsets ($106/122\,\unit{doy}$) are within given ranges by empirical observations \citep{TB:Kolari2007,IVL:Karlsson2018}.
+The found start of growing season ($106/122\,\unit{doy}$) is within given ranges by empirical observations \citep{TB:Kolari2007,IVL:Karlsson2018}.
 Based on gridded observational temperature and snow depth, we found a half a month shorter growing season for downy birch and perennial grassland in 2018 compared to 2019, respectively.
-The timing and emerging trend in the onset of growing season may differ substantially between local observation and common thermal or latitudinal estimations.
-Species-specific methods to estimators show a later onset for European silver birch and an earlier for Norway spruce, compared to the thermal growing season \citep{IVL:Karlsson2018}.
+The timing and emerging trend in the start of growing season may differ substantially between local observation and common thermal or latitudinal estimations.
+Species-specific methods show a later start for European silver birch and an earlier for Norway spruce, compared to the thermal growing season \citep{IVL:Karlsson2018}.
 With respect to ongoing climate change, a clear positive trend in length ($5.2\,\unit{d\,decade^{-1}}$) almost equally distributed between earlier start ($2.9\,\unit{days\,decade^{-1}}$) and later end ($2.3\,\unit{d\,decade^{-1}}$) could be deducted from the thermal growing season at Svanhovd.
@@
 
-Although high and more frequent ozone peak concentrations were observed in 2018, ozone concentrations were not significantly elevated by statistical means.
+Although high and more frequent ozone peak concentrations were observed in 2018, ozone concentrations were, in general, not significantly elevated by statistical means.
 It therefor occurs that weather conditions and constantly high ozone concentration are more important in risk assessment using flux-based metrics like $\mathrm{POD_y}$ than episodes of peak concentrations over shorter periods of time.
@@
 For herbal and grassland species (similar for crops) a strong dependency on phenological state, e.g.
-high susceptibility to ozone in the reproductive state, were found \citep{EP:Bassin2004}.
+high susceptibility to ozone in the reproductive state, was found \citep{EP:Bassin2004}.
 
@@
 We assume northern Fennoscandic vegetation to be more tolerant towards cold temperatures and less tolerant to heat.
-Based on probability density functions (PDF) of observed light and temperature, we assume that an acclimation to increase the enclosed area between the respective response functions and the PDF.
+Based on probability density functions (PDF) of observed light and temperature, we assume that an acclimation ought to increase the enclosed area between the respective response functions and the PDF.
 These hypothetical species or PFTs still need verification by experimental data.
@@
 \item usage of proper growing season $\Rightarrow$ reduced uptake with respect to the above,
-\item an earlier assumed onset of growing season $\Rightarrow$ increased uptake,
+\item an earlier assumed start of growing season $\Rightarrow$ increased uptake,
 \item drought conditions (SWP) $\Rightarrow$ reduced uptake,
@@
 This indicates a considerable underestimation of ozone risk for subarctic species, especially for downy birch and perennial grassland.
-However, taking bespoke length and timing of growing season into account the opposite might be the case under certain conditions, e.g.
+However, taking bespoke length and timing of growing season into account the opposite is be the case under certain conditions, e.g.
 a later start of the growing season with respect to the ozone spring peak for downy birch and perennial grassland.
@@
 The impact of the timing of the growing season can be of the same order of magnitude as the acclimation of response functions but points into the opposite direction.
-Especially, the trend towards an earlier onset of the growing season in northern Fennoscandia with respect to the ozone spring peak induces large uncertainties for future risk assessment and includes potential higher ozone damage risk.
+Especially, the trend towards an earlier start of the growing season in northern Fennoscandia with respect to the ozone spring peak induces large uncertainties for future risk assessment and includes potential higher ozone damage risk.
 However, the decline of ozone concentrations in arctic spring is partly caused by the removal of ozone through active vegetation at the start of the growing season.
-It is therefor not clear if the peak concentrations of ozone experienced by vegetation in the future with an earlier onset of the growing season, will be reduced compared to present day.
-Hence, a simulated earlier onset of growing season with present day ozone concentration might overestimate ozone uptake and hence ozone induced risk of damage.
+It is therefor not clear if the peak concentrations of ozone experienced by vegetation in the future with an earlier start of the growing season, will be reduced compared to present day.
+Hence, a simulated earlier start of growing season with present day ozone concentration might overestimate ozone uptake and hence ozone induced risk of damage.
 
@@
 For Norway spruce, the results are compatible with the default parameterization estimates and are contained within the range of systematic uncertainties deduced from our bespoke response function ($2.5-3\,\unit{\%}$).
-These results have to be treated with care, because the biomass reduction functions have been evaluated based on central European species or PFTs and might as well be subject to biases.
+These results have to be treated with care, because the biomass reduction functions were established based on central European species or PFTs and might as well be subject to biases.
 
@@
 At present, the $\mathrm{DO_3SE}$ model does not account for any ozone inflicted reduction of stomatal conductance throughout the growing season and is not coupled to photosynthesis.
-Hence, the presented $\mathrm{POD_y}$ may be regarded as an estimated upper limit.
-
-Beyond the risk assessment of ozone induced damage, our hypothetical bespoke temperature response function may have important indications for land-surface modeling, where problems in productivity of species especially in the Arctic regions occur due to PFTs which are not suitable for the climate.
-We pointed out with respect to default perennial grassland parameterization that that did not show any substantial stomatal conductance in 2019 which we characterized as a normal year.
+Hence, the presented $\mathrm{POD_y}$ may be regarded as an estimated upper limit.\\
+
+Beyond the risk assessment of ozone induced damage, our hypothetical bespoke temperature response functions, once varified by observations, can have important implications on land-surface modeling in global models, where problems in productivity of species especially in the Arctic regions occur due to PFTs which are not suitable for the climate.
+We pointed out, with respect to default perennial grassland parameterization, that these perennial grasslands did not show any substantial stomatal conductance in 2019 characterized as a normal year.
 In terms of GPP this would mean unrealistically low carbon sequestration.

### 2021-03-26 VegOzone.tex
Fix acronym

@@
 To measure the risk of damage inflicted to vegetation through ozone, different metrics have been defined.
-Two main types can be differentiated, concentration-based and exposure-based metrics, e.g., Accumulated Over a Threshold $\mathrm{x}$ ($\mathrm{AOT_x}$) and Phyto-toxic Ozone Dose over a threshold $\mathrm{y}$ ($\mathrm{POD_y}$).
+Two main types can be differentiated, concentration-based and exposure-based metrics, e.g., Accumulated Ozone Exposure over a threshold of $\mathrm{x}$ ($\mathrm{AOT_x}$) and Phyto-toxic Ozone Dose over a threshold $\mathrm{y}$ ($\mathrm{POD_y}$).
 Due to their simplicity, concentration-based metrics are still widely used, despite comparisons between the two methods (e.g.

### 2021-04-08 VegOzone.tex
Replace surface and ambient air ozone with near-surface ozone

@@
 Here we assess the importance of characterization of key meteorological variables and air pollution profiles on the modeled susceptibility of subarctic vegetation to ozone.
-Changes in climate (e.g.
-temperature, water availability) will manifest in both direct and indirect effects influencing the species' sensitivity to air pollution.
-A likely increase in frequency of persistent heatwaves in the future, will inflict heat related stress while generating favorable conditions for the formation of ambient air ozone, e.g.
+Future climate changes will manifest in both direct and indirect effects influencing the species' exposure to air pollution.
+A likely increase in frequency of persistent heatwaves in the future, will inflict heat related stress while generating favorable conditions for the formation of near-surface ozone, e.g.
 due to wildfires.
-We have monitored ambient air ozone concentrations at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd in Northern Norway in the vegetation period 2018/19.
+We have monitored near-surface ozone concentrations at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd in Northern Norway in the vegetation period 2018/19.
 Due to data acquisition problems, ozone concentrations for two weeks in July 2018 were missing from our record.
-We present a reconstruction based on probability density function with respect to the Swedish and Finnish atmospheric monitoring sites and compare with available surface ozone products for the region.
+We present a reconstruction based on probability density function with respect to the Swedish and Finnish atmospheric monitoring sites and compare with available near-surface ozone products for the region.
 Over all, ozone concentrations did not differ significantly between the two years.
@@
 The crop yield loss studies indicate a threat to food security in rapidly developing countries, e.g., in East and South-East Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018}, where ozone concentrations related to enhanced pollutant emissions from the transport, residential, power generation and industrial sectors have increased in the past few decades.
-Even though long-term (pre 1950) observations of surface ozone concentrations are scarce; evidence from trend analysis of those few sites suggests that since the industrial revolution, background concentrations in the northern hemisphere have at least doubled and continued to increase until relatively recently \citep[Chapter 2]{IPCC2013}.
+Even though long-term (pre 1950) observations of near-surface ozone concentrations are scarce; evidence from trend analysis of those few sites suggests that since the industrial revolution, background concentrations in the northern hemisphere have at least doubled and continued to increase until relatively recently \citep[Chapter 2]{IPCC2013}.
 Recent trend analyses indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric ozone started to level-off or even decline \citep{ESA:Cooper2014, ACP:Wespes2018,ESA:Gaudel2018}.
@@
 %
-with the hourly averaged ambient air ozone flux through the stomata Eq.~(\ref{eq:flux_stomata}) and a flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}\,s^{-1}}.
+with the hourly averaged near-surface ozone flux through the stomata Eq.~(\ref{eq:flux_stomata}) and a flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}\,s^{-1}}.
 The flux threshold $\mathrm{y}$ is representing the detoxification potential of the plant and is typically only exceeded during daylight hours (i.e.
@@
 
-Within the scope of this study to assess probable effects of conditions unique to northern Fennoscandia on ozone uptake by plants and resultant enhanced susceptibility to ozone inflicted damage, we conducted ambient air ozone concentration measurements at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd in the growing seasons of 2018/19 in collaboration with the Norwegian Institute for Air Research (NILU).
+Within the scope of this study to assess probable effects of conditions unique to northern Fennoscandia on ozone uptake by plants and resultant enhanced susceptibility to ozone inflicted damage, we conducted near-surface ozone concentration measurements at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd in the growing seasons of 2018/19 in collaboration with the Norwegian Institute for Air Research (NILU).
 In 2018, we have observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden at Svanhovd.
@@
 The observation site at NIBIO Environment Centre Svanhovd and all related data are presented in Section~\ref{sec:data}.
-In Section~\ref{sec:stats}, we derive climatologies for surface ozone concentration at Svanhovd and northern Fennoscandia based on in situ observations and also compute climatologies of surface temperature, precipitation, and global radiation for Svanhovd.
+In Section~\ref{sec:stats}, we derive climatologies for near-surface ozone concentration at Svanhovd and northern Fennoscandia based on in situ observations and also compute climatologies of near-surface temperature, precipitation, and global radiation for Svanhovd.
 Based on these, we evaluate anomalies of environmental key variables in 2018/19 by statistical means.
-In addition, the derived surface ozone concentration climatology for northern Fennoscandia will be compared with available ozone reanalysis products.
+In addition, the derived near-surface ozone concentration climatology for northern Fennoscandia will be compared with available ozone reanalysis products.
 We use the $\mathrm{DO_3SE}$ model to estimate the ozone uptake by natural and semi-natural vegetation to investigate main environmental drivers for the observed ozone damage in 2018 (Section~\ref{sec:do3se}).
@@
 \label{subsec:other_data}
-To determine the significance of the ozone conditions in 2018, we derive a surface ozone concentration climatology for Svanvik.
+To determine the significance of the ozone conditions in 2018, we derive a near-surface ozone concentration climatology for Svanvik.
 From 1986--1996 ozone monitoring has been operated by NILU at Svanhovd, but no ozone measurements have been conducted for the past $14\,\unit{years}$.
@@
 
-\subsubsection{Surface ozone observations}
+\subsubsection{Near-surface ozone observations}
 \label{subsubsec:ebas}
-All surface ozone observations used in this study are available from the EBAS database operated by \citet{NILU_EBAS}.
+All near-surface ozone observations used in this study are available from the EBAS database operated by \citet{NILU_EBAS}.
 In Table~\ref{tab:ebas_obs}, we give an overview over the exact locations and time range of operational data taking.
-All stations with available long-term observations are located on higher ground than Svanhovd which will result in higher surface ozone concentrations in comparison as ozone abundance, in general, decreases with decreasing altitude in the absence of local precursors \citep[e.g.,][]{AB:Klingberg2009}.
+All stations with available long-term observations are located on higher ground than Svanhovd which will result in higher near-surface ozone concentrations in comparison as ozone abundance, in general, decreases with decreasing altitude in the absence of local precursors \citep[e.g.,][]{AB:Klingberg2009}.
 The periphery vegetation at all sites is similar and consists mainly of pine forests, birch, and heath.
@@
 The temporal as well as spatial resolution of these reanalysis products is rather coarse: 3-hourly and $0.75^\circ\times 0.75^\circ$ or roughly $29.3\,\unit{km}\times 83.4\,\unit{km}$ at the location of Svanvik, respectively.
-In addition, higher resolution surface ozone reanalysis for whole Europe is available from Copernicus service for regional air quality control (note: since May 31st 2020 through the CAMS Atmosphere Data Store (ADS)).
+In addition, higher resolution near-surface ozone reanalysis for whole Europe is available from Copernicus service for regional air quality control (note: since May 31st 2020 through the CAMS Atmosphere Data Store (ADS)).
 The regional air quality model reanalysis ensemble (CAMS regional Ensemble) is based on nine European state-of-the-art numerical air quality models \citep{Copernicus_Regio_AIRQ}.
@@
 The MACC reanalysis assimilates only satellite derived tropospheric column ozone (TC), while CAMSRA also includes ozone profiles from satellite retrievals.
-Most importantly for our study, in situ observations from ozone surface station networks are only assimilated in the CAMS Regional Ensemble.
+Most importantly for our study, in situ observations from ozone near-surface station networks are only assimilated in the CAMS Regional Ensemble.
 The quality with respect to observations in northern Fennoscandia will be discussed in Section~\ref{subsubsec:clim_ozone}.
@@
 As representative for the sites, the nearest neighboring grid point has been chosen.
-The ECMWF products do not reproduce observed surface ozone well.
+The ECMWF products do not reproduce observed near-surface ozone well.
 The MACC reanalysis (Fig.~\ref{fig:ozone_climatologies}a) is in general too low ($(9\pm 7)\,\unit{ppb}$) and displays an erroneous seasonal cycle.
@@
 As representative for the sites, the nearest neighboring grid point has been chosen.
-The ECMWF products do not reproduce observed surface ozone well.
+The ECMWF products do not reproduce observed near-surface ozone well.
 The MACC reanalysis is in general too low and displays an erroneous seasonal cycle.
@@
 All the above factors make it necessary to look not only at anomalies but statistically assess ozone concentration residuals at Svanvik as well as at Esrange and Pallas in comparison.
-Since the climatology of latter is more reliable, we will start with surface ozone concentration residuals \chem{\Delta[O_3]} at Esrange and Pallas in 2018.
+Since the climatology of latter is more reliable, we will start with near-surface ozone concentration residuals \chem{\Delta[O_3]} at Esrange and Pallas in 2018.
 As data for 2019 has not yet been available from EBAS (last access, February 2020), we refrain from looking into these.
@@
 with $\beta = \frac{T_\text{max}-T_\text{opt}}{T_\text{opt}-T_\text{min}}$ a mixed reciprocal polynomial function of temperature.
-The shape parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various species or PFT and $T_\mathrm{air}$ is the surface air temperature.
+The shape parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various species or PFT and $T_\mathrm{air}$ is the near-surface air temperature.
 All temperatures are defined in units of \unit{^\circ C}.
@@
 
-Because the available reanalysis products for surface ozone have their limitations, we deem it necessary to reconstruct the missing data in 2018 for the purpose of $\mathrm{DO_3SE}$ modeling by applying a Reynolds's decomposition of the form:
+Because the available reanalysis products for near-surface ozone have their limitations, we deem it necessary to reconstruct the missing data in 2018 for the purpose of $\mathrm{DO_3SE}$ modeling by applying a Reynolds's decomposition of the form:
 \begin{equation}
@@
 
-We found representative ozone climatologies (multi-annual means) for northern Fennoscandia and Svanhovd based on surface ozone monitoring.
+We found representative ozone climatologies (multi-annual means) for northern Fennoscandia and Svanhovd based on near-surface ozone monitoring.
 To account for missing data from our 2018 record, we proposed a routine for gap filling of ozone observations following a Reynolds's decomposition into anomalies and climatology.

### 2021-04-08 VegOzone.tex
Updated abstract

@@
 \begin{abstract}
-Here we assess the importance of characterization of key meteorological variables and air pollution profiles on the modeled susceptibility of subarctic vegetation to ozone.
+Here we assess the importance of characterization of key meteorological variables and subarctic biomes on the modeled susceptibility of subarctic vegetation to ozone used in pollution and climate risk assessments.
 Future climate changes will manifest in both direct and indirect effects influencing the species' exposure to air pollution.
 A likely increase in frequency of persistent heatwaves in the future, will inflict heat related stress while generating favorable conditions for the formation of near-surface ozone, e.g.
-due to wildfires.
-We have monitored near-surface ozone concentrations at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd in Northern Norway in the vegetation period 2018/19.
+due to wildfires.\\
+We have monitored near-surface ozone concentrations at a high latitude site in Northern Norway during the growing season 2018/19.
 Due to data acquisition problems, ozone concentrations for two weeks in July 2018 were missing from our record.
-We present a reconstruction based on probability density function with respect to the Swedish and Finnish atmospheric monitoring sites and compare with available near-surface ozone products for the region.
-Over all, ozone concentrations did not differ significantly between the two years.
-While temperatures and global irradiance diverged substantially from the multi annual mean precipitation varied only to some extend.
+We present a reconstruction based on probability density function (PDF) and with respect to the Swedish and Finnish atmospheric monitoring sites.
+We review available near-surface ozone reanalysis products for the region and find only CAMS regional Ensemble satisfactory.
+In terms of anomalies from a multi-annual mean, ozone concentrations were only significantly enhanced in April/May 2018 before the start of the growing season.
+Temperatures and global irradiance showed consecutively significant positive anomalies during Spring and Summer 2018.
+In 2019 only early spring and late summer were significantly warmer, while colder during summer.
+Precipitation was lower in both years, but displayed a large variability.\\
 We investigate the difference in uptake of ozone using the $\mathrm{DO_3SE}$ model, for typical vegetation at the location.
 We find that bespoke parameterization for subarctic vegetation is essential for an accurate estimation of Phyto-toxic Ozone Dose (POD) and an assessment of damage risk.
-Accurate estimates of the start and length of growing season are also crucial.
-We find that temperature and its derivative vapor pressure deficit are likely the most important climate variables while the variability of ozone concentration within climatologically reasonable limits plays a minor role.
-The use of subarctic parameterization suggests an improvement in risk assessment as POD values are increased in 2018; corroborated by evidence of visible injury in the Svanhovd ozone garden.
+Accurate estimates of the start and length of growing season are also crucial (up to $15\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ deviation in $\mathrm{POD_1}$).\\
+We find that temperature and its derivative vapor pressure deficit are likely the most important climate variables while the variability of ozone concentration within climatological reasonable limits plays a minor role.
+The use of subarctic parameterization suggests an improvement in risk assessment as POD values are increased compared to standard expectations; corroborated by evidence of visible injury in the Svanhovd ozone garden.
+We estimate a total biomass reduction in Norway spruce of $2.5-3\,\unit{\%}$, downy birch by $15.5-17.4\,\unit{\%}$ and an above ground biomass reduction in perannial grassland of $12.4-13.7\,\unit{\%}$.
 \end{abstract}

### 2021-04-08 VegOzone.tex
Checked with Grammarly

@@
 Future climate changes will manifest in both direct and indirect effects influencing the species' exposure to air pollution.
-A likely increase in frequency of persistent heatwaves in the future, will inflict heat related stress while generating favorable conditions for the formation of near-surface ozone, e.g.
+A likely increase in the frequency of persistent heatwaves in the future will inflict heat-related stress while generating favorable conditions for the formation of near-surface ozone, e.g.
 due to wildfires.\\
@@
 Due to data acquisition problems, ozone concentrations for two weeks in July 2018 were missing from our record.
-We present a reconstruction based on probability density function (PDF) and with respect to the Swedish and Finnish atmospheric monitoring sites.
+We present a reconstruction based on a probability density function (PDF) and concidering the Swedish and Finnish atmospheric monitoring sites.
 We review available near-surface ozone reanalysis products for the region and find only CAMS regional Ensemble satisfactory.
@@
 In 2019 only early spring and late summer were significantly warmer, while colder during summer.
-Precipitation was lower in both years, but displayed a large variability.\\
+Precipitation was lower in both years but displayed a large variability.\\
 We investigate the difference in uptake of ozone using the $\mathrm{DO_3SE}$ model, for typical vegetation at the location.
@@
 We find that temperature and its derivative vapor pressure deficit are likely the most important climate variables while the variability of ozone concentration within climatological reasonable limits plays a minor role.
-The use of subarctic parameterization suggests an improvement in risk assessment as POD values are increased compared to standard expectations; corroborated by evidence of visible injury in the Svanhovd ozone garden.
-We estimate a total biomass reduction in Norway spruce of $2.5-3\,\unit{\%}$, downy birch by $15.5-17.4\,\unit{\%}$ and an above ground biomass reduction in perannial grassland of $12.4-13.7\,\unit{\%}$.
+The use of subarctic parameterization suggests an improvement in risk assessment as POD values are increased compared to standard expectations; corroborated by the evidence of visible injury in the Svanhovd ozone garden.
+We estimate a total biomass reduction in Norway spruce of $2.5-3\,\unit{\%}$, downy birch by $15.5-17.4\,\unit{\%}$ and an above-ground biomass reduction in perennial grassland of $12.4-13.7\,\unit{\%}$.
 \end{abstract}

### 2021-04-08 VegOzone.tex
First paragraph in intro edited and change near-surface ozone to ground level ozone

@@
 Future climate changes will manifest in both direct and indirect effects influencing the species' exposure to air pollution.
-A likely increase in the frequency of persistent heatwaves in the future will inflict heat-related stress while generating favorable conditions for the formation of near-surface ozone, e.g.
+A likely increase in the frequency of persistent heatwaves in the future will inflict heat-related stress while generating favorable conditions for the formation of ground level ozone, e.g.
 due to wildfires.\\
-We have monitored near-surface ozone concentrations at a high latitude site in Northern Norway during the growing season 2018/19.
+We have monitored ground level ozone concentrations at a high latitude site in Northern Norway during the growing season 2018/19.
 Due to data acquisition problems, ozone concentrations for two weeks in July 2018 were missing from our record.
 We present a reconstruction based on a probability density function (PDF) and concidering the Swedish and Finnish atmospheric monitoring sites.
-We review available near-surface ozone reanalysis products for the region and find only CAMS regional Ensemble satisfactory.
+We review available ground level ozone reanalysis products for the region and find only CAMS regional Ensemble satisfactory.
 In terms of anomalies from a multi-annual mean, ozone concentrations were only significantly enhanced in April/May 2018 before the start of the growing season.
@@
 The crop yield loss studies indicate a threat to food security in rapidly developing countries, e.g., in East and South-East Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018}, where ozone concentrations related to enhanced pollutant emissions from the transport, residential, power generation and industrial sectors have increased in the past few decades.
-Even though long-term (pre 1950) observations of near-surface ozone concentrations are scarce; evidence from trend analysis of those few sites suggests that since the industrial revolution, background concentrations in the northern hemisphere have at least doubled and continued to increase until relatively recently \citep[Chapter 2]{IPCC2013}.
-Recent trend analyses indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric ozone started to level-off or even decline \citep{ESA:Cooper2014, ACP:Wespes2018,ESA:Gaudel2018}.
+Even though long-term (pre 1950) observations of ground level ozone concentrations are scarce; evidence from trend analysis of those few sites suggests that since the industrial revolution, background concentrations in the northern hemisphere have at least doubled and continued to increase until relatively recently \citep[Chapter 2]{IPCC2013}.
+Recent trend analyses indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric ozone started to level-off or even decline in Europe \citep{ESA:Cooper2014, ACP:Wespes2018,ESA:Gaudel2018}.
 This can be attributed to a successful implementation of air quality regulations coupled with economic restructuring which saw a move from coal to gas, reducing the number of episodes of peak concentrations especially in summertime over Europe and North America \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
-However, there are indications that an increasing frequency of thermal stress accompanying climate change will reduce vegetation's' capacity for ozone uptake and therefor scotch these achievements \citep{NCC:Lin2020}.\\
+However, there are indications that an increase in the frequency of heatwave induced thermal stress will reduce the vegetations' capacity for ozone uptake and therefor diminish these achievements \citep{NCC:Lin2020}.\\
 
@@
 %
-with the hourly averaged near-surface ozone flux through the stomata Eq.~(\ref{eq:flux_stomata}) and a flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}\,s^{-1}}.
+with the hourly averaged ground level ozone flux through the stomata Eq.~(\ref{eq:flux_stomata}) and a flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}\,s^{-1}}.
 The flux threshold $\mathrm{y}$ is representing the detoxification potential of the plant and is typically only exceeded during daylight hours (i.e.
@@
 
-Within the scope of this study to assess probable effects of conditions unique to northern Fennoscandia on ozone uptake by plants and resultant enhanced susceptibility to ozone inflicted damage, we conducted near-surface ozone concentration measurements at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd in the growing seasons of 2018/19 in collaboration with the Norwegian Institute for Air Research (NILU).
+Within the scope of this study to assess probable effects of conditions unique to northern Fennoscandia on ozone uptake by plants and resultant enhanced susceptibility to ozone inflicted damage, we conducted ground level ozone concentration measurements at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd in the growing seasons of 2018/19 in collaboration with the Norwegian Institute for Air Research (NILU).
 In 2018, we have observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden at Svanhovd.
@@
 The observation site at NIBIO Environment Centre Svanhovd and all related data are presented in Section~\ref{sec:data}.
-In Section~\ref{sec:stats}, we derive climatologies for near-surface ozone concentration at Svanhovd and northern Fennoscandia based on in situ observations and also compute climatologies of near-surface temperature, precipitation, and global radiation for Svanhovd.
+In Section~\ref{sec:stats}, we derive climatologies for ground level ozone concentration at Svanhovd and northern Fennoscandia based on in situ observations and also compute climatologies of near-surface temperature, precipitation, and global radiation for Svanhovd.
 Based on these, we evaluate anomalies of environmental key variables in 2018/19 by statistical means.
-In addition, the derived near-surface ozone concentration climatology for northern Fennoscandia will be compared with available ozone reanalysis products.
+In addition, the derived ground level ozone concentration climatology for northern Fennoscandia will be compared with available ozone reanalysis products.
 We use the $\mathrm{DO_3SE}$ model to estimate the ozone uptake by natural and semi-natural vegetation to investigate main environmental drivers for the observed ozone damage in 2018 (Section~\ref{sec:do3se}).
@@
 \label{subsec:other_data}
-To determine the significance of the ozone conditions in 2018, we derive a near-surface ozone concentration climatology for Svanvik.
+To determine the significance of the ozone conditions in 2018, we derive a ground level ozone concentration climatology for Svanvik.
 From 1986--1996 ozone monitoring has been operated by NILU at Svanhovd, but no ozone measurements have been conducted for the past $14\,\unit{years}$.
@@
 
-\subsubsection{Near-surface ozone observations}
+\subsubsection{Ground level ozone observations}
 \label{subsubsec:ebas}
-All near-surface ozone observations used in this study are available from the EBAS database operated by \citet{NILU_EBAS}.
+All ground level ozone observations used in this study are available from the EBAS database operated by \citet{NILU_EBAS}.
 In Table~\ref{tab:ebas_obs}, we give an overview over the exact locations and time range of operational data taking.
-All stations with available long-term observations are located on higher ground than Svanhovd which will result in higher near-surface ozone concentrations in comparison as ozone abundance, in general, decreases with decreasing altitude in the absence of local precursors \citep[e.g.,][]{AB:Klingberg2009}.
+All stations with available long-term observations are located on higher ground than Svanhovd which will result in higher ground level ozone concentrations in comparison as ozone abundance, in general, decreases with decreasing altitude in the absence of local precursors \citep[e.g.,][]{AB:Klingberg2009}.
 The periphery vegetation at all sites is similar and consists mainly of pine forests, birch, and heath.
@@
 The temporal as well as spatial resolution of these reanalysis products is rather coarse: 3-hourly and $0.75^\circ\times 0.75^\circ$ or roughly $29.3\,\unit{km}\times 83.4\,\unit{km}$ at the location of Svanvik, respectively.
-In addition, higher resolution near-surface ozone reanalysis for whole Europe is available from Copernicus service for regional air quality control (note: since May 31st 2020 through the CAMS Atmosphere Data Store (ADS)).
+In addition, higher resolution ground level ozone reanalysis for whole Europe is available from Copernicus service for regional air quality control (note: since May 31st 2020 through the CAMS Atmosphere Data Store (ADS)).
 The regional air quality model reanalysis ensemble (CAMS regional Ensemble) is based on nine European state-of-the-art numerical air quality models \citep{Copernicus_Regio_AIRQ}.
@@
 As representative for the sites, the nearest neighboring grid point has been chosen.
-The ECMWF products do not reproduce observed near-surface ozone well.
+The ECMWF products do not reproduce observed ground level ozone well.
 The MACC reanalysis (Fig.~\ref{fig:ozone_climatologies}a) is in general too low ($(9\pm 7)\,\unit{ppb}$) and displays an erroneous seasonal cycle.
@@
 As representative for the sites, the nearest neighboring grid point has been chosen.
-The ECMWF products do not reproduce observed near-surface ozone well.
+The ECMWF products do not reproduce observed ground level ozone well.
 The MACC reanalysis is in general too low and displays an erroneous seasonal cycle.
@@
 All the above factors make it necessary to look not only at anomalies but statistically assess ozone concentration residuals at Svanvik as well as at Esrange and Pallas in comparison.
-Since the climatology of latter is more reliable, we will start with near-surface ozone concentration residuals \chem{\Delta[O_3]} at Esrange and Pallas in 2018.
+Since the climatology of latter is more reliable, we will start with ground level ozone concentration residuals \chem{\Delta[O_3]} at Esrange and Pallas in 2018.
 As data for 2019 has not yet been available from EBAS (last access, February 2020), we refrain from looking into these.
@@
 
-Because the available reanalysis products for near-surface ozone have their limitations, we deem it necessary to reconstruct the missing data in 2018 for the purpose of $\mathrm{DO_3SE}$ modeling by applying a Reynolds's decomposition of the form:
+Because the available reanalysis products for ground level ozone have their limitations, we deem it necessary to reconstruct the missing data in 2018 for the purpose of $\mathrm{DO_3SE}$ modeling by applying a Reynolds's decomposition of the form:
 \begin{equation}
@@
 
-We found representative ozone climatologies (multi-annual means) for northern Fennoscandia and Svanhovd based on near-surface ozone monitoring.
+We found representative ozone climatologies (multi-annual means) for northern Fennoscandia and Svanhovd based on ground level ozone monitoring.
 To account for missing data from our 2018 record, we proposed a routine for gap filling of ozone observations following a Reynolds's decomposition into anomalies and climatology.
-We also assessed the quality of available global reanalysis (ECMWF MACC and CAM) and regional model reanalysis (Copernicus) near-surface ozone concentrations products and conclude that only the Copernicus product can sufficiently reproduce observed ozone concentrations in northern Fennoscandia.
+We also assessed the quality of available global reanalysis (ECMWF MACC and CAM) and regional model reanalysis (Copernicus) ground level ozone concentrations products and conclude that only the Copernicus product can sufficiently reproduce observed ozone concentrations in northern Fennoscandia.
 

### 2021-04-08 VegOzone.tex
Included comments by Ane

@@
 
-In the annual cycle in Boreal and temperate climates, a distinct maximum of ground level ozone concentrations typically occurs in spring followed by lower concentrations throughout the summer.
+In the annual cycle in Boreal and also temperate climates, a distinct maximum of ground level ozone concentrations typically occurs in spring followed by lower concentrations throughout the summer.
 Several conditions unique to the Arctic are stimulating the formation of this spring peak.
@@
 This leads to a build up of ozone and its precursors in the course of winter.
-Come spring precursors are photochemical activated again, while snow cover typically continues to prevail and suppress ozone dry deposition.
+Come spring precursors are photochemically activated again, while snow cover typically continues to prevail and suppress ozone dry deposition.
 On the other hand, very high ozone concentrations occurring over shorter periods of time can often be attributed to deep tropopause folding events over the Arctic during winter.
@@
 
-The start of growing season (GS) in northern Fennoscandia has been shifting to earlier dates in response to the general warming trend \citep[e.g.,]{GCB:Menzel2006,RS:Hogda2013,IJB:Karlsen2007} and thereby converging with the period of the ozone spring peak.
+The start of growing season (GS) in northern Fennoscandia has been shifting to earlier dates in response to the general warming trend \citep[e.g.]{GCB:Menzel2006,RS:Hogda2013,IJB:Karlsen2007} and thereby converging with the period of the ozone spring peak.
 At the same time the growing season is also becoming longer.
-In conjunction with midnight sun conditions (almost $24\,\unit{h}$ photosynthetic activity), a longer growing season, is prolonging the time in which vegetation can accumulate ozone.
-For a natural and semi-natural vegetation already subject to rapid climatic changes, these factors could promote a higher potential risk for northern vegetation to ozone induced damage in the not so distant future.
+A longer growing season is prolonging the time in which vegetation can accumulate ozone.
+For a natural and semi-natural vegetation which is already subject to rapid climatic changes known as Arctic amplification \citep{AMAP2012,IPCC2013}, these factors could promote a higher potential risk for northern vegetation to ozone induced damage in the not so distant future.
 
@@
 The higher the stomatal conductance, the higher the potential for ozone uptake.
-Stomatal conductance has been empirically linked to environmental factors such as air temperature $T$, photosynthetic photon flux density (PPFD), vapor pressure deficit (VPD), and soil water potential (SWP) as well as photosynthesis itself.
+Stomatal conductance has been empirically linked to environmental factors such as air temperature $T$, photosynthetic photon flux density (PPFD), vapor pressure deficit (VPD), and soil water potential (SWP) as well as photosynthesis itself \cite[e.g.]{PTRS:Jarvis1976, BallBerry1987, Emberson2000, ICP:MappingManual2017}.
 Results from open top chamber (OTC) experiments on downy birch (\emph{Betula pubescens}) and mountain birch (\emph{Betula pubescens toruosa}), native to subarctic regions, indicated reductions in both biomass, in root:shoot ratio, and visible leaf damage under elevated ozone treatment ($\left<\chem{[O_3]}\right> = 36-54\,\unit{ppb}$) \citep{Amb:Manninen2009}.
-Although considered to be more ozone tolerant due to an absence of visible injuries \citep{Amb:Girgzdiene2009}, \citet{Amb:Manninen2009} also found chlorophyll:carotenoid ratio and polyamines reductions in Scots pine (\emph{Pinus sylvestris}).\\
+Although considered to be more ozone tolerant due to an absence of visible injuries \citep{Amb:Girgzdiene2009}, \citet{Amb:Manninen2009} also found chlorophyll:carotenoid ratio and polyamines reductions in Scots pine (\emph{Pinus sylvestris}), indicationd a susceptibility to ozone in these species.\\
 
 To measure the risk of damage inflicted to vegetation through ozone, different metrics have been defined.
-Two main types can be differentiated, concentration-based and exposure-based metrics, e.g., Accumulated Ozone Exposure over a threshold of $\mathrm{x}$ ($\mathrm{AOT_x}$) and Phyto-toxic Ozone Dose over a threshold $\mathrm{y}$ ($\mathrm{POD_y}$).
-Due to their simplicity, concentration-based metrics are still widely used, despite comparisons between the two methods (e.g.
-$\mathrm{AOT_{40}}$ with $\mathrm{POD_3}$) show that exposure-based metrics are better suited to represent the actual risk to vegetation from ozone, especially in the more climatically extreme parts of Europe \citep{GCB:Mills2010,ICP:MappingManual2017,iFJuran2018}.
-Most of these previous studies focused on the Mediterranean region.
-
-The concentration-based metric $\mathrm{AOT_x}$ (in units of $\unit{ppb}$) is defined as:
+Two main types can be differentiated, exposure-based and uptake-based metrics, e.g., Accumulated Ozone Exposure over a threshold of $\mathrm{x}$ ($\mathrm{AOT_x}$) and Phyto-toxic Ozone Dose over a threshold $\mathrm{y}$ ($\mathrm{POD_y}$).
+Due to their simplicity, exposure-based metrics are still widely used.
+Comparisons between the two methods (e.g.
+$\mathrm{AOT_{40}}$ with $\mathrm{POD_3}$) show that uptake-based metrics are better suited to represent the actual risk to vegetation from ozone, especially in the climatically extreme parts of Europe \citep{GCB:Mills2010,ICP:MappingManual2017,iF:Juran2018} (e.g., Mediterranean, Baltics/Fennoscandia).
+
+The exposure-based metric $\mathrm{AOT_x}$ (in units of $\unit{ppb}$) is defined as:
 \begin{equation}
@@
 \end{equation}
-with the ambient ozone concentration $\chem{[O_3]}$. The  time over which the integration is calculated varies, but is usually defined from $8\,\unit{am}-8\,\unit{pm}$.
-
-For computing the $\mathrm{POD_y}$, an estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf.
-Details specific to the $\mathrm{DO_3SE}$ model are given in Section~\ref{sec:do3se}.
+with the ambient ozone concentration $\chem{[O_3]}$.
+The threshold exeedence is typically integrated from $8\,\mathrm{am}-8\,\mathrm{pm}$ daily and $3\,\unit{months}$ (May--July) annually \citep{EUdirective2002}.
+
+For computing the $\mathrm{POD_y}$, an estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf and environmental factors affecting stomatal conductance as decribed further in Section~\ref{sec:do3se}.
 $\mathrm{POD_y}$ is then calculated according to:
@@
 
-Within the scope of this study to assess probable effects of conditions unique to northern Fennoscandia on ozone uptake by plants and resultant enhanced susceptibility to ozone inflicted damage, we conducted ground level ozone concentration measurements at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd in the growing seasons of 2018/19 in collaboration with the Norwegian Institute for Air Research (NILU).
-In 2018, we have observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden at Svanhovd.
+Within the scope of this study to assess probable effects of conditions unique to northern Fennoscandia on ozone uptake by plants and the resultant enhanced susceptibility to ozone inflicted damage, we conducted ground level ozone concentration measurements at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd in the growing seasons of 2018/19 in collaboration with the Norwegian Institute for Air Research (NILU).
+In 2018, we observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden at Svanhovd.
 In contrast, no such damage was found on the clovers in 2019, although it was found on the sensitive tobacco cultivar.
 This indicates that the vegetation may have been more affected by ozone in 2018 than in 2019.
-In terms of $\mathrm{AOT_{40}}$, the threshold for protection of vegetation of $3000\,\unit{ppb}$ (integrated over the growing season) given by the EU directive has been well exceeded at many ozone monitoring sites in Southern Norway in the 2018 growing season \citep{NILU2019} but not at Svanhovd.
+In terms of $\mathrm{AOT_{40}}$, the threshold for protection of vegetation of $3000\,\unit{ppb\,h}$ (integrated over the growing season) given by the EU~directive \citep{EUdirective2002} has been well exceeded at many ozone monitoring sites in Southern Norway in the 2018 growing season \citep{NILU2019} but not at Svanhovd.
 
@@
 A total of 2079 forest fires of different size were registered in Norway in 2018, twice as many as in the preceding years 2016/17 \citep{DSB2019}, last accessed April 2020).
-In Sweden, about 500 fires had been reported (five times more than in a usual summer) and an estimated total of $25000\,\unit{hectare}$ burned down in the course of the most extensive fires in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) in younger history \citep{SOU2019}.
-Coincident peak \chem{[O_3]} are found in ozone monitoring data from Svanhovd but also over whole northern Fennoscandia in July (see Fig.~\ref{fig:data_svanvik_2018_2019} or Fig.~\ref{fig:ozone_reconstruction_2018_07}).
-Like in the 2003 drought period, elevated ozone in Europe was likely promoted by a combination of various other factors as well, such as biogenic VOC emissions, reduced cloud and hence increased solar radiation, faster photochemical reactions, and reduced dry deposition and turbulent mixing due to stagnant weather conditions \citep{JGR:Solberg2018}.\\
+In Sweden, about 500 fires had been reported (five times more than in a usual summer) and an estimated total of $25000\,\unit{hectare}$ burned down in the course of what became known as the most extensive fires in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) in younger history \citep{SOU2019}.
+Coincident peak \chem{[O_3]} is found in ozone monitoring data from Svanhovd but also over whole northern Fennoscandia in July (see Fig.~\ref{fig:data_svanvik_2018_2019} or Fig.~\ref{fig:ozone_reconstruction_2018_07}).
+In the 2003 dought periode, elevated ozone in Europe was promoted by a combination of various factors such as reduced cloud (increased solar radiation), reduced dry deposition and turbulent mixing due to the stagnant weather conditions, and increased BVOC emissions \citep{JGR:Solberg2018}.
+It is likely that these factors also contributed in 2018.
 
@@
 In this study, conditions at Svanhovd for the growing seasons 2018/19 shall serve as reference.
-It is important to understand the implications of using these more generic parameterization for more extreme climate regions, here we explore how risk estimates might be over or underestimated using generic vs bespoke subarctic parameterization.
+It is important to understand the implications of using these more generic parameterization for more extreme climate regions.
+Here we explore how risk estimates might be over or underestimated using generic vs bespoke subarctic parameterization.
 we also explore the role \chem{[O_3]} variability and changing climate variables to understand how this risk might alter in the future with the advent of climate change.
 This allows both an understanding of how the threat from ozone may change over time but also which are the key aspects of plant physiology that might determine risk (and hence which parameters might require special attention to get right) both for pollution impact assessment, but also more generally for describing gas exchange response to changing climates that will influence bio-chemical cycling.
-This might be one key to solve long-standing issues of coupled earth system models (ESM) which generally lack PFTs representing subarctic conditions in their land surface model.\\
+This might be one key to solve long-standing issues of coupled earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface model \citep{GMD:Poulter2015,JAMES:Lawrence2019}.\\
 

### 2021-04-08 VegOzone.tex
Checked with Grammaly

@@
 
-Ground level ozone is a highly toxic pollutant known to cause damage to a variety of ecosystems around the World \citep{PT:Emberson2020}.
-At continuously high concentration levels, ozone negatively affects life expectancy of humans \citep{WHO2008} and is a cause of visible injury, photosynthetic damage, early senescence as well as programmed cell death of plants \citep{PCE:Kangasjarvi2005}.
+ground-level ozone is a highly toxic pollutant known to cause damage to a variety of ecosystems around the World \citep{PT:Emberson2020}.
+At continuously high concentration levels, ozone negatively affects the life expectancy of humans \citep{WHO2008} and is a cause of visible injury, photosynthetic damage, early senescence as well as programmed cell death of plants \citep{PCE:Kangasjarvi2005}.
 Annual global yield losses of four major crops (wheat, rice, maize, and soybean) of $3-15\,\unit{\%}$ \citep{PJ:Ainsworth2017} and a suggested loss in primary production in forestry of $7\,\unit{\%}$ \citep{GCB:Wittig2009,EP:Matyssek2012} has been attributed to ozone.
-The crop yield loss studies indicate a threat to food security in rapidly developing countries, e.g., in East and South-East Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018}, where ozone concentrations related to enhanced pollutant emissions from the transport, residential, power generation and industrial sectors have increased in the past few decades.
-Even though long-term (pre 1950) observations of ground level ozone concentrations are scarce; evidence from trend analysis of those few sites suggests that since the industrial revolution, background concentrations in the northern hemisphere have at least doubled and continued to increase until relatively recently \citep[Chapter 2]{IPCC2013}.
-Recent trend analyses indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric ozone started to level-off or even decline in Europe \citep{ESA:Cooper2014, ACP:Wespes2018,ESA:Gaudel2018}.
+The crop yield loss studies indicate a threat to food security in rapidly developing countries, e.g., in East and South-East Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018}, where ozone concentrations related to enhanced pollutant emissions from the transport, residential, power generation, and industrial sectors have increased in the past few decades.
+Even though long-term (pre-1950) observations of ground-level ozone concentrations are scarce; evidence from trend analysis of those few sites suggests that since the industrial revolution, background concentrations in the northern hemisphere have at least doubled and continued to increase until relatively recently \citep[Chapter 2]{IPCC2013}.
+Recent trend analyses indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric ozone started to level off or even decline in Europe \citep{ESA:Cooper2014, ACP:Wespes2018,ESA:Gaudel2018}.
 This can be attributed to a successful implementation of air quality regulations coupled with economic restructuring which saw a move from coal to gas, reducing the number of episodes of peak concentrations especially in summertime over Europe and North America \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
-However, there are indications that an increase in the frequency of heatwave induced thermal stress will reduce the vegetations' capacity for ozone uptake and therefor diminish these achievements \citep{NCC:Lin2020}.\\
-
-Ground level ozone is formed in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) and hydrocarbons known as volatile organic substances (VOCs) in the presence of nitrogen oxides (\chem{NO_x}).
-Despite a relatively short tropospheric life-time of approximately $22\,\unit{days}$, ranging from a few days in the tropical boundary layer to up to $1\,\unit{year}$ in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}, ozone is subject to advection over long distances.
-Episodes of high ozone concentrations in the absence of local pollution sources can therefor often be attributed to long-range transport of both ozone and its precursors.
+However, there are indications that an increase in the frequency of heatwave induced thermal stress will reduce the vegetations' capacity for ozone uptake and therefore diminish these achievements \citep{NCC:Lin2020}.\\
+
+Ground-level ozone is formed in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) and hydrocarbons are known as volatile organic substances (VOCs) in the presence of nitrogen oxides (\chem{NO_x}).
+Despite a relatively short tropospheric lifetime of approximately $22\,\unit{days}$, ranging from a few days in the tropical boundary layer to up to $1\,\unit{year}$ in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}, ozone is subject to advection over long distances.
+Episodes of high ozone concentrations in the absence of local pollution sources can therefore often be attributed to the long-range transport of both ozone and its precursors.
 In northern Fennoscandia, episodes of enhanced ozone concentrations in 2003 and 2006 have been traced back to ozone precursor emissions related to forest fires in southern and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}.
-In addition, hydrocarbons are also emitted by vegetation itself in form of terpenes and monoterpenes, so called biogenic volatile substances (BVOCs).
+Besides, hydrocarbons are also emitted by vegetation itself in form of terpenes and monoterpenes, so-called biogenic volatile substances (BVOCs).
 These emissions are associated with, e.g., thermal stress, defense against herbivores, or even attraction of pollinators \citep{TPS:Penuelas2003}.
 
-In the annual cycle in Boreal and also temperate climates, a distinct maximum of ground level ozone concentrations typically occurs in spring followed by lower concentrations throughout the summer.
+In the annual cycle in Boreal and also temperate climates, a distinct maximum of ground-level ozone concentrations typically occurs in spring followed by lower concentrations throughout the summer.
 Several conditions unique to the Arctic are stimulating the formation of this spring peak.
 Major pathways of ozone and its precursors' removal from the boundary layer include dry deposition on vegetated surfaces and photochemistry \citep{RG:Clifton2020}.
-As dry deposition to bare ground or snow and ice covered surfaces was found to be low \citep{ACP:Helmig2007}, both pathways of removal are substantially suppressed during polar night \citep{AE:Monks2000}.
-This leads to a build up of ozone and its precursors in the course of winter.
+As dry deposition to bare ground or snow and ice-covered surfaces was found to be low \citep{ACP:Helmig2007}, both pathways of removal are substantially suppressed during polar night \citep{AE:Monks2000}.
+This leads to a build-up of ozone and its precursors in winter.
 Come spring precursors are photochemically activated again, while snow cover typically continues to prevail and suppress ozone dry deposition.
-On the other hand, very high ozone concentrations occurring over shorter periods of time can often be attributed to deep tropopause folding events over the Arctic during winter.
+On the other hand, very high ozone concentrations occurring over shorter periods can often be attributed to deep tropopause folding events over the Arctic during winter.
 These events facilitate an intrusion of stratospheric air masses which are enriched in ozone \citep{JGR:Skerlak2015}.
@@
 
-The start of growing season (GS) in northern Fennoscandia has been shifting to earlier dates in response to the general warming trend \citep[e.g.]{GCB:Menzel2006,RS:Hogda2013,IJB:Karlsen2007} and thereby converging with the period of the ozone spring peak.
-At the same time the growing season is also becoming longer.
+The start of the growing season (GS) in northern Fennoscandia has been shifting to earlier dates in response to the general warming trend \citep[e.g.]{GCB:Menzel2006,RS:Hogda2013,IJB:Karlsen2007} and thereby converging with the period of the ozone spring peak.
+At the same time, the growing season is also becoming longer.
 A longer growing season is prolonging the time in which vegetation can accumulate ozone.
-For a natural and semi-natural vegetation which is already subject to rapid climatic changes known as Arctic amplification \citep{AMAP2012,IPCC2013}, these factors could promote a higher potential risk for northern vegetation to ozone induced damage in the not so distant future.
-
-Ozone acts as an oxidative stress to plants.
+For natural and semi-natural vegetation which is already subject to rapid climatic changes known as Arctic amplification \citep{AMAP2012,IPCC2013}, these factors could promote a higher potential risk for northern vegetation to ozone-induced damage in the not so distant future.
+
+Ozone acts as oxidative stress to plants.
 Its main action is imposed through reactions occurring in the cell walls and cell membranes of mesophyll cells inside the leaves.
-Ozone enters the leaves through stomata, that are kept open to allow for gas exchange allowing the entry of \chem{CO_2} for photosynthesis and loss of \chem{H_2O} vapor via the plant transpiration stream.
-Stomatal aperture, and hence stomatal conductance to these gases (including ozone) will vary over the course of the day and growing season primarily to balance \chem{CO_2} uptake against \chem{H_2O} vapor loss.
+Ozone enters the leaves through stomata, which are kept open to allow for gas exchange allowing the entry of \chem{CO_2} for photosynthesis and loss of \chem{H_2O} vapor via the plant transpiration stream.
+Stomatal aperture, and hence stomatal conductance to these gases (including ozone) will vary over the day and growing season primarily to balance \chem{CO_2} uptake against \chem{H_2O} vapor loss.
 The higher the stomatal conductance, the higher the potential for ozone uptake.
 Stomatal conductance has been empirically linked to environmental factors such as air temperature $T$, photosynthetic photon flux density (PPFD), vapor pressure deficit (VPD), and soil water potential (SWP) as well as photosynthesis itself \cite[e.g.]{PTRS:Jarvis1976, BallBerry1987, Emberson2000, ICP:MappingManual2017}.
-Results from open top chamber (OTC) experiments on downy birch (\emph{Betula pubescens}) and mountain birch (\emph{Betula pubescens toruosa}), native to subarctic regions, indicated reductions in both biomass, in root:shoot ratio, and visible leaf damage under elevated ozone treatment ($\left<\chem{[O_3]}\right> = 36-54\,\unit{ppb}$) \citep{Amb:Manninen2009}.
-Although considered to be more ozone tolerant due to an absence of visible injuries \citep{Amb:Girgzdiene2009}, \citet{Amb:Manninen2009} also found chlorophyll:carotenoid ratio and polyamines reductions in Scots pine (\emph{Pinus sylvestris}), indicationd a susceptibility to ozone in these species.\\
+Results from open-top chamber (OTC) experiments on downy birch (\emph{Betula pubescens}) and mountain birch (\emph{Betula pubescens toruosa}), native to subarctic regions, indicated reductions in both biomass, in root:shoot ratio, and visible leaf damage under elevated ozone treatment ($\left<\chem{[O_3]}\right> = 36-54\,\unit{ppb}$) \citep{Amb:Manninen2009}.
+Although considered to be more ozone tolerant due to an absence of visible injuries \citep{Amb:Girgzdiene2009}, \citet{Amb:Manninen2009} also found chlorophyll:carotenoid ratio and polyamines reductions in Scots pine (\emph{Pinus sylvestris}), indicating susceptibility to ozone in these species.\\
 
@@
 
-For computing the $\mathrm{POD_y}$, an estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf and environmental factors affecting stomatal conductance as decribed further in Section~\ref{sec:do3se}.
+For computing the $\mathrm{POD_y}$, an estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf and environmental factors affecting stomatal conductance as described further in Section~\ref{sec:do3se}.
 $\mathrm{POD_y}$ is then calculated according to:
@@
 %
-with the hourly averaged ground level ozone flux through the stomata Eq.~(\ref{eq:flux_stomata}) and a flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}\,s^{-1}}.
+with the hourly averaged ground-level ozone flux through the stomata Eq.~(\ref{eq:flux_stomata}) and a flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}\,s^{-1}}.
 The flux threshold $\mathrm{y}$ is representing the detoxification potential of the plant and is typically only exceeded during daylight hours (i.e.
@@
 
-Within the scope of this study to assess probable effects of conditions unique to northern Fennoscandia on ozone uptake by plants and the resultant enhanced susceptibility to ozone inflicted damage, we conducted ground level ozone concentration measurements at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd in the growing seasons of 2018/19 in collaboration with the Norwegian Institute for Air Research (NILU).
+Within the scope of this study to assess probable effects of conditions unique to northern Fennoscandia on ozone uptake by plants and the resultant enhanced susceptibility to ozone inflicted damage, we conducted ground-level ozone concentration measurements at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd in the growing seasons of 2018/19 in collaboration with the Norwegian Institute for Air Research (NILU).
 In 2018, we observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden at Svanhovd.
@@
 In the north (including Finnmark), it was amongst the warmest on record -- favorable conditions for planetary boundary layer ozone formation.
-An unusually weak and northward shifted jet stream allowed for a persistent high pressure system above northern Europe, including Fennoscandia which blocked the pressure systems for several consecutive months.
+An unusually weak and northward shifted jet stream allowed for a persistent high-pressure system above northern Europe, including Fennoscandia which blocked the pressure systems for several consecutive months.
 In the period May -- July, southern Norway had temperatures $4\,\unit{^\circ C}$ above normal.
 While southern Norway had only about $60\,\unit{\%}$ of normal precipitation, northern Norway as a whole had close to normal precipitation, but with local variations.
-Thermal stress on vegetation was exceptional not only in large parts of Fennoscandia but also in much of Europe, where the influence of the high pressure system extended even over a period of five months (April/May, July--September).
+Thermal stress on vegetation was exceptional not only in large parts of Fennoscandia but also in much of Europe, where the influence of the high-pressure system extended even over five months (April/May, July--September).
 These conditions gave rise to massive forest fires in different parts of Europe and thus an increase in ozone precursors.
 Boreal wildfires emit in addition to \chem{CO_2} also \chem{CO} ($\chem{[CO]}/\chem{[CO_2]} \propto 6-13\,\unit{\%}$) and \chem{VOCs} ($\chem{[VOCs]}/\chem{[CO_2]} \propto 0.5-1.5\,\unit{\%}$) \citep{AE:Cofer1990}.
-A total of 2079 forest fires of different size were registered in Norway in 2018, twice as many as in the preceding years 2016/17 \citep{DSB2019}, last accessed April 2020).
-In Sweden, about 500 fires had been reported (five times more than in a usual summer) and an estimated total of $25000\,\unit{hectare}$ burned down in the course of what became known as the most extensive fires in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) in younger history \citep{SOU2019}.
+A total of 2079 forest fires of different sizes were registered in Norway in 2018, twice as many as in the preceding years 2016/17 \citep{DSB2019}, last accessed April 2020).
+In Sweden, about 500 fires had been reported (five times more than in a usual summer), and an estimated total of $25000\,\unit{hectare}$ burned down in the course of what became known as the most extensive fires in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) in younger history \citep{SOU2019}.
 Coincident peak \chem{[O_3]} is found in ozone monitoring data from Svanhovd but also over whole northern Fennoscandia in July (see Fig.~\ref{fig:data_svanvik_2018_2019} or Fig.~\ref{fig:ozone_reconstruction_2018_07}).
-In the 2003 dought periode, elevated ozone in Europe was promoted by a combination of various factors such as reduced cloud (increased solar radiation), reduced dry deposition and turbulent mixing due to the stagnant weather conditions, and increased BVOC emissions \citep{JGR:Solberg2018}.
-It is likely that these factors also contributed in 2018.
+In the 2003 drought period, elevated ozone in Europe was promoted by a combination of various factors such as reduced cloud (increased solar radiation), reduced dry deposition and turbulent mixing due to the stagnant weather conditions, and increased BVOC emissions \citep{JGR:Solberg2018}.
+It is likely that these factors also contributed to enhanced \chem{[O_3]} in 2018.
 
 The methods described above rely on accurate representation of plant physiological response to climate variables (i.e.
-the species or plant functional type specific responses used to parameterize risk assessment models) and are most often parameterized using data for less extreme climates, i.e., temperate or continental climates; and to a lesser extent, Boreal and Mediterranean climates.
+the species or plant functional type-specific responses used to parameterize risk assessment models) and are most often parameterized using data for less extreme climates, i.e., temperate or continental climates; and to a lesser extent, Boreal and Mediterranean climates.
 Parameterization of these models for more extreme bio-geographical regions such as subarctic climates has not been performed and relies on assuming that less extreme Boreal parameterization will be applicable.
-In this study, conditions at Svanhovd for the growing seasons 2018/19 shall serve as reference.
-It is important to understand the implications of using these more generic parameterization for more extreme climate regions.
+In this study, conditions at Svanhovd for the growing seasons 2018/19 shall serve as a reference.
+It is important to understand the implications of using these more generic parameterizations for more extreme climate regions.
 Here we explore how risk estimates might be over or underestimated using generic vs bespoke subarctic parameterization.
 we also explore the role \chem{[O_3]} variability and changing climate variables to understand how this risk might alter in the future with the advent of climate change.
-This allows both an understanding of how the threat from ozone may change over time but also which are the key aspects of plant physiology that might determine risk (and hence which parameters might require special attention to get right) both for pollution impact assessment, but also more generally for describing gas exchange response to changing climates that will influence bio-chemical cycling.
+This allows both an understanding of how the threat from ozone may change over time but also which are the key aspects of plant physiology that might determine risk (and hence which parameters might require special attention to get right) both for pollution impact assessment, but also more generally for describing gas exchange response to changing climates that will influence biochemical cycling.
 This might be one key to solve long-standing issues of coupled earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface model \citep{GMD:Poulter2015,JAMES:Lawrence2019}.\\
@@
 The observation site at NIBIO Environment Centre Svanhovd and all related data are presented in Section~\ref{sec:data}.
-In Section~\ref{sec:stats}, we derive climatologies for ground level ozone concentration at Svanhovd and northern Fennoscandia based on in situ observations and also compute climatologies of near-surface temperature, precipitation, and global radiation for Svanhovd.
+In Section~\ref{sec:stats}, we derive climatologies for ground-level ozone concentration at Svanhovd and northern Fennoscandia based on in situ observations and also compute climatologies of near-surface temperature, precipitation, and global radiation for Svanhovd.
 Based on these, we evaluate anomalies of environmental key variables in 2018/19 by statistical means.
-In addition, the derived ground level ozone concentration climatology for northern Fennoscandia will be compared with available ozone reanalysis products.
+The derived ground-level ozone concentration climatology for northern Fennoscandia will also be compared with available ground-level ozone reanalysis products.
 We use the $\mathrm{DO_3SE}$ model to estimate the ozone uptake by natural and semi-natural vegetation to investigate main environmental drivers for the observed ozone damage in 2018 (Section~\ref{sec:do3se}).
 In this context, we introduce the$\mathrm{DO_3SE}$ model equations and derive bespoke parameterizations for subarctic species to assess systematic uncertainties arising from the use of default parameterizations.
-In Section~\ref{sec:conc} finally, we will summarize our results, draw conclusions, and give an outlook on further work.
+In Section~\ref{sec:conc} finally, we will summarize our results, conclude, and give an outlook on further work.
 
@@
 Other data used for cross-evaluation of the ozone monitoring data taken at Svanhovd is described in Section~\ref{subsec:other_data}.
-We compare with observational data from other monitoring sites in northern Fennoscandia (Section~\ref{subsubsec:ebas}) and use these to evaluate reanalysis products for ground level ozone (Section~\ref{subsubsec:ozone_rea}).
+We compare with observational data from other monitoring sites in northern Fennoscandia (Section~\ref{subsubsec:ebas}) and use these to evaluate reanalysis products for ground-level ozone (Section~\ref{subsubsec:ozone_rea}).
 
@@
 \label{subsec:other_data}
-To determine the significance of the ozone conditions in 2018, we derive a ground level ozone concentration climatology for Svanvik.
+To determine the significance of the ozone conditions in 2018, we derive a ground-level ozone concentration climatology for Svanvik.
 From 1986--1996 ozone monitoring has been operated by NILU at Svanhovd, but no ozone measurements have been conducted for the past $14\,\unit{years}$.
@@
 
-\subsubsection{Ground level ozone observations}
+\subsubsection{Ground-level ozone observations}
 \label{subsubsec:ebas}
-All ground level ozone observations used in this study are available from the EBAS database operated by \citet{NILU_EBAS}.
+All ground-level ozone observations used in this study are available from the EBAS database operated by \citet{NILU_EBAS}.
 In Table~\ref{tab:ebas_obs}, we give an overview over the exact locations and time range of operational data taking.
-All stations with available long-term observations are located on higher ground than Svanhovd which will result in higher ground level ozone concentrations in comparison as ozone abundance, in general, decreases with decreasing altitude in the absence of local precursors \citep[e.g.,][]{AB:Klingberg2009}.
+All stations with available long-term observations are located on higher ground than Svanhovd which will result in higher ground-level ozone concentrations in comparison as ozone abundance, in general, decreases with decreasing altitude in the absence of local precursors \citep[e.g.,][]{AB:Klingberg2009}.
 The periphery vegetation at all sites is similar and consists mainly of pine forests, birch, and heath.
@@
 The temporal as well as spatial resolution of these reanalysis products is rather coarse: 3-hourly and $0.75^\circ\times 0.75^\circ$ or roughly $29.3\,\unit{km}\times 83.4\,\unit{km}$ at the location of Svanvik, respectively.
-In addition, higher resolution ground level ozone reanalysis for whole Europe is available from Copernicus service for regional air quality control (note: since May 31st 2020 through the CAMS Atmosphere Data Store (ADS)).
+In addition, higher resolution ground-level ozone reanalysis for whole Europe is available from Copernicus service for regional air quality control (note: since May 31st 2020 through the CAMS Atmosphere Data Store (ADS)).
 The regional air quality model reanalysis ensemble (CAMS regional Ensemble) is based on nine European state-of-the-art numerical air quality models \citep{Copernicus_Regio_AIRQ}.
@@
 As representative for the sites, the nearest neighboring grid point has been chosen.
-The ECMWF products do not reproduce observed ground level ozone well.
+The ECMWF products do not reproduce observed ground-level ozone well.
 The MACC reanalysis (Fig.~\ref{fig:ozone_climatologies}a) is in general too low ($(9\pm 7)\,\unit{ppb}$) and displays an erroneous seasonal cycle.
@@
 As representative for the sites, the nearest neighboring grid point has been chosen.
-The ECMWF products do not reproduce observed ground level ozone well.
+The ECMWF products do not reproduce observed ground-level ozone well.
 The MACC reanalysis is in general too low and displays an erroneous seasonal cycle.
@@
 All the above factors make it necessary to look not only at anomalies but statistically assess ozone concentration residuals at Svanvik as well as at Esrange and Pallas in comparison.
-Since the climatology of latter is more reliable, we will start with ground level ozone concentration residuals \chem{\Delta[O_3]} at Esrange and Pallas in 2018.
+Since the climatology of latter is more reliable, we will start with ground-level ozone concentration residuals \chem{\Delta[O_3]} at Esrange and Pallas in 2018.
 As data for 2019 has not yet been available from EBAS (last access, February 2020), we refrain from looking into these.
@@
 
-Because the available reanalysis products for ground level ozone have their limitations, we deem it necessary to reconstruct the missing data in 2018 for the purpose of $\mathrm{DO_3SE}$ modeling by applying a Reynolds's decomposition of the form:
+Because the available reanalysis products for ground-level ozone have their limitations, we deem it necessary to reconstruct the missing data in 2018 for the purpose of $\mathrm{DO_3SE}$ modeling by applying a Reynolds's decomposition of the form:
 \begin{equation}
@@
 We identified $f_\mathrm{T}$ as abnormally low, being the limiting factor of stomatal conductance.
-Therefor, we propose bespoke parameterizations which may describe an actual acclimation of these species to the climatic conditions in the target region.
+Therefore, we propose bespoke parameterizations which may describe an actual acclimation of these species to the climatic conditions in the target region.
 
@@
 For Norway spruce (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
-Norway spruce is the least sensitive to an acclimation in temperature and therefor the least biased.
+Norway spruce is the least sensitive to an acclimation in temperature and therefore the least biased.
 Norway spruce displays a medium sensitivity to the light threshold, e.g.
@@
 
-We found representative ozone climatologies (multi-annual means) for northern Fennoscandia and Svanhovd based on ground level ozone monitoring.
+We found representative ozone climatologies (multi-annual means) for northern Fennoscandia and Svanhovd based on ground-level ozone monitoring.
 To account for missing data from our 2018 record, we proposed a routine for gap filling of ozone observations following a Reynolds's decomposition into anomalies and climatology.
-We also assessed the quality of available global reanalysis (ECMWF MACC and CAM) and regional model reanalysis (Copernicus) ground level ozone concentrations products and conclude that only the Copernicus product can sufficiently reproduce observed ozone concentrations in northern Fennoscandia.
+We also assessed the quality of available global reanalysis (ECMWF MACC and CAM) and regional model reanalysis (Copernicus) ground-level ozone concentrations products and conclude that only the Copernicus product can sufficiently reproduce observed ozone concentrations in northern Fennoscandia.
 
 Although high and more frequent ozone peak concentrations were observed in 2018, ozone concentrations were, in general, not significantly elevated by statistical means.
-It therefor occurs that weather conditions and constantly high ozone concentration are more important in risk assessment using flux-based metrics like $\mathrm{POD_y}$ than episodes of peak concentrations over shorter periods of time.
+It therefore occurs that weather conditions and constantly high ozone concentration are more important in risk assessment using flux-based metrics like $\mathrm{POD_y}$ than episodes of peak concentrations over shorter periods of time.
 Consequently, acute cellular damage caused by very high fluxes over a short period of time is not accounted for properly in $\mathrm{POD_y}$.
@@
 However, the decline of ozone concentrations in arctic spring is partly caused by the removal of ozone through active vegetation at the start of the growing season.
-It is therefor not clear if the peak concentrations of ozone experienced by vegetation in the future with an earlier start of the growing season, will be reduced compared to present day.
+It is therefore not clear if the peak concentrations of ozone experienced by vegetation in the future with an earlier start of the growing season, will be reduced compared to present day.
 Hence, a simulated earlier start of growing season with present day ozone concentration might overestimate ozone uptake and hence ozone induced risk of damage.

### 2021-04-08 VegOzone.tex
Comments by Ane

@@
 
+%%
 \section{Data}
@@
 The foundation of the NIBIO Environment Centre Svanhovd dates back to the 1930s \citep{NIBIO_Svanhovd}.
-The center focuses on research of the special environment of the region, but also comprises long-term observations of various agrometeorological variables, i.a.
+The center focuses on research of the special environment of the region, but also comprises long-term observations of various agrometeorological variables, i.e.
 $2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths.
 The locations of the atmospheric monitoring site and the ozone garden are marked in an aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}a).
-Svanhovd is located south of the settlement of Svanvik, where as the border to Russia, the Pasvik river, lies only a few hundred meters to the east.\\
+Svanhovd is located south of the settlement of Svanvik and the border to Russia, the Pasvik river, lies only a few hundred meters to the east.\\
 
@@
 The plants had to be secured against herbivores with a wire-mesh fence.
-The plants shown are approximately $6-12\,\unit{cm}$ in height.}
+The plants shown are approximately $6-12\,\unit{cm}$ high.}
   \label{fig:svanhovd_research_station}
@@
 An ozone garden was installed in the course of our project at the NIBIO Environment Centre Svanhovd and was likely the northern-most ozone garden to date.
-An ozone garden consists of selected plant species of which one cultivar is more sensitive to ozone than the other.
+An ozone garden consists of selected plant species which are more sensitive to ozone and display visible injuris.
 Cultivated species were, e.g.
-clover, tobacco, and beans.
+clover, tobacco, and potatoes.
 As shown in Fig.~\ref{fig:svanhovd_research_station}b), the plants had to be protected from herbivores with a wire-mesh fence.
-In 2018, we have observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden.
+In 2018, we observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden.
 In contrast, no such damage was found on the clovers in 2019, although it was found on the sensitive tobacco cultivar.
@@
 
-For the purpose of relating observed damage on ozone sensitive plants to local ozone concentrations, an ozone monitor had been installed during the growing seasons 2018/19 in addition to existing monitors of atmospheric key variables (temperature, pressure, precipitation).
-All variables have been obtained through \href{luftkvalitet.no}{luftkvalitet.no} operated by \citet{NILU_AIRQ}.
-Ozone monitoring data has also been added to EBAS database operated by \citet{NILU_EBAS}.
-
+For the purpose of relating observed damage on ozone sensitive plants to local ozone concentrations, an ozone monitor had been installed during the growing seasons 2018/19 in addition to existing monitors of atmospheric key variables.
+All variables were obtained through \href{luftkvalitet.no}{luftkvalitet.no} operated by \citet{NILU_AIRQ}.
+Our ozone monitoring data has also been added to EBAS database operated by \citet{NILU_EBAS} and is thus openly accessable.
 Long-term data sets of agrometeorological variables including temperature and precipitation are available from September 1992 to present day \citep[LandbruksMeteorologiske Tjeneste][note the station name here is Pasvik]{LMT_NIBIO}.
-
 All relevant data for the growing season 2018/19 are shown in Fig.~\ref{fig:data_svanvik_2018_2019}.
@@
  As can be seen in Fig.~\ref{fig:data_svanvik_2018_2019}a), \chem{[O_3]} peaks in spring (April/May) and reaches its minimum in late summer (July/August).
-The spring peak has not been captured completely in 2019, for data acquisition started later than in 2018.
+The spring peak has not been captured completely in 2019, since data acquisition started later than in 2018.
 In summer 2018 (June--August), high ozone concentrations ($\chem{[O_3]} > 40\,\unit{ppb}$) were recorded 50 times.
 The highest summer ozone concentration ($\chem{[O_3]} = 50.2\,\unit{ppb}$) was measured on July 25.
-This coincides with the period of the most extensive forest fires in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) which occurred from July 12--29 and destroyed about $18000\,\unit{hectare}$ of forest \citep{SOU2019}.
+This coincides with the period of the most extensive forest fires in central Sweden which occurred from July 12--29 \citep{SOU2019}.
 However, due to the above mentioned data acquisition problems, we missed most of the corresponding ozone data for this event and most likely also the peak ozone concentration.
 In contrast, ozone concentration only rose 18 times above the threshold of $40\,\unit{ppb}$ during the summer of 2019.
-
-Hourly averaged $2\,\unit{m}$ temperatures below $-20\,\unit{^\circ C}$ were observed in winter (January/February), while high temperatures (above $20\,\unit{^\circ C}$) occurred in summer (July), though more regularly in 2018 than in 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}b).
-In 2018, temperature regularly rose above freezing only in May, while in 2019 this occurred already early in March/April.
-2019 saw a very cold November with temperatures regularly below $-10\unit{^\circ C}$.
-
-More rain events with accumulated daily precipitation $\sum_d \mathrm{Precip}$ above $10\,\unit{mm}$ occurred in the summer of 2018 compared to 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}c).
-$\sum_d \mathrm{Precip} \ge 20\,\unit{mm}$ was only breached twice in 2018 and once in 2019.
-
-In 2018, global irradiance $Q_0$ (Fig.~\ref{fig:data_svanvik_2018_2019}d) displays more sequentially high values in spring (May) and summer (July) compared to 2019, while June 2019 saw more consecutive high irradiance than 2018.
-In both years, the maximum recorded $Q_0$ was roughly $750\,\unit{W\,m^{-2}}$ reached only in June.
+Hourly averaged $2\,\unit{m}$ temperatures above $20\,\unit{^\circ C}$ occurred more regularly in July 2018 than in 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}b).
+In 2018, spring temperature regularly rose above freezing only in May, while in 2019 this occurred already early in March/April.
+More rain events with accumulated daily precipitation ($\sum_d \mathrm{Precip}$) above $10\,\unit{mm}$ occurred in the summer of 2018 compared to 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}c).
+Qualitatively, global irradiance $Q_0$ (Fig.~\ref{fig:data_svanvik_2018_2019}d) was higher in spring (May) and summer (July) 2018 compared to 2019, while June 2019 showed higher irradiance than 2018.
+A high global irradiance in most cases is the results of a low cloud fraction.
+In both years, the maximum recorded $Q_0$ was roughly $750\,\unit{W\,m^{-2}}$ reached in June as naturally expected.
 
@@
   \caption{Observational data from atmospheric monitoring at Svanhovd in 2018/19.
-The hatched areas indicate missing ozone monitoring data.
+The hatched areas indicate periods without ozone monitoring data.
 (a) Hourly averaged ozone concentration; (b) hourly averaged temperature; (c) daily accumulated precipitation; (d) hourly averaged global irradiance.}
@@
 To determine the significance of the ozone conditions in 2018, we derive a ground-level ozone concentration climatology for Svanvik.
-From 1986--1996 ozone monitoring has been operated by NILU at Svanhovd, but no ozone measurements have been conducted for the past $14\,\unit{years}$.
+From 1986--1996 ozone monitoring has been operated continously by NILU at Svanhovd, but no ozone measurements have been conducted until the start of our project.
 Due to probable changes in northern hemispheric tropospheric background ozone and a reduction of summertime peak values related to ozone precursors which fell under air quality regulations in Europe, we will cross-calibrate our data with long-term ozone observations at other sites in northern Fennoscandia.
 In regard to ozone, available observational data are sparse in this region.
-An overview over the geographical locations of conducted atmospheric monitoring in the past and present is given in Fig.~\ref{fig:station_map_fennoscandia}.
+An overview over the geographical locations of conducted ozone monitoring in the past and present is given in Fig.~\ref{fig:station_map_fennoscandia}.
 The data are described in Section~\ref{subsubsec:ebas}.
@@
 The periphery vegetation at all sites is similar and consists mainly of pine forests, birch, and heath.
-However, at close vicinity it varies substantially.
+However, at close vicinity the vegetation varies substantially.
 Ozone measurements at Pallas are taken on top of mount Sammaltunturi $100\,\unit{m}$ above tree line (pine forest) and mainly consists of low vascular plants, mosses, and lichen \citep{BER:Hatakka2003}.
@@
   \end{tabular}
-  \belowtable{$^\dagger$ Exclusive ozone monitoring growing seasons 2018/19 operated by NILU.} % Table Footnotes
+  \belowtable{$^\dagger$ Exclusive ozone monitoring in growing seasons 2018/19 for the present study operated by NILU.} % Table Footnotes
 \end{table}
@@
   \includegraphics[width=12cm]{fig04}
-  \caption{Time series of ozone observations in northern Fennoscandia (Tab.~\ref{tab:ebas_obs}).
+  \caption{Time series of ozone observations in northern Fennoscandia (Tab.~\ref{tab:ebas_obs}) (1986--2019).
 Data taken from EBAS.
-The hatched areas indicate time periods with insufficient quality control \citep{NILU2003}.}
+The hatched areas indicate time periods with insufficient quality control \citep{NILU2003}.
+{\bf TODO: Redo plot with data from 2019 for Esrange and Pallas?}}
   \label{fig:ozone_timesseries_fenoscandic_obs}
@@
 The covered time periods differ and no data are available for before the turn of the millennium.
-The CAMSRA and CAMS Regional Air Quality ensemble mean are available in near real time, but only former covers a period of sufficient length for climate analysis (2003 -- present).
+The CAMSRA and CAMS Regional Air Quality ensemble mean are available in near real time, but only the former covers a period of sufficient length for climate analysis (2003 -- present).
 All reanalyses use different but, at that time, latest versions of the operational weather forecast system (OpenIFS) from ECMWF.

### 2021-04-08 VegOzone.tex
Update ozone monitoring information

@@
     \middlehline
-    Esrange    & SWE     & SE0013R & 67.83           & 21.07             & 475            & 1991 -- present\\
+    Esrange    & SWE     & SE0013R & 67.83           & 21.07             & 475            & 1991 -- 2018$^1$\\
     Janiskoski & RUS     & RU0001R & 68.93           & 28.85             & 118            & 1995 -- 1997\\
@@
     Karasjok   & NOR     & NO0055R & 69.467          & 25.217            & 333            & 1988 -- 1997\\
-    Pallas     & FIN     & FI0096G & 67.97           & 24.12             & 565            & 1995 -- present\\
-    Svanvik    & NOR     & NO0047R & 69.45           & 30.03             & 30             & 1986 -- 1996$^\dagger$\\
+    Pallas     & FIN     & FI0096G & 67.97           & 24.12             & 565            & 1995 -- present$^2$\\
+    Svanvik    & NOR     & NO0047R & 69.45           & 30.03             & 30             & 1986 -- 1996$^3$\\
     \bottomhline
   \end{tabular}
-  \belowtable{$^\dagger$ Exclusive ozone monitoring in growing seasons 2018/19 for the present study operated by NILU.} % Table Footnotes
+  \belowtable{$^1$ Data available until the end of 2018.\\
+    $^2$ Data available until the end of 2019.\\
+    $^3$ Exclusive ozone monitoring in growing seasons 2018/19 for the present study operated by NILU.} % Table Footnotes
 \end{table}
@@
 The hatched areas indicate time periods with insufficient quality control \citep{NILU2003}.
-{\bf TODO: Redo plot with data from 2019 for Esrange and Pallas?}}
+{\bf TODO: updata with recent Pallas data}}
   \label{fig:ozone_timesseries_fenoscandic_obs}

### 2021-04-09 VegOzone.tex
Some more comment by Ane and rewriting of subsec ozone

@@
   \end{tabular}
-  \belowtable{$^1$ Data available until the end of 2018.\\
-    $^2$ Data available until the end of 2019.\\
+  \belowtable{$^1$ Data available on EBAS until the end of 2018.\\
+    $^2$ Data available on EBAS until the end of 2019.\\
     $^3$ Exclusive ozone monitoring in growing seasons 2018/19 for the present study operated by NILU.} % Table Footnotes
@@
   \caption{Time series of ozone observations in northern Fennoscandia (Tab.~\ref{tab:ebas_obs}) (1986--2019).
-Data taken from EBAS.
-The hatched areas indicate time periods with insufficient quality control \citep{NILU2003}.
-{\bf TODO: updata with recent Pallas data}}
+Data available from EBAS.
+The hatched areas indicate time periods with insufficient quality control according to \citet{NILU2003}.}
   \label{fig:ozone_timesseries_fenoscandic_obs}
@@
 In this section, we will evaluate atmospheric conditions in the growing seasons of 2018/19 by statistical means.
-We will derive climatologies for both, temperature, precipitation, global irradiance, and ozone for Svanvik (Section~\ref{subsec:climatologies}).
+We will derive climatologies for temperature, precipitation, global irradiance, and ozone for Svanvik (Section~\ref{subsec:climatologies}).
 The ozone measurements will be put into a climatological as well as regional context.
@@
 The diagram is based on climatological data from Svanvik (1992--2012) downloaded from LandbruksMeteorologiske Tjeneste.
-Monthly averaged temperatures (red line) are displayed with standard error of mean as error band.
+Monthly averaged $2\,\unit{m}$ temperatures (red line) are displayed with standard error of mean (SE) as error band.
+We chose SE in this case due to the large interannual variability in temperature reflected in the standard deviation.
 Averaged monthly accumulated precipitation (blue bars) is shown with standard deviation of mean as error bars.
 Months with average temperatures below freezing are denoted with a star.
-As can be seen, temperatures stay below freezing for 5 consecutive months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for K\"{o}ppens climate classification of a regular subarctic climate (Dfc).
+As can be seen, temperatures stay below freezing for 5 consecutive months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for K\"{o}ppens climate classification of a regular subarctic climate (Dfc) \citep{SD:Beck2018}.
 The highest monthly average temperature is $(13.1\pm 1.1)\,\unit{^\circ C}$ in July and the lowest $(-12.8\pm 2.0)\,\unit{^\circ C}$ in February.
@@
 The average annual accumulated precipitation given with standard error of mean is $(383\pm 86)\,\unit{mm}$.
-Precipitation shows a large variability throughout the years.
+Precipitation shows a large interannual variability.
 
@@
 The multi-annual means of daily mean ozone, for simplicity here without standard deviations, at observation sites in northern Fennoscandia (refer to Table~\ref{tab:ebas_obs} for details) are shown in Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs}.
-All climatologies peak in spring, with peak values of about $48\,\unit{ppb}$ in late April.
+All climatologies except for Svanvik show peak values of about $48\,\unit{ppb}$ in late April.
+The spring peak values at Svanvik are lower because of the much lower altitude compared to the other monitoring sites.
 The annual average ozone concentration \chem{\left<[O_3]\right>} declines throughout the growing season until it reaches its minimum in late summer/fall ($24-30\,\unit{ppb}$).
 The turning points lie approximately in June and November.
-The ozone climatology lags about 3 months behind the temperature climatology.
 At Svanvik, \chem{\left<[O_3]\right>} is $6.6\,\unit{ppb}$ lower than on sites at Jergul/Karasjok (NOR), Esrange (SWE), and Pallas (FIN).
-The large spread in case of Svanvik and Janiskoski can be explained due to the lower overall statistics.
+The large spread in case of Svanvik and Janiskoski is due to the lower number of observations.
 
@@
 
-In Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}, the 2 dimensional density distributions for the thus defined northern Fennoscandia (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}a) and Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) are shown.
+In Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}, the two dimensional density distributions for the combined northern Fennoscandia data (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}a) and Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) are shown.
 Darker colors indicate higher probability to observe these values.
-On top of the density distributions, $10\,\unit{days}$ averages of daily mean \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, daily maximum \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}}, and daily minimum \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, min}} are displayed together with the $1 \sigma$ uncertainties, respectively.
-In addition, \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} is also shown with standard error of mean (albeit small and hardly visible).
+On top of the density distributions, $10\,\unit{days}$ averages of daily mean \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} are displayed together with $1 \sigma$ uncertainties and SE, respectively.
 Splines have been fitted through the data for easier inter-comparison with reanalysis data.
-
 From the northern Fennoscandic climatology, one can deduce several features.
-The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, min}} peaks about $20\,\unit{days}$ earlier than the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, and \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}}.
-With respect to the periodicity of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, it takes about $7.5\,\unit{months}$ from minimum concentration to maximum concentration, while from maximum to minimum it only takes about $4.5\,\unit{months}$.
-The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}} is at $52\,\unit{ppb}$ in late April and the minimum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, min}} occurs at $15\,\unit{ppb}$ in the beginning of September.
-The steep rise of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} in spring is in coincidence with increasing temperatures as well as increase in global irradiance.
-The decrease in \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}} begins about one month ahead of the estimated start of the greening season.
-Based on data from SeNorge.no and the agricultural rule of thumb that the daily average temperature has to be above $5\,\unit{^\circ C}$ for $5\,\unit{days}$ consecutively, we find $\left<G_\mathrm{begin}\right> = (150 \pm 14)\,\unit{days of the year}$ at Svanvik (Supplement Fig~\ref{fig:greening_season_change_Svanvik}), which translates into the end of May.
+The build-up of the ozone spring peak takes about $7.5\,\unit{months}$ from minimum to maximum.
+A sharp increase starts one month ahead of the estimated begin of greening $G_\mathrm{begin}$ (see Section~\ref{subsec:do3se_parameters} for details) and is in coincidence with rising temperatures and global irradiance.
+We find day of the year (\unit{doy}) $\left<G_\mathrm{begin}\right> = (150 \pm 14)$ for Svanvik (Supplement Fig~\ref{fig:greening_season_change_Svanvik}), which translates into the end of May.
 This finding is in line with more extensive studies \citep[][e.g.,]{IJB:Karlsen2007,AFM:Linderholm2006}.
-
-The density distribution for Svanvik and the respective \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, mean}}, and \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}, max}} (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) display a similar pattern as the northern Fennoscandic climatology.
+The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} is reached in late April ($100-120\,\unit{doy}$).
+The depletion of ozone from maximum to minimum takes only $4.5\,\unit{months}$.
+The minimum occurs at $25\,\unit{ppb}$ in the beginning of September ($240-250\,\unit{doy}$).
+
+The density distribution for Svanvik and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) display a similar pattern as the northern Fennoscandic climatology.
 The spring peak apparently occurs slightly earlier (in the beginning of April).
-As Svanvik is located at lower altitude than the other sites, the earlier decline of the ozone peak may point to a stronger ozone uptake by vegetation.
-\chem{CO_2} uptake coniferous trees can occur as early as day of the year (\unit{doy}) 100 \citep{TB:Kolari2007, TP:Wallin2013}.
+As Svanvik is located at lower altitude than the other sites, the earlier decline of the ozone peak may point to a stronger influence of vegetation on the removal of ozone.
+\chem{CO_2} uptake of coniferous trees can occur as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013}.
 The average ozone concentration at Svanvik is lower than at the other sites, as already mentioned, which is in league with a decrease in tropospheric ozone with decreasing altitude.
@@
   \includegraphics[width=8.3cm]{fig06}
-  \caption{Derivation of ozone climatologies.
-Density distributions are shown together with multi-annual mean of daily ozone concentrations \chem{[O_3]}.
+  \caption{Ozone climatologies.
+Fennoscandia combines all available data from Jergul/Karasjok, Esrange, and Pallas.
+The densitiy distributions are shown together with a multi-annual mean of daily ozone concentrations \chem{[O_3]} (climatology).
 Splines have been fitted though daily mean \chem{[O_3]}.

### 2021-04-09 VegOzone.tex
Rename CAMSRA

@@
 Details of reanalysis data sets referred to in this work are listed in Table~\ref{tab:ozone_rea}.
-There are two global reanalysis products available at the European Centre for Medium-Range Weather Forecasts (ECMWF) that include atmospheric tracers, in particular ozone (MACC and CAMSRA) \citep{ACP:Inness2013, ACP:Inness2019}.
-The temporal as well as spatial resolution of these reanalysis products is rather coarse: 3-hourly and $0.75^\circ\times 0.75^\circ$ or roughly $29.3\,\unit{km}\times 83.4\,\unit{km}$ at the location of Svanvik, respectively.
-In addition, higher resolution ground-level ozone reanalysis for whole Europe is available from Copernicus service for regional air quality control (note: since May 31st 2020 through the CAMS Atmosphere Data Store (ADS)).
-The regional air quality model reanalysis ensemble (CAMS regional Ensemble) is based on nine European state-of-the-art numerical air quality models \citep{Copernicus_Regio_AIRQ}.
-The ensemble mean is at fairly high spatial and temporal resolution compared to the global reanalysis from ECMWF: $0.1^\circ\times 0.1^\circ$ or roughly $3.9\,\unit{km}\times 11.1\,\unit{km}$ and 1-hourly, respectively.
-The covered time periods differ and no data are available for before the turn of the millennium.
-The CAMSRA and CAMS Regional Air Quality ensemble mean are available in near real time, but only the former covers a period of sufficient length for climate analysis (2003 -- present).
+There are two global reanalysis products available at the European Centre for Medium-Range Weather Forecasts (ECMWF) that include atmospheric tracers, in particular ozone (MACC and CAMS reanalysis - CAMSRA).
+The CAMS reanalysis and CAMS Regional Air Quality ensemble mean are available in near real time, but only the former covers a period of sufficient length for climate analysis (2003 -- present).
 All reanalyses use different but, at that time, latest versions of the operational weather forecast system (OpenIFS) from ECMWF.
 They differ substantially in the assimilated observational ozone data.
-The MACC reanalysis assimilates only satellite derived tropospheric column ozone (TC), while CAMSRA also includes ozone profiles from satellite retrievals.
+The MACC reanalysis assimilates only satellite derived tropospheric column ozone (TC), while CAMS reanalysis also includes ozone profiles from satellite retrievals.
 Most importantly for our study, in situ observations from ozone near-surface station networks are only assimilated in the CAMS Regional Ensemble.
 The quality with respect to observations in northern Fennoscandia will be discussed in Section~\ref{subsubsec:clim_ozone}.
-Without too much foreshadowing, the results prevented us from fetching the whole CAMSRA data.
+Without too much foreshadowing, the results prevented us from fetching the whole CAMS reanalysis data.
 
@@
 
-The density distribution for Svanvik and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) display a similar pattern as the northern Fennoscandic climatology.
+The density distribution for Svanvik and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} display a similar pattern as the northern Fennoscandic climatology (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b).
 The spring peak apparently occurs slightly earlier (in the beginning of April).
 As Svanvik is located at lower altitude than the other sites, the earlier decline of the ozone peak may point to a stronger influence of vegetation on the removal of ozone.
-\chem{CO_2} uptake of coniferous trees can occur as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013}.
+\chem{CO_2} uptake in coniferous trees can occur as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013}.
 The average ozone concentration at Svanvik is lower than at the other sites, as already mentioned, which is in league with a decrease in tropospheric ozone with decreasing altitude.
@@
   \includegraphics[width=8.3cm]{fig07}
-  \caption{Comparison of climatologies derived from observational data (EBAS) with various ozone reanalysis products: (a) ECMWF: MACC reanalysis; (b) ECMWF: CAMS reanalysis; (c) Copernicus: Regional model reanalysis ensemble mean.
+  \caption{Comparison of climatologies derived from observational data (EBAS) with various ozone reanalysis products: (a) ECMWF: MACC reanalysis; (b) ECMWF: CAMS reanalysis; (c) Copernicus: CAMS Regional model reanalysis ensemble mean.
 As representative for the sites, the nearest neighboring grid point has been chosen.
@@
 The CAMS reanalysis matches better but does not reproduce the seasonality in northern Fennoscandia.
-The ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in northern Fennoscandia, though the modeled ozone concentrations are slightly too low.}
+The ensemble mean of the Copernicus CAMS regional model reanalysis does reproduce the seasonal cycle in northern Fennoscandia, though the modeled ozone concentrations are slightly too low.}
   \label{fig:ozone_climatologies}
@@
 The effective changes from the MACC reanalysis to CAMSRA have been reported and discussed by \citet{ACP:Inness2019} on a global level.
-Amongst others, an assimilation of ozone profiles from satellite retrieval and an upgraded ozone chemistry scheme have lead to the enhanced performance of latter.
+Amongst others, an assimilation of ozone profiles from satellite retrieval and an upgraded ozone chemistry scheme have lead to the enhanced performance of the latter.
 The higher spatial resolution of the regional air quality models used in the CAMS Regional Ensemble compared to the global reanalyses does capture more localized depletion and peak episodes in \chem{[O_3]} better and enhances the capability of the models with respect to daily and seasonal cycles in ozone.
-Latter is, of course, also affected by upgrades in the OpenIFS used as meteorological forcing.
+The latter is, of course, also affected by upgrades in the OpenIFS used as meteorological forcing.
 However, we can assume that the main driver for the strikingly different results are the assimilated observational ozone data.

### 2021-04-10 VegOzone.tex
Reformulate reanalysis part and add that to conclusions.

@@
 We present a reconstruction based on a probability density function (PDF) and concidering the Swedish and Finnish atmospheric monitoring sites.
-We review available ground level ozone reanalysis products for the region and find only CAMS regional Ensemble satisfactory.
+We review available ground level ozone reanalysis products for the region and find only CAMS Regional Air Quality reanalysis ensemble satisfactory.
 In terms of anomalies from a multi-annual mean, ozone concentrations were only significantly enhanced in April/May 2018 before the start of the growing season.
@@
 Details of reanalysis data sets referred to in this work are listed in Table~\ref{tab:ozone_rea}.
-There are two global reanalysis products available at the European Centre for Medium-Range Weather Forecasts (ECMWF) that include atmospheric tracers, in particular ozone (MACC and CAMS reanalysis - CAMSRA).
-The CAMS reanalysis and CAMS Regional Air Quality ensemble mean are available in near real time, but only the former covers a period of sufficient length for climate analysis (2003 -- present).
+There are two global reanalysis products available at the European Centre for Medium-Range Weather Forecasts (ECMWF) that include atmospheric tracers, in particular ozone, the Monitoring Atmospheric Composition and Climate (MACC) and the newer Copernicus Atmosphere Monitoring Service reanalysis (CAMSRA).
+From Copernicus member institutions a CAMS Regional Air Quality (CAMSRAQ) reanalysis ensemble is available for Europe.
+The CAMSRA and CAMSRAQ are available in near real time, but only the former covers a period of sufficient length for climate analysis (2003 -- present).
 All reanalyses use different but, at that time, latest versions of the operational weather forecast system (OpenIFS) from ECMWF.
 They differ substantially in the assimilated observational ozone data.
-The MACC reanalysis assimilates only satellite derived tropospheric column ozone (TC), while CAMS reanalysis also includes ozone profiles from satellite retrievals.
-Most importantly for our study, in situ observations from ozone near-surface station networks are only assimilated in the CAMS Regional Ensemble.
+The MACC reanalysis assimilates only satellite derived tropospheric column ozone, while CAMSRA also includes ozone profiles from satellite retrievals.
+Most importantly for our study, in situ observations from ozone near-surface station networks are only assimilated in the CAMSRAQ reanalysis ensemble.
 The quality with respect to observations in northern Fennoscandia will be discussed in Section~\ref{subsubsec:clim_ozone}.
-Without too much foreshadowing, the results prevented us from fetching the whole CAMS reanalysis data.
+In summary, the results prevented us from using any of the global reanalysis data.
 
 \begin{table}[t]
-  \caption{Ozone related reanalysis products used in this study: \href{https://www.ecmwf.int/en/newsletter/158/meteorology/new-cams-global-reanalysis-atmospheric-composition}{Global reanalysis (ECMWF)}; \href{https://www.regional.atmosphere.copernicus.eu/}{ensemble mean of regional reanalysis for Europe (Copernicus)}.}
+  \caption{Global/refional ozone reanalysis products used in this study.}
   \label{tab:ozone_rea}
@@
 CAMSRA & ECMWF & $0.75^\circ \times 0.75^\circ$ & 3-hourly & 2003 -- 2012 $^\dagger$ & ERA5 / OPS $^\ddagger$ & satellite $^\blacktriangledown$\\
-CAMS Regional Ensemble & Copernicus & $0.1^\circ \times 0.1^\circ$ & 1-hourly & 2014 -- 2018 $^\dagger$ & OPS $^\star$ & in situ $^\vartriangle$\\
+CAMSRAQ & Copernicus & $0.1^\circ \times 0.1^\circ$ & 1-hourly & 2014 -- 2018 $^\dagger$ & OPS $^\star$ & in situ $^\vartriangle$\\
 \bottomhline
@@
 
-Finally, we shall look at the comparison of climatologies derived from observational ozone data and the various reanalysis products (Fig.~\ref{fig:ozone_climatologies}).
+Finally, we shall have a look at the climatologies derived from observational ozone data and the various reanalysis products (Fig.~\ref{fig:ozone_climatologies}).
 As representative for the sites, the nearest neighboring grid point has been chosen.
@@
 Despite displaying only one ozone peak, it still does not reproduce the seasonality in northern Fennoscandia well.
-The CAMSRA spring peak appears to lag behind by $1\,\unit{month}$ and is $5\,\unit{ppb}$ lower, whereas the minimum occurs in January compared to August/September.
+The CAMSRA spring peak lags behind observations by $1\,\unit{month}$ and is $5\,\unit{ppb}$ too low, whereas the minimum occurs in January compared to August/September.
 The annual amplitude ($(26\pm 1)\,\unit{ppb}$) is larger than in the climatology derived from observations ($19\,\unit{ppb}$).
-Modeled \chem{[O_3]} at Svanvik are even slightly higher than at Pallas.
-In contrast, the ensemble mean of the Copernicus regional model reanalysis does reproduce the seasonal cycle in northern Fennoscandia quite well (Fig.~\ref{fig:ozone_climatologies}c), but is on average still slightly too low ($(2.8\pm 0.5)\,\unit{ppb}$).
+The modeled \chem{[O_3]} at Svanvik are slightly higher than at Pallas which might be due to the height at which surface ozone has been evaluated in the model.
+In contrast, the ensemble mean of the CAMSRAQ reproduces the seasonal cycle in northern Fennoscandia quite well (Fig.~\ref{fig:ozone_climatologies}c), but showss on average a slight low-bias ($(2.8\pm 0.5)\,\unit{ppb}$).
 
@@
   \includegraphics[width=8.3cm]{fig07}
-  \caption{Comparison of climatologies derived from observational data (EBAS) with various ozone reanalysis products: (a) ECMWF: MACC reanalysis; (b) ECMWF: CAMS reanalysis; (c) Copernicus: CAMS Regional model reanalysis ensemble mean.
+  \caption{Comparison of climatologies derived from observational data (EBAS) with various ozone reanalysis products: (a) MACC reanalysis; (b) CAMSRA; (c) CAMSRAQ reanalysis ensemble mean.
 As representative for the sites, the nearest neighboring grid point has been chosen.
-The ECMWF products do not reproduce observed ground-level ozone well.
-The MACC reanalysis is in general too low and displays an erroneous seasonal cycle.
-The CAMS reanalysis matches better but does not reproduce the seasonality in northern Fennoscandia.
-The ensemble mean of the Copernicus CAMS regional model reanalysis does reproduce the seasonal cycle in northern Fennoscandia, though the modeled ozone concentrations are slightly too low.}
+The globale products (a), (b) do not reproduce observed ground-level ozone well.
+The MACC reanalysis (a) is in general too low and displays an erroneous seasonal cycle.
+The CAMSRA (b) matches better, but does not reproduce the seasonality in northern Fennoscandia.
+The ensemble mean of CAMSRAQ reanalysis ensemble reproduces the seasonal cycle in northern Fennoscandia, though the modeled ozone concentrations are slightly too low.}
   \label{fig:ozone_climatologies}
@@
 There are many probable explanations for the discovered differences in the climatologies derived from the three reanalyses.
-The effective changes from the MACC reanalysis to CAMSRA have been reported and discussed by \citet{ACP:Inness2019} on a global level.
+The model enhancements from MACC reanalysis to CAMSRA have been reported and discussed by \citet{ACP:Inness2019} on a global level.
 Amongst others, an assimilation of ozone profiles from satellite retrieval and an upgraded ozone chemistry scheme have lead to the enhanced performance of the latter.
-The higher spatial resolution of the regional air quality models used in the CAMS Regional Ensemble compared to the global reanalyses does capture more localized depletion and peak episodes in \chem{[O_3]} better and enhances the capability of the models with respect to daily and seasonal cycles in ozone.
+The higher spatial resolution of the regional air quality models used in the CAMSRAQ ensemble compared to the global reanalyses captures small scale depletion and peak episodes in \chem{[O_3]} better and enhances the capability of the models with respect to daily and seasonal cycles in ozone.
 The latter is, of course, also affected by upgrades in the OpenIFS used as meteorological forcing.
 However, we can assume that the main driver for the strikingly different results are the assimilated observational ozone data.
-Satellite retrievals resolve \chem{[O_3]} at the surface rather poorly and hence do not constrain the global models well enough \citet{ACP:Andersson2017}.
+Satellite retrievals resolve \chem{[O_3]} at the surface rather poorly and hence do not constrain the global models well enough \citep{ACP:Andersson2017}.
 
@@
 \label{sec:conc}
+We have studied the ozone seasonality for northern Fennoscandia and the quality of its representation in available global and regional reanalysis products for Europe.
+We found that both global reanalysis products of ECWMF fail to reproduce observed ozone paterns.
+Only the CAMSRAQ ensemble reproduces the observed ozone seasonality well, although with a slight low-bias remaining.
+This confirms that high spartial resosultion and assimilation of in situ observations of ground-level ozone concentrations are a must to constrain reanalysis related to air quality and risk assement ozone damage to vegetation, especially in high latitudes.
+
 We have exemplarily studied the effect of acclimations of biomes to a subarctic climate on phyto-toxic ozone uptake.

### 2021-04-10 VegOzone.tex
Last comment so far by Ane

@@
 \label{subsubsec:anomal_tpq}
-In Fig.~\ref{fig:anomalies_svanvik}a), temperature anomalies at Svanvik are shown as percentage of days warmer/colder than $\pm 1\,\sigma$ from the climatological mean for each month.
+In Fig.~\ref{fig:anomalies_svanvik}a), temperature anomalies at Svanvik are shown as percentage of days warmer or colder than $\pm 1\,\sigma$ from the climatological mean for each month.
 Negative deviations from climatology are displayed as negative percentage.
@@
 The summer of 2018 was significantly warmer than average.
-Especially in May and July more than $40\,\unit{\%}$ of days were significantly warmer than normal temperature.
-Temperatures continued to be significantly warmer throughout fall into early winter (August--November) with the exception of October.
-March 2018 was unusually cold.
-The summer of 2019 was fairly average, however significantly colder in July, while August was both significantly colder and warmer.
-Temperatures in April were significantly warmer than average.
+Especially in May and July more than $40\,\unit{\%}$ of the days were significantly warmer.
+Significantly warmer days continued to occure in all months (August--November) with the exception of October.
+March 2018 had many unusually cold days.
+The summer of 2019 was fairly average, however significantly colder days in July occured, while August had both significantly colder and warmer days.
+April 2019 had significantly warmer than average days.
 
@@
   \includegraphics[width=12cm]{fig08}
-  \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as percentages of days deviating from climatological mean by means of standard deviation for each month.
+  \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as percentages of days significantly deviating from climatological mean by means of standard deviation for each month.
 Negative deviations from climatology are shown as negative percentage.
@@
 The hatched area between the dashed lines indicates statistical significance under the assumption of a normal distribution ($15.9\,\unit{\%}$, $30.8\,\unit{\%}$ for Precip.).
-(a) Temperature, days warmer/colder than $\pm 1\,\sigma$; (b) precipitation, days wetter/drier than $\pm \frac{1}{2}\,\sigma$; (c) global irradiance, days lighter/darker than $\pm 1\,\sigma$; (d) ozone, days higher/lower than $\pm 1\,\sigma$.
-The July 2018 ozone data corrected for data acquisition problems is indicated with the star (refer Section~\ref{subsec:ozone_reco} for details).
-The growing season of 2018 was significantly warmer, brighter, and slightly drier than average, especially in May and July.
-Ozone was significantly elevated in May.
-The growing season of 2019 was fairly average but significantly warmer than average in April and colder in July/August.
-No significant elevation in ozone was found.}
+(a) Temperature; (b) precipitation; (c) global irradiance; (d) ozone.}
   \label{fig:anomalies_svanvik}
@@
 
-%\begin{figure}[t]
-  %\includegraphics[width=8.3cm]{temperature_signific_1}
-  %\includegraphics[width=8.3cm]{fig08}
-  %\caption{Temperature anomalies at Svanvik.
-Percentage of days warmer/colder than $\pm 1\,\sigma$ from climatological mean displayed for each month.
-Negative deviations from climatology are shown as negative percentage.
-The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
-The summer of 2018 was significantly warmer than average, especially May and July.
-The summer of 2019 was fairly average but significantly warmer than average in April and colder in July/August.}
-  %\label{fig:plot_temperature_anomalies_svanvik}
-%\end{figure}
-
 In a similar way, the precipitation anomalies at Svanvik are displayed in Fig.~\ref{fig:anomalies_svanvik}b).
-However, the percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ from the climatological mean is shown for each month, due to the larger variability in observed precipitation too few days showed a significance of $1 \sigma$ or higher.
-As above, negative deviations from the climatology are shown as negative percentage and annual positive/negative deviations are indicated in the respective corners.
+Due to a large interannual variability in observed precipitation, only a few days deviated significantly from the climatology on the $1 \sigma$ level.
+Therefore, we lowered the $\sigma$ level constraint.
+The percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ from the climatological mean is shown for each month.
+Negative deviations from the climatology are shown as negative percentage and annual positive/negative deviations are indicated in the respective corners.
 The hatched area between the dashed lines indicates the expected percentage of values falling above $\pm\frac{1}{2}\,\sigma$ if a normal distribution is assumed ($30.8\,\unit{\%}$).
 Unlike for the temperature anomalies, the picture is not as clear.
-While March 2018 was unusually wet, summer and fall (May/July, September/October) were significantly drier than average.
-2019 was rather average throughout the year, with drier than usual fall (September/October) and a wet December.
-
-%\begin{figure}[t]
-  %\includegraphics[width=8.3cm]{precipitation_signific_1}
-  %\includegraphics[width=8.3cm]{fig09}
-  %\caption{Precipitation anomalies at Svanvik.
-Percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ from climatological mean displayed for each month.
-Negative deviations from climatology are shown as negative percentage.
-The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The hatched area between the dashed lines indicates the expected percentage of values falling above $\pm\frac{1}{2}\,\sigma$ if a normal distribution is assumed ($30.8\,\unit{\%}$).
-Summer was significantly dry only in 2018, while fall was significantly dry in both years.}
-  %\label{fig:plot_precipitation_anomalies_svanvik}
-%\end{figure}
+While March 2018 had a significant number of days wetter than unusual, summer and fall (May/July, September/October) had a significantly number of days drier than average.
+Overall, 2019 was rather average throughout the year, but with a significant number of days  drier than normal (September/October) and a wetter December.
 
 Similarly, the global irradiance anomalies at Svanvik are presented in Fig.~\ref{fig:anomalies_svanvik}c).
-The percentage of days above/below $\pm 1\,\sigma$ from the climatological mean is displayed for each month.
+The percentage of days above or below $\pm 1\,\sigma$ from the climatological mean is displayed for each month.
 The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed.
-In summer 2018 (May/July), global irradiance was significantly higher than on average, while 2019 was generally an average year with a slightly darker spring (April/May).
-
-%\begin{figure}[t]
-  %\includegraphics[width=8.3cm]{global_rad_signific_1}
-  %\includegraphics[width=8.3cm]{fig10}
-  %\caption{Global irradiance anomalies at Svanvik.
-Percentage of days above/below $\pm 1\,\sigma$ from climatological mean displayed for each month.
-Negative deviations from climatology are shown as negative percentage.
-The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
-In summer 2018 (May/July), global irradiance was significantly higher than on average, while 2019 had significantly less than normal irradiance in April/May and more in June.}
-  %\label{fig:global_rad_signific}
-%\end{figure}
-
-In summary, 2018 was significantly warmer, drier, and lighter than the climatological average, while 2019 was a rather average year.
+In summer 2018 (May/July), global irradiance showed a significant number of days with higher than average irradiance, while 2019 was generally an average year with a significant number of days darker in spring (April/May).
+
+In summary, 2018 had a significant number of days which were warmer, drier, and brighter than the climatological average, while 2019 was a rather average year.
 
@@
 
-If the data collected at Svanvik had been complete in 2018 and the climatology had extended into the resent past, we could have assessed the significance of ozone anomalies in the same way as for temperature, precipitation, and global irradiance.
+If the data collected at Svanvik had been complete in 2018 and the climatology had extended into the recent past, we could have assessed the significance of ozone anomalies in the same way as for temperature, precipitation, and global irradiance.
 In Section~\ref{subsubsec:clim_ozone}, we described the construction of climatologies for northern Fennoscandia based on long-term observations at Esrange, Jergul/Karasjok, and Pallas.

### 2021-04-10 VegOzone.tex
Checked with Grammarly

@@
 In our assessment of primary drivers of ozone damage risk on vegetation in the subarctic, we focus on the 2018 growing season which was subject to an extended heatwave and its subsequent growing season of 2019.
-The NIBIO operated Environment Centre Svanhovd which lies close by the settlement of Svanvik in the Pasvik valley serves as main reference.
+The NIBIO operated Environment Centre Svanhovd which lies close by the settlement of Svanvik in the Pasvik valley serves as the main reference.
 Throughout the paper, we will use Svanhovd synonymously for the atmospheric monitoring site at Svanvik/Pasvik.
@@
 The foundation of the NIBIO Environment Centre Svanhovd dates back to the 1930s \citep{NIBIO_Svanhovd}.
-The center focuses on research of the special environment of the region, but also comprises long-term observations of various agrometeorological variables, i.e.
+The center focuses on the research of the special environment of the region, but also comprises long-term observations of various agrometeorological variables, i.e.
 $2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths.
-The locations of the atmospheric monitoring site and the ozone garden are marked in an aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}a).
+The locations of the atmospheric monitoring site and the ozone garden are marked in aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}a).
 Svanhovd is located south of the settlement of Svanvik and the border to Russia, the Pasvik river, lies only a few hundred meters to the east.\\
@@
 An ozone garden was installed in the course of our project at the NIBIO Environment Centre Svanhovd and was likely the northern-most ozone garden to date.
-An ozone garden consists of selected plant species which are more sensitive to ozone and display visible injuris.
+An ozone garden consists of selected plant species that are more sensitive to ozone and display visible injuries.
 Cultivated species were, e.g.
@@
 
-For the purpose of relating observed damage on ozone sensitive plants to local ozone concentrations, an ozone monitor had been installed during the growing seasons 2018/19 in addition to existing monitors of atmospheric key variables.
+To relate observed damage on ozone sensitive plants to local ozone concentrations, an ozone monitor had been installed during the growing seasons 2018/19 in addition to existing monitors of atmospheric key variables.
 All variables were obtained through \href{luftkvalitet.no}{luftkvalitet.no} operated by \citet{NILU_AIRQ}.
-Our ozone monitoring data has also been added to EBAS database operated by \citet{NILU_EBAS} and is thus openly accessable.
-Long-term data sets of agrometeorological variables including temperature and precipitation are available from September 1992 to present day \citep[LandbruksMeteorologiske Tjeneste][note the station name here is Pasvik]{LMT_NIBIO}.
+Our ozone monitoring data has also been added to the EBAS database operated by \citet{NILU_EBAS} and is thus openly accessible.
+Long-term data sets of agrometeorological variables including temperature and precipitation are available from September 1992 to the present day \citep[LandbruksMeteorologiske Tjeneste][note the station name here is Pasvik]{LMT_NIBIO}.
 All relevant data for the growing season 2018/19 are shown in Fig.~\ref{fig:data_svanvik_2018_2019}.
 Ozone concentrations \chem{[O_3]} measured in $2\,\unit{m}$ height above ground are averaged hourly.
-The hatched areas mark times when no ozone data was recorded.
+The hatched areas mark times when no ozone data were recorded.
 Note, while the downtime during winter was planned, missing data in two weeks of July 2018 (July 9--23) were due to problems in data acquisition.
@@
 This coincides with the period of the most extensive forest fires in central Sweden which occurred from July 12--29 \citep{SOU2019}.
-However, due to the above mentioned data acquisition problems, we missed most of the corresponding ozone data for this event and most likely also the peak ozone concentration.
+However, due to the above-mentioned data acquisition problems, we missed most of the corresponding ozone data for this event and most likely also the peak ozone concentration.
 In contrast, ozone concentration only rose 18 times above the threshold of $40\,\unit{ppb}$ during the summer of 2019.
@@
 Qualitatively, global irradiance $Q_0$ (Fig.~\ref{fig:data_svanvik_2018_2019}d) was higher in spring (May) and summer (July) 2018 compared to 2019, while June 2019 showed higher irradiance than 2018.
-A high global irradiance in most cases is the results of a low cloud fraction.
+A high global irradiance in most cases is the result of a low cloud fraction.
 In both years, the maximum recorded $Q_0$ was roughly $750\,\unit{W\,m^{-2}}$ reached in June as naturally expected.
@@
 To determine the significance of the ozone conditions in 2018, we derive a ground-level ozone concentration climatology for Svanvik.
-From 1986--1996 ozone monitoring has been operated continously by NILU at Svanhovd, but no ozone measurements have been conducted until the start of our project.
+From 1986--1996 ozone monitoring has been operated continuously by NILU at Svanhovd, but no ozone measurements have been conducted until the start of our project.
 Due to probable changes in northern hemispheric tropospheric background ozone and a reduction of summertime peak values related to ozone precursors which fell under air quality regulations in Europe, we will cross-calibrate our data with long-term ozone observations at other sites in northern Fennoscandia.
-In regard to ozone, available observational data are sparse in this region.
-An overview over the geographical locations of conducted ozone monitoring in the past and present is given in Fig.~\ref{fig:station_map_fennoscandia}.
+Concerning ozone, available observational data are sparse in this region.
+An overview of the geographical locations of conducted ozone monitoring in the past and present is given in Fig.~\ref{fig:station_map_fennoscandia}.
 The data are described in Section~\ref{subsubsec:ebas}.
-As additional complement to observational data which are not available at all times and spatially sparse, we also evaluate various ozone reanalysis products (Section~\ref{subsubsec:ozone_rea}).
+As an additional complement to observational data which are not available at all times and spatially sparse, we also evaluate various ozone reanalysis products (Section~\ref{subsubsec:ozone_rea}).
 
@@
   \caption{Cap of the North.
-Locations of past and present ozone observation sites in northern Fennoscandia used within this study.
+Locations of past and present ozone observation sites in northern Fennoscandia were used in this study.
 For more details see Tab.~\ref{tab:ebas_obs}.
-The same color coding for each station as displayed herein will be used in all following figures.}
+The same color coding for each station as displayed herein will be used in all the following figures.}
   \label{fig:station_map_fennoscandia}
@@
 All ground-level ozone observations used in this study are available from the EBAS database operated by \citet{NILU_EBAS}.
-In Table~\ref{tab:ebas_obs}, we give an overview over the exact locations and time range of operational data taking.
+In Table~\ref{tab:ebas_obs}, we give an overview of the exact locations and time range of operational data taking.
 All stations with available long-term observations are located on higher ground than Svanhovd which will result in higher ground-level ozone concentrations in comparison as ozone abundance, in general, decreases with decreasing altitude in the absence of local precursors \citep[e.g.,][]{AB:Klingberg2009}.
 The periphery vegetation at all sites is similar and consists mainly of pine forests, birch, and heath.
-However, at close vicinity the vegetation varies substantially.
-Ozone measurements at Pallas are taken on top of mount Sammaltunturi $100\,\unit{m}$ above tree line (pine forest) and mainly consists of low vascular plants, mosses, and lichen \citep{BER:Hatakka2003}.
+However, in close vicinity, the vegetation varies substantially.
+Ozone measurements at Pallas are taken on top of mount Sammaltunturi $100\,\unit{m}$ above treeline (pine forest) and mainly consists of low vascular plants, mosses, and lichen \citep{BER:Hatakka2003}.
 At Svanhovd, the ozone measurements are taken amidst a meadow surrounded by agriculturally managed herbaceous plants and grass (c.c.
@@
 Data from before the 1990s are available solely from Svanvik and Karasjok.
-These data, however, do not follow the high quality standard procedures implemented nowadays and have to be treated with care \citep{NILU2003}.
-Especially, ozone monitors did not undergo a regular re-calibration.
+These data, however, do not follow the high-quality standard procedures implemented nowadays and have to be treated with care \citep{NILU2003}.
+Especially, ozone monitors did not undergo regular re-calibration.
 This probably leads to drifts in the observed data and may impose a false periodicity.
-In 1997, the ozone monitoring at Jergul was relocated downstream the river Karasjokka to a site close by Karasjok.
-Although the same equipment has been used, this might have introduced a breaking point in the combined data series, denoted as Jergul/Karasjok in the following, which would be of a problem if this was the sole data to compare with.
-Only few data have been taken upstream the Pasvik river at Janiskovski, hence these will not be taken into further consideration, but are shown for completion.
+In 1997, the ozone monitoring at Jergul was relocated downstream of the river Karasjokka to a site close by Karasjok.
+Although the same equipment has been used, this might have introduced a breaking point in the combined data series, denoted as Jergul/Karasjok in the following, which would be a problem if this was the sole data to compare with.
+Only a few data have been taken upstream the Pasvik river at Janiskovski, hence these will not be taken into further consideration, but are shown for completeness.
 As can be seen in Fig.~\ref{fig:ozone_timesseries_fenoscandic_obs}, there is an overlap of about $4\,\unit{years}$ between data taken at Esrange, Jergul/Karasjok, and Svanvik.
-The hatched areas indicate time periods with insufficient quality control as mentioned above.
+The hatched areas indicate periods with insufficient quality control as mentioned above.
 
@@
   \caption{Time series of ozone observations in northern Fennoscandia (Tab.~\ref{tab:ebas_obs}) (1986--2019).
-Data available from EBAS.
-The hatched areas indicate time periods with insufficient quality control according to \citet{NILU2003}.}
+Data are available from EBAS.
+The hatched areas indicate periods with insufficient quality control according to \citet{NILU2003}.}
   \label{fig:ozone_timesseries_fenoscandic_obs}
@@
 The CAMSRA and CAMSRAQ are available in near real time, but only the former covers a period of sufficient length for climate analysis (2003 -- present).
-All reanalyses use different but, at that time, latest versions of the operational weather forecast system (OpenIFS) from ECMWF.
+All reanalyses use different but, at that time, the latest versions of the operational weather forecast system (OpenIFS) from ECMWF.
 They differ substantially in the assimilated observational ozone data.
-The MACC reanalysis assimilates only satellite derived tropospheric column ozone, while CAMSRA also includes ozone profiles from satellite retrievals.
+The MACC reanalysis assimilates only satellite-derived tropospheric column ozone, while CAMSRA also includes ozone profiles from satellite retrievals.
 Most importantly for our study, in situ observations from ozone near-surface station networks are only assimilated in the CAMSRAQ reanalysis ensemble.
-The quality with respect to observations in northern Fennoscandia will be discussed in Section~\ref{subsubsec:clim_ozone}.
+The quality concerning observations in northern Fennoscandia will be discussed in Section~\ref{subsubsec:clim_ozone}.
 In summary, the results prevented us from using any of the global reanalysis data.
@@
 \belowtable{$^\dagger$ Subset of the data used in this study.
-The full data set extends into near real time; $^\ddagger$ ERA5 (2003--2016), OPS (later); $^\star$ EURAD uses WRF for downscaling of operational IFS; $^\triangledown$ MLS, OMI - tropospheric column; $^\blacktriangledown$ SCIAMACHY, MIPAS, MLS, OMI, GOME2, SBUV2 - tropospheric column + profile; $^\vartriangle$ METEO France NRT.} % Table Footnotes
+The full data set extends into near real-time; $^\ddagger$ ERA5 (2003--2016), OPS (later); $^\star$ EURAD uses WRF for downscaling of operational IFS; $^\triangledown$ MLS, OMI - tropospheric column; $^\blacktriangledown$ SCIAMACHY, MIPAS, MLS, OMI, GOME2, SBUV2 - tropospheric column + profile; $^\vartriangle$ METEO France NRT.} % Table Footnotes
 \end{table}
@@
 The location of Svanvik at $69.45\,\unit{^\circ N}$ suggests a subarctic climate.
-The derived climate diagram following Walter and Lieth shown in Fig.~\ref{fig:climatediagram} supports this very well.
+The derived climate diagram following Walter and Lieth which is shown in Fig.~\ref{fig:climatediagram} supports this very well.
 The diagram is based on climatological data from Svanvik (1992--2012) downloaded from LandbruksMeteorologiske Tjeneste.
-Monthly averaged $2\,\unit{m}$ temperatures (red line) are displayed with standard error of mean (SE) as error band.
+Monthly averaged $2\,\unit{m}$ temperatures (red line) are displayed with a standard error of the mean (SE) as error band.
 We chose SE in this case due to the large interannual variability in temperature reflected in the standard deviation.
@@
 The highest monthly average temperature is $(13.1\pm 1.1)\,\unit{^\circ C}$ in July and the lowest $(-12.8\pm 2.0)\,\unit{^\circ C}$ in February.
-In the time period 1992--2012, the coldest measured temperature was $-45.2\,\unit{^\circ C}$ observed January 27, 1999, while the highest temperature on July 16 the same year was $29.4\,\unit{^\circ C}$.
+In the period 1992--2012, the coldest measured temperature was $-45.2\,\unit{^\circ C}$ observed January 27, 1999, while the highest temperature on July 16 the same year was $29.4\,\unit{^\circ C}$.
 
@@
 The diagram is based on climatological data for Svanvik/Pasvik (1992--2012) downloaded from LandbruksMeteorologiske Tjeneste.
-Monthly averaged temperatures (red line) are displayed with standard error of mean as error band.
-Averaged monthly accumulated precipitation (blue bars) is shown with standard deviation of mean as error bars.
+Monthly averaged temperatures (red line) are displayed with standard error of the mean as error band.
+Averaged monthly accumulated precipitation (blue bars) is shown with a standard deviation of mean as error bars.
 Months with average temperatures below freezing are denoted with a star.}
@@
 At Svanvik, \chem{\left<[O_3]\right>} is $6.6\,\unit{ppb}$ lower than on sites at Jergul/Karasjok (NOR), Esrange (SWE), and Pallas (FIN).
-The large spread in case of Svanvik and Janiskoski is due to the lower number of observations.
+The large spread in the case of Svanvik and Janiskoski is due to the lower number of observations.
 
@@
 Splines have been fitted through the data for easier inter-comparison with reanalysis data.
-From the northern Fennoscandic climatology, one can deduce several features.
+From the northern Fennoscandia climatology, one can deduce several features.
 The build-up of the ozone spring peak takes about $7.5\,\unit{months}$ from minimum to maximum.
@@
 
-The density distribution for Svanvik and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} display a similar pattern as the northern Fennoscandic climatology (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b).
-The spring peak apparently occurs slightly earlier (in the beginning of April).
-As Svanvik is located at lower altitude than the other sites, the earlier decline of the ozone peak may point to a stronger influence of vegetation on the removal of ozone.
+The density distribution for Svanvik and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} display a similar pattern as the northern Fennoscandia climatology (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b).
+The spring peak occurs slightly earlier (at the beginning of April).
+As Svanvik is located at a lower altitude than the other sites, the earlier decline of the ozone peak may point to a stronger influence of vegetation on the removal of ozone.
 \chem{CO_2} uptake in coniferous trees can occur as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013}.
 The average ozone concentration at Svanvik is lower than at the other sites, as already mentioned, which is in league with a decrease in tropospheric ozone with decreasing altitude.
-In July--September, ozone is occasionally almost completely depleted during night time.
+In July--September, ozone is occasionally almost completely depleted during nighttime.
 
@@
 Fennoscandia combines all available data from Jergul/Karasjok, Esrange, and Pallas.
-The densitiy distributions are shown together with a multi-annual mean of daily ozone concentrations \chem{[O_3]} (climatology).
-Splines have been fitted though daily mean \chem{[O_3]}.
+The density distributions are shown together with a multi-annual mean of daily ozone concentrations \chem{[O_3]} (climatology).
+Splines have been fitted through daily mean \chem{[O_3]}.
 The climatology of daily mean ozone is shown together with $1\,\sigma$ standard deviation and standard error of the mean.
@@
 The MACC reanalysis (Fig.~\ref{fig:ozone_climatologies}a) is in general too low ($(9\pm 7)\,\unit{ppb}$) and displays an erroneous seasonal cycle.
-Ozone concentrations peak already in March followed by a second peak in July.
+Ozone concentrations peak already in March followed by the second peak in July.
 Also CAMSRA matches the observed ozone climatology poorly (Fig.~\ref{fig:ozone_climatologies}b).
@@
 The modeled \chem{[O_3]} at Svanvik are slightly higher than at Pallas which might be due to the height at which surface ozone has been evaluated in the model.
-In contrast, the ensemble mean of the CAMSRAQ reproduces the seasonal cycle in northern Fennoscandia quite well (Fig.~\ref{fig:ozone_climatologies}c), but showss on average a slight low-bias ($(2.8\pm 0.5)\,\unit{ppb}$).
+In contrast, the ensemble mean of the CAMSRAQ reproduces the seasonal cycle in northern Fennoscandia quite well (Fig.~\ref{fig:ozone_climatologies}c), but shows on average a slight low-bias ($(2.8\pm 0.5)\,\unit{ppb}$).
 
@@
   \caption{Comparison of climatologies derived from observational data (EBAS) with various ozone reanalysis products: (a) MACC reanalysis; (b) CAMSRA; (c) CAMSRAQ reanalysis ensemble mean.
-As representative for the sites, the nearest neighboring grid point has been chosen.
-The globale products (a), (b) do not reproduce observed ground-level ozone well.
+As the representative for the sites, the nearest neighboring grid point has been chosen.
+The global products (a), (b) do not reproduce observed ground-level ozone well.
 The MACC reanalysis (a) is in general too low and displays an erroneous seasonal cycle.
-The CAMSRA (b) matches better, but does not reproduce the seasonality in northern Fennoscandia.
+The CAMSRA (b) matches better but does not reproduce the seasonality in northern Fennoscandia.
 The ensemble mean of CAMSRAQ reanalysis ensemble reproduces the seasonal cycle in northern Fennoscandia, though the modeled ozone concentrations are slightly too low.}
@@
 The model enhancements from MACC reanalysis to CAMSRA have been reported and discussed by \citet{ACP:Inness2019} on a global level.
-Amongst others, an assimilation of ozone profiles from satellite retrieval and an upgraded ozone chemistry scheme have lead to the enhanced performance of the latter.
-The higher spatial resolution of the regional air quality models used in the CAMSRAQ ensemble compared to the global reanalyses captures small scale depletion and peak episodes in \chem{[O_3]} better and enhances the capability of the models with respect to daily and seasonal cycles in ozone.
+Amongst others, assimilation of ozone profiles from satellite retrieval and an upgraded ozone chemistry scheme has lead to the enhanced performance of the latter.
+The higher spatial resolution of the regional air quality models used in the CAMSRAQ ensemble compared to the global reanalyses captures small scale depletion and peak episodes in \chem{[O_3]} better and enhances the capability of the models concerning daily and seasonal cycles in ozone.
 The latter is, of course, also affected by upgrades in the OpenIFS used as meteorological forcing.
-However, we can assume that the main driver for the strikingly different results are the assimilated observational ozone data.
+However, we can assume that the main driver for the strikingly different results is the assimilated observational ozone data.
 Satellite retrievals resolve \chem{[O_3]} at the surface rather poorly and hence do not constrain the global models well enough \citep{ACP:Andersson2017}.
@@
 \label{subsubsec:anomal_tpq}
-In Fig.~\ref{fig:anomalies_svanvik}a), temperature anomalies at Svanvik are shown as percentage of days warmer or colder than $\pm 1\,\sigma$ from the climatological mean for each month.
-Negative deviations from climatology are displayed as negative percentage.
+In Fig.~\ref{fig:anomalies_svanvik}a), temperature anomalies at Svanvik are shown as a percentage of days warmer or colder than $\pm 1\,\sigma$ from the climatological mean for each month.
+Negative deviations from climatology are displayed as a negative percentage.
 The full annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
@@
 Especially in May and July more than $40\,\unit{\%}$ of the days were significantly warmer.
-Significantly warmer days continued to occure in all months (August--November) with the exception of October.
+Significantly warmer days continued to occur in all months (August--November) except for October.
 March 2018 had many unusually cold days.
-The summer of 2019 was fairly average, however significantly colder days in July occured, while August had both significantly colder and warmer days.
+The summer of 2019 was fairly average, however, significantly colder days in July occurred, while August had both significantly colder and warmer days.
 April 2019 had significantly warmer than average days.
@@
   \includegraphics[width=12cm]{fig08}
-  \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as percentages of days significantly deviating from climatological mean by means of standard deviation for each month.
-Negative deviations from climatology are shown as negative percentage.
+  \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as percentages of days significantly deviating from climatological mean through standard deviation for each month.
+Negative deviations from the climatology are shown as a negative percentage.
 The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
@@
 
-In a similar way, the precipitation anomalies at Svanvik are displayed in Fig.~\ref{fig:anomalies_svanvik}b).
-Due to a large interannual variability in observed precipitation, only a few days deviated significantly from the climatology on the $1 \sigma$ level.
+Similarly, the precipitation anomalies at Svanvik are displayed in Fig.~\ref{fig:anomalies_svanvik}b).
+Due to large interannual variability in observed precipitation, only a few days deviated significantly from the climatology on the $1 \sigma$ level.
 Therefore, we lowered the $\sigma$ level constraint.
 The percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ from the climatological mean is shown for each month.
-Negative deviations from the climatology are shown as negative percentage and annual positive/negative deviations are indicated in the respective corners.
+Negative deviations from the climatology are shown as a negative percentage and annual positive/negative deviations are indicated in the respective corners.
 The hatched area between the dashed lines indicates the expected percentage of values falling above $\pm\frac{1}{2}\,\sigma$ if a normal distribution is assumed ($30.8\,\unit{\%}$).
-Unlike for the temperature anomalies, the picture is not as clear.
-While March 2018 had a significant number of days wetter than unusual, summer and fall (May/July, September/October) had a significantly number of days drier than average.
+Unlike the temperature anomalies, the picture is not as clear.
+While March 2018 had a significant number of days wetter than unusual, summer and fall (May/July, September/October) had a significant number of days drier than average.
 Overall, 2019 was rather average throughout the year, but with a significant number of days drier than normal (September/October) and a wetter December.
@@
 
-In summary, 2018 had a significant number of days which were warmer, drier, and brighter than the climatological average, while 2019 was a rather average year.
+In summary, 2018 had a significant number of days that were warmer, drier, and brighter than the climatological average, while 2019 was a rather average year.
 
@@
 For Svanvik this assessment is based on rather uncertain observations from 1986--1996.
-We also discussed the general patterns, similarities and differences.
-We shall now investigate \chem{\Delta[O_3]} in 2018/19 with respect to these climatologies in more detail.
+We also discussed the general patterns, similarities, and differences.
+We shall now investigate \chem{\Delta[O_3]} in 2018/19 for these climatologies in more detail.
 
@@
 All the above factors make it necessary to look not only at anomalies but statistically assess ozone concentration residuals at Svanvik as well as at Esrange and Pallas in comparison.
-Since the climatology of latter is more reliable, we will start with ground-level ozone concentration residuals \chem{\Delta[O_3]} at Esrange and Pallas in 2018.
+Since the climatology of the latter is more reliable, we will start with ground-level ozone concentration residuals \chem{\Delta[O_3]} at Esrange and Pallas in 2018.
 As data for 2019 has not yet been available from EBAS (last access, February 2020), we refrain from looking into these.
@@
 Supplement Figure~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik} displays the probability density functions of daily mean ozone concentration residuals for Svanvik based on data taken in 2018/19 with respect to the climatology derived for northern Fennoscandia (denoted NF in Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}a) and the climatology derived for Svanvik (Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b).
-As previously mentioned, \chem{\left<[O_3]\right>} at Svanvik are $(6.6\pm 7.1)\,\unit{ppb}$ lower than the northern Fennoscandic climatology.
+As previously mentioned, \chem{\left<[O_3]\right>} at Svanvik are $(6.6\pm 7.1)\,\unit{ppb}$ lower than the northern Fennoscandia climatology.
 For 2018 and 2019, we find $\Delta\chem{\left<[O_3]\right>} = (-3.8\pm 5.6)\,\unit{ppb}$ and $\Delta\chem{\left<[O_3]\right>} = (-5.2\pm 5.0)\,\unit{ppb}$, respectively.
@@
 This could mean either that 2018 had been more exceptional at Svanvik or the derived climatology may need a correction for increased tropospheric background ozone concentration.
-Based on the meteorological conditions, we assume 2019 as average year which leads to a constant bias of about $1.2\,\unit{ppb}$.
+Based on the meteorological conditions, we assume 2019 as an average year which leads to a constant bias of about $1.2\,\unit{ppb}$.
 Daily \chem{\Delta_{present-past}\left<[O_3]\right>} from Esrange and Pallas, wherein present is defined as the average of 2010--2018 and past as the respective beginning of data taking until 1997, do not reveal a clear trend: We find for Esrange $(1.5\pm 3.1)\,\unit{ppb}$ and for Pallas $(-2.5\pm 3.7)\,\unit{ppb}$ \citep{NILU2003, BER:Ruoho-Airola2015}.
@@
 In Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}, the test statistics for Svanvik are displayed.
-From Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}a), we can deduce, that the residuals between the Svanvik climatology and the northern Fennoscandic climatology are significant on the $2\sigma$ level for at least half of the data, while the residuals are less certain in 2018 and 2019.
-In Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}b) the same is shown for the years 2018/19 with respect to the climatology from Svanvik.
+From Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}a), we can deduce, that the residuals between the Svanvik climatology and the northern Fennoscandia climatology are significant on the $2\sigma$ level for at least half of the data, while the residuals are less certain in 2018 and 2019.
+In Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}b) the same is shown for the years 2018/19 for the climatology from Svanvik.
 We find that less than $50\,\unit{\%}$ of ozone concentration residuals are significant on the $1\sigma$ level.
@@
 In Fig.~\ref{fig:anomalies_svanvik}d), the resulting percentage of days above/below $\pm 1\,\sigma$ from the bias-corrected climatological mean is displayed for each month.
-Negative deviations from climatology are shown as negative percentage.
-Note that missing bars for certain months are both due to data availability (see Section~\ref{subsubsec:atmo_svanvik} for details about the time range of data taking at Svanhovd) and anomalies not meeting the criteria of $\pm 1\sigma$ deviation from climatological mean.
+Negative deviations from climatology are shown as a negative percentage.
+Note that missing bars for certain months are both due to data availability (see Section~\ref{subsubsec:atmo_svanvik} for details about the time range of data taking at Svanhovd) and anomalies not meeting the criteria of $\pm 1\sigma$ deviation from the climatological mean.
 In April/May 2018, ozone was significantly enhanced, while 2019 was generally an average year.
@@
 July 2018 has not been corrected for the missing data due to data acquisition problems.
-The corrected value is, however, indicated by the star (refer to Section~\ref{subsec:ozone_reco} for details of the used method for gap filling).
-
-%\begin{figure}[t]
-  %\includegraphics[width=8.3cm]{ozone_signific_1}
-  %\includegraphics[width=8.3cm]{fig11}
-  %\caption{Ozone concentration anomalies at Svanvik.
-Ozone climatology has been corrected for an constant negative bias of $-1.2\,\unit{ppb}$.
-Percentage of days above/below $\pm 1\,\sigma$ from climatological mean displayed for each month.
-Negative deviations from climatology are shown as negative percentage.
-The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
-July 2018 has not been corrected for the missing data due to data acquisition problems.
-The corrected value is indicated by the star (refer to Section~\ref{subsec:ozone_reco} for details).
-In April/May 2018, ozone was significantly higher than on average, while 2019 was generally an average year.
-In both years ozone concentrations were significantly below average in November.}
-  %\label{fig:ozone_signific}
-%\end{figure}
+The corrected value is, however, indicated by the star (refer to Section~\ref{subsec:ozone_reco} for details of the used method for gap-filling).
 
@@
 Due to the high correlation with Svanvik (Appendix~Fig.~\ref{fig:density_distribution}), we decide on Pallas as reference station for the reconstruction.
-We correct the Fennoscandic climatology for the found time lag \chem{\left<[O_3]\right>_{hourly,\,t-3}} by shifting the respective time series back in time by $3\,\unit{h}$ and use it to derive a projection for Svanvik:
+We correct the Fennoscandia climatology for the found time lag \chem{\left<[O_3]\right>_{hourly,\,t-3}} by shifting the respective time series back in time by $3\,\unit{h}$ and use it to derive a projection for Svanvik:
 \begin{equation}
@@
 
-To estimate potential biases in assessment of ozone induced risk on Fennoscandic vegetation, we defined hypothetical temperature response functions representing different levels of assumed acclimation to cold climates.
-We assume northern Fennoscandic vegetation to be more tolerant towards cold temperatures and less tolerant to heat.
+To estimate potential biases in assessment of ozone induced risk on Fennoscandia vegetation, we defined hypothetical temperature response functions representing different levels of assumed acclimation to cold climates.
+We assume northern Fennoscandia vegetation to be more tolerant towards cold temperatures and less tolerant to heat.
 Based on probability density functions (PDF) of observed light and temperature, we assume that an acclimation ought to increase the enclosed area between the respective response functions and the PDF.
@@
 In our study, we do not explicitly account for this natural distribution of individual plant traits, but the chosen hypothetical, bespoke response functions might reflect this diversity within a given species or PFT.
-The resilience of northern Fennoscandic plant communities to climate change and predictions in this regard strongly depends on both the existing range of acclimations as well as the actual type of acclimation.
+The resilience of northern Fennoscandia plant communities to climate change and predictions in this regard strongly depends on both the existing range of acclimations as well as the actual type of acclimation.
 

### 2021-04-16 supplements/VegOzone_supplement.tex
Fix size of figures
Drop AOT40 completely

@@
 
-\begin{figure}[t]
+\begin{figure*}[t]
   \includegraphics[width=12cm]{test_fennoscandia}
@@
   \label{fig:sunlight_fennoscandia}
-\end{figure}
+\end{figure*}
 
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs}
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{ozone_climatology_fenoscandic_obs}
   \caption{Multi-annual mean of daily mean ozone at observation sites in Fennoscandia.
@@
   \label{fig:ozone_climatology_fenoscandic_obs}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{ozone_fennoscandic_obs_aot40}
-  \caption{Integrated AOT40 for various sites in northern Fennoscandia over the course of 33 years.
+%\begin{figure*}[t]
+%  \includegraphics[width=12cm]{ozone_fennoscandic_obs_aot40}
+%  \caption{Integrated AOT40 for various sites in northern Fennoscandia over the course of 33 years.
 For simplicity, we integrated hourly ozone between $1\,\unit{am}-11\,\unit{pm}$ in any case, although a night time light intensity $> 50\,\unit{W\,m^{-2}}$ is only measured during midnight sun conditions.

### 2021-04-16 VegOzone.tex
Add reference regarding CAMSRA
Comments and suggestions by Lisa Section 1.

@@
 
-ground-level ozone is a highly toxic pollutant known to cause damage to a variety of ecosystems around the World \citep{PT:Emberson2020}.
-At continuously high concentration levels, ozone negatively affects the life expectancy of humans \citep{WHO2008} and is a cause of visible injury, photosynthetic damage, early senescence as well as programmed cell death of plants \citep{PCE:Kangasjarvi2005}.
+ground-level ozone is a highly toxic pollutant known to cause damage to of humans health \citep{WHO2008} and to a variety of ecosystems around the World \citep{PT:Emberson2020}.
+Ozone has been found to be a cause of visible injury, photosynthetic damage, early senescence as well as programmed cell death of plants \citep{PCE:Kangasjarvi2005}.
 Annual global yield losses of four major crops (wheat, rice, maize, and soybean) of $3-15\,\unit{\%}$ \citep{PJ:Ainsworth2017} and a suggested loss in primary production in forestry of $7\,\unit{\%}$ \citep{GCB:Wittig2009,EP:Matyssek2012} has been attributed to ozone.
@@
 This can be attributed to a successful implementation of air quality regulations coupled with economic restructuring which saw a move from coal to gas, reducing the number of episodes of peak concentrations especially in summertime over Europe and North America \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
-However, there are indications that an increase in the frequency of heatwave induced thermal stress will reduce the vegetations' capacity for ozone uptake and therefore diminish these achievements \citep{NCC:Lin2020}.\\
-
-Ground-level ozone is formed in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) and hydrocarbons are known as volatile organic substances (VOCs) in the presence of nitrogen oxides (\chem{NO_x}).
-Despite a relatively short tropospheric lifetime of approximately $22\,\unit{days}$, ranging from a few days in the tropical boundary layer to up to $1\,\unit{year}$ in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}, ozone is subject to advection over long distances.
+However, changes in environmental conditions associated with climate change, such as an increase in the frequency of heatwaves, could negate the effectiveness of emission reductions in relation to ozone impacts and air quality\citep{NCC:Lin2020}.
+
+Ground-level ozone is formed in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) and hydrocarbons are known as volatile organic compounds (VOCs) in the presence of nitrogen oxides (\chem{NO_x}).
+Despite a relatively short tropospheric lifetime of approximately $22\,\unit{days}$, ranging from a few days in the tropical boundary layer to up to $1\,\unit{year}$ in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}, ozone and is precursors are subject to advection over long distances.
 Episodes of high ozone concentrations in the absence of local pollution sources can therefore often be attributed to the long-range transport of both ozone and its precursors.
 In northern Fennoscandia, episodes of enhanced ozone concentrations in 2003 and 2006 have been traced back to ozone precursor emissions related to forest fires in southern and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}.
-Besides, hydrocarbons are also emitted by vegetation itself in form of terpenes and monoterpenes, so-called biogenic volatile substances (BVOCs).
-These emissions are associated with, e.g., thermal stress, defense against herbivores, or even attraction of pollinators \citep{TPS:Penuelas2003}.
+Besides, hydrocarbons are also emitted by vegetation itself in the form of terpenes and monoterpenes, so-called biogenic volatile substances (BVOCs).
+These emissions are thought to be a response to other factors, e.g., thermal stress, defense against herbivores, or even attraction of pollinators \citep{TPS:Penuelas2003}.
 
@@
 As dry deposition to bare ground or snow and ice-covered surfaces was found to be low \citep{ACP:Helmig2007}, both pathways of removal are substantially suppressed during polar night \citep{AE:Monks2000}.
-This leads to a build-up of ozone and its precursors in winter.
+This leads to a build-up of ozone and its precursors during winter.
 Come spring precursors are photochemically activated again, while snow cover typically continues to prevail and suppress ozone dry deposition.
@@
 Its main action is imposed through reactions occurring in the cell walls and cell membranes of mesophyll cells inside the leaves.
-Ozone enters the leaves through stomata, which are kept open to allow for gas exchange allowing the entry of \chem{CO_2} for photosynthesis and loss of \chem{H_2O} vapor via the plant transpiration stream.
+Ozone enters the leaves through stomata, leaf pores that enable gas exchange, allowing the entry of \chem{CO_2} for photosynthesis and loss of \chem{H_2O} vapor via the plant transpiration stream.
 Stomatal aperture, and hence stomatal conductance to these gases (including ozone) will vary over the day and growing season primarily to balance \chem{CO_2} uptake against \chem{H_2O} vapor loss.
@@
 
-To measure the risk of damage inflicted to vegetation through ozone, different metrics have been defined.
-Two main types can be differentiated, exposure-based and uptake-based metrics, e.g., Accumulated Ozone Exposure over a threshold of $\mathrm{x}$ ($\mathrm{AOT_x}$) and Phyto-toxic Ozone Dose over a threshold $\mathrm{y}$ ($\mathrm{POD_y}$).
-Due to their simplicity, exposure-based metrics are still widely used.
-Comparisons between the two methods (e.g.
-$\mathrm{AOT_{40}}$ with $\mathrm{POD_3}$) show that uptake-based metrics are better suited to represent the actual risk to vegetation from ozone, especially in the climatically extreme parts of Europe \citep{GCB:Mills2010,ICP:MappingManual2017,iF:Juran2018} (e.g., Mediterranean, Baltics/Fennoscandia).
-
-The exposure-based metric $\mathrm{AOT_x}$ (in units of $\unit{ppb}$) is defined as:
-\begin{equation}
-  \mathrm{AOT_x} = \int{\mathrm{max(\chem{[O_3]}-x)}\cdot \mathrm{dt}},
-  \label{eq:aot}
-\end{equation}
-with the ambient ozone concentration $\chem{[O_3]}$.
-The threshold exeedence is typically integrated from $8\,\mathrm{am}-8\,\mathrm{pm}$ daily and $3\,\unit{months}$ (May--July) annually \citep{EUdirective2002}.
-
-For computing the $\mathrm{POD_y}$, an estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf and environmental factors affecting stomatal conductance as described further in Section~\ref{sec:do3se}.
+A substantial body of evidence exists that suggests flux-based metrics, that relate stomatal ozone uptake to vegetation damage, are biologically more relevant for risk assessments than exposure-based metrics.
+This is because they are able to account particular species characteristics (i.e.
+physiology and phenology) as well as environmental conditions that can decouple the relationship between ozone concentration, ozone uptake, and consequent damage \citep{PT:Emberson2020}.
+These flux based metrics have been shown to be better suited to represent the actual risk to vegetation from ozone, especially in the climatically extreme parts of Europe \citep{EP:Simpson2007,GCB:Mills2010,ICP:MappingManual2017}.
+Most of these previous studie have focussed on the Mediterranean where soil moisture deficits are thought to be most influential in decoupling concentration from uptake.
+Relatively few studies have explored the situation in Northern European climates where low temperatures may be considered more likely to influence ozone uptake \citep[e.g.]{iF:Juran2018}.
+
+To estimate the flux-based metric, also referred to as the Phytotoxic Ozone Dose over a threshold y ($\mathrm{POD_y}$), an estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf and environmental factors affecting stomatal conductance as described further in Section~\ref{sec:do3se}.
 $\mathrm{POD_y}$ is then calculated according to:
@@
 %
-with the hourly averaged ground-level ozone flux through the stomata Eq.~(\ref{eq:flux_stomata}) and a flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}\,s^{-1}}.
-The flux threshold $\mathrm{y}$ is representing the detoxification potential of the plant and is typically only exceeded during daylight hours (i.e.
+with the hourly averaged stomatal ozone flux $\Phi_\mathrm{sto}$ (see Eq.~(\ref{eq:flux_stomata}) and a stomatal ozone flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}\,s^{-1}}.
+The flux threshold $\mathrm{y}$ represents the detoxification potential of the plant and is typically only exceeded during daylight hours (i.e.
 when global radiation is above $50\,\unit{W\,m^{-2}}$).
-
-Finally, exceedance of the flux-based critical level (CL) is calculated by:
+$\mathrm{POD_y}$ is given in \unit{mmol\,m^{-2}} over the growing season.
+
+The $\mathrm{POD_y}$ metric is currently used in risk assessments under the United Nations Economic Commission for Europe (UNECE) Long Range Transboundary Air Pollution (LRTAP) Convention to identify those locations across Europe where vegetation is at risk from ozone.
+This is with a view to developing effects-based emission reduction policy that can target those ozone precursor emissions that are most influential in causing damage.
+To achieve this the concept of critical levels (CLs) has been developed and applied \citep{Maas2016}.
+Exceedance of the CL is used to identify those areas across Europe that would benefit from targeted Europe-wide emission reductions.
+Methods to estimate $\mathrm{POD_y}$ and CLs are defined by the UNECE LRTAP Convention and described in a Mapping Manual \citep{ICP:MappingManual2017}.
+The CL is calculated by:
 %
@@
 
-Within the scope of this study to assess probable effects of conditions unique to northern Fennoscandia on ozone uptake by plants and the resultant enhanced susceptibility to ozone inflicted damage, we conducted ground-level ozone concentration measurements at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd in the growing seasons of 2018/19 in collaboration with the Norwegian Institute for Air Research (NILU).
-In 2018, we observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden at Svanhovd.
-In contrast, no such damage was found on the clovers in 2019, although it was found on the sensitive tobacco cultivar.
-This indicates that the vegetation may have been more affected by ozone in 2018 than in 2019.
-In terms of $\mathrm{AOT_{40}}$, the threshold for protection of vegetation of $3000\,\unit{ppb\,h}$ (integrated over the growing season) given by the EU~directive \citep{EUdirective2002} has been well exceeded at many ozone monitoring sites in Southern Norway in the 2018 growing season \citep{NILU2019} but not at Svanhovd.
+An application of the Mapping Manual method was made by \citet{GCB:Mills2011} where exceedance of the CL was related to observations of ozone damage to a clover biomonitor.
+This study clearly demonstrated the improved ability of flux-based metrics (in comparison to exposure-based metrics) to identify the geographical distribution of the risk of vegetation to ozone damage and also showed that the ozone damage could extend into more northerly regions of Europe.
+A similar biomonitoring study was conducted at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd to assess whether ozone was capable of inducing damage to vegetation growing under northern Fennoscandia conditions.
+In 2018, we indeed observed ozone damage on clover and tobacco.
+In contrast, no such damage was found on the clovers in 2019.
 
@@
 Coincident peak \chem{[O_3]} is found in ozone monitoring data from Svanhovd but also over whole northern Fennoscandia in July (see Fig.~\ref{fig:data_svanvik_2018_2019} or Fig.~\ref{fig:ozone_reconstruction_2018_07}).
-In the 2003 drought period, elevated ozone in Europe was promoted by a combination of various factors such as reduced cloud (increased solar radiation), reduced dry deposition and turbulent mixing due to the stagnant weather conditions, and increased BVOC emissions \citep{JGR:Solberg2018}.
+In the 2003 drought period, elevated ozone in Europe was promoted by a combination of various factors such as wildfires, reduced cloud cover (increased solar radiation), reduced dry deposition and turbulent mixing due to the stagnant weather conditions, and increased BVOC emissions \citep{JGR:Solberg2018}.
 It is likely that these factors also contributed to enhanced \chem{[O_3]} in 2018.
@@
 Parameterization of these models for more extreme bio-geographical regions such as subarctic climates has not been performed and relies on assuming that less extreme Boreal parameterization will be applicable.
-In this study, conditions at Svanhovd for the growing seasons 2018/19 shall serve as a reference.
+In this study, conditions at Svanhovd for the growing seasons 2018/19 shall serve as a reference for present and probable future conditions in northern Fennoscandia.
 It is important to understand the implications of using these more generic parameterizations for more extreme climate regions.
 Here we explore how risk estimates might be over or underestimated using generic vs bespoke subarctic parameterization.
-we also explore the role \chem{[O_3]} variability and changing climate variables to understand how this risk might alter in the future with the advent of climate change.
-This allows both an understanding of how the threat from ozone may change over time but also which are the key aspects of plant physiology that might determine risk (and hence which parameters might require special attention to get right) both for pollution impact assessment, but also more generally for describing gas exchange response to changing climates that will influence biochemical cycling.
-This might be one key to solve long-standing issues of coupled earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface model \citep{GMD:Poulter2015,JAMES:Lawrence2019}.\\
+We also explore the role of \chem{[O_3]} variability and changing climate variables to understand how this risk might alter in the future with the advent of climate change.
+This allows an understanding of how the threat from ozone may change over time and which are the key aspects of plant physiology that might determine risk.
+Hence, we determin which parameters require special attention to define accurately for pollution impact assessment and more generally for describing gas exchange response to changing climates that will influence biochemical cycling.
+This might be one key to solve long-standing issues of coupled earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface models \citep{GMD:Poulter2015,JAMES:Lawrence2019}.\\
 
@@
 In Section~\ref{sec:stats}, we derive climatologies for ground-level ozone concentration at Svanhovd and northern Fennoscandia based on in situ observations and also compute climatologies of near-surface temperature, precipitation, and global radiation for Svanhovd.
-Based on these, we evaluate anomalies of environmental key variables in 2018/19 by statistical means.
+Based on these, we evaluate anomalies of key environmental variables in 2018/19 by statistical means.
 The derived ground-level ozone concentration climatology for northern Fennoscandia will also be compared with available ground-level ozone reanalysis products.
-We use the $\mathrm{DO_3SE}$ model to estimate the ozone uptake by natural and semi-natural vegetation to investigate main environmental drivers for the observed ozone damage in 2018 (Section~\ref{sec:do3se}).
-In this context, we introduce the$\mathrm{DO_3SE}$ model equations and derive bespoke parameterizations for subarctic species to assess systematic uncertainties arising from the use of default parameterizations.
-In Section~\ref{sec:conc} finally, we will summarize our results, conclude, and give an outlook on further work.
+We use the $\mathrm{DO_3SE}$ model to estimate the ozone uptake by natural and semi-natural vegetation to investigate the main environmental drivers for the observed ozone damage in 2018 (Section~\ref{sec:do3se}).
+In this context, we introduce the $\mathrm{DO_3SE}$ model equations and derive bespoke parameterizations for subarctic species to assess systematic uncertainties arising from the use of default generic parameterizations established from less climatically extreme conditions.
+In Section~\ref{sec:conc} finally, we will summarize our results and discuss how this research can support improvements in future assessments of pollution risk and more generally bio-geochemical cycling.
 
@@
 Satellite retrievals resolve \chem{[O_3]} at the surface rather poorly and hence do not constrain the global models well enough \citep{ACP:Andersson2017}.
+\citet{ACPD:Barten2020} reported pronounced biases for CAMSRA in the high Arctic (e.g., Summit, Greenland) and attribute this to dry deposition to the ocean.
 
@@
 \begin{table}[t]
-  \caption{Bespoke temperature parameterizations. MM refers to mapping manual \citep{GCB:Mills2010,ICP:MappingManual2017}.}
+  \caption{Bespoke temperature parameterizations. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
   \label{tab:sensitivity_tests_temp}
@@
 \begin{table}[t]
-  \caption{Bespoke light parameterizations. MM refers to mapping manual \citep{GCB:Mills2010,ICP:MappingManual2017}.}
+  \caption{Bespoke light parameterizations. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
   \label{tab:sensitivity_tests_light}
@@
 \begin{table}[t]
-  \caption{Bespoke start and end of growing season in units of \unit{doy}. MM refers to mapping manual \citep{GCB:Mills2010,ICP:MappingManual2017}.}
+  \caption{Bespoke start and end of growing season in units of \unit{doy}. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
   \label{tab:sensitivity_tests_gs}

### 2021-04-16 VegOzone.tex
Comments and suggestions from Lisa Sec 2-3

@@
 
-In our assessment of primary drivers of ozone damage risk on vegetation in the subarctic, we focus on the 2018 growing season which was subject to an extended heatwave and its subsequent growing season of 2019.
-The NIBIO operated Environment Centre Svanhovd which lies close by the settlement of Svanvik in the Pasvik valley serves as the main reference.
+In our assessment of primary drivers of ozone damage risk on vegetation in the subarctic, we focus on comparing the anomalous warm, bright and dry 2018 growing season with the more climatically average 2019 season to understand the role these environmental conditions may have played in influencing the sensitivity of vegetation to ozone damage.
+The NIBIO operated Environment Centre Svanhovd which lies close by the settlement of Svanvik in the Pasvik valley serves as the main reference for conditions in northern Fennoscandia.
 Throughout the paper, we will use Svanhovd synonymously for the atmospheric monitoring site at Svanvik/Pasvik.
@@
 Other data used for cross-evaluation of the ozone monitoring data taken at Svanhovd is described in Section~\ref{subsec:other_data}.
-We compare with observational data from other monitoring sites in northern Fennoscandia (Section~\ref{subsubsec:ebas}) and use these to evaluate reanalysis products for ground-level ozone (Section~\ref{subsubsec:ozone_rea}).
+We, in particular, compare with observational data from other monitoring sites in northern Fennoscandia (Section~\ref{subsubsec:ebas}) and use these to evaluate reanalysis products for ground-level ozone (Section~\ref{subsubsec:ozone_rea}).
 
@@
 The locations of the atmospheric monitoring site and the ozone garden are marked in aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}a).
-Svanhovd is located south of the settlement of Svanvik and the border to Russia, the Pasvik river, lies only a few hundred meters to the east.\\
+Svanhovd is located south of the settlement of Svanvik, where the Pasvik river, which is the border to Russia, lies only a few hundred meters to the east.\\
 
@@
 
-An ozone garden was installed in the course of our project at the NIBIO Environment Centre Svanhovd and was likely the northern-most ozone garden to date.
+An ozone garden was installed over the course of our project at the NIBIO Environment Centre Svanhovd and was likely the northern-most ozone garden to date {\bf TODO: Reference?}.
 An ozone garden consists of selected plant species that are more sensitive to ozone and display visible injuries.
@@
 All relevant data for the growing season 2018/19 are shown in Fig.~\ref{fig:data_svanvik_2018_2019}.
-Ozone concentrations \chem{[O_3]} measured in $2\,\unit{m}$ height above ground are averaged hourly.
+Ozone concentrations \chem{[O_3]} measured at $2\,\unit{m}$ height above ground are averaged hourly.
 The hatched areas mark times when no ozone data were recorded.
@@
 However, due to the above-mentioned data acquisition problems, we missed most of the corresponding ozone data for this event and most likely also the peak ozone concentration.
+We present a reconstruction concerning the missing data in Section~\ref{subsec:ozone_reco}.
 In contrast, ozone concentration only rose 18 times above the threshold of $40\,\unit{ppb}$ during the summer of 2019.
@@
 \label{subsec:other_data}
-To determine the significance of the ozone conditions in 2018, we derive a ground-level ozone concentration climatology for Svanvik.
+To determine the significance of the higher \chem{[O_3]} in 2018 in relation to regional averages and trends, we derive a ground-level ozone concentration climatology for Svanvik.
 From 1986--1996 ozone monitoring has been operated continuously by NILU at Svanhovd, but no ozone measurements have been conducted until the start of our project.
-Due to probable changes in northern hemispheric tropospheric background ozone and a reduction of summertime peak values related to ozone precursors which fell under air quality regulations in Europe, we will cross-calibrate our data with long-term ozone observations at other sites in northern Fennoscandia.
+Due to probable changes in northern hemispheric tropospheric background ozone and a reduction of summertime peak values related to ozone precursors which fell under air quality regulations in Europe, we cross-calibrate our data with long-term ozone observations at other sites in northern Fennoscandia.
 Concerning ozone, available observational data are sparse in this region.
@@
 The data are described in Section~\ref{subsubsec:ebas}.
-As an additional complement to observational data which are not available at all times and spatially sparse, we also evaluate various ozone reanalysis products (Section~\ref{subsubsec:ozone_rea}).
+As an additional complement to observational data which are not available at all times and spatially sparse, we also evaluate \chem{[O_3]} derived from a variety of reanalysis products (Section~\ref{subsubsec:ozone_rea}).
 
@@
   \caption{Cap of the North.
-Locations of past and present ozone observation sites in northern Fennoscandia were used in this study.
+Locations of past and present (end of 2018) ozone observation sites in northern Fennoscandia were used in this study.
 For more details see Tab.~\ref{tab:ebas_obs}.
@@
 In Table~\ref{tab:ebas_obs}, we give an overview of the exact locations and time range of operational data taking.
-All stations with available long-term observations are located on higher ground than Svanhovd which will result in higher ground-level ozone concentrations in comparison as ozone abundance, in general, decreases with decreasing altitude in the absence of local precursors \citep[e.g.,][]{AB:Klingberg2009}.
+All stations with available long-term observations are located on higher ground than Svanhovd which will result in relatively higher ground-level ozone concentrations in comparison as ozone abundance, in general, decreases with decreasing altitude in the absence of local precursors \citep[e.g.,][]{AB:Klingberg2009}.
 The periphery vegetation at all sites is similar and consists mainly of pine forests, birch, and heath.
-However, in close vicinity, the vegetation varies substantially.
+However, at higher resolution the vegetation varies substantially.
 Ozone measurements at Pallas are taken on top of mount Sammaltunturi $100\,\unit{m}$ above treeline (pine forest) and mainly consists of low vascular plants, mosses, and lichen \citep{BER:Hatakka2003}.
-At Svanhovd, the ozone measurements are taken amidst a meadow surrounded by agriculturally managed herbaceous plants and grass (c.c.
-Fig.\ref{fig:svanhovd_research_station}).
+At Svanhovd, the ozone measurements are taken amidst a meadow surrounded by agriculturally managed herbaceous plants and grass adhering to the WMO standards {\bf TODO: Reference \url{https://library.wmo.int/index.php?id=12407&lvl=notice_display}} (Fig.\ref{fig:svanhovd_research_station}).
 
@@
 In 1997, the ozone monitoring at Jergul was relocated downstream of the river Karasjokka to a site close by Karasjok.
-Although the same equipment has been used, this might have introduced a breaking point in the combined data series, denoted as Jergul/Karasjok in the following, which would be a problem if this was the sole data to compare with.
+Although the same equipment has been used, this might have introduced a breaking point in the combined data series, denoted as Jergul/Karasjok in the following, which would be particularly problematic if these were the sole data for comparison.
 Only a few data have been taken upstream the Pasvik river at Janiskovski, hence these will not be taken into further consideration, but are shown for completeness.
@@
   \caption{Time series of ozone observations in northern Fennoscandia (Tab.~\ref{tab:ebas_obs}) (1986--2019).
-Data are available from EBAS.
+Data from EBAS until December 2018.
 The hatched areas indicate periods with insufficient quality control according to \citet{NILU2003}.}
@@
 Details of reanalysis data sets referred to in this work are listed in Table~\ref{tab:ozone_rea}.
-There are two global reanalysis products available at the European Centre for Medium-Range Weather Forecasts (ECMWF) that include atmospheric tracers, in particular ozone, the Monitoring Atmospheric Composition and Climate (MACC) and the newer Copernicus Atmosphere Monitoring Service reanalysis (CAMSRA).
-From Copernicus member institutions a CAMS Regional Air Quality (CAMSRAQ) reanalysis ensemble is available for Europe.
+
+There are two global reanalysis products available at the European Centre for Medium-Range Weather Forecasts (ECMWF) that include atmospheric tracers, in particular ozone, the Monitoring Atmospheric Composition and Climate (MACC) and the newer Copernicus Atmosphere Monitoring Service reanalysis (CAMSRA) \citep{ACP:Inness2013, ACP:Inness2019}.
+The temporal as well as spatial resolution of these reanalysis products is rather coarse: 3-hourly and $0.75\,\unit{^\circ}\times0.75\,\unit{^\circ}$ or roughly $29.3\,\unit{km}\times 83.4\unit{km}$ at the location of Svanvik.
+From Copernicus member institutions, higher resolution suface ozone ranalysis data are available for Europe (CAMS Regional Air Quality (CAMSRAQ) reanalysis ensemble).
+The CAMSRAQ is based on nine European state-of-the-art numerical air quality models.
+The ensemble mean is at fairly high spatial and temporal resolution compared to the global reanalysis: $0.1\,\unit{^\circ}\times 0.1\,\unit{^\circ}$ or roughly $3.9\,\unit{km}\times 11.1\,\unit{km}$ and 1-hourly, respectively.
+The periodes covered differ and no data is available before the turn of the millennium.
 The CAMSRA and CAMSRAQ are available in near real time, but only the former covers a period of sufficient length for climate analysis (2003 -- present).
@@
 The quality concerning observations in northern Fennoscandia will be discussed in Section~\ref{subsubsec:clim_ozone}.
-In summary, the results prevented us from using any of the global reanalysis data.
+Due to the found substantial discrepancies between reanalysis data and site observations in northern Fennoscandia, we refrain from using any of the global reanalysis data.
 
@@
 \label{sec:stats}
-In this section, we will evaluate atmospheric conditions in the growing seasons of 2018/19 by statistical means.
-We will derive climatologies for temperature, precipitation, global irradiance, and ozone for Svanvik (Section~\ref{subsec:climatologies}).
+To improve our understanding and determine the risk of vegetation from high \chem{[O_3]} in subarctic biomes in a changing climate, we first assess the climate conditions and evaluate the growing seasons 2018/19 concerning anomalies from multi-annual means (climatologies).
+We derive climatologies for temperature, precipitation, global irradiance, and ozone for Svanvik.
 The ozone measurements will be put into a climatological as well as regional context.
-We will look at anomalies of the atmospheric key variable mentioned above (Section~\ref{subsec:anomalies}).
 
@@
 
-In Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}, the two dimensional density distributions for the combined northern Fennoscandia data (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}a) and Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) are shown.
+In Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}, the two dimensional density distributions of \chem{[O_3]} for the combined northern Fennoscandia data (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}a) and Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) are shown.
 Darker colors indicate higher probability to observe these values.
@@
 The build-up of the ozone spring peak takes about $7.5\,\unit{months}$ from minimum to maximum.
-A sharp increase starts one month ahead of the estimated begin of greening $G_\mathrm{begin}$ (see Section~\ref{subsec:do3se_parameters} for details) and is in coincidence with rising temperatures and global irradiance.
-We find day of the year (\unit{doy}) $\left<G_\mathrm{begin}\right> = (150 \pm 14)$ for Svanvik (Supplement Fig~\ref{fig:greening_season_change_Svanvik}), which translates into the end of May.
+A sharp increase starts one month ahead of the estimated begin of greening $G_\mathrm{begin}$ (see Section~\ref{subsec:do3se_parameters} for details) and is in coincident with rising temperatures and global irradiance.
+We find day of the year (\unit{doy}) $\left<G_\mathrm{begin}\right> = (150 \pm 14)$ for Svanvik (Supplement Fig~\ref{fig:greening_season_change_Svanvik}), which translates to the end of May.
 This finding is in line with more extensive studies \citep[][e.g.,]{IJB:Karlsen2007,AFM:Linderholm2006}.
@@
 As Svanvik is located at a lower altitude than the other sites, the earlier decline of the ozone peak may point to a stronger influence of vegetation on the removal of ozone.
-\chem{CO_2} uptake in coniferous trees can occur as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013}.
+\chem{CO_2} uptake by coniferous trees can occur as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013}.
 The average ozone concentration at Svanvik is lower than at the other sites, as already mentioned, which is in league with a decrease in tropospheric ozone with decreasing altitude.
@@
 Finally, we shall have a look at the climatologies derived from observational ozone data and the various reanalysis products (Fig.~\ref{fig:ozone_climatologies}).
-As representative for the sites, the nearest neighboring grid point has been chosen.
-The ECMWF products do not reproduce observed ground-level ozone well.
+To represent the sites, the nearest neighboring grid point of the \chem{[O_3]} reanalysis products is chosen.
+The global products do not reproduce observed ground-level \chem{[O_3]} well.
+This is important for European wide and global risk assessment that rely on these data.
 The MACC reanalysis (Fig.~\ref{fig:ozone_climatologies}a) is in general too low ($(9\pm 7)\,\unit{ppb}$) and displays an erroneous seasonal cycle.
@@
   \includegraphics[width=8.3cm]{fig07}
-  \caption{Comparison of climatologies derived from observational data (EBAS) with various ozone reanalysis products: (a) MACC reanalysis; (b) CAMSRA; (c) CAMSRAQ reanalysis ensemble mean.
-As the representative for the sites, the nearest neighboring grid point has been chosen.
+  \caption{Comparison of climatologies derived from observational data (gray band) with three ozone reanalysis products: (a) MACC reanalysis; (b) CAMSRA; (c) CAMSRAQ reanalysis ensemble mean.
+The sites, represented by the nearest neighboring grid point of the \chem{[O_3]} reanalysis products, are displayed in different markers and colors.
 The global products (a), (b) do not reproduce observed ground-level ozone well.
@@
 The model enhancements from MACC reanalysis to CAMSRA have been reported and discussed by \citet{ACP:Inness2019} on a global level.
-Amongst others, assimilation of ozone profiles from satellite retrieval and an upgraded ozone chemistry scheme has lead to the enhanced performance of the latter.
+Amongst others, assimilation of ozone profiles from satellite retrieval and an upgraded ozone chemistry scheme has lead to the enhanced performance of the latter, but a considerable bias still remains \citep{GMD:Huijnen2020}.
+In particular, \citet{ACPD:Barten2020} reported a pronounced bias for CAMSRA in the high Arctic (e.g., Summit, Greenland) and attribute this bias to dry deposition to the ocean.
 The higher spatial resolution of the regional air quality models used in the CAMSRAQ ensemble compared to the global reanalyses captures small scale depletion and peak episodes in \chem{[O_3]} better and enhances the capability of the models concerning daily and seasonal cycles in ozone.
@@
 Satellite retrievals resolve \chem{[O_3]} at the surface rather poorly and hence do not constrain the global models well enough \citep{ACP:Andersson2017}.
-\citet{ACPD:Barten2020} reported pronounced biases for CAMSRA in the high Arctic (e.g., Summit, Greenland) and attribute this to dry deposition to the ocean.
 
@@
 Using these established climatologies, we look at the anomalies in 2018/19 and discuss the differences between the two years.
+As weather extremes like in 2018 are likely to become more frequent in occurrence with climate change, this will help improving our understanding of the role which these climate variables have on determining the sensitivity of subarctic vegetation to \chem{[O3]}.
 
@@
 Therefore, we lowered the $\sigma$ level constraint.
-The percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ from the climatological mean is shown for each month.
+The percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ from the climatological mean are shown for each month.
 Negative deviations from the climatology are shown as a negative percentage and annual positive/negative deviations are indicated in the respective corners.
@@
 In the following, we will use the interchangeable terms anomalies and residuals to distinguish between a climatological (anomalies) and a statistical (residuals) point of view.
-All the above factors make it necessary to look not only at anomalies but statistically assess ozone concentration residuals at Svanvik as well as at Esrange and Pallas in comparison.
+All the above factors make it necessary to look not only at anomalies but statistically also to assess ozone concentration residuals at Svanvik as well as at Esrange and Pallas in comparison.
 Since the climatology of the latter is more reliable, we will start with ground-level ozone concentration residuals \chem{\Delta[O_3]} at Esrange and Pallas in 2018.
@@
 In Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}, the probability density functions of daily mean ozone concentration residuals for Esrange and Pallas are shown.
-Both stations display a clear deviation from the climatological mean, however, not significant.
+Both stations display a clear, but not significant, deviation from the climatological mean.
 On average, \chem{[O_3]} had been elevated by $(1.9\pm 5.7)\,\unit{ppb}$ and $(1.9\pm 5.4)\,\unit{ppb}$, respectively.
@@
 This could mean either that 2018 had been more exceptional at Svanvik or the derived climatology may need a correction for increased tropospheric background ozone concentration.
-Based on the meteorological conditions, we assume 2019 as an average year which leads to a constant bias of about $1.2\,\unit{ppb}$.
+Based on the meteorological conditions, we assume 2019 as an average year which leads to an estimated constant bias of about $1.2\,\unit{ppb}$ in \chem{[O_3]}.
 Daily \chem{\Delta_{present-past}\left<[O_3]\right>} from Esrange and Pallas, wherein present is defined as the average of 2010--2018 and past as the respective beginning of data taking until 1997, do not reveal a clear trend: We find for Esrange $(1.5\pm 3.1)\,\unit{ppb}$ and for Pallas $(-2.5\pm 3.7)\,\unit{ppb}$ \citep{NILU2003, BER:Ruoho-Airola2015}.
@@
 Negative deviations from climatology are shown as a negative percentage.
-Note that missing bars for certain months are both due to data availability (see Section~\ref{subsubsec:atmo_svanvik} for details about the time range of data taking at Svanhovd) and anomalies not meeting the criteria of $\pm 1\sigma$ deviation from the climatological mean.
+Note that missing bars for certain months are both due to data availability (see Fig~\ref{fig:data_svanvik_2018_2019} for details concerning Svanvik) and anomalies not meeting the criteria of $\pm 1\sigma$ deviation from the climatological mean.
 In April/May 2018, ozone was significantly enhanced, while 2019 was generally an average year.

### 2021-04-17 VegOzone.tex
Add Lisa's comments and suggestions until section 4.4

@@
 \label{sec:do3se}
-To investigate which of the environmental factors lead to the observed ozone damage on clover and tobacco in 2018 in contrast to 2019, when no such damage was found on the clovers but only on the sensitive tobacco cultivar, we model ozone uptake of different natural and semi-natural vegetation using the $\mathrm{DO_3SE}$ model and develop bespoke subarctic parameterizations for common plant species.
-We first give an account of the $\mathrm{DO_3SE}$ model, input data, and the used method for deriving the bespoke parameters.
+To investigate which of the environmental factors might be most influential in determining \chem{O_3} damage; such as that observed on clover and tobaccos in 2018, we model ozone uptake of different natural and semi-natural vegetation using the $\mathrm{DO_3SE}$ model and develop bespoke subarctic parameterizations for common plant species.
+We first give an account of the $\mathrm{DO_3SE}$ model, input data, and the used methodology for deriving the bespoke parameters.
 Then we will present and discuss the modeling results.
@@
 \label{subsec:model_describtion}
-The opening of stomata is governed by different species-specific and environmental factors and can be described by a multiplicative model \citep{PTRS:Jarvis1976,ICP:MappingManual2017}:
-\begin{equation}
-  g^k_\text{sto} = g^k_\text{max} \cdot f^k_\text{phen} \cdot f^k_\text{light} \cdot \max{\left\{f^k_\text{min}, f^k_\mathrm{T} \cdot f^k_\mathrm{VPD} \cdot f^k_\text{SW}\right\}}.
+The opening of stomata is governed by different species-specific and environmental factors and can be described by a multiplicative model \citep{PTRS:Jarvis1976,Emberson2000,ICP:MappingManual2017}:
+\begin{equation}
+  g^k_\text{sto} = g^k_\text{max} \cdot f^k_\text{phen} \cdot f^k_\text{light} \cdot \max{\left\{f^k_\text{min}, f^k_\mathrm{T} \cdot f^k_\mathrm{VPD} \cdot f^k_\text{SWP}\right\}}.
   \label{eq:stomatal}
@@
 Where $g_\mathrm{max}$ is the specific-specific maximum stomatal conductance value which is then modified by seasonal and environmental factors that vary within a range $0-1$.
-They are empirically determined and account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_\mathrm{T}$), water vapor pressure deficit ($f_\mathrm{VPD}$), and soil water content ($f_\text{SW}$).
+They are empirically determined and account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_\mathrm{T}$), water vapor pressure deficit ($f_\mathrm{VPD}$), and soil water potential ($f_\text{SWP}$).
 All factors differ with species or plant functional type (PFT) denoted with $k$.
@@
 %
-where $\alpha_\mathrm{light}$ is a slope parameter describing how fast stomata open or close at low light intensities.
+where $\alpha_\mathrm{light}$ is a slope parameter describing the extent of stomatal opening at low light intensities.
 
@@
 \begin{equation}
-  f_\mathrm{SW} = \text{min}\left\{1, \text{max}\left\{f_\mathrm{min}, \frac{(1 - f_\mathrm{min})\cdot(\mathrm{SWP_{min}} - \mathrm{SWP})}{\mathrm{SWP_{min}} - \mathrm{SWP_{max}}} + f_\mathrm{min} \right\} \right\}
+  f_\mathrm{SWP} = \text{min}\left\{1, \text{max}\left\{f_\mathrm{min}, \frac{(1 - f_\mathrm{min})\cdot(\mathrm{SWP_{min}} - \mathrm{SWP})}{\mathrm{SWP_{min}} - \mathrm{SWP_{max}}} + f_\mathrm{min} \right\} \right\}
     \label{eq:fsw}
@@
 %
-where SWP is the soil water potential across the root zone and $\mathrm{SWP_{min}}$ and $\mathrm{SWP_{max}}$ are the parameters describing the $f_\mathrm{SW}$ relationship.
+where SWP is the soil water potential across the root zone and $\mathrm{SWP_{min}}$ and $\mathrm{SWP_{max}}$ are the parameters describing the $f_\mathrm{SWP}$ relationship.
 
@@
 \item Several consecutive hours of missing data (23 or less) were filled by taking the average of the corresponding hour the day before, and the day after; and repeating this for each missing hour of data.
-If data was unavailable from that hour of the previous day, then only the value from the day after was used and vice versa.
+If data were unavailable from that hour of the previous day, then only the value from the day after was used and vice versa.
 Interpolated values were not used in calculating averages.
@@
 We derive hourly climatologies for individual stations and combine Esrange/Pallas to derive a climatology for northern Fennoscandia for July (Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}a).
-We compute temporal correlations between Svanvik and the other stations in northern Fennoscandia to estimate a time lag (Supplement Fig.~\ref{fig:time_lag_correlation}.
+We compute temporal correlations between Svanvik and the other stations in northern Fennoscandia to estimate a time lag (Supplement Fig.~\ref{fig:time_lag_correlation}).
 We find $3\,\unit{h}$ for Esrange and Pallas with respect to Svanvik.
-Due to the high correlation with Svanvik (Appendix~Fig.~\ref{fig:density_distribution}), we decide on Pallas as reference station for the reconstruction.
+Due to the high correlation with Svanvik (Appendix~Fig.~\ref{fig:density_distribution}), we decide on Pallas as the reference station for the reconstruction.
 We correct the Fennoscandia climatology for the found time lag \chem{\left<[O_3]\right>_{hourly,\,t-3}} by shifting the respective time series back in time by $3\,\unit{h}$ and use it to derive a projection for Svanvik:
@@
 
-From Fig.~\ref{fig:svanhovd_research_station}, we roughly estimated land use types for an approximately $1\times 1\,\unit{km}$ patch centered at the NIBIO Environment Centre Svanhovd : $48\,\unit{\%}$ grassland, $32.5\,\unit{\%}$ deciduous, $12.5\,\unit{\%}$ coniferous, $2\,\unit{\%}$ urban.
-
-Default parameters for these species or PFTs are derived from \citet{ICP:MappingManual2017} and will be referred to as \emph{mapping manual} abbreviated as MM in the following.
+The main species or PFTs, we deduced on-site at the NIBIO Environment Centre Svanhovd and from Fig.~\ref{fig:svanhovd_research_station}, are perennial grassland, downy birch, and Norway spruce.
+Default parameters for these species or PFTs are derived from \citet{ICP:MappingManual2017} {\bf TODO: find added reference by Lisa} and will be referred to as \emph{mapping manual} abbreviated as MM in the following.
 A comprehensive list of model parameters can be found in Supplement~Table~S1.
@@
 
-We hypothesize that perennial vegetation will likely prioritize their growth or fast adolescence during a short, subarctic growing season and hence their carbon sequestration.
-Because the $\mathrm{DO_3SE}$ model currently does not simulate $A_\mathrm{n}$, we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$.
+We hypothesize that perennial vegetation will likely prioritize their growth or fast adolescenceand and hence their carbon sequestration, during a short, subarctic growing season.
+Because the mapping manual version of the $\mathrm{DO_3SE}$ model currently does not simulate $A_\mathrm{n}$, we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$.
+We further assume that $f_\mathrm{VPD}$ and $f_\mathrm{SWP}$ suffice our biomes.
 The procedure is graphically illustrated in Fig.~\ref{fig:f_temp_grassland} for grassland.
 We aim to increase the enclosed area between probability density function (PDF) and the respective response function.
-We show two climatological reference periods 1990s and 2000s to indicate that the vegetation is subject to ongoing climate change and acclimation may be changing.
+We show two climatological reference periods 1990s and 2000s to indicate that the vegetation is subject to ongoing climate change and hence continuous acclimation.
 Resulting response functions for Norway spruce (Picea abies) and downy birch are displayed in Appendix~Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}.
@@
 We define two different hypothetical temperature acclimation: \emph{Boreal} and \emph{cold}.
+The most extreme of the two acclimations is \emph{boreal}.
 For either, we assume that the hypothetical species are more tolerant to cold conditions and shift $f_\mathrm{T}$ towards colder temperatures accordingly.
@@
     \middlehline
-    \multirow{3}{*}{Downy birch} & MM & 5 & 20 & \multirow{3}{*}{100}\\
-    & Boreal & 0 & 10\\
-    & Cold & 5 & 15\\
+    \multirow{3}{*}{Downy birch} & MM & 5 & 20 & 100\\
+    & Boreal & 0 & 10 & 100\\
+    & Cold & 5 & 15 & 100\\
     \middlehline
-    \multirow{3}{*}{Norway spruce} & MM & 0 & 20 & \multirow{3}{*}{100}\\
-    & Boreal & 0 & 10 & \\
-    & Cold & 0 & 15 & \\
+    \multirow{3}{*}{Norway spruce} & MM & 0 & 20 & 100\\
+    & Boreal & 0 & 10 & 100\\
+    & Cold & 0 & 15 & 100\\
     \middlehline
@@
 GS (May--August) climatological averages and standard deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time) are shown for (a) Downy birch; (b) Norway spruce; (c) perennial grassland.
-For simplicity, $f_\mathrm{SW}$ is neglected in this calculation.
+For simplicity, $f_\mathrm{SWP}$ is neglected in this calculation.
 }
@@
 By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{doy}$ for 2018 and \unit{doy} 106 in 2019.
-$A_\text{end}$ amounts to \unit{doy} 261 and 274, respectively, and shall be used for all vegetation alike.
+$A_\text{end}$ amounts to \unit{doy} 261 and 274, respectively, and is the value used for all PFTs.
 The resulting growing season for Norway spruce in 2019 was one month longer then in 2018.
@@
 For both years, these dates coincident with the first snow-free day at the meteorological weather station at Øvre Neiden (Sør-Varanger, NOR) which is relatively close-by.
-From the latitude model typically used for localizing $\mathrm{DO_3SE}$, we get $A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$ indicating an about $2\,\unit{months}$ longer growing season.
+For comparison, we would get $A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$ from the latitude model typically used for localizing $\mathrm{DO_3SE}$.
+By comparison, the latitude model indicates about a $2\,\unit{month}$ longer growing season.
 Due to a lack of quantitative field observation, we assume a latency period of about $1\,\unit{month}$ after snow-melt for perennial grassland.
@@
 As Svanhovd is located in the bed of the Pasvik river, we assume a sandy loam texture at $60\,\unit{cm}$ depth.
-Further, leaves of downy birch in northern Fennoscandia are smaller than southern specimens'.
-Here, we have measured their dimension and found $(3.0\pm 0.5)\,\unit{cm}$.
-\citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area.
+The estimate of leaf boundary layer resistence is a function of windspeed which will be influenced by the tree height and leaf dimensions.
+Leaves of downy birch in northern Fennoscandia are smaller than southern specimens' with leaf dimensions of $(3.0\pm 0.5)\,\unit{cm}$.
+In addition, \citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area.
+Both these factors will affect leaf boundary layer resistance, the smaller leaves would tend to reduce leaf boundary layer resistances whilst the shorter trees would reduce canopy roughness and increase canopy boundary layer resistances.
 

### 2021-04-18 VegOzone.tex
Improve describtion regarding optimization metric

@@
 The main species or PFTs, we deduced on-site at the NIBIO Environment Centre Svanhovd and from Fig.~\ref{fig:svanhovd_research_station}, are perennial grassland, downy birch, and Norway spruce.
-Default parameters for these species or PFTs are derived from \citet{ICP:MappingManual2017} {\bf TODO: find added reference by Lisa} and will be referred to as \emph{mapping manual} abbreviated as MM in the following.
+Default parameters for these species or PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2010,ICP:MappingManual2017} and will be referred to as \emph{mapping manual} abbreviated as MM in the following.
 A comprehensive list of model parameters can be found in Supplement~Table~S1.
@@
 To quantify the resulting bespoke parameterizations, we propose the following metric.
-For a generic GS (May--August), we compute climatological average and standard deviation of $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal})) at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time).
+For a generic GS (May--August), we compute climatological average and standard deviation of $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal})) at noon (averaged over $11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning (averaged over $5-9\,\unit{am}$ local time).
 We presume that relative stomatal conductance of around noon (highest light intensity) and in the early morning (lower intensities) are good proxies for \chem{CO_2} uptake efficiency.
@@
   \caption{Proposed metric to test bespoke response functions.
-GS (May--August) climatological averages and standard deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time) are shown for (a) Downy birch; (b) Norway spruce; (c) perennial grassland.
+GS (May--August) climatological averages and standard deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon (averaged over $11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning (averaged over $5-9\,\unit{am}$ local time) are shown for (a) Downy birch; (b) Norway spruce; (c) perennial grassland.
 For simplicity, $f_\mathrm{SWP}$ is neglected in this calculation.
@@
 We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for three natural/semi-natural vegetation types, downy birch, Norway spruce, and perennial grassland using both MM parameterizations and a variety of bespoke response function in $f_\mathrm{T}$ and $f_\mathrm{light}$, and localized GS (described in detail in Section~\ref{subsec:do3se_parameters}).
-The results are shown comprehensively in Fig.~\ref{fig:pody_rel}.
-For comparison, critical levels for a $4\,\unit{\%}$ (silver birch), $2\,\unit{\%}$ (Norway spruce), and $10\,\unit{\%}$ (perennial grassland, above ground) reduction in biomass \citep{ICP:MappingManual2017,ESPR:Hayes2021} are shown as horizontal dashed lines.
+The results are shown in Fig.~\ref{fig:pody_rel}.
+For comparison, critical levels for a $4\,\unit{\%}$ (silver birch), $2\,\unit{\%}$ (Norway spruce) \citep{ICP:MappingManual2017}, and $10\,\unit{\%}$ (perennial grassland, above ground) \citep{ESPR:Hayes2021} reduction in biomass are shown as horizontal dashed lines.
 Different years are color coded as previously.

### 2021-04-18 VegOzone.tex
Rewrite the last paragraph of section 4

@@
 physiology and phenology) as well as environmental conditions that can decouple the relationship between ozone concentration, ozone uptake, and consequent damage \citep{PT:Emberson2020}.
-These flux based metrics have been shown to be better suited to represent the actual risk to vegetation from ozone, especially in the climatically extreme parts of Europe \citep{EP:Simpson2007,GCB:Mills2010,ICP:MappingManual2017}.
+These flux based metrics have been shown to be better suited to represent the actual risk to vegetation from ozone, especially in the climatically extreme parts of Europe \citep{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017}.
 Most of these previous studie have focussed on the Mediterranean where soil moisture deficits are thought to be most influential in decoupling concentration from uptake.
@@
 An application of the Mapping Manual method was made by \citet{GCB:Mills2011} where exceedance of the CL was related to observations of ozone damage to a clover biomonitor.
-This study clearly demonstrated the improved ability of flux-based metrics (in comparison to exposure-based metrics) to identify the geographical distribution of the risk of vegetation to ozone damage and also showed that the ozone damage could extend into more northerly regions of Europe.
+This study clearly demonstrated the improved ability of flux-based metrics (in comparison to exposure-based metrics) to identify the geographical distribution of the risk of ozone damage and also showed that the ozone damage could extend into more northerly regions of Europe.
 A similar biomonitoring study was conducted at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd to assess whether ozone was capable of inducing damage to vegetation growing under northern Fennoscandia conditions.
@@
 The main species or PFTs, we deduced on-site at the NIBIO Environment Centre Svanhovd and from Fig.~\ref{fig:svanhovd_research_station}, are perennial grassland, downy birch, and Norway spruce.
-Default parameters for these species or PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2010,ICP:MappingManual2017} and will be referred to as \emph{mapping manual} abbreviated as MM in the following.
+Default parameters for these species or PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as \emph{mapping manual} abbreviated as MM in the following.
 A comprehensive list of model parameters can be found in Supplement~Table~S1.
@@
 The results depicted in Fig.~\ref{fig:pody_rel} allow for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric for ozone induced damage in subarctic vegetation.
-In summary, we identify the following main effects on the model predicted ozone uptake with respect to the MM default parameterization:
+In summary, we identify the following main effects of the bespoke parameterizations of the $DO_3SE$ model with respect to the MM default parameterization on ozone uptake:
 \begin{itemize}
   \itemsep0pt
-\item Increasing cold tolerance/heat intolerance $\Rightarrow$ increased uptake regardless of individual year,
-\item usage of proper growing season $\Rightarrow$ reduced uptake with respect to the above,
+\item Increasing cold tolerance/heat intolerance $\Rightarrow$ increased uptake in both probed years,
+\item usage of proper growing season $\Rightarrow$ reduced uptake with respect to the increase reported above,
 \item an earlier assumed start of GS $\Rightarrow$ increased uptake,
 \item drought conditions (SWP) $\Rightarrow$ reduced uptake,
-\item lower/higher light threshold $\Rightarrow$ increased/reduced uptake,
-\item bespoke light threshold $\Rightarrow$ asymmetric response in POD,
+\item lower/higher extent of stomatal opening at low light intensities $\Rightarrow$ increased/reduced uptake,
+\item symmetric variation of the extent of stomatal opening at low light intensities $\Rightarrow$ asymmetric response in POD,
 \item temperature acclimation $\Rightarrow$ amplifying effect on light threshold and drought effects.
 \end{itemize}
-The exact order of magnitude of these effects varies for individual species or PFTs and between years, but the predicted ozone uptake for the bespoke parameterization is in any case larger than for the MM default parameterizations and of the same order of magnitude as the variability between the years studied here.
-Although our temperature acclimations are hypothetical, this indicates a possibly considerable underestimation of ozone risk for subarctic species.
-Especially, an earlier GS due to climate change, that is overlapping with the ozone spring peak, has the potential to increase the ozone damage risk for vegetation with an acclimation to cold climates.
+The exact order of magnitude of these effects varies for individual species or PFTs and between years, but the predicted ozone uptake for the bespoke parameterization is always larger than for the MM default parameterizations and of the same order of magnitude as the variability between the years studied here.
+Although our temperature acclimations are hypothetical, they suggest considerable underestimations of ozone risk for subarctic species when compared to generic MM parameterisations.
+In particular, an earlier GS due to climate change, that is overlapping with the ozone spring peak, has the potential to increase the risk of ozone damage for vegetation with an acclimation to cold climates.
 Drought effects (SWP) do only matter if a bespoke GS is not taken into account.
@@
 SWP induced reduction in ozone uptake is only effective in 2018 and without bespoke GS.
-The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to the variation of the light threshold of $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
+The temperature acclimation dependent bias in ozone uptake deduced for bespoke GS ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+The maximum systematic deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of the extent of stomatal opening at low light intensities amount to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
 
 For Norway spruce (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
-Norway spruce is the least sensitive to an acclimation in temperature and therefore the least biased.
-Norway spruce displays a medium sensitivity to the light threshold, e.g.
+Norway spruce is the least sensitive to an acclimation in temperature and therefore the least biased compared to the generic default MM parameterization.
+Norway spruce displays a medium sensitivity to the variation in the light threshold, e.g.
 $\Delta\mathrm{POD_1} = ^{+1.7}_{-1.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of boreal type without bespoke GS in 2018.
 Taking the bespoke GS into account, we find a higher ozone uptake in 2019 for all bespoke temperature response functions.
-The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to the variation of the light threshold of $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
+The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+The maximum systematic deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of the extent of stomatal opening at low light intensities amount to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
 
@@
 Perennial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant without bespoke GS.
-The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to the variation of the light threshold of $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
+The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+The maximum systematic deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of the extent of stomatal opening at low light intensities amount to $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
 
@@
 
-The estimated biomass reduction in \unit{\%} for the bespoke temperature response function of type boreal in accordance to \citet{ICP:MappingManual2017, ESPR:Hayes2021} is shown in Table~\ref{tab:biomass_reduction}.
-The estimated uncertainty range corresponds to the deviation given by the $\pm 20\,\unit{\%}$ variations of the light threshold and SWP=on.
-For downy birch and perennial grassland, substantial reductions in total biomass are predicted for any of the two years.
-Norway spruce is least affected.
-The systematic uncertainties for downy birch are one order of magnitude larger than for Norway spruce and perennial grassland.
-For all species or PFTs the interannual variability is larger than the systematic uncertainties estimated for a specific acclimation and is of the same order of magnitude as the divergence between the hypothetical acclimations.
-For comparison, we averaged the corresponding biomass reduction for the MM temperature and light response functions with both SWP on/off but with bespoke GS over both years and find a non negligible, which means above systematic uncertainties, low-bias for both downy birch and perennial grassland with respect to the MM default parameterization.
-For Norway spruce, the default MM parameterization average lies within the given systematic uncertainty range, this may reflect the more comprehensive research on this species or PFT.
+We estimated the biomass reductions in accordance to the CLs in \citet{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme temperature acclimation (\emph{boreal}) for each bespoke PFT.
+The estimates are listed in Table~\ref{tab:biomass_reduction}.
+The reported range in uncertainty corresponds to the maximum deviation defined by the $\pm 20\,\unit{\%}$ variation of the extent of stomatal opening at low light intensities with SWP=on.
+For downy birch and perennial grassland, substantial reductions in total biomass in comparison with the default generic MM parameterizations are predicted for both years.
+Norway spruce is the least affected and the magnitude of total biomass reduction is independent of the choice of the parameterization within given uncertainties.
+This may reflect the more comprehensive research on this species or PFT.
+The systematic deviation for downy birch is one order of magnitude larger than for Norway spruce and perennial grassland.
+
+From the difference between 2018 and 2019, we infer a substantial interannual variability.
+The interannual variability is estimated to be larger than our reported systematic deviations and can be of the same order of magnitude as the effect of the temperature acclimation.
+This means that the choice of temperature acclimation has the potential to either revoke or amplify the risk of ozone damage for vegetation.
+However, the risk of ozone damage in downy birch and perennial grassland in comparison with the generic default MM parametrizations is increased, regardlessly.
 
 \begin{table}[t]
-  \caption{Estimated total biomass reduction in \unit{\%} for bespoke temperature response function of type boreal with bespoke GS and SWP=off.
-The uncertainty ranges reported here correspond to maximum a divergence by varying the light threshold and turning SWP sensitivity on in $\mathrm{DO_3SE}$.
-For comparison, the corresponding biomass reduction for the MM temperature and light response functions, SWP on/off, and bespoke GS averaged over both years is shown with its standard deviation.}
+  \caption{Estimated total biomass reduction in \unit{\%} for temperature acclimation \emph{boreal} with bespoke GS and SWP=off.
+The uncertainty ranges reported here correspond to maximum a divergence deducted from varying the extent of stomatal opening at low light intensities and SWP=on.
+For comparison, the corresponding biomass reduction for the generic default MM averaged over both years is shown.
+The reported standard deviation is computed from all sensitivity simulations ($\pm 20\,\unit{\%}$ in extent of stomatal opening at low light, SWP on/off, and bespoke GS).}
   \label{tab:biomass_reduction}
@@
 \item drought conditions (SWP) $\Rightarrow$ reduced uptake,
-\item lower/higher light threshold $\Rightarrow$ increased/reduced uptake,
-\item bespoke light threshold $\Rightarrow$ asymmetric response in POD,
+\item lower/higher extent of stomatal opening at low light intensities $\Rightarrow$ increased/reduced uptake,
+\item symmetric variation of the extent of stomatal opening at low light intensities $\Rightarrow$ asymmetric response in POD,
 \item temperature acclimation $\Rightarrow$ amplifying effect on light threshold and drought.

### 2021-04-18 VegOzone.tex
Added comments regarding conclusions and discussion

@@
 The exact order of magnitude of these effects varies for individual species or PFTs and between years, but the predicted ozone uptake for the bespoke parameterization is always larger than for the MM default parameterizations and of the same order of magnitude as the variability between the years studied here.
-Although our temperature acclimations are hypothetical, they suggest considerable underestimations of ozone risk for subarctic species when compared to generic MM parameterisations.
+Although our temperature acclimations are hypothetical, they suggest considerable underestimations of ozone risk for subarctic species when compared to generic MM parameterizations.
 In particular, an earlier GS due to climate change, that is overlapping with the ozone spring peak, has the potential to increase the risk of ozone damage for vegetation with an acclimation to cold climates.
@@
 \label{sec:conc}
-We have studied the ozone seasonality for northern Fennoscandia and the quality of its representation in available global and regional reanalysis products for Europe.
-We found that both global reanalysis products of ECWMF fail to reproduce observed ozone paterns.
-Only the CAMSRAQ ensemble reproduces the observed ozone seasonality well, although with a slight low-bias remaining.
-This confirms that high spartial resosultion and assimilation of in situ observations of ground-level ozone concentrations are a must to constrain reanalysis related to air quality and risk assement ozone damage to vegetation, especially in high latitudes.
-
-We have exemplarily studied the effect of acclimations of biomes to a subarctic climate on phyto-toxic ozone uptake.
-We also assessed implications on potential ozone damage risks arising at the advent of climate change.
+
+We have studied the effect of acclimations of biomes to a subarctic climate on phyto-toxic ozone uptake.
+The comparison between 2018 and 2019 conditions allowed us to consider these acclimations and their influence on vegetation sensitivity to \chem{O_3} in light of future changes of key climate variables as may occur under climate change (e.g.
+increase of frequency and extent of heatwaves).
 We found that conditions for ozone formation were more favorable in the 2018 growing season than in 2019, with 2018 being significantly warmer and less cloudy in spring and early summer.
 Accordingly, peak ozone concentrations occurred more frequently and at higher levels in 2018.
-This was in particular caused by the extended heatwave in spring and early summer and associated, extensive wildfires in Central Sweden (in particular in mid--end July).
-
-We found that the length and timing of the growing season is crucial for risk assessment based on accumulated ozone flux metrics ($\mathrm{POD_y}$).
+This was particularly a result of the extended heatwave in spring and early summer and associated, extensive wildfires in Central Sweden (in particular in mid--end July).
+
+We found that the length and timing of the growing season are crucial for risk assessment based on accumulated ozone flux metrics ($\mathrm{POD_y}$).
 Based on a MODIS photosynthesis product, we estimated the growing season for Norway spruce in 2018 to be at least one month shorter than in 2019.
-The found start of growing season ($106/122\,\unit{doy}$) is within given ranges by empirical observations \citep{TB:Kolari2007,IVL:Karlsson2018}.
+The found start of growing season in 2018 ($122\,\unit{doy}$) and 2019 ($106\,\unit{doy}$) are within given ranges by empirical observations \citep{TB:Kolari2007,IVL:Karlsson2018}.
 Based on gridded observational temperature and snow depth, we found a half a month shorter growing season for downy birch and perennial grassland in 2018 compared to 2019, respectively.
-The timing and emerging trend in the start of growing season may differ substantially between local observation and common thermal or latitudinal estimations.
+The timing and emerging trend in the start of the growing season may differ substantially between local observation and common thermal or latitudinal estimations.
 Species-specific methods show a later start for European silver birch and an earlier for Norway spruce, compared to the thermal growing season \citep{IVL:Karlsson2018}.
@@
 To account for missing data from our 2018 record, we proposed a routine for gap filling of ozone observations following a Reynolds's decomposition into anomalies and climatology.
-We also assessed the quality of available global reanalysis (ECMWF MACC and CAM) and regional model reanalysis (Copernicus) ground-level ozone concentrations products and conclude that only the Copernicus product can sufficiently reproduce observed ozone concentrations in northern Fennoscandia.
-
-Although high and more frequent ozone peak concentrations were observed in 2018, ozone concentrations were, in general, not significantly elevated by statistical means.
-It therefore occurs that weather conditions and constantly high ozone concentration are more important in risk assessment using flux-based metrics like $\mathrm{POD_y}$ than episodes of peak concentrations over shorter periods of time.
-Consequently, acute cellular damage caused by very high fluxes over a short period of time is not accounted for properly in $\mathrm{POD_y}$.
+In the course, we also assessed the quality of available global (MACC and CAMS) and regional (CAMS Regional Air Quality (CAMSRAQ) ensemble) reanalysis products for Europe concerning ozone.
+We have studied the ozone seasonality for northern Fennoscandia and the quality of its representation in available global and regional reanalysis products for Europe.
+We confirm that both global reanalysis products do not reproduce the observed ozone patterns in northern Fennoscandia \citep{GMD:Huijnen2020,ACPD:Barten2020}.
+Only the CAMSRAQ ensemble reproduces the observed ozone seasonality, although with a remaining low-bias.
+This confirms that high spatial resolution and assimilation of in situ observations of ground-level ozone concentrations are a must to constrain reanalysis products related to air quality and risk assessment ozone damage to vegetation, especially in high latitudes.
+
+Ozone peak concentrations were observed more frequently in 2018 than in 2019.
+In 2018, we found significantly enhanced \chem{[O_3]} particularly in April and May, but not during the active wildfires in July.
+This suggests that weather conditions and high ozone concentration at the beginning of the growing season are more important in risk assessment using flux-based metrics than episodes of peak concentrations over shorter periods.
+Because these peak \chem{[O_3]} may coincident with environmental conditions that limit ozone uptake and hence do not lead to damage.
+Consequently, acute cellular damage caused by very high fluxes over a short period is not accounted for properly in $\mathrm{POD_y}$.
 As pointed out by \citet{AE:Musselman2006}, a higher temporal resolution, e.g.
-susceptibility to ozone depending on leaf age, in the associated damage functions might give a more differentiate picture.
-For herbal and grassland species (similar for crops) a strong dependency on phenological state, e.g.
+susceptibility to ozone depending on leaf age, in the associated damage functions might give a different picture.
+For herbal and grassland species (similar for crops) a strong dependency on the phenological state, e.g.
 high susceptibility to ozone in the reproductive state, was found \citep{EP:Bassin2004}.
 
-To estimate potential biases in assessment of ozone induced risk on Fennoscandia vegetation, we defined hypothetical temperature response functions representing different levels of assumed acclimation to cold climates.
+To estimate potential biases in the assessment of ozone-induced risk on Fennoscandia vegetation, we defined hypothetical temperature response functions representing different levels of assumed acclimation to cold climates.
 We assume northern Fennoscandia vegetation to be more tolerant towards cold temperatures and less tolerant to heat.
@@
 These hypothetical species or PFTs still need verification by experimental data.
-Our results, however, allowed for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric.
-In particular, we identified the following main effects on the model predicted ozone uptake with respect to the default parameterization:
-\begin{itemize}
+
+In summary, our key findings are:
+\begin{enumerate}
    \itemsep0pt
-\item Increasing cold tolerance/heat intolerance $\Rightarrow$ increased uptake regardless of individual year,
-\item usage of proper growing season $\Rightarrow$ reduced uptake with respect to the above,
-\item an earlier assumed start of growing season $\Rightarrow$ increased uptake,
-\item drought conditions (SWP) $\Rightarrow$ reduced uptake,
-\item lower/higher extent of stomatal opening at low light intensities $\Rightarrow$ increased/reduced uptake,
-\item symmetric variation of the extent of stomatal opening at low light intensities $\Rightarrow$ asymmetric response in POD,
-\item temperature acclimation $\Rightarrow$ amplifying effect on light threshold and drought.
-\end{itemize}
-The exact order of magnitude of these effects varies for individual species or PFTs and between years.
-Only accounting for acclimation, the predicted ozone uptake is typically larger than for default parameterizations.
-This indicates a considerable underestimation of ozone risk for subarctic species, especially for downy birch and perennial grassland.
-However, taking bespoke length and timing of growing season into account the opposite is be the case under certain conditions, e.g.
-a later start of the growing season with respect to the ozone spring peak for downy birch and perennial grassland.
-Simulations of Norway spruce with bespoke (shorter than default) growing season show only little sensitivity to acclimations and inter-annual variation of meteorological conditions.
-We have not assessed the effect of VPD response function here, but our climatological data show that conditions at Svanhovd are shifting towards a more VPD limited regime.
-
-The impact of the timing of the growing season can be of the same order of magnitude as the acclimation of response functions but points into the opposite direction.
-Especially, the trend towards an earlier start of the growing season in northern Fennoscandia with respect to the ozone spring peak induces large uncertainties for future risk assessment and includes potential higher ozone damage risk.
-However, the decline of ozone concentrations in arctic spring is partly caused by the removal of ozone through active vegetation at the start of the growing season.
-It is therefore not clear if the peak concentrations of ozone experienced by vegetation in the future with an earlier start of the growing season, will be reduced compared to present day.
-Hence, a simulated earlier start of growing season with present day ozone concentration might overestimate ozone uptake and hence ozone induced risk of damage.
+ \item the bespoke parameterizations for subarctic species are important (especially concerning light and temperature and the climate effect on the growing season) to be able to assess \chem{O_3} uptake,
+ \item the parameterizations defined in the mapping manual for European regional risk assessment would appear to not adequately capture the physiological responses of subarctic physiology that are important determinants of \chem{[O_3]} sensitivity under prevailing environmental conditions and are likely to substantially underestimate risk,
+ \item the ozone reanalysis products tend to underestimate \chem{[O_3]} in Fennoscandia, and
+ \item that the climatic conditions promoting ozone damage (particularly extended heatwaves and the timing of the growing season concerning the ozone spring peak) and which are likely to become more frequent under climate change are likely to further enhance the sensitivity of subarctic vegetation to ozone.
+\end{enumerate}
+
+However, the decline of the ozone spring peak is partly caused by dry deposition of ozone on vegetation.
+It, therefore, remains unclear whether an earlier start of the growing season has the potential to reduce the ozone spring peak concentration compared to present-day and hence reduce the damage risk of ozone in the future.
+We have not assessed the effect of the plants' response to VPD here, but our climatological data show that conditions at Svanhovd are shifting towards a more VPD limited regime.
 
@@
 For Norway spruce, the results are compatible with the default parameterization estimates and are contained within the range of systematic uncertainties deduced from our bespoke response function ($2.5-3\,\unit{\%}$).
-These results have to be treated with care, because the biomass reduction functions were established based on central European species or PFTs and might as well be subject to biases.
-
-At the same time, intra-species variability even at the same location is typically non negligible \citep{EP:Bassin2004,Amb:Girgzdiene2009}.
+These results have to be treated with care because the biomass reduction functions were not established based on subarctic species and might as well be subject to biases.
+
+At the same time, intra-species variability even at the same location is typically non-negligible \citep{EP:Bassin2004,Amb:Girgzdiene2009}.
 This may explain contradictory results regarding the ozone sensitivity of natural vegetation in northern Europe.
-While \citet{FS:Subramanian2014} reported a modeled reduction of net primary production in Norway spruce ($4.3-15.5\,\unit{\%}$) and birch ($1.4-4.3\,\unit{\%}$) under elevated ozone in Sweden, Lithuanian forest decidous trees were observed to be more susceptible to visible damage by ozone than Norway spruce \citep{Amb:Girgzdiene2009}.
-Open-top chamber (OTC) experiments performed in northern Finland where Scots pine and downy birch were exposed to elevated ozone concentrations showed biomass reductions of both downy birch and mountain birch seedlings after one-season exposure \citep{Amb:Manninen2009}.
+While \citet{FS:Subramanian2014} reported a modeled reduction of net primary production in Norway spruce ($4.3-15.5\,\unit{\%}$) and birch ($1.4-4.3\,\unit{\%}$) under elevated ozone in Sweden, Lithuanian forest deciduous trees were observed to be more susceptible to visible damage by ozone than Norway spruce \citep{Amb:Girgzdiene2009}.
+Open-top chamber (OTC) experiments performed in northern Finland where Scots pine and downy birch were exposed to elevated ozone concentrations showed biomass reductions of both downy birch and mountain birch seedlings after one season of exposure \citep{Amb:Manninen2009}.
 The leaves of the mountain birch in the high ozone treatments showed necrotic stipples as early as June.
@@
 Cumulative ozone uptake has been reported to reduce stomatal conductance significantly and independently from photosynthesis \citep[e.g.]{BGS:Lombardozzi2012}, with photosynthesis being more strongly affected.
-This might be due competing effects, as an ozone induced sluggishness \citep{SR:Hoshika2015} renders stomata inapt to fully close and leads to higher stomatal conductance than expected and thus higher ozone uptake as well as transpiration.
-At present, the $\mathrm{DO_3SE}$ model does not account for any ozone inflicted reduction of stomatal conductance throughout the growing season and is not coupled to photosynthesis.
+This might be due competing effects, as an ozone-induced sluggishness \citep{SR:Hoshika2015} renders stomata unable to fully close and leads to higher stomatal conductance than expected and thus higher ozone uptake as well as transpiration.
+At present, the MM version of the $\mathrm{DO_3SE}$ model does not account for any ozone-inflicted reduction of stomatal conductance throughout the growing season and is not coupled to photosynthesis.
 Hence, the presented $\mathrm{POD_y}$ may be regarded as an estimated upper limit.\\
 
-Beyond the risk assessment of ozone induced damage, our hypothetical bespoke temperature response functions, once varified by observations, can have important implications on land-surface modeling in global models, where problems in productivity of species especially in the Arctic regions occur due to PFTs which are not suitable for the climate.
+Beyond the risk assessment of ozone-induced damage, our hypothetical bespoke temperature response functions, once verified by observations, can have important implications on land-surface modeling in global models, where problems in productivity of species especially in the Arctic regions occur due to PFTs which are not suitable for the climate.
 We pointed out, with respect to default perennial grassland parameterization, that these perennial grasslands did not show any substantial stomatal conductance in 2019 characterized as a normal year.
 In terms of GPP this would mean unrealistically low carbon sequestration.
-An automation of the here proposed PDF-based acclimation using machine learning techniques could overcome these issues in the future.
+Automation of the here proposed PDF-based acclimation using machine learning techniques could overcome these issues in the future.
 

### 2021-04-18 VegOzone.tex
Spell-checked

@@
 Due to data acquisition problems, ozone concentrations for two weeks in July 2018 were missing from our record.
-We present a reconstruction based on a probability density function (PDF) and concidering the Swedish and Finnish atmospheric monitoring sites.
+We present a reconstruction based on a probability density function (PDF) and considering the Swedish and Finnish atmospheric monitoring sites.
 We review available ground level ozone reanalysis products for the region and find only CAMS Regional Air Quality reanalysis ensemble satisfactory.
@@
 These flux based metrics have been shown to be better suited to represent the actual risk to vegetation from ozone, especially in the climatically extreme parts of Europe \citep{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017}.
-Most of these previous studie have focussed on the Mediterranean where soil moisture deficits are thought to be most influential in decoupling concentration from uptake.
+Most of these previous studies have focused on the Mediterranean where soil moisture deficits are thought to be most influential in decoupling concentration from uptake.
 Relatively few studies have explored the situation in Northern European climates where low temperatures may be considered more likely to influence ozone uptake \citep[e.g.]{iF:Juran2018}.
 
-To estimate the flux-based metric, also referred to as the Phytotoxic Ozone Dose over a threshold y ($\mathrm{POD_y}$), an estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf and environmental factors affecting stomatal conductance as described further in Section~\ref{sec:do3se}.
+To estimate the flux-based metric, also referred to as the Phyto-toxic Ozone Dose over a threshold y ($\mathrm{POD_y}$), an estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf and environmental factors affecting stomatal conductance as described further in Section~\ref{sec:do3se}.
 $\mathrm{POD_y}$ is then calculated according to:
@@
 
-An application of the Mapping Manual method was made by \citet{GCB:Mills2011} where exceedance of the CL was related to observations of ozone damage to a clover biomonitor.
+An application of the Mapping Manual method was made by \citet{GCB:Mills2011} where exceedance of the CL was related to observations of ozone damage to a clover bio-monitor.
 This study clearly demonstrated the improved ability of flux-based metrics (in comparison to exposure-based metrics) to identify the geographical distribution of the risk of ozone damage and also showed that the ozone damage could extend into more northerly regions of Europe.
-A similar biomonitoring study was conducted at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd to assess whether ozone was capable of inducing damage to vegetation growing under northern Fennoscandia conditions.
+A similar bio-monitoring study was conducted at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd to assess whether ozone was capable of inducing damage to vegetation growing under northern Fennoscandia conditions.
 In 2018, we indeed observed ozone damage on clover and tobacco.
@@
 This allows an understanding of how the threat from ozone may change over time and which are the key aspects of plant physiology that might determine risk.
-Hence, we determin which parameters require special attention to define accurately for pollution impact assessment and more generally for describing gas exchange response to changing climates that will influence biochemical cycling.
+Hence, we determine which parameters require special attention to define accurately for pollution impact assessment and more generally for describing gas exchange response to changing climates that will influence biochemical cycling.
 This might be one key to solve long-standing issues of coupled earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface models \citep{GMD:Poulter2015,JAMES:Lawrence2019}.\\
@@
 The temporal as well as spatial resolution of these reanalysis products is rather coarse: 3-hourly and $0.75\,\unit{^\circ}\times0.75\,\unit{^\circ}$ or roughly $29.3\,\unit{km}\times 83.4\unit{km}$ at the location of Svanvik.
-From Copernicus member institutions, higher resolution suface ozone ranalysis data are available for Europe (CAMS Regional Air Quality (CAMSRAQ) reanalysis ensemble).
+From Copernicus member institutions, higher resolution surface ozone reanalysis data are available for Europe (CAMS Regional Air Quality (CAMSRAQ) reanalysis ensemble).
 The CAMSRAQ is based on nine European state-of-the-art numerical air quality models.
 The ensemble mean is at fairly high spatial and temporal resolution compared to the global reanalysis: $0.1\,\unit{^\circ}\times 0.1\,\unit{^\circ}$ or roughly $3.9\,\unit{km}\times 11.1\,\unit{km}$ and 1-hourly, respectively.
-The periodes covered differ and no data is available before the turn of the millennium.
+The periods covered differ and no data is available before the turn of the millennium.
 The CAMSRA and CAMSRAQ are available in near real time, but only the former covers a period of sufficient length for climate analysis (2003 -- present).
@@
 \begin{table}[t]
-  \caption{Global/refional ozone reanalysis products used in this study.}
+  \caption{Global/regional ozone reanalysis products used in this study.}
   \label{tab:ozone_rea}
@@
 
-We hypothesize that perennial vegetation will likely prioritize their growth or fast adolescenceand and hence their carbon sequestration, during a short, subarctic growing season.
+We hypothesize that perennial vegetation will likely prioritize their growth or fast adolescence and hence their carbon sequestration, during a short, subarctic growing season.
 Because the mapping manual version of the $\mathrm{DO_3SE}$ model currently does not simulate $A_\mathrm{n}$, we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$.
@@
 As Svanhovd is located in the bed of the Pasvik river, we assume a sandy loam texture at $60\,\unit{cm}$ depth.
-The estimate of leaf boundary layer resistence is a function of windspeed which will be influenced by the tree height and leaf dimensions.
+The estimate of leaf boundary layer resistance is a function of wind speed which will be influenced by the tree height and leaf dimensions.
 Leaves of downy birch in northern Fennoscandia are smaller than southern specimens' with leaf dimensions of $(3.0\pm 0.5)\,\unit{cm}$.
@@
 This means that the choice of temperature acclimation has the potential to either revoke or amplify the risk of ozone damage for vegetation.
-However, the risk of ozone damage in downy birch and perennial grassland in comparison with the generic default MM parametrizations is increased, regardlessly.
+However, the risk of ozone damage in downy birch and perennial grassland in comparison with the generic default MM parameterizations is increased, regardless.
 
@@
   Tore Flatlandsmo Berglen for hourly pressure data from Svanhovd.
-  This work was financed by the Research Council of Norway (NRC) grand No. xxxxx.
+  This work was financed by the Research Council of Norway (NRC) grand No. 268073.
 \end{acknowledgements}

### 2021-04-19 VegOzone.tex
Improve wording
Abstract merged with Lisa's suggestions

@@
 \begin{abstract}
-Here we assess the importance of characterization of key meteorological variables and subarctic biomes on the modeled susceptibility of subarctic vegetation to ozone used in pollution and climate risk assessments.
-Future climate changes will manifest in both direct and indirect effects influencing the species' exposure to air pollution.
-A likely increase in the frequency of persistent heatwaves in the future will inflict heat-related stress while generating favorable conditions for the formation of ground level ozone, e.g.
-due to wildfires.\\
-We have monitored ground level ozone concentrations at a high latitude site in Northern Norway during the growing season 2018/19.
-Due to data acquisition problems, ozone concentrations for two weeks in July 2018 were missing from our record.
-We present a reconstruction based on a probability density function (PDF) and considering the Swedish and Finnish atmospheric monitoring sites.
-We review available ground level ozone reanalysis products for the region and find only CAMS Regional Air Quality reanalysis ensemble satisfactory.
-In terms of anomalies from a multi-annual mean, ozone concentrations were only significantly enhanced in April/May 2018 before the start of the growing season.
-Temperatures and global irradiance showed consecutively significant positive anomalies during Spring and Summer 2018.
-In 2019 only early spring and late summer were significantly warmer, while colder during summer.
-Precipitation was lower in both years but displayed a large variability.\\
-We investigate the difference in uptake of ozone using the $\mathrm{DO_3SE}$ model, for typical vegetation at the location.
-We find that bespoke parameterization for subarctic vegetation is essential for an accurate estimation of Phyto-toxic Ozone Dose (POD) and an assessment of damage risk.
-Accurate estimates of the start and length of growing season are also crucial (up to $15\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ deviation in $\mathrm{POD_1}$).\\
-We find that temperature and its derivative vapor pressure deficit are likely the most important climate variables while the variability of ozone concentration within climatological reasonable limits plays a minor role.
-The use of subarctic parameterization suggests an improvement in risk assessment as POD values are increased compared to standard expectations; corroborated by the evidence of visible injury in the Svanhovd ozone garden.
-We estimate a total biomass reduction in Norway spruce of $2.5-3\,\unit{\%}$, downy birch by $15.5-17.4\,\unit{\%}$ and an above-ground biomass reduction in perennial grassland of $12.4-13.7\,\unit{\%}$.
+The subarctic region is likely to be at threat from the combined impacts of climate change and ozone pollution due to its unique atmospheric mixing characteristics (which lead to peak ozone concentrations in spring), the influence of the jet stream on regional climate, and its unique vegetation bio-types that are acclimatized to extremes of cold and midnight sun.
+Unfortunately, to date, only a very few studies have been conducted to assess the response of vegetation to variations in climate and air pollution profiles, at least by comparison with other parts of Europe.
+This study looks to fill this knowledge gap by examining trends in climate variables and ozone pollution that have occurred over the past few decades.
+These climatological and ozone data are used to assess the extent to which two recent years (2018 and 2019) deviate from climatic and pollution norms and how these current (and potentially more frequent future) deviations may influence ozone damage to subarctic vegetation.
+To this end, the quality of available global and regional ozone reanalysis products for Europe is assessed.
+The analysis suggests that only the CAMS Regional Air Quality (CAMSRAQ) reanalysis ensemble sufficiently reproduces the observed seasonal cycle while systematically underestimating \chem{O_3} concentration in the region.
+We find that 2018 is an anomalous warm and bright year (in particular in spring and early summer) with higher than average \chem{O_3} concentrations in April/May and a high number of episodic peak \chem{O_3} concentration above $40\,\unit{ppb}$ in June--August, the latter in part attributable to an increase in forest fires in the Northern Hemisphere as well as warmer and sunnier conditions.
+We apply a flux-metric to determine ozone risk and damage to vegetation as a function of ozone concentration, climate variables, and species-specific physiology.
+We find that parameterization of this flux-metric for subarctic vegetation bio-types increases estimates of ozone-induced reduction of total biomass by between $2.5$ and $17.4\,\unit{\%}$ suggesting that the UNECE LRTAP assessments that use generic parameterizations would underestimate ozone risk in this region.
+We conclude that appropriate parameterization for the growing season, and plant physiological responses to light and temperature are particularly important for accurate modeling of seasonal gas exchange and that efforts should be targeted towards accurately defining subarctic bio-type physiological response to climatic variables to improve regional and global scale biogeochemical cycling under current and future climates.
 \end{abstract}
@@
 Ozone measurements at Pallas are taken on top of mount Sammaltunturi $100\,\unit{m}$ above treeline (pine forest) and mainly consists of low vascular plants, mosses, and lichen \citep{BER:Hatakka2003}.
-At Svanhovd, the ozone measurements are taken amidst a meadow surrounded by agriculturally managed herbaceous plants and grass adhering to the WMO standards {\bf TODO: Reference \url{https://library.wmo.int/index.php?id=12407&lvl=notice_display}} (Fig.\ref{fig:svanhovd_research_station}).
+At Svanhovd, the ozone measurements are taken amidst a meadow surrounded by agriculturally managed herbaceous plants and grass adhering to the \citet{WMOGuide2018} standards (Fig.\ref{fig:svanhovd_research_station}).
 
@@
 This can be partly explained by very high light response factors (over $f_\mathrm{light}>0.8$) in the beginning of the GS in these simulations.
-Whether this is a bug or a feature of the $\mathrm{DO_3SE}$ model is not yet clear.
+Whether this is a bug or a feature of the $\mathrm{DO_3SE}$ model is not clear.
 SWP induced reduction in ozone uptake is only effective in 2018 and without bespoke GS.
@@
 \dataavailability{All data is available from public databases or through institutional access.
-$\mathrm{DO_3SE}$ modeling results can be made available upon request {\bf TODO: Upload to an achieve together with processed, gap filled input data?}.} %% use this section when having only data sets available
+$\mathrm{DO_3SE}$ modeling results can be made available upon request {\bf TODO: Upload to an achieve together with processed, gap filled input data on zenodo?}.} %% use this section when having only data sets available
 
@@
 
-\sampleavailability{TEXT} %% use this section when having geoscientific samples available
+%\sampleavailability{TEXT} %% use this section when having geoscientific samples available
 

### 2021-04-19 VegOzone.tex
Rewrite contributions

@@
 \authorcontribution{SF has written the manuscript, collected and processed ozone and environmental data, and performed all statistical analyses.
-AVV helped in framing the research article, has performed on-site observation of vegetation damage induced by ozone, advised with regard to plant physiological processes and contributed text to the introduction.
-LE has contributed to discussions, setup and evaluation of $\mathrm{DO_3SE}$ modeling results, and development of the PDF-based temperature acclimation.
-CO has collected species or PFT parameters from literature, performed all $\mathrm{DO_3SE}$ simulations and initial graphical checks of the results.
-ALEB contributed with her experience regarding subarctic vegetation in Finnmark.
-FS contributed with a summary of the 2018 meteorological conditions.
+AVV contributed to framing this research article, has performed the on-site observation of vegetation damage induced by ozone, and advision in plant physiological processes.
+LE has contributed to framing of the this article, $\mathrm{DO_3SE}$ modeling, and development of the PDF-based temperature acclimation.
+CO has collected species or PFT parameters from literature, performed all $\mathrm{DO_3SE}$ simulations and initial validation of the results.
+AE contributed with her experience regarding subarctic vegetation in Finnmark.
+FS contributed with an assessment of the 2018 meteorological conditions.
 TB gave guidance in a broader research sense.

### 2021-05-05 VegOzone.tex
Improved fig04 and changed caption accordingly.
Update to latest version of coprnicus style
Fix Aud's name
Restructure and refine abstract

@@
 % Ocean Science (os)
+% Polarforschung - Journal of the German Society for Polar Research (polf)
 % Primate Biology (pb)
@@
 \Author[2]{Ane Victoria}{Vollsnes}
-\Author[2]{Aud E. B.}{Eriksen}
+\Author[2]{Aud}{Eriksen}
 \Author[3]{Lisa}{Emberson}
@@
 
-%% If an author is deceased, please mark the respective author name(s) with a dagger, e.g. "\Author[2,$\dag$]{Anton}{Aman}", and add a further "\affil[$\dag$]{deceased, 1 July 2019}".
+%% If an author is deceased, please mark the respective author name(s) with a dagger, e.g. "\Author[2,$\dag$]{Anton}{Smith}", and add a further "\affil[$\dag$]{deceased, 1 July 2019}".
 
 %% If authors contributed equally, please mark the respective author names with an asterisk, e.g.
-"\Author[2,*]{Anton}{Aman}" and "\Author[3,*]{Bradley}{Bman}" and add a further affiliation: "\affil[*]{These authors contributed equally to this work.}".
+"\Author[2,*]{Anton}{Smith}" and "\Author[3,*]{Bradley}{Miller}" and add a further affiliation: "\affil[*]{These authors contributed equally to this work.}".
 
@@
 \begin{abstract}
-The subarctic region is likely to be at threat from the combined impacts of climate change and ozone pollution due to its unique atmospheric mixing characteristics (which lead to peak ozone concentrations in spring), the influence of the jet stream on regional climate, and its unique vegetation bio-types that are acclimatized to extremes of cold and midnight sun.
-Unfortunately, to date, only a very few studies have been conducted to assess the response of vegetation to variations in climate and air pollution profiles, at least by comparison with other parts of Europe.
-This study looks to fill this knowledge gap by examining trends in climate variables and ozone pollution that have occurred over the past few decades.
-These climatological and ozone data are used to assess the extent to which two recent years (2018 and 2019) deviate from climatic and pollution norms and how these current (and potentially more frequent future) deviations may influence ozone damage to subarctic vegetation.
-To this end, the quality of available global and regional ozone reanalysis products for Europe is assessed.
-The analysis suggests that only the CAMS Regional Air Quality (CAMSRAQ) reanalysis ensemble sufficiently reproduces the observed seasonal cycle while systematically underestimating \chem{O_3} concentration in the region.
-We find that 2018 is an anomalous warm and bright year (in particular in spring and early summer) with higher than average \chem{O_3} concentrations in April/May and a high number of episodic peak \chem{O_3} concentration above $40\,\unit{ppb}$ in June--August, the latter in part attributable to an increase in forest fires in the Northern Hemisphere as well as warmer and sunnier conditions.
-We apply a flux-metric to determine ozone risk and damage to vegetation as a function of ozone concentration, climate variables, and species-specific physiology.
-We find that parameterization of this flux-metric for subarctic vegetation bio-types increases estimates of ozone-induced reduction of total biomass by between $2.5$ and $17.4\,\unit{\%}$ suggesting that the UNECE LRTAP assessments that use generic parameterizations would underestimate ozone risk in this region.
-We conclude that appropriate parameterization for the growing season, and plant physiological responses to light and temperature are particularly important for accurate modeling of seasonal gas exchange and that efforts should be targeted towards accurately defining subarctic bio-type physiological response to climatic variables to improve regional and global scale biogeochemical cycling under current and future climates.
+  The unique vegetation biomes of the subarctic region acclimatized to extremes of cold and midnight sun are likely to be at threat from the combined impacts of climate change and increasing ozone (\chem{O_3}) concentrations.
+The atmospheric and climatic characteristics of the subarctic are known to lead to pronounced peak \chem{O_3} concentrations in spring.
+To date, only a few studies have been conducted to assess the response of subarctic vegetation to variations in climate and air pollution, at least by comparison with other parts of Europe.
+This study looks to fill this knowledge gap by examining trends in climate variables and \chem{O_3} concentrations that have occurred over the past few decades.
+These climatological and \chem{O_3} data are used to assess the extent to which two recent years (2018 and 2019) deviate from climatic and \chem{O_3} concentration norms and how these current (and potentially more frequent future) deviations may influence \chem{O_3} damage to subarctic vegetation.
+We find that 2018 was an anomalous warm and bright year (in particular in spring and early summer) with higher than average \chem{O_3} concentrations in April/May and a high number of episodic peak \chem{O_3} concentration above $40\,\unit{ppb}$ in June--August, the latter in part attributable to an increase in forest fires in the Northern Hemisphere as well as warmer and sunnier conditions.
+We apply a flux-metric to determine \chem{O_3} risk and damage to vegetation as a function of \chem{O_3} concentration, climate variables, and species-specific physiology.
+We find that bespoke parameterizations of plant fuctional types (PFTs) for subarctic vegetation bio-types increase estimates of ozone-induced reduction of total biomass by $2.5$ to $17.4\,\unit{\%}$ in this flux-metric.
+Our study, suggests that assessments that use generic parameterizations are likely to underestimate the risk of \chem{O_3} damage in this region.
+We conclude that appropriate parameterization for the growing season, and plant physiological responses to light and temperature are particularly important for accurate modeling of seasonal gas exchange.
+Efforts should be targeted towards accurately defining subarctic bio-type physiological response to climatic variables to improve regional and global scale biogeochemical cycling under current and future climates.
 \end{abstract}
 
-%\copyrightstatement{TEXT}
+%\copyrightstatement{TEXT}  %% This section is optional and can be used for copyright transfers.
 
@@
 Data from EBAS until December 2018.
-The hatched areas indicate periods with insufficient quality control according to \citet{NILU2003}.}
+The hatched areas indicate periods with insufficient quality control according to \citet{NILU2003}.
+(a) Esrange (SWE); (b) Pallas (FIN); (c) Jergul/Karasjok; (d) Svanvik; (e) Janiskoski.}
   \label{fig:ozone_timesseries_fenoscandic_obs}

### 2021-05-05 VegOzone.tex
Refined Section 1. Checked with grammarly

@@
 We apply a flux-metric to determine \chem{O_3} risk and damage to vegetation as a function of \chem{O_3} concentration, climate variables, and species-specific physiology.
-We find that bespoke parameterizations of plant fuctional types (PFTs) for subarctic vegetation bio-types increase estimates of ozone-induced reduction of total biomass by $2.5$ to $17.4\,\unit{\%}$ in this flux-metric.
+We find that bespoke parameterizations of plant functional types (PFTs) for subarctic vegetation bio-types increase estimates of ozone-induced reduction of total biomass by $2.5$ to $17.4\,\unit{\%}$ in this flux-metric.
 Our study, suggests that assessments that use generic parameterizations are likely to underestimate the risk of \chem{O_3} damage in this region.
-We conclude that appropriate parameterization for the growing season, and plant physiological responses to light and temperature are particularly important for accurate modeling of seasonal gas exchange.
+We conclude that appropriate parameterization for the growing season and plant physiological responses to light and temperature are particularly important for accurate modeling of seasonal gas exchange.
 Efforts should be targeted towards accurately defining subarctic bio-type physiological response to climatic variables to improve regional and global scale biogeochemical cycling under current and future climates.
@@
 
-ground-level ozone is a highly toxic pollutant known to cause damage to of humans health \citep{WHO2008} and to a variety of ecosystems around the World \citep{PT:Emberson2020}.
-Ozone has been found to be a cause of visible injury, photosynthetic damage, early senescence as well as programmed cell death of plants \citep{PCE:Kangasjarvi2005}.
+Ground-level ozone (\chem{O_3}) is a highly toxic pollutant known to cause damage to human health \citep{WHO2008} and to a variety of ecosystems around the World \citep{PT:Emberson2020}.
+Ozone is a cause of visible injury, photosynthetic damage, early senescence as well as programmed cell death of plants \citep{PCE:Kangasjarvi2005}.
 Annual global yield losses of four major crops (wheat, rice, maize, and soybean) of $3-15\,\unit{\%}$ \citep{PJ:Ainsworth2017} and a suggested loss in primary production in forestry of $7\,\unit{\%}$ \citep{GCB:Wittig2009,EP:Matyssek2012} has been attributed to ozone.
 The crop yield loss studies indicate a threat to food security in rapidly developing countries, e.g., in East and South-East Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018}, where ozone concentrations related to enhanced pollutant emissions from the transport, residential, power generation, and industrial sectors have increased in the past few decades.
-Even though long-term (pre-1950) observations of ground-level ozone concentrations are scarce; evidence from trend analysis of those few sites suggests that since the industrial revolution, background concentrations in the northern hemisphere have at least doubled and continued to increase until relatively recently \citep[Chapter 2]{IPCC2013}.
-Recent trend analyses indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric ozone started to level off or even decline in Europe \citep{ESA:Cooper2014, ACP:Wespes2018,ESA:Gaudel2018}.
-This can be attributed to a successful implementation of air quality regulations coupled with economic restructuring which saw a move from coal to gas, reducing the number of episodes of peak concentrations especially in summertime over Europe and North America \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
-However, changes in environmental conditions associated with climate change, such as an increase in the frequency of heatwaves, could negate the effectiveness of emission reductions in relation to ozone impacts and air quality\citep{NCC:Lin2020}.
-
-Ground-level ozone is formed in complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) and hydrocarbons are known as volatile organic compounds (VOCs) in the presence of nitrogen oxides (\chem{NO_x}).
-Despite a relatively short tropospheric lifetime of approximately $22\,\unit{days}$, ranging from a few days in the tropical boundary layer to up to $1\,\unit{year}$ in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}, ozone and is precursors are subject to advection over long distances.
+Even though long-term (pre-1950) observations of ground-level ozone concentrations are scarce; evidence from trend analysis of those few sites suggests that since the industrial revolution, background concentrations in the northern hemisphere have at least doubled and continued to increase \citep[Chapter 2]{IPCC2013}.
+Recent trend analyses for Europe indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric ozone started to level off or even decline \citep{ESA:Cooper2014, ACP:Wespes2018,ESA:Gaudel2018}.
+This can be attributed to a successful implementation of air quality regulations coupled with economic restructuring, reducing the number of episodes of peak concentrations especially in summertime over Europe and North America \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
+However, changes in environmental conditions associated with climate change, such as an increase in the frequency of heatwaves, could negate the effectiveness of emission reductions in relation to ozone impacts and air quality \citep{NCC:Lin2020}.
+
+Complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) and hydrocarbons known as volatile organic compounds (VOCs) in the presence of nitrogen oxides (\chem{NO_x}) lead to the formation of \chem{O_3} at ground-level.
+Despite a relatively short average tropospheric lifetime of approximately $22\,\unit{days}$, ranging from a few days in the tropical boundary layer to up to $1\,\unit{year}$ in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}, ozone and is precursors are subject to advection over long distances.
 Episodes of high ozone concentrations in the absence of local pollution sources can therefore often be attributed to the long-range transport of both ozone and its precursors.
 In northern Fennoscandia, episodes of enhanced ozone concentrations in 2003 and 2006 have been traced back to ozone precursor emissions related to forest fires in southern and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}.
-Besides, hydrocarbons are also emitted by vegetation itself in the form of terpenes and monoterpenes, so-called biogenic volatile substances (BVOCs).
-These emissions are thought to be a response to other factors, e.g., thermal stress, defense against herbivores, or even attraction of pollinators \citep{TPS:Penuelas2003}.
-
-In the annual cycle in Boreal and also temperate climates, a distinct maximum of ground-level ozone concentrations typically occurs in spring followed by lower concentrations throughout the summer.
+Besides VOCs from anthropogenic sources, hydrocarbons are also emitted by vegetation itself in the form of terpenes and monoterpenes, so-called biogenic volatile substances (BVOCs).
+These emissions are thought to be a response to factors, such as thermal stress, defense against herbivores, or even attraction of pollinators \citep{TPS:Penuelas2003}.
+
+In the annual cycle in Boreal climates, a distinct maximum of ground-level ozone concentrations typically occurs in spring followed by lower concentrations throughout the summer.
 Several conditions unique to the Arctic are stimulating the formation of this spring peak.
@@
 Results from open-top chamber (OTC) experiments on downy birch (\emph{Betula pubescens}) and mountain birch (\emph{Betula pubescens toruosa}), native to subarctic regions, indicated reductions in both biomass, in root:shoot ratio, and visible leaf damage under elevated ozone treatment ($\left<\chem{[O_3]}\right> = 36-54\,\unit{ppb}$) \citep{Amb:Manninen2009}.
-Although considered to be more ozone tolerant due to an absence of visible injuries \citep{Amb:Girgzdiene2009}, \citet{Amb:Manninen2009} also found chlorophyll:carotenoid ratio and polyamines reductions in Scots pine (\emph{Pinus sylvestris}), indicating susceptibility to ozone in these species.\\
+Although Scots pine (\emph{Pinus sylvestris}) is considered to be more ozone tolerant due to an absence of visible injuries \citep{Amb:Girgzdiene2009}, \citet{Amb:Manninen2009} found chlorophyll:carotenoid ratio and polyamines reductions under elevated ozone concentrations indicating susceptibility to ozone also in these species.\\
 
 A substantial body of evidence exists that suggests flux-based metrics, that relate stomatal ozone uptake to vegetation damage, are biologically more relevant for risk assessments than exposure-based metrics.
-This is because they are able to account particular species characteristics (i.e.
+This is because they can account for particular species characteristics (i.e.
 physiology and phenology) as well as environmental conditions that can decouple the relationship between ozone concentration, ozone uptake, and consequent damage \citep{PT:Emberson2020}.
-These flux based metrics have been shown to be better suited to represent the actual risk to vegetation from ozone, especially in the climatically extreme parts of Europe \citep{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017}.
+These flux-based metrics are better suited to represent the actual risk to vegetation from ozone, especially in the climatically extreme parts of Europe \citep{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017}.
 Most of these previous studies have focused on the Mediterranean where soil moisture deficits are thought to be most influential in decoupling concentration from uptake.
@@
 %
-with the hourly averaged stomatal ozone flux $\Phi_\mathrm{sto}$ (see Eq.~(\ref{eq:flux_stomata}) and a stomatal ozone flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}\,s^{-1}}.
+with the hourly averaged stomatal ozone flux $\Phi_\mathrm{sto}$ (see Eq.~(\ref{eq:flux_stomata})) and a stomatal ozone flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}\,s^{-1}}.
 The flux threshold $\mathrm{y}$ represents the detoxification potential of the plant and is typically only exceeded during daylight hours (i.e.
@@
 The $\mathrm{POD_y}$ metric is currently used in risk assessments under the United Nations Economic Commission for Europe (UNECE) Long Range Transboundary Air Pollution (LRTAP) Convention to identify those locations across Europe where vegetation is at risk from ozone.
-This is with a view to developing effects-based emission reduction policy that can target those ozone precursor emissions that are most influential in causing damage.
+This convention aims to develope an effects-based emission reduction policy that can target those ozone precursor emissions that are most influential in causing damage.
 To achieve this the concept of critical levels (CLs) has been developed and applied \citep{Maas2016}.
@@
 An application of the Mapping Manual method was made by \citet{GCB:Mills2011} where exceedance of the CL was related to observations of ozone damage to a clover bio-monitor.
-This study clearly demonstrated the improved ability of flux-based metrics (in comparison to exposure-based metrics) to identify the geographical distribution of the risk of ozone damage and also showed that the ozone damage could extend into more northerly regions of Europe.
-A similar bio-monitoring study was conducted at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd to assess whether ozone was capable of inducing damage to vegetation growing under northern Fennoscandia conditions.
+This study demonstrated the improved ability of flux-based metrics (in comparison to exposure-based metrics) to identify the geographical distribution of the risk of ozone damage and also showed that the ozone damage could extend into more northerly regions of Europe.
+A similar bio-monitoring study was conducted at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd to assess whether local ozone concentrations were capable of inducing damage to vegetation growing under northern Fennoscandia conditions.
 In 2018, we indeed observed ozone damage on clover and tobacco.
@@
 According to a report by the Norwegian Meteorological institute \citep{MetNOR2019}, the summer of 2018 was the warmest and driest ever recorded in eastern, western, and southern Norway.
-In the north (including Finnmark), it was amongst the warmest on record -- favorable conditions for planetary boundary layer ozone formation.
-An unusually weak and northward shifted jet stream allowed for a persistent high-pressure system above northern Europe, including Fennoscandia which blocked the pressure systems for several consecutive months.
+In the north (including Finnmark), it was amongst the warmest on record -- favorable conditions for ground-level ozone formation.
+An unusually weak and northward shifted jet stream allowed for a persistent high-pressure system above northern Europe, including Fennoscandia which blocked the low-pressure systems for several consecutive months.
 In the period May -- July, southern Norway had temperatures $4\,\unit{^\circ C}$ above normal.
@@
 These conditions gave rise to massive forest fires in different parts of Europe and thus an increase in ozone precursors.
-Boreal wildfires emit in addition to \chem{CO_2} also \chem{CO} ($\chem{[CO]}/\chem{[CO_2]} \propto 6-13\,\unit{\%}$) and \chem{VOCs} ($\chem{[VOCs]}/\chem{[CO_2]} \propto 0.5-1.5\,\unit{\%}$) \citep{AE:Cofer1990}.
+Boreal wildfires emit in addition to \chem{CO_2} also \chem{CO} ($\chem{[CO]}/\chem{[CO_2]} \propto 6-13\,\unit{\%}$) and VOCs ($\chem{[VOCs]}/\chem{[CO_2]} \propto 0.5-1.5\,\unit{\%}$) \citep{AE:Cofer1990}.
 A total of 2079 forest fires of different sizes were registered in Norway in 2018, twice as many as in the preceding years 2016/17 \citep{DSB2019}, last accessed April 2020).
-In Sweden, about 500 fires had been reported (five times more than in a usual summer), and an estimated total of $25000\,\unit{hectare}$ burned down in the course of what became known as the most extensive fires in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) in younger history \citep{SOU2019}.
+In Sweden, about 500 fires had been reported (five times more than in a usual summer), and an estimated total of $25000\,\unit{hectare}$ burned down in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) \citep{SOU2019}.
 Coincident peak \chem{[O_3]} is found in ozone monitoring data from Svanhovd but also over whole northern Fennoscandia in July (see Fig.~\ref{fig:data_svanvik_2018_2019} or Fig.~\ref{fig:ozone_reconstruction_2018_07}).
-In the 2003 drought period, elevated ozone in Europe was promoted by a combination of various factors such as wildfires, reduced cloud cover (increased solar radiation), reduced dry deposition and turbulent mixing due to the stagnant weather conditions, and increased BVOC emissions \citep{JGR:Solberg2018}.
-It is likely that these factors also contributed to enhanced \chem{[O_3]} in 2018.
-
-The methods described above rely on accurate representation of plant physiological response to climate variables (i.e.
-the species or plant functional type-specific responses used to parameterize risk assessment models) and are most often parameterized using data for less extreme climates, i.e., temperate or continental climates; and to a lesser extent, Boreal and Mediterranean climates.
+During the 2003 drought period, elevated ozone in Europe was promoted by a combination of various factors such as wildfires, reduced cloud cover (increased solar radiation), reduced dry deposition and turbulent mixing due to the stagnant weather conditions, and increased BVOC emissions \citep{JGR:Solberg2018}.
+It is likely that these factors also contributed to enhanced \chem{[O_3]} in 2018.\\
+
+The described methods for risk assessment rely on accurate representation of plant physiological response to key climate variables and are most often parameterized using data for less extreme climates, i.e., temperate or continental climates; and to a lesser extent, Boreal and Mediterranean climates.
 Parameterization of these models for more extreme bio-geographical regions such as subarctic climates has not been performed and relies on assuming that less extreme Boreal parameterization will be applicable.
-In this study, conditions at Svanhovd for the growing seasons 2018/19 shall serve as a reference for present and probable future conditions in northern Fennoscandia.
 It is important to understand the implications of using these more generic parameterizations for more extreme climate regions.
-Here we explore how risk estimates might be over or underestimated using generic vs bespoke subarctic parameterization.
+We explore how risk estimates might be over or underestimated using generic vs bespoke subarctic parameterization.
 We also explore the role of \chem{[O_3]} variability and changing climate variables to understand how this risk might alter in the future with the advent of climate change.
-This allows an understanding of how the threat from ozone may change over time and which are the key aspects of plant physiology that might determine risk.
-Hence, we determine which parameters require special attention to define accurately for pollution impact assessment and more generally for describing gas exchange response to changing climates that will influence biochemical cycling.
-This might be one key to solve long-standing issues of coupled earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface models \citep{GMD:Poulter2015,JAMES:Lawrence2019}.\\
+To this end, conditions at Svanhovd for the growing seasons 2018/19 shall serve as a reference for present and probable future conditions in northern Fennoscandia.
+This allows an understanding of how the threat from ozone may change in the future and which are the key aspects of plant physiology that might determine the potential risk.
+Hence, we determine which parameters require special attention to define accurately for pollution impact assessment and more generally for describing gas exchange response to a changing climate that will influence biochemical cycling.
+This might also be one key to solve long-standing issues of earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface models \citep{GMD:Poulter2015,JAMES:Lawrence2019}.\\
 
 The observation site at NIBIO Environment Centre Svanhovd and all related data are presented in Section~\ref{sec:data}.
-In Section~\ref{sec:stats}, we derive climatologies for ground-level ozone concentration at Svanhovd and northern Fennoscandia based on in situ observations and also compute climatologies of near-surface temperature, precipitation, and global radiation for Svanhovd.
-Based on these, we evaluate anomalies of key environmental variables in 2018/19 by statistical means.
-The derived ground-level ozone concentration climatology for northern Fennoscandia will also be compared with available ground-level ozone reanalysis products.
+In Section~\ref{sec:stats}, we derive multi-annual means (referred to as climatologies) for ground-level ozone concentration and environmental key variables (temperature, precipitation, and global irradiance) based on in situ observations and evaluate the deviation of 2018/19 from the norm (referred to as anomalies).
 We use the $\mathrm{DO_3SE}$ model to estimate the ozone uptake by natural and semi-natural vegetation to investigate the main environmental drivers for the observed ozone damage in 2018 (Section~\ref{sec:do3se}).
 In this context, we introduce the $\mathrm{DO_3SE}$ model equations and derive bespoke parameterizations for subarctic species to assess systematic uncertainties arising from the use of default generic parameterizations established from less climatically extreme conditions.
-In Section~\ref{sec:conc} finally, we will summarize our results and discuss how this research can support improvements in future assessments of pollution risk and more generally bio-geochemical cycling.
+In Section~\ref{sec:conc}, we will summarize our results and discuss how this research can support improvements in future assessments of pollution risk and more generally bio-geochemical cycling.
 

### 2021-05-06 VegOzone.tex
Restructure Section 2 - not spell checked

@@
 
-The described methods for risk assessment rely on accurate representation of plant physiological response to key climate variables and are most often parameterized using data for less extreme climates, i.e., temperate or continental climates; and to a lesser extent, Boreal and Mediterranean climates.
+The described methods for risk assessment rely on accurate representation of plant physiological response to key environmental variables and are most often parameterized using data for less extreme climates, i.e., temperate or continental climates; and to a lesser extent, Boreal and Mediterranean climates.
 Parameterization of these models for more extreme bio-geographical regions such as subarctic climates has not been performed and relies on assuming that less extreme Boreal parameterization will be applicable.
@@
 The observation site at NIBIO Environment Centre Svanhovd and all related data are presented in Section~\ref{sec:data}.
-In Section~\ref{sec:stats}, we derive multi-annual means (referred to as climatologies) for ground-level ozone concentration and environmental key variables (temperature, precipitation, and global irradiance) based on in situ observations and evaluate the deviation of 2018/19 from the norm (referred to as anomalies).
+In Section~\ref{sec:stats}, we derive multi-annual means (referred to as climatologies) for ground-level ozone concentration and key environmental variables (temperature, precipitation, and global irradiance) based on in situ observations and evaluate the deviation of 2018/19 from the norm (referred to as anomalies).
 We use the $\mathrm{DO_3SE}$ model to estimate the ozone uptake by natural and semi-natural vegetation to investigate the main environmental drivers for the observed ozone damage in 2018 (Section~\ref{sec:do3se}).
@@
 %%
-\section{Data}
+\section{Data acquisition of ozone and key environmental variables}
 \label{sec:data}
 
-In our assessment of primary drivers of ozone damage risk on vegetation in the subarctic, we focus on comparing the anomalous warm, bright and dry 2018 growing season with the more climatically average 2019 season to understand the role these environmental conditions may have played in influencing the sensitivity of vegetation to ozone damage.
-The NIBIO operated Environment Centre Svanhovd which lies close by the settlement of Svanvik in the Pasvik valley serves as the main reference for conditions in northern Fennoscandia.
-Throughout the paper, we will use Svanhovd synonymously for the atmospheric monitoring site at Svanvik/Pasvik.
-The available meteorological monitoring data are described in Section~\ref{subsec:atmo_svanvik}.
-Other data used for cross-evaluation of the ozone monitoring data taken at Svanhovd is described in Section~\ref{subsec:other_data}.
-We, in particular, compare with observational data from other monitoring sites in northern Fennoscandia (Section~\ref{subsubsec:ebas}) and use these to evaluate reanalysis products for ground-level ozone (Section~\ref{subsubsec:ozone_rea}).
-
-\subsection{NIBIO Environment Centre Svanhovd and atmospheric monitoring}
-\label{subsec:atmo_svanvik}
+In our assessment of primary drivers of ozone damage risk on vegetation in the subarctic, we focus on conditions during the 2018/19 growing season at the NIBIO operated Environment Centre Svanhovd.
+To this end, we collected ozone and atmospheric monitoring data taken at Svanhovd from different distributors.
+The main reasons for this spread are varring update frequencies and length of available data sets.
+In terms of ozone, we also use observational data from monitoring sites in northern Finland and Sweden for cross-evaluation.
+Svanhovd synonymously refers to the atmospheric monitoring site at Svanvik/Pasvik throughout the paper.
+
+%\subsection{NIBIO Environment Centre Svanhovd and bio-monitoring}
+%\label{subsec:atmo_svanvik}
 The foundation of the NIBIO Environment Centre Svanhovd dates back to the 1930s \citep{NIBIO_Svanhovd}.
@@
 $2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths.
-The locations of the atmospheric monitoring site and the ozone garden are marked in aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}a).
-Svanhovd is located south of the settlement of Svanvik, where the Pasvik river, which is the border to Russia, lies only a few hundred meters to the east.\\
+
+To relate observed visible damage on ozone sensitive plants to local ozone concentration (\chem{[O_3]}), both an ozone bio-monitor (referred to as ozone garden) and a conventional ozone monitor using the existing atmospheric monitoring infrastructure had been installed during the 2018/19 growing season at the NIBIO centre.
+The locations of the atmospheric monitoring site and the ozone garden are marked in the aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}a).
+Svanhovd is located south of the settlement of Svanvik, where the Pasvik river, which is the border to Russia, lies only a few hundred meters to the east.
+This ozone garden was likely the northern-most to date {\bf TODO: Reference?}.
+An ozone garden consists of selected plant species that are very sensitive to ozone and are likely to display visible injuries.
+Cultivated species were, e.g.
+clover, tobacco, and potatoes.
+As shown in Fig.~\ref{fig:svanhovd_research_station}b), the plants had to be protected from herbivores with a wire-mesh fence.
+In 2018, we observed qualitatively visible ozone damage on semi-natural vegetation (clover) and crops (tobacco).
+In contrast, no such visible damage was found on the clovers in 2019, although it was observed on the sensitive tobacco cultivar.
+This indicates that the vegetation may have been more affected by ozone in 2018 than in 2019.
+
+All measured atmospheric variables including ozone for the 2018/19 growing season were obtained through \href{luftkvalitet.no}{luftkvalitet.no} operated by \citet{NILU_AIRQ}.
+The ozone monitoring data have been added to the EBAS database operated by \citet{NILU_EBAS} and are thus openly accessible.
+All long-term ozone observation data described in Section~\ref{subsec:ozone_data} are also obtained from EBAS.
+The long-term data sets of agrometeorological variables including temperature and precipitation are available from September 1992 to the present day \citep[LandbruksMeteorologiske Tjeneste][note the station name here is Pasvik]{LMT_NIBIO}.
 
@@
 
-An ozone garden was installed over the course of our project at the NIBIO Environment Centre Svanhovd and was likely the northern-most ozone garden to date {\bf TODO: Reference?}.
-An ozone garden consists of selected plant species that are more sensitive to ozone and display visible injuries.
-Cultivated species were, e.g.
-clover, tobacco, and potatoes.
-As shown in Fig.~\ref{fig:svanhovd_research_station}b), the plants had to be protected from herbivores with a wire-mesh fence.
-In 2018, we observed ozone damage on semi-natural vegetation (clover) and crops (tobacco) in the ozone garden.
-In contrast, no such damage was found on the clovers in 2019, although it was found on the sensitive tobacco cultivar.
-This indicates that the vegetation may have been more affected by ozone in 2018 than in 2019.
-
-To relate observed damage on ozone sensitive plants to local ozone concentrations, an ozone monitor had been installed during the growing seasons 2018/19 in addition to existing monitors of atmospheric key variables.
-All variables were obtained through \href{luftkvalitet.no}{luftkvalitet.no} operated by \citet{NILU_AIRQ}.
-Our ozone monitoring data has also been added to the EBAS database operated by \citet{NILU_EBAS} and is thus openly accessible.
-Long-term data sets of agrometeorological variables including temperature and precipitation are available from September 1992 to the present day \citep[LandbruksMeteorologiske Tjeneste][note the station name here is Pasvik]{LMT_NIBIO}.
-All relevant data for the growing season 2018/19 are shown in Fig.~\ref{fig:data_svanvik_2018_2019}.
-Ozone concentrations \chem{[O_3]} measured at $2\,\unit{m}$ height above ground are averaged hourly.
+\subsection{2018/19 growing season}
+\label{subsec:growing_season_1819}
+
+All relevant data for the growing season 2018/19 are shown as timeseries in Fig.~\ref{fig:data_svanvik_2018_2019}.
+Ozone concentrations measured at $2\,\unit{m}$ height above ground are averaged hourly.
 The hatched areas mark times when no ozone data were recorded.
@@
 This coincides with the period of the most extensive forest fires in central Sweden which occurred from July 12--29 \citep{SOU2019}.
-However, due to the above-mentioned data acquisition problems, we missed most of the corresponding ozone data for this event and most likely also the peak ozone concentration.
-We present a reconstruction concerning the missing data in Section~\ref{subsec:ozone_reco}.
-In contrast, ozone concentration only rose 18 times above the threshold of $40\,\unit{ppb}$ during the summer of 2019.
+However, due to the above-mentioned data acquisition problems, we missed most of the corresponding ozone data for this event and most likely also the peak \chem{[O_3]}.
+In contrast, \chem{[O_3]} only rose 18 times above the threshold of $40\,\unit{ppb}$ during the summer of 2019.
+For further analysis (Section~\ref{sec:stats}) and modeling study with the $\mathrm{DO_3SE}$ model (Section~\ref{sec:do3se}), we reconstructed the missing data.
+The used gap filling method will be described in detail in a companion paper to be submitted.
+
 Hourly averaged $2\,\unit{m}$ temperatures above $20\,\unit{^\circ C}$ occurred more regularly in July 2018 than in 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}b).
@@
 More rain events with accumulated daily precipitation ($\sum_d \mathrm{Precip}$) above $10\,\unit{mm}$ occurred in the summer of 2018 compared to 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}c).
-Qualitatively, global irradiance $Q_0$ (Fig.~\ref{fig:data_svanvik_2018_2019}d) was higher in spring (May) and summer (July) 2018 compared to 2019, while June 2019 showed higher irradiance than 2018.
+Qualitatively, global irradiance ($Q_0$) displayed in Fig.~\ref{fig:data_svanvik_2018_2019}d was higher in May and July 2018 compared to 2019, while June 2019 showed higher irradiance than 2018.
 A high global irradiance in most cases is the result of a low cloud fraction.
@@
 The hatched areas indicate periods without ozone monitoring data.
-(a) Hourly averaged ozone concentration; (b) hourly averaged temperature; (c) daily accumulated precipitation; (d) hourly averaged global irradiance.}
+(a) Hourly averaged \chem{[O_3]}; (b) hourly averaged temperature; (c) daily accumulated precipitation; (d) hourly averaged global irradiance.}
   \label{fig:data_svanvik_2018_2019}
@@
 
-\subsection{Other data}
-\label{subsec:other_data}
-To determine the significance of the higher \chem{[O_3]} in 2018 in relation to regional averages and trends, we derive a ground-level ozone concentration climatology for Svanvik.
-From 1986--1996 ozone monitoring has been operated continuously by NILU at Svanhovd, but no ozone measurements have been conducted until the start of our project.
-Due to probable changes in northern hemispheric tropospheric background ozone and a reduction of summertime peak values related to ozone precursors which fell under air quality regulations in Europe, we cross-calibrate our data with long-term ozone observations at other sites in northern Fennoscandia.
-Concerning ozone, available observational data are sparse in this region.
-An overview of the geographical locations of conducted ozone monitoring in the past and present is given in Fig.~\ref{fig:station_map_fennoscandia}.
-The data are described in Section~\ref{subsubsec:ebas}.
-As an additional complement to observational data which are not available at all times and spatially sparse, we also evaluate \chem{[O_3]} derived from a variety of reanalysis products (Section~\ref{subsubsec:ozone_rea}).
-
+\subsection{Ozone data}
+\label{subsec:ozone_data}
+To determine the significance of the higher \chem{[O_3]} in 2018 in relation to regional averages and trends, we derive a ground-level \chem{[O_3]} climatology for Svanvik (Section~\ref{sec:stats}).
+From 1986--1996, ozone monitoring has been operated continuously by NILU at Svanhovd, but no ozone measurements have been conducted before the 2018/19 growing season.
+Due to probable changes in northern hemispheric tropospheric background ozone and a reduction of summertime peak values related to ozone precursors which fell under air quality regulations in Europe, we cross-calibrate our data with long-term ozone observations at sites in northern Finland and Sweden.
+
+%\subsubsection{Ground-level ozone observations}
+%\label{subsubsec:ebas}
 \begin{figure}[t]
@@
 
-\subsubsection{Ground-level ozone observations}
-\label{subsubsec:ebas}
-All ground-level ozone observations used in this study are available from the EBAS database operated by \citet{NILU_EBAS}.
-In Table~\ref{tab:ebas_obs}, we give an overview of the exact locations and time range of operational data taking.
-All stations with available long-term observations are located on higher ground than Svanhovd which will result in relatively higher ground-level ozone concentrations in comparison as ozone abundance, in general, decreases with decreasing altitude in the absence of local precursors \citep[e.g.,][]{AB:Klingberg2009}.
+As indicated above, all ground-level ozone observations used in this study are available from the EBAS database operated by \citet{NILU_EBAS}.
+A comprehensive overview over all past and present ozone observation sites in northern Fennoscandia is given in Table~\ref{tab:ebas_obs} and Fig.~\ref{fig:station_map_fennoscandia}, respectively.
+All stations with available long-term observations are located at higher altitude than Svanhovd which will result in relatively higher ground-level \chem{[O_3]} in comparison.
+In the absence of local precursors, this is due to the stratification of the atmosphere leading to an increase in ozone abundance with altitude \citep[e.g.,][]{AB:Klingberg2009}.
 The periphery vegetation at all sites is similar and consists mainly of pine forests, birch, and heath.
-However, at higher resolution the vegetation varies substantially.
-Ozone measurements at Pallas are taken on top of mount Sammaltunturi $100\,\unit{m}$ above treeline (pine forest) and mainly consists of low vascular plants, mosses, and lichen \citep{BER:Hatakka2003}.
+However, at higher spatial resolution the vegetation varies substantially.
+Ozone measurements at Pallas are taken on top of mount Sammaltunturi $100\,\unit{m}$ above treeline (pine forest) and the surrounding vegetation mainly consists of low vascular plants, mosses, and lichen \citep{BER:Hatakka2003}.
 At Svanhovd, the ozone measurements are taken amidst a meadow surrounded by agriculturally managed herbaceous plants and grass adhering to the \citet{WMOGuide2018} standards (Fig.\ref{fig:svanhovd_research_station}).
@@
 
-\subsubsection{Ozone reanalysis}
-\label{subsubsec:ozone_rea}
-
-Details of reanalysis data sets referred to in this work are listed in Table~\ref{tab:ozone_rea}.
-
-There are two global reanalysis products available at the European Centre for Medium-Range Weather Forecasts (ECMWF) that include atmospheric tracers, in particular ozone, the Monitoring Atmospheric Composition and Climate (MACC) and the newer Copernicus Atmosphere Monitoring Service reanalysis (CAMSRA) \citep{ACP:Inness2013, ACP:Inness2019}.
-The temporal as well as spatial resolution of these reanalysis products is rather coarse: 3-hourly and $0.75\,\unit{^\circ}\times0.75\,\unit{^\circ}$ or roughly $29.3\,\unit{km}\times 83.4\unit{km}$ at the location of Svanvik.
-From Copernicus member institutions, higher resolution surface ozone reanalysis data are available for Europe (CAMS Regional Air Quality (CAMSRAQ) reanalysis ensemble).
-The CAMSRAQ is based on nine European state-of-the-art numerical air quality models.
-The ensemble mean is at fairly high spatial and temporal resolution compared to the global reanalysis: $0.1\,\unit{^\circ}\times 0.1\,\unit{^\circ}$ or roughly $3.9\,\unit{km}\times 11.1\,\unit{km}$ and 1-hourly, respectively.
-The periods covered differ and no data is available before the turn of the millennium.
-The CAMSRA and CAMSRAQ are available in near real time, but only the former covers a period of sufficient length for climate analysis (2003 -- present).
-All reanalyses use different but, at that time, the latest versions of the operational weather forecast system (OpenIFS) from ECMWF.
-They differ substantially in the assimilated observational ozone data.
-The MACC reanalysis assimilates only satellite-derived tropospheric column ozone, while CAMSRA also includes ozone profiles from satellite retrievals.
-Most importantly for our study, in situ observations from ozone near-surface station networks are only assimilated in the CAMSRAQ reanalysis ensemble.
-The quality concerning observations in northern Fennoscandia will be discussed in Section~\ref{subsubsec:clim_ozone}.
-Due to the found substantial discrepancies between reanalysis data and site observations in northern Fennoscandia, we refrain from using any of the global reanalysis data.
-
-\begin{table}[t]
-  \caption{Global/regional ozone reanalysis products used in this study.}
-  \label{tab:ozone_rea}
-\begin{tabular}{llcclcc}
-\tophline
-Name & Provider & \multicolumn{2}{c}{Resolution} & Time period & Meteorological forcing & \chem{O_3} assimilation\\
-&              & spatial & temporal & & \\
-\middlehline
-MACC & ECMWF & $0.75^\circ \times 0.75^\circ$ & 3-hourly & 2003 -- 2012 & OPS & satellite $^\triangledown$\\
-CAMSRA & ECMWF & $0.75^\circ \times 0.75^\circ$ & 3-hourly & 2003 -- 2012 $^\dagger$ & ERA5 / OPS $^\ddagger$ & satellite $^\blacktriangledown$\\
-CAMSRAQ & Copernicus & $0.1^\circ \times 0.1^\circ$ & 1-hourly & 2014 -- 2018 $^\dagger$ & OPS $^\star$ & in situ $^\vartriangle$\\
-\bottomhline
-\end{tabular}
-\belowtable{$^\dagger$ Subset of the data used in this study.
-The full data set extends into near real-time; $^\ddagger$ ERA5 (2003--2016), OPS (later); $^\star$ EURAD uses WRF for downscaling of operational IFS; $^\triangledown$ MLS, OMI - tropospheric column; $^\blacktriangledown$ SCIAMACHY, MIPAS, MLS, OMI, GOME2, SBUV2 - tropospheric column + profile; $^\vartriangle$ METEO France NRT.} % Table Footnotes
-\end{table}
-
-\section{Statistical analysis of environmental key variables}
+\section{Statistical analysis of key environmental variables}
 \label{sec:stats}
@@
 The spring peak values at Svanvik are lower because of the much lower altitude compared to the other monitoring sites.
-The annual average ozone concentration \chem{\left<[O_3]\right>} declines throughout the growing season until it reaches its minimum in late summer/fall ($24-30\,\unit{ppb}$).
+The annual average \chem{[O_3]} \chem{\left<[O_3]\right>} declines throughout the growing season until it reaches its minimum in late summer/fall ($24-30\,\unit{ppb}$).
 The turning points lie approximately in June and November.
@@
 \chem{CO_2} uptake by coniferous trees can occur as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013}.
-The average ozone concentration at Svanvik is lower than at the other sites, as already mentioned, which is in league with a decrease in tropospheric ozone with decreasing altitude.
+The average \chem{[O_3]} at Svanvik is lower than at the other sites, as already mentioned, which is in league with a decrease in tropospheric ozone with decreasing altitude.
 In July--September, ozone is occasionally almost completely depleted during nighttime.
@@
 Fennoscandia combines all available data from Jergul/Karasjok, Esrange, and Pallas.
-The density distributions are shown together with a multi-annual mean of daily ozone concentrations \chem{[O_3]} (climatology).
+The density distributions are shown together with a multi-annual mean of daily \chem{[O_3]} (climatology).
 Splines have been fitted through daily mean \chem{[O_3]}.
@@
 The CAMSRA (b) matches better but does not reproduce the seasonality in northern Fennoscandia.
-The ensemble mean of CAMSRAQ reanalysis ensemble reproduces the seasonal cycle in northern Fennoscandia, though the modeled ozone concentrations are slightly too low.}
+The ensemble mean of CAMSRAQ reanalysis ensemble reproduces the seasonal cycle in northern Fennoscandia, though the modeled \chem{[O_3]}s are slightly too low.}
   \label{fig:ozone_climatologies}
@@
 We have studied the effect of acclimations of biomes to a subarctic climate on phyto-toxic ozone uptake.
-The comparison between 2018 and 2019 conditions allowed us to consider these acclimations and their influence on vegetation sensitivity to \chem{O_3} in light of future changes of key climate variables as may occur under climate change (e.g.
+The comparison between 2018 and 2019 conditions allowed us to consider these acclimations and their influence on vegetation sensitivity to \chem{O_3} in light of future changes of key environmental variables as may occur under climate change (e.g.
 increase of frequency and extent of heatwaves).

### 2021-05-06 VegOzone.tex
Restructured sec 3 - texts has not been fully revised yet

@@
 %%
-\section{Data acquisition of ozone and key environmental variables}
+\section{Data acquisition}
 \label{sec:data}
@@
 For further analysis (Section~\ref{sec:stats}) and modeling study with the $\mathrm{DO_3SE}$ model (Section~\ref{sec:do3se}), we reconstructed the missing data.
-The used gap filling method will be described in detail in a companion paper to be submitted.
+The used gap-filling method will be described in detail in a companion paper to be submitted.
 
@@
 More rain events with accumulated daily precipitation ($\sum_d \mathrm{Precip}$) above $10\,\unit{mm}$ occurred in the summer of 2018 compared to 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}c).
-Qualitatively, global irradiance ($Q_0$) displayed in Fig.~\ref{fig:data_svanvik_2018_2019}d was higher in May and July 2018 compared to 2019, while June 2019 showed higher irradiance than 2018.
+Qualitatively, global irradiance ($Q_0$) displayed in Fig.~\ref{fig:data_svanvik_2018_2019}d) was higher in May and July 2018 compared to 2019, while June 2019 showed higher irradiance than 2018.
 A high global irradiance in most cases is the result of a low cloud fraction.
@@
 
-\section{Statistical analysis of key environmental variables}
+\section{Statistical analysis}
 \label{sec:stats}
-To improve our understanding and determine the risk of vegetation from high \chem{[O_3]} in subarctic biomes in a changing climate, we first assess the climate conditions and evaluate the growing seasons 2018/19 concerning anomalies from multi-annual means (climatologies).
-We derive climatologies for temperature, precipitation, global irradiance, and ozone for Svanvik.
-The ozone measurements will be put into a climatological as well as regional context.
+In this section, we assess the climate conditions and multi-annual mean (referred to as climatology in thhe following) annual cycle of gound-level ozone concentrations at Svanhovd.
+To this end, we derive climatologies for temperature, precipitation, global irradiance, and ozone.
+We evaluate statistical significance of divergances from the norm in these variables for the growing seasons 2018/19 (referred to as anomalies).
+We perform a bias correction on the Svanvik ozone climatology with respect to long-term series of \chem{[O_3]} at the sites in northern Finland and Sweden.
 
@@
 The derived climate diagram following Walter and Lieth which is shown in Fig.~\ref{fig:climatediagram} supports this very well.
-The diagram is based on climatological data from Svanvik (1992--2012) downloaded from LandbruksMeteorologiske Tjeneste.
+The diagram is based on climatological data from the atmospheric monitoring in Svanvik (1992--2012).
 Monthly averaged $2\,\unit{m}$ temperatures (red line) are displayed with a standard error of the mean (SE) as error band.
@@
 In the period 1992--2012, the coldest measured temperature was $-45.2\,\unit{^\circ C}$ observed January 27, 1999, while the highest temperature on July 16 the same year was $29.4\,\unit{^\circ C}$.
-
-Winter and spring (November--April, except for March) are relatively dry ($\sum_m \mathrm{Precip} < 20\,\unit{mm}$).
-The driest month is January with $\sum_m \mathrm{Precip} = (16.7\pm 3.0)\,\unit{mm}$.
-The most precipitation occurs in the summer months, with a highest monthly averaged accumulated precipitation $\sum_m \mathrm{Precip} = (58.5\pm 9.2)\,\unit{mm}$ in August.
-The average annual accumulated precipitation given with standard error of mean is $(383\pm 86)\,\unit{mm}$.
-Precipitation shows a large interannual variability.
-
-A climatology derived for global irradiance $Q_0$ is shown in Appendix Fig.~\ref{fig:global_rad_clim}a) for both, climatological average of hourly maximum $\left<Q_0^\mathrm{max}\right>$, minimum $\left<Q_0^\mathrm{min}\right>$, and mean $\left<Q_0^\mathrm{mean}\right>$ irradiance observed at Svanvik.
-In agreement with the geometrically derived sunshine duration shown in Supplement Fig.~\ref{fig:sunlight_fennoscandia}, $\left<Q_0^\mathrm{max}\right>$ approaches zero in December/January commonly known as polar night and reaches its maximum in June/July commonly referred to as midnight sun conditions.
-The highest observed $\left<Q_0^\mathrm{max}\right>$ is about $800\,\unit{W\,m^{-2}}$, while on average $\left<Q_0^\mathrm{mean}\right>$ it is about $450\,\unit{W\,m^{-2}}$ (regularly reached in May--July).
-The maximum of the climatological minimum $\left<Q_0^\mathrm{min}\right>$ lies at about $200\,\unit{W\,m^{-2}}$.
 
@@
 \end{figure}
+
+The average accumulated monthly precipitation $\sum_d\mathrm{Precip}$ indicates that winter and spring (November--April, except for March) are relatively dry ($\sum_d\mathrm{Precip} < 20\,\unit{mm}$).
+The driest month is January with $\sum_d\mathrm{Precip} = (16.7\pm 3.0)\,\unit{mm}$.
+The most precipitation occurs in the summer months, with a $\sum\mathrm{Precip}_d = (58.5\pm 9.2)\,\unit{mm}$ in August.
+The average annual accumulated precipitation ($\sum_m\mathrm{Precip}$) given with standard error of mean is $(383\pm 86)\,\unit{mm}$.
+Precipitation shows a large interannual variability.
+
+The derived climatology for global irradiance ($Q_0$) at Svanvik is shown in Appendix Fig.~\ref{fig:global_rad_clim}a).
+Shown are the average of hourly maximum $\left<Q_0^\mathrm{max}\right>$, minimum $\left<Q_0^\mathrm{min}\right>$, and mean $\left<Q_0^\mathrm{mean}\right>$.
+In agreement with the geometrically derived sunshine duration displayed in Supplement Fig.~\ref{fig:sunlight_fennoscandia}, $\left<Q_0^\mathrm{max}\right>$ approaches zero in December/January commonly known as polar night and reaches its maximum in June/July commonly referred to as midnight sun conditions.
+The highest observed $\left<Q_0^\mathrm{max}\right>$ is about $800\,\unit{W\,m^{-2}}$, while on average $\left<Q_0^\mathrm{mean}\right>$ it is about $450\,\unit{W\,m^{-2}}$ (regularly reached in May--July).
+The maximum of the climatological minimum $\left<Q_0^\mathrm{min}\right>$ lies at about $200\,\unit{W\,m^{-2}}$.
 
@@
 
-Finally, we shall have a look at the climatologies derived from observational ozone data and the various reanalysis products (Fig.~\ref{fig:ozone_climatologies}).
-To represent the sites, the nearest neighboring grid point of the \chem{[O_3]} reanalysis products is chosen.
-The global products do not reproduce observed ground-level \chem{[O_3]} well.
-This is important for European wide and global risk assessment that rely on these data.
-The MACC reanalysis (Fig.~\ref{fig:ozone_climatologies}a) is in general too low ($(9\pm 7)\,\unit{ppb}$) and displays an erroneous seasonal cycle.
-Ozone concentrations peak already in March followed by the second peak in July.
-Also CAMSRA matches the observed ozone climatology poorly (Fig.~\ref{fig:ozone_climatologies}b).
-Despite displaying only one ozone peak, it still does not reproduce the seasonality in northern Fennoscandia well.
-The CAMSRA spring peak lags behind observations by $1\,\unit{month}$ and is $5\,\unit{ppb}$ too low, whereas the minimum occurs in January compared to August/September.
-The annual amplitude ($(26\pm 1)\,\unit{ppb}$) is larger than in the climatology derived from observations ($19\,\unit{ppb}$).
-The modeled \chem{[O_3]} at Svanvik are slightly higher than at Pallas which might be due to the height at which surface ozone has been evaluated in the model.
-In contrast, the ensemble mean of the CAMSRAQ reproduces the seasonal cycle in northern Fennoscandia quite well (Fig.~\ref{fig:ozone_climatologies}c), but shows on average a slight low-bias ($(2.8\pm 0.5)\,\unit{ppb}$).
-
-\begin{figure}[t]
-  %\includegraphics[width=8.3cm]{ozone_climatologies_portrait}
-  \includegraphics[width=8.3cm]{fig07}
-  \caption{Comparison of climatologies derived from observational data (gray band) with three ozone reanalysis products: (a) MACC reanalysis; (b) CAMSRA; (c) CAMSRAQ reanalysis ensemble mean.
-The sites, represented by the nearest neighboring grid point of the \chem{[O_3]} reanalysis products, are displayed in different markers and colors.
-The global products (a), (b) do not reproduce observed ground-level ozone well.
-The MACC reanalysis (a) is in general too low and displays an erroneous seasonal cycle.
-The CAMSRA (b) matches better but does not reproduce the seasonality in northern Fennoscandia.
-The ensemble mean of CAMSRAQ reanalysis ensemble reproduces the seasonal cycle in northern Fennoscandia, though the modeled \chem{[O_3]}s are slightly too low.}
-  \label{fig:ozone_climatologies}
-\end{figure}
-
-There are many probable explanations for the discovered differences in the climatologies derived from the three reanalyses.
-The model enhancements from MACC reanalysis to CAMSRA have been reported and discussed by \citet{ACP:Inness2019} on a global level.
-Amongst others, assimilation of ozone profiles from satellite retrieval and an upgraded ozone chemistry scheme has lead to the enhanced performance of the latter, but a considerable bias still remains \citep{GMD:Huijnen2020}.
-In particular, \citet{ACPD:Barten2020} reported a pronounced bias for CAMSRA in the high Arctic (e.g., Summit, Greenland) and attribute this bias to dry deposition to the ocean.
-The higher spatial resolution of the regional air quality models used in the CAMSRAQ ensemble compared to the global reanalyses captures small scale depletion and peak episodes in \chem{[O_3]} better and enhances the capability of the models concerning daily and seasonal cycles in ozone.
-The latter is, of course, also affected by upgrades in the OpenIFS used as meteorological forcing.
-However, we can assume that the main driver for the strikingly different results is the assimilated observational ozone data.
-Satellite retrievals resolve \chem{[O_3]} at the surface rather poorly and hence do not constrain the global models well enough \citep{ACP:Andersson2017}.
-
 \subsection{2018/19 anomalies}
@@
 Using these established climatologies, we look at the anomalies in 2018/19 and discuss the differences between the two years.
-As weather extremes like in 2018 are likely to become more frequent in occurrence with climate change, this will help improving our understanding of the role which these climate variables have on determining the sensitivity of subarctic vegetation to \chem{[O3]}.
+As weather extremes like in 2018 are likely to become more frequent with climate change, this will help to improve our understanding of a possible future increase in vulnerability of subarctic vegetation to ozone.
 
@@
 In both years, ozone concentrations were significantly below average in November.
-July 2018 has not been corrected for the missing data due to data acquisition problems.
-The corrected value is, however, indicated by the star (refer to Section~\ref{subsec:ozone_reco} for details of the used method for gap-filling).
+July 2018 represented by the color bar has not been corrected for the missing data due to data acquisition problems.
+The corrected value is indicated by the star.
+A detailed describtion of the gap-filling based on a Reynolds's decomposition method will be published seperately.
 

### 2021-05-06 VegOzone.tex
Move more text to GapFilling

@@
 
-Because the available reanalysis products for ground-level ozone have their limitations, we deem it necessary to reconstruct the missing data in 2018 for the purpose of $\mathrm{DO_3SE}$ modeling by applying a Reynolds's decomposition of the form:
+For the missing ozone data, we applyed a Reynolds's decomposition of the form:
 \begin{equation}
@@
 with climatology \chem{\left<[O_3]\right>} and anomalies \chem{\Delta[O_3]}.
-The exact procedure will be described in the following.
-We derive hourly climatologies for individual stations and combine Esrange/Pallas to derive a climatology for northern Fennoscandia for July (Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}a).
-We compute temporal correlations between Svanvik and the other stations in northern Fennoscandia to estimate a time lag (Supplement Fig.~\ref{fig:time_lag_correlation}).
-We find $3\,\unit{h}$ for Esrange and Pallas with respect to Svanvik.
-Due to the high correlation with Svanvik (Appendix~Fig.~\ref{fig:density_distribution}), we decide on Pallas as the reference station for the reconstruction.
-We correct the Fennoscandia climatology for the found time lag \chem{\left<[O_3]\right>_{hourly,\,t-3}} by shifting the respective time series back in time by $3\,\unit{h}$ and use it to derive a projection for Svanvik:
-\begin{equation}
-  X = \frac{\chem{\left<[O_3]\right>_{hourly}^{Svanvik}}}{\chem{\left<[O_3]\right>_{hourly,\,t-3}}}.
-  \label{eq:mapping}
-\end{equation}
-As an intermediate step, we apply Eq.~(\ref{eq:reynolds_decomp}) to derive hourly anomalies for each station $i$
-\begin{equation}
-  \chem{\Delta[O_3]_{hourly}}^i = \chem{[O_3]_{hourly}}^i-\chem{\left<[O_3]\right>_{hourly}},
-\end{equation}
-with $i \in \{\text{Esrange, Pallas, Svanvik\}}$.
-In Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}b), these hourly ozone concentration anomalies are shown.
-We also correct the ozone anomalies for Pallas for the time lag \chem{\Delta[O_3]_{hourly,\,t-3}^{Pallas}} and use the mapping function in Eq.~(\ref{eq:mapping}) to derive reconstructed anomalies for the missing values at Svanvik:
-\begin{equation}
-    \chem{\Delta [O_3]_{hourly}^{Svanvik,\,reco}} = \chem{\Delta[O_3]_{hourly,\,t-3}^{Pallas}} \cdot X
-\end{equation}
-Finally, we add the bias corrected Svanvik ozone climatology and derive a reconstructed ozone concentration time series \chem{[O_3]_{hourly}^{Svanvik,\,reco}}.
-In Appendix Fig.~\ref{fig:ozone_reconstruction_2018_07}c), the result is shown together with the observed data and the Copernicus Regional Model Reanalysis at the nearest grid point for comparison.
+Observed \chem{[O_3]} in 2018 at Pallas served as the reference.
+The exact procedure will be described in an compagnion paper.
 

### 2021-05-07 VegOzone.tex
Fix typo and move figure

@@
 
-\begin{figure*}[t]
-  %\includegraphics[width=12cm]{ozone_timeseries_fenoscandic_obs.png}
-  \includegraphics[width=12cm]{fig04}
-  \caption{Time series of ozone observations in northern Fennoscandia (Tab.~\ref{tab:ebas_obs}) (1986--2019).
-Data from EBAS until December 2018.
-The hatched areas indicate periods with insufficient quality control according to \citet{NILU2003}.
-(a) Esrange (SWE); (b) Pallas (FIN); (c) Jergul/Karasjok; (d) Svanvik; (e) Janiskoski.}
-  \label{fig:ozone_timesseries_fenoscandic_obs}
-\end{figure*}
-
 \section{Statistical analysis}
 \label{sec:stats}
-In this section, we assess the climate conditions and multi-annual mean (referred to as climatology in thhe following) annual cycle of gound-level ozone concentrations at Svanhovd.
+In this section, we assess the climate conditions and multi-annual mean (referred to as climatology in the following) annual cycle of gound-level ozone concentrations at Svanhovd.
 To this end, we derive climatologies for temperature, precipitation, global irradiance, and ozone.
@@
 
-We found representative ozone climatologies (multi-annual means) for northern Fennoscandia and Svanhovd based on ground-level ozone monitoring.
-To account for missing data from our 2018 record, we proposed a routine for gap filling of ozone observations following a Reynolds's decomposition into anomalies and climatology.
-In the course, we also assessed the quality of available global (MACC and CAMS) and regional (CAMS Regional Air Quality (CAMSRAQ) ensemble) reanalysis products for Europe concerning ozone.
-We have studied the ozone seasonality for northern Fennoscandia and the quality of its representation in available global and regional reanalysis products for Europe.
-We confirm that both global reanalysis products do not reproduce the observed ozone patterns in northern Fennoscandia \citep{GMD:Huijnen2020,ACPD:Barten2020}.
-Only the CAMSRAQ ensemble reproduces the observed ozone seasonality, although with a remaining low-bias.
-This confirms that high spatial resolution and assimilation of in situ observations of ground-level ozone concentrations are a must to constrain reanalysis products related to air quality and risk assessment ozone damage to vegetation, especially in high latitudes.
-
 Ozone peak concentrations were observed more frequently in 2018 than in 2019.

### 2021-05-08 VegOzone.tex
Tune text in Section 2

@@
 In our assessment of primary drivers of ozone damage risk on vegetation in the subarctic, we focus on conditions during the 2018/19 growing season at the NIBIO operated Environment Centre Svanhovd.
-To this end, we collected ozone and atmospheric monitoring data taken at Svanhovd from different distributors.
-The main reasons for this spread are varring update frequencies and length of available data sets.
-In terms of ozone, we also use observational data from monitoring sites in northern Finland and Sweden for cross-evaluation.
-Svanhovd synonymously refers to the atmospheric monitoring site at Svanvik/Pasvik throughout the paper.
-
-%\subsection{NIBIO Environment Centre Svanhovd and bio-monitoring}
-%\label{subsec:atmo_svanvik}
+We also set these into a climatological and regional context.
+To this end, we collected atmospheric and ozone monitoring data taken at Svanhovd and additional ozone monitoring data from observation sites in northern Finland and Sweden for cross-evaluation.
+Varring update frequencies and length of available data sets made it necessary to fetch the data from various sources (further details below).
+Svanhovd synonymously refers to the atmospheric monitoring site Svanvik/Pasvik throughout the paper.\\
+
 The foundation of the NIBIO Environment Centre Svanhovd dates back to the 1930s \citep{NIBIO_Svanhovd}.
-The center focuses on the research of the special environment of the region, but also comprises long-term observations of various agrometeorological variables, i.e.
+The NIBIO centre Svanhovd is located south of the settlement of Svanvik, where the Pasvik river, which is the border to Russia, lies only a few hundred meters to the east.
+It focuses on the research of the special environment of the region, but also comprises long-term observations of various agrometeorological variables, i.e.
 $2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths.
 
-To relate observed visible damage on ozone sensitive plants to local ozone concentration (\chem{[O_3]}), both an ozone bio-monitor (referred to as ozone garden) and a conventional ozone monitor using the existing atmospheric monitoring infrastructure had been installed during the 2018/19 growing season at the NIBIO centre.
+To relate local ozone concentration (\chem{[O_3]}) with visible damage on ozone sensitive plants, both an ozone bio-monitor (referred to as ozone garden) and a conventional ozone monitor using the existing atmospheric monitoring infrastructure had been installed during the 2018/19 growing season at the NIBIO centre.
 The locations of the atmospheric monitoring site and the ozone garden are marked in the aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}a).
-Svanhovd is located south of the settlement of Svanvik, where the Pasvik river, which is the border to Russia, lies only a few hundred meters to the east.
 This ozone garden was likely the northern-most to date {\bf TODO: Reference?}.
-An ozone garden consists of selected plant species that are very sensitive to ozone and are likely to display visible injuries.
+An ozone garden consists of selected plant species that are sensitive to ozone and are likely to display visible injuries.
 Cultivated species were, e.g.
@@
 All measured atmospheric variables including ozone for the 2018/19 growing season were obtained through \href{luftkvalitet.no}{luftkvalitet.no} operated by \citet{NILU_AIRQ}.
-The ozone monitoring data have been added to the EBAS database operated by \citet{NILU_EBAS} and are thus openly accessible.
+The ozone monitoring data have been also added to the EBAS database operated by \citet{NILU_EBAS} and are thus openly accessible.
 All long-term ozone observation data described in Section~\ref{subsec:ozone_data} are also obtained from EBAS.
@@
 As can be seen in Fig.~\ref{fig:data_svanvik_2018_2019}a), \chem{[O_3]} peaks in spring (April/May) and reaches its minimum in late summer (July/August).
-The spring peak has not been captured completely in 2019, since data acquisition started later than in 2018.
+The spring peak has not been captured completely in 2019, since data acquisition only started in late April.
 In summer 2018 (June--August), high ozone concentrations ($\chem{[O_3]} > 40\,\unit{ppb}$) were recorded 50 times.
@@
 From 1986--1996, ozone monitoring has been operated continuously by NILU at Svanhovd, but no ozone measurements have been conducted before the 2018/19 growing season.
-Due to probable changes in northern hemispheric tropospheric background ozone and a reduction of summertime peak values related to ozone precursors which fell under air quality regulations in Europe, we cross-calibrate our data with long-term ozone observations at sites in northern Finland and Sweden.
+Due to probable changes in northern hemispheric tropospheric background \chem{[O_3]} and a reduction of summertime peak values related to ozone precursors which fell under air quality regulations in Europe, we cross-calibrate our data with long-term ozone observations at sites in northern Finland and Sweden.
 
@@
 In 1997, the ozone monitoring at Jergul was relocated downstream of the river Karasjokka to a site close by Karasjok.
-Although the same equipment has been used, this might have introduced a breaking point in the combined data series, denoted as Jergul/Karasjok in the following, which would be particularly problematic if these were the sole data for comparison.
-Only a few data have been taken upstream the Pasvik river at Janiskovski, hence these will not be taken into further consideration, but are shown for completeness.
-As can be seen in Fig.~\ref{fig:ozone_timesseries_fenoscandic_obs}, there is an overlap of about $4\,\unit{years}$ between data taken at Esrange, Jergul/Karasjok, and Svanvik.
-The hatched areas indicate periods with insufficient quality control as mentioned above.
+Although the same equipment has been used, this might have introduced a baseline shift in the combined data series, denoted as Jergul/Karasjok, which would be particularly problematic if these were the sole data for comparison.
+Only a few data have been taken upstream the Pasvik river at Janiskovski, hence these will not be taken into further consideration, but are listed for completeness.
+There is an overlap of about $4\,\unit{years}$ between data taken at Esrange, Jergul/Karasjok, and Svanvik (Fig.~\ref{fig:ozone_timesseries_fenoscandic_obs}).
 
@@
 \label{sec:stats}
-In this section, we assess the climate conditions and multi-annual mean (referred to as climatology in the following) annual cycle of gound-level ozone concentrations at Svanhovd.
-To this end, we derive climatologies for temperature, precipitation, global irradiance, and ozone.
-We evaluate statistical significance of divergances from the norm in these variables for the growing seasons 2018/19 (referred to as anomalies).
-We perform a bias correction on the Svanvik ozone climatology with respect to long-term series of \chem{[O_3]} at the sites in northern Finland and Sweden.
+In this section, we assess the climate conditions.
+To this end, we compute multi-annual mean (referred to as climatology) of environmental key variables (temperature, precipitation, global irradiance) and gound-level \chem{[O_3]} at Svanhovd.
+We evaluate the statistical significance of divergances from the norm in these variables for the growing seasons 2018/19 (referred to as anomalies) and perform a bias correction on the ozone climatology at Svanvik accounting for the increase in tropospheric background \chem{[O_3]}.
+For this, we cross-calibrate with long-term observations of \chem{[O_3]} at the sites in northern Finland and Sweden.
 
@@
 The location of Svanvik at $69.45\,\unit{^\circ N}$ suggests a subarctic climate.
-The derived climate diagram following Walter and Lieth which is shown in Fig.~\ref{fig:climatediagram} supports this very well.
-The diagram is based on climatological data from the atmospheric monitoring in Svanvik (1992--2012).
+The derived climate diagram based on climatological data from the atmospheric monitoring in Svanvik (1992--2012) (Fig.~\ref{fig:climatediagram}) supports this very well.
 Monthly averaged $2\,\unit{m}$ temperatures (red line) are displayed with a standard error of the mean (SE) as error band.
 We chose SE in this case due to the large interannual variability in temperature reflected in the standard deviation.
-Averaged monthly accumulated precipitation (blue bars) is shown with standard deviation of mean as error bars.
+Averaged monthly accumulated precipitation (blue bars) is shown with standard deviation as error bars.
 Months with average temperatures below freezing are denoted with a star.
-As can be seen, temperatures stay below freezing for 5 consecutive months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for K\"{o}ppens climate classification of a regular subarctic climate (Dfc) \citep{SD:Beck2018}.
+As can be seen, temperatures stay below freezing for 5 consecutive months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for K\"{o}ppen's climate classification of a regular subarctic climate (Dfc) \citep[][e.g.]{SD:Beck2018}.
 The highest monthly average temperature is $(13.1\pm 1.1)\,\unit{^\circ C}$ in July and the lowest $(-12.8\pm 2.0)\,\unit{^\circ C}$ in February.
-In the period 1992--2012, the coldest measured temperature was $-45.2\,\unit{^\circ C}$ observed January 27, 1999, while the highest temperature on July 16 the same year was $29.4\,\unit{^\circ C}$.
+In the period 1992--2012, the coldest measured temperature was $-45.2\,\unit{^\circ C}$ observed January 27, 1999, while the highest temperature ($29.4\,\unit{^\circ C}$) occurred July 16 the same year.
 
@@
   \caption{Climate diagram following Walter and Lieth.
-The diagram is based on climatological data for Svanvik/Pasvik (1992--2012) downloaded from LandbruksMeteorologiske Tjeneste.
+The diagram is based on climatological data for Svanvik/Pasvik (1992--2012).
 Monthly averaged temperatures (red line) are displayed with standard error of the mean as error band.
@@
 
-The derived climatology for global irradiance ($Q_0$) at Svanvik is shown in Appendix Fig.~\ref{fig:global_rad_clim}a).
-Shown are the average of hourly maximum $\left<Q_0^\mathrm{max}\right>$, minimum $\left<Q_0^\mathrm{min}\right>$, and mean $\left<Q_0^\mathrm{mean}\right>$.
-In agreement with the geometrically derived sunshine duration displayed in Supplement Fig.~\ref{fig:sunlight_fennoscandia}, $\left<Q_0^\mathrm{max}\right>$ approaches zero in December/January commonly known as polar night and reaches its maximum in June/July commonly referred to as midnight sun conditions.
-The highest observed $\left<Q_0^\mathrm{max}\right>$ is about $800\,\unit{W\,m^{-2}}$, while on average $\left<Q_0^\mathrm{mean}\right>$ it is about $450\,\unit{W\,m^{-2}}$ (regularly reached in May--July).
-The maximum of the climatological minimum $\left<Q_0^\mathrm{min}\right>$ lies at about $200\,\unit{W\,m^{-2}}$.
+The derived climatology for global irradiance ($Q_0$) at Svanvik (Appendix Fig.~\ref{fig:global_rad_clim}a)) in agreement with the geometrically derived sunshine duration (Supplement Fig.~\ref{fig:sunlight_fennoscandia}), approaches zero in December/January (polar night) and reaches its maximum in June/July (midnight sun conditions).
+The highest observed averaged daily maximum $\left<Q_0^\mathrm{max}\right>$ is about $800\,\unit{W\,m^{-2}}$, while on averaged daily mean $\left<Q_0^\mathrm{mean}\right>$ it is about $450\,\unit{W\,m^{-2}}$ (regularly reached in May--July).
+The maximum of the averaged daily minimum $\left<Q_0^\mathrm{min}\right>$ lies at about $200\,\unit{W\,m^{-2}}$.
 
@@
 The \chem{\left<\Delta\left<[O_3]\right>\right>} for Svanvik in 2018 is slightly larger than for Esrange and Pallas.
-This could mean either that 2018 had been more exceptional at Svanvik or the derived climatology may need a correction for increased tropospheric background ozone concentration.
+This could mean either that 2018 had been more exceptional at Svanvik or the derived climatology may need a correction for increased tropospheric background \chem{[O_3]}.
 Based on the meteorological conditions, we assume 2019 as an average year which leads to an estimated constant bias of about $1.2\,\unit{ppb}$ in \chem{[O_3]}.

### 2021-05-08 VegOzone.tex
Rephrase Section 3.1

@@
 From 1986--1996, ozone monitoring has been operated continuously by NILU at Svanhovd, but no ozone measurements have been conducted before the 2018/19 growing season.
-Due to probable changes in northern hemispheric tropospheric background \chem{[O_3]} and a reduction of summertime peak values related to ozone precursors which fell under air quality regulations in Europe, we cross-calibrate our data with long-term ozone observations at sites in northern Finland and Sweden.
 
@@
 \label{subsubsec:clim_ozone}
-In the following, we will look at derived ozone climatologies from different perspectives.
-The multi-annual means of daily mean ozone, for simplicity here without standard deviations, at observation sites in northern Fennoscandia (refer to Table~\ref{tab:ebas_obs} for details) are shown in Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs}.
-All climatologies except for Svanvik show peak values of about $48\,\unit{ppb}$ in late April.
-The spring peak values at Svanvik are lower because of the much lower altitude compared to the other monitoring sites.
-The annual average \chem{[O_3]} \chem{\left<[O_3]\right>} declines throughout the growing season until it reaches its minimum in late summer/fall ($24-30\,\unit{ppb}$).
+In the following, we will derived a ozone climatology (\chem{\left<[O_3]\right>}) for Svanvik and for northern Fennoscadia.
+To this end, we first calculate a pairwise correlation between the different stations.
+The correlation of \chem{[O_3]} between Esrange, Pallas, and Jergul/Karasjok is fairly high ($r^2\approx 0.8$, Appendix Fig.~\ref{fig:density_distribution}a, b).
+Therefore, we select these data to derive a generallized ozone climatology for northern Fennoscandia.
+The correlation between Svanvik with any of the active monitoring sites in Esrange and Pallas is rather low, 0.4 and 0.6, respectively.
+Thus, conditions at Svanvik cannot directly be represented by measurements taken at either Esrange or Pallas.
+
+In Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}, two dimensional density distributions of \chem{[O_3]} for the combined northern Fennoscandia data (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}a) and Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) are shown.
+Darker colors indicate higher probability to observe these values.
+On top of the density distributions, a $10\,\unit{days}$ average of daily mean \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} is displayed together with $1 \sigma$ uncertainties and SE, respectively.
+Splines have been fitted through the data for easier inter-comparison.
+
+The generalized Fennoscandic \chem{\left<[O_3]\right>} peaks at $48\,\unit{ppb}$ in late April (spring peak).
+From the northern Fennoscandia climatology, we deduce several features.
+\chem{\left<[O_3]\right>} declines throughout the growing season until a minimum is reached in late summer/fall ($24-30\,\unit{ppb}$).
 The turning points lie approximately in June and November.
-At Svanvik, \chem{\left<[O_3]\right>} is $6.6\,\unit{ppb}$ lower than on sites at Jergul/Karasjok (NOR), Esrange (SWE), and Pallas (FIN).
-The large spread in the case of Svanvik and Janiskoski is due to the lower number of observations.
-
-The correlation of \chem{[O_3]} between Esrange, Pallas, and Jergul/Karasjok is fairly high ($r^2\approx 0.8$, Appendix Fig.~\ref{fig:density_distribution}a, b), hence we can put those data series together to derive a general ozone climatology for northern Fennoscandia.
-The correlation between Svanvik and Esrange and Pallas is rather low, 0.4 and 0.6, respectively.
-Thus, Svanvik cannot directly be represented by any measurements taken at either Esrange or Pallas.
-We shall come back to this in Section~\ref{subsec:ozone_reco}.
-
-In Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}, the two dimensional density distributions of \chem{[O_3]} for the combined northern Fennoscandia data (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}a) and Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) are shown.
-Darker colors indicate higher probability to observe these values.
-On top of the density distributions, $10\,\unit{days}$ averages of daily mean \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} are displayed together with $1 \sigma$ uncertainties and SE, respectively.
-Splines have been fitted through the data for easier inter-comparison with reanalysis data.
-From the northern Fennoscandia climatology, one can deduce several features.
 The build-up of the ozone spring peak takes about $7.5\,\unit{months}$ from minimum to maximum.
-A sharp increase starts one month ahead of the estimated begin of greening $G_\mathrm{begin}$ (see Section~\ref{subsec:do3se_parameters} for details) and is in coincident with rising temperatures and global irradiance.
-We find day of the year (\unit{doy}) $\left<G_\mathrm{begin}\right> = (150 \pm 14)$ for Svanvik (Supplement Fig~\ref{fig:greening_season_change_Svanvik}), which translates to the end of May.
-This finding is in line with more extensive studies \citep[][e.g.,]{IJB:Karlsen2007,AFM:Linderholm2006}.
+The pronounced increase in \chem{\left<[O_3]\right>} starts on average one month ahead of the estimated begin of greening ($G_\mathrm{begin}$) (see Section~\ref{subsec:do3se_parameters} for details) and coincidents with rising temperatures and global irradiance.
+For comparison, we found $\left<G_\mathrm{begin}\right> = (150 \pm 14)$ in units of day of the year (\unit{doy}) for Svanvik (Supplement Fig~\ref{fig:greening_season_change_Svanvik}), which translates to the end of May.
+This finding is in line with more extensive studies of on-site observations and satellite retrievals \citep[][e.g.,]{IJB:Karlsen2007,AFM:Linderholm2006}.
 The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} is reached in late April ($100-120\,\unit{doy}$).
@@
 The density distribution for Svanvik and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} display a similar pattern as the northern Fennoscandia climatology (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b).
-The spring peak occurs slightly earlier (at the beginning of April).
-As Svanvik is located at a lower altitude than the other sites, the earlier decline of the ozone peak may point to a stronger influence of vegetation on the removal of ozone.
-\chem{CO_2} uptake by coniferous trees can occur as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013}.
-The average \chem{[O_3]} at Svanvik is lower than at the other sites, as already mentioned, which is in league with a decrease in tropospheric ozone with decreasing altitude.
+The spring peak occurs slightly earlier (at the beginning of April) and amounts to $40\,\unit{ppb}$ while the annual average \chem{\left<[O_3]\right>} is $6.6\,\unit{ppb}$ lower than the Fennoscandic climatology.
+This is in line with a general decrease in tropospheric ozone with decreasing altitude as Svanvik is located at a lower altitude than the other sites.
+The earlier decline in \chem{\left<[O_3\right>]} may point to an influence of coniferous vegetation on the removal of ozone since\chem{CO_2} uptake by coniferous trees can occur as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013}.
 In July--September, ozone is occasionally almost completely depleted during nighttime.
+This may point to ozone uptake by vegetation even at low light intensities during midnight sun conditions in combination with stable planetary boundary layer conditions preventing mixing of ozone rich air.
+
+\chem{\left<[O_3]\right>} is likely low-biased at Svanvik due to an increase in northern hemispheric tropospheric background \chem{[O_3]} since the 1990s.
+To account for this, we derive a bias correction factor by cross-calibrating with the timeseries taken at Esrange and Pallas (Section~\ref{subsec:anomalies}).
 

### 2021-05-08 supplements/VegOzone_supplement.tex
Move figure to supplement

@@
 \setcounter{figure}{0}
+
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{ozone_timeseries_fennoscandic_obs.png}
+  %\includegraphics[width=12cm]{fig04}
+  \caption{Time series of ozone observations in northern Fennoscandia (Tab.~\ref{tab:ebas_obs}) (1986--2019).
+Data from EBAS until December 2018.
+The hatched areas indicate periods with insufficient quality control according to \citet{NILU2003}.
+(a) Esrange (SWE); (b) Janiskoski (RUS); (c) Jergul/Karasjok (NOR); (d) Pallas (FIN); (e) Svanvik (NOR).}
+  \label{fig:ozone_timesseries_fenoscandic_obs}
+\end{figure*}
 

### 2021-05-09 VegOzone.tex
Refine introduction

@@
 The observation site at NIBIO Environment Centre Svanhovd and all related data are presented in Section~\ref{sec:data}.
-In Section~\ref{sec:stats}, we derive multi-annual means (referred to as climatologies) for ground-level ozone concentration and key environmental variables (temperature, precipitation, and global irradiance) based on in situ observations and evaluate the deviation of 2018/19 from the norm (referred to as anomalies).
-We use the $\mathrm{DO_3SE}$ model to estimate the ozone uptake by natural and semi-natural vegetation to investigate the main environmental drivers for the observed ozone damage in 2018 (Section~\ref{sec:do3se}).
-In this context, we introduce the $\mathrm{DO_3SE}$ model equations and derive bespoke parameterizations for subarctic species to assess systematic uncertainties arising from the use of default generic parameterizations established from less climatically extreme conditions.
+In Section~\ref{sec:stats}, we derive multi-annual means for ground-level ozone concentration and key environmental variables (temperature, precipitation, and global irradiance) based on in situ observations and evaluate the deviation of 2018/19 from the norm.
+We use the $\mathrm{DO_3SE}$ model to estimate the ozone uptake by natural and semi-natural vegetation to investigate the main environmental drivers for projected ozone damage risk in 2018 (Section~\ref{sec:do3se}).
+In this end, we derive bespoke parameterizations for subarctic species to assess systematic uncertainties arising from the use of default generic parameterizations established from less climatically extreme conditions.
 In Section~\ref{sec:conc}, we will summarize our results and discuss how this research can support improvements in future assessments of pollution risk and more generally bio-geochemical cycling.

### 2021-05-09 supplements/VegOzone_supplement.tex
Remove figures

@@
 
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{test_fennoscandia}
-  \caption{Maximum daily sun shine duration based on geometrical calculations at different sites in Fennoscandia.
+%\begin{figure*}[t]
+%  \includegraphics[width=12cm]{test_fennoscandia}
+%  \caption{Maximum daily sun shine duration based on geometrical calculations at different sites in Fennoscandia.
 Midnight sun conditions at Jergul/Karasjok and Svanvik prevail from the end of May until the end of July, while at Esrange and Pallas they only prevail from the beginning of June until mid of July.}
-  \label{fig:sunlight_fennoscandia}
-\end{figure*}
+%  \label{fig:sunlight_fennoscandia}
+%\end{figure*}
 
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{ozone_climatology_fenoscandic_obs}
-  \caption{Multi-annual mean of daily mean ozone at observation sites in Fennoscandia.
+%\begin{figure*}[t]
+%  \includegraphics[width=12cm]{ozone_climatology_fenoscandic_obs}
+%  \caption{Multi-annual mean of daily mean ozone at observation sites in Fennoscandia.
 The large spread in case of Svanvik and Janiskoski is due to the lower statistics.
@@
 The annual average ozone concentration \chem{\left<[O_3]\right>} at Svanvik is $6.6\,\unit{ppb}$ lower than at the sites at Jergul/Karasjok (NOR), Esrange (SWE), and Pallas (FIN).}
-  \label{fig:ozone_climatology_fenoscandic_obs}
-\end{figure*}
+%  \label{fig:ozone_climatology_fenoscandic_obs}
+%\end{figure*}
 

### 2021-05-09 VegOzone.tex
Fix text in section 2.1

@@
 We use the $\mathrm{DO_3SE}$ model to estimate the ozone uptake by natural and semi-natural vegetation to investigate the main environmental drivers for projected ozone damage risk in 2018 (Section~\ref{sec:do3se}).
-In this end, we derive bespoke parameterizations for subarctic species to assess systematic uncertainties arising from the use of default generic parameterizations established from less climatically extreme conditions.
+In this end, we erive bespoke parameterizations for subarctic species to assess systematic uncertainties arising from the use of default generic parameterizations established from less climatically extreme conditions.
 In Section~\ref{sec:conc}, we will summarize our results and discuss how this research can support improvements in future assessments of pollution risk and more generally bio-geochemical cycling.
@@
 As shown in Fig.~\ref{fig:svanhovd_research_station}b), the plants had to be protected from herbivores with a wire-mesh fence.
-In 2018, we observed qualitatively visible ozone damage on semi-natural vegetation (clover) and crops (tobacco).
+In 2018, we qualitatively observed visible ozone damage on semi-natural vegetation (clover) and crops (tobacco).
 In contrast, no such visible damage was found on the clovers in 2019, although it was observed on the sensitive tobacco cultivar.
@@
 For further analysis (Section~\ref{sec:stats}) and modeling study with the $\mathrm{DO_3SE}$ model (Section~\ref{sec:do3se}), we reconstructed the missing data.
-The used gap-filling method will be described in detail in a companion paper to be submitted.
+The used gap-filling method will be described in detail in a companion paper.
 
@@
 A high global irradiance in most cases is the result of a low cloud fraction.
-In both years, the maximum recorded $Q_0$ was roughly $750\,\unit{W\,m^{-2}}$ reached in June as naturally expected.
+In both years, the maximum recorded $Q_0$ was roughly $750\,\unit{W\,m^{-2}}$ reached in June.
 
@@
 The driest month is January with $\sum_d\mathrm{Precip} = (16.7\pm 3.0)\,\unit{mm}$.
-The most precipitation occurs in the summer months, with a $\sum\mathrm{Precip}_d = (58.5\pm 9.2)\,\unit{mm}$ in August.
+The most precipitation occurs in the summer months, with a $\sum_d\mathrm{Precip} = (58.5\pm 9.2)\,\unit{mm}$ in August.
 The average annual accumulated precipitation ($\sum_m\mathrm{Precip}$) given with standard error of mean is $(383\pm 86)\,\unit{mm}$.

### 2021-05-09 VegOzone.tex
Refine section 2.2.

@@
 \label{subsec:ozone_data}
-To determine the significance of the higher \chem{[O_3]} in 2018 in relation to regional averages and trends, we derive a ground-level \chem{[O_3]} climatology for Svanvik (Section~\ref{sec:stats}).
 From 1986--1996, ozone monitoring has been operated continuously by NILU at Svanhovd, but no ozone measurements have been conducted before the 2018/19 growing season.
+To determine the significance of the \chem{[O_3]} in 2018/19 in relation to regional trends, we derive a ground-level \chem{[O_3]} climatology for Svanvik and northern Fennoscandia in general (Section~\ref{sec:stats}).
 
@@
 
-As indicated above, all ground-level ozone observations used in this study are available from the EBAS database operated by \citet{NILU_EBAS}.
 A comprehensive overview over all past and present ozone observation sites in northern Fennoscandia is given in Table~\ref{tab:ebas_obs} and Fig.~\ref{fig:station_map_fennoscandia}, respectively.
-All stations with available long-term observations are located at higher altitude than Svanhovd which will result in relatively higher ground-level \chem{[O_3]} in comparison.
-In the absence of local precursors, this is due to the stratification of the atmosphere leading to an increase in ozone abundance with altitude \citep[e.g.,][]{AB:Klingberg2009}.
+All stations with available long-term observations are located at higher altitude than Svanhovd.
+In the absence of local precursors, the stratification of the atmosphere usually leads to a higher abundances in ozone at higher altitude \citep[e.g.,][]{AB:Klingberg2009}.
 The periphery vegetation at all sites is similar and consists mainly of pine forests, birch, and heath.
-However, at higher spatial resolution the vegetation varies substantially.
+However, on closer inspection the vegetation differs substantially.
 Ozone measurements at Pallas are taken on top of mount Sammaltunturi $100\,\unit{m}$ above treeline (pine forest) and the surrounding vegetation mainly consists of low vascular plants, mosses, and lichen \citep{BER:Hatakka2003}.
 At Svanhovd, the ozone measurements are taken amidst a meadow surrounded by agriculturally managed herbaceous plants and grass adhering to the \citet{WMOGuide2018} standards (Fig.\ref{fig:svanhovd_research_station}).
+In conjunction with the location at lower altitude, this results in generally lower \chem{[O_3]} at Svanhovd compared to Pallas.
 

### 2021-05-09 VegOzone.tex
Refine intro to section 3.1

@@
 
-Data from before the 1990s are available solely from Svanvik and Karasjok.
+Data from before the 1990s are available solely from Svanvik and Jergul.
 These data, however, do not follow the high-quality standard procedures implemented nowadays and have to be treated with care \citep{NILU2003}.
@@
 In this section, we assess the climate conditions.
-To this end, we compute multi-annual mean (referred to as climatology) of environmental key variables (temperature, precipitation, global irradiance) and gound-level \chem{[O_3]} at Svanhovd.
-We evaluate the statistical significance of divergances from the norm in these variables for the growing seasons 2018/19 (referred to as anomalies) and perform a bias correction on the ozone climatology at Svanvik accounting for the increase in tropospheric background \chem{[O_3]}.
-For this, we cross-calibrate with long-term observations of \chem{[O_3]} at the sites in northern Finland and Sweden.
+To this end, we compute multi-annual mean (referred to as climatology in the following) of environmental key variables (temperature, precipitation, global irradiance) and gound-level \chem{[O_3]} at Svanhovd.
+For the growing seasons 2018/19, we evaluate the statistical significance of divergances from the norm in these variables (referred to as anomalies).
+We perform a bias correction accounting for the increase in tropospheric background \chem{[O_3]} and cross-calibrate the derived climatology for Svanvik with long-term observations of \chem{[O_3]} at Pallas and Esrange.
 

### 2021-05-09 VegOzone.tex
Refine global irradiance paragraph

@@
 
-The average accumulated monthly precipitation $\sum_d\mathrm{Precip}$ indicates that winter and spring (November--April, except for March) are relatively dry ($\sum_d\mathrm{Precip} < 20\,\unit{mm}$).
+The average accumulated monthly precipitation ($\sum_d\mathrm{Precip}$) indicates that winter and spring (November--April, except for March) are relatively dry ($\sum_d\mathrm{Precip} < 20\,\unit{mm}$).
 The driest month is January with $\sum_d\mathrm{Precip} = (16.7\pm 3.0)\,\unit{mm}$.
@@
 
-The derived climatology for global irradiance ($Q_0$) at Svanvik (Appendix Fig.~\ref{fig:global_rad_clim}a)) in agreement with the geometrically derived sunshine duration (Supplement Fig.~\ref{fig:sunlight_fennoscandia}), approaches zero in December/January (polar night) and reaches its maximum in June/July (midnight sun conditions).
-The highest observed averaged daily maximum $\left<Q_0^\mathrm{max}\right>$ is about $800\,\unit{W\,m^{-2}}$, while on averaged daily mean $\left<Q_0^\mathrm{mean}\right>$ it is about $450\,\unit{W\,m^{-2}}$ (regularly reached in May--July).
-The maximum of the averaged daily minimum $\left<Q_0^\mathrm{min}\right>$ lies at about $200\,\unit{W\,m^{-2}}$.
+The derived climatology for global irradiance ($Q_0$) at Svanvik (Appendix Fig.~\ref{fig:global_rad_clim}a)) approaches zero in December/January (polar night) and reaches its maximum in June/July (midnight sun conditions).
+The highest observed averaged daily maximum ($\left<Q_0^\mathrm{max}\right>$) amounts to $800\,\unit{W\,m^{-2}}$, while the averaged daily mean ($\left<Q_0^\mathrm{mean}\right>$) is $450\,\unit{W\,m^{-2}}$ (regularly reached in May--July).
+The maximum of the averaged daily minimum ($\left<Q_0^\mathrm{min}\right>$) lies at about $200\,\unit{W\,m^{-2}}$.
 

### 2021-05-09 VegOzone.tex
Refine Section 3.1

@@
 For the growing seasons 2018/19, we evaluate the statistical significance of divergances from the norm in these variables (referred to as anomalies).
-We perform a bias correction accounting for the increase in tropospheric background \chem{[O_3]} and cross-calibrate the derived climatology for Svanvik with long-term observations of \chem{[O_3]} at Pallas and Esrange.
+We perform a bias correction accounting for the increase in tropospheric background \chem{[O_3]} and cross-calibrate the derived climatology for Svanvik with long-term observations of \chem{[O_3]} at Esrange and Pallas.
 
@@
 \label{subsubsec:clim_ozone}
-In the following, we will derived a ozone climatology (\chem{\left<[O_3]\right>}) for Svanvik and for northern Fennoscadia.
+In the following, we derive an ozone climatology (\chem{\left<[O_3]\right>}) for Svanvik and for northern Fennoscadia.
 To this end, we first calculate a pairwise correlation between the different stations.
 The correlation of \chem{[O_3]} between Esrange, Pallas, and Jergul/Karasjok is fairly high ($r^2\approx 0.8$, Appendix Fig.~\ref{fig:density_distribution}a, b).
-Therefore, we select these data to derive a generallized ozone climatology for northern Fennoscandia.
-The correlation between Svanvik with any of the active monitoring sites in Esrange and Pallas is rather low, 0.4 and 0.6, respectively.
+Therefore, we select these data to compute the generallized ozone climatology for northern Fennoscandia.
+The correlation between Svanvik with any of the active monitoring sites in 2018 (Esrange and Pallas) is rather low, 0.4 and 0.6, respectively.
 Thus, conditions at Svanvik cannot directly be represented by measurements taken at either Esrange or Pallas.
@@
 Darker colors indicate higher probability to observe these values.
-On top of the density distributions, a $10\,\unit{days}$ average of daily mean \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} is displayed together with $1 \sigma$ uncertainties and SE, respectively.
+On top of the density distributions, a $10\,\unit{days}$ average of daily mean (\chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}}) is displayed together with $1 \sigma$ uncertainties and SE, respectively.
 Splines have been fitted through the data for easier inter-comparison.
 
-The generalized Fennoscandic \chem{\left<[O_3]\right>} peaks at $48\,\unit{ppb}$ in late April (spring peak).
-From the northern Fennoscandia climatology, we deduce several features.
-\chem{\left<[O_3]\right>} declines throughout the growing season until a minimum is reached in late summer/fall ($24-30\,\unit{ppb}$).
+The generalized northern Fennoscandic ozone climatology peaks at $48\,\unit{ppb}$  in late April (day of the year (doy) $100-120$) and declines throughout the growing season until a minimum is reached in late summer/fall ($24-30\,\unit{ppb}$).
 The turning points lie approximately in June and November.
 The build-up of the ozone spring peak takes about $7.5\,\unit{months}$ from minimum to maximum.
-The pronounced increase in \chem{\left<[O_3]\right>} starts on average one month ahead of the estimated begin of greening ($G_\mathrm{begin}$) (see Section~\ref{subsec:do3se_parameters} for details) and coincidents with rising temperatures and global irradiance.
-For comparison, we found $\left<G_\mathrm{begin}\right> = (150 \pm 14)$ in units of day of the year (\unit{doy}) for Svanvik (Supplement Fig~\ref{fig:greening_season_change_Svanvik}), which translates to the end of May.
-This finding is in line with more extensive studies of on-site observations and satellite retrievals \citep[][e.g.,]{IJB:Karlsen2007,AFM:Linderholm2006}.
-The maximum of \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} is reached in late April ($100-120\,\unit{doy}$).
+The most pronounced increase in \chem{\left<[O_3]\right>} coincidents with rising temperatures and global irradiance.
 The depletion of ozone from maximum to minimum takes only $4.5\,\unit{months}$.
@@
 
-The density distribution for Svanvik and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} display a similar pattern as the northern Fennoscandia climatology (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b).
-The spring peak occurs slightly earlier (at the beginning of April) and amounts to $40\,\unit{ppb}$ while the annual average \chem{\left<[O_3]\right>} is $6.6\,\unit{ppb}$ lower than the Fennoscandic climatology.
-This is in line with a general decrease in tropospheric ozone with decreasing altitude as Svanvik is located at a lower altitude than the other sites.
-The earlier decline in \chem{\left<[O_3\right>]} may point to an influence of coniferous vegetation on the removal of ozone since\chem{CO_2} uptake by coniferous trees can occur as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013}.
+The \chem{[O_3]} density distribution and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} for Svanvik display a similar pattern as the generallized northern Fennoscandia climatology (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b).
+The spring peak occurs slightly earlier ($100\,\unit{doy}$) and amounts to $40\,\unit{ppb}$ while the annual average \chem{\left<[O_3]\right>} is $6.6\,\unit{ppb}$ lower than the Fennoscandic climatology.
+This is in line with a general decrease in tropospheric ozone with decreasing altitude.
+The earlier decline in \chem{\left<[O_3]\right>} may point to an influence of coniferous vegetation on the removal of ozone since \chem{CO_2} uptake by coniferous trees can occur as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013} compared to an estimated average begin of greening ($\left<G_\mathrm{start}\right>$) on $(150\pm 14)\,\unit{doy}$ at Svanvik (see Section~\ref{subsec:do3se_parameters} for details).
 In July--September, ozone is occasionally almost completely depleted during nighttime.
-This may point to ozone uptake by vegetation even at low light intensities during midnight sun conditions in combination with stable planetary boundary layer conditions preventing mixing of ozone rich air.
-
-\chem{\left<[O_3]\right>} is likely low-biased at Svanvik due to an increase in northern hemispheric tropospheric background \chem{[O_3]} since the 1990s.
+This hint to ozone uptake by vegetation even at low light intensities during midnight sun conditions in combination with stable planetary boundary layer conditions preventing mixing of ozone rich air.
+The derived \chem{\left<[O_3]\right>} is likely biased towards lower \chem{[O_3]} due to an increase in northern hemispheric tropospheric background \chem{[O_3]} since the 1990s.
 To account for this, we derive a bias correction factor by cross-calibrating with the timeseries taken at Esrange and Pallas (Section~\ref{subsec:anomalies}).

### 2021-05-09 VegOzone.tex
Rephrase gpp co2 uptake sentence

@@
 This is in line with a general decrease in tropospheric ozone with decreasing altitude.
-The earlier decline in \chem{\left<[O_3]\right>} may point to an influence of coniferous vegetation on the removal of ozone since \chem{CO_2} uptake by coniferous trees can occur as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013} compared to an estimated average begin of greening ($\left<G_\mathrm{start}\right>$) on $(150\pm 14)\,\unit{doy}$ at Svanvik (see Section~\ref{subsec:do3se_parameters} for details).
+The earlier decline in \chem{\left<[O_3]\right>} may point to an influence of coniferous vegetation on the removal of ozone since \chem{CO_2} uptake by coniferous trees has been observed as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013} compared to an estimated average begin of greening ($\left<G_\mathrm{start}\right>$) on $(150\pm 14)\,\unit{doy}$ at Svanvik (see Section~\ref{subsec:do3se_parameters} for details).
 In July--September, ozone is occasionally almost completely depleted during nighttime.

### 2021-05-09 VegOzone.tex
First revision of sec 3.2.2

@@
 \label{subsubsec:anomal_ozone}
-As we discussed in the preceding Section~\ref{subsubsec:anomal_tpq}, meteorological conditions in 2018 were significantly different from the climatological average.
-How does this translate into observed ozone anomalies \chem{\Delta[O_3]}?
-
-If the data collected at Svanvik had been complete in 2018 and the climatology had extended into the recent past, we could have assessed the significance of ozone anomalies in the same way as for temperature, precipitation, and global irradiance.
-In Section~\ref{subsubsec:clim_ozone}, we described the construction of climatologies for northern Fennoscandia based on long-term observations at Esrange, Jergul/Karasjok, and Pallas.
-For Svanvik this assessment is based on rather uncertain observations from 1986--1996.
-We also discussed the general patterns, similarities, and differences.
-We shall now investigate \chem{\Delta[O_3]} in 2018/19 for these climatologies in more detail.
-
-In the following, we will use the interchangeable terms anomalies and residuals to distinguish between a climatological (anomalies) and a statistical (residuals) point of view.
-All the above factors make it necessary to look not only at anomalies but statistically also to assess ozone concentration residuals at Svanvik as well as at Esrange and Pallas in comparison.
-Since the climatology of the latter is more reliable, we will start with ground-level ozone concentration residuals \chem{\Delta[O_3]} at Esrange and Pallas in 2018.
-As data for 2019 has not yet been available from EBAS (last access, February 2020), we refrain from looking into these.
-
-In Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}, the probability density functions of daily mean ozone concentration residuals for Esrange and Pallas are shown.
-Both stations display a clear, but not significant, deviation from the climatological mean.
+To assess the significance of 2018/19 anomalies, we estimate the bias due to the change in the tropospheric background \chem{[O_3]} since the 1990s and account for the missing data in June 2018.
+To reconstruct these data, we performed a Reynold's decomposition (Eq.~(\ref{eq:reynolds_decomp})) and cross-calibrate with ozone obserations at Esrange and Pallas (compagnion paper).
+In the following, we will describe the bias-correction.
+
+In 2018, the probability density function of ozone concentration anomalies (\chem{\Delta[O_3]}) for Esrange and Pallas deviate clearly, but not significantly from the climatological mean (Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}).
 On average, \chem{[O_3]} had been elevated by $(1.9\pm 5.7)\,\unit{ppb}$ and $(1.9\pm 5.4)\,\unit{ppb}$, respectively.
@@
 
-Supplement Figure~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik} displays the probability density functions of daily mean ozone concentration residuals for Svanvik based on data taken in 2018/19 with respect to the climatology derived for northern Fennoscandia (denoted NF in Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}a) and the climatology derived for Svanvik (Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b).
-As previously mentioned, \chem{\left<[O_3]\right>} at Svanvik are $(6.6\pm 7.1)\,\unit{ppb}$ lower than the northern Fennoscandia climatology.
-For 2018 and 2019, we find $\Delta\chem{\left<[O_3]\right>} = (-3.8\pm 5.6)\,\unit{ppb}$ and $\Delta\chem{\left<[O_3]\right>} = (-5.2\pm 5.0)\,\unit{ppb}$, respectively.
-While the probability density of the residuals between the two climatologies is skewed towards negative values, the skew of 2018/19 is positive.
-With respect to the climatology for Svanvik (Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}b), the probability density functions average for 2018/19 lies at $\Delta\chem{\left<[O_3]\right>} = (2.6\pm 5.8)\,\unit{ppb}$ and $\Delta\chem{\left<[O_3]\right>} = (1.2\pm 5.0)\,\unit{ppb}$, respectively.
-Neither of the \chem{\Delta\left<[O_3]\right>} is significant by statistical means, but daily average ozone concentrations are slightly elevated in 2018 compared to 2019.
-
-The \chem{\left<\Delta\left<[O_3]\right>\right>} for Svanvik in 2018 is slightly larger than for Esrange and Pallas.
-This could mean either that 2018 had been more exceptional at Svanvik or the derived climatology may need a correction for increased tropospheric background \chem{[O_3]}.
-Based on the meteorological conditions, we assume 2019 as an average year which leads to an estimated constant bias of about $1.2\,\unit{ppb}$ in \chem{[O_3]}.
-Daily \chem{\Delta_{present-past}\left<[O_3]\right>} from Esrange and Pallas, wherein present is defined as the average of 2010--2018 and past as the respective beginning of data taking until 1997, do not reveal a clear trend: We find for Esrange $(1.5\pm 3.1)\,\unit{ppb}$ and for Pallas $(-2.5\pm 3.7)\,\unit{ppb}$ \citep{NILU2003, BER:Ruoho-Airola2015}.
-Former is compatible with our assumed bias for the climatology derived for Svanvik.
-
-To further investigate the ozone residuals, we conduct a Student's~t-test ($t = \frac{\chem{\Delta[O_3]}}{\sigma}$), assuming the same sample uncertainty $\sigma$ in both, climatology and the year in question.
+Similarly, we find a skew towards higher \chem{[O_3]} for Svanvik in 2018, while 2019 follows a normal distribution rather well (Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}).
+In conjunction with the meteorological conditions, we therefore can assume that 2019 had been a normal year and the annual average anomally in 2019 can be used for bias correction of the Svanvik climatology (\chem{\left<\Delta[O_3]\right>} = $(1.2\pm 5.0)\,\unit{ppb}$).
+To further strengthen this, we look at trends in ozone observations at Esrange and Pallas and compute the average difference between recent years and the mid 1990s (\chem{\Delta_{present-past}\left<[O_3]\right>}).
+Present is defined as the average over 2010--2018 and past as the time between the respective beginning of data taking until 1997.
+We do not find a clear trend.
+For Esrange \chem{\Delta_{present-past}\left<[O_3]\right>} amounts to $(1.5\pm 3.1)\,\unit{ppb}$ and for Pallas to $(-2.5\pm 3.7)\,\unit{ppb}$ \citep[][for comparison]{NILU2003, BER:Ruoho-Airola2015}.
+The observed trend at Esrange is compatible with our assumed bias correction for the Svanvik ozone climatology.
+
+%To further investigate the ozone residuals, we conduct a Student's~t-test ($t = \frac{\chem{\Delta[O_3]}}{\sigma}$), assuming the same sample uncertainty $\sigma$ in both, climatology and the year in question.
 Results for Esrange and Pallas are shown in Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test}.
@@
 
-In accordance to the above findings, we adjust the derived ozone climatology for Svanvik by adding a constant bias correction factor $\Delta\chem{[O_3]_\mathrm{bias-corr}} = 1.2\,\unit{ppb}$ and estimate the ozone concentration anomalies for 2018/19.
-In Fig.~\ref{fig:anomalies_svanvik}d), the resulting percentage of days above/below $\pm 1\,\sigma$ from the bias-corrected climatological mean is displayed for each month.
+In accordance to the above findings, we adjust the derived ozone climatology for Svanvik by adding a constant bias correction factor $\Delta\chem{[O_3]_\mathrm{bias-corr}} = 1.2\,\unit{ppb}$ and assess the monthly significance of the ozone concentration anomalies in 2018/19 (Fig.~\ref{fig:anomalies_svanvik}d)).
+The resulting percentage of days above/below $\pm 1\,\sigma$ from the bias-corrected climatological mean is displayed for each month in Fig~\ref{fig:data_svanvik_2018_2019}.
 Negative deviations from climatology are shown as a negative percentage.
-Note that missing bars for certain months are both due to data availability (see Fig~\ref{fig:data_svanvik_2018_2019} for details concerning Svanvik) and anomalies not meeting the criteria of $\pm 1\sigma$ deviation from the climatological mean.
-In April/May 2018, ozone was significantly enhanced, while 2019 was generally an average year.
-In both years, ozone concentrations were significantly below average in November.
-July 2018 represented by the color bar has not been corrected for the missing data due to data acquisition problems.
-The corrected value is indicated by the star.
-A detailed describtion of the gap-filling based on a Reynolds's decomposition method will be published seperately.
-
-To conclude, ozone concentrations in 2018 have been elevated compared to the respective climatologies, however, on average not significantly by statistical means.
+Note that missing bars for certain months are both due to data availability and anomalies not meeting the criteria for significance.
+The the star indicates resulting anomalies in after the reconstruction of missing values in July.
+In April/May 2018, ozone was significantly enhanced, while no significant enhancement is found in 2019.
+In both years, ozone concentrations were significantly lower than average in November.
+
+%To conclude, ozone concentrations in 2018 have been elevated compared to the respective climatologies, however, on average not significantly by statistical means.
 At Esrange and Pallas at least $25\,\unit{\%}$ (3.~quantile) of anomalies are significant on a $3\sigma$ level or more, while at Svanvik the 3.~quantile lies at about $2\sigma$.

### 2021-05-09 VegOzone.tex
Second revision of Sec 3.2.2

@@
 Similarly, we find a skew towards higher \chem{[O_3]} for Svanvik in 2018, while 2019 follows a normal distribution rather well (Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}).
-In conjunction with the meteorological conditions, we therefore can assume that 2019 had been a normal year and the annual average anomally in 2019 can be used for bias correction of the Svanvik climatology (\chem{\left<\Delta[O_3]\right>} = $(1.2\pm 5.0)\,\unit{ppb}$).
+In conjunction with the meteorological conditions, we therefore can assume that 2019 had been a normal year and that the annual average anomally in 2019 can be used for the bias correction of the Svanvik climatology (\chem{\left<\Delta[O_3]\right>}~=~$(1.2\pm 5.0)\,\unit{ppb}$).
 To further strengthen this, we look at trends in ozone observations at Esrange and Pallas and compute the average difference between recent years and the mid 1990s (\chem{\Delta_{present-past}\left<[O_3]\right>}).
@@
 Negative deviations from climatology are shown as a negative percentage.
-Note that missing bars for certain months are both due to data availability and anomalies not meeting the criteria for significance.
-The the star indicates resulting anomalies in after the reconstruction of missing values in July.
+Note that missing bars for certain months are both due to data availability and anomalies not meeting the $1\sigma$ criteria.
+The the star indicates resulting anomalies after the reconstruction of missing data in July.
 In April/May 2018, ozone was significantly enhanced, while no significant enhancement is found in 2019.
 In both years, ozone concentrations were significantly lower than average in November.
+These results mean, that the extensive forest fires in Sweden did not lead to a significant enhancement in \chem{[O_3]} at Svanhovd in 2018.
 

### 2021-05-09 VegOzone.tex
Move model describtion and gap-filling to appendix

@@
 \label{sec:do3se}
-To investigate which of the environmental factors might be most influential in determining \chem{O_3} damage; such as that observed on clover and tobaccos in 2018, we model ozone uptake of different natural and semi-natural vegetation using the $\mathrm{DO_3SE}$ model and develop bespoke subarctic parameterizations for common plant species.
-We first give an account of the $\mathrm{DO_3SE}$ model, input data, and the used methodology for deriving the bespoke parameters.
-Then we will present and discuss the modeling results.
-
-\subsection{Model description}
-\label{subsec:model_describtion}
-The opening of stomata is governed by different species-specific and environmental factors and can be described by a multiplicative model \citep{PTRS:Jarvis1976,Emberson2000,ICP:MappingManual2017}:
-\begin{equation}
-  g^k_\text{sto} = g^k_\text{max} \cdot f^k_\text{phen} \cdot f^k_\text{light} \cdot \max{\left\{f^k_\text{min}, f^k_\mathrm{T} \cdot f^k_\mathrm{VPD} \cdot f^k_\text{SWP}\right\}}.
-  \label{eq:stomatal}
-\end{equation}
-%
-Where $g_\mathrm{max}$ is the specific-specific maximum stomatal conductance value which is then modified by seasonal and environmental factors that vary within a range $0-1$.
-They are empirically determined and account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_\mathrm{T}$), water vapor pressure deficit ($f_\mathrm{VPD}$), and soil water potential ($f_\text{SWP}$).
-All factors differ with species or plant functional type (PFT) denoted with $k$.
-
-The temperature adjustment function is defined as
-%
-\begin{equation}
-  f_\mathrm{T} = \frac{T_\text{air}-T_\text{min}}{T_\text{opt}-T_\text{min}} \cdot \left(\frac{T_{\text{max}}-T_\text{air}}{T_{\text{max}}-T_\text{opt}}\right)^\beta,
-  \label{eq:f_temp}
-\end{equation}
-with $\beta = \frac{T_\text{max}-T_\text{opt}}{T_\text{opt}-T_\text{min}}$ a mixed reciprocal polynomial function of temperature.
-The shape parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various species or PFT and $T_\mathrm{air}$ is the near-surface air temperature.
-All temperatures are defined in units of \unit{^\circ C}.
-
-The water vapor pressure deficit function is
-%
-\begin{equation}
-  f_\mathrm{VPD} = f_\text{min}+(1-f_\text{min}) \cdot \frac{D_\text{min} - \text{VPD}}{D_\text{min}-D_\text{max}}
-  \label{eq:f_vpd}
-\end{equation}
-%
-where VPD is the leaf to air vapor pressure deficit in \unit{kPa} with $f_\text{min}$, $D_\text{min}$, $D_\text{max}$ describing the relative stomatal conductance to changes in vapor pressure deficit.
-
-The wavelength band $400-700\,\unit{nm}$ plant chlorophyll responds to is called photosynthetic active radiation (PAR).
-Its integral is the photosynthetic photon flux density (PPFD).
-The relationship between relative $g_\mathrm{sto}$ and PPFD is given by
-%
-\begin{equation}
-  f_\text{light} = 1-\text{exp}(-\alpha_\text{light}\cdot \text{PPFD})
-  \label{eq:flight}
-\end{equation}
-%
-where $\alpha_\mathrm{light}$ is a slope parameter describing the extent of stomatal opening at low light intensities.
-
-The $\mathrm{DO_3SE}$ model as described in \citet{ACP:Bueker2012} is used to simulate SWP across a species or PFT specific root depth according to the Penman--Monteith energy balance method that drives water cycling through the soil--plant--atmosphere system.
-\citet{ACP:Bueker2012} discuss several formulations available to parameterize the reduction in stomata conductance due to water content in the soil.
-Here we use the relationship between relative $g_\mathrm{sto}$ and soil water potential (SWP) given by:
-%
-\begin{equation}
-  f_\mathrm{SWP} = \text{min}\left\{1, \text{max}\left\{f_\mathrm{min}, \frac{(1 - f_\mathrm{min})\cdot(\mathrm{SWP_{min}} - \mathrm{SWP})}{\mathrm{SWP_{min}} - \mathrm{SWP_{max}}} + f_\mathrm{min} \right\} \right\}
-    \label{eq:fsw}
-\end{equation}
-%
-where SWP is the soil water potential across the root zone and $\mathrm{SWP_{min}}$ and $\mathrm{SWP_{max}}$ are the parameters describing the $f_\mathrm{SWP}$ relationship.
-
-To compute $\mathrm{POD_y}$, we estimate the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf.
-The terms $r_c$ (leaf surface resistance) and $r_b$ (quasi-laminar resistance) allow for the deposition of \chem{O_3} to the leaf and the fraction that is taken up via the stomata.
-%
-\begin{equation}
-  \Phi_\mathrm{sto} = \chem{[O_3]} \cdot \frac{u(z_1) \cdot g_\mathrm{sto} \cdot r_c}{r_b + r_c}.
-  \label{eq:flux_stomata}
-\end{equation}
-%
-The quasi-laminar boundary layer resistance is calculated by
-%
-\begin{equation}
-  r_b = 1.3 \cdot 150 \cdot \sqrt{\frac{L}{u(z_1)}}.
-  \label{eq:quasi_laminar_resist}
-\end{equation}
-%
-Where $L$ is the cross-wind leaf dimension, $u(z_1)$ the wind speed at height $z_1$, and the factor $1.3$ accounts for the diffusivity between heat and \chem{O_3}.
-
-\subsection{Input data and gap filling methodology}
-\label{subsec:ozone_reco}
-The $\mathrm{DO_3SE}$ model requires hourly, continuous meteorological observations.
-In addition to variables covered in Section~\ref{sec:data}, $2\,\unit{m}$ wind $u_\text{2m}$ and vapor pressure deficit VPD are needed.
-VPD has been calculated from observed $T_\text{2m}$ and relative humidity
-\begin{equation}
-  \text{VPD} = P_s(T) \cdot \left(1-\frac{\text{relHum}}{100}\right),
-\end{equation}
-with saturation vapor pressure $P_s(T)$ in \unit{hPa} derived from Arden Buck equation~\citep{JAP:Buck1981, Buck2012}.
-
-The following gap filling methodology was devised for the meteorological input data:
-\begin{itemize}
-\item Single hours of missing data were filled by taking the average of the hourly values coming before, and after, the missing value.
-\item Several consecutive hours of missing data (23 or less) were filled by taking the average of the corresponding hour the day before, and the day after; and repeating this for each missing hour of data.
-If data were unavailable from that hour of the previous day, then only the value from the day after was used and vice versa.
-Interpolated values were not used in calculating averages.
-\item Data gaps longer than $24\,\unit{h}$ were filled using weekly diurnal averages.
-i.e.
-an average was calculated using the corresponding hour throughout the week before and after.
-Interpolated values were not used in calculating averages.
-\end{itemize}
-
-For the missing ozone data, we applyed a Reynolds's decomposition of the form:
-\begin{equation}
-  \chem{[O_3]} = \chem{\left<[O_3]\right>} + \chem{\Delta[O_3]},
-  \label{eq:reynolds_decomp}
-\end{equation}
-with climatology \chem{\left<[O_3]\right>} and anomalies \chem{\Delta[O_3]}.
-Observed \chem{[O_3]} in 2018 at Pallas served as the reference.
-The exact procedure will be described in an compagnion paper.
+To investigate the difference between 2018 and 2019 concerning stomatal ozone uptake and to assess which of the environmental factors might be most influential in determining \chem{O_3} damage risk, we model ozone uptake of different natural and semi-natural vegetation using the $\mathrm{DO_3SE}$ model and develop bespoke subarctic parameterizations for common plant species.
+We the $\mathrm{DO_3SE}$ model and gapp-filling methods are described in Appendix~\ref{sec}.
+First, we give an account of the methodology for deriving bespoke parameters.
+Then, we will present and discuss the modeling results.
 
@@
 
-\section{DO3SE parameterization}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{figB1}
+\section{DO3SE model}
+\label{appendix:do3se_model}
+
+\subsection{Model description}
+\label{subsec:model_describtion}
+The opening of stomata is governed by different species-specific and environmental factors and can be described by a multiplicative model \citep{PTRS:Jarvis1976,Emberson2000,ICP:MappingManual2017}:
+\begin{equation}
+  g^k_\text{sto} = g^k_\text{max} \cdot f^k_\text{phen} \cdot f^k_\text{light} \cdot \max{\left\{f^k_\text{min}, f^k_\mathrm{T} \cdot f^k_\mathrm{VPD} \cdot f^k_\text{SWP}\right\}}.
+  \label{eq:stomatal}
+\end{equation}
+%
+Where $g_\mathrm{max}$ is the specific-specific maximum stomatal conductance value which is then modified by seasonal and environmental factors that vary within a range $0-1$.
+They are empirically determined and account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_\mathrm{T}$), water vapor pressure deficit ($f_\mathrm{VPD}$), and soil water potential ($f_\text{SWP}$).
+All factors differ with species or plant functional type (PFT) denoted with $k$.
+
+The temperature adjustment function is defined as
+%
+\begin{equation}
+  f_\mathrm{T} = \frac{T_\text{air}-T_\text{min}}{T_\text{opt}-T_\text{min}} \cdot \left(\frac{T_{\text{max}}-T_\text{air}}{T_{\text{max}}-T_\text{opt}}\right)^\beta,
+  \label{eq:f_temp}
+\end{equation}
+with $\beta = \frac{T_\text{max}-T_\text{opt}}{T_\text{opt}-T_\text{min}}$ a mixed reciprocal polynomial function of temperature.
+The shape parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various species or PFT and $T_\mathrm{air}$ is the near-surface air temperature.
+All temperatures are defined in units of \unit{^\circ C}.
+
+The water vapor pressure deficit function is
+%
+\begin{equation}
+  f_\mathrm{VPD} = f_\text{min}+(1-f_\text{min}) \cdot \frac{D_\text{min} - \text{VPD}}{D_\text{min}-D_\text{max}}
+  \label{eq:f_vpd}
+\end{equation}
+%
+where VPD is the leaf to air vapor pressure deficit in \unit{kPa} with $f_\text{min}$, $D_\text{min}$, $D_\text{max}$ describing the relative stomatal conductance to changes in vapor pressure deficit.
+
+The wavelength band $400-700\,\unit{nm}$ plant chlorophyll responds to is called photosynthetic active radiation (PAR).
+Its integral is the photosynthetic photon flux density (PPFD).
+The relationship between relative $g_\mathrm{sto}$ and PPFD is given by
+%
+\begin{equation}
+  f_\text{light} = 1-\text{exp}(-\alpha_\text{light}\cdot \text{PPFD})
+  \label{eq:flight}
+\end{equation}
+%
+where $\alpha_\mathrm{light}$ is a slope parameter describing the extent of stomatal opening at low light intensities.
+
+The $\mathrm{DO_3SE}$ model as described in \citet{ACP:Bueker2012} is used to simulate SWP across a species or PFT specific root depth according to the Penman--Monteith energy balance method that drives water cycling through the soil--plant--atmosphere system.
+\citet{ACP:Bueker2012} discuss several formulations available to parameterize the reduction in stomata conductance due to water content in the soil.
+Here we use the relationship between relative $g_\mathrm{sto}$ and soil water potential (SWP) given by:
+%
+\begin{equation}
+  f_\mathrm{SWP} = \text{min}\left\{1, \text{max}\left\{f_\mathrm{min}, \frac{(1 - f_\mathrm{min})\cdot(\mathrm{SWP_{min}} - \mathrm{SWP})}{\mathrm{SWP_{min}} - \mathrm{SWP_{max}}} + f_\mathrm{min} \right\} \right\}
+    \label{eq:fsw}
+\end{equation}
+%
+where SWP is the soil water potential across the root zone and $\mathrm{SWP_{min}}$ and $\mathrm{SWP_{max}}$ are the parameters describing the $f_\mathrm{SWP}$ relationship.
+
+To compute $\mathrm{POD_y}$, we estimate the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf.
+The terms $r_c$ (leaf surface resistance) and $r_b$ (quasi-laminar resistance) allow for the deposition of \chem{O_3} to the leaf and the fraction that is taken up via the stomata.
+%
+\begin{equation}
+  \Phi_\mathrm{sto} = \chem{[O_3]} \cdot \frac{u(z_1) \cdot g_\mathrm{sto} \cdot r_c}{r_b + r_c}.
+  \label{eq:flux_stomata}
+\end{equation}
+%
+The quasi-laminar boundary layer resistance is calculated by
+%
+\begin{equation}
+  r_b = 1.3 \cdot 150 \cdot \sqrt{\frac{L}{u(z_1)}}.
+  \label{eq:quasi_laminar_resist}
+\end{equation}
+%
+Where $L$ is the cross-wind leaf dimension, $u(z_1)$ the wind speed at height $z_1$, and the factor $1.3$ accounts for the diffusivity between heat and \chem{O_3}.
+
+\subsection{Input data and gap filling methodology}
+\label{subsec:ozone_reco}
+The $\mathrm{DO_3SE}$ model requires hourly, continuous meteorological observations.
+In addition to variables covered in Section~\ref{sec:data}, $2\,\unit{m}$ wind $u_\text{2m}$ and vapor pressure deficit VPD are needed.
+VPD has been calculated from observed $T_\text{2m}$ and relative humidity
+\begin{equation}
+  \text{VPD} = P_s(T) \cdot \left(1-\frac{\text{relHum}}{100}\right),
+\end{equation}
+with saturation vapor pressure $P_s(T)$ in \unit{hPa} derived from Arden Buck equation~\citep{JAP:Buck1981, Buck2012}.
+
+The following gap filling methodology was devised for the meteorological input data:
+\begin{itemize}
+\item Single hours of missing data were filled by taking the average of the hourly values coming before, and after, the missing value.
+\item Several consecutive hours of missing data (23 or less) were filled by taking the average of the corresponding hour the day before, and the day after; and repeating this for each missing hour of data.
+If data were unavailable from that hour of the previous day, then only the value from the day after was used and vice versa.
+Interpolated values were not used in calculating averages.
+\item Data gaps longer than $24\,\unit{h}$ were filled using weekly diurnal averages.
+i.e.
+an average was calculated using the corresponding hour throughout the week before and after.
+Interpolated values were not used in calculating averages.
+\end{itemize}
+
+For the missing ozone data, we applyed a Reynolds's decomposition of the form:
+\begin{equation}
+  \chem{[O_3]} = \chem{\left<[O_3]\right>} + \chem{\Delta[O_3]},
+  \label{eq:reynolds_decomp}
+\end{equation}
+with climatology \chem{\left<[O_3]\right>} and anomalies \chem{\Delta[O_3]}.
+Observed \chem{[O_3]} in 2018 at Pallas served as the reference.
+The exact procedure will be described in an compagnion paper.
+
+\subsection{Bespoke parameterizations}
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{figB1}
   \caption{Estimated begin (BGS) and end (EGS) of growing season for coniferous trees from MODIS Aqua/Terra net photosynthesis (PSN) product.
@@
   \label{fig:modis_Psn}
-\end{figure*}
+\end{figure}
 
@@
   %\includegraphics[width=8.3cm]{javis_funcs_rad_hist_evergreen}
-  \includegraphics[width=12cm]{figB2}
+  \includegraphics[width=8.3cm]{figB2}
 \caption{Construction of bespoke response functions for Norway spruce.
@@
   %\includegraphics[width=8.3cm]{javis_funcs_rad_hist_birch}
-  \includegraphics[width=12cm]{figB3}
+  \includegraphics[width=8.3cm]{figB3}
 \caption{Construction of bespoke response functions for downy birch.

### 2021-05-10 VegOzone.tex
Fix broken hyperkink
Write about modis retrieved start and end of greening season.

@@
 To investigate the difference between 2018 and 2019 concerning stomatal ozone uptake and to assess which of the environmental factors might be most influential in determining \chem{O_3} damage risk, we model ozone uptake of different natural and semi-natural vegetation using the $\mathrm{DO_3SE}$ model and develop bespoke subarctic parameterizations for common plant species.
-We the $\mathrm{DO_3SE}$ model and gapp-filling methods are described in Appendix~\ref{sec}.
+We the $\mathrm{DO_3SE}$ model and gapp-filling methods are described in Appendix~\ref{appendix:do3se_model}.
 First, we give an account of the methodology for deriving bespoke parameters.
@@
 
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{figA3}
-  \caption{Reconstruction of missing \chem{[O_3]} observation data at Svanvik in July 2018 based on derived hourly climatologies of northern Fennoscandia, Svanvik, and data Pallas.
+%\begin{figure*}[t]
+%  \includegraphics[width=12cm]{figA3}
+%  \caption{Reconstruction of missing \chem{[O_3]} observation data at Svanvik in July 2018 based on derived hourly climatologies of northern Fennoscandia, Svanvik, and data Pallas.
 (a) Time series; (b) anomalies; (c) reconstruction.}
-  \label{fig:ozone_reconstruction_2018_07}
-\end{figure*}
+%  \label{fig:ozone_reconstruction_2018_07}
+%\end{figure*}
 
@@
 
+To assess the $G_\mathrm{start}$  and $G_\mathrm{end}$ of the greening season for coniferous trees at Svanhovd in 2018/19, we used the net photosynthesis product ($A_\mathrm{net}$) of MODIS AQUA/TERRA over a $1\times 1\,\unit{km}$ area centered at Svanhovd.
+MODIS data indicate a higher photosynthetic activity in 2018 than in 2019.
+As shown in Fig.~\ref{fig:modis_Psn} we fitted a second order polynomal function of the general form
+%
+\begin{equation}
+A_\mathrm{net}(t) =  -m_0\cdot(t-m_1)^2+m_2,
+\end{equation}
+%
+with the form parameters $m_i$ through the data. Numerically, we retrieved the roots as $G_\mathrm{start}$ 122/106 and $G_\mathrm{end}$ 261/274 for 2018 and 2019, respectively.
+
 \begin{figure}[t]
   \includegraphics[width=8.3cm]{figB1}
-  \caption{Estimated begin (BGS) and end (EGS) of growing season for coniferous trees from MODIS Aqua/Terra net photosynthesis (PSN) product.
+  \caption{Estimated $G_\mathrm{start}$ and $G_\mathrm{end}$ of growing season for coniferous trees from MODIS Aqua/Terra net photosynthesis (PSN) product.
 A $1\times 1\,\unit{km}$ area around Svanhovd was selected.

### 2021-05-10 VegOzone.tex
Add text to appendices

@@
 
+As described in Section~\ref{subsec:do3se_parameters}, we developed bespoke parameterizations for natural and semi-natural vegetation at Svanhovd.
+Here, we show the temperature response and light response functions for coniferous and deciduous trees (Figs.~\ref{fig:f_temp_spruce}-\ref{fig:f_temp_birch}).
+
 \begin{figure}[t]
@@
   \includegraphics[width=8.3cm]{figB2}
-\caption{Construction of bespoke response functions for Norway spruce.
+\caption{Construction of bespoke response functions for coniferous trees.
 (a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively.
@@
 
+\subsection{Initial $\mathrm{DO_3SE}$ modeling results with mapping manual parameters}
+To demonstrate the necessary acclimation of the mapping manual parameters, we show the initial results of stomatal conductance at leaf-level ($G_\mathrm{sto}^\mathrm{leaf}$) and $\mathrm{POD_y}$ over \unit{doy} for both years and all species or PFTs (Fig.~\ref{fig:pody_mm_composit}).
+Observed \chem{[O_3]} is also indicated.
+From Fig.~\ref{fig:pody_mm_composit}f) it is apparent that the mapping manual parameterized grassland would not have been able to grow in 2019.
+
 \begin{figure*}[t]
@@
 \chem{[O_3]} are plotted on the same axis and scales as $G_\text{sto}^\text{leaf}$ but in units of $\unit{ppb}$.
-(a, b) Downy birch; (c, d) Norway spruce; (e, f) perennial grassland.
+(a, b) deciduous tree; (c, d) coniferous tree; (e, f) perennial grassland.
 From left to right: 2018, 2019.}
@@
 \end{figure*}
+
+\clearpage
 

### 2021-05-10 VegOzone.tex
Change naming from Norway spruce to coniferous

@@
 
-The main species or PFTs, we deduced on-site at the NIBIO Environment Centre Svanhovd and from Fig.~\ref{fig:svanhovd_research_station}, are perennial grassland, downy birch, and Norway spruce.
+The main species or PFTs, we deduced on-site at the NIBIO Environment Centre Svanhovd and from Fig.~\ref{fig:svanhovd_research_station}, are perennial grassland, birch, and pine.
 Default parameters for these species or PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as \emph{mapping manual} abbreviated as MM in the following.
+We base our parameterization of coniferous trees on boreal Norway spruce from the MM.
+Birch will be referred to as deciduous tree in the following.
 A comprehensive list of model parameters can be found in Supplement~Table~S1.
@@
 We show two climatological reference periods 1990s and 2000s to indicate that the vegetation is subject to ongoing climate change and hence continuous acclimation.
-Resulting response functions for Norway spruce (Picea abies) and downy birch are displayed in Appendix~Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}.
+Resulting response functions for coniferous and deciduous trees are displayed in Appendix~Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}.
 
@@
 We invert Eq.~(\ref{eq:inverse_function}) for $\alpha$ and find $\alpha'(\gamma')$.
-The resulting functions for perennial grassland are shown in Fig.~\ref{fig:f_temp_grassland}b) (for downy birch and Norway spruce refer Appendix~Figs.~\ref{fig:f_temp_spruce}b)--\ref{fig:f_temp_birch}b)).
+The resulting functions for perennial grassland are shown in Fig.~\ref{fig:f_temp_grassland}b) (for deciduous and coniferous trees refer Appendix~Figs.~\ref{fig:f_temp_spruce}b)--\ref{fig:f_temp_birch}b)).
 All derived parameters are tabulated in Table~\ref{tab:sensitivity_tests_light}.
@@
     \middlehline
-    \multirow{3}{*}{Downy birch} & MM & 5 & 20 & 100\\
+    \multirow{3}{*}{Deciduous tree} & MM & 5 & 20 & 100\\
     & Boreal & 0 & 10 & 100\\
@@
     \middlehline
-    \multirow{3}{*}{Norway spruce} & MM & 0 & 20 & 100\\
+    \multirow{3}{*}{Coniferous tree} & MM & 0 & 20 & 100\\
     & Boreal & 0 & 10 & 100\\
@@
     \middlehline
-    \multirow{3}{*}{Downy birch} & MM & 0.004 \\
+    \multirow{3}{*}{Deciduous tree} & MM & 0.004 \\
     & PPFD0.8 & 0.005\\
@@
     \middlehline
-    \multirow{3}{*}{Norway spruce} & MM & 0.006 \\
+    \multirow{3}{*}{Coniferous tree} & MM & 0.006 \\
     & PPFD0.8 & 0.008  \\
@@
 From the results shown in Fig.~\ref{fig:javis_func_opt_mean}, we identify boreal, PPFD0.8 as the best acclimation for all hypothetical species or PFTs alike.
-Overall, Norway spruce displays the least deviation between the different types of parameterizations, while the divergence for perennial grassland is substantial.
+As expected due to the small adjustments, coniferous trees display the lowest deviation between the different types of parameterizations, while the divergence for perennial grassland is substantial as response to the stronger proposed temperature acclimation.
 
@@
   \caption{Proposed metric to test bespoke response functions.
-GS (May--August) climatological averages and standard deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon (averaged over $11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning (averaged over $5-9\,\unit{am}$ local time) are shown for (a) Downy birch; (b) Norway spruce; (c) perennial grassland.
+GS (May--August) climatological averages and standard deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon (averaged over $11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning (averaged over $5-9\,\unit{am}$ local time) are shown for (a) deciduous tree; (b) coniferous tree; (c) perennial grassland.
 For simplicity, $f_\mathrm{SWP}$ is neglected in this calculation.
@@
 $A_\text{end}$ amounts to \unit{doy} 261 and 274, respectively, and is the value used for all PFTs.
-The resulting growing season for Norway spruce in 2019 was one month longer then in 2018.
+The resulting growing season for coniferous trees in 2019 was one month longer then in 2018.
 For estimating $A_\text{start}$ of downy birch, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Fig.~\ref{fig:greening_season_change_Svanvik}).
@@
     \middlehline
-    \multirow{2}{*}{Downy birch} & 2018 & 129$^*$ & 261$^\dagger$ \\
+    \multirow{2}{*}{Deciduous tree} & 2018 & 129$^*$ & 261$^\dagger$ \\
     & 2019 & 130$^*$ & 274$^\dagger$ \\
-    \multirow{2}{*}{Norway spruce} & 2018 & 122$^\dagger$ & 261$^\dagger$ \\
+    \multirow{2}{*}{Coniferous tree} & 2018 & 122$^\dagger$ & 261$^\dagger$ \\
     & 2019 & 106$^\dagger$ & 274$^\dagger$ \\
@@
 
-We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for three natural/semi-natural vegetation types, downy birch, Norway spruce, and perennial grassland using both MM parameterizations and a variety of bespoke response function in $f_\mathrm{T}$ and $f_\mathrm{light}$, and localized GS (described in detail in Section~\ref{subsec:do3se_parameters}).
+We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using both MM parameterizations and a variety of bespoke response function in $f_\mathrm{T}$ and $f_\mathrm{light}$, and localized GS (described in detail in Section~\ref{subsec:do3se_parameters}).
 The results are shown in Fig.~\ref{fig:pody_rel}.
@@
 
-For Norway spruce (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
-Norway spruce is the least sensitive to an acclimation in temperature and therefore the least biased compared to the generic default MM parameterization.
-Norway spruce displays a medium sensitivity to the variation in the light threshold, e.g.
+For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
+Coniferous trees' acclimation to prevailing temperatures was already good from the beginning and therefore the least biased of the generic MM parameterization.
+Coniferous trees display a medium sensitivity to the variation in the light threshold, e.g.
 $\Delta\mathrm{POD_1} = ^{+1.7}_{-1.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of boreal type without bespoke GS in 2018.
@@
 Bespoke parameterizations include temperature response functions for boreal, cold and light thresholds (PPFD0.8, PPFD1.2) at the end of a MM GS (lighter hue and dotted lines) and the estimated GS for Svanhovd (2018, 2019).
-(a) Downy birch; (b) Norway spruce; (c) Perennial grassland.
+(a) Deciduous tree; (b) coniferous tree; (c) perennial grassland.
 The dashed horizontal lines indicate critical levels defined in \citet{ICP:MappingManual2017,ESPR:Hayes2021} for each species, respectively.}
@@
 For downy birch and perennial grassland, substantial reductions in total biomass in comparison with the default generic MM parameterizations are predicted for both years.
-Norway spruce is the least affected and the magnitude of total biomass reduction is independent of the choice of the parameterization within given uncertainties.
-This may reflect the more comprehensive research on this species or PFT.
-The systematic deviation for downy birch is one order of magnitude larger than for Norway spruce and perennial grassland.
+Coniferous trees are the least affected and the magnitude of total biomass reduction is independent of the choice of the parameterization within given uncertainties.
+This may reflect the more comprehensive research on this species or PFT feeding into the generic MM parameterization.
+The systematic deviation for deciduous trees is one order of magnitude larger than for coniferous trees and perennial grassland.
 
@@
 Year & \multicolumn{4}{c}{PFT}\\
-& Downy birch & Norway spruce & \multicolumn{2}{c}{Perennial grassland}\\
+& Deciduous tree& Coniferous tree & \multicolumn{2}{c}{Perennial grassland}\\
 \middlehline
@@
 We found that the length and timing of the growing season are crucial for risk assessment based on accumulated ozone flux metrics ($\mathrm{POD_y}$).
-Based on a MODIS photosynthesis product, we estimated the growing season for Norway spruce in 2018 to be at least one month shorter than in 2019.
+Based on a MODIS photosynthesis product, we estimated the growing season for coniferous trees in 2018 to be at least one month shorter than in 2019.
 The found start of growing season in 2018 ($122\,\unit{doy}$) and 2019 ($106\,\unit{doy}$) are within given ranges by empirical observations \citep{TB:Kolari2007,IVL:Karlsson2018}.
@@
 Based on the methods described in \citet{ICP:MappingManual2017}, we estimated biomass reductions due to ozone uptake for both 2018 and 2019 and find significant reductions for downy birch ($15.5-17.4\,\unit{\%}$) and perennial grassland (above ground, $12.4-13.7\,\unit{\%}$) with respect to estimates based on $\mathrm{POD_1}$ from default parameterizations.
-For Norway spruce, the results are compatible with the default parameterization estimates and are contained within the range of systematic uncertainties deduced from our bespoke response function ($2.5-3\,\unit{\%}$).
+For coniferous trees, the results are compatible with the default parameterization estimates and are contained within the range of systematic uncertainties deduced from our bespoke response function ($2.5-3\,\unit{\%}$).
 These results have to be treated with care because the biomass reduction functions were not established based on subarctic species and might as well be subject to biases.
@@
 This may explain contradictory results regarding the ozone sensitivity of natural vegetation in northern Europe.
-While \citet{FS:Subramanian2014} reported a modeled reduction of net primary production in Norway spruce ($4.3-15.5\,\unit{\%}$) and birch ($1.4-4.3\,\unit{\%}$) under elevated ozone in Sweden, Lithuanian forest deciduous trees were observed to be more susceptible to visible damage by ozone than Norway spruce \citep{Amb:Girgzdiene2009}.
+While \citet{FS:Subramanian2014} reported a modeled reduction of net primary production in coniferous ($4.3-15.5\,\unit{\%}$) and birch ($1.4-4.3\,\unit{\%}$) under elevated ozone in Sweden, Lithuanian forest deciduous trees were observed to be more susceptible to visible damage by ozone than Norway spruce \citep{Amb:Girgzdiene2009}.
 Open-top chamber (OTC) experiments performed in northern Finland where Scots pine and downy birch were exposed to elevated ozone concentrations showed biomass reductions of both downy birch and mountain birch seedlings after one season of exposure \citep{Amb:Manninen2009}.

### 2021-05-10 VegOzone.tex
Change from downy birch to decideous tree

@@
 Based on the time series of observed \chem{CO_2} uptake and temperature at two observation sites in Southern ($\rightarrow$ \emph{cold}) and Northern ($\rightarrow$ \emph{boreal}) Finland \citep{TB:Kolari2007}, we estimate $10\,\unit{^\circ C}\le T_\mathrm{opt} \le 15\,\unit{^\circ C}$.
-We assume a similar acclimation for downy birch.
+We assume a similar acclimation for deciduous trees.
 All bespoke temperature parameters are listed in Table~\ref{tab:sensitivity_tests_temp}.\\
@@
 The resulting growing season for coniferous trees in 2019 was one month longer then in 2018.
-For estimating $A_\text{start}$ of downy birch, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Fig.~\ref{fig:greening_season_change_Svanvik}).
+For estimating $A_\text{start}$ of deciduous trees, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Fig.~\ref{fig:greening_season_change_Svanvik}).
 We find 129 and 130, respectively.
@@
 
-For downy birch (Fig.~\ref{fig:pody_rel}a) all simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
-Downy birch displays the largest variance of ozone uptake among all species or PFTs ($10-31\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
+For deciduous trees (Fig.~\ref{fig:pody_rel}a) all simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
+Deciduous trees display the largest variance of ozone uptake among all species or PFTs ($10-31\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
 In particular, the sensitivity to the light threshold is especially pronounced, e.g.
@@
 For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations with bespoke GS stay below the CL.
-Perennial grassland shows the smallest ozone uptake, but a similarly large bias as downy birch.
+Perennial grassland shows the smallest ozone uptake, but a similarly large bias as deciduous trees.
 We also find a reversal for predicted ozone damage risk in $\mathrm{POD_1}$ between 2018 and 2019 for the boreal type.
@@
 The reported range in uncertainty corresponds to the maximum deviation defined by the $\pm 20\,\unit{\%}$ variation of the extent of stomatal opening at low light intensities with SWP=on.
-For downy birch and perennial grassland, substantial reductions in total biomass in comparison with the default generic MM parameterizations are predicted for both years.
+For deciduous trees and perennial grassland, substantial reductions in total biomass in comparison with the default generic MM parameterizations are predicted for both years.
 Coniferous trees are the least affected and the magnitude of total biomass reduction is independent of the choice of the parameterization within given uncertainties.
@@
 This means that the choice of temperature acclimation has the potential to either revoke or amplify the risk of ozone damage for vegetation.
-However, the risk of ozone damage in downy birch and perennial grassland in comparison with the generic default MM parameterizations is increased, regardless.
+However, the risk of ozone damage in deciduous trees and perennial grassland in comparison with the generic MM parameterizations is increased, regardless.
 
@@
 The found start of growing season in 2018 ($122\,\unit{doy}$) and 2019 ($106\,\unit{doy}$) are within given ranges by empirical observations \citep{TB:Kolari2007,IVL:Karlsson2018}.
-Based on gridded observational temperature and snow depth, we found a half a month shorter growing season for downy birch and perennial grassland in 2018 compared to 2019, respectively.
+Based on gridded observational temperature and snow depth, we found a half a month shorter growing season for deciduous trees and perennial grassland in 2018 compared to 2019, respectively.
 The timing and emerging trend in the start of the growing season may differ substantially between local observation and common thermal or latitudinal estimations.
@@
 
-Based on the methods described in \citet{ICP:MappingManual2017}, we estimated biomass reductions due to ozone uptake for both 2018 and 2019 and find significant reductions for downy birch ($15.5-17.4\,\unit{\%}$) and perennial grassland (above ground, $12.4-13.7\,\unit{\%}$) with respect to estimates based on $\mathrm{POD_1}$ from default parameterizations.
+Based on the methods described in \citet{ICP:MappingManual2017}, we estimated biomass reductions due to ozone uptake for both 2018 and 2019 and find significant reductions for deciduous trees ($15.5-17.4\,\unit{\%}$) and perennial grassland (above ground, $12.4-13.7\,\unit{\%}$) with respect to estimates based on $\mathrm{POD_1}$ from default parameterizations.
 For coniferous trees, the results are compatible with the default parameterization estimates and are contained within the range of systematic uncertainties deduced from our bespoke response function ($2.5-3\,\unit{\%}$).
@@
   \includegraphics[width=8.3cm]{figB3}
-\caption{Construction of bespoke response functions for downy birch.
+\caption{Construction of bespoke response functions for deciduous trees.
 (a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively.

### 2021-05-10 VegOzone.tex
Fix typo in table
Include gamma(0.5) value in table
Fix labeling of critical levels

@@
   \label{tab:sensitivity_tests_light}
-  \begin{tabular}{llr}
+  \begin{tabular}{llcr}
     \tophline
-    Species & type & $\alpha$\\
+    Species & type & $\alpha$ & $\gamma(0.5)$\\
     \middlehline
-    \multirow{3}{*}{Deciduous tree} & MM & 0.004 \\
-    & PPFD0.8 & 0.005\\
-    & PPFD1.2 & 0.004\\
+    \multirow{3}{*}{Deciduous tree} & MM & 0.004 & 165.035\\
+    & PPFD0.8 & 0.005 & 132.028\\
+    & PPFD1.2 & 0.003 & 198.042\\
     \middlehline
-    \multirow{3}{*}{Coniferous tree} & MM & 0.006 \\
-    & PPFD0.8 & 0.008  \\
-    & PPFD1.2 & 0.005  \\
+    \multirow{3}{*}{Coniferous tree} & MM & 0.006 & 115.525\\
+    & PPFD0.8 & 0.008 & 92.420\\
+    & PPFD1.2 & 0.005 & 138.629\\
     \middlehline
-    \multirow{3}{*}{Perennial grassland} & MM & 0.011\\
-    & PPFD0.8 & 0.014 \\
-    & PPFD1.2 & 0.009 \\
+    \multirow{3}{*}{Perennial grassland} & MM & 0.011 & 63.013\\
+    & PPFD0.8 & 0.014 & 50.411\\
+    & PPFD1.2 & 0.009 & 75.616\\
     \bottomhline
@@
 The results are shown in Fig.~\ref{fig:pody_rel}.
-For comparison, critical levels for a $4\,\unit{\%}$ (silver birch), $2\,\unit{\%}$ (Norway spruce) \citep{ICP:MappingManual2017}, and $10\,\unit{\%}$ (perennial grassland, above ground) \citep{ESPR:Hayes2021} reduction in biomass are shown as horizontal dashed lines.
+For comparison, critical levels for a $4\,\unit{\%}$ (deciduous), $2\,\unit{\%}$ (coniferous) \citep[refer to silver birch and Norway spruce in][]{ICP:MappingManual2017}, and $10\,\unit{\%}$ (perennial grassland, above ground) \citep{ESPR:Hayes2021} reduction in biomass are shown as horizontal dashed lines.
 Different years are color coded as previously.

### 2021-05-10 VegOzone.tex
Fix typo of DO3SE

@@
 The results depicted in Fig.~\ref{fig:pody_rel} allow for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric for ozone induced damage in subarctic vegetation.
-In summary, we identify the following main effects of the bespoke parameterizations of the $DO_3SE$ model with respect to the MM default parameterization on ozone uptake:
+In summary, we identify the following main effects of the bespoke parameterizations of the $\mathrm{DO_3SE}$ model with respect to the MM default parameterization on ozone uptake:
 \begin{itemize}

### 2021-05-10 VegOzone.tex
Fix use of term greening

@@
 This is in line with a general decrease in tropospheric ozone with decreasing altitude.
-The earlier decline in \chem{\left<[O_3]\right>} may point to an influence of coniferous vegetation on the removal of ozone since \chem{CO_2} uptake by coniferous trees has been observed as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013} compared to an estimated average begin of greening ($\left<G_\mathrm{start}\right>$) on $(150\pm 14)\,\unit{doy}$ at Svanvik (see Section~\ref{subsec:do3se_parameters} for details).
+The earlier decline in \chem{\left<[O_3]\right>} may point to an influence of coniferous vegetation on the removal of ozone since \chem{CO_2} uptake by coniferous trees has been observed as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013} compared to an estimated average begin of growing season ($\left<G_\mathrm{start}\right>$) on $(150\pm 14)\,\unit{doy}$ at Svanvik (see Section~\ref{subsec:do3se_parameters} for details).
 In July--September, ozone is occasionally almost completely depleted during nighttime.
@@
 
-To assess the $G_\mathrm{start}$  and $G_\mathrm{end}$ of the greening season for coniferous trees at Svanhovd in 2018/19, we used the net photosynthesis product ($A_\mathrm{net}$) of MODIS AQUA/TERRA over a $1\times 1\,\unit{km}$ area centered at Svanhovd.
+To assess the $G_\mathrm{start}$  and $G_\mathrm{end}$ of the growing season for coniferous trees at Svanhovd in 2018/19, we used the net photosynthesis product ($A_\mathrm{net}$) of MODIS AQUA/TERRA over a $1\times 1\,\unit{km}$ area centered at Svanhovd.
 MODIS data indicate a higher photosynthetic activity in 2018 than in 2019.

### 2021-05-10 VegOzone.tex
Fix naming convention boreal -> subarctic

@@
 These results mean, that the extensive forest fires in Sweden did not lead to a significant enhancement in \chem{[O_3]} at Svanhovd in 2018.
+However, there are indications that the sites further south at Esrange and Pallas experienced more days with elevated ozone in 2018 than Svanvik.
 
@@
 
-We define two different hypothetical temperature acclimation: \emph{Boreal} and \emph{cold}.
-The most extreme of the two acclimations is \emph{boreal}.
+We define two different hypothetical temperature acclimation: \emph{subarctic} and \emph{cold}.
+The most extreme of the two acclimations is \emph{subarctic}.
 For either, we assume that the hypothetical species are more tolerant to cold conditions and shift $f_\mathrm{T}$ towards colder temperatures accordingly.
 We construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to MM.
-In the same way, \emph{boreal} refers to a hypothetical species which is very tolerant to cold but is sensitive to high temperatures, and most efficient at cool temperatures ($T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000).
+In the same way, \emph{subarctic} refers to a hypothetical species which is very tolerant to cold but is sensitive to high temperatures, and most efficient at cool temperatures ($T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000).
 For perennial grassland the resulting response functions are depicted together with the PDF of temperature in Fig.~\ref{fig:f_temp_grassland}a).
@@
 Norway spruce was experimentally found to be active already at rather low air temperatures and could reach $60\,\unit{\%}$ photosynthetic activity as early as \unit{doy}~100 \citep{TB:Kolari2007, TP:Wallin2013}.
-Based on the time series of observed \chem{CO_2} uptake and temperature at two observation sites in Southern ($\rightarrow$ \emph{cold}) and Northern ($\rightarrow$ \emph{boreal}) Finland \citep{TB:Kolari2007}, we estimate $10\,\unit{^\circ C}\le T_\mathrm{opt} \le 15\,\unit{^\circ C}$.
+Based on the time series of observed \chem{CO_2} uptake and temperature at two observation sites in Southern ($\rightarrow$ \emph{cold}) and Northern ($\rightarrow$ \emph{subarctic}) Finland \citep{TB:Kolari2007}, we estimate $10\,\unit{^\circ C}\le T_\mathrm{opt} \le 15\,\unit{^\circ C}$.
 We assume a similar acclimation for deciduous trees.
@@
     \multirow{3}{*}{Deciduous tree} & MM & 5 & 20 & 100\\
-    & Boreal & 0 & 10 & 100\\
     & Cold & 5 & 15 & 100\\
+    & Subarctic & 0 & 10 & 100\\
     \middlehline
     \multirow{3}{*}{Coniferous tree} & MM & 0 & 20 & 100\\
-    & Boreal & 0 & 10 & 100\\
     & Cold & 0 & 15 & 100\\
+    & Subarctic & 0 & 10 & 100\\
     \middlehline
     \multirow{3}{*}{Perennial grassland} & MM & 10 & 24 & 36\\
-    & Boreal & 0 & 10 & 24 \\
     & Cold & 5 & 15 & 36\\
+    & Subarctic & 0 & 10 & 24\\
     \bottomhline
@@
 For simplicity, we neglect the dependency on SWP in this regard.
-From the results shown in Fig.~\ref{fig:javis_func_opt_mean}, we identify boreal, PPFD0.8 as the best acclimation for all hypothetical species or PFTs alike.
+From the results shown in Fig.~\ref{fig:javis_func_opt_mean}, we identify \emph{subarctic}, PPFD0.8 as the best acclimation for all hypothetical species or PFTs alike.
 As expected due to the small adjustments, coniferous trees display the lowest deviation between the different types of parameterizations, while the divergence for perennial grassland is substantial as response to the stronger proposed temperature acclimation.
@@
 In particular, the sensitivity to the light threshold is especially pronounced, e.g.
-$\Delta\mathrm{POD_1} = ^{+2.5}_{-3.6}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of the boreal type without bespoke GS in 2019.
+$\Delta\mathrm{POD_1} = ^{+2.5}_{-3.6}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of the subarctic type without bespoke GS in 2019.
 Without bespoke GS, 2018 displays a higher ozone damage risk than 2019.
-However, taking the later start and overall shorter GS in 2018 into account, this is reversed for the boreal acclimation type.
+However, taking the later start and overall shorter GS in 2018 into account, this is reversed for the \emph{subarctic} acclimation type.
 This can be partly explained by very high light response factors (over $f_\mathrm{light}>0.8$) in the beginning of the GS in these simulations.
@@
 Coniferous trees display a medium sensitivity to the variation in the light threshold, e.g.
-$\Delta\mathrm{POD_1} = ^{+1.7}_{-1.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of boreal type without bespoke GS in 2018.
+$\Delta\mathrm{POD_1} = ^{+1.7}_{-1.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of subarctic type without bespoke GS in 2018.
 Taking the bespoke GS into account, we find a higher ozone uptake in 2019 for all bespoke temperature response functions.
-The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+The overall bias deduced from bespoke GS for subarctic and cold types with respect to MM ranges between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
 The maximum systematic deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of the extent of stomatal opening at low light intensities amount to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
@@
 Perennial grassland shows the smallest ozone uptake, but a similarly large bias as deciduous trees.
-We also find a reversal for predicted ozone damage risk in $\mathrm{POD_1}$ between 2018 and 2019 for the boreal type.
+We also find a reversal for predicted ozone damage risk in $\mathrm{POD_1}$ between 2018 and 2019 for the subarctic type.
 Perennial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant without bespoke GS.
-The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+The overall bias deduced from bespoke GS for subarctic and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
 The maximum systematic deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of the extent of stomatal opening at low light intensities amount to $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
@@
   \caption{$\mathrm{DO_3SE}$ modeling results for $\mathrm{POD_1}$ with default parameterization (MM) and bespoke parameters for Svanhovd.
-Bespoke parameterizations include temperature response functions for boreal, cold and light thresholds (PPFD0.8, PPFD1.2) at the end of a MM GS (lighter hue and dotted lines) and the estimated GS for Svanhovd (2018, 2019).
+Bespoke parameterizations include temperature response functions for subarctic, cold and light thresholds (PPFD0.8, PPFD1.2) at the end of a MM GS (lighter hue and dotted lines) and the estimated GS for Svanhovd (2018, 2019).
 (a) Deciduous tree; (b) coniferous tree; (c) perennial grassland.
@@
 
-We estimated the biomass reductions in accordance to the CLs in \citet{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme temperature acclimation (\emph{boreal}) for each bespoke PFT.
+We estimated the biomass reductions in accordance to the CLs in \citet{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme temperature acclimation (\emph{subarctic}) for each bespoke PFT.
 The estimates are listed in Table~\ref{tab:biomass_reduction}.
@@
 \begin{table}[t]
-  \caption{Estimated total biomass reduction in \unit{\%} for temperature acclimation \emph{boreal} with bespoke GS and SWP=off.
+  \caption{Estimated total biomass reduction in \unit{\%} for temperature acclimation \emph{subarctic} with bespoke GS and SWP=off.
 The uncertainty ranges reported here correspond to maximum a divergence deducted from varying the extent of stomatal opening at low light intensities and SWP=on.

### 2021-05-10 VegOzone.tex
Fix phrasing
Fix position of appendix figures
Switch figure - need to rename them again
Fix table legend and table refs

@@
 Locations of past and present (end of 2018) ozone observation sites in northern Fennoscandia were used in this study.
-For more details see Tab.~\ref{tab:ebas_obs}.
+For more details see Table~\ref{tab:ebas_obs}.
 The same color coding for each station as displayed herein will be used in all the following figures.}
@@
   %\includegraphics[width=8.3cm]{javis_funcs_rad_hist_grassland}
-  \includegraphics[width=8.3cm]{fig09}
+  \includegraphics[width=8.3cm]{fig07fin}
 \caption{Construction of bespoke response functions for grassland.
@@
 Another important factor for ozone exposure metrics is the timing and length of the GS.
-Start ($A_\text{start}$) and end ($A_\text{end}$) of GS for each PFT are given in units of \unit{doy} (Tab.~\ref{tab:sensitivity_tests_gs}) and estimated as follows.
+Start ($A_\text{start}$) and end ($A_\text{end}$) of GS for each PFT are given in units of \unit{doy} (Table~\ref{tab:sensitivity_tests_gs}) and estimated as follows.
 Evergreen needle leaf trees show photosynthetic activity already at low temperatures \citep{TP:Wallin2013, TB:Kolari2007}.
@@
 By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{doy}$ for 2018 and \unit{doy} 106 in 2019.
-$A_\text{end}$ amounts to \unit{doy} 261 and 274, respectively, and is the value used for all PFTs.
+$A_\text{end}$ amounts to \unit{doy} 261 and 274, respectively.
+This value will be used for all PFTs alike.
 The resulting growing season for coniferous trees in 2019 was one month longer then in 2018.
@@
   \includegraphics[width=8.3cm]{fig11}
-  \caption{$\mathrm{DO_3SE}$ modeling results for $\mathrm{POD_1}$ with default parameterization (MM) and bespoke parameters for Svanhovd.
-Bespoke parameterizations include temperature response functions for subarctic, cold and light thresholds (PPFD0.8, PPFD1.2) at the end of a MM GS (lighter hue and dotted lines) and the estimated GS for Svanhovd (2018, 2019).
-(a) Deciduous tree; (b) coniferous tree; (c) perennial grassland.
-The dashed horizontal lines indicate critical levels defined in \citet{ICP:MappingManual2017,ESPR:Hayes2021} for each species, respectively.}
+  \caption{$\mathrm{DO_3SE}$ modeling $\mathrm{POD_1}$ results for (a) deciduous trees, (b) coniferous trees, and (c) perennial grasslands at Svanhovd in 2018 (pink) and 2019 (purple).
+The results varied depending on parameterizations of temperature response functions (MM, Cold, and Subarctic, details in Table~\ref{tab:sensitivity_tests_temp}), growing season (MM (squares) or bespoke (circles), details in Table~\ref{tab:sensitivity_tests_gs}), light response functions (first, second and third point attached by a line are depicting the MM, PPFD0.8 and PPFD1.2, details in Table~\ref{tab:sensitivity_tests_light}), and the effect of taking SWP into account (filled symbols without and open symbols with effects of drought kept in the model).
+Critical levels for ozone damage \citet{ICP:MappingManual2017,ESPR:Hayes2021} are given as dashed horizontal lines.
+Note that the horizontal axis is only categorical..}
   \label{fig:pody_rel}
 \end{figure}
+
+DO3SE modelling POD1 results (Mills et al., 2017; Hayes et al., 2021) are given as dashed horizontal lines. Note that the horizontal axis is only categorical.
 
@@
 \appendixfigures  %% needs to be added in front of appendix figures
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{figA1}
+\begin{figure}[th]
+  \includegraphics[width=8.3cm]{figA1}
   \caption{Observed global irradiance at Svanvik.
@@
   \label{fig:global_rad_clim}
-\end{figure*}
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{figA2}
+\end{figure}
+
+\begin{figure}[th]
+  \includegraphics[width=8.3cm]{figA2}
   \caption{Probability densities and correlation coefficient for \chem{[O_3]} between different sites in northern Fennoscandia.
@@
   \label{fig:density_distribution}
-\end{figure*}
+\end{figure}
 
@@
 
-\begin{figure}[t]
+\begin{figure}[th]
   \includegraphics[width=8.3cm]{figB1}

### 2021-05-10 VegOzone.tex
Fix dot at the end of caption
Remove remnand text

@@
 GS (May--August) climatological averages and standard deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon (averaged over $11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning (averaged over $5-9\,\unit{am}$ local time) are shown for (a) deciduous tree; (b) coniferous tree; (c) perennial grassland.
-For simplicity, $f_\mathrm{SWP}$ is neglected in this calculation.
-}
+For simplicity, $f_\mathrm{SWP}$ is neglected in this calculation.}
   \label{fig:javis_func_opt_mean}
@@
 Critical levels for ozone damage \citet{ICP:MappingManual2017,ESPR:Hayes2021} are given as dashed horizontal lines.
-Note that the horizontal axis is only categorical..}
+Note that the horizontal axis is only categorical.}
   \label{fig:pody_rel}
 \end{figure}
-
-DO3SE modelling POD1 results (Mills et al., 2017; Hayes et al., 2021) are given as dashed horizontal lines. Note that the horizontal axis is only categorical.
 

### 2021-05-10 VegOzone.tex
Rename and remove figures

@@
   %\includegraphics[width=8.3cm]{climatediagram}
-  \includegraphics[width=8.3cm]{fig05}
+  \includegraphics[width=8.3cm]{fig04}
   \caption{Climate diagram following Walter and Lieth.
@@
   %\includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_spline}
-  \includegraphics[width=8.3cm]{fig06}
+  \includegraphics[width=8.3cm]{fig05}
   \caption{Ozone climatologies.
@@
 \begin{figure}[t]
-  \includegraphics[width=12cm]{fig08}
+  \includegraphics[width=12cm]{fig06}
   \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as percentages of days significantly deviating from climatological mean through standard deviation for each month.
@@
   %\includegraphics[width=8.3cm]{javis_funcs_rad_hist_grassland}
-  \includegraphics[width=8.3cm]{fig07fin}
+  \includegraphics[width=8.3cm]{fig07}
 \caption{Construction of bespoke response functions for grassland.
@@
   %\includegraphics[width=8.3cm]{javis_func_opt_mean}
-  \includegraphics[width=8.3cm]{fig10}
+  \includegraphics[width=8.3cm]{fig08}
   \caption{Proposed metric to test bespoke response functions.
-GS (May--August) climatological averages and standard deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon (averaged over $11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning (averaged over $5-9\,\unit{am}$ local time) are shown for (a) deciduous tree; (b) coniferous tree; (c) perennial grassland.
-For simplicity, $f_\mathrm{SWP}$ is neglected in this calculation.}
+GS (May--August) climatological averages and standard deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon (averaged over $11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning (averaged over $5-9\,\unit{am}$ local time) are shown for (a) deciduous trees; (b) coniferous trees; (c) perennial grassland.
+For $f_\mathrm{SWP}$ is connected to a plant hydraulic model within $\mathrm{DO_3SE}$ it left out in this calculation.}
   \label{fig:javis_func_opt_mean}
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{fig11}
+  \includegraphics[width=8.3cm]{fig09}
   \caption{$\mathrm{DO_3SE}$ modeling $\mathrm{POD_1}$ results for (a) deciduous trees, (b) coniferous trees, and (c) perennial grasslands at Svanhovd in 2018 (pink) and 2019 (purple).

### 2021-05-10 VegOzone.tex
Clean up supplement

@@
 Only a few data have been taken upstream the Pasvik river at Janiskovski, hence these will not be taken into further consideration, but are listed for completeness.
-There is an overlap of about $4\,\unit{years}$ between data taken at Esrange, Jergul/Karasjok, and Svanvik (Fig.~\ref{fig:ozone_timesseries_fenoscandic_obs}).
+There is an overlap of about $4\,\unit{years}$ between data taken at Esrange, Jergul/Karasjok, and Svanvik (Supplement~Fig.~\ref{fig:ozone_timesseries_fenoscandic_obs}).
 
@@
 
-In 2018, the probability density function of ozone concentration anomalies (\chem{\Delta[O_3]}) for Esrange and Pallas deviate clearly, but not significantly from the climatological mean (Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}).
+In 2018, the probability density function of ozone concentration anomalies (\chem{\Delta[O_3]}) for Esrange and Pallas deviate clearly, but not significantly from the climatological mean (Supplement~Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}).
 On average, \chem{[O_3]} had been elevated by $(1.9\pm 5.7)\,\unit{ppb}$ and $(1.9\pm 5.4)\,\unit{ppb}$, respectively.
@@
 
-Similarly, we find a skew towards higher \chem{[O_3]} for Svanvik in 2018, while 2019 follows a normal distribution rather well (Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}).
+Similarly, we find a skew towards higher \chem{[O_3]} for Svanvik in 2018, while 2019 follows a normal distribution rather well (Supplement~Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}).
 In conjunction with the meteorological conditions, we therefore can assume that 2019 had been a normal year and that the annual average anomally in 2019 can be used for the bias correction of the Svanvik climatology (\chem{\left<\Delta[O_3]\right>}~=~$(1.2\pm 5.0)\,\unit{ppb}$).
@@
 The resulting growing season for coniferous trees in 2019 was one month longer then in 2018.
-For estimating $A_\text{start}$ of deciduous trees, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Fig.~\ref{fig:greening_season_change_Svanvik}).
+For estimating $A_\text{start}$ of deciduous trees, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Supplement~Fig.~\ref{fig:greening_season_change_Svanvik}).
 We find 129 and 130, respectively.

### 2021-05-10 supplements/VegOzone_supplement.tex
Remove figure
Clean up supplement

@@
-\graphicspath{{pictures/}}          % graphics
+\graphicspath{{figures_supplement/}}          % graphics
 \renewcommand{\thefigure}{S\arabic{figure}}
@@
 \setcounter{page}{1}
+\setcounter{section}{1}
 \setcounter{figure}{0}
 
-\begin{figure*}[t]
+\section*{Supplement}
+
+\begin{figure*}[th]
   \includegraphics[width=12cm]{ozone_timeseries_fennoscandic_obs.png}
-  %\includegraphics[width=12cm]{fig04}
   \caption{Time series of ozone observations in northern Fennoscandia (Tab.~\ref{tab:ebas_obs}) (1986--2019).
@@
 
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{greening_season_change_Svanvik}
-  \caption{Estimated shift and prolongation of greening season at Svanhovd over the past 6 decades based on data from \citet{SeNorge}.}
-  \label{fig:greening_season_change_Svanvik}
-\end{figure*}
-
-\begin{figure*}[t]
+\begin{figure*}[th]
   \includegraphics[width=12cm]{ozone_climatology_fenoscandic_obs_residuals}
@@
 
-\begin{figure*}[t]
+\begin{figure*}[th]
   \includegraphics[width=12cm]{ozone_climatology_fenoscandic_obs_residuals-Svanvik}
@@
 
-\begin{figure*}[t]
+\begin{figure*}[th]
   \includegraphics[width=12cm]{ozone_climatology_fenoscandic_obs_test}
@@
 
-\begin{figure*}[t]
+\begin{figure*}[th]
   \includegraphics[width=12cm]{ozone_climatology_fenoscandic_obs_test-Svanvik}
@@
 
+\begin{figure*}[th]
+  \centering
+  \includegraphics[width=1\textwidth]{supplements/supplement_do3se-parameterization_v2}
+\label{tab:do3se_mm_param}
+\end{figure*}
+
 \begin{figure*}[t]
-  \includegraphics[width=12cm]{ozone_observation_timelag_svanvik}
-  \caption{Temporal correlation of \chem{[O_3]} data at Svanvik with other ozone monitoring stations in northern Fennoscandia.
+  \includegraphics[width=12cm]{greening_season_change_Svanvik}
+  \caption{Estimated shift and prolongation of growing season at Svanhovd over the past 6 decades based on data from \citet{SeNorge}.}
+  \label{fig:greening_season_change_Svanvik}
+\end{figure*}
+
+%\begin{figure*}[t]
+%  \includegraphics[width=12cm]{ozone_observation_timelag_svanvik}
+%  \caption{Temporal correlation of \chem{[O_3]} data at Svanvik with other ozone monitoring stations in northern Fennoscandia.
 A negative lag means that Svanvik lags behind, while a positive lag mean the other station lags behind.
 The highest correlation with Pallas/Esrange is found at a time lag of $3\,\unit{h}$, for Jergul/Karasjok at $1\,\unit{h}$, and for Janiskoski at $-20\,\unit{h}$.}
-  \label{fig:time_lag_correlation}
-\end{figure*}
+%  \label{fig:time_lag_correlation}
+%\end{figure*}
 
@@
 %\end{figure*}
-
-\begin{figure*}[t]
-  \centering
-  \includegraphics[width=1\textwidth]{supplements/supplement_do3se-parameterization_v2}
-\label{tab:do3se_mm_param}
-\end{figure*}

### 2021-05-10 VegOzone.tex
Spell-checked

@@
 The $\mathrm{POD_y}$ metric is currently used in risk assessments under the United Nations Economic Commission for Europe (UNECE) Long Range Transboundary Air Pollution (LRTAP) Convention to identify those locations across Europe where vegetation is at risk from ozone.
-This convention aims to develope an effects-based emission reduction policy that can target those ozone precursor emissions that are most influential in causing damage.
+This convention aims to develop an effects-based emission reduction policy that can target those ozone precursor emissions that are most influential in causing damage.
 To achieve this the concept of critical levels (CLs) has been developed and applied \citep{Maas2016}.
@@
 We use the $\mathrm{DO_3SE}$ model to estimate the ozone uptake by natural and semi-natural vegetation to investigate the main environmental drivers for projected ozone damage risk in 2018 (Section~\ref{sec:do3se}).
-In this end, we erive bespoke parameterizations for subarctic species to assess systematic uncertainties arising from the use of default generic parameterizations established from less climatically extreme conditions.
-In Section~\ref{sec:conc}, we will summarize our results and discuss how this research can support improvements in future assessments of pollution risk and more generally bio-geochemical cycling.
+In this end, we derive bespoke parameterizations for subarctic species to assess systematic uncertainties arising from the use of default generic parameterizations established from less climatically extreme conditions.
+In Section~\ref{sec:conc}, we will summarize our results and discuss how this research can support improvements in future assessments of pollution risk and more generally bio-biogeochemical cycling.
 
@@
 To this end, we collected atmospheric and ozone monitoring data taken at Svanhovd and additional ozone monitoring data from observation sites in northern Finland and Sweden for cross-evaluation.
-Varring update frequencies and length of available data sets made it necessary to fetch the data from various sources (further details below).
+Varying update frequencies and length of available data sets made it necessary to fetch the data from various sources (further details below).
 Svanhovd synonymously refers to the atmospheric monitoring site Svanvik/Pasvik throughout the paper.\\
@@
 
-All relevant data for the growing season 2018/19 are shown as timeseries in Fig.~\ref{fig:data_svanvik_2018_2019}.
+All relevant data for the growing season 2018/19 are shown as time series in Fig.~\ref{fig:data_svanvik_2018_2019}.
 Ozone concentrations measured at $2\,\unit{m}$ height above ground are averaged hourly.
@@
 In this section, we assess the climate conditions.
-To this end, we compute multi-annual mean (referred to as climatology in the following) of environmental key variables (temperature, precipitation, global irradiance) and gound-level \chem{[O_3]} at Svanhovd.
-For the growing seasons 2018/19, we evaluate the statistical significance of divergances from the norm in these variables (referred to as anomalies).
+To this end, we compute multi-annual mean (referred to as climatology in the following) of environmental key variables (temperature, precipitation, global irradiance) and ground-level \chem{[O_3]} at Svanhovd.
+For the growing seasons 2018/19, we evaluate the statistical significance of divergences from the norm in these variables (referred to as anomalies).
 We perform a bias correction accounting for the increase in tropospheric background \chem{[O_3]} and cross-calibrate the derived climatology for Svanvik with long-term observations of \chem{[O_3]} at Esrange and Pallas.
@@
 \label{subsubsec:clim_ozone}
-In the following, we derive an ozone climatology (\chem{\left<[O_3]\right>}) for Svanvik and for northern Fennoscadia.
+In the following, we derive an ozone climatology (\chem{\left<[O_3]\right>}) for Svanvik and for northern Fennoscandia.
 To this end, we first calculate a pairwise correlation between the different stations.
 The correlation of \chem{[O_3]} between Esrange, Pallas, and Jergul/Karasjok is fairly high ($r^2\approx 0.8$, Appendix Fig.~\ref{fig:density_distribution}a, b).
-Therefore, we select these data to compute the generallized ozone climatology for northern Fennoscandia.
+Therefore, we select these data to compute the generalized ozone climatology for northern Fennoscandia.
 The correlation between Svanvik with any of the active monitoring sites in 2018 (Esrange and Pallas) is rather low, 0.4 and 0.6, respectively.
@@
 The build-up of the ozone spring peak takes about $7.5\,\unit{months}$ from minimum to maximum.
-The most pronounced increase in \chem{\left<[O_3]\right>} coincidents with rising temperatures and global irradiance.
+The most pronounced increase in \chem{\left<[O_3]\right>} coincident with rising temperatures and global irradiance.
 The depletion of ozone from maximum to minimum takes only $4.5\,\unit{months}$.
@@
 
-The \chem{[O_3]} density distribution and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} for Svanvik display a similar pattern as the generallized northern Fennoscandia climatology (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b).
+The \chem{[O_3]} density distribution and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} for Svanvik display a similar pattern as the generalized northern Fennoscandia climatology (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b).
 The spring peak occurs slightly earlier ($100\,\unit{doy}$) and amounts to $40\,\unit{ppb}$ while the annual average \chem{\left<[O_3]\right>} is $6.6\,\unit{ppb}$ lower than the Fennoscandic climatology.
@@
 The derived \chem{\left<[O_3]\right>} is likely biased towards lower \chem{[O_3]} due to an increase in northern hemispheric tropospheric background \chem{[O_3]} since the 1990s.
-To account for this, we derive a bias correction factor by cross-calibrating with the timeseries taken at Esrange and Pallas (Section~\ref{subsec:anomalies}).
+To account for this, we derive a bias correction factor by cross-calibrating with the time series taken at Esrange and Pallas (Section~\ref{subsec:anomalies}).
 
@@
 To assess the significance of 2018/19 anomalies, we estimate the bias due to the change in the tropospheric background \chem{[O_3]} since the 1990s and account for the missing data in June 2018.
-To reconstruct these data, we performed a Reynold's decomposition (Eq.~(\ref{eq:reynolds_decomp})) and cross-calibrate with ozone obserations at Esrange and Pallas (compagnion paper).
+To reconstruct these data, we performed a Reynolds decomposition (Eq.~(\ref{eq:reynolds_decomp})) and cross-calibrate with ozone observations at Esrange and Pallas (companion paper).
 In the following, we will describe the bias-correction.
@@
 Similarly, we find a skew towards higher \chem{[O_3]} for Svanvik in 2018, while 2019 follows a normal distribution rather well (Supplement~Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}).
-In conjunction with the meteorological conditions, we therefore can assume that 2019 had been a normal year and that the annual average anomally in 2019 can be used for the bias correction of the Svanvik climatology (\chem{\left<\Delta[O_3]\right>}~=~$(1.2\pm 5.0)\,\unit{ppb}$).
+In conjunction with the meteorological conditions, we therefore can assume that 2019 had been a normal year and that the annual average anomaly in 2019 can be used for the bias correction of the Svanvik climatology (\chem{\left<\Delta[O_3]\right>}~=~$(1.2\pm 5.0)\,\unit{ppb}$).
 To further strengthen this, we look at trends in ozone observations at Esrange and Pallas and compute the average difference between recent years and the mid 1990s (\chem{\Delta_{present-past}\left<[O_3]\right>}).
@@
 To investigate the difference between 2018 and 2019 concerning stomatal ozone uptake and to assess which of the environmental factors might be most influential in determining \chem{O_3} damage risk, we model ozone uptake of different natural and semi-natural vegetation using the $\mathrm{DO_3SE}$ model and develop bespoke subarctic parameterizations for common plant species.
-We the $\mathrm{DO_3SE}$ model and gapp-filling methods are described in Appendix~\ref{appendix:do3se_model}.
+We the $\mathrm{DO_3SE}$ model and gap-filling methods are described in Appendix~\ref{appendix:do3se_model}.
 First, we give an account of the methodology for deriving bespoke parameters.
@@
 Default parameters for these species or PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as \emph{mapping manual} abbreviated as MM in the following.
-We base our parameterization of coniferous trees on boreal Norway spruce from the MM.
+We base our parameterization of coniferous trees on Boreal Norway spruce from the MM.
 Birch will be referred to as deciduous tree in the following.
@@
 
-For the missing ozone data, we applyed a Reynolds's decomposition of the form:
+For the missing ozone data, we applied a Reynolds decomposition of the form:
 \begin{equation}
@@
 Observed \chem{[O_3]} in 2018 at Pallas served as the reference.
-The exact procedure will be described in an compagnion paper.
+The exact procedure will be described in an companion paper.
 
@@
 MODIS data indicate a higher photosynthetic activity in 2018 than in 2019.
-As shown in Fig.~\ref{fig:modis_Psn} we fitted a second order polynomal function of the general form
+As shown in Fig.~\ref{fig:modis_Psn} we fitted a second order polynomial function of the general form
 %
@@
 \authorcontribution{SF has written the manuscript, collected and processed ozone and environmental data, and performed all statistical analyses.
-AVV contributed to framing this research article, has performed the on-site observation of vegetation damage induced by ozone, and advision in plant physiological processes.
+AVV contributed to framing this research article, has performed the on-site observation of vegetation damage induced by ozone, and advising in plant physiological processes.
 LE has contributed to framing of the this article, $\mathrm{DO_3SE}$ modeling, and development of the PDF-based temperature acclimation.

### 2021-05-10 VegOzone.tex
Refine section 2 and 3

@@
 In Sweden, about 500 fires had been reported (five times more than in a usual summer), and an estimated total of $25000\,\unit{hectare}$ burned down in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) \citep{SOU2019}.
-Coincident peak \chem{[O_3]} is found in ozone monitoring data from Svanhovd but also over whole northern Fennoscandia in July (see Fig.~\ref{fig:data_svanvik_2018_2019} or Fig.~\ref{fig:ozone_reconstruction_2018_07}).
+Coincident peak \chem{[O_3]} is found in ozone monitoring data from Svanhovd (Fig.~\ref{fig:data_svanvik_2018_2019} but also over whole northern Fennoscandia in July.
 During the 2003 drought period, elevated ozone in Europe was promoted by a combination of various factors such as wildfires, reduced cloud cover (increased solar radiation), reduced dry deposition and turbulent mixing due to the stagnant weather conditions, and increased BVOC emissions \citep{JGR:Solberg2018}.
@@
 We use the $\mathrm{DO_3SE}$ model to estimate the ozone uptake by natural and semi-natural vegetation to investigate the main environmental drivers for projected ozone damage risk in 2018 (Section~\ref{sec:do3se}).
-In this end, we derive bespoke parameterizations for subarctic species to assess systematic uncertainties arising from the use of default generic parameterizations established from less climatically extreme conditions.
-In Section~\ref{sec:conc}, we will summarize our results and discuss how this research can support improvements in future assessments of pollution risk and more generally bio-biogeochemical cycling.
+To this end, we derive bespoke parameterizations for subarctic species to assess systematic uncertainties arising from the use of generic parameterizations established from less climatically extreme conditions.
+In Section~\ref{sec:conc}, we will summarize our results and discuss how this research can support improvements in future assessments of pollution risk and more generally biogeochemical cycling.
 
@@
 A high global irradiance in most cases is the result of a low cloud fraction.
-In both years, the maximum recorded $Q_0$ was roughly $750\,\unit{W\,m^{-2}}$ reached in June.
+In both years, the maximum recorded $Q_0$ was $750\,\unit{W\,m^{-2}}$ and reached in June.
 
@@
 In the absence of local precursors, the stratification of the atmosphere usually leads to a higher abundances in ozone at higher altitude \citep[e.g.,][]{AB:Klingberg2009}.
-The periphery vegetation at all sites is similar and consists mainly of pine forests, birch, and heath.
+The periphery vegetation at all sites is similar and consists mainly of pine forests, birch, perennial grassland, and heath.
 However, on closer inspection the vegetation differs substantially.
-Ozone measurements at Pallas are taken on top of mount Sammaltunturi $100\,\unit{m}$ above treeline (pine forest) and the surrounding vegetation mainly consists of low vascular plants, mosses, and lichen \citep{BER:Hatakka2003}.
+Ozone measurements at Pallas are taken on top of mount Sammaltunturi, $100\,\unit{m}$ above treeline (pine forest), and the surrounding vegetation mainly consists of low vascular plants, mosses, and lichen \citep{BER:Hatakka2003}.
 At Svanhovd, the ozone measurements are taken amidst a meadow surrounded by agriculturally managed herbaceous plants and grass adhering to the \citet{WMOGuide2018} standards (Fig.\ref{fig:svanhovd_research_station}).
@@
 Especially, ozone monitors did not undergo regular re-calibration.
-This probably leads to drifts in the observed data and may impose a false periodicity.
+This probably leads to drifts in the observed data and may impose a false trend.
 In 1997, the ozone monitoring at Jergul was relocated downstream of the river Karasjokka to a site close by Karasjok.
@@
 The location of Svanvik at $69.45\,\unit{^\circ N}$ suggests a subarctic climate.
-The derived climate diagram based on climatological data from the atmospheric monitoring in Svanvik (1992--2012) (Fig.~\ref{fig:climatediagram}) supports this very well.
+The derived climate diagram based on climatological data (1992--2012) (Fig.~\ref{fig:climatediagram}) supports this very well.
 Monthly averaged $2\,\unit{m}$ temperatures (red line) are displayed with a standard error of the mean (SE) as error band.
@@
 Months with average temperatures below freezing are denoted with a star.
-As can be seen, temperatures stay below freezing for 5 consecutive months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for K\"{o}ppen's climate classification of a regular subarctic climate (Dfc) \citep[][e.g.]{SD:Beck2018}.
+As can be seen, temperatures stay below freezing for 5 consecutive months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for K\"{o}ppen's climate classification of a regular subarctic climate (Dfc) \citep[e.g.][]{SD:Beck2018}.
 The highest monthly average temperature is $(13.1\pm 1.1)\,\unit{^\circ C}$ in July and the lowest $(-12.8\pm 2.0)\,\unit{^\circ C}$ in February.
@@
 The derived climatology for global irradiance ($Q_0$) at Svanvik (Appendix Fig.~\ref{fig:global_rad_clim}a)) approaches zero in December/January (polar night) and reaches its maximum in June/July (midnight sun conditions).
-The highest observed averaged daily maximum ($\left<Q_0^\mathrm{max}\right>$) amounts to $800\,\unit{W\,m^{-2}}$, while the averaged daily mean ($\left<Q_0^\mathrm{mean}\right>$) is $450\,\unit{W\,m^{-2}}$ (regularly reached in May--July).
+The highest observed averaged daily maximum ($\left<Q_0^\mathrm{max}\right>$) amounts to $800\,\unit{W\,m^{-2}}$, while the maximum of the averaged daily mean ($\left<Q_0^\mathrm{mean}\right>$) is $450\,\unit{W\,m^{-2}}$ (regularly reached in May--July).
 The maximum of the averaged daily minimum ($\left<Q_0^\mathrm{min}\right>$) lies at about $200\,\unit{W\,m^{-2}}$.
@@
 To this end, we first calculate a pairwise correlation between the different stations.
-The correlation of \chem{[O_3]} between Esrange, Pallas, and Jergul/Karasjok is fairly high ($r^2\approx 0.8$, Appendix Fig.~\ref{fig:density_distribution}a, b).
+The correlation of \chem{[O_3]} between Esrange, Pallas, and Jergul/Karasjok is fairly high ($r^2\approx 0.8$, Appendix Fig.~\ref{fig:density_distribution}).
 Therefore, we select these data to compute the generalized ozone climatology for northern Fennoscandia.
@@
 
-In Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}, two dimensional density distributions of \chem{[O_3]} for the combined northern Fennoscandia data (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}a) and Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) are shown.
+In Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}, two dimensional density distributions of \chem{[O_3]} for the combined data (Esrange, Jergul/Karasjok, Pallas) (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}a) and Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) are shown.
 Darker colors indicate higher probability to observe these values.
@@
 
-The generalized northern Fennoscandic ozone climatology peaks at $48\,\unit{ppb}$  in late April (day of the year (doy) $100-120$) and declines throughout the growing season until a minimum is reached in late summer/fall ($24-30\,\unit{ppb}$).
+The generalized northern Fennoscandic ozone climatology peaks at $48\,\unit{ppb}$  in late April (day of the year (doy) $100-120$) and declines throughout the growing season until a minimum is reached in the beginning of September ($240-250\,\unit{doy}$) and amounts to $24-30\,\unit{ppb}$.
 The turning points lie approximately in June and November.
 The build-up of the ozone spring peak takes about $7.5\,\unit{months}$ from minimum to maximum.
-The most pronounced increase in \chem{\left<[O_3]\right>} coincident with rising temperatures and global irradiance.
+The most pronounced increase in \chem{\left<[O_3]\right>} coincides with rising temperatures and global irradiance.
 The depletion of ozone from maximum to minimum takes only $4.5\,\unit{months}$.
-The minimum occurs at $25\,\unit{ppb}$ in the beginning of September ($240-250\,\unit{doy}$).
 
@@
 This is in line with a general decrease in tropospheric ozone with decreasing altitude.
-The earlier decline in \chem{\left<[O_3]\right>} may point to an influence of coniferous vegetation on the removal of ozone since \chem{CO_2} uptake by coniferous trees has been observed as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013} compared to an estimated average begin of growing season ($\left<G_\mathrm{start}\right>$) on $(150\pm 14)\,\unit{doy}$ at Svanvik (see Section~\ref{subsec:do3se_parameters} for details).
+The earlier decline in \chem{\left<[O_3]\right>} may point to an influence of coniferous vegetation on the removal of ozone since \chem{CO_2} uptake by coniferous trees has been observed as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013}.
+The estimated average begin of growing season for deciduous trees ($\left<G_\mathrm{start}\right>_\mathrm{decid}$) is $(150\pm 14)\,\unit{doy}$ at Svanvik (see Section~\ref{subsec:do3se_parameters} for details).
 In July--September, ozone is occasionally almost completely depleted during nighttime.
-This hint to ozone uptake by vegetation even at low light intensities during midnight sun conditions in combination with stable planetary boundary layer conditions preventing mixing of ozone rich air.
+This hints to ozone uptake by vegetation even at low light intensities during midnight sun conditions in combination with stable planetary boundary layer conditions preventing mixing of ozone rich air.
 The derived \chem{\left<[O_3]\right>} is likely biased towards lower \chem{[O_3]} due to an increase in northern hemispheric tropospheric background \chem{[O_3]} since the 1990s.
@@
 \label{subsec:anomalies}
-Using these established climatologies, we look at the anomalies in 2018/19 and discuss the differences between the two years.
-As weather extremes like in 2018 are likely to become more frequent with climate change, this will help to improve our understanding of a possible future increase in vulnerability of subarctic vegetation to ozone.
+In the following, we look at the 2018/19 anomalies and discuss the differences between the two years.
+We also conduct a bias correction for the Svanvik \chem{\left<[O_3]\right>.} As weather extremes like in 2018 are likely to become more frequent with climate change, this will help to assess possible future changes in the vulnerability of subarctic vegetation to ozone.
 
@@
 \label{subsubsec:anomal_tpq}
-In Fig.~\ref{fig:anomalies_svanvik}a), temperature anomalies at Svanvik are shown as a percentage of days warmer or colder than $\pm 1\,\sigma$ from the climatological mean for each month.
+In Fig.~\ref{fig:anomalies_svanvik}a), temperature anomalies for Svanvik are shown as a percentage of days warmer or colder than $\pm 1\,\sigma$ from the climatological mean for each month.
 Negative deviations from climatology are displayed as a negative percentage.
-The full annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
+The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
 The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
-The summer of 2018 was significantly warmer than average.
+We find, that the summer of 2018 was significantly warmer than average.
 Especially in May and July more than $40\,\unit{\%}$ of the days were significantly warmer.
@@
 March 2018 had many unusually cold days.
-The summer of 2019 was fairly average, however, significantly colder days in July occurred, while August had both significantly colder and warmer days.
+The summer of 2019 was fairly average, however, significantly colder days occurred in July, while August had both significantly colder and warmer days.
 April 2019 had significantly warmer than average days.
@@
 
-Similarly, the precipitation anomalies at Svanvik are displayed in Fig.~\ref{fig:anomalies_svanvik}b).
-Due to large interannual variability in observed precipitation, only a few days deviated significantly from the climatology on the $1 \sigma$ level.
-Therefore, we lowered the $\sigma$ level constraint.
-The percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ from the climatological mean are shown for each month.
-Negative deviations from the climatology are shown as a negative percentage and annual positive/negative deviations are indicated in the respective corners.
-The hatched area between the dashed lines indicates the expected percentage of values falling above $\pm\frac{1}{2}\,\sigma$ if a normal distribution is assumed ($30.8\,\unit{\%}$).
+Similarly, the precipitation anomalies for Svanvik are displayed in Fig.~\ref{fig:anomalies_svanvik}b).
+Due to the large interannual variability in observed precipitation, only a few days deviated significantly from the climatology on the $1 \sigma$ level.
+Therefore, we lowered the $\sigma$ level constraint to $\pm \frac{1}{2}\,\sigma$.
+The percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ are shown for each month.
+Positive deviation refers to days wetter than average and negative deviation to drier than average.
+The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm\frac{1}{2}\,\sigma$ if a normal distribution is assumed ($30.8\,\unit{\%}$).
 Unlike the temperature anomalies, the picture is not as clear.
 While March 2018 had a significant number of days wetter than unusual, summer and fall (May/July, September/October) had a significant number of days drier than average.
-Overall, 2019 was rather average throughout the year, but with a significant number of days drier than normal (September/October) and a wetter December.
-
-Similarly, the global irradiance anomalies at Svanvik are presented in Fig.~\ref{fig:anomalies_svanvik}c).
+2019 was rather average throughout the year, but with a significant number of days drier than normal in September/October and wetter in December.
+
+The global irradiance anomalies for Svanvik are presented in Fig.~\ref{fig:anomalies_svanvik}c).
 The percentage of days above or below $\pm 1\,\sigma$ from the climatological mean is displayed for each month.
@@
 To assess the significance of 2018/19 anomalies, we estimate the bias due to the change in the tropospheric background \chem{[O_3]} since the 1990s and account for the missing data in June 2018.
-To reconstruct these data, we performed a Reynolds decomposition (Eq.~(\ref{eq:reynolds_decomp})) and cross-calibrate with ozone observations at Esrange and Pallas (companion paper).
-In the following, we will describe the bias-correction.
-
-In 2018, the probability density function of ozone concentration anomalies (\chem{\Delta[O_3]}) for Esrange and Pallas deviate clearly, but not significantly from the climatological mean (Supplement~Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}).
+To reconstruct these missing data, we performed a Reynolds decomposition (Eq.~(\ref{eq:reynolds_decomp})) and cross-calibrate with ozone observations at Esrange and Pallas (companion paper).
+In the following, we will describe the bias-correction and discuss the significance of \chem{[O_3]} enhancements in 2018.
+
+In 2018, the probability density functions (PDFs) of ozone concentration anomalies (\chem{\Delta[O_3]}) for Esrange and Pallas deviate clearly, but not significantly from the climatological mean (Supplement~Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}).
 On average, \chem{[O_3]} had been elevated by $(1.9\pm 5.7)\,\unit{ppb}$ and $(1.9\pm 5.4)\,\unit{ppb}$, respectively.
@@
 
-Similarly, we find a skew towards higher \chem{[O_3]} for Svanvik in 2018, while 2019 follows a normal distribution rather well (Supplement~Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}).
-In conjunction with the meteorological conditions, we therefore can assume that 2019 had been a normal year and that the annual average anomaly in 2019 can be used for the bias correction of the Svanvik climatology (\chem{\left<\Delta[O_3]\right>}~=~$(1.2\pm 5.0)\,\unit{ppb}$).
-To further strengthen this, we look at trends in ozone observations at Esrange and Pallas and compute the average difference between recent years and the mid 1990s (\chem{\Delta_{present-past}\left<[O_3]\right>}).
-Present is defined as the average over 2010--2018 and past as the time between the respective beginning of data taking until 1997.
+We find a similar skew towards higher \chem{[O_3]} for Svanvik in 2018, while 2019 follows a normal distribution rather well (Supplement~Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}).
+In conjunction with the meteorological conditions, we therefore assume that 2019 had been a normal year and the annual average anomaly in 2019 (\chem{\left<\Delta[O_3]\right>}~=~$(1.2\pm 5.0)\,\unit{ppb}$) can be used for the bias correction.
+We also consult the trends in ozone observations at Esrange and Pallas to strengthen this assumption and compute the average difference between recent years and the mid 1990s (\chem{\Delta_{present-past}\left<[O_3]\right>}).
+Herein, present is defined as the average over 2010--2018 and past as the time between the respective beginning of data taking until 1997.
 We do not find a clear trend.
 For Esrange \chem{\Delta_{present-past}\left<[O_3]\right>} amounts to $(1.5\pm 3.1)\,\unit{ppb}$ and for Pallas to $(-2.5\pm 3.7)\,\unit{ppb}$ \citep[][for comparison]{NILU2003, BER:Ruoho-Airola2015}.
-The observed trend at Esrange is compatible with our assumed bias correction for the Svanvik ozone climatology.
+Only the observed trend at Esrange is compatible with our assumed bias correction for the Svanvik ozone climatology.
 
@@
 
-In accordance to the above findings, we adjust the derived ozone climatology for Svanvik by adding a constant bias correction factor $\Delta\chem{[O_3]_\mathrm{bias-corr}} = 1.2\,\unit{ppb}$ and assess the monthly significance of the ozone concentration anomalies in 2018/19 (Fig.~\ref{fig:anomalies_svanvik}d)).
-The resulting percentage of days above/below $\pm 1\,\sigma$ from the bias-corrected climatological mean is displayed for each month in Fig~\ref{fig:data_svanvik_2018_2019}.
+In accordance to the above findings, we adjust the derived ozone climatology for Svanvik by adding a constant bias correction factor $\Delta\chem{[O_3]_\mathrm{bias-corr}} = 1.2\,\unit{ppb}$ and assess the monthly significance of the ozone concentration anomalies in 2018/19.
+The resulting percentage of days above/below $\pm 1\,\sigma$ from the bias-corrected climatological mean is displayed for each month in Fig.~\ref{fig:anomalies_svanvik}d).
 Negative deviations from climatology are shown as a negative percentage.
 Note that missing bars for certain months are both due to data availability and anomalies not meeting the $1\sigma$ criteria.
-The the star indicates resulting anomalies after the reconstruction of missing data in July.
+The star indicates results after the reconstruction of missing data in July.
 In April/May 2018, ozone was significantly enhanced, while no significant enhancement is found in 2019.
 In both years, ozone concentrations were significantly lower than average in November.
-These results mean, that the extensive forest fires in Sweden did not lead to a significant enhancement in \chem{[O_3]} at Svanhovd in 2018.
-However, there are indications that the sites further south at Esrange and Pallas experienced more days with elevated ozone in 2018 than Svanvik.
+These results indicate that the extensive forest fires in Sweden did not contribute to a significant enhancement in \chem{[O_3]} at Svanhovd in 2018, but the stagnant meteorological conditions in April/May in combination with a late start of the growing season.
+However, there are indications that the sites further south at Esrange and Pallas experienced more days of elevated ozone in 2018 than Svanvik.
 
@@
 \label{sec:do3se}
-To investigate the difference between 2018 and 2019 concerning stomatal ozone uptake and to assess which of the environmental factors might be most influential in determining \chem{O_3} damage risk, we model ozone uptake of different natural and semi-natural vegetation using the $\mathrm{DO_3SE}$ model and develop bespoke subarctic parameterizations for common plant species.
+In the following, we investigate the difference between 2018 and 2019 concerning stomatal ozone uptake and assess which of the environmental factors might be most influential in determining \chem{O_3} damage risk.
+We model ozone uptake of different natural and semi-natural vegetation using the $\mathrm{DO_3SE}$ model and develop bespoke subarctic parameterizations for common plant species.
 We the $\mathrm{DO_3SE}$ model and gap-filling methods are described in Appendix~\ref{appendix:do3se_model}.

### 2021-05-10 VegOzone.tex
Change order in appenix in accordance to occurance in text

@@
 We model ozone uptake of different natural and semi-natural vegetation using the $\mathrm{DO_3SE}$ model and develop bespoke subarctic parameterizations for common plant species.
-We the $\mathrm{DO_3SE}$ model and gap-filling methods are described in Appendix~\ref{appendix:do3se_model}.
-First, we give an account of the methodology for deriving bespoke parameters.
-Then, we will present and discuss the modeling results.
+First, we give an account of the methodology for deriving bespoke parameters, after which we present and discuss the modeling results.
+The $\mathrm{DO_3SE}$ model and conducted gap-filling methods are described in Appendix~\ref{appendix:do3se_model}.
 
@@
 
-The main species or PFTs, we deduced on-site at the NIBIO Environment Centre Svanhovd and from Fig.~\ref{fig:svanhovd_research_station}, are perennial grassland, birch, and pine.
+We deduced the dominant species on-site at the NIBIO Environment Centre Svanhovd and from Fig.~\ref{fig:svanhovd_research_station} and found perennial grassland, birch (generalized as deciduous trees), and Scots pine (generalized as coniferous trees).
 Default parameters for these species or PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as \emph{mapping manual} abbreviated as MM in the following.
-We base our parameterization of coniferous trees on Boreal Norway spruce from the MM.
-Birch will be referred to as deciduous tree in the following.
+We base our parameterization of coniferous trees on the MM's Boreal Norway spruce and deciduous trees are derived from MM's silver birch.
 A comprehensive list of model parameters can be found in Supplement~Table~S1.
@@
 
-\subsection{Bespoke parameterizations}
-
-To assess the $G_\mathrm{start}$  and $G_\mathrm{end}$ of the growing season for coniferous trees at Svanhovd in 2018/19, we used the net photosynthesis product ($A_\mathrm{net}$) of MODIS AQUA/TERRA over a $1\times 1\,\unit{km}$ area centered at Svanhovd.
-MODIS data indicate a higher photosynthetic activity in 2018 than in 2019.
-As shown in Fig.~\ref{fig:modis_Psn} we fitted a second order polynomial function of the general form
-%
-\begin{equation}
-A_\mathrm{net}(t) =  -m_0\cdot(t-m_1)^2+m_2,
-\end{equation}
-%
-with the form parameters $m_i$ through the data. Numerically, we retrieved the roots as $G_\mathrm{start}$ 122/106 and $G_\mathrm{end}$ 261/274 for 2018 and 2019, respectively.
-
-\begin{figure}[th]
-  \includegraphics[width=8.3cm]{figB1}
-  \caption{Estimated $G_\mathrm{start}$ and $G_\mathrm{end}$ of growing season for coniferous trees from MODIS Aqua/Terra net photosynthesis (PSN) product.
-A $1\times 1\,\unit{km}$ area around Svanhovd was selected.
-Daily averaged data for both 2018 and 2019 has been fitted with a quadratic polynomial function.
-The numerically computed root yields: BGS 122/106 day of year (\unit{doy}) and EGS 261/274 day of year for 2018/2019, respectively.}
-  \label{fig:modis_Psn}
-\end{figure}
-
-As described in Section~\ref{subsec:do3se_parameters}, we developed bespoke parameterizations for natural and semi-natural vegetation at Svanhovd.
-Here, we show the temperature response and light response functions for coniferous and deciduous trees (Figs.~\ref{fig:f_temp_spruce}-\ref{fig:f_temp_birch}).
-
-\begin{figure}[t]
-  %\includegraphics[width=8.3cm]{javis_funcs_temp_hist_evergreen}\\
-  %\includegraphics[width=8.3cm]{javis_funcs_rad_hist_evergreen}
-  \includegraphics[width=8.3cm]{figB2}
-\caption{Construction of bespoke response functions for coniferous trees.
-(a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively.
-Original mapping manual parameterization is shown in comparison as solid line.
-Note that $Q_0$ has been truncated to $0.006$.
-PPFD0.8 and PPFD1.2 refer to $\alpha$ values increasing/decreasing PPFD at $f_\mathrm{light}=0.5$ by $\pm 20\,\%$, respectively.}
-\label{fig:f_temp_spruce}
-\end{figure}
-
-\begin{figure}[t]
-  %\includegraphics[width=8.3cm]{javis_funcs_temp_hist_birch}\\
-  %\includegraphics[width=8.3cm]{javis_funcs_rad_hist_birch}
-  \includegraphics[width=8.3cm]{figB3}
-\caption{Construction of bespoke response functions for deciduous trees.
-(a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively.
-Original mapping manual parameterization is shown in comparison as solid line.
-Note that $Q_0$ has been truncated to $0.006$.
-PPFD0.8 and PPFD1.2 refer to $\alpha$ values increasing/decreasing PPFD at $f_\mathrm{light}=0.5$ by $\pm 20\,\%$, respectively.}
-\label{fig:f_temp_birch}
-\end{figure}
-
 \subsection{Initial $\mathrm{DO_3SE}$ modeling results with mapping manual parameters}
@@
 \begin{figure*}[t]
-  \includegraphics[width=12cm]{figB4}
+  \includegraphics[width=12cm]{figB1}
   \caption{$\mathrm{DO_3SE}$ modeling results for mapping manual default parameterization.
@@
 \end{figure*}
+
+\subsection{Bespoke parameterizations}
+
+To assess the $G_\mathrm{start}$  and $G_\mathrm{end}$ of the growing season for coniferous trees at Svanhovd in 2018/19, we used the net photosynthesis product ($A_\mathrm{net}$) of MODIS AQUA/TERRA over a $1\times 1\,\unit{km}$ area centered at Svanhovd.
+MODIS data indicate a higher photosynthetic activity in 2018 than in 2019.
+As shown in Fig.~\ref{fig:modis_Psn} we fitted a second order polynomial function of the general form
+%
+\begin{equation}
+A_\mathrm{net}(t) =  -m_0\cdot(t-m_1)^2+m_2,
+\end{equation}
+%
+with the form parameters $m_i$ through the data. Numerically, we retrieved the roots as $G_\mathrm{start}$ 122/106 and $G_\mathrm{end}$ 261/274 for 2018 and 2019, respectively.
+
+\begin{figure}[th]
+  \includegraphics[width=8.3cm]{figB2}
+  \caption{Estimated $G_\mathrm{start}$ and $G_\mathrm{end}$ of growing season for coniferous trees from MODIS Aqua/Terra net photosynthesis (PSN) product.
+A $1\times 1\,\unit{km}$ area around Svanhovd was selected.
+Daily averaged data for both 2018 and 2019 has been fitted with a quadratic polynomial function.
+The numerically computed root yields: BGS 122/106 day of year (\unit{doy}) and EGS 261/274 day of year for 2018/2019, respectively.}
+  \label{fig:modis_Psn}
+\end{figure}
+
+As described in Section~\ref{subsec:do3se_parameters}, we developed bespoke parameterizations for natural and semi-natural vegetation at Svanhovd.
+Here, we show the temperature response and light response functions for coniferous and deciduous trees (Figs.~\ref{fig:f_temp_spruce}-\ref{fig:f_temp_birch}).
+
+\begin{figure}[t]
+  %\includegraphics[width=8.3cm]{javis_funcs_temp_hist_evergreen}\\
+  %\includegraphics[width=8.3cm]{javis_funcs_rad_hist_evergreen}
+  \includegraphics[width=8.3cm]{figB3}
+\caption{Construction of bespoke response functions for coniferous trees.
+(a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively.
+Original mapping manual parameterization is shown in comparison as solid line.
+Note that $Q_0$ has been truncated to $0.006$.
+PPFD0.8 and PPFD1.2 refer to $\alpha$ values increasing/decreasing PPFD at $f_\mathrm{light}=0.5$ by $\pm 20\,\%$, respectively.}
+\label{fig:f_temp_spruce}
+\end{figure}
+
+\begin{figure}[t]
+  %\includegraphics[width=8.3cm]{javis_funcs_temp_hist_birch}\\
+  %\includegraphics[width=8.3cm]{javis_funcs_rad_hist_birch}
+  \includegraphics[width=8.3cm]{figB4}
+\caption{Construction of bespoke response functions for deciduous trees.
+(a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively.
+Original mapping manual parameterization is shown in comparison as solid line.
+Note that $Q_0$ has been truncated to $0.006$.
+PPFD0.8 and PPFD1.2 refer to $\alpha$ values increasing/decreasing PPFD at $f_\mathrm{light}=0.5$ by $\pm 20\,\%$, respectively.}
+\label{fig:f_temp_birch}
+\end{figure}
 

### 2021-05-10 VegOzone.tex
Refine Section 4.1

@@
 We base our parameterization of coniferous trees on the MM's Boreal Norway spruce and deciduous trees are derived from MM's silver birch.
-A comprehensive list of model parameters can be found in Supplement~Table~S1.
-Initial test simulations with the MM PFTs showed unrealistically low $g_\mathrm{sto}$ in 2019 particularly for perennial grassland (Appendix~Fig.~\ref{fig:pody_mm_composit}).
-Only during an extended warm period in late July did we find substantial stomatal conductance.
-In ecological terms this would mean that there was almost no growth of grass in the summer of 2019 -- a prediction which is easily falsified by reality.
-We identified $f_\mathrm{T}$ as abnormally low, being the limiting factor of stomatal conductance.
-Therefore, we propose bespoke parameterizations which may describe an actual acclimation of these species to the climatic conditions in the target region.
-
+For a comprehensive list of model parameters consult Supplement~Table~S1.
+
+Initial test simulations with the MM PFTs revealed an unrealistically low stomatal conductance at leaf-level ($G_\mathrm{sto}^\mathrm{leaf}$) in 2019 particularly for perennial grassland (Appendix~Fig.~\ref{fig:pody_mm_composit}).
+Substantial stomatal conductance occured only during an extended warm period in late July.
+In ecological terms, this means that there was almost no growth of grass in the summer of 2019 -- a prediction which is easily falsified by reality.
+We identified $f_\mathrm{T}$ as abnormally low, being the limiting factor of stomatal conductance in perennial grassland.
+Therefore, we saw the necessarity to propose bespoke parameterizations describing an acclimation of these species to the climatic conditions in the target region.
 Note, however, that these parameterizations represent hypothetical PFTs or species that have yet to be verified by experiments.
 
-We hypothesize that perennial vegetation will likely prioritize their growth or fast adolescence and hence their carbon sequestration, during a short, subarctic growing season.
-Because the mapping manual version of the $\mathrm{DO_3SE}$ model currently does not simulate $A_\mathrm{n}$, we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$.
+We hypothesize that perennial vegetation will likely prioritize their growth or fast adolescence and hence their carbon sequestration, during the short, subarctic growing season.
+Because the mapping manual version of the $\mathrm{DO_3SE}$ model currently does not simulate net photosynthesis ($A_\mathrm{n}$), we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune the temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (see Appendix~\ref{appendix:do3se_model} for details).
 We further assume that $f_\mathrm{VPD}$ and $f_\mathrm{SWP}$ suffice our biomes.
@@
 We show two climatological reference periods 1990s and 2000s to indicate that the vegetation is subject to ongoing climate change and hence continuous acclimation.
-Resulting response functions for coniferous and deciduous trees are displayed in Appendix~Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}.
+Resulting response functions for coniferous and deciduous trees are displayed in Appendix~Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}, respectively.
 
 We define two different hypothetical temperature acclimation: \emph{subarctic} and \emph{cold}.
-The most extreme of the two acclimations is \emph{subarctic}.
 For either, we assume that the hypothetical species are more tolerant to cold conditions and shift $f_\mathrm{T}$ towards colder temperatures accordingly.
@@
 In the same way, \emph{subarctic} refers to a hypothetical species which is very tolerant to cold but is sensitive to high temperatures, and most efficient at cool temperatures ($T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000).
-For perennial grassland the resulting response functions are depicted together with the PDF of temperature in Fig.~\ref{fig:f_temp_grassland}a).
-
-Norway spruce was experimentally found to be active already at rather low air temperatures and could reach $60\,\unit{\%}$ photosynthetic activity as early as \unit{doy}~100 \citep{TB:Kolari2007, TP:Wallin2013}.
-Based on the time series of observed \chem{CO_2} uptake and temperature at two observation sites in Southern ($\rightarrow$ \emph{cold}) and Northern ($\rightarrow$ \emph{subarctic}) Finland \citep{TB:Kolari2007}, we estimate $10\,\unit{^\circ C}\le T_\mathrm{opt} \le 15\,\unit{^\circ C}$.
-We assume a similar acclimation for deciduous trees.
-All bespoke temperature parameters are listed in Table~\ref{tab:sensitivity_tests_temp}.\\
-
-For an acclimation with respect to light conditions, we hypothesize that the light threshold for a $50\,\unit{\%}$ opening of plants' stomata may differ in subarctic species.
-We analytically derive the inverse function $f_{\mathrm{light},k}^{-1}$ for each PFT or species $k$
+The most extreme of the two acclimation types is \emph{subarctic}.
+For perennial grassland the resulting response functions are depicted together with the temperature PDF in Fig.~\ref{fig:f_temp_grassland}a).
+
+Norway spruce was experimentally found to be active already at rather low air temperatures and can reach $60\,\unit{\%}$ photosynthetic activity as early as \unit{doy}~100 \citep{TB:Kolari2007, TP:Wallin2013}.
+Based on the time series of \chem{CO_2} uptake and temperature at observation sites in southern ($\rightarrow$ \emph{cold}) and northern ($\rightarrow$ \emph{subarctic}) Finland \citep{TB:Kolari2007}, we estimate the optimal temperature interval $10\,\unit{^\circ C}\le T_\mathrm{opt} \le 15\,\unit{^\circ C}$.
+We assume a similar temperature acclimation for deciduous trees.
+All bespoke temperature parameters are listed in Table~\ref{tab:sensitivity_tests_temp}.
+
+With respect to light conditions, we hypothesize that the light threshold for a $50\,\unit{\%}$ opening of plants' stomata may differ in subarctic species compared to species in less extreme climates.
+Analytically, we derive the inverse function $f_{\mathrm{light},k}^{-1}$ of Eq.~(\ref{eq:flight}) for each PFT or species $k$
 \begin{equation}
@@
 
-As indicated above, a better acclimation will result in a higher average $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$.
+As indicated above, a better acclimation to prevailing climate conditions will result in a higher average $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$.
 To quantify the resulting bespoke parameterizations, we propose the following metric.
 For a generic GS (May--August), we compute climatological average and standard deviation of $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal})) at noon (averaged over $11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning (averaged over $5-9\,\unit{am}$ local time).
-We presume that relative stomatal conductance of around noon (highest light intensity) and in the early morning (lower intensities) are good proxies for \chem{CO_2} uptake efficiency.
+We presume that relative stomatal conductance around noon (highest light intensity) and in the early morning (lower intensities) are good proxies for \chem{CO_2} uptake efficiency.
 Our arbitrary optimization target is $60\,\unit{\%}$.
 Additional constrains are a small deviation between noon and morning and a low standard deviation indicating a higher robustness to interannual variability of growing conditions.
-For simplicity, we neglect the dependency on SWP in this regard.
+For simplicity, we neglect the dependency on SWP in this metric.
 From the results shown in Fig.~\ref{fig:javis_func_opt_mean}, we identify \emph{subarctic}, PPFD0.8 as the best acclimation for all hypothetical species or PFTs alike.
-As expected due to the small adjustments, coniferous trees display the lowest deviation between the different types of parameterizations, while the divergence for perennial grassland is substantial as response to the stronger proposed temperature acclimation.
+As expected due to the small adjustments, coniferous trees display the lowest deviation between the different types of acclimation, while the divergence for perennial grassland is substantial as response to the stronger proposed temperature acclimation.
 
@@
 Start ($A_\text{start}$) and end ($A_\text{end}$) of GS for each PFT are given in units of \unit{doy} (Table~\ref{tab:sensitivity_tests_gs}) and estimated as follows.
-Evergreen needle leaf trees show photosynthetic activity already at low temperatures \citep{TP:Wallin2013, TB:Kolari2007}.
+Coniferous trees show photosynthetic activity already at low temperatures \citep{TP:Wallin2013, TB:Kolari2007}.
 We use a MODIS (Aqua/Terra) retrieved net photosynthesis product for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine $A_\text{start}$ and $A_\text{end}$.
-Net photosynthesis data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}).
+The net photosynthesis data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}).
 By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{doy}$ for 2018 and \unit{doy} 106 in 2019.
@@
 We find 129 and 130, respectively.
-For both years, these dates coincident with the first snow-free day at the meteorological weather station at Øvre Neiden (Sør-Varanger, NOR) which is relatively close-by.
+For both years, these dates coincide with the first snow-free day at the meteorological weather station at Øvre~Neiden (Sør-Varanger, NOR) which is relatively close-by.
 For comparison, we would get $A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$ from the latitude model typically used for localizing $\mathrm{DO_3SE}$.
-By comparison, the latitude model indicates about a $2\,\unit{month}$ longer growing season.
-Due to a lack of quantitative field observation, we assume a latency period of about $1\,\unit{month}$ after snow-melt for perennial grassland.
+The latitude model indicates a $2\,\unit{month}$ longer growing season.
+Due to a lack of quantitative field observation, we assume a latency period of $1\,\unit{month}$ after snow-melt for perennial grassland.
 
@@
 In addition, \citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area.
-Both these factors will affect leaf boundary layer resistance, the smaller leaves would tend to reduce leaf boundary layer resistances whilst the shorter trees would reduce canopy roughness and increase canopy boundary layer resistances.
+Both factors will affect leaf boundary layer resistance, the smaller leaves would tend to reduce leaf boundary layer resistances whilst the shorter trees would reduce canopy roughness and increase canopy boundary layer resistances and hence the uptake of \chem{O_3}.
 
@@
 
-We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using both MM parameterizations and a variety of bespoke response function in $f_\mathrm{T}$ and $f_\mathrm{light}$, and localized GS (described in detail in Section~\ref{subsec:do3se_parameters}).
+We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for the three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using the generic MM parameterization and our bespoke temperaure (\emph{cold}, \emph{subarctic}) and light (PPFD0.8, PPFD1.2) acclimation types, and localized GS.
 The results are shown in Fig.~\ref{fig:pody_rel}.
 For comparison, critical levels for a $4\,\unit{\%}$ (deciduous), $2\,\unit{\%}$ (coniferous) \citep[refer to silver birch and Norway spruce in][]{ICP:MappingManual2017}, and $10\,\unit{\%}$ (perennial grassland, above ground) \citep{ESPR:Hayes2021} reduction in biomass are shown as horizontal dashed lines.
-Different years are color coded as previously.
+2018 is color-coded in pink and 2019 in purple.
 Simulations with MM GS, canopy, and leaf parameters are indicated with squared markers, bespoke GS with circles.
-The annotation $\pm$ refers to bespoke light response function threshold ($+\rightarrow\mathrm{PPFD1.2}$, $-\rightarrow\mathrm{PPFD0.8}$), while bespoke temperature response functions are indicated by categories on the x-axis.
+The annotation $\pm$ refers to the bespoke light response function ($+\rightarrow\mathrm{PPFD1.2}$, $-\rightarrow\mathrm{PPFD0.8}$), while bespoke temperature response functions are indicated by categories on the x-axis.
 Closed/open symbols represent SWP sensitivity switched off/on in the model.

### 2021-05-11 VegOzone.tex
Include reference to icp veg gap filling method

@@
 For further analysis (Section~\ref{sec:stats}) and modeling study with the $\mathrm{DO_3SE}$ model (Section~\ref{sec:do3se}), we reconstructed the missing data.
-The used gap-filling method will be described in detail in a companion paper.
+The implemented gap-filling method adheres to \citet[][Section~12.5]{ICP:MappingManualSBB2020} and will be described in detail in a companion paper.
 
@@
 To assess the significance of 2018/19 anomalies, we estimate the bias due to the change in the tropospheric background \chem{[O_3]} since the 1990s and account for the missing data in June 2018.
-To reconstruct these missing data, we performed a Reynolds decomposition (Eq.~(\ref{eq:reynolds_decomp})) and cross-calibrate with ozone observations at Esrange and Pallas (companion paper).
+To reconstruct these missing data, we performed a Reynolds decomposition (Eq.~(\ref{eq:reynolds_decomp})) and cross-calibrate with ozone observations at Esrange and Pallas \citep{ICP:MappingManualSBB2020}.
 In the following, we will describe the bias-correction and discuss the significance of \chem{[O_3]} enhancements in 2018.
@@
 First, we give an account of the methodology for deriving bespoke parameters, after which we present and discuss the modeling results.
-The $\mathrm{DO_3SE}$ model and conducted gap-filling methods are described in Appendix~\ref{appendix:do3se_model}.
+The $\mathrm{DO_3SE}$ model and conducted gap-filling methods adhering to \citet{ICP:MappingManualSBB2020} are summarized in Appendix~\ref{appendix:do3se_model}.
 

### 2021-05-27 VegOzone.tex
Remove not used figures
Change title
Update affiliation

@@
 %\title{The 2018 heatwave and its implications on ozone induced damage on vegetation in a subarctic climate}
-\title{Characterizing subarctic biomes for land surface modeling of pollution and climate risk}
+\title{Characterizing subarctic biomes for land surface modeling of ozone pollution and climate risk}
 
@@
 \affil[2]{Department of Biosciences, University of Oslo, Oslo, Norway}
-\affil[3]{Department of Environment and Geography, University of York, UK}
+\affil[3]{Department of Environment and Geography, Stockholm Environment Institute, University of York, UK}
 
@@
 
-%\begin{figure*}[t]
-%  \includegraphics[width=12cm]{figA3}
-%  \caption{Reconstruction of missing \chem{[O_3]} observation data at Svanvik in July 2018 based on derived hourly climatologies of northern Fennoscandia, Svanvik, and data Pallas.
-(a) Time series; (b) anomalies; (c) reconstruction.}
-%  \label{fig:ozone_reconstruction_2018_07}
-%\end{figure*}
-
 \clearpage

### 2021-09-22 VegOzone.tex
Adjust acknowledgment
Update from Overleaf

--- 2021-05-10 VegOzone.tex
@@
 %\title{The 2018 heatwave and its implications on ozone induced damage on vegetation in a subarctic climate}
-\title{Characterizing subarctic biomes for land surface modeling of pollution and climate risk}
+\title{Characterizing subarctic European vegetation for land surface modeling of ozone pollution and climate risk}
 
@@
 
-\Author[1]{Stefanie}{Falk}
-\Author[2]{Ane Victoria}{Vollsnes}
-\Author[2]{Aud}{Eriksen}
+\Author[1,4]{Stefanie}{Falk}
+\Author[2]{Ane V.}{Vollsnes}
+\Author[2]{Aud B.}{Eriksen}
 \Author[3]{Lisa}{Emberson}
@@
 \affil[3]{Department of Environment and Geography, University of York, UK}
+\affil[4]{Department of Geography, Ludwig Maximilians Universität, Munich, Germany}
 
@@
 \begin{abstract}
-  The unique vegetation biomes of the subarctic region acclimatized to extremes of cold and midnight sun are likely to be at threat from the combined impacts of climate change and increasing ozone (\chem{O_3}) concentrations.
-The atmospheric and climatic characteristics of the subarctic are known to lead to pronounced peak \chem{O_3} concentrations in spring.
-To date, only a few studies have been conducted to assess the response of subarctic vegetation to variations in climate and air pollution, at least by comparison with other parts of Europe.
-This study looks to fill this knowledge gap by examining trends in climate variables and \chem{O_3} concentrations that have occurred over the past few decades.
-These climatological and \chem{O_3} data are used to assess the extent to which two recent years (2018 and 2019) deviate from climatic and \chem{O_3} concentration norms and how these current (and potentially more frequent future) deviations may influence \chem{O_3} damage to subarctic vegetation.
-We find that 2018 was an anomalous warm and bright year (in particular in spring and early summer) with higher than average \chem{O_3} concentrations in April/May and a high number of episodic peak \chem{O_3} concentration above $40\,\unit{ppb}$ in June--August, the latter in part attributable to an increase in forest fires in the Northern Hemisphere as well as warmer and sunnier conditions.
-We apply a flux-metric to determine \chem{O_3} risk and damage to vegetation as a function of \chem{O_3} concentration, climate variables, and species-specific physiology.
-We find that bespoke parameterizations of plant functional types (PFTs) for subarctic vegetation bio-types increase estimates of ozone-induced reduction of total biomass by $2.5$ to $17.4\,\unit{\%}$ in this flux-metric.
-Our study, suggests that assessments that use generic parameterizations are likely to underestimate the risk of \chem{O_3} damage in this region.
+  The unique vegetation of the subarctic region acclimatized to extremes of cold and midnight sun are likely to be at threat from the combined impacts of climate change and increasing ozone concentrations (\chem{[O_3]}).
+The atmospheric and climatic characteristics of the subarctic are known to lead to pronounced peak \chem{[O_3]} in spring.
+To date, only a few studies have been conducted to assess the response of subarctic vegetation to variations in climate and air pollution.
+This study looks to fill this knowledge gap by examining essential climate variables, in particular \chem{[O_3]}, over the past few decades.
+These data are used to assess the extent to which two recent years (2018 and 2019) deviate from climatic and \chem{[O_3]} norms and how these potentially more frequent future deviations may influence ozone damage to subarctic vegetation.
+We find that 2018 was an anomalously warm and bright year (in particular in spring and early summer) with higher than average \chem{O_3} concentrations in April/May and a high number of episodic peak \chem{O_3} concentration above $40\,\unit{ppb}$ in June--August, the latter in part attributable to forest fires in the Northern Hemisphere as well as warmer and sunnier conditions.
+We apply the integrated flux-metric Phytotoxic Ozone Dose (POD) to determine ozone risk and damage to vegetation as a function of \chem{[O_3]}, climate variables, and species-specific physiology.
+We find that bespoke parameterizations of plant functional types (PFTs) for subarctic vegetation bio-types result in a ozone-induced biomass loss of $2.5$ to $17.4\,\unit{\%}$.
+Our study, suggests that assessments that use generic parameterizations are likely to underestimate the risk of ozone damage in this region.
 We conclude that appropriate parameterization for the growing season and plant physiological responses to light and temperature are particularly important for accurate modeling of seasonal gas exchange.
-Efforts should be targeted towards accurately defining subarctic bio-type physiological response to climatic variables to improve regional and global scale biogeochemical cycling under current and future climates.
+Efforts should be targeted towards accurately defining subarctic ecotypes' physiological response to essential climate variables to improve regional and global scale biogeochemical cycling under current and future climates.
 \end{abstract}
@@
 Coincident peak \chem{[O_3]} is found in ozone monitoring data from Svanhovd (Fig.~\ref{fig:data_svanvik_2018_2019} but also over whole northern Fennoscandia in July.
+Therefore conditions at Svanhovd for the growing seasons 2018/19 shall serve as a reference for present and probable future conditions in northern Fennoscandia in our study.
+
 During the 2003 drought period, elevated ozone in Europe was promoted by a combination of various factors such as wildfires, reduced cloud cover (increased solar radiation), reduced dry deposition and turbulent mixing due to the stagnant weather conditions, and increased BVOC emissions \citep{JGR:Solberg2018}.
@@
 We explore how risk estimates might be over or underestimated using generic vs bespoke subarctic parameterization.
-We also explore the role of \chem{[O_3]} variability and changing climate variables to understand how this risk might alter in the future with the advent of climate change.
-To this end, conditions at Svanhovd for the growing seasons 2018/19 shall serve as a reference for present and probable future conditions in northern Fennoscandia.
+We also explore the effects of the interannual variability of \chem{[O_3]} and essential climate variables (sensu World Meteorological Organization) on future risk of vegetation damage.
 This allows an understanding of how the threat from ozone may change in the future and which are the key aspects of plant physiology that might determine the potential risk.
@@
 
-The foundation of the NIBIO Environment Centre Svanhovd dates back to the 1930s \citep{NIBIO_Svanhovd}.
-The NIBIO centre Svanhovd is located south of the settlement of Svanvik, where the Pasvik river, which is the border to Russia, lies only a few hundred meters to the east.
-It focuses on the research of the special environment of the region, but also comprises long-term observations of various agrometeorological variables, i.e.
+The NIBIO centre Svanhovd focuses on the research of the special environment of the region, but also comprises long-term observations of various agrometeorological variables, i.e.
 $2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths.
@@
 The locations of the atmospheric monitoring site and the ozone garden are marked in the aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}a).
-This ozone garden was likely the northern-most to date {\bf TODO: Reference?}.
 An ozone garden consists of selected plant species that are sensitive to ozone and are likely to display visible injuries.
@@
   \caption{Cap of the North.
-Locations of past and present (end of 2018) ozone observation sites in northern Fennoscandia were used in this study.
+Locations of past and present (end of 2018) ozone observation sites in northern Fennoscandia used in this study.
 For more details see Table~\ref{tab:ebas_obs}.
-The same color coding for each station as displayed herein will be used in all the following figures.}
+The same color coding for each station as displayed herein will be used in all the following figures.
+Figure adapted from \citet{ACP:Falk2021}.
+\textbf{TODO: Add map copy right}}
   \label{fig:station_map_fennoscandia}
@@
 At Svanhovd, the ozone measurements are taken amidst a meadow surrounded by agriculturally managed herbaceous plants and grass adhering to the \citet{WMOGuide2018} standards (Fig.\ref{fig:svanhovd_research_station}).
-In conjunction with the location at lower altitude, this results in generally lower \chem{[O_3]} at Svanhovd compared to Pallas.
 
 Data from before the 1990s are available solely from Svanvik and Jergul.
-These data, however, do not follow the high-quality standard procedures implemented nowadays and have to be treated with care \citep{NILU2003}.
+These data do not follow the high-quality standard procedures implemented nowadays and have to be treated with care \citep{NILU2003}.
 Especially, ozone monitors did not undergo regular re-calibration.
-This probably leads to drifts in the observed data and may impose a false trend.
-In 1997, the ozone monitoring at Jergul was relocated downstream of the river Karasjokka to a site close by Karasjok.
+This probably lead to drifts in the observed data and may impose a false trend.
+In 1997, the ozone monitoring at Jergul was relocated downstream of the river Karasjohka to a site close by Karasjok.
 Although the same equipment has been used, this might have introduced a baseline shift in the combined data series, denoted as Jergul/Karasjok, which would be particularly problematic if these were the sole data for comparison.
@@
 \begin{table}[t]
-  \caption{Past and present ozone observation sites in northern Fennoscandia. Data available on \href{http://ebas.nilu.no/}{EBAS}.}
+  \caption{Past and present ozone observation sites in northern Fennoscandia. Data available on \href{http://ebas.nilu.no/}{EBAS}. Table adapted from \citet{ACP:Falk2021}}
   \label{tab:ebas_obs}
@@
 \label{subsubsec:anomal_ozone}
+\textbf{TODO: Remove all content which is part of the technical note}
 To assess the significance of 2018/19 anomalies, we estimate the bias due to the change in the tropospheric background \chem{[O_3]} since the 1990s and account for the missing data in June 2018.
@@
 On average, \chem{[O_3]} had been elevated by $(1.9\pm 5.7)\,\unit{ppb}$ and $(1.9\pm 5.4)\,\unit{ppb}$, respectively.
-Both distributions are slightly skewed towards positive deviations with $\frac{1}{4}$ of days (3.
-quantile, q3) falling above $5.8\,\unit{ppb}$ and $5.6\,\unit{ppb}$, respectively.
+Both resemble normal distributions.
 
@@
 In the following, we investigate the difference between 2018 and 2019 concerning stomatal ozone uptake and assess which of the environmental factors might be most influential in determining \chem{O_3} damage risk.
-We model ozone uptake of different natural and semi-natural vegetation using the $\mathrm{DO_3SE}$ model and develop bespoke subarctic parameterizations for common plant species.
+We model ozone uptake of natural and semi-natural vegetation using the $\mathrm{DO_3SE}$ model and develop bespoke subarctic parameterizations for common plant species.
 First, we give an account of the methodology for deriving bespoke parameters, after which we present and discuss the modeling results.
@@
 We deduced the dominant species on-site at the NIBIO Environment Centre Svanhovd and from Fig.~\ref{fig:svanhovd_research_station} and found perennial grassland, birch (generalized as deciduous trees), and Scots pine (generalized as coniferous trees).
-Default parameters for these species or PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as \emph{mapping manual} abbreviated as MM in the following.
-We base our parameterization of coniferous trees on the MM's Boreal Norway spruce and deciduous trees are derived from MM's silver birch.
-For a comprehensive list of model parameters consult Supplement~Table~S1.
+Default parameters for these species or PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as \emph{mapping manual} (MM) parameterizations in the following.
+We base our parameterization of coniferous trees on the MM's Boreal Norway spruce and deciduous trees on the MM's silver birch.
+For a comprehensive list of model parameters, consult Supplement~Table~S1.
 
@@
 We identified $f_\mathrm{T}$ as abnormally low, being the limiting factor of stomatal conductance in perennial grassland.
-Therefore, we saw the necessarity to propose bespoke parameterizations describing an acclimation of these species to the climatic conditions in the target region.
-Note, however, that these parameterizations represent hypothetical PFTs or species that have yet to be verified by experiments.
-
-We hypothesize that perennial vegetation will likely prioritize their growth or fast adolescence and hence their carbon sequestration, during the short, subarctic growing season.
-Because the mapping manual version of the $\mathrm{DO_3SE}$ model currently does not simulate net photosynthesis ($A_\mathrm{n}$), we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune the temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (see Appendix~\ref{appendix:do3se_model} for details).
-We further assume that $f_\mathrm{VPD}$ and $f_\mathrm{SWP}$ suffice our biomes.
+Therefore, we saw the necessity to propose bespoke parameterizations describing an acclimation of of the PFTs to the climatic conditions in the target region.
+Note, however, that these parameterizations are hypothetical and have yet to be verified by experiments.
+
+We assume that perennial vegetation will likely maximize carbon acquisition during the short subarctic growing season.
+Because the mapping manual version of the $\mathrm{DO_3SE}$ model currently does not simulate net photosynthesis ($A_\mathrm{n}$), we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune the temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (see Appendix~\ref{appendix:do3se_model} for details) to acclimate the growth potential to the local climate.
+We further assume that $f_\mathrm{VPD}$ and $f_\mathrm{SWP}$ suit our biomes.
 The procedure is graphically illustrated in Fig.~\ref{fig:f_temp_grassland} for grassland.
-We aim to increase the enclosed area between probability density function (PDF) and the respective response function.
+We aim to increase the overlapping area between the probability density function (PDF) and the respective response function.
 We show two climatological reference periods 1990s and 2000s to indicate that the vegetation is subject to ongoing climate change and hence continuous acclimation.
-Resulting response functions for coniferous and deciduous trees are displayed in Appendix~Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}, respectively.
-
-We define two different hypothetical temperature acclimation: \emph{subarctic} and \emph{cold}.
+The resulting response functions for coniferous and deciduous trees are displayed in Appendix~Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}, respectively.
+
+We define two different hypothetical temperature acclimations: \emph{subarctic} and \emph{cold}.
 For either, we assume that the hypothetical species are more tolerant to cold conditions and shift $f_\mathrm{T}$ towards colder temperatures accordingly.
@@
 The most extreme of the two acclimation types is \emph{subarctic}.
-For perennial grassland the resulting response functions are depicted together with the temperature PDF in Fig.~\ref{fig:f_temp_grassland}a).
+For perennial grassland the resulting response functions are depicted together with a five month growing season temperature PDF of the two periods 1992-2000 and 2011-2019 in Fig.~\ref{fig:f_temp_grassland}a).
 
@@
 Based on the time series of \chem{CO_2} uptake and temperature at observation sites in southern ($\rightarrow$ \emph{cold}) and northern ($\rightarrow$ \emph{subarctic}) Finland \citep{TB:Kolari2007}, we estimate the optimal temperature interval $10\,\unit{^\circ C}\le T_\mathrm{opt} \le 15\,\unit{^\circ C}$.
-We assume a similar temperature acclimation for deciduous trees.
+We assume similar temperature acclimations for deciduous trees.
 All bespoke temperature parameters are listed in Table~\ref{tab:sensitivity_tests_temp}.
 
-With respect to light conditions, we hypothesize that the light threshold for a $50\,\unit{\%}$ opening of plants' stomata may differ in subarctic species compared to species in less extreme climates.
+With respect to light conditions, we assume that the light threshold for a $50\,\unit{\%}$ opening of plants' stomata may differ in subarctic species compared to species in less extreme climates.
 Analytically, we derive the inverse function $f_{\mathrm{light},k}^{-1}$ of Eq.~(\ref{eq:flight}) for each PFT or species $k$
@@
 As indicated above, a better acclimation to prevailing climate conditions will result in a higher average $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$.
-To quantify the resulting bespoke parameterizations, we propose the following metric.
-For a generic GS (May--August), we compute climatological average and standard deviation of $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal})) at noon (averaged over $11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning (averaged over $5-9\,\unit{am}$ local time).
-We presume that relative stomatal conductance around noon (highest light intensity) and in the early morning (lower intensities) are good proxies for \chem{CO_2} uptake efficiency.
-Our arbitrary optimization target is $60\,\unit{\%}$.
-Additional constrains are a small deviation between noon and morning and a low standard deviation indicating a higher robustness to interannual variability of growing conditions.
-For simplicity, we neglect the dependency on SWP in this metric.
-From the results shown in Fig.~\ref{fig:javis_func_opt_mean}, we identify \emph{subarctic}, PPFD0.8 as the best acclimation for all hypothetical species or PFTs alike.
-As expected due to the small adjustments, coniferous trees display the lowest deviation between the different types of acclimation, while the divergence for perennial grassland is substantial as response to the stronger proposed temperature acclimation.
+To compare the bespoke parameterizations, we propose the following metric.
+We calculate an average relative stomatal conductance $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal}) with $f_\mathrm{SWP}=1$) at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time) for the whole record of meteorological data and apply mean and standard deviation.
+We presume that $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ around noon (highest light intensity) is a good proxy for \chem{CO_2} uptake efficiency (growth) which should be maximized.
+In additional, a small deviation between noon and morning and a low standard deviation would indicate a higher robustness to variability in growing conditions.
+From the results shown in Fig.~\ref{fig:javis_func_opt_mean}, we identify \emph{subarctic}, PPFD0.8 as the best acclimation for all PFTs, based on the criteria.
+As expected due to the small adjustments, coniferous trees display the lowest deviation between the different types of acclimation, while the divergence for perennial grassland is substantial as a response to the stronger proposed temperature acclimation.
 
@@
 GS (May--August) climatological averages and standard deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon (averaged over $11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning (averaged over $5-9\,\unit{am}$ local time) are shown for (a) deciduous trees; (b) coniferous trees; (c) perennial grassland.
-For $f_\mathrm{SWP}$ is connected to a plant hydraulic model within $\mathrm{DO_3SE}$ it left out in this calculation.}
+}
   \label{fig:javis_func_opt_mean}
@@
 Start ($A_\text{start}$) and end ($A_\text{end}$) of GS for each PFT are given in units of \unit{doy} (Table~\ref{tab:sensitivity_tests_gs}) and estimated as follows.
-Coniferous trees show photosynthetic activity already at low temperatures \citep{TP:Wallin2013, TB:Kolari2007}.
-We use a MODIS (Aqua/Terra) retrieved net photosynthesis product for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine $A_\text{start}$ and $A_\text{end}$.
-The net photosynthesis data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}).
+We use a MODIS (Aqua/Terra) retrieved net photosynthesis product \textbf{TODO: reference} for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine $A_\text{start}$ and $A_\text{end}$.
+The net photosynthesis vs.
+time of year data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}).
 By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{doy}$ for 2018 and \unit{doy} 106 in 2019.
@@
 This value will be used for all PFTs alike.
-The resulting growing season for coniferous trees in 2019 was one month longer then in 2018.
+The resulting growing season for coniferous trees in 2019 was one month longer than in 2018.
 For estimating $A_\text{start}$ of deciduous trees, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Supplement~Fig.~\ref{fig:greening_season_change_Svanvik}).
 We find 129 and 130, respectively.
-For both years, these dates coincide with the first snow-free day at the meteorological weather station at Øvre~Neiden (Sør-Varanger, NOR) which is relatively close-by.
-For comparison, we would get $A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$ from the latitude model typically used for localizing $\mathrm{DO_3SE}$.
-The latitude model indicates a $2\,\unit{month}$ longer growing season.
-Due to a lack of quantitative field observation, we assume a latency period of $1\,\unit{month}$ after snow-melt for perennial grassland.
+For both years, these dates coincide with the first snow-free day at the closest inland meteorological weather station at Øvre~Neiden (Sør-Varanger, NOR).
+For comparison, the default values ($A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$) indicating a substantially longer growing season.
+Due to a lack of quantitative field observation, we assume a latency period of $1\,\unit{month}$ after snow-melt for perennial grassland \textbf{TODO: Reference}.
 
 \begin{table}[t]
-  \caption{Bespoke start and end of growing season in units of \unit{doy}. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
+  \caption{Start and end of growing season in units of \unit{doy}.}
   \label{tab:sensitivity_tests_gs}
-  \begin{tabular}{llcc}
+  \begin{tabular}{lcll}
     \tophline
@@
     \middlehline
+    Default$^\star$ & -- & 100 & 307\\
     \multirow{2}{*}{Deciduous tree} & 2018 & 129$^*$ & 261$^\dagger$ \\
@@
   \belowtable{$^\dagger$ MODIS (Aqua/Terra) net photosynthesis product;\\
+    $^\star$ \citet{ACP:Simpson2012},\\
     $^*$ $5\,\unit{days}-5\,\unit{^\circ C}$-rule, $T_\mathrm{air}$ from \href{seNorge.no}{seNorge.no};\\
@@
 
-As pointed out by \citet{ACP:Bueker2012}, a very important factor in deriving a proper $\mathrm{POD_y}$ is the soil texture or how much water the soil can hold.
-As Svanhovd is located in the bed of the Pasvik river, we assume a sandy loam texture at $60\,\unit{cm}$ depth.
+Soil texture influences water availability and thus $\mathrm{POD_y}$ calculations in $\mathrm{DO_3SE}$ \citep{ACP:Bueker2012}.
+The soil at Svanhovd is characterized in an ICP Forest plot survey as gleyic cambisols according to the Food and Agriculture Organization of the United Nations (FAO) soil classification system and as eluviated dystric brunisol in the ICP Forest soil classification system (Volkmar Timmerman, pers.
+comm.) Within the $\mathrm{DO_3SE}$ soil parameterization, the sandy loam texture describes the upper $60\,\unit{cm}$, where the plant roots are found, of this soil type best.
+
 The estimate of leaf boundary layer resistance is a function of wind speed which will be influenced by the tree height and leaf dimensions.
-Leaves of downy birch in northern Fennoscandia are smaller than southern specimens' with leaf dimensions of $(3.0\pm 0.5)\,\unit{cm}$.
+Leaves of downy birch in northern Fennoscandia are probably smaller than the $5\,\unit{cm}$ from the MM parameterization of silver birch leaves.
+A sample of downy birch leaves collected at Svanhovd had an average length of $(3.0\pm 0.5)\,\unit{cm}$.
 In addition, \citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area.
 Both factors will affect leaf boundary layer resistance, the smaller leaves would tend to reduce leaf boundary layer resistances whilst the shorter trees would reduce canopy roughness and increase canopy boundary layer resistances and hence the uptake of \chem{O_3}.
-
+In addition, one Scots pine forest at Svanhovd has been studied as a part of the ICP Forest mapping.
+Heights were measured in 2004, when the stand was 90 years old.
+Mean tree height was $10.1\,\unit{m}$ and maximum height was $16.2\,\unit{m}$ \textbf{TODO:(Volkmar Timmerman, priv.
+comm.)}.
+Canopy height and leaf size are adjusted to local values in the models with bespoke growing season (GS bespoke) in the following.
+%%%
 \subsection{$\mathrm{POD_1}$ results and implications on ozone impact}
@@
   Tore Flatlandsmo Berglen for hourly pressure data from Svanhovd.
-  This work was financed by the Research Council of Norway (NRC) grand No. 268073.
+  This work was supported by the Research Council of Norway (NRC) grand No. 268073.
 \end{acknowledgements}

### 2021-09-27 VegOzone.tex
Add  changes from Ane to Section 4 and 5

@@
 Canopy height and leaf size are adjusted to local values in the models with bespoke growing season (GS bespoke) in the following.
-%%%
+
 \subsection{$\mathrm{POD_1}$ results and implications on ozone impact}
@@
 
-We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for the three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using the generic MM parameterization and our bespoke temperaure (\emph{cold}, \emph{subarctic}) and light (PPFD0.8, PPFD1.2) acclimation types, and localized GS.
-The results are shown in Fig.~\ref{fig:pody_rel}.
-For comparison, critical levels for a $4\,\unit{\%}$ (deciduous), $2\,\unit{\%}$ (coniferous) \citep[refer to silver birch and Norway spruce in][]{ICP:MappingManual2017}, and $10\,\unit{\%}$ (perennial grassland, above ground) \citep{ESPR:Hayes2021} reduction in biomass are shown as horizontal dashed lines.
-2018 is color-coded in pink and 2019 in purple.
-Simulations with MM GS, canopy, and leaf parameters are indicated with squared markers, bespoke GS with circles.
-The annotation $\pm$ refers to the bespoke light response function ($+\rightarrow\mathrm{PPFD1.2}$, $-\rightarrow\mathrm{PPFD0.8}$), while bespoke temperature response functions are indicated by categories on the x-axis.
-Closed/open symbols represent SWP sensitivity switched off/on in the model.
-
-The results depicted in Fig.~\ref{fig:pody_rel} allow for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric for ozone induced damage in subarctic vegetation.
-In summary, we identify the following main effects of the bespoke parameterizations of the $\mathrm{DO_3SE}$ model with respect to the MM default parameterization on ozone uptake:
-\begin{itemize}
-  \itemsep0pt
-\item Increasing cold tolerance/heat intolerance $\Rightarrow$ increased uptake in both probed years,
-\item usage of proper growing season $\Rightarrow$ reduced uptake with respect to the increase reported above,
-\item an earlier assumed start of GS $\Rightarrow$ increased uptake,
-\item drought conditions (SWP) $\Rightarrow$ reduced uptake,
-\item lower/higher extent of stomatal opening at low light intensities $\Rightarrow$ increased/reduced uptake,
-\item symmetric variation of the extent of stomatal opening at low light intensities $\Rightarrow$ asymmetric response in POD,
-\item temperature acclimation $\Rightarrow$ amplifying effect on light threshold and drought effects.
-\end{itemize}
-The exact order of magnitude of these effects varies for individual species or PFTs and between years, but the predicted ozone uptake for the bespoke parameterization is always larger than for the MM default parameterizations and of the same order of magnitude as the variability between the years studied here.
-Although our temperature acclimations are hypothetical, they suggest considerable underestimations of ozone risk for subarctic species when compared to generic MM parameterizations.
-In particular, an earlier GS due to climate change, that is overlapping with the ozone spring peak, has the potential to increase the risk of ozone damage for vegetation with an acclimation to cold climates.
-Drought effects (SWP) do only matter if a bespoke GS is not taken into account.
-We will look at the specific species or PFTs in more detail in the following.
-
-For deciduous trees (Fig.~\ref{fig:pody_rel}a) all simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
-Deciduous trees display the largest variance of ozone uptake among all species or PFTs ($10-31\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
-In particular, the sensitivity to the light threshold is especially pronounced, e.g.
-$\Delta\mathrm{POD_1} = ^{+2.5}_{-3.6}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of the subarctic type without bespoke GS in 2019.
-Without bespoke GS, 2018 displays a higher ozone damage risk than 2019.
-However, taking the later start and overall shorter GS in 2018 into account, this is reversed for the \emph{subarctic} acclimation type.
-This can be partly explained by very high light response factors (over $f_\mathrm{light}>0.8$) in the beginning of the GS in these simulations.
-Whether this is a bug or a feature of the $\mathrm{DO_3SE}$ model is not clear.
-SWP induced reduction in ozone uptake is only effective in 2018 and without bespoke GS.
-The temperature acclimation dependent bias in ozone uptake deduced for bespoke GS ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
-The maximum systematic deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of the extent of stomatal opening at low light intensities amount to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
+We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for the three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using the generic MM parameterization and our bespoke temperature (\emph{cold}, \emph{subarctic}) and light (PPFD0.8, PPFD1.2) response functions.
+We also locally adapted leaf size for deciduous trees and the tree height together with the bespoke GS.
+The results displayed in Fig.~\ref{fig:pody_rel} allow for an assessment of systematic uncertainties arising from the use of generic parameterizations established from less climatically extreme conditions.
+For comparison, critical levels for reduction in biomass are shown as horizontal dashed lines \citep[deciduous forest $4\,\unit{\%}$, coniferous forest $2\,\unit{\%}$, and grasslands $10\,\unit{\%}$ biomass reduction,][]{ICP:MappingManual2017,ESPR:Hayes2021}.
+We identify that an increase in cold tolerance leads to higher ozone uptake in both probed years, while the usage of locally adapted GS causes a reduction.
+The earlier and longer GS in 2019 leads to increased ozone uptake compared to 2018 despite more episodes of elevated \chem{[O_3]} in 2018.
+Due to the shape of the light response function, a variation of the extent of stomatal opening at low light intensities leads to an asymmetric response in $\mathrm{POD_1}$ and can cause a higher sensitivity to drought conditions.
+Similarly, the temperature acclimation to cold or subarctic conditions can lead to an amplification of effects by droughts.
+
+The magnitude of these effects varies between PFTs as well as years, but the predicted ozone uptake for the bespoke parameterization is always larger than for the MM default parameterizations and of the same order of magnitude as the variability between the years studied here.
+Although our temperature acclimations are hypothetical, they suggest considerable underestimations of ozone risk for subarctic species when relying on generic MM parameterizations.
+We will look at the specific PFTs in more detail in the following.
+
+For deciduous trees (Fig.~\ref{fig:pody_rel}a) all $\mathrm{POD_1}$ simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
+Deciduous trees display the largest variation of ozone uptake among all PFTs ($10-31\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
+In particular, the sensitivity to the light parameterization is pronounced.
+With MM GS, $\mathrm{POD_1}$ values were higher in 2018 than in 2019, throughout our analyses.
+However, taking the later start and overall shorter GS in 2018 into account (bespoke GS), this is reversed for the \emph{subarctic} acclimation type.
+This can be explained by favorable growing conditions in 2019.
+Low soil water potential reduced the $\mathrm{POD_1}$ values in 2018 in some cases when using the MM growing season, but did not or only slightly when using bespoke growing season.
+The temperature acclimation-dependent bias in ozone uptake deduced for bespoke GS ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+The maximum systematic deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of the extent of stomatal opening at low light intensities amount to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.
 
 For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
-Coniferous trees' acclimation to prevailing temperatures was already good from the beginning and therefore the least biased of the generic MM parameterization.
-Coniferous trees display a medium sensitivity to the variation in the light threshold, e.g.
-$\Delta\mathrm{POD_1} = ^{+1.7}_{-1.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of subarctic type without bespoke GS in 2018.
-Taking the bespoke GS into account, we find a higher ozone uptake in 2019 for all bespoke temperature response functions.
-The overall bias deduced from bespoke GS for subarctic and cold types with respect to MM ranges between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
-The maximum systematic deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of the extent of stomatal opening at low light intensities amount to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
+Coniferous trees' parameterization was already well acclimated to prevailing temperatures from the beginning and therefore the least biased of the generic MM parameterizations.
+Taking the bespoke GS and canopy height into account, we find a higher ozone uptake in 2019 than in 2018 for both bespoke temperature response functions.
+The overall bias deduced from bespoke GS for subarctic and cold types with respect to MM range between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+The maximum systematic deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of the extent of stomatal opening at low light intensities is medium and amounts to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.
 
 For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations with bespoke GS stay below the CL.
-Perennial grassland shows the smallest ozone uptake, but a similarly large bias as deciduous trees.
-We also find a reversal for predicted ozone damage risk in $\mathrm{POD_1}$ between 2018 and 2019 for the subarctic type.
+Perennial grassland shows the smallest ozone uptake, but a similarly large response to the temperature parameterizations as deciduous trees.
+Again, we find a reversal for predicted ozone damage risk in $\mathrm{POD_1}$ between 2018 and 2019 for the subarctic type.
 Perennial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant without bespoke GS.
 The overall bias deduced from bespoke GS for subarctic and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+The lower maximum temperature tolerance in the subarctic compared to the cold temperature parameterization probably caused the lower $\mathrm{POD_1}$ values calculated for 2018 in subarctic compared to cold climate, whereas this was not seen for 2019.
 The maximum systematic deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of the extent of stomatal opening at low light intensities amount to $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
@@
   \caption{$\mathrm{DO_3SE}$ modeling $\mathrm{POD_1}$ results for (a) deciduous trees, (b) coniferous trees, and (c) perennial grasslands at Svanhovd in 2018 (pink) and 2019 (purple).
-The results varied depending on parameterizations of temperature response functions (MM, Cold, and Subarctic, details in Table~\ref{tab:sensitivity_tests_temp}), growing season (MM (squares) or bespoke (circles), details in Table~\ref{tab:sensitivity_tests_gs}), light response functions (first, second and third point attached by a line are depicting the MM, PPFD0.8 and PPFD1.2, details in Table~\ref{tab:sensitivity_tests_light}), and the effect of taking SWP into account (filled symbols without and open symbols with effects of drought kept in the model).
-Critical levels for ozone damage \citet{ICP:MappingManual2017,ESPR:Hayes2021} are given as dashed horizontal lines.
+The results varied depending on parameterizations of temperature response functions (MM, Cold, and Subarctic, details in Table~\ref{tab:sensitivity_tests_temp}), growing season (GS MM (squares) or GS bespoke (circles), details in Table~\ref{tab:sensitivity_tests_gs}), light response functions (first, second and third point attached by a line are depicting the MM, PPFD0.8 and PPFD1.2, details in Table~\ref{tab:sensitivity_tests_light}), and the effect of taking SWP into account (filled symbols without and open symbols with effects of drought kept in the model).
+Critical levels for ozone damage \citep{ICP:MappingManual2017,ESPR:Hayes2021} are given as dashed horizontal lines.
 Note that the horizontal axis is only categorical.}
@@
 The reported range in uncertainty corresponds to the maximum deviation defined by the $\pm 20\,\unit{\%}$ variation of the extent of stomatal opening at low light intensities with SWP=on.
-For deciduous trees and perennial grassland, substantial reductions in total biomass in comparison with the default generic MM parameterizations are predicted for both years.
-Coniferous trees are the least affected and the magnitude of total biomass reduction is independent of the choice of the parameterization within given uncertainties.
-This may reflect the more comprehensive research on this species or PFT feeding into the generic MM parameterization.
+For deciduous trees and perennial grassland, total biomass was reduced in comparison with the default generic MM parameterizations for both years.
+\textbf{TODO: Discussion about CL for grassland!} Coniferous trees are the least affected and the magnitude of total biomass reduction is independent of the choice of the parameterization within given uncertainties.
+This may reflect that the temperature parameterization in the MM is almost never a limiting factor for stomatal opening.
 The systematic deviation for deciduous trees is one order of magnitude larger than for coniferous trees and perennial grassland.
 
+\textbf{TODO: Check if the legnth effect central Europe vs Svanvik is propperly taken care of.}
 From the difference between 2018 and 2019, we infer a substantial interannual variability.
@@
   \caption{Estimated total biomass reduction in \unit{\%} for temperature acclimation \emph{subarctic} with bespoke GS and SWP=off.
-The uncertainty ranges reported here correspond to maximum a divergence deducted from varying the extent of stomatal opening at low light intensities and SWP=on.
+The uncertainty ranges reported here correspond to maximum a divergence deducted from varying $f_\mathrm{light}$ and SWP=on.
 For comparison, the corresponding biomass reduction for the generic default MM averaged over both years is shown.
@@
 & Deciduous tree& Coniferous tree & \multicolumn{2}{c}{Perennial grassland}\\
+& & & total & above ground\\
 \middlehline
-2018 & $15.5^{+(1.9...2.1)}_{-(0.8...1.5)}$ & $2.5^{+0.2}_{-0.2}$ & $9.7^{+0.2}_{-0.2}$ & $^*12.4^{+0.2}_{-0.2}$\\
-2019 & $17.4^{+2.5}_{-1.8}$ & $3.0^{+0.2}_{-0.3}$ & $10.5^{+(0.1...0.2)}_{-(0.0...0.1)}$ & $^*13.7^{+(0.1...0.3)}_{-(0.1...0.3)}$\\
+2018 & $15.5^{+(1.9...2.1)}_{-(0.8...1.5)}$ & $2.5^{+0.2}_{-0.2}$ & $9.7^{+0.2}_{-0.2}$ & $12.4^{+0.2}_{-0.2}$\\
+2019 & $17.4^{+2.5}_{-1.8}$ & $3.0^{+0.2}_{-0.3}$ & $10.5^{+(0.1...0.2)}_{-(0.0...0.1)}$ & $13.7^{+(0.1...0.3)}_{-(0.1...0.3)}$\\
 \middlehline
-$\left<\mathrm{MM}\right>$ & $11.2\pm 1.1$ & $2.31\pm 0.04$ & $7.5\pm 0.9$ & $^*9.6\pm 1.4$\\
+$\left<\mathrm{MM}\right>$ & $11.2\pm 1.1$ & $2.31\pm 0.04$ & $7.5\pm 0.9$ & $9.6\pm 1.4$\\
 \bottomhline
 \end{tabular}
-\belowtable{$^*$ above ground biomass} % Table Footnotes
+%\belowtable{$^*$ above ground biomass} % Table Footnotes
 \end{table}
@@
 
-We have studied the effect of acclimations of biomes to a subarctic climate on phyto-toxic ozone uptake.
+\textbf{TODO: Decide on terms acclimation vs parameterization etc} We have studied the effect of model acclimations of some common vegetation types to a subarctic climate on phytotoxic ozone uptake.
 The comparison between 2018 and 2019 conditions allowed us to consider these acclimations and their influence on vegetation sensitivity to \chem{O_3} in light of future changes of key environmental variables as may occur under climate change (e.g.
@@
 This was particularly a result of the extended heatwave in spring and early summer and associated, extensive wildfires in Central Sweden (in particular in mid--end July).
+\textbf{TODO: Check explanation regarding significance.}
 
@@
 Based on gridded observational temperature and snow depth, we found a half a month shorter growing season for deciduous trees and perennial grassland in 2018 compared to 2019, respectively.
-The timing and emerging trend in the start of the growing season may differ substantially between local observation and common thermal or latitudinal estimations.
-Species-specific methods show a later start for European silver birch and an earlier for Norway spruce, compared to the thermal growing season \citep{IVL:Karlsson2018}.
-With respect to ongoing climate change, a clear positive trend in length ($5.2\,\unit{d\,decade^{-1}}$) almost equally distributed between earlier start ($2.9\,\unit{days\,decade^{-1}}$) and later end ($2.3\,\unit{d\,decade^{-1}}$) could be deducted from the thermal growing season at Svanhovd.
+With respect to ongoing climate change, a clear positive trend in length ($5.2\,\unit{d\,decade^{-1}}$) almost equally distributed between earlier start ($2.9\,\unit{days\,decade^{-1}}$) and later end ($2.3\,\unit{d\,decade^{-1}}$) could be deducted from the thermal growing season at Svanhovd (Supplement Fig.~\ref{fig:greening_season_change_Svanvik}).
+\textbf{TODO: Move S6 into appendix.}
 
 Ozone peak concentrations were observed more frequently in 2018 than in 2019.
-In 2018, we found significantly enhanced \chem{[O_3]} particularly in April and May, but not during the active wildfires in July.
-This suggests that weather conditions and high ozone concentration at the beginning of the growing season are more important in risk assessment using flux-based metrics than episodes of peak concentrations over shorter periods.
-Because these peak \chem{[O_3]} may coincident with environmental conditions that limit ozone uptake and hence do not lead to damage.
-Consequently, acute cellular damage caused by very high fluxes over a short period is not accounted for properly in $\mathrm{POD_y}$.
-As pointed out by \citet{AE:Musselman2006}, a higher temporal resolution, e.g.
-susceptibility to ozone depending on leaf age, in the associated damage functions might give a different picture.
-For herbal and grassland species (similar for crops) a strong dependency on the phenological state, e.g.
-high susceptibility to ozone in the reproductive state, was found \citep{EP:Bassin2004}.
+Nevertheless, \chem{POD_1} values were very similar for both years.
+This suggests that the overall length of the growing season is more important in risk assessment using flux-based metrics than episodes of peak concentrations.
+These peak \chem{[O_3]} often coincide with environmental conditions that limit ozone uptake the model.
+However, acute cellular damage caused by very high episodic ozone fluxes are not accounted for in $\mathrm{POD_y}$.
+That might explain why ozone damage was observed at Svanhovd to a larger extent in 2018 than in 2019.
+As pointed out by \citet{AE:Musselman2006}, a susceptibility to ozone depending on leaf age in the associated damage functions might give a different picture.
+Forbs and perennial grassland have been found to be more susceptible in the reproductive state \citep{EP:Bassin2004}.
 

### 2021-09-27 VegOzone.tex
Add changes to section 5 - not finalized

@@
 
-To estimate potential biases in the assessment of ozone-induced risk on Fennoscandia vegetation, we defined hypothetical temperature response functions representing different levels of assumed acclimation to cold climates.
+To estimate potential biases in the assessment of ozone-induced risk on Fennoscandic vegetation, we defined hypothetical temperature response functions representing different levels of assumed acclimation to cold climates.
 We assume northern Fennoscandia vegetation to be more tolerant towards cold temperatures and less tolerant to heat.
 Based on probability density functions (PDF) of observed light and temperature, we assume that an acclimation ought to increase the enclosed area between the respective response functions and the PDF.
-These hypothetical species or PFTs still need verification by experimental data.
+These hypothetical parameterisations of species or PFTs still need verification by experimental data.
 
@@
  \item the bespoke parameterizations for subarctic species are important (especially concerning light and temperature and the climate effect on the growing season) to be able to assess \chem{O_3} uptake,
- \item the parameterizations defined in the mapping manual for European regional risk assessment would appear to not adequately capture the physiological responses of subarctic physiology that are important determinants of \chem{[O_3]} sensitivity under prevailing environmental conditions and are likely to substantially underestimate risk,
- \item the ozone reanalysis products tend to underestimate \chem{[O_3]} in Fennoscandia, and
+ \item the parameterizations defined in the mapping manual for European regional risk assessment appear to not adequately capture the physiological responses of subarctic physiology that are important determinants of \chem{[O_3]} sensitivity and are likely to substantially underestimate risk, and
  \item that the climatic conditions promoting ozone damage (particularly extended heatwaves and the timing of the growing season concerning the ozone spring peak) and which are likely to become more frequent under climate change are likely to further enhance the sensitivity of subarctic vegetation to ozone.
@@
 
-Based on the methods described in \citet{ICP:MappingManual2017}, we estimated biomass reductions due to ozone uptake for both 2018 and 2019 and find significant reductions for deciduous trees ($15.5-17.4\,\unit{\%}$) and perennial grassland (above ground, $12.4-13.7\,\unit{\%}$) with respect to estimates based on $\mathrm{POD_1}$ from default parameterizations.
-For coniferous trees, the results are compatible with the default parameterization estimates and are contained within the range of systematic uncertainties deduced from our bespoke response function ($2.5-3\,\unit{\%}$).
-These results have to be treated with care because the biomass reduction functions were not established based on subarctic species and might as well be subject to biases.
+Based on the methods described in \citet{ICP:MappingManual2017}, we estimated biomass reductions due to ozone uptake for both 2018 and 2019 and found significant reductions for deciduous trees ($15.5-17.4\,\unit{\%}$) and perennial grassland (above ground, $12.4-13.7\,\unit{\%}$) with respect to estimates based on $\mathrm{POD_1}$ from default parameterizations.
+For coniferous trees, the adjusted parameterizations are comparable with the default parameterization estimates and are contained within the range of systematic uncertainties deduced from our bespoke response function ($2.5-3\,\unit{\%}$).
+These results have to be treated with care because the biomass reduction functions were not established based on subarctic species.
 
@@
 Cumulative ozone uptake has been reported to reduce stomatal conductance significantly and independently from photosynthesis \citep[e.g.]{BGS:Lombardozzi2012}, with photosynthesis being more strongly affected.
-This might be due competing effects, as an ozone-induced sluggishness \citep{SR:Hoshika2015} renders stomata unable to fully close and leads to higher stomatal conductance than expected and thus higher ozone uptake as well as transpiration.
+This might be due to competing effects, as an ozone-induced sluggishness \citep{SR:Hoshika2015} renders stomata unable to fully close and leads to higher stomatal conductance than expected and thus higher ozone uptake as well as transpiration.
 At present, the MM version of the $\mathrm{DO_3SE}$ model does not account for any ozone-inflicted reduction of stomatal conductance throughout the growing season and is not coupled to photosynthesis.

### 2021-09-27 VegOzone.tex
Add reference to Korhonen

@@
 For comparison, the default values ($A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$) indicating a substantially longer growing season.
-Due to a lack of quantitative field observation, we assume a latency period of $1\,\unit{month}$ after snow-melt for perennial grassland \textbf{TODO: Reference}.
+Due to a lack of quantitative field observation, we assume a latency period of $1\,\unit{month}$ after snow-melt for perennial grassland \citep[][Supplement Fig.~S1, Rovaniemi]{FCR:Korhonen2018}.
 

### 2021-09-27 VegOzone.tex
Update abstract

@@
 \begin{abstract}
-  The unique vegetation of the subarctic region acclimatized to extremes of cold and midnight sun are likely to be at threat from the combined impacts of climate change and increasing ozone concentrations (\chem{[O_3]}).
+  The unique vegetation of the subarctic region acclimatized to extremes of cold and midnight sun are likely to be at threat from the combined impacts of climate change and increasing ozone concentrations \chem{[O_3]}.
 The atmospheric and climatic characteristics of the subarctic are known to lead to pronounced peak \chem{[O_3]} in spring.
@@
 These data are used to assess the extent to which two recent years (2018 and 2019) deviate from climatic and \chem{[O_3]} norms and how these potentially more frequent future deviations may influence ozone damage to subarctic vegetation.
-We find that 2018 was an anomalously warm and bright year (in particular in spring and early summer) with higher than average \chem{O_3} concentrations in April/May and a high number of episodic peak \chem{O_3} concentration above $40\,\unit{ppb}$ in June--August, the latter in part attributable to forest fires in the Northern Hemisphere as well as warmer and sunnier conditions.
+We find that 2018 was an anomalously warm and bright year (in particular in spring and early summer) with higher than average \chem{[O_3]} in April/May and a high number of episodic peak \chem{[O_3]} above $40\,\unit{ppb}$ in June--August, the latter in part attributable to forest fires in the Northern Hemisphere as well as warmer and sunnier conditions.
 We apply the integrated flux-metric Phytotoxic Ozone Dose (POD) to determine ozone risk and damage to vegetation as a function of \chem{[O_3]}, climate variables, and species-specific physiology.
+Our study, suggests that assessments that use generic parameterizations are likely to underestimate the risk of ozone damage in this region.
 We find that bespoke parameterizations of plant functional types (PFTs) for subarctic vegetation bio-types result in a ozone-induced biomass loss of $2.5$ to $17.4\,\unit{\%}$.
-Our study, suggests that assessments that use generic parameterizations are likely to underestimate the risk of ozone damage in this region.
-We conclude that appropriate parameterization for the growing season and plant physiological responses to light and temperature are particularly important for accurate modeling of seasonal gas exchange.
 Efforts should be targeted towards accurately defining subarctic ecotypes' physiological response to essential climate variables to improve regional and global scale biogeochemical cycling under current and future climates.

### 2021-09-27 VegOzone.tex
Remove sentences and add heyes reference

@@
 Recent trend analyses for Europe indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric ozone started to level off or even decline \citep{ESA:Cooper2014, ACP:Wespes2018,ESA:Gaudel2018}.
-This can be attributed to a successful implementation of air quality regulations coupled with economic restructuring, reducing the number of episodes of peak concentrations especially in summertime over Europe and North America \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
+This can be attributed to a successful implementation of air quality regulations coupled with economic restructuring, reducing the number of episodes of peak concentrations especially in summertime \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
 However, changes in environmental conditions associated with climate change, such as an increase in the frequency of heatwaves, could negate the effectiveness of emission reductions in relation to ozone impacts and air quality \citep{NCC:Lin2020}.
@@
 Complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) and hydrocarbons known as volatile organic compounds (VOCs) in the presence of nitrogen oxides (\chem{NO_x}) lead to the formation of \chem{O_3} at ground-level.
-Despite a relatively short average tropospheric lifetime of approximately $22\,\unit{days}$, ranging from a few days in the tropical boundary layer to up to $1\,\unit{year}$ in the upper troposphere \citep{JGR:Stevenson2005,ACP:Young2013}, ozone and is precursors are subject to advection over long distances.
+Despite a relatively short average tropospheric lifetime of approximately $22\,\unit{days}$ \citep{JGR:Stevenson2005,ACP:Young2013}, ozone and is precursors are subject to advection over long distances.
 Episodes of high ozone concentrations in the absence of local pollution sources can therefore often be attributed to the long-range transport of both ozone and its precursors.
@@
 The spring peak is more pronounced in northern Fennoscandia than in more southern locations but shows a  high interannual variability \citep{AB:Klingberg2009, BER:Klingberg2019}.
-A cause for this is the complex interplay between atmospheric dynamics and chemistry \citep{AE:Laurila1996,BER:Hatakka2003} as described above as well as anthropogenic and biogenic activity \citep{AE:Rummukainen1996,AE:Simpson2002,QJRMS:Galbally2007,NGS:Schnell2009}.
+A cause for this is the complex interplay between atmospheric dynamics and chemistry \citep{AE:Laurila1996,BER:Hatakka2003} as well as anthropogenic and biogenic activity \citep{AE:Rummukainen1996,AE:Simpson2002,QJRMS:Galbally2007,NGS:Schnell2009}.
 \citet{ACP:Andersson2017} showed in a modeling study focused on Fennoscandia that the variability of ozone concentrations in winter can be attributed mainly to changes in atmospheric background (transport) of ozone, while summertime abundance is mostly affected by emissions of precursors in the rest of Europe.
@@
 For natural and semi-natural vegetation which is already subject to rapid climatic changes known as Arctic amplification \citep{AMAP2012,IPCC2013}, these factors could promote a higher potential risk for northern vegetation to ozone-induced damage in the not so distant future.
+Recently, \citet{ESPR:Hayes2021} showed that the highest sensitivity to future \chem{[O_3]} variability is expected in summer when vegetation is most productive.
 

### 2021-09-27 VegOzone.tex
Rework introduction

@@
 Most of these previous studies have focused on the Mediterranean where soil moisture deficits are thought to be most influential in decoupling concentration from uptake.
-Relatively few studies have explored the situation in Northern European climates where low temperatures may be considered more likely to influence ozone uptake \citep[e.g.]{iF:Juran2018}.
-
-To estimate the flux-based metric, also referred to as the Phyto-toxic Ozone Dose over a threshold y ($\mathrm{POD_y}$), an estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf and environmental factors affecting stomatal conductance as described further in Section~\ref{sec:do3se}.
+Relatively few studies have explored the situation in Northern European climates \citep[e.g.][]{iF:Juran2018}.
+
+To estimate the flux-based metric, also referred to as the Phyto-toxic Ozone Dose over a threshold y ($\mathrm{POD_y}$), an estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf and environmental factors affecting stomatal conductance.
 $\mathrm{POD_y}$ is then calculated according to:
@@
 This study demonstrated the improved ability of flux-based metrics (in comparison to exposure-based metrics) to identify the geographical distribution of the risk of ozone damage and also showed that the ozone damage could extend into more northerly regions of Europe.
-A similar bio-monitoring study was conducted at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd to assess whether local ozone concentrations were capable of inducing damage to vegetation growing under northern Fennoscandia conditions.
+We conducted a similar bio-monitoring study at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd to assess whether local ozone concentrations were capable of inducing damage to vegetation growing under northern Fennoscandia conditions.
 In 2018, we indeed observed ozone damage on clover and tobacco.
@@
 Boreal wildfires emit in addition to \chem{CO_2} also \chem{CO} ($\chem{[CO]}/\chem{[CO_2]} \propto 6-13\,\unit{\%}$) and VOCs ($\chem{[VOCs]}/\chem{[CO_2]} \propto 0.5-1.5\,\unit{\%}$) \citep{AE:Cofer1990}.
-A total of 2079 forest fires of different sizes were registered in Norway in 2018, twice as many as in the preceding years 2016/17 \citep{DSB2019}, last accessed April 2020).
+A total of 2079 forest fires were registered in Norway in 2018, twice as many as in the preceding years 2016/17 \citep{DSB2019}, last accessed April 2020).
 In Sweden, about 500 fires had been reported (five times more than in a usual summer), and an estimated total of $25000\,\unit{hectare}$ burned down in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) \citep{SOU2019}.
-Coincident peak \chem{[O_3]} is found in ozone monitoring data from Svanhovd (Fig.~\ref{fig:data_svanvik_2018_2019} but also over whole northern Fennoscandia in July.
-Therefore conditions at Svanhovd for the growing seasons 2018/19 shall serve as a reference for present and probable future conditions in northern Fennoscandia in our study.
-
-During the 2003 drought period, elevated ozone in Europe was promoted by a combination of various factors such as wildfires, reduced cloud cover (increased solar radiation), reduced dry deposition and turbulent mixing due to the stagnant weather conditions, and increased BVOC emissions \citep{JGR:Solberg2018}.
-It is likely that these factors also contributed to enhanced \chem{[O_3]} in 2018.\\
+Coincident peak \chem{[O_3]} is found in ozone monitoring data (Fig.~\ref{fig:data_svanvik_2018_2019}) in July.
+It is very likely that elevated ozone in Europe was promoted by a combination of various factors such as wildfires, reduced cloud cover (increased solar radiation), reduced dry deposition and turbulent mixing due to the stagnant weather conditions, and increased BVOC emissions like during the 2003 drought period,\citep{JGR:Solberg2018}.
+Therefore conditions at Svanhovd for the growing seasons 2018/19 shall serve as a reference for present and probable future conditions in northern Fennoscandia in our study.\\
 
@@
 The observation site at NIBIO Environment Centre Svanhovd and all related data are presented in Section~\ref{sec:data}.
-In Section~\ref{sec:stats}, we derive multi-annual means for ground-level ozone concentration and key environmental variables (temperature, precipitation, and global irradiance) based on in situ observations and evaluate the deviation of 2018/19 from the norm.
+\textbf{TODO: Rewrite!} In Section~\ref{sec:stats}, we derive multi-annual means for ground-level ozone concentration and key environmental variables (temperature, precipitation, and global irradiance) based on in situ observations and evaluate the deviation of 2018/19 from the norm.
 We use the $\mathrm{DO_3SE}$ model to estimate the ozone uptake by natural and semi-natural vegetation to investigate the main environmental drivers for projected ozone damage risk in 2018 (Section~\ref{sec:do3se}).

### 2021-09-27 VegOzone.tex
Rework section 2 intro

@@
 In our assessment of primary drivers of ozone damage risk on vegetation in the subarctic, we focus on conditions during the 2018/19 growing season at the NIBIO operated Environment Centre Svanhovd.
-We also set these into a climatological and regional context.
-To this end, we collected atmospheric and ozone monitoring data taken at Svanhovd and additional ozone monitoring data from observation sites in northern Finland and Sweden for cross-evaluation.
-Varying update frequencies and length of available data sets made it necessary to fetch the data from various sources (further details below).
+We also set these into a climatological context.
+To this end, we collected atmospheric and ozone monitoring data taken at Svanhovd over the past $35\,\unit{years}$.
 Svanhovd synonymously refers to the atmospheric monitoring site Svanvik/Pasvik throughout the paper.\\
@@
 
-To relate local ozone concentration (\chem{[O_3]}) with visible damage on ozone sensitive plants, both an ozone bio-monitor (referred to as ozone garden) and a conventional ozone monitor using the existing atmospheric monitoring infrastructure had been installed during the 2018/19 growing season at the NIBIO centre.
-The locations of the atmospheric monitoring site and the ozone garden are marked in the aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}a).
+All measured atmospheric variables including ozone for the 2018/19 growing season were obtained through \href{luftkvalitet.no}{luftkvalitet.no} operated by \citet{NILU_AIRQ}.
+The ozone monitoring data have been also added to the EBAS database operated by \citet{NILU_EBAS} and are thus openly accessible.
+Long-term ozone observation data (1986--1996) were obtained from EBAS.
+The long-term data sets of agrometeorological variables including temperature and precipitation are available from September 1992 to the present day \citep[LandbruksMeteorologiske Tjeneste][note the station name here is Pasvik]{LMT_NIBIO}.
+
+To relate local ozone concentration \chem{[O_3]} with visible damage on ozone sensitive plants, both an ozone bio-monitor (referred to as ozone garden) and a conventional ozone monitor using the existing atmospheric monitoring infrastructure had been installed during the 2018/19 growing season at the NIBIO centre.
+We will report \chem{[O_3]} as volume mixing ratios (VMR) in \unit{ppb}.
+The locations of the atmospheric monitoring site and the ozone garden are marked in the aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}a.
 An ozone garden consists of selected plant species that are sensitive to ozone and are likely to display visible injuries.
@@
 clover, tobacco, and potatoes.
-As shown in Fig.~\ref{fig:svanhovd_research_station}b), the plants had to be protected from herbivores with a wire-mesh fence.
+As shown in Fig.~\ref{fig:svanhovd_research_station}b, the plants had to be protected from herbivores with a wire-mesh fence.
 In 2018, we qualitatively observed visible ozone damage on semi-natural vegetation (clover) and crops (tobacco).
 In contrast, no such visible damage was found on the clovers in 2019, although it was observed on the sensitive tobacco cultivar.
-This indicates that the vegetation may have been more affected by ozone in 2018 than in 2019.
-
-All measured atmospheric variables including ozone for the 2018/19 growing season were obtained through \href{luftkvalitet.no}{luftkvalitet.no} operated by \citet{NILU_AIRQ}.
-The ozone monitoring data have been also added to the EBAS database operated by \citet{NILU_EBAS} and are thus openly accessible.
-All long-term ozone observation data described in Section~\ref{subsec:ozone_data} are also obtained from EBAS.
-The long-term data sets of agrometeorological variables including temperature and precipitation are available from September 1992 to the present day \citep[LandbruksMeteorologiske Tjeneste][note the station name here is Pasvik]{LMT_NIBIO}.
+This indicates that the vegetation was more affected by ozone in 2018 than in 2019.
 

### 2021-09-27 VegOzone.tex
Restructure Section 2

@@
 Long-term ozone observation data (1986--1996) were obtained from EBAS.
+Note that data from before the 1990s do not follow the high-quality standard procedures implemented nowadays and have to be treated with care \citep{NILU2003}.
+Especially, ozone monitors did not undergo regular re-calibration.
+This probably lead to drifts in the observed data and may impose a false trend.
 The long-term data sets of agrometeorological variables including temperature and precipitation are available from September 1992 to the present day \citep[LandbruksMeteorologiske Tjeneste][note the station name here is Pasvik]{LMT_NIBIO}.
@@
 As shown in Fig.~\ref{fig:svanhovd_research_station}b, the plants had to be protected from herbivores with a wire-mesh fence.
-In 2018, we qualitatively observed visible ozone damage on semi-natural vegetation (clover) and crops (tobacco).
-In contrast, no such visible damage was found on the clovers in 2019, although it was observed on the sensitive tobacco cultivar.
-This indicates that the vegetation was more affected by ozone in 2018 than in 2019.
 
@@
 
-As can be seen in Fig.~\ref{fig:data_svanvik_2018_2019}a), \chem{[O_3]} peaks in spring (April/May) and reaches its minimum in late summer (July/August).
+As can be seen in Fig.~\ref{fig:data_svanvik_2018_2019}a, \chem{[O_3]} peaks in spring (April/May) and reaches its minimum in late summer (July/August).
 The spring peak has not been captured completely in 2019, since data acquisition only started in late April.
@@
 In contrast, \chem{[O_3]} only rose 18 times above the threshold of $40\,\unit{ppb}$ during the summer of 2019.
-For further analysis (Section~\ref{sec:stats}) and modeling study with the $\mathrm{DO_3SE}$ model (Section~\ref{sec:do3se}), we reconstructed the missing data.
-The used gap-filling method will be described in detail in a companion paper.
+A reconstruction method for gap-filling the data has been presented in \citet{ACPD:Falk2021}.
 
@@
 In 2018, spring temperature regularly rose above freezing only in May, while in 2019 this occurred already early in March/April.
-More rain events with accumulated daily precipitation ($\sum_d \mathrm{Precip}$) above $10\,\unit{mm}$ occurred in the summer of 2018 compared to 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}c).
-Qualitatively, global irradiance ($Q_0$) displayed in Fig.~\ref{fig:data_svanvik_2018_2019}d) was higher in May and July 2018 compared to 2019, while June 2019 showed higher irradiance than 2018.
+More rain events with accumulated daily precipitation ($\sum_d \mathrm{Precip}$) above $10\,\unit{mm}$ occurred in the summer of 2018 compared to 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}c.
+Qualitatively, global irradiance ($Q_0$) displayed in Fig.~\ref{fig:data_svanvik_2018_2019}d was higher in May and July 2018 compared to 2019, while June 2019 showed higher irradiance than 2018.
 A high global irradiance in most cases is the result of a low cloud fraction.
 In both years, the maximum recorded $Q_0$ was $750\,\unit{W\,m^{-2}}$ and reached in June.
+
+In 2018, we qualitatively observed visible ozone damage on semi-natural vegetation (clover) and crops (tobacco).
+In contrast, no such visible damage was found on the clovers in 2019, although it was observed on the sensitive tobacco cultivar.
+This indicates that the vegetation was more affected by ozone in 2018 than in 2019.
 
@@
 
-\subsection{Ozone data}
-\label{subsec:ozone_data}
-From 1986--1996, ozone monitoring has been operated continuously by NILU at Svanhovd, but no ozone measurements have been conducted before the 2018/19 growing season.
-To determine the significance of the \chem{[O_3]} in 2018/19 in relation to regional trends, we derive a ground-level \chem{[O_3]} climatology for Svanvik and northern Fennoscandia in general (Section~\ref{sec:stats}).
-
-%\subsubsection{Ground-level ozone observations}
-%\label{subsubsec:ebas}
-\begin{figure}[t]
-  %\includegraphics[width=8.3cm]{station_map_fennoscandia}
-  \includegraphics[width=8.3cm]{fig03}
-  \caption{Cap of the North.
-Locations of past and present (end of 2018) ozone observation sites in northern Fennoscandia used in this study.
-For more details see Table~\ref{tab:ebas_obs}.
-The same color coding for each station as displayed herein will be used in all the following figures.
-Figure adapted from \citet{ACP:Falk2021}.
-\textbf{TODO: Add map copy right}}
-  \label{fig:station_map_fennoscandia}
-\end{figure}
-
-A comprehensive overview over all past and present ozone observation sites in northern Fennoscandia is given in Table~\ref{tab:ebas_obs} and Fig.~\ref{fig:station_map_fennoscandia}, respectively.
-All stations with available long-term observations are located at higher altitude than Svanhovd.
-In the absence of local precursors, the stratification of the atmosphere usually leads to a higher abundances in ozone at higher altitude \citep[e.g.,][]{AB:Klingberg2009}.
-The periphery vegetation at all sites is similar and consists mainly of pine forests, birch, perennial grassland, and heath.
-However, on closer inspection the vegetation differs substantially.
-Ozone measurements at Pallas are taken on top of mount Sammaltunturi, $100\,\unit{m}$ above treeline (pine forest), and the surrounding vegetation mainly consists of low vascular plants, mosses, and lichen \citep{BER:Hatakka2003}.
-At Svanhovd, the ozone measurements are taken amidst a meadow surrounded by agriculturally managed herbaceous plants and grass adhering to the \citet{WMOGuide2018} standards (Fig.\ref{fig:svanhovd_research_station}).
-
-Data from before the 1990s are available solely from Svanvik and Jergul.
-These data do not follow the high-quality standard procedures implemented nowadays and have to be treated with care \citep{NILU2003}.
-Especially, ozone monitors did not undergo regular re-calibration.
-This probably lead to drifts in the observed data and may impose a false trend.
-In 1997, the ozone monitoring at Jergul was relocated downstream of the river Karasjohka to a site close by Karasjok.
-Although the same equipment has been used, this might have introduced a baseline shift in the combined data series, denoted as Jergul/Karasjok, which would be particularly problematic if these were the sole data for comparison.
-Only a few data have been taken upstream the Pasvik river at Janiskovski, hence these will not be taken into further consideration, but are listed for completeness.
-There is an overlap of about $4\,\unit{years}$ between data taken at Esrange, Jergul/Karasjok, and Svanvik (Supplement~Fig.~\ref{fig:ozone_timesseries_fenoscandic_obs}).
-
-\begin{table}[t]
-  \caption{Past and present ozone observation sites in northern Fennoscandia. Data available on \href{http://ebas.nilu.no/}{EBAS}. Table adapted from \citet{ACP:Falk2021}}
-  \label{tab:ebas_obs}
-  \begin{tabular}{lllllrl}
-    \tophline
-    Name       & Country & ID      & \multicolumn{3}{c}{Location} & Operational\\
-    &         &         & lat             & lon               & alt            &\\
-    &         &         & (\unit{^\circ N}) & (\unit{^\circ E})  & (\unit{m})     &\\
-    \middlehline
-    Esrange    & SWE     & SE0013R & 67.83           & 21.07             & 475            & 1991 -- 2018$^1$\\
-    Janiskoski & RUS     & RU0001R & 68.93           & 28.85             & 118            & 1995 -- 1997\\
-    Jergul     & NOR     & NO0030R & 69.45           & 24.60             & 255            & 1997 -- 2011\\
-    Karasjok   & NOR     & NO0055R & 69.467          & 25.217            & 333            & 1988 -- 1997\\
-    Pallas     & FIN     & FI0096G & 67.97           & 24.12             & 565            & 1995 -- present$^2$\\
-    Svanvik    & NOR     & NO0047R & 69.45           & 30.03             & 30             & 1986 -- 1996$^3$\\
-    \bottomhline
-  \end{tabular}
-  \belowtable{$^1$ Data available on EBAS until the end of 2018.\\
-    $^2$ Data available on EBAS until the end of 2019.\\
-    $^3$ Exclusive ozone monitoring in growing seasons 2018/19 for the present study operated by NILU.} % Table Footnotes
-\end{table}
-
 \section{Statistical analysis}
@@
 \label{subsubsec:clim_temp_prec}
-The location of Svanvik at $69.45\,\unit{^\circ N}$ suggests a subarctic climate.
+The location of Svanvik ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m}$) suggests a subarctic climate.
 The derived climate diagram based on climatological data (1992--2012) (Fig.~\ref{fig:climatediagram}) supports this very well.

### 2021-09-27 VegOzone.tex
Remove subsection and add reference to gap-filling paper

@@
 
-\subsection{2018/19 growing season}
-\label{subsec:growing_season_1819}
-
 All relevant data for the growing season 2018/19 are shown as time series in Fig.~\ref{fig:data_svanvik_2018_2019}.
@@
 For the growing seasons 2018/19, we evaluate the statistical significance of divergences from the norm in these variables (referred to as anomalies).
-We perform a bias correction accounting for the increase in tropospheric background \chem{[O_3]} and cross-calibrate the derived climatology for Svanvik with long-term observations of \chem{[O_3]} at Esrange and Pallas.
+We performed a bias correction accounting for the increase in tropospheric background \chem{[O_3]} and cross-calibrate the derived climatology for Svanvik with long-term observations of \chem{[O_3]} at other sites in northern Fennoscandia as described in \citet{ACPD:Falk2021}.
 

### 2021-09-27 VegOzone.tex
Remove subsections

@@
 
-\subsubsection{Temperature, precipitation, and global irradiance}
-\label{subsubsec:clim_temp_prec}
 The location of Svanvik ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m}$) suggests a subarctic climate.
@@
 
-\subsubsection{Ozone}
-\label{subsubsec:clim_ozone}
-In the following, we derive an ozone climatology (\chem{\left<[O_3]\right>}) for Svanvik and for northern Fennoscandia.
-To this end, we first calculate a pairwise correlation between the different stations.
-The correlation of \chem{[O_3]} between Esrange, Pallas, and Jergul/Karasjok is fairly high ($r^2\approx 0.8$, Appendix Fig.~\ref{fig:density_distribution}).
-Therefore, we select these data to compute the generalized ozone climatology for northern Fennoscandia.
-The correlation between Svanvik with any of the active monitoring sites in 2018 (Esrange and Pallas) is rather low, 0.4 and 0.6, respectively.
-Thus, conditions at Svanvik cannot directly be represented by measurements taken at either Esrange or Pallas.
-
-In Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}, two dimensional density distributions of \chem{[O_3]} for the combined data (Esrange, Jergul/Karasjok, Pallas) (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}a) and Svanvik (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b) are shown.
+The derived ozone climatology (\chem{\left<[O_3]\right>}) for Svanvik is shown in Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}.
 Darker colors indicate higher probability to observe these values.
@@
 Splines have been fitted through the data for easier inter-comparison.
-
-The generalized northern Fennoscandic ozone climatology peaks at $48\,\unit{ppb}$  in late April (day of the year (doy) $100-120$) and declines throughout the growing season until a minimum is reached in the beginning of September ($240-250\,\unit{doy}$) and amounts to $24-30\,\unit{ppb}$.
-The turning points lie approximately in June and November.
-The build-up of the ozone spring peak takes about $7.5\,\unit{months}$ from minimum to maximum.
-The most pronounced increase in \chem{\left<[O_3]\right>} coincides with rising temperatures and global irradiance.
-The depletion of ozone from maximum to minimum takes only $4.5\,\unit{months}$.
-
-The \chem{[O_3]} density distribution and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} for Svanvik display a similar pattern as the generalized northern Fennoscandia climatology (Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_spline}b).
+The \chem{[O_3]} density distribution and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} for Svanvik display a pronounced seasonal cycle.
 The spring peak occurs slightly earlier ($100\,\unit{doy}$) and amounts to $40\,\unit{ppb}$ while the annual average \chem{\left<[O_3]\right>} is $6.6\,\unit{ppb}$ lower than the Fennoscandic climatology.
@@
 
-\subsubsection{Temperature, precipitation, and global irradiance}
-\label{subsubsec:anomal_tpq}
 In Fig.~\ref{fig:anomalies_svanvik}a), temperature anomalies for Svanvik are shown as a percentage of days warmer or colder than $\pm 1\,\sigma$ from the climatological mean for each month.
@@
 
-\subsubsection{Ozone}
-\label{subsubsec:anomal_ozone}
 \textbf{TODO: Remove all content which is part of the technical note}

### 2021-09-27 VegOzone.tex
Refracture Section 4

@@
 For the growing seasons 2018/19, we evaluate the statistical significance of divergences from the norm in these variables (referred to as anomalies).
-We performed a bias correction accounting for the increase in tropospheric background \chem{[O_3]} and cross-calibrate the derived climatology for Svanvik with long-term observations of \chem{[O_3]} at other sites in northern Fennoscandia as described in \citet{ACPD:Falk2021}.
+We performed a bias correction accounting for the increase in tropospheric background \chem{[O_3]} and cross-calibrated the derived climatology for Svanvik with long-term observations of \chem{[O_3]} at other sites in northern Fennoscandia as described in \citet{ACPD:Falk2021}.
 
@@
 
+%Ozone
 The derived ozone climatology (\chem{\left<[O_3]\right>}) for Svanvik is shown in Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}.
@@
 On top of the density distributions, a $10\,\unit{days}$ average of daily mean (\chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}}) is displayed together with $1 \sigma$ uncertainties and SE, respectively.
-Splines have been fitted through the data for easier inter-comparison.
+A spline was fitted through the data for comparison.
 The \chem{[O_3]} density distribution and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} for Svanvik display a pronounced seasonal cycle.
-The spring peak occurs slightly earlier ($100\,\unit{doy}$) and amounts to $40\,\unit{ppb}$ while the annual average \chem{\left<[O_3]\right>} is $6.6\,\unit{ppb}$ lower than the Fennoscandic climatology.
-This is in line with a general decrease in tropospheric ozone with decreasing altitude.
-The earlier decline in \chem{\left<[O_3]\right>} may point to an influence of coniferous vegetation on the removal of ozone since \chem{CO_2} uptake by coniferous trees has been observed as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013}.
+The spring peak occurs on $100\,\unit{doy}$ and amounts to $40\,\unit{ppb}$ while the annual average \chem{\left<[O_3]\right>} is $28\,\unit{ppb}$.
+The early decline in \chem{\left<[O_3]\right>} may point to an influence of coniferous vegetation on the removal of ozone.
+Coniferous trees have been observed \chem{CO_2} uptake as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013}.
 The estimated average begin of growing season for deciduous trees ($\left<G_\mathrm{start}\right>_\mathrm{decid}$) is $(150\pm 14)\,\unit{doy}$ at Svanvik (see Section~\ref{subsec:do3se_parameters} for details).
@@
 This hints to ozone uptake by vegetation even at low light intensities during midnight sun conditions in combination with stable planetary boundary layer conditions preventing mixing of ozone rich air.
-The derived \chem{\left<[O_3]\right>} is likely biased towards lower \chem{[O_3]} due to an increase in northern hemispheric tropospheric background \chem{[O_3]} since the 1990s.
-To account for this, we derive a bias correction factor by cross-calibrating with the time series taken at Esrange and Pallas (Section~\ref{subsec:anomalies}).
 
@@
   \includegraphics[width=8.3cm]{fig05}
-  \caption{Ozone climatologies.
-Fennoscandia combines all available data from Jergul/Karasjok, Esrange, and Pallas.
-The density distributions are shown together with a multi-annual mean of daily \chem{[O_3]} (climatology).
-Splines have been fitted through daily mean \chem{[O_3]}.
-The climatology of daily mean ozone is shown together with $1\,\sigma$ standard deviation and standard error of the mean.
-Since \chem{[O_3]} are strongly correlated for sites at Jergul/Karasjok (NOR), Esrange (SWE), and Pallas (FIN) (see Fig.~\ref{fig:density_distribution}), data of these have been averaged together to derive a climatology for northern Fennoscandia.
-(a) Northern Fennoscandia; (b) Svanvik.}
+  \caption{Svanhovd ozone climatology.
+The density distribution is shown together with a multi-annual mean of daily \chem{[O_3]} (climatology).
+A spline has been fitted through daily mean \chem{[O_3]}.
+The climatology of daily mean ozone is shown together with $1\,\sigma$ standard deviation and standard error of the mean.}
   \label{fig:ozone_climatology_fenoscandic_obs_spline}
@@
 In the following, we look at the 2018/19 anomalies and discuss the differences between the two years.
-We also conduct a bias correction for the Svanvik \chem{\left<[O_3]\right>.} As weather extremes like in 2018 are likely to become more frequent with climate change, this will help to assess possible future changes in the vulnerability of subarctic vegetation to ozone.
-
-In Fig.~\ref{fig:anomalies_svanvik}a), temperature anomalies for Svanvik are shown as a percentage of days warmer or colder than $\pm 1\,\sigma$ from the climatological mean for each month.
+As weather extremes like in 2018 are likely to become more frequent with climate change, this will help to assess possible future changes in the vulnerability of subarctic vegetation to ozone.\\
+
+In Fig.~\ref{fig:anomalies_svanvik}a, temperature anomalies for Svanvik are shown as a percentage of days warmer or colder than $\pm 1\,\sigma$ from the climatology for each month.
 Negative deviations from climatology are displayed as a negative percentage.
@@
 Especially in May and July more than $40\,\unit{\%}$ of the days were significantly warmer.
-Significantly warmer days continued to occur in all months (August--November) except for October.
+Significantly more warmer days continued to occur in all months (August--November) except for October.
 March 2018 had many unusually cold days.
-The summer of 2019 was fairly average, however, significantly colder days occurred in July, while August had both significantly colder and warmer days.
+The summer of 2019 was fairly average, however, significantly more colder days occurred in July, while August had both significantly colder and warmer days.
 April 2019 had significantly warmer than average days.
@@
 
-Similarly, the precipitation anomalies for Svanvik are displayed in Fig.~\ref{fig:anomalies_svanvik}b).
+Similarly, the precipitation anomalies for Svanvik are displayed in Fig.~\ref{fig:anomalies_svanvik}b.
 Due to the large interannual variability in observed precipitation, only a few days deviated significantly from the climatology on the $1 \sigma$ level.
@@
 Unlike the temperature anomalies, the picture is not as clear.
-While March 2018 had a significant number of days wetter than unusual, summer and fall (May/July, September/October) had a significant number of days drier than average.
+While March 2018 had a significant number of days wetter than usual, summer and fall (May/July, September/October) had a significant number of days drier than average.
 2019 was rather average throughout the year, but with a significant number of days drier than normal in September/October and wetter in December.
 
-The global irradiance anomalies for Svanvik are presented in Fig.~\ref{fig:anomalies_svanvik}c).
+The global irradiance anomalies for Svanvik are presented in Fig.~\ref{fig:anomalies_svanvik}c.
 The percentage of days above or below $\pm 1\,\sigma$ from the climatological mean is displayed for each month.
-The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed.
 In summer 2018 (May/July), global irradiance showed a significant number of days with higher than average irradiance, while 2019 was generally an average year with a significant number of days darker in spring (April/May).
 
-In summary, 2018 had a significant number of days that were warmer, drier, and brighter than the climatological average, while 2019 was a rather average year.
-
-\textbf{TODO: Remove all content which is part of the technical note}
-To assess the significance of 2018/19 anomalies, we estimate the bias due to the change in the tropospheric background \chem{[O_3]} since the 1990s and account for the missing data in June 2018.
-To reconstruct these missing data, we performed a Reynolds decomposition (Eq.~(\ref{eq:reynolds_decomp})) and cross-calibrate with ozone observations at Esrange and Pallas (companion paper).
-In the following, we will describe the bias-correction and discuss the significance of \chem{[O_3]} enhancements in 2018.
-
-In 2018, the probability density functions (PDFs) of ozone concentration anomalies (\chem{\Delta[O_3]}) for Esrange and Pallas deviate clearly, but not significantly from the climatological mean (Supplement~Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}).
-On average, \chem{[O_3]} had been elevated by $(1.9\pm 5.7)\,\unit{ppb}$ and $(1.9\pm 5.4)\,\unit{ppb}$, respectively.
-Both resemble normal distributions.
-
-We find a similar skew towards higher \chem{[O_3]} for Svanvik in 2018, while 2019 follows a normal distribution rather well (Supplement~Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}).
-In conjunction with the meteorological conditions, we therefore assume that 2019 had been a normal year and the annual average anomaly in 2019 (\chem{\left<\Delta[O_3]\right>}~=~$(1.2\pm 5.0)\,\unit{ppb}$) can be used for the bias correction.
-We also consult the trends in ozone observations at Esrange and Pallas to strengthen this assumption and compute the average difference between recent years and the mid 1990s (\chem{\Delta_{present-past}\left<[O_3]\right>}).
-Herein, present is defined as the average over 2010--2018 and past as the time between the respective beginning of data taking until 1997.
-We do not find a clear trend.
-For Esrange \chem{\Delta_{present-past}\left<[O_3]\right>} amounts to $(1.5\pm 3.1)\,\unit{ppb}$ and for Pallas to $(-2.5\pm 3.7)\,\unit{ppb}$ \citep[][for comparison]{NILU2003, BER:Ruoho-Airola2015}.
-Only the observed trend at Esrange is compatible with our assumed bias correction for the Svanvik ozone climatology.
-
-%To further investigate the ozone residuals, we conduct a Student's~t-test ($t = \frac{\chem{\Delta[O_3]}}{\sigma}$), assuming the same sample uncertainty $\sigma$ in both, climatology and the year in question.
-Results for Esrange and Pallas are shown in Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test}.
-This t-test clearly shows, that at both sites the residuals are significant on the $1\sigma$ level for at least $50\,\unit{\%}$.
-In Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}, the test statistics for Svanvik are displayed.
-From Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}a), we can deduce, that the residuals between the Svanvik climatology and the northern Fennoscandia climatology are significant on the $2\sigma$ level for at least half of the data, while the residuals are less certain in 2018 and 2019.
-In Supplement Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}b) the same is shown for the years 2018/19 for the climatology from Svanvik.
-We find that less than $50\,\unit{\%}$ of ozone concentration residuals are significant on the $1\sigma$ level.
-
-In accordance to the above findings, we adjust the derived ozone climatology for Svanvik by adding a constant bias correction factor $\Delta\chem{[O_3]_\mathrm{bias-corr}} = 1.2\,\unit{ppb}$ and assess the monthly significance of the ozone concentration anomalies in 2018/19.
-The resulting percentage of days above/below $\pm 1\,\sigma$ from the bias-corrected climatological mean is displayed for each month in Fig.~\ref{fig:anomalies_svanvik}d).
+%Ozone
+We adjusted the ozone climatology for Svanvik \citep{ACPD:Falk2021} and assess the monthly significance of the ozone concentration anomalies in 2018/19.
+The resulting percentage of days above/below $\pm 1\,\sigma$ from the bias-corrected climatological mean is displayed for each month in Fig.~\ref{fig:anomalies_svanvik}d.
 Negative deviations from climatology are shown as a negative percentage.
@@
 These results indicate that the extensive forest fires in Sweden did not contribute to a significant enhancement in \chem{[O_3]} at Svanhovd in 2018, but the stagnant meteorological conditions in April/May in combination with a late start of the growing season.
-However, there are indications that the sites further south at Esrange and Pallas experienced more days of elevated ozone in 2018 than Svanvik.
-
-%To conclude, ozone concentrations in 2018 have been elevated compared to the respective climatologies, however, on average not significantly by statistical means.
-At Esrange and Pallas at least $25\,\unit{\%}$ (3.~quantile) of anomalies are significant on a $3\sigma$ level or more, while at Svanvik the 3.~quantile lies at about $2\sigma$.
-This means that in 2018 the stations at Esrange and Pallas saw more days with elevated ozone than was monitored at Svanvik.
+
+In summary, 2018 had a significant number of days that were warmer, drier, and brighter than the climatological average, while 2019 was a rather average year.
+Ozone was not significantly enhanced during the growing season in 2018, but a substantially higher number of peak \chem{[O_3]} were observed in 2018 than in 2019.
 

### 2021-09-27 VegOzone.tex
Remove reference to gap filling
Replace figures

@@
 
-%\graphicspath{{pictures/}}
-\graphicspath{{figures/}}
+\graphicspath{{pictures/}}
+%\graphicspath{{figures/}}
 
@@
 \begin{figure}[t]
-  %\includegraphics[width=8.3cm]{svanhovd_researchstation}\\
-  %\includegraphics[width=8.3cm]{IMG_7739.JPG}
-  \includegraphics[width=8.3cm]{fig01}
+  \centering
+  (a)\\
+  \includegraphics[width=8.3cm]{svanhovd_researchstation}\\
+  (b)\\
+  \includegraphics[width=8.3cm]{IMG_7739.JPG}
+  %\includegraphics[width=8.3cm]{fig01}
   \caption{NIBIO Environment Centre Svanhovd close by the settlement of Svanvik, Norway.
@@
 \begin{figure*}[t]
-  %\includegraphics[width=12cm]{data_svanvik_2018_2019}
-  \includegraphics[width=12cm]{fig02}
+  \includegraphics[width=12cm]{data_svanvik_2018_2019}
+  %\includegraphics[width=12cm]{fig02}
   \caption{Observational data from atmospheric monitoring at Svanhovd in 2018/19.
@@
 \begin{figure}[t]
-  %\includegraphics[width=8.3cm]{climatediagram}
-  \includegraphics[width=8.3cm]{fig04}
+  \includegraphics[width=8.3cm]{climatediagram}
+  %\includegraphics[width=8.3cm]{fig04}
   \caption{Climate diagram following Walter and Lieth.
@@
 \begin{figure}[t]
-  %\includegraphics[width=8.3cm]{ozone_climatology_fenoscandic_obs_spline}
-  \includegraphics[width=8.3cm]{fig05}
+  \includegraphics[width=8.3cm]{ozone_climatology_svanvik_obs_spline}
+  %\includegraphics[width=8.3cm]{fig05}
   \caption{Svanhovd ozone climatology.
@@
 \begin{figure}[t]
-  \includegraphics[width=12cm]{fig06}
+  %\includegraphics[width=12cm]{fig06}
+  \centering
+  (a)\hspace{8.3cm}(b)\\
+  \includegraphics[width=8.3cm]{temperature_signific_1.pdf}
+  \includegraphics[width=8.3cm]{precipitation_signific_1.pdf}\\
+  (c)\hspace{8.3cm}(d)\\
+  \includegraphics[width=8.3cm]{global_rad_signific_1.pdf}
+  \includegraphics[width=8.3cm]{ozone_signific_1.pdf}
   \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as percentages of days significantly deviating from climatological mean through standard deviation for each month.
@@
 First, we give an account of the methodology for deriving bespoke parameters, after which we present and discuss the modeling results.
-The $\mathrm{DO_3SE}$ model and conducted gap-filling methods are described in Appendix~\ref{appendix:do3se_model}.
+The relevant parts of $\mathrm{DO_3SE}$ model are summarized in Appendix~\ref{appendix:do3se_model}.
 
@@
 \begin{figure}[t]
-  %\includegraphics[width=8.3cm]{javis_funcs_temp_hist_grassland}\\
-  %\includegraphics[width=8.3cm]{javis_funcs_rad_hist_grassland}
-  \includegraphics[width=8.3cm]{fig07}
+  \centering
+  (a)\\
+  \includegraphics[width=8.3cm]{javis_funcs_temp_hist_grassland}\\
+  (b)\\
+  \includegraphics[width=8.3cm]{javis_funcs_light_hist_grassland}
+  %\includegraphics[width=8.3cm]{fig07}
 \caption{Construction of bespoke response functions for grassland.
@@
 \begin{figure}[t]
-  %\includegraphics[width=8.3cm]{javis_func_opt_mean}
-  \includegraphics[width=8.3cm]{fig08}
+  \includegraphics[width=8.3cm]{javis_func_opt_mean}
+  %\includegraphics[width=8.3cm]{fig08}
   \caption{Proposed metric to test bespoke response functions.
@@
 The resulting growing season for coniferous trees in 2019 was one month longer than in 2018.
-For estimating $A_\text{start}$ of deciduous trees, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Supplement~Fig.~\ref{fig:greening_season_change_Svanvik}).
+For estimating $A_\text{start}$ of deciduous trees, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Appendix~Fig.~\ref{fig:greening_season_change_Svanvik}).
 We find 129 and 130, respectively.
@@
 For comparison, the default values ($A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$) indicating a substantially longer growing season.
-Due to a lack of quantitative field observation, we assume a latency period of $1\,\unit{month}$ after snow-melt for perennial grassland \citep[][Supplement Fig.~S1, Rovaniemi]{FCR:Korhonen2018}.
+Due to a lack of quantitative field observation, we assume a latency period of $1\,\unit{month}$ after snow-melt for perennial grassland \citep[][Supplement~Fig.~S1, Rovaniemi]{FCR:Korhonen2018}.
 
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{fig09}
+  \centering
+  (a)\\
+  \includegraphics[width=8.3cm]{do3se_results_pod_v3_Birch.pdf}\\
+  (b)\\
+  \includegraphics[width=8.3cm]{do3se_results_pod_v3_Spruce.pdf}\\
+  (c)\\
+  \includegraphics[width=8.3cm]{do3se_results_pod_v3_Grassland.pdf}
+  %\includegraphics[width=8.3cm]{fig09}
   \caption{$\mathrm{DO_3SE}$ modeling $\mathrm{POD_1}$ results for (a) deciduous trees, (b) coniferous trees, and (c) perennial grasslands at Svanhovd in 2018 (pink) and 2019 (purple).
@@
 Based on gridded observational temperature and snow depth, we found a half a month shorter growing season for deciduous trees and perennial grassland in 2018 compared to 2019, respectively.
-With respect to ongoing climate change, a clear positive trend in length ($5.2\,\unit{d\,decade^{-1}}$) almost equally distributed between earlier start ($2.9\,\unit{days\,decade^{-1}}$) and later end ($2.3\,\unit{d\,decade^{-1}}$) could be deducted from the thermal growing season at Svanhovd (Supplement Fig.~\ref{fig:greening_season_change_Svanvik}).
+With respect to ongoing climate change, a clear positive trend in length ($5.2\,\unit{d\,decade^{-1}}$) almost equally distributed between earlier start ($2.9\,\unit{days\,decade^{-1}}$) and later end ($2.3\,\unit{d\,decade^{-1}}$) could be deducted from the thermal growing season at Svanhovd (Appendix Fig.~\ref{fig:greening_season_change_Svanvik}).
 \textbf{TODO: Move S6 into appendix.}
@@
 \appendix
-\section{Derived climatologies}    %% Appendix A
-
-%\subsection{Temperature, precipitation, and global irradiance}     %% Appendix A1, A2, etc.
-
-\appendixfigures  %% needs to be added in front of appendix figures
-\begin{figure}[th]
-  \includegraphics[width=8.3cm]{figA1}
-  \caption{Observed global irradiance at Svanvik.
-(a) Climatology (hourly maxima, mean, and minima); deviation from climatology (b) 2018; (c) 2019.
-The cyan lines indicate the $1\,\sigma$ level derived from climatology.}
-  \label{fig:global_rad_clim}
-\end{figure}
-
-\begin{figure}[th]
-  \includegraphics[width=8.3cm]{figA2}
-  \caption{Probability densities and correlation coefficient for \chem{[O_3]} between different sites in northern Fennoscandia.
-Jergul/Karasjok is well correlated with Esrange, and Pallas.
-Svanvik displays highest correlation with Pallas.
-Note: The stripe at $26\,\unit{ppb}$ is an artifact from plotting.
-(a) Jergul/Karasjok--Esrange; (b) Jergul/Karasjok--Pallas; (c) Svanvik--Esrange; (d) Svanvik--Pallas.}
-  \label{fig:density_distribution}
-\end{figure}
-
-\clearpage
+\section{Growing season}
+\label{appendix:growing_season}
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{greening_season_change_Svanvik}
+  \caption{Estimated shift and prolongation of growing season at Svanhovd over the past 6 decades based on data from \citet{SeNorge}.}
+  \label{fig:greening_season_change_Svanvik}
+\end{figure*}
 
@@
 \end{equation}
-with climatology \chem{\left<[O_3]\right>} and anomalies \chem{\Delta[O_3]}.
-Observed \chem{[O_3]} in 2018 at Pallas served as the reference.
-The exact procedure will be described in an companion paper.
+with climatology \chem{\left<[O_3]\right>} and anomalies \chem{\Delta[O_3]} as described in \citet{ACPD:Falk2021}.
 
@@
 \begin{figure*}[t]
-  \includegraphics[width=12cm]{figB1}
+  %\includegraphics[width=12cm]{figB1}
+  \includegraphics[width=12cm]{DO3SE_results_pody_gsto_o3.png}
   \caption{$\mathrm{DO_3SE}$ modeling results for mapping manual default parameterization.
@@
 \begin{figure}[th]
-  \includegraphics[width=8.3cm]{figB2}
+  %\includegraphics[width=8.3cm]{figB2}
+  \includegraphics[width=8.3cm]{modis_Psn.pdf}
   \caption{Estimated $G_\mathrm{start}$ and $G_\mathrm{end}$ of growing season for coniferous trees from MODIS Aqua/Terra net photosynthesis (PSN) product.
@@
 \begin{figure}[t]
-  %\includegraphics[width=8.3cm]{javis_funcs_temp_hist_evergreen}\\
-  %\includegraphics[width=8.3cm]{javis_funcs_rad_hist_evergreen}
-  \includegraphics[width=8.3cm]{figB3}
+  \centering
+  (a)\\
+  \includegraphics[width=8.3cm]{javis_funcs_temp_hist_coniferous}\\
+  (b)\\
+  \includegraphics[width=8.3cm]{javis_funcs_light_hist_coniferous.pdf}
+  %\includegraphics[width=8.3cm]{figB3}
 \caption{Construction of bespoke response functions for coniferous trees.
@@
 \begin{figure}[t]
-  %\includegraphics[width=8.3cm]{javis_funcs_temp_hist_birch}\\
-  %\includegraphics[width=8.3cm]{javis_funcs_rad_hist_birch}
-  \includegraphics[width=8.3cm]{figB4}
+  \centering
+  (a)\\
+  \includegraphics[width=8.3cm]{javis_funcs_temp_hist_deciduous.pdf}\\
+  (b)\\
+  \includegraphics[width=8.3cm]{javis_funcs_light_hist_deciduous.pdf}
+  %\includegraphics[width=8.3cm]{figB4}
 \caption{Construction of bespoke response functions for deciduous trees.
@@
 
-\input{supplements/VegOzone_supplement}
-
 \end{document}

### 2021-09-27 VegOzone.tex
Placing of appendix figure
Rearange text

@@
 Initial test simulations with the MM PFTs revealed an unrealistically low stomatal conductance at leaf-level ($G_\mathrm{sto}^\mathrm{leaf}$) in 2019 particularly for perennial grassland (Appendix~Fig.~\ref{fig:pody_mm_composit}).
-Substantial stomatal conductance occured only during an extended warm period in late July.
+Substantial stomatal conductance occurred only during an extended warm period in late July.
 In ecological terms, this means that there was almost no growth of grass in the summer of 2019 -- a prediction which is easily falsified by reality.
@@
 Because the mapping manual version of the $\mathrm{DO_3SE}$ model currently does not simulate net photosynthesis ($A_\mathrm{n}$), we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune the temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (see Appendix~\ref{appendix:do3se_model} for details) to acclimate the growth potential to the local climate.
-We further assume that $f_\mathrm{VPD}$ and $f_\mathrm{SWP}$ suit our biomes.
-The procedure is graphically illustrated in Fig.~\ref{fig:f_temp_grassland} for grassland.
+We further assume that $f_\mathrm{VPD}$ and $f_\mathrm{SWP}$ suit our vegetation types.
 We aim to increase the overlapping area between the probability density function (PDF) and the respective response function.
-We show two climatological reference periods 1990s and 2000s to indicate that the vegetation is subject to ongoing climate change and hence continuous acclimation.
-The resulting response functions for coniferous and deciduous trees are displayed in Appendix~Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}, respectively.
 
@@
 In the same way, \emph{subarctic} refers to a hypothetical species which is very tolerant to cold but is sensitive to high temperatures, and most efficient at cool temperatures ($T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000).
+We show two climatological reference periods 1990s and 2000s to indicate that the vegetation is subject to ongoing climate change and hence continuous acclimation.
 The most extreme of the two acclimation types is \emph{subarctic}.
-For perennial grassland the resulting response functions are depicted together with a five month growing season temperature PDF of the two periods 1992-2000 and 2011-2019 in Fig.~\ref{fig:f_temp_grassland}a).
+For perennial grassland the resulting response functions are depicted together with a five month growing season temperature PDF of the two periods 1992-2000 and 2011-2019 in Fig.~\ref{fig:f_temp_grassland}a.
+
+The resulting response functions for coniferous and deciduous trees are displayed in Appendix~Figs.~\ref{fig:f_temp_spruce} and \ref{fig:f_temp_birch}, respectively.
 
@@
 \label{appendix:growing_season}
-\begin{figure*}[t]
+
+\begin{figure*}[th]
   \includegraphics[width=12cm]{greening_season_change_Svanvik}

### 2021-09-28 VegOzone.tex
Restructure text

@@
 In summary, 2018 had a significant number of days that were warmer, drier, and brighter than the climatological average, while 2019 was a rather average year.
-Ozone was not significantly enhanced during the growing season in 2018, but a substantially higher number of peak \chem{[O_3]} were observed in 2018 than in 2019.
+Ozone was not significantly enhanced during the GS 2018/19, but a substantially higher number of peak \chem{[O_3]} were observed in 2018 than in 2019.
 
@@
 In the following, we investigate the difference between 2018 and 2019 concerning stomatal ozone uptake and assess which of the environmental factors might be most influential in determining \chem{O_3} damage risk.
-We model ozone uptake of natural and semi-natural vegetation using the $\mathrm{DO_3SE}$ model and develop bespoke subarctic parameterizations for common plant species.
+We model ozone uptake of natural and semi-natural vegetation using the $\mathrm{DO_3SE}$ model \citep{ACP:Bueker2012} and develop bespoke parameterizations for common plant species in our focus area.
 First, we give an account of the methodology for deriving bespoke parameters, after which we present and discuss the modeling results.
-The relevant parts of $\mathrm{DO_3SE}$ model are summarized in Appendix~\ref{appendix:do3se_model}.
+The relevant equations of the $\mathrm{DO_3SE}$ model are summarized in Appendix~\ref{appendix:do3se_model}.
 
@@
 We deduced the dominant species on-site at the NIBIO Environment Centre Svanhovd and from Fig.~\ref{fig:svanhovd_research_station} and found perennial grassland, birch (generalized as deciduous trees), and Scots pine (generalized as coniferous trees).
-Default parameters for these species or PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as \emph{mapping manual} (MM) parameterizations in the following.
-We base our parameterization of coniferous trees on the MM's Boreal Norway spruce and deciduous trees on the MM's silver birch.
-For a comprehensive list of model parameters, consult Supplement~Table~S1.
+Default parameters for these species or PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as mapping manual (MM) parameterizations in the following.
+We base our parameterization of coniferous trees on the MM's Boreal Norway spruce, deciduous trees on the MM's silver birch, and perennial grassland on MM's perennial grassland for central Europe.
+For a comprehensive list of model default parameters, consult Supplement~Table~S1.
 
@@
 We identified $f_\mathrm{T}$ as abnormally low, being the limiting factor of stomatal conductance in perennial grassland.
-Therefore, we saw the necessity to propose bespoke parameterizations describing an acclimation of of the PFTs to the climatic conditions in the target region.
+Therefore, we saw the necessity to propose bespoke parameterizations describing an acclimation of the default PFTs to the climatic conditions in the focus area.
 Note, however, that these parameterizations are hypothetical and have yet to be verified by experiments.
@@
 Because the mapping manual version of the $\mathrm{DO_3SE}$ model currently does not simulate net photosynthesis ($A_\mathrm{n}$), we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune the temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (see Appendix~\ref{appendix:do3se_model} for details) to acclimate the growth potential to the local climate.
-We further assume that $f_\mathrm{VPD}$ and $f_\mathrm{SWP}$ suit our vegetation types.
-We aim to increase the overlapping area between the probability density function (PDF) and the respective response function.
-
-We define two different hypothetical temperature acclimations: \emph{subarctic} and \emph{cold}.
-For either, we assume that the hypothetical species are more tolerant to cold conditions and shift $f_\mathrm{T}$ towards colder temperatures accordingly.
+Further, we presume that $f_\mathrm{VPD}$ and $f_\mathrm{SWP}$ suit our vegetation types and no acclimation is necessary for these.
+
+We define two different hypothetical temperature acclimations: subarctic and cold.
 We construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to MM.
-In the same way, \emph{subarctic} refers to a hypothetical species which is very tolerant to cold but is sensitive to high temperatures, and most efficient at cool temperatures ($T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000).
-We show two climatological reference periods 1990s and 2000s to indicate that the vegetation is subject to ongoing climate change and hence continuous acclimation.
+This is accomplished by moving $T_\mathrm{opt}$ towards cooler temperatures while keeping the other parameters fixed to default values.
+In the same way, \emph{subarctic} constructed to represent a species which is very tolerant to cold but sensitive to high temperatures, and most efficient at cool temperatures.
+Therefore, we shift $T\mathrm{min}$ and $T\mathrm{max}$ to colder temperatures and choose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000.
+In other words, we aim to increase the overlapping area between the $f_\mathrm{temp}$ and the temperature PDF of the GS.
 The most extreme of the two acclimation types is \emph{subarctic}.
-For perennial grassland the resulting response functions are depicted together with a five month growing season temperature PDF of the two periods 1992-2000 and 2011-2019 in Fig.~\ref{fig:f_temp_grassland}a.
-
-The resulting response functions for coniferous and deciduous trees are displayed in Appendix~Figs.~\ref{fig:f_temp_spruce} and \ref{fig:f_temp_birch}, respectively.
-
-Norway spruce was experimentally found to be active already at rather low air temperatures and can reach $60\,\unit{\%}$ photosynthetic activity as early as \unit{doy}~100 \citep{TB:Kolari2007, TP:Wallin2013}.
-Based on the time series of \chem{CO_2} uptake and temperature at observation sites in southern ($\rightarrow$ \emph{cold}) and northern ($\rightarrow$ \emph{subarctic}) Finland \citep{TB:Kolari2007}, we estimate the optimal temperature interval $10\,\unit{^\circ C}\le T_\mathrm{opt} \le 15\,\unit{^\circ C}$.
-We assume similar temperature acclimations for deciduous trees.
-All bespoke temperature parameters are listed in Table~\ref{tab:sensitivity_tests_temp}.
-
-With respect to light conditions, we assume that the light threshold for a $50\,\unit{\%}$ opening of plants' stomata may differ in subarctic species compared to species in less extreme climates.
-Analytically, we derive the inverse function $f_{\mathrm{light},k}^{-1}$ of Eq.~(\ref{eq:flight}) for each PFT or species $k$
+In Fig.~\ref{fig:f_temp_grassland}a, we show the resulting $f_\mathrm{temp}$ for perennial grassland together with temperature PDFs deduced from five months GS in two climatological periods.
+The periods 1990s and 2000s have been chosen to indicate that the vegetation is subject to ongoing climate change Thus, the predominant acclimation at present is not known.
+
+The resulting response functions for coniferous and deciduous trees are displayed in Appendix~Figs.~\ref{fig:f_temp_spruce}a and \ref{fig:f_temp_birch}a, respectively.
+We base our temperature acclimation of coniferous trees on experimental results on Norway spruce which were found to be active already at rather low air temperatures and can reach $60\,\unit{\%}$ photosynthetic activity as early as \unit{doy}~100 \citep{TB:Kolari2007, TP:Wallin2013}.
+In particular, we use the time series of \chem{CO_2} uptake and temperature at observation sites in southern ($\rightarrow$ \emph{cold}) and northern ($\rightarrow$ \emph{subarctic}) Finland by \citet{TB:Kolari2007} to estimate the optimal temperature interval ($10\,\unit{^\circ C}\le T_\mathrm{opt} \le 15\,\unit{^\circ C}$).
+We presume a similar temperature acclimation for deciduous trees.
+All bespoke temperature parameters are listed in Table~\ref{tab:sensitivity_tests_temp}.\\
+
+Regarding the opening extent of stomata at low light conditions, we look at the PPFD threshold for a $50\,\unit{\%}$ opening of plants' stomata.
+We assume that this threshold may differ in subarctic species compared to species in less extreme climates.
+We derive the inverse function $f_{\mathrm{light},k}^{-1}$ of Eq.~(\ref{eq:flight}) analytically for each PFT or species $k$
 \begin{equation}
@@
 \end{equation}
-First, we solve Eq.~(\ref{eq:inverse_function}) for the MM default value $\alpha_\mathrm{MM}^k$ for each PFT or species $k$ at $0.5$ 
+First, we solve Eq.~(\ref{eq:inverse_function}) for the MM default value $\alpha_\mathrm{MM}^k$ at $0.5$ 
 \begin{equation}
@@
 \end{equation}
-and define a variation $\gamma' = \gamma \cdot \eta$ with $\eta \in \{0.8, 1.2\}$.
-We invert Eq.~(\ref{eq:inverse_function}) for $\alpha$ and find $\alpha'(\gamma')$.
-The resulting functions for perennial grassland are shown in Fig.~\ref{fig:f_temp_grassland}b) (for deciduous and coniferous trees refer Appendix~Figs.~\ref{fig:f_temp_spruce}b)--\ref{fig:f_temp_birch}b)).
+and define a variation $\gamma' = \gamma \cdot \eta$ with $\eta \in \{0.8, 1.2\}$ and invert Eq.~(\ref{eq:inverse_function}) for $\alpha$ to find $\alpha'(\gamma')$.
+The resulting functions for perennial grassland (PPFD0.8, PPFD1.2) are shown in Fig.~\ref{fig:f_temp_grassland}b) (for deciduous and coniferous trees refer Appendix~Figs.~\ref{fig:f_temp_spruce}b and \ref{fig:f_temp_birch}b).
 All derived parameters are tabulated in Table~\ref{tab:sensitivity_tests_light}.
@@
 
-As indicated above, a better acclimation to prevailing climate conditions will result in a higher average $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$.
-To compare the bespoke parameterizations, we propose the following metric.
+As indicated above, a better acclimation to prevailing climate conditions is characterized by a higher average $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$.
+To compare our bespoke parameterizations, we propose the following metric.
 We calculate an average relative stomatal conductance $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal}) with $f_\mathrm{SWP}=1$) at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time) for the whole record of meteorological data and apply mean and standard deviation.
-We presume that $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ around noon (highest light intensity) is a good proxy for \chem{CO_2} uptake efficiency (growth) which should be maximized.
+We presume that $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ around noon (highest light intensity) is a good proxy for \chem{CO_2} uptake efficiency (growth).
 In additional, a small deviation between noon and morning and a low standard deviation would indicate a higher robustness to variability in growing conditions.
-From the results shown in Fig.~\ref{fig:javis_func_opt_mean}, we identify \emph{subarctic}, PPFD0.8 as the best acclimation for all PFTs, based on the criteria.
-As expected due to the small adjustments, coniferous trees display the lowest deviation between the different types of acclimation, while the divergence for perennial grassland is substantial as a response to the stronger proposed temperature acclimation.
+Based on the criteria, we identify \emph{subarctic}--PPFD0.8 as the best parameterization for all PFTs (Fig.~\ref{fig:javis_func_opt_mean}).
+As expected due to the small adjustments, coniferous trees display the smallest differences between the different types of acclimation, while the differences for perennial grassland are substantial as a response to the stronger proposed temperature acclimation.
 

### 2021-09-28 VegOzone.tex
Restructure text

@@
 Start ($A_\text{start}$) and end ($A_\text{end}$) of GS for each PFT are given in units of \unit{doy} (Table~\ref{tab:sensitivity_tests_gs}) and estimated as follows.
-We use a MODIS (Aqua/Terra) retrieved net photosynthesis product \textbf{TODO: reference} for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine $A_\text{start}$ and $A_\text{end}$.
+We use a MODIS (Aqua/Terra) retrieved net photosynthesis product for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine $A_\text{start}$ and $A_\text{end}$.
 The net photosynthesis vs.
@@
 For both years, these dates coincide with the first snow-free day at the closest inland meteorological weather station at Øvre~Neiden (Sør-Varanger, NOR).
-For comparison, the default values ($A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$) indicating a substantially longer growing season.
-Due to a lack of quantitative field observation, we assume a latency period of $1\,\unit{month}$ after snow-melt for perennial grassland \citep[][Supplement~Fig.~S1, Rovaniemi]{FCR:Korhonen2018}.
+For comparison, the default values for central Europe ($A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$) indicating a substantially longer growing season.
+Due to a lack of quantitative field observation at Svanhovd, we assume a latency period of $1\,\unit{month}$ after snow-melt for perennial grassland.
+This assumption is supported by observations in Rovaniemi, Finland \citep[][Supplement~Fig.~S1]{FCR:Korhonen2018}.
 
@@
 Soil texture influences water availability and thus $\mathrm{POD_y}$ calculations in $\mathrm{DO_3SE}$ \citep{ACP:Bueker2012}.
-The soil at Svanhovd is characterized in an ICP Forest plot survey as gleyic cambisols according to the Food and Agriculture Organization of the United Nations (FAO) soil classification system and as eluviated dystric brunisol in the ICP Forest soil classification system (Volkmar Timmerman, pers.
-comm.) Within the $\mathrm{DO_3SE}$ soil parameterization, the sandy loam texture describes the upper $60\,\unit{cm}$, where the plant roots are found, of this soil type best.
+The soil at Svanhovd is characterized in an ICP Forest plot survey as gleyic cambisols according to the Food and Agriculture Organization of the United Nations (FAO) soil classification system and as eluviated dystric brunisol in the ICP Forest soil classification system (V.
+Timmerman, personal communication, 2021 May 10) Within the $\mathrm{DO_3SE}$ soil parameterization, the sandy loam texture describes the upper $60\,\unit{cm}$, where the plant roots are found, of this soil type best.
 
@@
 Heights were measured in 2004, when the stand was 90 years old.
-Mean tree height was $10.1\,\unit{m}$ and maximum height was $16.2\,\unit{m}$ \textbf{TODO:(Volkmar Timmerman, priv.
-comm.)}.
-Canopy height and leaf size are adjusted to local values in the models with bespoke growing season (GS bespoke) in the following.
+Mean tree height was $10.1\,\unit{m}$ and maximum height was $16.2\,\unit{m}$ (V.
+Timmerman, personal communication, 2021 May 10).
 
@@
 We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for the three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using the generic MM parameterization and our bespoke temperature (\emph{cold}, \emph{subarctic}) and light (PPFD0.8, PPFD1.2) response functions.
-We also locally adapted leaf size for deciduous trees and the tree height together with the bespoke GS.
+Canopy height and leaf size are adjusted to local values in the model simulations with bespoke growing season (GS besp.).
 The results displayed in Fig.~\ref{fig:pody_rel} allow for an assessment of systematic uncertainties arising from the use of generic parameterizations established from less climatically extreme conditions.
 For comparison, critical levels for reduction in biomass are shown as horizontal dashed lines \citep[deciduous forest $4\,\unit{\%}$, coniferous forest $2\,\unit{\%}$, and grasslands $10\,\unit{\%}$ biomass reduction,][]{ICP:MappingManual2017,ESPR:Hayes2021}.
-We identify that an increase in cold tolerance leads to higher ozone uptake in both probed years, while the usage of locally adapted GS causes a reduction.
+We identify that an increase in cold tolerance (MM $\rightarrow$ cold $\rightarrow$ subarctic) leads to higher ozone uptake in both probed years, while the usage of locally adapted GS causes a reduction.
 The earlier and longer GS in 2019 leads to increased ozone uptake compared to 2018 despite more episodes of elevated \chem{[O_3]} in 2018.

### 2021-09-28 VegOzone.tex
Restructure and include remark on grassland CL

@@
 Therefore, we shift $T\mathrm{min}$ and $T\mathrm{max}$ to colder temperatures and choose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000.
-In other words, we aim to increase the overlapping area between the $f_\mathrm{temp}$ and the temperature PDF of the GS.
+In other words, we aim to increase the overlapping area between $f_\mathrm{temp}$ and the temperature PDF of the GS.
 The most extreme of the two acclimation types is \emph{subarctic}.
@@
 The magnitude of these effects varies between PFTs as well as years, but the predicted ozone uptake for the bespoke parameterization is always larger than for the MM default parameterizations and of the same order of magnitude as the variability between the years studied here.
-Although our temperature acclimations are hypothetical, they suggest considerable underestimations of ozone risk for subarctic species when relying on generic MM parameterizations.
-We will look at the specific PFTs in more detail in the following.
+Our bespoke temperature parameterizations suggest considerable underestimations of ozone risk for subarctic species when relying on the generic MM parameterizations.
 
@@
 With MM GS, $\mathrm{POD_1}$ values were higher in 2018 than in 2019, throughout our analyses.
-However, taking the later start and overall shorter GS in 2018 into account (bespoke GS), this is reversed for the \emph{subarctic} acclimation type.
-This can be explained by favorable growing conditions in 2019.
-Low soil water potential reduced the $\mathrm{POD_1}$ values in 2018 in some cases when using the MM growing season, but did not or only slightly when using bespoke growing season.
-The temperature acclimation-dependent bias in ozone uptake deduced for bespoke GS ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
-The maximum systematic deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of the extent of stomatal opening at low light intensities amount to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.
+However, taking the bespoke GS into account, this is reversed for the \emph{subarctic} acclimation type.
+This can be explained by more favorable growing conditions in 2019.
+Low soil water potential reduced the $\mathrm{POD_1}$ values in 2018 in some cases when using the MM growing season, but did not or only slightly when using bespoke GS.
+The temperature acclimation-dependent difference in ozone uptake deduced for bespoke GS ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+The maximum deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of $f_\mathrm{light}$ amount to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.
 
 For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
-Coniferous trees' parameterization was already well acclimated to prevailing temperatures from the beginning and therefore the least biased of the generic MM parameterizations.
+Coniferous trees' parameterization was already well acclimated to prevailing temperatures from the start and therefore differs least.
 Taking the bespoke GS and canopy height into account, we find a higher ozone uptake in 2019 than in 2018 for both bespoke temperature response functions.
-The overall bias deduced from bespoke GS for subarctic and cold types with respect to MM range between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
-The maximum systematic deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of the extent of stomatal opening at low light intensities is medium and amounts to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.
+The overall difference deduced from bespoke GS for subarctic and cold types with respect to MM range between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+The maximum deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of $f_\mathrm{light}$ and amounts to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.
 
@@
 Perennial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant without bespoke GS.
-The overall bias deduced from bespoke GS for subarctic and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+The overall difference deduced from bespoke GS for subarctic and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
 The lower maximum temperature tolerance in the subarctic compared to the cold temperature parameterization probably caused the lower $\mathrm{POD_1}$ values calculated for 2018 in subarctic compared to cold climate, whereas this was not seen for 2019.
-The maximum systematic deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of the extent of stomatal opening at low light intensities amount to $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
+The maximum systematic deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of $f_\mathrm{light}$ amount to $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
 
@@
 
-We estimated the biomass reductions in accordance to the CLs in \citet{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme temperature acclimation (\emph{subarctic}) for each bespoke PFT.
+We estimated the biomass reductions in accordance to the CLs \citep{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme temperature parameterization (\emph{subarctic}) for each bespoke species or PFT.
 The estimates are listed in Table~\ref{tab:biomass_reduction}.
-The reported range in uncertainty corresponds to the maximum deviation defined by the $\pm 20\,\unit{\%}$ variation of the extent of stomatal opening at low light intensities with SWP=on.
-For deciduous trees and perennial grassland, total biomass was reduced in comparison with the default generic MM parameterizations for both years.
-\textbf{TODO: Discussion about CL for grassland!} Coniferous trees are the least affected and the magnitude of total biomass reduction is independent of the choice of the parameterization within given uncertainties.
+The reported range in uncertainty corresponds to the maximum deviation defined by the $\pm 20\,\unit{\%}$ variation of $f_\mathrm{light}$ with SWP=on.
+Total biomass of deciduous trees and perennial grassland was more strongly reduced in comparison with the default MM parameterizations for both years.
+
+The CL for perennial grassland is placed higher than for the other species.
+From an agro-economical point of view in our focus area, a loss of biomass of more than $10\,\unit{\%}$ can be more severe compared to more productive areas in central Europe and may need adjustment.
+
+Coniferous trees are the least affected and the magnitude of total biomass reduction is independent of the choice of the parameterization within given uncertainties.
 This may reflect that the temperature parameterization in the MM is almost never a limiting factor for stomatal opening.

### 2021-10-03 VegOzone.tex
Restructure and elaborate on conclusions

@@
 
-\textbf{TODO: Decide on terms acclimation vs parameterization etc} We have studied the effect of model acclimations of some common vegetation types to a subarctic climate on phytotoxic ozone uptake.
-The comparison between 2018 and 2019 conditions allowed us to consider these acclimations and their influence on vegetation sensitivity to \chem{O_3} in light of future changes of key environmental variables as may occur under climate change (e.g.
-increase of frequency and extent of heatwaves).
+We have studied the effect of bespoke parameterizations of common vegetation types (coniferous and deciduous trees, perennial grassland) to a subarctic climate on $\mathrm{POD_y}$.
+The comparison between meteorological conditions in 2018 and 2019 and their divergence from climatology allowed us to assess the influence of key environmental variables such as temperature, PPFD, and precipitation on vegetation susceptibility to \chem{O_3} damage in light of future changes as may occur under climate change.
 We found that conditions for ozone formation were more favorable in the 2018 growing season than in 2019, with 2018 being significantly warmer and less cloudy in spring and early summer.
 Accordingly, peak ozone concentrations occurred more frequently and at higher levels in 2018.
-This was particularly a result of the extended heatwave in spring and early summer and associated, extensive wildfires in Central Sweden (in particular in mid--end July).
-\textbf{TODO: Check explanation regarding significance.}
-
-We found that the length and timing of the growing season are crucial for risk assessment based on accumulated ozone flux metrics ($\mathrm{POD_y}$).
-Based on a MODIS photosynthesis product, we estimated the growing season for coniferous trees in 2018 to be at least one month shorter than in 2019.
-The found start of growing season in 2018 ($122\,\unit{doy}$) and 2019 ($106\,\unit{doy}$) are within given ranges by empirical observations \citep{TB:Kolari2007,IVL:Karlsson2018}.
-Based on gridded observational temperature and snow depth, we found a half a month shorter growing season for deciduous trees and perennial grassland in 2018 compared to 2019, respectively.
-With respect to ongoing climate change, a clear positive trend in length ($5.2\,\unit{d\,decade^{-1}}$) almost equally distributed between earlier start ($2.9\,\unit{days\,decade^{-1}}$) and later end ($2.3\,\unit{d\,decade^{-1}}$) could be deducted from the thermal growing season at Svanhovd (Appendix Fig.~\ref{fig:greening_season_change_Svanvik}).
-\textbf{TODO: Move S6 into appendix.}
-
-Ozone peak concentrations were observed more frequently in 2018 than in 2019.
-Nevertheless, \chem{POD_1} values were very similar for both years.
-This suggests that the overall length of the growing season is more important in risk assessment using flux-based metrics than episodes of peak concentrations.
-These peak \chem{[O_3]} often coincide with environmental conditions that limit ozone uptake the model.
-However, acute cellular damage caused by very high episodic ozone fluxes are not accounted for in $\mathrm{POD_y}$.
-That might explain why ozone damage was observed at Svanhovd to a larger extent in 2018 than in 2019.
-As pointed out by \citet{AE:Musselman2006}, a susceptibility to ozone depending on leaf age in the associated damage functions might give a different picture.
-Forbs and perennial grassland have been found to be more susceptible in the reproductive state \citep{EP:Bassin2004}.
-
-To estimate potential biases in the assessment of ozone-induced risk on Fennoscandic vegetation, we defined hypothetical temperature response functions representing different levels of assumed acclimation to cold climates.
-We assume northern Fennoscandia vegetation to be more tolerant towards cold temperatures and less tolerant to heat.
-Based on probability density functions (PDF) of observed light and temperature, we assume that an acclimation ought to increase the enclosed area between the respective response functions and the PDF.
-These hypothetical parameterisations of species or PFTs still need verification by experimental data.
-
-In summary, our key findings are:
-\begin{enumerate}
-   \itemsep0pt
- \item the bespoke parameterizations for subarctic species are important (especially concerning light and temperature and the climate effect on the growing season) to be able to assess \chem{O_3} uptake,
- \item the parameterizations defined in the mapping manual for European regional risk assessment appear to not adequately capture the physiological responses of subarctic physiology that are important determinants of \chem{[O_3]} sensitivity and are likely to substantially underestimate risk, and
- \item that the climatic conditions promoting ozone damage (particularly extended heatwaves and the timing of the growing season concerning the ozone spring peak) and which are likely to become more frequent under climate change are likely to further enhance the sensitivity of subarctic vegetation to ozone.
-\end{enumerate}
-
-However, the decline of the ozone spring peak is partly caused by dry deposition of ozone on vegetation.
-It, therefore, remains unclear whether an earlier start of the growing season has the potential to reduce the ozone spring peak concentration compared to present-day and hence reduce the damage risk of ozone in the future.
-We have not assessed the effect of the plants' response to VPD here, but our climatological data show that conditions at Svanhovd are shifting towards a more VPD limited regime.
-
-Based on the methods described in \citet{ICP:MappingManual2017}, we estimated biomass reductions due to ozone uptake for both 2018 and 2019 and found significant reductions for deciduous trees ($15.5-17.4\,\unit{\%}$) and perennial grassland (above ground, $12.4-13.7\,\unit{\%}$) with respect to estimates based on $\mathrm{POD_1}$ from default parameterizations.
-For coniferous trees, the adjusted parameterizations are comparable with the default parameterization estimates and are contained within the range of systematic uncertainties deduced from our bespoke response function ($2.5-3\,\unit{\%}$).
-These results have to be treated with care because the biomass reduction functions were not established based on subarctic species.
-
-At the same time, intra-species variability even at the same location is typically non-negligible \citep{EP:Bassin2004,Amb:Girgzdiene2009}.
-This may explain contradictory results regarding the ozone sensitivity of natural vegetation in northern Europe.
-While \citet{FS:Subramanian2014} reported a modeled reduction of net primary production in coniferous ($4.3-15.5\,\unit{\%}$) and birch ($1.4-4.3\,\unit{\%}$) under elevated ozone in Sweden, Lithuanian forest deciduous trees were observed to be more susceptible to visible damage by ozone than Norway spruce \citep{Amb:Girgzdiene2009}.
-Open-top chamber (OTC) experiments performed in northern Finland where Scots pine and downy birch were exposed to elevated ozone concentrations showed biomass reductions of both downy birch and mountain birch seedlings after one season of exposure \citep{Amb:Manninen2009}.
-The leaves of the mountain birch in the high ozone treatments showed necrotic stipples as early as June.
-As mountain birch very rapidly develops during spring and early summer, compensation for damaged leaves is difficult for this species.
-In contrast to mountain birches, northern Scots pine of local origin only showed invisible needle-level \chem{O_3} impacts after three years of exposure.
-There is no field data comparing the ozone sensitivity of birches and Scots pine from northern Europe.
-In our study, we do not explicitly account for this natural distribution of individual plant traits, but the chosen hypothetical, bespoke response functions might reflect this diversity within a given species or PFT.
-The resilience of northern Fennoscandia plant communities to climate change and predictions in this regard strongly depends on both the existing range of acclimations as well as the actual type of acclimation.
-
-Cumulative ozone uptake has been reported to reduce stomatal conductance significantly and independently from photosynthesis \citep[e.g.]{BGS:Lombardozzi2012}, with photosynthesis being more strongly affected.
-This might be due to competing effects, as an ozone-induced sluggishness \citep{SR:Hoshika2015} renders stomata unable to fully close and leads to higher stomatal conductance than expected and thus higher ozone uptake as well as transpiration.
-At present, the MM version of the $\mathrm{DO_3SE}$ model does not account for any ozone-inflicted reduction of stomatal conductance throughout the growing season and is not coupled to photosynthesis.
-Hence, the presented $\mathrm{POD_y}$ may be regarded as an estimated upper limit.\\
-
-Beyond the risk assessment of ozone-induced damage, our hypothetical bespoke temperature response functions, once verified by observations, can have important implications on land-surface modeling in global models, where problems in productivity of species especially in the Arctic regions occur due to PFTs which are not suitable for the climate.
+This was particularly a result of the extended heatwave in spring and early summer and associated, extensive wildfires in Central Sweden.
+However, ozone concentrations during the growing season were not significantly enhanced and \chem{POD_1} values were very similar for both years.
+
+Based on a MODIS photosynthesis product \citep{MODIS_PSN}, we estimated the growing season for coniferous trees in 2018 to be at least one month shorter than in 2019.
+The deduced start of growing season in 2018 ($122\,\unit{doy}$) and 2019 ($106\,\unit{doy}$) are within given ranges by empirical observations \citep{TB:Kolari2007,IVL:Karlsson2018}.
+We determined the start of the growing season for deciduous trees based on gridded observational temperature and for perennial grassland on snow depth data.
+For perennial grassland, we assumed a one month latency period after snow melt which is supported by observational data from Finland \citep{FCR:Korhonen2018}.
+With respect to ongoing climate change, a clear positive trend emerged in length ($5.2\,\unit{d\,decade^{-1}}$) of the growing season which is almost equally distributed between earlier start ($2.9\,\unit{days\,decade^{-1}}$) and later end ($2.3\,\unit{d\,decade^{-1}}$) (Appendix Fig.~\ref{fig:greening_season_change_Svanvik}).
+
+This suggests that the overall length of the growing season is more crucial in risk assessment when using flux-based metrics than episodes of peak concentrations as these peak \chem{[O_3]} often coincide with environmental conditions that limit ozone uptake in the $\mathrm{DO_3SE}$ model.
+Visible damage caused by peak ozone uptake and a dependence of ozone sensibility on phenology are not accounted for in $\mathrm{POD_y}$.
+But open-top chamber (OTC) experiments performed in northern Finland where Scots pine and downy birch seedlings were exposed to elevated ozone concentrations showed a reduction in biomass and reproduction with visible damage explicitly attributed to peak \chem{O_3} concentrations and fast phenological development at high growth rate \citep{Amb:Manninen2009}.
+In particular, forbs and perennial grassland have been found to be more susceptible to ozone-induced damage in the reproductive state \citep{EP:Bassin2004}.
+That might explain why we observed ozone damage to a larger extent in 2018 than in 2019.
+An associated damage function to ozone depending on leaf age as proposed by \citet{AE:Musselman2006} or phenological state might improve the predictions.
+
+At the same time, intra-species variability even at the same location is typically non-negligible \citep{EP:Bassin2004,Amb:Girgzdiene2009} but not accounted for in the models.
+This may also explain some contradictory results regarding ozone sensitivity of natural vegetation in northern Europe.
+\citet{FS:Subramanian2014} reported a higher, modeled reduction of net primary production in conifers ($4.3-15.5\,\unit{\%}$) than birch ($1.4-4.3\,\unit{\%}$) under elevated ozone in Sweden, while \citet{Amb:Girgzdiene2009} observed more visible damage on deciduous trees than on Norway spruce in Lithuania.
+But visible damage and biomass reduction may not occur at the same rate and hence may not be interchangeable.
+In our study, we do not explicitly account for this natural distribution of individual plant traits, but our bespoke response functions reflect the diversity within a given species.
+In particular, we assumed two tolerance levels to cold temperatures leading to a higher efficiency in photosynthesis at cool temperatures than for central European species.
+To this end, we computed a PDF from observed temperature at the NIBIO environment centre Svanhovd (May--September, 1992--2000).
+We adjusted the parameters in the temperature response function of stomatal conductance to increase the enclosed area with the PDF.
+Assuming a proportionality between $g_\mathrm{sto}$ and $A_\mathrm{n}$, this is equivalent to optimizing carbon uptake.
+Similarly, we deduced a PPFD PDF and varied the light response function to test a higher or lower extent of stomatal opening at low light intensities.
+These parameterizations are hypothetical and need verification by experimental data.
+We found that soil water potential under 2018/19 meteorological conditions was negligible when considering the bespoke growing season.
+We have not assessed the effect of the plants' response to VPD in detail, but our climatological data indicated that conditions at Svanhovd are shifting towards a more VPD limited regime.
+Our key findings are: (1) Our bespoke parameterizations for subarctic species are better suited to describe vegetation adapted to subarctic conditions for instance allowing for stomata opening during a larger part of the growing season, (2) the parameterizations defined in the mapping manual for European regional risk assessment appear to not adequately capture the physiological responses of subarctic vegetation that are important determinants of \chem{[O_3]} susceptibility and are likely to underestimate ozone-damage risk, and (3) climatic conditions that are promoting a longer growing season are likely to increase the ozone-damage risk in flux-based metrics.
+
+In this regard, vegetation might become more exposed to higher ozone concentrations in early phenological stages due to an overlap with ozone spring peak \citep{EP:Karlsson2007}.
+However, the decline of this ozone spring peak is partly caused by uptake of vegetation.
+It, therefore, remains unclear whether an earlier start of the growing season will increase the exposure of vegetation to ozone or lead to a compensation due to an earlier decline of the peak ozone in the future.
+\citet{ESPR:Hayes2021} pointed out that the highest susceptibility of vegetation to future ozone variability is likely to occur in summer when vegetation is most productive.
+The actual resilience of northern Fennoscandia plant communities to climate change and its predictions strongly depend on both the range of existing acclimations as well as the actual type.
+
+Following the methods described in \citet{ICP:MappingManual2017}, we estimated biomass reductions due to ozone uptake for both 2018 and 2019 and found substantial reductions for deciduous trees ($15.5-17.4\,\unit{\%}$) and perennial grassland (above ground, $12.4-13.7\,\unit{\%}$) in both years.
+These reductions are higher than for the default parameterizations, $11.2\,\unit{\%}$ (deciduous) and $9.6\,\unit{\%}$ (perennial grassland).
+For coniferous trees, the estimated reduction in biomass differs insignificantly between our bespoke parameterizations ($2.5-3\,\unit{\%}$) and the default ($2.31\,\unit{\%}$).
+This indicates a higher modeled robustness to ozone damage in coniferous trees.
+All these results have to be treated with care because the biomass reduction functions were established based on species in less extreme climates.
+In particular, CLs for deciduous and coniferous trees were breach in all years and fora ll parameterizations.
+For perennial grassland, the CL was not breached if a bespoke GS was assumed, but the mapping manual defined CL of $10\,\unit{\%}$ loss in biomass might not be transferable to subarctic regions with considerable constrains on productivity leading to more severe economical consequences.
+
+Beyond the risk assessment of ozone-induced damage, our bespoke temperature response functions, once verified by observations, can have important implications on land-surface modeling in global models, where problems in productivity of species especially in the Arctic regions occur due to PFTs which are not suitable for the climate.
 We pointed out, with respect to default perennial grassland parameterization, that these perennial grasslands did not show any substantial stomatal conductance in 2019 characterized as a normal year.
-In terms of GPP this would mean unrealistically low carbon sequestration.
+In terms of GPP this would mean unrealistically low carbon uptake.
 Automation of the here proposed PDF-based acclimation using machine learning techniques could overcome these issues in the future.

### 2021-10-04 VegOzone.tex
Elaborate on title
Refine abstract.

@@
 
-%\title{The 2018 heatwave and its implications on ozone induced damage on vegetation in a subarctic climate}
-\title{Characterizing subarctic European vegetation for land surface modeling of ozone pollution and climate risk}
+\title{Characterizing subarctic European vegetation's response on key environmental variables for land surface modeling of ozone pollution and risk assessment}
 
@@
 To date, only a few studies have been conducted to assess the response of subarctic vegetation to variations in climate and air pollution.
-This study looks to fill this knowledge gap by examining essential climate variables, in particular \chem{[O_3]}, over the past few decades.
+This study looks to fill this knowledge gap by examining essential climate variables, in particular ozone, over the past few decades.
 These data are used to assess the extent to which two recent years (2018 and 2019) deviate from climatic and \chem{[O_3]} norms and how these potentially more frequent future deviations may influence ozone damage to subarctic vegetation.
-We find that 2018 was an anomalously warm and bright year (in particular in spring and early summer) with higher than average \chem{[O_3]} in April/May and a high number of episodic peak \chem{[O_3]} above $40\,\unit{ppb}$ in June--August, the latter in part attributable to forest fires in the Northern Hemisphere as well as warmer and sunnier conditions.
-We apply the integrated flux-metric Phytotoxic Ozone Dose (POD) to determine ozone risk and damage to vegetation as a function of \chem{[O_3]}, climate variables, and species-specific physiology.
+We find that 2018 was an anomalously warm and bright year (in particular in spring and early summer) with higher than average \chem{[O_3]} in April/May and a high number of episodic peak values in ozone volume mixing ratio (VMR) above $40\,\unit{ppb}$ in June--August, the latter in part attributable to forest fires in the Northern Hemisphere as well as warmer and sunnier conditions.
+We apply the integrated flux-metric Phytotoxic Ozone Dose (POD) to determine ozone risk and damage to vegetation as a function of \chem{[O_3]}, meteorological variables, and species-specific physiology.
 Our study, suggests that assessments that use generic parameterizations are likely to underestimate the risk of ozone damage in this region.
-We find that bespoke parameterizations of plant functional types (PFTs) for subarctic vegetation bio-types result in a ozone-induced biomass loss of $2.5$ to $17.4\,\unit{\%}$.
+We find that bespoke parameterizations of plant functional types (PFTs) for subarctic vegetation bio-types result in an ozone-induced biomass loss of $2.5$ to $17.4\,\unit{\%}$.
+For some species this loss is up to $6\,\unit{\%}$ larger than projected from generic parameterizations.
 Efforts should be targeted towards accurately defining subarctic ecotypes' physiological response to essential climate variables to improve regional and global scale biogeochemical cycling under current and future climates.

### 2021-10-04 VegOzone.tex
Elaborate on phrasing

@@
 We find that 2018 was an anomalously warm and bright year (in particular in spring and early summer) with higher than average \chem{[O_3]} in April/May and a high number of episodic peak values in ozone volume mixing ratio (VMR) above $40\,\unit{ppb}$ in June--August, the latter in part attributable to forest fires in the Northern Hemisphere as well as warmer and sunnier conditions.
-We apply the integrated flux-metric Phytotoxic Ozone Dose (POD) to determine ozone risk and damage to vegetation as a function of \chem{[O_3]}, meteorological variables, and species-specific physiology.
+We apply the integrated flux-metric Phytotoxic Ozone Dose (POD) to determine ozone risk and damage to vegetation as a function of \chem{[O_3]}, environmental factors, and species-specific physiology.
 Our study, suggests that assessments that use generic parameterizations are likely to underestimate the risk of ozone damage in this region.
@@
 Ground-level ozone (\chem{O_3}) is a highly toxic pollutant known to cause damage to human health \citep{WHO2008} and to a variety of ecosystems around the World \citep{PT:Emberson2020}.
-Ozone is a cause of visible injury, photosynthetic damage, early senescence as well as programmed cell death of plants \citep{PCE:Kangasjarvi2005}.
-Annual global yield losses of four major crops (wheat, rice, maize, and soybean) of $3-15\,\unit{\%}$ \citep{PJ:Ainsworth2017} and a suggested loss in primary production in forestry of $7\,\unit{\%}$ \citep{GCB:Wittig2009,EP:Matyssek2012} has been attributed to ozone.
+Ozone is a cause of reduction of photosynthesis, visible foliar injury, early senescence and programmed cell death of plants \citep{PCE:Kangasjarvi2005}.
+Annual global yield losses of four major crops (wheat, rice, maize, and soybean) of $3-15\,\unit{\%}$ have been reported \citep{PJ:Ainsworth2017} and a suggested loss in primary production in forestry of $7\,\unit{\%}$ \citep{GCB:Wittig2009,EP:Matyssek2012} has been attributed to ozone.
 The crop yield loss studies indicate a threat to food security in rapidly developing countries, e.g., in East and South-East Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018}, where ozone concentrations related to enhanced pollutant emissions from the transport, residential, power generation, and industrial sectors have increased in the past few decades.
-Even though long-term (pre-1950) observations of ground-level ozone concentrations are scarce; evidence from trend analysis of those few sites suggests that since the industrial revolution, background concentrations in the northern hemisphere have at least doubled and continued to increase \citep[Chapter 2]{IPCC2013}.
-Recent trend analyses for Europe indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric ozone started to level off or even decline \citep{ESA:Cooper2014, ACP:Wespes2018,ESA:Gaudel2018}.
+Even though long-term (pre-1950) observations of ground-level ozone concentrations are scarce; evidence from trend analysis of those few sites suggests that since the industrial revolution, background concentrations in the northern hemisphere have at least doubled and continued to increase \citep{IPCC2013}.
+Recent trend analyses for Europe indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric background ozone started to level off or even decline \citep{ESA:Cooper2014, ACP:Wespes2018,ESA:Gaudel2018}.
 This can be attributed to a successful implementation of air quality regulations coupled with economic restructuring, reducing the number of episodes of peak concentrations especially in summertime \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
@@
 
-Complex photochemical cycles involving precursor gases such as carbon monoxide (\chem{CO}) and hydrocarbons known as volatile organic compounds (VOCs) in the presence of nitrogen oxides (\chem{NO_x}) lead to the formation of \chem{O_3} at ground-level.
+The complex photochemical cycles leading to the formation of \chem{O_3} at ground-level involve precursor gases such as carbon monoxide (\chem{CO}) and hydrocarbons known as volatile organic compounds (VOCs) in the presence of nitrogen oxides (\chem{NO_x}).
 Despite a relatively short average tropospheric lifetime of approximately $22\,\unit{days}$ \citep{JGR:Stevenson2005,ACP:Young2013}, ozone and is precursors are subject to advection over long distances.
 Episodes of high ozone concentrations in the absence of local pollution sources can therefore often be attributed to the long-range transport of both ozone and its precursors.
-In northern Fennoscandia, episodes of enhanced ozone concentrations in 2003 and 2006 have been traced back to ozone precursor emissions related to forest fires in southern and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}.
+In northern Fennoscandia, episodes of elevated ozone concentrations in 2003 and 2006 have been traced back to ozone precursor emissions related to forest fires in southern and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}.
 Besides VOCs from anthropogenic sources, hydrocarbons are also emitted by vegetation itself in the form of terpenes and monoterpenes, so-called biogenic volatile substances (BVOCs).
@@
 The spring peak is more pronounced in northern Fennoscandia than in more southern locations but shows a high interannual variability \citep{AB:Klingberg2009, BER:Klingberg2019}.
-A cause for this is the complex interplay between atmospheric dynamics and chemistry \citep{AE:Laurila1996,BER:Hatakka2003} as well as anthropogenic and biogenic activity \citep{AE:Rummukainen1996,AE:Simpson2002,QJRMS:Galbally2007,NGS:Schnell2009}.
+A cause for this is the complex interplay between atmospheric chemistry and dynamics \citep{AE:Laurila1996,BER:Hatakka2003} as well as anthropogenic and biogenic activity \citep{AE:Rummukainen1996,AE:Simpson2002,QJRMS:Galbally2007,NGS:Schnell2009}.
 \citet{ACP:Andersson2017} showed in a modeling study focused on Fennoscandia that the variability of ozone concentrations in winter can be attributed mainly to changes in atmospheric background (transport) of ozone, while summertime abundance is mostly affected by emissions of precursors in the rest of Europe.
@@
 Results from open-top chamber (OTC) experiments on downy birch (\emph{Betula pubescens}) and mountain birch (\emph{Betula pubescens toruosa}), native to subarctic regions, indicated reductions in both biomass, in root:shoot ratio, and visible leaf damage under elevated ozone treatment ($\left<\chem{[O_3]}\right> = 36-54\,\unit{ppb}$) \citep{Amb:Manninen2009}.
-Although Scots pine (\emph{Pinus sylvestris}) is considered to be more ozone tolerant due to an absence of visible injuries \citep{Amb:Girgzdiene2009}, \citet{Amb:Manninen2009} found chlorophyll:carotenoid ratio and polyamines reductions under elevated ozone concentrations indicating susceptibility to ozone also in these species.\\
+Though, Scots pine (\emph{Pinus sylvestris}) is considered to be more ozone tolerant due to an absence of visible injuries \citep{Amb:Girgzdiene2009}, \citet{Amb:Manninen2009} found chlorophyll:carotenoid ratio and polyamines reductions under elevated ozone concentrations indicating susceptibility to ozone also in these species.\\
 
@@
 
-To estimate the flux-based metric, also referred to as the Phyto-toxic Ozone Dose over a threshold y ($\mathrm{POD_y}$), an estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf and environmental factors affecting stomatal conductance.
+To estimate the flux-based metric, also referred to as the Phytotoxic Ozone Dose over a threshold y ($\mathrm{POD_y}$), an estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf and environmental factors affecting stomatal conductance.
 $\mathrm{POD_y}$ is then calculated according to:
@@
 when global radiation is above $50\,\unit{W\,m^{-2}}$).
-$\mathrm{POD_y}$ is given in \unit{mmol\,m^{-2}} over the growing season.
+$\mathrm{POD_y}$ is given in \unit{mmol\,m^{-2}} integrated over the growing season.
 
@@
 This study demonstrated the improved ability of flux-based metrics (in comparison to exposure-based metrics) to identify the geographical distribution of the risk of ozone damage and also showed that the ozone damage could extend into more northerly regions of Europe.
-We conducted a similar bio-monitoring study at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd to assess whether local ozone concentrations were capable of inducing damage to vegetation growing under northern Fennoscandia conditions.
+For our study, we conducted a similar bio-monitoring study at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd to assess whether local ozone concentrations were capable of inducing damage to vegetation growing under northern Fennoscandia conditions.
 In 2018, we indeed observed ozone damage on clover and tobacco.
 In contrast, no such damage was found on the clovers in 2019.
+This indicates that environmental factors in 2018 differed from 2019.
 
@@
 An unusually weak and northward shifted jet stream allowed for a persistent high-pressure system above northern Europe, including Fennoscandia which blocked the low-pressure systems for several consecutive months.
-In the period May -- July, southern Norway had temperatures $4\,\unit{^\circ C}$ above normal.
-While southern Norway had only about $60\,\unit{\%}$ of normal precipitation, northern Norway as a whole had close to normal precipitation, but with local variations.
+In the period May -- July, southern Norway experienced temperatures $4\,\unit{^\circ C}$ above normal and saw only about $60\,\unit{\%}$ of normal precipitation.Northern Norway as a whole had close to normal precipitation, but with local variations.
 Thermal stress on vegetation was exceptional not only in large parts of Fennoscandia but also in much of Europe, where the influence of the high-pressure system extended even over five months (April/May, July--September).
 These conditions gave rise to massive forest fires in different parts of Europe and thus an increase in ozone precursors.
-Boreal wildfires emit in addition to \chem{CO_2} also \chem{CO} ($\chem{[CO]}/\chem{[CO_2]} \propto 6-13\,\unit{\%}$) and VOCs ($\chem{[VOCs]}/\chem{[CO_2]} \propto 0.5-1.5\,\unit{\%}$) \citep{AE:Cofer1990}.
 A total of 2079 forest fires were registered in Norway in 2018, twice as many as in the preceding years 2016/17 \citep{DSB2019}, last accessed April 2020).
 In Sweden, about 500 fires had been reported (five times more than in a usual summer), and an estimated total of $25000\,\unit{hectare}$ burned down in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) \citep{SOU2019}.
+Boreal wildfires are found to emit ozone precursors \chem{CO} ($\chem{[CO]}/\chem{[CO_2]} \propto 6-13\,\unit{\%}$) and VOCs ($\chem{[VOCs]}/\chem{[CO_2]} \propto 0.5-1.5\,\unit{\%}$) \citep{AE:Cofer1990}.
 Coincident peak \chem{[O_3]} is found in ozone monitoring data (Fig.~\ref{fig:data_svanvik_2018_2019}) in July.
-It is very likely that elevated ozone in Europe was promoted by a combination of various factors such as wildfires, reduced cloud cover (increased solar radiation), reduced dry deposition and turbulent mixing due to the stagnant weather conditions, and increased BVOC emissions like during the 2003 drought period,\citep{JGR:Solberg2018}.
-Therefore conditions at Svanhovd for the growing seasons 2018/19 shall serve as a reference for present and probable future conditions in northern Fennoscandia in our study.\\
+It is very likely that elevated ozone, like during the 2003 drought period \citep{JGR:Solberg2018}, was promoted by a combination of various factors such as wildfires, reduced cloud cover (increased solar radiation), reduced dry deposition and turbulent mixing due to the stagnant weather conditions, and increased BVOC emissions.
+As such events may become more frequent in the future, conditions at Svanhovd for the growing seasons 2018/19 shall serve as a reference for present and probable future conditions in northern Fennoscandia.\\
 
@@
 It is important to understand the implications of using these more generic parameterizations for more extreme climate regions.
-We explore how risk estimates might be over or underestimated using generic vs bespoke subarctic parameterization.
-We also explore the effects of the interannual variability of \chem{[O_3]} and essential climate variables (sensu World Meteorological Organization) on future risk of vegetation damage.
+In our study, we explore how risk might be over- or underestimated using generic vs bespoke subarctic parameterization.
+We also investigate the effects of the interannual variability of essential climate variables (sensu World Meteorological Organization) on future risk of vegetation damage.
 This allows an understanding of how the threat from ozone may change in the future and which are the key aspects of plant physiology that might determine the potential risk.
 Hence, we determine which parameters require special attention to define accurately for pollution impact assessment and more generally for describing gas exchange response to a changing climate that will influence biochemical cycling.
-This might also be one key to solve long-standing issues of earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface models \citep{GMD:Poulter2015,JAMES:Lawrence2019}.\\
-
-The observation site at NIBIO Environment Centre Svanhovd and all related data are presented in Section~\ref{sec:data}.
-\textbf{TODO: Rewrite!} In Section~\ref{sec:stats}, we derive multi-annual means for ground-level ozone concentration and key environmental variables (temperature, precipitation, and global irradiance) based on in situ observations and evaluate the deviation of 2018/19 from the norm.
-We use the $\mathrm{DO_3SE}$ model to estimate the ozone uptake by natural and semi-natural vegetation to investigate the main environmental drivers for projected ozone damage risk in 2018 (Section~\ref{sec:do3se}).
-To this end, we derive bespoke parameterizations for subarctic species to assess systematic uncertainties arising from the use of generic parameterizations established from less climatically extreme conditions.
-In Section~\ref{sec:conc}, we will summarize our results and discuss how this research can support improvements in future assessments of pollution risk and more generally biogeochemical cycling.
+This might also be one key to solve long-standing issues of earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface models \citep{GMD:Poulter2015,JAMES:Lawrence2019}.
+
+The observation site at NIBIO Environment Centre Svanhovd and available observational data are presented in Section~\ref{sec:data}.
+From these data, we derive multi-annual means of essential climate variables (temperature, precipitation, global irradiance, and ozone) and evaluate the deviation of 2018/19 from this norm (Section~\ref{sec:stats}).
+In Section~\ref{sec:do3se}, we use the $\mathrm{DO_3SE}$ model to estimate the ozone uptake by natural and semi-natural vegetation to investigate the main environmental drivers for projected ozone damage risk in 2018.
+To this end, we derive bespoke parameterizations for subarctic species to assess over- and underestimations arising from the use of generic parameterizations established from less climatically extreme conditions.
+We will summarize our results in Section~\ref{sec:conc} and discuss how this research can support improvements in future assessments of pollution risk and more generally biogeochemical cycling.
 
@@
 
-In our assessment of primary drivers of ozone damage risk on vegetation in the subarctic, we focus on conditions during the 2018/19 growing season at the NIBIO operated Environment Centre Svanhovd.
-We also set these into a climatological context.
+In our assessment of primary drivers of ozone damage risk on vegetation in the subarctic, we focus on conditions at the NIBIO operated Environment Centre Svanhovd (Sør-Varanger, NOR).
+The NIBIO centre Svanhovd focuses on the research of the special environment of the region, but also comprises long-term observations of various agrometeorological variables, i.e.
+$2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths.
+We investigate the 2018/19 growing season in a climatological context.
 To this end, we collected atmospheric and ozone monitoring data taken at Svanhovd over the past $35\,\unit{years}$.
@@
 
-The NIBIO centre Svanhovd focuses on the research of the special environment of the region, but also comprises long-term observations of various agrometeorological variables, i.e.
-$2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths.
-
-All measured atmospheric variables including ozone for the 2018/19 growing season were obtained through \href{luftkvalitet.no}{luftkvalitet.no} operated by \citet{NILU_AIRQ}.
+All measured essential climate variables including ozone for the 2018/19 growing season were obtained through \href{luftkvalitet.no}{luftkvalitet.no} operated by \citet{NILU_AIRQ} (last accessed May 2020).
 The ozone monitoring data have been also added to the EBAS database operated by \citet{NILU_EBAS} and are thus openly accessible.
@@
 Especially, ozone monitors did not undergo regular re-calibration.
-This probably lead to drifts in the observed data and may impose a false trend.
+This probably led to drifts in the observed data and may impose a false trend which shall not be of concern for us.
 The long-term data sets of agrometeorological variables including temperature and precipitation are available from September 1992 to the present day \citep[LandbruksMeteorologiske Tjeneste][note the station name here is Pasvik]{LMT_NIBIO}.
@@
 To relate local ozone concentration \chem{[O_3]} with visible damage on ozone sensitive plants, both an ozone bio-monitor (referred to as ozone garden) and a conventional ozone monitor using the existing atmospheric monitoring infrastructure had been installed during the 2018/19 growing season at the NIBIO centre.
-We will report \chem{[O_3]} as volume mixing ratios (VMR) in \unit{ppb}.
+We report \chem{[O_3]} as volume mixing ratios (VMR) in \unit{ppb} for easier comparison with exposure-based risk assessment.
 The locations of the atmospheric monitoring site and the ozone garden are marked in the aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}a.
@@
 As shown in Fig.~\ref{fig:svanhovd_research_station}b, the plants had to be protected from herbivores with a wire-mesh fence.
+In 2018, we qualitatively observed visible ozone damage on semi-natural vegetation (clover) and crops (tobacco).
+In contrast, no such visible damage was found on the clovers in 2019, although it was observed on the sensitive tobacco cultivar.
+This indicates that the vegetation was more affected by ozone in 2018 than in 2019.
 
@@
 However, due to the above-mentioned data acquisition problems, we missed most of the corresponding ozone data for this event and most likely also the peak \chem{[O_3]}.
+A method for gap-filling these data has been presented in \citet{ACPD:Falk2021}.
 In contrast, \chem{[O_3]} only rose 18 times above the threshold of $40\,\unit{ppb}$ during the summer of 2019.
-A reconstruction method for gap-filling the data has been presented in \citet{ACPD:Falk2021}.
 
@@
 In 2018, spring temperature regularly rose above freezing only in May, while in 2019 this occurred already early in March/April.
-More rain events with accumulated daily precipitation ($\sum_d \mathrm{Precip}$) above $10\,\unit{mm}$ occurred in the summer of 2018 compared to 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}c.
-Qualitatively, global irradiance ($Q_0$) displayed in Fig.~\ref{fig:data_svanvik_2018_2019}d was higher in May and July 2018 compared to 2019, while June 2019 showed higher irradiance than 2018.
+More rain events with accumulated daily precipitation ($\sum_d \mathrm{Precip}$) above $10\,\unit{mm}$ occurred in the summer of 2018 compared to 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}c).
+Qualitatively, global irradiance ($Q_0$) displayed in Fig.~\ref{fig:data_svanvik_2018_2019}d was higher in May and July 2018 compared to 2019, while June 2019 showed higher irradiance than in 2018.
 A high global irradiance in most cases is the result of a low cloud fraction.
 In both years, the maximum recorded $Q_0$ was $750\,\unit{W\,m^{-2}}$ and reached in June.
-
-In 2018, we qualitatively observed visible ozone damage on semi-natural vegetation (clover) and crops (tobacco).
-In contrast, no such visible damage was found on the clovers in 2019, although it was observed on the sensitive tobacco cultivar.
-This indicates that the vegetation was more affected by ozone in 2018 than in 2019.
 
@@
 \label{sec:stats}
-In this section, we assess the climate conditions.
-To this end, we compute multi-annual mean (referred to as climatology in the following) of environmental key variables (temperature, precipitation, global irradiance) and ground-level \chem{[O_3]} at Svanhovd.
-For the growing seasons 2018/19, we evaluate the statistical significance of divergences from the norm in these variables (referred to as anomalies).
-We performed a bias correction accounting for the increase in tropospheric background \chem{[O_3]} and cross-calibrated the derived climatology for Svanvik with long-term observations of \chem{[O_3]} at other sites in northern Fennoscandia as described in \citet{ACPD:Falk2021}.
+In this section, we assess the climate conditions at Svanhovd.
+To this end, we compute a multi-annual mean (referred to as climatology) for each environmental key variable (temperature, precipitation, global irradiance) and ground-level ozone.
+We evaluate the statistical significance of divergences from the norm in these variables (referred to as anomalies) in 2018/19.
+Concerning ozone, we performed a bias correction accounting for an increase in tropospheric background \chem{[O_3]} and cross-calibrated the climatology with long-term observations at other sites in northern Fennoscandia as described in \citet{ACPD:Falk2021}.
 
@@
 
-The location of Svanvik ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m}$) suggests a subarctic climate.
-The derived climate diagram based on climatological data (1992--2012) (Fig.~\ref{fig:climatediagram}) supports this very well.
+The location of the NIBIO environment centre Svanhovd ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m}$) suggests a subarctic climate.
+The 1992--2012 climatologies of temperature and precipitation are shown in a combined climate diagram (Fig.~\ref{fig:climatediagram}).
 Monthly averaged $2\,\unit{m}$ temperatures (red line) are displayed with a standard error of the mean (SE) as error band.
@@
 The highest monthly average temperature is $(13.1\pm 1.1)\,\unit{^\circ C}$ in July and the lowest $(-12.8\pm 2.0)\,\unit{^\circ C}$ in February.
-In the period 1992--2012, the coldest measured temperature was $-45.2\,\unit{^\circ C}$ observed January 27, 1999, while the highest temperature ($29.4\,\unit{^\circ C}$) occurred July 16 the same year.
+The coldest measured temperature was $-45.2\,\unit{^\circ C}$ in January 27, 1999, while the highest temperature ($29.4\,\unit{^\circ C}$) occurred in July 16 the same year.
 
@@
 The average annual accumulated precipitation ($\sum_m\mathrm{Precip}$) given with standard error of mean is $(383\pm 86)\,\unit{mm}$.
-Precipitation shows a large interannual variability.
-
-The derived climatology for global irradiance ($Q_0$) at Svanvik (Appendix Fig.~\ref{fig:global_rad_clim}a)) approaches zero in December/January (polar night) and reaches its maximum in June/July (midnight sun conditions).
+Precipitation shows a considerable interannual variability.
+
+The climatology for global irradiance ($Q_0$, not shown) approaches zero in December/January (polar night) and reaches its maximum in June/July (midnight sun conditions).
 The highest observed averaged daily maximum ($\left<Q_0^\mathrm{max}\right>$) amounts to $800\,\unit{W\,m^{-2}}$, while the maximum of the averaged daily mean ($\left<Q_0^\mathrm{mean}\right>$) is $450\,\unit{W\,m^{-2}}$ (regularly reached in May--July).
@@
 %Ozone
-The derived ozone climatology (\chem{\left<[O_3]\right>}) for Svanvik is shown in Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}.
+The derived ozone climatology (\chem{\left<[O_3]\right>}) (1986--1996) is shown in Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}.
 Darker colors indicate higher probability to observe these values.
 On top of the density distributions, a $10\,\unit{days}$ average of daily mean (\chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}}) is displayed together with $1 \sigma$ uncertainties and SE, respectively.
-A spline was fitted through the data for comparison.
-The \chem{[O_3]} density distribution and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} for Svanvik display a pronounced seasonal cycle.
-The spring peak occurs on $100\,\unit{doy}$ and amounts to $40\,\unit{ppb}$ while the annual average \chem{\left<[O_3]\right>} is $28\,\unit{ppb}$.
-The early decline in \chem{\left<[O_3]\right>} may point to an influence of coniferous vegetation on the removal of ozone.
-Coniferous trees have been observed \chem{CO_2} uptake as early as \unit{doy} 100 \citep{TB:Kolari2007, TP:Wallin2013}.
-The estimated average begin of growing season for deciduous trees ($\left<G_\mathrm{start}\right>_\mathrm{decid}$) is $(150\pm 14)\,\unit{doy}$ at Svanvik (see Section~\ref{subsec:do3se_parameters} for details).
-In July--September, ozone is occasionally almost completely depleted during nighttime.
+A spline was fitted through the data to guide the eye.
+The \chem{[O_3]} density distribution and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} display the expected, pronounced seasonal cycle.
+The spring peak occurs on average on $100\,\unit{doy}$ and amounts to $40\,\unit{ppb}$ while the annual average \chem{\left<[O_3]\right>} is $28\,\unit{ppb}$.
+The decline in \chem{\left<[O_3]\right>} coincides with the average begin of \chem{CO_2} uptake by coniferous trees \citep{TB:Kolari2007, TP:Wallin2013}.
+In July--September, ozone is occasionally almost completely depleted.
 This hints to ozone uptake by vegetation even at low light intensities during midnight sun conditions in combination with stable planetary boundary layer conditions preventing mixing of ozone rich air.
@@
 %Ozone
-We adjusted the ozone climatology for Svanvik \citep{ACPD:Falk2021} and assess the monthly significance of the ozone concentration anomalies in 2018/19.
-The resulting percentage of days above/below $\pm 1\,\sigma$ from the bias-corrected climatological mean is displayed for each month in Fig.~\ref{fig:anomalies_svanvik}d.
-Negative deviations from climatology are shown as a negative percentage.
-Note that missing bars for certain months are both due to data availability and anomalies not meeting the $1\sigma$ criteria.
-The star indicates results after the reconstruction of missing data in July.
-In April/May 2018, ozone was significantly enhanced, while no significant enhancement is found in 2019.
+We use the bias-corrected and cross-calibrated ozone climatology \citep{ACPD:Falk2021} and assess the monthly significance of the ozone concentration anomalies in 2018/19.
+The resulting percentage of days above/below $\pm 1\,\sigma$ is displayed for each month in Fig.~\ref{fig:anomalies_svanvik}d.
+Note that missing bars for months in winter and spring are both due to the temporal extend of our data and anomalies not meeting the $1\sigma$ criteria.
+The star indicates the reconstructed data in July.
+In April/May 2018, ozone was significantly elevated, while no significant enhancement is found in 2019.
 In both years, ozone concentrations were significantly lower than average in November.
-These results indicate that the extensive forest fires in Sweden did not contribute to a significant enhancement in \chem{[O_3]} at Svanhovd in 2018, but the stagnant meteorological conditions in April/May in combination with a late start of the growing season.
-
-In summary, 2018 had a significant number of days that were warmer, drier, and brighter than the climatological average, while 2019 was a rather average year.
+These results suggest that stagnant meteorological conditions in April/May in combination with a late start of the growing season contributed more to elevated ozone than the extensive forest fires in Sweden that did not significantly enhance \chem{[O_3]} in 2018.
+
+In summary, 2018 had a significant number of days that were warmer, drier, and brighter than the climatology, while 2019 was a rather average year.
 Ozone was not significantly enhanced during the GS 2018/19, but a substantially higher number of peak \chem{[O_3]} were observed in 2018 than in 2019.

### 2021-10-04 VegOzone.tex
Refine text

@@
 In our assessment of primary drivers of ozone damage risk on vegetation in the subarctic, we focus on conditions at the NIBIO operated Environment Centre Svanhovd (Sør-Varanger, NOR).
-The NIBIO centre Svanhovd focuses on the research of the special environment of the region, but also comprises long-term observations of various agrometeorological variables, i.e.
+The NIBIO Environment Centre Svanhovd focuses on the research of the special environment of the region, but also comprises long-term observations of various agrometeorological variables, i.e.
 $2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths.
@@
 
-The location of the NIBIO environment centre Svanhovd ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m}$) suggests a subarctic climate.
+The location of the NIBIO Environment Centre Svanhovd ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m}$) suggests a subarctic climate.
 The 1992--2012 climatologies of temperature and precipitation are shown in a combined climate diagram (Fig.~\ref{fig:climatediagram}).
@@
 \label{sec:do3se}
-In the following, we investigate the difference between 2018 and 2019 concerning stomatal ozone uptake and assess which of the environmental factors might be most influential in determining \chem{O_3} damage risk.
+We investigate the difference between 2018 and 2019 concerning stomatal ozone uptake and assess which of the environmental factors might be most influential in determining \chem{O_3} damage risk.
 We model ozone uptake of natural and semi-natural vegetation using the $\mathrm{DO_3SE}$ model \citep{ACP:Bueker2012} and develop bespoke parameterizations for common plant species in our focus area.
 First, we give an account of the methodology for deriving bespoke parameters, after which we present and discuss the modeling results.
-The relevant equations of the $\mathrm{DO_3SE}$ model are summarized in Appendix~\ref{appendix:do3se_model}.
+The relevant components of the $\mathrm{DO_3SE}$ model are summarized in Appendix~\ref{appendix:do3se_model}.
 
@@
 We deduced the dominant species on-site at the NIBIO Environment Centre Svanhovd and from Fig.~\ref{fig:svanhovd_research_station} and found perennial grassland, birch (generalized as deciduous trees), and Scots pine (generalized as coniferous trees).
-Default parameters for these species or PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as mapping manual (MM) parameterizations in the following.
-We base our parameterization of coniferous trees on the MM's Boreal Norway spruce, deciduous trees on the MM's silver birch, and perennial grassland on MM's perennial grassland for central Europe.
-For a comprehensive list of model default parameters, consult Supplement~Table~S1.
+Generic parameters for these PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as mapping manual (MM) parameterizations in the following.
+Our parameterization of coniferous trees is based on the MM's Boreal Norway spruce, deciduous trees on the MM's silver birch, and perennial grassland on MM's perennial grassland for central Europe.
+For a comprehensive list of generic MM model parameters, consult Supplement~Table~S1.
 
 Initial test simulations with the MM PFTs revealed an unrealistically low stomatal conductance at leaf-level ($G_\mathrm{sto}^\mathrm{leaf}$) in 2019 particularly for perennial grassland (Appendix~Fig.~\ref{fig:pody_mm_composit}).
-Substantial stomatal conductance occurred only during an extended warm period in late July.
+Substantial stomatal conductance in this species occurred only during an extended warm period in late July.
 In ecological terms, this means that there was almost no growth of grass in the summer of 2019 -- a prediction which is easily falsified by reality.
 We identified $f_\mathrm{T}$ as abnormally low, being the limiting factor of stomatal conductance in perennial grassland.
-Therefore, we saw the necessity to propose bespoke parameterizations describing an acclimation of the default PFTs to the climatic conditions in the focus area.
+Therefore, we saw the necessity to propose bespoke parameterizations describing an acclimation of the generic PFTs to the climatic conditions in the focus area.
 Note, however, that these parameterizations are hypothetical and have yet to be verified by experiments.
 
-We assume that perennial vegetation will likely maximize carbon acquisition during the short subarctic growing season.
-Because the mapping manual version of the $\mathrm{DO_3SE}$ model currently does not simulate net photosynthesis ($A_\mathrm{n}$), we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune the temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (see Appendix~\ref{appendix:do3se_model} for details) to acclimate the growth potential to the local climate.
+We presume that perennial vegetation will likely maximize carbon acquisition during the short subarctic growing season.
+Because the MM version of the $\mathrm{DO_3SE}$ model currently does not simulate net photosynthesis ($A_\mathrm{n}$), we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune the temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ to acclimate the growth potential to the local climate.
+For details regarding these functions see Appendix~\ref{appendix:do3se_model}.
 Further, we presume that $f_\mathrm{VPD}$ and $f_\mathrm{SWP}$ suit our vegetation types and no acclimation is necessary for these.
@@
 We construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to MM.
-This is accomplished by moving $T_\mathrm{opt}$ towards cooler temperatures while keeping the other parameters fixed to default values.
-In the same way, \emph{subarctic} constructed to represent a species which is very tolerant to cold but sensitive to high temperatures, and most efficient at cool temperatures.
+This is accomplished by moving $T_\mathrm{opt}$ towards cooler temperatures while keeping the other parameters fixed to MM values.
+In the same way, \emph{subarctic} is constructed to represent a species which is very tolerant to cold but sensitive to high temperatures, and most efficient at cool temperatures.
 Therefore, we shift $T\mathrm{min}$ and $T\mathrm{max}$ to colder temperatures and choose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000.
-In other words, we aim to increase the overlapping area between $f_\mathrm{temp}$ and the temperature PDF of the GS.
+In other words, we aim to increase the overlapping area between $f_\mathrm{temp}$ and the temperature PDF during the GS.
 The most extreme of the two acclimation types is \emph{subarctic}.
-In Fig.~\ref{fig:f_temp_grassland}a, we show the resulting $f_\mathrm{temp}$ for perennial grassland together with temperature PDFs deduced from five months GS in two climatological periods.
-The periods 1990s and 2000s have been chosen to indicate that the vegetation is subject to ongoing climate change Thus, the predominant acclimation at present is not known.
+In Fig.~\ref{fig:f_temp_grassland}a, we show the resulting $f_\mathrm{temp}$ for perennial grassland together with temperature PDFs deduced from the five months GS in two climatological periods.
+The periods 1990s and 2000s have been chosen to indicate that the vegetation has been subject to climate change.
+Thus, the predominant acclimation at present is not known.
 
@@
 
-Regarding the opening extent of stomata at low light conditions, we look at the PPFD threshold for a $50\,\unit{\%}$ opening of plants' stomata.
-We assume that this threshold may differ in subarctic species compared to species in less extreme climates.
+Regarding the acclimation of $f_\mathrm{light}$, we presume that the opening extent of stomata at low light conditions differs in subarctic species compared to species in less extreme climates.
+We, therefore, look at the PPFD threshold for a $50\,\unit{\%}$ opening of plants' stomata.
 We derive the inverse function $f_{\mathrm{light},k}^{-1}$ of Eq.~(\ref{eq:flight}) analytically for each PFT or species $k$
@@
 \end{equation}
-and define a variation $\gamma' = \gamma \cdot \eta$ with $\eta \in \{0.8, 1.2\}$ and invert Eq.~(\ref{eq:inverse_function}) for $\alpha$ to find $\alpha'(\gamma')$.
-The resulting functions for perennial grassland (PPFD0.8, PPFD1.2) are shown in Fig.~\ref{fig:f_temp_grassland}b) (for deciduous and coniferous trees refer Appendix~Figs.~\ref{fig:f_temp_spruce}b and \ref{fig:f_temp_birch}b).
+and define a variation $\gamma' = \gamma \cdot \eta$ with $\eta \in \{0.8, 1.2\}$.
+By inverting Eq.~(\ref{eq:inverse_function}) for $\alpha$, we find $\alpha'(\gamma')$.
+The resulting functions for perennial grassland (PPFD0.8, PPFD1.2) are shown in Fig.~\ref{fig:f_temp_grassland}b) (for deciduous and coniferous trees refer Appendix~Figs.~\ref{fig:f_temp_spruce}b and \ref{fig:f_temp_birch}b.
 All derived parameters are tabulated in Table~\ref{tab:sensitivity_tests_light}.
@@
 To compare our bespoke parameterizations, we propose the following metric.
-We calculate an average relative stomatal conductance $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal}) with $f_\mathrm{SWP}=1$) at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time) for the whole record of meteorological data and apply mean and standard deviation.
+We calculate an average relative stomatal conductance $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal}) with $f_\mathrm{SWP}=1$) at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time) for the whole record of key environmental factors' data and apply mean and standard deviation.
 We presume that $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ around noon (highest light intensity) is a good proxy for \chem{CO_2} uptake efficiency (growth).
-In additional, a small deviation between noon and morning and a low standard deviation would indicate a higher robustness to variability in growing conditions.
-Based on the criteria, we identify \emph{subarctic}--PPFD0.8 as the best parameterization for all PFTs (Fig.~\ref{fig:javis_func_opt_mean}).
-As expected due to the small adjustments, coniferous trees display the smallest differences between the different types of acclimation, while the differences for perennial grassland are substantial as a response to the stronger proposed temperature acclimation.
+In addition, a small deviation between noon and morning and a low standard deviation would indicate a higher robustness to variability in growing conditions.
+Based on these criteria, we identify \emph{subarctic}--PPFD0.8 as the best parameterization for all PFTs (Fig.~\ref{fig:javis_func_opt_mean}).
+As expected due to the small adjustments, coniferous trees display the smallest differences between the different parameterizations, while the differences for perennial grassland are substantial as a response to the proposed temperature acclimation.
 
@@
 
-Another important factor for ozone exposure metrics is the timing and length of the GS.
+Another important factor is the timing and length of the GS.
 Start ($A_\text{start}$) and end ($A_\text{end}$) of GS for each PFT are given in units of \unit{doy} (Table~\ref{tab:sensitivity_tests_gs}) and estimated as follows.
-We use a MODIS (Aqua/Terra) retrieved net photosynthesis product for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine $A_\text{start}$ and $A_\text{end}$.
+We use a MODIS (Aqua/Terra) retrieved net photosynthesis product \citep{MODIS_PSN} for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine $A_\text{start}$ and $A_\text{end}$.
 The net photosynthesis vs.
@@
 This value will be used for all PFTs alike.
-The resulting growing season for coniferous trees in 2019 was one month longer than in 2018.
+The resulting growing season for coniferous trees in 2019 is, hence, one month longer than in 2018.
 For estimating $A_\text{start}$ of deciduous trees, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Appendix~Fig.~\ref{fig:greening_season_change_Svanvik}).
@@
 For both years, these dates coincide with the first snow-free day at the closest inland meteorological weather station at Øvre~Neiden (Sør-Varanger, NOR).
-For comparison, the default values for central Europe ($A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$) indicating a substantially longer growing season.
+For comparison, the generic values for central Europe ($A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$) indicating a substantially longer growing season.
 Due to a lack of quantitative field observation at Svanhovd, we assume a latency period of $1\,\unit{month}$ after snow-melt for perennial grassland.
@@
     \middlehline
-    Default$^\star$ & -- & 100 & 307\\
-    \multirow{2}{*}{Deciduous tree} & 2018 & 129$^*$ & 261$^\dagger$ \\
-    & 2019 & 130$^*$ & 274$^\dagger$ \\
-    \multirow{2}{*}{Coniferous tree} & 2018 & 122$^\dagger$ & 261$^\dagger$ \\
-    & 2019 & 106$^\dagger$ & 274$^\dagger$ \\
-    \multirow{2}{*}{Perennial grassland} & 2018 & 159$^\ddagger$ & 261$^\dagger$\\
-    & 2019 & 161$^\ddagger$ & 274$^\dagger$ \\
+    Default$^b$ & -- & 100 & 307\\
+    \multirow{2}{*}{Deciduous tree} & 2018 & 129$^*$ & 261$^a$ \\
+    & 2019 & 130$^c$ & 274$^a$ \\
+    \multirow{2}{*}{Coniferous tree} & 2018 & 122$^a$ & 261$^a$ \\
+    & 2019 & 106$^a$ & 274$^a$ \\
+    \multirow{2}{*}{Perennial grassland} & 2018 & 159$^d$ & 261$^a$\\
+    & 2019 & 161$^d$ & 274$^a$ \\
     \bottomhline
   \end{tabular}
-  \belowtable{$^\dagger$ MODIS (Aqua/Terra) net photosynthesis product;\\
-    $^\star$ \citet{ACP:Simpson2012},\\
-    $^*$ $5\,\unit{days}-5\,\unit{^\circ C}$-rule, $T_\mathrm{air}$ from \href{seNorge.no}{seNorge.no};\\
-    $^\ddagger$ One month after snow melt; reference station Øvre Neiden (Sør-Varanger, NOR)} % Table Footnotes
+  \belowtable{$^a$ MODIS (Aqua/Terra) net photosynthesis product;\\
+    $^b$ \citet{ACP:Simpson2012},\\
+    $^c$ $5\,\unit{days}-5\,\unit{^\circ C}$-rule, $T_\mathrm{air}$ from \href{seNorge.no}{seNorge.no};\\
+    $^d$ One month after snow melt; reference station Øvre Neiden (Sør-Varanger, NOR)} % Table Footnotes
 \end{table}
@@
 The soil at Svanhovd is characterized in an ICP Forest plot survey as gleyic cambisols according to the Food and Agriculture Organization of the United Nations (FAO) soil classification system and as eluviated dystric brunisol in the ICP Forest soil classification system (V.
-Timmerman, personal communication, 2021 May 10) Within the $\mathrm{DO_3SE}$ soil parameterization, the sandy loam texture describes the upper $60\,\unit{cm}$, where the plant roots are found, of this soil type best.
+Timmerman, personal communication, May 2021) Within the $\mathrm{DO_3SE}$ soil parameterization, the sandy loam texture describes the upper $60\,\unit{cm}$, where the plant roots are found, of this soil type best.
 
 The estimate of leaf boundary layer resistance is a function of wind speed which will be influenced by the tree height and leaf dimensions.
-Leaves of downy birch in northern Fennoscandia are probably smaller than the $5\,\unit{cm}$ from the MM parameterization of silver birch leaves.
-A sample of downy birch leaves collected at Svanhovd had an average length of $(3.0\pm 0.5)\,\unit{cm}$.
-In addition, \citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area.
-Both factors will affect leaf boundary layer resistance, the smaller leaves would tend to reduce leaf boundary layer resistances whilst the shorter trees would reduce canopy roughness and increase canopy boundary layer resistances and hence the uptake of \chem{O_3}.
-In addition, one Scots pine forest at Svanhovd has been studied as a part of the ICP Forest mapping.
+A sample of downy birch leaves collected at Svanhovd had an average length of $(3.0\pm 0.5)\,\unit{cm}$ which is smaller than in the MM parameterization.
+\citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area.
+One Scots pine forest at Svanhovd has been studied as a part of the ICP Forest mapping.
 Heights were measured in 2004, when the stand was 90 years old.
@@
 
-We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for the three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using the generic MM parameterization and our bespoke temperature (\emph{cold}, \emph{subarctic}) and light (PPFD0.8, PPFD1.2) response functions.
+We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for the three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using the generic MM parameterization and our bespoke temperature (\emph{cold}, \emph{subarctic}) and light (PPFD0.8, PPFD1.2) parameterizations.
 Canopy height and leaf size are adjusted to local values in the model simulations with bespoke growing season (GS besp.).
-The results displayed in Fig.~\ref{fig:pody_rel} allow for an assessment of systematic uncertainties arising from the use of generic parameterizations established from less climatically extreme conditions.
-For comparison, critical levels for reduction in biomass are shown as horizontal dashed lines \citep[deciduous forest $4\,\unit{\%}$, coniferous forest $2\,\unit{\%}$, and grasslands $10\,\unit{\%}$ biomass reduction,][]{ICP:MappingManual2017,ESPR:Hayes2021}.
+The results displayed in Fig.~\ref{fig:pody_rel} allow for an assessment of over- or underestimation of ozone risk arising from the use of generic parameterizations established from less climatically extreme conditions.
+Critical levels for reduction in biomass \citep[deciduous forest $4\,\unit{\%}$, coniferous forest $2\,\unit{\%}$, and grasslands $10\,\unit{\%}$ biomass reduction,][]{ICP:MappingManual2017,ESPR:Hayes2021} are shown as horizontal dashed lines.
 We identify that an increase in cold tolerance (MM $\rightarrow$ cold $\rightarrow$ subarctic) leads to higher ozone uptake in both probed years, while the usage of locally adapted GS causes a reduction.
 The earlier and longer GS in 2019 leads to increased ozone uptake compared to 2018 despite more episodes of elevated \chem{[O_3]} in 2018.
-Due to the shape of the light response function, a variation of the extent of stomatal opening at low light intensities leads to an asymmetric response in $\mathrm{POD_1}$ and can cause a higher sensitivity to drought conditions.
-Similarly, the temperature acclimation to cold or subarctic conditions can lead to an amplification of effects by droughts.
-
-The magnitude of these effects varies between PFTs as well as years, but the predicted ozone uptake for the bespoke parameterization is always larger than for the MM default parameterizations and of the same order of magnitude as the variability between the years studied here.
+Due to the shape of $f_\mathrm{light}$, a symmetric variation leads to an asymmetric response in $\mathrm{POD_1}$.
+We find that an opening of stomata at lower light intensities can cause a higher sensitivity to drought conditions.
+Similarly, a temperature acclimation to cold or subarctic conditions can lead to an amplification of effects by droughts.
+
+The magnitude of these effects varies between PFTs as well as years, but the predicted ozone uptake for the bespoke parameterization is always larger than for the MM parameterizations and of the same order of magnitude as the variability between the years studied here.
 Our bespoke temperature parameterizations suggest considerable underestimations of ozone risk for subarctic species when relying on the generic MM parameterizations.
@@
 With MM GS, $\mathrm{POD_1}$ values were higher in 2018 than in 2019, throughout our analyses.
-However, taking the bespoke GS into account, this is reversed for the \emph{subarctic} acclimation type.
+However, taking the bespoke GS into account, this is reversed for the subarctic temperature acclimation.
 This can be explained by more favorable growing conditions in 2019.
@@
 For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
-Coniferous trees' parameterization was already well acclimated to prevailing temperatures from the start and therefore differs least.
-Taking the bespoke GS and canopy height into account, we find a higher ozone uptake in 2019 than in 2018 for both bespoke temperature response functions.
-The overall difference deduced from bespoke GS for subarctic and cold types with respect to MM range between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+The MM Boreal Norway spruce arameterization was already well acclimated to prevailing temperatures and therefore differs least from our bespoke parameterizations.
+Taking the bespoke GS and canopy height into account, we find a higher ozone uptake in 2019 than in 2018.
+The overall difference deduced from bespoke GS for subarctic and cold parameterizations with respect to MM range between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
 The maximum deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of $f_\mathrm{light}$ and amounts to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.
@@
 The lower maximum temperature tolerance in the subarctic compared to the cold temperature parameterization probably caused the lower $\mathrm{POD_1}$ values calculated for 2018 in subarctic compared to cold climate, whereas this was not seen for 2019.
-The maximum systematic deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of $f_\mathrm{light}$ amount to $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
+The maximum deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of $f_\mathrm{light}$ amount to $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
 
@@
 
-We estimated the biomass reductions in accordance to the CLs \citep{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme temperature parameterization (\emph{subarctic}) for each bespoke species or PFT.
+We estimated the biomass reductions in accordance to the CLs \citep{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme temperature parameterization (\emph{subarctic}) for each bespoke PFT.
 The estimates are listed in Table~\ref{tab:biomass_reduction}.
-The reported range in uncertainty corresponds to the maximum deviation defined by the $\pm 20\,\unit{\%}$ variation of $f_\mathrm{light}$ with SWP=on.
-Total biomass of deciduous trees and perennial grassland was more strongly reduced in comparison with the default MM parameterizations for both years.
-
-The CL for perennial grassland is placed higher than for the other species.
+The reported range corresponds to the maximum deviation defined by the $\pm 20\,\unit{\%}$ variation of $f_\mathrm{light}$ with SWP=on.
+Total biomass of deciduous trees and perennial grassland was substentially reduced in comparison with the default MM parameterizations in both years.
+The CL for perennial grassland is higher than for the other species.
 From an agro-economical point of view in our focus area, a loss of biomass of more than $10\,\unit{\%}$ can be more severe compared to more productive areas in central Europe and may need adjustment.
 
-Coniferous trees are the least affected and the magnitude of total biomass reduction is independent of the choice of the parameterization within given uncertainties.
+Coniferous trees are the least affected and the magnitude of total biomass reduction is almost independent of the choice of the parameterization within given uncertainties.
 This may reflect that the temperature parameterization in the MM is almost never a limiting factor for stomatal opening.
-The systematic deviation for deciduous trees is one order of magnitude larger than for coniferous trees and perennial grassland.
-
-\textbf{TODO: Check if the legnth effect central Europe vs Svanvik is propperly taken care of.}
+
 From the difference between 2018 and 2019, we infer a substantial interannual variability.
-The interannual variability is estimated to be larger than our reported systematic deviations and can be of the same order of magnitude as the effect of the temperature acclimation.
+The interannual variability appears to be large and is of the same order of magnitude as our reported deviations compared to MM due to temperature and light acclimations.
 This means that the choice of temperature acclimation has the potential to either revoke or amplify the risk of ozone damage for vegetation.
-However, the risk of ozone damage in deciduous trees and perennial grassland in comparison with the generic MM parameterizations is increased, regardless.
 
@@
 In particular, we assumed two tolerance levels to cold temperatures leading to a higher efficiency in photosynthesis at cool temperatures than for central European species.
-To this end, we computed a PDF from observed temperature at the NIBIO environment centre Svanhovd (May--September, 1992--2000).
+To this end, we computed a PDF from observed temperature at the NIBIO Environment Centre Svanhovd (May--September, 1992--2000).
 We adjusted the parameters in the temperature response function of stomatal conductance to increase the enclosed area with the PDF.
@@
 They are empirically determined and account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_\mathrm{T}$), water vapor pressure deficit ($f_\mathrm{VPD}$), and soil water potential ($f_\text{SWP}$).
-All factors differ with species or plant functional type (PFT) denoted with $k$.
+All factors differ with plant functional type (PFT) denoted with $k$.
 
@@
 with $\beta = \frac{T_\text{max}-T_\text{opt}}{T_\text{opt}-T_\text{min}}$ a mixed reciprocal polynomial function of temperature.
-The shape parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various species or PFT and $T_\mathrm{air}$ is the near-surface air temperature.
+The shape parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various PFT and $T_\mathrm{air}$ is the near-surface air temperature.
 All temperatures are defined in units of \unit{^\circ C}.
@@
 
-The $\mathrm{DO_3SE}$ model as described in \citet{ACP:Bueker2012} is used to simulate SWP across a species or PFT specific root depth according to the Penman--Monteith energy balance method that drives water cycling through the soil--plant--atmosphere system.
+The $\mathrm{DO_3SE}$ model as described in \citet{ACP:Bueker2012} is used to simulate SWP across a PFT specific root depth according to the Penman--Monteith energy balance method that drives water cycling through the soil--plant--atmosphere system.
 \citet{ACP:Bueker2012} discuss several formulations available to parameterize the reduction in stomata conductance due to water content in the soil.
@@
 \subsection{Initial $\mathrm{DO_3SE}$ modeling results with mapping manual parameters}
-To demonstrate the necessary acclimation of the mapping manual parameters, we show the initial results of stomatal conductance at leaf-level ($G_\mathrm{sto}^\mathrm{leaf}$) and $\mathrm{POD_y}$ over \unit{doy} for both years and all species or PFTs (Fig.~\ref{fig:pody_mm_composit}).
+To demonstrate the necessary acclimation of the mapping manual parameters, we show the initial results of stomatal conductance at leaf-level ($G_\mathrm{sto}^\mathrm{leaf}$) and $\mathrm{POD_y}$ over \unit{doy} for both years and all PFTs (Fig.~\ref{fig:pody_mm_composit}).
 Observed \chem{[O_3]} is also indicated.
@@
 LE has contributed to framing of the this article, $\mathrm{DO_3SE}$ modeling, and development of the PDF-based temperature acclimation.
-CO has collected species or PFT parameters from literature, performed all $\mathrm{DO_3SE}$ simulations and initial validation of the results.
+CO has collected PFT parameters from literature, performed all $\mathrm{DO_3SE}$ simulations and validation of the results.
 AE contributed with her experience regarding subarctic vegetation in Finnmark.

### 2021-10-04 VegOzone.tex
Remove comment on uploade to zenodo

@@
 \dataavailability{All data is available from public databases or through institutional access.
-$\mathrm{DO_3SE}$ modeling results can be made available upon request {\bf TODO: Upload to an achieve together with processed, gap filled input data on zenodo?}.} %% use this section when having only data sets available
+$\mathrm{DO_3SE}$ modeling results can be made available upon request.} %% use this section when having only data sets available
 

### 2021-10-07 VegOzone.tex
Suggestion by Connie

@@
 
-\title{Characterizing subarctic European vegetation's response on key environmental variables for land surface modeling of ozone pollution and risk assessment}
+\title{Characterizing the response of subarctic European vegetation to key environmental variables for land surface modeling of ozone pollution and risk assessment}
 
@@
 We apply the integrated flux-metric Phytotoxic Ozone Dose (POD) to determine ozone risk and damage to vegetation as a function of \chem{[O_3]}, environmental factors, and species-specific physiology.
-Our study, suggests that assessments that use generic parameterizations are likely to underestimate the risk of ozone damage in this region.
+Our study suggests that assessments that use generic parameterizations are likely to underestimate the risk of ozone damage in this region.
 We find that bespoke parameterizations of plant functional types (PFTs) for subarctic vegetation bio-types result in an ozone-induced biomass loss of $2.5$ to $17.4\,\unit{\%}$.
@@
 The complex photochemical cycles leading to the formation of \chem{O_3} at ground-level involve precursor gases such as carbon monoxide (\chem{CO}) and hydrocarbons known as volatile organic compounds (VOCs) in the presence of nitrogen oxides (\chem{NO_x}).
-Despite a relatively short average tropospheric lifetime of approximately $22\,\unit{days}$ \citep{JGR:Stevenson2005,ACP:Young2013}, ozone and is precursors are subject to advection over long distances.
+Despite a relatively short average tropospheric lifetime of approximately $22\,\unit{days}$ \citep{JGR:Stevenson2005,ACP:Young2013}, ozone and its precursors are subject to advection over long distances.
 Episodes of high ozone concentrations in the absence of local pollution sources can therefore often be attributed to the long-range transport of both ozone and its precursors.

### 2021-10-08 VegOzone.tex
Change doys' position

@@
 
-\correspondence{Stefanie Falk (stefanie.falk@geo.uio.no)}
+\correspondence{Stefanie Falk (stefanie.falk@lmu.de)}
 
@@
 The \chem{[O_3]} density distribution and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} display the expected, pronounced seasonal cycle.
-The spring peak occurs on average on $100\,\unit{doy}$ and amounts to $40\,\unit{ppb}$ while the annual average \chem{\left<[O_3]\right>} is $28\,\unit{ppb}$.
+The spring peak occurs on average on day of the year (\unit{doy}) $100$ and amounts to $40\,\unit{ppb}$ while the annual average \chem{\left<[O_3]\right>} is $28\,\unit{ppb}$.
 The decline in \chem{\left<[O_3]\right>} coincides with the average begin of \chem{CO_2} uptake by coniferous trees \citep{TB:Kolari2007, TP:Wallin2013}.
@@
 Another important factor is the timing and length of the GS.
-Start ($A_\text{start}$) and end ($A_\text{end}$) of GS for each PFT are given in units of \unit{doy} (Table~\ref{tab:sensitivity_tests_gs}) and estimated as follows.
+Start ($A_\text{start}$) and end ($A_\text{end}$) of GS for each PFT are given in \unit{doy} (Table~\ref{tab:sensitivity_tests_gs}) and estimated as follows.
 We use a MODIS (Aqua/Terra) retrieved net photosynthesis product \citep{MODIS_PSN} for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine $A_\text{start}$ and $A_\text{end}$.
@@
 time of year data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}).
-By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{doy}$ for 2018 and \unit{doy} 106 in 2019.
-$A_\text{end}$ amounts to \unit{doy} 261 and 274, respectively.
+By calculating the roots of the fitted polynomial, we derive $A_\text{start}$ as \unit{doy}~122 for 2018 and \unit{doy}~106 in 2019.
+$A_\text{end}$ amounts to \unit{doy}~261 and 274, respectively.
 This value will be used for all PFTs alike.
@@
 For estimating $A_\text{start}$ of deciduous trees, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Appendix~Fig.~\ref{fig:greening_season_change_Svanvik}).
-We find 129 and 130, respectively.
+We find \unit{doy}~129 and 130, respectively.
 For both years, these dates coincide with the first snow-free day at the closest inland meteorological weather station at Øvre~Neiden (Sør-Varanger, NOR).
-For comparison, the generic values for central Europe ($A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$) indicating a substantially longer growing season.
+For comparison, the generic dates for central Europe ($A_\mathrm{start} = 100$ and $A_\mathrm{end} = 307$) indicating a substantially longer growing season.
 Due to a lack of quantitative field observation at Svanhovd, we assume a latency period of $1\,\unit{month}$ after snow-melt for perennial grassland.
@@
 \begin{table}[t]
-  \caption{Start and end of growing season in units of \unit{doy}.}
+  \caption{Start and end of growing season in \unit{doy}.}
   \label{tab:sensitivity_tests_gs}
@@
 
-We have studied the effect of bespoke parameterizations of common vegetation types (coniferous and deciduous trees, perennial grassland) to a subarctic climate on $\mathrm{POD_y}$.
+To summarize, we have studied the effect of bespoke parameterizations of common vegetation types (coniferous and deciduous trees, perennial grassland) to a subarctic climate on $\mathrm{POD_y}$.
 The comparison between meteorological conditions in 2018 and 2019 and their divergence from climatology allowed us to assess the influence of key environmental variables such as temperature, PPFD, and precipitation on vegetation susceptibility to \chem{O_3} damage in light of future changes as may occur under climate change.
@@
 Based on a MODIS photosynthesis product \citep{MODIS_PSN}, we estimated the growing season for coniferous trees in 2018 to be at least one month shorter than in 2019.
-The deduced start of growing season in 2018 ($122\,\unit{doy}$) and 2019 ($106\,\unit{doy}$) are within given ranges by empirical observations \citep{TB:Kolari2007,IVL:Karlsson2018}.
-We determined the start of the growing season for deciduous trees based on gridded observational temperature and for perennial grassland on snow depth data.
+The determined start of growing season in 2018 ($\unit{doy}~122$) and 2019 ($\unit{doy}~106$) lies within the range given by observations in Fennoscandia \citep{TB:Kolari2007,IVL:Karlsson2018}.
+We deduced the start of the growing season for deciduous trees based on gridded observational temperature and for perennial grassland on snow depth data.
 For perennial grassland, we assumed a one month latency period after snow melt which is supported by observational data from Finland \citep{FCR:Korhonen2018}.
@@
   \caption{$\mathrm{DO_3SE}$ modeling results for mapping manual default parameterization.
-$\mathrm{POD_y}$ is shown over day of year (\unit{doy}), March--October.
+$\mathrm{POD_y}$ is shown over \unit{doy}, March--October.
 A flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) has been chosen.
@@
 Daily averaged data for both 2018 and 2019 has been fitted with a quadratic polynomial function.
-The numerically computed root yields: BGS 122/106 day of year (\unit{doy}) and EGS 261/274 day of year for 2018/2019, respectively.}
+The numerically computed root yields: BGS \unit{doy}~122/106 and EGS \unit{doy}~261/274 for 2018/2019, respectively.}
   \label{fig:modis_Psn}

### 2021-10-08 VegOzone.tex
Restructure conclusions

@@
 
-To summarize, we have studied the effect of bespoke parameterizations of common vegetation types (coniferous and deciduous trees, perennial grassland) to a subarctic climate on $\mathrm{POD_y}$.
+To summarize, we have developed bespoke parameterizations of common vegetation types (coniferous and deciduous trees, perennial grassland) to a subarctic climate and studied their effect on $\mathrm{POD_y}$.
 The comparison between meteorological conditions in 2018 and 2019 and their divergence from climatology allowed us to assess the influence of key environmental variables such as temperature, PPFD, and precipitation on vegetation susceptibility to \chem{O_3} damage in light of future changes as may occur under climate change.
@@
 However, ozone concentrations during the growing season were not significantly enhanced and \chem{POD_1} values were very similar for both years.
+This suggests that the overall length of the growing season is more crucial in risk assessment when using flux-based metrics than episodes of peak concentrations as these peak \chem{[O_3]} often coincide with environmental conditions that limit ozone uptake in the $\mathrm{DO_3SE}$ model.
 
@@
 We deduced the start of the growing season for deciduous trees based on gridded observational temperature and for perennial grassland on snow depth data.
-For perennial grassland, we assumed a one month latency period after snow melt which is supported by observational data from Finland \citep{FCR:Korhonen2018}.
+For perennial grassland, we assumed a one month latency period after snow melt which is supported by observational data from Rovaniemi (Finland) \citep{FCR:Korhonen2018}.
 With respect to ongoing climate change, a clear positive trend emerged in length ($5.2\,\unit{d\,decade^{-1}}$) of the growing season which is almost equally distributed between earlier start ($2.9\,\unit{days\,decade^{-1}}$) and later end ($2.3\,\unit{d\,decade^{-1}}$) (Appendix Fig.~\ref{fig:greening_season_change_Svanvik}).
 
-This suggests that the overall length of the growing season is more crucial in risk assessment when using flux-based metrics than episodes of peak concentrations as these peak \chem{[O_3]} often coincide with environmental conditions that limit ozone uptake in the $\mathrm{DO_3SE}$ model.
-Visible damage caused by peak ozone uptake and a dependence of ozone sensibility on phenology are not accounted for in $\mathrm{POD_y}$.
-But open-top chamber (OTC) experiments performed in northern Finland where Scots pine and downy birch seedlings were exposed to elevated ozone concentrations showed a reduction in biomass and reproduction with visible damage explicitly attributed to peak \chem{O_3} concentrations and fast phenological development at high growth rate \citep{Amb:Manninen2009}.
+ The $\mathrm{DO_3SE}$ modeling results are in contrast to our observation of more visible damage in 2018 compared to 2019.
+However, visible damage caused by peak ozone uptake and a dependence of ozone sensibility on phenology are not accounted for in $\mathrm{POD_y}$.
+In this regard, open-top chamber (OTC) experiments performed in northern Finland where Scots pine and downy birch seedlings were exposed to elevated ozone concentrations attributed a reduction in biomass and reproduction with visible damage explicitly to peak \chem{O_3} concentrations and fast phenological development at high growth rate \citep{Amb:Manninen2009}.
 In particular, forbs and perennial grassland have been found to be more susceptible to ozone-induced damage in the reproductive state \citep{EP:Bassin2004}.
@@
 But visible damage and biomass reduction may not occur at the same rate and hence may not be interchangeable.
+
 In our study, we do not explicitly account for this natural distribution of individual plant traits, but our bespoke response functions reflect the diversity within a given species.
@@
 Similarly, we deduced a PPFD PDF and varied the light response function to test a higher or lower extent of stomatal opening at low light intensities.
-These parameterizations are hypothetical and need verification by experimental data.
+Remark that these parameterizations are hypothetical and need verification by experimental data.
 We found that soil water potential under 2018/19 meteorological conditions was negligible when considering the bespoke growing season.
 We have not assessed the effect of the plants' response to VPD in detail, but our climatological data indicated that conditions at Svanhovd are shifting towards a more VPD limited regime.
-Our key findings are: (1) Our bespoke parameterizations for subarctic species are better suited to describe vegetation adapted to subarctic conditions for instance allowing for stomata opening during a larger part of the growing season, (2) the parameterizations defined in the mapping manual for European regional risk assessment appear to not adequately capture the physiological responses of subarctic vegetation that are important determinants of \chem{[O_3]} susceptibility and are likely to underestimate ozone-damage risk, and (3) climatic conditions that are promoting a longer growing season are likely to increase the ozone-damage risk in flux-based metrics.
+
+Our key findings are: 
+\begin{enumerate}
+\item Our bespoke parameterizations for subarctic species are better suited to describe vegetation adapted to subarctic conditions for instance allowing for stomata opening during a larger part of the growing season,
+\item the parameterizations defined in the mapping manual for European regional risk assessment appear to not adequately capture the physiological responses of subarctic vegetation that are important determinants of \chem{[O_3]} susceptibility and are likely to underestimate ozone-damage risk, and
+\item climatic conditions that are promoting a longer growing season are likely to increase the ozone-damage risk in flux-based metrics.
+\end{enumerate}
 
@@
 \citet{ESPR:Hayes2021} pointed out that the highest susceptibility of vegetation to future ozone variability is likely to occur in summer when vegetation is most productive.
-The actual resilience of northern Fennoscandia plant communities to climate change and its predictions strongly depend on both the range of existing acclimations as well as the actual type.
+The actual resilience of northern Fennoscandia plant communities to climate change and its predictions strongly depend on both the range of existing acclimations as well as the actual type acclimation.
 
@@
 All these results have to be treated with care because the biomass reduction functions were established based on species in less extreme climates.
-In particular, CLs for deciduous and coniferous trees were breach in all years and fora ll parameterizations.
+In particular, CLs for deciduous and coniferous trees were breached in all years and forall parameterizations.
 For perennial grassland, the CL was not breached if a bespoke GS was assumed, but the mapping manual defined CL of $10\,\unit{\%}$ loss in biomass might not be transferable to subarctic regions with considerable constrains on productivity leading to more severe economical consequences.

### 2021-10-13 VegOzone.tex
Change running title
Implement suggestions by Lisa

@@
 
-\runningtitle{Subarctic biomes for land surface modeling}
+\runningtitle{Response of subarctic European vegetation to key environmental variables}
 
@@
 \authorcontribution{SF has written the manuscript, collected and processed ozone and environmental data, and performed all statistical analyses.
-AVV contributed to framing this research article, has performed the on-site observation of vegetation damage induced by ozone, and advising in plant physiological processes.
-LE has contributed to framing of the this article, $\mathrm{DO_3SE}$ modeling, and development of the PDF-based temperature acclimation.
-CO has collected PFT parameters from literature, performed all $\mathrm{DO_3SE}$ simulations and validation of the results.
+AVV contributed to framing this research article, has performed the on-site observation of vegetation damage induced by ozone, and provided advice on plant physiological processes.
+LE has contributed to framing of the this article, $\mathrm{DO_3SE}$ modeling, and development of the PDF-based temperature acclimation assessment.
+CO has collected PFT parameters from the literature, performed all $\mathrm{DO_3SE}$ simulations and validation.
 AE contributed with her experience regarding subarctic vegetation in Finnmark.

### 2021-10-14 VegOzone.tex
Include Ane's suggestions

@@
 Ground-level ozone (\chem{O_3}) is a highly toxic pollutant known to cause damage to human health \citep{WHO2008} and to a variety of ecosystems around the World \citep{PT:Emberson2020}.
-Ozone is a cause of reduction of photosynthesis, visible foliar injury, early senescence and programmed cell death of plants \citep{PCE:Kangasjarvi2005}.
+Ozone causes reduced photosynthesis, visible foliar injury, early senescence and programmed cell death of plants \citep{PCE:Kangasjarvi2005}.
 Annual global yield losses of four major crops (wheat, rice, maize, and soybean) of $3-15\,\unit{\%}$ have been reported \citep{PJ:Ainsworth2017} and a suggested loss in primary production in forestry of $7\,\unit{\%}$ \citep{GCB:Wittig2009,EP:Matyssek2012} has been attributed to ozone.
@@
 However, changes in environmental conditions associated with climate change, such as an increase in the frequency of heatwaves, could negate the effectiveness of emission reductions in relation to ozone impacts and air quality \citep{NCC:Lin2020}.
+Tropospheric background ozone concentrations derived from satellite instruments range between $50-65\,\unit{ppb}$ over Fennoscandia but strongly depend on season \citep{ESA:Cooper2014}.
 
@@
 Major pathways of ozone and its precursors' removal from the boundary layer include dry deposition on vegetated surfaces and photochemistry \citep{RG:Clifton2020}.
-As dry deposition to bare ground or snow and ice-covered surfaces was found to be low \citep{ACP:Helmig2007}, both pathways of removal are substantially suppressed during polar night \citep{AE:Monks2000}.
-This leads to a build-up of ozone and its precursors during winter.
+Dry deposition to bare ground or snow and ice-covered surfaces was found to be low \citep{ACP:Helmig2007}.
+This causes a  substantial suppression of ozone removal \citep{AE:Monks2000} and leads to a build-up of ozone and its precursors during winter.
 Come spring precursors are photochemically activated again, while snow cover typically continues to prevail and suppress ozone dry deposition.
@@
 The spring peak is more pronounced in northern Fennoscandia than in more southern locations but shows a high interannual variability \citep{AB:Klingberg2009, BER:Klingberg2019}.
-A cause for this is the complex interplay between atmospheric chemistry and dynamics \citep{AE:Laurila1996,BER:Hatakka2003} as well as anthropogenic and biogenic activity \citep{AE:Rummukainen1996,AE:Simpson2002,QJRMS:Galbally2007,NGS:Schnell2009}.
 \citet{ACP:Andersson2017} showed in a modeling study focused on Fennoscandia that the variability of ozone concentrations in winter can be attributed mainly to changes in atmospheric background (transport) of ozone, while summertime abundance is mostly affected by emissions of precursors in the rest of Europe.
-Meteorology plays an important role throughout the year.
 
@@
 %
-with the hourly averaged stomatal ozone flux $\Phi_\mathrm{sto}$ (see Eq.~(\ref{eq:flux_stomata})) and a stomatal ozone flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}\,s^{-1}}.
+with the hourly averaged stomatal ozone flux $\Phi_\mathrm{sto}$ (see Appendix Eq.~(\ref{eq:flux_stomata}) and a stomatal ozone flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}\,s^{-1}}.
 The flux threshold $\mathrm{y}$ represents the detoxification potential of the plant and is typically only exceeded during daylight hours (i.e.
@@
 This convention aims to develop an effects-based emission reduction policy that can target those ozone precursor emissions that are most influential in causing damage.
-To achieve this the concept of critical levels (CLs) has been developed and applied \citep{Maas2016}.
+To achieve this, the concept of critical levels (CLs) has been developed and applied \citep{Maas2016}.
 Exceedance of the CL is used to identify those areas across Europe that would benefit from targeted Europe-wide emission reductions.
@@
 This study demonstrated the improved ability of flux-based metrics (in comparison to exposure-based metrics) to identify the geographical distribution of the risk of ozone damage and also showed that the ozone damage could extend into more northerly regions of Europe.
-For our study, we conducted a similar bio-monitoring study at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd to assess whether local ozone concentrations were capable of inducing damage to vegetation growing under northern Fennoscandia conditions.
+For our study, we conducted a similar bio-monitoring study at Svanhovd, a site with agrometeorological and air pollution monitoring, to assess whether local ozone concentrations were capable of inducing damage to vegetation growing under northern Fennoscandia conditions.
 In 2018, we indeed observed ozone damage on clover and tobacco.
@@
 An unusually weak and northward shifted jet stream allowed for a persistent high-pressure system above northern Europe, including Fennoscandia which blocked the low-pressure systems for several consecutive months.
-In the period May -- July, southern Norway experienced temperatures $4\,\unit{^\circ C}$ above normal and saw only about $60\,\unit{\%}$ of normal precipitation.Northern Norway as a whole had close to normal precipitation, but with local variations.
+In the period May -- July, southern Norway experienced temperatures $4\,\unit{^\circ C}$ above normal and saw only about $60\,\unit{\%}$ of normal precipitation.
+Northern Norway as a whole had close to normal precipitation, one of the causes of the high rate of forest fires in that area.
 Thermal stress on vegetation was exceptional not only in large parts of Fennoscandia but also in much of Europe, where the influence of the high-pressure system extended even over five months (April/May, July--September).
@@
 It is very likely that elevated ozone, like during the 2003 drought period \citep{JGR:Solberg2018}, was promoted by a combination of various factors such as wildfires, reduced cloud cover (increased solar radiation), reduced dry deposition and turbulent mixing due to the stagnant weather conditions, and increased BVOC emissions.
-As such events may become more frequent in the future, conditions at Svanhovd for the growing seasons 2018/19 shall serve as a reference for present and probable future conditions in northern Fennoscandia.\\
-
-The described methods for risk assessment rely on accurate representation of plant physiological response to key environmental variables and are most often parameterized using data for less extreme climates, i.e., temperate or continental climates; and to a lesser extent, Boreal and Mediterranean climates.
+As such events may become more frequent in the future, conditions at Svanhovd for the growing seasons 2018 shall serve as a reference for probable future conditions in northern Fennoscandia.
+Conditions in 2019 represent present climate.\\
+
+The described methods for risk assessment rely on accurate representation of plant physiological responses to key environmental variables and are most often parameterized using data for less extreme climates, i.e., temperate or continental climates; and to a lesser extent, Boreal and Mediterranean climates.
 Parameterization of these models for more extreme bio-geographical regions such as subarctic climates has not been performed and relies on assuming that less extreme Boreal parameterization will be applicable.
@@
 This allows an understanding of how the threat from ozone may change in the future and which are the key aspects of plant physiology that might determine the potential risk.
-Hence, we determine which parameters require special attention to define accurately for pollution impact assessment and more generally for describing gas exchange response to a changing climate that will influence biochemical cycling.
+Hence, we determine which parameters require special attention for more accurate pollution impact assessment and more generally for describing gas exchange response to a changing climate that will influence biogeochemical cycling.
 This might also be one key to solve long-standing issues of earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface models \citep{GMD:Poulter2015,JAMES:Lawrence2019}.
 
-The observation site at NIBIO Environment Centre Svanhovd and available observational data are presented in Section~\ref{sec:data}.
-From these data, we derive multi-annual means of essential climate variables (temperature, precipitation, global irradiance, and ozone) and evaluate the deviation of 2018/19 from this norm (Section~\ref{sec:stats}).
+The observation site Svanhovd and available observational data are presented in Section~\ref{sec:data}.
+From these data, we derive multi-annual averages of essential climate variables (temperature, precipitation, global irradiance, and ozone) and evaluate the deviation of 2018/19 from this norm (Section~\ref{sec:stats}).
 In Section~\ref{sec:do3se}, we use the $\mathrm{DO_3SE}$ model to estimate the ozone uptake by natural and semi-natural vegetation to investigate the main environmental drivers for projected ozone damage risk in 2018.
@@
 
-In our assessment of primary drivers of ozone damage risk on vegetation in the subarctic, we focus on conditions at the NIBIO operated Environment Centre Svanhovd (Sør-Varanger, NOR).
-The NIBIO Environment Centre Svanhovd focuses on the research of the special environment of the region, but also comprises long-term observations of various agrometeorological variables, i.e.
-$2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths.
-We investigate the 2018/19 growing season in a climatological context.
-To this end, we collected atmospheric and ozone monitoring data taken at Svanhovd over the past $35\,\unit{years}$.
-Svanhovd synonymously refers to the atmospheric monitoring site Svanvik/Pasvik throughout the paper.\\
-
-All measured essential climate variables including ozone for the 2018/19 growing season were obtained through \href{luftkvalitet.no}{luftkvalitet.no} operated by \citet{NILU_AIRQ} (last accessed May 2020).
-The ozone monitoring data have been also added to the EBAS database operated by \citet{NILU_EBAS} and are thus openly accessible.
+We performed ozone concentration measurements in the subarctic, during the growing seasons of 2018 and 2019.
+Our chosen location was Svanhovd ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m\,a.s.l.}$) for several reasons.
+There were ozone measurements there in the past, there is an agrometeorological station measuring many of the climate variables we needed, e.g.
+$2\,\unit{m}$ temperature, precipitation, global irradiance and soil temperature at different depths.
+We collected atmospheric monitoring and ozone monitoring data taken at Svanhovd over the past $35\,\unit{years}$.
+There is also ongoing national surveillance of air pollution (\chem{NO_x}, \chem{SO_x}, PM, radiological species).
+Most of these measurements are performed by the Norwegian Institute for Air Research (NILU) and they also performed our ozone measurements.\\
+
+All measured essential climate variables for the 2018/19 growing season were obtained through \href{luftkvalitet.no}{luftkvalitet.no} operated by \citet{NILU_AIRQ} (last accessed May 2020).
+Our ozone monitoring data have been added to luftkvalitet.no and the EBAS database operated by \citet{NILU_EBAS} and are thus openly accessible.
 Long-term ozone observation data (1986--1996) were obtained from EBAS.
@@
 Especially, ozone monitors did not undergo regular re-calibration.
-This probably led to drifts in the observed data and may impose a false trend which shall not be of concern for us.
+This probably led to drifts in the observed data and may impose a false trend which shall not be of concern for us as we are interested in the seasonality.
 The long-term data sets of agrometeorological variables including temperature and precipitation are available from September 1992 to the present day \citep[LandbruksMeteorologiske Tjeneste][note the station name here is Pasvik]{LMT_NIBIO}.
 
-To relate local ozone concentration \chem{[O_3]} with visible damage on ozone sensitive plants, both an ozone bio-monitor (referred to as ozone garden) and a conventional ozone monitor using the existing atmospheric monitoring infrastructure had been installed during the 2018/19 growing season at the NIBIO centre.
+To relate local ozone concentration \chem{[O_3]} with visible damage on ozone sensitive plants, both an ozone bio-monitor (referred to as ozone garden) and a conventional ozone monitor using the existing atmospheric monitoring infrastructure had been installed during the 2018/19 growing season at Svanhovd.
 We report \chem{[O_3]} as volume mixing ratios (VMR) in \unit{ppb} for easier comparison with exposure-based risk assessment.
@@
 Cultivated species were, e.g.
-clover, tobacco, and potatoes.
+clovers (\emph{Trifolium repens}, \emph{Trifolium pratense}), tobacco (\emph{Nicotiana tabacum}, cultivars Bel-W3, Bel-B and Bel-C), and potato (\emph{Solanum tuberosum}).
 As shown in Fig.~\ref{fig:svanhovd_research_station}b, the plants had to be protected from herbivores with a wire-mesh fence.
 In 2018, we qualitatively observed visible ozone damage on semi-natural vegetation (clover) and crops (tobacco).
-In contrast, no such visible damage was found on the clovers in 2019, although it was observed on the sensitive tobacco cultivar.
+In contrast, no such visible damage was found on the clovers in 2019, although it was observed on the sensitive tobacco cultivar (Bel-W3).
 This indicates that the vegetation was more affected by ozone in 2018 than in 2019.
@@
 In this section, we assess the climate conditions at Svanhovd.
-To this end, we compute a multi-annual mean (referred to as climatology) for each environmental key variable (temperature, precipitation, global irradiance) and ground-level ozone.
+To this end, we compute a multi-annual average (referred to as climatology) for each environmental key variable (temperature, precipitation, global irradiance) and ground-level ozone.
 We evaluate the statistical significance of divergences from the norm in these variables (referred to as anomalies) in 2018/19.
@@
 
-The location of the NIBIO Environment Centre Svanhovd ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m}$) suggests a subarctic climate.
+Svanhovd is located within the subarctic climate zone.
 The 1992--2012 climatologies of temperature and precipitation are shown in a combined climate diagram (Fig.~\ref{fig:climatediagram}).
@@
 The average accumulated monthly precipitation ($\sum_d\mathrm{Precip}$) indicates that winter and spring (November--April, except for March) are relatively dry ($\sum_d\mathrm{Precip} < 20\,\unit{mm}$).
+March precipitation is primarily snow and will influence the start of the growing season.
 The driest month is January with $\sum_d\mathrm{Precip} = (16.7\pm 3.0)\,\unit{mm}$.
@@
 The climatology for global irradiance ($Q_0$, not shown) approaches zero in December/January (polar night) and reaches its maximum in June/July (midnight sun conditions).
-The highest observed averaged daily maximum ($\left<Q_0^\mathrm{max}\right>$) amounts to $800\,\unit{W\,m^{-2}}$, while the maximum of the averaged daily mean ($\left<Q_0^\mathrm{mean}\right>$) is $450\,\unit{W\,m^{-2}}$ (regularly reached in May--July).
-The maximum of the averaged daily minimum ($\left<Q_0^\mathrm{min}\right>$) lies at about $200\,\unit{W\,m^{-2}}$.
+We select the highest $Q_0$ for each day (noon) and average each over the whole observation period.
+Thus, we find the highest averaged daily maximum ($\left<Q_0^\mathrm{max}\right>$) which amounts to $800\,\unit{W\,m^{-2}}$.
+Similarly, we diurnal mean $Q_0$ (including night) and find the maximum of the averaged daily mean ($\left<Q_0^\mathrm{mean}\right>$) $450\,\unit{W\,m^{-2}}$ (regularly reached in May--July).
+By selecting the lowest daily $Q_0$ (including night), we compute the maximum of the averaged daily minimum ($\left<Q_0^\mathrm{min}\right>$) about $200\,\unit{W\,m^{-2}}$.
+This indicates that a considerable amount of light is also available during night.
 
@@
   \caption{Svanhovd ozone climatology.
-The density distribution is shown together with a multi-annual mean of daily \chem{[O_3]} (climatology).
+The density distribution is shown together with a multi-annual average of daily \chem{[O_3]} (climatology).
 A spline has been fitted through daily mean \chem{[O_3]}.
@@
 
-We deduced the dominant species on-site at the NIBIO Environment Centre Svanhovd and from Fig.~\ref{fig:svanhovd_research_station} and found perennial grassland, birch (generalized as deciduous trees), and Scots pine (generalized as coniferous trees).
+We deduced the dominant species on-site at Svanhovd and from Fig.~\ref{fig:svanhovd_research_station} and found perennial grassland, birch (generalized as deciduous trees), and Scots pine (generalized as coniferous trees).
 Generic parameters for these PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as mapping manual (MM) parameterizations in the following.
@@
 In particular, we assumed two tolerance levels to cold temperatures leading to a higher efficiency in photosynthesis at cool temperatures than for central European species.
-To this end, we computed a PDF from observed temperature at the NIBIO Environment Centre Svanhovd (May--September, 1992--2000).
+To this end, we computed a PDF from observed temperature at Svanhovd (May--September, 1992--2000).
 We adjusted the parameters in the temperature response function of stomatal conductance to increase the enclosed area with the PDF.
@@
 Daily averaged data for both 2018 and 2019 has been fitted with a quadratic polynomial function.
-The numerically computed root yields: BGS \unit{doy}~122/106 and EGS \unit{doy}~261/274 for 2018/2019, respectively.}
+The numerically computed root yields: BGS \unit{doy}~122/106 and EGS \unit{doy}~261/274 for 2018/19, respectively.}
   \label{fig:modis_Psn}
@@
 \begin{acknowledgements}
+  We thank the NIBIO Environment Centre Svanhovd for their support.
   Bj{\o}rg Rognerud for processing SeNorge.no data with respect for begin of growing season.

### 2021-10-14 VegOzone.tex
Conclude suggestios by Ane and Aud

@@
 The climatology for global irradiance ($Q_0$, not shown) approaches zero in December/January (polar night) and reaches its maximum in June/July (midnight sun conditions).
-We select the highest $Q_0$ for each day (noon) and average each over the whole observation period.
-Thus, we find the highest averaged daily maximum ($\left<Q_0^\mathrm{max}\right>$) which amounts to $800\,\unit{W\,m^{-2}}$.
-Similarly, we diurnal mean $Q_0$ (including night) and find the maximum of the averaged daily mean ($\left<Q_0^\mathrm{mean}\right>$) $450\,\unit{W\,m^{-2}}$ (regularly reached in May--July).
-By selecting the lowest daily $Q_0$ (including night), we compute the maximum of the averaged daily minimum ($\left<Q_0^\mathrm{min}\right>$) about $200\,\unit{W\,m^{-2}}$.
-This indicates that a considerable amount of light is also available during night.
+We deduce the corresponding highest $Q_0$ to $800\,\unit{W\,m^{-2}}$.
+The maximum of the lowest $Q_0$ (including night) amounts to $200\,\unit{W\,m^{-2}}$ indicating that a considerable amount of light is available under midnight sun conditions.
 
 %Ozone
-The derived ozone climatology (\chem{\left<[O_3]\right>}) (1986--1996) is shown in Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}.
+The daily mean ozone climatology (\chem{\left<[O_3]\right>}) (1986--1996) is shown in Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}.
 Darker colors indicate higher probability to observe these values.
@@
 The spring peak occurs on average on day of the year (\unit{doy}) $100$ and amounts to $40\,\unit{ppb}$ while the annual average \chem{\left<[O_3]\right>} is $28\,\unit{ppb}$.
-The decline in \chem{\left<[O_3]\right>} coincides with the average begin of \chem{CO_2} uptake by coniferous trees \citep{TB:Kolari2007, TP:Wallin2013}.
-In July--September, ozone is occasionally almost completely depleted.
+The decline in \chem{\left<[O_3]\right>} coincides with the average beginning of \chem{CO_2} uptake by coniferous trees \citep{TB:Kolari2007, TP:Wallin2013}.
+In July--September (\unit{doy}~182--273), ozone is occasionally almost completely depleted.
 This hints to ozone uptake by vegetation even at low light intensities during midnight sun conditions in combination with stable planetary boundary layer conditions preventing mixing of ozone rich air.
@@
 Therefore, we shift $T\mathrm{min}$ and $T\mathrm{max}$ to colder temperatures and choose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000.
-In other words, we aim to increase the overlapping area between $f_\mathrm{temp}$ and the temperature PDF during the GS.
+In other words, we aim to increase the overlapping area between $f_\mathrm{temp}$ and the temperature probability density function (PDF) during the GS.
 The most extreme of the two acclimation types is \emph{subarctic}.
@@
 Regarding the acclimation of $f_\mathrm{light}$, we presume that the opening extent of stomata at low light conditions differs in subarctic species compared to species in less extreme climates.
-We, therefore, look at the PPFD threshold for a $50\,\unit{\%}$ opening of plants' stomata.
-We derive the inverse function $f_{\mathrm{light},k}^{-1}$ of Eq.~(\ref{eq:flight}) analytically for each PFT or species $k$
+We, therefore, adjust the light sensitivity of stomata such that the PPFD value needed to reach $50\,\unit{\%}$ opening is varied $\pm 20\,\unit{\%}$.
+For this we derive the inverse function $f_{\mathrm{light},k}^{-1}$ of Eq.~(\ref{eq:flight}) analytically for each PFT or species $k$
 \begin{equation}
@@
 Start ($A_\text{start}$) and end ($A_\text{end}$) of GS for each PFT are given in \unit{doy} (Table~\ref{tab:sensitivity_tests_gs}) and estimated as follows.
-We use a MODIS (Aqua/Terra) retrieved net photosynthesis product \citep{MODIS_PSN} for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine $A_\text{start}$ and $A_\text{end}$.
+For coniferous trees, we use a MODIS (Aqua/Terra) retrieved net photosynthesis product \citep{MODIS_PSN} for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine $A_\text{start}$ and $A_\text{end}$.
 The net photosynthesis vs.
@@
 The soil at Svanhovd is characterized in an ICP Forest plot survey as gleyic cambisols according to the Food and Agriculture Organization of the United Nations (FAO) soil classification system and as eluviated dystric brunisol in the ICP Forest soil classification system (V.
-Timmerman, personal communication, May 2021) Within the $\mathrm{DO_3SE}$ soil parameterization, the sandy loam texture describes the upper $60\,\unit{cm}$, where the plant roots are found, of this soil type best.
+Timmerman, personal communication, June 2020) Within the $\mathrm{DO_3SE}$ soil parameterization, the sandy loam texture describes the upper $60\,\unit{cm}$, where the plant roots are found, of this soil type best.
 
@@
 Mean tree height was $10.1\,\unit{m}$ and maximum height was $16.2\,\unit{m}$ (V.
-Timmerman, personal communication, 2021 May 10).
+Timmerman, personal communication, June 2020).
+We used $13.5\,\unit{m}$ height for both deciduous and coniferous trees.
 
@@
 
-The magnitude of these effects varies between PFTs as well as years, but the predicted ozone uptake for the bespoke parameterization is always larger than for the MM parameterizations and of the same order of magnitude as the variability between the years studied here.
+The magnitude of these effects varies between PFTs as well as years, but the predicted ozone uptake for the bespoke temperature parameterization is always larger than for the MM parameterizations and of the same order of magnitude as the variability between the years studied here.
 Our bespoke temperature parameterizations suggest considerable underestimations of ozone risk for subarctic species when relying on the generic MM parameterizations.
@@
 For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
-The MM Boreal Norway spruce arameterization was already well acclimated to prevailing temperatures and therefore differs least from our bespoke parameterizations.
+The MM Boreal Norway spruce parameterization was already well acclimated to the prevailing temperatures and therefore differs the least from our bespoke parameterizations.
 Taking the bespoke GS and canopy height into account, we find a higher ozone uptake in 2019 than in 2018.
@@
 The maximum deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of $f_\mathrm{light}$ amount to $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
+
+From the difference between 2018 and 2019, we infer a substantial interannual variability.
+The maximum deviation from the generic default parameterizations due to our bespoke temperature and light response functions is of the same order of magnitude as the interannual variability.
 
@@
 The CL for perennial grassland is higher than for the other species.
-From an agro-economical point of view in our focus area, a loss of biomass of more than $10\,\unit{\%}$ can be more severe compared to more productive areas in central Europe and may need adjustment.
+From an agro-economical point of view in our focus area, where grass is cut only once a year, a loss of biomass of more than $10\,\unit{\%}$ can have severer consequences compared to more productive areas in central Europe and may need adjustment.
 
@@
 This may reflect that the temperature parameterization in the MM is almost never a limiting factor for stomatal opening.
-
-From the difference between 2018 and 2019, we infer a substantial interannual variability.
-The interannual variability appears to be large and is of the same order of magnitude as our reported deviations compared to MM due to temperature and light acclimations.
-This means that the choice of temperature acclimation has the potential to either revoke or amplify the risk of ozone damage for vegetation.
 
@@
 In particular, forbs and perennial grassland have been found to be more susceptible to ozone-induced damage in the reproductive state \citep{EP:Bassin2004}.
-That might explain why we observed ozone damage to a larger extent in 2018 than in 2019.
+That might explain why we observed damage in the ozone garden to a larger extent in 2018 than in 2019.
 An associated damage function to ozone depending on leaf age as proposed by \citet{AE:Musselman2006} or phenological state might improve the predictions.
 
-At the same time, intra-species variability even at the same location is typically non-negligible \citep{EP:Bassin2004,Amb:Girgzdiene2009} but not accounted for in the models.
-This may also explain some contradictory results regarding ozone sensitivity of natural vegetation in northern Europe.
+At the same time, intra-species variability even at the same location is typically non-negligible \citep{EP:Bassin2004} but not accounted for in the models.
+The literature results regarding ozone sensitivity of natural vegetation in northern Europe are contratictive.
 \citet{FS:Subramanian2014} reported a higher, modeled reduction of net primary production in conifers ($4.3-15.5\,\unit{\%}$) than birch ($1.4-4.3\,\unit{\%}$) under elevated ozone in Sweden, while \citet{Amb:Girgzdiene2009} observed more visible damage on deciduous trees than on Norway spruce in Lithuania.
@@
 In our study, we do not explicitly account for this natural distribution of individual plant traits, but our bespoke response functions reflect the diversity within a given species.
-In particular, we assumed two tolerance levels to cold temperatures leading to a higher efficiency in photosynthesis at cool temperatures than for central European species.
+In particular, we assumed two tolerance levels to cold temperatures than for central European species.
 To this end, we computed a PDF from observed temperature at Svanhovd (May--September, 1992--2000).
@@
 \citet{ESPR:Hayes2021} pointed out that the highest susceptibility of vegetation to future ozone variability is likely to occur in summer when vegetation is most productive.
-The actual resilience of northern Fennoscandia plant communities to climate change and its predictions strongly depend on both the range of existing acclimations as well as the actual type acclimation.
+The actual resilience of northern Fennoscandia plant communities to climate change strongly depend on both the range of existing acclimations as well as the actual type acclimation.
+This means vegetation which is acclimated to cold temperatures and little to heat- and drought stress might suffer more strongly from a more frequent occurance of heatwaves than plants which have a higher temperature tolerance.
 
@@
 This indicates a higher modeled robustness to ozone damage in coniferous trees.
-All these results have to be treated with care because the biomass reduction functions were established based on species in less extreme climates.
-In particular, CLs for deciduous and coniferous trees were breached in all years and forall parameterizations.
+All these results have to be treated with care because the biomass reduction functions were established based on studies from less extreme climates.
+In particular, CLs for deciduous and coniferous trees were breached in all years and for all parameterizations.
 For perennial grassland, the CL was not breached if a bespoke GS was assumed, but the mapping manual defined CL of $10\,\unit{\%}$ loss in biomass might not be transferable to subarctic regions with considerable constrains on productivity leading to more severe economical consequences.
@@
 
-\authorcontribution{SF has written the manuscript, collected and processed ozone and environmental data, and performed all statistical analyses.
-AVV contributed to framing this research article, has performed the on-site observation of vegetation damage induced by ozone, and provided advice on plant physiological processes.
-LE has contributed to framing of the this article, $\mathrm{DO_3SE}$ modeling, and development of the PDF-based temperature acclimation assessment.
+\authorcontribution{All authors contributed to framing this research article, writing, and conclusions.
+SF has written the manuscript, collected and processed ozone and environmental data, and performed all statistical analyses.
+AVV has performed the on-site observation of vegetation damage induced by ozone, provided advice on plant physiological processes, collected exisiting literature on subarctic vegetation, contributed significantly to framing and proof reading this research article.
+LE has contributed with expertize in $\mathrm{DO_3SE}$ modeling and development of the PDF-based temperature acclimation assessment.
 CO has collected PFT parameters from the literature, performed all $\mathrm{DO_3SE}$ simulations and validation.
 AE contributed with her experience regarding subarctic vegetation in Finnmark.
-FS contributed with an assessment of the 2018 meteorological conditions.
-TB gave guidance in a broader research sense.
-All authors contributed to discussions, writing, and conclusions.} %% this section is mandatory
+FS contributed with an assessment of the 2018 meteorological conditions and writing the proposal for founding this project.
+TB gave valuable guidance in a broader research sense.} %% this section is mandatory
 
@@
 \begin{acknowledgements}
-  We thank the NIBIO Environment Centre Svanhovd for their support.
-  Bj{\o}rg Rognerud for processing SeNorge.no data with respect for begin of growing season.
-  Tore Flatlandsmo Berglen for hourly pressure data from Svanhovd.
-  This work was supported by the Research Council of Norway (NRC) grand No. 268073.
+  We thank NILU for performing the ozone measurements at Svanhovd for us, Bj{\o}rg Rognerud (Department of Geosciences, UiO) for processing SeNorge.no data with respect to beginning of growing season, Tore Flatlandsmo Berglen (NILU) for hourly pressure data from Svanhovd, NIBIO Environment Center Svanhovd for establishing and running the ozone garden both years, and Volkmar Timmermann (NIBIO) for sharing ICP Forest data from Svanhovd.
+This work was supported by the Research Council of Norway (Grant No.
+268073).
 \end{acknowledgements}

### 2021-10-15 VegOzone.tex
Check abstract with grammarly

@@
 \affil[3]{Department of Environment and Geography, University of York, UK}
-\affil[4]{Department of Geography, Ludwig Maximilians Universität, Munich, Germany}
+\affil[4]{Department of Geography, Ludwig Maximilian University of Munich, Munich, Germany}
 
@@
 The atmospheric and climatic characteristics of the subarctic are known to lead to pronounced peak \chem{[O_3]} in spring.
-To date, only a few studies have been conducted to assess the response of subarctic vegetation to variations in climate and air pollution.
+To date, only a few studies assessed the response of subarctic vegetation to variations in climate and air pollution.
 This study looks to fill this knowledge gap by examining essential climate variables, in particular ozone, over the past few decades.
-These data are used to assess the extent to which two recent years (2018 and 2019) deviate from climatic and \chem{[O_3]} norms and how these potentially more frequent future deviations may influence ozone damage to subarctic vegetation.
-We find that 2018 was an anomalously warm and bright year (in particular in spring and early summer) with higher than average \chem{[O_3]} in April/May and a high number of episodic peak values in ozone volume mixing ratio (VMR) above $40\,\unit{ppb}$ in June--August, the latter in part attributable to forest fires in the Northern Hemisphere as well as warmer and sunnier conditions.
+We evaluate the extent to which two recent years (2018 and 2019) deviate from climatic and \chem{[O_3]} norms and how these potentially more frequent future deviations may influence ozone damage to subarctic vegetation.
+We find that 2018 was an anomalously warm and bright year, particularly in spring and early summer.
+Higher than average \chem{[O_3]} was observed in April/May while frequent episodes of ozone volume mixing ratios (VMRs) above $40\,\unit{ppb}$ occurred in June--August.
+These episodes are in part attributable to forest fires in the Northern Hemisphere and warmer and sunnier conditions.
 We apply the integrated flux-metric Phytotoxic Ozone Dose (POD) to determine ozone risk and damage to vegetation as a function of \chem{[O_3]}, environmental factors, and species-specific physiology.
-Our study suggests that assessments that use generic parameterizations are likely to underestimate the risk of ozone damage in this region.
+Our study suggests that using generic parameterizations in assessments likely leads to underestimating the risk of ozone damage in this region.
 We find that bespoke parameterizations of plant functional types (PFTs) for subarctic vegetation bio-types result in an ozone-induced biomass loss of $2.5$ to $17.4\,\unit{\%}$.
-For some species this loss is up to $6\,\unit{\%}$ larger than projected from generic parameterizations.
-Efforts should be targeted towards accurately defining subarctic ecotypes' physiological response to essential climate variables to improve regional and global scale biogeochemical cycling under current and future climates.
+For some species, this loss is up to $6\,\unit{\%}$ larger than projected from generic parameterizations.
+Efforts should be targeted towards accurately defining subarctic vegetations' physiological response to essential climate variables.
+That could help to improve regional and global scale biogeochemical cycling under current and future climates.
 \end{abstract}

### 2021-10-15 VegOzone.tex
Check intro duction with grammarly

@@
 Ground-level ozone (\chem{O_3}) is a highly toxic pollutant known to cause damage to human health \citep{WHO2008} and to a variety of ecosystems around the World \citep{PT:Emberson2020}.
-Ozone causes reduced photosynthesis, visible foliar injury, early senescence and programmed cell death of plants \citep{PCE:Kangasjarvi2005}.
+Ozone causes reduced photosynthesis, visible foliar injury, early senescence, and programmed cell death of plants \citep{PCE:Kangasjarvi2005}.
 Annual global yield losses of four major crops (wheat, rice, maize, and soybean) of $3-15\,\unit{\%}$ have been reported \citep{PJ:Ainsworth2017} and a suggested loss in primary production in forestry of $7\,\unit{\%}$ \citep{GCB:Wittig2009,EP:Matyssek2012} has been attributed to ozone.
@@
 This can be attributed to a successful implementation of air quality regulations coupled with economic restructuring, reducing the number of episodes of peak concentrations especially in summertime \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
-However, changes in environmental conditions associated with climate change, such as an increase in the frequency of heatwaves, could negate the effectiveness of emission reductions in relation to ozone impacts and air quality \citep{NCC:Lin2020}.
+However, changes in environmental conditions associated with climate change, such as an increase in the frequency of heatwaves, could negate the effectiveness of emission reductions concerning ozone impacts and air quality \citep{NCC:Lin2020}.
 Tropospheric background ozone concentrations derived from satellite instruments range between $50-65\,\unit{ppb}$ over Fennoscandia but strongly depend on season \citep{ESA:Cooper2014}.
@@
 As such events may become more frequent in the future, conditions at Svanhovd for the growing seasons 2018 shall serve as a reference for probable future conditions in northern Fennoscandia.
-Conditions in 2019 represent present climate.\\
+Conditions in 2019 represent the present climate.\\
 

### 2021-10-15 VegOzone.tex
Checked section 2 with grammerly

@@
 We performed ozone concentration measurements in the subarctic, during the growing seasons of 2018 and 2019.
-Our chosen location was Svanhovd ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m\,a.s.l.}$) for several reasons.
-There were ozone measurements there in the past, there is an agrometeorological station measuring many of the climate variables we needed, e.g.
-$2\,\unit{m}$ temperature, precipitation, global irradiance and soil temperature at different depths.
-We collected atmospheric monitoring and ozone monitoring data taken at Svanhovd over the past $35\,\unit{years}$.
-There is also ongoing national surveillance of air pollution (\chem{NO_x}, \chem{SO_x}, PM, radiological species).
-Most of these measurements are performed by the Norwegian Institute for Air Research (NILU) and they also performed our ozone measurements.\\
+Our chosen location was Svanhovd ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m\,a.s.l.}$) for several reasons: (1) Ozone had been monitored there in the past, (2) an agrometeorological station is measuring, e.g.
+$2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths, and (3) a national surveillance of air pollution (\chem{NO_x}, \chem{SO_x}, PM, radiological species) is ongoing.
+Most of these measurements are performed by the Norwegian Institute for Air Research (NILU), who also conducted ozone measurements on our behalf.
+In addition, we collected atmospheric monitoring and ozone monitoring data taken at Svanhovd over the past $35\,\unit{years}$.\\
 
@@
 Especially, ozone monitors did not undergo regular re-calibration.
-This probably led to drifts in the observed data and may impose a false trend which shall not be of concern for us as we are interested in the seasonality.
+That probably led to drifts in the observed data and may impose a false trend which shall not be of concern for us as we are interested in the seasonality.
 The long-term data sets of agrometeorological variables including temperature and precipitation are available from September 1992 to the present day \citep[LandbruksMeteorologiske Tjeneste][note the station name here is Pasvik]{LMT_NIBIO}.
@@
 (a) Atmospheric monitoring site and ozone garden have been marked.
-Aerial photography \copyright Norges Kartverk; (b) Clover in the Svanhovd ozone garden.
+Aerial photography \copyright Norges Kartverk; (b) Clovers in the Svanhovd ozone garden.
 The plants had to be secured against herbivores with a wire-mesh fence.
@@
 Qualitatively, global irradiance ($Q_0$) displayed in Fig.~\ref{fig:data_svanvik_2018_2019}d was higher in May and July 2018 compared to 2019, while June 2019 showed higher irradiance than in 2018.
-A high global irradiance in most cases is the result of a low cloud fraction.
+A high global irradiance, in most cases, is the result of a low cloud fraction.
 In both years, the maximum recorded $Q_0$ was $750\,\unit{W\,m^{-2}}$ and reached in June.
@@
 Svanhovd is located within the subarctic climate zone.
-The 1992--2012 climatologies of temperature and precipitation are shown in a combined climate diagram (Fig.~\ref{fig:climatediagram}).
+Climatologies of temperature and precipitation derived for 1992--2012 are shown in a combined climate diagram (Fig.~\ref{fig:climatediagram}).
 Monthly averaged $2\,\unit{m}$ temperatures (red line) are displayed with a standard error of the mean (SE) as error band.
@@
 The highest monthly average temperature is $(13.1\pm 1.1)\,\unit{^\circ C}$ in July and the lowest $(-12.8\pm 2.0)\,\unit{^\circ C}$ in February.
-The coldest measured temperature was $-45.2\,\unit{^\circ C}$ in January 27, 1999, while the highest temperature ($29.4\,\unit{^\circ C}$) occurred in July 16 the same year.
+The coldest measured temperature was $-45.2\,\unit{^\circ C}$ in January 27, 1999, while the highest temperature ($29.4\,\unit{^\circ C}$) occurred on July 16 the same year.
 
@@
 Especially in May and July more than $40\,\unit{\%}$ of the days were significantly warmer.
-Significantly more warmer days continued to occur in all months (August--November) except for October.
+A significant number of warmer days continued to occur in all months (August--November) except for October.
 March 2018 had many unusually cold days.
-The summer of 2019 was fairly average, however, significantly more colder days occurred in July, while August had both significantly colder and warmer days.
+The summer of 2019 was fairly average, however, more significantly colder days occurred in July, while August had both significantly colder and warmer days.
 April 2019 had significantly warmer than average days.
@@
 The resulting percentage of days above/below $\pm 1\,\sigma$ is displayed for each month in Fig.~\ref{fig:anomalies_svanvik}d.
-Note that missing bars for months in winter and spring are both due to the temporal extend of our data and anomalies not meeting the $1\sigma$ criteria.
+Note that missing bars for months in winter and spring are both due to the temporal extent of our data and anomalies not meeting the $1\sigma$ criteria.
 The star indicates the reconstructed data in July.

### 2021-10-15 VegOzone.tex
Checked with grammarly

@@
 A longer growing season is prolonging the time in which vegetation can accumulate ozone.
-For natural and semi-natural vegetation which is already subject to rapid climatic changes known as Arctic amplification \citep{AMAP2012,IPCC2013}, these factors could promote a higher potential risk for northern vegetation to ozone-induced damage in the not so distant future.
+For natural and semi-natural vegetation that is already subject to rapid climatic changes known as Arctic amplification \citep{AMAP2012,IPCC2013}, these factors could promote a higher potential risk for northern vegetation to ozone-induced damage in the not so distant future.
 Recently, \citet{ESPR:Hayes2021} showed that the highest sensitivity to future \chem{[O_3]} variability is expected in summer when vegetation is most productive.
@@
 Substantial stomatal conductance in this species occurred only during an extended warm period in late July.
-In ecological terms, this means that there was almost no growth of grass in the summer of 2019 -- a prediction which is easily falsified by reality.
+In ecological terms, this means that there was almost no growth of grass in the summer of 2019 -- a prediction that is easily falsified by reality.
 We identified $f_\mathrm{T}$ as abnormally low, being the limiting factor of stomatal conductance in perennial grassland.
@@
 This is accomplished by moving $T_\mathrm{opt}$ towards cooler temperatures while keeping the other parameters fixed to MM values.
-In the same way, \emph{subarctic} is constructed to represent a species which is very tolerant to cold but sensitive to high temperatures, and most efficient at cool temperatures.
+In the same way, \emph{subarctic} is constructed to represent a species that is very tolerant to cold but sensitive to high temperatures, and most efficient at cool temperatures.
 Therefore, we shift $T\mathrm{min}$ and $T\mathrm{max}$ to colder temperatures and choose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000.
@@
 We presume that $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ around noon (highest light intensity) is a good proxy for \chem{CO_2} uptake efficiency (growth).
-In addition, a small deviation between noon and morning and a low standard deviation would indicate a higher robustness to variability in growing conditions.
+In addition, a small deviation between noon and morning and a low standard deviation would indicate higher robustness to variability in growing conditions.
 Based on these criteria, we identify \emph{subarctic}--PPFD0.8 as the best parameterization for all PFTs (Fig.~\ref{fig:javis_func_opt_mean}).
@@
 The net photosynthesis vs.
-time of year data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}).
+time of year data can be approximated by a second-order polynomial function (Fig.~\ref{fig:modis_Psn}).
 By calculating the roots of the fitted polynomial, we derive $A_\text{start}$ as \unit{doy}~122 for 2018 and \unit{doy}~106 in 2019.
@@
 The estimate of leaf boundary layer resistance is a function of wind speed which will be influenced by the tree height and leaf dimensions.
-A sample of downy birch leaves collected at Svanhovd had an average length of $(3.0\pm 0.5)\,\unit{cm}$ which is smaller than in the MM parameterization.
+A sample of downy birch leaves collected at Svanhovd had an average length of $(3.0\pm 0.5)\,\unit{cm}$ that is smaller than in the MM parameterization.
 \citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area.
 One Scots pine forest at Svanhovd has been studied as a part of the ICP Forest mapping.
-Heights were measured in 2004, when the stand was 90 years old.
+Heights were measured in 2004 when the stand was 90 years old.
 Mean tree height was $10.1\,\unit{m}$ and maximum height was $16.2\,\unit{m}$ (V.
@@
 We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for the three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using the generic MM parameterization and our bespoke temperature (\emph{cold}, \emph{subarctic}) and light (PPFD0.8, PPFD1.2) parameterizations.
-Canopy height and leaf size are adjusted to local values in the model simulations with bespoke growing season (GS besp.).
+Canopy height and leaf size are adjusted to local values in the model simulations with the bespoke growing season (GS besp.).
 The results displayed in Fig.~\ref{fig:pody_rel} allow for an assessment of over- or underestimation of ozone risk arising from the use of generic parameterizations established from less climatically extreme conditions.
@@
 Due to the shape of $f_\mathrm{light}$, a symmetric variation leads to an asymmetric response in $\mathrm{POD_1}$.
-We find that an opening of stomata at lower light intensities can cause a higher sensitivity to drought conditions.
+We find that an opening of stomata at lower light intensities can cause higher sensitivity to drought conditions.
 Similarly, a temperature acclimation to cold or subarctic conditions can lead to an amplification of effects by droughts.
@@
 The reported range corresponds to the maximum deviation defined by the $\pm 20\,\unit{\%}$ variation of $f_\mathrm{light}$ with SWP=on.
-Total biomass of deciduous trees and perennial grassland was substentially reduced in comparison with the default MM parameterizations in both years.
+The total biomass of deciduous trees and perennial grassland was substantially reduced in comparison with the default MM parameterizations in both years.
 The CL for perennial grassland is higher than for the other species.
-From an agro-economical point of view in our focus area, where grass is cut only once a year, a loss of biomass of more than $10\,\unit{\%}$ can have severer consequences compared to more productive areas in central Europe and may need adjustment.
+From an agro-economical point of view in our focus area, where the grass is cut only once a year, a loss of biomass of more than $10\,\unit{\%}$ can have severer consequences compared to more productive areas in central Europe and may need adjustment.
 
 Coniferous trees are the least affected and the magnitude of total biomass reduction is almost independent of the choice of the parameterization within given uncertainties.
-This may reflect that the temperature parameterization in the MM is almost never a limiting factor for stomatal opening.
+This may reflect that the temperature parameterization in the MM is rarely a limiting factor for stomatal opening.
 
@@
 We deduced the start of the growing season for deciduous trees based on gridded observational temperature and for perennial grassland on snow depth data.
-For perennial grassland, we assumed a one month latency period after snow melt which is supported by observational data from Rovaniemi (Finland) \citep{FCR:Korhonen2018}.
-With respect to ongoing climate change, a clear positive trend emerged in length ($5.2\,\unit{d\,decade^{-1}}$) of the growing season which is almost equally distributed between earlier start ($2.9\,\unit{days\,decade^{-1}}$) and later end ($2.3\,\unit{d\,decade^{-1}}$) (Appendix Fig.~\ref{fig:greening_season_change_Svanvik}).
+For perennial grassland, we assumed a one month latency period after snow melt that is supported by observational data from Rovaniemi (Finland) \citep{FCR:Korhonen2018}.
+With respect to ongoing climate change, a clear positive trend emerged in length ($5.2\,\unit{d\,decade^{-1}}$) of the growing season that is almost equally distributed between earlier start ($2.9\,\unit{days\,decade^{-1}}$) and later end ($2.3\,\unit{d\,decade^{-1}}$) (Appendix Fig.~\ref{fig:greening_season_change_Svanvik}).
 
@@
 The actual resilience of northern Fennoscandia plant communities to climate change strongly depend on both the range of existing acclimations as well as the actual type acclimation.
-This means vegetation which is acclimated to cold temperatures and little to heat- and drought stress might suffer more strongly from a more frequent occurance of heatwaves than plants which have a higher temperature tolerance.
+This means vegetation that is acclimated to cold temperatures and little to heat- and drought stress might suffer more strongly from a more frequent occurance of heatwaves than plants which have a higher temperature tolerance.
 
@@
 %
-Where $g_\mathrm{max}$ is the specific-specific maximum stomatal conductance value which is then modified by seasonal and environmental factors that vary within a range $0-1$.
+Where $g_\mathrm{max}$ is the specific-specific maximum stomatal conductance value that is then modified by seasonal and environmental factors that vary within a range $0-1$.
 They are empirically determined and account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_\mathrm{T}$), water vapor pressure deficit ($f_\mathrm{VPD}$), and soil water potential ($f_\text{SWP}$).

### 2021-10-15 VegOzone.tex
Checked

@@
     $^c$ $5\,\unit{days}-5\,\unit{^\circ C}$-rule, $T_\mathrm{air}$ from \href{seNorge.no}{seNorge.no};\\
-    $^d$ One month after snow melt; reference station Øvre Neiden (Sør-Varanger, NOR)} % Table Footnotes
+    $^d$ One month after snowmelt; reference station Øvre Neiden (Sør-Varanger, NOR)} % Table Footnotes
 \end{table}
@@
 We deduced the start of the growing season for deciduous trees based on gridded observational temperature and for perennial grassland on snow depth data.
-For perennial grassland, we assumed a one month latency period after snow melt that is supported by observational data from Rovaniemi (Finland) \citep{FCR:Korhonen2018}.
+For perennial grassland, we assumed an one-month latency period after snowmelt that is supported by observational data from Rovaniemi (Finland) \citep{FCR:Korhonen2018}.
 With respect to ongoing climate change, a clear positive trend emerged in length ($5.2\,\unit{d\,decade^{-1}}$) of the growing season that is almost equally distributed between earlier start ($2.9\,\unit{days\,decade^{-1}}$) and later end ($2.3\,\unit{d\,decade^{-1}}$) (Appendix Fig.~\ref{fig:greening_season_change_Svanvik}).
@@
  The $\mathrm{DO_3SE}$ modeling results are in contrast to our observation of more visible damage in 2018 compared to 2019.
-However, visible damage caused by peak ozone uptake and a dependence of ozone sensibility on phenology are not accounted for in $\mathrm{POD_y}$.
+However, visible damage caused by peak ozone uptake and dependence of ozone sensibility on phenology are not accounted for in $\mathrm{POD_y}$.
 In this regard, open-top chamber (OTC) experiments performed in northern Finland where Scots pine and downy birch seedlings were exposed to elevated ozone concentrations attributed a reduction in biomass and reproduction with visible damage explicitly to peak \chem{O_3} concentrations and fast phenological development at high growth rate \citep{Amb:Manninen2009}.
-In particular, forbs and perennial grassland have been found to be more susceptible to ozone-induced damage in the reproductive state \citep{EP:Bassin2004}.
+In particular, forbs and perennial grassland are more susceptible to ozone-induced damage in the reproductive state \citep{EP:Bassin2004}.
 That might explain why we observed damage in the ozone garden to a larger extent in 2018 than in 2019.
@@
 At the same time, intra-species variability even at the same location is typically non-negligible \citep{EP:Bassin2004} but not accounted for in the models.
-The literature results regarding ozone sensitivity of natural vegetation in northern Europe are contratictive.
+The literature results regarding the ozone sensitivity of natural vegetation in northern Europe are contradictive.
 \citet{FS:Subramanian2014} reported a higher, modeled reduction of net primary production in conifers ($4.3-15.5\,\unit{\%}$) than birch ($1.4-4.3\,\unit{\%}$) under elevated ozone in Sweden, while \citet{Amb:Girgzdiene2009} observed more visible damage on deciduous trees than on Norway spruce in Lithuania.
@@
 In particular, we assumed two tolerance levels to cold temperatures than for central European species.
-To this end, we computed a PDF from observed temperature at Svanhovd (May--September, 1992--2000).
+To this end, we computed a PDF from observed temperature at Svanhovd (May--September 1992--2000).
 We adjusted the parameters in the temperature response function of stomatal conductance to increase the enclosed area with the PDF.
@@
 In this regard, vegetation might become more exposed to higher ozone concentrations in early phenological stages due to an overlap with ozone spring peak \citep{EP:Karlsson2007}.
-However, the decline of this ozone spring peak is partly caused by uptake of vegetation.
-It, therefore, remains unclear whether an earlier start of the growing season will increase the exposure of vegetation to ozone or lead to a compensation due to an earlier decline of the peak ozone in the future.
+However, the decline of this ozone spring peak is partly caused by the uptake of vegetation.
+It, therefore, remains unclear whether an earlier start of the growing season will increase the exposure of vegetation to ozone or lead to compensation due to an earlier decline of the peak ozone in the future.
 \citet{ESPR:Hayes2021} pointed out that the highest susceptibility of vegetation to future ozone variability is likely to occur in summer when vegetation is most productive.
-The actual resilience of northern Fennoscandia plant communities to climate change strongly depend on both the range of existing acclimations as well as the actual type acclimation.
-This means vegetation that is acclimated to cold temperatures and little to heat- and drought stress might suffer more strongly from a more frequent occurance of heatwaves than plants which have a higher temperature tolerance.
+The actual resilience of northern Fennoscandia plant communities to climate change strongly depends on both the range of existing acclimations as well as the actual type of acclimation.
+This means vegetation that is acclimated to cold temperatures and little to heat- and drought stress might suffer more strongly from a more frequent occurence of heatwaves than plants that have a higher temperature tolerance.
 
@@
 In particular, CLs for deciduous and coniferous trees were breached in all years and for all parameterizations.
-For perennial grassland, the CL was not breached if a bespoke GS was assumed, but the mapping manual defined CL of $10\,\unit{\%}$ loss in biomass might not be transferable to subarctic regions with considerable constrains on productivity leading to more severe economical consequences.
+For perennial grassland, the CL was not breached if a bespoke GS was assumed, but the mapping manual defined CL of $10\,\unit{\%}$ loss in biomass might not be transferable to subarctic regions with considerable constraints on productivity leading to more severe economical consequences.
 
 Beyond the risk assessment of ozone-induced damage, our bespoke temperature response functions, once verified by observations, can have important implications on land-surface modeling in global models, where problems in productivity of species especially in the Arctic regions occur due to PFTs which are not suitable for the climate.
-We pointed out, with respect to default perennial grassland parameterization, that these perennial grasslands did not show any substantial stomatal conductance in 2019 characterized as a normal year.
+We pointed out, concerning default perennial grassland parameterization, that these perennial grasslands did not show any substantial stomatal conductance in 2019 characterized as a normal year.
 In terms of GPP this would mean unrealistically low carbon uptake.

### 2021-10-15 VegOzone.tex
Checked author contrib

@@
 SF has written the manuscript, collected and processed ozone and environmental data, and performed all statistical analyses.
-AVV has performed the on-site observation of vegetation damage induced by ozone, provided advice on plant physiological processes, collected exisiting literature on subarctic vegetation, contributed significantly to framing and proof reading this research article.
-LE has contributed with expertize in $\mathrm{DO_3SE}$ modeling and development of the PDF-based temperature acclimation assessment.
+AVV has conducted the on-site observation of vegetation damage induced by ozone, provided advice on plant physiological processes, collected existing literature on subarctic vegetation, contributed significantly to framing and proofreading this research article.
+LE has contributed with expertise in $\mathrm{DO_3SE}$ modeling and development of the PDF-based temperature acclimation assessment.
 CO has collected PFT parameters from the literature, performed all $\mathrm{DO_3SE}$ simulations and validation.
 AE contributed with her experience regarding subarctic vegetation in Finnmark.
-FS contributed with an assessment of the 2018 meteorological conditions and writing the proposal for founding this project.
+FS contributed with an assessment of the 2018 meteorological conditions and writing the proposal for funding this project.
 TB gave valuable guidance in a broader research sense.} %% this section is mandatory

### 2021-10-15 VegOzone.tex
Checked acknowledgements

@@
 \begin{acknowledgements}
-  We thank NILU for performing the ozone measurements at Svanhovd for us, Bj{\o}rg Rognerud (Department of Geosciences, UiO) for processing SeNorge.no data with respect to beginning of growing season, Tore Flatlandsmo Berglen (NILU) for hourly pressure data from Svanhovd, NIBIO Environment Center Svanhovd for establishing and running the ozone garden both years, and Volkmar Timmermann (NIBIO) for sharing ICP Forest data from Svanhovd.
+  We thank NILU for performing the ozone measurements at Svanhovd for us, Bj{\o}rg Rognerud (Department of Geosciences, UiO) for processing SeNorge.no data with respect to the beginning of the growing season, Tore Flatlandsmo Berglen (NILU) for hourly pressure data from Svanhovd, NIBIO Environment Center Svanhovd for establishing and running the ozone garden both years, and Volkmar Timmermann (NIBIO) for sharing ICP Forest data from Svanhovd.
 This work was supported by the Research Council of Norway (Grant No.

### 2021-10-15 VegOzone.tex
Improve  abstract
Upgrade author contributions

@@
 Efforts should be targeted towards accurately defining subarctic vegetations' physiological response to essential climate variables.
-That could help to improve regional and global scale biogeochemical cycling under current and future climates.
+Our method could help to improve regional and global scale biogeochemical cycling under current and future climates.
 \end{abstract}
@@
 
-\authorcontribution{All authors contributed to framing this research article, writing, and conclusions.
-SF has written the manuscript, collected and processed ozone and environmental data, and performed all statistical analyses.
-AVV has conducted the on-site observation of vegetation damage induced by ozone, provided advice on plant physiological processes, collected existing literature on subarctic vegetation, contributed significantly to framing and proofreading this research article.
-LE has contributed with expertise in $\mathrm{DO_3SE}$ modeling and development of the PDF-based temperature acclimation assessment.
+\authorcontribution{All authors contributed to conceptualization of this research article and commented on the manuscript.
+SF has prepared the original draft, collected and processed ozone and environmental data, and performed all statistical analyses.
+AVV has conducted the on-site observation of vegetation damage induced by ozone, provided advice on plant physiological processes, collected existing literature on subarctic vegetation, contributed significantly proofreading this research article.
+LE has contributed with expertise in $\mathrm{DO_3SE}$ modeling and suggestion of the PDF-based temperature acclimation methodology.
 CO has collected PFT parameters from the literature, performed all $\mathrm{DO_3SE}$ simulations and validation.
 AE contributed with her experience regarding subarctic vegetation in Finnmark.
-FS contributed with an assessment of the 2018 meteorological conditions and writing the proposal for funding this project.
-TB gave valuable guidance in a broader research sense.} %% this section is mandatory
+FS contributed with an assessment of the 2018 meteorological conditions.
+TB gave valuable guidance in a broader research sense.
+Funding acquisition for the project: AVV, AE, FS, TB.} %% this section is mandatory
 

### 2021-11-24 review/response-RC1.tex
Add draft for answers to referees.

% Authors' response
\documentclass{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{multirow}
\thispagestyle{empty}

\begin{document}
\section*{Authors' response}
To bg-2021-260-RC1 (05 Nov 2021):
We thank the anonymous referee \#1 for their comments.
We will address the \textbf{major comments} raised regarding the structure and readability of the manuscript in our next revision. We will refine the language to reduce ambiguity.\\
As the manuscript will, hence, undergo major structural changes, we have to refrain from addressing all \textbf{specific comments} in detail at this point. In the final revision of the manuscript, we will take all specific comments which are still relevant into consideration.

Here, we shall give a brief response to questions where it seems appropriate.

\subsection*{Specific comments} 
\begin{itemize}

    \item {\color{blue} }

\end{itemize}

\end{document}

### 2021-11-24 review/response-RC2.tex
Add draft for answers to referees.

% Authors' response
\documentclass{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{multirow}
\thispagestyle{empty}

\begin{document}
\section*{Authors' response}
To bg-2021-260-RC2 (23 Nov 2021):
We thank the anonymous referee \#2 for their very constructive comments and useful suggestions.
We will address the \textbf{major comments} raised regarding the structure and readability of the manuscript in our next revision. We will refine the language to reduce ambiguity and supplement our statements with appropriate citations.\\
As the manuscript will, hence, undergo major structural changes, we have to refrain from addressing all \textbf{specific comments} in detail at this point. In the final revision of the manuscript, we will take all specific and technical comments which are still relevant into consideration.

Here, we shall briefly respond to all issues where it seems appropriate.
\subsection*{General comments}
\begin{itemize}

    \item {\color{blue} "[...] main figures is missing [...]"} The referee is referring to Fig.~8 which is apparently absent from their version of the manuscript. We downloaded the pre-print from the associated discussion page (\href{https://bg.copernicus.org/preprints/bg-2021-260/}{bg-2021-260}) and cannot confirm Fig.~8 to be missing. It can be found on page~23 (just before the list of references). We, however, acknowledge that the current placement (after the main body of text) is rather unfortunate. This shall be fixed in the final typesetting process!

\end{itemize}

\subsection*{Specific comments}
\begin{itemize}

    \item {\color{blue} "[...] I find the introduction to be a bit longer than would be most effective for conveying this information. It would be helpful to break it into subsections [...]. [...]content of [...] lines 115--134 is important for contextualizing [...] could be better emphasized [...] or [...] opening paragraph for [...] section~2 [...]."} Thank you for this constructive comment. We will condense our introduction and put subsections in place where necessary. As suggested, we move the discussion of meteorology to Section~2.
    \item {\color{blue} "Please correct the text in line 105. [...] This doesn't seem to be a correct descriptor for Equation~2 [...]"} Thank you for pointing out this carelessness which should not have gone unnoticed.
    \item {\color{blue} "$f_\mathrm{T}$ hasn't been defined yet." "Is $f_\mathrm{T} = f_\mathrm{temp}$?"} These two issues are related. Of course, $f_\mathrm{T}$ is the same as $f_\mathrm{temp}$ and we should not have used both at the same time. We remove this ambiguity.
    \item {\color{blue} "Line 317-318 -- Mention that this is PPFD~0.8 ans 1.2 in parentheses [...]"} We add the suggested information to clarify our naming convention.
    \item {\color{blue} "Line 321 appears to substitute for $f_\mathrm{light}$" in Equation 4, not for as described in the text."} Indeed this is sentence ambiguous, and should read: "First, we solve Eq.~(3) for the MM default values of $\alpha_\mathrm{MM}$ at $f_\mathrm{light}=0.5$" We should, perhaps, clarify that we use $\gamma$ synonymously with PPFD in this paragraph or substitute $\gamma$ with PPFD.
    \item {\color{blue} "Line 331 -- Is the low standard deviation specific to nighttime?"} No, we have selected our sample (May--October), so that the vegetation will not experience "nighttime" between 5~am and 9~am. See, f.e. \href{https://www.timeanddate.com/sun/@777232}{TimeAndDate.com/Svanvik}.
    \item {\color{blue} "I find the conclusion that the standard deviation indicating "higher robustness to variability in growing conditions" to sweeping relative to the evidence precented. [...]"} This might be a misunderstanding due to inaccuracies in language? The robustness is not a conclusion but a presumption. We have not explicitly stated the associated standard deviation of our sampled essential climate variables which we could easily do to strengthen our point. We assume that the essential climate variables (excluding ozone) vary substantially in our 20 years time frame -- meaning having a large spread. A low standard deviation in stomatal conductance, therefore, suggests that stomatal conductance would to be relatively unaffected which we call "robustness". We will make this more clear.
    \item {\color{blue}}

\end{itemize}

\subsection*{Technical comments}
\begin{itemize}

    \item {\color{blue} d}

\end{itemize}

\end{document}

### 2021-11-25 review/response-RC2.tex
Add remark about do3se model description.

@@
 As the manuscript will, hence, undergo major structural changes, we have to refrain from addressing all \textbf{specific comments} in detail at this point.
-In the final revision of the manuscript, we will take all specific and technical comments which are still relevant into consideration.
-
-Here, we shall briefly respond to all issues where it seems appropriate.
+In the final revision of the manuscript, we will take all specific and technical comments which are still relevant into consideration.\\
+In the following, we shall briefly respond to all issues where it seems appropriate.
 \subsection*{General comments}
@@
 
-    \item {\color{blue} "[...] main figures is missing [...]"} The referee is referring to Fig.~8 which is apparently absent from their version of the manuscript.
+    \item {\color{blue} "[...] main figure is missing [...]"} The referee is referring to Fig.~8 which is apparently absent from their version of the manuscript.
 We downloaded the pre-print from the associated discussion page (\href{https://bg.copernicus.org/preprints/bg-2021-260/}{bg-2021-260}) and cannot confirm Fig.~8 to be missing.
@@
 [...] This doesn't seem to be a correct descriptor for Equation~2 [...]"} Thank you for pointing out this carelessness which should not have gone unnoticed.
+    \item {\color{blue} "Section~4 - Please expand the description of the $\mathrm{DO_3SE}$ model"} A comprehensive model description can be found in Appendix~B.
+As no modifications to the $\mathrm{DO_3SE}$ model were made (except for the change in parameters), we had concluded that an excessive model description would distract the reader.
+In particular, relationships between plant physiological parameters and meteorological observations are described in Appendix~B1.
+Model "high-level" input and its preparation is described in Appendix~B2.
+We agree that important information is therefore missing from the main text.
+In the course of restructuring, we include a condensed model description in Section~4.
     \item {\color{blue} "$f_\mathrm{T}$ hasn't been defined yet." "Is $f_\mathrm{T} = f_\mathrm{temp}$?"} These two issues are related.
@@
     \item {\color{blue} "Line 317-318 -- Mention that this is PPFD~0.8 ans 1.2 in parentheses [...]"} We add the suggested information to clarify our naming convention.
-    \item {\color{blue} "Line 321 appears to substitute for $f_\mathrm{light}$" in Equation 4, not for as described in the text."} Indeed this is sentence ambiguous, and should read: "First, we solve Eq.~(3) for the MM default values of $\alpha_\mathrm{MM}$ at $f_\mathrm{light}=0.5$" We should, perhaps, clarify that we use $\gamma$ synonymously with PPFD in this paragraph or substitute $\gamma$ with PPFD.
+    \item {\color{blue} "Line 321 appears to substitute for $f_\mathrm{light}$" in Equation 4, not for as described in the text."} Indeed this sentence is ambiguous, and should read: "First, we solve Eq.~(3) for the MM default values of $\alpha_\mathrm{MM}$ at $f_\mathrm{light}=0.5$" We should, perhaps, clarify that we use $\gamma$ synonymously with PPFD in this paragraph or substitute $\gamma$ with PPFD.
     \item {\color{blue} "Line 331 -- Is the low standard deviation specific to nighttime?"} No, we have selected our sample (May--October), so that the vegetation will not experience "nighttime" between 5~am and 9~am.
@@
 The robustness is not a conclusion but a presumption.
-We have not explicitly stated the associated standard deviation of our sampled essential climate variables which we could easily do to strengthen our point.
-We assume that the essential climate variables (excluding ozone) vary substantially in our 20 years time frame -- meaning having a large spread.
-A low standard deviation in stomatal conductance, therefore, suggests that stomatal conductance would to be relatively unaffected which we call "robustness".
-We will make this more clear.
-    \item {\color{blue}}
+We assume that the essential climate variables (excluding ozone) vary substantially in our sampled 20 years.
+Though, we have not explicitly shown the associated standard deviation of this subsample of our data.
+We could do this to strengthen our point.
+A low standard deviation in stomatal conductance, therefore, suggests that stomatal conductance is relatively unaffected by these variations.
+We refer to this as "robustness".
+We will make this more clear in the next revision.
+Or remove the whole point as suggested.
+    \item {\color{blue} "Line 332 -- "subarctic-PPFD0.8" is "best" relative to what?"} In this case "best" refers to the highest relative stomatal conductance and smallest standard deviation compared to the other probed combinations of parameter.
+We rephrase the sentence accordingly.
+    \item {\color{blue}"Figure~8 appears to be missing from the document."} As mentioned earlier, the placement of the figure on page 23 (after the main body of text) is unfortunate.
+The figure is not missing (at least as far as we can see), but might, therefore, have slipped the vigilant eye of the referee.
 
@@
 \subsection*{Technical comments}
+We will appropriately address all comments raised, but only address selected items at this point.
 \begin{itemize}
-
-    \item {\color{blue} d}
+    \item {\color{blue} Concerning Figures 3--5} Thank you for the useful suggestions to further improve our figures.
+We will take them into consideration.
+We may kindly ask to review Figure~8 as well.
+    \item {\color{blue} I haven't previously seen this method for identifying the start and end points of photosynthesis, and (noting that I'm not an ecologist) it seems elegant in concept and approach.} Thank you.
+This was a pragmatic, data-driven approach.
+To our knowledge, most satellite data based studies on start and end of growing season focus on correlations with essential climate variables to identify processes and relationships.
+    \item {\color{blue} "[...] please clarify what leaf length you are using [...]"} We have conducted two sets of simulations, one with leaf dimensions, tree height, and growing season as defined in the mapping manual and one with leaf dimensions of deciduous trees adjusted to the measured size.
+We have not adjusted the leaf dimensions of coniferous trees and grassland.
+We will make this clearer.
+    \item {\color{blue} "Line 457 -- do you mean the response to SWP, rather SWP itself being negligible?"} Of course! We apologize for the sloppy language.
 

### 2021-11-29 review/response-RC2.tex
Include answers to most remaining questions

@@
 We thank the anonymous referee \#2 for their very constructive comments and useful suggestions.
-We will address the \textbf{major comments} raised regarding the structure and readability of the manuscript in our next revision.
-We will refine the language to reduce ambiguity and supplement our statements with appropriate citations.\\
-As the manuscript will, hence, undergo major structural changes, we have to refrain from addressing all \textbf{specific comments} in detail at this point.
+We will address all \textbf{general comments} raised regarding the structure and readability of the manuscript in our next revision.
+We will refine the language to reduce ambiguity and supplement our statements with appropriate citations where indicated by the referees.\\
+As the manuscript will, hence, undergo major structural changes, we can only addressing some \textbf{specific comments} in detail at this point.
 In the final revision of the manuscript, we will take all specific and technical comments which are still relevant into consideration.\\
-In the following, we shall briefly respond to all issues where it seems appropriate.
+In the following, we shall respond to all issues where reasonable at this point.
 \subsection*{General comments}
@@
 It would be helpful to break it into subsections [...].
-[...]content of [...] lines 115--134 is important for contextualizing [...] could be better emphasized [...] or [...] opening paragraph for [...] section~2 [...]."} Thank you for this constructive comment.
+[...]content of [...] lines 115--134 is important for contextualizing [...] could be better emphasized [...] or [...] opening paragraph for [...] section~2 [...]."} Thank you for this very constructive comment.
 We will condense our introduction and put subsections in place where necessary.
 As suggested, we move the discussion of meteorology to Section~2.
+    \item {\color{blue}"Line 102 -- The concept of critical loads is much older than 2016 [...] please include an earlier citations [...]"} Indeed the concept is much older (1970s).
+Critical load is a term defined by the UNECE's Convention on Long Range Transboundary Air Pollution (LRTAP) and is generally associated with wet deposition of N and S.
+For ozone (and other pollutant gases (i.e.
+not in solution) we refer to 'critical levels'.
+Either concept is the level or load of the pollutant below which damage or injury would not be expected to occur to the receptor in question (e.g.
+forests, crops, grasslands etc...).
+We will include an appropriate citation.
     \item {\color{blue} "Please correct the text in line 105.
 [...] This doesn't seem to be a correct descriptor for Equation~2 [...]"} Thank you for pointing out this carelessness which should not have gone unnoticed.
+    \item {\color{blue} "In the paragraph that starts at line 241, the meaning of the word "significant" is unclear and varies within the paragraph.
+[...] it should be stated early on in this section.
+[...] tie back what is said to the objective of the paper."} Thank you.
+We will restructure the section to make the definition of the term "significance" easier assessable.
+The results shall be put into context with the paper's objective, either in-section or in a "results" Section.
     \item {\color{blue} "Section~4 - Please expand the description of the $\mathrm{DO_3SE}$ model"} A comprehensive model description can be found in Appendix~B.
@@
 We remove this ambiguity.
+    \item {\color{blue} "Line 304 (and Line 237-328) Again "extreme" is ambiguous here and seems to differ in definition from previous use [...]."} We will carefully evaluate the use of the term and define it appropriately.
+    \item {\color{blue} "Lines 306-307 -- Please clarify which time frames are being compared related to what is meant by "subject to climate change".
+In other words, are you comparing 1990s to 2000s?
+Or 1990s + 2000s with implicit impacts of climate relative to the preindustrial era?"} In the context of our study, the time frame for climate change is 2010s compared to early 1990s.
     \item {\color{blue} "Line 317-318 -- Mention that this is PPFD~0.8 ans 1.2 in parentheses [...]"} We add the suggested information to clarify our naming convention.
@@
 We rephrase the sentence accordingly.
+    \item {\color{blue}"Is there a citation or justification for the 1 nmol/m2/s flux threshold?"} This threshold (amongst others) has been introduced for modeling ozone induced damage on forest as part of the ICP Vegetation taskforce in the framework of LRTAP.
+"A uniform $\mathrm{O_3}$ flux threshold of $Y=1 \mathrm{nmol~m^{-2}~s^{-1}~PLA}$ was adopted for use in $\mathrm{POD_Y SPEC}$ for all tree species at the $\mathrm{O_3}$ Critical Levels workshop in Madrid, November 2016, based on data and analyses as presented in Büker et al.
+(2015).
+For the majority of tree species, this threshold fulfilled the recommendations that the confidence interval of the intercept includes 100\% and that the $R_2$ value is within 2\% of the maximum $R_2$ value (Büker et al., 2015)." (Mapping Manual; Mills et al.
+(2017)) We will include this in the revised manuscript.
     \item {\color{blue}"Figure~8 appears to be missing from the document."} As mentioned earlier, the placement of the figure on page 23 (after the main body of text) is unfortunate.
 The figure is not missing (at least as far as we can see), but might, therefore, have slipped the vigilant eye of the referee.
+    \item {\color{blue}"Lines 367-372 -- "leads to" seems strong, given that there are many variables changing [...]"} We change the sentence accordingly.
+    \item {\color{blue}"Where us ut shown that temperature acclimation relates to the amplification of drought effects?"} This is shown as part of Fig.~8 in which we summarize all sensitivity studies.
+Drought only affects the ozone dose in 2018 and not in 2019.
+The effect is stronger the more acclimated to cold temperatures our parameterizations are.
 
@@
 To our knowledge, most satellite data based studies on start and end of growing season focus on correlations with essential climate variables to identify processes and relationships.
+    \item {\color{blue}"Sentence in lines 339--340: "This value will be used for all PFTs alike".
+Please specify -- $\mathrm{A_{end}}$?
+In this paragraph please explain why it is acceptable to use the same $\mathrm{A_{end}}$ for all PFTs while $\mathrm{A_{start}}$ is specific."} It is still difficult to predict when deciduous vegetation will start the abscission process.
+For practical reasons, we set the date for all PFTs alike, although it could differ in reality.
+Coniferous trees are capable of photosynthesis as long as there is light which is the hard limit for all species alike.
+Grassland, on the other hand, will in most cases stop photosynthesis after it has been covered by snow.
     \item {\color{blue} "[...] please clarify what leaf length you are using [...]"} We have conducted two sets of simulations, one with leaf dimensions, tree height, and growing season as defined in the mapping manual and one with leaf dimensions of deciduous trees adjusted to the measured size.

### 2021-11-29 review/response-RC2.tex
Fix typo

@@
 We will refine the language to reduce ambiguity and supplement our statements with appropriate citations where indicated by the referees.\\
-As the manuscript will, hence, undergo major structural changes, we can only addressing some \textbf{specific comments} in detail at this point.
-In the final revision of the manuscript, we will take all specific and technical comments which are still relevant into consideration.\\
+As the manuscript will, hence, undergo major structural changes, we can not address all \textbf{specific / technical comments} in detail at this point.
+In the next revision of the manuscript, we will take all specific and technical comments which are still relevant into consideration.\\
 In the following, we shall respond to all issues where reasonable at this point.
+
 \subsection*{General comments}

### 2021-11-29 review/response-RC2.tex
Fine tuning.

--- 2021-11-29 review/response-RC2.tex
@@
 We will address all \textbf{general comments} raised regarding the structure and readability of the manuscript in our next revision.
-We will refine the language to reduce ambiguity and supplement our statements with appropriate citations where indicated by the referees.\\
-As the manuscript will, hence, undergo major structural changes, we can only addressing some \textbf{specific comments} in detail at this point.
-In the final revision of the manuscript, we will take all specific and technical comments which are still relevant into consideration.\\
-In the following, we shall respond to all issues where reasonable at this point.
+We will refine the language to reduce ambiguity and supplement our statements with appropriate citations where indicated by the referee.\\
+As the manuscript will, hence, undergo major structural changes, we can not address all \textbf{specific / technical comments} in detail at this point.
+In the next revision of the manuscript, we will take all specific and technical comments which are still relevant into consideration.\\
+In the following, we shall respond to all issues where reasonable.
+
 \subsection*{General comments}
@@
 We will condense our introduction and put subsections in place where necessary.
-As suggested, we move the discussion of meteorology to Section~2.
+As suggested, we'd move the discussion of meteorology to Section~2.
     \item {\color{blue}"Line 102 -- The concept of critical loads is much older than 2016 [...] please include an earlier citations [...]"} Indeed the concept is much older (1970s).
@@
 [...] it should be stated early on in this section.
+[...] Ideally it will pertain to a
+    significance test.
 [...] tie back what is said to the objective of the paper."} Thank you.
-We will restructure the section to make the definition of the term "significance" easier assessable.
+We will restructure the section to make the definition of the term "significance" easier assessable and unambiguous.
 The results shall be put into context with the paper's objective, either in-section or in a "results" Section.
+Regarding the significance test, we deliberately refrained from using p-values as their interpretation is often ill defined (for a discussion see \href{https://doi.org/10.1080/00031305.2016.1154108}{statement of the American Statistical Association (2016)}).
     \item {\color{blue} "Section~4 - Please expand the description of the $\mathrm{DO_3SE}$ model"} A comprehensive model description can be found in Appendix~B.
@@
 We agree that important information is therefore missing from the main text.
-In the course of restructuring, we include a condensed model description in Section~4.
+In the course of restructuring, we'd include a concise model description in Section~4.
     \item {\color{blue} "$f_\mathrm{T}$ hasn't been defined yet." "Is $f_\mathrm{T} = f_\mathrm{temp}$?"} These two issues are related.
@@
 Or 1990s + 2000s with implicit impacts of climate relative to the preindustrial era?"} In the context of our study, the time frame for climate change is 2010s compared to early 1990s.
-    \item {\color{blue} "Line 317-318 -- Mention that this is PPFD~0.8 ans 1.2 in parentheses [...]"} We add the suggested information to clarify our naming convention.
+    \item {\color{blue} "Line 317-318 -- Mention that this is PPFD~0.8 and 1.2 in parentheses [...]"} We add the suggested information to clarify our naming convention.
     \item {\color{blue} "Line 321 appears to substitute for $f_\mathrm{light}$" in Equation 4, not for as described in the text."} Indeed this sentence is ambiguous, and should read: "First, we solve Eq.~(3) for the MM default values of $\alpha_\mathrm{MM}$ at $f_\mathrm{light}=0.5$" We should, perhaps, clarify that we use $\gamma$ synonymously with PPFD in this paragraph or substitute $\gamma$ with PPFD.
@@
 We assume that the essential climate variables (excluding ozone) vary substantially in our sampled 20 years.
-Though, we have not explicitly shown the associated standard deviation of this subsample of our data.
+Though, we have not explicitly shown the associated standard deviation of this particular subsample (5--9 am and 11 am--1 pm).
 We could do this to strengthen our point.
@@
     \item {\color{blue}"Figure~8 appears to be missing from the document."} As mentioned earlier, the placement of the figure on page 23 (after the main body of text) is unfortunate.
-The figure is not missing (at least as far as we can see), but might, therefore, have slipped the vigilant eye of the referee.
+The figure is not missing (at least as far as we can tell), but might, therefore, have slipped the vigilant eye of the referee.
     \item {\color{blue}"Lines 367-372 -- "leads to" seems strong, given that there are many variables changing [...]"} We change the sentence accordingly.
-    \item {\color{blue}"Where us ut shown that temperature acclimation relates to the amplification of drought effects?"} This is shown as part of Fig.~8 in which we summarize all sensitivity studies.
-Drought only affects the ozone dose in 2018 and not in 2019.
+    \item {\color{blue}"Where is it shown that temperature acclimation relates to the amplification of drought effects?"} This is shown as part of Fig.~8 in which we summarize all sensitivity studies.
+Drought affects the ozone dose in 2018 only and not in 2019.
 The effect is stronger the more acclimated to cold temperatures our parameterizations are.
@@
 For practical reasons, we set the date for all PFTs alike, although it could differ in reality.
-Coniferous trees are capable of photosynthesis as long as there is light which is the hard limit for all species alike.
+Coniferous trees are capable of photosynthesis as long as there is light which is the hard limit for all species alike in our focus area.
 Grassland, on the other hand, will in most cases stop photosynthesis after it has been covered by snow.
-    \item {\color{blue} "[...] please clarify what leaf length you are using [...]"} We have conducted two sets of simulations, one with leaf dimensions, tree height, and growing season as defined in the mapping manual and one with leaf dimensions of deciduous trees adjusted to the measured size.
+    \item {\color{blue} "[...] please clarify what leaf length you are using [...]"} We have conducted two sets of simulations, one with leaf dimensions, tree height, and growing season as defined in the mapping manual and one with leaf dimensions of deciduous trees and hight of trees adjusted according to observations.
 We have not adjusted the leaf dimensions of coniferous trees and grassland.

### 2021-11-30 review/response-RC1.tex
Minor adjustments

@@
 We thank the anonymous referee \#1 for their comments.
-We will address the \textbf{major comments} raised regarding the structure and readability of the manuscript in our next revision. We will refine the language to reduce ambiguity.\\
-As the manuscript will, hence, undergo major structural changes, we have to refrain from addressing all \textbf{specific comments} in detail at this point.
-In the final revision of the manuscript, we will take all specific comments which are still relevant into consideration.
+We will address all \textbf{general comments} raised regarding the structure and readability of the manuscript in our next revision.
+We will refine the language to reduce ambiguity and supplement our statements with appropriate citations where indicated by the referee.\\
+As the manuscript will, hence, undergo major structural changes, we can not address all \textbf{specific / technical comments} in detail at this point.
+In the next revision of the manuscript, we will take all specific and technical comments which are still relevant into consideration.\\
+In the following, we shall respond to all issues where reasonable at this point.
 
-Here, we shall give a brief response to questions where it seems appropriate.
+\subsection*{Genral comments} 
+\begin{itemize}
+
+    \item {\color{blue} "the [...] study is highly speculative and not based on any experimental evidence [...]"} We are aware that our work is not complemented by experimental evidence and did not claim otherwise.
+We welcome relevant observational data to further improve parameterizations of Arctic and subarctic species.
+We ought to stress that we do not own the "true" parameterization for subarctic vegetation but assess the uncertainty arising from the use of generic parameterizations.
+    \item {\color{blue} "Many claims sound superficial [...], such as [...] the alleged deviation of the years 2018 and 2019 from the site's climatology [...]"} We strongly disagree with this statement.
+We assessed the particular years 2018 and 2019 by statistical means with respect to our derived climatologies.
+We acknowledge that we have not used common tools such as the infamous \emph{p-value} (for a discussion see \href{https://doi.org/10.1080/00031305.2016.1154108}{statement of the American Statistical Association (2016)}).
+    \item {\color{blue} "The manuscript is too long and the organization is confused  [...] no clear division between methodological section and results section"} We will address this in the next revision.
+
+\end{itemize}
 

### 2021-12-14 review/response-RC1.tex
Include comments and start to respond.

@@
 We thank the anonymous referee \#1 for their comments.
-We will address all \textbf{general comments} raised regarding the structure and readability of the manuscript in our next revision.
+We will address all \textbf{general comments} raised regarding the structure and readability of the manuscript in our next revision or resubmission.
 We will refine the language to reduce ambiguity and supplement our statements with appropriate citations where indicated by the referee.\\
 As the manuscript will, hence, undergo major structural changes, we can not address all \textbf{specific / technical comments} in detail at this point.
-In the next revision of the manuscript, we will take all specific and technical comments which are still relevant into consideration.\\
-In the following, we shall respond to all issues where reasonable at this point.
+We will take all specific and technical comments which are still relevant into consideration.\\
+In the following, we shall give a brief respond to all relevant issues.
 
-\subsection*{Genral comments} 
+\subsection*{General comments} 
 \begin{itemize}
@@
 
-    \item {\color{blue} }
+\item {\color{blue}Lines 51-52.
+"and leads to a build-up of ozone and its precursors during winter.” Are you sure?
+How can BVOC accumulate in the atmosphere if the vegetation is covered by snow?
+Please add some citations to support this claim.
+Much more credible is the following explanation based on stratospheric intrusions.}
+We agree with the referee that a vegetation covered by snow would not release BVOCs to the atmosphere.
+We implicitly assume this.
+Therefore, we did not explicitly that BVOCs are not among the ozone precursors that accumulate in the curse of winter.
+If this connection is not obvious to the readers, we will rephrase the sentence accordingly: "[...] ozone and its non-biogenic precursors [...]"
+
+\item {\color{blue}Line 63. "the time in which vegetation can accumulate ozone.” 
+It sounds very bad written this way: vegetation does not accumulate ozone because ozone it is not bioaccumulative. Did you mean the dose?}
+Thank you for pointing this out. We will rephrase accordingly and refer to the accumulated ozone dose.
+
+\item {\color{blue}Line 77.
+"($\left<[\mathrm{O_3}]\right>$ = 36-54ppb)” Please explain the formalism.
+What do different brackets mean?} The bracket notation commonly refers to the concentration of a chemical species.
+Parts-per-billion (ppb) is a volume mixing ratio (VMR) and strictly speaking no concentration.
+Though, they are often used interchangablly because measured VMR has been reported in concentrations (e.g.
+$\mu$gm$^{-3}$) by simply multiplying with a factor 2.
+
+\item {\color{blue}Line 81.
+"A substantial body of evidence exists that suggests flux-based metrics, that relate stomatal ozone uptake to vegetation damage, are biologically more relevant for risk assessments than exposure-based metrics.”
+Well, please cite at least some works of this "substantial body”}
+
+\item {\color{blue}Figure 1. This figure was never referred in the text.}
+
+\item {\color{blue}Figure 2.
+It does not seem to me that the O3 concentrations of 2019 are different from those of 2018.
+The spring peak could even be identical (although unknown, because in 2019 O3 measurements started about 20 days after the spring peak)}
+
+\item {\color{blue}Line 146. Please make clear the acronym PFT on first use}
+
+\item {\color{blue}Line 165. "luftkvalitet.no” What is it? And EBAS? Please make them clear.}
+
+\item {\color{blue}Line 182.
+"This indicates that the vegetation was more affected by ozone in 2018 than in 2019.” Being affected by visible symptoms does not necessarily mean having suffered biomass or productivity reduction.}
+
+\item {\color{blue}Line 190. "high ozone concentrations ([O3] > 40ppb” It is strange to read that O3 concentrations above 40 ppb are "high” concentrations.}
+
+\item {\color{blue}Line 194. "A method for gapfilling data has been presented in Falk et al. (2021).”
+Ok, but was it then applied to this work? Please write it.}
+
+\item {\color{blue}Line 205. "We evaluate the statistical significance of divergences from the norm in these variables (referred to as anomalies) in 2018/19”
+I suspect a misuse of the locution "statistical significance”. How was this significance assessed? Which statistical test was applied? What is the level of significance?}
+
+\item {\color{blue}Lines 206-208. I do not understand. Please, explicit the methodology.}
+
+\item {\color{blue}Line 213.
+"Averaged monthly accumulated precipitation (blue bars) is shown with standard deviation” It is not consistent to show SE once and STDEV the other time.
+The use of SE is more appropriate when estimating averages.}
+
+\item {\color{blue}Line 228. "Darker colors indicate higher probability to observe these values.” }
+
+\item {\color{blue}Line 229.
+"On top of the density distributions, a 10 days average of daily mean (h[O3]i10d) is displayed together with 1sigma uncertainties and SE, respectively” What does it mean?
+It is not clear to me.
+Why show a probability density if you are plotting a multiannual average?
+Or does the line represent the median instead?}
+
+\item {\color{blue}Line 232.
+"The decline in h[O3]i coincides with the average beginning of CO2 uptake by coniferous trees (Kolari et al., 2007; Wallin et al., 2013)” I didn't know that evergreens only uptake CO2 starting in May.
+I was convinced they always did.
+Is it true?
+Doesn't that contradict what you wrote in line 309 ("We base our temperature acclimation of coniferous trees on experimental results on Norway spruce which were found to be active already at rather low air temperatures and can reach 60\% photosynthetic activity as early as doy 100 (Kolariet al., 2007; Wallin et al., 2013).")?
+Here you state that photosynthesis is already active at DOY 100 and is at 60\% of its maximum!}
+
+\item {\color{blue}Line 233.
+"In July--September (doy 182--273), ozone is occasionally almost completely depleted.
+This hints to ozone uptake by vegetation even at low light intensities during midnight sun conditions in combination with stable planetary boundary layer conditions preventing mixing of ozone rich air.”
+I don't understand the connection. What does the night uptake have to do with the
+occasionally complete ozone depletion?}
+
+\item {\color{blue}Line 244. "if a normal distribution is assumed” 
+Are you sure that the distribution is normal and not lognormal or something else? There are some literature on the type of statistical distributions for variable such as Temperature, Rain, etc ...
+Moreover, looking at your Figure 6b the distribution of the irradiance seems to be a
+Poisson distribution.}
+
+\item {\color{blue}Figure 5, caption.
+"dashed lines indicates statistical significance” Statistical significance of what?
+By means of what test was it obtained, at what alpha level?
+And what are the numbers on the top right of each graph?}
+
+\item {\color{blue}Line 251.
+"deviated significantly from the climatology on the 1 sigma level.” Here the standard deviation is used as reference for the significance.
+But the significance of the deviation should be statistically tested in another way.}
+
+\item {\color{blue}Line 262.
+"We use the bias-corrected and cross-calibrated ozone climatology (Falk et al., 2021) and assess the monthly significance of the ozone concentration anomalies in 2018/19.”
+"Bias-corrected cross-calibrated” ozone? What is it? And what is the "significance” of the concentration anomalies? Please explain.}
+
+\item {\color{blue}Line 267.
+"Further, we presume that fVPD and fSWP suit our vegetation types and no acclimation is necessary for these.” This statement is questionable, because in cold conditions VPD can be high (you also told
+it in the conclusions) and the water in the soil can be limiting because partially unavailable due to freezing or other.}
+
+\item {\color{blue}Line 272.
+"but a substantially higher number of peak [O3] were observed in 2018 than in 2019.” How can you tell it if O3 measurements for all months of March, April and July are missing in 2019?
+I don't seem to see any differences between 2018 and 2019}
+
+\item {\color{blue}Line 291 " Note, however, that these parameterizations are hypothetical and have yet to be verified by experiments.” Figure 6a.
+Looking at the graph I understand that you assume an adaptation of the subarctic grasslands to the temperature distribution of the last decade (climate already changed) and not to the historical temperature distribution at your site.
+Is it reasonable to hypothesize such a rapid adaptation of vegetation to the new climate conditions?}
+
+\item {\color{blue}Line 298.
+"We construct cold as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to MM.
+This is accomplished by moving Topt towards cooler temperatures while keeping the other parameters fixed to MM values”.
+From Figure 6a and Table 1 I see that for the "cold” parameterization not only Topt was moved, but also Tmin for (e.g. for grassland).}
+
+\item {\color{blue}Figure 7.
+The gstom/gmax ratio in the subarctic parameterization with PPFD0.8 is greater in the morning than at noon.
+How then the choice of PPFD08 is explained?
+Please comment on this in the text.}
+\item {\color{blue}Line 334. I don't understand how we can say that the differences are "substantial". I don't see much difference between deciduous trees (a) and grassland (c), sorry.}
+
+\item {\color{blue}Line 336.
+Using net photosynthesis to calculate leaf emergence is not completely justified.
+Leaves are likely to be present and active well before gross photosynthesis equals heterotrophic respiration (eg.
+soil respiration).
+Gross photosynthesis should be used to calculate Astart and Aend instead.}
+
+\item {\color{blue}Line 354.
+"A sample of downy birch leaves collected at Svanhovd had an average length of (3.0±0.5)cm” Were top-canopy leaves sampled?
+How many leaves were collected to get +- 0.5 cm standard error?}
+
+\item {\color{blue}Line 355. "We used 13.5m height”
+Why was this value chosen? What is the meaning of a height between the average tree
+height and the maximum tree heigh? Perhaps it would have been more reasonable to use
+the average height.}
+
+\item {\color{blue}Line 360 and following.
+POD1 was calculated by gap filling the data, right?
+Because there is a lot of data missing in the middle of the season.
+Or were POD1 compensated for missing data?
+If so, how?
+Please confirm it by writing it in the text.}
+
+\item {\color{blue}Line 369. "Due to the shape of flight, a symmetric variation”.
+Symmetric variation of what?}
+
+\item {\color{blue}Line 370.
+"We find that an opening of stomata at lower light intensities can cause higher sensitivity to drought conditions.” Please, explain where we can see this.
+Graph 8 is not clear at all to me.
+And then, "sensitivity” of what?
+Of plants?
+Of POD1?}
+
+\item {\color{blue}Line 373.
+"The magnitude of these effects varies between PFTs as well as years, but the predicted ozone uptake for the bespoke temperature parameterization is always larger than for the MM parameterizations and of the same order of magnitude as the variability between the years studied here.”
+What effects? "Of the same order of magnitude as the interannual variability...”: can you conclude it by comparing only two years? The same for line 402}
+
+\item {\color{blue}Table 4.  Have the percentage of reduction been calculated taking into account pre-industrial concentrations as prescribed by the MM?
+What are meaning of the superscripts? And, above all, why some superscripts indicate a range (e.g. 1.9 ... 2.1)? I did not understand how the stdev of the MM estimation was calculated, sorry.}
+
+\item {\color{blue}Line 416. "we have developed bespoke parameterizations”
+it seems a bit strong statement to me, you have not developed any new tailored
+parameterization, you have only hypothesized one. There is no one experiment nor
+comparison with experimental results in your work.}
+
+\item {\color{blue}Line 417.
+"The comparison between meteorological conditions in 2018 and 2019 and their divergence from climatology allowed us to assess the influence of key environmental variables such as temperature, PPFD, and precipitation on vegetation susceptibility to O3 damage in light of future changes as may occur under climate change” I did not understand where all this "divergence with the climatological average" of these two years alone lies, sorry.}
+
+\item {\color{blue}Line 432.
+"With respect to ongoing climate change, a clear positive trend emerged in length (5.2d decade-1) of the growing season that is almost equally distributed between earlier start (2.9 days decade-1) and later end (2.3d decade-1) (Appendix Fig.
+A1).” How did you figure it out?
+Have you been doing retrospective MODIS analysis for 30 years?
+Or do you have a publication to quote?}
+
+\item {\color{blue}Line 435 and following. "visible damage”
+Visible damage and POD can be totally unrelated, as demonstrated by some research
+conducted on agricultural species. I recommend caution in stating that the O3 peaks
+causing the visible symptoms can result in a biomass reduction (damage).}
+
+\item {\color{blue}Line 441. Does "damage” mean "visible leaf symptoms”? Or does it mean biomass reduction?}
+
+\item {\color{blue}Line 456.
+"We found that soil water potential under 2018/19 meteorological conditions was negligible” What does it mean?
+That there was no water in the soil (SWP were negligible) or that the effect on the POD of the presence or absence of SWP in the calculation was negligible?
+Please clarify.}
+
+\item {\color{blue}Line 461. "better suited” Point 1 is questionable.
+Also point 2 is questionable. How can you say that the MM parameterization does not
+capture the plant physiology of subactic vegetation if no comparisons with physiological measurements taken on subarctic vegetation are presented?}
+
+\item {\color{blue}Line 469. "However, the decline of this ozone spring peak is partly caused by the uptake of vegetation" Are you sure? Please cite a reference.}
+
+\item {\color{blue}Line 491. "Automation of the here proposed PDF-based acclimation using machine learning techniques could overcome these issues in the future”
+What does it mean? Please explain. Make an example.}
+
+\item {\color{blue}Figure A1. How was the length of the growing seasons in the various years identified? By satellite? Other method? What does the gray band represent?}
+
+\item {\color{blue}Line 511.
+"with fmin, Dmin, Dmax describing the relative stomatal conductance to changes in vapor pressure deficit.” It is not clear.
+Please, clarify what D and fmin are, and their units.}
+
+\item {\color{blue}Line 517.
+"The DO3SE model as described in Büker et al.
+(2012) is used to simulate SWP0 across a PFT specific root depth according to the Penman--Monteith energy balance method that drives water cycling through the soil--plant--atmosphere system” I cannot understand how the P-M energy balance is used in DO3SE to derive the SWP.
+Please explain in detail.}
+
+\item {\color{blue}Line 525.
+"the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf” At what height was the O3 concentration measured?
+If it was not measured at the top of the canopy (10m for trees or 10 cm for grassland), how was the O3 concentration at the top canopy calculated?
+Please explain in detail.}
+
+\item {\color{blue}Line 526.  What does rc represent? Is it the cuticular resistance or the bulk canopy resistance? What is its value?}
+
+\item {\color{blue}Line 528.  Can you explain where that formula for calculating the flux comes from? Why is there u(z1) in? And what is the z1 height?}
+
+\item {\color{blue}Line 531. What is the z1 height? Where is it?}
+
+\item {\color{blue}Line 535. Wind speed at 2 m: what is it used for? Please explain}
+
+\item {\color{blue}Line 550. Please explicitly describe the method used to gap-fill O3 concentrations because it could be crucial.}
+
+\item {\color{blue}Section B1. The description of fPHEN is missing. Please, provide it. Again, how do you calculate the day-to-day SWP on your site? Please describe it in detail.}
+
+\item {\color{blue}Line 554. "From Fig. B1f) it is apparent that the mapping manual parameterized grassland would not have been able to grow in 2019.” 
+It does not seem to me that gstom has been reseted at all. If this is the case, the
+premises of the work appear weak.}
 

### 2021-12-15 review/response-RC1.tex
Continue response.

@@
 It sounds very bad written this way: vegetation does not accumulate ozone because ozone it is not bioaccumulative. Did you mean the dose?}
-Thank you for pointing this out. We will rephrase accordingly and refer to the accumulated ozone dose.
+Thank you for pointing this out. We will rephrase accordingly and refer instead the accumulated ozone dose.
 
@@
 Parts-per-billion (ppb) is a volume mixing ratio (VMR) and strictly speaking no concentration.
-Though, they are often used interchangablly because measured VMR has been reported in concentrations (e.g.
-$\mu$gm$^{-3}$) by simply multiplying with a factor 2.
+Though, they are often used synonymously because measured VMR is reported in concentrations by simply multiplying with a factor 2.
+So, perhaps, we should use another terminology instead, e.g.
+$\chi_{O_3}$ (\href{https://www.e-education.psu.edu/meteo300/node/534}{e-education.psu.edu}).
 
@@
 "A substantial body of evidence exists that suggests flux-based metrics, that relate stomatal ozone uptake to vegetation damage, are biologically more relevant for risk assessments than exposure-based metrics.”
-Well, please cite at least some works of this "substantial body”}
+Well, please cite at least some works of this "substantial body”} Thank you, we will include relevant citations.
 
 \item {\color{blue}Figure 1. This figure was never referred in the text.}
+This Figure has been referred to in Sect.~2 L177/179.
 
@@
 The spring peak could even be identical (although unknown, because in 2019 O3 measurements started about 20 days after the spring peak)}
+It is remarkable how the referee can infer from Fig.~2 that ozone concentrations do not differ between 2018 and 2019.
+Based on the observed visible damage on ozone sensitive vegetation in 2018 but not in 2019, our working hypothesis was that either ozone or one of the other important environmental variables may have differed.
+In Sect.~2, we therefore show whether and how they differ.
+We found that there are but only small differences.\\
+Regarding the spring peak, until the beginning of May most of the vegetation in the subarctic is still either covered by snow (see Table~3 for the dates) or bud burst has not yet occurred.
+Hence, it was unlikely for non-coniferous trees to experience the ozone spring peak in 2018/19.
+We will come back to coniferous trees later.
 
 \item {\color{blue}Line 146. Please make clear the acronym PFT on first use}
+Plant functional type (PFT) was introduced in the abstract. Following the common code of conduct, we ought to introduce it also in the main matter. Thank you for pointing this out.
 
 \item {\color{blue}Line 165. "luftkvalitet.no” What is it? And EBAS? Please make them clear.}
+"luftkvalitet.no” is the webinterface to the database of air quality measurements in Norway operated by NILU.
+EBAS is a database operated by NILU as well.
+It comprises global air quality measurements, e.g.
+trace gases, pollutants, particular matter from different networks.
+We will elaborate on this.
 
 \item {\color{blue}Line 182.
-"This indicates that the vegetation was more affected by ozone in 2018 than in 2019.” Being affected by visible symptoms does not necessarily mean having suffered biomass or productivity reduction.}
+"This indicates that the vegetation was more affected by ozone in 2018 than in 2019.” Being affected by visible symptoms does not necessarily mean having suffered biomass or productivity reduction.} It is correct that biomass reduction and visible damage due to ozone are not the same.
+We discuss this briefly in Sec.~5 but will elaborate on it to make the difference clearer.
 
 \item {\color{blue}Line 190. "high ozone concentrations ([O3] > 40ppb” It is strange to read that O3 concentrations above 40 ppb are "high” concentrations.}
+40\,ppb is not particularly high. It has been the threshold for the toxic accumulated ozone exposure (AOT) on ecosystems since its introduction at the UN ECE-Workshop in Egham in 1992.
 
@@
 Ok, but was it then applied to this work? Please write it.}
+We will rephrase: "A method for gapfilling data has been presented in Falk et al. (2021) and applied in this work."
 
@@
 I suspect a misuse of the locution "statistical significance”. How was this significance assessed? Which statistical test was applied? What is the level of significance?}
+We will elaborate on the description the devised hypothesis test.
+First, we compute the excess number of days with residuals larger than $1\sigma$.
+This means, we calculate the daily mean and std for 21 years $\times$ 24 hours (264 data points).
+From this, we find the residuals in 2018/19 and divide by std.
+In the next step, we explicitly assume that these are normal distributed - we will elaborate on this assumption in the next revision and check the underlying distributions (but see also Fig.~4).
+Due to the variability in the data, we choose a low sigma level for the significance test ($1\sigma$).
+By grouping the data into months, we find the number of days with observations in each month.
+We count the number of days with residuals exceeding $1\sigma$ and compute the ratio with the number of days of observation.
+If the data was normal distributed, we would expect ~18\% of data in this interval.
+If the percentage of days exceeding $1\sigma$ is larger, we state that we find a significant excess of days with high ozone.
 
 \item {\color{blue}Lines 206-208. I do not understand. Please, explicit the methodology.}
+We will rephrase and include a brief description of the method.
+We use a Reynolds decomposition on the 1986-1996 data from Svanvik and data from all other long-term monitoring stations in Fennoscandia (1992-2012) to establish multi-annual means.
+We then remove the annual cycle from the July 2018 observations at Pallas (reference station) to find the anomalies.
+We have shown that these correlate best with observations at Svanvik.
+We find the time lag between the observations and shift the time series of anomalies accordingly.
+We then add the climatology of Svanvik to retrieve a reconstruction of the missing data.
+As 2019 does not deviated much from the norm by means of temperature, precipitation, and irradiance, we assume it for a normal year.
+From this, we find a probable offset of the historical ozone climatology compared to present day and correct for this.
+We compared this reconstruction with high resolution air quality model reanalysis and found a similar accuracy (residuals compared to observation).
+For details, please refer the cited work Falk et al (2021).
 
@@
 The use of SE is more appropriate when estimating averages.}
+We will change the figure accordingly.
 
-\item {\color{blue}Line 228. "Darker colors indicate higher probability to observe these values.” }
-
-\item {\color{blue}Line 229.
+\item {\color{blue}Line 228.
+"Darker colors indicate higher probability to observe these values.”\\Line 229.
 "On top of the density distributions, a 10 days average of daily mean (h[O3]i10d) is displayed together with 1sigma uncertainties and SE, respectively” What does it mean?
@@
 Why show a probability density if you are plotting a multiannual average?
-Or does the line represent the median instead?}
+Or does the line represent the median instead?} In Fig.~4, we show the distribution of ozone data in 2-dimensions.
+The dotes represent the daily averages (but only every 10th day is shown).
+We will rephrase the caption to correct for this.
+We show the daily average to infer that not all daily mean ozone are strictly normal distributed (tails in one or the other direction) but it is also visually deductible that most data fall in the $1\sigma$ band justifying the chosen confidence intervals for a normal distribution.
+We shall, however, elaborate on this by statistical means.
 
@@
 Here you state that photosynthesis is already active at DOY 100 and is at 60\% of its maximum!}
+It is true that coniferous can, in principle, take up $\mathrm{CO_2}$ also in winter and the fact that they can suffer from frost drought shows that.
+With respect to the spring peak that was mentioned earlier, this means that coniferous trees may experience these elevated ozone concentrations regularly.
+Concerning the contradiction in the statements -- if the conditions are right, coniferous trees can quickly reach 60\% of their max photosynthesis.
+Conditions to keep photosynthesis up would occur from doy 100 onwards.
+To make sure that we do not falsely accumulate too much ozone from the spring peak periode (when coniferous trees show on-off photosynthesis), we consulted satellite observations to extrapolate the dates when vegetations' photosynthesis (including coniferous trees) at Svanhovd rose over the detectable threshold.
 
@@
 I don't understand the connection. What does the night uptake have to do with the
-occasionally complete ozone depletion?}
+occasionally complete ozone depletion?} This was actually a working hypothesis and should be removed from the text.
 
@@
 Poisson distribution.}
+The data shown in Fig.6 are not the same as used to calculate the results for Fig.5 because the data selection is different.
+But we should of course check and describe the underlying distribution of the data selection used in Fig.5.
 
@@
 And what are the numbers on the top right of each graph?}
+Regarding the test, see response to "L205" above.
+The number at the top right are described in the caption: "The annual positive/negative deviations are indicated in the respective corners (right upper/lower)." This is the percentage of days on an annual basis that exceed the $1\sigma$ level.
+This indicates the overall deviation from a normal year.
+Only temperatures in 2018 were abnormally high compared to the climatology on the $1\sigma$ level.
 
@@
 But the significance of the deviation should be statistically tested in another way.}
+Regarding the test, see response to "L205" above.
+Our hypothesis test is that you would expect $\approx18\%$ of days exceeding the $1\sigma$ level (normal distribution).
+If we observe a higher percentage of days, we find an exceedence from the expectation.
+If this is not proper in terms of statistical terminology we can remove the term "significant" and call it exceedence (in case this is deemed a better term).
 
@@
 "Bias-corrected cross-calibrated” ozone? What is it? And what is the "significance” of the concentration anomalies? Please explain.}
+Please see response to "Lines 206-208.".
+Cross-calibrated means that we used data from other monitoring stations in the region to reconstruct the missing data at Svanvik.
+Regarding "significance", see explanation above.
 
@@
 it in the conclusions) and the water in the soil can be limiting because partially unavailable due to freezing or other.}
+The referee is correct that VPD can be high due to cold conditions and thus become a limiting factor.
+We did not find any unexpected or abnormal behavior of stomatal conductance with respect to VPD which indicated that the VPD limitation on stomatal conductance might be more general applicable for high VPD caused by either drought or cold.
+We are not stating that the response to VPD or SWP could differ but in subarctic vegetation we have to assume that they are fair enough.
+Water availability due to frozen soil is most likely not represented in the $\mathrm{DO_3SE}$ model and would be worthwhile investigating further.
+(Lisa?)
 
@@
 I don't seem to see any differences between 2018 and 2019}
+As we have indicated, the measurement down time in March and April was planned because most vegetation (in particular the one in the ozone garden) was covered by snow and hence not photosynthetically active (as the referee has noted, too in their comment regarding BVOC emission).
+"but a substantially higher number of peak [O3] were observed in 2018 than in 2019.” refers back to Sect.~2 L188-195.
+Peak concentrations were calculated as exceedence over 40\,ppb in summer, explicitly excluding the spring peak.
+Even with two weeks of data missing July 2018, we found 50 elevated ozone event in 2018 compared to 18 times in 2019.
+This should qualify as "substantially higher number".
 
@@
 Looking at the graph I understand that you assume an adaptation of the subarctic grasslands to the temperature distribution of the last decade (climate already changed) and not to the historical temperature distribution at your site.
-Is it reasonable to hypothesize such a rapid adaptation of vegetation to the new climate conditions?}
+Is it reasonable to hypothesize such a rapid adaptation of vegetation to the new climate conditions?} Indeed it is not reasonable for long-lived species such as trees or coniferous dwarf scrubs and depends on the ecological turn-over periode.
+Competition would lead to a decrease in number of individuals of those species which natural acclimation range differs strongly from the new norm.
+If we look at a grassland as sum of all species the whole grassland may rapidly adapt.
+(Does this make sense Ane?
+Would Vigdis' results back this up?)
 
@@
 From Figure 6a and Table 1 I see that for the "cold” parameterization not only Topt was moved, but also Tmin for (e.g. for grassland).}
+Indeed, we will correct the text.
 

### 2021-12-16 review/response-RC1.tex
Include further responds

@@
 We assessed the particular years 2018 and 2019 by statistical means with respect to our derived climatologies.
-We acknowledge that we have not used common tools such as the infamous \emph{p-value} (for a discussion see \href{https://doi.org/10.1080/00031305.2016.1154108}{statement of the American Statistical Association (2016)}).
+We acknowledge that we have not used common tools such as the \emph{p-value} (for a discussion see \href{https://doi.org/10.1080/00031305.2016.1154108}{statement of the American Statistical Association (2016)}).
     \item {\color{blue} "The manuscript is too long and the organization is confused  [...] no clear division between methodological section and results section"} We will address this in the next revision.
@@
 \item {\color{blue}Lines 51-52.
-"and leads to a build-up of ozone and its precursors during winter.” Are you sure?
+"and leads to a build-up of ozone and its precursors during winter." Are you sure?
 How can BVOC accumulate in the atmosphere if the vegetation is covered by snow?
@@
 We implicitly assume this.
-Therefore, we did not explicitly that BVOCs are not among the ozone precursors that accumulate in the curse of winter.
-If this connection is not obvious to the readers, we will rephrase the sentence accordingly: "[...] ozone and its non-biogenic precursors [...]"
+Therefore, we did not explicitly that BVOCs are not among the ozone precursors that accumulate in the course of winter.
+We will rephrase the sentence accordingly: "[...] ozone and its non-biogenic precursors [...]"
 
-\item {\color{blue}Line 63. "the time in which vegetation can accumulate ozone.” 
+\item {\color{blue}Line 63. "the time in which vegetation can accumulate ozone." 
 It sounds very bad written this way: vegetation does not accumulate ozone because ozone it is not bioaccumulative. Did you mean the dose?}
@@
 \item {\color{blue}Line 77.
-"($\left<[\mathrm{O_3}]\right>$ = 36-54ppb)” Please explain the formalism.
+"($\left<[\mathrm{O_3}]\right>$ = 36-54ppb)" Please explain the formalism.
 What do different brackets mean?} The bracket notation commonly refers to the concentration of a chemical species.
 Parts-per-billion (ppb) is a volume mixing ratio (VMR) and strictly speaking no concentration.
-Though, they are often used synonymously because measured VMR is reported in concentrations by simply multiplying with a factor 2.
+Though, they are often used synonymously because measured VMR is reported in concentrations (e.g.
+$\mathrm{\mu g\,m^{-3}}$) by multiplying with a factor 2.
 So, perhaps, we should use another terminology instead, e.g.
@@
 \item {\color{blue}Line 81.
-"A substantial body of evidence exists that suggests flux-based metrics, that relate stomatal ozone uptake to vegetation damage, are biologically more relevant for risk assessments than exposure-based metrics.”
-Well, please cite at least some works of this "substantial body”} Thank you, we will include relevant citations.
+"A substantial body of evidence exists that suggests flux-based metrics, that relate stomatal ozone uptake to vegetation damage, are biologically more relevant for risk assessments than exposure-based metrics."
+Well, please cite at least some works of this "substantial body"} Thank you, we will include relevant citations.
 
@@
 The spring peak could even be identical (although unknown, because in 2019 O3 measurements started about 20 days after the spring peak)}
-It is remarkable how the referee can infer from Fig.~2 that ozone concentrations do not differ between 2018 and 2019.
 Based on the observed visible damage on ozone sensitive vegetation in 2018 but not in 2019, our working hypothesis was that either ozone or one of the other important environmental variables may have differed.
@@
 
-\item {\color{blue}Line 165. "luftkvalitet.no” What is it? And EBAS? Please make them clear.}
-"luftkvalitet.no” is the webinterface to the database of air quality measurements in Norway operated by NILU.
+\item {\color{blue}Line 165. "luftkvalitet.no" What is it? And EBAS? Please make them clear.}
+"luftkvalitet.no" is the webinterface to the database of air quality measurements in Norway operated by NILU.
 EBAS is a database operated by NILU as well.
@@
 trace gases, pollutants, particular matter from different networks.
-We will elaborate on this.
+We will clarify this.
 
 \item {\color{blue}Line 182.
-"This indicates that the vegetation was more affected by ozone in 2018 than in 2019.” Being affected by visible symptoms does not necessarily mean having suffered biomass or productivity reduction.} It is correct that biomass reduction and visible damage due to ozone are not the same.
+"This indicates that the vegetation was more affected by ozone in 2018 than in 2019." Being affected by visible symptoms does not necessarily mean having suffered biomass or productivity reduction.} It is correct that biomass reduction and visible damage due to ozone are not the same.
 We discuss this briefly in Sec.~5 but will elaborate on it to make the difference clearer.
 
-\item {\color{blue}Line 190. "high ozone concentrations ([O3] > 40ppb” It is strange to read that O3 concentrations above 40 ppb are "high” concentrations.}
-40\,ppb is not particularly high. It has been the threshold for the toxic accumulated ozone exposure (AOT) on ecosystems since its introduction at the UN ECE-Workshop in Egham in 1992.
+\item {\color{blue}Line 190. "high ozone concentrations ([O3] $>$ 40ppb" It is strange to read that O3 concentrations above 40 ppb are "high" concentrations.}
+Indeed 40\,ppb is not particularly "high" compared to peak ozone pollution downwind of densely populated areas, e.g.
+in Eastern Asia.
+We will remove the term "high".
+However, ozone concentrations of $40\,\mathrm{ppb}$ have been chosen as critical level for the accumulated ozone exposure (AOT) toxic to ecosystems.
+This metric was first introduced at the UN/ECE-Workshop in Egham in 1992.
+For the protection of crops and semi-natural vegetation (limit to 5\% yield reduction) and forest different targets exist (see e.g.
+\href{https://www.eea.europa.eu/publications/TOP08-98/page008.html}{EU directive}).
 
-\item {\color{blue}Line 194. "A method for gapfilling data has been presented in Falk et al. (2021).”
+\item {\color{blue}Line 194. "A method for gapfilling data has been presented in Falk et al. (2021)."
 Ok, but was it then applied to this work? Please write it.}
@@
 
-\item {\color{blue}Line 205. "We evaluate the statistical significance of divergences from the norm in these variables (referred to as anomalies) in 2018/19”
-I suspect a misuse of the locution "statistical significance”. How was this significance assessed? Which statistical test was applied? What is the level of significance?}
+\item {\color{blue}Line 205. "We evaluate the statistical significance of divergences from the norm in these variables (referred to as anomalies) in 2018/19"
+I suspect a misuse of the locution "statistical significance". How was this significance assessed? Which statistical test was applied? What is the level of significance?}
 We will elaborate on the description the devised hypothesis test.
@@
 \item {\color{blue}Line 213.
-"Averaged monthly accumulated precipitation (blue bars) is shown with standard deviation” It is not consistent to show SE once and STDEV the other time.
+"Averaged monthly accumulated precipitation (blue bars) is shown with standard deviation" It is not consistent to show SE once and STDEV the other time.
 The use of SE is more appropriate when estimating averages.}
@@
 \item {\color{blue}Line 228.
-"Darker colors indicate higher probability to observe these values.”\\Line 229.
-"On top of the density distributions, a 10 days average of daily mean (h[O3]i10d) is displayed together with 1sigma uncertainties and SE, respectively” What does it mean?
+"Darker colors indicate higher probability to observe these values."\\Line 229.
+"On top of the density distributions, a 10 days average of daily mean (h[O3]i10d) is displayed together with 1sigma uncertainties and SE, respectively" What does it mean?
 It is not clear to me.
@@
 \item {\color{blue}Line 232.
-"The decline in h[O3]i coincides with the average beginning of CO2 uptake by coniferous trees (Kolari et al., 2007; Wallin et al., 2013)” I didn't know that evergreens only uptake CO2 starting in May.
+"The decline in h[O3]i coincides with the average beginning of CO2 uptake by coniferous trees (Kolari et al., 2007; Wallin et al., 2013)" I didn't know that evergreens only uptake CO2 starting in May.
 I was convinced they always did.
@@
 "In July--September (doy 182--273), ozone is occasionally almost completely depleted.
-This hints to ozone uptake by vegetation even at low light intensities during midnight sun conditions in combination with stable planetary boundary layer conditions preventing mixing of ozone rich air.”
+This hints to ozone uptake by vegetation even at low light intensities during midnight sun conditions in combination with stable planetary boundary layer conditions preventing mixing of ozone rich air."
 I don't understand the connection. What does the night uptake have to do with the
@@
 
-\item {\color{blue}Line 244. "if a normal distribution is assumed” 
+\item {\color{blue}Line 244. "if a normal distribution is assumed" 
 Are you sure that the distribution is normal and not lognormal or something else? There are some literature on the type of statistical distributions for variable such as Temperature, Rain, etc ...
@@
 \item {\color{blue}Figure 5, caption.
-"dashed lines indicates statistical significance” Statistical significance of what?
+"dashed lines indicates statistical significance" Statistical significance of what?
 By means of what test was it obtained, at what alpha level?
@@
 \item {\color{blue}Line 251.
-"deviated significantly from the climatology on the 1 sigma level.” Here the standard deviation is used as reference for the significance.
+"deviated significantly from the climatology on the 1 sigma level." Here the standard deviation is used as reference for the significance.
 But the significance of the deviation should be statistically tested in another way.}
@@
 \item {\color{blue}Line 262.
-"We use the bias-corrected and cross-calibrated ozone climatology (Falk et al., 2021) and assess the monthly significance of the ozone concentration anomalies in 2018/19.”
-"Bias-corrected cross-calibrated” ozone? What is it? And what is the "significance” of the concentration anomalies? Please explain.}
+"We use the bias-corrected and cross-calibrated ozone climatology (Falk et al., 2021) and assess the monthly significance of the ozone concentration anomalies in 2018/19."
+"Bias-corrected cross-calibrated" ozone? What is it? And what is the "significance" of the concentration anomalies? Please explain.}
 Please see response to "Lines 206-208.".
@@
 \item {\color{blue}Line 267.
-"Further, we presume that fVPD and fSWP suit our vegetation types and no acclimation is necessary for these.” This statement is questionable, because in cold conditions VPD can be high (you also told
+"Further, we presume that fVPD and fSWP suit our vegetation types and no acclimation is necessary for these." This statement is questionable, because in cold conditions VPD can be high (you also told
 it in the conclusions) and the water in the soil can be limiting because partially unavailable due to freezing or other.}
@@
 \item {\color{blue}Line 272.
-"but a substantially higher number of peak [O3] were observed in 2018 than in 2019.” How can you tell it if O3 measurements for all months of March, April and July are missing in 2019?
+"but a substantially higher number of peak [O3] were observed in 2018 than in 2019." How can you tell it if O3 measurements for all months of March, April and July are missing in 2019?
 I don't seem to see any differences between 2018 and 2019}
 As we have indicated, the measurement down time in March and April was planned because most vegetation (in particular the one in the ozone garden) was covered by snow and hence not photosynthetically active (as the referee has noted, too in their comment regarding BVOC emission).
-"but a substantially higher number of peak [O3] were observed in 2018 than in 2019.” refers back to Sect.~2 L188-195.
+"but a substantially higher number of peak [O3] were observed in 2018 than in 2019." refers back to Sect.~2 L188-195.
 Peak concentrations were calculated as exceedence over 40\,ppb in summer, explicitly excluding the spring peak.
@@
 
-\item {\color{blue}Line 291 " Note, however, that these parameterizations are hypothetical and have yet to be verified by experiments.” Figure 6a.
+\item {\color{blue}Line 291 " Note, however, that these parameterizations are hypothetical and have yet to be verified by experiments." Figure 6a.
 Looking at the graph I understand that you assume an adaptation of the subarctic grasslands to the temperature distribution of the last decade (climate already changed) and not to the historical temperature distribution at your site.
@@
 "We construct cold as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to MM.
-This is accomplished by moving Topt towards cooler temperatures while keeping the other parameters fixed to MM values”.
-From Figure 6a and Table 1 I see that for the "cold” parameterization not only Topt was moved, but also Tmin for (e.g. for grassland).}
+This is accomplished by moving Topt towards cooler temperatures while keeping the other parameters fixed to MM values".
+From Figure 6a and Table 1 I see that for the "cold" parameterization not only Topt was moved, but also Tmin for (e.g. for grassland).}
 Indeed, we will correct the text.
@@
 Please comment on this in the text.}
+PPFD0.8 was chosen to maximize the relative stomatal conductance at noon and in the morning. Please refer to L326--334. We will elaborate on the text and make it clearer.
+
 \item {\color{blue}Line 334. I don't understand how we can say that the differences are "substantial". I don't see much difference between deciduous trees (a) and grassland (c), sorry.}
+There seems to be a misunderstanding?
+The referee refers to the sentence "As expected due to the small adjustments, coniferous trees display the smallest differences between the different parameterizations, while the differences for perennial grassland are substantial as a response to the proposed temperature acclimation." Figure~7 is meant to be read for each species separately and from right to left.
+It comprises the mean and standard deviation of relative stomatal conductance computed from hourly meteorological data over 21~years.
+Panel (c) displays perennial grassland.
+Furthest to the right, the results using the mapping manual parameterization are shown.
+The relative stomatal conductance around noon when the highest values are expected is rather low (about 40\%).
+With our adjusted parameterization the relative stomatal conductance around noon reaches about 70--80\%.
+The standard deviation reflects the variability of growing conditions and should not be interpreted as statistical measure for to separate one distribution from the other.
+Even more data would not reduce the standard deviation in this case.
+We will remove the term "substantial" and make our interpretation of the Figure clearer in the text.
 
@@
 Gross photosynthesis should be used to calculate Astart and Aend instead.}
+Thank you for pointing this out.
+It is correct that net photosynthesis does not capture the real photosynthesis.
+However, gross photosynthesis is not among the available products from MODIS satellites.
+Instead, we use gross primary production which is including maintenance respiration.
+According to the product website (\href{http://www.ntsg.umt.edu/project/modis/mod17.php}{MOD17}, last accessed Dec.
+2021) it is computed from observed absorbed photosynthetically active radiation (APAR), a species specific conversion efficiency parameter $\epsilon(\mathrm{temperature, water stress})$, temperature, and water stress.
+In Fig.~\ref{fig:modis_gpp} the estimated start and end of growing season from MODIS (Aqua/Terra) GPP over a $1\times 1$\,km patch centered at Svanhovd is shown.
+They differ only slightly from the estimates using net photosynthesis.
+The start of the growing season is shifted by one day (later) and the end by one/two days (later).
+Considering the temporal resolution of 8\,days of the satellite product this difference is negligible.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.75\textwidth]{./modis_Gpp.png}
+    \caption{Estimated start and end of growing season from MODIS (Aqua/Terra) GPP over a $1\times 1$\,km patch centered at Svanhovd.
+The average start in 2018 amounts to doy 123 and to 107 in 2019.
+The end of the growing season in 2018 is doy 262 and in 2019 276.}
+    \label{fig:modis_gpp}
+\end{figure}
 
 \item {\color{blue}Line 354.
-"A sample of downy birch leaves collected at Svanhovd had an average length of (3.0±0.5)cm” Were top-canopy leaves sampled?
+"A sample of downy birch leaves collected at Svanhovd had an average length of (3.0±0.5)cm" Were top-canopy leaves sampled?
 How many leaves were collected to get +- 0.5 cm standard error?}
+Ane?
 
-\item {\color{blue}Line 355. "We used 13.5m height”
+\item {\color{blue}Line 355. "We used 13.5m height"
 Why was this value chosen? What is the meaning of a height between the average tree
@@
 the average height.}
+We suppose, the referee refers to the 10.1\,m average height of one particular Scots pine forest which has been measured in 2004.
+We assumed that after 14 year 10.1\,m might not be the average height of this particular forest anymore.
+In the absence of more recent measurements, we found it more reasonable to choose 13.5\,m the average tree height in the whole area in 2004 which included all tree species and ages.
+We will include this in the text.
 
@@
 Please confirm it by writing it in the text.}
+In Appendix~B2, we write that the "$\mathrm{DO_3SE}$ model requires hourly, continuous meteorologische observations." This means that all input data, including ozone, have to be gap filled before POD1 can be computed.
+The gap filling method for all data but ozone are described in appendix~B2.
+The used gap filling method for ozone has been summarized above (response to "Lines 206--208") and published in Falk et al.
+(2021).
+We shall restructure the text to make this more clear.
 
-\item {\color{blue}Line 369. "Due to the shape of flight, a symmetric variation”.
+\item {\color{blue}Line 369. "Due to the shape of flight, a symmetric variation".
 Symmetric variation of what?}
+This refers to the variation of PPFD by $\pm 20\%$ at a stomatal opening of 50\%. due to the exponential shape of $f_\mathrm{light}$.
 
 \item {\color{blue}Line 370.
-"We find that an opening of stomata at lower light intensities can cause higher sensitivity to drought conditions.” Please, explain where we can see this.
+"We find that an opening of stomata at lower light intensities can cause higher sensitivity to drought conditions." Please, explain where we can see this.
 Graph 8 is not clear at all to me.
-And then, "sensitivity” of what?
+And then, "sensitivity" of what?
 Of plants?
@@
 \item {\color{blue}Line 373.
-"The magnitude of these effects varies between PFTs as well as years, but the predicted ozone uptake for the bespoke temperature parameterization is always larger than for the MM parameterizations and of the same order of magnitude as the variability between the years studied here.”
-What effects? "Of the same order of magnitude as the interannual variability...”: can you conclude it by comparing only two years? The same for line 402}
+"The magnitude of these effects varies between PFTs as well as years, but the predicted ozone uptake for the bespoke temperature parameterization is always larger than for the MM parameterizations and of the same order of magnitude as the variability between the years studied here."
+What effects? "Of the same order of magnitude as the interannual variability...": can you conclude it by comparing only two years? The same for line 402}
 
@@
 
-\item {\color{blue}Line 416. "we have developed bespoke parameterizations”
+\item {\color{blue}Line 416. "we have developed bespoke parameterizations"
 it seems a bit strong statement to me, you have not developed any new tailored
@@
 \item {\color{blue}Line 417.
-"The comparison between meteorological conditions in 2018 and 2019 and their divergence from climatology allowed us to assess the influence of key environmental variables such as temperature, PPFD, and precipitation on vegetation susceptibility to O3 damage in light of future changes as may occur under climate change” I did not understand where all this "divergence with the climatological average" of these two years alone lies, sorry.}
+"The comparison between meteorological conditions in 2018 and 2019 and their divergence from climatology allowed us to assess the influence of key environmental variables such as temperature, PPFD, and precipitation on vegetation susceptibility to O3 damage in light of future changes as may occur under climate change" I did not understand where all this "divergence with the climatological average" of these two years alone lies, sorry.}
 
@@
 "With respect to ongoing climate change, a clear positive trend emerged in length (5.2d decade-1) of the growing season that is almost equally distributed between earlier start (2.9 days decade-1) and later end (2.3d decade-1) (Appendix Fig.
-A1).” How did you figure it out?
+A1)." How did you figure it out?
 Have you been doing retrospective MODIS analysis for 30 years?
@@
 
-\item {\color{blue}Line 435 and following. "visible damage”
+\item {\color{blue}Line 435 and following. "visible damage"
 Visible damage and POD can be totally unrelated, as demonstrated by some research
@@
 
-\item {\color{blue}Line 441. Does "damage” mean "visible leaf symptoms”? Or does it mean biomass reduction?}
+\item {\color{blue}Line 441. Does "damage" mean "visible leaf symptoms"? Or does it mean biomass reduction?}
 
 \item {\color{blue}Line 456.
-"We found that soil water potential under 2018/19 meteorological conditions was negligible” What does it mean?
+"We found that soil water potential under 2018/19 meteorological conditions was negligible" What does it mean?
 That there was no water in the soil (SWP were negligible) or that the effect on the POD of the presence or absence of SWP in the calculation was negligible?
@@
 
-\item {\color{blue}Line 461. "better suited” Point 1 is questionable.
+\item {\color{blue}Line 461. "better suited" Point 1 is questionable.
 Also point 2 is questionable. How can you say that the MM parameterization does not
@@
 
-\item {\color{blue}Line 491. "Automation of the here proposed PDF-based acclimation using machine learning techniques could overcome these issues in the future”
+\item {\color{blue}Line 491. "Automation of the here proposed PDF-based acclimation using machine learning techniques could overcome these issues in the future"
 What does it mean? Please explain. Make an example.}
@@
 \item {\color{blue}Line 511.
-"with fmin, Dmin, Dmax describing the relative stomatal conductance to changes in vapor pressure deficit.” It is not clear.
+"with fmin, Dmin, Dmax describing the relative stomatal conductance to changes in vapor pressure deficit." It is not clear.
 Please, clarify what D and fmin are, and their units.}
@@
 "The DO3SE model as described in Büker et al.
-(2012) is used to simulate SWP0 across a PFT specific root depth according to the Penman--Monteith energy balance method that drives water cycling through the soil--plant--atmosphere system” I cannot understand how the P-M energy balance is used in DO3SE to derive the SWP.
+(2012) is used to simulate SWP0 across a PFT specific root depth according to the Penman--Monteith energy balance method that drives water cycling through the soil--plant--atmosphere system" I cannot understand how the P-M energy balance is used in DO3SE to derive the SWP.
 Please explain in detail.}
@@
 \item {\color{blue}Line 525.
-"the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf” At what height was the O3 concentration measured?
+"the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf" At what height was the O3 concentration measured?
 If it was not measured at the top of the canopy (10m for trees or 10 cm for grassland), how was the O3 concentration at the top canopy calculated?
@@
 
-\item {\color{blue}Line 554. "From Fig. B1f) it is apparent that the mapping manual parameterized grassland would not have been able to grow in 2019.” 
+\item {\color{blue}Line 554. "From Fig. B1f) it is apparent that the mapping manual parameterized grassland would not have been able to grow in 2019." 
 It does not seem to me that gstom has been reseted at all. If this is the case, the

### 2021-12-16 review/response-RC1.tex
Small additions

@@
 (2021).
-We shall restructure the text to make this more clear.
+We shall restructure the text to make the description of the $\mathrm{DO_3SE}$ model and its input data clearer.
 
@@
 Symmetric variation of what?}
-This refers to the variation of PPFD by $\pm 20\%$ at a stomatal opening of 50\%. due to the exponential shape of $f_\mathrm{light}$.
+This refers to the variation of PPFD at a stomatal opening of 50\% by $\pm 20\%$.
 
@@
 Of POD1?}
+Figure~8 is indeed very complex and we will elaborate on the explanation and interpretation in the text. Sensitivity of POD1 is indicated.
 

### 2021-12-17 review/response-RC1.tex
Finish first draft

@@
 We will address all \textbf{general comments} raised regarding the structure and readability of the manuscript in our next revision or resubmission.
+We consider to separate the manuscript in Introduction--Methods--Results--Conclusions and will improve on the "methods" Section to make our results more comprehensible.
 We will refine the language to reduce ambiguity and supplement our statements with appropriate citations where indicated by the referee.\\
@@
 Symmetric variation of what?}
-This refers to the variation of PPFD at a stomatal opening of 50\% by $\pm 20\%$.
+This refers to the variation of PPFD at a stomatal opening of 50\% by $\pm 20\%$ and the resulting response in computed POD1. We will make this clearer in the text.
 
@@
 Of POD1?}
-Figure~8 is indeed very complex and we will elaborate on the explanation and interpretation in the text. Sensitivity of POD1 is indicated.
+Figure~8 is indeed very complex and we will elaborate on the explanation and interpretation in the text to make it more accessible to the readers.
+You are right that this sentence is not clear and does not communicate what we intended to say.
+We refer to the sensitivity of modeled POD1.
+If we take a look at, e.g.
+Fig.~8a subarctic parameterazation and 2018 (left hand side of the upper panel).
+Open symbols represent a model simulation where SWP was taken into account.
+Closed symbols where this effect was switched off.
+The '--' symbolizes PPFD0.8 (earlier opening), the '+' PPFD1.2 (later opening).
+We shall have a look at the circles connected by a solide line representing the simulations with a GS which we have determined from temperatures.
+If the stomata open earlier / close later more ozone can be taken up (longer opening time), hence POD1 is larger than for the unchanged $f_\mathrm{light}$.
+Now, if we take SWP into account, we find that only the PPFD0.8 is affected.
+Therefore, we conclude that conditions that negatively affect SWP (referred to as droughts) reduce the uptake of ozone (POD1) when an earlier opening / later closing of stomata is considered.
+Hence, POD1 is more sensitive to drought conditions.
 
@@
 What effects? "Of the same order of magnitude as the interannual variability...": can you conclude it by comparing only two years? The same for line 402}
+In Section~3, we tried to show that 2019 is representative for a normal year while 2018 is more extreme.
+We should evaluate our data with this question in mind and confirm that the difference between 2018 and 2019 is indeed representative for the interannual variability.
+Depending on the outcome, we will rephrase the sentences and substitute the term "interannual variability".
 
@@
 What are meaning of the superscripts? And, above all, why some superscripts indicate a range (e.g. 1.9 ... 2.1)? I did not understand how the stdev of the MM estimation was calculated, sorry.}
+We will make this clearer in the next version of the manuscript.
+We used the relationship between biomass reduction and POD1 unchanged from the MM.
+This, however, is to some degree questionable because subarctic vegetation could be affected more or less by the same ozone dose than the central European species.
+Apparently, the caption of the table is not complete and parts of the explanation missing.
+We explain how we derive the uncertainty ranges in L405--407 but this may not be clear enough.
+From Fig.~8, we deduced that SWP is negligible in most cases (solid and open symbols show the same POD1).
+Hence, we computed the biomass reduction based on simulations with SWP effects on POD1 turned off and MM $f_\mathrm{light}$.
+The uncertainties reported are actually the differences and not standard deviations.
+The differences are calculated between the simulation with MM $f_\mathrm{light}$ and "bespoke" GS and the simulations with PPFD0.8 (larger POD1 $\rightarrow$ larger biomass reduction) and PPFD1.2 (smaller POD1 $\rightarrow$ less biomass reduction).
+The sub- and superscripts show the asymmetry introduced by the different $f_\mathrm{light}$ parameterizations.
+The range represents the additional uncertainty from the choice in GS.
 
@@
 comparison with experimental results in your work.}
+We will substitute the term to appropriately reflect our intensions and work.
 
@@
 "The comparison between meteorological conditions in 2018 and 2019 and their divergence from climatology allowed us to assess the influence of key environmental variables such as temperature, PPFD, and precipitation on vegetation susceptibility to O3 damage in light of future changes as may occur under climate change" I did not understand where all this "divergence with the climatological average" of these two years alone lies, sorry.}
+We will make this clearer in the revision of the manuscript.
 
@@
 Or do you have a publication to quote?}
+We have derived the number from an analysis of the thermal start and end of the GS based on the gridded temperature data from SeNorge.no provided by the Norwegian weather service and shown in Appendix Fig.~9.
+We may remove this part eventually or rephrase the sentence to include this information, support it with citations of relevant works, and include this analysis in a methods section.
 
@@
 causing the visible symptoms can result in a biomass reduction (damage).}
+You are absolutely right. We intended to state that visible damage and biomass reduction are not necessarily related in this paragraph. We agree that this is not clear and we will improve on this.
 
 \item {\color{blue}Line 441. Does "damage" mean "visible leaf symptoms"? Or does it mean biomass reduction?}
+See above. We will carefully rephrase the text with respect to the term "damage" and clearly distinguish between visible damage and biomass reduction.
 
@@
 Please clarify.}
+The latter is correct. Thank you for pointing out this ambiguity.
 
@@
 capture the plant physiology of subactic vegetation if no comparisons with physiological measurements taken on subarctic vegetation are presented?}
+We will adresse this point by reframing the storyline of the manuscript towards a proposed  adjustment of existing stomatal conductance parameterizations based on localized climatological data.
 
 \item {\color{blue}Line 469. "However, the decline of this ozone spring peak is partly caused by the uptake of vegetation" Are you sure? Please cite a reference.}
+We will check if there is one citable reference stating exactly this.
+This statement is based on a simple deduction on how ozone is removed from the atmosphere by dry deposition which is described as a network of resistances.
+In general, snow and ice surfaces are found to have a low resistance to ozone , hence less ozone can be taken up by the land surface which is a major sink in absence of photochemical reactions (winter).
+Therefore, existing ozone accumulates in the subarctic winter boundary layer.
+The same goes for non-biogenic precursors.
+In spring these are chemically reactivated and increase the ozone production.
+At the same time coniferous trees might also start to produce BVOCs.
+At the same time, most of the vegetation is still covered by snow and ice and dry deposition remains low.
+In addition to the intrusion of stratospheric ozone, this causes the enhancement of ozone at the surface (spring peak).
+Now, with snowmelt and bud burst in spring, ozone dry deposition increases.
+Simultaneously, photochemical destruction of ozone also increases.
+As only part of ozone actually enters the stomata, we stated "\textbf{partly} caused by the uptake of vegetation".
 
@@
 What does it mean? Please explain. Make an example.}
+The proposed method to use climate data for adjusting the parameterization of stomatal conductance can be formulated as optimization problem (maximizing the enclosed temperature PDF area) which would make it possible to derive new parameters in an more automized way.
+Now, considering enough stomatal conductance data were available from observation in well known climate conditions these data could be used to evaluate our method and train a model to find the optimal stomatal conductance parameterization based on climate data alone.
+As climate data is more readily available compared to actual measurment of stomatal conductance this could help to improve modeling of stomatal conductance globally.
 
-\item {\color{blue}Figure A1. How was the length of the growing seasons in the various years identified? By satellite? Other method? What does the gray band represent?}
+\item {\color{blue}Figure A1.
+How was the length of the growing seasons in the various years identified?
+By satellite?
+Other method?
+What does the gray band represent?} See response to "Line~432".
+The gray band represents the standard deviation.
+We will either remove or elaborate on this analysis.
 
@@
 
-\item {\color{blue}Line 517.
+\item {\color{blue}
+\begin{itemize}
+    \item Line 517.
 "The DO3SE model as described in Büker et al.
 (2012) is used to simulate SWP0 across a PFT specific root depth according to the Penman--Monteith energy balance method that drives water cycling through the soil--plant--atmosphere system" I cannot understand how the P-M energy balance is used in DO3SE to derive the SWP.
-Please explain in detail.}
+Please explain in detail.
 
-\item {\color{blue}Line 525.
+    \item Line 525.
 "the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf" At what height was the O3 concentration measured?
 If it was not measured at the top of the canopy (10m for trees or 10 cm for grassland), how was the O3 concentration at the top canopy calculated?
-Please explain in detail.}
+Please explain in detail.
 
-\item {\color{blue}Line 526.  What does rc represent? Is it the cuticular resistance or the bulk canopy resistance? What is its value?}
+    \item Line 526.  What does rc represent? Is it the cuticular resistance or the bulk canopy resistance? What is its value?
 
-\item {\color{blue}Line 528.  Can you explain where that formula for calculating the flux comes from? Why is there u(z1) in? And what is the z1 height?}
+    \item Line 528.  Can you explain where that formula for calculating the flux comes from? Why is there u(z1) in? And what is the z1 height?
 
-\item {\color{blue}Line 531. What is the z1 height? Where is it?}
+    \item Line 531. What is the z1 height? Where is it?
 
-\item {\color{blue}Line 535. Wind speed at 2 m: what is it used for? Please explain}
+    \item Line 535. Wind speed at 2 m: what is it used for? Please explain
+
+    \item Section B1. The description of fPHEN is missing. Please, provide it. Again, how do you calculate the day-to-day SWP on your site? Please describe it in detail.
+\end{itemize} }
+We will make the description of the $\mathrm{DO_3SE}$ model more comprehensible and comprehensive.
+However, in the interest of the main focus of this manuscript it may not be possible to give a full recap of all details concerning this well established model.
 
 \item {\color{blue}Line 550. Please explicitly describe the method used to gap-fill O3 concentrations because it could be crucial.}
-
-\item {\color{blue}Section B1. The description of fPHEN is missing. Please, provide it. Again, how do you calculate the day-to-day SWP on your site? Please describe it in detail.}
+See response to "Lines 206--208" above.
 
 \item {\color{blue}Line 554. "From Fig. B1f) it is apparent that the mapping manual parameterized grassland would not have been able to grow in 2019." 
-It does not seem to me that gstom has been reseted at all. If this is the case, the
-premises of the work appear weak.}
+It does not seem to me that gstom has been reseted at all. If this is the case, the premises of the work appear weak.}
+We indeed have no observational data to validate the suggested method at tis point.
 

### 2021-12-20 review/response-RC1.tex
Include comments by Frode and Lisa.

@@
 We will refine the language to reduce ambiguity and supplement our statements with appropriate citations where indicated by the referee.\\
-As the manuscript will, hence, undergo major structural changes, we can not address all \textbf{specific / technical comments} in detail at this point.
+As the manuscript will, hence, undergo major structural changes, we will not address all \textbf{specific / technical comments} in detail at this point.
 We will take all specific and technical comments which are still relevant into consideration.\\
@@
 
-    \item {\color{blue} "the [...] study is highly speculative and not based on any experimental evidence [...]"} We are aware that our work is not complemented by experimental evidence and did not claim otherwise.
-We welcome relevant observational data to further improve parameterizations of Arctic and subarctic species.
-We ought to stress that we do not own the "true" parameterization for subarctic vegetation but assess the uncertainty arising from the use of generic parameterizations.
+    \item {\color{blue} "the [...] study is highly speculative and not based on any experimental evidence [...]"} In the revision of our work, we will emphasize that the main objective of our work is to explore the possibilities of readily accessible data (i.e.
+long-term meteorological temperature observations, $\mathrm{CO_2}$ flux estimations from remote sensing) to provide alternatives to site-specific data that is needed to parameterize plant physiology in models such as the $\mathrm{DO_3SE}$ model used in our study.
     \item {\color{blue} "Many claims sound superficial [...], such as [...] the alleged deviation of the years 2018 and 2019 from the site's climatology [...]"} We strongly disagree with this statement.
@@
 We implicitly assume this.
-Therefore, we did not explicitly that BVOCs are not among the ozone precursors that accumulate in the course of winter.
+Therefore, we did not explicitly state that BVOCs are not among the ozone precursors that accumulate in the course of winter.
 We will rephrase the sentence accordingly: "[...] ozone and its non-biogenic precursors [...]"
@@
 The referee is correct that VPD can be high due to cold conditions and thus become a limiting factor.
-We did not find any unexpected or abnormal behavior of stomatal conductance with respect to VPD which indicated that the VPD limitation on stomatal conductance might be more general applicable for high VPD caused by either drought or cold.
-We are not stating that the response to VPD or SWP could differ but in subarctic vegetation we have to assume that they are fair enough.
-Water availability due to frozen soil is most likely not represented in the $\mathrm{DO_3SE}$ model and would be worthwhile investigating further.
-(Lisa?)
+We did not find any unexpected or abnormal behavior of stomatal conductance with respect to VPD.
+This indicates that the parameterized VPD limitation on stomatal conductance might be applicable for high VPD caused by either drought or cold.
+The response of subarctic species to high VPD or SWP could, however, differ.
+We have to assume that these parameterizations are fair enough.
+Different thresholds for $f_\mathrm{VPD}$ could have been tested in analogon to $f_\mathrm{light}$.
+Water (un)availability due to frozen soil is currently not represented in the $\mathrm{DO_3SE}$ model and would be worthwhile investigating further.
 
@@
 The standard deviation reflects the variability of growing conditions and should not be interpreted as statistical measure for to separate one distribution from the other.
-Even more data would not reduce the standard deviation in this case.
+Even more data would perhaps not reduce the standard deviation in this case.
 We will remove the term "substantial" and make our interpretation of the Figure clearer in the text.
@@
 
-\begin{figure}
+\begin{figure}[!th]
     \centering
@@
 capture the plant physiology of subactic vegetation if no comparisons with physiological measurements taken on subarctic vegetation are presented?}
-We will adresse this point by reframing the storyline of the manuscript towards a proposed  adjustment of existing stomatal conductance parameterizations based on localized climatological data.
+We will adresse this point by reframing the storyline of the manuscript towards a new method to derive stomatal conductance parameterizations based on climatological and remotely sensed data.
 
 \item {\color{blue}Line 469. "However, the decline of this ozone spring peak is partly caused by the uptake of vegetation" Are you sure? Please cite a reference.}
-We will check if there is one citable reference stating exactly this.
-This statement is based on a simple deduction on how ozone is removed from the atmosphere by dry deposition which is described as a network of resistances.
+We will rephrase this sentence to "[...] could partly be caused [...]". And may give a recap of ozone removal, as follows:
+\emph{ozone is removed from the atmosphere by dry deposition which is described as a network of resistances.
 In general, snow and ice surfaces are found to have a low resistance to ozone , hence less ozone can be taken up by the land surface which is a major sink in absence of photochemical reactions (winter).
@@
 Simultaneously, photochemical destruction of ozone also increases.
-As only part of ozone actually enters the stomata, we stated "\textbf{partly} caused by the uptake of vegetation".
+Only part of ozone actually enters the stomata.}
 
@@
 Now, considering enough stomatal conductance data were available from observation in well known climate conditions these data could be used to evaluate our method and train a model to find the optimal stomatal conductance parameterization based on climate data alone.
-As climate data is more readily available compared to actual measurment of stomatal conductance this could help to improve modeling of stomatal conductance globally.
+As climate data is more readily available compared to actual measurement of stomatal conductance this could help to improve modeling of stomatal conductance globally.
 
@@
 Please explain in detail.
+    {\color{black} This is described in Bücker et al (2012) in detail.
+We will include an improved, short description of how the $\mathrm{DO_3SE}$ model incorporates the PM model to estimate evapotranspiration and hence water loss from the soil.
+However, for full details of this method it is best for readers to refer to Bücker et al.
+(2012).}
 
@@
 Please explain in detail.
+    {\color{black} The $\mathrm{DO_3SE}$ model was used to estimate the difference in $[\mathrm{O_3}]$} between a reference height above the canopy and the canopy height.
+This employs the deposition component of the $\mathrm{DO_3SE}$ model that estimates the transfer of mass (i.e.
+ozone) as a function of windspeed, convection, surface roughness, and vegetation $[\mathrm{O_3}]$ sink.
+We will add additional detail in the paper to make this clear.
 
     \item Line 526.  What does rc represent? Is it the cuticular resistance or the bulk canopy resistance? What is its value?
+    {\color{black}This represents cuticular resistance - we will include its value in the revision of the paper.}
 
     \item Line 528.  Can you explain where that formula for calculating the flux comes from? Why is there u(z1) in? And what is the z1 height?
+    {\color{black}This comes from the $\mathrm{DO_3SE}$ model - we will add a suitable reference.}
 
     \item Line 531. What is the z1 height? Where is it?
+    {\color{black}See above.}
 
     \item Line 535. Wind speed at 2 m: what is it used for? Please explain
+    {\color{black}See above.}
 
     \item Section B1. The description of fPHEN is missing. Please, provide it. Again, how do you calculate the day-to-day SWP on your site? Please describe it in detail.
+    {\color{black}See above.}
+
 \end{itemize} }

### 2021-12-20 review/response-RC1.tex
Remove double spaces
Include bra-ket notation description

@@
 We acknowledge that we have not used common tools such as the \emph{p-value} (for a discussion see \href{https://doi.org/10.1080/00031305.2016.1154108}{statement of the American Statistical Association (2016)}).
-    \item {\color{blue} "The manuscript is too long and the organization is confused  [...] no clear division between methodological section and results section"} We will address this in the next revision.
+    \item {\color{blue} "The manuscript is too long and the organization is confused [...] no clear division between methodological section and results section"} We will address this in the next revision by fracturing the manuscript clearly into separate sections containing methods, results, and discussion.
 
@@
 "($\left<[\mathrm{O_3}]\right>$ = 36-54ppb)" Please explain the formalism.
-What do different brackets mean?} The bracket notation commonly refers to the concentration of a chemical species.
+What do different brackets mean?} The bracket "$[X]$" notation commonly refers to the concentration of a chemical species X.
 Parts-per-billion (ppb) is a volume mixing ratio (VMR) and strictly speaking no concentration.
@@
 $\chi_{O_3}$ (\href{https://www.e-education.psu.edu/meteo300/node/534}{e-education.psu.edu}).
+The "$\left<A\right>$" notation is derived from Dirac's "bra-ket" formulism in quantum mechanics, e.g.
+$\left<\Psi |A|\Psi \right>\rightarrow \left<A\right>$ "this expression gives the expectation value, or mean or average value, of the observable represented by operator A for the physical system in the state $\left|\Psi \right>$ " (\href{https://en.wikipedia.org/wiki/Bra%E2%80%93ket_notation}{wikipedia}).
 
@@
 \end{itemize} }
-We will make the description of the $\mathrm{DO_3SE}$ model more comprehensible and comprehensive.
+In summary, we will make the description of the $\mathrm{DO_3SE}$ model more comprehensible and comprehensive.
 However, in the interest of the main focus of this manuscript it may not be possible to give a full recap of all details concerning this well established model.

### 2021-12-20 review/response-RC1.tex
Some rephrasing

@@
 The spring peak could even be identical (although unknown, because in 2019 O3 measurements started about 20 days after the spring peak)}
-Based on the observed visible damage on ozone sensitive vegetation in 2018 but not in 2019, our working hypothesis was that either ozone or one of the other important environmental variables may have differed.
-In Sect.~2, we therefore show whether and how they differ.
-We found that there are but only small differences.\\
+In Sect.~2, we assess in detail whether and how those two years differ. We will make this analysis more comprehensible and comprehensive in the next revision.\\
 Regarding the spring peak, until the beginning of May most of the vegetation in the subarctic is still either covered by snow (see Table~3 for the dates) or bud burst has not yet occurred.
-Hence, it was unlikely for non-coniferous trees to experience the ozone spring peak in 2018/19.
+Hence, it was unlikely for deciduous vegetation and scrubs or other short vegetation to experience the ozone spring peak in 2018/19.
 We will come back to coniferous trees later.
@@
 "luftkvalitet.no" is the webinterface to the database of air quality measurements in Norway operated by NILU.
-EBAS is a database operated by NILU as well.
+EBAS is a database with webinterface operated by NILU as well.
 It comprises global air quality measurements, e.g.
@@
 We will remove the term "high".
-However, ozone concentrations of $40\,\mathrm{ppb}$ have been chosen as critical level for the accumulated ozone exposure (AOT) toxic to ecosystems.
+However, ozone concentrations of $40\,\mathrm{ppb}$ have been chosen as critical level for the metric accumulated over a threshold (AOT) to gage ozone exposure toxic to ecosystems.
 This metric was first introduced at the UN/ECE-Workshop in Egham in 1992.
@@
 I suspect a misuse of the locution "statistical significance". How was this significance assessed? Which statistical test was applied? What is the level of significance?}
-We will elaborate on the description the devised hypothesis test.
-First, we compute the excess number of days with residuals larger than $1\sigma$.
+We will elaborate on the hypothesis test in the next revision and may show the residuals as violin plots to increase the transparency of our method and make the data more assessable to the reader.\\
+In the recent version of the manuscript, we computed the excess number of days with residuals larger than $1\sigma$ first.
 This means, we calculate the daily mean and std for 21 years $\times$ 24 hours (264 data points).
@@
 In the next step, we explicitly assume that these are normal distributed - we will elaborate on this assumption in the next revision and check the underlying distributions (but see also Fig.~4).
-Due to the variability in the data, we choose a low sigma level for the significance test ($1\sigma$).
-By grouping the data into months, we find the number of days with observations in each month.
+Due to natural variability (diurnal, annual) in the data, we chose a $1\sigma$ level for the significance test.
+Because we are not looking at repeated experiments with well constraint conditions, a much higher sigma level is not necessarily meaning full.
+But the diurnal cycle could be perhaps removed to smooth the data.
+Second, by grouping the data into months, we find the number of days with observations in each month.
 We count the number of days with residuals exceeding $1\sigma$ and compute the ratio with the number of days of observation.
-If the data was normal distributed, we would expect ~18\% of data in this interval.
+If this data was normal distributed, we would expect ~18\% of data in this interval.
 If the percentage of days exceeding $1\sigma$ is larger, we state that we find a significant excess of days with high ozone.
@@
 We will rephrase and include a brief description of the method.
-We use a Reynolds decomposition on the 1986-1996 data from Svanvik and data from all other long-term monitoring stations in Fennoscandia (1992-2012) to establish multi-annual means.
+We use a Reynolds decomposition on the 1986-1996 data from Svanvik and data from all long-term monitoring stations in Fennoscandia (1992-2012, Jergul-Karasjok, Pallas, Esrange) to establish multi-annual means.
 We then remove the annual cycle from the July 2018 observations at Pallas (reference station) to find the anomalies.
@@
 We find the time lag between the observations and shift the time series of anomalies accordingly.
-We then add the climatology of Svanvik to retrieve a reconstruction of the missing data.
+We then add the climatology of Svanvik to retrieve a reconstruction of the missing data.\\
 As 2019 does not deviated much from the norm by means of temperature, precipitation, and irradiance, we assume it for a normal year.
 From this, we find a probable offset of the historical ozone climatology compared to present day and correct for this.
-We compared this reconstruction with high resolution air quality model reanalysis and found a similar accuracy (residuals compared to observation).
+We compared this reconstruction with high resolution air quality model reanalysis and found a similar accuracy compared to observation.
 For details, please refer the cited work Falk et al (2021).
@@
 Here you state that photosynthesis is already active at DOY 100 and is at 60\% of its maximum!}
-It is true that coniferous can, in principle, take up $\mathrm{CO_2}$ also in winter and the fact that they can suffer from frost drought shows that.
+It is true that coniferous trees can, in principle, take up $\mathrm{CO_2}$ also in winter and the fact that they can suffer from something called frost drought shows that.
 With respect to the spring peak that was mentioned earlier, this means that coniferous trees may experience these elevated ozone concentrations regularly.
-Concerning the contradiction in the statements -- if the conditions are right, coniferous trees can quickly reach 60\% of their max photosynthesis.
-Conditions to keep photosynthesis up would occur from doy 100 onwards.
+Concerning the contradiction in the statements -- if the conditions are right, coniferous trees can quickly reach 60\% of their max photosynthesis but stop photosynthesis the days after.
+Conditions to keep up this levels of photosynthesis, however, would only occur from doy 100 onwards.
 To make sure that we do not falsely accumulate too much ozone from the spring peak periode (when coniferous trees show on-off photosynthesis), we consulted satellite observations to extrapolate the dates when vegetations' photosynthesis (including coniferous trees) at Svanhovd rose over the detectable threshold.

### 2021-12-20 review/response-RC3.tex
Init response to review 3

--- 2021-12-20 review/response-RC1.tex
@@
 \section*{Authors' response}
-To bg-2021-260-RC1 (05 Nov 2021):
-We thank the anonymous referee \#1 for their comments.
+To bg-2021-260-RC3 (30 Nov 2021):
+We thank the anonymous referee \#3 for their comments.
 We will address all \textbf{general comments} raised regarding the structure and readability of the manuscript in our next revision or resubmission.
@@
 
-\subsection*{General comments} 
+\subsection*{Major comments} 
 \begin{itemize}
 
-    \item {\color{blue} "the [...] study is highly speculative and not based on any experimental evidence [...]"} In the revision of our work, we will emphasize that the main objective of our work is to explore the possibilities of readily accessible data (i.e.
-long-term meteorological temperature observations, $\mathrm{CO_2}$ flux estimations from remote sensing) to provide alternatives to site-specific data that is needed to parameterize plant physiology in models such as the $\mathrm{DO_3SE}$ model used in our study.
-    \item {\color{blue} "Many claims sound superficial [...], such as [...] the alleged deviation of the years 2018 and 2019 from the site's climatology [...]"} We strongly disagree with this statement.
-We assessed the particular years 2018 and 2019 by statistical means with respect to our derived climatologies.
-We acknowledge that we have not used common tools such as the \emph{p-value} (for a discussion see \href{https://doi.org/10.1080/00031305.2016.1154108}{statement of the American Statistical Association (2016)}).
-    \item {\color{blue} "The manuscript is too long and the organization is confused  [...] no clear division between methodological section and results section"} We will address this in the next revision.
+    \item {\color{blue}Concerning the length of the paper, I really cannot see the benefit of all the work on long-
+    term ozone and meteorology given in Sections 2-3. Isn't the main point of this paper to
+    illustrate differences in vegetation characteristics between plants at nearly 70 degrees N
+    and the Mapping Manual (MM)? That doesn't need a long analysis of climate differences or
+    ozone trends.}
+
+    \item {\color{blue}Concerning the "default" growing seasons (from DOY 100 to 307) from the MM against
+    which the "bespoke" are compared, these are just wrong for this latitude. The MM
+    suggests a latitude function that gives DOY 129 to 268.}
+
+    \item {\color{blue}Some of the basic equations look wrong to me, see my comments on Appendix B below.}
+
+    \item {\color{blue}On p21 the author's first "key findings" state that their "bespoke parameterizations for
+    subarctic species are better suited..." than the MM ones. This is a strong statement given
+    that the authors admit on L293 that their bespoke suggestions are "hypothetical and have
+    yet to be verified by observations".}
 
@@
 
-\subsection*{Specific comments} 
+\subsection*{Other comments} 
 \begin{itemize}
 
-\item {\color{blue}Lines 51-52.
-"and leads to a build-up of ozone and its precursors during winter." Are you sure?
-How can BVOC accumulate in the atmosphere if the vegetation is covered by snow?
-Please add some citations to support this claim.
-Much more credible is the following explanation based on stratospheric intrusions.}
-We agree with the referee that a vegetation covered by snow would not release BVOCs to the atmosphere.
-We implicitly assume this.
-Therefore, we did not explicitly state that BVOCs are not among the ozone precursors that accumulate in the course of winter.
-We will rephrase the sentence accordingly: "[...] ozone and its non-biogenic precursors [...]"
-
-\item {\color{blue}Line 63. "the time in which vegetation can accumulate ozone." 
-It sounds very bad written this way: vegetation does not accumulate ozone because ozone it is not bioaccumulative. Did you mean the dose?}
-Thank you for pointing this out. We will rephrase accordingly and refer instead the accumulated ozone dose.
-
-\item {\color{blue}Line 77.
-"($\left<[\mathrm{O_3}]\right>$ = 36-54ppb)" Please explain the formalism.
-What do different brackets mean?} The bracket notation commonly refers to the concentration of a chemical species.
-Parts-per-billion (ppb) is a volume mixing ratio (VMR) and strictly speaking no concentration.
-Though, they are often used synonymously because measured VMR is reported in concentrations (e.g.
-$\mathrm{\mu g\,m^{-3}}$) by multiplying with a factor 2.
-So, perhaps, we should use another terminology instead, e.g.
-$\chi_{O_3}$ (\href{https://www.e-education.psu.edu/meteo300/node/534}{e-education.psu.edu}).
-
-\item {\color{blue}Line 81.
-"A substantial body of evidence exists that suggests flux-based metrics, that relate stomatal ozone uptake to vegetation damage, are biologically more relevant for risk assessments than exposure-based metrics."
-Well, please cite at least some works of this "substantial body"} Thank you, we will include relevant citations.
-
-\item {\color{blue}Figure 1. This figure was never referred in the text.}
-This Figure has been referred to in Sect.~2 L177/179.
-
-\item {\color{blue}Figure 2.
-It does not seem to me that the O3 concentrations of 2019 are different from those of 2018.
-The spring peak could even be identical (although unknown, because in 2019 O3 measurements started about 20 days after the spring peak)}
-Based on the observed visible damage on ozone sensitive vegetation in 2018 but not in 2019, our working hypothesis was that either ozone or one of the other important environmental variables may have differed.
-In Sect.~2, we therefore show whether and how they differ.
-We found that there are but only small differences.\\
-Regarding the spring peak, until the beginning of May most of the vegetation in the subarctic is still either covered by snow (see Table~3 for the dates) or bud burst has not yet occurred.
-Hence, it was unlikely for non-coniferous trees to experience the ozone spring peak in 2018/19.
-We will come back to coniferous trees later.
-
-\item {\color{blue}Line 146. Please make clear the acronym PFT on first use}
-Plant functional type (PFT) was introduced in the abstract. Following the common code of conduct, we ought to introduce it also in the main matter. Thank you for pointing this out.
-
-\item {\color{blue}Line 165. "luftkvalitet.no" What is it? And EBAS? Please make them clear.}
-"luftkvalitet.no" is the webinterface to the database of air quality measurements in Norway operated by NILU.
-EBAS is a database operated by NILU as well.
-It comprises global air quality measurements, e.g.
-trace gases, pollutants, particular matter from different networks.
-We will clarify this.
-
-\item {\color{blue}Line 182.
-"This indicates that the vegetation was more affected by ozone in 2018 than in 2019." Being affected by visible symptoms does not necessarily mean having suffered biomass or productivity reduction.} It is correct that biomass reduction and visible damage due to ozone are not the same.
-We discuss this briefly in Sec.~5 but will elaborate on it to make the difference clearer.
-
-\item {\color{blue}Line 190. "high ozone concentrations ([O3] $>$ 40ppb" It is strange to read that O3 concentrations above 40 ppb are "high" concentrations.}
-Indeed 40\,ppb is not particularly "high" compared to peak ozone pollution downwind of densely populated areas, e.g.
-in Eastern Asia.
-We will remove the term "high".
-However, ozone concentrations of $40\,\mathrm{ppb}$ have been chosen as critical level for the accumulated ozone exposure (AOT) toxic to ecosystems.
-This metric was first introduced at the UN/ECE-Workshop in Egham in 1992.
-For the protection of crops and semi-natural vegetation (limit to 5\% yield reduction) and forest different targets exist (see e.g.
-\href{https://www.eea.europa.eu/publications/TOP08-98/page008.html}{EU directive}).
-
-\item {\color{blue}Line 194. "A method for gapfilling data has been presented in Falk et al. (2021)."
-Ok, but was it then applied to this work? Please write it.}
-We will rephrase: "A method for gapfilling data has been presented in Falk et al. (2021) and applied in this work."
-
-\item {\color{blue}Line 205. "We evaluate the statistical significance of divergences from the norm in these variables (referred to as anomalies) in 2018/19"
-I suspect a misuse of the locution "statistical significance". How was this significance assessed? Which statistical test was applied? What is the level of significance?}
-We will elaborate on the description the devised hypothesis test.
-First, we compute the excess number of days with residuals larger than $1\sigma$.
-This means, we calculate the daily mean and std for 21 years $\times$ 24 hours (264 data points).
-From this, we find the residuals in 2018/19 and divide by std.
-In the next step, we explicitly assume that these are normal distributed - we will elaborate on this assumption in the next revision and check the underlying distributions (but see also Fig.~4).
-Due to the variability in the data, we choose a low sigma level for the significance test ($1\sigma$).
-By grouping the data into months, we find the number of days with observations in each month.
-We count the number of days with residuals exceeding $1\sigma$ and compute the ratio with the number of days of observation.
-If the data was normal distributed, we would expect ~18\% of data in this interval.
-If the percentage of days exceeding $1\sigma$ is larger, we state that we find a significant excess of days with high ozone.
-
-\item {\color{blue}Lines 206-208. I do not understand. Please, explicit the methodology.}
-We will rephrase and include a brief description of the method.
-We use a Reynolds decomposition on the 1986-1996 data from Svanvik and data from all other long-term monitoring stations in Fennoscandia (1992-2012) to establish multi-annual means.
-We then remove the annual cycle from the July 2018 observations at Pallas (reference station) to find the anomalies.
-We have shown that these correlate best with observations at Svanvik.
-We find the time lag between the observations and shift the time series of anomalies accordingly.
-We then add the climatology of Svanvik to retrieve a reconstruction of the missing data.
-As 2019 does not deviated much from the norm by means of temperature, precipitation, and irradiance, we assume it for a normal year.
-From this, we find a probable offset of the historical ozone climatology compared to present day and correct for this.
-We compared this reconstruction with high resolution air quality model reanalysis and found a similar accuracy (residuals compared to observation).
-For details, please refer the cited work Falk et al (2021).
-
-\item {\color{blue}Line 213.
-"Averaged monthly accumulated precipitation (blue bars) is shown with standard deviation" It is not consistent to show SE once and STDEV the other time.
-The use of SE is more appropriate when estimating averages.}
-We will change the figure accordingly.
-
-\item {\color{blue}Line 228.
-"Darker colors indicate higher probability to observe these values."\\Line 229.
-"On top of the density distributions, a 10 days average of daily mean (h[O3]i10d) is displayed together with 1sigma uncertainties and SE, respectively" What does it mean?
-It is not clear to me.
-Why show a probability density if you are plotting a multiannual average?
-Or does the line represent the median instead?} In Fig.~4, we show the distribution of ozone data in 2-dimensions.
-The dotes represent the daily averages (but only every 10th day is shown).
-We will rephrase the caption to correct for this.
-We show the daily average to infer that not all daily mean ozone are strictly normal distributed (tails in one or the other direction) but it is also visually deductible that most data fall in the $1\sigma$ band justifying the chosen confidence intervals for a normal distribution.
-We shall, however, elaborate on this by statistical means.
-
-\item {\color{blue}Line 232.
-"The decline in h[O3]i coincides with the average beginning of CO2 uptake by coniferous trees (Kolari et al., 2007; Wallin et al., 2013)" I didn't know that evergreens only uptake CO2 starting in May.
-I was convinced they always did.
-Is it true?
-Doesn't that contradict what you wrote in line 309 ("We base our temperature acclimation of coniferous trees on experimental results on Norway spruce which were found to be active already at rather low air temperatures and can reach 60\% photosynthetic activity as early as doy 100 (Kolariet al., 2007; Wallin et al., 2013).")?
-Here you state that photosynthesis is already active at DOY 100 and is at 60\% of its maximum!}
-It is true that coniferous can, in principle, take up $\mathrm{CO_2}$ also in winter and the fact that they can suffer from frost drought shows that.
-With respect to the spring peak that was mentioned earlier, this means that coniferous trees may experience these elevated ozone concentrations regularly.
-Concerning the contradiction in the statements -- if the conditions are right, coniferous trees can quickly reach 60\% of their max photosynthesis.
-Conditions to keep photosynthesis up would occur from doy 100 onwards.
-To make sure that we do not falsely accumulate too much ozone from the spring peak periode (when coniferous trees show on-off photosynthesis), we consulted satellite observations to extrapolate the dates when vegetations' photosynthesis (including coniferous trees) at Svanhovd rose over the detectable threshold.
-
-\item {\color{blue}Line 233.
-"In July--September (doy 182--273), ozone is occasionally almost completely depleted.
-This hints to ozone uptake by vegetation even at low light intensities during midnight sun conditions in combination with stable planetary boundary layer conditions preventing mixing of ozone rich air."
-I don't understand the connection. What does the night uptake have to do with the
-occasionally complete ozone depletion?} This was actually a working hypothesis and should be removed from the text.
-
-\item {\color{blue}Line 244. "if a normal distribution is assumed" 
-Are you sure that the distribution is normal and not lognormal or something else? There are some literature on the type of statistical distributions for variable such as Temperature, Rain, etc ...
-Moreover, looking at your Figure 6b the distribution of the irradiance seems to be a
-Poisson distribution.}
-The data shown in Fig.6 are not the same as used to calculate the results for Fig.5 because the data selection is different.
-But we should of course check and describe the underlying distribution of the data selection used in Fig.5.
-
-\item {\color{blue}Figure 5, caption.
-"dashed lines indicates statistical significance" Statistical significance of what?
-By means of what test was it obtained, at what alpha level?
-And what are the numbers on the top right of each graph?}
-Regarding the test, see response to "L205" above.
-The number at the top right are described in the caption: "The annual positive/negative deviations are indicated in the respective corners (right upper/lower)." This is the percentage of days on an annual basis that exceed the $1\sigma$ level.
-This indicates the overall deviation from a normal year.
-Only temperatures in 2018 were abnormally high compared to the climatology on the $1\sigma$ level.
-
-\item {\color{blue}Line 251.
-"deviated significantly from the climatology on the 1 sigma level." Here the standard deviation is used as reference for the significance.
-But the significance of the deviation should be statistically tested in another way.}
-Regarding the test, see response to "L205" above.
-Our hypothesis test is that you would expect $\approx18\%$ of days exceeding the $1\sigma$ level (normal distribution).
-If we observe a higher percentage of days, we find an exceedence from the expectation.
-If this is not proper in terms of statistical terminology we can remove the term "significant" and call it exceedence (in case this is deemed a better term).
-
-\item {\color{blue}Line 262.
-"We use the bias-corrected and cross-calibrated ozone climatology (Falk et al., 2021) and assess the monthly significance of the ozone concentration anomalies in 2018/19."
-"Bias-corrected cross-calibrated" ozone? What is it? And what is the "significance" of the concentration anomalies? Please explain.}
-Please see response to "Lines 206-208.".
-Cross-calibrated means that we used data from other monitoring stations in the region to reconstruct the missing data at Svanvik.
-Regarding "significance", see explanation above.
-
-\item {\color{blue}Line 267.
-"Further, we presume that fVPD and fSWP suit our vegetation types and no acclimation is necessary for these." This statement is questionable, because in cold conditions VPD can be high (you also told
-it in the conclusions) and the water in the soil can be limiting because partially unavailable due to freezing or other.}
-The referee is correct that VPD can be high due to cold conditions and thus become a limiting factor.
-We did not find any unexpected or abnormal behavior of stomatal conductance with respect to VPD.
-This indicates that the parameterized VPD limitation on stomatal conductance might be applicable for high VPD caused by either drought or cold.
-The response of subarctic species to high VPD or SWP could, however, differ.
-We have to assume that these parameterizations are fair enough.
-Different thresholds for $f_\mathrm{VPD}$ could have been tested in analogon to $f_\mathrm{light}$.
-Water (un)availability due to frozen soil is currently not represented in the $\mathrm{DO_3SE}$ model and would be worthwhile investigating further.
-
-\item {\color{blue}Line 272.
-"but a substantially higher number of peak [O3] were observed in 2018 than in 2019." How can you tell it if O3 measurements for all months of March, April and July are missing in 2019?
-I don't seem to see any differences between 2018 and 2019}
-As we have indicated, the measurement down time in March and April was planned because most vegetation (in particular the one in the ozone garden) was covered by snow and hence not photosynthetically active (as the referee has noted, too in their comment regarding BVOC emission).
-"but a substantially higher number of peak [O3] were observed in 2018 than in 2019." refers back to Sect.~2 L188-195.
-Peak concentrations were calculated as exceedence over 40\,ppb in summer, explicitly excluding the spring peak.
-Even with two weeks of data missing July 2018, we found 50 elevated ozone event in 2018 compared to 18 times in 2019.
-This should qualify as "substantially higher number".
-
-\item {\color{blue}Line 291 " Note, however, that these parameterizations are hypothetical and have yet to be verified by experiments." Figure 6a.
-Looking at the graph I understand that you assume an adaptation of the subarctic grasslands to the temperature distribution of the last decade (climate already changed) and not to the historical temperature distribution at your site.
-Is it reasonable to hypothesize such a rapid adaptation of vegetation to the new climate conditions?} Indeed it is not reasonable for long-lived species such as trees or coniferous dwarf scrubs and depends on the ecological turn-over periode.
-Competition would lead to a decrease in number of individuals of those species which natural acclimation range differs strongly from the new norm.
-If we look at a grassland as sum of all species the whole grassland may rapidly adapt.
-(Does this make sense Ane?
-Would Vigdis' results back this up?)
-
-\item {\color{blue}Line 298.
-"We construct cold as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to MM.
-This is accomplished by moving Topt towards cooler temperatures while keeping the other parameters fixed to MM values".
-From Figure 6a and Table 1 I see that for the "cold" parameterization not only Topt was moved, but also Tmin for (e.g. for grassland).}
-Indeed, we will correct the text.
-
-\item {\color{blue}Figure 7.
-The gstom/gmax ratio in the subarctic parameterization with PPFD0.8 is greater in the morning than at noon.
-How then the choice of PPFD08 is explained?
-Please comment on this in the text.}
-PPFD0.8 was chosen to maximize the relative stomatal conductance at noon and in the morning. Please refer to L326--334. We will elaborate on the text and make it clearer.
-
-\item {\color{blue}Line 334. I don't understand how we can say that the differences are "substantial". I don't see much difference between deciduous trees (a) and grassland (c), sorry.}
-There seems to be a misunderstanding?
-The referee refers to the sentence "As expected due to the small adjustments, coniferous trees display the smallest differences between the different parameterizations, while the differences for perennial grassland are substantial as a response to the proposed temperature acclimation." Figure~7 is meant to be read for each species separately and from right to left.
-It comprises the mean and standard deviation of relative stomatal conductance computed from hourly meteorological data over 21~years.
-Panel (c) displays perennial grassland.
-Furthest to the right, the results using the mapping manual parameterization are shown.
-The relative stomatal conductance around noon when the highest values are expected is rather low (about 40\%).
-With our adjusted parameterization the relative stomatal conductance around noon reaches about 70--80\%.
-The standard deviation reflects the variability of growing conditions and should not be interpreted as statistical measure for to separate one distribution from the other.
-Even more data would perhaps not reduce the standard deviation in this case.
-We will remove the term "substantial" and make our interpretation of the Figure clearer in the text.
-
-\item {\color{blue}Line 336.
-Using net photosynthesis to calculate leaf emergence is not completely justified.
-Leaves are likely to be present and active well before gross photosynthesis equals heterotrophic respiration (eg.
-soil respiration).
-Gross photosynthesis should be used to calculate Astart and Aend instead.}
-Thank you for pointing this out.
-It is correct that net photosynthesis does not capture the real photosynthesis.
-However, gross photosynthesis is not among the available products from MODIS satellites.
-Instead, we use gross primary production which is including maintenance respiration.
-According to the product website (\href{http://www.ntsg.umt.edu/project/modis/mod17.php}{MOD17}, last accessed Dec.
-2021) it is computed from observed absorbed photosynthetically active radiation (APAR), a species specific conversion efficiency parameter $\epsilon(\mathrm{temperature, water stress})$, temperature, and water stress.
-In Fig.~\ref{fig:modis_gpp} the estimated start and end of growing season from MODIS (Aqua/Terra) GPP over a $1\times 1$\,km patch centered at Svanhovd is shown.
-They differ only slightly from the estimates using net photosynthesis.
-The start of the growing season is shifted by one day (later) and the end by one/two days (later).
-Considering the temporal resolution of 8\,days of the satellite product this difference is negligible.
-
-\begin{figure}[!th]
-    \centering
-    \includegraphics[width=0.75\textwidth]{./modis_Gpp.png}
-    \caption{Estimated start and end of growing season from MODIS (Aqua/Terra) GPP over a $1\times 1$\,km patch centered at Svanhovd.
-The average start in 2018 amounts to doy 123 and to 107 in 2019.
-The end of the growing season in 2018 is doy 262 and in 2019 276.}
-    \label{fig:modis_gpp}
-\end{figure}
-
-\item {\color{blue}Line 354.
-"A sample of downy birch leaves collected at Svanhovd had an average length of (3.0±0.5)cm" Were top-canopy leaves sampled?
-How many leaves were collected to get +- 0.5 cm standard error?}
-Ane?
-
-\item {\color{blue}Line 355. "We used 13.5m height"
-Why was this value chosen? What is the meaning of a height between the average tree
-height and the maximum tree heigh? Perhaps it would have been more reasonable to use
-the average height.}
-We suppose, the referee refers to the 10.1\,m average height of one particular Scots pine forest which has been measured in 2004.
-We assumed that after 14 year 10.1\,m might not be the average height of this particular forest anymore.
-In the absence of more recent measurements, we found it more reasonable to choose 13.5\,m the average tree height in the whole area in 2004 which included all tree species and ages.
-We will include this in the text.
-
-\item {\color{blue}Line 360 and following.
-POD1 was calculated by gap filling the data, right?
-Because there is a lot of data missing in the middle of the season.
-Or were POD1 compensated for missing data?
-If so, how?
-Please confirm it by writing it in the text.}
-In Appendix~B2, we write that the "$\mathrm{DO_3SE}$ model requires hourly, continuous meteorologische observations." This means that all input data, including ozone, have to be gap filled before POD1 can be computed.
-The gap filling method for all data but ozone are described in appendix~B2.
-The used gap filling method for ozone has been summarized above (response to "Lines 206--208") and published in Falk et al.
-(2021).
-We shall restructure the text to make the description of the $\mathrm{DO_3SE}$ model and its input data clearer.
-
-\item {\color{blue}Line 369. "Due to the shape of flight, a symmetric variation".
-Symmetric variation of what?}
-This refers to the variation of PPFD at a stomatal opening of 50\% by $\pm 20\%$ and the resulting response in computed POD1. We will make this clearer in the text.
-
-\item {\color{blue}Line 370.
-"We find that an opening of stomata at lower light intensities can cause higher sensitivity to drought conditions." Please, explain where we can see this.
-Graph 8 is not clear at all to me.
-And then, "sensitivity" of what?
-Of plants?
-Of POD1?}
-Figure~8 is indeed very complex and we will elaborate on the explanation and interpretation in the text to make it more accessible to the readers.
-You are right that this sentence is not clear and does not communicate what we intended to say.
-We refer to the sensitivity of modeled POD1.
-If we take a look at, e.g.
-Fig.~8a subarctic parameterazation and 2018 (left hand side of the upper panel).
-Open symbols represent a model simulation where SWP was taken into account.
-Closed symbols where this effect was switched off.
-The '--' symbolizes PPFD0.8 (earlier opening), the '+' PPFD1.2 (later opening).
-We shall have a look at the circles connected by a solide line representing the simulations with a GS which we have determined from temperatures.
-If the stomata open earlier / close later more ozone can be taken up (longer opening time), hence POD1 is larger than for the unchanged $f_\mathrm{light}$.
-Now, if we take SWP into account, we find that only the PPFD0.8 is affected.
-Therefore, we conclude that conditions that negatively affect SWP (referred to as droughts) reduce the uptake of ozone (POD1) when an earlier opening / later closing of stomata is considered.
-Hence, POD1 is more sensitive to drought conditions.
-
-\item {\color{blue}Line 373.
-"The magnitude of these effects varies between PFTs as well as years, but the predicted ozone uptake for the bespoke temperature parameterization is always larger than for the MM parameterizations and of the same order of magnitude as the variability between the years studied here."
-What effects? "Of the same order of magnitude as the interannual variability...": can you conclude it by comparing only two years? The same for line 402}
-In Section~3, we tried to show that 2019 is representative for a normal year while 2018 is more extreme.
-We should evaluate our data with this question in mind and confirm that the difference between 2018 and 2019 is indeed representative for the interannual variability.
-Depending on the outcome, we will rephrase the sentences and substitute the term "interannual variability".
-
-\item {\color{blue}Table 4.  Have the percentage of reduction been calculated taking into account pre-industrial concentrations as prescribed by the MM?
-What are meaning of the superscripts? And, above all, why some superscripts indicate a range (e.g. 1.9 ... 2.1)? I did not understand how the stdev of the MM estimation was calculated, sorry.}
-We will make this clearer in the next version of the manuscript.
-We used the relationship between biomass reduction and POD1 unchanged from the MM.
-This, however, is to some degree questionable because subarctic vegetation could be affected more or less by the same ozone dose than the central European species.
-Apparently, the caption of the table is not complete and parts of the explanation missing.
-We explain how we derive the uncertainty ranges in L405--407 but this may not be clear enough.
-From Fig.~8, we deduced that SWP is negligible in most cases (solid and open symbols show the same POD1).
-Hence, we computed the biomass reduction based on simulations with SWP effects on POD1 turned off and MM $f_\mathrm{light}$.
-The uncertainties reported are actually the differences and not standard deviations.
-The differences are calculated between the simulation with MM $f_\mathrm{light}$ and "bespoke" GS and the simulations with PPFD0.8 (larger POD1 $\rightarrow$ larger biomass reduction) and PPFD1.2 (smaller POD1 $\rightarrow$ less biomass reduction).
-The sub- and superscripts show the asymmetry introduced by the different $f_\mathrm{light}$ parameterizations.
-The range represents the additional uncertainty from the choice in GS.
-
-\item {\color{blue}Line 416. "we have developed bespoke parameterizations"
-it seems a bit strong statement to me, you have not developed any new tailored
-parameterization, you have only hypothesized one. There is no one experiment nor
-comparison with experimental results in your work.}
-We will substitute the term to appropriately reflect our intensions and work.
-
-\item {\color{blue}Line 417.
-"The comparison between meteorological conditions in 2018 and 2019 and their divergence from climatology allowed us to assess the influence of key environmental variables such as temperature, PPFD, and precipitation on vegetation susceptibility to O3 damage in light of future changes as may occur under climate change" I did not understand where all this "divergence with the climatological average" of these two years alone lies, sorry.}
-We will make this clearer in the revision of the manuscript.
-
-\item {\color{blue}Line 432.
-"With respect to ongoing climate change, a clear positive trend emerged in length (5.2d decade-1) of the growing season that is almost equally distributed between earlier start (2.9 days decade-1) and later end (2.3d decade-1) (Appendix Fig.
-A1)." How did you figure it out?
-Have you been doing retrospective MODIS analysis for 30 years?
-Or do you have a publication to quote?}
-We have derived the number from an analysis of the thermal start and end of the GS based on the gridded temperature data from SeNorge.no provided by the Norwegian weather service and shown in Appendix Fig.~9.
-We may remove this part eventually or rephrase the sentence to include this information, support it with citations of relevant works, and include this analysis in a methods section.
-
-\item {\color{blue}Line 435 and following. "visible damage"
-Visible damage and POD can be totally unrelated, as demonstrated by some research
-conducted on agricultural species. I recommend caution in stating that the O3 peaks
-causing the visible symptoms can result in a biomass reduction (damage).}
-You are absolutely right. We intended to state that visible damage and biomass reduction are not necessarily related in this paragraph. We agree that this is not clear and we will improve on this.
-
-\item {\color{blue}Line 441. Does "damage" mean "visible leaf symptoms"? Or does it mean biomass reduction?}
-See above. We will carefully rephrase the text with respect to the term "damage" and clearly distinguish between visible damage and biomass reduction.
-
-\item {\color{blue}Line 456.
-"We found that soil water potential under 2018/19 meteorological conditions was negligible" What does it mean?
-That there was no water in the soil (SWP were negligible) or that the effect on the POD of the presence or absence of SWP in the calculation was negligible?
-Please clarify.}
-The latter is correct. Thank you for pointing out this ambiguity.
-
-\item {\color{blue}Line 461. "better suited" Point 1 is questionable.
-Also point 2 is questionable. How can you say that the MM parameterization does not
-capture the plant physiology of subactic vegetation if no comparisons with physiological measurements taken on subarctic vegetation are presented?}
-We will adresse this point by reframing the storyline of the manuscript towards a new method to derive stomatal conductance parameterizations based on climatological and remotely sensed data.
-
-\item {\color{blue}Line 469. "However, the decline of this ozone spring peak is partly caused by the uptake of vegetation" Are you sure? Please cite a reference.}
-We will rephrase this sentence to "[...] could partly be caused [...]". And may give a recap of ozone removal, as follows:
-\emph{ozone is removed from the atmosphere by dry deposition which is described as a network of resistances.
-In general, snow and ice surfaces are found to have a low resistance to ozone , hence less ozone can be taken up by the land surface which is a major sink in absence of photochemical reactions (winter).
-Therefore, existing ozone accumulates in the subarctic winter boundary layer.
-The same goes for non-biogenic precursors.
-In spring these are chemically reactivated and increase the ozone production.
-At the same time coniferous trees might also start to produce BVOCs.
-At the same time, most of the vegetation is still covered by snow and ice and dry deposition remains low.
-In addition to the intrusion of stratospheric ozone, this causes the enhancement of ozone at the surface (spring peak).
-Now, with snowmelt and bud burst in spring, ozone dry deposition increases.
-Simultaneously, photochemical destruction of ozone also increases.
-Only part of ozone actually enters the stomata.}
-
-\item {\color{blue}Line 491. "Automation of the here proposed PDF-based acclimation using machine learning techniques could overcome these issues in the future"
-What does it mean? Please explain. Make an example.}
-The proposed method to use climate data for adjusting the parameterization of stomatal conductance can be formulated as optimization problem (maximizing the enclosed temperature PDF area) which would make it possible to derive new parameters in an more automized way.
-Now, considering enough stomatal conductance data were available from observation in well known climate conditions these data could be used to evaluate our method and train a model to find the optimal stomatal conductance parameterization based on climate data alone.
-As climate data is more readily available compared to actual measurement of stomatal conductance this could help to improve modeling of stomatal conductance globally.
-
-\item {\color{blue}Figure A1.
-How was the length of the growing seasons in the various years identified?
-By satellite?
-Other method?
-What does the gray band represent?} See response to "Line~432".
-The gray band represents the standard deviation.
-We will either remove or elaborate on this analysis.
-
-\item {\color{blue}Line 511.
-"with fmin, Dmin, Dmax describing the relative stomatal conductance to changes in vapor pressure deficit." It is not clear.
-Please, clarify what D and fmin are, and their units.}
-
-\item {\color{blue}
-\begin{itemize}
-    \item Line 517.
-"The DO3SE model as described in Büker et al.
-(2012) is used to simulate SWP0 across a PFT specific root depth according to the Penman--Monteith energy balance method that drives water cycling through the soil--plant--atmosphere system" I cannot understand how the P-M energy balance is used in DO3SE to derive the SWP.
-Please explain in detail.
-    {\color{black} This is described in Bücker et al (2012) in detail.
-We will include an improved, short description of how the $\mathrm{DO_3SE}$ model incorporates the PM model to estimate evapotranspiration and hence water loss from the soil.
-However, for full details of this method it is best for readers to refer to Bücker et al.
-(2012).}
-
-    \item Line 525.
-"the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf" At what height was the O3 concentration measured?
-If it was not measured at the top of the canopy (10m for trees or 10 cm for grassland), how was the O3 concentration at the top canopy calculated?
-Please explain in detail.
-    {\color{black} The $\mathrm{DO_3SE}$ model was used to estimate the difference in $[\mathrm{O_3}]$} between a reference height above the canopy and the canopy height.
-This employs the deposition component of the $\mathrm{DO_3SE}$ model that estimates the transfer of mass (i.e.
-ozone) as a function of windspeed, convection, surface roughness, and vegetation $[\mathrm{O_3}]$ sink.
-We will add additional detail in the paper to make this clear.
-
-    \item Line 526.  What does rc represent? Is it the cuticular resistance or the bulk canopy resistance? What is its value?
-    {\color{black}This represents cuticular resistance - we will include its value in the revision of the paper.}
-
-    \item Line 528.  Can you explain where that formula for calculating the flux comes from? Why is there u(z1) in? And what is the z1 height?
-    {\color{black}This comes from the $\mathrm{DO_3SE}$ model - we will add a suitable reference.}
-
-    \item Line 531. What is the z1 height? Where is it?
-    {\color{black}See above.}
-
-    \item Line 535. Wind speed at 2 m: what is it used for? Please explain
-    {\color{black}See above.}
-
-    \item Section B1. The description of fPHEN is missing. Please, provide it. Again, how do you calculate the day-to-day SWP on your site? Please describe it in detail.
-    {\color{black}See above.}
-
-\end{itemize} }
-We will make the description of the $\mathrm{DO_3SE}$ model more comprehensible and comprehensive.
-However, in the interest of the main focus of this manuscript it may not be possible to give a full recap of all details concerning this well established model.
-
-\item {\color{blue}Line 550. Please explicitly describe the method used to gap-fill O3 concentrations because it could be crucial.}
-See response to "Lines 206--208" above.
-
-\item {\color{blue}Line 554. "From Fig. B1f) it is apparent that the mapping manual parameterized grassland would not have been able to grow in 2019." 
-It does not seem to me that gstom has been reseted at all. If this is the case, the premises of the work appear weak.}
-We indeed have no observational data to validate the suggested method at tis point.
+\item {\color{blue}b}
 

### 2021-12-20 review/response-RC3.tex
Finish initial formatation

@@
 
-\item {\color{blue}b}
+\item {\color{blue}}
+
+\item {\color{blue}L14. one can ask if an underestimate of 6\% is a significant problem. I would guess that the overall uncertainty of POD calculations is far greater than this!}
+
+\item {\color{blue}L22.
+Better to say "estimated" rather than reported.
+These numbers do not represent actual measured yield losses.
+Also, as this paper is mainly discussing DO3SE-type approaches to POD, it would have been relevant to cite the DO3SE-based estimates of the Mills et al 2018 paper (below) which also made yield estimates for four crops.}
+
+\item {\color{blue}L29. A more recent citation would be Tarasick et al 2019.}
+
+\item {\color{blue}L36.
+Are these numbers for tropospheric ozone relevant to this paper?
+Vegetation is affected by near-surface ozone, and levels of this are much lower than 50-65 ppb over the Nordic countries.}
+
+\item {\color{blue}L39-40.
+Why "despite"?
+22 days is a very long time!
+This time-scale is not so relevant though; it is the lifetime of ozone in the boundary layer that matters, and that is about 1 week (Hov et al., 1978, or as can be estimated by typical deposition velocities and boundary layer depths).
+Solberg et al.
+2008 also show how residence times of less than 7 days are usual for ozone episodes.}
+
+\item {\color{blue}L48.
+Spring peaks are seen across large swathes of the northern hemisphere, for example at Mace Head, caused by many factors (Monks et al., 2009).
+In fact, Monks et al cite Winkler (1988) that north of 60N, the maximum in ozone reduces in amplitude and shifts to later months.}
+
+\item {\color{blue}L85.
+I don't think Simpson et al 2007 or Mills et al 2011, 2017 focused on the Mediterranean?
+These were pan-European studies, that also stressed that POD as a metric showed that risks in e.g.
+northern Europe were likely higher than the earlier AOT40 metrics showed.
+The ICP-vegetation type metrics have in fact been based upon data from countries across Europe, and with more data from central and northern than from southern European countries.}
+
+\item {\color{blue}L89. Continuing the above, the studies from Karlsson et al have had a strong focus on Scandinavian vegetation.}
+
+\item {\color{blue}L96. Define PLA}
+
+\item {\color{blue}L97. I don't think you need this about typically exceeded under daylight hours. POD calculations follow DO3SE methods, and the flux is zero at night by definition.}
+
+\item {\color{blue}L103.
+The Maas et al book isn't so easily available (and no web-access address is given anyway).
+It can be good to cite some of the other key papers behind ozone CLs, e.g.
+Fuhrer et al.
+(1997), Mills et al 2011 (below).}
+
+\item {\color{blue}L105. "The CL is calculated by"? Do you mean exceedance of CL?}
+
+\item {\color{blue}L110.
+"For our study...".
+This confused me.
+Are the authors reporting results from the present manuscript, or from some unpublished and uncited study?
+The sentences starting here seem to be out of place.}
+
+\item {\color{blue}L127. Do you mean non-methane VOC (NMVOC)?.}
+
+\item {\color{blue}L128. Why brackets in "[O3]"?}
+
+\item {\color{blue}L131-132. I didn't quite understand "shall serve as a reference". Do you mean as an example or possible future?}
+
+\item {\color{blue}L141. Why "sensu" here. I had to look up the word, and read that it is "used especially in technical taxonomic references". Use simpler English.}
+
+\item {\color{blue}L141. In any case what does "(sensu World Meteorological Organisation)" mean in this sentence?}
+
+\item {\color{blue}L165-167. Mangled sentences? Why "Last accessed"? There is no url here. Placement of (a) and (b) is strange.}
+
+\item {\color{blue}L170-171.
+Why deal with data that may introduce a false trend at all?
+Why wouldn't the problems influence seasonality?
+I would guess that weather and climate conditions may well have influenced the frequency of quality-assurance checks.}
+
+\item {\color{blue}L172. Here it says that agromet variables are available from 1992, so, again, why use problematic O3 data from the 1980s?}
+
+\item {\color{blue}L214, Fig. 3. The stars are very small on the figure. Make these more obvious. Or just skip them - doesn't the temperature curve give enough information?}
+
+\item {\color{blue}L215-224.
+Do we need to be told all of these details with plus/minus limits?
+These are just climatalogical values and can be seen well enough in Fig.
+3.
+The numbers are not used for anything.}
+
+\item {\color{blue}L255-256. Why "deduce"? Now the numbers given, 800 W/m2 and 200 W/m2, are very rounded - are these measurements or assumptions or limits?
+L276. I would say hypothetical or local rather than bespoke.}
+
+\item {\color{blue}L280. Again, why "deduce"?}
+
+\item {\color{blue}L281-282. The better ref here for the generic PFTs is the 2017 Mapping Manual. Or what did you use from Simpson et al 2007 or Mills et al. 2011 that isn't in Mills et al 2017?}
+
+\item {\color{blue}L286.
+First, what is Gstoleaf compared to gsto?
+On L553 it is stated that Gsto(leaf) is at leaf-level, which suggests that same as gsto in eqn (B1).
+Then, you claim that Fig.
+B1 shows low Gsto(leaf), but it shows enormous values!
+According to Fig.
+B1, Gstoleaf is ca.
+100 mmole/m2/s, this is 1.0e5 nmole/m2/s!
+This makes no sense.}
+
+\item {\color{blue}L292.
+The authors say that their fT system has not been verified, but haven't ecosystem models (e.g.
+JULES, CLM, LPJ-GUESS) parameterized such vegetation already?
+I would have thought that there was something to learn from such models.}
+
+\item {\color{blue}L344.
+As noted above, why compare your Astart (129, 130) with the MM value of 100 which is valid for 50 degrees N?
+The MM suggests a latitude function, which gives a start date of 129 for your location - i.e.
+almost perfect, and your comment was misleading.
+Why would you consider an agricultural criteria for forests anyway?
+Why not apply the MODIS response for deciduous trees too?}
+
+\item {\color{blue}Table 3: Again, the "Defaults" given here are incorrect for this latitude. The correct ref should anyway be the MM.}
+
+\item {\color{blue}L345.
+"Due to the lack of quantitative field observation...".
+I didn't follow the logic here.
+Table 4.
+Why average the MM but not the bespoke?
+Give the MM for both years - this is useful information.}
+
+\item {\color{blue}L488. Give references for your statements that global land-surface models have problems in the Arctic regions.}
+
+\item {\color{blue}L624-625. Mangled ref. And why source googlebooks, when this report is easily available from www.emep.int?}
+
+\item {\color{blue}References: check upper-lower case, in e.g. AMBIO, grennfelt etc.}
+
+\item {\color{blue}Appendix B: DO3SE model
+The equations used here, and especially around L524-532, have several problems:
+\begin{itemize}
+\item they do not account for differences between canopy and leaf-scale conductances
+\item they make no mention of how the ozone is calculated at the height of the vegetation from the 2m measured values.
+\item Equation B6 looks wrong. What is the u(z1) term doing here?
+\item What is "z1" in eqn. B7?
+\item L531 the factor 1.3 accounts for the "difference in" diffusivity.
+\item Figure B2 - does MODIS only respond to coniferous trees in this 1x1 km2 grid? No other vegetation?
+\end{itemize}}
 

### 2022-01-05 review/response-RC3.tex
Respond to major comments

@@
 
-    \item {\color{blue}Concerning the length of the paper, I really cannot see the benefit of all the work on long-
-    term ozone and meteorology given in Sections 2-3. Isn't the main point of this paper to
-    illustrate differences in vegetation characteristics between plants at nearly 70 degrees N
-    and the Mapping Manual (MM)? That doesn't need a long analysis of climate differences or
-    ozone trends.}
+    \item {\color{blue}Concerning the length of the paper, I really cannot see the benefit of all the work on long-term ozone and meteorology given in Sections 2-3.
+Isn't the main point of this paper to illustrate differences in vegetation characteristics between plants at nearly 70 degrees N and the Mapping Manual (MM)?
+That doesn't need a long analysis of climate differences or ozone trends.} A detailed analysis of long-term ozone and meteorological conditions is important for to modeling of ozone risk in the subarctic.
+Our sample period only extends over two growing seasons (2018/2019).
+To assess wether or not the visible damage could correspond to an unusual high ozone dose in 2018, we need to set this particular year into a climatological context.
+Upon restructuring the manuscript, we may remove or condense this part.
 
-    \item {\color{blue}Concerning the "default" growing seasons (from DOY 100 to 307) from the MM against
-    which the "bespoke" are compared, these are just wrong for this latitude. The MM
-    suggests a latitude function that gives DOY 129 to 268.}
+    \item {\color{blue}Concerning the "default" growing seasons (from DOY 100 to 307) from the MM against which the "bespoke" are compared, these are just wrong for this latitude.
+The MM suggests a latitude function that gives DOY 129 to 268.} The referee is right regarding the latitude model at $70\,^\circ$N and we could have used these instead from the start.
+However, as we state in L344, these start and end DOYs for grassland refer to central Europe.
 
-    \item {\color{blue}Some of the basic equations look wrong to me, see my comments on Appendix B below.}
+    \item {\color{blue}Some of the basic equations look wrong to me, see my comments on Appendix B below.} We will double check the equations make the description of the $\mathrm{DO_3SE}$ model more comprehensible and comprehensive.
+However, in the interest of the main focus of this manuscript it may not be possible to give a full recap of all details concerning this well established model.
 
-    \item {\color{blue}On p21 the author's first "key findings" state that their "bespoke parameterizations for
-    subarctic species are better suited..." than the MM ones. This is a strong statement given
-    that the authors admit on L293 that their bespoke suggestions are "hypothetical and have
-    yet to be verified by observations".}
+    \item {\color{blue}On p21 the author's first "key findings" state that their "bespoke parameterizations for subarctic species are better suited..." than the MM ones.
+This is a strong statement given that the authors admit on L293 that their bespoke suggestions are "hypothetical and have yet to be verified by observations".} Thank you for pointing this out.
+We will rephrase our key findings accordingly.
 
@@
 \begin{itemize}
-
-\item {\color{blue}}
 

### 2022-01-05 review/response-RC1.tex
Include answer from Ane regarding leaves

@@
 How many leaves were collected to get +- 0.5 cm standard error?}
-Ane?
+The birch (B.
+pubescence) leaves were collected from the outer canopy (good light exposure) of tall trees at two sites in Finnmark.
+The leaves were collected late in July so they were fully expanded.
+They were collected by hand, so at about $2\,\mathrm{m}$ height.
+As the sun is at a low elevation angle for most of the time in the growing season in this area, light exposure at the top of the tree is not expected to be so different from lower leaves as long as they grow in a part of the tree that is not shadowed by other trees.
+At the first site (Karasjok), three leaves were collected from each of five adult trees.
+At the second site (Svanvik), five leaves were collected from one tree.
+The leaves were pressed and dried before scanning for the area and shape determination through image analysis.
+ImageJ was used for thresholding the images to give silhouettes of the leaves.
+After scaling the images, the program was used for finding minimum Feret diameter.
+This measure finds the largest width of a leaf at a $90^\circ$ angle to the length of the leaf (the Feret diameter).
+With this method we found that the 20 leaves had a mean width of $3.014\,\mathrm{cm}$ and a standard deviation of $0.4996\,\mathrm{cm}$, given as $(3.0\pm0.5)\,\mathrm{cm}$ in the manuscript.
 

### 2022-01-05 review/response-RC3.tex
Continue with specific questions

@@
 Isn't the main point of this paper to illustrate differences in vegetation characteristics between plants at nearly 70 degrees N and the Mapping Manual (MM)?
-That doesn't need a long analysis of climate differences or ozone trends.} A detailed analysis of long-term ozone and meteorological conditions is important for to modeling of ozone risk in the subarctic.
+That doesn't need a long analysis of climate differences or ozone trends.} A detailed analysis of long-term ozone and meteorological conditions is important for to modeling and understanding of past, present, and future ozone risk in the subarctic.
 Our sample period only extends over two growing seasons (2018/2019).
-To assess wether or not the visible damage could correspond to an unusual high ozone dose in 2018, we need to set this particular year into a climatological context.
-Upon restructuring the manuscript, we may remove or condense this part.
+To assess wether or not the visible damage that we observed in 2018 could correspond to an unusual high ozone dose, we had to to set this particular year into a climatological context.
+Upon restructuring the manuscript, we may condense this part.
 
@@
 
-\item {\color{blue}L14. one can ask if an underestimate of 6\% is a significant problem. I would guess that the overall uncertainty of POD calculations is far greater than this!}
+\item {\color{blue}L14.
+one can ask if an underestimate of 6\% is a significant problem.
+I would guess that the overall uncertainty of POD calculations is far greater than this!} To assess the underlying uncertainty caused by choice of the parameterizations that are used to compute POD, we have varied the temperature and light response paramterizations and looked at the influence of SWP in our target region.
+Within this parameter space the uncertainty due to temperature and light response is $\approx3\,\%$ for deciduous trees.
+(For which we found the $6\,\%$ underestimate.) Hence, we conclude that the estimated underestimate is significant (albeit not in a strict statistical sense).
 
@@
 These numbers do not represent actual measured yield losses.
-Also, as this paper is mainly discussing DO3SE-type approaches to POD, it would have been relevant to cite the DO3SE-based estimates of the Mills et al 2018 paper (below) which also made yield estimates for four crops.}
+Also, as this paper is mainly discussing DO3SE-type approaches to POD, it would have been relevant to cite the DO3SE-based estimates of the Mills et al 2018 paper (below) which also made yield estimates for four crops.} Thank you for pointing this out.
+We will rephrase and include the relevant citations.
 
 \item {\color{blue}L29. A more recent citation would be Tarasick et al 2019.}
+Tanks you. We will include the citation.
 
@@
 Vegetation is affected by near-surface ozone, and levels of this are much lower than 50-65 ppb over the Nordic countries.}
+These numbers are indeed not relevant for the monitoring station at Svanhovd which is situated at low altitude. We will remove the reference.
 
@@
 Solberg et al.
-2008 also show how residence times of less than 7 days are usual for ozone episodes.}
+2008 also show how residence times of less than 7 days are usual for ozone episodes.} Thank you for pointing this out.
+We will rephrase the sentence to include the more relevant time scales.
+Wehen citing 22 days, we had the long-range transport (trans-continental) in mind.
 
@@
 In fact, Monks et al cite Winkler (1988) that north of 60N, the maximum in ozone reduces in amplitude and shifts to later months.}
+Thank you for pointing this out. We will rephrase the sentence accordingly.
 
@@
 The ICP-vegetation type metrics have in fact been based upon data from countries across Europe, and with more data from central and northern than from southern European countries.}
+We will carefully check the cited works and rephrase the paragraph accordingly.
 
 \item {\color{blue}L89. Continuing the above, the studies from Karlsson et al have had a strong focus on Scandinavian vegetation.}
+See above.
 
 \item {\color{blue}L96. Define PLA}
+We will include the definition "per leave area" in the revision.
 
-\item {\color{blue}L97. I don't think you need this about typically exceeded under daylight hours. POD calculations follow DO3SE methods, and the flux is zero at night by definition.}
+\item {\color{blue}L97.
+I don't think you need this about typically exceeded under daylight hours.
+POD calculations follow DO3SE methods, and the flux is zero at night by definition.} We will remove it accordingly.
 
@@
 Fuhrer et al.
-(1997), Mills et al 2011 (below).}
+(1997), Mills et al 2011 (below).} We will include the given references in addition.
 
-\item {\color{blue}L105. "The CL is calculated by"? Do you mean exceedance of CL?}
+\item {\color{blue}L105. "The CL is calculated by"? Do you mean exceedance of CL?} Of course, we mean the exeedance.
 
@@
 The sentences starting here seem to be out of place.}
+The ozone monitoring at Svanhovd was originally meant as a supplement for another study in our project. Hence, the confusing formulation. We will rephrase the sentence.
 
-\item {\color{blue}L127. Do you mean non-methane VOC (NMVOC)?.}
+\item {\color{blue}L127. Do you mean non-methane VOC (NMVOC)?.} Yes, we mean non-methane VOCs.
 
 \item {\color{blue}L128. Why brackets in "[O3]"?}
+The bracket "$[X]$" notation commonly refers to the concentration of a chemical species X.
+Parts-per-billion (ppb) is a volume mixing ratio (VMR) and strictly speaking no concentration.
+Though, they are often used synonymously because measured VMR is reported in concentrations (e.g.
+$\mathrm{\mu g\,m^{-3}}$) by multiplying with a factor 2.
+So, perhaps, we should use another terminology instead, e.g.
+$\chi_{O_3}$ (\href{https://www.e-education.psu.edu/meteo300/node/534}{e-education.psu.edu}).
 
 \item {\color{blue}L131-132. I didn't quite understand "shall serve as a reference". Do you mean as an example or possible future?}
+Exactly. "Example" may be the right term in this context.
 
 \item {\color{blue}L141. Why "sensu" here. I had to look up the word, and read that it is "used especially in technical taxonomic references". Use simpler English.}
+We will rephrase.
 
 \item {\color{blue}L141. In any case what does "(sensu World Meteorological Organisation)" mean in this sentence?}
+In general it means "in the sense of".
+But in this context, it could also be read as "refer to" for further reading.
+"Essential climate Variables" is a term defined and commonly used by the WMO that might not be well known in the plant physiology community.
 
 \item {\color{blue}L165-167. Mangled sentences? Why "Last accessed"? There is no url here. Placement of (a) and (b) is strange.}
+This was supposed to be in the bibliography. We will check the bibtex citation styles to fix the (a) and (b) placements.
 
@@
 Why wouldn't the problems influence seasonality?
-I would guess that weather and climate conditions may well have influenced the frequency of quality-assurance checks.}
-
-\item {\color{blue}L172. Here it says that agromet variables are available from 1992, so, again, why use problematic O3 data from the 1980s?}
+I would guess that weather and climate conditions may well have influenced the frequency of quality-assurance checks.\\
+L172. Here it says that agromet variables are available from 1992, so, again, why use problematic O3 data from the 1980s?}
+These two questions are related.
+Consulting the ozone monitoring report by Solberg et al.
+(2003), we find an inaccuracy in the text which should read "[...] from before 1997 [...]".
+The systematic uncertainty in the pre 1997 ozone date is of the order of $10\,\%$ in AOT.
+Which means that the uncertainty in each individual measurement is lower but in the order of the trend in background ozone.
+This concerns the whole historical monitoring period at Svanhovd.
+Combining all data results in a fair confidence in the observed seasonality rather than each individual measurement.
+From this we find, that 2019 was a rather usual year in terms of ozone and weather.
+Stressing that the $6\,\%$ underestimation of biomass loss for deciduous trees is rather the norm than an exception.
+Without relaying on an ozone climatology, we could not make this assessment.
 
 \item {\color{blue}L214, Fig. 3. The stars are very small on the figure. Make these more obvious. Or just skip them - doesn't the temperature curve give enough information?}
+We will revise the figure accordingly.
 
@@
 The numbers are not used for anything.}
+We consider removing these in the next revision.
 
-\item {\color{blue}L255-256. Why "deduce"? Now the numbers given, 800 W/m2 and 200 W/m2, are very rounded - are these measurements or assumptions or limits?
-L276. I would say hypothetical or local rather than bespoke.}
+\item {\color{blue}L255-256. Why "deduce"? Now the numbers given, 800 W/m2 and 200 W/m2, are very rounded - are these measurements or assumptions or limits?}
+We assume the referee refers to L225-226? We agree that "deduce" is not the right term in this context and will rephrase. These are limits based on the observations.
+
+\item {\color{blue}L276. I would say hypothetical or local rather than bespoke.}
+Thank you for pointing this out. We will change the term accordingly.
 
 \item {\color{blue}L280. Again, why "deduce"?}
+See above.
 

### 2022-01-05 review/response-RC1.tex
Fix formatimg mistake

@@
 Please explain in detail.
-    {\color{black} The $\mathrm{DO_3SE}$ model was used to estimate the difference in $[\mathrm{O_3}]$} between a reference height above the canopy and the canopy height.
+    {\color{black}The $\mathrm{DO_3SE}$ model was used to estimate the difference in $[\mathrm{O_3}]$ between a reference height above the canopy and the canopy height.
 This employs the deposition component of the $\mathrm{DO_3SE}$ model that estimates the transfer of mass (i.e.
 ozone) as a function of windspeed, convection, surface roughness, and vegetation $[\mathrm{O_3}]$ sink.
-We will add additional detail in the paper to make this clear.
+We will add additional detail in the paper to make this clear.}
 

### 2022-01-05 review/response-RC3.tex
Finish first draft.

@@
 \item {\color{blue}L281-282. The better ref here for the generic PFTs is the 2017 Mapping Manual. Or what did you use from Simpson et al 2007 or Mills et al. 2011 that isn't in Mills et al 2017?}
+This was probably a copy and past mistake as all values here were taken from the MM except for LAI.
 
@@
 This makes no sense.}
+$g_\mathrm{sto}$ is per leaf area while $G_\mathrm{sto}$ is per $\mathrm{m}^2$.
+Hence, $G_\mathrm{sto}$ at leaf level is $g_\mathrm{sto}$ multiplied by the leaf area of the whole plant.
+(Please confirm, Lisa.)
 
@@
 I would have thought that there was something to learn from such models.}
+Thanks for pointing this out, we have to double check with the available literature.
 
@@
 Why would you consider an agricultural criteria for forests anyway?
-Why not apply the MODIS response for deciduous trees too?}
+Why not apply the MODIS response for deciduous trees too?} The agricultural rule refers to the bud burst of birch trees.
+We will make this clearer.
+You are right about the choice of $A_\mathrm{start}$ for a central European latitude being misleading.
+We definition, we ran the model with a "default" parameterization for grassland and intended to use the latitude model -- which did not take the latitude dependency into consideration automatically which we only found out later.
+We consider updating the default configuration accordingly.
 
 \item {\color{blue}Table 3: Again, the "Defaults" given here are incorrect for this latitude. The correct ref should anyway be the MM.}
+See above.
 
@@
 "Due to the lack of quantitative field observation...".
-I didn't follow the logic here.
-Table 4.
-Why average the MM but not the bespoke?
-Give the MM for both years - this is useful information.}
+I didn't follow the logic here.} We qualitatively checked the start of the GS by looking at the live stream from a nearby webcam.
+This cannot be used for a quantitative analysis.
+Hence, we take the measurement of snow depth at a meteorological station into consideration instead.
+From this observation, we inferred a latency period of about one month after snow melt for the gras to sprout.
+We will rephrase this sentence to make this clearer.
+
+\item {\color{blue}Table 4. Why average the MM but not the bespoke? Give the MM for both years - this is useful information.}
+Thanks for pointing this out. We will change the table accordingly.
 
 \item {\color{blue}L488. Give references for your statements that global land-surface models have problems in the Arctic regions.}
+We will include references in the next revision.
 
 \item {\color{blue}L624-625. Mangled ref. And why source googlebooks, when this report is easily available from www.emep.int?}
+Thanks. We will check our list of references.
 
 \item {\color{blue}References: check upper-lower case, in e.g. AMBIO, grennfelt etc.}
+Thanks. We will check our list of references.
 
@@
 \item they make no mention of how the ozone is calculated at the height of the vegetation from the 2m measured values.
+{\color{black}The $\mathrm{DO_3SE}$ model was used to estimate the difference in $[\mathrm{O_3}]$ between a reference height above the canopy and the canopy height.
+This employs the deposition component of the $\mathrm{DO_3SE}$ model that estimates the transfer of mass (i.e.
+ozone) as a function of windspeed, convection, surface roughness, and vegetation $[\mathrm{O_3}]$ sink.
+We will add additional detail in the paper to make this clear.}
 \item Equation B6 looks wrong. What is the u(z1) term doing here?
+{\color{black}We will double check the equations.}
 \item What is "z1" in eqn. B7?
+{\color{black}This comes from the $\mathrm{DO_3SE}$ model - we will add a suitable reference.}
 \item L531 the factor 1.3 accounts for the "difference in" diffusivity.
+{\color{black}We will rephrase this sentence.}
 \item Figure B2 - does MODIS only respond to coniferous trees in this 1x1 km2 grid? No other vegetation?
+\item  {\color{black}Estimated land use (from areal pictures available at kartverket.no): 48\% grassland, 32.5\% deciduous, 12.5\% coniferous, 2\% urban.
+We will include this information in the figure caption.}
 \end{itemize}}

### 2022-01-07 review/response-RC3.tex
Include suggestions by Lisa

@@
 Isn't the main point of this paper to illustrate differences in vegetation characteristics between plants at nearly 70 degrees N and the Mapping Manual (MM)?
-That doesn't need a long analysis of climate differences or ozone trends.} A detailed analysis of long-term ozone and meteorological conditions is important for to modeling and understanding of past, present, and future ozone risk in the subarctic.
-Our sample period only extends over two growing seasons (2018/2019).
-To assess wether or not the visible damage that we observed in 2018 could correspond to an unusual high ozone dose, we had to to set this particular year into a climatological context.
-Upon restructuring the manuscript, we may condense this part.
+That doesn't need a long analysis of climate differences or ozone trends.} You are right and upon restructuring the manuscript, we will address this matter by condensing this part.
+We deemed a detailed analysis of long-term ozone and meteorological conditions important for modeling and understanding of past, present, and future ozone risk in the subarctic.
+Our specific sample period extends only over two growing seasons (2018/2019).
+To assess wether or not any of the observed visible damage in 2018 could correspond to an unusual high ozone dose, we had to set this particular year into a climatological context.
 
     \item {\color{blue}Concerning the "default" growing seasons (from DOY 100 to 307) from the MM against which the "bespoke" are compared, these are just wrong for this latitude.
-The MM suggests a latitude function that gives DOY 129 to 268.} The referee is right regarding the latitude model at $70\,^\circ$N and we could have used these instead from the start.
-However, as we state in L344, these start and end DOYs for grassland refer to central Europe.
+The MM suggests a latitude function that gives DOY 129 to 268.} In retrospective, this was a fault in the model set up and will be corrected.
+However, we clearly state in L344, that these start and end DOYs for grassland are valid only to central Europe.
 
-    \item {\color{blue}Some of the basic equations look wrong to me, see my comments on Appendix B below.} We will double check the equations make the description of the $\mathrm{DO_3SE}$ model more comprehensible and comprehensive.
+    \item {\color{blue}Some of the basic equations look wrong to me, see my comments on Appendix B below.} We will double check the equations (see comments below) and make the description of the $\mathrm{DO_3SE}$ model more comprehensible and comprehensive.
 However, in the interest of the main focus of this manuscript it may not be possible to give a full recap of all details concerning this well established model.
@@
 one can ask if an underestimate of 6\% is a significant problem.
-I would guess that the overall uncertainty of POD calculations is far greater than this!} To assess the underlying uncertainty caused by choice of the parameterizations that are used to compute POD, we have varied the temperature and light response paramterizations and looked at the influence of SWP in our target region.
-Within this parameter space the uncertainty due to temperature and light response is $\approx3\,\%$ for deciduous trees.
+I would guess that the overall uncertainty of POD calculations is far greater than this!} To address the underlying uncertainty caused by the choice of parameterizations that are used to compute POD, we varied the temperature and light response paramterizations (accounting for the right start and end of the growing season) and looked at the influence of SWP on POD in our target region.
+Within this parameter space the uncertainty due to temperature and light response is, e.g., $\sim3\,\%$ for deciduous trees.
 (For which we found the $6\,\%$ underestimate.) Hence, we conclude that the estimated underestimate is significant (albeit not in a strict statistical sense).
+We will add a discussion about the methodological uncertainty of POD in the next revision.
 
@@
 \item {\color{blue}L29. A more recent citation would be Tarasick et al 2019.}
-Tanks you. We will include the citation.
+Thank you. We will include the citation.
 
@@
 \item {\color{blue}L89. Continuing the above, the studies from Karlsson et al have had a strong focus on Scandinavian vegetation.}
-See above.
+Our focus lies in particular on vegetation in the subarctic. Whereas works by Karlsson et al. focused mainly on vegetation further south in the boreal zone.
 
@@
 Hence, $G_\mathrm{sto}$ at leaf level is $g_\mathrm{sto}$ multiplied by the leaf area of the whole plant.
-(Please confirm, Lisa.)
+(Please confirm, Lisa.
+- I don't think that there was much of a difference, but is the referee trying to say that something is wrong with the order of magnitude?
+That's the number and units given by the model.)
 
@@
 I would have thought that there was something to learn from such models.}
-Thanks for pointing this out, we have to double check with the available literature.
+Thanks for pointing this out.
+It depends on what type of $g_\mathrm{sto}$ module ecosystems models use.
+Generally, these will use coupled photosynthesis-stomatal conductance models which do not require the definition of an $f_\mathrm{T}$ function.
 
@@
 You are right about the choice of $A_\mathrm{start}$ for a central European latitude being misleading.
-We definition, we ran the model with a "default" parameterization for grassland and intended to use the latitude model -- which did not take the latitude dependency into consideration automatically which we only found out later.
-We consider updating the default configuration accordingly.
+This was a fault in the model set up and will be corrected.
 

### 2022-01-12 review/response-RC1.tex
Include comments by Ane

@@
 long-term meteorological temperature observations, $\mathrm{CO_2}$ flux estimations from remote sensing) to provide alternatives to site-specific data that is needed to parameterize plant physiology in models such as the $\mathrm{DO_3SE}$ model used in our study.
-    \item {\color{blue} "Many claims sound superficial [...], such as [...] the alleged deviation of the years 2018 and 2019 from the site's climatology [...]"} We strongly disagree with this statement.
-We assessed the particular years 2018 and 2019 by statistical means with respect to our derived climatologies.
-We acknowledge that we have not used common tools such as the \emph{p-value} (for a discussion see \href{https://doi.org/10.1080/00031305.2016.1154108}{statement of the American Statistical Association (2016)}).
+    \item {\color{blue} "Many claims sound superficial [...], such as [...] the alleged deviation of the years 2018 and 2019 from the site's climatology [...]"} We acknowledge that we have not used tools such as the \emph{p-value} (for a discussion see \href{https://doi.org/10.1080/00031305.2016.1154108}{statement of the American Statistical Association (2016)}).
+We will make our statistical approach clearer where necessary (see also more detailed responses below) and replace terms such as "significantly different" where significance has not been shown appropriately.
     \item {\color{blue} "The manuscript is too long and the organization is confused [...] no clear division between methodological section and results section"} We will address this in the next revision by fracturing the manuscript clearly into separate sections containing methods, results, and discussion.
@@
 "($\left<[\mathrm{O_3}]\right>$ = 36-54ppb)" Please explain the formalism.
-What do different brackets mean?} The bracket "$[X]$" notation commonly refers to the concentration of a chemical species X.
+What do different brackets mean?} We will make this notation and formalism clearer in the text.
+The "$[X]$" notation commonly refers to the concentration of a chemical species X.
 Parts-per-billion (ppb) is a volume mixing ratio (VMR) and strictly speaking no concentration.
-Though, they are often used synonymously because measured VMR is reported in concentrations (e.g.
-$\mathrm{\mu g\,m^{-3}}$) by multiplying with a factor 2.
-So, perhaps, we should use another terminology instead, e.g.
+Though, if you calculate the amount of ozone per volume from ppb to $\mathrm{\mu g\,m^{-3}}$ meter, assuming a pressure of 1 atmosphere, temperature of 298K and use the ideal gas law, you get a factor $\sim 2$.
+Therefore, concentrations and VMRs are often used synonymously.
+\\We could also use another terminology, e.g.
 $\chi_{O_3}$ (\href{https://www.e-education.psu.edu/meteo300/node/534}{e-education.psu.edu}).
-The "$\left<A\right>$" notation is derived from Dirac's "bra-ket" formulism in quantum mechanics, e.g.
+\\The "$\left<A\right>$" notation is derived from Dirac's "bra-ket" formulism in quantum mechanics, e.g.
 $\left<\Psi |A|\Psi \right>\rightarrow \left<A\right>$ "this expression gives the expectation value, or mean or average value, of the observable represented by operator A for the physical system in the state $\left|\Psi \right>$ " (\href{https://en.wikipedia.org/wiki/Bra%E2%80%93ket_notation}{wikipedia}).
@@
 This means, we calculate the daily mean and std for 21 years $\times$ 24 hours (264 data points).
-From this, we find the residuals in 2018/19 and divide by std.
-In the next step, we explicitly assume that these are normal distributed - we will elaborate on this assumption in the next revision and check the underlying distributions (but see also Fig.~4).
+From this, we find the residuals in 2018/19.
+In the next step, we explicitly assume that these are normal distributed -- we will elaborate on this assumption in the next revision and check the underlying distributions (but see also Fig.~4).
 Due to natural variability (diurnal, annual) in the data, we chose a $1\sigma$ level for the significance test.
@@
 But the diurnal cycle could be perhaps removed to smooth the data.
-Second, by grouping the data into months, we find the number of days with observations in each month.
+Second, by grouping the data into months, we find number of days with observations sufficiently diverging from the mean in each month.
 We count the number of days with residuals exceeding $1\sigma$ and compute the ratio with the number of days of observation.
 If this data was normal distributed, we would expect ~18\% of data in this interval.
-If the percentage of days exceeding $1\sigma$ is larger, we state that we find a significant excess of days with high ozone.
+If the percentage of days exceeding $1\sigma$ is larger, we state that we find a prominent excess of days with high ozone.
 
@@
 I don't seem to see any differences between 2018 and 2019}
-As we have indicated, the measurement down time in March and April was planned because most vegetation (in particular the one in the ozone garden) was covered by snow and hence not photosynthetically active (as the referee has noted, too in their comment regarding BVOC emission).
-"but a substantially higher number of peak [O3] were observed in 2018 than in 2019." refers back to Sect.~2 L188-195.
+As we have indicated, the measurement down time in March and April was planned because most vegetation (in particular the one in the ozone garden) was covered by snow and hence not photosynthetically active (as the referee remarked in their comment regarding BVOC emission).
+"but a substantially higher number of peak [O3] were observed in 2018 than in 2019." refers back to Sec.~2 L188-195.
 Peak concentrations were calculated as exceedence over 40\,ppb in summer, explicitly excluding the spring peak.
@@
 This should qualify as "substantially higher number".
+This explanation will be added in the text for clarification.
 
@@
 Looking at the graph I understand that you assume an adaptation of the subarctic grasslands to the temperature distribution of the last decade (climate already changed) and not to the historical temperature distribution at your site.
-Is it reasonable to hypothesize such a rapid adaptation of vegetation to the new climate conditions?} Indeed it is not reasonable for long-lived species such as trees or coniferous dwarf scrubs and depends on the ecological turn-over periode.
-Competition would lead to a decrease in number of individuals of those species which natural acclimation range differs strongly from the new norm.
-If we look at a grassland as sum of all species the whole grassland may rapidly adapt.
-(Does this make sense Ane?
-Would Vigdis' results back this up?)
+Is it reasonable to hypothesize such a rapid adaptation of vegetation to the new climate conditions?} Both \emph{acclimation} of individual plants during their life-time and adaptation of a plant population (\emph{evolutionary process}) can contribute to the vegetation's stomatal conductance response to temperature (e.g., \href{https://bg.copernicus.org/articles/8/1453/2011/}{Juan et al.; 2011}).
+The evolutionary process would lead to a decrease in number of individuals of those species which natural acclimation range differs strongly from the new norm.
+As "perennial grassland" in our model consists of a range of species, the grassland as a whole may rapidly adapt.\\
+In addition, there is the phenomenon of \emph{epigenetics}, exemplified by Norway spruce trees growing from seeds.
+The temperatures experienced during seed maturation (climate at the site where the mother tree actually stands) affect the new trees' response to environmental factors determining bud set in autumn.
+This means that the seedlings may be adapted to the site where they are produced, although the genetic background of the parents predicts a different response (e.g.
+\href{https://link.springer.com/article/10.1007%2Fs11295-009-0231-z}{Skrøppa et al.; 2010}).
 
@@
 It does not seem to me that gstom has been reseted at all. If this is the case, the premises of the work appear weak.}
-We indeed have no observational data to validate the suggested method at tis point.
+As mentioned earlier, in the revision of our work, we will emphasize that the main objective of our work is to explore the possibilities of readily accessible data (i.e.
+long-term meteorological temperature observations, $\mathrm{CO_2}$ flux estimations from remote sensing) to provide alternatives to site-specific data that is needed to parameterize plant physiology in models such as the $\mathrm{DO_3SE}$ model used in our study.
 

### 2022-01-13 review/response-RC3.tex
Minor changes by Frode
Include comments by Ane

@@
 Isn't the main point of this paper to illustrate differences in vegetation characteristics between plants at nearly 70 degrees N and the Mapping Manual (MM)?
-That doesn't need a long analysis of climate differences or ozone trends.} You are right and upon restructuring the manuscript, we will address this matter by condensing this part.
+That doesn't need a long analysis of climate differences or ozone trends.}
+    You are right and upon restructuring the manuscript, we will address this matter by condensing this part.
 We deemed a detailed analysis of long-term ozone and meteorological conditions important for modeling and understanding of past, present, and future ozone risk in the subarctic.
@@
     \item {\color{blue}Concerning the "default" growing seasons (from DOY 100 to 307) from the MM against which the "bespoke" are compared, these are just wrong for this latitude.
-The MM suggests a latitude function that gives DOY 129 to 268.} In retrospective, this was a fault in the model set up and will be corrected.
-However, we clearly state in L344, that these start and end DOYs for grassland are valid only to central Europe.
+The MM suggests a latitude function that gives DOY 129 to 268.}
+    In retrospective, this was a fault in the model set up and will be corrected. However, we clearly state in L344, that these start and end DOYs for grassland are valid only to central Europe.
 
-    \item {\color{blue}Some of the basic equations look wrong to me, see my comments on Appendix B below.} We will double check the equations (see comments below) and make the description of the $\mathrm{DO_3SE}$ model more comprehensible and comprehensive.
+    \item {\color{blue}Some of the basic equations look wrong to me, see my comments on Appendix B below.}
+    We will double check the equations (see comments below) and make the description of the $\mathrm{DO_3SE}$ model more comprehensible and comprehensive.
 However, in the interest of the main focus of this manuscript it may not be possible to give a full recap of all details concerning this well established model.
@@
     \item {\color{blue}On p21 the author's first "key findings" state that their "bespoke parameterizations for subarctic species are better suited..." than the MM ones.
-This is a strong statement given that the authors admit on L293 that their bespoke suggestions are "hypothetical and have yet to be verified by observations".} Thank you for pointing this out.
-We will rephrase our key findings accordingly.
+This is a strong statement given that the authors admit on L293 that their bespoke suggestions are "hypothetical and have yet to be verified by observations".}
+    Thank you for pointing this out. We will rephrase our key findings accordingly.
 
@@
 Within this parameter space the uncertainty due to temperature and light response is, e.g., $\sim3\,\%$ for deciduous trees.
-(For which we found the $6\,\%$ underestimate.) Hence, we conclude that the estimated underestimate is significant (albeit not in a strict statistical sense).
-We will add a discussion about the methodological uncertainty of POD in the next revision.
+(For which we found the $6\,\%$ underestimate.)
+rom Table 4 we see that the deciduous tree biomass loss is $11.2\,\%$ with MM settings and $17.4\,\%$ with bespoke settings for temperature and GS.
+The difference is 6 percentage points, but the relative increase between loosing $11\,\%$ and $17.4\,\%$ is $55\,\%$.
+Hence, we conclude that the estimated underestimate is significant (albeit not in a strict statistical sense).
+We will add a discussion about the methodological uncertainty of POD in the next revision and may rephrase the sentence.
 
@@
 \item {\color{blue}L96. Define PLA}
-We will include the definition "per leave area" in the revision.
+PLA stands for "projected leave area" indicating that the area is measured on only one side of the leaf.
+It also means that any two-sided effects of leaves need to be accounted for by doubling the leaf area in calculations.
 
@@
 \item {\color{blue}L128. Why brackets in "[O3]"?}
-The bracket "$[X]$" notation commonly refers to the concentration of a chemical species X.
+The "$[X]$" notation commonly refers to the concentration of a chemical species X.
 Parts-per-billion (ppb) is a volume mixing ratio (VMR) and strictly speaking no concentration.
-Though, they are often used synonymously because measured VMR is reported in concentrations (e.g.
-$\mathrm{\mu g\,m^{-3}}$) by multiplying with a factor 2.
-So, perhaps, we should use another terminology instead, e.g.
+Though, if you calculate the amount of ozone per volume from ppb to $\mathrm{\mu g\,m^{-3}}$ meter, assuming a pressure of 1 atmosphere, temperature of 298K and use the ideal gas law, you get a factor $\sim 2$.
+Therefore, concentrations and VMRs are often used synonymously.
+\\We could also use another terminology, e.g.
 $\chi_{O_3}$ (\href{https://www.e-education.psu.edu/meteo300/node/534}{e-education.psu.edu}).
+\\The "$\left<A\right>$" notation is derived from Dirac's "bra-ket" formulism in quantum mechanics, e.g.
+$\left<\Psi |A|\Psi \right>\rightarrow \left<A\right>$ "this expression gives the expectation value, or mean or average value, of the observable represented by operator A for the physical system in the state $\left|\Psi \right>$ " (\href{https://en.wikipedia.org/wiki/Bra%E2%80%93ket_notation}{wikipedia}).
 
@@
 These two questions are related.
-Consulting the ozone monitoring report by Solberg et al.
-(2003), we find an inaccuracy in the text which should read "[...] from before 1997 [...]".
+Re-Consulting the ozone monitoring report by Solberg et al.
+(2003), we find an inaccuracy in our manuscript text which should read "[...] from before 1997 [...]".
 The systematic uncertainty in the pre 1997 ozone date is of the order of $10\,\%$ in AOT.
@@
 This concerns the whole historical monitoring period at Svanhovd.
-Combining all data results in a fair confidence in the observed seasonality rather than each individual measurement.
+Combining all data gives a good confidence in the observed seasonality rather than each individual measurement.
 From this we find, that 2019 was a rather usual year in terms of ozone and weather.
-Stressing that the $6\,\%$ underestimation of biomass loss for deciduous trees is rather the norm than an exception.
-Without relaying on an ozone climatology, we could not make this assessment.
+Without relying on an ozone climatology, we could not make this assessment.
+In \href{https://acp.copernicus.org/articles/21/15647/2021/}{Falk et al.
+(2021)} we have shown that the correlation between data taken at Svanvik and Pallas in Finnland during the overlapping period in the 90s is $r^2=0.61$ which is good given the different altitudes and surroundings.
 
@@
 \item {\color{blue}L281-282. The better ref here for the generic PFTs is the 2017 Mapping Manual. Or what did you use from Simpson et al 2007 or Mills et al. 2011 that isn't in Mills et al 2017?}
-This was probably a copy and past mistake as all values here were taken from the MM except for LAI.
+This was probably a copy and paste mistake as all values here were taken from the MM except for LAI.
 
@@
 This makes no sense.}
-$g_\mathrm{sto}$ is per leaf area while $G_\mathrm{sto}$ is per $\mathrm{m}^2$.
+$g_\mathrm{sto}$ is the stomatal conductance per projected leaf area while $G_\mathrm{sto}$ is per $\mathrm{m}^2$.
 Hence, $G_\mathrm{sto}$ at leaf level is $g_\mathrm{sto}$ multiplied by the leaf area of the whole plant.
-(Please confirm, Lisa.
-- I don't think that there was much of a difference, but is the referee trying to say that something is wrong with the order of magnitude?
-That's the number and units given by the model.)
 
@@
 Why not apply the MODIS response for deciduous trees too?} The agricultural rule refers to the bud burst of birch trees.
+For a thorough discussion of the use of accumulated mean temperatures in estimating the start and end of the growing season in Scandinavia see e.g.
+\href{https://pub.epsilon.slu.se/3910/1/SFS194.pdf}{Morén and Perttu (1994)}.
 We will make this clearer.
-You are right about the choice of $A_\mathrm{start}$ for a central European latitude being misleading.
+\\The choice of $A_\mathrm{start}$ for a central European latitude is indeed misleading.
 This was a fault in the model set up and will be corrected.

### 2022-01-13 review/response-RC3.tex
Checked with grammaly

@@
 We will address all \textbf{general comments} raised regarding the structure and readability of the manuscript in our next revision or resubmission.
-We consider to separate the manuscript in Introduction--Methods--Results--Conclusions and will improve on the "methods" Section to make our results more comprehensible.
-We will refine the language to reduce ambiguity and supplement our statements with appropriate citations where indicated by the referee.\\
+We consider separating the manuscript in Introduction--Methods--Results--Conclusions and will improve on the "methods" Section to make our results more comprehensible.
+We will refine the language to reduce ambiguity and supplement our statements with appropriate citations where those were indicated by the referee.\\
 As the manuscript will, hence, undergo major structural changes, we will not address all \textbf{specific / technical comments} in detail at this point.
 We will take all specific and technical comments which are still relevant into consideration.\\
-In the following, we shall give a brief respond to all relevant issues.
+In the following, we shall give a brief response to all relevant issues.
 
@@
     You are right and upon restructuring the manuscript, we will address this matter by condensing this part.
-We deemed a detailed analysis of long-term ozone and meteorological conditions important for modeling and understanding of past, present, and future ozone risk in the subarctic.
+We deemed a detailed analysis of long-term ozone and meteorological conditions important for modeling and understanding past, present, and future ozone risk in the subarctic.
 Our specific sample period extends only over two growing seasons (2018/2019).
-To assess wether or not any of the observed visible damage in 2018 could correspond to an unusual high ozone dose, we had to set this particular year into a climatological context.
+To assess whether or not any of the observed visible damage in 2018 could correspond to an unusual high ozone dose, we had to set this particular year into a climatological context.
 
@@
 The MM suggests a latitude function that gives DOY 129 to 268.}
-    In retrospective, this was a fault in the model set up and will be corrected. However, we clearly state in L344, that these start and end DOYs for grassland are valid only to central Europe.
+    In retrospect, this was a fault in the model setup and will be corrected. However, we clearly state in L344, that these start and end DOYs for grassland are valid only to central Europe.
 
     \item {\color{blue}Some of the basic equations look wrong to me, see my comments on Appendix B below.}
-    We will double check the equations (see comments below) and make the description of the $\mathrm{DO_3SE}$ model more comprehensible and comprehensive.
-However, in the interest of the main focus of this manuscript it may not be possible to give a full recap of all details concerning this well established model.
+    We will double-check the equations (see comments below) and make the description of the $\mathrm{DO_3SE}$ model more comprehensible and comprehensive.
+However, in the interest of the main focus of this manuscript, it may not be possible to give a full recap of all details concerning this well-established model.
 
@@
     Thank you for pointing this out. We will rephrase our key findings accordingly.
-
 \end{itemize}
@@
 
-\item {\color{blue}L14.
-one can ask if an underestimate of 6\% is a significant problem.
-I would guess that the overall uncertainty of POD calculations is far greater than this!} To address the underlying uncertainty caused by the choice of parameterizations that are used to compute POD, we varied the temperature and light response paramterizations (accounting for the right start and end of the growing season) and looked at the influence of SWP on POD in our target region.
-Within this parameter space the uncertainty due to temperature and light response is, e.g., $\sim3\,\%$ for deciduous trees.
+\item {\color{blue}L14. one can ask if an underestimate of 6\% is a significant problem. I would guess that the overall uncertainty of POD calculations is far greater than this!}
+To address the underlying uncertainty caused by the choice of parameterizations that are used to compute POD, we varied the temperature and light response paramterizations (accounting for the right start and end of the growing season) and looked at the influence of SWP on POD in our target region.
+Within this parameter space, the uncertainty due to temperature and light response is, e.g., $\sim3\,\%$ for deciduous trees.
 (For which we found the $6\,\%$ underestimate.)
-rom Table 4 we see that the deciduous tree biomass loss is $11.2\,\%$ with MM settings and $17.4\,\%$ with bespoke settings for temperature and GS.
-The difference is 6 percentage points, but the relative increase between loosing $11\,\%$ and $17.4\,\%$ is $55\,\%$.
+From Table 4 we see that the deciduous tree biomass loss is $11.2\,\%$ with MM settings and $17.4\,\%$ with bespoke settings for temperature and GS.
+The difference is 6 percentage points, but the relative increase between losing $11\,\%$ and $17.4\,\%$ is $55\,\%$.
 Hence, we conclude that the estimated underestimate is significant (albeit not in a strict statistical sense).
@@
 These numbers do not represent actual measured yield losses.
-Also, as this paper is mainly discussing DO3SE-type approaches to POD, it would have been relevant to cite the DO3SE-based estimates of the Mills et al 2018 paper (below) which also made yield estimates for four crops.} Thank you for pointing this out.
-We will rephrase and include the relevant citations.
+Also, as this paper is mainly discussing DO3SE-type approaches to POD, it would have been relevant to cite the DO3SE-based estimates of the Mills et al 2018 paper (below) which also made yield estimates for four crops.}
+Thank you for pointing this out. We will rephrase and include the relevant citations.
 
@@
 Vegetation is affected by near-surface ozone, and levels of this are much lower than 50-65 ppb over the Nordic countries.}
-These numbers are indeed not relevant for the monitoring station at Svanhovd which is situated at low altitude. We will remove the reference.
+These numbers are indeed not relevant for the monitoring station at Svanhovd which is situated at a low altitude. We will remove the reference.
 
@@
 Solberg et al.
-2008 also show how residence times of less than 7 days are usual for ozone episodes.} Thank you for pointing this out.
-We will rephrase the sentence to include the more relevant time scales.
-Wehen citing 22 days, we had the long-range transport (trans-continental) in mind.
+2008 also show how residence times of less than 7 days are usual for ozone episodes.}
+Thank you for pointing this out. We will rephrase the sentence to include the more relevant time scales. When citing 22 days, we had the long-range transport (trans-continental) in mind.
 
@@
 
-\item {\color{blue}L97.
-I don't think you need this about typically exceeded under daylight hours.
-POD calculations follow DO3SE methods, and the flux is zero at night by definition.} We will remove it accordingly.
+\item {\color{blue}L97. I don't think you need this about typically exceeded under daylight hours. POD calculations follow DO3SE methods, and the flux is zero at night by definition.}
+We will remove it accordingly.
 
@@
 Fuhrer et al.
-(1997), Mills et al 2011 (below).} We will include the given references in addition.
+(1997), Mills et al 2011 (below).}
+We will include the given references in addition.
 
-\item {\color{blue}L105. "The CL is calculated by"? Do you mean exceedance of CL?} Of course, we mean the exeedance.
+\item {\color{blue}L105. "The CL is calculated by"? Do you mean exceedance of CL?}
+Of course, we mean exceedance.
 
@@
 
-\item {\color{blue}L127. Do you mean non-methane VOC (NMVOC)?.} Yes, we mean non-methane VOCs.
+\item {\color{blue}L127. Do you mean non-methane VOC (NMVOC)?.} 
+Yes, we mean non-methane VOCs.
 
@@
 Parts-per-billion (ppb) is a volume mixing ratio (VMR) and strictly speaking no concentration.
-Though, if you calculate the amount of ozone per volume from ppb to $\mathrm{\mu g\,m^{-3}}$ meter, assuming a pressure of 1 atmosphere, temperature of 298K and use the ideal gas law, you get a factor $\sim 2$.
+Though, if you calculate the amount of ozone per volume from ppb to $\mathrm{\mu g\,m^{-3}}$ meter, assuming a pressure of 1 atmosphere, the temperature of 298K and use the ideal gas law, you get a factor $\sim 2$.
 Therefore, concentrations and VMRs are often used synonymously.
@@
 $\chi_{O_3}$ (\href{https://www.e-education.psu.edu/meteo300/node/534}{e-education.psu.edu}).
-\\The "$\left<A\right>$" notation is derived from Dirac's "bra-ket" formulism in quantum mechanics, e.g.
+\\The "$\left<A\right>$" notation is derived from Dirac's "bra-ket" formalism in quantum mechanics, e.g.
 $\left<\Psi |A|\Psi \right>\rightarrow \left<A\right>$ "this expression gives the expectation value, or mean or average value, of the observable represented by operator A for the physical system in the state $\left|\Psi \right>$ " (\href{https://en.wikipedia.org/wiki/Bra%E2%80%93ket_notation}{wikipedia}).
@@
 \item {\color{blue}L141. In any case what does "(sensu World Meteorological Organisation)" mean in this sentence?}
-In general it means "in the sense of".
+In general, it means "in the sense of".
 But in this context, it could also be read as "refer to" for further reading.
-"Essential climate Variables" is a term defined and commonly used by the WMO that might not be well known in the plant physiology community.
+"Essential Climate Variables" is a term defined and commonly used by the WMO that might not be well known in the plant physiology community.
 
@@
 (2003), we find an inaccuracy in our manuscript text which should read "[...] from before 1997 [...]".
-The systematic uncertainty in the pre 1997 ozone date is of the order of $10\,\%$ in AOT.
-Which means that the uncertainty in each individual measurement is lower but in the order of the trend in background ozone.
+The systematic uncertainty in the pre-1997 ozone date is of the order of $10\,\%$ in AOT.
+This means that the uncertainty in each measurement is lower but in the order of the trend in background ozone.
 This concerns the whole historical monitoring period at Svanhovd.
-Combining all data gives a good confidence in the observed seasonality rather than each individual measurement.
-From this we find, that 2019 was a rather usual year in terms of ozone and weather.
-Without relying on an ozone climatology, we could not make this assessment.
+Combining all data gives good confidence in the observed seasonality rather than each measurement.
+From this, we find, that 2019 was a rather usual year in terms of ozone and weather.
+Without relying on ozone climatology, we could not make this assessment.
 In \href{https://acp.copernicus.org/articles/21/15647/2021/}{Falk et al.
@@
 \item {\color{blue}L255-256. Why "deduce"? Now the numbers given, 800 W/m2 and 200 W/m2, are very rounded - are these measurements or assumptions or limits?}
-We assume the referee refers to L225-226? We agree that "deduce" is not the right term in this context and will rephrase. These are limits based on the observations.
+Do we assume the referee refers to L225-226? We agree that "deduce" is not the right term in this context and will rephrase. These are limits based on the observations.
 
@@
 Why would you consider an agricultural criteria for forests anyway?
-Why not apply the MODIS response for deciduous trees too?} The agricultural rule refers to the bud burst of birch trees.
+Why not apply the MODIS response for deciduous trees too?}
+The agricultural rule refers to the bud burst of birch trees.
 For a thorough discussion of the use of accumulated mean temperatures in estimating the start and end of the growing season in Scandinavia see e.g.
@@
 \\The choice of $A_\mathrm{start}$ for a central European latitude is indeed misleading.
-This was a fault in the model set up and will be corrected.
+This was a fault in the model setup and will be corrected.
 
@@
 
-\item {\color{blue}L345.
-"Due to the lack of quantitative field observation...".
-I didn't follow the logic here.} We qualitatively checked the start of the GS by looking at the live stream from a nearby webcam.
-This cannot be used for a quantitative analysis.
-Hence, we take the measurement of snow depth at a meteorological station into consideration instead.
-From this observation, we inferred a latency period of about one month after snow melt for the gras to sprout.
+\item {\color{blue}L345. "Due to the lack of quantitative field observation...". I didn't follow the logic here.}
+We qualitatively checked the start of the GS by looking at the live stream from a nearby webcam.
+This cannot be used for quantitative analysis.
+Hence, snow depth measured at a meteorological station is taken into consideration instead.
+From this observation, we inferred a latency period of about one month after snowmelt for the grass to sprout.
 We will rephrase this sentence to make this clearer.
@@
 This employs the deposition component of the $\mathrm{DO_3SE}$ model that estimates the transfer of mass (i.e.
-ozone) as a function of windspeed, convection, surface roughness, and vegetation $[\mathrm{O_3}]$ sink.
+ozone) as a function of wind speed, convection, surface roughness, and vegetation $[\mathrm{O_3}]$ sink.
 We will add additional detail in the paper to make this clear.}

### 2022-01-13 review/response-RC1.tex
Checked with grammaly

@@
 We will address all \textbf{general comments} raised regarding the structure and readability of the manuscript in our next revision or resubmission.
-We consider to separate the manuscript in Introduction--Methods--Results--Conclusions and will improve on the "methods" Section to make our results more comprehensible.
-We will refine the language to reduce ambiguity and supplement our statements with appropriate citations where indicated by the referee.\\
+We consider separating the manuscript in Introduction--Methods--Results--Conclusions and will improve on the "methods" Section to make our results more comprehensible.
+We will refine the language to reduce ambiguity and supplement our statements with appropriate citations that were indicated by the referee.\\
 As the manuscript will, hence, undergo major structural changes, we will not address all \textbf{specific / technical comments} in detail at this point.
 We will take all specific and technical comments which are still relevant into consideration.\\
-In the following, we shall give a brief respond to all relevant issues.
-
+In the following, we shall give a brief response to all relevant issues.
 \subsection*{General comments} 
@@
 
-    \item {\color{blue} "the [...] study is highly speculative and not based on any experimental evidence [...]"} In the revision of our work, we will emphasize that the main objective of our work is to explore the possibilities of readily accessible data (i.e.
+    \item {\color{blue} "the [...] study is highly speculative and not based on any experimental evidence [...]"} 
+    In the revision of our work, we will emphasize that the main objective of our work is to explore the possibilities of readily accessible data (i.e.
 long-term meteorological temperature observations, $\mathrm{CO_2}$ flux estimations from remote sensing) to provide alternatives to site-specific data that is needed to parameterize plant physiology in models such as the $\mathrm{DO_3SE}$ model used in our study.
-    \item {\color{blue} "Many claims sound superficial [...], such as [...] the alleged deviation of the years 2018 and 2019 from the site's climatology [...]"} We acknowledge that we have not used tools such as the \emph{p-value} (for a discussion see \href{https://doi.org/10.1080/00031305.2016.1154108}{statement of the American Statistical Association (2016)}).
+    \item {\color{blue} "Many claims sound superficial [...], such as [...] the alleged deviation of the years 2018 and 2019 from the site's climatology [...]"}
+    We acknowledge that we have not used tools such as the \emph{p-value} (for a discussion see \href{https://doi.org/10.1080/00031305.2016.1154108}{statement of the American Statistical Association (2016)}).
 We will make our statistical approach clearer where necessary (see also more detailed responses below) and replace terms such as "significantly different" where significance has not been shown appropriately.
-    \item {\color{blue} "The manuscript is too long and the organization is confused [...] no clear division between methodological section and results section"} We will address this in the next revision by fracturing the manuscript clearly into separate sections containing methods, results, and discussion.
+    \item {\color{blue} "The manuscript is too long and the organization is confused [...] no clear division between methodological section and results section"} 
+    We will address this in the next revision by fracturing the manuscript clearly into separate sections containing methods, results, and discussion.
 
@@
 Much more credible is the following explanation based on stratospheric intrusions.}
-We agree with the referee that a vegetation covered by snow would not release BVOCs to the atmosphere.
+We agree with the referee that vegetation that is covered by snow would not release BVOCs into the atmosphere.
 We implicitly assume this.
-Therefore, we did not explicitly state that BVOCs are not among the ozone precursors that accumulate in the course of winter.
+Therefore, we did not explicitly state that BVOCs are not among the ozone precursors that accumulate in winter.
 We will rephrase the sentence accordingly: "[...] ozone and its non-biogenic precursors [...]"
@@
 It sounds very bad written this way: vegetation does not accumulate ozone because ozone it is not bioaccumulative. Did you mean the dose?}
-Thank you for pointing this out. We will rephrase accordingly and refer instead the accumulated ozone dose.
+Thank you for pointing this out. We will rephrase accordingly and refer instead to the accumulated ozone dose.
 
-\item {\color{blue}Line 77.
-"($\left<[\mathrm{O_3}]\right>$ = 36-54ppb)" Please explain the formalism.
-What do different brackets mean?} We will make this notation and formalism clearer in the text.
+\item {\color{blue}Line 77. "($\left<[\mathrm{O_3}]\right>$ = 36-54ppb)" Please explain the formalism. What do different brackets mean?} 
+We will make this notation and formalism clearer in the text.
 The "$[X]$" notation commonly refers to the concentration of a chemical species X.
 Parts-per-billion (ppb) is a volume mixing ratio (VMR) and strictly speaking no concentration.
-Though, if you calculate the amount of ozone per volume from ppb to $\mathrm{\mu g\,m^{-3}}$ meter, assuming a pressure of 1 atmosphere, temperature of 298K and use the ideal gas law, you get a factor $\sim 2$.
+Though, if you calculate the amount of ozone per volume from ppb to $\mathrm{\mu g\,m^{-3}}$ meter, assuming a pressure of 1 atmosphere, the temperature of 298K and use the ideal gas law, you get a factor $\sim 2$.
 Therefore, concentrations and VMRs are often used synonymously.
@@
 "A substantial body of evidence exists that suggests flux-based metrics, that relate stomatal ozone uptake to vegetation damage, are biologically more relevant for risk assessments than exposure-based metrics."
-Well, please cite at least some works of this "substantial body"} Thank you, we will include relevant citations.
+Well, please cite at least some works of this "substantial body"} 
+Thank you, we will include relevant citations.
 
@@
 \item {\color{blue}Line 165. "luftkvalitet.no" What is it? And EBAS? Please make them clear.}
-"luftkvalitet.no" is the webinterface to the database of air quality measurements in Norway operated by NILU.
-EBAS is a database with webinterface operated by NILU as well.
+"luftkvalitet.no" is the web interface to the database of air quality measurements in Norway operated by NILU.
+EBAS is a database with a web interface operated by NILU as well.
 It comprises global air quality measurements, e.g.
@@
 \item {\color{blue}Line 182.
-"This indicates that the vegetation was more affected by ozone in 2018 than in 2019." Being affected by visible symptoms does not necessarily mean having suffered biomass or productivity reduction.} It is correct that biomass reduction and visible damage due to ozone are not the same.
-We discuss this briefly in Sec.~5 but will elaborate on it to make the difference clearer.
+"This indicates that the vegetation was more affected by ozone in 2018 than in 2019." Being affected by visible symptoms does not necessarily mean having suffered biomass or productivity reduction.}
+It is correct that biomass reduction and visible damage due to ozone are not the same. We discuss this briefly in Sect.~5 but will elaborate on it to make the difference clearer.
 
@@
 We will remove the term "high".
-However, ozone concentrations of $40\,\mathrm{ppb}$ have been chosen as critical level for the metric accumulated over a threshold (AOT) to gage ozone exposure toxic to ecosystems.
+However, ozone concentrations of $40\,\mathrm{ppb}$ have been chosen as a critical level for the metric accumulated over a threshold (AOT) to gauge ozone exposure toxic to ecosystems.
 This metric was first introduced at the UN/ECE-Workshop in Egham in 1992.
@@
 Ok, but was it then applied to this work? Please write it.}
-We will rephrase: "A method for gapfilling data has been presented in Falk et al. (2021) and applied in this work."
+We will rephrase: "A method for gap-filling data has been presented in Falk et al. (2021) and applied in this work."
 
@@
 In the recent version of the manuscript, we computed the excess number of days with residuals larger than $1\sigma$ first.
-This means, we calculate the daily mean and std for 21 years $\times$ 24 hours (264 data points).
+This means we calculate the daily mean and std for 21 years $\times$ 24 hours (264 data points).
 From this, we find the residuals in 2018/19.
-In the next step, we explicitly assume that these are normal distributed -- we will elaborate on this assumption in the next revision and check the underlying distributions (but see also Fig.~4).
+In the next step, we explicitly assume that these are normally distributed -- we will elaborate on this assumption in the next revision and check the underlying distributions (but see also Fig.~4).
 Due to natural variability (diurnal, annual) in the data, we chose a $1\sigma$ level for the significance test.
@@
 But the diurnal cycle could be perhaps removed to smooth the data.
-Second, by grouping the data into months, we find number of days with observations sufficiently diverging from the mean in each month.
+Second, by grouping the data into months, we find the number of days with observations sufficiently diverging from the mean in each month.
 We count the number of days with residuals exceeding $1\sigma$ and compute the ratio with the number of days of observation.
-If this data was normal distributed, we would expect ~18\% of data in this interval.
+If this data were normally distributed, we would expect ~18\% of data in this interval.
 If the percentage of days exceeding $1\sigma$ is larger, we state that we find a prominent excess of days with high ozone.
@@
 We then add the climatology of Svanvik to retrieve a reconstruction of the missing data.\\
-As 2019 does not deviated much from the norm by means of temperature, precipitation, and irradiance, we assume it for a normal year.
-From this, we find a probable offset of the historical ozone climatology compared to present day and correct for this.
-We compared this reconstruction with high resolution air quality model reanalysis and found a similar accuracy compared to observation.
-For details, please refer the cited work Falk et al (2021).
+As 2019 does not deviate much from the norm by means of temperature, precipitation, and irradiance, we assume it for a normal year.
+From this, we find a probable offset of the historical ozone climatology compared to the present day and correct for this.
+We compared this reconstruction with high-resolution air quality model reanalysis and found a similar accuracy compared to observation.
+For details, please refer to the cited work Falk et al (2021).
 
@@
 Why show a probability density if you are plotting a multiannual average?
-Or does the line represent the median instead?} In Fig.~4, we show the distribution of ozone data in 2-dimensions.
+Or does the line represent the median instead?}
+In Fig.~4, we show the distribution of ozone data in 2-dimensions.
 The dotes represent the daily averages (but only every 10th day is shown).
-We will rephrase the caption to correct for this.
-We show the daily average to infer that not all daily mean ozone are strictly normal distributed (tails in one or the other direction) but it is also visually deductible that most data fall in the $1\sigma$ band justifying the chosen confidence intervals for a normal distribution.
+We will rephrase the caption to correct this.
+We show the daily average to infer that not all daily mean ozone is strictly normal distributed (tails in one or the other direction) but it is also visually deductible that most data fall in the $1\sigma$ band justifying the chosen confidence intervals for a normal distribution.
 We shall, however, elaborate on this by statistical means.
@@
 Here you state that photosynthesis is already active at DOY 100 and is at 60\% of its maximum!}
-It is true that coniferous trees can, in principle, take up $\mathrm{CO_2}$ also in winter and the fact that they can suffer from something called frost drought shows that.
-With respect to the spring peak that was mentioned earlier, this means that coniferous trees may experience these elevated ozone concentrations regularly.
+Coniferous trees can indeed, in principle, take up $\mathrm{CO_2}$ also in winter and the fact that they can suffer from something called frost drought shows that.
+Concerning the spring peak that was mentioned earlier, this means that coniferous trees may experience these elevated ozone concentrations regularly.
 Concerning the contradiction in the statements -- if the conditions are right, coniferous trees can quickly reach 60\% of their max photosynthesis but stop photosynthesis the days after.
-Conditions to keep up this levels of photosynthesis, however, would only occur from doy 100 onwards.
-To make sure that we do not falsely accumulate too much ozone from the spring peak periode (when coniferous trees show on-off photosynthesis), we consulted satellite observations to extrapolate the dates when vegetations' photosynthesis (including coniferous trees) at Svanhovd rose over the detectable threshold.
+Conditions to keep up these levels of photosynthesis, however, would only occur from doy 100 onwards.
+To make sure that we do not falsely accumulate too much ozone from the spring peak period (when coniferous trees show on-off photosynthesis), we consulted satellite observations to extrapolate the dates when vegetations' photosynthesis (including coniferous trees) at Svanhovd rose over the detectable threshold.
 
@@
 I don't understand the connection. What does the night uptake have to do with the
-occasionally complete ozone depletion?} This was actually a working hypothesis and should be removed from the text.
+occasionally complete ozone depletion?} 
+This was a working hypothesis and should be removed from the text.
 
@@
 And what are the numbers on the top right of each graph?}
-Regarding the test, see response to "L205" above.
-The number at the top right are described in the caption: "The annual positive/negative deviations are indicated in the respective corners (right upper/lower)." This is the percentage of days on an annual basis that exceed the $1\sigma$ level.
+Regarding the test, see the response to "L205" above.
+The number at the top right is described in the caption: "The annual positive/negative deviations are indicated in the respective corners (right upper/lower)." This is the percentage of days on an annual basis that exceed the $1\sigma$ level.
 This indicates the overall deviation from a normal year.
@@
 But the significance of the deviation should be statistically tested in another way.}
-Regarding the test, see response to "L205" above.
+Regarding the test, see the response to "L205" above.
 Our hypothesis test is that you would expect $\approx18\%$ of days exceeding the $1\sigma$ level (normal distribution).
-If we observe a higher percentage of days, we find an exceedence from the expectation.
-If this is not proper in terms of statistical terminology we can remove the term "significant" and call it exceedence (in case this is deemed a better term).
+If we observe a higher percentage of days, we find an exceedance from the expectation.
+If this is not proper in terms of statistical terminology we can remove the term "significant" and call it exceedance (in case this is deemed a better term).
 
@@
 "Bias-corrected cross-calibrated" ozone? What is it? And what is the "significance" of the concentration anomalies? Please explain.}
-Please see response to "Lines 206-208.".
+Please see the response to "Lines 206-208.".
 Cross-calibrated means that we used data from other monitoring stations in the region to reconstruct the missing data at Svanvik.
-Regarding "significance", see explanation above.
+Regarding "significance", see the explanation above.
 
@@
 The referee is correct that VPD can be high due to cold conditions and thus become a limiting factor.
-We did not find any unexpected or abnormal behavior of stomatal conductance with respect to VPD.
+We did not find any unexpected or abnormal behavior of stomatal conductance concerning VPD.
 This indicates that the parameterized VPD limitation on stomatal conductance might be applicable for high VPD caused by either drought or cold.
@@
 I don't seem to see any differences between 2018 and 2019}
-As we have indicated, the measurement down time in March and April was planned because most vegetation (in particular the one in the ozone garden) was covered by snow and hence not photosynthetically active (as the referee remarked in their comment regarding BVOC emission).
-"but a substantially higher number of peak [O3] were observed in 2018 than in 2019." refers back to Sec.~2 L188-195.
-Peak concentrations were calculated as exceedence over 40\,ppb in summer, explicitly excluding the spring peak.
-Even with two weeks of data missing July 2018, we found 50 elevated ozone event in 2018 compared to 18 times in 2019.
-This should qualify as "substantially higher number".
-This explanation will be added in the text for clarification.
+As we have indicated, the measurement downtime in March and April was planned because most vegetation (in particular the one in the ozone garden) was covered by snow and hence not photosynthetically active (as the referee remarked in their comment regarding BVOC emission).
+"but a substantially higher number of peak [O3] were observed in 2018 than in 2019." refers back to Sect.~2 L188-195.
+Peak concentrations were calculated as exceedance over 40\,ppb in summer, explicitly excluding the spring peak.
+Even with two weeks of data missing July 2018, we found 50 elevated ozone events in 2018 compared to 18 times in 2019.
+This should qualify as a "substantially higher number".
+This explanation will be added to the text for clarification.
 
@@
 Looking at the graph I understand that you assume an adaptation of the subarctic grasslands to the temperature distribution of the last decade (climate already changed) and not to the historical temperature distribution at your site.
-Is it reasonable to hypothesize such a rapid adaptation of vegetation to the new climate conditions?} Both \emph{acclimation} of individual plants during their life-time and adaptation of a plant population (\emph{evolutionary process}) can contribute to the vegetation's stomatal conductance response to temperature (e.g., \href{https://bg.copernicus.org/articles/8/1453/2011/}{Juan et al.; 2011}).
-The evolutionary process would lead to a decrease in number of individuals of those species which natural acclimation range differs strongly from the new norm.
+Is it reasonable to hypothesize such a rapid adaptation of vegetation to the new climate conditions?}
+Both \emph{acclimation} of individual plants during their lifetime and adaptation of a plant population (\emph{evolutionary process}) can contribute to the vegetation's stomatal conductance response to temperature (e.g., \href{https://bg.copernicus.org/articles/8/1453/2011/}{Juan et al.; 2011}).
+The evolutionary process would lead to a decrease in the number of individuals of those species whose natural acclimation range differs strongly from the new norm.
 As "perennial grassland" in our model consists of a range of species, the grassland as a whole may rapidly adapt.\\
-In addition, there is the phenomenon of \emph{epigenetics}, exemplified by Norway spruce trees growing from seeds.
-The temperatures experienced during seed maturation (climate at the site where the mother tree actually stands) affect the new trees' response to environmental factors determining bud set in autumn.
+In addition, there is the phenomenon of \emph{epigenetics}, exemplified by Norway spruce trees grown from seeds.
+The temperatures experienced during seed maturation (climate at the site where the mother tree stands) affect the new trees' response to environmental factors determining bud set in autumn.
 This means that the seedlings may be adapted to the site where they are produced, although the genetic background of the parents predicts a different response (e.g.
@@
 The relative stomatal conductance around noon when the highest values are expected is rather low (about 40\%).
-With our adjusted parameterization the relative stomatal conductance around noon reaches about 70--80\%.
-The standard deviation reflects the variability of growing conditions and should not be interpreted as statistical measure for to separate one distribution from the other.
-Even more data would perhaps not reduce the standard deviation in this case.
+With our adjusted parameterization, the relative stomatal conductance around noon reaches about 70--80\%.
+The standard deviation reflects the variability of growing conditions and should not be interpreted as a statistical measure to separate one distribution from the other.
+Even more, data would perhaps not reduce the standard deviation in this case.
 We will remove the term "substantial" and make our interpretation of the Figure clearer in the text.
@@
 Thank you for pointing this out.
-It is correct that net photosynthesis does not capture the real photosynthesis.
+It is correct that net photosynthesis does not capture real photosynthesis.
 However, gross photosynthesis is not among the available products from MODIS satellites.
@@
 According to the product website (\href{http://www.ntsg.umt.edu/project/modis/mod17.php}{MOD17}, last accessed Dec.
-2021) it is computed from observed absorbed photosynthetically active radiation (APAR), a species specific conversion efficiency parameter $\epsilon(\mathrm{temperature, water stress})$, temperature, and water stress.
-In Fig.~\ref{fig:modis_gpp} the estimated start and end of growing season from MODIS (Aqua/Terra) GPP over a $1\times 1$\,km patch centered at Svanhovd is shown.
+2021) it is computed from observed absorbed photosynthetically active radiation (APAR), a species-specific conversion efficiency parameter $\epsilon(\mathrm{temperature, water stress})$, temperature, and water stress.
+In Fig.~\ref{fig:modis_gpp} the estimated start and end of the growing season from MODIS (Aqua/Terra) GPP over a $1\times 1$\,km patch centered at Svanhovd is shown.
 They differ only slightly from the estimates using net photosynthesis.
@@
 ImageJ was used for thresholding the images to give silhouettes of the leaves.
-After scaling the images, the program was used for finding minimum Feret diameter.
+After scaling the images, the program was used for finding the minimum Feret diameter.
 This measure finds the largest width of a leaf at a $90^\circ$ angle to the length of the leaf (the Feret diameter).
@@
 We suppose, the referee refers to the 10.1\,m average height of one particular Scots pine forest which has been measured in 2004.
-We assumed that after 14 year 10.1\,m might not be the average height of this particular forest anymore.
+We assumed that after 14 years 10.1\,m might not be the average height of this particular forest anymore.
 In the absence of more recent measurements, we found it more reasonable to choose 13.5\,m the average tree height in the whole area in 2004 which included all tree species and ages.
@@
 Please confirm it by writing it in the text.}
-In Appendix~B2, we write that the "$\mathrm{DO_3SE}$ model requires hourly, continuous meteorologische observations." This means that all input data, including ozone, have to be gap filled before POD1 can be computed.
-The gap filling method for all data but ozone are described in appendix~B2.
-The used gap filling method for ozone has been summarized above (response to "Lines 206--208") and published in Falk et al.
+In Appendix~B2, we write that the "$\mathrm{DO_3SE}$ model requires hourly, continuous meteorological observations." This means that all input data, including ozone, have to be gap-filled before POD1 can be computed.
+The gap-filling method for all data but ozone is described in appendix~B2.
+The used gap-filling method for ozone has been summarized above (response to "Lines 206--208") and published in Falk et al.
 (2021).
@@
 If we take a look at, e.g.
-Fig.~8a subarctic parameterazation and 2018 (left hand side of the upper panel).
+Fig.~8a subarctic parameterization and 2018 (left-hand side of the upper panel).
 Open symbols represent a model simulation where SWP was taken into account.
@@
 The '--' symbolizes PPFD0.8 (earlier opening), the '+' PPFD1.2 (later opening).
-We shall have a look at the circles connected by a solide line representing the simulations with a GS which we have determined from temperatures.
+We shall have a look at the circles connected by a solid line representing the simulations with a GS which we have determined from temperatures.
 If the stomata open earlier / close later more ozone can be taken up (longer opening time), hence POD1 is larger than for the unchanged $f_\mathrm{light}$.
@@
 What effects? "Of the same order of magnitude as the interannual variability...": can you conclude it by comparing only two years? The same for line 402}
-In Section~3, we tried to show that 2019 is representative for a normal year while 2018 is more extreme.
-We should evaluate our data with this question in mind and confirm that the difference between 2018 and 2019 is indeed representative for the interannual variability.
+In Section~3, we tried to show that 2019 is representative of a normal year while 2018 is more extreme.
+We should evaluate our data with this question in mind and confirm that the difference between 2018 and 2019 is indeed representative of the interannual variability.
 Depending on the outcome, we will rephrase the sentences and substitute the term "interannual variability".
@@
 We used the relationship between biomass reduction and POD1 unchanged from the MM.
-This, however, is to some degree questionable because subarctic vegetation could be affected more or less by the same ozone dose than the central European species.
-Apparently, the caption of the table is not complete and parts of the explanation missing.
+This, however, is to some degree questionable because subarctic vegetation could be affected more or less by the same ozone dose as the central European species.
+The caption of the table is not complete and parts of the explanation missing.
 We explain how we derive the uncertainty ranges in L405--407 but this may not be clear enough.
@@
 Hence, we computed the biomass reduction based on simulations with SWP effects on POD1 turned off and MM $f_\mathrm{light}$.
-The uncertainties reported are actually the differences and not standard deviations.
+The uncertainties reported are the differences and not standard deviations.
 The differences are calculated between the simulation with MM $f_\mathrm{light}$ and "bespoke" GS and the simulations with PPFD0.8 (larger POD1 $\rightarrow$ larger biomass reduction) and PPFD1.2 (smaller POD1 $\rightarrow$ less biomass reduction).
@@
 comparison with experimental results in your work.}
-We will substitute the term to appropriately reflect our intensions and work.
+We will substitute the term to appropriately reflect our intentions and work.
 
@@
 causing the visible symptoms can result in a biomass reduction (damage).}
-You are absolutely right. We intended to state that visible damage and biomass reduction are not necessarily related in this paragraph. We agree that this is not clear and we will improve on this.
+You are right. We intended to state that visible damage and biomass reduction are not necessarily related in this paragraph. We agree that this is not clear and we will improve on this.
 
 \item {\color{blue}Line 441. Does "damage" mean "visible leaf symptoms"? Or does it mean biomass reduction?}
-See above. We will carefully rephrase the text with respect to the term "damage" and clearly distinguish between visible damage and biomass reduction.
+See above. We will carefully rephrase the text concerning the term "damage" and clearly distinguish between visible damage and biomass reduction.
 
@@
 capture the plant physiology of subactic vegetation if no comparisons with physiological measurements taken on subarctic vegetation are presented?}
-We will adresse this point by reframing the storyline of the manuscript towards a new method to derive stomatal conductance parameterizations based on climatological and remotely sensed data.
+We will address this point by reframing the storyline of the manuscript towards a new method to derive stomatal conductance parameterizations based on climatological and remotely sensed data.
 
@@
 \emph{ozone is removed from the atmosphere by dry deposition which is described as a network of resistances.
-In general, snow and ice surfaces are found to have a low resistance to ozone , hence less ozone can be taken up by the land surface which is a major sink in absence of photochemical reactions (winter).
+In general, snow and ice surfaces are found to have a low resistance to ozone, hence less ozone can be taken up by the land surface which is a major sink in absence of photochemical reactions (winter).
 Therefore, existing ozone accumulates in the subarctic winter boundary layer.
 The same goes for non-biogenic precursors.
-In spring these are chemically reactivated and increase the ozone production.
-At the same time coniferous trees might also start to produce BVOCs.
+In spring these are chemically reactivated and increase ozone production.
+At the same time, coniferous trees might also start to produce BVOCs.
 At the same time, most of the vegetation is still covered by snow and ice and dry deposition remains low.
@@
 Simultaneously, photochemical destruction of ozone also increases.
-Only part of ozone actually enters the stomata.}
+Only part of ozone enters the stomata.}
 
@@
 What does it mean? Please explain. Make an example.}
-The proposed method to use climate data for adjusting the parameterization of stomatal conductance can be formulated as optimization problem (maximizing the enclosed temperature PDF area) which would make it possible to derive new parameters in an more automized way.
-Now, considering enough stomatal conductance data were available from observation in well known climate conditions these data could be used to evaluate our method and train a model to find the optimal stomatal conductance parameterization based on climate data alone.
-As climate data is more readily available compared to actual measurement of stomatal conductance this could help to improve modeling of stomatal conductance globally.
+The proposed method to use climate data for adjusting the parameterization of stomatal conductance can be formulated as an optimization problem (maximizing the enclosed temperature PDF area) which would make it possible to derive new parameters in a more automized way.
+Now, considering enough stomatal conductance data were available from observation in well-known climate conditions these data could be used to evaluate our method and train a model to find the optimal stomatal conductance parameterization based on climate data alone.
+As climate data is more readily available compared to the actual measurement of stomatal conductance this could help to improve modeling of stomatal conductance globally.
 
-\item {\color{blue}Figure A1.
-How was the length of the growing seasons in the various years identified?
-By satellite?
-Other method?
-What does the gray band represent?} See response to "Line~432".
-The gray band represents the standard deviation.
-We will either remove or elaborate on this analysis.
+\item {\color{blue}Figure A1. How was the length of the growing seasons in the various years identified? By satellite? Other method? What does the gray band represent?} 
+See response to "Line~432". The gray band represents the standard deviation. We will either remove or elaborate on this analysis.
 
@@
 We will include an improved, short description of how the $\mathrm{DO_3SE}$ model incorporates the PM model to estimate evapotranspiration and hence water loss from the soil.
-However, for full details of this method it is best for readers to refer to Bücker et al.
+However, for full details of this method, readers should refer to Bücker et al.
 (2012).}
@@
 In summary, we will make the description of the $\mathrm{DO_3SE}$ model more comprehensible and comprehensive.
-However, in the interest of the main focus of this manuscript it may not be possible to give a full recap of all details concerning this well established model.
+However, in the interest of the main focus of this manuscript, it may not be possible to give a full recap of all details concerning this well-established model.
 

### 2022-03-31 VegOzone.tex
Change title
Seperate figures.
Update abstract

@@
 %\usepackage{rotating}
-
+\usepackage[nomarkers,figuresonly]{endfloat}
 \graphicspath{{pictures/}}
@@
 
-\title{Characterizing the response of subarctic European vegetation to key environmental variables for land surface modeling of ozone pollution and risk assessment}
+\title{Parameterization of the responses of subarctic European vegetation to key environmental variables for ozone risk assessment}
 
@@
 To date, only a few studies assessed the response of subarctic vegetation to variations in climate and air pollution.
-This study looks to fill this knowledge gap by examining essential climate variables, in particular ozone, over the past few decades.
-We evaluate the extent to which two recent years (2018 and 2019) deviate from climatic and \chem{[O_3]} norms and how these potentially more frequent future deviations may influence ozone damage to subarctic vegetation.
+With no bespoke parameterization for stomatal response to essential climate variables available, this study looks to fill this knowledge gap by examining an adjustment of existing, Javis type boreal parameterizations, based on local climate condition.
+To this end, we vary the temperature response function according to the climatological temperature distribution during the subarctic growing season.
+We evaluate the extent to which two recent years (2018 and 2019) deviate from climatic and \chem{[O_3]} norms and how these potentially more frequent future deviations may influence the projected ozone damage to subarctic vegetation.
 We find that 2018 was an anomalously warm and bright year, particularly in spring and early summer.
@@
 We apply the integrated flux-metric Phytotoxic Ozone Dose (POD) to determine ozone risk and damage to vegetation as a function of \chem{[O_3]}, environmental factors, and species-specific physiology.
-Our study suggests that using generic parameterizations in assessments likely leads to underestimating the risk of ozone damage in this region.
-We find that bespoke parameterizations of plant functional types (PFTs) for subarctic vegetation bio-types result in an ozone-induced biomass loss of $2.5$ to $17.4\,\unit{\%}$.
-For some species, this loss is up to $6\,\unit{\%}$ larger than projected from generic parameterizations.
+Our study suggests that using standard boreal parameterizations in assessments likely leads to underestimating the risk of ozone damage in this region.
+We find that local climate adjusted parameterizations of plant functional types (PFTs) for subarctic vegetation result in an ozone-induced biomass loss of $2.5$ to $17.4\,\unit{\%}$.
+For some species, this loss is up to $xx\,\unit{\%}$ larger than projected from standard boreal parameterizations used in general assessments.
 Efforts should be targeted towards accurately defining subarctic vegetations' physiological response to essential climate variables.
-Our method could help to improve regional and global scale biogeochemical cycling under current and future climates.
 \end{abstract}

### 2022-03-31 VegOzone.tex
Rephrase Introduction

@@
 To date, only a few studies assessed the response of subarctic vegetation to variations in climate and air pollution.
-With no bespoke parameterization for stomatal response to essential climate variables available, this study looks to fill this knowledge gap by examining an adjustment of existing, Javis type boreal parameterizations, based on local climate condition.
+This study looks to fill this knowledge gap by examining the uncertainty in risk assessments which are usinge more general boreal paramterizations.
+With no bespoke parameterization for stomatal response to essential climate variables available, we propose an adjustment of existing, Javis type boreal parameterizations, based on local climate condition.
 To this end, we vary the temperature response function according to the climatological temperature distribution during the subarctic growing season.
@@
 
+\newpage
+
 %\copyrightstatement{TEXT}  %% This section is optional and can be used for copyright transfers.
@@
 
-Ground-level ozone (\chem{O_3}) is a highly toxic pollutant known to cause damage to human health \citep{WHO2008} and to a variety of ecosystems around the World \citep{PT:Emberson2020}.
+Ground-level ozone (\chem{O_3}) is a highly toxic pollutant known to cause damage to a variety of ecosystems around the World \citep{PT:Emberson2020}.
 Ozone causes reduced photosynthesis, visible foliar injury, early senescence, and programmed cell death of plants \citep{PCE:Kangasjarvi2005}.
+Hence affecting the global carbon cycle \citep{BGS:Lombardozzi2013} and crop yields.
 Annual global yield losses of four major crops (wheat, rice, maize, and soybean) of $3-15\,\unit{\%}$ have been reported \citep{PJ:Ainsworth2017} and a suggested loss in primary production in forestry of $7\,\unit{\%}$ \citep{GCB:Wittig2009,EP:Matyssek2012} has been attributed to ozone.
-The crop yield loss studies indicate a threat to food security in rapidly developing countries, e.g., in East and South-East Asia \citep{GCB:Tang2013,NCC:Tai2014,AE:Chuwah2015,GCB:Mills2018}, where ozone concentrations related to enhanced pollutant emissions from the transport, residential, power generation, and industrial sectors have increased in the past few decades.
 Even though long-term (pre-1950) observations of ground-level ozone concentrations are scarce; evidence from trend analysis of those few sites suggests that since the industrial revolution, background concentrations in the northern hemisphere have at least doubled and continued to increase \citep{IPCC2013}.
@@
 
-The complex photochemical cycles leading to the formation of \chem{O_3} at ground-level involve precursor gases such as carbon monoxide (\chem{CO}) and hydrocarbons known as volatile organic compounds (VOCs) in the presence of nitrogen oxides (\chem{NO_x}).
-Despite a relatively short average tropospheric lifetime of approximately $22\,\unit{days}$ \citep{JGR:Stevenson2005,ACP:Young2013}, ozone and its precursors are subject to advection over long distances.
-Episodes of high ozone concentrations in the absence of local pollution sources can therefore often be attributed to the long-range transport of both ozone and its precursors.
-In northern Fennoscandia, episodes of elevated ozone concentrations in 2003 and 2006 have been traced back to ozone precursor emissions related to forest fires in southern and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}.
-Besides VOCs from anthropogenic sources, hydrocarbons are also emitted by vegetation itself in the form of terpenes and monoterpenes, so-called biogenic volatile substances (BVOCs).
-These emissions are thought to be a response to factors, such as thermal stress, defense against herbivores, or even attraction of pollinators \citep{TPS:Penuelas2003}.
-
-In the annual cycle in Boreal climates, a distinct maximum of ground-level ozone concentrations typically occurs in spring followed by lower concentrations throughout the summer.
-Several conditions unique to the Arctic are stimulating the formation of this spring peak.
-Major pathways of ozone and its precursors' removal from the boundary layer include dry deposition on vegetated surfaces and photochemistry \citep{RG:Clifton2020}.
-Dry deposition to bare ground or snow and ice-covered surfaces was found to be low \citep{ACP:Helmig2007}.
-This causes a  substantial suppression of ozone removal \citep{AE:Monks2000} and leads to a build-up of ozone and its precursors during winter.
-Come spring precursors are photochemically activated again, while snow cover typically continues to prevail and suppress ozone dry deposition.
-On the other hand, very high ozone concentrations occurring over shorter periods can often be attributed to deep tropopause folding events over the Arctic during winter.
-These events facilitate an intrusion of stratospheric air masses which are enriched in ozone \citep{JGR:Skerlak2015}.
-The spring peak is more pronounced in northern Fennoscandia than in more southern locations but shows a high interannual variability \citep{AB:Klingberg2009, BER:Klingberg2019}.
-\citet{ACP:Andersson2017} showed in a modeling study focused on Fennoscandia that the variability of ozone concentrations in winter can be attributed mainly to changes in atmospheric background (transport) of ozone, while summertime abundance is mostly affected by emissions of precursors in the rest of Europe.
+In the annual cycle, a distinct maximum of ground-level ozone concentrations typically occurs in spring followed by lower concentrations throughout the summer.
+This ozone spring peak is more pronounced in northern Fennoscandia than in more southern locations but shows a high interannual variability \citep{AB:Klingberg2009, BER:Klingberg2019}.
+\citet{ACP:Andersson2017} showed in a modeling study focused on Fennoscandia that the variability of ozone concentrations in winter can be mainly attributed to changes in atmospheric background (transport) of ozone, while summertime abundance is mostly affected by emissions of precursors in the rest of Europe.
+
+The precursor gases involved in the formation of ground-level ozone are carbon monoxide (\chem{CO}) and hydrocarbons known as volatile organic compounds (VOCs) in the presence of nitrogen oxides (\chem{NO_x}).
+In northern Fennoscandia, episodes of elevated summertime ozone concentrations in 2003 and 2006 have been traced back to ozone precursor emissions related to forest fires in southern and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}.
+photosynthetically active vegetation in the form of terpenes and monoterpenes, so-called biogenic volatile substances (BVOCs) and soil.
+These emissions are thought to be a response to factors, such as thermal stress, defense against herbivores \citep{Ecosystems:Ghimire2021}, or even attraction of pollinators \citep{TPS:Penuelas2003}.
+\citet{Ecosystems:Ghimire2021} found no clear evidence that general warming would increase the BVOC emissions in subarctic vegetation, while a reduction of herbivory would decrease them by $44-72\,\%$.
 
@@
 
-A substantial body of evidence exists that suggests flux-based metrics, that relate stomatal ozone uptake to vegetation damage, are biologically more relevant for risk assessments than exposure-based metrics.
-This is because they can account for particular species characteristics (i.e.
+Flux-based metrics, that relate stomatal ozone uptake to vegetation damage, are biologically more relevant for risk assessments than exposure-based metrics because they can account for particular species characteristics (i.e.
 physiology and phenology) as well as environmental conditions that can decouple the relationship between ozone concentration, ozone uptake, and consequent damage \citep{PT:Emberson2020}.
 These flux-based metrics are better suited to represent the actual risk to vegetation from ozone, especially in the climatically extreme parts of Europe \citep{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017}.
-Most of these previous studies have focused on the Mediterranean where soil moisture deficits are thought to be most influential in decoupling concentration from uptake.
-Relatively few studies have explored the situation in Northern European climates \citep[e.g.][]{iF:Juran2018}.
+But no bespoke parameterization for subarctic vegetation is available.
 
@@
 %
-with the hourly averaged stomatal ozone flux $\Phi_\mathrm{sto}$ (see Appendix Eq.~(\ref{eq:flux_stomata}) and a stomatal ozone flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}\,s^{-1}}.
+with the hourly averaged stomatal ozone flux $\Phi_\mathrm{sto}$ (see Appendix Eq.~(\ref{eq:flux_stomata}) and a stomatal ozone flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{Projected Leaf Area (PLA)}\,s^{-1}}.
 The flux threshold $\mathrm{y}$ represents the detoxification potential of the plant and is typically only exceeded during daylight hours (i.e.
@@
 This convention aims to develop an effects-based emission reduction policy that can target those ozone precursor emissions that are most influential in causing damage.
-To achieve this, the concept of critical levels (CLs) has been developed and applied \citep{Maas2016}.
+To achieve this, the concept of critical levels (CLs) has been adopted and applied \citep{Maas2016}.
 Exceedance of the CL is used to identify those areas across Europe that would benefit from targeted Europe-wide emission reductions.
-Methods to estimate $\mathrm{POD_y}$ and CLs are defined by the UNECE LRTAP Convention and described in a Mapping Manual \citep{ICP:MappingManual2017}.
-The CL is calculated by:
+Methods to estimate $\mathrm{POD_y}$ and CLs are defined by the UNECE LRTAP Convention and described in a Mapping Manual \citep[most recent version]{ICP:MappingManual2017}.
+The exceedance of the CL is calculated by:
 %
@@
 This study demonstrated the improved ability of flux-based metrics (in comparison to exposure-based metrics) to identify the geographical distribution of the risk of ozone damage and also showed that the ozone damage could extend into more northerly regions of Europe.
-For our study, we conducted a similar bio-monitoring study at Svanhovd, a site with agrometeorological and air pollution monitoring, to assess whether local ozone concentrations were capable of inducing damage to vegetation growing under northern Fennoscandia conditions.
-In 2018, we indeed observed ozone damage on clover and tobacco.
-In contrast, no such damage was found on the clovers in 2019.
-This indicates that environmental factors in 2018 differed from 2019.
-
-According to a report by the Norwegian Meteorological institute \citep{MetNOR2019}, the summer of 2018 was the warmest and driest ever recorded in eastern, western, and southern Norway.
-In the north (including Finnmark), it was amongst the warmest on record -- favorable conditions for ground-level ozone formation.
-An unusually weak and northward shifted jet stream allowed for a persistent high-pressure system above northern Europe, including Fennoscandia which blocked the low-pressure systems for several consecutive months.
-In the period May -- July, southern Norway experienced temperatures $4\,\unit{^\circ C}$ above normal and saw only about $60\,\unit{\%}$ of normal precipitation.
-Northern Norway as a whole had close to normal precipitation, one of the causes of the high rate of forest fires in that area.
-Thermal stress on vegetation was exceptional not only in large parts of Fennoscandia but also in much of Europe, where the influence of the high-pressure system extended even over five months (April/May, July--September).
-These conditions gave rise to massive forest fires in different parts of Europe and thus an increase in ozone precursors.
-A total of 2079 forest fires were registered in Norway in 2018, twice as many as in the preceding years 2016/17 \citep{DSB2019}, last accessed April 2020).
-In Sweden, about 500 fires had been reported (five times more than in a usual summer), and an estimated total of $25000\,\unit{hectare}$ burned down in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) \citep{SOU2019}.
-Boreal wildfires are found to emit ozone precursors \chem{CO} ($\chem{[CO]}/\chem{[CO_2]} \propto 6-13\,\unit{\%}$) and VOCs ($\chem{[VOCs]}/\chem{[CO_2]} \propto 0.5-1.5\,\unit{\%}$) \citep{AE:Cofer1990}.
-Coincident peak \chem{[O_3]} is found in ozone monitoring data (Fig.~\ref{fig:data_svanvik_2018_2019}) in July.
-It is very likely that elevated ozone, like during the 2003 drought period \citep{JGR:Solberg2018}, was promoted by a combination of various factors such as wildfires, reduced cloud cover (increased solar radiation), reduced dry deposition and turbulent mixing due to the stagnant weather conditions, and increased BVOC emissions.
-As such events may become more frequent in the future, conditions at Svanhovd for the growing seasons 2018 shall serve as a reference for probable future conditions in northern Fennoscandia.
-Conditions in 2019 represent the present climate.\\
 
@@
 It is important to understand the implications of using these more generic parameterizations for more extreme climate regions.
-In our study, we explore how risk might be over- or underestimated using generic vs bespoke subarctic parameterization.
-We also investigate the effects of the interannual variability of essential climate variables (sensu World Meteorological Organization) on future risk of vegetation damage.
+In our study, we explore how risk might be over- or underestimated using Boreal parameterization and explore an adjustment of Javis type temperature response functions based on climate data.
 This allows an understanding of how the threat from ozone may change in the future and which are the key aspects of plant physiology that might determine the potential risk.
-Hence, we determine which parameters require special attention for more accurate pollution impact assessment and more generally for describing gas exchange response to a changing climate that will influence biogeochemical cycling.
-This might also be one key to solve long-standing issues of earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface models \citep{GMD:Poulter2015,JAMES:Lawrence2019}.
-
-The observation site Svanhovd and available observational data are presented in Section~\ref{sec:data}.
-From these data, we derive multi-annual averages of essential climate variables (temperature, precipitation, global irradiance, and ozone) and evaluate the deviation of 2018/19 from this norm (Section~\ref{sec:stats}).
-In Section~\ref{sec:do3se}, we use the $\mathrm{DO_3SE}$ model to estimate the ozone uptake by natural and semi-natural vegetation to investigate the main environmental drivers for projected ozone damage risk in 2018.
-To this end, we derive bespoke parameterizations for subarctic species to assess over- and underestimations arising from the use of generic parameterizations established from less climatically extreme conditions.
-We will summarize our results in Section~\ref{sec:conc} and discuss how this research can support improvements in future assessments of pollution risk and more generally biogeochemical cycling.
+Our results may hint a solution to solve long-standing issues of earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface models and display over- or underproductivity in the Arctic \citep{GMD:Poulter2015,JAMES:Lawrence2019}.
+
+Data and methods including the DO3SE model are described in Section 2.
+The model experiments and derived parameterizations are summarized in Section 3.
+In Section 4, we present our modelling results.
+We discuss our results and give an outlook on how these could improve future assessments of pollution risk.
 
@@
 
+For our study, we conducted a similar bio-monitoring study at Svanhovd, a site with agrometeorological and air pollution monitoring, to assess whether local ozone concentrations were capable of inducing damage to vegetation growing under northern Fennoscandia conditions.
+In 2018, we indeed observed ozone damage on clover and tobacco.
+In contrast, no such damage was found on the clovers in 2019.
+This indicates that environmental factors in 2018 differed from 2019.
+
+According to a report by the Norwegian Meteorological institute \citep{MetNOR2019}, the summer of 2018 was the warmest and driest ever recorded in eastern, western, and southern Norway.
+In the north (including Finnmark), it was amongst the warmest on record -- favorable conditions for ground-level ozone formation.
+An unusually weak and northward shifted jet stream allowed for a persistent high-pressure system above northern Europe, including Fennoscandia which blocked the low-pressure systems for several consecutive months.
+In the period May -- July, southern Norway experienced temperatures $4\,\unit{^\circ C}$ above normal and saw only about $60\,\unit{\%}$ of normal precipitation.
+Northern Norway as a whole had close to normal precipitation, one of the causes of the high rate of forest fires in that area.
+Thermal stress on vegetation was exceptional not only in large parts of Fennoscandia but also in much of Europe, where the influence of the high-pressure system extended even over five months (April/May, July--September).
+These conditions gave rise to massive forest fires in different parts of Europe and thus an increase in ozone precursors.
+A total of 2079 forest fires were registered in Norway in 2018, twice as many as in the preceding years 2016/17 \citep{DSB2019}, last accessed April 2020).
+In Sweden, about 500 fires had been reported (five times more than in a usual summer), and an estimated total of $25000\,\unit{hectare}$ burned down in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) \citep{SOU2019}.
+Boreal wildfires are found to emit ozone precursors \chem{CO} ($\chem{[CO]}/\chem{[CO_2]} \propto 6-13\,\unit{\%}$) and VOCs ($\chem{[VOCs]}/\chem{[CO_2]} \propto 0.5-1.5\,\unit{\%}$) \citep{AE:Cofer1990}.
+Coincident peak \chem{[O_3]} is found in ozone monitoring data (Fig.~\ref{fig:data_svanvik_2018_2019}) in July.
+It is very likely that elevated ozone, like during the 2003 drought period \citep{JGR:Solberg2018}, was promoted by a combination of various factors such as wildfires, reduced cloud cover (increased solar radiation), reduced dry deposition and turbulent mixing due to the stagnant weather conditions, and increased BVOC emissions.
+As such events may become more frequent in the future, conditions at Svanhovd for the growing seasons 2018 shall serve as a reference for probable future conditions in northern Fennoscandia.
+Conditions in 2019 represent the present climate.\\
+
 To summarize, we have developed bespoke parameterizations of common vegetation types (coniferous and deciduous trees, perennial grassland) to a subarctic climate and studied their effect on $\mathrm{POD_y}$.

### 2022-04-01 VegOzone.tex
Restructure section2

@@
 This study looks to fill this knowledge gap by examining the uncertainty in risk assessments which are usinge more general boreal paramterizations.
-With no bespoke parameterization for stomatal response to essential climate variables available, we propose an adjustment of existing, Javis type boreal parameterizations, based on local climate condition.
+With no bespoke parameterization for stomatal response to essential climate variables available, we propose an adjustment of existing, Boreal parameterizations, based on local climate condition.
 To this end, we vary the temperature response function according to the climatological temperature distribution during the subarctic growing season.
@@
 
-The described methods for risk assessment rely on accurate representation of plant physiological responses to key environmental variables and are most often parameterized using data for less extreme climates, i.e., temperate or continental climates; and to a lesser extent, Boreal and Mediterranean climates.
+The described methods for risk assessment rely on accurate representation of plant physiological responses to key environmental variables and are most often parameterized using data for less extreme climates, i.e., temperate, continental, or Boreal climates.
 Parameterization of these models for more extreme bio-geographical regions such as subarctic climates has not been performed and relies on assuming that less extreme Boreal parameterization will be applicable.
-It is important to understand the implications of using these more generic parameterizations for more extreme climate regions.
-In our study, we explore how risk might be over- or underestimated using Boreal parameterization and explore an adjustment of Javis type temperature response functions based on climate data.
-This allows an understanding of how the threat from ozone may change in the future and which are the key aspects of plant physiology that might determine the potential risk.
-Our results may hint a solution to solve long-standing issues of earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface models and display over- or underproductivity in the Arctic \citep{GMD:Poulter2015,JAMES:Lawrence2019}.
-
-Data and methods including the DO3SE model are described in Section 2.
+It is important to understand the implications of using these in the subarctic.
+In our study, we explore how risk might be over- or underestimated using Boreal parameterization and explore an adjustment of temperature and light response functions based on climate data.
+This allows an understanding of which are the key aspects of plant physiology that might determine the potential risk and also how the threat from ozone may change in the future.
+Our results may hint a solution to solve long-standing issues of earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface models and display over- or underproductivity in the Arctic \citep{GMD:Poulter2015,JAMES:Lawrence2019}.\\
+
+Data and methods including the DO3SE model and the localization of response function are described in Section~\ref{sec:dam}.
 The model experiments and derived parameterizations are summarized in Section 3.
@@
 %%
-\section{Data acquisition}
-\label{sec:data}
-
-We performed ozone concentration measurements in the subarctic, during the growing seasons of 2018 and 2019.
-Our chosen location was Svanhovd ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m\,a.s.l.}$) for several reasons: (1) Ozone had been monitored there in the past, (2) an agrometeorological station is measuring, e.g.
+\section{Data \& methods}
+\label{sec:dam}
+
+\subsection{DO3SE model}
+\label{sec:do3se}
+To investigate the \chem{O_3} damage risk in the subarctic, we model the ozone dose of natural (coniferous and deciduous trees) and semi-natural (grassland) vegetation using the Deposition of Ozone for Stomatal Exchange ($\mathrm{DO_3SE}$) model \citep{ACP:Bueker2012} which computes estimates of the total and stomatal deposition of \chem{O_3} to selected land-cover types and plant species according to UNECE LRTAP (Long-Range Transboundary Air Pollution) methodologies for effects-based risk assessment.
+
+\subsubsection{Model description}
+\label{subsec:model_describtion}
+The opening of stomata is governed by different species-specific and environmental factors.
+In $\mathrm{DO_3SE}$ it is described by a multiplicative empirical model \citep{PTRS:Jarvis1976,Emberson2000,ICP:MappingManual2017}:
+\begin{equation}
+  g^k_\text{sto} = g^k_\text{max} \cdot f^k_\text{phen} \cdot f^k_\text{light} \cdot \max{\left\{f^k_\text{min}, f^k_\mathrm{T} \cdot f^k_\mathrm{VPD} \cdot f^k_\text{SWP}\right\}}.
+  \label{eq:stomatal}
+\end{equation}
+%
+Where $g_\mathrm{max}$ is a species-specific maximum stomatal conductance value that is then modified by seasonal and environmental factors that vary within a range $0-1$.
+These factors are empirically determined and account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_\mathrm{T}$), water vapor pressure deficit ($f_\mathrm{VPD}$), and soil water potential ($f_\text{SWP}$).
+All factors differ with plant functional type (PFT) denoted with $k$.
+
+The phenology ($f_\mathrm{phen}$) is described by a stepwise defined linear function that allows for a increase of stomatal conductance in spring (start of growing season) and a decrease in fall (end of growing season).
+Furthermore, mid-season dips can also be accounted for.
+
+The temperature adjustment function is defined as
+%
+\begin{equation}
+  f_\mathrm{T} = \frac{T_\text{air}-T_\text{min}}{T_\text{opt}-T_\text{min}} \cdot \left(\frac{T_{\text{max}}-T_\text{air}}{T_{\text{max}}-T_\text{opt}}\right)^\beta,
+  \label{eq:f_temp}
+\end{equation}
+with $\beta = \frac{T_\text{max}-T_\text{opt}}{T_\text{opt}-T_\text{min}}$ a mixed reciprocal polynomial function of temperature.
+The shape parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various PFT and $T_\mathrm{air}$ is the near-surface air temperature.
+All temperatures are defined in units of \unit{^\circ C}.
+
+The water vapor pressure deficit function is
+%
+\begin{equation}
+  f_\mathrm{VPD} = f_\text{min}+(1-f_\text{min}) \cdot \frac{D_\text{min} - \text{VPD}}{D_\text{min}-D_\text{max}}
+  \label{eq:f_vpd}
+\end{equation}
+%
+where VPD is the leaf to air vapor pressure deficit in \unit{kPa} with $f_\text{min}$, $D_\text{min}$, $D_\text{max}$ describing the relative stomatal conductance to changes in vapor pressure deficit.
+
+The wavelength band ($400-700\,\unit{nm}$) plant chlorophyll responds to is called photosynthetic active radiation (PAR).
+Its integral is the photosynthetic photon flux density (PPFD).
+The relationship between relative $g_\mathrm{sto}$ and PPFD is given by
+%
+\begin{equation}
+  f_\text{light} = 1-\text{exp}(-\alpha_\text{light}\cdot \text{PPFD})
+  \label{eq:flight}
+\end{equation}
+%
+where $\alpha_\mathrm{light}$ is a slope parameter describing the extent of stomatal opening at low light intensities.
+
+The $\mathrm{DO_3SE}$ model as described in \citet{ACP:Bueker2012} is used to simulate SWP across a PFT specific root depth according to the Penman--Monteith energy balance method that drives water cycling through the soil--plant--atmosphere system.
+The stomatal response to changes in SWP ($f_\mathrm{SWP}$) are, thus, given by:
+%
+\begin{equation}
+  f_\mathrm{SWP} = \text{min}\left\{1, \text{max}\left\{f_\mathrm{min}, \frac{(1 - f_\mathrm{min})\cdot(\mathrm{SWP_{min}} - \mathrm{SWP})}{\mathrm{SWP_{min}} - \mathrm{SWP_{max}}} + f_\mathrm{min} \right\} \right\}
+    \label{eq:fsw}
+\end{equation}
+%
+where SWP is the soil water potential across the root zone and $\mathrm{SWP_{min}}$ and $\mathrm{SWP_{max}}$ are the parameters describing the $f_\mathrm{SWP}$ relationship.
+
+$\Phi_\mathrm{sto}$ in Eq.~(\ref{eq:pod}) as computed by the $\mathrm{DO_3SE}$ model is a fraction of the total \chem{O_3} deposition flux.
+A standard resistance analogous approach \citep{AE:Wesely1989} is used to transport \chem{O_3} from the atmospheric reference height to the vegetated surface.
+The calculation of aerodynamic $R_a$ and quasi-laminar boundary layer resistance $R_b$ follow \citet{WASP:Simpson2003,ACP:Simpson2012}.
+The surface resistance $R_\mathrm{surf}$ consists of stomatal $r_\mathrm{sto}$ and non-stomatal canopy resistances that includes external plant surfaces $r_\mathrm{ext}$, aerodynamic within-canopy $R_\mathrm{inc}$ and ground surfaces or soil resistances $R_\mathrm{gs}$ \citep{WASP:Simpson2003,ACP:Simpson2012}.
+Under the assumption that \chem{[O_3]} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf, $r_\mathrm{sto}$ and $r_\mathrm{ext}$ are calculated per leaf/needle area (note lower case $r$).
+%
+\begin{equation}
+  R_\mathrm{sur} = \frac{1}{\frac{\mathrm{LAI}}{r_\mathrm{sto}}+\frac{\mathrm{SAI}}{r_\mathrm{ext}}+\frac{1}{R_\mathrm{inc}+R_\mathrm{gs}}}
+\end{equation}
+%
+Stomatal resistance is the inverse of stomatal conductance $g_\mathrm{sto}$ in Eq.~(\ref{eq:stomatal}).
+
+\subsubsection{Input data and gap filling methodology}
+\label{subsec:ozone_reco}
+The $\mathrm{DO_3SE}$ model requires hourly, continuous ozone and meteorological observations of $2\,\unit{m}$ temperature $T_\mathrm{2m}$, global irradiation $Q_0$, precipitation, $2\,\unit{m}$ wind $u_\text{2m}$, and vapor pressure deficit VPD.
+VPD has been calculated from observed $T_\text{2m}$ and relative humidity
+\begin{equation}
+  \text{VPD} = P_s(T) \cdot \left(1-\frac{\text{relHum}}{100}\right),
+\end{equation}
+with saturation vapor pressure $P_s(T)$ in \unit{hPa} derived from Arden Buck equation~\citep{JAP:Buck1981, Buck2012}.
+
+For our study, we chose Svanhovd ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m\,a.s.l.}$).
+Reasons for our choice: (1) Ozone had been monitored there in the past, (2) an agrometeorological station is measuring, e.g.
 $2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths, and (3) a national surveillance of air pollution (\chem{NO_x}, \chem{SO_x}, PM, radiological species) is ongoing.
-Most of these measurements are performed by the Norwegian Institute for Air Research (NILU), who also conducted ozone measurements on our behalf.
-In addition, we collected atmospheric monitoring and ozone monitoring data taken at Svanhovd over the past $35\,\unit{years}$.\\
-
-All measured essential climate variables for the 2018/19 growing season were obtained through \href{luftkvalitet.no}{luftkvalitet.no} operated by \citet{NILU_AIRQ} (last accessed May 2020).
-Our ozone monitoring data have been added to luftkvalitet.no and the EBAS database operated by \citet{NILU_EBAS} and are thus openly accessible.
+Most of these measurements are performed by the Norwegian Institute for Air Research (NILU), who also conducted ozone measurements on our behalf during the 2018/19 growing season.
+In addition, we collected atmospheric monitoring and ozone monitoring data taken at Svanhovd over the past $35\,\unit{years}$.
+
+All measured essential climate variables for the 2018/19 growing season were provided by NILU through the web portal \href{luftkvalitet.no}{luftkvalitet.no} (last accessed May 2020).
+The ozone monitoring data have been added to a database for air-quality monitoring called EBAS as well operated by NILU and are thus openly accessible.
 Long-term ozone observation data (1986--1996) were obtained from EBAS.
 Note that data from before the 1990s do not follow the high-quality standard procedures implemented nowadays and have to be treated with care \citep{NILU2003}.
-Especially, ozone monitors did not undergo regular re-calibration.
-That probably led to drifts in the observed data and may impose a false trend which shall not be of concern for us as we are interested in the seasonality.
+Because ozone monitors did not undergo regular re-calibration drifts in the observed data occurred imposing false trends.
+This does not concern the general ozone seasonality which we are particularly interested in.
 The long-term data sets of agrometeorological variables including temperature and precipitation are available from September 1992 to the present day \citep[LandbruksMeteorologiske Tjeneste][note the station name here is Pasvik]{LMT_NIBIO}.
@@
 In contrast, no such visible damage was found on the clovers in 2019, although it was observed on the sensitive tobacco cultivar (Bel-W3).
-This indicates that the vegetation was more affected by ozone in 2018 than in 2019.
 
@@
 \end{figure*}
+
+The following gap filling methodology was devised for the meteorological input data:
+\begin{itemize}
+\item Single hours of missing data were filled by taking the average of the hourly values before, and after, the missing value.
+\item Several consecutive hours of missing data (23 or less) were filled by taking the average of the corresponding hour the day before, and the day after; and repeating this for each missing hour of data.
+If data were unavailable from that hour of the previous day, then only the value from the day after was used and vice versa.
+Interpolated values were not used in calculating averages.
+\item Data gaps longer than $24\,\unit{h}$ were filled using weekly diurnal averages.
+i.e.
+an average was calculated using the corresponding hour throughout the week before and after.
+Interpolated values were not used in calculating averages.
+\end{itemize}
+
+For the missing ozone data, we applied a Reynolds decomposition of the form:
+\begin{equation}
+  \chem{[O_3]} = \chem{\left<[O_3]\right>} + \chem{\Delta[O_3]},
+  \label{eq:reynolds_decomp}
+\end{equation}
+with derived ozone climatology \chem{\left<[O_3]\right>} and anomalies (deviation from the norm) \chem{\Delta[O_3]}.
+We derived a generalized climatology for northern Fennoscandia from several multi-year records and for our site and determined the station with the highest correlation coefficient to our site.
+For this site, we computed the anomalies compared to the climatology for the dates of missing ozone data at our site.
+These anomalies were than added to the climatology at out site and corrected for an increase in tropospheric background ozone since the early 1990s.
+For more details refer to as \citet{ACPD:Falk2021}.
+%%%
 
@@
 
-\section{DO3SE modeling}
-\label{sec:do3se}
-We investigate the difference between 2018 and 2019 concerning stomatal ozone uptake and assess which of the environmental factors might be most influential in determining \chem{O_3} damage risk.
-We model ozone uptake of natural and semi-natural vegetation using the $\mathrm{DO_3SE}$ model \citep{ACP:Bueker2012} and develop bespoke parameterizations for common plant species in our focus area.
-First, we give an account of the methodology for deriving bespoke parameters, after which we present and discuss the modeling results.
-The relevant components of the $\mathrm{DO_3SE}$ model are summarized in Appendix~\ref{appendix:do3se_model}.
-
-\subsection{Model parameters}
-\label{subsec:do3se_parameters}
+\section{Model parameter localization}
+\label{sec:do3se_parameters}
 
@@
 Therefore, we shift $T\mathrm{min}$ and $T\mathrm{max}$ to colder temperatures and choose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000.
-In other words, we aim to increase the overlapping area between $f_\mathrm{temp}$ and the temperature probability density function (PDF) during the GS.
+In other words, we aim to increase the overlapping area between $f_\mathrm{T}$ and the temperature probability density function (PDF) during the GS.
 The most extreme of the two acclimation types is \emph{subarctic}.
-In Fig.~\ref{fig:f_temp_grassland}a, we show the resulting $f_\mathrm{temp}$ for perennial grassland together with temperature PDFs deduced from the five months GS in two climatological periods.
+In Fig.~\ref{fig:f_temp_grassland}a, we show the resulting $f_\mathrm{T}$ for perennial grassland together with temperature PDFs deduced from the five months GS in two climatological periods.
 The periods 1990s and 2000s have been chosen to indicate that the vegetation has been subject to climate change.
@@
 
-\subsection{Model description}
-\label{subsec:model_describtion}
-The opening of stomata is governed by different species-specific and environmental factors and can be described by a multiplicative model \citep{PTRS:Jarvis1976,Emberson2000,ICP:MappingManual2017}:
-\begin{equation}
-  g^k_\text{sto} = g^k_\text{max} \cdot f^k_\text{phen} \cdot f^k_\text{light} \cdot \max{\left\{f^k_\text{min}, f^k_\mathrm{T} \cdot f^k_\mathrm{VPD} \cdot f^k_\text{SWP}\right\}}.
-  \label{eq:stomatal}
-\end{equation}
-%
-Where $g_\mathrm{max}$ is the specific-specific maximum stomatal conductance value that is then modified by seasonal and environmental factors that vary within a range $0-1$.
-They are empirically determined and account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_\mathrm{T}$), water vapor pressure deficit ($f_\mathrm{VPD}$), and soil water potential ($f_\text{SWP}$).
-All factors differ with plant functional type (PFT) denoted with $k$.
-
-The temperature adjustment function is defined as
-%
-\begin{equation}
-  f_\mathrm{T} = \frac{T_\text{air}-T_\text{min}}{T_\text{opt}-T_\text{min}} \cdot \left(\frac{T_{\text{max}}-T_\text{air}}{T_{\text{max}}-T_\text{opt}}\right)^\beta,
-  \label{eq:f_temp}
-\end{equation}
-with $\beta = \frac{T_\text{max}-T_\text{opt}}{T_\text{opt}-T_\text{min}}$ a mixed reciprocal polynomial function of temperature.
-The shape parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various PFT and $T_\mathrm{air}$ is the near-surface air temperature.
-All temperatures are defined in units of \unit{^\circ C}.
-
-The water vapor pressure deficit function is
-%
-\begin{equation}
-  f_\mathrm{VPD} = f_\text{min}+(1-f_\text{min}) \cdot \frac{D_\text{min} - \text{VPD}}{D_\text{min}-D_\text{max}}
-  \label{eq:f_vpd}
-\end{equation}
-%
-where VPD is the leaf to air vapor pressure deficit in \unit{kPa} with $f_\text{min}$, $D_\text{min}$, $D_\text{max}$ describing the relative stomatal conductance to changes in vapor pressure deficit.
-
-The wavelength band $400-700\,\unit{nm}$ plant chlorophyll responds to is called photosynthetic active radiation (PAR).
-Its integral is the photosynthetic photon flux density (PPFD).
-The relationship between relative $g_\mathrm{sto}$ and PPFD is given by
-%
-\begin{equation}
-  f_\text{light} = 1-\text{exp}(-\alpha_\text{light}\cdot \text{PPFD})
-  \label{eq:flight}
-\end{equation}
-%
-where $\alpha_\mathrm{light}$ is a slope parameter describing the extent of stomatal opening at low light intensities.
-
-The $\mathrm{DO_3SE}$ model as described in \citet{ACP:Bueker2012} is used to simulate SWP across a PFT specific root depth according to the Penman--Monteith energy balance method that drives water cycling through the soil--plant--atmosphere system.
-\citet{ACP:Bueker2012} discuss several formulations available to parameterize the reduction in stomata conductance due to water content in the soil.
-Here we use the relationship between relative $g_\mathrm{sto}$ and soil water potential (SWP) given by:
-%
-\begin{equation}
-  f_\mathrm{SWP} = \text{min}\left\{1, \text{max}\left\{f_\mathrm{min}, \frac{(1 - f_\mathrm{min})\cdot(\mathrm{SWP_{min}} - \mathrm{SWP})}{\mathrm{SWP_{min}} - \mathrm{SWP_{max}}} + f_\mathrm{min} \right\} \right\}
-    \label{eq:fsw}
-\end{equation}
-%
-where SWP is the soil water potential across the root zone and $\mathrm{SWP_{min}}$ and $\mathrm{SWP_{max}}$ are the parameters describing the $f_\mathrm{SWP}$ relationship.
-
-To compute $\mathrm{POD_y}$, we estimate the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf.
-The terms $r_c$ (leaf surface resistance) and $r_b$ (quasi-laminar resistance) allow for the deposition of \chem{O_3} to the leaf and the fraction that is taken up via the stomata.
-%
-\begin{equation}
-  \Phi_\mathrm{sto} = \chem{[O_3]} \cdot \frac{u(z_1) \cdot g_\mathrm{sto} \cdot r_c}{r_b + r_c}.
-  \label{eq:flux_stomata}
-\end{equation}
-%
-The quasi-laminar boundary layer resistance is calculated by
-%
-\begin{equation}
-  r_b = 1.3 \cdot 150 \cdot \sqrt{\frac{L}{u(z_1)}}.
-  \label{eq:quasi_laminar_resist}
-\end{equation}
-%
-Where $L$ is the cross-wind leaf dimension, $u(z_1)$ the wind speed at height $z_1$, and the factor $1.3$ accounts for the diffusivity between heat and \chem{O_3}.
-
-\subsection{Input data and gap filling methodology}
-\label{subsec:ozone_reco}
-The $\mathrm{DO_3SE}$ model requires hourly, continuous meteorological observations.
-In addition to variables covered in Section~\ref{sec:data}, $2\,\unit{m}$ wind $u_\text{2m}$ and vapor pressure deficit VPD are needed.
-VPD has been calculated from observed $T_\text{2m}$ and relative humidity
-\begin{equation}
-  \text{VPD} = P_s(T) \cdot \left(1-\frac{\text{relHum}}{100}\right),
-\end{equation}
-with saturation vapor pressure $P_s(T)$ in \unit{hPa} derived from Arden Buck equation~\citep{JAP:Buck1981, Buck2012}.
-
-The following gap filling methodology was devised for the meteorological input data:
-\begin{itemize}
-\item Single hours of missing data were filled by taking the average of the hourly values coming before, and after, the missing value.
-\item Several consecutive hours of missing data (23 or less) were filled by taking the average of the corresponding hour the day before, and the day after; and repeating this for each missing hour of data.
-If data were unavailable from that hour of the previous day, then only the value from the day after was used and vice versa.
-Interpolated values were not used in calculating averages.
-\item Data gaps longer than $24\,\unit{h}$ were filled using weekly diurnal averages.
-i.e.
-an average was calculated using the corresponding hour throughout the week before and after.
-Interpolated values were not used in calculating averages.
-\end{itemize}
-
-For the missing ozone data, we applied a Reynolds decomposition of the form:
-\begin{equation}
-  \chem{[O_3]} = \chem{\left<[O_3]\right>} + \chem{\Delta[O_3]},
-  \label{eq:reynolds_decomp}
-\end{equation}
-with climatology \chem{\left<[O_3]\right>} and anomalies \chem{\Delta[O_3]} as described in \citet{ACPD:Falk2021}.
-
 \subsection{Initial $\mathrm{DO_3SE}$ modeling results with mapping manual parameters}

### 2022-04-06 VegOzone.tex
Restructuring

@@
 To date, only a few studies assessed the response of subarctic vegetation to variations in climate and air pollution.
-This study looks to fill this knowledge gap by examining the uncertainty in risk assessments which are usinge more general boreal paramterizations.
-With no bespoke parameterization for stomatal response to essential climate variables available, we propose an adjustment of existing, Boreal parameterizations, based on local climate condition.
-To this end, we vary the temperature response function according to the climatological temperature distribution during the subarctic growing season.
+This study looks to fill this knowledge gap by examining the uncertainty in risk assessments that are using more general boreal paramterizations.
+With no bespoke parameterization of stomatal response of subarctic vegetation to essential climate variables available, we propose an adjustment of existing, Boreal parameterizations, based on local climate condition.
+To this end, we vary temperature response functions according to the climatological temperature distribution within the growing season at our site.
 We evaluate the extent to which two recent years (2018 and 2019) deviate from climatic and \chem{[O_3]} norms and how these potentially more frequent future deviations may influence the projected ozone damage to subarctic vegetation.
@@
 Annual global yield losses of four major crops (wheat, rice, maize, and soybean) of $3-15\,\unit{\%}$ have been reported \citep{PJ:Ainsworth2017} and a suggested loss in primary production in forestry of $7\,\unit{\%}$ \citep{GCB:Wittig2009,EP:Matyssek2012} has been attributed to ozone.
-Even though long-term (pre-1950) observations of ground-level ozone concentrations are scarce; evidence from trend analysis of those few sites suggests that since the industrial revolution, background concentrations in the northern hemisphere have at least doubled and continued to increase \citep{IPCC2013}.
+Even though long-term (pre-1950) observations of ground-level ozone concentrations are scarce; evidence from trend analysis of those few sites suggests that since the industrial revolution, background concentrations in the northern hemisphere have at least doubled and continued to increase \citep{IPCC2013} until recently.
 Recent trend analyses for Europe indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric background ozone started to level off or even decline \citep{ESA:Cooper2014, ACP:Wespes2018,ESA:Gaudel2018}.
@@
 In northern Fennoscandia, episodes of elevated summertime ozone concentrations in 2003 and 2006 have been traced back to ozone precursor emissions related to forest fires in southern and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}.
-photosynthetically active vegetation in the form of terpenes and monoterpenes, so-called biogenic volatile substances (BVOCs) and soil.
+Besides VOCs from anthropogenic sources, hydrocarbons are also emitted by photosynthetically active vegetation in the form of terpenes and monoterpenes, so-called biogenic volatile substances (BVOCs) and soil.
 These emissions are thought to be a response to factors, such as thermal stress, defense against herbivores \citep{Ecosystems:Ghimire2021}, or even attraction of pollinators \citep{TPS:Penuelas2003}.
@@
 At the same time, the growing season is also becoming longer.
-A longer growing season is prolonging the time in which vegetation can accumulate ozone.
+A longer growing season is prolonging the time in which vegetation can accumulate an ozone dose.
 For natural and semi-natural vegetation that is already subject to rapid climatic changes known as Arctic amplification \citep{AMAP2012,IPCC2013}, these factors could promote a higher potential risk for northern vegetation to ozone-induced damage in the not so distant future.
@@
 The higher the stomatal conductance, the higher the potential for ozone uptake.
-Stomatal conductance has been empirically linked to environmental factors such as air temperature $T$, photosynthetic photon flux density (PPFD), vapor pressure deficit (VPD), and soil water potential (SWP) as well as photosynthesis itself \cite[e.g.]{PTRS:Jarvis1976, BallBerry1987, Emberson2000, ICP:MappingManual2017}.
-Results from open-top chamber (OTC) experiments on downy birch (\emph{Betula pubescens}) and mountain birch (\emph{Betula pubescens toruosa}), native to subarctic regions, indicated reductions in both biomass, in root:shoot ratio, and visible leaf damage under elevated ozone treatment ($\left<\chem{[O_3]}\right> = 36-54\,\unit{ppb}$) \citep{Amb:Manninen2009}.
+Stomatal conductance has been empirically linked to environmental factors such as air temperature $T$, photosynthetic photon flux density (PPFD), vapor pressure deficit (VPD), and soil water potential (SWP) as well as photosynthesis itself \citep[e.g.,][]{PTRS:Jarvis1976, BallBerry1987, Emberson2000, ICP:MappingManual2017}.
+Results from open-top chamber (OTC) experiments on downy birch (\emph{Betula pubescens}) and mountain birch (\emph{Betula pubescens toruosa}), native to subarctic regions, indicated reductions in both biomass, root:shoot ratio, and visible leaf damage under elevated ozone treatment ($\left<\chem{[O_3]}\right> = 36-54\,\unit{ppb}$) \citep{Amb:Manninen2009}.
 Though, Scots pine (\emph{Pinus sylvestris}) is considered to be more ozone tolerant due to an absence of visible injuries \citep{Amb:Girgzdiene2009}, \citet{Amb:Manninen2009} found chlorophyll:carotenoid ratio and polyamines reductions under elevated ozone concentrations indicating susceptibility to ozone also in these species.\\
 
-Flux-based metrics, that relate stomatal ozone uptake to vegetation damage, are biologically more relevant for risk assessments than exposure-based metrics because they can account for particular species characteristics (i.e.
+To estimate the potential ozone damage risk, we use a flux-based metric, also referred to as the Phytotoxic Ozone Dose over a threshold y ($\mathrm{POD_y}$).
+Flux-based metrics, that relate the accumulated dose of ozone to vegetation damage, are biologically more relevant for risk assessments than exposure-based metrics because they can account for particular species characteristics (i.e.
 physiology and phenology) as well as environmental conditions that can decouple the relationship between ozone concentration, ozone uptake, and consequent damage \citep{PT:Emberson2020}.
@@
 But no bespoke parameterization for subarctic vegetation is available.
-
-To estimate the flux-based metric, also referred to as the Phytotoxic Ozone Dose over a threshold y ($\mathrm{POD_y}$), an estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf and environmental factors affecting stomatal conductance.
+An estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy $h$ represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf and environmental factors affecting stomatal conductance.
 $\mathrm{POD_y}$ is then calculated according to:
@@
 \begin{equation}
-  \mathrm{POD_y} = \int{(\Phi_\mathrm{sto}-y)\cdot \mathrm{dt}},
+  \mathrm{POD_y} = \int{(\Phi_\mathrm{sto}-y)\cdot \mathrm{d}t},
   \label{eq:pod}
@@
 %
-with the hourly averaged stomatal ozone flux $\Phi_\mathrm{sto}$ (see Appendix Eq.~(\ref{eq:flux_stomata}) and a stomatal ozone flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{Projected Leaf Area (PLA)}\,s^{-1}}.
-The flux threshold $\mathrm{y}$ represents the detoxification potential of the plant and is typically only exceeded during daylight hours (i.e.
+with the hourly averaged stomatal ozone flux $\Phi_\mathrm{sto}$ (see Eq.~(\ref{eq:flux_stomata})) and a stomatal ozone flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}^{-1}\,s^{-1}}.
+Herein PLA stands for Projected~Leaf~Area.
+The timestep $\mathrm{d}t$ is usually 1~hour.
+The flux threshold~$\mathrm{y}$ represents the detoxification potential of the plant and is typically only exceeded during daylight hours (i.e.
 when global radiation is above $50\,\unit{W\,m^{-2}}$).
@@
 The described methods for risk assessment rely on accurate representation of plant physiological responses to key environmental variables and are most often parameterized using data for less extreme climates, i.e., temperate, continental, or Boreal climates.
-Parameterization of these models for more extreme bio-geographical regions such as subarctic climates has not been performed and relies on assuming that less extreme Boreal parameterization will be applicable.
+Parameterization for more extreme bio-geographical regions such as subarctic climates has not been performed and relies on assuming that less extreme Boreal parameterization will be applicable.
 It is important to understand the implications of using these in the subarctic.
 In our study, we explore how risk might be over- or underestimated using Boreal parameterization and explore an adjustment of temperature and light response functions based on climate data.
-This allows an understanding of which are the key aspects of plant physiology that might determine the potential risk and also how the threat from ozone may change in the future.
-Our results may hint a solution to solve long-standing issues of earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface models and display over- or underproductivity in the Arctic \citep{GMD:Poulter2015,JAMES:Lawrence2019}.\\
-
-Data and methods including the DO3SE model and the localization of response function are described in Section~\ref{sec:dam}.
-The model experiments and derived parameterizations are summarized in Section 3.
-In Section 4, we present our modelling results.
-We discuss our results and give an outlook on how these could improve future assessments of pollution risk.
+This allows an understanding of which are the key aspects of modelled plant physiology that might determine the potential risk and also how the threat from ozone may change in the future.
+In addition, our results may hint a solution to solve long-standing issues of earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface models and display over- or underproductivity in the Arctic \citep{GMD:Poulter2015,JAMES:Lawrence2019}.\\
+
+Section~\ref{sec:dam}, data and methods, includes a brief summary of the major components of the used model and its key drivers.
+Further more we propose a method of localization of response functions based on climate data and start and end of the growing season from satellite observed gross primary productivity (GPP).
+In Section~\ref{sec:do3se_results}, we present our modelling results.
+We discuss our results and give an outlook on how these could improve future assessments of pollution risk in Section~\ref{sec:discussion}.
 
 %%
-\section{Data \& methods}
+\section{Methods and data}
 \label{sec:dam}
+In this section, we first describe the used model and its key components and input data and gap filling procedure (Section~\ref{sec:do3se}).
+After which, we put the conditions during the growing season 2018/19 into a climatological context (Section~\ref{sec:stats}).
+Finally, we describe our proposed method for localization of existing Boreal parameterizations to a subarctic climate (Section~\ref{sec:localization}).
 
@@
 Stomatal resistance is the inverse of stomatal conductance $g_\mathrm{sto}$ in Eq.~(\ref{eq:stomatal}).
-
-\subsubsection{Input data and gap filling methodology}
-\label{subsec:ozone_reco}
-The $\mathrm{DO_3SE}$ model requires hourly, continuous ozone and meteorological observations of $2\,\unit{m}$ temperature $T_\mathrm{2m}$, global irradiation $Q_0$, precipitation, $2\,\unit{m}$ wind $u_\text{2m}$, and vapor pressure deficit VPD.
+Accounting for these the effective stomatal flux $\Phi_\mathrm{sto}$ at the vegetation height $h$ is defined as
+%
+\begin{equation}
+  \Phi_\mathrm{sto} = \chem{[O_3]}(h) \cdot g_\mathrm{sto}(h).
+  \label{eq:flux_stomata}
+\end{equation}
+A linear gradient of \chem{[O_3]} towards the surface has been adopted in the $\mathrm{DO_3SE}$ model based on detailed studies using the chemical transport model EMEP \citep{ACP:Simpson2012,ICP:MappingManual2017}.
+
+\subsubsection{Gap filling methodology}
+\label{subsec:gap_filling}
+The $\mathrm{DO_3SE}$ model requires hourly, continuous input data. Therfore, the following gap filling methodology was devised for the meteorological observations:
+\begin{itemize}
+\item Single hours of missing data were filled by taking the average of the hourly values before, and after, the missing value.
+\item Several consecutive hours of missing data (23 or less) were filled by taking the average of the corresponding hour the day before, and the day after; and repeating this for each missing hour of data.
+If data were unavailable from that hour of the previous day, then only the value from the day after was used and vice versa.
+Interpolated values were not used in calculating averages.
+\item Data gaps longer than $24\,\unit{h}$ were filled using weekly diurnal averages.
+i.e.
+an average was calculated using the corresponding hour throughout the week before and after.
+Interpolated values were not used in calculating averages.
+\end{itemize}
+
+To fill gaps in the ozone data, we applied a methodology based on a Reynolds decomposition of the form:
+\begin{equation}
+  \chem{[O_3]} = \chem{\left<[O_3]\right>} + \chem{\Delta[O_3]},
+  \label{eq:reynolds_decomp}
+\end{equation}
+with an ozone climatology \chem{\left<[O_3]\right>} and ozone anomalies (deviation from the norm) \chem{\Delta[O_3]}.
+We computed a generalized climatology for northern Fennoscandia from long-term records of different close by sites and for our site.
+We determined the station with the highest correlation coefficient compared to Svanhovd.
+For this site, we computed the anomalies with respect to the generalized climatology for the gap period.
+These anomalies were than added to the climatology at out site and corrected for an increase in present-day tropospheric background ozone compared to the early 1990s.
+For more details refer to \citet{ACPD:Falk2021}.
+
+\subsubsection{Input data}
+\label{subsec:data}
+The required input data consists of observed ozone concentration and meteorological observations of $2\,\unit{m}$ temperature $T_\mathrm{2m}$, global irradiation $Q_0$, precipitation, $2\,\unit{m}$ wind $u_\text{2m}$, and vapor pressure deficit VPD.
 VPD has been calculated from observed $T_\text{2m}$ and relative humidity
@@
 
-For our study, we chose Svanhovd ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m\,a.s.l.}$).
-Reasons for our choice: (1) Ozone had been monitored there in the past, (2) an agrometeorological station is measuring, e.g.
+As site for our study, we chose Svanhovd ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m\,a.s.l.}$).
+Reasons for our choice: (1) Ambient ozone concentration had been monitored there in the past, (2) an agrometeorological station is measuring, e.g.
 $2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths, and (3) a national surveillance of air pollution (\chem{NO_x}, \chem{SO_x}, PM, radiological species) is ongoing.
@@
 
-All relevant data for the growing season 2018/19 are shown as time series in Fig.~\ref{fig:data_svanvik_2018_2019}.
-Ozone concentrations measured at $2\,\unit{m}$ height above ground are averaged hourly.
-The hatched areas mark times when no ozone data were recorded.
-Note, while the downtime during winter was planned, missing data in two weeks of July 2018 (July 9--23) were due to problems in data acquisition.
-
-As can be seen in Fig.~\ref{fig:data_svanvik_2018_2019}a, \chem{[O_3]} peaks in spring (April/May) and reaches its minimum in late summer (July/August).
-The spring peak has not been captured completely in 2019, since data acquisition only started in late April.
-In summer 2018 (June--August), high ozone concentrations ($\chem{[O_3]} > 40\,\unit{ppb}$) were recorded 50 times.
-The highest summer ozone concentration ($\chem{[O_3]} = 50.2\,\unit{ppb}$) was measured on July 25.
-This coincides with the period of the most extensive forest fires in central Sweden which occurred from July 12--29 \citep{SOU2019}.
-However, due to the above-mentioned data acquisition problems, we missed most of the corresponding ozone data for this event and most likely also the peak \chem{[O_3]}.
-A method for gap-filling these data has been presented in \citet{ACPD:Falk2021}.
-In contrast, \chem{[O_3]} only rose 18 times above the threshold of $40\,\unit{ppb}$ during the summer of 2019.
-
-Hourly averaged $2\,\unit{m}$ temperatures above $20\,\unit{^\circ C}$ occurred more regularly in July 2018 than in 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}b).
-In 2018, spring temperature regularly rose above freezing only in May, while in 2019 this occurred already early in March/April.
-More rain events with accumulated daily precipitation ($\sum_d \mathrm{Precip}$) above $10\,\unit{mm}$ occurred in the summer of 2018 compared to 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}c).
-Qualitatively, global irradiance ($Q_0$) displayed in Fig.~\ref{fig:data_svanvik_2018_2019}d was higher in May and July 2018 compared to 2019, while June 2019 showed higher irradiance than in 2018.
-A high global irradiance, in most cases, is the result of a low cloud fraction.
-In both years, the maximum recorded $Q_0$ was $750\,\unit{W\,m^{-2}}$ and reached in June.
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{data_svanvik_2018_2019}
-  %\includegraphics[width=12cm]{fig02}
-  \caption{Observational data from atmospheric monitoring at Svanhovd in 2018/19.
-The hatched areas indicate periods without ozone monitoring data.
-(a) Hourly averaged \chem{[O_3]}; (b) hourly averaged temperature; (c) daily accumulated precipitation; (d) hourly averaged global irradiance.}
-  \label{fig:data_svanvik_2018_2019}
-\end{figure*}
-
-The following gap filling methodology was devised for the meteorological input data:
-\begin{itemize}
-\item Single hours of missing data were filled by taking the average of the hourly values before, and after, the missing value.
-\item Several consecutive hours of missing data (23 or less) were filled by taking the average of the corresponding hour the day before, and the day after; and repeating this for each missing hour of data.
-If data were unavailable from that hour of the previous day, then only the value from the day after was used and vice versa.
-Interpolated values were not used in calculating averages.
-\item Data gaps longer than $24\,\unit{h}$ were filled using weekly diurnal averages.
-i.e.
-an average was calculated using the corresponding hour throughout the week before and after.
-Interpolated values were not used in calculating averages.
-\end{itemize}
-
-For the missing ozone data, we applied a Reynolds decomposition of the form:
-\begin{equation}
-  \chem{[O_3]} = \chem{\left<[O_3]\right>} + \chem{\Delta[O_3]},
-  \label{eq:reynolds_decomp}
-\end{equation}
-with derived ozone climatology \chem{\left<[O_3]\right>} and anomalies (deviation from the norm) \chem{\Delta[O_3]}.
-We derived a generalized climatology for northern Fennoscandia from several multi-year records and for our site and determined the station with the highest correlation coefficient to our site.
-For this site, we computed the anomalies compared to the climatology for the dates of missing ozone data at our site.
-These anomalies were than added to the climatology at out site and corrected for an increase in tropospheric background ozone since the early 1990s.
-For more details refer to as \citet{ACPD:Falk2021}.
 %%%
 
-\section{Statistical analysis}
-\label{sec:stats}
-In this section, we assess the climate conditions at Svanhovd.
-To this end, we compute a multi-annual average (referred to as climatology) for each environmental key variable (temperature, precipitation, global irradiance) and ground-level ozone.
-We evaluate the statistical significance of divergences from the norm in these variables (referred to as anomalies) in 2018/19.
-Concerning ozone, we performed a bias correction accounting for an increase in tropospheric background \chem{[O_3]} and cross-calibrated the climatology with long-term observations at other sites in northern Fennoscandia as described in \citet{ACPD:Falk2021}.
-
-\subsection{Derived climatologies}
-\label{subsec:climatologies}
-
-Svanhovd is located within the subarctic climate zone.
-Climatologies of temperature and precipitation derived for 1992--2012 are shown in a combined climate diagram (Fig.~\ref{fig:climatediagram}).
-Monthly averaged $2\,\unit{m}$ temperatures (red line) are displayed with a standard error of the mean (SE) as error band.
-We chose SE in this case due to the large interannual variability in temperature reflected in the standard deviation.
-Averaged monthly accumulated precipitation (blue bars) is shown with standard deviation as error bars.
-Months with average temperatures below freezing are denoted with a star.
-As can be seen, temperatures stay below freezing for 5 consecutive months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for K\"{o}ppen's climate classification of a regular subarctic climate (Dfc) \citep[e.g.][]{SD:Beck2018}.
-The highest monthly average temperature is $(13.1\pm 1.1)\,\unit{^\circ C}$ in July and the lowest $(-12.8\pm 2.0)\,\unit{^\circ C}$ in February.
-The coldest measured temperature was $-45.2\,\unit{^\circ C}$ in January 27, 1999, while the highest temperature ($29.4\,\unit{^\circ C}$) occurred on July 16 the same year.
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{climatediagram}
-  %\includegraphics[width=8.3cm]{fig04}
-  \caption{Climate diagram following Walter and Lieth.
-The diagram is based on climatological data for Svanvik/Pasvik (1992--2012).
-Monthly averaged temperatures (red line) are displayed with standard error of the mean as error band.
-Averaged monthly accumulated precipitation (blue bars) is shown with a standard deviation of mean as error bars.
-Months with average temperatures below freezing are denoted with a star.}
-  \label{fig:climatediagram}
-\end{figure}
-
-The average accumulated monthly precipitation ($\sum_d\mathrm{Precip}$) indicates that winter and spring (November--April, except for March) are relatively dry ($\sum_d\mathrm{Precip} < 20\,\unit{mm}$).
-March precipitation is primarily snow and will influence the start of the growing season.
-The driest month is January with $\sum_d\mathrm{Precip} = (16.7\pm 3.0)\,\unit{mm}$.
-The most precipitation occurs in the summer months, with a $\sum_d\mathrm{Precip} = (58.5\pm 9.2)\,\unit{mm}$ in August.
-The average annual accumulated precipitation ($\sum_m\mathrm{Precip}$) given with standard error of mean is $(383\pm 86)\,\unit{mm}$.
-Precipitation shows a considerable interannual variability.
-
-The climatology for global irradiance ($Q_0$, not shown) approaches zero in December/January (polar night) and reaches its maximum in June/July (midnight sun conditions).
-We deduce the corresponding highest $Q_0$ to $800\,\unit{W\,m^{-2}}$.
-The maximum of the lowest $Q_0$ (including night) amounts to $200\,\unit{W\,m^{-2}}$ indicating that a considerable amount of light is available under midnight sun conditions.
-
-%Ozone
-The daily mean ozone climatology (\chem{\left<[O_3]\right>}) (1986--1996) is shown in Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}.
-Darker colors indicate higher probability to observe these values.
-On top of the density distributions, a $10\,\unit{days}$ average of daily mean (\chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}}) is displayed together with $1 \sigma$ uncertainties and SE, respectively.
-A spline was fitted through the data to guide the eye.
-The \chem{[O_3]} density distribution and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} display the expected, pronounced seasonal cycle.
-The spring peak occurs on average on day of the year (\unit{doy}) $100$ and amounts to $40\,\unit{ppb}$ while the annual average \chem{\left<[O_3]\right>} is $28\,\unit{ppb}$.
-The decline in \chem{\left<[O_3]\right>} coincides with the average beginning of \chem{CO_2} uptake by coniferous trees \citep{TB:Kolari2007, TP:Wallin2013}.
-In July--September (\unit{doy}~182--273), ozone is occasionally almost completely depleted.
-This hints to ozone uptake by vegetation even at low light intensities during midnight sun conditions in combination with stable planetary boundary layer conditions preventing mixing of ozone rich air.
-
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_climatology_svanvik_obs_spline}
-  %\includegraphics[width=8.3cm]{fig05}
-  \caption{Svanhovd ozone climatology.
-The density distribution is shown together with a multi-annual average of daily \chem{[O_3]} (climatology).
-A spline has been fitted through daily mean \chem{[O_3]}.
-The climatology of daily mean ozone is shown together with $1\,\sigma$ standard deviation and standard error of the mean.}
-  \label{fig:ozone_climatology_fenoscandic_obs_spline}
-\end{figure}
-
-\subsection{2018/19 anomalies}
-\label{subsec:anomalies}
-In the following, we look at the 2018/19 anomalies and discuss the differences between the two years.
-As weather extremes like in 2018 are likely to become more frequent with climate change, this will help to assess possible future changes in the vulnerability of subarctic vegetation to ozone.\\
-
-In Fig.~\ref{fig:anomalies_svanvik}a, temperature anomalies for Svanvik are shown as a percentage of days warmer or colder than $\pm 1\,\sigma$ from the climatology for each month.
-Negative deviations from climatology are displayed as a negative percentage.
-The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
-We find, that the summer of 2018 was significantly warmer than average.
-Especially in May and July more than $40\,\unit{\%}$ of the days were significantly warmer.
-A significant number of warmer days continued to occur in all months (August--November) except for October.
-March 2018 had many unusually cold days.
-The summer of 2019 was fairly average, however, more significantly colder days occurred in July, while August had both significantly colder and warmer days.
-April 2019 had significantly warmer than average days.
-
-\begin{figure}[t]
-  %\includegraphics[width=12cm]{fig06}
-  \centering
-  (a)\hspace{8.3cm}(b)\\
-  \includegraphics[width=8.3cm]{temperature_signific_1.pdf}
-  \includegraphics[width=8.3cm]{precipitation_signific_1.pdf}\\
-  (c)\hspace{8.3cm}(d)\\
-  \includegraphics[width=8.3cm]{global_rad_signific_1.pdf}
-  \includegraphics[width=8.3cm]{ozone_signific_1.pdf}
-  \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as percentages of days significantly deviating from climatological mean through standard deviation for each month.
-Negative deviations from the climatology are shown as a negative percentage.
-The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The hatched area between the dashed lines indicates statistical significance under the assumption of a normal distribution ($15.9\,\unit{\%}$, $30.8\,\unit{\%}$ for Precip.).
-(a) Temperature; (b) precipitation; (c) global irradiance; (d) ozone.}
-  \label{fig:anomalies_svanvik}
-\end{figure}
-
-Similarly, the precipitation anomalies for Svanvik are displayed in Fig.~\ref{fig:anomalies_svanvik}b.
-Due to the large interannual variability in observed precipitation, only a few days deviated significantly from the climatology on the $1 \sigma$ level.
-Therefore, we lowered the $\sigma$ level constraint to $\pm \frac{1}{2}\,\sigma$.
-The percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ are shown for each month.
-Positive deviation refers to days wetter than average and negative deviation to drier than average.
-The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm\frac{1}{2}\,\sigma$ if a normal distribution is assumed ($30.8\,\unit{\%}$).
-Unlike the temperature anomalies, the picture is not as clear.
-While March 2018 had a significant number of days wetter than usual, summer and fall (May/July, September/October) had a significant number of days drier than average.
-2019 was rather average throughout the year, but with a significant number of days drier than normal in September/October and wetter in December.
-
-The global irradiance anomalies for Svanvik are presented in Fig.~\ref{fig:anomalies_svanvik}c.
-The percentage of days above or below $\pm 1\,\sigma$ from the climatological mean is displayed for each month.
-In summer 2018 (May/July), global irradiance showed a significant number of days with higher than average irradiance, while 2019 was generally an average year with a significant number of days darker in spring (April/May).
-
-%Ozone
-We use the bias-corrected and cross-calibrated ozone climatology \citep{ACPD:Falk2021} and assess the monthly significance of the ozone concentration anomalies in 2018/19.
-The resulting percentage of days above/below $\pm 1\,\sigma$ is displayed for each month in Fig.~\ref{fig:anomalies_svanvik}d.
-Note that missing bars for months in winter and spring are both due to the temporal extent of our data and anomalies not meeting the $1\sigma$ criteria.
-The star indicates the reconstructed data in July.
-In April/May 2018, ozone was significantly elevated, while no significant enhancement is found in 2019.
-In both years, ozone concentrations were significantly lower than average in November.
-These results suggest that stagnant meteorological conditions in April/May in combination with a late start of the growing season contributed more to elevated ozone than the extensive forest fires in Sweden that did not significantly enhance \chem{[O_3]} in 2018.
-
-In summary, 2018 had a significant number of days that were warmer, drier, and brighter than the climatology, while 2019 was a rather average year.
-Ozone was not significantly enhanced during the GS 2018/19, but a substantially higher number of peak \chem{[O_3]} were observed in 2018 than in 2019.
-
-\section{Model parameter localization}
-\label{sec:do3se_parameters}
-
-We deduced the dominant species on-site at Svanhovd and from Fig.~\ref{fig:svanhovd_research_station} and found perennial grassland, birch (generalized as deciduous trees), and Scots pine (generalized as coniferous trees).
+\subsection{Model parameter localization}
+\label{sec:localization}
+
+We deduced the dominant species on-site at Svanhovd from Fig.~\ref{fig:svanhovd_research_station} and found perennial grassland, birch (generalized as deciduous trees), and Scots pine (generalized as coniferous trees).
 Generic parameters for these PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as mapping manual (MM) parameterizations in the following.
@@
 
-\subsection{$\mathrm{POD_1}$ results and implications on ozone impact}
-\label{subsec:do3se_results}
+\section{Climate and weather conditions}
+\label{sec:stats}
+
+\subsection{Weather conditions in the growing season 2018/19}
+\label{subsec:weather}
+
+All relevant data for the growing season 2018/19 are shown as time series in Fig.~\ref{fig:data_svanvik_2018_2019}.
+Ozone concentrations measured at $2\,\unit{m}$ height above ground are averaged hourly.
+The hatched areas mark times when no ozone data were recorded.
+Note, while the downtime during winter was planned, missing data in two weeks of July 2018 (July 9--23) were due to problems in data acquisition.
+
+As can be seen in Fig.~\ref{fig:data_svanvik_2018_2019}a, \chem{[O_3]} peaks in spring (April/May) and reaches its minimum in late summer (July/August).
+The spring peak has not been captured completely in 2019, since data acquisition only started in late April.
+In summer 2018 (June--August), high ozone concentrations ($\chem{[O_3]} > 40\,\unit{ppb}$) were recorded 50 times.
+The highest summer ozone concentration ($\chem{[O_3]} = 50.2\,\unit{ppb}$) was measured on July 25.
+This coincides with the period of the most extensive forest fires in central Sweden which occurred from July 12--29 \citep{SOU2019}.
+However, due to the above-mentioned data acquisition problems, we missed most of the corresponding ozone data for this event and most likely also the peak \chem{[O_3]}.
+A method for gap-filling these data has been presented in \citet{ACPD:Falk2021}.
+In contrast, \chem{[O_3]} only rose 18 times above the threshold of $40\,\unit{ppb}$ during the summer of 2019.
+
+Hourly averaged $2\,\unit{m}$ temperatures above $20\,\unit{^\circ C}$ occurred more regularly in July 2018 than in 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}b).
+In 2018, spring temperature regularly rose above freezing only in May, while in 2019 this occurred already early in March/April.
+More rain events with accumulated daily precipitation ($\sum_d \mathrm{Precip}$) above $10\,\unit{mm}$ occurred in the summer of 2018 compared to 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}c).
+Qualitatively, global irradiance ($Q_0$) displayed in Fig.~\ref{fig:data_svanvik_2018_2019}d was higher in May and July 2018 compared to 2019, while June 2019 showed higher irradiance than in 2018.
+A high global irradiance, in most cases, is the result of a low cloud fraction.
+In both years, the maximum recorded $Q_0$ was $750\,\unit{W\,m^{-2}}$ and reached in June.
+
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{data_svanvik_2018_2019}
+  %\includegraphics[width=12cm]{fig02}
+  \caption{Observational data from atmospheric monitoring at Svanhovd in 2018/19.
+The hatched areas indicate periods without ozone monitoring data.
+(a) Hourly averaged \chem{[O_3]}; (b) hourly averaged temperature; (c) daily accumulated precipitation; (d) hourly averaged global irradiance.}
+  \label{fig:data_svanvik_2018_2019}
+\end{figure*}
+
+\subsection{Climate (1990-present)}
+\label{subsec:climatologies}
+
+Svanhovd is located within the subarctic climate zone.
+Climatologies of temperature and precipitation derived for 1992--2012 are shown in a combined climate diagram (Fig.~\ref{fig:climatediagram}).
+Monthly averaged $2\,\unit{m}$ temperatures (red line) are displayed with a standard error of the mean (SE) as error band.
+We chose SE in this case due to the large interannual variability in temperature reflected in the standard deviation.
+Averaged monthly accumulated precipitation (blue bars) is shown with standard deviation as error bars.
+Months with average temperatures below freezing are denoted with a star.
+As can be seen, temperatures stay below freezing for 5 consecutive months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for K\"{o}ppen's climate classification of a regular subarctic climate (Dfc) \citep[e.g.][]{SD:Beck2018}.
+The highest monthly average temperature is $(13.1\pm 1.1)\,\unit{^\circ C}$ in July and the lowest $(-12.8\pm 2.0)\,\unit{^\circ C}$ in February.
+The coldest measured temperature was $-45.2\,\unit{^\circ C}$ in January 27, 1999, while the highest temperature ($29.4\,\unit{^\circ C}$) occurred on July 16 the same year.
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{climatediagram}
+  %\includegraphics[width=8.3cm]{fig04}
+  \caption{Climate diagram following Walter and Lieth.
+The diagram is based on climatological data for Svanvik/Pasvik (1992--2012).
+Monthly averaged temperatures (red line) are displayed with standard error of the mean as error band.
+Averaged monthly accumulated precipitation (blue bars) is shown with a standard deviation of mean as error bars.
+Months with average temperatures below freezing are denoted with a star.}
+  \label{fig:climatediagram}
+\end{figure}
+
+The average accumulated monthly precipitation ($\sum_d\mathrm{Precip}$) indicates that winter and spring (November--April, except for March) are relatively dry ($\sum_d\mathrm{Precip} < 20\,\unit{mm}$).
+March precipitation is primarily snow and will influence the start of the growing season.
+The driest month is January with $\sum_d\mathrm{Precip} = (16.7\pm 3.0)\,\unit{mm}$.
+The most precipitation occurs in the summer months, with a $\sum_d\mathrm{Precip} = (58.5\pm 9.2)\,\unit{mm}$ in August.
+The average annual accumulated precipitation ($\sum_m\mathrm{Precip}$) given with standard error of mean is $(383\pm 86)\,\unit{mm}$.
+Precipitation shows a considerable interannual variability.
+
+The climatology for global irradiance ($Q_0$, not shown) approaches zero in December/January (polar night) and reaches its maximum in June/July (midnight sun conditions).
+We deduce the corresponding highest $Q_0$ to $800\,\unit{W\,m^{-2}}$.
+The maximum of the lowest $Q_0$ (including night) amounts to $200\,\unit{W\,m^{-2}}$ indicating that a considerable amount of light is available under midnight sun conditions.
+
+%Ozone
+The daily mean ozone climatology (\chem{\left<[O_3]\right>}) (1986--1996) is shown in Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}.
+Darker colors indicate higher probability to observe these values.
+On top of the density distributions, a $10\,\unit{days}$ average of daily mean (\chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}}) is displayed together with $1 \sigma$ uncertainties and SE, respectively.
+A spline was fitted through the data to guide the eye.
+The \chem{[O_3]} density distribution and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} display the expected, pronounced seasonal cycle.
+The spring peak occurs on average on day of the year (\unit{doy}) $100$ and amounts to $40\,\unit{ppb}$ while the annual average \chem{\left<[O_3]\right>} is $28\,\unit{ppb}$.
+The decline in \chem{\left<[O_3]\right>} coincides with the average beginning of \chem{CO_2} uptake by coniferous trees \citep{TB:Kolari2007, TP:Wallin2013}.
+In July--September (\unit{doy}~182--273), ozone is occasionally almost completely depleted.
+This hints to ozone uptake by vegetation even at low light intensities during midnight sun conditions in combination with stable planetary boundary layer conditions preventing mixing of ozone rich air.
+
+\begin{figure}[t]
+  \includegraphics[width=8.3cm]{ozone_climatology_svanvik_obs_spline}
+  %\includegraphics[width=8.3cm]{fig05}
+  \caption{Svanhovd ozone climatology.
+The density distribution is shown together with a multi-annual average of daily \chem{[O_3]} (climatology).
+A spline has been fitted through daily mean \chem{[O_3]}.
+The climatology of daily mean ozone is shown together with $1\,\sigma$ standard deviation and standard error of the mean.}
+  \label{fig:ozone_climatology_fenoscandic_obs_spline}
+\end{figure}
+
+\subsection{Derivation from climate norm in the 2018/19 growing season}
+\label{subsec:anomalies}
+In the following, we look at the 2018/19 anomalies and discuss the differences between the two years.
+As weather extremes like in 2018 are likely to become more frequent with climate change, this will help to assess possible future changes in the vulnerability of subarctic vegetation to ozone.\\
+
+In Fig.~\ref{fig:anomalies_svanvik}a, temperature anomalies for Svanvik are shown as a percentage of days warmer or colder than $\pm 1\,\sigma$ from the climatology for each month.
+Negative deviations from climatology are displayed as a negative percentage.
+The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
+The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
+We find, that the summer of 2018 was significantly warmer than average.
+Especially in May and July more than $40\,\unit{\%}$ of the days were significantly warmer.
+A significant number of warmer days continued to occur in all months (August--November) except for October.
+March 2018 had many unusually cold days.
+The summer of 2019 was fairly average, however, more significantly colder days occurred in July, while August had both significantly colder and warmer days.
+April 2019 had significantly warmer than average days.
+
+\begin{figure}[t]
+  %\includegraphics[width=12cm]{fig06}
+  \centering
+  (a)\hspace{8.3cm}(b)\\
+  \includegraphics[width=8.3cm]{temperature_signific_1.pdf}
+  \includegraphics[width=8.3cm]{precipitation_signific_1.pdf}\\
+  (c)\hspace{8.3cm}(d)\\
+  \includegraphics[width=8.3cm]{global_rad_signific_1.pdf}
+  \includegraphics[width=8.3cm]{ozone_signific_1.pdf}
+  \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as percentages of days significantly deviating from climatological mean through standard deviation for each month.
+Negative deviations from the climatology are shown as a negative percentage.
+The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
+The hatched area between the dashed lines indicates statistical significance under the assumption of a normal distribution ($15.9\,\unit{\%}$, $30.8\,\unit{\%}$ for Precip.).
+(a) Temperature; (b) precipitation; (c) global irradiance; (d) ozone.}
+  \label{fig:anomalies_svanvik}
+\end{figure}
+
+Similarly, the precipitation anomalies for Svanvik are displayed in Fig.~\ref{fig:anomalies_svanvik}b.
+Due to the large interannual variability in observed precipitation, only a few days deviated significantly from the climatology on the $1 \sigma$ level.
+Therefore, we lowered the $\sigma$ level constraint to $\pm \frac{1}{2}\,\sigma$.
+The percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ are shown for each month.
+Positive deviation refers to days wetter than average and negative deviation to drier than average.
+The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm\frac{1}{2}\,\sigma$ if a normal distribution is assumed ($30.8\,\unit{\%}$).
+Unlike the temperature anomalies, the picture is not as clear.
+While March 2018 had a significant number of days wetter than usual, summer and fall (May/July, September/October) had a significant number of days drier than average.
+2019 was rather average throughout the year, but with a significant number of days drier than normal in September/October and wetter in December.
+
+The global irradiance anomalies for Svanvik are presented in Fig.~\ref{fig:anomalies_svanvik}c.
+The percentage of days above or below $\pm 1\,\sigma$ from the climatological mean is displayed for each month.
+In summer 2018 (May/July), global irradiance showed a significant number of days with higher than average irradiance, while 2019 was generally an average year with a significant number of days darker in spring (April/May).
+
+%Ozone
+We use the bias-corrected and cross-calibrated ozone climatology \citep{ACPD:Falk2021} and assess the monthly significance of the ozone concentration anomalies in 2018/19.
+The resulting percentage of days above/below $\pm 1\,\sigma$ is displayed for each month in Fig.~\ref{fig:anomalies_svanvik}d.
+Note that missing bars for months in winter and spring are both due to the temporal extent of our data and anomalies not meeting the $1\sigma$ criteria.
+The star indicates the reconstructed data in July.
+In April/May 2018, ozone was significantly elevated, while no significant enhancement is found in 2019.
+In both years, ozone concentrations were significantly lower than average in November.
+These results suggest that stagnant meteorological conditions in April/May in combination with a late start of the growing season contributed more to elevated ozone than the extensive forest fires in Sweden that did not significantly enhance \chem{[O_3]} in 2018.
+
+In summary, 2018 had a significant number of days that were warmer, drier, and brighter than the climatology, while 2019 was a rather average year.
+Ozone was not significantly enhanced during the GS 2018/19, but a substantially higher number of peak \chem{[O_3]} were observed in 2018 than in 2019.
+
+\section{Modeling results and implications on ozone risk assessments}
+\label{sec:do3se_results}
 

### 2022-04-07 VegOzone.tex
Change figure

@@
 % Archives Animal Breeding (aab)
-% ASTRA Proceedings (ap)
 % Atmospheric Chemistry and Physics (acp)
@@
 % E&G Quaternary Science Journal (egqsj)
+% EGUsphere (egusphere) | This is only for EGUsphere preprints submitted without relation to an EGU journal.
 % European Journal of Mineralogy (ejm)
@@
 % Proceedings of the International Association of Hydrological Sciences (piahs)
+% Safety of Nuclear Waste Disposal (sand)
 % Scientific Drilling (sd)
@@
 %\usepackage{rotating}
-\usepackage[nomarkers,figuresonly]{endfloat}
+
+%\usepackage[nomarkers,figuresonly]{endfloat}
 \graphicspath{{pictures/}}
@@
 
-We deduced the dominant species on-site at Svanhovd from Fig.~\ref{fig:svanhovd_research_station} and found perennial grassland, birch (generalized as deciduous trees), and Scots pine (generalized as coniferous trees).
+We deduced the dominant species on-site at Svanhovd from Fig.~\ref{fig:svanhovd_research_station}a and found perennial grassland, birch (generalized as deciduous trees), and Scots pine (generalized as coniferous trees).
 Generic parameters for these PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as mapping manual (MM) parameterizations in the following.
@@
 
-Svanhovd is located within the subarctic climate zone.
-Climatologies of temperature and precipitation derived for 1992--2012 are shown in a combined climate diagram (Fig.~\ref{fig:climatediagram}).
-Monthly averaged $2\,\unit{m}$ temperatures (red line) are displayed with a standard error of the mean (SE) as error band.
-We chose SE in this case due to the large interannual variability in temperature reflected in the standard deviation.
-Averaged monthly accumulated precipitation (blue bars) is shown with standard deviation as error bars.
+Climatologies of temperature and precipitation for Svanhovd were derived for the period 1992--2012 and are shown in a combined climate diagram (Fig.~\ref{fig:climatediagram}).
+Monthly averaged $2\,\unit{m}$ temperatures (red line) are displayed with a standard deviation as error band.
+Averaged monthly accumulated precipitation (blue bars) is shown without standard deviation.
 Months with average temperatures below freezing are denoted with a star.
@@
 \begin{figure}[t]
-  \includegraphics[width=8.3cm]{climatediagram}
+  \includegraphics[width=12cm, clip, trim=1cm 0 1cm 1cm]{climatediagram_box}
   %\includegraphics[width=8.3cm]{fig04}
-  \caption{Climate diagram following Walter and Lieth.
-The diagram is based on climatological data for Svanvik/Pasvik (1992--2012).
-Monthly averaged temperatures (red line) are displayed with standard error of the mean as error band.
-Averaged monthly accumulated precipitation (blue bars) is shown with a standard deviation of mean as error bars.
-Months with average temperatures below freezing are denoted with a star.}
+  \caption{Climate diagram. The diagram is based on climatological data for Svanvik/Pasvik (1992--2012). (a) Monthly temperatures. (b) Monthly accumulated precipitation.}
   \label{fig:climatediagram}

### 2022-04-11 VegOzone.tex
Update plots

@@
 We evaluate the extent to which two recent years (2018 and 2019) deviate from climatic and \chem{[O_3]} norms and how these potentially more frequent future deviations may influence the projected ozone damage to subarctic vegetation.
-We find that 2018 was an anomalously warm and bright year, particularly in spring and early summer.
+We find that 2018 was an anomalously warm and less cloudy year, particularly in spring and early summer.
 Higher than average \chem{[O_3]} was observed in April/May while frequent episodes of ozone volume mixing ratios (VMRs) above $40\,\unit{ppb}$ occurred in June--August.
@@
 These anomalies were than added to the climatology at out site and corrected for an increase in present-day tropospheric background ozone compared to the early 1990s.
-For more details refer to \citet{ACPD:Falk2021}.
+For more details refer to \citet{ACP:Falk2021}.
 
@@
 However, due to the above-mentioned data acquisition problems, we missed most of the corresponding ozone data for this event and most likely also the peak \chem{[O_3]}.
-A method for gap-filling these data has been presented in \citet{ACPD:Falk2021}.
+A method for gap-filling these data has been presented in \citet{ACP:Falk2021}.
 In contrast, \chem{[O_3]} only rose 18 times above the threshold of $40\,\unit{ppb}$ during the summer of 2019.
@@
   \includegraphics[width=12cm]{data_svanvik_2018_2019}
-  %\includegraphics[width=12cm]{fig02}
+  %\includegraphics[width=12cm]{fig02}.
   \caption{Observational data from atmospheric monitoring at Svanhovd in 2018/19.
 The hatched areas indicate periods without ozone monitoring data.
-(a) Hourly averaged \chem{[O_3]}; (b) hourly averaged temperature; (c) daily accumulated precipitation; (d) hourly averaged global irradiance.}
+(a) Hourly averaged ozone VMR; (b) hourly averaged temperature; (c) daily accumulated precipitation; (d) hourly averaged global irradiance.}
   \label{fig:data_svanvik_2018_2019}
@@
 
-\begin{figure}[t]
-  \includegraphics[width=12cm, clip, trim=1cm 0 1cm 1cm]{climatediagram_box}
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{climatediagram_box}
   %\includegraphics[width=8.3cm]{fig04}
-  \caption{Climate diagram. The diagram is based on climatological data for Svanvik/Pasvik (1992--2012). (a) Monthly temperatures. (b) Monthly accumulated precipitation.}
+  \caption{Climate diagram based on meteorological data for Svanvik/Pasvik (1992--2012).
+Ozone climatology based on ozone monitoring data at NIBIO Environment Centre Svanhovd(1986--2003).
+A box indicates the upper/lower quartile, whiskers the $1.5\times$ Interquartile Range (IQR), and circles outliers.
+The median is marked by a colored horizontal line and the mean by a triangle.
+Notches indicate the confidence interval around the median.
+(a) Monthly temperatures; (b) Monthly accumulated precipitation; (c) Monthly global radiation; (d) Monthly ozone VMR}
   \label{fig:climatediagram}
-\end{figure}
+\end{figure*}
 
@@
 %Ozone
-The daily mean ozone climatology (\chem{\left<[O_3]\right>}) (1986--1996) is shown in Figure~\ref{fig:ozone_climatology_fenoscandic_obs_spline}.
+The daily mean ozone climatology (\chem{\left<[O_3]\right>}) (1986--1996) is shown in Figure~\ref{fig:climatediagram}d.
 Darker colors indicate higher probability to observe these values.
@@
 
-\begin{figure}[t]
-  \includegraphics[width=8.3cm]{ozone_climatology_svanvik_obs_spline}
-  %\includegraphics[width=8.3cm]{fig05}
-  \caption{Svanhovd ozone climatology.
-The density distribution is shown together with a multi-annual average of daily \chem{[O_3]} (climatology).
-A spline has been fitted through daily mean \chem{[O_3]}.
-The climatology of daily mean ozone is shown together with $1\,\sigma$ standard deviation and standard error of the mean.}
-  \label{fig:ozone_climatology_fenoscandic_obs_spline}
-\end{figure}
-
-\subsection{Derivation from climate norm in the 2018/19 growing season}
+\subsection{2018/19 growing season anomalies}
 \label{subsec:anomalies}
@@
 
-\begin{figure}[t]
+\begin{figure*}[t]
   %\includegraphics[width=12cm]{fig06}
   \centering
-  (a)\hspace{8.3cm}(b)\\
-  \includegraphics[width=8.3cm]{temperature_signific_1.pdf}
-  \includegraphics[width=8.3cm]{precipitation_signific_1.pdf}\\
-  (c)\hspace{8.3cm}(d)\\
-  \includegraphics[width=8.3cm]{global_rad_signific_1.pdf}
-  \includegraphics[width=8.3cm]{ozone_signific_1.pdf}
+  (a)\hspace{6cm}(b)\\
+  \includegraphics[width=6cm]{temperature_signific_1.pdf}
+  \includegraphics[width=6cm]{precipitation_signific_1.pdf}\\
+  (c)\hspace{6cm}(d)\\
+  \includegraphics[width=6cm]{global_rad_signific_1.pdf}
+  \includegraphics[width=6cm]{ozone_signific_1.pdf}
   \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as percentages of days significantly deviating from climatological mean through standard deviation for each month.
@@
   \label{fig:anomalies_svanvik}
-\end{figure}
+\end{figure*}
 
@@
 %Ozone
-We use the bias-corrected and cross-calibrated ozone climatology \citep{ACPD:Falk2021} and assess the monthly significance of the ozone concentration anomalies in 2018/19.
+We use the bias-corrected and cross-calibrated ozone climatology \citep{ACP:Falk2021} and assess the monthly significance of the ozone concentration anomalies in 2018/19.
 The resulting percentage of days above/below $\pm 1\,\sigma$ is displayed for each month in Fig.~\ref{fig:anomalies_svanvik}d.
@@
 
-\section{Modeling results and implications on ozone risk assessments}
+\section{Results}
+\label{sec:results}
+
+\subsection{Parameter localization}
+\label{subsec:res_local}
+
+\section{Implications on ozone risk assessments}
 \label{sec:do3se_results}

### 2022-04-13 VegOzone.tex
Rearange and rwrite last paragraph in intro

@@
 
-Section~\ref{sec:dam}, data and methods, includes a brief summary of the major components of the used model and its key drivers.
-Further more we propose a method of localization of response functions based on climate data and start and end of the growing season from satellite observed gross primary productivity (GPP).
-In Section~\ref{sec:do3se_results}, we present our modelling results.
-We discuss our results and give an outlook on how these could improve future assessments of pollution risk in Section~\ref{sec:discussion}.
+In Section~\ref{sec:dam}, we give a brief summary of the key components and  drivers of the $\mathrm{DO_3SE}$ model that is used to assess the ozone risk.
+To study the influence of selected parameters on ozone risk in the subarctic, we propose a method to localize temperature response functions based on climate data and start and end of the growing season from satellite observed gross primary productivity (GPP).
+For the later discussion of results, we summarize climate and weather conditions of the 2018/19 growing season at our research site in Section~\ref{sec:stats}.
+Results of the localization and ozone risk modelling are presented in Section~\ref{sec:results}.
+In Section~\ref{sec:conc}, we discuss these results concerning prevailing conditions and give an outlook on how future assessments of pollution risk in the subarctic could be improved.
 
 %%
-\section{Methods and data}
+\section{Methods}
 \label{sec:dam}
-In this section, we first describe the used model and its key components and input data and gap filling procedure (Section~\ref{sec:do3se}).
-After which, we put the conditions during the growing season 2018/19 into a climatological context (Section~\ref{sec:stats}).
-Finally, we describe our proposed method for localization of existing Boreal parameterizations to a subarctic climate (Section~\ref{sec:localization}).
+In the following, we briefly describe the model used for assessing the ozone risk, its key components, input data, and gap filling procedure (Section~\ref{sec:do3se}).
+To study the impact of model parameters, we propose a method for parameter localization to a subarctic climate based on existing Boreal parameterizations and estimate the start and end of the growing season from satellite derived GPP (Section~\ref{sec:localization}).
 
@@
 \label{sec:stats}
+Here, we put the conditions during the growing season 2018/19 into a climatological context.
 
@@
 Notches indicate the confidence interval around the median.
-(a) Monthly temperatures; (b) Monthly accumulated precipitation; (c) Monthly global radiation; (d) Monthly ozone VMR}
+(a) Temperature; (b) accumulated precipitation; (c) radiant energy; (d) ozone VMR}
   \label{fig:climatediagram}
@@
 \begin{figure*}[t]
-  %\includegraphics[width=12cm]{fig06}
+
   \centering
-  (a)\hspace{6cm}(b)\\
-  \includegraphics[width=6cm]{temperature_signific_1.pdf}
-  \includegraphics[width=6cm]{precipitation_signific_1.pdf}\\
-  (c)\hspace{6cm}(d)\\
-  \includegraphics[width=6cm]{global_rad_signific_1.pdf}
-  \includegraphics[width=6cm]{ozone_signific_1.pdf}
-  \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as percentages of days significantly deviating from climatological mean through standard deviation for each month.
-Negative deviations from the climatology are shown as a negative percentage.
-The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The hatched area between the dashed lines indicates statistical significance under the assumption of a normal distribution ($15.9\,\unit{\%}$, $30.8\,\unit{\%}$ for Precip.).
-(a) Temperature; (b) precipitation; (c) global irradiance; (d) ozone.}
+  \includegraphics[width=12cm]{anomalies}
+  \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as difference from median for each month.
+The hatched area between indicates the IQR.
+(a) Temperature; (b) precipitation; (c) radiant energy; (d) ozone VMR.}
   \label{fig:anomalies_svanvik}
@@
 
-\section{Implications on ozone risk assessments}
-\label{sec:do3se_results}
+\subsection{Implications on ozone risk assessments}
+\label{subsec:do3se_results}
 

### 2022-04-13 VegOzone.tex
Include info from soilgrids.org

@@
 
-To estimate the potential ozone damage risk, we use a flux-based metric, also referred to as the Phytotoxic Ozone Dose over a threshold y ($\mathrm{POD_y}$).
+To estimate the potential ozone risk, we use a flux-based metric, also referred to as the Phytotoxic Ozone Dose over a threshold y ($\mathrm{POD_y}$).
 Flux-based metrics, that relate the accumulated dose of ozone to vegetation damage, are biologically more relevant for risk assessments than exposure-based metrics because they can account for particular species characteristics (i.e.
@@
 Soil texture influences water availability and thus $\mathrm{POD_y}$ calculations in $\mathrm{DO_3SE}$ \citep{ACP:Bueker2012}.
-The soil at Svanhovd is characterized in an ICP Forest plot survey as gleyic cambisols according to the Food and Agriculture Organization of the United Nations (FAO) soil classification system and as eluviated dystric brunisol in the ICP Forest soil classification system (V.
-Timmerman, personal communication, June 2020) Within the $\mathrm{DO_3SE}$ soil parameterization, the sandy loam texture describes the upper $60\,\unit{cm}$, where the plant roots are found, of this soil type best.
+The soil at Svanhovd is characterized in an ICP~Forest plot survey as gleyic cambisols according to the Food and Agriculture Organization of the United Nations (FAO) soil classification system and as eluviated dystric brunisol in the ICP~Forest soil classification system (V.
+Timmerman, personal communication, June 2020).
+The clay:sand:silt content is estimated to range from 12\,\%:61\,\%:27\,\% at the top layer to 16\,\%:59\,\%:25\,\% in about $1\,\unit{m}$ depth (\href{https://soilgrids.org/}{soilgrids.org}, last accessed April 2022 for location $30.0334\,\unit{^\circ E}$, $69.4549\,\unit{^\circ N}$) Within the $\mathrm{DO_3SE}$ soil parameterization, the sandy loam texture describes the upper $60\,\unit{cm}$, where the plant roots are found, of this soil type best.
 
@@
 Perennial grassland shows the smallest ozone uptake, but a similarly large response to the temperature parameterizations as deciduous trees.
-Again, we find a reversal for predicted ozone damage risk in $\mathrm{POD_1}$ between 2018 and 2019 for the subarctic type.
+Again, we find a reversal for predicted ozone risk in $\mathrm{POD_1}$ between 2018 and 2019 for the subarctic type.
 Perennial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant without bespoke GS.

### 2022-04-13 VegOzone.tex
Harmonize se of key environmental variables

@@
 These episodes are in part attributable to forest fires in the Northern Hemisphere and warmer and sunnier conditions.
-We apply the integrated flux-metric Phytotoxic Ozone Dose (POD) to determine ozone risk and damage to vegetation as a function of \chem{[O_3]}, environmental factors, and species-specific physiology.
+We apply the integrated flux-metric Phytotoxic Ozone Dose (POD) to determine ozone risk and damage to vegetation as a function of \chem{[O_3]}, key environmental variables, and species-specific physiology.
 Our study suggests that using standard boreal parameterizations in assessments likely leads to underestimating the risk of ozone damage in this region.
@@
 The higher the stomatal conductance, the higher the potential for ozone uptake.
-Stomatal conductance has been empirically linked to environmental factors such as air temperature $T$, photosynthetic photon flux density (PPFD), vapor pressure deficit (VPD), and soil water potential (SWP) as well as photosynthesis itself \citep[e.g.,][]{PTRS:Jarvis1976, BallBerry1987, Emberson2000, ICP:MappingManual2017}.
+Stomatal conductance has been empirically linked to key environmental variables such as air temperature $T$, photosynthetic photon flux density (PPFD), vapor pressure deficit (VPD), and soil water potential (SWP) as well as photosynthesis itself \citep[e.g.,][]{PTRS:Jarvis1976, BallBerry1987, Emberson2000, ICP:MappingManual2017}.
 Results from open-top chamber (OTC) experiments on downy birch (\emph{Betula pubescens}) and mountain birch (\emph{Betula pubescens toruosa}), native to subarctic regions, indicated reductions in both biomass, root:shoot ratio, and visible leaf damage under elevated ozone treatment ($\left<\chem{[O_3]}\right> = 36-54\,\unit{ppb}$) \citep{Amb:Manninen2009}.
@@
 Flux-based metrics, that relate the accumulated dose of ozone to vegetation damage, are biologically more relevant for risk assessments than exposure-based metrics because they can account for particular species characteristics (i.e.
-physiology and phenology) as well as environmental conditions that can decouple the relationship between ozone concentration, ozone uptake, and consequent damage \citep{PT:Emberson2020}.
+physiology and phenology) as well as key environmental variables that can decouple the relationship between ozone concentration, ozone uptake, and consequent damage \citep{PT:Emberson2020}.
 These flux-based metrics are better suited to represent the actual risk to vegetation from ozone, especially in the climatically extreme parts of Europe \citep{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017}.
 But no bespoke parameterization for subarctic vegetation is available.
-An estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy $h$ represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf and environmental factors affecting stomatal conductance.
+An estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy $h$ represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf and key environmental variables affecting stomatal conductance.
 $\mathrm{POD_y}$ is then calculated according to:
@@
 \label{subsec:model_describtion}
-The opening of stomata is governed by different species-specific and environmental factors.
+The opening of stomata is governed by different species-specific factors and response to key environmental variables.
 In $\mathrm{DO_3SE}$ it is described by a multiplicative empirical model \citep{PTRS:Jarvis1976,Emberson2000,ICP:MappingManual2017}:
@@
 %
-Where $g_\mathrm{max}$ is a species-specific maximum stomatal conductance value that is then modified by seasonal and environmental factors that vary within a range $0-1$.
+Where $g_\mathrm{max}$ is a species-specific maximum stomatal conductance value that is then modified by seasonal and response functions to key environmental variables that vary within a range $0-1$.
 These factors are empirically determined and account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_\mathrm{T}$), water vapor pressure deficit ($f_\mathrm{VPD}$), and soil water potential ($f_\text{SWP}$).
@@
 To compare our bespoke parameterizations, we propose the following metric.
-We calculate an average relative stomatal conductance $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal}) with $f_\mathrm{SWP}=1$) at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time) for the whole record of key environmental factors' data and apply mean and standard deviation.
+We calculate an average relative stomatal conductance $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal}) with $f_\mathrm{SWP}=1$) at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time) for the whole record of key environmental variables' data and apply mean and standard deviation.
 We presume that $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ around noon (highest light intensity) is a good proxy for \chem{CO_2} uptake efficiency (growth).
@@
 In contrast, no such damage was found on the clovers in 2019.
-This indicates that environmental factors in 2018 differed from 2019.
+This indicates that key environmental variables in 2018 differed from 2019.
 

### 2022-04-13 VegOzone.tex
Resturcture and edit Section 2

@@
 Results from open-top chamber (OTC) experiments on downy birch (\emph{Betula pubescens}) and mountain birch (\emph{Betula pubescens toruosa}), native to subarctic regions, indicated reductions in both biomass, root:shoot ratio, and visible leaf damage under elevated ozone treatment ($\left<\chem{[O_3]}\right> = 36-54\,\unit{ppb}$) \citep{Amb:Manninen2009}.
-Though, Scots pine (\emph{Pinus sylvestris}) is considered to be more ozone tolerant due to an absence of visible injuries \citep{Amb:Girgzdiene2009}, \citet{Amb:Manninen2009} found chlorophyll:carotenoid ratio and polyamines reductions under elevated ozone concentrations indicating susceptibility to ozone also in these species.\\
+Though, Scots pine (\emph{Pinus sylvestris}) is considered to be more ozone tolerant due to an absence of visible injuries \citep{Amb:Girgzdiene2009}, \citet{Amb:Manninen2009} found chlorophyll:carotenoid ratio and polyamines reductions under elevated ozone concentrations indicating susceptibility to ozone also in these species.
 
@@
 In the following, we briefly describe the model used for assessing the ozone risk, its key components, input data, and gap filling procedure (Section~\ref{sec:do3se}).
-To study the impact of model parameters, we propose a method for parameter localization to a subarctic climate based on existing Boreal parameterizations and estimate the start and end of the growing season from satellite derived GPP (Section~\ref{sec:localization}).
+To study the impact of model parameters, we propose a method for parameter localization to a subarctic climate based on existing Boreal parameterizations and local climate conditions and estimate the start and end of the growing season from satellite derived GPP (Section~\ref{sec:localization}).
 
@@
 \label{sec:do3se}
-To investigate the \chem{O_3} damage risk in the subarctic, we model the ozone dose of natural (coniferous and deciduous trees) and semi-natural (grassland) vegetation using the Deposition of Ozone for Stomatal Exchange ($\mathrm{DO_3SE}$) model \citep{ACP:Bueker2012} which computes estimates of the total and stomatal deposition of \chem{O_3} to selected land-cover types and plant species according to UNECE LRTAP (Long-Range Transboundary Air Pollution) methodologies for effects-based risk assessment.
+In our investigation of the ozone risk in the subarctic, we use the well established Deposition of Ozone for Stomatal Exchange ($\mathrm{DO_3SE}$) model \citep{ACP:Bueker2012} which computes estimates of the total and stomatal deposition of \chem{O_3} to selected land-cover types and plant species according to UNECE Long-Range Transboundary Air Pollution (LRTAP) methodologies for effects-based risk assessment.
 
@@
 \label{subsec:model_describtion}
-The opening of stomata is governed by different species-specific factors and response to key environmental variables.
-In $\mathrm{DO_3SE}$ it is described by a multiplicative empirical model \citep{PTRS:Jarvis1976,Emberson2000,ICP:MappingManual2017}:
-\begin{equation}
-  g^k_\text{sto} = g^k_\text{max} \cdot f^k_\text{phen} \cdot f^k_\text{light} \cdot \max{\left\{f^k_\text{min}, f^k_\mathrm{T} \cdot f^k_\mathrm{VPD} \cdot f^k_\text{SWP}\right\}}.
+
+In the $\mathrm{DO_3SE}$ model, the stomatal flux $\Phi_\mathrm{sto}$ in Eq.~(\ref{eq:pod}) is a fraction of the total \chem{O_3} deposition flux.
+A standard resistance analogous approach \citep{AE:Wesely1989} is used to transport \chem{O_3} from the atmospheric reference height $h_0$ to the vegetated surface.
+The calculation of aerodynamic $R_a$ and quasi-laminar boundary layer resistance $R_b$ follow \citet{WASP:Simpson2003,ACP:Simpson2012}.
+The surface resistance $R_\mathrm{surf}$ consists of stomatal $r_\mathrm{sto}$ and non-stomatal canopy resistances that include external plant surfaces $r_\mathrm{ext}$, aerodynamic within-canopy $R_\mathrm{inc}$ and ground surfaces or soil resistances $R_\mathrm{gs}$ \citep{WASP:Simpson2003,ACP:Simpson2012}.
+Under the assumption that the ozone concentration \chem{[O_3]} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf, $r_\mathrm{sto}$ and $r_\mathrm{ext}$ are calculated per leaf/needle area (note lower case $r$).
+%
+\begin{equation}
+  R_\mathrm{sur} = \frac{1}{\frac{\mathrm{LAI}}{r_\mathrm{sto}}+\frac{\mathrm{SAI}}{r_\mathrm{ext}}+\frac{1}{R_\mathrm{inc}+R_\mathrm{gs}}}
+\end{equation}
+%
+Stomatal resistance $r_\mathrm{sto}$ is the inverse of stomatal conductance $g_\mathrm{sto}$ in Eq.~(\ref{eq:stomatal}).
+Accounting for these the effective stomatal flux $\Phi_\mathrm{sto}$ at the vegetation height $h$ is defined as
+%
+\begin{equation}
+  \Phi_\mathrm{sto} = \chem{[O_3]}(h) \cdot g_\mathrm{sto}(h).
+  \label{eq:flux_stomata}
+\end{equation}
+A linear gradient reducing \chem{[O_3]} from $h_0$ to $h$ was adopted in the $\mathrm{DO_3SE}$ model based on detailed transport studies using the chemical transport model EMEP \citep{ACP:Simpson2012,ICP:MappingManual2017}.
+
+The opening of stomata or stomatal conductance $g_\mathrm{sto}$ is governed by different species-specific factors and response to key environmental variables.
+In the $\mathrm{DO_3SE}$ model, it is described by a multiplicative empirical model \citep{PTRS:Jarvis1976,Emberson2000,ICP:MappingManual2017}:
+\begin{equation}
+  g_\text{sto} = g_\text{max} \cdot f_\text{phen} \cdot f_\text{light} \cdot \max{\left\{f_\text{min}, f_\mathrm{T} \cdot f_\mathrm{VPD} \cdot f_\text{SWP}\right\}}.
   \label{eq:stomatal}
@@
 %
-Where $g_\mathrm{max}$ is a species-specific maximum stomatal conductance value that is then modified by seasonal and response functions to key environmental variables that vary within a range $0-1$.
+Where $g_\mathrm{max}$ is a species-specific maximum stomatal conductance value that is modified by factors that vary within a range of $0-1$.
 These factors are empirically determined and account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_\mathrm{T}$), water vapor pressure deficit ($f_\mathrm{VPD}$), and soil water potential ($f_\text{SWP}$).
-All factors differ with plant functional type (PFT) denoted with $k$.
-
-The phenology ($f_\mathrm{phen}$) is described by a stepwise defined linear function that allows for a increase of stomatal conductance in spring (start of growing season) and a decrease in fall (end of growing season).
-Furthermore, mid-season dips can also be accounted for.
-
-The temperature adjustment function is defined as
+All factors differ with plant functional type (PFT).
+
+The leaf phenology factor ($f_\mathrm{phen}$) is a stepwise defined linear function that allows for an increase of stomatal conductance in spring (start of growing season) and a decrease in fall (end of growing season).
+Mid anthesis and mid season dips in stomata conductance can also be accounted for but are not relevant in the subarctic.
+
+The temperature response function is defined as
 %
@@
 %
-where VPD is the leaf to air vapor pressure deficit in \unit{kPa} with $f_\text{min}$, $D_\text{min}$, $D_\text{max}$ describing the relative stomatal conductance to changes in vapor pressure deficit.
-
-The wavelength band ($400-700\,\unit{nm}$) plant chlorophyll responds to is called photosynthetic active radiation (PAR).
-Its integral is the photosynthetic photon flux density (PPFD).
-The relationship between relative $g_\mathrm{sto}$ and PPFD is given by
+where VPD is the leaf to air vapor pressure deficit in \unit{kPa} with a minimum conductance $f_\text{min}$.
+$D_\text{min}$, $D_\text{max}$ describe the relative stomatal conductance to changes in vapor pressure deficit.
+
+The light response function $f_\mathrm{light}$ is given by
 %
@@
 where $\alpha_\mathrm{light}$ is a slope parameter describing the extent of stomatal opening at low light intensities.
-
-The $\mathrm{DO_3SE}$ model as described in \citet{ACP:Bueker2012} is used to simulate SWP across a PFT specific root depth according to the Penman--Monteith energy balance method that drives water cycling through the soil--plant--atmosphere system.
+The wavelength band $400-700\,\unit{nm}$ to which plant chlorophyll responds to is called photosynthetic active radiation (PAR).
+Its integral is the photosynthetic photon flux density (PPFD).
+
+In the $\mathrm{DO_3SE}$ model, Soil Water Potential (SWP) across a PFT specific root depth is simulated according to the Penman--Monteith energy balance method that drives water cycling through the soil--plant--atmosphere system \citep{ACP:Bueker2012}.
 The stomatal response to changes in SWP ($f_\mathrm{SWP}$) are, thus, given by:
@@
 %
-where SWP is the soil water potential across the root zone and $\mathrm{SWP_{min}}$ and $\mathrm{SWP_{max}}$ are the parameters describing the $f_\mathrm{SWP}$ relationship.
-
-$\Phi_\mathrm{sto}$ in Eq.~(\ref{eq:pod}) as computed by the $\mathrm{DO_3SE}$ model is a fraction of the total \chem{O_3} deposition flux.
-A standard resistance analogous approach \citep{AE:Wesely1989} is used to transport \chem{O_3} from the atmospheric reference height to the vegetated surface.
-The calculation of aerodynamic $R_a$ and quasi-laminar boundary layer resistance $R_b$ follow \citet{WASP:Simpson2003,ACP:Simpson2012}.
-The surface resistance $R_\mathrm{surf}$ consists of stomatal $r_\mathrm{sto}$ and non-stomatal canopy resistances that includes external plant surfaces $r_\mathrm{ext}$, aerodynamic within-canopy $R_\mathrm{inc}$ and ground surfaces or soil resistances $R_\mathrm{gs}$ \citep{WASP:Simpson2003,ACP:Simpson2012}.
-Under the assumption that \chem{[O_3]} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf, $r_\mathrm{sto}$ and $r_\mathrm{ext}$ are calculated per leaf/needle area (note lower case $r$).
-%
-\begin{equation}
-  R_\mathrm{sur} = \frac{1}{\frac{\mathrm{LAI}}{r_\mathrm{sto}}+\frac{\mathrm{SAI}}{r_\mathrm{ext}}+\frac{1}{R_\mathrm{inc}+R_\mathrm{gs}}}
-\end{equation}
-%
-Stomatal resistance is the inverse of stomatal conductance $g_\mathrm{sto}$ in Eq.~(\ref{eq:stomatal}).
-Accounting for these the effective stomatal flux $\Phi_\mathrm{sto}$ at the vegetation height $h$ is defined as
-%
-\begin{equation}
-  \Phi_\mathrm{sto} = \chem{[O_3]}(h) \cdot g_\mathrm{sto}(h).
-  \label{eq:flux_stomata}
-\end{equation}
-A linear gradient of \chem{[O_3]} towards the surface has been adopted in the $\mathrm{DO_3SE}$ model based on detailed studies using the chemical transport model EMEP \citep{ACP:Simpson2012,ICP:MappingManual2017}.
+where $\mathrm{SWP_{min}}$ and $\mathrm{SWP_{max}}$ are the parameters describing the $f_\mathrm{SWP}$ relationship.
 
@@
 \label{subsec:gap_filling}
-The $\mathrm{DO_3SE}$ model requires hourly, continuous input data. Therfore, the following gap filling methodology was devised for the meteorological observations:
+The $\mathrm{DO_3SE}$ model requires hourly, continuous input data. Therefor, the following gap filling methodology was devised for the meteorological drivers of the model:
 \begin{itemize}
@@
 
-To fill gaps in the ozone data, we applied a methodology based on a Reynolds decomposition of the form:
+To fill gaps in the ozone data, we applied the methodology described in \citet{ACP:Falk2021} that is based on a Reynolds decomposition of the form:
 \begin{equation}
@@
 \end{equation}
-with an ozone climatology \chem{\left<[O_3]\right>} and ozone anomalies (deviation from the norm) \chem{\Delta[O_3]}.
+with an ozone climatology (multiannual norm) \chem{\left<[O_3]\right>} and ozone anomalies (deviation from the norm) \chem{\Delta[O_3]}.
 We computed a generalized climatology for northern Fennoscandia from long-term records of different close by sites and for our site.
-We determined the station with the highest correlation coefficient compared to Svanhovd.
-For this site, we computed the anomalies with respect to the generalized climatology for the gap period.
-These anomalies were than added to the climatology at out site and corrected for an increase in present-day tropospheric background ozone compared to the early 1990s.
+We determined the station with the highest correlation coefficient compared to our site.
+For this site, we calculated the anomalies with respect to the generalized climatology for the gap period.
+These anomalies were than added to the climatology of our site and corrected for an increase in present-day tropospheric background ozone compared to the early 1990s.
 For more details refer to \citet{ACP:Falk2021}.
 
-\subsubsection{Input data}
+\subsubsection{Input data and observation site}
 \label{subsec:data}
-The required input data consists of observed ozone concentration and meteorological observations of $2\,\unit{m}$ temperature $T_\mathrm{2m}$, global irradiation $Q_0$, precipitation, $2\,\unit{m}$ wind $u_\text{2m}$, and vapor pressure deficit VPD.
-VPD has been calculated from observed $T_\text{2m}$ and relative humidity
-\begin{equation}
-  \text{VPD} = P_s(T) \cdot \left(1-\frac{\text{relHum}}{100}\right),
-\end{equation}
-with saturation vapor pressure $P_s(T)$ in \unit{hPa} derived from Arden Buck equation~\citep{JAP:Buck1981, Buck2012}.
-
-As site for our study, we chose Svanhovd ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m\,a.s.l.}$).
+The required input data consists of ozone concentration $\chem{[O_3]}$, $2\,\unit{m}$ temperature $T_\mathrm{2m}$, global irradiation $Q_0$, precipitation, $2\,\unit{m}$ wind $u_\mathrm{2m}$, and vapor pressure deficit VPD.
+VPD has been calculated from observed $T_\mathrm{2m}$ and relative humidity
+\begin{equation}
+  \mathrm{VPD} = P_s(T) \cdot \left(1-\frac{\mathrm{relHum}}{100}\right),
+\end{equation}
+with saturation vapor pressure $P_s(T)$ in \unit{hPa} from Arden~Buck~equation~\citep{JAP:Buck1981, Buck2012}.
+
+As site for our study, we chose Svanhovd in northern Norway ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m\,a.s.l.}$).
 Reasons for our choice: (1) Ambient ozone concentration had been monitored there in the past, (2) an agrometeorological station is measuring, e.g.
-$2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths, and (3) a national surveillance of air pollution (\chem{NO_x}, \chem{SO_x}, PM, radiological species) is ongoing.
+$2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths, and (3) a national surveillance of air pollution (\chem{NO_x}, \chem{SO_x}, particular matter (PM), radiological species) is ongoing.
 Most of these measurements are performed by the Norwegian Institute for Air Research (NILU), who also conducted ozone measurements on our behalf during the 2018/19 growing season.
-In addition, we collected atmospheric monitoring and ozone monitoring data taken at Svanhovd over the past $35\,\unit{years}$.
-
-All measured essential climate variables for the 2018/19 growing season were provided by NILU through the web portal \href{luftkvalitet.no}{luftkvalitet.no} (last accessed May 2020).
+
+All essential climate variables for the 2018/19 growing season were provided by NILU through the web portal \href{luftkvalitet.no}{luftkvalitet.no} (last accessed May 2020).
 The ozone monitoring data have been added to a database for air-quality monitoring called EBAS as well operated by NILU and are thus openly accessible.
@@
 Note that data from before the 1990s do not follow the high-quality standard procedures implemented nowadays and have to be treated with care \citep{NILU2003}.
-Because ozone monitors did not undergo regular re-calibration drifts in the observed data occurred imposing false trends.
-This does not concern the general ozone seasonality which we are particularly interested in.
+Because ozone monitors did not undergo regular re-calibration drifts in the observed data occurred imposing false trends which are small but in the same order of magnitude as the expected trends in tropospheric background ozone.
+This does not concern the ozone seasonality which we are particularly interested in.
 The long-term data sets of agrometeorological variables including temperature and precipitation are available from September 1992 to the present day \citep[LandbruksMeteorologiske Tjeneste][note the station name here is Pasvik]{LMT_NIBIO}.
 
-To relate local ozone concentration \chem{[O_3]} with visible damage on ozone sensitive plants, both an ozone bio-monitor (referred to as ozone garden) and a conventional ozone monitor using the existing atmospheric monitoring infrastructure had been installed during the 2018/19 growing season at Svanhovd.
-We report \chem{[O_3]} as volume mixing ratios (VMR) in \unit{ppb} for easier comparison with exposure-based risk assessment.
+To relate local ozone concentration \chem{[O_3]} with visible damage on ozone sensitive plants, an ozone bio-monitor (referred to as ozone garden) has been installed at Svanhovd in the 2018/19 GS.
 The locations of the atmospheric monitoring site and the ozone garden are marked in the aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}a.
@@
 %%%
-
 \subsection{Model parameter localization}
@@
 We deduced the dominant species on-site at Svanhovd from Fig.~\ref{fig:svanhovd_research_station}a and found perennial grassland, birch (generalized as deciduous trees), and Scots pine (generalized as coniferous trees).
-Generic parameters for these PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as mapping manual (MM) parameterizations in the following.
-Our parameterization of coniferous trees is based on the MM's Boreal Norway spruce, deciduous trees on the MM's silver birch, and perennial grassland on MM's perennial grassland for central Europe.
+Basic parameters for these PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as mapping manual (MM) parameterizations in the following.
+For coniferous trees, we base our localization of parameters on MM's Boreal Norway spruce.
+Our deciduous tree paramterization is based on MM's silver birch and perennial grassland on MM's perennial grassland for Central Europe.
 For a comprehensive list of generic MM model parameters, consult Supplement~Table~S1.
 
-Initial test simulations with the MM PFTs revealed an unrealistically low stomatal conductance at leaf-level ($G_\mathrm{sto}^\mathrm{leaf}$) in 2019 particularly for perennial grassland (Appendix~Fig.~\ref{fig:pody_mm_composit}).
+\subsubsection{Temperature acclimation}
+
+Initial test simulations with the MM PFTs revealed an unrealistically low stomatal conductance at leaf-level ($G_\mathrm{sto}^\mathrm{leaf}$) in 2019 for perennial grassland (Appendix~Fig.~\ref{fig:pody_mm_composit}).
 Substantial stomatal conductance in this species occurred only during an extended warm period in late July.
 In ecological terms, this means that there was almost no growth of grass in the summer of 2019 -- a prediction that is easily falsified by reality.
-We identified $f_\mathrm{T}$ as abnormally low, being the limiting factor of stomatal conductance in perennial grassland.
-Therefore, we saw the necessity to propose bespoke parameterizations describing an acclimation of the generic PFTs to the climatic conditions in the focus area.
-Note, however, that these parameterizations are hypothetical and have yet to be verified by experiments.
+We identified $f_\mathrm{T}$ as abnormally low throughout the season, being the limiting factor of stomatal conductance in perennial grassland.
+Note, these parameterizations are hypothetical and have yet to be verified by experiments.
 
@@
 Because the MM version of the $\mathrm{DO_3SE}$ model currently does not simulate net photosynthesis ($A_\mathrm{n}$), we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune the temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ to acclimate the growth potential to the local climate.
-For details regarding these functions see Appendix~\ref{appendix:do3se_model}.
+For details regarding these functions see Section~\ref{subsec:model_describtion}.
 Further, we presume that $f_\mathrm{VPD}$ and $f_\mathrm{SWP}$ suit our vegetation types and no acclimation is necessary for these.
+But acknowledge that these are very important factors as shown by~\citet{ACP:Bueker2012}.
 
@@
 We construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to MM.
-This is accomplished by moving $T_\mathrm{opt}$ towards cooler temperatures while keeping the other parameters fixed to MM values.
+This is accomplished by moving $f_\mathrm{T}$ towards cooler temperatures but keep $T_\mathrm{max}$ at its MM value.
 In the same way, \emph{subarctic} is constructed to represent a species that is very tolerant to cold but sensitive to high temperatures, and most efficient at cool temperatures.
-Therefore, we shift $T\mathrm{min}$ and $T\mathrm{max}$ to colder temperatures and choose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000.
+Therefore, we shift $f\mathrm{T}$ to even colder temperatures and choose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000.
 In other words, we aim to increase the overlapping area between $f_\mathrm{T}$ and the temperature probability density function (PDF) during the GS.
@@
 In Fig.~\ref{fig:f_temp_grassland}a, we show the resulting $f_\mathrm{T}$ for perennial grassland together with temperature PDFs deduced from the five months GS in two climatological periods.
-The periods 1990s and 2000s have been chosen to indicate that the vegetation has been subject to climate change.
+The periods 1990s and 2000s have been chosen to indicate that the vegetation has been subject to climate change already.
 Thus, the predominant acclimation at present is not known.
@@
 We presume a similar temperature acclimation for deciduous trees.
-All bespoke temperature parameters are listed in Table~\ref{tab:sensitivity_tests_temp}.\\
-
+All bespoke temperature parameters are listed in Table~\ref{tab:sensitivity_tests_temp}.
+
+\subsubsection{Light acclimation}
+TODO:
 Regarding the acclimation of $f_\mathrm{light}$, we presume that the opening extent of stomata at low light conditions differs in subarctic species compared to species in less extreme climates.
@@
 
+\subsubsection{Growing season estimate}
+
 Another important factor is the timing and length of the GS.

### 2022-04-19 VegOzone.tex
Restructure Methodes and results

@@
 For the later discussion of results, we summarize climate and weather conditions of the 2018/19 growing season at our research site in Section~\ref{sec:stats}.
-Results of the localization and ozone risk modelling are presented in Section~\ref{sec:results}.
+Results of the localization, the ozone risk modelling, and the climate perspective are presented in Section~\ref{sec:results}.
 In Section~\ref{sec:conc}, we discuss these results concerning prevailing conditions and give an outlook on how future assessments of pollution risk in the subarctic could be improved.
@@
 In contrast, no such visible damage was found on the clovers in 2019, although it was observed on the sensitive tobacco cultivar (Bel-W3).
+Note that the extent of visible damage is not equivalent to a quantitative reduction of above/below ground biomass of a species.
 
@@
 
-We deduced the dominant species on-site at Svanhovd from Fig.~\ref{fig:svanhovd_research_station}a and found perennial grassland, birch (generalized as deciduous trees), and Scots pine (generalized as coniferous trees).
-Basic parameters for these PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as mapping manual (MM) parameterizations in the following.
-For coniferous trees, we base our localization of parameters on MM's Boreal Norway spruce.
-Our deciduous tree paramterization is based on MM's silver birch and perennial grassland on MM's perennial grassland for Central Europe.
-For a comprehensive list of generic MM model parameters, consult Supplement~Table~S1.
-
-\subsubsection{Temperature acclimation}
-
-Initial test simulations with the MM PFTs revealed an unrealistically low stomatal conductance at leaf-level ($G_\mathrm{sto}^\mathrm{leaf}$) in 2019 for perennial grassland (Appendix~Fig.~\ref{fig:pody_mm_composit}).
-Substantial stomatal conductance in this species occurred only during an extended warm period in late July.
-In ecological terms, this means that there was almost no growth of grass in the summer of 2019 -- a prediction that is easily falsified by reality.
+We deduced the dominant species on-site at Svanhovd and from Fig.~\ref{fig:svanhovd_research_station}a and found perennial grassland, birch (generalized as deciduous trees), and Scots pine (generalized as coniferous trees).
+Basic parameters for these PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as mapping manual (MM) parameterizations.
+The localization of parameters for coniferous trees is based on MM's Boreal Norway spruce, deciduous trees on MM's silver birch, and perennial grassland on MM's perennial grassland for Central Europe.
+For a comprehensive list of MM parameters, consult Supplement~Table~S1.
+In the following, we suggest localizations of key environmental parameters to a subarctic climate based on meteorological data to assess over-/underestimation of ozone damage by using MM Boreal parameterizations.
+
+\subsubsection{Temperature and light response}
+
+Initial simulations with the MM parameterizations for perennial grassland revealed an unexpected low stomatal conductance at leaf-level ($G_\mathrm{sto}^\mathrm{leaf}$) in 2019 (Appendix~Fig.~\ref{fig:pody_mm_composit}).
+Substantial stomatal conductance occurred only during an extended warm period in late July.
+In ecological terms, this means that there would have been almost no growth of grass in the summer of 2019 -- a model prediction that is falsified by reality.
 We identified $f_\mathrm{T}$ as abnormally low throughout the season, being the limiting factor of stomatal conductance in perennial grassland.
-Note, these parameterizations are hypothetical and have yet to be verified by experiments.
-
-We presume that perennial vegetation will likely maximize carbon acquisition during the short subarctic growing season.
-Because the MM version of the $\mathrm{DO_3SE}$ model currently does not simulate net photosynthesis ($A_\mathrm{n}$), we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune the temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ to acclimate the growth potential to the local climate.
-For details regarding these functions see Section~\ref{subsec:model_describtion}.
-Further, we presume that $f_\mathrm{VPD}$ and $f_\mathrm{SWP}$ suit our vegetation types and no acclimation is necessary for these.
-But acknowledge that these are very important factors as shown by~\citet{ACP:Bueker2012}.
-
-We define two different hypothetical temperature acclimations: subarctic and cold.
-We construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to MM.
-This is accomplished by moving $f_\mathrm{T}$ towards cooler temperatures but keep $T_\mathrm{max}$ at its MM value.
+
+To localize the temperature response function $f_\mathrm{T}$, we divide the temperature dataset into two subsets: 1992--2000 (near past) and 2011--2019 (present).
+From these two subsets, we select dates within the GS (May--September) only and plot the probability density function (PDF) as histogram with $1\,\unit{^\circ C}$ binning.
+We overlay the histograms with $f_\mathrm{T}$ to graphically determine the goodness of fit of the MM parameterization from the overlapping area.
+In Fig.~\ref{fig:f_temp_grassland}a, we show the $f_\mathrm{T}$ for perennial grassland together with the temperature PDFs for near past and present climate conditions.
+The MM parameterization is displayed as a solid line.
+A comparison between the two PDFs indicates that the temperature distribution has been shifted towards higher temperatures reducing the number of days below $5\,\unit{^\circ C}$.
+It is unclear how fast the species composition of perennial grassland responds to these changes.
+Thus, the predominant acclimation at present is not known.
+
+The idea is to define a localized temperature response function by increasing the overlapping area between $f_\mathrm{T}$ and the temperature PDF during the GS.
+To this end, we define two different hypothetical target species: subarctic and cold.
+We construct \emph{cold} as representative for a species that is more tolerant to cold temperatures than MM, but slightly less efficient at warm temperatures than MM.
+This is accomplished by moving $T_\mathrm{min}$ and $T_\mathrm{opt}$ towards cooler temperatures but keep $T_\mathrm{max}$ at its MM value.
 In the same way, \emph{subarctic} is constructed to represent a species that is very tolerant to cold but sensitive to high temperatures, and most efficient at cool temperatures.
-Therefore, we shift $f\mathrm{T}$ to even colder temperatures and choose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000.
-In other words, we aim to increase the overlapping area between $f_\mathrm{T}$ and the temperature probability density function (PDF) during the GS.
-The most extreme of the two acclimation types is \emph{subarctic}.
-In Fig.~\ref{fig:f_temp_grassland}a, we show the resulting $f_\mathrm{T}$ for perennial grassland together with temperature PDFs deduced from the five months GS in two climatological periods.
-The periods 1990s and 2000s have been chosen to indicate that the vegetation has been subject to climate change already.
-Thus, the predominant acclimation at present is not known.
-
-The resulting response functions for coniferous and deciduous trees are displayed in Appendix~Figs.~\ref{fig:f_temp_spruce}a and \ref{fig:f_temp_birch}a, respectively.
-We base our temperature acclimation of coniferous trees on experimental results on Norway spruce which were found to be active already at rather low air temperatures and can reach $60\,\unit{\%}$ photosynthetic activity as early as \unit{doy}~100 \citep{TB:Kolari2007, TP:Wallin2013}.
-In particular, we use the time series of \chem{CO_2} uptake and temperature at observation sites in southern ($\rightarrow$ \emph{cold}) and northern ($\rightarrow$ \emph{subarctic}) Finland by \citet{TB:Kolari2007} to estimate the optimal temperature interval ($10\,\unit{^\circ C}\le T_\mathrm{opt} \le 15\,\unit{^\circ C}$).
-We presume a similar temperature acclimation for deciduous trees.
-All bespoke temperature parameters are listed in Table~\ref{tab:sensitivity_tests_temp}.
-
-\subsubsection{Light acclimation}
-TODO:
-Regarding the acclimation of $f_\mathrm{light}$, we presume that the opening extent of stomata at low light conditions differs in subarctic species compared to species in less extreme climates.
-We, therefore, adjust the light sensitivity of stomata such that the PPFD value needed to reach $50\,\unit{\%}$ opening is varied $\pm 20\,\unit{\%}$.
-For this we derive the inverse function $f_{\mathrm{light},k}^{-1}$ of Eq.~(\ref{eq:flight}) analytically for each PFT or species $k$
+Therefore, we shift both $T\mathrm{min}$ and $T_\mathrm{max}$ to colder temperatures and choose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000.
+The most extreme of the two localizations compared to MM is \emph{subarctic}.
+
+For localizing $f_\mathrm{light}$, we presume that the opening extent of stomata at low light conditions differs in subarctic species compared to species in less extreme climates.
+We, therefore, adjust the modelled light sensitivity of stomata such that the PPFD value needed to reach $50\,\unit{\%}$ opening is varied by $\pm 20\,\unit{\%}$.
+For this, we derive the inverse function $f_{\mathrm{light},k}^{-1}$ of Eq.~(\ref{eq:flight}) for each PFT $k$ analytically
+%
 \begin{equation}
@@
 \end{equation}
-First, we solve Eq.~(\ref{eq:inverse_function}) for the MM default value $\alpha_\mathrm{MM}^k$ at $0.5$ 
+%
+First, we solve Eq.~(\ref{eq:inverse_function}) for the MM default value of $\alpha_\mathrm{MM}^k$ at $0.5$ ($50\,\%$ opening)
 \begin{equation}
@@
 \end{equation}
-and define a variation $\gamma' = \gamma \cdot \eta$ with $\eta \in \{0.8, 1.2\}$.
-By inverting Eq.~(\ref{eq:inverse_function}) for $\alpha$, we find $\alpha'(\gamma')$.
-The resulting functions for perennial grassland (PPFD0.8, PPFD1.2) are shown in Fig.~\ref{fig:f_temp_grassland}b) (for deciduous and coniferous trees refer Appendix~Figs.~\ref{fig:f_temp_spruce}b and \ref{fig:f_temp_birch}b.
-All derived parameters are tabulated in Table~\ref{tab:sensitivity_tests_light}.
+and define a variation $\gamma' = \gamma \cdot \eta$ with $\eta \in \{0.8, 1.2\}$. By solving Eq.~(\ref{eq:inverse_function}) for $\alpha$, we find $\alpha'(\gamma')$.
+
+In Fig.~\ref{fig:f_temp_grassland}b, we show the PPFD PDF for near past and present together with the $f_\mathrm{light}$ for perennial grassland. The MM parameterization is displayed as solid line.
 
@@
 \end{figure}
+
+Similarly, we defined localized $f_\mathrm{T}$ and $f_\mathrm{light}$ for coniferous and deciduous trees.
+The resulting response functions are shown in Appendix~Figs.~\ref{fig:f_temp_spruce} and \ref{fig:f_temp_birch}, respectively.
+
+Coniferous trees are able to start photosynthesis as soon as the leaf temperature reaches a certain threshold \citep{TB:Kolari2007} and are, hence, found to be active already at rather low air temperatures and can reach $60\,\unit{\%}$ photosynthetic activity as early as \unit{doy}~100 but cease activity in less favorable conditions \citep{TB:Kolari2007, TP:Wallin2013}.
+
+To determine the temperature localization for coniferous trees, we use the time series of \chem{CO_2} uptake and temperature at observation sites in southern ($\rightarrow$ \emph{cold}) and northern ($\rightarrow$ \emph{subarctic}) Finland by \citet{TB:Kolari2007}.
+We estimate an optimal temperature interval ($10\,\unit{^\circ C}\le T_\mathrm{opt} \le 15\,\unit{^\circ C}$).
+For simplicity, we assume the same interval for deciduous trees.
+
+The resulting parameterizations are presented in detail in Section~\ref{subsec:res_local}.
+
+\subsubsection{Growing season estimate}
+\label{subsec:gs_est}
+
+Another important factor for the accumulation of ozone dose is the timing and length of the GS.
+
+TODO: estimated as follows.
+For coniferous trees, we use a MODIS (Aqua/Terra) retrieved net photosynthesis product \citep{MODIS_PSN} for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine $A_\text{start}$ and $A_\text{end}$.
+The net photosynthesis vs.
+time of year data can be approximated by a second-order polynomial function (Fig.~\ref{fig:modis_Psn}).
+By calculating the roots of the fitted polynomial, we derive $A_\text{start}$ as \unit{doy}~122 for 2018 and \unit{doy}~106 in 2019.
+$A_\text{end}$ amounts to \unit{doy}~261 and 274, respectively.
+This value will be used for all PFTs alike.
+The resulting growing season for coniferous trees in 2019 is, hence, one month longer than in 2018.
+For estimating $A_\text{start}$ of deciduous trees, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Appendix~Fig.~\ref{fig:greening_season_change_Svanvik}).
+We find \unit{doy}~129 and 130, respectively.
+For both years, these dates coincide with the first snow-free day at the closest inland meteorological weather station at Øvre~Neiden (Sør-Varanger, NOR).
+For comparison, the generic dates for central Europe ($A_\mathrm{start} = 100$ and $A_\mathrm{end} = 307$) indicating a substantially longer growing season.
+Due to a lack of quantitative field observation at Svanhovd, we assume a latency period of $1\,\unit{month}$ after snow-melt for perennial grassland.
+This assumption is supported by observations in Rovaniemi, Finland \citep[][Supplement~Fig.~S1]{FCR:Korhonen2018}.
+
+\subsubsection{Soil conditions}
+\label{subsec:soil}
+
+Soil texture influences water availability and thus $\mathrm{POD_y}$ calculations in $\mathrm{DO_3SE}$ \citep{ACP:Bueker2012}.
+The soil at Svanhovd is characterized in an ICP~Forest plot survey as gleyic cambisols according to the Food and Agriculture Organization of the United Nations (FAO) soil classification system and as eluviated dystric brunisol in the ICP~Forest soil classification system (V.
+Timmerman, personal communication, June 2020).
+The clay:sand:silt content is estimated to range from 12\,\%:61\,\%:27\,\% at the top layer to 16\,\%:59\,\%:25\,\% in about $1\,\unit{m}$ depth (\href{https://soilgrids.org/}{soilgrids.org}, last accessed April 2022 for location $30.0334\,\unit{^\circ E}$, $69.4549\,\unit{^\circ N}$) Within the $\mathrm{DO_3SE}$ soil parameterization, the sandy loam texture describes the upper $60\,\unit{cm}$, where the plant roots are found, of this soil type best.
+
+The estimate of leaf boundary layer resistance is a function of wind speed which will be influenced by the tree height and leaf dimensions.
+A sample of downy birch leaves collected at Svanhovd had an average length of $(3.0\pm 0.5)\,\unit{cm}$ that is smaller than in the MM parameterization.
+\citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area.
+One Scots pine forest at Svanhovd has been studied as a part of the ICP Forest mapping.
+Heights were measured in 2004 when the stand was 90 years old.
+Mean tree height was $10.1\,\unit{m}$ and maximum height was $16.2\,\unit{m}$ (V.
+Timmerman, personal communication, June 2020).
+We used $13.5\,\unit{m}$ height for both deciduous and coniferous trees.
+
+\section{Results}
+\label{sec:results}
+
+\subsection{Parameter localization}
+\label{subsec:res_local}
+
+All bespoke temperature parameters are listed in Table~\ref{tab:sensitivity_tests_temp}.
+
+(for deciduous and coniferous trees refer Appendix~Figs.~\ref{fig:f_temp_spruce}b and \ref{fig:f_temp_birch}b)
+
+All derived parameters are tabulated in Table~\ref{tab:sensitivity_tests_light}.
 
@@
 
+We presume that perennial vegetation will likely maximize carbon acquisition during the short subarctic growing season.
+Because the MM version of the $\mathrm{DO_3SE}$ model currently does not simulate net photosynthesis ($A_\mathrm{n}$), we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune the temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ to acclimate the growth potential to the local climate.
+For details regarding these functions see Section~\ref{subsec:model_describtion}.
+Further, we presume that $f_\mathrm{VPD}$ and $f_\mathrm{SWP}$ suit our vegetation types and no acclimation is necessary for these.
+But acknowledge that these are very important factors as shown by~\citet{ACP:Bueker2012}.
+
 As indicated above, a better acclimation to prevailing climate conditions is characterized by a higher average $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$.
@@
   \caption{Proposed metric to test bespoke response functions.
-GS (May--August) climatological averages and standard deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon (averaged over $11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning (averaged over $5-9\,\unit{am}$ local time) are shown for (a) deciduous trees; (b) coniferous trees; (c) perennial grassland.
-}
+GS (May--August) climatological averages and standard deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon (averaged over $11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning (averaged over $5-9\,\unit{am}$ local time) are shown for (a) deciduous trees; (b) coniferous trees; (c) perennial grassland.}
   \label{fig:javis_func_opt_mean}
@@
 
-\subsubsection{Growing season estimate}
-
-Another important factor is the timing and length of the GS.
-Start ($A_\text{start}$) and end ($A_\text{end}$) of GS for each PFT are given in \unit{doy} (Table~\ref{tab:sensitivity_tests_gs}) and estimated as follows.
-For coniferous trees, we use a MODIS (Aqua/Terra) retrieved net photosynthesis product \citep{MODIS_PSN} for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine $A_\text{start}$ and $A_\text{end}$.
-The net photosynthesis vs.
-time of year data can be approximated by a second-order polynomial function (Fig.~\ref{fig:modis_Psn}).
-By calculating the roots of the fitted polynomial, we derive $A_\text{start}$ as \unit{doy}~122 for 2018 and \unit{doy}~106 in 2019.
-$A_\text{end}$ amounts to \unit{doy}~261 and 274, respectively.
-This value will be used for all PFTs alike.
-The resulting growing season for coniferous trees in 2019 is, hence, one month longer than in 2018.
-For estimating $A_\text{start}$ of deciduous trees, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Appendix~Fig.~\ref{fig:greening_season_change_Svanvik}).
-We find \unit{doy}~129 and 130, respectively.
-For both years, these dates coincide with the first snow-free day at the closest inland meteorological weather station at Øvre~Neiden (Sør-Varanger, NOR).
-For comparison, the generic dates for central Europe ($A_\mathrm{start} = 100$ and $A_\mathrm{end} = 307$) indicating a substantially longer growing season.
-Due to a lack of quantitative field observation at Svanhovd, we assume a latency period of $1\,\unit{month}$ after snow-melt for perennial grassland.
-This assumption is supported by observations in Rovaniemi, Finland \citep[][Supplement~Fig.~S1]{FCR:Korhonen2018}.
+Start ($A_\text{start}$) and end ($A_\text{end}$) of the GS for each PFT are usually given in \unit{doy} (Table~\ref{tab:sensitivity_tests_gs})
 
@@
 
-Soil texture influences water availability and thus $\mathrm{POD_y}$ calculations in $\mathrm{DO_3SE}$ \citep{ACP:Bueker2012}.
-The soil at Svanhovd is characterized in an ICP~Forest plot survey as gleyic cambisols according to the Food and Agriculture Organization of the United Nations (FAO) soil classification system and as eluviated dystric brunisol in the ICP~Forest soil classification system (V.
-Timmerman, personal communication, June 2020).
-The clay:sand:silt content is estimated to range from 12\,\%:61\,\%:27\,\% at the top layer to 16\,\%:59\,\%:25\,\% in about $1\,\unit{m}$ depth (\href{https://soilgrids.org/}{soilgrids.org}, last accessed April 2022 for location $30.0334\,\unit{^\circ E}$, $69.4549\,\unit{^\circ N}$) Within the $\mathrm{DO_3SE}$ soil parameterization, the sandy loam texture describes the upper $60\,\unit{cm}$, where the plant roots are found, of this soil type best.
-
-The estimate of leaf boundary layer resistance is a function of wind speed which will be influenced by the tree height and leaf dimensions.
-A sample of downy birch leaves collected at Svanhovd had an average length of $(3.0\pm 0.5)\,\unit{cm}$ that is smaller than in the MM parameterization.
-\citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area.
-One Scots pine forest at Svanhovd has been studied as a part of the ICP Forest mapping.
-Heights were measured in 2004 when the stand was 90 years old.
-Mean tree height was $10.1\,\unit{m}$ and maximum height was $16.2\,\unit{m}$ (V.
-Timmerman, personal communication, June 2020).
-We used $13.5\,\unit{m}$ height for both deciduous and coniferous trees.
-
-\section{Climate and weather conditions}
-\label{sec:stats}
-Here, we put the conditions during the growing season 2018/19 into a climatological context.
-
-\subsection{Weather conditions in the growing season 2018/19}
-\label{subsec:weather}
-
-All relevant data for the growing season 2018/19 are shown as time series in Fig.~\ref{fig:data_svanvik_2018_2019}.
-Ozone concentrations measured at $2\,\unit{m}$ height above ground are averaged hourly.
-The hatched areas mark times when no ozone data were recorded.
-Note, while the downtime during winter was planned, missing data in two weeks of July 2018 (July 9--23) were due to problems in data acquisition.
-
-As can be seen in Fig.~\ref{fig:data_svanvik_2018_2019}a, \chem{[O_3]} peaks in spring (April/May) and reaches its minimum in late summer (July/August).
-The spring peak has not been captured completely in 2019, since data acquisition only started in late April.
-In summer 2018 (June--August), high ozone concentrations ($\chem{[O_3]} > 40\,\unit{ppb}$) were recorded 50 times.
-The highest summer ozone concentration ($\chem{[O_3]} = 50.2\,\unit{ppb}$) was measured on July 25.
-This coincides with the period of the most extensive forest fires in central Sweden which occurred from July 12--29 \citep{SOU2019}.
-However, due to the above-mentioned data acquisition problems, we missed most of the corresponding ozone data for this event and most likely also the peak \chem{[O_3]}.
-A method for gap-filling these data has been presented in \citet{ACP:Falk2021}.
-In contrast, \chem{[O_3]} only rose 18 times above the threshold of $40\,\unit{ppb}$ during the summer of 2019.
-
-Hourly averaged $2\,\unit{m}$ temperatures above $20\,\unit{^\circ C}$ occurred more regularly in July 2018 than in 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}b).
-In 2018, spring temperature regularly rose above freezing only in May, while in 2019 this occurred already early in March/April.
-More rain events with accumulated daily precipitation ($\sum_d \mathrm{Precip}$) above $10\,\unit{mm}$ occurred in the summer of 2018 compared to 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}c).
-Qualitatively, global irradiance ($Q_0$) displayed in Fig.~\ref{fig:data_svanvik_2018_2019}d was higher in May and July 2018 compared to 2019, while June 2019 showed higher irradiance than in 2018.
-A high global irradiance, in most cases, is the result of a low cloud fraction.
-In both years, the maximum recorded $Q_0$ was $750\,\unit{W\,m^{-2}}$ and reached in June.
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{data_svanvik_2018_2019}
-  %\includegraphics[width=12cm]{fig02}.
-  \caption{Observational data from atmospheric monitoring at Svanhovd in 2018/19.
-The hatched areas indicate periods without ozone monitoring data.
-(a) Hourly averaged ozone VMR; (b) hourly averaged temperature; (c) daily accumulated precipitation; (d) hourly averaged global irradiance.}
-  \label{fig:data_svanvik_2018_2019}
-\end{figure*}
-
-\subsection{Climate (1990-present)}
-\label{subsec:climatologies}
-
-Climatologies of temperature and precipitation for Svanhovd were derived for the period 1992--2012 and are shown in a combined climate diagram (Fig.~\ref{fig:climatediagram}).
-Monthly averaged $2\,\unit{m}$ temperatures (red line) are displayed with a standard deviation as error band.
-Averaged monthly accumulated precipitation (blue bars) is shown without standard deviation.
-Months with average temperatures below freezing are denoted with a star.
-As can be seen, temperatures stay below freezing for 5 consecutive months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for K\"{o}ppen's climate classification of a regular subarctic climate (Dfc) \citep[e.g.][]{SD:Beck2018}.
-The highest monthly average temperature is $(13.1\pm 1.1)\,\unit{^\circ C}$ in July and the lowest $(-12.8\pm 2.0)\,\unit{^\circ C}$ in February.
-The coldest measured temperature was $-45.2\,\unit{^\circ C}$ in January 27, 1999, while the highest temperature ($29.4\,\unit{^\circ C}$) occurred on July 16 the same year.
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{climatediagram_box}
-  %\includegraphics[width=8.3cm]{fig04}
-  \caption{Climate diagram based on meteorological data for Svanvik/Pasvik (1992--2012).
-Ozone climatology based on ozone monitoring data at NIBIO Environment Centre Svanhovd(1986--2003).
-A box indicates the upper/lower quartile, whiskers the $1.5\times$ Interquartile Range (IQR), and circles outliers.
-The median is marked by a colored horizontal line and the mean by a triangle.
-Notches indicate the confidence interval around the median.
-(a) Temperature; (b) accumulated precipitation; (c) radiant energy; (d) ozone VMR}
-  \label{fig:climatediagram}
-\end{figure*}
-
-The average accumulated monthly precipitation ($\sum_d\mathrm{Precip}$) indicates that winter and spring (November--April, except for March) are relatively dry ($\sum_d\mathrm{Precip} < 20\,\unit{mm}$).
-March precipitation is primarily snow and will influence the start of the growing season.
-The driest month is January with $\sum_d\mathrm{Precip} = (16.7\pm 3.0)\,\unit{mm}$.
-The most precipitation occurs in the summer months, with a $\sum_d\mathrm{Precip} = (58.5\pm 9.2)\,\unit{mm}$ in August.
-The average annual accumulated precipitation ($\sum_m\mathrm{Precip}$) given with standard error of mean is $(383\pm 86)\,\unit{mm}$.
-Precipitation shows a considerable interannual variability.
-
-The climatology for global irradiance ($Q_0$, not shown) approaches zero in December/January (polar night) and reaches its maximum in June/July (midnight sun conditions).
-We deduce the corresponding highest $Q_0$ to $800\,\unit{W\,m^{-2}}$.
-The maximum of the lowest $Q_0$ (including night) amounts to $200\,\unit{W\,m^{-2}}$ indicating that a considerable amount of light is available under midnight sun conditions.
-
-%Ozone
-The daily mean ozone climatology (\chem{\left<[O_3]\right>}) (1986--1996) is shown in Figure~\ref{fig:climatediagram}d.
-Darker colors indicate higher probability to observe these values.
-On top of the density distributions, a $10\,\unit{days}$ average of daily mean (\chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}}) is displayed together with $1 \sigma$ uncertainties and SE, respectively.
-A spline was fitted through the data to guide the eye.
-The \chem{[O_3]} density distribution and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} display the expected, pronounced seasonal cycle.
-The spring peak occurs on average on day of the year (\unit{doy}) $100$ and amounts to $40\,\unit{ppb}$ while the annual average \chem{\left<[O_3]\right>} is $28\,\unit{ppb}$.
-The decline in \chem{\left<[O_3]\right>} coincides with the average beginning of \chem{CO_2} uptake by coniferous trees \citep{TB:Kolari2007, TP:Wallin2013}.
-In July--September (\unit{doy}~182--273), ozone is occasionally almost completely depleted.
-This hints to ozone uptake by vegetation even at low light intensities during midnight sun conditions in combination with stable planetary boundary layer conditions preventing mixing of ozone rich air.
-
-\subsection{2018/19 growing season anomalies}
-\label{subsec:anomalies}
-In the following, we look at the 2018/19 anomalies and discuss the differences between the two years.
-As weather extremes like in 2018 are likely to become more frequent with climate change, this will help to assess possible future changes in the vulnerability of subarctic vegetation to ozone.\\
-
-In Fig.~\ref{fig:anomalies_svanvik}a, temperature anomalies for Svanvik are shown as a percentage of days warmer or colder than $\pm 1\,\sigma$ from the climatology for each month.
-Negative deviations from climatology are displayed as a negative percentage.
-The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
-We find, that the summer of 2018 was significantly warmer than average.
-Especially in May and July more than $40\,\unit{\%}$ of the days were significantly warmer.
-A significant number of warmer days continued to occur in all months (August--November) except for October.
-March 2018 had many unusually cold days.
-The summer of 2019 was fairly average, however, more significantly colder days occurred in July, while August had both significantly colder and warmer days.
-April 2019 had significantly warmer than average days.
-
-\begin{figure*}[t]
-
-  \centering
-  \includegraphics[width=12cm]{anomalies}
-  \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as difference from median for each month.
-The hatched area between indicates the IQR.
-(a) Temperature; (b) precipitation; (c) radiant energy; (d) ozone VMR.}
-  \label{fig:anomalies_svanvik}
-\end{figure*}
-
-Similarly, the precipitation anomalies for Svanvik are displayed in Fig.~\ref{fig:anomalies_svanvik}b.
-Due to the large interannual variability in observed precipitation, only a few days deviated significantly from the climatology on the $1 \sigma$ level.
-Therefore, we lowered the $\sigma$ level constraint to $\pm \frac{1}{2}\,\sigma$.
-The percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ are shown for each month.
-Positive deviation refers to days wetter than average and negative deviation to drier than average.
-The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm\frac{1}{2}\,\sigma$ if a normal distribution is assumed ($30.8\,\unit{\%}$).
-Unlike the temperature anomalies, the picture is not as clear.
-While March 2018 had a significant number of days wetter than usual, summer and fall (May/July, September/October) had a significant number of days drier than average.
-2019 was rather average throughout the year, but with a significant number of days drier than normal in September/October and wetter in December.
-
-The global irradiance anomalies for Svanvik are presented in Fig.~\ref{fig:anomalies_svanvik}c.
-The percentage of days above or below $\pm 1\,\sigma$ from the climatological mean is displayed for each month.
-In summer 2018 (May/July), global irradiance showed a significant number of days with higher than average irradiance, while 2019 was generally an average year with a significant number of days darker in spring (April/May).
-
-%Ozone
-We use the bias-corrected and cross-calibrated ozone climatology \citep{ACP:Falk2021} and assess the monthly significance of the ozone concentration anomalies in 2018/19.
-The resulting percentage of days above/below $\pm 1\,\sigma$ is displayed for each month in Fig.~\ref{fig:anomalies_svanvik}d.
-Note that missing bars for months in winter and spring are both due to the temporal extent of our data and anomalies not meeting the $1\sigma$ criteria.
-The star indicates the reconstructed data in July.
-In April/May 2018, ozone was significantly elevated, while no significant enhancement is found in 2019.
-In both years, ozone concentrations were significantly lower than average in November.
-These results suggest that stagnant meteorological conditions in April/May in combination with a late start of the growing season contributed more to elevated ozone than the extensive forest fires in Sweden that did not significantly enhance \chem{[O_3]} in 2018.
-
-In summary, 2018 had a significant number of days that were warmer, drier, and brighter than the climatology, while 2019 was a rather average year.
-Ozone was not significantly enhanced during the GS 2018/19, but a substantially higher number of peak \chem{[O_3]} were observed in 2018 than in 2019.
-
-\section{Results}
-\label{sec:results}
-
-\subsection{Parameter localization}
-\label{subsec:res_local}
-
 \subsection{Implications on ozone risk assessments}
@@
 
+\subsection{Climate and weather conditions}
+\label{sec:stats}
+Here, we put the conditions during the growing season 2018/19 into a climatological context.
+
+\subsubsection{Weather conditions in the growing season 2018/19}
+\label{subsec:weather}
+
+All relevant data for the growing season 2018/19 are shown as time series in Fig.~\ref{fig:data_svanvik_2018_2019}.
+Ozone concentrations measured at $2\,\unit{m}$ height above ground are averaged hourly.
+The hatched areas mark times when no ozone data were recorded.
+Note, while the downtime during winter was planned, missing data in two weeks of July 2018 (July 9--23) were due to problems in data acquisition.
+
+As can be seen in Fig.~\ref{fig:data_svanvik_2018_2019}a, \chem{[O_3]} peaks in spring (April/May) and reaches its minimum in late summer (July/August).
+The spring peak has not been captured completely in 2019, since data acquisition only started in late April.
+In summer 2018 (June--August), high ozone concentrations ($\chem{[O_3]} > 40\,\unit{ppb}$) were recorded 50 times.
+The highest summer ozone concentration ($\chem{[O_3]} = 50.2\,\unit{ppb}$) was measured on July 25.
+This coincides with the period of the most extensive forest fires in central Sweden which occurred from July 12--29 \citep{SOU2019}.
+However, due to the above-mentioned data acquisition problems, we missed most of the corresponding ozone data for this event and most likely also the peak \chem{[O_3]}.
+A method for gap-filling these data has been presented in \citet{ACP:Falk2021}.
+In contrast, \chem{[O_3]} only rose 18 times above the threshold of $40\,\unit{ppb}$ during the summer of 2019.
+
+Hourly averaged $2\,\unit{m}$ temperatures above $20\,\unit{^\circ C}$ occurred more regularly in July 2018 than in 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}b).
+In 2018, spring temperature regularly rose above freezing only in May, while in 2019 this occurred already early in March/April.
+More rain events with accumulated daily precipitation ($\sum_d \mathrm{Precip}$) above $10\,\unit{mm}$ occurred in the summer of 2018 compared to 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}c).
+Qualitatively, global irradiance ($Q_0$) displayed in Fig.~\ref{fig:data_svanvik_2018_2019}d was higher in May and July 2018 compared to 2019, while June 2019 showed higher irradiance than in 2018.
+A high global irradiance, in most cases, is the result of a low cloud fraction.
+In both years, the maximum recorded $Q_0$ was $750\,\unit{W\,m^{-2}}$ and reached in June.
+
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{data_svanvik_2018_2019}
+  %\includegraphics[width=12cm]{fig02}.
+  \caption{Observational data from atmospheric monitoring at Svanhovd in 2018/19.
+The hatched areas indicate periods without ozone monitoring data.
+(a) Hourly averaged ozone VMR; (b) hourly averaged temperature; (c) daily accumulated precipitation; (d) hourly averaged global irradiance.}
+  \label{fig:data_svanvik_2018_2019}
+\end{figure*}
+
+\subsubsection{Climate (1990-present)}
+\label{subsec:climatologies}
+
+Climatologies of temperature and precipitation for Svanhovd were derived for the period 1992--2012 and are shown in a combined climate diagram (Fig.~\ref{fig:climatediagram}).
+Monthly averaged $2\,\unit{m}$ temperatures (red line) are displayed with a standard deviation as error band.
+Averaged monthly accumulated precipitation (blue bars) is shown without standard deviation.
+Months with average temperatures below freezing are denoted with a star.
+As can be seen, temperatures stay below freezing for 5 consecutive months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for K\"{o}ppen's climate classification of a regular subarctic climate (Dfc) \citep[e.g.][]{SD:Beck2018}.
+The highest monthly average temperature is $(13.1\pm 1.1)\,\unit{^\circ C}$ in July and the lowest $(-12.8\pm 2.0)\,\unit{^\circ C}$ in February.
+The coldest measured temperature was $-45.2\,\unit{^\circ C}$ in January 27, 1999, while the highest temperature ($29.4\,\unit{^\circ C}$) occurred on July 16 the same year.
+
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{climatediagram_box}
+  %\includegraphics[width=8.3cm]{fig04}
+  \caption{Climate diagram based on meteorological data for Svanvik/Pasvik (1992--2012).
+Ozone climatology based on ozone monitoring data at NIBIO Environment Centre Svanhovd(1986--2003).
+A box indicates the upper/lower quartile, whiskers the $1.5\times$ Interquartile Range (IQR), and circles outliers.
+The median is marked by a colored horizontal line and the mean by a triangle.
+Notches indicate the confidence interval around the median.
+(a) Temperature; (b) accumulated precipitation; (c) radiant energy; (d) ozone VMR}
+  \label{fig:climatediagram}
+\end{figure*}
+
+The average accumulated monthly precipitation ($\sum_d\mathrm{Precip}$) indicates that winter and spring (November--April, except for March) are relatively dry ($\sum_d\mathrm{Precip} < 20\,\unit{mm}$).
+March precipitation is primarily snow and will influence the start of the growing season.
+The driest month is January with $\sum_d\mathrm{Precip} = (16.7\pm 3.0)\,\unit{mm}$.
+The most precipitation occurs in the summer months, with a $\sum_d\mathrm{Precip} = (58.5\pm 9.2)\,\unit{mm}$ in August.
+The average annual accumulated precipitation ($\sum_m\mathrm{Precip}$) given with standard error of mean is $(383\pm 86)\,\unit{mm}$.
+Precipitation shows a considerable interannual variability.
+
+The climatology for global irradiance ($Q_0$, not shown) approaches zero in December/January (polar night) and reaches its maximum in June/July (midnight sun conditions).
+We deduce the corresponding highest $Q_0$ to $800\,\unit{W\,m^{-2}}$.
+The maximum of the lowest $Q_0$ (including night) amounts to $200\,\unit{W\,m^{-2}}$ indicating that a considerable amount of light is available under midnight sun conditions.
+
+%Ozone
+The daily mean ozone climatology (\chem{\left<[O_3]\right>}) (1986--1996) is shown in Figure~\ref{fig:climatediagram}d.
+Darker colors indicate higher probability to observe these values.
+On top of the density distributions, a $10\,\unit{days}$ average of daily mean (\chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}}) is displayed together with $1 \sigma$ uncertainties and SE, respectively.
+A spline was fitted through the data to guide the eye.
+The \chem{[O_3]} density distribution and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} display the expected, pronounced seasonal cycle.
+The spring peak occurs on average on day of the year (\unit{doy}) $100$ and amounts to $40\,\unit{ppb}$ while the annual average \chem{\left<[O_3]\right>} is $28\,\unit{ppb}$.
+The decline in \chem{\left<[O_3]\right>} coincides with the average beginning of \chem{CO_2} uptake by coniferous trees \citep{TB:Kolari2007, TP:Wallin2013}.
+In July--September (\unit{doy}~182--273), ozone is occasionally almost completely depleted.
+This hints to ozone uptake by vegetation even at low light intensities during midnight sun conditions in combination with stable planetary boundary layer conditions preventing mixing of ozone rich air.
+
+\subsection{2018/19 growing season anomalies}
+\label{subsec:anomalies}
+In the following, we look at the 2018/19 anomalies and discuss the differences between the two years.
+As weather extremes like in 2018 are likely to become more frequent with climate change, this will help to assess possible future changes in the vulnerability of subarctic vegetation to ozone.\\
+
+In Fig.~\ref{fig:anomalies_svanvik}a, temperature anomalies for Svanvik are shown as a percentage of days warmer or colder than $\pm 1\,\sigma$ from the climatology for each month.
+Negative deviations from climatology are displayed as a negative percentage.
+The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
+The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
+We find, that the summer of 2018 was significantly warmer than average.
+Especially in May and July more than $40\,\unit{\%}$ of the days were significantly warmer.
+A significant number of warmer days continued to occur in all months (August--November) except for October.
+March 2018 had many unusually cold days.
+The summer of 2019 was fairly average, however, more significantly colder days occurred in July, while August had both significantly colder and warmer days.
+April 2019 had significantly warmer than average days.
+
+\begin{figure*}[t]
+
+  \centering
+  \includegraphics[width=12cm]{anomalies}
+  \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as difference from median for each month.
+The hatched area between indicates the IQR.
+(a) Temperature; (b) precipitation; (c) radiant energy; (d) ozone VMR.}
+  \label{fig:anomalies_svanvik}
+\end{figure*}
+
+Similarly, the precipitation anomalies for Svanvik are displayed in Fig.~\ref{fig:anomalies_svanvik}b.
+Due to the large interannual variability in observed precipitation, only a few days deviated significantly from the climatology on the $1 \sigma$ level.
+Therefore, we lowered the $\sigma$ level constraint to $\pm \frac{1}{2}\,\sigma$.
+The percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ are shown for each month.
+Positive deviation refers to days wetter than average and negative deviation to drier than average.
+The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm\frac{1}{2}\,\sigma$ if a normal distribution is assumed ($30.8\,\unit{\%}$).
+Unlike the temperature anomalies, the picture is not as clear.
+While March 2018 had a significant number of days wetter than usual, summer and fall (May/July, September/October) had a significant number of days drier than average.
+2019 was rather average throughout the year, but with a significant number of days drier than normal in September/October and wetter in December.
+
+The global irradiance anomalies for Svanvik are presented in Fig.~\ref{fig:anomalies_svanvik}c.
+The percentage of days above or below $\pm 1\,\sigma$ from the climatological mean is displayed for each month.
+In summer 2018 (May/July), global irradiance showed a significant number of days with higher than average irradiance, while 2019 was generally an average year with a significant number of days darker in spring (April/May).
+
+%Ozone
+We use the bias-corrected and cross-calibrated ozone climatology \citep{ACP:Falk2021} and assess the monthly significance of the ozone concentration anomalies in 2018/19.
+The resulting percentage of days above/below $\pm 1\,\sigma$ is displayed for each month in Fig.~\ref{fig:anomalies_svanvik}d.
+Note that missing bars for months in winter and spring are both due to the temporal extent of our data and anomalies not meeting the $1\sigma$ criteria.
+The star indicates the reconstructed data in July.
+In April/May 2018, ozone was significantly elevated, while no significant enhancement is found in 2019.
+In both years, ozone concentrations were significantly lower than average in November.
+These results suggest that stagnant meteorological conditions in April/May in combination with a late start of the growing season contributed more to elevated ozone than the extensive forest fires in Sweden that did not significantly enhance \chem{[O_3]} in 2018.
+
+In summary, 2018 had a significant number of days that were warmer, drier, and brighter than the climatology, while 2019 was a rather average year.
+Ozone was not significantly enhanced during the GS 2018/19, but a substantially higher number of peak \chem{[O_3]} were observed in 2018 than in 2019.
+
 \conclusions[Discussion and conclusions]

### 2022-04-20 VegOzone.tex
Deduce -> determine

@@
 
-We deduced the dominant species on-site at Svanhovd and from Fig.~\ref{fig:svanhovd_research_station}a and found perennial grassland, birch (generalized as deciduous trees), and Scots pine (generalized as coniferous trees).
+We determined the dominant species on-site at Svanhovd and from Fig.~\ref{fig:svanhovd_research_station}a and found perennial grassland, birch (generalized as deciduous trees), and Scots pine (generalized as coniferous trees).
 Basic parameters for these PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as mapping manual (MM) parameterizations.
@@
 Low soil water potential reduced the $\mathrm{POD_1}$ values in 2018 in some cases when using the MM growing season, but did not or only slightly when using bespoke GS.
-The temperature acclimation-dependent difference in ozone uptake deduced for bespoke GS ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
-The maximum deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of $f_\mathrm{light}$ amount to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.
+The temperature acclimation-dependent difference in ozone uptake determined for bespoke GS ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+The maximum deviations determined from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of $f_\mathrm{light}$ amount to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.
 
@@
 Taking the bespoke GS and canopy height into account, we find a higher ozone uptake in 2019 than in 2018.
-The overall difference deduced from bespoke GS for subarctic and cold parameterizations with respect to MM range between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
-The maximum deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of $f_\mathrm{light}$ and amounts to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.
+The overall difference determined from bespoke GS for subarctic and cold parameterizations with respect to MM range between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+The maximum deviations determined from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of $f_\mathrm{light}$ and amounts to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.
 
@@
 Perennial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant without bespoke GS.
-The overall difference deduced from bespoke GS for subarctic and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+The overall difference determined from bespoke GS for subarctic and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
 The lower maximum temperature tolerance in the subarctic compared to the cold temperature parameterization probably caused the lower $\mathrm{POD_1}$ values calculated for 2018 in subarctic compared to cold climate, whereas this was not seen for 2019.
-The maximum deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of $f_\mathrm{light}$ amount to $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
+The maximum deviations determined from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of $f_\mathrm{light}$ amount to $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
 
@@
 The climatology for global irradiance ($Q_0$, not shown) approaches zero in December/January (polar night) and reaches its maximum in June/July (midnight sun conditions).
-We deduce the corresponding highest $Q_0$ to $800\,\unit{W\,m^{-2}}$.
+We determine the corresponding highest $Q_0$ to $800\,\unit{W\,m^{-2}}$.
 The maximum of the lowest $Q_0$ (including night) amounts to $200\,\unit{W\,m^{-2}}$ indicating that a considerable amount of light is available under midnight sun conditions.
@@
 The determined start of growing season in 2018 ($\unit{doy}~122$) and 2019 ($\unit{doy}~106$) lies within the range given by observations in Fennoscandia \citep{TB:Kolari2007,IVL:Karlsson2018}.
-We deduced the start of the growing season for deciduous trees based on gridded observational temperature and for perennial grassland on snow depth data.
+We determined the start of the growing season for deciduous trees based on gridded observational temperature and for perennial grassland on snow depth data.
 For perennial grassland, we assumed an one-month latency period after snowmelt that is supported by observational data from Rovaniemi (Finland) \citep{FCR:Korhonen2018}.
@@
 Assuming a proportionality between $g_\mathrm{sto}$ and $A_\mathrm{n}$, this is equivalent to optimizing carbon uptake.
-Similarly, we deduced a PPFD PDF and varied the light response function to test a higher or lower extent of stomatal opening at low light intensities.
+Similarly, we determined a PPFD PDF and varied the light response function to test a higher or lower extent of stomatal opening at low light intensities.
 Remark that these parameterizations are hypothetical and need verification by experimental data.

### 2022-04-20 VegOzone.tex
GS revisited

@@
 
-We determined the dominant species on-site at Svanhovd and from Fig.~\ref{fig:svanhovd_research_station}a and found perennial grassland, birch (generalized as deciduous trees), and Scots pine (generalized as coniferous trees).
+We determined the dominant species on-site at Svanhovd from Fig.~\ref{fig:svanhovd_research_station}a and found perennial grassland, birch (generalized as deciduous trees), and Scots pine (generalized as coniferous trees).
 Basic parameters for these PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as mapping manual (MM) parameterizations.
@@
 For a comprehensive list of MM parameters, consult Supplement~Table~S1.
-In the following, we suggest localizations of key environmental parameters to a subarctic climate based on meteorological data to assess over-/underestimation of ozone damage by using MM Boreal parameterizations.
+In the following, we suggest localizations of key environmental parameters to a subarctic climate based on meteorological data to assess over-/underestimation of ozone damage when using MM Boreal parameterizations in ozone risk assessments.
 
@@
 We identified $f_\mathrm{T}$ as abnormally low throughout the season, being the limiting factor of stomatal conductance in perennial grassland.
-
-To localize the temperature response function $f_\mathrm{T}$, we divide the temperature dataset into two subsets: 1992--2000 (near past) and 2011--2019 (present).
-From these two subsets, we select dates within the GS (May--September) only and plot the probability density function (PDF) as histogram with $1\,\unit{^\circ C}$ binning.
-We overlay the histograms with $f_\mathrm{T}$ to graphically determine the goodness of fit of the MM parameterization from the overlapping area.
-In Fig.~\ref{fig:f_temp_grassland}a, we show the $f_\mathrm{T}$ for perennial grassland together with the temperature PDFs for near past and present climate conditions.
-The MM parameterization is displayed as a solid line.
-A comparison between the two PDFs indicates that the temperature distribution has been shifted towards higher temperatures reducing the number of days below $5\,\unit{^\circ C}$.
+This suggested that a localization of the parameterization to a subarctic climate was necessary.
+
+To localize the temperature response function $f_\mathrm{T}$, we divided the temperature dataset into two subsets: 1992--2000 (near past) and 2011--2019 (present).
+From these two subsets, we selected only dates within the GS (May--September) and plotted the probability density functions (PDFs) as histograms with $1\,\unit{^\circ C}$ binning (Fig.~\ref{fig:f_temp_grassland}a).
+A comparison between the two PDFs indicates that the temperature distribution has shifted towards higher temperatures reducing the number of days below $5\,\unit{^\circ C}$.
 It is unclear how fast the species composition of perennial grassland responds to these changes.
 Thus, the predominant acclimation at present is not known.
-
-The idea is to define a localized temperature response function by increasing the overlapping area between $f_\mathrm{T}$ and the temperature PDF during the GS.
-To this end, we define two different hypothetical target species: subarctic and cold.
-We construct \emph{cold} as representative for a species that is more tolerant to cold temperatures than MM, but slightly less efficient at warm temperatures than MM.
-This is accomplished by moving $T_\mathrm{min}$ and $T_\mathrm{opt}$ towards cooler temperatures but keep $T_\mathrm{max}$ at its MM value.
-In the same way, \emph{subarctic} is constructed to represent a species that is very tolerant to cold but sensitive to high temperatures, and most efficient at cool temperatures.
-Therefore, we shift both $T\mathrm{min}$ and $T_\mathrm{max}$ to colder temperatures and choose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000.
+Further, we overlaid these histograms with the MM $f_\mathrm{T}$.
+A low overlapping area suggests a low agreement with local climate conditions.
+
+Hence, the idea was to define localized temperature response functions by increasing the overlapping area between $f_\mathrm{T}$ and the temperature PDF during the GS.
+To this end, we defined two different hypothetical target species: subarctic and cold.
+We constructed \emph{cold} as representative for a species that is more tolerant to cold temperatures than MM, but slightly less efficient at warm temperatures than MM.
+This was accomplished by moving $T_\mathrm{min}$ and $T_\mathrm{opt}$ towards cooler temperatures but $T_\mathrm{max}$ was kept at its MM value.
+In the same way, \emph{subarctic} was constructed to represent a species that is very tolerant to cold but sensitive to high temperatures, and most efficient at cool temperatures.
+Therefore, we shifted both $T\mathrm{min}$ and $T_\mathrm{max}$ to colder temperatures and chose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000.
 The most extreme of the two localizations compared to MM is \emph{subarctic}.
 
-For localizing $f_\mathrm{light}$, we presume that the opening extent of stomata at low light conditions differs in subarctic species compared to species in less extreme climates.
-We, therefore, adjust the modelled light sensitivity of stomata such that the PPFD value needed to reach $50\,\unit{\%}$ opening is varied by $\pm 20\,\unit{\%}$.
+For localizing $f_\mathrm{light}$, we presumed that the opening extent of stomata at low light conditions differs in subarctic species compared to species in less extreme climates.
+We, therefore, adjusted the modelled light sensitivity of stomata such that the PPFD value needed to reach $50\,\unit{\%}$ opening is varied by $\pm 20\,\unit{\%}$.
 For this, we derive the inverse function $f_{\mathrm{light},k}^{-1}$ of Eq.~(\ref{eq:flight}) for each PFT $k$ analytically
@@
 
-Coniferous trees are able to start photosynthesis as soon as the leaf temperature reaches a certain threshold \citep{TB:Kolari2007} and are, hence, found to be active already at rather low air temperatures and can reach $60\,\unit{\%}$ photosynthetic activity as early as \unit{doy}~100 but cease activity in less favorable conditions \citep{TB:Kolari2007, TP:Wallin2013}.
-
-To determine the temperature localization for coniferous trees, we use the time series of \chem{CO_2} uptake and temperature at observation sites in southern ($\rightarrow$ \emph{cold}) and northern ($\rightarrow$ \emph{subarctic}) Finland by \citet{TB:Kolari2007}.
-We estimate an optimal temperature interval ($10\,\unit{^\circ C}\le T_\mathrm{opt} \le 15\,\unit{^\circ C}$).
-For simplicity, we assume the same interval for deciduous trees.
+Coniferous trees are known to start photosynthesis as soon as the leaf temperature reaches a certain threshold \citep{TB:Kolari2007} and are, thus, already active at low air temperatures and can reach $60\,\unit{\%}$ photosynthetic activity as early as \unit{doy}~100 but cease activity in less favorable conditions \citep{TB:Kolari2007, TP:Wallin2013}.
+
+To determine the temperature localization for coniferous trees, we used the time series of \chem{CO_2} uptake and temperature at observation sites in southern ($\rightarrow$ \emph{cold}) and northern ($\rightarrow$ \emph{subarctic}) Finland by \citet{TB:Kolari2007}.
+From these, we estimated an optimal temperature interval ($10\,\unit{^\circ C}\le T_\mathrm{opt} \le 15\,\unit{^\circ C}$).
+The same interval was assumed for deciduous trees.
 
@@
 
-Another important factor for the accumulation of ozone dose is the timing and length of the GS.
-
-TODO: estimated as follows.
-For coniferous trees, we use a MODIS (Aqua/Terra) retrieved net photosynthesis product \citep{MODIS_PSN} for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine $A_\text{start}$ and $A_\text{end}$.
-The net photosynthesis vs.
-time of year data can be approximated by a second-order polynomial function (Fig.~\ref{fig:modis_Psn}).
-By calculating the roots of the fitted polynomial, we derive $A_\text{start}$ as \unit{doy}~122 for 2018 and \unit{doy}~106 in 2019.
+The accumulation of ozone dose is strongly depended on the timing and length of the GS.
+\citet{ICP:MappingManual2017} suggest a simple latitude model for beech, birch, and Norway spruce for the localization of GS:
+%
+\begin{align}
+  A_\mathrm{start} &= 105 + (LAT-50\,\unit{^\circ N})\cdot 1.5\\
+  A_\mathrm{end} &= 297 - (LAT-50\,\unit{^\circ N})\cdot 2
+\end{align}
+%
+At Svanhovd, this yields $A_\mathrm{start}=135\,\unit{doy}$ and $A_\mathrm{end}=257\,\unit{doy}$. We verified these as follows.
+
+To determine $A_\text{start}$ and $A_\text{end}$ for coniferous trees, we used a MODIS (Aqua/Terra) retrieved GPP product \citep{MODIS_PSN} on a $1\times 1\,\unit{km}$ patch centered at Svanhovd.
+GPP over time follows approximately a second-order polynomial function (Fig.~\ref{fig:modis_gpp}).
+By calculating the roots of the fitted polynomial, we found $A_\text{start}$ as \unit{doy}~122 for 2018 and \unit{doy}~106 in 2019.
 $A_\text{end}$ amounts to \unit{doy}~261 and 274, respectively.
-This value will be used for all PFTs alike.
-The resulting growing season for coniferous trees in 2019 is, hence, one month longer than in 2018.
-For estimating $A_\text{start}$ of deciduous trees, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Appendix~Fig.~\ref{fig:greening_season_change_Svanvik}).
-We find \unit{doy}~129 and 130, respectively.
+These $A_\mathrm{end}$ value will be used for all PFTs alike.
+The resulting growing season for coniferous trees in 2019 was one month longer than in 2018.
+For $A_\text{start}$ of deciduous trees, we use temperature-degree-days (above $5\,\unit{^\circ C}$) on gridded temperature data from SeNorge.no (Appendix~Fig.~\ref{fig:greening_season_change_Svanvik}).
+We found \unit{doy}~129 and 130, respectively.
 For both years, these dates coincide with the first snow-free day at the closest inland meteorological weather station at Øvre~Neiden (Sør-Varanger, NOR).
-For comparison, the generic dates for central Europe ($A_\mathrm{start} = 100$ and $A_\mathrm{end} = 307$) indicating a substantially longer growing season.
-Due to a lack of quantitative field observation at Svanhovd, we assume a latency period of $1\,\unit{month}$ after snow-melt for perennial grassland.
+These dates are very similar to the latitude model results.
+For perennial grassland, we assume a latency period of $1\,\unit{month}$ after snow-melt.
 This assumption is supported by observations in Rovaniemi, Finland \citep[][Supplement~Fig.~S1]{FCR:Korhonen2018}.
+All results are tabulated in Table~\ref{tab:sensitivity_tests_gs}.
+
+\begin{table}[t]
+  \caption{Start and end of growing season in \unit{doy}.}
+  \label{tab:sensitivity_tests_gs}
+  \begin{tabular}{lcll}
+    \tophline
+    Species & Year & $A_\mathrm{start}$ & $A_\mathrm{end}$\\
+    \middlehline
+    Default$^b$ & -- & 135 & 257\\
+    \multirow{2}{*}{Deciduous tree} & 2018 & 129$^*$ & 261$^a$ \\
+    & 2019 & 130$^c$ & 274$^a$ \\
+    \multirow{2}{*}{Coniferous tree} & 2018 & 122$^a$ & 261$^a$ \\
+    & 2019 & 106$^a$ & 274$^a$ \\
+    \multirow{2}{*}{Perennial grassland} & 2018 & 159$^d$ & 261$^a$\\
+    & 2019 & 161$^d$ & 274$^a$ \\
+    \bottomhline
+  \end{tabular}
+  \belowtable{$^a$ MODIS (Aqua/Terra) GPP product;\\
+    $^b$ \citet{ICP:MappingManual2017},\\
+    $^c$ $5\,\unit{days}-5\,\unit{^\circ C}$-rule, $T_\mathrm{air}$ from \href{seNorge.no}{seNorge.no};\\
+    $^d$ One month after snowmelt; reference station Øvre Neiden (Sør-Varanger, NOR)} % Table Footnotes
+\end{table}
 
@@
 
-\begin{table}[t]
-  \caption{Start and end of growing season in \unit{doy}.}
-  \label{tab:sensitivity_tests_gs}
-  \begin{tabular}{lcll}
-    \tophline
-    Species & Year & $A_\mathrm{start}$ & $A_\mathrm{end}$\\
-    \middlehline
-    Default$^b$ & -- & 100 & 307\\
-    \multirow{2}{*}{Deciduous tree} & 2018 & 129$^*$ & 261$^a$ \\
-    & 2019 & 130$^c$ & 274$^a$ \\
-    \multirow{2}{*}{Coniferous tree} & 2018 & 122$^a$ & 261$^a$ \\
-    & 2019 & 106$^a$ & 274$^a$ \\
-    \multirow{2}{*}{Perennial grassland} & 2018 & 159$^d$ & 261$^a$\\
-    & 2019 & 161$^d$ & 274$^a$ \\
-    \bottomhline
-  \end{tabular}
-  \belowtable{$^a$ MODIS (Aqua/Terra) net photosynthesis product;\\
-    $^b$ \citet{ACP:Simpson2012},\\
-    $^c$ $5\,\unit{days}-5\,\unit{^\circ C}$-rule, $T_\mathrm{air}$ from \href{seNorge.no}{seNorge.no};\\
-    $^d$ One month after snowmelt; reference station Øvre Neiden (Sør-Varanger, NOR)} % Table Footnotes
-\end{table}
-
 \subsection{Implications on ozone risk assessments}
@@
 MODIS data indicate a higher photosynthetic activity in 2018 than in 2019.
-As shown in Fig.~\ref{fig:modis_Psn} we fitted a second order polynomial function of the general form
+As shown in Fig.~\ref{fig:modis_gpp} we fitted a second order polynomial function of the general form
 %
@@
   %\includegraphics[width=8.3cm]{figB2}
-  \includegraphics[width=8.3cm]{modis_Psn.pdf}
-  \caption{Estimated $G_\mathrm{start}$ and $G_\mathrm{end}$ of growing season for coniferous trees from MODIS Aqua/Terra net photosynthesis (PSN) product.
+  \includegraphics[width=8.3cm]{modis_gpp}
+  \caption{Estimated $G_\mathrm{start}$ and $G_\mathrm{end}$ of growing season for coniferous trees from MODIS Aqua/Terra gross primary productivity (GPP) product.
 A $1\times 1\,\unit{km}$ area around Svanhovd was selected.
@@
 The numerically computed root yields: BGS \unit{doy}~122/106 and EGS \unit{doy}~261/274 for 2018/19, respectively.}
-  \label{fig:modis_Psn}
+  \label{fig:modis_gpp}
 \end{figure}

### 2022-04-20 VegOzone.tex
Add figure and finish other props

@@
 
-\subsubsection{Soil conditions}
+\subsubsection{Other properties}
 \label{subsec:soil}
@@
 
-The estimate of leaf boundary layer resistance is a function of wind speed which will be influenced by the tree height and leaf dimensions.
-A sample of downy birch leaves collected at Svanhovd had an average length of $(3.0\pm 0.5)\,\unit{cm}$ that is smaller than in the MM parameterization.
-\citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area.
+The quasi-laminar boundary layer resistance $R_b$ is a function of wind speed $u$ which is influenced by the tree height and leaf dimensions.
+For comparison with MM dimensions, we collected a sample of birch leaves.
+The birch (B.
+pubescence) leaves were collected from the outer canopy (good light exposure) of tall trees at two sites in Finnmark.
+The leaves were collected late in July so they were fully expanded.
+They were collected by hand, so at about $2\,\mathrm{m}$ height.
+As the sun is at a low elevation angle for most of the time in the growing season in this area, light exposure at the top of the tree is not expected to be different from lower leaves as long as they grow in a part of the tree that is not shadowed by other trees.
+At the first site (Karasjok), three leaves were collected from each of five adult trees.
+At the second site (Svanvik), five leaves were collected from one tree.
+The leaves were pressed and dried before scanning for the area and shape determination through image analysis.
+ImageJ was used for thresholding the images to give silhouettes of the leaves.
+After scaling the images, the program was used for finding the minimum Feret diameter.
+This measure finds the largest width of a leaf at a $90^\circ$ angle to the length of the leaf (the Feret diameter).
+With this method we found that the 20 leaves had a mean width of $(3.0\pm 0.5)\,\unit{cm}$ that is smaller than in the MM parameterization.
+
+A report by \citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the Svanhovd area.
 One Scots pine forest at Svanhovd has been studied as a part of the ICP Forest mapping.
@@
 Timmerman, personal communication, June 2020).
-We used $13.5\,\unit{m}$ height for both deciduous and coniferous trees.
+We used $13.5\,\unit{m}$ height for both deciduous and coniferous trees as this number better represents a larger region and diversity of species.
 

### 2022-04-21 VegOzone.tex
Revise Section 2 and include intro to section3

@@
 
-The resulting parameterizations are presented in detail in Section~\ref{subsec:res_local}.
+All resulting temperature response function localizations are tabulated in Table~\ref{tab:sensitivity_tests_temp} and light response function in Table~\ref{tab:sensitivity_tests_light}.
+In Section~\ref{subsec:res_local}, we compare relative $g_\mathrm{sto}$ for the different localizations based on the long term meteorological data.
+
+\begin{table}[t]
+  \caption{Bespoke temperature parameterizations. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
+  \label{tab:sensitivity_tests_temp}
+  \begin{tabular}{llrrr}
+    \tophline
+    Species & type & $T_\mathrm{min}$ & $T_\mathrm{opt}$ & $T_\mathrm{max}$ \\
+    \middlehline
+    \multirow{3}{*}{Deciduous tree} & MM & 5 & 20 & 100\\
+    & Cold & 5 & 15 & 100\\
+    & Subarctic & 0 & 10 & 100\\
+    \middlehline
+    \multirow{3}{*}{Coniferous tree} & MM & 0 & 20 & 100\\
+    & Cold & 0 & 15 & 100\\
+    & Subarctic & 0 & 10 & 100\\
+    \middlehline
+    \multirow{3}{*}{Perennial grassland} & MM & 10 & 24 & 36\\
+    & Cold & 5 & 15 & 36\\
+    & Subarctic & 0 & 10 & 24\\
+    \bottomhline
+    \end{tabular}
+\end{table}
+
+\begin{table}[t]
+  \caption{Bespoke light parameterizations. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
+  \label{tab:sensitivity_tests_light}
+  \begin{tabular}{llcr}
+    \tophline
+    Species & type & $\alpha$ & $\gamma(0.5)$\\
+    \middlehline
+    \multirow{3}{*}{Deciduous tree} & MM & 0.004 & 165.035\\
+    & PPFD0.8 & 0.005 & 132.028\\
+    & PPFD1.2 & 0.003 & 198.042\\
+    \middlehline
+    \multirow{3}{*}{Coniferous tree} & MM & 0.006 & 115.525\\
+    & PPFD0.8 & 0.008 & 92.420\\
+    & PPFD1.2 & 0.005 & 138.629\\
+    \middlehline
+    \multirow{3}{*}{Perennial grassland} & MM & 0.011 & 63.013\\
+    & PPFD0.8 & 0.014 & 50.411\\
+    & PPFD1.2 & 0.009 & 75.616\\
+    \bottomhline
+    \end{tabular}
+\end{table}
 
@@
 
-The quasi-laminar boundary layer resistance $R_b$ is a function of wind speed $u$ which is influenced by the tree height and leaf dimensions.
-For comparison with MM dimensions, we collected a sample of birch leaves.
-The birch (B.
-pubescence) leaves were collected from the outer canopy (good light exposure) of tall trees at two sites in Finnmark.
-The leaves were collected late in July so they were fully expanded.
-They were collected by hand, so at about $2\,\mathrm{m}$ height.
+The quasi-laminar boundary layer resistance $R_b$ is a function of wind speed $u$ and influenced by the tree height and leaf dimensions.
+For comparison with MM, we collected a sample of birch (B.
+pubescence) leaves from the outer canopy (good light exposure) of tall trees at two sites in Finnmark.
+The leaves were collected late in July, thus, they were fully expanded.
+They were collected by hand at about $2\,\mathrm{m}$ height.
 As the sun is at a low elevation angle for most of the time in the growing season in this area, light exposure at the top of the tree is not expected to be different from lower leaves as long as they grow in a part of the tree that is not shadowed by other trees.
-At the first site (Karasjok), three leaves were collected from each of five adult trees.
+At the first site (Karasjok), three leaves were collected from five adult trees each.
 At the second site (Svanvik), five leaves were collected from one tree.
 The leaves were pressed and dried before scanning for the area and shape determination through image analysis.
-ImageJ was used for thresholding the images to give silhouettes of the leaves.
+ImageJ (\href{https://imagej.nih.gov/ij/}{see webpage}) was used for thresholding the images to give silhouettes of the leaves.
 After scaling the images, the program was used for finding the minimum Feret diameter.
 This measure finds the largest width of a leaf at a $90^\circ$ angle to the length of the leaf (the Feret diameter).
-With this method we found that the 20 leaves had a mean width of $(3.0\pm 0.5)\,\unit{cm}$ that is smaller than in the MM parameterization.
-
-A report by \citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the Svanhovd area.
+With this method we found that the 20 leaves had a mean width of $(3.0\pm 0.5)\,\unit{cm}$ that is smaller than in the MM parameterization ($5\,\unit{cm}$, Supplement~Table~S1).
+
+A report by \citet[][p.~52]{NINA2004} indicates an average tree height of $13.5\,\unit{m}$ in the Svanhovd area.
 One Scots pine forest at Svanhovd has been studied as a part of the ICP Forest mapping.
-Heights were measured in 2004 when the stand was 90 years old.
-Mean tree height was $10.1\,\unit{m}$ and maximum height was $16.2\,\unit{m}$ (V.
+Heights were measured in 2004 when the stand was 90~years old.
+The mean tree height was $10.1\,\unit{m}$ and maximum height was $16.2\,\unit{m}$ (V.
 Timmerman, personal communication, June 2020).
-We used $13.5\,\unit{m}$ height for both deciduous and coniferous trees as this number better represents a larger region and diversity of species.
+We used $13.5\,\unit{m}$ height for both deciduous and coniferous trees as this height is sampled from a larger and more diverse population of species.
 
@@
 \label{sec:results}
+In this section, we first compute the average stomatal conductance $\left<g_\mathrm{sto}\right>$ relative to $g_\mathrm{max}$ for each PFT from the long term meteorological data and compare the different localizations with MM (Section~\ref{subsec:res_local}).
+We use the $\mathrm{DO_3SE}$ model to compute the ozone dose and projected biomass loss due to ozone for the two recent GS 2018/19 for each species and localization (Section~\ref{subsec:do3se_results}).
+Finally, we put the results into a climatological perspective (Section~\ref{subsec:climatologies}).
 
@@
 All derived parameters are tabulated in Table~\ref{tab:sensitivity_tests_light}.
-
-\begin{table}[t]
-  \caption{Bespoke temperature parameterizations. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
-  \label{tab:sensitivity_tests_temp}
-  \begin{tabular}{llrrr}
-    \tophline
-    Species & type & $T_\mathrm{min}$ & $T_\mathrm{opt}$ & $T_\mathrm{max}$ \\
-    \middlehline
-    \multirow{3}{*}{Deciduous tree} & MM & 5 & 20 & 100\\
-    & Cold & 5 & 15 & 100\\
-    & Subarctic & 0 & 10 & 100\\
-    \middlehline
-    \multirow{3}{*}{Coniferous tree} & MM & 0 & 20 & 100\\
-    & Cold & 0 & 15 & 100\\
-    & Subarctic & 0 & 10 & 100\\
-    \middlehline
-    \multirow{3}{*}{Perennial grassland} & MM & 10 & 24 & 36\\
-    & Cold & 5 & 15 & 36\\
-    & Subarctic & 0 & 10 & 24\\
-    \bottomhline
-    \end{tabular}
-\end{table}
-
-\begin{table}[t]
-  \caption{Bespoke light parameterizations. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
-  \label{tab:sensitivity_tests_light}
-  \begin{tabular}{llcr}
-    \tophline
-    Species & type & $\alpha$ & $\gamma(0.5)$\\
-    \middlehline
-    \multirow{3}{*}{Deciduous tree} & MM & 0.004 & 165.035\\
-    & PPFD0.8 & 0.005 & 132.028\\
-    & PPFD1.2 & 0.003 & 198.042\\
-    \middlehline
-    \multirow{3}{*}{Coniferous tree} & MM & 0.006 & 115.525\\
-    & PPFD0.8 & 0.008 & 92.420\\
-    & PPFD1.2 & 0.005 & 138.629\\
-    \middlehline
-    \multirow{3}{*}{Perennial grassland} & MM & 0.011 & 63.013\\
-    & PPFD0.8 & 0.014 & 50.411\\
-    & PPFD1.2 & 0.009 & 75.616\\
-    \bottomhline
-    \end{tabular}
-\end{table}
 

### 2022-04-21 VegOzone.tex
Move reference to Appendix figures
Finish first part of section 3

@@
 The same interval was assumed for deciduous trees.
+For deciduous and coniferous trees refer to Appendix~Figs.~\ref{fig:f_temp_spruce}b and \ref{fig:f_temp_birch}b), respectively.
 
@@
 The clay:sand:silt content is estimated to range from 12\,\%:61\,\%:27\,\% at the top layer to 16\,\%:59\,\%:25\,\% in about $1\,\unit{m}$ depth (\href{https://soilgrids.org/}{soilgrids.org}, last accessed April 2022 for location $30.0334\,\unit{^\circ E}$, $69.4549\,\unit{^\circ N}$) Within the $\mathrm{DO_3SE}$ soil parameterization, the sandy loam texture describes the upper $60\,\unit{cm}$, where the plant roots are found, of this soil type best.
+We presume that no localization of the water vapor pressure deficite $f_\mathrm{VPD}$ and soil water potential $f_\mathrm{SWP}$ response functions is necessary for our species.
+But acknowledge that these are very important factors in the modeling of ozone dose as pointed out by~\citet{ACP:Bueker2012}.
 
@@
 
-All bespoke temperature parameters are listed in Table~\ref{tab:sensitivity_tests_temp}.
-
-(for deciduous and coniferous trees refer Appendix~Figs.~\ref{fig:f_temp_spruce}b and \ref{fig:f_temp_birch}b)
-
-All derived parameters are tabulated in Table~\ref{tab:sensitivity_tests_light}.
-
 We presume that perennial vegetation will likely maximize carbon acquisition during the short subarctic growing season.
-Because the MM version of the $\mathrm{DO_3SE}$ model currently does not simulate net photosynthesis ($A_\mathrm{n}$), we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune the temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ to acclimate the growth potential to the local climate.
-For details regarding these functions see Section~\ref{subsec:model_describtion}.
-Further, we presume that $f_\mathrm{VPD}$ and $f_\mathrm{SWP}$ suit our vegetation types and no acclimation is necessary for these.
-But acknowledge that these are very important factors as shown by~\citet{ACP:Bueker2012}.
-
-As indicated above, a better acclimation to prevailing climate conditions is characterized by a higher average $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$.
-To compare our bespoke parameterizations, we propose the following metric.
-We calculate an average relative stomatal conductance $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal}) with $f_\mathrm{SWP}=1$) at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time) for the whole record of key environmental variables' data and apply mean and standard deviation.
-We presume that $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ around noon (highest light intensity) is a good proxy for \chem{CO_2} uptake efficiency (growth).
-In addition, a small deviation between noon and morning and a low standard deviation would indicate higher robustness to variability in growing conditions.
-Based on these criteria, we identify \emph{subarctic}--PPFD0.8 as the best parameterization for all PFTs (Fig.~\ref{fig:javis_func_opt_mean}).
-As expected due to the small adjustments, coniferous trees display the smallest differences between the different parameterizations, while the differences for perennial grassland are substantial as a response to the proposed temperature acclimation.
+Because the MM version of the $\mathrm{DO_3SE}$ model currently does not simulate net photosynthesis $A_\mathrm{n}$, we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011}.
+A higher average $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ ratio, thus, indicates a higher growth potential in the local climate.
+
+To compare our localized parameterizations, we propose the following metric.
+From the long term of meteorological data, we select to distinct periods noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and morning ($5-9\,\unit{am}$ local time).
+For these, we calculate the average relative stomatal conductance $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal}) and assume $f_\mathrm{SWP}=1$).
+We presume that $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ around noon (the highest light intensity) is a good proxy for \chem{CO_2} uptake efficiency (growth).
+In addition, a small deviation between noon and morning and a low standard deviation indicates higher robustness of this localization to variability in growing conditions.
+From left to right in Fig.~\ref{fig:javis_func_opt_mean}, results using the subarctic, cold, and MM $f_\mathrm{T}$ are shown marked by the alternating background color.
+Within each of these groups $f_\mathrm{light}$ categories are indicated by PPFD0.8 and PPFD1.2, respectively.
+The standard deviation is denoted with error bars.
+
+Based on these criteria, we identify \emph{subarctic}--PPFD0.8 as the parameterization with the highest score for all PFTs.
+The differences between the localized parameterizations and the MM are smallest for coniferous trees.
+This indicates that the MM Boreal parameterization is well suited also for subarctic climate at Svanhovd.
+For perennial grassland, the difference compared to MM is largest suggesting that the MM Central European parameterization is not appropriate in a subarctic climate.
 
@@
 \end{figure}
-
-Start ($A_\text{start}$) and end ($A_\text{end}$) of the GS for each PFT are usually given in \unit{doy} (Table~\ref{tab:sensitivity_tests_gs})
 

### 2022-04-21 VegOzone.tex
Rephrase decideous tree paragraph

@@
 
-We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for the three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using the generic MM parameterization and our bespoke temperature (\emph{cold}, \emph{subarctic}) and light (PPFD0.8, PPFD1.2) parameterizations.
-Canopy height and leaf size are adjusted to local values in the model simulations with the bespoke growing season (GS besp.).
-The results displayed in Fig.~\ref{fig:pody_rel} allow for an assessment of over- or underestimation of ozone risk arising from the use of generic parameterizations established from less climatically extreme conditions.
+We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for the three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using the MM parameterization and our localized temperature (\emph{cold}, \emph{subarctic}) and light (PPFD0.8, PPFD1.2) parameterizations.
+
+The results displayed in Fig.~\ref{fig:pody_rel} allow for an assessment of over- or underestimation of ozone risk arising from the use of MM parameterizations established from less climatically extreme conditions.
 Critical levels for reduction in biomass \citep[deciduous forest $4\,\unit{\%}$, coniferous forest $2\,\unit{\%}$, and grasslands $10\,\unit{\%}$ biomass reduction,][]{ICP:MappingManual2017,ESPR:Hayes2021} are shown as horizontal dashed lines.
+The different background colors indicate the temperature response function localization.
+The light response function localization is denoted by $+$ (PPFD1.2) and $-$ (PPFD0.8).
+Circles indicate simulations with localized GS, canopy height, and leaf dimensions and squares simulations with MM GS.
+With open symbols the effect on $\mathrm{POD_y}$ of taking SWP into account is displayed.
+
 We identify that an increase in cold tolerance (MM $\rightarrow$ cold $\rightarrow$ subarctic) leads to higher ozone uptake in both probed years, while the usage of locally adapted GS causes a reduction.
-The earlier and longer GS in 2019 leads to increased ozone uptake compared to 2018 despite more episodes of elevated \chem{[O_3]} in 2018.
+The earlier and longer GS in 2019 leads to increased ozone uptake compared to 2018 despite more episodes of elevated \chem{[O_3]} in 2018 (see Section~\ref{subsec:anomalies} for details).
 Due to the shape of $f_\mathrm{light}$, a symmetric variation leads to an asymmetric response in $\mathrm{POD_1}$.
-We find that an opening of stomata at lower light intensities can cause higher sensitivity to drought conditions.
-Similarly, a temperature acclimation to cold or subarctic conditions can lead to an amplification of effects by droughts.
-
-The magnitude of these effects varies between PFTs as well as years, but the predicted ozone uptake for the bespoke temperature parameterization is always larger than for the MM parameterizations and of the same order of magnitude as the variability between the years studied here.
-Our bespoke temperature parameterizations suggest considerable underestimations of ozone risk for subarctic species when relying on the generic MM parameterizations.
+We find that an opening of stomata at lower light intensities will cause a reduction of $\mathrm{POD_1}$ due to drought conditions.
+This is seen for both $f_\mathrm{T}$ localization in 2018 but not 2019 (compare PPFD0.8 and PPFD1.2 in Fig.~\ref{fig:pody_rel}a, subarctic, 2018, with SWP effect on).
 
 For deciduous trees (Fig.~\ref{fig:pody_rel}a) all $\mathrm{POD_1}$ simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
-Deciduous trees display the largest variation of ozone uptake among all PFTs ($10-31\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
-In particular, the sensitivity to the light parameterization is pronounced.
-With MM GS, $\mathrm{POD_1}$ values were higher in 2018 than in 2019, throughout our analyses.
-However, taking the bespoke GS into account, this is reversed for the subarctic temperature acclimation.
+Among all PFTs, deciduous trees display the largest sensitivity in $\mathrm{POD_1}$.
+Compared to MM $\Delta\mathrm{POD_1}=(10-31)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
+In particular, the sensitivity to the $f_\mathrm{light}$ is pronounced (extent of $\angle(+,0,-)$).
+With MM GS, $\mathrm{POD_1}$ values were higher in 2018 than in 2019.
+However, taking the localized GS into account, this is reversed for the subarctic temperature response function.
 This can be explained by more favorable growing conditions in 2019.
-Low soil water potential reduced the $\mathrm{POD_1}$ values in 2018 in some cases when using the MM growing season, but did not or only slightly when using bespoke GS.
-The temperature acclimation-dependent difference in ozone uptake determined for bespoke GS ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
-The maximum deviations determined from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of $f_\mathrm{light}$ amount to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.
+Low soil water potential reduced the $\mathrm{POD_1}$ values in 2018 in some cases when using the MM GS, but did not or only slightly when using the localized GS.
+This indicates (frost) drought conditions early in the regular GS in 2018 when the ground was still covered in snow.
+The difference in $\mathrm{POD_1}$ due to the localization of $f_\mathrm{T}$ for the localized GS ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+The maximum difference due to the $f_\mathrm{light}$ localization for the PPFD0.8 (superscript) and PPFD1.2 (subscript) amount to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.
 
@@
 
-From the difference between 2018 and 2019, we infer a substantial interannual variability.
-The maximum deviation from the generic default parameterizations due to our bespoke temperature and light response functions is of the same order of magnitude as the interannual variability.
+From the difference between 2018 and 2019, we conclude that the interannual variability of ozone dose is non-negligible.
+The maximum difference in $\mathrm{POD_1}$ between simulations with the MM parameterizations and the localized temperature and light response functions is of the same order of magnitude as the interannual difference.
+The magnitude of all described effects on $\mathrm{POD_1}$ varies between PFTs as well as years, but the predicted ozone uptake for the localized $f_\mathrm{T}$ is always larger than for the MM parameterizations.
+This suggests underestimation of ozone risk for subarctic species when using the MM parameterizations.
 

### 2022-04-24 VegOzone.tex
Fix formating issue in formula

@@
 In the same way, \emph{subarctic} was constructed to represent a species that is very tolerant to cold but sensitive to high temperatures, and most efficient at cool temperatures.
-Therefore, we shifted both $T\mathrm{min}$ and $T_\mathrm{max}$ to colder temperatures and chose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000.
+Therefore, we shifted both $T_\mathrm{min}$ and $T_\mathrm{max}$ to colder temperatures and chose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000.
 The most extreme of the two localizations compared to MM is \emph{subarctic}.

### 2022-04-24 VegOzone.tex
Rewrite climate section

@@
 Stomatal conductance has been empirically linked to key environmental variables such as air temperature $T$, photosynthetic photon flux density (PPFD), vapor pressure deficit (VPD), and soil water potential (SWP) as well as photosynthesis itself \citep[e.g.,][]{PTRS:Jarvis1976, BallBerry1987, Emberson2000, ICP:MappingManual2017}.
-Results from open-top chamber (OTC) experiments on downy birch (\emph{Betula pubescens}) and mountain birch (\emph{Betula pubescens toruosa}), native to subarctic regions, indicated reductions in both biomass, root:shoot ratio, and visible leaf damage under elevated ozone treatment ($\left<\chem{[O_3]}\right> = 36-54\,\unit{ppb}$) \citep{Amb:Manninen2009}.
+Results from open-top chamber (OTC) experiments on downy birch (\emph{Betula pubescens}) and mountain birch (\emph{Betula pubescens toruosa}), native to subarctic regions, indicated reductions in both biomass, root:shoot ratio, and visible leaf damage under elevated ozone treatment ($\left<\chem{\chi_{O_3}}\right> = 36-54\,\unit{ppb}$) \citep{Amb:Manninen2009}.
 Though, Scots pine (\emph{Pinus sylvestris}) is considered to be more ozone tolerant due to an absence of visible injuries \citep{Amb:Girgzdiene2009}, \citet{Amb:Manninen2009} found chlorophyll:carotenoid ratio and polyamines reductions under elevated ozone concentrations indicating susceptibility to ozone also in these species.
@@
 %
-First, we solve Eq.~(\ref{eq:inverse_function}) for the MM default value of $\alpha_\mathrm{MM}^k$ at $0.5$ ($50\,\%$ opening)
+First, we solved Eq.~(\ref{eq:inverse_function}) for the MM default value of $\alpha_\mathrm{MM}^k$ at $0.5$ ($50\,\%$ opening)
 \begin{equation}
@@
 \end{equation}
-and define a variation $\gamma' = \gamma \cdot \eta$ with $\eta \in \{0.8, 1.2\}$. By solving Eq.~(\ref{eq:inverse_function}) for $\alpha$, we find $\alpha'(\gamma')$.
+and defined a variation $\gamma' = \gamma \cdot \eta$ with $\eta \in \{0.8, 1.2\}$. By solving Eq.~(\ref{eq:inverse_function}) for $\alpha$, we find $\alpha'(\gamma')$.
 
@@
 The resulting response functions are shown in Appendix~Figs.~\ref{fig:f_temp_spruce} and \ref{fig:f_temp_birch}, respectively.
-
-Coniferous trees are known to start photosynthesis as soon as the leaf temperature reaches a certain threshold \citep{TB:Kolari2007} and are, thus, already active at low air temperatures and can reach $60\,\unit{\%}$ photosynthetic activity as early as \unit{doy}~100 but cease activity in less favorable conditions \citep{TB:Kolari2007, TP:Wallin2013}.
-
+Coniferous trees are known to start photosynthesis as soon as the leaf temperature reaches a certain threshold \citep{TB:Kolari2007}.
+Thus, coniferous trees are already active at low air temperatures and can reach $60\,\unit{\%}$ photosynthetic activity as early as \unit{doy}~100 but cease activity in less favorable conditions \citep{TB:Kolari2007, TP:Wallin2013}.
 To determine the temperature localization for coniferous trees, we used the time series of \chem{CO_2} uptake and temperature at observation sites in southern ($\rightarrow$ \emph{cold}) and northern ($\rightarrow$ \emph{subarctic}) Finland by \citet{TB:Kolari2007}.
 From these, we estimated an optimal temperature interval ($10\,\unit{^\circ C}\le T_\mathrm{opt} \le 15\,\unit{^\circ C}$).
-The same interval was assumed for deciduous trees.
-For deciduous and coniferous trees refer to Appendix~Figs.~\ref{fig:f_temp_spruce}b and \ref{fig:f_temp_birch}b), respectively.
+For simplicity, the same interval was assumed for deciduous trees.
 
 All resulting temperature response function localizations are tabulated in Table~\ref{tab:sensitivity_tests_temp} and light response function in Table~\ref{tab:sensitivity_tests_light}.
-In Section~\ref{subsec:res_local}, we compare relative $g_\mathrm{sto}$ for the different localizations based on the long term meteorological data.
+In Section~\ref{subsec:res_local}, we compare relative $g_\mathrm{sto}$ based on Eq.~(\ref{eq:stomatal}) for the different localizations using the long term meteorological data.
 
@@
 
-A report by \citet[][p.~52]{NINA2004} indicates an average tree height of $13.5\,\unit{m}$ in the Svanhovd area.
+A report by \citet[][p.~52]{NINA2004} indicated an average tree height of $13.5\,\unit{m}$ in the Svanhovd area.
 One Scots pine forest at Svanhovd has been studied as a part of the ICP Forest mapping.
-Heights were measured in 2004 when the stand was 90~years old.
+There, heights were measured in 2004 when the stand was 90~years old.
 The mean tree height was $10.1\,\unit{m}$ and maximum height was $16.2\,\unit{m}$ (V.
 Timmerman, personal communication, June 2020).
-We used $13.5\,\unit{m}$ height for both deciduous and coniferous trees as this height is sampled from a larger and more diverse population of species.
+We used $13.5\,\unit{m}$ for both deciduous and coniferous trees as this height is sampled from a larger and more diverse population of species.
 
@@
 \label{sec:results}
-In this section, we first compute the average stomatal conductance $\left<g_\mathrm{sto}\right>$ relative to $g_\mathrm{max}$ for each PFT from the long term meteorological data and compare the different localizations with MM (Section~\ref{subsec:res_local}).
+In this section, we first compute the average stomatal conductance $\left<g_\mathrm{sto}\right>$ relative to $g_\mathrm{max}$ for each PFT from the long term meteorological data according to Eq.~(\ref{eq:stomatal})and compare the different localizations with MM (Section~\ref{subsec:res_local}).
 We use the $\mathrm{DO_3SE}$ model to compute the ozone dose and projected biomass loss due to ozone for the two recent GS 2018/19 for each species and localization (Section~\ref{subsec:do3se_results}).
@@
 The differences between the localized parameterizations and the MM are smallest for coniferous trees.
-This indicates that the MM Boreal parameterization is well suited also for subarctic climate at Svanhovd.
-For perennial grassland, the difference compared to MM is largest suggesting that the MM Central European parameterization is not appropriate in a subarctic climate.
+This indicates that the MM Boreal parameterization is well suited also for the subarctic climate at Svanhovd.
+For perennial grassland, the differences compared to MM are largest suggesting that the MM Central European parameterization is not appropriate in a subarctic climate.
+Deciduous trees display also a notable increase in mean stomatal conductance.
 
@@
 The results displayed in Fig.~\ref{fig:pody_rel} allow for an assessment of over- or underestimation of ozone risk arising from the use of MM parameterizations established from less climatically extreme conditions.
-Critical levels for reduction in biomass \citep[deciduous forest $4\,\unit{\%}$, coniferous forest $2\,\unit{\%}$, and grasslands $10\,\unit{\%}$ biomass reduction,][]{ICP:MappingManual2017,ESPR:Hayes2021} are shown as horizontal dashed lines.
+Critical levels for reduction in biomass \citep[deciduous forest $4\,\unit{\%}$, coniferous forest $2\,\unit{\%}$, and grasslands $10\,\unit{\%}$ biomass reduction adopted from][]{ICP:MappingManual2017,ESPR:Hayes2021} are shown as horizontal dashed lines.
 The different background colors indicate the temperature response function localization.
 The light response function localization is denoted by $+$ (PPFD1.2) and $-$ (PPFD0.8).
-Circles indicate simulations with localized GS, canopy height, and leaf dimensions and squares simulations with MM GS.
-With open symbols the effect on $\mathrm{POD_y}$ of taking SWP into account is displayed.
-
-We identify that an increase in cold tolerance (MM $\rightarrow$ cold $\rightarrow$ subarctic) leads to higher ozone uptake in both probed years, while the usage of locally adapted GS causes a reduction.
-The earlier and longer GS in 2019 leads to increased ozone uptake compared to 2018 despite more episodes of elevated \chem{[O_3]} in 2018 (see Section~\ref{subsec:anomalies} for details).
+Circles indicate simulations with localized GS, canopy height, and leaf dimensions, whereas squares indicate simulations with MM GS.
+With open symbols, we'll display the effect on $\mathrm{POD_y}$ if SWP is taken into account.
+
+We identified that an increasing cold tolerance (MM $\rightarrow$ cold $\rightarrow$ subarctic) leads to higher ozone uptake in both probed years, while using locally adapted GS causes a reduction.
+Therefore, an earlier and longer GS in 2019 leads to an increased ozone uptake compared to 2018 despite more episodes of elevated \chem{[O_3]} in 2018 (see Section~\ref{subsec:anomalies} for details).
+
+An earlier opening of stomata at low light intensities (PPFD0.8) will cause an increase in $\mathrm{POD_1}$ while a later opening (PPFD1.2) will reduce $\mathrm{POD_1}$.
+The conditions in 2018 reduce $\mathrm{POD_1}$ more strongly for the PPDF0.8 light response function localization indicating an increase in sensitivity to drought stress (compare PPFD0.8 and PPFD1.2 in Fig.~\ref{fig:pody_rel}a, subarctic, 2018, with SWP effect on).
 Due to the shape of $f_\mathrm{light}$, a symmetric variation leads to an asymmetric response in $\mathrm{POD_1}$.
-We find that an opening of stomata at lower light intensities will cause a reduction of $\mathrm{POD_1}$ due to drought conditions.
-This is seen for both $f_\mathrm{T}$ localization in 2018 but not 2019 (compare PPFD0.8 and PPFD1.2 in Fig.~\ref{fig:pody_rel}a, subarctic, 2018, with SWP effect on).
 
@@
 Low soil water potential reduced the $\mathrm{POD_1}$ values in 2018 in some cases when using the MM GS, but did not or only slightly when using the localized GS.
-This indicates (frost) drought conditions early in the regular GS in 2018 when the ground was still covered in snow.
+This indicates (frost) drought conditions early in the regular GS in 2018 when the ground was still covered in snow in reality.
 The difference in $\mathrm{POD_1}$ due to the localization of $f_\mathrm{T}$ for the localized GS ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
-The maximum difference due to the $f_\mathrm{light}$ localization for the PPFD0.8 (superscript) and PPFD1.2 (subscript) amount to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.
+The maximum difference in $\mathrm{POD_1}$ due to the $f_\mathrm{light}$ localization for the PPFD0.8 (superscript) and PPFD1.2 (subscript) amount to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.
 
 For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
-The MM Boreal Norway spruce parameterization was already well acclimated to the prevailing temperatures and therefore differs the least from our bespoke parameterizations.
-Taking the bespoke GS and canopy height into account, we find a higher ozone uptake in 2019 than in 2018.
-The overall difference determined from bespoke GS for subarctic and cold parameterizations with respect to MM range between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
-The maximum deviations determined from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of $f_\mathrm{light}$ and amounts to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.
-
-For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations with bespoke GS stay below the CL.
-Perennial grassland shows the smallest ozone uptake, but a similarly large response to the temperature parameterizations as deciduous trees.
-Again, we find a reversal for predicted ozone risk in $\mathrm{POD_1}$ between 2018 and 2019 for the subarctic type.
-Perennial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant without bespoke GS.
-The overall difference determined from bespoke GS for subarctic and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
-The lower maximum temperature tolerance in the subarctic compared to the cold temperature parameterization probably caused the lower $\mathrm{POD_1}$ values calculated for 2018 in subarctic compared to cold climate, whereas this was not seen for 2019.
-The maximum deviations determined from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of $f_\mathrm{light}$ amount to $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
-
-From the difference between 2018 and 2019, we conclude that the interannual variability of ozone dose is non-negligible.
-The maximum difference in $\mathrm{POD_1}$ between simulations with the MM parameterizations and the localized temperature and light response functions is of the same order of magnitude as the interannual difference.
-The magnitude of all described effects on $\mathrm{POD_1}$ varies between PFTs as well as years, but the predicted ozone uptake for the localized $f_\mathrm{T}$ is always larger than for the MM parameterizations.
-This suggests underestimation of ozone risk for subarctic species when using the MM parameterizations.
+The MM Boreal Norway spruce parameterization was already well acclimated to the prevailing temperatures and therefore differs the least from our localized parameterizations.
+Taking the localized GS and canopy height into account, we find a higher ozone uptake in 2019 than in 2018.
+The maximum difference in $\mathrm{POD_1}$ determined between localized GS for subarctic and cold parameterizations with respect to MM ranges between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+The maximum difference determined from PPFD0.8 (superscript) and PPFD1.2 (subscript) $f_\mathrm{light}$ amounts to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.
+
+For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations with localize GS stay below the CL.
+Perennial grassland shows the smallest ozone dose, but a similarly large response to the temperature parameterization localizations as deciduous trees.
+Again, we find a reversal for predicted ozone risk in 2018 and 2019 for the subarctic type.
+Perennial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant without localized GS (see above).
+The overall difference in $\mathrm{POD_1}$ determined from the localized GS for subarctic and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+The lower maximum temperature tolerance probably caused the lower $\mathrm{POD_1}$ values calculated for 2018 using the subarctic compared to cold localization, whereas this was not seen for 2019.
+The maximum difference in $\mathrm{POD_1}$ determined from the PPFD0.8 (superscript) and PPFD1.2 (subscript) variation of $f_\mathrm{light}$ amount to $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
+
+The maximum difference in $\mathrm{POD_1}$ between simulations with the MM parameterizations and the localized temperature and light response functions is of the same order of magnitude as the interannual difference and should not be neglected.
+The magnitude of all described effects on $\mathrm{POD_1}$ varies between PFTs as well as years, but the predicted ozone dose for the localized $f_\mathrm{T}$ is always larger than for the MM parameterizations.
+This indicates a probable underestimation of ozone risk for subarctic species when using the MM parameterizations.
 
@@
 
-We estimated the biomass reductions in accordance to the CLs \citep{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme temperature parameterization (\emph{subarctic}) for each bespoke PFT.
+We estimated the biomass reductions in accordance to the CLs \citep{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme temperature parameterization localization (\emph{subarctic}) for each PFT.
 The estimates are listed in Table~\ref{tab:biomass_reduction}.
-The reported range corresponds to the maximum deviation defined by the $\pm 20\,\unit{\%}$ variation of $f_\mathrm{light}$ with SWP=on.
-The total biomass of deciduous trees and perennial grassland was substantially reduced in comparison with the default MM parameterizations in both years.
+The reported range corresponds to the maximum difference in $\mathrm{POD_1}$ defined by the $\pm 20\,\unit{\%}$ variation of $f_\mathrm{light}$ with SWP=on.
+The total biomass of deciduous trees and perennial grassland was notablly reduced in comparison with the MM parameterization in both years.
 The CL for perennial grassland is higher than for the other species.
@@
 Coniferous trees are the least affected and the magnitude of total biomass reduction is almost independent of the choice of the parameterization within given uncertainties.
-This may reflect that the temperature parameterization in the MM is rarely a limiting factor for stomatal opening.
+This may reflect that the temperature parameterization in the MM is rarely a limiting factor for stomatal conductance.
 
@@
 
-\subsection{Climate and weather conditions}
+\subsection{Climate and weather conditions in the growing season 2018/19}
 \label{sec:stats}
-Here, we put the conditions during the growing season 2018/19 into a climatological context.
-
-\subsubsection{Weather conditions in the growing season 2018/19}
-\label{subsec:weather}
-
+Here, we put the weather conditions during the growing season 2018/19 into a climatological context.
 All relevant data for the growing season 2018/19 are shown as time series in Fig.~\ref{fig:data_svanvik_2018_2019}.
@@
 
-As can be seen in Fig.~\ref{fig:data_svanvik_2018_2019}a, \chem{[O_3]} peaks in spring (April/May) and reaches its minimum in late summer (July/August).
+\subsubsection{Weather conditions}
+\label{subsec:weather}
+
+As can be seen in Fig.~\ref{fig:data_svanvik_2018_2019}a, \chem{\chi_{O_3}} peaks in spring (April/May) and reaches its minimum in late summer (July/August).
 The spring peak has not been captured completely in 2019, since data acquisition only started in late April.
-In summer 2018 (June--August), high ozone concentrations ($\chem{[O_3]} > 40\,\unit{ppb}$) were recorded 50 times.
-The highest summer ozone concentration ($\chem{[O_3]} = 50.2\,\unit{ppb}$) was measured on July 25.
-This coincides with the period of the most extensive forest fires in central Sweden which occurred from July 12--29 \citep{SOU2019}.
-However, due to the above-mentioned data acquisition problems, we missed most of the corresponding ozone data for this event and most likely also the peak \chem{[O_3]}.
-A method for gap-filling these data has been presented in \citet{ACP:Falk2021}.
-In contrast, \chem{[O_3]} only rose 18 times above the threshold of $40\,\unit{ppb}$ during the summer of 2019.
+In summer 2018 (June--August), high ozone concentrations ($\chem{\chi_{O_3}} > 40\,\unit{ppb}$) were recorded 50 times.
+The highest summer ozone VMR ($\chem{\chi_{O_3}} = 50.2\,\unit{ppb}$) was measured on July~25.
+This coincides with the period of the most extensive forest fires in central Sweden which occurred from July~12--29 \citep{SOU2019}.
+However, due to the above-mentioned data acquisition problems, we missed most of the corresponding ozone data for this event and most likely also the peak \chem{\chi_{O_3}}.
+A method for gap-filling these data has been presented in \citet{ACP:Falk2021} and explained in Section~\ref{subsec:gap_filling}.
+In contrast, \chem{\chi_{O_3}} only rose 18 times above the threshold of $40\,\unit{ppb}$ during the summer of 2019.
 
@@
 
-Climatologies of temperature and precipitation for Svanhovd were derived for the period 1992--2012 and are shown in a combined climate diagram (Fig.~\ref{fig:climatediagram}).
-Monthly averaged $2\,\unit{m}$ temperatures (red line) are displayed with a standard deviation as error band.
-Averaged monthly accumulated precipitation (blue bars) is shown without standard deviation.
-Months with average temperatures below freezing are denoted with a star.
-As can be seen, temperatures stay below freezing for 5 consecutive months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for K\"{o}ppen's climate classification of a regular subarctic climate (Dfc) \citep[e.g.][]{SD:Beck2018}.
+Climatologies (multi-year seasonal averages) of temperature, precipitation, and radiant energy $Q_\mathrm{e}$ for Svanhovd were derived for the period 1992--2012 and are shown as box plots in Fig.~\ref{fig:climatediagram}.
+Therein, the colored horizontal lines indicate the median and the triangle the mean value for the respective month.
+A box denotes the interquartile range (IQR).
+Observations exceeding 1.5~times the IQR (whiskers) are usually considered outliers but ought to be interpreted differently in a climate diagram, namely as events deviating extraordinarily from the norm.
+
+Temperatures usually stay below freezing for 5 consecutive months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for a K\"{o}ppen's climate classification of a regular subarctic climate (Dfc) \citep[e.g.][]{SD:Beck2018}.
 The highest monthly average temperature is $(13.1\pm 1.1)\,\unit{^\circ C}$ in July and the lowest $(-12.8\pm 2.0)\,\unit{^\circ C}$ in February.
-The coldest measured temperature was $-45.2\,\unit{^\circ C}$ in January 27, 1999, while the highest temperature ($29.4\,\unit{^\circ C}$) occurred on July 16 the same year.
+The coldest ever recorded temperature was $-45.2\,\unit{^\circ C}$ on January~27, 1999, while the highest temperature ($29.4\,\unit{^\circ C}$) occurred on July~16 the same year.
+As can be seen mean and median are well aligned in the center of the IQR for most months meaning that temperature anomalies are symmetrically distributed.
+
+The average accumulated monthly precipitation ($\sum_d\mathrm{Precip}$) indicates that winter and spring (November--April, except for March) are relatively dry ($\sum_d\mathrm{Precip} < 20\,\unit{mm}$).
+March precipitation is primarily snow and will influence the start of the growing season.
+The driest month is January with $\sum_d\mathrm{Precip} = (16.7\pm 3.0)\,\unit{mm}$.
+The most precipitation occurs in summer, with a $\sum_d\mathrm{Precip} = (58.5\pm 9.2)\,\unit{mm}$ in August.
+The average annual accumulated precipitation ($\sum_m\mathrm{Precip}$) given with standard error of mean is $(383\pm 86)\,\unit{mm}$.
+Mean and median precipitation are less well aligned compared to temperature and point to a considerable interannual variability.
+
+The climatology for global radiant energy $Q_e$ in \unit{J\,m^{-2}} (the integral of $Q_0$) is zero in November--January (polar night) and reaches its maximum in May--July (midnight sun conditions).
+Mean and median are well aligned.
+Only June displays a slightly skewed distribution towards higher $Q_e$.
+
+%Ozone
+The ozone VMR climatology (\chem{\chi_{O_3}}) (1986--1996) is shown in Figure~\ref{fig:climatediagram}d.
+The climatological spring peak occurs in April (around \unit{doy} 100) and amounts to $40\,\unit{ppb}$ while the annual average \chem{\chi_{O_3}} was $28\,\unit{ppb}$.
+The decline in \chem{\chi_{O_3}} coincides with the average beginning of \chem{CO_2} uptake by coniferous trees \citep{TB:Kolari2007, TP:Wallin2013}.
+Mean and median are well aligned and centered indicating a symmetric distribution for each month.
+The highest recorded \chem{\chi_{O_3}} occur in June most likely due to forest fires as mentioned earlier.
 
@@
 
-The average accumulated monthly precipitation ($\sum_d\mathrm{Precip}$) indicates that winter and spring (November--April, except for March) are relatively dry ($\sum_d\mathrm{Precip} < 20\,\unit{mm}$).
-March precipitation is primarily snow and will influence the start of the growing season.
-The driest month is January with $\sum_d\mathrm{Precip} = (16.7\pm 3.0)\,\unit{mm}$.
-The most precipitation occurs in the summer months, with a $\sum_d\mathrm{Precip} = (58.5\pm 9.2)\,\unit{mm}$ in August.
-The average annual accumulated precipitation ($\sum_m\mathrm{Precip}$) given with standard error of mean is $(383\pm 86)\,\unit{mm}$.
-Precipitation shows a considerable interannual variability.
-
-The climatology for global irradiance ($Q_0$, not shown) approaches zero in December/January (polar night) and reaches its maximum in June/July (midnight sun conditions).
-We determine the corresponding highest $Q_0$ to $800\,\unit{W\,m^{-2}}$.
-The maximum of the lowest $Q_0$ (including night) amounts to $200\,\unit{W\,m^{-2}}$ indicating that a considerable amount of light is available under midnight sun conditions.
+\subsubsection{2018/19 growing season anomalies}
+\label{subsec:anomalies}
+
+In climate science, a tool to compare specific years anomalies -- the deviation from the norm --  are used.
+In Figure~\ref{fig:anomalies_svanvik}, we show anomalies for both temperature, precipitation, radiant energy, and ozone VMR.
+We defined anomalies as median deviation from the climatological median.
+The hatched area indicates the respective IQR.
+
+Temperature anomalies (Fig.~\ref{fig:anomalies_svanvik}a) show enhanced temperatures through out summer (May--September) 2018 breaching the IQR two times while March was unusually cold.
+No notably enhanced temperatures occurred in 2019 indicating that 2019 was an average year.
+
+Similarly, the precipitation anomalies are displayed in Fig.~\ref{fig:anomalies_svanvik}b.
+March 2018 saw more precipitation than usual but mainly as snow.
+May and October were unusually dry while August was unusually wet.
+2019 was rather average throughout the year, but with unusually wet in June and dry in July.
+
+Radiant energy anomalies are presented in Fig.~\ref{fig:anomalies_svanvik}c.
+In summer 2018 (May/July), $Q_e$ was extraordinarily higher than usual, while 2019 was generally an average year with a rather dark spring (April/May) and late summer (August).
 
 %Ozone
-The daily mean ozone climatology (\chem{\left<[O_3]\right>}) (1986--1996) is shown in Figure~\ref{fig:climatediagram}d.
-Darker colors indicate higher probability to observe these values.
-On top of the density distributions, a $10\,\unit{days}$ average of daily mean (\chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}}) is displayed together with $1 \sigma$ uncertainties and SE, respectively.
-A spline was fitted through the data to guide the eye.
-The \chem{[O_3]} density distribution and the \chem{\left<[O_3]\right>_\mathrm{10\,\unit{d}}} display the expected, pronounced seasonal cycle.
-The spring peak occurs on average on day of the year (\unit{doy}) $100$ and amounts to $40\,\unit{ppb}$ while the annual average \chem{\left<[O_3]\right>} is $28\,\unit{ppb}$.
-The decline in \chem{\left<[O_3]\right>} coincides with the average beginning of \chem{CO_2} uptake by coniferous trees \citep{TB:Kolari2007, TP:Wallin2013}.
-In July--September (\unit{doy}~182--273), ozone is occasionally almost completely depleted.
-This hints to ozone uptake by vegetation even at low light intensities during midnight sun conditions in combination with stable planetary boundary layer conditions preventing mixing of ozone rich air.
-
-\subsection{2018/19 growing season anomalies}
-\label{subsec:anomalies}
-In the following, we look at the 2018/19 anomalies and discuss the differences between the two years.
-As weather extremes like in 2018 are likely to become more frequent with climate change, this will help to assess possible future changes in the vulnerability of subarctic vegetation to ozone.\\
-
-In Fig.~\ref{fig:anomalies_svanvik}a, temperature anomalies for Svanvik are shown as a percentage of days warmer or colder than $\pm 1\,\sigma$ from the climatology for each month.
-Negative deviations from climatology are displayed as a negative percentage.
-The annual positive/negative deviations are indicated in the respective corners (right upper/right lower).
-The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$).
-We find, that the summer of 2018 was significantly warmer than average.
-Especially in May and July more than $40\,\unit{\%}$ of the days were significantly warmer.
-A significant number of warmer days continued to occur in all months (August--November) except for October.
-March 2018 had many unusually cold days.
-The summer of 2019 was fairly average, however, more significantly colder days occurred in July, while August had both significantly colder and warmer days.
-April 2019 had significantly warmer than average days.
+We use the bias-corrected and cross-calibrated ozone climatology \citep{ACP:Falk2021} and assess the monthly significance of the ozone concentration anomalies in 2018/19.
+The resulting percentage of days above/below $\pm 1\,\sigma$ is displayed for each month in Fig.~\ref{fig:anomalies_svanvik}d.
+Note that missing bars for months in winter and spring are both due to the temporal extent of our data and anomalies not meeting the $1\sigma$ criteria.
+The star indicates the reconstructed data in July.
+In April/May 2018, ozone was significantly elevated, while no significant enhancement is found in 2019.
+In both years, ozone concentrations were significantly lower than average in November.
+These results suggest that stagnant meteorological conditions in April/May in combination with a late start of the growing season contributed more to elevated ozone than the extensive forest fires in Sweden that did not significantly enhance \chem{\chi_{O_3}} in 2018.
+
+In summary, 2018 had a significant number of days that were warmer, drier, and brighter than the climatology, while 2019 was a rather average year.
+Ozone was not significantly enhanced during the GS 2018/19, but a substantially higher number of peak \chem{\chi_{O_3}} were observed in 2018 than in 2019.
 
@@
 \end{figure*}
-
-Similarly, the precipitation anomalies for Svanvik are displayed in Fig.~\ref{fig:anomalies_svanvik}b.
-Due to the large interannual variability in observed precipitation, only a few days deviated significantly from the climatology on the $1 \sigma$ level.
-Therefore, we lowered the $\sigma$ level constraint to $\pm \frac{1}{2}\,\sigma$.
-The percentage of days wetter/drier than $\pm \frac{1}{2}\,\sigma$ are shown for each month.
-Positive deviation refers to days wetter than average and negative deviation to drier than average.
-The hatched area between the dashed lines indicates the expected percentage of values falling above/below $\pm\frac{1}{2}\,\sigma$ if a normal distribution is assumed ($30.8\,\unit{\%}$).
-Unlike the temperature anomalies, the picture is not as clear.
-While March 2018 had a significant number of days wetter than usual, summer and fall (May/July, September/October) had a significant number of days drier than average.
-2019 was rather average throughout the year, but with a significant number of days drier than normal in September/October and wetter in December.
-
-The global irradiance anomalies for Svanvik are presented in Fig.~\ref{fig:anomalies_svanvik}c.
-The percentage of days above or below $\pm 1\,\sigma$ from the climatological mean is displayed for each month.
-In summer 2018 (May/July), global irradiance showed a significant number of days with higher than average irradiance, while 2019 was generally an average year with a significant number of days darker in spring (April/May).
-
-%Ozone
-We use the bias-corrected and cross-calibrated ozone climatology \citep{ACP:Falk2021} and assess the monthly significance of the ozone concentration anomalies in 2018/19.
-The resulting percentage of days above/below $\pm 1\,\sigma$ is displayed for each month in Fig.~\ref{fig:anomalies_svanvik}d.
-Note that missing bars for months in winter and spring are both due to the temporal extent of our data and anomalies not meeting the $1\sigma$ criteria.
-The star indicates the reconstructed data in July.
-In April/May 2018, ozone was significantly elevated, while no significant enhancement is found in 2019.
-In both years, ozone concentrations were significantly lower than average in November.
-These results suggest that stagnant meteorological conditions in April/May in combination with a late start of the growing season contributed more to elevated ozone than the extensive forest fires in Sweden that did not significantly enhance \chem{[O_3]} in 2018.
-
-In summary, 2018 had a significant number of days that were warmer, drier, and brighter than the climatology, while 2019 was a rather average year.
-Ozone was not significantly enhanced during the GS 2018/19, but a substantially higher number of peak \chem{[O_3]} were observed in 2018 than in 2019.
 
@@
 Boreal wildfires are found to emit ozone precursors \chem{CO} ($\chem{[CO]}/\chem{[CO_2]} \propto 6-13\,\unit{\%}$) and VOCs ($\chem{[VOCs]}/\chem{[CO_2]} \propto 0.5-1.5\,\unit{\%}$) \citep{AE:Cofer1990}.
-Coincident peak \chem{[O_3]} is found in ozone monitoring data (Fig.~\ref{fig:data_svanvik_2018_2019}) in July.
+Coincident peak \chem{\chi_{O_3}} is found in ozone monitoring data (Fig.~\ref{fig:data_svanvik_2018_2019}) in July.
 It is very likely that elevated ozone, like during the 2003 drought period \citep{JGR:Solberg2018}, was promoted by a combination of various factors such as wildfires, reduced cloud cover (increased solar radiation), reduced dry deposition and turbulent mixing due to the stagnant weather conditions, and increased BVOC emissions.
@@
 However, ozone concentrations during the growing season were not significantly enhanced and \chem{POD_1} values were very similar for both years.
-This suggests that the overall length of the growing season is more crucial in risk assessment when using flux-based metrics than episodes of peak concentrations as these peak \chem{[O_3]} often coincide with environmental conditions that limit ozone uptake in the $\mathrm{DO_3SE}$ model.
+This suggests that the overall length of the growing season is more crucial in risk assessment when using flux-based metrics than episodes of peak concentrations as these peak \chem{\chi_{O_3}} often coincide with environmental conditions that limit ozone uptake in the $\mathrm{DO_3SE}$ model.
 
@@
 \item Our bespoke parameterizations for subarctic species are better suited to describe vegetation adapted to subarctic conditions for instance allowing for stomata opening during a larger part of the growing season,
-\item the parameterizations defined in the mapping manual for European regional risk assessment appear to not adequately capture the physiological responses of subarctic vegetation that are important determinants of \chem{[O_3]} susceptibility and are likely to underestimate ozone-damage risk, and
+\item the parameterizations defined in the mapping manual for European regional risk assessment appear to not adequately capture the physiological responses of subarctic vegetation that are important determinants of \chem{\chi_{O_3}} susceptibility and are likely to underestimate ozone-damage risk, and
 \item climatic conditions that are promoting a longer growing season are likely to increase the ozone-damage risk in flux-based metrics.
@@
 To demonstrate the necessary acclimation of the mapping manual parameters, we show the initial results of stomatal conductance at leaf-level ($G_\mathrm{sto}^\mathrm{leaf}$) and $\mathrm{POD_y}$ over \unit{doy} for both years and all PFTs (Fig.~\ref{fig:pody_mm_composit}).
-Observed \chem{[O_3]} is also indicated.
+Observed \chem{\chi_{O_3}} is also indicated.
 From Fig.~\ref{fig:pody_mm_composit}f) it is apparent that the mapping manual parameterized grassland would not have been able to grow in 2019.
@@
 A flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) has been chosen.
-\chem{[O_3]} are plotted on the same axis and scales as $G_\text{sto}^\text{leaf}$ but in units of $\unit{ppb}$.
+\chem{\chi_{O_3}} are plotted on the same axis and scales as $G_\text{sto}^\text{leaf}$ but in units of $\unit{ppb}$.
 (a, b) deciduous tree; (c, d) coniferous tree; (e, f) perennial grassland.

### 2022-04-24 VegOzone.tex
finish rewriting climatology section

@@
 %Ozone
-We use the bias-corrected and cross-calibrated ozone climatology \citep{ACP:Falk2021} and assess the monthly significance of the ozone concentration anomalies in 2018/19.
-The resulting percentage of days above/below $\pm 1\,\sigma$ is displayed for each month in Fig.~\ref{fig:anomalies_svanvik}d.
-Note that missing bars for months in winter and spring are both due to the temporal extent of our data and anomalies not meeting the $1\sigma$ criteria.
-The star indicates the reconstructed data in July.
-In April/May 2018, ozone was significantly elevated, while no significant enhancement is found in 2019.
-In both years, ozone concentrations were significantly lower than average in November.
-These results suggest that stagnant meteorological conditions in April/May in combination with a late start of the growing season contributed more to elevated ozone than the extensive forest fires in Sweden that did not significantly enhance \chem{\chi_{O_3}} in 2018.
-
-In summary, 2018 had a significant number of days that were warmer, drier, and brighter than the climatology, while 2019 was a rather average year.
-Ozone was not significantly enhanced during the GS 2018/19, but a substantially higher number of peak \chem{\chi_{O_3}} were observed in 2018 than in 2019.
+We used the bias-corrected and cross-calibrated ozone climatology \citep{ACP:Falk2021} to assess the ozone anomalies in 2018/19.
+Note that we have not recorded ozone during winter (refer to Section~\ref{subsec:weather}).
+In April/May 2018, ozone was unusually elevated and stayed higher than usual through the whole GS.
+The highest anomaly in 2019 occurred in September at the end of the GS.
+The forest fires in Sweden in July enhanced ozone notably at Svanhovd.
+
+In summary, 2018 was warmer, drier, and brighter than usual, while 2019 was a rather average year.
+Ozone was more enhanced in 2018 than 2019 meaning that the drought conditions in summer 2018 contributed to lower $\mathrm{POD_1}$ according to our $\mathrm{DO_3SE}$ modeling results.
 

### 2022-04-24 VegOzone.tex
Typo

@@
 
-Temperature anomalies (Fig.~\ref{fig:anomalies_svanvik}a) show enhanced temperatures through out summer (May--September) 2018 breaching the IQR two times while March was unusually cold.
+Temperature anomalies (Fig.~\ref{fig:anomalies_svanvik}a) show enhanced temperatures throughout summer (May--September) 2018 breaching the IQR two times while March was unusually cold.
 No notably enhanced temperatures occurred in 2019 indicating that 2019 was an average year.

### 2022-04-28 VegOzone.tex
Add TODO
Add central Europe to table
Update abstract

@@
   The unique vegetation of the subarctic region acclimatized to extremes of cold and midnight sun are likely to be at threat from the combined impacts of climate change and increasing ozone concentrations \chem{[O_3]}.
-The atmospheric and climatic characteristics of the subarctic are known to lead to pronounced peak \chem{[O_3]} in spring.
 To date, only a few studies assessed the response of subarctic vegetation to variations in climate and air pollution.
-This study looks to fill this knowledge gap by examining the uncertainty in risk assessments that are using more general boreal paramterizations.
-With no bespoke parameterization of stomatal response of subarctic vegetation to essential climate variables available, we propose an adjustment of existing, Boreal parameterizations, based on local climate condition.
+This study looks to fill this knowledge gap by examining the uncertainty in risk assessments that are using Boreal paramterizations.
+With no localized parameterization of stomatal response of subarctic vegetation to essential climate variables available, we propose an adjustment of existing, Boreal parameterizations, based on local climate condition.
 To this end, we vary temperature response functions according to the climatological temperature distribution within the growing season at our site.
-We evaluate the extent to which two recent years (2018 and 2019) deviate from climatic and \chem{[O_3]} norms and how these potentially more frequent future deviations may influence the projected ozone damage to subarctic vegetation.
+We evaluate the extent to which two recent years (2018 and 2019) deviate from climatic and \chem{[O_3]} norms and how these potentially more frequent future deviations may influence the projected ozone damage risk to subarctic vegetation.
 We find that 2018 was an anomalously warm and less cloudy year, particularly in spring and early summer.
@@
 We apply the integrated flux-metric Phytotoxic Ozone Dose (POD) to determine ozone risk and damage to vegetation as a function of \chem{[O_3]}, key environmental variables, and species-specific physiology.
-Our study suggests that using standard boreal parameterizations in assessments likely leads to underestimating the risk of ozone damage in this region.
-We find that local climate adjusted parameterizations of plant functional types (PFTs) for subarctic vegetation result in an ozone-induced biomass loss of $2.5$ to $17.4\,\unit{\%}$.
-For some species, this loss is up to $xx\,\unit{\%}$ larger than projected from standard boreal parameterizations used in general assessments.
+Our study suggests that using standard Boreal parameterizations in assessments likely leads to underestimating the risk of ozone damage in this region.
+We find that local climate adjusted parameterizations for subarctic vegetation of plant functional types (PFTs) result in an ozone-induced biomass loss of $2.5$ to $17.4\,\unit{\%}$.
+For some species, this loss is up to $55\,\unit{\%}$ larger than projected from Boreal parameterizations used in general assessments.
 Efforts should be targeted towards accurately defining subarctic vegetations' physiological response to essential climate variables.
@@
     \middlehline
-    Default$^b$ & -- & 135 & 257\\
+    Central Europe$^b$ & -- & 100 & 307\\
+    Latitude model ($70\,\unit{^\circ N}$)$^b$ & -- & 135 & 257\\
     \multirow{2}{*}{Deciduous tree} & 2018 & 129$^*$ & 261$^a$ \\
@@
 
+TODO: Compare with dry deposition velociteis from Finland.
+
 For our study, we conducted a similar bio-monitoring study at Svanhovd, a site with agrometeorological and air pollution monitoring, to assess whether local ozone concentrations were capable of inducing damage to vegetation growing under northern Fennoscandia conditions.

### 2022-04-28 VegOzone.tex
Rephrase abstract

@@
 This study looks to fill this knowledge gap by examining the uncertainty in risk assessments that are using Boreal paramterizations.
-With no localized parameterization of stomatal response of subarctic vegetation to essential climate variables available, we propose an adjustment of existing, Boreal parameterizations, based on local climate condition.
+With no parameterization of stomatal response of subarctic vegetation available, we propose an adjustment of existing Boreal parameterizations based on local climate condition.
 To this end, we vary temperature response functions according to the climatological temperature distribution within the growing season at our site.
-We evaluate the extent to which two recent years (2018 and 2019) deviate from climatic and \chem{[O_3]} norms and how these potentially more frequent future deviations may influence the projected ozone damage risk to subarctic vegetation.
-We find that 2018 was an anomalously warm and less cloudy year, particularly in spring and early summer.
-Higher than average \chem{[O_3]} was observed in April/May while frequent episodes of ozone volume mixing ratios (VMRs) above $40\,\unit{ppb}$ occurred in June--August.
+Thus, we refer to our approach as localization of stomatal response functions.
+With these, we apply the integrated flux-metric Phytotoxic Ozone Dose (POD) to determine ozone risk and damage to vegetation as a function of \chem{[O_3]}, key environmental variables, and species-specific physiology.
+In addition, we evaluate the extent to which two recent years (2018 and 2019) deviate from climatic and \chem{[O_3]} norms and find that 2018 was an anomalously warm and less cloudy year, particularly in spring and early summer.
+Higher than average \chem{[O_3]} was observed in April/May while frequent episodes of ozone volume mixing ratios (VMRs) exceeding $40\,\unit{ppb}$ occurred in June--August.
 These episodes are in part attributable to forest fires in the Northern Hemisphere and warmer and sunnier conditions.
-We apply the integrated flux-metric Phytotoxic Ozone Dose (POD) to determine ozone risk and damage to vegetation as a function of \chem{[O_3]}, key environmental variables, and species-specific physiology.
-Our study suggests that using standard Boreal parameterizations in assessments likely leads to underestimating the risk of ozone damage in this region.
-We find that local climate adjusted parameterizations for subarctic vegetation of plant functional types (PFTs) result in an ozone-induced biomass loss of $2.5$ to $17.4\,\unit{\%}$.
+In contrast, 2019 was an average year.
+Our modeling study suggests that using standard Boreal parameterizations in assessments likely leads to underestimating the risk of ozone damage in this region for some species.
+We find that localized parameterizations for subarctic vegetation result in an ozone-induced biomass loss of $2.5$ to $17.4\,\unit{\%}$.
 For some species, this loss is up to $55\,\unit{\%}$ larger than projected from Boreal parameterizations used in general assessments.

### 2022-04-28 VegOzone.tex
Move parts of text to methods section and rephrase

@@
 
-To estimate the potential ozone risk, we use a flux-based metric, also referred to as the Phytotoxic Ozone Dose over a threshold y ($\mathrm{POD_y}$).
-Flux-based metrics, that relate the accumulated dose of ozone to vegetation damage, are biologically more relevant for risk assessments than exposure-based metrics because they can account for particular species characteristics (i.e.
+To estimate the potential ozone risk on subarctic vegetation, we use a flux-based metric, also referred to as the Phytotoxic Ozone Dose over a threshold y ($\mathrm{POD_y}$).
+Flux-based metrics, that relate the accumulated dose of ozone to a reduction in biomass, are biologically more relevant for risk assessments than exposure-based metrics because they can account for particular species characteristics (i.e.
 physiology and phenology) as well as key environmental variables that can decouple the relationship between ozone concentration, ozone uptake, and consequent damage \citep{PT:Emberson2020}.
@@
 But no bespoke parameterization for subarctic vegetation is available.
+
+The $\mathrm{POD_y}$ metric is currently used in risk assessments under the United Nations Economic Commission for Europe (UNECE) Long Range Transboundary Air Pollution (LRTAP) Convention to identify those locations across Europe where vegetation is at risk from ozone.
+This convention aims to develop an effects-based emission reduction policy that can target those ozone precursor emissions that are most influential in causing damage.
+To achieve this, the concept of critical levels (CLs) has been adopted and applied \citep{Maas2016}.
+Exceedance of the CL is used to identify those areas across Europe that would benefit from targeted Europe-wide emission reductions.
+Methods to estimate $\mathrm{POD_y}$ and CLs are defined by the UNECE LRTAP Convention and described in a Mapping Manual \citep[most recent version]{ICP:MappingManual2017}.
+The exceedance of the CL is calculated by:
+%
+\begin{equation}
+  \mathrm{CL_{exeed}} = \mathrm{POD_y} - \mathrm{CL}.
+\end{equation}
+%
+
+An application of the Mapping Manual method was made by \citet{GCB:Mills2011} where exceedance of the CL was related to observations of ozone damage to a clover bio-monitor.
+This study demonstrated the improved ability of flux-based metrics (in comparison to exposure-based metrics) to identify the geographical distribution of the risk of ozone damage and also showed that the ozone damage could extend into more northerly regions of Europe.
+
+The described methods for risk assessment rely on accurate representation of plant physiological responses to key environmental variables and are most often parameterized using data for less extreme climates, i.e., temperate, continental, or Boreal climates.
+Parameterization for more extreme bio-geographical regions such as subarctic climates has not been performed and relies on assuming that less extreme Boreal parameterization will be applicable.
+It is important to understand the implications of using these in the subarctic.
+In our study, we explore how risk might be over- or underestimated using Boreal parameterization and explore an adjustment of temperature and light response functions based on climate data.
+This allows an understanding of which are the key aspects of modelled plant physiology that might determine the potential risk and also how the threat from ozone may change in the future.
+In addition, our results may hint a solution to solve long-standing issues of earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface models and display over- or underproductivity in the Arctic \citep{GMD:Poulter2015,JAMES:Lawrence2019}.\\
+
+In Section~\ref{sec:dam}, we give a brief summary of the key components and  drivers of the $\mathrm{DO_3SE}$ model that is used to assess the ozone risk.
+To study the influence of selected parameters on ozone risk in the subarctic, we propose a method to localize temperature response functions based on climate data and start and end of the growing season from satellite observed gross primary productivity (GPP).
+For the later discussion of results, we summarize climate and weather conditions of the 2018/19 growing season at our research site in Section~\ref{sec:stats}.
+Results of the localization, the ozone risk modelling, and the climate perspective are presented in Section~\ref{sec:results}.
+In Section~\ref{sec:conc}, we discuss these results concerning prevailing conditions and give an outlook on how future assessments of pollution risk in the subarctic could be improved.
+
+%%
+\section{Methods}
+\label{sec:dam}
+In the following, we briefly describe the model used for assessing the ozone risk, its key components, input data, and gap filling procedure (Section~\ref{sec:do3se}).
+To study the impact of model parameters, we propose a method for parameter localization to a subarctic climate based on existing Boreal parameterizations and local climate conditions and estimate the start and end of the growing season from satellite derived GPP (Section~\ref{sec:localization}).
+
+\subsection{DO3SE model}
+\label{sec:do3se}
+In our investigation of the ozone risk in the subarctic, we use the well established Deposition of Ozone for Stomatal Exchange ($\mathrm{DO_3SE}$) model \citep{ACP:Bueker2012} which computes estimates of the total and stomatal deposition of \chem{O_3} to selected land-cover types and plant species according to UNECE Long-Range Transboundary Air Pollution (LRTAP) methodologies for effects-based risk assessment.
+
+\subsubsection{Model description}
+\label{subsec:model_describtion}
+
 An estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy $h$ represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf and key environmental variables affecting stomatal conductance.
@@
 $\mathrm{POD_y}$ is given in \unit{mmol\,m^{-2}} integrated over the growing season.
-
-The $\mathrm{POD_y}$ metric is currently used in risk assessments under the United Nations Economic Commission for Europe (UNECE) Long Range Transboundary Air Pollution (LRTAP) Convention to identify those locations across Europe where vegetation is at risk from ozone.
-This convention aims to develop an effects-based emission reduction policy that can target those ozone precursor emissions that are most influential in causing damage.
-To achieve this, the concept of critical levels (CLs) has been adopted and applied \citep{Maas2016}.
-Exceedance of the CL is used to identify those areas across Europe that would benefit from targeted Europe-wide emission reductions.
-Methods to estimate $\mathrm{POD_y}$ and CLs are defined by the UNECE LRTAP Convention and described in a Mapping Manual \citep[most recent version]{ICP:MappingManual2017}.
-The exceedance of the CL is calculated by:
-%
-\begin{equation}
-  \mathrm{CL_{exeed}} = \mathrm{POD_y} - \mathrm{CL}.
-\end{equation}
-%
-
-An application of the Mapping Manual method was made by \citet{GCB:Mills2011} where exceedance of the CL was related to observations of ozone damage to a clover bio-monitor.
-This study demonstrated the improved ability of flux-based metrics (in comparison to exposure-based metrics) to identify the geographical distribution of the risk of ozone damage and also showed that the ozone damage could extend into more northerly regions of Europe.
-
-The described methods for risk assessment rely on accurate representation of plant physiological responses to key environmental variables and are most often parameterized using data for less extreme climates, i.e., temperate, continental, or Boreal climates.
-Parameterization for more extreme bio-geographical regions such as subarctic climates has not been performed and relies on assuming that less extreme Boreal parameterization will be applicable.
-It is important to understand the implications of using these in the subarctic.
-In our study, we explore how risk might be over- or underestimated using Boreal parameterization and explore an adjustment of temperature and light response functions based on climate data.
-This allows an understanding of which are the key aspects of modelled plant physiology that might determine the potential risk and also how the threat from ozone may change in the future.
-In addition, our results may hint a solution to solve long-standing issues of earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface models and display over- or underproductivity in the Arctic \citep{GMD:Poulter2015,JAMES:Lawrence2019}.\\
-
-In Section~\ref{sec:dam}, we give a brief summary of the key components and  drivers of the $\mathrm{DO_3SE}$ model that is used to assess the ozone risk.
-To study the influence of selected parameters on ozone risk in the subarctic, we propose a method to localize temperature response functions based on climate data and start and end of the growing season from satellite observed gross primary productivity (GPP).
-For the later discussion of results, we summarize climate and weather conditions of the 2018/19 growing season at our research site in Section~\ref{sec:stats}.
-Results of the localization, the ozone risk modelling, and the climate perspective are presented in Section~\ref{sec:results}.
-In Section~\ref{sec:conc}, we discuss these results concerning prevailing conditions and give an outlook on how future assessments of pollution risk in the subarctic could be improved.
-
-%%
-\section{Methods}
-\label{sec:dam}
-In the following, we briefly describe the model used for assessing the ozone risk, its key components, input data, and gap filling procedure (Section~\ref{sec:do3se}).
-To study the impact of model parameters, we propose a method for parameter localization to a subarctic climate based on existing Boreal parameterizations and local climate conditions and estimate the start and end of the growing season from satellite derived GPP (Section~\ref{sec:localization}).
-
-\subsection{DO3SE model}
-\label{sec:do3se}
-In our investigation of the ozone risk in the subarctic, we use the well established Deposition of Ozone for Stomatal Exchange ($\mathrm{DO_3SE}$) model \citep{ACP:Bueker2012} which computes estimates of the total and stomatal deposition of \chem{O_3} to selected land-cover types and plant species according to UNECE Long-Range Transboundary Air Pollution (LRTAP) methodologies for effects-based risk assessment.
-
-\subsubsection{Model description}
-\label{subsec:model_describtion}
 

### 2022-04-28 VegOzone.tex
REphrase Intro

@@
 In Section~\ref{sec:dam}, we give a brief summary of the key components and  drivers of the $\mathrm{DO_3SE}$ model that is used to assess the ozone risk.
-To study the influence of selected parameters on ozone risk in the subarctic, we propose a method to localize temperature response functions based on climate data and start and end of the growing season from satellite observed gross primary productivity (GPP).
-For the later discussion of results, we summarize climate and weather conditions of the 2018/19 growing season at our research site in Section~\ref{sec:stats}.
-Results of the localization, the ozone risk modelling, and the climate perspective are presented in Section~\ref{sec:results}.
+To study the influence of selected parameters on ozone risk in the subarctic, we propose a method to localize temperature response functions based on climate data and to infer start and end of the growing season from satellite observed gross primary productivity (GPP).
+Results of the ozone risk modelling are presented in Section~\ref{sec:results}.
+For evaluating these results, we assess the climate anomaly of weather conditions in the 2018/19 GS at our research site in Section~\ref{sec:stats}.
 In Section~\ref{sec:conc}, we discuss these results concerning prevailing conditions and give an outlook on how future assessments of pollution risk in the subarctic could be improved.

### 2022-04-28 VegOzone.tex
Small rephrasing

@@
 In the following, we briefly describe the model used for assessing the ozone risk, its key components, input data, and gap filling procedure (Section~\ref{sec:do3se}).
-To study the impact of model parameters, we propose a method for parameter localization to a subarctic climate based on existing Boreal parameterizations and local climate conditions and estimate the start and end of the growing season from satellite derived GPP (Section~\ref{sec:localization}).
+To study the impact of model parameters, we propose a method for parameter localization to a subarctic climate based on existing Boreal parameterizations and local climate conditions.
+We estimate the start and end of the growing season from satellite derived GPP (Section~\ref{sec:localization}).
 
@@
 \label{sec:do3se}
-In our investigation of the ozone risk in the subarctic, we use the well established Deposition of Ozone for Stomatal Exchange ($\mathrm{DO_3SE}$) model \citep{ACP:Bueker2012} which computes estimates of the total and stomatal deposition of \chem{O_3} to selected land-cover types and plant species according to UNECE Long-Range Transboundary Air Pollution (LRTAP) methodologies for effects-based risk assessment.
+We use the well established Deposition of Ozone for Stomatal Exchange ($\mathrm{DO_3SE}$) model \citep{ACP:Bueker2012} which computes estimates of the total and stomatal deposition of \chem{O_3} to selected land-cover types and plant species according to UNECE Long-Range Transboundary Air Pollution (LRTAP) methodologies for effects-based risk assessment.
 
@@
 
-The opening of stomata or stomatal conductance $g_\mathrm{sto}$ is governed by different species-specific factors and response to key environmental variables.
-In the $\mathrm{DO_3SE}$ model, it is described by a multiplicative empirical model \citep{PTRS:Jarvis1976,Emberson2000,ICP:MappingManual2017}:
+The opening of stomata referred to as stomatal conductance $g_\mathrm{sto}$ is governed by different species-specific factors and responds to key environmental variables.
+In the $\mathrm{DO_3SE}$ model, $g_\mathrm{sto}$ is described by a multiplicative empirical model \citep{PTRS:Jarvis1976,Emberson2000,ICP:MappingManual2017}:
 \begin{equation}
@@
 The leaf phenology factor ($f_\mathrm{phen}$) is a stepwise defined linear function that allows for an increase of stomatal conductance in spring (start of growing season) and a decrease in fall (end of growing season).
-Mid anthesis and mid season dips in stomata conductance can also be accounted for but are not relevant in the subarctic.
+Mid anthesis and mid season dips in stomata conductance can also be accounted for but are not relevant in this study.
 
@@
 with an ozone climatology (multiannual norm) \chem{\left<[O_3]\right>} and ozone anomalies (deviation from the norm) \chem{\Delta[O_3]}.
-We computed a generalized climatology for northern Fennoscandia from long-term records of different close by sites and for our site.
-We determined the station with the highest correlation coefficient compared to our site.
-For this site, we calculated the anomalies with respect to the generalized climatology for the gap period.
-These anomalies were than added to the climatology of our site and corrected for an increase in present-day tropospheric background ozone compared to the early 1990s.
-For more details refer to \citet{ACP:Falk2021}.
+\citet{ACP:Falk2021} computed a generalized climatology for northern Fennoscandia from long-term records of different close by sites and determined the station with the highest correlation coefficient compared to the site with the gap.
+For this site, the anomalies with respect to the generalized climatology were calculated for the gap period.
+These anomalies were than added to the climatology of the site with the gap and corrected for an increase in present-day tropospheric background ozone compared to the early 1990s.
+For more details and a comparison with regional air quality models refer to \citet{ACP:Falk2021}.
 
@@
 \end{equation}
-with saturation vapor pressure $P_s(T)$ in \unit{hPa} from Arden~Buck~equation~\citep{JAP:Buck1981, Buck2012}.
-
-As site for our study, we chose Svanhovd in northern Norway ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m\,a.s.l.}$).
-Reasons for our choice: (1) Ambient ozone concentration had been monitored there in the past, (2) an agrometeorological station is measuring, e.g.
+with saturation vapor pressure $P_s(T)$ in \unit{hPa} from the Arden~Buck~equation~\citep{JAP:Buck1981, Buck2012}.
+
+For our study, we chose Svanhovd in northern Norway ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m\,a.s.l.}$).
+Reasons for our choice were: (1) Ambient ozone concentration had been monitored there in the past, (2) an agrometeorological station is measuring, e.g.
 $2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths, and (3) a national surveillance of air pollution (\chem{NO_x}, \chem{SO_x}, particular matter (PM), radiological species) is ongoing.

### 2022-04-28 VegOzone.tex
Small rephrasing in Section 2

@@
 
-To determine $A_\text{start}$ and $A_\text{end}$ for coniferous trees, we used a MODIS (Aqua/Terra) retrieved GPP product \citep{MODIS_PSN} on a $1\times 1\,\unit{km}$ patch centered at Svanhovd.
+For coniferous trees, we used a MODIS (Aqua/Terra) retrieved GPP product \citep{MODIS_PSN} on a $1\times 1\,\unit{km}$ patch centered at Svanhovd.
 GPP over time follows approximately a second-order polynomial function (Fig.~\ref{fig:modis_gpp}).
@@
 $A_\text{end}$ amounts to \unit{doy}~261 and 274, respectively.
-These $A_\mathrm{end}$ value will be used for all PFTs alike.
+This $A_\mathrm{end}$ value will be used for all PFTs alike.
 The resulting growing season for coniferous trees in 2019 was one month longer than in 2018.
@@
 We found \unit{doy}~129 and 130, respectively.
-For both years, these dates coincide with the first snow-free day at the closest inland meteorological weather station at Øvre~Neiden (Sør-Varanger, NOR).
-These dates are very similar to the latitude model results.
+For both years, these dates coincide with the first snow-free day at the closest inland meteorological weather station at Øvre~Neiden (Sør-Varanger, NOR) and are very similar to the latitude model predictions.
 For perennial grassland, we assume a latency period of $1\,\unit{month}$ after snow-melt.
@@
 The clay:sand:silt content is estimated to range from 12\,\%:61\,\%:27\,\% at the top layer to 16\,\%:59\,\%:25\,\% in about $1\,\unit{m}$ depth (\href{https://soilgrids.org/}{soilgrids.org}, last accessed April 2022 for location $30.0334\,\unit{^\circ E}$, $69.4549\,\unit{^\circ N}$) Within the $\mathrm{DO_3SE}$ soil parameterization, the sandy loam texture describes the upper $60\,\unit{cm}$, where the plant roots are found, of this soil type best.
-We presume that no localization of the water vapor pressure deficite $f_\mathrm{VPD}$ and soil water potential $f_\mathrm{SWP}$ response functions is necessary for our species.
+
+We did not perform a localization of the water vapor pressure deficite $f_\mathrm{VPD}$ and soil water potential $f_\mathrm{SWP}$ response functions.
 But acknowledge that these are very important factors in the modeling of ozone dose as pointed out by~\citet{ACP:Bueker2012}.
@@
 ImageJ (\href{https://imagej.nih.gov/ij/}{see webpage}) was used for thresholding the images to give silhouettes of the leaves.
-After scaling the images, the program was used for finding the minimum Feret diameter.
+After scaling the images, the program was used to find the minimum Feret diameter.
 This measure finds the largest width of a leaf at a $90^\circ$ angle to the length of the leaf (the Feret diameter).
@@
 Timmerman, personal communication, June 2020).
-We used $13.5\,\unit{m}$ for both deciduous and coniferous trees as this height is sampled from a larger and more diverse population of species.
+We used $13.5\,\unit{m}$ for both deciduous and coniferous trees as this height is found from a larger and more diverse population of species.
 

### 2022-04-28 VegOzone.tex
Rephrase

@@
 \label{sec:results}
-In this section, we first compute the average stomatal conductance $\left<g_\mathrm{sto}\right>$ relative to $g_\mathrm{max}$ for each PFT from the long term meteorological data according to Eq.~(\ref{eq:stomatal})and compare the different localizations with MM (Section~\ref{subsec:res_local}).
-We use the $\mathrm{DO_3SE}$ model to compute the ozone dose and projected biomass loss due to ozone for the two recent GS 2018/19 for each species and localization (Section~\ref{subsec:do3se_results}).
+In this section, we first compute the average stomatal conductance $\left<g_\mathrm{sto}\right>$ relative to $g_\mathrm{max}$ for each PFT from the long term meteorological data according to Eq.~(\ref{eq:stomatal}) and compare the different localizations to MM (Section~\ref{subsec:res_local}).
+We use the $\mathrm{DO_3SE}$ model to compute the ozone dose and projected biomass loss due to ozone for the two recent GSs 2018/19 for each species and localization (Section~\ref{subsec:do3se_results}).
 Finally, we put the results into a climatological perspective (Section~\ref{subsec:climatologies}).
@@
 
-To compare our localized parameterizations, we propose the following metric.
-From the long term of meteorological data, we select to distinct periods noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and morning ($5-9\,\unit{am}$ local time).
-For these, we calculate the average relative stomatal conductance $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal}) and assume $f_\mathrm{SWP}=1$).
-We presume that $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ around noon (the highest light intensity) is a good proxy for \chem{CO_2} uptake efficiency (growth).
+We propose the following metric.
+From the long term of meteorological data, we select two distinct periods: noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and morning ($5-9\,\unit{am}$ local time) and calculate the average relative stomatal conductance $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal}) assuming $f_\mathrm{SWP}=1$).
+Presumably, $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ around noon (the highest light intensity) is a good proxy for \chem{CO_2} uptake efficiency (growth).
 In addition, a small deviation between noon and morning and a low standard deviation indicates higher robustness of this localization to variability in growing conditions.
-From left to right in Fig.~\ref{fig:javis_func_opt_mean}, results using the subarctic, cold, and MM $f_\mathrm{T}$ are shown marked by the alternating background color.
+From left to right in Fig.~\ref{fig:javis_func_opt_mean}, results using the subarctic, cold, and MM $f_\mathrm{T}$ are shown marked by an alternating background color.
 Within each of these groups $f_\mathrm{light}$ categories are indicated by PPFD0.8 and PPFD1.2, respectively.
@@
 
-Based on these criteria, we identify \emph{subarctic}--PPFD0.8 as the parameterization with the highest score for all PFTs.
+Based on the above criteria, we identify \emph{subarctic}--PPFD0.8 as the parameterization with the highest score.
 The differences between the localized parameterizations and the MM are smallest for coniferous trees.
-This indicates that the MM Boreal parameterization is well suited also for the subarctic climate at Svanhovd.
+This indicates that the MM Boreal parameterization is well suited for the subarctic climate at Svanhovd.
 For perennial grassland, the differences compared to MM are largest suggesting that the MM Central European parameterization is not appropriate in a subarctic climate.
-Deciduous trees display also a notable increase in mean stomatal conductance.
+Deciduous trees display also a notable increase in mean stomatal conductance using our localizations.
 

### 2022-04-28 VegOzone.tex
Rephrase to include CE  GS logically

@@
 
-We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for the three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using the MM parameterization and our localized temperature (\emph{cold}, \emph{subarctic}) and light (PPFD0.8, PPFD1.2) parameterizations.
-
-The results displayed in Fig.~\ref{fig:pody_rel} allow for an assessment of over- or underestimation of ozone risk arising from the use of MM parameterizations established from less climatically extreme conditions.
+We modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per PLA for the three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using the MM parameterization and our localized temperature (\emph{cold}, \emph{subarctic}) and light (PPFD0.8, PPFD1.2) parameterizations.
+
+The results are comprehensively displayed in Fig.~\ref{fig:pody_rel}.
 Critical levels for reduction in biomass \citep[deciduous forest $4\,\unit{\%}$, coniferous forest $2\,\unit{\%}$, and grasslands $10\,\unit{\%}$ biomass reduction adopted from][]{ICP:MappingManual2017,ESPR:Hayes2021} are shown as horizontal dashed lines.
 The different background colors indicate the temperature response function localization.
-The light response function localization is denoted by $+$ (PPFD1.2) and $-$ (PPFD0.8).
-Circles indicate simulations with localized GS, canopy height, and leaf dimensions, whereas squares indicate simulations with MM GS.
-With open symbols, we'll display the effect on $\mathrm{POD_y}$ if SWP is taken into account.
+The light response function localizations are denoted by $+$ (PPFD1.2) and $-$ (PPFD0.8).
+Circles indicate simulations with localized GS, canopy height, and leaf dimensions, whereas squares indicate simulations with MM GS for Central Europe.
+We chose this to demonstrate the impact of a projected prolonged GS in the future.
+With open symbols, the effect on $\mathrm{POD_y}$ if SWP is taken into account is displayed.
 

### 2022-05-07 VegOzone.tex
Rephrase and mark passages that need rephrasing after new results

@@
 \begin{table}[t]
-  \caption{Start and end of growing season in \unit{doy}.}
+  \caption{Start and end of growing season in \unit{doy}. Central Europe is shown for comparison.}
   \label{tab:sensitivity_tests_gs}
@@
     Latitude model ($70\,\unit{^\circ N}$)$^b$ & -- & 135 & 257\\
-    \multirow{2}{*}{Deciduous tree} & 2018 & 129$^*$ & 261$^a$ \\
+    \multirow{3}{*}{Deciduous tree} & 2018 & 129$^*$ & 261$^a$ \\
     & 2019 & 130$^c$ & 274$^a$ \\
+    & 2100 & 116$^e$ & 286$^e$ \\
     \multirow{2}{*}{Coniferous tree} & 2018 & 122$^a$ & 261$^a$ \\
@@
     $^c$ $5\,\unit{days}-5\,\unit{^\circ C}$-rule, $T_\mathrm{air}$ from \href{seNorge.no}{seNorge.no};\\
-    $^d$ One month after snowmelt; reference station Øvre Neiden (Sør-Varanger, NOR)} % Table Footnotes
+    $^d$ One month after snowmelt; reference station Øvre Neiden (Sør-Varanger, NOR)\\
+    $^e$ Future projection based on linear regression of $^c$ (Fig.~\ref{appendix:growing_season})
+    } % Table Footnotes
 \end{table}
@@
 The light response function localizations are denoted by $+$ (PPFD1.2) and $-$ (PPFD0.8).
-Circles indicate simulations with localized GS, canopy height, and leaf dimensions, whereas squares indicate simulations with MM GS for Central Europe.
-We chose this to demonstrate the impact of a projected prolonged GS in the future.
-With open symbols, the effect on $\mathrm{POD_y}$ if SWP is taken into account is displayed.
-
-We identified that an increasing cold tolerance (MM $\rightarrow$ cold $\rightarrow$ subarctic) leads to higher ozone uptake in both probed years, while using locally adapted GS causes a reduction.
-Therefore, an earlier and longer GS in 2019 leads to an increased ozone uptake compared to 2018 despite more episodes of elevated \chem{[O_3]} in 2018 (see Section~\ref{subsec:anomalies} for details).
-
-An earlier opening of stomata at low light intensities (PPFD0.8) will cause an increase in $\mathrm{POD_1}$ while a later opening (PPFD1.2) will reduce $\mathrm{POD_1}$.
-The conditions in 2018 reduce $\mathrm{POD_1}$ more strongly for the PPDF0.8 light response function localization indicating an increase in sensitivity to drought stress (compare PPFD0.8 and PPFD1.2 in Fig.~\ref{fig:pody_rel}a, subarctic, 2018, with SWP effect on).
-Due to the shape of $f_\mathrm{light}$, a symmetric variation leads to an asymmetric response in $\mathrm{POD_1}$.
-
-For deciduous trees (Fig.~\ref{fig:pody_rel}a) all $\mathrm{POD_1}$ simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
-Among all PFTs, deciduous trees display the largest sensitivity in $\mathrm{POD_1}$.
-Compared to MM $\Delta\mathrm{POD_1}=(10-31)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
-In particular, the sensitivity to the $f_\mathrm{light}$ is pronounced (extent of $\angle(+,0,-)$).
-With MM GS, $\mathrm{POD_1}$ values were higher in 2018 than in 2019.
+Circles indicate simulations with localized GS, canopy height, and leaf dimensions, TODO: (whereas squares indicate simulations with MM GS for Central Europe.
+We chose this to demonstrate the impact of a projected prolonged GS in the future.) With open symbols, the effect on $\mathrm{POD_y}$ with SWP taken into account is displayed.
+
+We identified in general that an increasing cold tolerance represented by a shift in the temperature response function (MM $<$ cold $<$ subarctic) leads to an increase in $\mathrm{POD_1}$ in both probed years.
+TODO: (while using locally adapted GS causes a reduction.) An earlier start and overall longer GS in 2019 leads to higher $\mathrm{POD_1}$ compared to 2018 despite the occurrence of more episodes of elevated \chem{O_3} in 2018 than in 2019 (see Section~\ref{subsec:anomalies} for details).
+Taking into account SWP, $\mathrm{POD_1}$ was effectively only reduced for deciduous trees in 2018, while in all other cases simulation results with and without SWP effects were identical.
+An earlier opening of stomata at low light intensities (PPFD0.8) caused an increase in $\mathrm{POD_1}$ compared to MM while a later opening (PPFD1.2) reduced $\mathrm{POD_1}$.
+Due to the shape of $f_\mathrm{light}$, a symmetric variation can lead to an asymmetric response in $\mathrm{POD_1}$.
+
+For deciduous trees (Fig.~\ref{fig:pody_rel}a) all $\mathrm{POD_1}$ simulations exceed the CL by TODO: ($5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$) and display the largest spread in $\mathrm{POD_1}$ with respect to $f_\mathrm{light}$.
+In 2019, the lowest estimated $\mathrm{POD_1}$ (MM, PPFD1.2) amounts to $10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, while the highest estimate (subarctic, PPFD0.8) reaches about $20\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
+In 2018, this uncertainty range is much smaller ($(12-18)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
+In particular, the impact of $f_\mathrm{light}$ is pronounced (extent of $\angle(+,0,-)$).
+The maximum difference in $\mathrm{POD_1}$ due to the $f_\mathrm{light}$ localization for the PPFD0.8 (superscript) and PPFD1.2 (subscript) amount to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.
+The difference in $\mathrm{POD_1}$ compared to MM in response to the localization of $f_\mathrm{T}$ and GS ranges between $\Delta\mathrm{POD_1}=(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $\Delta\mathrm{POD_1}=(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+A low soil water potential caused a reduction in $\mathrm{POD_1}$ in 2018 in cases when using PPFD0.8.
+This indicates that drought conditions can reduce the ozone damage risk.
+TODO: (With MM GS, $\mathrm{POD_1}$ values were higher in 2018 than in 2019.
 However, taking the localized GS into account, this is reversed for the subarctic temperature response function.
-This can be explained by more favorable growing conditions in 2019.
-Low soil water potential reduced the $\mathrm{POD_1}$ values in 2018 in some cases when using the MM GS, but did not or only slightly when using the localized GS.
-This indicates (frost) drought conditions early in the regular GS in 2018 when the ground was still covered in snow in reality.
-The difference in $\mathrm{POD_1}$ due to the localization of $f_\mathrm{T}$ for the localized GS ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
-The maximum difference in $\mathrm{POD_1}$ due to the $f_\mathrm{light}$ localization for the PPFD0.8 (superscript) and PPFD1.2 (subscript) amount to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.
+This can be explained by more favorable growing conditions in 2019.)
 
 For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
-The MM Boreal Norway spruce parameterization was already well acclimated to the prevailing temperatures and therefore differs the least from our localized parameterizations.
-Taking the localized GS and canopy height into account, we find a higher ozone uptake in 2019 than in 2018.
+Our localization does not show much difference compared to MM indicating the the MM Boreal Norway spruce parameterization was already well-adjusted to the local subarctic climate.
+We find a higher $\mathrm{POD_1}$ in 2019 than in 2018.
 The maximum difference in $\mathrm{POD_1}$ determined between localized GS for subarctic and cold parameterizations with respect to MM ranges between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
@@
 For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations with localize GS stay below the CL.
-Perennial grassland shows the smallest ozone dose, but a similarly large response to the temperature parameterization localizations as deciduous trees.
+Perennial grassland shows the smallest $\mathrm{POD_1}$, but a similarly large response to the temperature parameterization localizations as deciduous trees.
 Again, we find a reversal for predicted ozone risk in 2018 and 2019 for the subarctic type.
-Perennial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant without localized GS (see above).
-The overall difference in $\mathrm{POD_1}$ determined from the localized GS for subarctic and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
-The lower maximum temperature tolerance probably caused the lower $\mathrm{POD_1}$ values calculated for 2018 using the subarctic compared to cold localization, whereas this was not seen for 2019.
+Perennial grassland shows the lowest sensitivity to the light threshold.
+TODO: (and SWP is only relevant without localized GS (see above).) The overall difference in $\mathrm{POD_1}$ determined from the localized GS for subarctic and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+The chosen lower maximum temperature in the subarctic parameterization probably caused the lower $\mathrm{POD_1}$ in 2018 (warm year) compared to the cold parameterization.
+This was not seen in 2019.
 The maximum difference in $\mathrm{POD_1}$ determined from the PPFD0.8 (superscript) and PPFD1.2 (subscript) variation of $f_\mathrm{light}$ amount to $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
 
-The maximum difference in $\mathrm{POD_1}$ between simulations with the MM parameterizations and the localized temperature and light response functions is of the same order of magnitude as the interannual difference and should not be neglected.
-The magnitude of all described effects on $\mathrm{POD_1}$ varies between PFTs as well as years, but the predicted ozone dose for the localized $f_\mathrm{T}$ is always larger than for the MM parameterizations.
-This indicates a probable underestimation of ozone risk for subarctic species when using the MM parameterizations.
+The maximum difference in $\mathrm{POD_1}$ between simulations with the MM parameterizations and the localized temperature and light response functions is of the same order of magnitude as the difference between the two probed years.
+The magnitude of all described effects on $\mathrm{POD_1}$ varies between PFTs as well as years, but is by construction always larger than estimated with the MM parameterization.
+This means that the ozone damage risk could be possibly underestimated even in normal years and establishing dedicated subarctic parameterizations are a must to properly assess ozone damage risk in these regions.
 
@@
 We estimated the biomass reductions in accordance to the CLs \citep{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme temperature parameterization localization (\emph{subarctic}) for each PFT.
-The estimates are listed in Table~\ref{tab:biomass_reduction}.
-The reported range corresponds to the maximum difference in $\mathrm{POD_1}$ defined by the $\pm 20\,\unit{\%}$ variation of $f_\mathrm{light}$ with SWP=on.
-The total biomass of deciduous trees and perennial grassland was notablly reduced in comparison with the MM parameterization in both years.
-The CL for perennial grassland is higher than for the other species.
-From an agro-economical point of view in our focus area, where the grass is cut only once a year, a loss of biomass of more than $10\,\unit{\%}$ can have severer consequences compared to more productive areas in central Europe and may need adjustment.
-
-Coniferous trees are the least affected and the magnitude of total biomass reduction is almost independent of the choice of the parameterization within given uncertainties.
-This may reflect that the temperature parameterization in the MM is rarely a limiting factor for stomatal conductance.
+The estimates for SWP=off are listed in Table~\ref{tab:biomass_reduction}.
+We defined an uncertainty range from the difference in biomass reduction caused by the variation of $f_\mathrm{light}$.
+Superscript refers to PPFD0.8 and subscript to PPFD1.2.
+The reported range corresponds to the effect of taking SWP into account or not.
+As SWP reduces $\mathrm{POD_1}$, biomass reduction is decreased when taking SWP into account.
+The biomass reduction ranges between $2.5\,\unit{\%}$ (coniferous, 2018) and $17.4\,\unit{\%}$ (deciduous, 2019).
+For deciduous trees this amounts to a reduction by up to $55\,\unit{\%}$ compared to MM.
+From an agro-economical point of view in our focus area, where the grass is cut only once a year, a loss of biomass of more than $10\,\unit{\%}$ can have severer consequences compared to more productive areas in central Europe.
+Therefore, CL may also need localization.
+Coniferous trees are the least affected PFT and the magnitude of total biomass reduction is almost independent of the choice of the parameterization within given uncertainties.
 
 \begin{table}[t]
-  \caption{Estimated total biomass reduction in \unit{\%} for temperature acclimation \emph{subarctic} with bespoke GS and SWP=off.
+  \caption{Estimated total biomass reduction in \unit{\%} for temperature acclimation \emph{subarctic} with localized GS and SWP=off.
 The uncertainty ranges reported here correspond to maximum a divergence deducted from varying $f_\mathrm{light}$ and SWP=on.
-For comparison, the corresponding biomass reduction for the generic default MM averaged over both years is shown.
-The reported standard deviation is computed from all sensitivity simulations ($\pm 20\,\unit{\%}$ in extent of stomatal opening at low light, SWP on/off, and bespoke GS).}
+For comparison, the corresponding biomass reduction for the MM parameterization averaged over both years is shown.
+In this case, standard deviation is computed from all sensitivity simulations ($\pm 20\,\unit{\%}$ in extent of stomatal opening at low light, SWP on/off, and localized GS).}
   \label{tab:biomass_reduction}

### 2022-05-07 VegOzone.tex
Rephrase climate, weather and anomalies

@@
 Hourly averaged $2\,\unit{m}$ temperatures above $20\,\unit{^\circ C}$ occurred more regularly in July 2018 than in 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}b).
-In 2018, spring temperature regularly rose above freezing only in May, while in 2019 this occurred already early in March/April.
+In 2018, spring temperature regularly rose above freezing only in May, while in 2019 this occurred already early in March/April indicating a later start of the GS in 2018.
 More rain events with accumulated daily precipitation ($\sum_d \mathrm{Precip}$) above $10\,\unit{mm}$ occurred in the summer of 2018 compared to 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}c).
@@
 
-Climatologies (multi-year seasonal averages) of temperature, precipitation, and radiant energy $Q_\mathrm{e}$ for Svanhovd were derived for the period 1992--2012 and are shown as box plots in Fig.~\ref{fig:climatediagram}.
-Therein, the colored horizontal lines indicate the median and the triangle the mean value for the respective month.
+Climatologies (typically multi-year monthly averages) of temperature, precipitation, and radiant energy $Q_\mathrm{e}$ for Svanhovd were derived for the period 1992--2012 and are shown as box plots in Fig.~\ref{fig:climatediagram}.
+Therein, the colored horizontal lines indicate the median and the triangles the mean value for the respective month.
 A box denotes the interquartile range (IQR).
@@
 The climatological spring peak occurs in April (around \unit{doy} 100) and amounts to $40\,\unit{ppb}$ while the annual average \chem{\chi_{O_3}} was $28\,\unit{ppb}$.
-The decline in \chem{\chi_{O_3}} coincides with the average beginning of \chem{CO_2} uptake by coniferous trees \citep{TB:Kolari2007, TP:Wallin2013}.
+The decline in \chem{\chi_{O_3}} roughly coincides with the beginning of \chem{CO_2} uptake by coniferous trees \citep{TB:Kolari2007, TP:Wallin2013}.
 Mean and median are well aligned and centered indicating a symmetric distribution for each month.
-The highest recorded \chem{\chi_{O_3}} occur in June most likely due to forest fires as mentioned earlier.
+The highest recorded \chem{\chi_{O_3}} occur in June most likely due to forest fires as mentioned above.
 
@@
   \caption{Climate diagram based on meteorological data for Svanvik/Pasvik (1992--2012).
-Ozone climatology based on ozone monitoring data at NIBIO Environment Centre Svanhovd(1986--2003).
+Ozone climatology based on ozone monitoring data (1986--1996).
 A box indicates the upper/lower quartile, whiskers the $1.5\times$ Interquartile Range (IQR), and circles outliers.
@@
 May and October were unusually dry while August was unusually wet.
-2019 was rather average throughout the year, but with unusually wet in June and dry in July.
+2019 was rather average throughout the year, but unusually wet in June and dry in July.
 
@@
 In summary, 2018 was warmer, drier, and brighter than usual, while 2019 was a rather average year.
-Ozone was more enhanced in 2018 than 2019 meaning that the drought conditions in summer 2018 contributed to lower $\mathrm{POD_1}$ according to our $\mathrm{DO_3SE}$ modeling results.
+Ozone was more enhanced in 2018 than 2019.
+This means that the drought conditions in summer 2018 contributed more strongly to lower $\mathrm{POD_1}$ than elevated $\chem{\chi_{O_3}}$ in our $\mathrm{DO_3SE}$ modeling results.
 

### 2022-05-07 VegOzone.tex
Move conclusions
Move text to conclusions and rephrase

@@
 For deciduous trees this amounts to a reduction by up to $55\,\unit{\%}$ compared to MM.
-From an agro-economical point of view in our focus area, where the grass is cut only once a year, a loss of biomass of more than $10\,\unit{\%}$ can have severer consequences compared to more productive areas in central Europe.
-Therefore, CL may also need localization.
 Coniferous trees are the least affected PFT and the magnitude of total biomass reduction is almost independent of the choice of the parameterization within given uncertainties.
@@
 
-In summary, 2018 was warmer, drier, and brighter than usual, while 2019 was a rather average year.
-Ozone was more enhanced in 2018 than 2019.
-This means that the drought conditions in summer 2018 contributed more strongly to lower $\mathrm{POD_1}$ than elevated $\chem{\chi_{O_3}}$ in our $\mathrm{DO_3SE}$ modeling results.
-
 \begin{figure*}[t]
@@
 
+We found that 2018 was warmer, drier, and brighter than usual, while 2019 was a rather average year.
+Ozone was more enhanced in 2018 than 2019 due to large forest fires in Central Sweden.
+We derived localized temperature and light response functions for modeling $\mathrm{POD_1}$ and biomass reduction using the $\mathrm{DO_3SE}$ model and found that vegetation that is sensitive to warm temperatures could be less effected by ozone stress under heat stress conditions.
+On the other hand, vegetation that thrives better in a warmer climate could be more effected by coincident elevated $\chi_{O_3}$.
+
+The qualitative bio-monitoring study which we conducted at Svanhovd along conventional ozone monitoring showed more visible ozone damage on clover and tobacco in 2018.
+In contrast, no such damage was found on the clovers in 2019.
+However, visible damage is not necessarily related to the modeled reductions in biomass.
+
+Our study shows that it could be essential for ozone risk assessments in the subarctic to take the proper plant physiology into account as the difference in biomass reduction compared to the usually used parameterizations could amount to $55\,\unit{\%}$ (deciduous trees).
+We found an above ground biomass reductuion in perennial grassland of well above $10\,\unit{\%}$.
+From an agro-economical point of view in our focus area, where the grass is cut only once a year, a loss of biomass of more than $10\,\unit{\%}$ can have severer consequences compared to more productive areas in central Europe.
+Therefore, CL may also need localization.
+
 TODO: Compare with dry deposition velociteis from Finland.
-
-For our study, we conducted a similar bio-monitoring study at Svanhovd, a site with agrometeorological and air pollution monitoring, to assess whether local ozone concentrations were capable of inducing damage to vegetation growing under northern Fennoscandia conditions.
-In 2018, we indeed observed ozone damage on clover and tobacco.
-In contrast, no such damage was found on the clovers in 2019.
-This indicates that key environmental variables in 2018 differed from 2019.
 

### 2022-05-10 VegOzone.tex
Rephrase discussion and add refrence

--- 2022-05-07 VegOzone.tex
@@
 
-TODO: Compare with dry deposition velociteis from Finland.
-
-For our study, we conducted a similar bio-monitoring study at Svanhovd, a site with agrometeorological and air pollution monitoring, to assess whether local ozone concentrations were capable of inducing damage to vegetation growing under northern Fennoscandia conditions.
-In 2018, we indeed observed ozone damage on clover and tobacco.
-In contrast, no such damage was found on the clovers in 2019.
-This indicates that key environmental variables in 2018 differed from 2019.
-
+We found that 2018 was warmer, slightly drier, and brighter than usual, while 2019 was a rather average year at Svanhovd.
 According to a report by the Norwegian Meteorological institute \citep{MetNOR2019}, the summer of 2018 was the warmest and driest ever recorded in eastern, western, and southern Norway.
-In the north (including Finnmark), it was amongst the warmest on record -- favorable conditions for ground-level ozone formation.
-An unusually weak and northward shifted jet stream allowed for a persistent high-pressure system above northern Europe, including Fennoscandia which blocked the low-pressure systems for several consecutive months.
-In the period May -- July, southern Norway experienced temperatures $4\,\unit{^\circ C}$ above normal and saw only about $60\,\unit{\%}$ of normal precipitation.
-Northern Norway as a whole had close to normal precipitation, one of the causes of the high rate of forest fires in that area.
-Thermal stress on vegetation was exceptional not only in large parts of Fennoscandia but also in much of Europe, where the influence of the high-pressure system extended even over five months (April/May, July--September).
+In the north (including Finnmark), it was amongst the warmest on record.
+Thermal stress on vegetation was exceptional not only in large parts of Fennoscandia but also in much of Europe, where the influence of the blocking high-pressure system extended even over five months (April/May, July--September).
 These conditions gave rise to massive forest fires in different parts of Europe and thus an increase in ozone precursors.
 A total of 2079 forest fires were registered in Norway in 2018, twice as many as in the preceding years 2016/17 \citep{DSB2019}, last accessed April 2020).
-In Sweden, about 500 fires had been reported (five times more than in a usual summer), and an estimated total of $25000\,\unit{hectare}$ burned down in central Sweden (G\"{a}vleborgs, J\"{a}mtlands, and Dalarnas l\"{a}n) \citep{SOU2019}.
-Boreal wildfires are found to emit ozone precursors \chem{CO} ($\chem{[CO]}/\chem{[CO_2]} \propto 6-13\,\unit{\%}$) and VOCs ($\chem{[VOCs]}/\chem{[CO_2]} \propto 0.5-1.5\,\unit{\%}$) \citep{AE:Cofer1990}.
-Coincident peak \chem{\chi_{O_3}} is found in ozone monitoring data (Fig.~\ref{fig:data_svanvik_2018_2019}) in July.
+In Sweden, about 500 fires had been reported (five times more than in a usual summer), and an estimated total of $25000\,\unit{hectare}$ burned down in central Sweden \citep{SOU2019}.
+Coincident peak \chem{\chi_{O_3}} is found in ozone monitoring data (Fig.~\ref{fig:data_svanvik_2018_2019}) in July and ozone was more enhanced in 2018 than 2019.
 It is very likely that elevated ozone, like during the 2003 drought period \citep{JGR:Solberg2018}, was promoted by a combination of various factors such as wildfires, reduced cloud cover (increased solar radiation), reduced dry deposition and turbulent mixing due to the stagnant weather conditions, and increased BVOC emissions.
-As such events may become more frequent in the future, conditions at Svanhovd for the growing seasons 2018 shall serve as a reference for probable future conditions in northern Fennoscandia.
-Conditions in 2019 represent the present climate.\\
-
-To summarize, we have developed bespoke parameterizations of common vegetation types (coniferous and deciduous trees, perennial grassland) to a subarctic climate and studied their effect on $\mathrm{POD_y}$.
-The comparison between meteorological conditions in 2018 and 2019 and their divergence from climatology allowed us to assess the influence of key environmental variables such as temperature, PPFD, and precipitation on vegetation susceptibility to \chem{O_3} damage in light of future changes as may occur under climate change.
-We found that conditions for ozone formation were more favorable in the 2018 growing season than in 2019, with 2018 being significantly warmer and less cloudy in spring and early summer.
-Accordingly, peak ozone concentrations occurred more frequently and at higher levels in 2018.
-This was particularly a result of the extended heatwave in spring and early summer and associated, extensive wildfires in Central Sweden.
-However, ozone concentrations during the growing season were not significantly enhanced and \chem{POD_1} values were very similar for both years.
-This suggests that the overall length of the growing season is more crucial in risk assessment when using flux-based metrics than episodes of peak concentrations as these peak \chem{\chi_{O_3}} often coincide with environmental conditions that limit ozone uptake in the $\mathrm{DO_3SE}$ model.
-
-Based on a MODIS photosynthesis product \citep{MODIS_PSN}, we estimated the growing season for coniferous trees in 2018 to be at least one month shorter than in 2019.
-The determined start of growing season in 2018 ($\unit{doy}~122$) and 2019 ($\unit{doy}~106$) lies within the range given by observations in Fennoscandia \citep{TB:Kolari2007,IVL:Karlsson2018}.
-We determined the start of the growing season for deciduous trees based on gridded observational temperature and for perennial grassland on snow depth data.
-For perennial grassland, we assumed an one-month latency period after snowmelt that is supported by observational data from Rovaniemi (Finland) \citep{FCR:Korhonen2018}.
-With respect to ongoing climate change, a clear positive trend emerged in length ($5.2\,\unit{d\,decade^{-1}}$) of the growing season that is almost equally distributed between earlier start ($2.9\,\unit{days\,decade^{-1}}$) and later end ($2.3\,\unit{d\,decade^{-1}}$) (Appendix Fig.~\ref{fig:greening_season_change_Svanvik}).
-
- The $\mathrm{DO_3SE}$ modeling results are in contrast to our observation of more visible damage in 2018 compared to 2019.
-However, visible damage caused by peak ozone uptake and dependence of ozone sensibility on phenology are not accounted for in $\mathrm{POD_y}$.
+As such events may become more frequent in the future, conditions at Svanhovd for the growing seasons 2018 could serve as a reference for probable future conditions in northern Fennoscandia.
+Conditions in 2019 represent the present climate.
+
+Taking a presumed adaption of vegetation to local climate into consideration, we derived localized temperature and light response functions for modeling $\mathrm{POD_1}$ and biomass reduction with the $\mathrm{DO_3SE}$ model.
+Our key findings are that vegetation that is more sensitive to warm temperatures would be less effected by coincident ozone and heat stress because stomata will close to prevent water loss.
+On the other hand, a warmer climate and hence prolonged GS would negatively affect vegetation adapted to the subarctic.
+It could be essential for ozone risk assessments in the subarctic to define plant physiology and traits properly.
+For comparison, $\mathrm{DO_3SE}$ model derived dry deposition velocities $V_\mathrm{d}$ for coniferous trees with any parameterization are comparable with observations by \citet{BER:Keronen2003} (refer to supplementary material) but might show slightly too high $V_\mathrm{d}$ in spring.
+As perennial grassland and deciduous trees are more strongly affected by our localization (increase in frequency of maximum $V_\mathrm{d}$ throughout the GS), flux measurements over these vegetation types in a subarctic climate would be needed to evaluate our proposed method of localization.
+
+We found that the difference in biomass reduction compared to Boreal parmeterizations \citep{ICP:MappingManual2017} could amount to up to $55\,\unit{\%}$ (deciduous trees).
+Considering perennial grassland with predicted above ground biomass reduction due to ozone of $12-14\,\unit{\%}$ in 2018/2019 ($29-43\,\unit{\%}$ difference).
+From an agro-economical point of view in our focus area, where the grass is cut only once a year, this predicted loss of biomass is large and could be more severe than in more productive areas in other parts of Europe.
+Therefore, the CL for protection of vegetation of $10\,\unit{\%}$ may need adaption.
+
+Along the conventional ozone monitoring at Svanhovd, we had installed a bio-monitoring with ozone sensitive species.
+Qualitatively, this showed visible ozone damage on clover and tobacco in 2018.
+In contrast, no such damage was found on the clovers in 2019.
+This might indicate a higher ozone damage risk under warmer climate conditions for these specific plant species in a subarctic climate.
+However, visible damage is not always correlated with a reduction in biomass \citep[and citations therein]{CRG:Felzer2007}
+and our $\mathrm{DO_3SE}$ modeling results are in contrast to these observations.
+Visible damage caused by peak ozone uptake and specific dependency of ozone sensibility on phenology stage are not accounted for in the $\mathrm{DO_3SE}$ model.
 In this regard, open-top chamber (OTC) experiments performed in northern Finland where Scots pine and downy birch seedlings were exposed to elevated ozone concentrations attributed a reduction in biomass and reproduction with visible damage explicitly to peak \chem{O_3} concentrations and fast phenological development at high growth rate \citep{Amb:Manninen2009}.
@@
 
-At the same time, intra-species variability even at the same location is typically non-negligible \citep{EP:Bassin2004} but not accounted for in the models.
-The literature results regarding the ozone sensitivity of natural vegetation in northern Europe are contradictive.
-\citet{FS:Subramanian2014} reported a higher, modeled reduction of net primary production in conifers ($4.3-15.5\,\unit{\%}$) than birch ($1.4-4.3\,\unit{\%}$) under elevated ozone in Sweden, while \citet{Amb:Girgzdiene2009} observed more visible damage on deciduous trees than on Norway spruce in Lithuania.
-But visible damage and biomass reduction may not occur at the same rate and hence may not be interchangeable.
-
-In our study, we do not explicitly account for this natural distribution of individual plant traits, but our bespoke response functions reflect the diversity within a given species.
-In particular, we assumed two tolerance levels to cold temperatures than for central European species.
-To this end, we computed a PDF from observed temperature at Svanhovd (May--September 1992--2000).
-We adjusted the parameters in the temperature response function of stomatal conductance to increase the enclosed area with the PDF.
-Assuming a proportionality between $g_\mathrm{sto}$ and $A_\mathrm{n}$, this is equivalent to optimizing carbon uptake.
-Similarly, we determined a PPFD PDF and varied the light response function to test a higher or lower extent of stomatal opening at low light intensities.
-Remark that these parameterizations are hypothetical and need verification by experimental data.
+Based on a MODIS GPP product \citep{MODIS_PSN}, we estimated the growing season for coniferous trees in 2018 to be at least one month shorter than in 2019.
+The determined start of growing season in 2018 ($\unit{doy}~122$) and 2019 ($\unit{doy}~106$) lies within the range given by observations in Fennoscandia \citep{TB:Kolari2007,IVL:Karlsson2018}.
+We determined the start of the growing season for deciduous trees based on gridded observational temperature and for perennial grassland on snow depth data.
+For perennial grassland, we assumed a one-month latency period after snowmelt that is supported by observational data from Rovaniemi (Finland) \citep{FCR:Korhonen2018}.
+
+TODO:
 We found that soil water potential under 2018/19 meteorological conditions was negligible when considering the bespoke growing season.
 We have not assessed the effect of the plants' response to VPD in detail, but our climatological data indicated that conditions at Svanhovd are shifting towards a more VPD limited regime.
-
-Our key findings are: 
-\begin{enumerate}
-\item Our bespoke parameterizations for subarctic species are better suited to describe vegetation adapted to subarctic conditions for instance allowing for stomata opening during a larger part of the growing season,
-\item the parameterizations defined in the mapping manual for European regional risk assessment appear to not adequately capture the physiological responses of subarctic vegetation that are important determinants of \chem{\chi_{O_3}} susceptibility and are likely to underestimate ozone-damage risk, and
-\item climatic conditions that are promoting a longer growing season are likely to increase the ozone-damage risk in flux-based metrics.
-\end{enumerate}
 

### 2022-05-12 VegOzone.tex
Update bib and finish discussion and appendix

@@
 Thermal stress on vegetation was exceptional not only in large parts of Fennoscandia but also in much of Europe, where the influence of the blocking high-pressure system extended even over five months (April/May, July--September).
-These conditions gave rise to massive forest fires in different parts of Europe and thus an increase in ozone precursors.
-A total of 2079 forest fires were registered in Norway in 2018, twice as many as in the preceding years 2016/17 \citep{DSB2019}, last accessed April 2020).
+These conditions gave rise to massive forest fires in different parts of Europe.
+A total of 2079 forest fires were registered in Norway in 2018, twice as many as in the preceding years 2016/17 \citep[last accessed April 2020]{DSB2019}.
 In Sweden, about 500 fires had been reported (five times more than in a usual summer), and an estimated total of $25000\,\unit{hectare}$ burned down in central Sweden \citep{SOU2019}.
-Coincident peak \chem{\chi_{O_3}} is found in ozone monitoring data (Fig.~\ref{fig:data_svanvik_2018_2019}) in July and ozone was more enhanced in 2018 than 2019.
+We found coincident peak \chem{\chi_{O_3}} in ozone monitoring data at Svanhovd in July.
+Overall, ozone was more enhanced in 2018 than 2019.
 It is very likely that elevated ozone, like during the 2003 drought period \citep{JGR:Solberg2018}, was promoted by a combination of various factors such as wildfires, reduced cloud cover (increased solar radiation), reduced dry deposition and turbulent mixing due to the stagnant weather conditions, and increased BVOC emissions.
-As such events may become more frequent in the future, conditions at Svanhovd for the growing seasons 2018 could serve as a reference for probable future conditions in northern Fennoscandia.
+As such events may become more frequent in the future, conditions at Svanhovd for the growing seasons 2018 could serve as a reference for probable future conditions in northern Fennoscandia and arising ozone-induced damage risk.
 Conditions in 2019 represent the present climate.
 
-Taking a presumed adaption of vegetation to local climate into consideration, we derived localized temperature and light response functions for modeling $\mathrm{POD_1}$ and biomass reduction with the $\mathrm{DO_3SE}$ model.
-Our key findings are that vegetation that is more sensitive to warm temperatures would be less effected by coincident ozone and heat stress because stomata will close to prevent water loss.
-On the other hand, a warmer climate and hence prolonged GS would negatively affect vegetation adapted to the subarctic.
-It could be essential for ozone risk assessments in the subarctic to define plant physiology and traits properly.
-For comparison, $\mathrm{DO_3SE}$ model derived dry deposition velocities $V_\mathrm{d}$ for coniferous trees with any parameterization are comparable with observations by \citet{BER:Keronen2003} (refer to supplementary material) but might show slightly too high $V_\mathrm{d}$ in spring.
-As perennial grassland and deciduous trees are more strongly affected by our localization (increase in frequency of maximum $V_\mathrm{d}$ throughout the GS), flux measurements over these vegetation types in a subarctic climate would be needed to evaluate our proposed method of localization.
-
-We found that the difference in biomass reduction compared to Boreal parmeterizations \citep{ICP:MappingManual2017} could amount to up to $55\,\unit{\%}$ (deciduous trees).
-Considering perennial grassland with predicted above ground biomass reduction due to ozone of $12-14\,\unit{\%}$ in 2018/2019 ($29-43\,\unit{\%}$ difference).
-From an agro-economical point of view in our focus area, where the grass is cut only once a year, this predicted loss of biomass is large and could be more severe than in more productive areas in other parts of Europe.
-Therefore, the CL for protection of vegetation of $10\,\unit{\%}$ may need adaption.
+Taking a presumed adaption of vegetation to local climate into consideration, we derived localized Jarvis-type temperature and light response functions for modeling $\mathrm{POD_1}$ and biomass reduction with the $\mathrm{DO_3SE}$ model.
+Our key findings are that vegetation that is more sensitive to warm temperatures would be less effected by coincident ozone and heat stress because stomata will close to prevent water loss whereas vegetation adapted to warmer climate zones could be under higher ozone damage risk under these conditions.
+On the other hand, a warmer climate and hence prolonged GS would negatively affect any vegetation using the $\mathrm{POD_y}$ metric.
+It is essential for ozone risk assessments in the subarctic to define plant physiology and traits such as phenological states properly.
+As no independent ozone flux measurements are available at Svanhovd, we could not fully validate our methodology at this point.
+A comparison of $\mathrm{DO_3SE}$ model derived dry deposition velocities $V_\mathrm{d}$ for coniferous trees with comprehensive records over Pine forests from Hyytilälä, Finland \citep{BER:Keronen2003} (refer to supplementary material) indicate a slightly too high $V_\mathrm{d}$ in spring for our most extreme localization (subarctic).
+No comparable data is available for northern Fennoscandia or deciduous tress and perennial grassland.
+Perennial grassland and deciduous trees displayed strong susceptibility to our localization methodology (increase in frequency of maximum $V_\mathrm{d}$ throughout the GS, see supplementary material) but are within the range given in the literature \citep[e.g.][]{JGR:Jacob1992,AE:Padro1996,RG:Clifton2020}.
+Independent flux measurements over these vegetation types in a subarctic climate are thus crucial for evaluation and validation.
 
@@
 This might indicate a higher ozone damage risk under warmer climate conditions for these specific plant species in a subarctic climate.
-However, visible damage is not always correlated with a reduction in biomass \citep[and citations therein]{CRG:Felzer2007}
-and our $\mathrm{DO_3SE}$ modeling results are in contrast to these observations.
-Visible damage caused by peak ozone uptake and specific dependency of ozone sensibility on phenology stage are not accounted for in the $\mathrm{DO_3SE}$ model.
+However, visible damage is not always correlated with a reduction in biomass \citep[and citations therein]{CRG:Felzer2007}.
+Also, our $\mathrm{DO_3SE}$ modeling results are in contrast to these observations at Svanhovd.
+Any visible damage caused by peak ozone flux through stomata or a specific dependency of plant sensibility to ozone on phenology stage are not accounted for in the $\mathrm{DO_3SE}$ model.
 In this regard, open-top chamber (OTC) experiments performed in northern Finland where Scots pine and downy birch seedlings were exposed to elevated ozone concentrations attributed a reduction in biomass and reproduction with visible damage explicitly to peak \chem{O_3} concentrations and fast phenological development at high growth rate \citep{Amb:Manninen2009}.
-In particular, forbs and perennial grassland are more susceptible to ozone-induced damage in the reproductive state \citep{EP:Bassin2004}.
+Forbs and perennial grassland are more susceptible to ozone-induced damage in the reproductive state \citep{EP:Bassin2004}.
 That might explain why we observed damage in the ozone garden to a larger extent in 2018 than in 2019.
@@
 
-Based on a MODIS GPP product \citep{MODIS_PSN}, we estimated the growing season for coniferous trees in 2018 to be at least one month shorter than in 2019.
+Based on the MODIS GPP product \citep{MODIS_PSN}, we estimated the growing season for coniferous trees in 2018 to be at least one month shorter than in 2019.
 The determined start of growing season in 2018 ($\unit{doy}~122$) and 2019 ($\unit{doy}~106$) lies within the range given by observations in Fennoscandia \citep{TB:Kolari2007,IVL:Karlsson2018}.
-We determined the start of the growing season for deciduous trees based on gridded observational temperature and for perennial grassland on snow depth data.
-For perennial grassland, we assumed a one-month latency period after snowmelt that is supported by observational data from Rovaniemi (Finland) \citep{FCR:Korhonen2018}.
-
-TODO:
-We found that soil water potential under 2018/19 meteorological conditions was negligible when considering the bespoke growing season.
-We have not assessed the effect of the plants' response to VPD in detail, but our climatological data indicated that conditions at Svanhovd are shifting towards a more VPD limited regime.
-
-In this regard, vegetation might become more exposed to higher ozone concentrations in early phenological stages due to an overlap with ozone spring peak \citep{EP:Karlsson2007}.
-However, the decline of this ozone spring peak is partly caused by the uptake of vegetation.
+We determined the start of the growing season for deciduous trees using growing-degree-days above $5\,\unit{^\circ C}$ on gridded observational temperature.
+The results coincide with the first snow free days in our target region.
+For the period $1994-2003$ \citet{IJB:Shutova2006} reported an observed average start of the GS for mountain birch at Svanhovd at $\unit{doy}~154$ and end of GS at $\unit{doy}~ 229$.
+These dates are well in line with our estimated average start and end of GS for the period $1960-2020$ (Fig.~\ref{fig:greening_season_change_Svanvik}).
+Our estimates for perennial grassland, are supported by observational data from Rovaniemi (Finland) \citep{FCR:Korhonen2018}.
+The end of GS inferred from the zero-crossing of MODIS observed GPP is in line with independent estimates from the gridded observational temperature.
+Our method differs from NDVI based determination of GS \citep{IGARSS:Hogda2001,IJB:Karlsen2007, RS:Hogda2013}.
+The \citet{IJB:Shutova2006} data indicate an earlier end of GS for birch compared to 2018/19.
+Therefore, $\mathrm{POD_1}$ for deciduous trees could be underestimated by $1.8\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $3.0\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019, respectively if a climatological GS is used.
+For perennial grassland the resulting underestimation would amount to $2.0\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $2.3\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+This demonstrates the importance of an accurate determination of GS for $\mathrm{POD_1}$ based risk assessments.
+
+As shown by \citet{ACP:Bueker2012}, soil water potential is a very important factor for modeling stomatal conductance in the $\mathrm{DO_3SE}$ model.
+Using our best estimate of the soil type at Svanhovd, we found that SWP under 2018/19 meteorological conditions was mainly negligible when considering the localized growing season.
+We have not assessed the effect of the plants' response to VPD in detail, but the climatological data indicated that conditions at Svanhovd are shifting towards a more VPD limited regime that could limit $\mathrm{POD_y}$ under drought stress.
+
+Due to a climate change-induced convergence of start of GS with the ozone spring peak \citep{EP:Karlsson2007}, vegetation might become more exposed to higher ozone concentrations in early phenological stages.
+However, the decline of this ozone spring peak is partly caused by dry deposition of ozone to vegetation itself.
 It, therefore, remains unclear whether an earlier start of the growing season will increase the exposure of vegetation to ozone or lead to compensation due to an earlier decline of the peak ozone in the future.
 \citet{ESPR:Hayes2021} pointed out that the highest susceptibility of vegetation to future ozone variability is likely to occur in summer when vegetation is most productive.
-The actual resilience of northern Fennoscandia plant communities to climate change strongly depends on both the range of existing acclimations as well as the actual type of acclimation.
-This means vegetation that is acclimated to cold temperatures and little to heat- and drought stress might suffer more strongly from a more frequent occurence of heatwaves than plants that have a higher temperature tolerance.
-
-Following the methods described in \citet{ICP:MappingManual2017}, we estimated biomass reductions due to ozone uptake for both 2018 and 2019 and found substantial reductions for deciduous trees ($15.5-17.4\,\unit{\%}$) and perennial grassland (above ground, $12.4-13.7\,\unit{\%}$) in both years.
-These reductions are higher than for the default parameterizations, $11.2\,\unit{\%}$ (deciduous) and $9.6\,\unit{\%}$ (perennial grassland).
-For coniferous trees, the estimated reduction in biomass differs insignificantly between our bespoke parameterizations ($2.5-3\,\unit{\%}$) and the default ($2.31\,\unit{\%}$).
-This indicates a higher modeled robustness to ozone damage in coniferous trees.
-All these results have to be treated with care because the biomass reduction functions were established based on studies from less extreme climates.
-In particular, CLs for deciduous and coniferous trees were breached in all years and for all parameterizations.
-For perennial grassland, the CL was not breached if a bespoke GS was assumed, but the mapping manual defined CL of $10\,\unit{\%}$ loss in biomass might not be transferable to subarctic regions with considerable constraints on productivity leading to more severe economical consequences.
-
-Beyond the risk assessment of ozone-induced damage, our bespoke temperature response functions, once verified by observations, can have important implications on land-surface modeling in global models, where problems in productivity of species especially in the Arctic regions occur due to PFTs which are not suitable for the climate.
-We pointed out, concerning default perennial grassland parameterization, that these perennial grasslands did not show any substantial stomatal conductance in 2019 characterized as a normal year.
-In terms of GPP this would mean unrealistically low carbon uptake.
-Automation of the here proposed PDF-based acclimation using machine learning techniques could overcome these issues in the future.
+The actual resilience of northern Fennoscandia plant communities to climate change strongly depends on both the range of acclimation and the type of acclimation.
+This means, vegetation that is acclimated to cold temperatures and less to heat- and drought stress might suffer more strongly from a more frequent occurrence of heatwaves than plants that have a higher temperature tolerance, but our results point to the opposite concerning the ozone stress.
+
+We found that the difference in biomass reduction compared to Boreal parmeterizations \citep{ICP:MappingManual2017} could amount to $55\,\unit{\%}$ in deciduous trees and $29-43\,\unit{\%}$ in perennial grassland.
+But the biomass reduction functions were established based on studies from less extreme climates, hence, these results ought to be treated with care.
+In particular, CLs for deciduous and coniferous trees were breached in both years and for all parameterizations.
+For perennial grassland, the CL was not breached if a localized GS was assumed, but the defined CL of $10\,\unit{\%}$ loss in biomass might not be transferable to subarctic regions with considerable constraints on productivity leading to more severe economic consequences.
+
+Beyond the risk assessment of ozone-induced damage, our localized Jarvis-type temperature response functions, once verified by observations, can have important implications on land-surface modeling still using such parameterizations in global models.
+Problems in productivity of species especially in the Arctic regions occur due to PFTs which are not suitable for the climate.
+A verification of our proposed methodology by independent flux and dry deposition measurements over birch forest and perennial grassland would be necessary to establish a climate data based localization of stomatal conductance parameterizations.
 
@@
   \includegraphics[width=12cm]{greening_season_change_Svanvik}
-  \caption{Estimated shift and prolongation of growing season at Svanhovd over the past 6 decades based on data from \citet{SeNorge}.}
+  \caption{Estimated shift and prolongation of growing season at Svanhovd over the past 6 decades based on gridded obervational temperature data from \citet{SeNorge}.}
   \label{fig:greening_season_change_Svanvik}
@@
 \subsection{Initial $\mathrm{DO_3SE}$ modeling results with mapping manual parameters}
-To demonstrate the necessary acclimation of the mapping manual parameters, we show the initial results of stomatal conductance at leaf-level ($G_\mathrm{sto}^\mathrm{leaf}$) and $\mathrm{POD_y}$ over \unit{doy} for both years and all PFTs (Fig.~\ref{fig:pody_mm_composit}).
+To demonstrate the necessary localization of the mapping manual parameters, we show the initial results of stomatal conductance at leaf-level ($G_\mathrm{sto}^\mathrm{leaf}$) and $\mathrm{POD_y}$ over \unit{doy} for both years and all PFTs (Fig.~\ref{fig:pody_mm_composit}).
 Observed \chem{\chi_{O_3}} is also indicated.
-From Fig.~\ref{fig:pody_mm_composit}f) it is apparent that the mapping manual parameterized grassland would not have been able to grow in 2019.
 
@@
 
-\subsection{Bespoke parameterizations}
-
-To assess the $G_\mathrm{start}$  and $G_\mathrm{end}$ of the growing season for coniferous trees at Svanhovd in 2018/19, we used the net photosynthesis product ($A_\mathrm{net}$) of MODIS AQUA/TERRA over a $1\times 1\,\unit{km}$ area centered at Svanhovd.
+\subsection{Localized parameterizations}
+
+To assess the $G_\mathrm{start}$  and $G_\mathrm{end}$ of the growing season for coniferous trees at Svanhovd in 2018/19, we used the gross photosynthesis product (GPP) of MODIS AQUA/TERRA over a $1\times 1\,\unit{km}$ area centered at Svanhovd.
 MODIS data indicate a higher photosynthetic activity in 2018 than in 2019.
-As shown in Fig.~\ref{fig:modis_gpp} we fitted a second order polynomial function of the general form
+As shown in Fig.~\ref{fig:modis_gpp}, we fitted a second order polynomial function of the general form
 %
@@
 %
-with the form parameters $m_i$ through the data. Numerically, we retrieved the roots as $G_\mathrm{start}$ 122/106 and $G_\mathrm{end}$ 261/274 for 2018 and 2019, respectively.
+with the form parameters $m_i$ to the data. Numerically, we retrieved the roots as $G_\mathrm{start}$ 122/106 and $G_\mathrm{end}$ 261/274 for 2018 and 2019, respectively.
 
@@
 A $1\times 1\,\unit{km}$ area around Svanhovd was selected.
-Daily averaged data for both 2018 and 2019 has been fitted with a quadratic polynomial function.
-The numerically computed root yields: BGS \unit{doy}~122/106 and EGS \unit{doy}~261/274 for 2018/19, respectively.}
+Daily averaged data for both 2018 and 2019 has been fitted with a quadratic polynomial function.}
   \label{fig:modis_gpp}
@@
 
-As described in Section~\ref{subsec:do3se_parameters}, we developed bespoke parameterizations for natural and semi-natural vegetation at Svanhovd.
-Here, we show the temperature response and light response functions for coniferous and deciduous trees (Figs.~\ref{fig:f_temp_spruce}-\ref{fig:f_temp_birch}).
+Here, we show the temperature and light response functions for coniferous and deciduous trees (Figs.~\ref{fig:f_temp_spruce}-\ref{fig:f_temp_birch}).
 
@@
   %\includegraphics[width=8.3cm]{figB3}
-\caption{Construction of bespoke response functions for coniferous trees.
+\caption{Construction of localized response functions for coniferous trees.
 (a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively.
 Original mapping manual parameterization is shown in comparison as solid line.
-Note that $Q_0$ has been truncated to $0.006$.
+Note that the y-axis has been truncated to $0.006$.
 PPFD0.8 and PPFD1.2 refer to $\alpha$ values increasing/decreasing PPFD at $f_\mathrm{light}=0.5$ by $\pm 20\,\%$, respectively.}

### 2022-05-12 supplements/VegOzone_supplement.tex
Tuning of table
Remove not used figures
Fix supplement

@@
-\graphicspath{{figures_supplement/}}          % graphics
+\documentclass[11pt,a4paper,english]{article} % document type and language
+
+\usepackage[english]{babel}
+\usepackage{tabularx}
+\usepackage{cancel}
+\usepackage{multirow}
+\usepackage{supertabular}
+\usepackage{algorithmic}
+\usepackage{algorithm}
+\usepackage{amsthm}
+\usepackage{float}
+\usepackage{subfig}
+\usepackage{rotating}
+
+\graphicspath{{../figures_supplement/}}          % graphics
 \renewcommand{\thefigure}{S\arabic{figure}}
@@
 
+\begin{document}
 \clearpage
@@
 \begin{figure*}[th]
-  \includegraphics[width=12cm]{ozone_timeseries_fennoscandic_obs.png}
-  \caption{Time series of ozone observations in northern Fennoscandia (Tab.~\ref{tab:ebas_obs}) (1986--2019).
-Data from EBAS until December 2018.
-The hatched areas indicate periods with insufficient quality control according to \citet{NILU2003}.
-(a) Esrange (SWE); (b) Janiskoski (RUS); (c) Jergul/Karasjok (NOR); (d) Pallas (FIN); (e) Svanvik (NOR).}
-  \label{fig:ozone_timesseries_fenoscandic_obs}
-\end{figure*}
-
-%\begin{figure*}[t]
-%  \includegraphics[width=12cm]{test_fennoscandia}
-%  \caption{Maximum daily sun shine duration based on geometrical calculations at different sites in Fennoscandia.
-Midnight sun conditions at Jergul/Karasjok and Svanvik prevail from the end of May until the end of July, while at Esrange and Pallas they only prevail from the beginning of June until mid of July.}
-%  \label{fig:sunlight_fennoscandia}
-%\end{figure*}
-
-%\begin{figure*}[t]
-%  \includegraphics[width=12cm]{ozone_climatology_fenoscandic_obs}
-%  \caption{Multi-annual mean of daily mean ozone at observation sites in Fennoscandia.
-The large spread in case of Svanvik and Janiskoski is due to the lower statistics.
-All climatologies peak in spring, with peak values in April.
-The annual average ozone concentration \chem{\left<[O_3]\right>} at Svanvik is $6.6\,\unit{ppb}$ lower than at the sites at Jergul/Karasjok (NOR), Esrange (SWE), and Pallas (FIN).}
-%  \label{fig:ozone_climatology_fenoscandic_obs}
-%\end{figure*}
-
-\begin{figure*}[th]
-  \includegraphics[width=12cm]{ozone_climatology_fenoscandic_obs_residuals}
-  \caption{Probability density functions of ozone concentration residuals. 2018 with respect to respective climatology for different sites in Fennoscandia. (a) Esrange; (b) Pallas.}
-  \label{fig:ozone_climatology_fenoscandic_obs_residuals}
+  \includegraphics[width=0.95\textwidth]{DO3SE_results_Vd_MM}
+  \caption{Time series of $\mathrm{DO_3SE}$ modeled ozone dry deposition velocities for mapping manual temperature and light response functions in GS 2018/19.
+(a) Deciduous trees; (b) coniferous trees; (c) perennial grassland.}
+  \label{fig:dose_results_vd_mm}
 \end{figure*}
@@
 \begin{figure*}[th]
-  \includegraphics[width=12cm]{ozone_climatology_fenoscandic_obs_residuals-Svanvik}
-  \caption{Probability density functions of ozone concentration residuals. 2018/19 observations at Svanhovd with respect to derived climatologies for (a) Northern Fennoscandia; (b) Svanvik.}
-  \label{fig:ozone_climatology_fenoscandic_obs_residuals-Svanvik}
+  \includegraphics[width=0.95\textwidth]{DO3SE_results_Vd_Cold}
+  \caption{Time series of $\mathrm{DO_3SE}$ modeled ozone dry deposition velocities for "cold" temperature and light response functions in GS 2018/19.
+(a) Deciduous trees; (b) coniferous trees; (c) perennial grassland.}
+  \label{fig:dose_results_vd_cold}
 \end{figure*}
@@
 \begin{figure*}[th]
-  \includegraphics[width=12cm]{ozone_climatology_fenoscandic_obs_test}
-  \caption{Student's t-test assuming same sample uncertainty in both, climatology and 2018 observations for different sites in Fennoscandia.
-\chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals} are significantly different from zero-hypothesis on the $1\,\sigma$ level.
-(a) Esrange; (b) Pallas.}
-  \label{fig:ozone_climatology_fenoscandic_obs_test}
+  \includegraphics[width=0.95\textwidth]{DO3SE_results_Vd_Boreal}
+  \caption{Time series of $\mathrm{DO_3SE}$ modeled ozone dry deposition velocities for "subarctic" temperature and light response functions in GS 2018/19.
+(a) Deciduous trees; (b) coniferous trees; (c) perennial grassland.}
+  \label{fig:dose_results_vd_subarctic}
 \end{figure*}
@@
 \begin{figure*}[th]
-  \includegraphics[width=12cm]{ozone_climatology_fenoscandic_obs_test-Svanvik}
-  \caption{Student's t-test assuming same sample uncertainty in both, climatology and 2018/19 observations at Svanhovd with respect to derived climatologies for (a) Northern Fennoscandia; (b) Svanvik.
-\chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}a) are significantly different from zero-hypothesis on the $2\,\sigma$,  $1\,\sigma$ level, respectively.
-\chem{\Delta[O_3]} in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_residuals}b) are not significantly different from zero-hypothesis.}
-  \label{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}
-\end{figure*}
-
-\begin{figure*}[th]
-  \centering
-  \includegraphics[width=1\textwidth]{supplements/supplement_do3se-parameterization_v2}
+  \hspace*{-3cm}
+  \includegraphics[width=1.5\textwidth]{supplement_do3se-parameterization_v2}
   \label{tab:do3se_mm_param}
@@
 
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{greening_season_change_Svanvik}
-  \caption{Estimated shift and prolongation of growing season at Svanhovd over the past 6 decades based on data from \citet{SeNorge}.}
-  \label{fig:greening_season_change_Svanvik}
-\end{figure*}
-
-%\begin{figure*}[t]
-%  \includegraphics[width=12cm]{ozone_observation_timelag_svanvik}
-%  \caption{Temporal correlation of \chem{[O_3]} data at Svanvik with other ozone monitoring stations in northern Fennoscandia.
-A negative lag means that Svanvik lags behind, while a positive lag mean the other station lags behind.
-The highest correlation with Pallas/Esrange is found at a time lag of $3\,\unit{h}$, for Jergul/Karasjok at $1\,\unit{h}$, and for Janiskoski at $-20\,\unit{h}$.}
-%  \label{fig:time_lag_correlation}
-%\end{figure*}
-
-%\begin{figure*}[t]
-%  \includegraphics[width=12cm]{ozone_fennoscandic_obs_aot40}
-%  \caption{Integrated AOT40 for various sites in northern Fennoscandia over the course of 33 years.
-For simplicity, we integrated hourly ozone between $1\,\unit{am}-11\,\unit{pm}$ in any case, although a night time light intensity $> 50\,\unit{W\,m^{-2}}$ is only measured during midnight sun conditions.
-Since night time \chem{[O_3]} is mostly below $40\,\unit{ppb}$, the such induced high bias in spring should be tolerable.
-The dashed red lines indicate the respective threshold given by the EU directive.
-(a) May 1 -- July 31 (b) April 1 -- September 30.}
-%  \label{fig:fennoscandic_aot40}
-%\end{figure*}
+\end{document}

### 2022-05-12 VegOzone.tex
Add reference to supplementary info

@@
 As no independent ozone flux measurements are available at Svanhovd, we could not fully validate our methodology at this point.
-A comparison of $\mathrm{DO_3SE}$ model derived dry deposition velocities $V_\mathrm{d}$ for coniferous trees with comprehensive records over Pine forests from Hyytilälä, Finland \citep{BER:Keronen2003} (refer to supplementary material) indicate a slightly too high $V_\mathrm{d}$ in spring for our most extreme localization (subarctic).
+A comparison of $\mathrm{DO_3SE}$ model derived dry deposition velocities $V_\mathrm{d}$ for coniferous trees with comprehensive records over Pine forests from Hyytilälä, Finland \citep{BER:Keronen2003} (refer to supplementary material Figs.~S1--S3) indicate a slightly too high $V_\mathrm{d}$ in spring for our most extreme localization (subarctic).
 No comparable data is available for northern Fennoscandia or deciduous tress and perennial grassland.
-Perennial grassland and deciduous trees displayed strong susceptibility to our localization methodology (increase in frequency of maximum $V_\mathrm{d}$ throughout the GS, see supplementary material) but are within the range given in the literature \citep[e.g.][]{JGR:Jacob1992,AE:Padro1996,RG:Clifton2020}.
+Perennial grassland and deciduous trees displayed strong susceptibility to our localization methodology (increase in frequency of maximum $V_\mathrm{d}$ throughout the GS, see supplementary material Figs.~S1--S3) but are within the range given in the literature \citep[e.g.][]{JGR:Jacob1992,AE:Padro1996,RG:Clifton2020}.
 Independent flux measurements over these vegetation types in a subarctic climate are thus crucial for evaluation and validation.

### 2022-05-12 VegOzone.tex
Small update

@@
 No comparable data is available for northern Fennoscandia or deciduous tress and perennial grassland.
-Perennial grassland and deciduous trees displayed strong susceptibility to our localization methodology (increase in frequency of maximum $V_\mathrm{d}$ throughout the GS, see supplementary material Figs.~S1--S3) but are within the range given in the literature \citep[e.g.][]{JGR:Jacob1992,AE:Padro1996,RG:Clifton2020}.
+Perennial grassland and deciduous trees displayed strong susceptibility to our localization methodology (increase in frequency of maximum $V_\mathrm{d}$ throughout the GS, see supplementary material Figs.~S1--S3) but are within the range ($v\mathrm{d} = (2-9)\,\unit{mm\,s^{-1}}$ for grassland) given in the literature \citep[e.g.][]{JGR:Jacob1992,AE:Padro1996,RG:Clifton2020}.
 Independent flux measurements over these vegetation types in a subarctic climate are thus crucial for evaluation and validation.

### 2022-05-30 VegOzone.tex
Update figures and caption

@@
   %\includegraphics[width=8.3cm]{fig09}
-  \caption{$\mathrm{DO_3SE}$ modeling $\mathrm{POD_1}$ results for (a) deciduous trees, (b) coniferous trees, and (c) perennial grasslands at Svanhovd in 2018 (pink) and 2019 (purple).
-The results varied depending on parameterizations of temperature response functions (MM, Cold, and Subarctic, details in Table~\ref{tab:sensitivity_tests_temp}), growing season (GS MM (squares) or GS bespoke (circles), details in Table~\ref{tab:sensitivity_tests_gs}), light response functions (first, second and third point attached by a line are depicting the MM, PPFD0.8 and PPFD1.2, details in Table~\ref{tab:sensitivity_tests_light}), and the effect of taking SWP into account (filled symbols without and open symbols with effects of drought kept in the model).
+  \caption{$\mathrm{DO_3SE}$ modeling $\mathrm{POD_1}$ results for (a) deciduous trees, (b) coniferous trees, and (c) perennial grasslands at Svanhovd in 2018 and 2019.
+The results vary depending on parameterizations of temperature response functions (MM, Cold, and Subarctic, details in Table~\ref{tab:sensitivity_tests_temp}), growing season (details in Table~\ref{tab:sensitivity_tests_gs}), light response functions (first, second and third point attached by a line are depicting the MM, PPFD0.8 and PPFD1.2, respectively, details in Table~\ref{tab:sensitivity_tests_light}), and taking SWP (effect of drought) into account.
 Critical levels for ozone damage \citep{ICP:MappingManual2017,ESPR:Hayes2021} are given as dashed horizontal lines.

### 2022-05-30 VegOzone.tex
Fix equation

@@
 \begin{equation}
-  \Phi_\mathrm{sto} = \chem{[O_3]}(h) \cdot g_\mathrm{sto}(h).
+  \Phi_\mathrm{sto} = \chem{[O_3]}(h) \cdot \frac{g_\mathrm{sto}}{g_\mathrm{sto}+g_\mathrm{ext}} \cdot \frac{1}{r_b+r_c}.
   \label{eq:flux_stomata}
 \end{equation}
-A linear gradient reducing \chem{[O_3]} from $h_0$ to $h$ was adopted in the $\mathrm{DO_3SE}$ model based on detailed transport studies using the chemical transport model EMEP \citep{ACP:Simpson2012,ICP:MappingManual2017}.
+A linear gradient adjusting \chem{[O_3]} from $h_0$ to $h$ was adopted in the $\mathrm{DO_3SE}$ model based on detailed transport studies using the chemical transport model EMEP \citep{ACP:Simpson2012,ICP:MappingManual2017}.
 

### 2022-05-30 VegOzone.tex
Fix all todos

@@
 The light response function localizations are denoted by $+$ (PPFD1.2) and $-$ (PPFD0.8).
-Circles indicate simulations with localized GS, canopy height, and leaf dimensions, TODO: (whereas squares indicate simulations with MM GS for Central Europe.
-We chose this to demonstrate the impact of a projected prolonged GS in the future.) With open symbols, the effect on $\mathrm{POD_y}$ with SWP taken into account is displayed.
-
-We identified in general that an increasing cold tolerance represented by a shift in the temperature response function (MM $<$ cold $<$ subarctic) leads to an increase in $\mathrm{POD_1}$ in both probed years.
-TODO: (while using locally adapted GS causes a reduction.) An earlier start and overall longer GS in 2019 leads to higher $\mathrm{POD_1}$ compared to 2018 despite the occurrence of more episodes of elevated \chem{O_3} in 2018 than in 2019 (see Section~\ref{subsec:anomalies} for details).
-Taking into account SWP, $\mathrm{POD_1}$ was effectively only reduced for deciduous trees in 2018, while in all other cases simulation results with and without SWP effects were identical.
-An earlier opening of stomata at low light intensities (PPFD0.8) caused an increase in $\mathrm{POD_1}$ compared to MM while a later opening (PPFD1.2) reduced $\mathrm{POD_1}$.
+Circles indicate simulations with localized GS, canopy height, and leaf dimensions.
+With open symbols, the effect on $\mathrm{POD_y}$ taking SWP (drought) into account is displayed.
+
+We identified in general that an increasing cold tolerance represented by a shift in the temperature response function (MM $<$ cold $<$ subarctic) leads to an increase in $\mathrm{POD_1}$ in both probed years compared to the MM parameterization.
+The more extreme the temperature response function localization the higher $\mathrm{POD_1}$.
+In 2018, however, coniferous trees and perennial grassland display a reduction in $\mathrm{POD_1}$ for the subarctic parameterization compared to cold in responds to the prescribed lower $g_\mathrm{sto}$ at higher temperatures.
+An earlier start and overall longer GS in 2019 leads to higher $\mathrm{POD_1}$ compared to 2018 despite the occurrence of more episodes of elevated \chem{O_3} in 2018 than in 2019 (see Section~\ref{subsec:anomalies} for details on the particular features of these two years).
+Taking into account SWP, $\mathrm{POD_1}$ was only notably reduced in 2018 (for deciduous trees).
+In all other model experiments SWP was hardly of relevance.
+An earlier opening of stomata at low light intensities (PPFD0.8) causes an increase in $\mathrm{POD_1}$ compared to MM while a later opening (PPFD1.2) reduces $\mathrm{POD_1}$.
 Due to the shape of $f_\mathrm{light}$, a symmetric variation can lead to an asymmetric response in $\mathrm{POD_1}$.
 
-For deciduous trees (Fig.~\ref{fig:pody_rel}a) all $\mathrm{POD_1}$ simulations exceed the CL by TODO: ($5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$) and display the largest spread in $\mathrm{POD_1}$ with respect to $f_\mathrm{light}$.
-In 2019, the lowest estimated $\mathrm{POD_1}$ (MM, PPFD1.2) amounts to $10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, while the highest estimate (subarctic, PPFD0.8) reaches about $20\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
-In 2018, this uncertainty range is much smaller ($(12-18)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
-In particular, the impact of $f_\mathrm{light}$ is pronounced (extent of $\angle(+,0,-)$).
-The maximum difference in $\mathrm{POD_1}$ due to the $f_\mathrm{light}$ localization for the PPFD0.8 (superscript) and PPFD1.2 (subscript) amount to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.
-The difference in $\mathrm{POD_1}$ compared to MM in response to the localization of $f_\mathrm{T}$ and GS ranges between $\Delta\mathrm{POD_1}=(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $\Delta\mathrm{POD_1}=(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
-A low soil water potential caused a reduction in $\mathrm{POD_1}$ in 2018 in cases when using PPFD0.8.
+For deciduous trees (Fig.~\ref{fig:pody_rel}a), all model experiments exceed the CL by ($5-16\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
+In 2019, the lowest $\mathrm{POD_1}$ estimate (MM, PPFD1.2) amounts to $10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and the highest (subarctic, PPFD0.8) $21\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
+In 2018, this range is smaller ($(12-18)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
+The sensitivity to $f_\mathrm{light}$ is largest in the subarctic parameterization and amounts to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+Herein, the difference of PPFD0.8 compared to MM $f_\mathrm{light}$ is denoted as superscript and for PPFD1.2 as subscript.
+The extent of the opening angle ($\angle(+,0,-)$) indicates the sensitivity visually.
+A low soil water potential caused a reduction in $\mathrm{POD_1}$ in 2018 in particular when using PPFD0.8.
 This indicates that drought conditions can reduce the ozone damage risk.
-TODO: (With MM GS, $\mathrm{POD_1}$ values were higher in 2018 than in 2019.
-However, taking the localized GS into account, this is reversed for the subarctic temperature response function.
-This can be explained by more favorable growing conditions in 2019.)
-
-For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
-Our localization does not show much difference compared to MM indicating the the MM Boreal Norway spruce parameterization was already well-adjusted to the local subarctic climate.
-We find a higher $\mathrm{POD_1}$ in 2019 than in 2018.
-The maximum difference in $\mathrm{POD_1}$ determined between localized GS for subarctic and cold parameterizations with respect to MM ranges between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
-The maximum difference determined from PPFD0.8 (superscript) and PPFD1.2 (subscript) $f_\mathrm{light}$ amounts to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2018 and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ for 2019.
-
-For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations with localize GS stay below the CL.
+
+For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL by $0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
+The different localizations do not show much impact compared to MM indicating that the MM's Boreal Norway spruce parameterization is well-suited for a subarctic climate.
+We find higher $\mathrm{POD_1}$ in 2019 than in 2018.
+The maximum difference in $\mathrm{POD_1}$ determined between subarctic and cold parameterizations with respect to MM amounts to $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+The maximum difference due to $f_\mathrm{light}$ amounts to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+
+For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations stay below the CL.
 Perennial grassland shows the smallest $\mathrm{POD_1}$, but a similarly large response to the temperature parameterization localizations as deciduous trees.
-Again, we find a reversal for predicted ozone risk in 2018 and 2019 for the subarctic type.
 Perennial grassland shows the lowest sensitivity to the light threshold.
-TODO: (and SWP is only relevant without localized GS (see above).) The overall difference in $\mathrm{POD_1}$ determined from the localized GS for subarctic and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
-The chosen lower maximum temperature in the subarctic parameterization probably caused the lower $\mathrm{POD_1}$ in 2018 (warm year) compared to the cold parameterization.
-This was not seen in 2019.
-The maximum difference in $\mathrm{POD_1}$ determined from the PPFD0.8 (superscript) and PPFD1.2 (subscript) variation of $f_\mathrm{light}$ amount to $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.
-
-The maximum difference in $\mathrm{POD_1}$ between simulations with the MM parameterizations and the localized temperature and light response functions is of the same order of magnitude as the difference between the two probed years.
-The magnitude of all described effects on $\mathrm{POD_1}$ varies between PFTs as well as years, but is by construction always larger than estimated with the MM parameterization.
-This means that the ozone damage risk could be possibly underestimated even in normal years and establishing dedicated subarctic parameterizations are a must to properly assess ozone damage risk in these regions.
+SWP is hardly relevant.
+The difference in $\mathrm{POD_1}$ for subarctic and cold types with respect to MM amounts to $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+The maximum difference in $\mathrm{POD_1}$ determined from the PPFD0.8 (superscript) and PPFD1.2 (subscript) variation of $f_\mathrm{light}$ amount to $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ (2018) and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ (2019), respectively.
+
+The magnitude of all described effects on $\mathrm{POD_1}$ varies between PFTs as well as years and is of the same order of magnitude.
+By construction, $\mathrm{POD_1}$ from localized parmeterizations is always larger than estimates with the MM parameterization.
+This means that the ozone damage risk could be possibly underestimated especially in more normal years like 2019 (see Section~\ref{sec:stats} for details).
+Establishing dedicated subarctic parameterizations are a must to properly assess ozone damage risk in these regions.
 

### 2022-05-30 VegOzone.tex
Editorial changes suggested by Lisa

@@
 \begin{abstract}
-  The unique vegetation of the subarctic region acclimatized to extremes of cold and midnight sun are likely to be at threat from the combined impacts of climate change and increasing ozone concentrations \chem{[O_3]}.
-To date, only a few studies assessed the response of subarctic vegetation to variations in climate and air pollution.
-This study looks to fill this knowledge gap by examining the uncertainty in risk assessments that are using Boreal paramterizations.
+  The unique vegetation of the subarctic region, acclimatized to extremes of cold and midnight sun, is likely to be at threat from the combined impacts of climate change and increasing ozone concentrations \chem{[O_3]}.
+To date, only a few studies have assessed the response of subarctic vegetation to variations in climate and air pollution.
+This study looks to fill this knowledge gap by examining the uncertainty in risk assessments that result from using Boreal rather than sub-Arctic paramterizations.
 With no parameterization of stomatal response of subarctic vegetation available, we propose an adjustment of existing Boreal parameterizations based on local climate condition.
-To this end, we vary temperature response functions according to the climatological temperature distribution within the growing season at our site.
-Thus, we refer to our approach as localization of stomatal response functions.
+To this end, we vary the temperature response function according to the climatological temperature distribution during the growing season at our site.
+We refer to this approach as a 'local climate derived parameterization' of stomatal response functions.
 With these, we apply the integrated flux-metric Phytotoxic Ozone Dose (POD) to determine ozone risk and damage to vegetation as a function of \chem{[O_3]}, key environmental variables, and species-specific physiology.
@@
 Higher than average \chem{[O_3]} was observed in April/May while frequent episodes of ozone volume mixing ratios (VMRs) exceeding $40\,\unit{ppb}$ occurred in June--August.
-These episodes are in part attributable to forest fires in the Northern Hemisphere and warmer and sunnier conditions.
+These episodes are in part attributable to forest fires in the Northern Hemisphere and warmer and sunnier conditions caused by an extensive blocking event.
+Such conditions might be more likely in the future under advancing climate change.
 In contrast, 2019 was an average year.
 Our modeling study suggests that using standard Boreal parameterizations in assessments likely leads to underestimating the risk of ozone damage in this region for some species.
-We find that localized parameterizations for subarctic vegetation result in an ozone-induced biomass loss of $2.5$ to $17.4\,\unit{\%}$.
-For some species, this loss is up to $55\,\unit{\%}$ larger than projected from Boreal parameterizations used in general assessments.
+We find that local climate derived parameterizations for subarctic vegetation result in an ozone-induced biomass loss of $2.5$ to $17.4\,\unit{\%}$.
+For some species, this loss is up to $55\,\unit{\%}$ larger than that projected from Boreal parameterizations used in general assessments.
 Efforts should be targeted towards accurately defining subarctic vegetations' physiological response to essential climate variables.
@@
 In Section~\ref{sec:dam}, we give a brief summary of the key components and  drivers of the $\mathrm{DO_3SE}$ model that is used to assess the ozone risk.
-To study the influence of selected parameters on ozone risk in the subarctic, we propose a method to localize temperature response functions based on climate data and to infer start and end of the growing season from satellite observed gross primary productivity (GPP).
+To study the influence of selected parameters on ozone risk in the subarctic, we propose a method to derive temperature response functions based on local climate and to infer start and end of the growing season from satellite observed gross primary productivity (GPP).
 Results of the ozone risk modelling are presented in Section~\ref{sec:results}.

### 2022-05-30 VegOzone.tex
Follow Lisa's suggestions for introduction and move part to discussion

@@
 We refer to this approach as a 'local climate derived parameterization' of stomatal response functions.
+We also study the impact of the opening extent of stomata at low light intensities.
 With these, we apply the integrated flux-metric Phytotoxic Ozone Dose (POD) to determine ozone risk and damage to vegetation as a function of \chem{[O_3]}, key environmental variables, and species-specific physiology.
@@
 Our modeling study suggests that using standard Boreal parameterizations in assessments likely leads to underestimating the risk of ozone damage in this region for some species.
-We find that local climate derived parameterizations for subarctic vegetation result in an ozone-induced biomass loss of $2.5$ to $17.4\,\unit{\%}$.
+We find that local climate derived parameterizations for subarctic vegetation result in an ozone-induced biomass loss of $2.5-17.4\,\unit{\%}$ compared to $2.3-11.2\,\%$.
 For some species, this loss is up to $55\,\unit{\%}$ larger than that projected from Boreal parameterizations used in general assessments.
@@
 Ozone causes reduced photosynthesis, visible foliar injury, early senescence, and programmed cell death of plants \citep{PCE:Kangasjarvi2005}.
-Hence affecting the global carbon cycle \citep{BGS:Lombardozzi2013} and crop yields.
-Annual global yield losses of four major crops (wheat, rice, maize, and soybean) of $3-15\,\unit{\%}$ have been reported \citep{PJ:Ainsworth2017} and a suggested loss in primary production in forestry of $7\,\unit{\%}$ \citep{GCB:Wittig2009,EP:Matyssek2012} has been attributed to ozone.
-Even though long-term (pre-1950) observations of ground-level ozone concentrations are scarce; evidence from trend analysis of those few sites suggests that since the industrial revolution, background concentrations in the northern hemisphere have at least doubled and continued to increase \citep{IPCC2013} until recently.
-Recent trend analyses for Europe indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric background ozone started to level off or even decline \citep{ESA:Cooper2014, ACP:Wespes2018,ESA:Gaudel2018}.
+Such ozone induced injury and damage can cause losses in crop \citep{PJ:Ainsworth2017}, forest \citep{GCB:Wittig2009,EP:Matyssek2012}, and productive grassland productivity and interfere with the global carbon cycle \citep{BGS:Lombardozzi2013}.
+Even though long-term (pre-1950) observations of ground-level ozone concentrations are scarce; evidence from trend analysis of those few sites for which such data do exist suggests that since the industrial revolution, background concentrations in the northern hemisphere have at least doubled and continued to increase \citep{IPCC2013} until recently.
+Recent trend analyses for Europe indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric background ozone concentrations started to level off or even decline \citep{ESA:Cooper2014, ACP:Wespes2018,ESA:Gaudel2018}.
 This can be attributed to a successful implementation of air quality regulations coupled with economic restructuring, reducing the number of episodes of peak concentrations especially in summertime \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
-However, changes in environmental conditions associated with climate change, such as an increase in the frequency of heatwaves, could negate the effectiveness of emission reductions concerning ozone impacts and air quality \citep{NCC:Lin2020}.
+However, changes in environmental conditions associated with climate change, such as an increase in the frequency of heatwaves, could negate the effectiveness of emission reductions on ozone concentrations and associated impacts \citep{NCC:Lin2020}.
+
 Tropospheric background ozone concentrations derived from satellite instruments range between $50-65\,\unit{ppb}$ over Fennoscandia but strongly depend on season \citep{ESA:Cooper2014}.
-
-In the annual cycle, a distinct maximum of ground-level ozone concentrations typically occurs in spring followed by lower concentrations throughout the summer.
+Annual ozone concentration profiles for Fennoscandia show a distinct maximum ground level-ozone concentration during the spring followed by lower concentrations throughout the summer.
 This ozone spring peak is more pronounced in northern Fennoscandia than in more southern locations but shows a high interannual variability \citep{AB:Klingberg2009, BER:Klingberg2019}.
-\citet{ACP:Andersson2017} showed in a modeling study focused on Fennoscandia that the variability of ozone concentrations in winter can be mainly attributed to changes in atmospheric background (transport) of ozone, while summertime abundance is mostly affected by emissions of precursors in the rest of Europe.
-
-The precursor gases involved in the formation of ground-level ozone are carbon monoxide (\chem{CO}) and hydrocarbons known as volatile organic compounds (VOCs) in the presence of nitrogen oxides (\chem{NO_x}).
-In northern Fennoscandia, episodes of elevated summertime ozone concentrations in 2003 and 2006 have been traced back to ozone precursor emissions related to forest fires in southern and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}.
+\citet{ACP:Andersson2017} showed in a modeling study focused on Fennoscandia that the variability of ozone concentrations in winter can be mainly attributed to changes in atmospheric background ozone concentrations determined by circulation patterns and atmospheric transport, whilst summertime abundance is mostly affected by emissions of precursors in the rest of Europe.
+For example, in northern Fennoscandia, episodes of elevated summertime ozone concentrations in 2003 and 2006 were attributed to ozone precursor emissions from forest fires in southern and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}.
+The precursor gases were predominantly carbon monoxide (\chem{CO}) and hydrocarbons known as volatile organic compounds (VOCs) in the presence of nitrogen oxides (\chem{NO_x}).
 Besides VOCs from anthropogenic sources, hydrocarbons are also emitted by photosynthetically active vegetation in the form of terpenes and monoterpenes, so-called biogenic volatile substances (BVOCs) and soil.
-These emissions are thought to be a response to factors, such as thermal stress, defense against herbivores \citep{Ecosystems:Ghimire2021}, or even attraction of pollinators \citep{TPS:Penuelas2003}.
-\citet{Ecosystems:Ghimire2021} found no clear evidence that general warming would increase the BVOC emissions in subarctic vegetation, while a reduction of herbivory would decrease them by $44-72\,\%$.
-
-The start of the growing season (GS) in northern Fennoscandia has been shifting to earlier dates in response to the general warming trend \citep[e.g.]{GCB:Menzel2006,RS:Hogda2013,IJB:Karlsen2007} and thereby converging with the period of the ozone spring peak.
+
+The growing season (GS) in northern Fennoscandia has been shifting and starts earlier in response to the general warming trend \citep[e.g.]{GCB:Menzel2006,RS:Hogda2013,IJB:Karlsen2007}.
+Its start is thereby converging with the period of the ozone spring peak.
 At the same time, the growing season is also becoming longer.
 A longer growing season is prolonging the time in which vegetation can accumulate an ozone dose.
-For natural and semi-natural vegetation that is already subject to rapid climatic changes known as Arctic amplification \citep{AMAP2012,IPCC2013}, these factors could promote a higher potential risk for northern vegetation to ozone-induced damage in the not so distant future.
+For natural and semi-natural vegetation types that is already subject to rapid climatic changes known as Arctic amplification \citep{AMAP2012,IPCC2013}, these factors could promote a higher potential risk for northern vegetation to ozone-induced damage in the not so distant future.
 Recently, \citet{ESPR:Hayes2021} showed that the highest sensitivity to future \chem{[O_3]} variability is expected in summer when vegetation is most productive.
@@
 Stomatal conductance has been empirically linked to key environmental variables such as air temperature $T$, photosynthetic photon flux density (PPFD), vapor pressure deficit (VPD), and soil water potential (SWP) as well as photosynthesis itself \citep[e.g.,][]{PTRS:Jarvis1976, BallBerry1987, Emberson2000, ICP:MappingManual2017}.
-Results from open-top chamber (OTC) experiments on downy birch (\emph{Betula pubescens}) and mountain birch (\emph{Betula pubescens toruosa}), native to subarctic regions, indicated reductions in both biomass, root:shoot ratio, and visible leaf damage under elevated ozone treatment ($\left<\chem{\chi_{O_3}}\right> = 36-54\,\unit{ppb}$) \citep{Amb:Manninen2009}.
-Though, Scots pine (\emph{Pinus sylvestris}) is considered to be more ozone tolerant due to an absence of visible injuries \citep{Amb:Girgzdiene2009}, \citet{Amb:Manninen2009} found chlorophyll:carotenoid ratio and polyamines reductions under elevated ozone concentrations indicating susceptibility to ozone also in these species.
+Results from open-top chamber (OTC) experiments on downy birch (\emph{Betula pubescens}) and mountain birch (\emph{Betula pubescens toruosa}), native to subarctic regions, indicated reductions in both biomass, root:shoot ratio, and visible leaf damage under elevated ozone treatments ($\left<\chem{\chi_{O_3}}\right> = 36-54\,\unit{ppb}$) \citep{Amb:Manninen2009}.
+Although, Scots pine (\emph{Pinus sylvestris}) is considered to be more ozone tolerant due to an absence of visible injuries \citep{Amb:Girgzdiene2009}, \citet{Amb:Manninen2009} found chlorophyll:carotenoid ratio and polyamine reductions under elevated ozone concentrations indicating susceptibility to ozone also in these species.
 
@@
 These flux-based metrics are better suited to represent the actual risk to vegetation from ozone, especially in the climatically extreme parts of Europe \citep{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017}.
-But no bespoke parameterization for subarctic vegetation is available.
-
-The $\mathrm{POD_y}$ metric is currently used in risk assessments under the United Nations Economic Commission for Europe (UNECE) Long Range Transboundary Air Pollution (LRTAP) Convention to identify those locations across Europe where vegetation is at risk from ozone.
+Currently, parameterizations are available for some species for particular climatic regions e.g.
+Boreal, temperate and Mediterranean; there is no parameterization for sub arctic vegetation.
+Given the unique conditions of the sub arctic (i.e.
+long duration of daylight in the summer and cold temperatures) it would be expected that this vegetation type has a particular physiological response to the prevailing climatic conditions different to boreal species.
+
+The $\mathrm{POD_y}$ metric is currently used within the United Nations Economic Commission for Europe (UNECE) Long Range Transboundary Air Pollution (LRTAP) Convention to identify those locations across Europe where vegetation is at risk from ozone.
 This convention aims to develop an effects-based emission reduction policy that can target those ozone precursor emissions that are most influential in causing damage.
 To achieve this, the concept of critical levels (CLs) has been adopted and applied \citep{Maas2016}.
-Exceedance of the CL is used to identify those areas across Europe that would benefit from targeted Europe-wide emission reductions.
 Methods to estimate $\mathrm{POD_y}$ and CLs are defined by the UNECE LRTAP Convention and described in a Mapping Manual \citep[most recent version]{ICP:MappingManual2017}.
@@
 
-The described methods for risk assessment rely on accurate representation of plant physiological responses to key environmental variables and are most often parameterized using data for less extreme climates, i.e., temperate, continental, or Boreal climates.
-Parameterization for more extreme bio-geographical regions such as subarctic climates has not been performed and relies on assuming that less extreme Boreal parameterization will be applicable.
-It is important to understand the implications of using these in the subarctic.
-In our study, we explore how risk might be over- or underestimated using Boreal parameterization and explore an adjustment of temperature and light response functions based on climate data.
-This allows an understanding of which are the key aspects of modelled plant physiology that might determine the potential risk and also how the threat from ozone may change in the future.
-In addition, our results may hint a solution to solve long-standing issues of earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface models and display over- or underproductivity in the Arctic \citep{GMD:Poulter2015,JAMES:Lawrence2019}.\\
+Flux-based methods of risk assessment rely on accurate representation of plant physiological responses to key environmental variables and are most often parameterized using data for less extreme climates, i.e., temperate, continental, or Boreal climates.
+Parameterization for more extreme bio-geographical regions such as subarctic climates has not been performed and relies on assuming that Boreal parameterization will be applicable.
+In our study, we explore how risk might be over- or underestimated using Boreal parameterization and explore a local climate derived temperature response function.
+We also study the impact of the opening extent of stomata at low light intensities (light response function).\\
 
@@
   %\includegraphics[width=8.3cm]{fig07}
-\caption{Construction of bespoke response functions for grassland.
+\caption{Construction of local climate derived response functions for grassland.
 (a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively.
@@
 \begin{table}[t]
-  \caption{Bespoke temperature parameterizations. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
+  \caption{Local climate derived temperature parameterizations. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
   \label{tab:sensitivity_tests_temp}
@@
 \begin{table}[t]
-  \caption{Bespoke light parameterizations. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
+  \caption{Local climate derived light parameterizations. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
   \label{tab:sensitivity_tests_light}
@@
   %\includegraphics[width=8.3cm]{fig08}
-  \caption{Proposed metric to test bespoke response functions.
+  \caption{Proposed metric to test local climate derived response functions.
 GS (May--August) climatological averages and standard deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon (averaged over $11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning (averaged over $5-9\,\unit{am}$ local time) are shown for (a) deciduous trees; (b) coniferous trees; (c) perennial grassland.}
@@
 
+..
+This allows an understanding of which are the key aspects of modelled plant physiology that might determine the potential risk and also how the threat from ozone may change in the future.
+In addition, our results may hint a solution to solve long-standing issues of earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface models and display over- or underproductivity in the Arctic \citep{GMD:Poulter2015,JAMES:Lawrence2019}.
+..
+
 We found that 2018 was warmer, slightly drier, and brighter than usual, while 2019 was a rather average year at Svanhovd.
@@
   %\includegraphics[width=8.3cm]{figB4}
-\caption{Construction of bespoke response functions for deciduous trees.
+\caption{Construction of local climate derived response functions for deciduous trees.
 (a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively.

### 2022-05-30 VegOzone.tex
Remove redudant text

@@
 We refer to this approach as a 'local climate derived parameterization' of stomatal response functions.
-We also study the impact of the opening extent of stomata at low light intensities.
+We also study the impact of the stomata aperture at low light intensities.
 With these, we apply the integrated flux-metric Phytotoxic Ozone Dose (POD) to determine ozone risk and damage to vegetation as a function of \chem{[O_3]}, key environmental variables, and species-specific physiology.
@@
 This ozone spring peak is more pronounced in northern Fennoscandia than in more southern locations but shows a high interannual variability \citep{AB:Klingberg2009, BER:Klingberg2019}.
-\citet{ACP:Andersson2017} showed in a modeling study focused on Fennoscandia that the variability of ozone concentrations in winter can be mainly attributed to changes in atmospheric background ozone concentrations determined by circulation patterns and atmospheric transport, whilst summertime abundance is mostly affected by emissions of precursors in the rest of Europe.
+\citet{ACP:Andersson2017} presented in a modeling study focused on Fennoscandia that the variability of ozone concentrations in winter can be mainly attributed to changes in atmospheric background ozone concentrations determined by circulation patterns and atmospheric transport, whilst summertime abundance is mostly affected by emissions of precursors in the rest of Europe.
 For example, in northern Fennoscandia, episodes of elevated summertime ozone concentrations in 2003 and 2006 were attributed to ozone precursor emissions from forest fires in southern and eastern Europe, respectively \citep{AE:Lindskog2007,EP:Karlsson2013}.
@@
 A longer growing season is prolonging the time in which vegetation can accumulate an ozone dose.
-For natural and semi-natural vegetation types that is already subject to rapid climatic changes known as Arctic amplification \citep{AMAP2012,IPCC2013}, these factors could promote a higher potential risk for northern vegetation to ozone-induced damage in the not so distant future.
+For natural and semi-natural vegetation types that are already subject to rapid climatic changes known as Arctic amplification \citep{AMAP2012,IPCC2013}, these factors could promote a higher potential risk to ozone-induced damage in the not so distant future.
 Recently, \citet{ESPR:Hayes2021} showed that the highest sensitivity to future \chem{[O_3]} variability is expected in summer when vegetation is most productive.
@@
 Ozone enters the leaves through stomata, leaf pores that enable gas exchange, allowing the entry of \chem{CO_2} for photosynthesis and loss of \chem{H_2O} vapor via the plant transpiration stream.
-Stomatal aperture, and hence stomatal conductance to these gases (including ozone) will vary over the day and growing season primarily to balance \chem{CO_2} uptake against \chem{H_2O} vapor loss.
+Stomata aperture and hence stomatal conductance to these gases (including ozone) will vary over the day and growing season primarily to balance \chem{CO_2} uptake against \chem{H_2O} vapor loss.
 The higher the stomatal conductance, the higher the potential for ozone uptake.
 Stomatal conductance has been empirically linked to key environmental variables such as air temperature $T$, photosynthetic photon flux density (PPFD), vapor pressure deficit (VPD), and soil water potential (SWP) as well as photosynthesis itself \citep[e.g.,][]{PTRS:Jarvis1976, BallBerry1987, Emberson2000, ICP:MappingManual2017}.
-Results from open-top chamber (OTC) experiments on downy birch (\emph{Betula pubescens}) and mountain birch (\emph{Betula pubescens toruosa}), native to subarctic regions, indicated reductions in both biomass, root:shoot ratio, and visible leaf damage under elevated ozone treatments ($\left<\chem{\chi_{O_3}}\right> = 36-54\,\unit{ppb}$) \citep{Amb:Manninen2009}.
-Although, Scots pine (\emph{Pinus sylvestris}) is considered to be more ozone tolerant due to an absence of visible injuries \citep{Amb:Girgzdiene2009}, \citet{Amb:Manninen2009} found chlorophyll:carotenoid ratio and polyamine reductions under elevated ozone concentrations indicating susceptibility to ozone also in these species.
 
@@
 Currently, parameterizations are available for some species for particular climatic regions e.g.
-Boreal, temperate and Mediterranean; there is no parameterization for sub arctic vegetation.
-Given the unique conditions of the sub arctic (i.e.
+Boreal, temperate and Mediterranean; there is no parameterization for subarctic vegetation.
+Given the unique conditions of the subarctic (i.e.
 long duration of daylight in the summer and cold temperatures) it would be expected that this vegetation type has a particular physiological response to the prevailing climatic conditions different to boreal species.
@@
 An application of the Mapping Manual method was made by \citet{GCB:Mills2011} where exceedance of the CL was related to observations of ozone damage to a clover bio-monitor.
-This study demonstrated the improved ability of flux-based metrics (in comparison to exposure-based metrics) to identify the geographical distribution of the risk of ozone damage and also showed that the ozone damage could extend into more northerly regions of Europe.
-
-Flux-based methods of risk assessment rely on accurate representation of plant physiological responses to key environmental variables and are most often parameterized using data for less extreme climates, i.e., temperate, continental, or Boreal climates.
-Parameterization for more extreme bio-geographical regions such as subarctic climates has not been performed and relies on assuming that Boreal parameterization will be applicable.
-In our study, we explore how risk might be over- or underestimated using Boreal parameterization and explore a local climate derived temperature response function.
-We also study the impact of the opening extent of stomata at low light intensities (light response function).\\
-
-In Section~\ref{sec:dam}, we give a brief summary of the key components and  drivers of the $\mathrm{DO_3SE}$ model that is used to assess the ozone risk.
-To study the influence of selected parameters on ozone risk in the subarctic, we propose a method to derive temperature response functions based on local climate and to infer start and end of the growing season from satellite observed gross primary productivity (GPP).
-Results of the ozone risk modelling are presented in Section~\ref{sec:results}.
-For evaluating these results, we assess the climate anomaly of weather conditions in the 2018/19 GS at our research site in Section~\ref{sec:stats}.
-In Section~\ref{sec:conc}, we discuss these results concerning prevailing conditions and give an outlook on how future assessments of pollution risk in the subarctic could be improved.
+This study demonstrated the improved ability of flux-based metrics (in comparison to exposure-based metrics) to identify the geographical distribution of the risk of ozone damage and also showed that the ozone damage could extend into more northerly regions of Europe.\\
+
+In Section~\ref{sec:dam}, we give a brief summary of the key components and drivers of the $\mathrm{DO_3SE}$ model that we used to assess the ozone risk.
+Subsequently, we propose a local climate derived temperature response function and explore the impact of the stomata aperture at low light intensities.
+Start and end of the growing season are inferred from satellite observed gross primary productivity (GPP).
+We explore how risk might be over- or underestimated using Boreal parameterization (Section~\ref{sec:results}).
+For evaluating these results in a climate perspective, we assess the climate anomalies of weather conditions and ozone concentrations in the 2018/19 GS at our research site (Section~\ref{sec:stats}).
+In Section~\ref{sec:conc}, we discuss our results concerning prevailing conditions and give an outlook on how future assessments of pollution risk in the subarctic could be improved.
 
@@
 
-The opening of stomata referred to as stomatal conductance $g_\mathrm{sto}$ is governed by different species-specific factors and responds to key environmental variables.
+The aperture of stomata referred to as stomatal conductance $g_\mathrm{sto}$ is governed by different species-specific factors and responds to key environmental variables.
 In the $\mathrm{DO_3SE}$ model, $g_\mathrm{sto}$ is described by a multiplicative empirical model \citep{PTRS:Jarvis1976,Emberson2000,ICP:MappingManual2017}:
@@
 %
-where $\alpha_\mathrm{light}$ is a slope parameter describing the extent of stomatal opening at low light intensities.
+where $\alpha_\mathrm{light}$ is a slope parameter describing the aperture of stomata at low light intensities.
 The wavelength band $400-700\,\unit{nm}$ to which plant chlorophyll responds to is called photosynthetic active radiation (PAR).
@@
 
-For localizing $f_\mathrm{light}$, we presumed that the opening extent of stomata at low light conditions differs in subarctic species compared to species in less extreme climates.
+For localizing $f_\mathrm{light}$, we presumed that the aperture of stomata at low light conditions differs in subarctic species compared to species in less extreme climates.
 We, therefore, adjusted the modelled light sensitivity of stomata such that the PPFD value needed to reach $50\,\unit{\%}$ opening is varied by $\pm 20\,\unit{\%}$.
@@
 In all other model experiments SWP was hardly of relevance.
-An earlier opening of stomata at low light intensities (PPFD0.8) causes an increase in $\mathrm{POD_1}$ compared to MM while a later opening (PPFD1.2) reduces $\mathrm{POD_1}$.
+A larger aperture of stomata at low light intensities (PPFD0.8) causes an increase in $\mathrm{POD_1}$ compared to MM while a smaller aperture (PPFD1.2) reduces $\mathrm{POD_1}$.
 Due to the shape of $f_\mathrm{light}$, a symmetric variation can lead to an asymmetric response in $\mathrm{POD_1}$.
@@
 Herein, the difference of PPFD0.8 compared to MM $f_\mathrm{light}$ is denoted as superscript and for PPFD1.2 as subscript.
-The extent of the opening angle ($\angle(+,0,-)$) indicates the sensitivity visually.
+The aperture angle ($\angle(+,0,-)$) indicates the sensitivity visually.
 A low soil water potential caused a reduction in $\mathrm{POD_1}$ in 2018 in particular when using PPFD0.8.
@@
 For comparison, the corresponding biomass reduction for the MM parameterization averaged over both years is shown.
-In this case, standard deviation is computed from all sensitivity simulations ($\pm 20\,\unit{\%}$ in extent of stomatal opening at low light, SWP on/off, and localized GS).}
+In this case, standard deviation is computed from all sensitivity simulations ($\pm 20\,\unit{\%}$ aperture of stomata at low light, SWP on/off, and localized GS).}
   \label{tab:biomass_reduction}
@@
 ..
+Results from open-top chamber (OTC) experiments on downy birch (\emph{Betula pubescens}) and mountain birch (\emph{Betula pubescens toruosa}), native to subarctic regions, indicated reductions in both biomass, root:shoot ratio, and visible leaf damage under elevated ozone treatments ($\left<\chem{\chi_{O_3}}\right> = 36-54\,\unit{ppb}$) \citep{Amb:Manninen2009}.
+Although, Scots pine (\emph{Pinus sylvestris}) is considered to be more ozone tolerant due to an absence of visible injuries \citep{Amb:Girgzdiene2009}, \citet{Amb:Manninen2009} found chlorophyll:carotenoid ratio and polyamine reductions under elevated ozone concentrations indicating susceptibility to ozone also in these species.
+..
 

### 2022-05-30 VegOzone.tex
Reforge methodes according to Lisa

@@
 \label{sec:dam}
-In the following, we briefly describe the model used for assessing the ozone risk, its key components, input data, and gap filling procedure (Section~\ref{sec:do3se}).
-To study the impact of model parameters, we propose a method for parameter localization to a subarctic climate based on existing Boreal parameterizations and local climate conditions.
+In the following, we briefly describe the $\mathrm{DO_3SE}$ model used for assessing the ozone risk, its key components, input data, and gap filling procedure (Section~\ref{sec:do3se}).
+To study the impact of model parameters, we propose a method to define a parameterization for a subarctic climate based on existing Boreal parameterizations and local climate conditions and vary the stomata aperture at low light intensities.
 We estimate the start and end of the growing season from satellite derived GPP (Section~\ref{sec:localization}).
@@
 \label{sec:do3se}
-We use the well established Deposition of Ozone for Stomatal Exchange ($\mathrm{DO_3SE}$) model \citep{ACP:Bueker2012} which computes estimates of the total and stomatal deposition of \chem{O_3} to selected land-cover types and plant species according to UNECE Long-Range Transboundary Air Pollution (LRTAP) methodologies for effects-based risk assessment.
+We use the well established Deposition of Ozone for Stomatal Exchange ($\mathrm{DO_3SE}$) model \citep{ACP:Bueker2012} which computes estimates of the total and stomatal deposition of \chem{O_3} to selected land-cover types and plant species according to UNECE LRTAP methodologies for effects-based risk assessment.
 
@@
 
-An estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy $h$ represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf and key environmental variables affecting stomatal conductance.
+In the $\mathrm{DO_3SE}$ model, the stomatal \chem{O_3} flux $\Phi_\mathrm{sto}$ in Eq.~(\ref{eq:pod}) is a fraction of the total \chem{O_3} deposition flux.
+A standard resistance analogous approach \citep{AE:Wesely1989} is used to transport \chem{O_3} from the atmospheric reference height $h_0$ to the vegetated surface.
+The calculation of aerodynamic $R_a$ and quasi-laminar boundary layer resistance $R_b$ follow \citet{WASP:Simpson2003,ACP:Simpson2012}.
+The surface resistance $R_\mathrm{surf}$ consists of stomatal $r_\mathrm{sto}$ and non-stomatal canopy resistances that include external plant surfaces $r_\mathrm{ext}$, aerodynamic within-canopy $R_\mathrm{inc}$ and ground surfaces or soil resistances $R_\mathrm{gs}$ \citep{WASP:Simpson2003,ACP:Simpson2012}.
+Under the assumption that the ozone concentration \chem{[O_3]} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf, $r_\mathrm{sto}$ and $r_\mathrm{ext}$ are calculated per leaf/needle area (note lower case $r$).
+%
+\begin{equation}
+  R_\mathrm{sur} = \frac{1}{\frac{\mathrm{LAI}}{r_\mathrm{sto}}+\frac{\mathrm{SAI}}{r_\mathrm{ext}}+\frac{1}{R_\mathrm{inc}+R_\mathrm{gs}}}
+\end{equation}
+%
+Stomatal resistance $r_\mathrm{sto}$ is the inverse of stomatal conductance $g_\mathrm{sto}$ in Eq.~(\ref{eq:stomatal}).
+
+The aperture of stomata referred to as stomatal conductance $g_\mathrm{sto}$ is governed by different species-specific factors and responds to key environmental variables.
+In the $\mathrm{DO_3SE}$ model, $g_\mathrm{sto}$ is described by a multiplicative empirical model \citep{PTRS:Jarvis1976,Emberson2000,ICP:MappingManual2017}:
+\begin{equation}
+  g_\text{sto} = g_\text{max} \cdot f_\text{phen} \cdot f_\text{light} \cdot \max{\left\{f_\text{min}, f_\mathrm{T} \cdot f_\mathrm{VPD} \cdot f_\text{SWP}\right\}}.
+  \label{eq:stomatal}
+\end{equation}
+%
+Where $g_\mathrm{max}$ is a species-specific maximum stomatal conductance value that is modified by factors that vary within a range of $0-1$.
+These factors are empirically determined and account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_\mathrm{T}$), water vapor pressure deficit ($f_\mathrm{VPD}$), and soil water potential ($f_\text{SWP}$).
+All factors differ with plant functional type (PFT).
+
+The leaf phenology factor ($f_\mathrm{phen}$) is a stepwise defined linear function that allows for an increase of stomatal conductance in spring (start of growing season) and a decrease in fall (end of growing season).
+
+The temperature response function is defined as
+%
+\begin{equation}
+  f_\mathrm{T} = \frac{T_\text{air}-T_\text{min}}{T_\text{opt}-T_\text{min}} \cdot \left(\frac{T_{\text{max}}-T_\text{air}}{T_{\text{max}}-T_\text{opt}}\right)^\beta,
+  \label{eq:f_temp}
+\end{equation}
+with $\beta = \frac{T_\text{max}-T_\text{opt}}{T_\text{opt}-T_\text{min}}$ a mixed reciprocal polynomial function of temperature.
+The shape parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various PFT and $T_\mathrm{air}$ is the near-surface air temperature.
+All temperatures are defined in units of \unit{^\circ C}.
+
+The water vapor pressure deficit function is
+%
+\begin{equation}
+  f_\mathrm{VPD} = f_\text{min}+(1-f_\text{min}) \cdot \frac{D_\text{min} - \text{VPD}}{D_\text{min}-D_\text{max}}
+  \label{eq:f_vpd}
+\end{equation}
+%
+where VPD is the leaf to air vapor pressure deficit in \unit{kPa} with a minimum conductance $f_\text{min}$.
+$D_\text{min}$ and $D_\text{max}$ describe the relative stomatal conductance to changes in vapor pressure deficit.
+
+The light response function $f_\mathrm{light}$ is given by
+%
+\begin{equation}
+  f_\text{light} = 1-\text{exp}(-\alpha_\text{light}\cdot \text{PPFD})
+  \label{eq:flight}
+\end{equation}
+%
+where $\alpha_\mathrm{light}$ is a slope parameter describing the aperture of stomata at low light intensities.
+The wavelength band $400-700\,\unit{nm}$ to which plant chlorophyll responds to is called photosynthetic active radiation (PAR).
+Its integral is the photosynthetic photon flux density (PPFD).
+
+In the $\mathrm{DO_3SE}$ model, Soil Water Potential (SWP) across a PFT specific root depth is simulated according to the Penman--Monteith energy balance method that drives water cycling through the soil--plant--atmosphere system \citep{ACP:Bueker2012}.
+The stomatal response to changes in SWP ($f_\mathrm{SWP}$) are, thus, given by:
+%
+\begin{equation}
+  f_\mathrm{SWP} = \text{min}\left\{1, \text{max}\left\{f_\mathrm{min}, \frac{(1 - f_\mathrm{min})\cdot(\mathrm{SWP_{min}} - \mathrm{SWP})}{\mathrm{SWP_{min}} - \mathrm{SWP_{max}}} + f_\mathrm{min} \right\} \right\}
+    \label{eq:fsw}
+\end{equation}
+%
+where $\mathrm{SWP_{min}}$ and $\mathrm{SWP_{max}}$ are the parameters describing the $f_\mathrm{SWP}$ relationship.
+
+Stomatal \chem{O_3} flux $\Phi_\mathrm{sto}$ for a sunlit upper canopy leaf (vegetation height $h$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy $h$ represents a reasonable estimate of the \chem{O_3} concentration at the upper surface of the laminar layer:
+%
+\begin{equation}
+  \Phi_\mathrm{sto} = \chem{[O_3]}(h) \cdot \frac{g_\mathrm{sto}}{g_\mathrm{sto}+g_\mathrm{ext}} \cdot \frac{1}{r_b+r_c}.
+  \label{eq:flux_stomata}
+\end{equation}
+
 $\mathrm{POD_y}$ is then calculated according to:
@@
 $\mathrm{POD_y}$ is given in \unit{mmol\,m^{-2}} integrated over the growing season.
-
-In the $\mathrm{DO_3SE}$ model, the stomatal flux $\Phi_\mathrm{sto}$ in Eq.~(\ref{eq:pod}) is a fraction of the total \chem{O_3} deposition flux.
-A standard resistance analogous approach \citep{AE:Wesely1989} is used to transport \chem{O_3} from the atmospheric reference height $h_0$ to the vegetated surface.
-The calculation of aerodynamic $R_a$ and quasi-laminar boundary layer resistance $R_b$ follow \citet{WASP:Simpson2003,ACP:Simpson2012}.
-The surface resistance $R_\mathrm{surf}$ consists of stomatal $r_\mathrm{sto}$ and non-stomatal canopy resistances that include external plant surfaces $r_\mathrm{ext}$, aerodynamic within-canopy $R_\mathrm{inc}$ and ground surfaces or soil resistances $R_\mathrm{gs}$ \citep{WASP:Simpson2003,ACP:Simpson2012}.
-Under the assumption that the ozone concentration \chem{[O_3]} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf, $r_\mathrm{sto}$ and $r_\mathrm{ext}$ are calculated per leaf/needle area (note lower case $r$).
-%
-\begin{equation}
-  R_\mathrm{sur} = \frac{1}{\frac{\mathrm{LAI}}{r_\mathrm{sto}}+\frac{\mathrm{SAI}}{r_\mathrm{ext}}+\frac{1}{R_\mathrm{inc}+R_\mathrm{gs}}}
-\end{equation}
-%
-Stomatal resistance $r_\mathrm{sto}$ is the inverse of stomatal conductance $g_\mathrm{sto}$ in Eq.~(\ref{eq:stomatal}).
-Accounting for these the effective stomatal flux $\Phi_\mathrm{sto}$ at the vegetation height $h$ is defined as
-%
-\begin{equation}
-  \Phi_\mathrm{sto} = \chem{[O_3]}(h) \cdot \frac{g_\mathrm{sto}}{g_\mathrm{sto}+g_\mathrm{ext}} \cdot \frac{1}{r_b+r_c}.
-  \label{eq:flux_stomata}
-\end{equation}
-A linear gradient adjusting \chem{[O_3]} from $h_0$ to $h$ was adopted in the $\mathrm{DO_3SE}$ model based on detailed transport studies using the chemical transport model EMEP \citep{ACP:Simpson2012,ICP:MappingManual2017}.
-
-The aperture of stomata referred to as stomatal conductance $g_\mathrm{sto}$ is governed by different species-specific factors and responds to key environmental variables.
-In the $\mathrm{DO_3SE}$ model, $g_\mathrm{sto}$ is described by a multiplicative empirical model \citep{PTRS:Jarvis1976,Emberson2000,ICP:MappingManual2017}:
-\begin{equation}
-  g_\text{sto} = g_\text{max} \cdot f_\text{phen} \cdot f_\text{light} \cdot \max{\left\{f_\text{min}, f_\mathrm{T} \cdot f_\mathrm{VPD} \cdot f_\text{SWP}\right\}}.
-  \label{eq:stomatal}
-\end{equation}
-%
-Where $g_\mathrm{max}$ is a species-specific maximum stomatal conductance value that is modified by factors that vary within a range of $0-1$.
-These factors are empirically determined and account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_\mathrm{T}$), water vapor pressure deficit ($f_\mathrm{VPD}$), and soil water potential ($f_\text{SWP}$).
-All factors differ with plant functional type (PFT).
-
-The leaf phenology factor ($f_\mathrm{phen}$) is a stepwise defined linear function that allows for an increase of stomatal conductance in spring (start of growing season) and a decrease in fall (end of growing season).
-Mid anthesis and mid season dips in stomata conductance can also be accounted for but are not relevant in this study.
-
-The temperature response function is defined as
-%
-\begin{equation}
-  f_\mathrm{T} = \frac{T_\text{air}-T_\text{min}}{T_\text{opt}-T_\text{min}} \cdot \left(\frac{T_{\text{max}}-T_\text{air}}{T_{\text{max}}-T_\text{opt}}\right)^\beta,
-  \label{eq:f_temp}
-\end{equation}
-with $\beta = \frac{T_\text{max}-T_\text{opt}}{T_\text{opt}-T_\text{min}}$ a mixed reciprocal polynomial function of temperature.
-The shape parameters $T_\text{min}$, $T_\text{max}$ and $T_\text{opt}$ are tabulated for various PFT and $T_\mathrm{air}$ is the near-surface air temperature.
-All temperatures are defined in units of \unit{^\circ C}.
-
-The water vapor pressure deficit function is
-%
-\begin{equation}
-  f_\mathrm{VPD} = f_\text{min}+(1-f_\text{min}) \cdot \frac{D_\text{min} - \text{VPD}}{D_\text{min}-D_\text{max}}
-  \label{eq:f_vpd}
-\end{equation}
-%
-where VPD is the leaf to air vapor pressure deficit in \unit{kPa} with a minimum conductance $f_\text{min}$.
-$D_\text{min}$, $D_\text{max}$ describe the relative stomatal conductance to changes in vapor pressure deficit.
-
-The light response function $f_\mathrm{light}$ is given by
-%
-\begin{equation}
-  f_\text{light} = 1-\text{exp}(-\alpha_\text{light}\cdot \text{PPFD})
-  \label{eq:flight}
-\end{equation}
-%
-where $\alpha_\mathrm{light}$ is a slope parameter describing the aperture of stomata at low light intensities.
-The wavelength band $400-700\,\unit{nm}$ to which plant chlorophyll responds to is called photosynthetic active radiation (PAR).
-Its integral is the photosynthetic photon flux density (PPFD).
-
-In the $\mathrm{DO_3SE}$ model, Soil Water Potential (SWP) across a PFT specific root depth is simulated according to the Penman--Monteith energy balance method that drives water cycling through the soil--plant--atmosphere system \citep{ACP:Bueker2012}.
-The stomatal response to changes in SWP ($f_\mathrm{SWP}$) are, thus, given by:
-%
-\begin{equation}
-  f_\mathrm{SWP} = \text{min}\left\{1, \text{max}\left\{f_\mathrm{min}, \frac{(1 - f_\mathrm{min})\cdot(\mathrm{SWP_{min}} - \mathrm{SWP})}{\mathrm{SWP_{min}} - \mathrm{SWP_{max}}} + f_\mathrm{min} \right\} \right\}
-    \label{eq:fsw}
-\end{equation}
-%
-where $\mathrm{SWP_{min}}$ and $\mathrm{SWP_{max}}$ are the parameters describing the $f_\mathrm{SWP}$ relationship.
 

### 2022-05-31 VegOzone.tex
Fine tuning

@@
 In the following, we briefly describe the $\mathrm{DO_3SE}$ model used for assessing the ozone risk, its key components, input data, and gap filling procedure (Section~\ref{sec:do3se}).
-To study the impact of model parameters, we propose a method to define a parameterization for a subarctic climate based on existing Boreal parameterizations and local climate conditions and vary the stomata aperture at low light intensities.
-We estimate the start and end of the growing season from satellite derived GPP (Section~\ref{sec:localization}).
+To study the impact of model parameters, we propose a method to define a parameterization for a subarctic climate based on existing Boreal parameterizations and local climate conditions, study the impact of stomata aperture at low light intensities, and estimate the start and end of the growing season from satellite derived GPP (Section~\ref{sec:localization}).
 
@@
 
-In the $\mathrm{DO_3SE}$ model, the stomatal \chem{O_3} flux $\Phi_\mathrm{sto}$ in Eq.~(\ref{eq:pod}) is a fraction of the total \chem{O_3} deposition flux.
+In the $\mathrm{DO_3SE}$ model, the stomatal \chem{O_3} flux $\Phi_\mathrm{sto}$ is a fraction of the total \chem{O_3} deposition flux $\Phi_\chem{O_3}$.
 A standard resistance analogous approach \citep{AE:Wesely1989} is used to transport \chem{O_3} from the atmospheric reference height $h_0$ to the vegetated surface.
 The calculation of aerodynamic $R_a$ and quasi-laminar boundary layer resistance $R_b$ follow \citet{WASP:Simpson2003,ACP:Simpson2012}.
-The surface resistance $R_\mathrm{surf}$ consists of stomatal $r_\mathrm{sto}$ and non-stomatal canopy resistances that include external plant surfaces $r_\mathrm{ext}$, aerodynamic within-canopy $R_\mathrm{inc}$ and ground surfaces or soil resistances $R_\mathrm{gs}$ \citep{WASP:Simpson2003,ACP:Simpson2012}.
-Under the assumption that the ozone concentration \chem{[O_3]} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf, $r_\mathrm{sto}$ and $r_\mathrm{ext}$ are calculated per leaf/needle area (note lower case $r$).
-%
-\begin{equation}
-  R_\mathrm{sur} = \frac{1}{\frac{\mathrm{LAI}}{r_\mathrm{sto}}+\frac{\mathrm{SAI}}{r_\mathrm{ext}}+\frac{1}{R_\mathrm{inc}+R_\mathrm{gs}}}
-\end{equation}
-%
-Stomatal resistance $r_\mathrm{sto}$ is the inverse of stomatal conductance $g_\mathrm{sto}$ in Eq.~(\ref{eq:stomatal}).
-
-The aperture of stomata referred to as stomatal conductance $g_\mathrm{sto}$ is governed by different species-specific factors and responds to key environmental variables.
+The surface resistance $R_\mathrm{surf}$ consists of stomatal $r_\mathrm{sto}$ and non-stomatal canopy resistances that include external plant surfaces $r_\mathrm{ext}$, aerodynamic within-canopy $R_\mathrm{inc}$, and ground surfaces or soil resistances $R_\mathrm{gs}$ \citep{WASP:Simpson2003,ACP:Simpson2012}.
+The lower case indicates that $r_\mathrm{sto}$ and $r_\mathrm{ext}$ are calculated per leaf/needle area.
+%
+\begin{equation}
+  R_\mathrm{sur} = \frac{1}{\frac{\mathrm{LAI}}{r_\mathrm{sto}}+\frac{\mathrm{SAI}}{r_\mathrm{ext}}+\frac{1}{R_\mathrm{inc}+R_\mathrm{gs}}},
+\end{equation}
+%
+with stomatal resistance $r_\mathrm{sto}$ as the inverse of the stomatal conductance $g_\mathrm{sto}$.
+
+The aperture of stomata referred to as $g_\mathrm{sto}$ is governed by different species-specific factors and responds to key environmental variables.
 In the $\mathrm{DO_3SE}$ model, $g_\mathrm{sto}$ is described by a multiplicative empirical model \citep{PTRS:Jarvis1976,Emberson2000,ICP:MappingManual2017}:
@@
 \begin{equation}
-  f_\mathrm{VPD} = f_\text{min}+(1-f_\text{min}) \cdot \frac{D_\text{min} - \text{VPD}}{D_\text{min}-D_\text{max}}
+  f_\mathrm{VPD} = f_\text{min}+(1-f_\text{min}) \cdot \frac{D_\text{min} - \text{VPD}}{D_\text{min}-D_\text{max}},
   \label{eq:f_vpd}
@@
 \begin{equation}
-  f_\text{light} = 1-\text{exp}(-\alpha_\text{light}\cdot \text{PPFD})
+  f_\text{light} = 1-\text{exp}(-\alpha_\text{light}\cdot \text{PPFD}),
   \label{eq:flight}
@@
 
-In the $\mathrm{DO_3SE}$ model, Soil Water Potential (SWP) across a PFT specific root depth is simulated according to the Penman--Monteith energy balance method that drives water cycling through the soil--plant--atmosphere system \citep{ACP:Bueker2012}.
-The stomatal response to changes in SWP ($f_\mathrm{SWP}$) are, thus, given by:
-%
-\begin{equation}
-  f_\mathrm{SWP} = \text{min}\left\{1, \text{max}\left\{f_\mathrm{min}, \frac{(1 - f_\mathrm{min})\cdot(\mathrm{SWP_{min}} - \mathrm{SWP})}{\mathrm{SWP_{min}} - \mathrm{SWP_{max}}} + f_\mathrm{min} \right\} \right\}
+In the $\mathrm{DO_3SE}$ model, soil water potential (SWP) across a PFT specific root depth is simulated according to the Penman--Monteith energy balance method that drives water cycling through the soil--plant--atmosphere system \citep{ACP:Bueker2012}.
+The stomatal response to changes in SWP $f_\mathrm{SWP}$ are given by:
+%
+\begin{equation}
+  f_\mathrm{SWP} = \text{min}\left\{1, \text{max}\left\{f_\mathrm{min}, \frac{(1 - f_\mathrm{min})\cdot(\mathrm{SWP_{min}} - \mathrm{SWP})}{\mathrm{SWP_{min}} - \mathrm{SWP_{max}}} + f_\mathrm{min} \right\} \right\},
     \label{eq:fsw}
@@
 
-Stomatal \chem{O_3} flux $\Phi_\mathrm{sto}$ for a sunlit upper canopy leaf (vegetation height $h$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy $h$ represents a reasonable estimate of the \chem{O_3} concentration at the upper surface of the laminar layer:
-%
-\begin{equation}
-  \Phi_\mathrm{sto} = \chem{[O_3]}(h) \cdot \frac{g_\mathrm{sto}}{g_\mathrm{sto}+g_\mathrm{ext}} \cdot \frac{1}{r_b+r_c}.
+$\Phi_\mathrm{sto}$ for a sunlit upper canopy leaf (vegetation height $h$) is then calculated under the assumption that the concentration of \chem{O_3} at the top of the canopy $h$ represents a reasonable estimate of the \chem{O_3} concentration at the upper surface of the laminar layer:
+%
+\begin{equation}
+  \Phi_\mathrm{sto}(t) = \chem{[O_3]}_h(t) \cdot \frac{g_\mathrm{sto}}{g_\mathrm{sto}+g_\mathrm{ext}} \cdot \frac{1}{r_b+r_c}.
   \label{eq:flux_stomata}
@@
 
-$\mathrm{POD_y}$ is then calculated according to:
-%
-\begin{equation}
-  \mathrm{POD_y} = \int{(\Phi_\mathrm{sto}-y)\cdot \mathrm{d}t},
+$\mathrm{POD_y}$ is then computed according to:
+%
+\begin{equation}
+  \mathrm{POD_y} = \int{(\Phi_\mathrm{sto}(t)-y)\cdot \mathrm{d}t},
   \label{eq:pod}
@@
 %
-with the hourly averaged stomatal ozone flux $\Phi_\mathrm{sto}$ (see Eq.~(\ref{eq:flux_stomata})) and a stomatal ozone flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}^{-1}\,s^{-1}}.
+with an hourly averaged stomatal ozone flux $\Phi_\mathrm{sto}(t)$ (Eq.~(\ref{eq:flux_stomata})) and a stomatal ozone flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}^{-1}\,s^{-1}}.
 Herein PLA stands for Projected~Leaf~Area.
@@
 when global radiation is above $50\,\unit{W\,m^{-2}}$).
-$\mathrm{POD_y}$ is given in \unit{mmol\,m^{-2}} integrated over the growing season.
+Finally, $\mathrm{POD_y}$ is given in \unit{mmol\,m^{-2}} integrated over the growing season.
 

### 2022-05-31 VegOzone.tex
Fine tuning and comments by Lisa

@@
 \label{subsec:gap_filling}
-The $\mathrm{DO_3SE}$ model requires hourly, continuous input data. Therefor, the following gap filling methodology was devised for the meteorological drivers of the model:
+The $\mathrm{DO_3SE}$ model requires hourly, continuous input data. Therefore, the following gap filling methodology was devised for the meteorological input data:
 \begin{itemize}
@@
 \label{subsec:data}
-The required input data consists of ozone concentration $\chem{[O_3]}$, $2\,\unit{m}$ temperature $T_\mathrm{2m}$, global irradiation $Q_0$, precipitation, $2\,\unit{m}$ wind $u_\mathrm{2m}$, and vapor pressure deficit VPD.
-VPD has been calculated from observed $T_\mathrm{2m}$ and relative humidity
-\begin{equation}
-  \mathrm{VPD} = P_s(T) \cdot \left(1-\frac{\mathrm{relHum}}{100}\right),
-\end{equation}
-with saturation vapor pressure $P_s(T)$ in \unit{hPa} from the Arden~Buck~equation~\citep{JAP:Buck1981, Buck2012}.
-
-For our study, we chose Svanhovd in northern Norway ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m\,a.s.l.}$).
+The required input data consists of ozone concentration $\chem{[O_3]}$, air temperature at $2\,\unit{m}$ $T_\mathrm{air}$, global irradiation $Q_0$, precipitation, $2\,\unit{m}$ wind $u_\mathrm{2m}$, and vapor pressure deficit VPD.
+VPD has been calculated from the Arden~Buck~equation~\citep{JAP:Buck1981, Buck2012}
+\begin{equation}
+  \mathrm{VPD} = P_s(T_\mathrm{air}) \cdot \left(1-\frac{\mathrm{relHum}}{100}\right),
+\end{equation}
+with saturation vapor pressure $P_s(T_\mathrm{air})$ in \unit{hPa}, observed $T_\mathrm{air}$, and relative humidity.
+
+For our study focuses on ozone impact on subarctic vegetation, we chose Svanhovd in northern Norway ($69.45\,\unit{^\circ N}$, $30.03\,\unit{^\circ E}$, $30\,\unit{m\,a.s.l.}$) as site.
 Reasons for our choice were: (1) Ambient ozone concentration had been monitored there in the past, (2) an agrometeorological station is measuring, e.g.
-$2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths, and (3) a national surveillance of air pollution (\chem{NO_x}, \chem{SO_x}, particular matter (PM), radiological species) is ongoing.
+$2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths, and (3) a national surveillance of air pollution monitoring (\chem{NO_x}, \chem{SO_x}, particular matter (PM), radiological species) is ongoing.
 Most of these measurements are performed by the Norwegian Institute for Air Research (NILU), who also conducted ozone measurements on our behalf during the 2018/19 growing season.
@@
 All essential climate variables for the 2018/19 growing season were provided by NILU through the web portal \href{luftkvalitet.no}{luftkvalitet.no} (last accessed May 2020).
-The ozone monitoring data have been added to a database for air-quality monitoring called EBAS as well operated by NILU and are thus openly accessible.
+The \chem{O_3} monitoring data have been added to a database for air-quality monitoring called EBAS operated by NILU and are thus openly accessible.
 Long-term ozone observation data (1986--1996) were obtained from EBAS.
 Note that data from before the 1990s do not follow the high-quality standard procedures implemented nowadays and have to be treated with care \citep{NILU2003}.
-Because ozone monitors did not undergo regular re-calibration drifts in the observed data occurred imposing false trends which are small but in the same order of magnitude as the expected trends in tropospheric background ozone.
+Because ozone monitors did not undergo regular re-calibration, drifts in the observed data occurred imposing false trends which are small but in the same order of magnitude as the expected trends in tropospheric background ozone.
 This does not concern the ozone seasonality which we are particularly interested in.
@@
 
-To relate local ozone concentration \chem{[O_3]} with visible damage on ozone sensitive plants, an ozone bio-monitor (referred to as ozone garden) has been installed at Svanhovd in the 2018/19 GS.
+To relate local ozone concentration \chem{[O_3]} with visible injuries on ozone sensitive plants, following \citet{GCB:Mills2011}, an ozone bio-monitor (referred to as ozone garden) has been installed at Svanhovd in the 2018/19 GS.
 The locations of the atmospheric monitoring site and the ozone garden are marked in the aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}a.
@@
 As shown in Fig.~\ref{fig:svanhovd_research_station}b, the plants had to be protected from herbivores with a wire-mesh fence.
-In 2018, we qualitatively observed visible ozone damage on semi-natural vegetation (clover) and crops (tobacco).
-In contrast, no such visible damage was found on the clovers in 2019, although it was observed on the sensitive tobacco cultivar (Bel-W3).
-Note that the extent of visible damage is not equivalent to a quantitative reduction of above/below ground biomass of a species.
+In 2018, we observed visible ozone injuries on semi-natural vegetation (clover) and crops (tobacco).
+In contrast, no such visible injuries were found on the clovers in 2019, although it was observed on the sensitive tobacco cultivar (Bel-W3).
+Note that the extent of visible injuries is, in general, not equivalent to a quantitative reduction of above/below ground biomass of a species.
 
@@
 For a comprehensive list of MM parameters, consult Supplement~Table~S1.
-In the following, we suggest localizations of key environmental parameters to a subarctic climate based on meteorological data to assess over-/underestimation of ozone damage when using MM Boreal parameterizations in ozone risk assessments.
+In the following, we suggest localizations of key environmental parameters to a subarctic climate based on meteorological data to assess over-/underestimation of ozone-induced reduction of biomass when using MM Boreal parameterizations in ozone risk assessments.
 
@@
 Along the conventional ozone monitoring at Svanhovd, we had installed a bio-monitoring with ozone sensitive species.
-Qualitatively, this showed visible ozone damage on clover and tobacco in 2018.
+Qualitatively, this showed visible ozone injuries on clover and tobacco in 2018.
 In contrast, no such damage was found on the clovers in 2019.
 This might indicate a higher ozone damage risk under warmer climate conditions for these specific plant species in a subarctic climate.
-However, visible damage is not always correlated with a reduction in biomass \citep[and citations therein]{CRG:Felzer2007}.
+However, visible injuries are not always correlated with a reduction in biomass \citep[and citations therein]{CRG:Felzer2007}.
 Also, our $\mathrm{DO_3SE}$ modeling results are in contrast to these observations at Svanhovd.
-Any visible damage caused by peak ozone flux through stomata or a specific dependency of plant sensibility to ozone on phenology stage are not accounted for in the $\mathrm{DO_3SE}$ model.
-In this regard, open-top chamber (OTC) experiments performed in northern Finland where Scots pine and downy birch seedlings were exposed to elevated ozone concentrations attributed a reduction in biomass and reproduction with visible damage explicitly to peak \chem{O_3} concentrations and fast phenological development at high growth rate \citep{Amb:Manninen2009}.
+Any visible injuries caused by peak ozone flux through stomata or a specific dependency of plant sensibility to ozone on phenology stage are not accounted for in the $\mathrm{DO_3SE}$ model.
+In this regard, open-top chamber (OTC) experiments performed in northern Finland where Scots pine and downy birch seedlings were exposed to elevated ozone concentrations attributed a reduction in biomass and reproduction with visible injuries explicitly to peak \chem{O_3} concentrations and fast phenological development at high growth rate \citep{Amb:Manninen2009}.
 Forbs and perennial grassland are more susceptible to ozone-induced damage in the reproductive state \citep{EP:Bassin2004}.
-That might explain why we observed damage in the ozone garden to a larger extent in 2018 than in 2019.
+That might explain why we observed visible injuries in the ozone garden to a larger extent in 2018 than in 2019.
 An associated damage function to ozone depending on leaf age as proposed by \citet{AE:Musselman2006} or phenological state might improve the predictions.

### 2022-05-31 VegOzone.tex
Refine and comments by Lisa

@@
 We determined the dominant species on-site at Svanhovd from Fig.~\ref{fig:svanhovd_research_station}a and found perennial grassland, birch (generalized as deciduous trees), and Scots pine (generalized as coniferous trees).
-Basic parameters for these PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as mapping manual (MM) parameterizations.
-The localization of parameters for coniferous trees is based on MM's Boreal Norway spruce, deciduous trees on MM's silver birch, and perennial grassland on MM's perennial grassland for Central Europe.
+Basic parameters for these PFTs are derived from Boreal paramterizations \citep{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as mapping manual (MM) parameterizations.
+The subarctic local climate derived parameterization for coniferous trees is based on MM's Boreal Norway spruce, deciduous trees on the MM's silver birch, and perennial grassland on MM's perennial grassland for Central Europe.
 For a comprehensive list of MM parameters, consult Supplement~Table~S1.
-In the following, we suggest localizations of key environmental parameters to a subarctic climate based on meteorological data to assess over-/underestimation of ozone-induced reduction of biomass when using MM Boreal parameterizations in ozone risk assessments.
 
@@
 
-Initial simulations with the MM parameterizations for perennial grassland revealed an unexpected low stomatal conductance at leaf-level ($G_\mathrm{sto}^\mathrm{leaf}$) in 2019 (Appendix~Fig.~\ref{fig:pody_mm_composit}).
+Initial simulations with the Central European (Boreal is not available) MM parameterizations for perennial grassland revealed an unexpectedly low leaf level stomatal conductance ($G_\mathrm{sto}^\mathrm{leaf}$) in 2019 (Appendix~Fig.~\ref{fig:pody_mm_composit}).
 Substantial stomatal conductance occurred only during an extended warm period in late July.
@@
 We identified $f_\mathrm{T}$ as abnormally low throughout the season, being the limiting factor of stomatal conductance in perennial grassland.
-This suggested that a localization of the parameterization to a subarctic climate was necessary.
-
-To localize the temperature response function $f_\mathrm{T}$, we divided the temperature dataset into two subsets: 1992--2000 (near past) and 2011--2019 (present).
+This suggested that deriving a local climate parameterization is necessary for appropriate modeling.
+
+To derive a subarctic parameterization for the temperature response function $f_\mathrm{T}$, we divided the temperature dataset into two subsets: 1992--2000 (near past) and 2011--2019 (present).
 From these two subsets, we selected only dates within the GS (May--September) and plotted the probability density functions (PDFs) as histograms with $1\,\unit{^\circ C}$ binning (Fig.~\ref{fig:f_temp_grassland}a).
-A comparison between the two PDFs indicates that the temperature distribution has shifted towards higher temperatures reducing the number of days below $5\,\unit{^\circ C}$.
-It is unclear how fast the species composition of perennial grassland responds to these changes.
+A comparison between the two PDFs indicates that during the past decade the temperature distribution has shifted towards higher temperatures reducing the number of days below $5\,\unit{^\circ C}$.
+It is unclear how fast the species composition or physiology via acclimation of perennial grassland responds to these changes.
 Thus, the predominant acclimation at present is not known.
@@
 
-Hence, the idea was to define localized temperature response functions by increasing the overlapping area between $f_\mathrm{T}$ and the temperature PDF during the GS.
-To this end, we defined two different hypothetical target species: subarctic and cold.
-We constructed \emph{cold} as representative for a species that is more tolerant to cold temperatures than MM, but slightly less efficient at warm temperatures than MM.
+We defined local climate derived parameterizations for two hypothetical species (\emph{subarctic} and \emph{cold}) by increasing the overlapping area between $f_\mathrm{T}$ and the temperature PDF during the GS.
+We constructed \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures than the MM species.
 This was accomplished by moving $T_\mathrm{min}$ and $T_\mathrm{opt}$ towards cooler temperatures but $T_\mathrm{max}$ was kept at its MM value.
@@
 Therefore, we shifted both $T_\mathrm{min}$ and $T_\mathrm{max}$ to colder temperatures and chose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000.
-The most extreme of the two localizations compared to MM is \emph{subarctic}.
+The most extreme of the two localizations compared to the MM is the \emph{subarctic}.
 
@@
 
-In Fig.~\ref{fig:f_temp_grassland}b, we show the PPFD PDF for near past and present together with the $f_\mathrm{light}$ for perennial grassland. The MM parameterization is displayed as solid line.
+In Fig.~\ref{fig:f_temp_grassland}b, we show the PPFD PDF for near past and present together with the $f_\mathrm{light}$ for perennial grassland.
+As no Boreal parameterization is available for perennial grassland, the Central European MM parameterization is displayed as solid line.
 
@@
 Coniferous trees are known to start photosynthesis as soon as the leaf temperature reaches a certain threshold \citep{TB:Kolari2007}.
-Thus, coniferous trees are already active at low air temperatures and can reach $60\,\unit{\%}$ photosynthetic activity as early as \unit{doy}~100 but cease activity in less favorable conditions \citep{TB:Kolari2007, TP:Wallin2013}.
+Thus, coniferous trees are already active at low air temperatures and can reach $60\,\unit{\%}$ photosynthetic activity as early as day of the year (\unit{doy})~100 but cease activity in less favorable conditions \citep{TB:Kolari2007, TP:Wallin2013}.
 To determine the temperature localization for coniferous trees, we used the time series of \chem{CO_2} uptake and temperature at observation sites in southern ($\rightarrow$ \emph{cold}) and northern ($\rightarrow$ \emph{subarctic}) Finland by \citet{TB:Kolari2007}.

### 2022-06-01 VegOzone.tex
Fine tuning and comment by Lisa

@@
 Coniferous trees are known to start photosynthesis as soon as the leaf temperature reaches a certain threshold \citep{TB:Kolari2007}.
-Thus, coniferous trees are already active at low air temperatures and can reach $60\,\unit{\%}$ photosynthetic activity as early as day of the year (\unit{doy})~100 but cease activity in less favorable conditions \citep{TB:Kolari2007, TP:Wallin2013}.
+Thus, coniferous trees are already active at low air temperatures and can reach $60\,\unit{\%}$ photosynthetic activity as early as day of the year~(\unit{doy})~100 but cease activity in less favorable conditions \citep{TB:Kolari2007, TP:Wallin2013}.
 To determine the temperature localization for coniferous trees, we used the time series of \chem{CO_2} uptake and temperature at observation sites in southern ($\rightarrow$ \emph{cold}) and northern ($\rightarrow$ \emph{subarctic}) Finland by \citet{TB:Kolari2007}.
@@
 The accumulation of ozone dose is strongly depended on the timing and length of the GS.
-\citet{ICP:MappingManual2017} suggest a simple latitude model for beech, birch, and Norway spruce for the localization of GS:
+\citet{ICP:MappingManual2017} suggest a simple latitude model for beech, birch, and Norway spruce to localize start and end of the GS:
 %
@@
 %
-At Svanhovd, this yields $A_\mathrm{start}=135\,\unit{doy}$ and $A_\mathrm{end}=257\,\unit{doy}$. We verified these as follows.
+At Svanhovd, this yields $A_\mathrm{start}=135~\unit{doy}$ and $A_\mathrm{end}=257~\unit{doy}$. We verified these as follows.
 
@@
 GPP over time follows approximately a second-order polynomial function (Fig.~\ref{fig:modis_gpp}).
-By calculating the roots of the fitted polynomial, we found $A_\text{start}$ as \unit{doy}~122 for 2018 and \unit{doy}~106 in 2019.
+By calculating the roots of the fitted polynomial, we found $A_\text{start}=122~\unit{doy}$ in 2018 and $A_\text{start}=106~\unit{doy}$ in 2019.
 $A_\text{end}$ amounts to \unit{doy}~261 and 274, respectively.
-This $A_\mathrm{end}$ value will be used for all PFTs alike.
+These $A_\mathrm{end}$ values will be used for all PFTs alike.
 The resulting growing season for coniferous trees in 2019 was one month longer than in 2018.
-For $A_\text{start}$ of deciduous trees, we use temperature-degree-days (above $5\,\unit{^\circ C}$) on gridded temperature data from SeNorge.no (Appendix~Fig.~\ref{fig:greening_season_change_Svanvik}).
+For $A_\text{start}$ of deciduous trees, we used temperature-degree-days (above $5\,\unit{^\circ C}$) on gridded temperature data from SeNorge.no (Appendix~Fig.~\ref{fig:greening_season_change_Svanvik}) which corresponds to the phenological timing of budburst for birch trees.
 We found \unit{doy}~129 and 130, respectively.
@@
 This assumption is supported by observations in Rovaniemi, Finland \citep[][Supplement~Fig.~S1]{FCR:Korhonen2018}.
+Assuming the linear trend for temperture-degree-days defined GS over the past 6~decades prevails, GS at Svanhovd could converge with GS in Central Europe.
 All results are tabulated in Table~\ref{tab:sensitivity_tests_gs}.
@@
 
-We did not perform a localization of the water vapor pressure deficite $f_\mathrm{VPD}$ and soil water potential $f_\mathrm{SWP}$ response functions.
+We did not  derive a local climate parameterization for the water vapour pressure deficit function $f_\mathrm{VPD}$ and soil water potential $f_\mathrm{SWP}$ response functions.
 But acknowledge that these are very important factors in the modeling of ozone dose as pointed out by~\citet{ACP:Bueker2012}.
@@
 They were collected by hand at about $2\,\mathrm{m}$ height.
-As the sun is at a low elevation angle for most of the time in the growing season in this area, light exposure at the top of the tree is not expected to be different from lower leaves as long as they grow in a part of the tree that is not shadowed by other trees.
+As the sun is at a low elevation angle for most of the time in the growing season in this area, light exposure at the top of the tree is not expected to be different from lower leaves as long as they grow in a part of the tree that is not shadowed by parts of the tree canopy.
 At the first site (Karasjok), three leaves were collected from five adult trees each.
@@
 The ozone VMR climatology (\chem{\chi_{O_3}}) (1986--1996) is shown in Figure~\ref{fig:climatediagram}d.
-The climatological spring peak occurs in April (around \unit{doy} 100) and amounts to $40\,\unit{ppb}$ while the annual average \chem{\chi_{O_3}} was $28\,\unit{ppb}$.
+The climatological spring peak occurs in April (around \unit{doy}~100) and amounts to $40\,\unit{ppb}$ while the annual average \chem{\chi_{O_3}} was $28\,\unit{ppb}$.
 The decline in \chem{\chi_{O_3}} roughly coincides with the beginning of \chem{CO_2} uptake by coniferous trees \citep{TB:Kolari2007, TP:Wallin2013}.

### 2022-06-01 VegOzone.tex
Lisa's comments

@@
 \label{sec:results}
-In this section, we first compute the average stomatal conductance $\left<g_\mathrm{sto}\right>$ relative to $g_\mathrm{max}$ for each PFT from the long term meteorological data according to Eq.~(\ref{eq:stomatal}) and compare the different localizations to MM (Section~\ref{subsec:res_local}).
-We use the $\mathrm{DO_3SE}$ model to compute the ozone dose and projected biomass loss due to ozone for the two recent GSs 2018/19 for each species and localization (Section~\ref{subsec:do3se_results}).
-Finally, we put the results into a climatological perspective (Section~\ref{subsec:climatologies}).
-
-\subsection{Parameter localization}
+In this section, we first estimate the multiannual average stomatal conductance $\left<g_\mathrm{sto}\right>$ relative to $g_\mathrm{max}$ for each PFT from the long term meteorological data according to Eq.~(\ref{eq:stomatal}) and compare the different local climate derived parameterizations to MM (Section~\ref{subsec:res_local}).
+We also use the $\mathrm{DO_3SE}$ model to compute the ozone dose and predicted biomass loss due to ozone for the two recent GSs 2018 and 2019 for each species and set of local parameters (Section~\ref{subsec:do3se_results}).
+Finally, we put the results on ozone damage risk in 2018 and 2019 into a climatological perspective meaning how common these conditions have been in the past.
+To this end, we analyze both the past seasonality and the deviation from the norm of temperature, precipitation radiant energy, and ozone for these two years (Section~\ref{subsec:climatologies}).
+
+\subsection{Local climate derived parameterization}
 \label{subsec:res_local}
@@
 Because the MM version of the $\mathrm{DO_3SE}$ model currently does not simulate net photosynthesis $A_\mathrm{n}$, we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011}.
-A higher average $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ ratio, thus, indicates a higher growth potential in the local climate.
-
-We propose the following metric.
-From the long term of meteorological data, we select two distinct periods: noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and morning ($5-9\,\unit{am}$ local time) and calculate the average relative stomatal conductance $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal}) assuming $f_\mathrm{SWP}=1$).
+A higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ ratio, thus, indicates a higher growth potential in the local climate.
+
+To compare the different parameterizations, we propose the following metric: From the selected set of meteorological data, we define two distinct periods in the diurnal cycle: noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and morning ($5-9\,\unit{am}$ local time) and calculate the $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ ratio (Eq.~(\ref{eq:stomatal}) assuming $f_\mathrm{SWP}=1$).
+That this choice is justified is shown in Section~\ref{subsec:do3se_results}.
 Presumably, $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ around noon (the highest light intensity) is a good proxy for \chem{CO_2} uptake efficiency (growth).
-In addition, a small deviation between noon and morning and a low standard deviation indicates higher robustness of this localization to variability in growing conditions.
-From left to right in Fig.~\ref{fig:javis_func_opt_mean}, results using the subarctic, cold, and MM $f_\mathrm{T}$ are shown marked by an alternating background color.
+In addition, a small deviation between noon and morning and a low standard deviation indicates a lower location induced variability in growing conditions.
+From left to right in Fig.~\ref{fig:javis_func_opt_mean}, results using the \emph{subarctic}, \emph{cold}, and MM $f_\mathrm{T}$ are shown highlighted by an alternating background color (linen-white-linen).
 Within each of these groups $f_\mathrm{light}$ categories are indicated by PPFD0.8 and PPFD1.2, respectively.
@@
 For perennial grassland, the differences compared to MM are largest suggesting that the MM Central European parameterization is not appropriate in a subarctic climate.
-Deciduous trees display also a notable increase in mean stomatal conductance using our localizations.
+Deciduous trees display also a notable increase in mean stomatal conductance using our local climate derived parameterizations.
 
@@
   %\includegraphics[width=8.3cm]{fig08}
-  \caption{Proposed metric to test local climate derived response functions.
+  \caption{Proposed metric to test local climate derived parameterizations for response functions.
 GS (May--August) climatological averages and standard deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon (averaged over $11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning (averaged over $5-9\,\unit{am}$ local time) are shown for (a) deciduous trees; (b) coniferous trees; (c) perennial grassland.}
@@
 
-We modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per PLA for the three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using the MM parameterization and our localized temperature (\emph{cold}, \emph{subarctic}) and light (PPFD0.8, PPFD1.2) parameterizations.
-
-The results are comprehensively displayed in Fig.~\ref{fig:pody_rel}.
-Critical levels for reduction in biomass \citep[deciduous forest $4\,\unit{\%}$, coniferous forest $2\,\unit{\%}$, and grasslands $10\,\unit{\%}$ biomass reduction adopted from][]{ICP:MappingManual2017,ESPR:Hayes2021} are shown as horizontal dashed lines.
-The different background colors indicate the temperature response function localization.
+We modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per PLA for the three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using the MM parameterization and our climate derived local temperature (\emph{cold}, \emph{subarctic}) and light (PPFD0.8, PPFD1.2) parameterizations.
+
+The results are displayed in Fig.~\ref{fig:pody_rel}.
+Critical levels for reduction in biomass \citep[deciduous forest $4\,\unit{\%}$, coniferous forest $2\,\unit{\%}$, and grasslands $10\,\unit{\%}$ reduction adopted from][]{ICP:MappingManual2017,ESPR:Hayes2021} are shown as horizontal dashed lines.
+The different background colors indicate the the local climate derived temperature response function.
 The light response function localizations are denoted by $+$ (PPFD1.2) and $-$ (PPFD0.8).
 Circles indicate simulations with localized GS, canopy height, and leaf dimensions.
-With open symbols, the effect on $\mathrm{POD_y}$ taking SWP (drought) into account is displayed.
-
-We identified in general that an increasing cold tolerance represented by a shift in the temperature response function (MM $<$ cold $<$ subarctic) leads to an increase in $\mathrm{POD_1}$ in both probed years compared to the MM parameterization.
+With open symbols, the effect on $\mathrm{POD_y}$ of any soil water stress is accounted for.
+
+We identified that, in general, an increasing cold tolerance represented by a shift in the temperature response function (MM~$<$~cold~$<$~subarctic) leads to an increase in $\mathrm{POD_1}$ in both investigated years (2018 and 2019) compared to the MM parameterization.
 The more extreme the temperature response function localization the higher $\mathrm{POD_1}$.
-In 2018, however, coniferous trees and perennial grassland display a reduction in $\mathrm{POD_1}$ for the subarctic parameterization compared to cold in responds to the prescribed lower $g_\mathrm{sto}$ at higher temperatures.
+In 2018, coniferous trees and perennial grassland display a reduction in $\mathrm{POD_1}$ for the subarctic parameterization compared to the cold paramerization in responds to the prescribed lower $g_\mathrm{sto}$ at higher temperatures.
+This is not seen in 2019.
 An earlier start and overall longer GS in 2019 leads to higher $\mathrm{POD_1}$ compared to 2018 despite the occurrence of more episodes of elevated \chem{O_3} in 2018 than in 2019 (see Section~\ref{subsec:anomalies} for details on the particular features of these two years).
@@
 In all other model experiments SWP was hardly of relevance.
-A larger aperture of stomata at low light intensities (PPFD0.8) causes an increase in $\mathrm{POD_1}$ compared to MM while a smaller aperture (PPFD1.2) reduces $\mathrm{POD_1}$.
+A larger aperture of stomata at low light intensities (PPFD0.8) causes an increase in $\mathrm{POD_1}$ compared to MM while a smaller aperture (PPFD1.2) is leading to reduced $\mathrm{POD_1}$.
 Due to the shape of $f_\mathrm{light}$, a symmetric variation can lead to an asymmetric response in $\mathrm{POD_1}$.
@@
 In 2019, the lowest $\mathrm{POD_1}$ estimate (MM, PPFD1.2) amounts to $10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and the highest (subarctic, PPFD0.8) $21\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
-In 2018, this range is smaller ($(12-18)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
+In 2018, this range is smaller $(12-18)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
 The sensitivity to $f_\mathrm{light}$ is largest in the subarctic parameterization and amounts to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
 Herein, the difference of PPFD0.8 compared to MM $f_\mathrm{light}$ is denoted as superscript and for PPFD1.2 as subscript.
-The aperture angle ($\angle(+,0,-)$) indicates the sensitivity visually.
 A low soil water potential caused a reduction in $\mathrm{POD_1}$ in 2018 in particular when using PPFD0.8.
-This indicates that drought conditions can reduce the ozone damage risk.
-
-For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL by $0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
-The different localizations do not show much impact compared to MM indicating that the MM's Boreal Norway spruce parameterization is well-suited for a subarctic climate.
+This indicates that water limiting conditions can reduce the ozone damage risk.
+
+For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL by $(0-10)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
+The different climate derived local parameterizations do not show much impact compared to MM indicating that the MM's Boreal Norway spruce parameterization is well-suited for a subarctic climate.
 We find higher $\mathrm{POD_1}$ in 2019 than in 2018.
@@
 By construction, $\mathrm{POD_1}$ from localized parmeterizations is always larger than estimates with the MM parameterization.
-This means that the ozone damage risk could be possibly underestimated especially in more normal years like 2019 (see Section~\ref{sec:stats} for details).
-Establishing dedicated subarctic parameterizations are a must to properly assess ozone damage risk in these regions.
+This means that the risk of ozone damage could be underestimated especially in more normal years like 2019 (see Section~\ref{sec:stats} for details).
+Establishing dedicated subarctic parameterizations are most likely required to properly assess risk.
 
@@
 
-We estimated the biomass reductions in accordance to the CLs \citep{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme temperature parameterization localization (\emph{subarctic}) for each PFT.
-The estimates for SWP=off are listed in Table~\ref{tab:biomass_reduction}.
+We estimated the biomass reductions in accordance to the CLs \citep{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme climate derived local temperature parameterization (subarctic) for each PFT.
+The estimates, assuming no soil water stress, are given in Table~\ref{tab:biomass_reduction}.
 We defined an uncertainty range from the difference in biomass reduction caused by the variation of $f_\mathrm{light}$.
 Superscript refers to PPFD0.8 and subscript to PPFD1.2.
-The reported range corresponds to the effect of taking SWP into account or not.
-As SWP reduces $\mathrm{POD_1}$, biomass reduction is decreased when taking SWP into account.
+Herein, the reported range indicated by dots corresponds to the limiting effect by soil water stress.
+Because soil water stress reduces $\mathrm{POD_1}$, biomass reduction is also reduced.
 The biomass reduction ranges between $2.5\,\unit{\%}$ (coniferous, 2018) and $17.4\,\unit{\%}$ (deciduous, 2019).
 For deciduous trees this amounts to a reduction by up to $55\,\unit{\%}$ compared to MM.
-Coniferous trees are the least affected PFT and the magnitude of total biomass reduction is almost independent of the choice of the parameterization within given uncertainties.
+Coniferous trees are the least affected PFT suggesting the Boreal MM parameterization is appropriate for this PFT.
 

### 2022-06-01 VegOzone.tex
Comments by Lisa

@@
 The results are displayed in Fig.~\ref{fig:pody_rel}.
-Critical levels for reduction in biomass \citep[deciduous forest $4\,\unit{\%}$, coniferous forest $2\,\unit{\%}$, and grasslands $10\,\unit{\%}$ reduction adopted from][]{ICP:MappingManual2017,ESPR:Hayes2021} are shown as horizontal dashed lines.
-The different background colors indicate the the local climate derived temperature response function.
-The light response function localizations are denoted by $+$ (PPFD1.2) and $-$ (PPFD0.8).
-Circles indicate simulations with localized GS, canopy height, and leaf dimensions.
-With open symbols, the effect on $\mathrm{POD_y}$ of any soil water stress is accounted for.
+Critical levels for reduction in biomass \citep[deciduous forest $4\,\unit{\%}$, coniferous forest $2\,\unit{\%}$, and grasslands $10\,\unit{\%}$ adopted from][]{ICP:MappingManual2017,ESPR:Hayes2021} are shown as horizontal dashed lines.
+The different background colors indicate the local climate derived temperature response function.
+The light response functions are denoted by $+$ (PPFD1.2) and $-$ (PPFD0.8).
+Open symbols indicate, the effect on $\mathrm{POD_y}$ of any soil water stress is accounted for.
+All model experiments have been performed with localized GS, canopy height, and leaf dimensions.
 
@@
 Taking into account SWP, $\mathrm{POD_1}$ was only notably reduced in 2018 (for deciduous trees).
-In all other model experiments SWP was hardly of relevance.
+In all other model experiments SWP has little relevance justifying the choice in Section~\ref{subsec:res_local}.
 A larger aperture of stomata at low light intensities (PPFD0.8) causes an increase in $\mathrm{POD_1}$ compared to MM while a smaller aperture (PPFD1.2) is leading to reduced $\mathrm{POD_1}$.
@@
 
-For deciduous trees (Fig.~\ref{fig:pody_rel}a), all model experiments exceed the CL by ($5-16\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
-In 2019, the lowest $\mathrm{POD_1}$ estimate (MM, PPFD1.2) amounts to $10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and the highest (subarctic, PPFD0.8) $21\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
+For deciduous trees (Fig.~\ref{fig:pody_rel}a), all model experiments exceed the CL by ($5-16)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
+In 2019, the lowest $\mathrm{POD_1}$ estimate (MM, PPFD1.2) amounts to $10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and the highest (subarctic, PPFD0.8) to $21\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
 In 2018, this range is smaller $(12-18)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
@@
 
-For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL by $(0-10)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
+For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL by $(0-5)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
 The different climate derived local parameterizations do not show much impact compared to MM indicating that the MM's Boreal Norway spruce parameterization is well-suited for a subarctic climate.
@@
 For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations stay below the CL.
-Perennial grassland shows the smallest $\mathrm{POD_1}$, but a similarly large response to the temperature parameterization localizations as deciduous trees.
-Perennial grassland shows the lowest sensitivity to the light threshold.
-SWP is hardly relevant.
+Perennial grassland has the smallest $\mathrm{POD_1}$, but a similarly large response to the local climate derived temperature parameterization as deciduous trees, and the lowest sensitivity to the light threshold.
+Soil water limitation has little relevance.
 The difference in $\mathrm{POD_1}$ for subarctic and cold types with respect to MM amounts to $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
@@
 
-The magnitude of all described effects on $\mathrm{POD_1}$ varies between PFTs as well as years and is of the same order of magnitude.
+The magnitude of all described effects on $\mathrm{POD_1}$ varies between PFTs and years.
 By construction, $\mathrm{POD_1}$ from localized parmeterizations is always larger than estimates with the MM parameterization.
@@
 \label{sec:stats}
-Here, we put the weather conditions during the growing season 2018/19 into a climatological context.
-All relevant data for the growing season 2018/19 are shown as time series in Fig.~\ref{fig:data_svanvik_2018_2019}.
+In the following section, we consider the climate local to the Svanhovd region to understand the characteristics that might influence the extent of stomatal ozone dose and hence risk.
+We conduct this analysis for the two years 2018 and 2019 and relate these years to seasonality to understand how risk is influence by climate.
+
+\subsubsection{Weather conditions}
+\label{subsec:weather}
+
+All relevant data for the growing seasons 2018 and 2019 are shown as time series in Fig.~\ref{fig:data_svanvik_2018_2019}.
 Ozone concentrations measured at $2\,\unit{m}$ height above ground are averaged hourly.
@@
 Note, while the downtime during winter was planned, missing data in two weeks of July 2018 (July 9--23) were due to problems in data acquisition.
-
-\subsubsection{Weather conditions}
-\label{subsec:weather}
 
@@
 However, due to the above-mentioned data acquisition problems, we missed most of the corresponding ozone data for this event and most likely also the peak \chem{\chi_{O_3}}.
-A method for gap-filling these data has been presented in \citet{ACP:Falk2021} and explained in Section~\ref{subsec:gap_filling}.
+These data have been reconstructed following the method presented in \citet{ACP:Falk2021} and briefly explained in Section~\ref{subsec:gap_filling}.
+Hence, $\mathrm{POD_1}$ estimates in Section~\ref{subsec:do3se_results} do already account for a realistic ozone profile during this period.
 In contrast, \chem{\chi_{O_3}} only rose 18 times above the threshold of $40\,\unit{ppb}$ during the summer of 2019.
@@
 \end{figure*}
+
+TODO: Some words to tie back to risk assment.
 
@@
 
-Taking a presumed adaption of vegetation to local climate into consideration, we derived localized Jarvis-type temperature and light response functions for modeling $\mathrm{POD_1}$ and biomass reduction with the $\mathrm{DO_3SE}$ model.
-Our key findings are that vegetation that is more sensitive to warm temperatures would be less effected by coincident ozone and heat stress because stomata will close to prevent water loss whereas vegetation adapted to warmer climate zones could be under higher ozone damage risk under these conditions.
-On the other hand, a warmer climate and hence prolonged GS would negatively affect any vegetation using the $\mathrm{POD_y}$ metric.
+Taking a presumed adaption of vegetation to local climate into consideration, we derived local parameterization for the Jarvis-type temperature and light response functions for modeling $\mathrm{POD_1}$ and biomass reduction with the $\mathrm{DO_3SE}$ model.
+Our key findings are that vegetation that is adapted to cold, subarctic conditions may be at higher risk of ozone damage than currently suggested by UNECE risk assessment methods as such vegetation can be physiologically active even under cold temperature conditions; the occurrence of spring ozone peaks enhance the likelihood of ozone uptake and potential damage.
+TODO: On the other hand, a warmer climate and hence prolonged GS would negatively affect any vegetation using the $\mathrm{POD_y}$ metric.
 It is essential for ozone risk assessments in the subarctic to define plant physiology and traits such as phenological states properly.

### 2022-06-01 VegOzone.tex
Suggested changes by Lisa

@@
 
-TODO: Some words to tie back to risk assment.
+\subsection{Implications on ozone damage risk}
+\label{subsec:clim_risk}
+
+As we have shown, 2019 was a rather normal year compared to 2018 with respect to present-day climate and ozone concentrations.
+In conjunction with the $\mathrm{DO_3SE}$ modeling results (Section~\ref{subsec:do3se_results}), this indicates that risk of ozone-induced damage in subarctic vegetation has been underestimated in assessments using the Boreal MM parameterizations.
+In future scenarios in which persistent heatwaves like in 2018 become more common, the risk for ozone-induced damage is probably reduced for vegetation adapted to cold growing conditions.
+A longer growing season promoted by climate change without weather extremes like in 2019 is likely to increase this risk.
 
@@
 \label{sec:conc}
-
-..
-This allows an understanding of which are the key aspects of modelled plant physiology that might determine the potential risk and also how the threat from ozone may change in the future.
-In addition, our results may hint a solution to solve long-standing issues of earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface models and display over- or underproductivity in the Arctic \citep{GMD:Poulter2015,JAMES:Lawrence2019}.
-..
-Results from open-top chamber (OTC) experiments on downy birch (\emph{Betula pubescens}) and mountain birch (\emph{Betula pubescens toruosa}), native to subarctic regions, indicated reductions in both biomass, root:shoot ratio, and visible leaf damage under elevated ozone treatments ($\left<\chem{\chi_{O_3}}\right> = 36-54\,\unit{ppb}$) \citep{Amb:Manninen2009}.
-Although, Scots pine (\emph{Pinus sylvestris}) is considered to be more ozone tolerant due to an absence of visible injuries \citep{Amb:Girgzdiene2009}, \citet{Amb:Manninen2009} found chlorophyll:carotenoid ratio and polyamine reductions under elevated ozone concentrations indicating susceptibility to ozone also in these species.
-..
 
@@
 Our key findings are that vegetation that is adapted to cold, subarctic conditions may be at higher risk of ozone damage than currently suggested by UNECE risk assessment methods as such vegetation can be physiologically active even under cold temperature conditions; the occurrence of spring ozone peaks enhance the likelihood of ozone uptake and potential damage.
-TODO: On the other hand, a warmer climate and hence prolonged GS would negatively affect any vegetation using the $\mathrm{POD_y}$ metric.
+On the other hand, a warmer climate and longer GS increase $\mathrm{POD_y}$ and hence risk.
 It is essential for ozone risk assessments in the subarctic to define plant physiology and traits such as phenological states properly.
 As no independent ozone flux measurements are available at Svanhovd, we could not fully validate our methodology at this point.
-A comparison of $\mathrm{DO_3SE}$ model derived dry deposition velocities $V_\mathrm{d}$ for coniferous trees with comprehensive records over Pine forests from Hyytilälä, Finland \citep{BER:Keronen2003} (refer to supplementary material Figs.~S1--S3) indicate a slightly too high $V_\mathrm{d}$ in spring for our most extreme localization (subarctic).
+Dry deposition velocities $V_\mathrm{d}$ for coniferous trees from comprehensive records over Pine forests from Hyytilälä, Finland \citep{BER:Keronen2003} yield summer maximum $V_\mathrm{d}\sim 11\,\unit{mm\,s^{-1}}$.
+The $\mathrm{DO_3SE}$ model derived summer maximum $V_\mathrm{d}\sim 9\,\unit{mm\,s^{-1}}$ with minimal differences between the parameterizations.
+The highest $V_\mathrm{d}$ is found in summer 2019 using the cold paramterization.
+The maximum observed spring $V_\mathrm{d}\sim 8\,\unit{mm\,s^{-1}}$ is slightly smaller compared to $V_\mathrm{d}\sim 10\,\unit{mm\,s^{-1}}$ from the $\mathrm{DO_3SE}$ model.
 No comparable data is available for northern Fennoscandia or deciduous tress and perennial grassland.
@@
 In contrast, no such damage was found on the clovers in 2019.
-This might indicate a higher ozone damage risk under warmer climate conditions for these specific plant species in a subarctic climate.
-However, visible injuries are not always correlated with a reduction in biomass \citep[and citations therein]{CRG:Felzer2007}.
-Also, our $\mathrm{DO_3SE}$ modeling results are in contrast to these observations at Svanhovd.
-Any visible injuries caused by peak ozone flux through stomata or a specific dependency of plant sensibility to ozone on phenology stage are not accounted for in the $\mathrm{DO_3SE}$ model.
+This might indicate a higher risk of ozone damage under warmer climate conditions for these specific plant species in a subarctic climate.
+However, visible injuries are not always correlated with a reduction in biomass \citep[and citations therein]{CRG:Felzer2007} and the $\mathrm{DO_3SE}$ model does not estimate visible injury which is often related to short periods of very high ozone uptake.
 In this regard, open-top chamber (OTC) experiments performed in northern Finland where Scots pine and downy birch seedlings were exposed to elevated ozone concentrations attributed a reduction in biomass and reproduction with visible injuries explicitly to peak \chem{O_3} concentrations and fast phenological development at high growth rate \citep{Amb:Manninen2009}.
+The \citet{Amb:Manninen2009} results from open-top chamber (OTC) experiments on downy birch (\emph{Betula pubescens}) and mountain birch (\emph{Betula pubescens toruosa}), native to subarctic regions, indicated reductions in both biomass, root:shoot ratio, and visible leaf damage under elevated ozone treatments ($\left<\chem{\chi_{O_3}}\right> = 36-54\,\unit{ppb}$).
+Although, Scots pine (\emph{Pinus sylvestris}) is considered to be more ozone tolerant due to an absence of visible injuries \citep{Amb:Girgzdiene2009}, \citet{Amb:Manninen2009} found chlorophyll:carotenoid ratio and polyamine reductions under elevated ozone concentrations indicating susceptibility to ozone also in these species.
 Forbs and perennial grassland are more susceptible to ozone-induced damage in the reproductive state \citep{EP:Bassin2004}.
 That might explain why we observed visible injuries in the ozone garden to a larger extent in 2018 than in 2019.
-An associated damage function to ozone depending on leaf age as proposed by \citet{AE:Musselman2006} or phenological state might improve the predictions.
+An associated damage function relating ozone susceptibility to leaf age as proposed by \citet{AE:Musselman2006} or phenological state might improve the predictions.
+Such a damage function could be different from $f_\mathrm{phen}$.
 
@@
 In particular, CLs for deciduous and coniferous trees were breached in both years and for all parameterizations.
-For perennial grassland, the CL was not breached if a localized GS was assumed, but the defined CL of $10\,\unit{\%}$ loss in biomass might not be transferable to subarctic regions with considerable constraints on productivity leading to more severe economic consequences.
-
-Beyond the risk assessment of ozone-induced damage, our localized Jarvis-type temperature response functions, once verified by observations, can have important implications on land-surface modeling still using such parameterizations in global models.
-Problems in productivity of species especially in the Arctic regions occur due to PFTs which are not suitable for the climate.
+For perennial grassland, the CL was not breached, but the defined CL of $10\,\unit{\%}$ loss in biomass might not be transferable to subarctic regions with considerable constraints on productivity leading to more severe economic consequences.
+
+Beyond the risk assessment of ozone-induced damage, our climate derived local parameterization Jarvis-type temperature response functions, once verified by observations, can have important implications on land-surface modeling still using only Boreal parameterizations in global models.
+Problems in estimating the productivity of species especially in the Arctic regions occur due to PFTs which are not suitable for the climate.
 A verification of our proposed methodology by independent flux and dry deposition measurements over birch forest and perennial grassland would be necessary to establish a climate data based localization of stomatal conductance parameterizations.

### 2022-06-01 VegOzone.tex
Rearange sections and remove refeences to localization

@@
 Therefore, we shifted both $T_\mathrm{min}$ and $T_\mathrm{max}$ to colder temperatures and chose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000.
-The most extreme of the two localizations compared to the MM is the \emph{subarctic}.
+The most extreme of the two local climate derived parameterizations compared to the MM is the \emph{subarctic}.
 
@@
 
-Similarly, we defined localized $f_\mathrm{T}$ and $f_\mathrm{light}$ for coniferous and deciduous trees.
+Similarly, we defined local climate derived $f_\mathrm{T}$ and localized $f_\mathrm{light}$ for coniferous and deciduous trees.
 The resulting response functions are shown in Appendix~Figs.~\ref{fig:f_temp_spruce} and \ref{fig:f_temp_birch}, respectively.
@@
 Thus, coniferous trees are already active at low air temperatures and can reach $60\,\unit{\%}$ photosynthetic activity as early as day of the year~(\unit{doy})~100 but cease activity in less favorable conditions \citep{TB:Kolari2007, TP:Wallin2013}.
-To determine the temperature localization for coniferous trees, we used the time series of \chem{CO_2} uptake and temperature at observation sites in southern ($\rightarrow$ \emph{cold}) and northern ($\rightarrow$ \emph{subarctic}) Finland by \citet{TB:Kolari2007}.
+To determine the local climate derived temperature parameterization for coniferous trees, we used the time series of \chem{CO_2} uptake and temperature at observation sites in southern ($\rightarrow$ \emph{cold}) and northern ($\rightarrow$ \emph{subarctic}) Finland by \citet{TB:Kolari2007}.
 From these, we estimated an optimal temperature interval ($10\,\unit{^\circ C}\le T_\mathrm{opt} \le 15\,\unit{^\circ C}$).
@@
 
-All resulting temperature response function localizations are tabulated in Table~\ref{tab:sensitivity_tests_temp} and light response function in Table~\ref{tab:sensitivity_tests_light}.
-In Section~\ref{subsec:res_local}, we compare relative $g_\mathrm{sto}$ based on Eq.~(\ref{eq:stomatal}) for the different localizations using the long term meteorological data.
+All resulting temperature response functions are tabulated in Table~\ref{tab:sensitivity_tests_temp} and light response function in Table~\ref{tab:sensitivity_tests_light}.
+In Section~\ref{subsec:res_local}, we compare relative $g_\mathrm{sto}$ based on Eq.~(\ref{eq:stomatal}) for the different local climate derived parameterizations using the long term meteorological data.
 
@@
 Based on the above criteria, we identify \emph{subarctic}--PPFD0.8 as the parameterization with the highest score.
-The differences between the localized parameterizations and the MM are smallest for coniferous trees.
+The differences between the derived parameterizations and the MM are smallest for coniferous trees.
 This indicates that the MM Boreal parameterization is well suited for the subarctic climate at Svanhovd.
@@
 
+\subsection{Climate and weather conditions in the growing season 2018/19}
+\label{sec:stats}
+In the following section, we consider the climate local to the Svanhovd region to understand the characteristics that might influence the extent of stomatal ozone dose and hence risk.
+We conduct this analysis for the two years 2018 and 2019 and relate these years to seasonality to understand how risk is influence by climate.
+
+\subsubsection{Weather conditions}
+\label{subsec:weather}
+
+All relevant data for the growing seasons 2018 and 2019 are shown as time series in Fig.~\ref{fig:data_svanvik_2018_2019}.
+Ozone concentrations measured at $2\,\unit{m}$ height above ground are averaged hourly.
+The hatched areas mark times when no ozone data were recorded.
+Note, while the downtime during winter was planned, missing data in two weeks of July 2018 (July 9--23) were due to problems in data acquisition.
+
+As can be seen in Fig.~\ref{fig:data_svanvik_2018_2019}a, \chem{\chi_{O_3}} peaks in spring (April/May) and reaches its minimum in late summer (July/August).
+The spring peak has not been captured completely in 2019, since data acquisition only started in late April.
+In summer 2018 (June--August), high ozone concentrations ($\chem{\chi_{O_3}} > 40\,\unit{ppb}$) were recorded 50 times.
+The highest summer ozone VMR ($\chem{\chi_{O_3}} = 50.2\,\unit{ppb}$) was measured on July~25.
+This coincides with the period of the most extensive forest fires in central Sweden which occurred from July~12--29 \citep{SOU2019}.
+However, due to the above-mentioned data acquisition problems, we missed most of the corresponding ozone data for this event and most likely also the peak \chem{\chi_{O_3}}.
+These data have been reconstructed following the method presented in \citet{ACP:Falk2021} and briefly explained in Section~\ref{subsec:gap_filling}.
+Hence, $\mathrm{POD_1}$ estimates in Section~\ref{subsec:do3se_results} do already account for a realistic ozone profile during this period.
+In contrast, \chem{\chi_{O_3}} only rose 18 times above the threshold of $40\,\unit{ppb}$ during the summer of 2019.
+
+Hourly averaged $2\,\unit{m}$ temperatures above $20\,\unit{^\circ C}$ occurred more regularly in July 2018 than in 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}b).
+In 2018, spring temperature regularly rose above freezing only in May, while in 2019 this occurred already early in March/April indicating a later start of the GS in 2018.
+More rain events with accumulated daily precipitation ($\sum_d \mathrm{Precip}$) above $10\,\unit{mm}$ occurred in the summer of 2018 compared to 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}c).
+Qualitatively, global irradiance ($Q_0$) displayed in Fig.~\ref{fig:data_svanvik_2018_2019}d was higher in May and July 2018 compared to 2019, while June 2019 showed higher irradiance than in 2018.
+A high global irradiance, in most cases, is the result of a low cloud fraction.
+In both years, the maximum recorded $Q_0$ was $750\,\unit{W\,m^{-2}}$ and reached in June.
+
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{data_svanvik_2018_2019}
+  %\includegraphics[width=12cm]{fig02}.
+  \caption{Observational data from atmospheric monitoring at Svanhovd in 2018/19.
+The hatched areas indicate periods without ozone monitoring data.
+(a) Hourly averaged ozone VMR; (b) hourly averaged temperature; (c) daily accumulated precipitation; (d) hourly averaged global irradiance.}
+  \label{fig:data_svanvik_2018_2019}
+\end{figure*}
+
+\subsubsection{Climate (1990-present)}
+\label{subsec:climatologies}
+
+Climatologies (typically multi-year monthly averages) of temperature, precipitation, and radiant energy $Q_\mathrm{e}$ for Svanhovd were derived for the period 1992--2012 and are shown as box plots in Fig.~\ref{fig:climatediagram}.
+Therein, the colored horizontal lines indicate the median and the triangles the mean value for the respective month.
+A box denotes the interquartile range (IQR).
+Observations exceeding 1.5~times the IQR (whiskers) are usually considered outliers but ought to be interpreted differently in a climate diagram, namely as events deviating extraordinarily from the norm.
+
+Temperatures usually stay below freezing for 5 consecutive months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for a K\"{o}ppen's climate classification of a regular subarctic climate (Dfc) \citep[e.g.][]{SD:Beck2018}.
+The highest monthly average temperature is $(13.1\pm 1.1)\,\unit{^\circ C}$ in July and the lowest $(-12.8\pm 2.0)\,\unit{^\circ C}$ in February.
+The coldest ever recorded temperature was $-45.2\,\unit{^\circ C}$ on January~27, 1999, while the highest temperature ($29.4\,\unit{^\circ C}$) occurred on July~16 the same year.
+As can be seen mean and median are well aligned in the center of the IQR for most months meaning that temperature anomalies are symmetrically distributed.
+
+The average accumulated monthly precipitation ($\sum_d\mathrm{Precip}$) indicates that winter and spring (November--April, except for March) are relatively dry ($\sum_d\mathrm{Precip} < 20\,\unit{mm}$).
+March precipitation is primarily snow and will influence the start of the growing season.
+The driest month is January with $\sum_d\mathrm{Precip} = (16.7\pm 3.0)\,\unit{mm}$.
+The most precipitation occurs in summer, with a $\sum_d\mathrm{Precip} = (58.5\pm 9.2)\,\unit{mm}$ in August.
+The average annual accumulated precipitation ($\sum_m\mathrm{Precip}$) given with standard error of mean is $(383\pm 86)\,\unit{mm}$.
+Mean and median precipitation are less well aligned compared to temperature and point to a considerable interannual variability.
+
+The climatology for global radiant energy $Q_e$ in \unit{J\,m^{-2}} (the integral of $Q_0$) is zero in November--January (polar night) and reaches its maximum in May--July (midnight sun conditions).
+Mean and median are well aligned.
+Only June displays a slightly skewed distribution towards higher $Q_e$.
+
+%Ozone
+The ozone VMR climatology (\chem{\chi_{O_3}}) (1986--1996) is shown in Figure~\ref{fig:climatediagram}d.
+The climatological spring peak occurs in April (around \unit{doy}~100) and amounts to $40\,\unit{ppb}$ while the annual average \chem{\chi_{O_3}} was $28\,\unit{ppb}$.
+The decline in \chem{\chi_{O_3}} roughly coincides with the beginning of \chem{CO_2} uptake by coniferous trees \citep{TB:Kolari2007, TP:Wallin2013}.
+Mean and median are well aligned and centered indicating a symmetric distribution for each month.
+The highest recorded \chem{\chi_{O_3}} occur in June most likely due to forest fires as mentioned above.
+
+\begin{figure*}[t]
+  \includegraphics[width=12cm]{climatediagram_box}
+  %\includegraphics[width=8.3cm]{fig04}
+  \caption{Climate diagram based on meteorological data for Svanvik/Pasvik (1992--2012).
+Ozone climatology based on ozone monitoring data (1986--1996).
+A box indicates the upper/lower quartile, whiskers the $1.5\times$ Interquartile Range (IQR), and circles outliers.
+The median is marked by a colored horizontal line and the mean by a triangle.
+Notches indicate the confidence interval around the median.
+(a) Temperature; (b) accumulated precipitation; (c) radiant energy; (d) ozone VMR}
+  \label{fig:climatediagram}
+\end{figure*}
+
+\subsubsection{2018/19 growing season anomalies}
+\label{subsec:anomalies}
+
+In climate science, a tool to compare specific years anomalies -- the deviation from the norm --  are used.
+In Figure~\ref{fig:anomalies_svanvik}, we show anomalies for both temperature, precipitation, radiant energy, and ozone VMR.
+We defined anomalies as median deviation from the climatological median.
+The hatched area indicates the respective IQR.
+
+Temperature anomalies (Fig.~\ref{fig:anomalies_svanvik}a) show enhanced temperatures throughout summer (May--September) 2018 breaching the IQR two times while March was unusually cold.
+No notably enhanced temperatures occurred in 2019 indicating that 2019 was an average year.
+
+Similarly, the precipitation anomalies are displayed in Fig.~\ref{fig:anomalies_svanvik}b.
+March 2018 saw more precipitation than usual but mainly as snow.
+May and October were unusually dry while August was unusually wet.
+2019 was rather average throughout the year, but unusually wet in June and dry in July.
+
+Radiant energy anomalies are presented in Fig.~\ref{fig:anomalies_svanvik}c.
+In summer 2018 (May/July), $Q_e$ was extraordinarily higher than usual, while 2019 was generally an average year with a rather dark spring (April/May) and late summer (August).
+
+%Ozone
+We used the bias-corrected and cross-calibrated ozone climatology \citep{ACP:Falk2021} to assess the ozone anomalies in 2018/19.
+Note that we have not recorded ozone during winter (refer to Section~\ref{subsec:weather}).
+In April/May 2018, ozone was unusually elevated and stayed higher than usual through the whole GS.
+The highest anomaly in 2019 occurred in September at the end of the GS.
+The forest fires in Sweden in July enhanced ozone notably at Svanhovd.
+
+\begin{figure*}[t]
+
+  \centering
+  \includegraphics[width=12cm]{anomalies}
+  \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as difference from median for each month.
+The hatched area between indicates the IQR.
+(a) Temperature; (b) precipitation; (c) radiant energy; (d) ozone VMR.}
+  \label{fig:anomalies_svanvik}
+\end{figure*}
+
 \subsection{Implications on ozone risk assessments}
@@
 We identified that, in general, an increasing cold tolerance represented by a shift in the temperature response function (MM~$<$~cold~$<$~subarctic) leads to an increase in $\mathrm{POD_1}$ in both investigated years (2018 and 2019) compared to the MM parameterization.
-The more extreme the temperature response function localization the higher $\mathrm{POD_1}$.
+The more extreme the local climate derived temperature response function the higher $\mathrm{POD_1}$.
 In 2018, coniferous trees and perennial grassland display a reduction in $\mathrm{POD_1}$ for the subarctic parameterization compared to the cold paramerization in responds to the prescribed lower $g_\mathrm{sto}$ at higher temperatures.
@@
 The magnitude of all described effects on $\mathrm{POD_1}$ varies between PFTs and years.
-By construction, $\mathrm{POD_1}$ from localized parmeterizations is always larger than estimates with the MM parameterization.
+By construction, $\mathrm{POD_1}$ from local climate derived parmeterizations is always larger than estimates with the MM parameterization.
 This means that the risk of ozone damage could be underestimated especially in more normal years like 2019 (see Section~\ref{sec:stats} for details).
@@
 
-\subsection{Climate and weather conditions in the growing season 2018/19}
-\label{sec:stats}
-In the following section, we consider the climate local to the Svanhovd region to understand the characteristics that might influence the extent of stomatal ozone dose and hence risk.
-We conduct this analysis for the two years 2018 and 2019 and relate these years to seasonality to understand how risk is influence by climate.
-
-\subsubsection{Weather conditions}
-\label{subsec:weather}
-
-All relevant data for the growing seasons 2018 and 2019 are shown as time series in Fig.~\ref{fig:data_svanvik_2018_2019}.
-Ozone concentrations measured at $2\,\unit{m}$ height above ground are averaged hourly.
-The hatched areas mark times when no ozone data were recorded.
-Note, while the downtime during winter was planned, missing data in two weeks of July 2018 (July 9--23) were due to problems in data acquisition.
-
-As can be seen in Fig.~\ref{fig:data_svanvik_2018_2019}a, \chem{\chi_{O_3}} peaks in spring (April/May) and reaches its minimum in late summer (July/August).
-The spring peak has not been captured completely in 2019, since data acquisition only started in late April.
-In summer 2018 (June--August), high ozone concentrations ($\chem{\chi_{O_3}} > 40\,\unit{ppb}$) were recorded 50 times.
-The highest summer ozone VMR ($\chem{\chi_{O_3}} = 50.2\,\unit{ppb}$) was measured on July~25.
-This coincides with the period of the most extensive forest fires in central Sweden which occurred from July~12--29 \citep{SOU2019}.
-However, due to the above-mentioned data acquisition problems, we missed most of the corresponding ozone data for this event and most likely also the peak \chem{\chi_{O_3}}.
-These data have been reconstructed following the method presented in \citet{ACP:Falk2021} and briefly explained in Section~\ref{subsec:gap_filling}.
-Hence, $\mathrm{POD_1}$ estimates in Section~\ref{subsec:do3se_results} do already account for a realistic ozone profile during this period.
-In contrast, \chem{\chi_{O_3}} only rose 18 times above the threshold of $40\,\unit{ppb}$ during the summer of 2019.
-
-Hourly averaged $2\,\unit{m}$ temperatures above $20\,\unit{^\circ C}$ occurred more regularly in July 2018 than in 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}b).
-In 2018, spring temperature regularly rose above freezing only in May, while in 2019 this occurred already early in March/April indicating a later start of the GS in 2018.
-More rain events with accumulated daily precipitation ($\sum_d \mathrm{Precip}$) above $10\,\unit{mm}$ occurred in the summer of 2018 compared to 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}c).
-Qualitatively, global irradiance ($Q_0$) displayed in Fig.~\ref{fig:data_svanvik_2018_2019}d was higher in May and July 2018 compared to 2019, while June 2019 showed higher irradiance than in 2018.
-A high global irradiance, in most cases, is the result of a low cloud fraction.
-In both years, the maximum recorded $Q_0$ was $750\,\unit{W\,m^{-2}}$ and reached in June.
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{data_svanvik_2018_2019}
-  %\includegraphics[width=12cm]{fig02}.
-  \caption{Observational data from atmospheric monitoring at Svanhovd in 2018/19.
-The hatched areas indicate periods without ozone monitoring data.
-(a) Hourly averaged ozone VMR; (b) hourly averaged temperature; (c) daily accumulated precipitation; (d) hourly averaged global irradiance.}
-  \label{fig:data_svanvik_2018_2019}
-\end{figure*}
-
-\subsubsection{Climate (1990-present)}
-\label{subsec:climatologies}
-
-Climatologies (typically multi-year monthly averages) of temperature, precipitation, and radiant energy $Q_\mathrm{e}$ for Svanhovd were derived for the period 1992--2012 and are shown as box plots in Fig.~\ref{fig:climatediagram}.
-Therein, the colored horizontal lines indicate the median and the triangles the mean value for the respective month.
-A box denotes the interquartile range (IQR).
-Observations exceeding 1.5~times the IQR (whiskers) are usually considered outliers but ought to be interpreted differently in a climate diagram, namely as events deviating extraordinarily from the norm.
-
-Temperatures usually stay below freezing for 5 consecutive months, while only 2 months breach $10\,\unit{^\circ C}$ regularly (July, August), satisfying the conditions for a K\"{o}ppen's climate classification of a regular subarctic climate (Dfc) \citep[e.g.][]{SD:Beck2018}.
-The highest monthly average temperature is $(13.1\pm 1.1)\,\unit{^\circ C}$ in July and the lowest $(-12.8\pm 2.0)\,\unit{^\circ C}$ in February.
-The coldest ever recorded temperature was $-45.2\,\unit{^\circ C}$ on January~27, 1999, while the highest temperature ($29.4\,\unit{^\circ C}$) occurred on July~16 the same year.
-As can be seen mean and median are well aligned in the center of the IQR for most months meaning that temperature anomalies are symmetrically distributed.
-
-The average accumulated monthly precipitation ($\sum_d\mathrm{Precip}$) indicates that winter and spring (November--April, except for March) are relatively dry ($\sum_d\mathrm{Precip} < 20\,\unit{mm}$).
-March precipitation is primarily snow and will influence the start of the growing season.
-The driest month is January with $\sum_d\mathrm{Precip} = (16.7\pm 3.0)\,\unit{mm}$.
-The most precipitation occurs in summer, with a $\sum_d\mathrm{Precip} = (58.5\pm 9.2)\,\unit{mm}$ in August.
-The average annual accumulated precipitation ($\sum_m\mathrm{Precip}$) given with standard error of mean is $(383\pm 86)\,\unit{mm}$.
-Mean and median precipitation are less well aligned compared to temperature and point to a considerable interannual variability.
-
-The climatology for global radiant energy $Q_e$ in \unit{J\,m^{-2}} (the integral of $Q_0$) is zero in November--January (polar night) and reaches its maximum in May--July (midnight sun conditions).
-Mean and median are well aligned.
-Only June displays a slightly skewed distribution towards higher $Q_e$.
-
-%Ozone
-The ozone VMR climatology (\chem{\chi_{O_3}}) (1986--1996) is shown in Figure~\ref{fig:climatediagram}d.
-The climatological spring peak occurs in April (around \unit{doy}~100) and amounts to $40\,\unit{ppb}$ while the annual average \chem{\chi_{O_3}} was $28\,\unit{ppb}$.
-The decline in \chem{\chi_{O_3}} roughly coincides with the beginning of \chem{CO_2} uptake by coniferous trees \citep{TB:Kolari2007, TP:Wallin2013}.
-Mean and median are well aligned and centered indicating a symmetric distribution for each month.
-The highest recorded \chem{\chi_{O_3}} occur in June most likely due to forest fires as mentioned above.
-
-\begin{figure*}[t]
-  \includegraphics[width=12cm]{climatediagram_box}
-  %\includegraphics[width=8.3cm]{fig04}
-  \caption{Climate diagram based on meteorological data for Svanvik/Pasvik (1992--2012).
-Ozone climatology based on ozone monitoring data (1986--1996).
-A box indicates the upper/lower quartile, whiskers the $1.5\times$ Interquartile Range (IQR), and circles outliers.
-The median is marked by a colored horizontal line and the mean by a triangle.
-Notches indicate the confidence interval around the median.
-(a) Temperature; (b) accumulated precipitation; (c) radiant energy; (d) ozone VMR}
-  \label{fig:climatediagram}
-\end{figure*}
-
-\subsubsection{2018/19 growing season anomalies}
-\label{subsec:anomalies}
-
-In climate science, a tool to compare specific years anomalies -- the deviation from the norm --  are used.
-In Figure~\ref{fig:anomalies_svanvik}, we show anomalies for both temperature, precipitation, radiant energy, and ozone VMR.
-We defined anomalies as median deviation from the climatological median.
-The hatched area indicates the respective IQR.
-
-Temperature anomalies (Fig.~\ref{fig:anomalies_svanvik}a) show enhanced temperatures throughout summer (May--September) 2018 breaching the IQR two times while March was unusually cold.
-No notably enhanced temperatures occurred in 2019 indicating that 2019 was an average year.
-
-Similarly, the precipitation anomalies are displayed in Fig.~\ref{fig:anomalies_svanvik}b.
-March 2018 saw more precipitation than usual but mainly as snow.
-May and October were unusually dry while August was unusually wet.
-2019 was rather average throughout the year, but unusually wet in June and dry in July.
-
-Radiant energy anomalies are presented in Fig.~\ref{fig:anomalies_svanvik}c.
-In summer 2018 (May/July), $Q_e$ was extraordinarily higher than usual, while 2019 was generally an average year with a rather dark spring (April/May) and late summer (August).
-
-%Ozone
-We used the bias-corrected and cross-calibrated ozone climatology \citep{ACP:Falk2021} to assess the ozone anomalies in 2018/19.
-Note that we have not recorded ozone during winter (refer to Section~\ref{subsec:weather}).
-In April/May 2018, ozone was unusually elevated and stayed higher than usual through the whole GS.
-The highest anomaly in 2019 occurred in September at the end of the GS.
-The forest fires in Sweden in July enhanced ozone notably at Svanhovd.
-
-\begin{figure*}[t]
-
-  \centering
-  \includegraphics[width=12cm]{anomalies}
-  \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as difference from median for each month.
-The hatched area between indicates the IQR.
-(a) Temperature; (b) precipitation; (c) radiant energy; (d) ozone VMR.}
-  \label{fig:anomalies_svanvik}
-\end{figure*}
-
-\subsection{Implications on ozone damage risk}
-\label{subsec:clim_risk}
-
 As we have shown, 2019 was a rather normal year compared to 2018 with respect to present-day climate and ozone concentrations.
-In conjunction with the $\mathrm{DO_3SE}$ modeling results (Section~\ref{subsec:do3se_results}), this indicates that risk of ozone-induced damage in subarctic vegetation has been underestimated in assessments using the Boreal MM parameterizations.
-In future scenarios in which persistent heatwaves like in 2018 become more common, the risk for ozone-induced damage is probably reduced for vegetation adapted to cold growing conditions.
+This indicates that risk of ozone-induced damage in subarctic vegetation has been previously underestimated in assessments using the Boreal MM parameterizations.
+In future scenarios in which persistent heatwaves like in 2018 may become more common, the risk for ozone-induced damage is probably reduced for vegetation adapted to cold growing conditions.
 A longer growing season promoted by climate change without weather extremes like in 2019 is likely to increase this risk.
@@
 No comparable data is available for northern Fennoscandia or deciduous tress and perennial grassland.
-Perennial grassland and deciduous trees displayed strong susceptibility to our localization methodology (increase in frequency of maximum $V_\mathrm{d}$ throughout the GS, see supplementary material Figs.~S1--S3) but are within the range ($v\mathrm{d} = (2-9)\,\unit{mm\,s^{-1}}$ for grassland) given in the literature \citep[e.g.][]{JGR:Jacob1992,AE:Padro1996,RG:Clifton2020}.
+Perennial grassland and deciduous trees displayed strong susceptibility to our methodology of local climate derived parameterization (increase in frequency of maximum $V_\mathrm{d}$ throughout the GS, see supplementary material Figs.~S1--S3) but are within the range ($v\mathrm{d} = (2-9)\,\unit{mm\,s^{-1}}$ for grassland) given in the literature \citep[e.g.][]{JGR:Jacob1992,AE:Padro1996,RG:Clifton2020}.
 Independent flux measurements over these vegetation types in a subarctic climate are thus crucial for evaluation and validation.
@@
 Problems in estimating the productivity of species especially in the Arctic regions occur due to PFTs which are not suitable for the climate.
-A verification of our proposed methodology by independent flux and dry deposition measurements over birch forest and perennial grassland would be necessary to establish a climate data based localization of stomatal conductance parameterizations.
+A verification of our proposed methodology by independent flux and dry deposition measurements over birch forest and perennial grassland would be necessary to establish a climate data based local parameterization of stomatal conductance.
 
@@
   %\includegraphics[width=8.3cm]{figB3}
-\caption{Construction of localized response functions for coniferous trees.
+\caption{Construction of local climate derived response functions for coniferous trees.
 (a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively.

### 2022-06-02 VegOzone.tex
Include photosynthesis acclimation reference

@@
 A comparison between the two PDFs indicates that during the past decade the temperature distribution has shifted towards higher temperatures reducing the number of days below $5\,\unit{^\circ C}$.
-It is unclear how fast the species composition or physiology via acclimation of perennial grassland responds to these changes.
+It is unclear how fast the species composition or physiology of perennial grassland responds to these changes.
 Thus, the predominant acclimation at present is not known.
@@
 The most extreme of the two local climate derived parameterizations compared to the MM is the \emph{subarctic}.
+This methodology is analogous to the optimal temperature acclimation experimentally found for photosynthesis \citep{PR:Way2013}.
 
@@
 The actual resilience of northern Fennoscandia plant communities to climate change strongly depends on both the range of acclimation and the type of acclimation.
-This means, vegetation that is acclimated to cold temperatures and less to heat- and drought stress might suffer more strongly from a more frequent occurrence of heatwaves than plants that have a higher temperature tolerance, but our results point to the opposite concerning the ozone stress.
+This means, vegetation that is acclimated to cold temperatures and less to heat- and drought stress might suffer more strongly from a more frequent occurrence of heatwaves than plants that have a higher temperature tolerance.
+Concerning the ozone stress, our results point to the opposite, namely that the closure of stomata at higher temperatures coinciding with high surface air ozone concentrations will prevent a high ozone dose in species acclimated to cold climates.
 

### 2022-06-02 VegOzone.tex
Fine tuning

@@
 To date, only a few studies have assessed the response of subarctic vegetation to variations in climate and air pollution.
-This study looks to fill this knowledge gap by examining the uncertainty in risk assessments that result from using Boreal rather than sub-Arctic paramterizations.
+This study looks to fill this knowledge gap by examining the uncertainty in risk assessments that result from using Boreal rather than subrctic paramterizations.
 With no parameterization of stomatal response of subarctic vegetation available, we propose an adjustment of existing Boreal parameterizations based on local climate condition.
@@
 We refer to this approach as a 'local climate derived parameterization' of stomatal response functions.
-We also study the impact of the stomata aperture at low light intensities.
+We also study the impact of the stomata aperture parameterization at low light intensities.
 With these, we apply the integrated flux-metric Phytotoxic Ozone Dose (POD) to determine ozone risk and damage to vegetation as a function of \chem{[O_3]}, key environmental variables, and species-specific physiology.
@@
 Our modeling study suggests that using standard Boreal parameterizations in assessments likely leads to underestimating the risk of ozone damage in this region for some species.
-We find that local climate derived parameterizations for subarctic vegetation result in an ozone-induced biomass loss of $2.5-17.4\,\unit{\%}$ compared to $2.3-11.2\,\%$.
+We find that local climate derived parameterizations for subarctic vegetation result in an ozone-induced biomass loss of $2.5-17.4\,\unit{\%}$.
 For some species, this loss is up to $55\,\unit{\%}$ larger than that projected from Boreal parameterizations used in general assessments.
@@
 
-The growing season (GS) in northern Fennoscandia has been shifting and starts earlier in response to the general warming trend \citep[e.g.]{GCB:Menzel2006,RS:Hogda2013,IJB:Karlsen2007}.
+The growing season (GS) in northern Fennoscandia has been shifting and starts earlier in response to the general warming trend \citep[e.g.,][]{GCB:Menzel2006,RS:Hogda2013,IJB:Karlsen2007}.
 Its start is thereby converging with the period of the ozone spring peak.
@@
 The higher the stomatal conductance, the higher the potential for ozone uptake.
-Stomatal conductance has been empirically linked to key environmental variables such as air temperature $T$, photosynthetic photon flux density (PPFD), vapor pressure deficit (VPD), and soil water potential (SWP) as well as photosynthesis itself \citep[e.g.,][]{PTRS:Jarvis1976, BallBerry1987, Emberson2000, ICP:MappingManual2017}.
+Stomatal conductance has been empirically linked to key environmental variables such as air temperature $T_\mathrm{air}$, photosynthetic photon flux density (PPFD), vapor pressure deficit (VPD), and soil water potential (SWP) as well as photosynthesis itself \citep[e.g.,][]{PTRS:Jarvis1976, BallBerry1987, Emberson2000, ICP:MappingManual2017}.
 
@@
 Given the unique conditions of the subarctic (i.e.
-long duration of daylight in the summer and cold temperatures) it would be expected that this vegetation type has a particular physiological response to the prevailing climatic conditions different to boreal species.
+long duration of daylight in the summer and cold temperatures) it would be expected that this vegetation type has a particular physiological response to the prevailing climatic conditions different to Boreal species.
 
@@
 
-An application of the Mapping Manual method was made by \citet{GCB:Mills2011} where exceedance of the CL was related to observations of ozone damage to a clover bio-monitor.
+An application of the Mapping Manual method was made by \citet{GCB:Mills2011} where exceedance of the CL was related to observations of visible ozone injuries and reduction of biomass of a clover bio-monitor.
 This study demonstrated the improved ability of flux-based metrics (in comparison to exposure-based metrics) to identify the geographical distribution of the risk of ozone damage and also showed that the ozone damage could extend into more northerly regions of Europe.\\
 
-In Section~\ref{sec:dam}, we give a brief summary of the key components and drivers of the $\mathrm{DO_3SE}$ model that we used to assess the ozone risk.
-Subsequently, we propose a local climate derived temperature response function and explore the impact of the stomata aperture at low light intensities.
+In Section~\ref{sec:dam}, we give a brief summary of the key components and meteorological input data of the $\mathrm{DO_3SE}$ model that we use to assess the ozone risk.
+Subsequently, we propose a local climate derived temperature response function and explore the impact of the stomata aperture parameterization at low light intensities.
 Start and end of the growing season are inferred from satellite observed gross primary productivity (GPP).
 We explore how risk might be over- or underestimated using Boreal parameterization (Section~\ref{sec:results}).
-For evaluating these results in a climate perspective, we assess the climate anomalies of weather conditions and ozone concentrations in the 2018/19 GS at our research site (Section~\ref{sec:stats}).
+In Section~\ref{subsec:res_local}, we evaluate the local climate derived parameterizations by means of average relative stomatal conductance.
+For evaluating the $\mathrm{DO_3SE}$ model results in a climate perspective, we assess the climate anomalies of weather conditions and ozone concentrations in the 2018/19 GS at our research site (Section~\ref{sec:stats}).
+In Section~\ref{subsec:do3se_results}, we assess the difference in predicted risk for ozone-induced damage for the local climate derived parameterizations compared to MM parameters.
 In Section~\ref{sec:conc}, we discuss our results concerning prevailing conditions and give an outlook on how future assessments of pollution risk in the subarctic could be improved.

### 2022-06-02 VegOzone.tex
Fine tuning

@@
 In Section~\ref{subsec:res_local}, we evaluate the local climate derived parameterizations by means of average relative stomatal conductance.
-For evaluating the $\mathrm{DO_3SE}$ model results in a climate perspective, we assess the climate anomalies of weather conditions and ozone concentrations in the 2018/19 GS at our research site (Section~\ref{sec:stats}).
+For evaluating the $\mathrm{DO_3SE}$ model results in a climate perspective, we assess the climate anomalies of weather conditions and ozone concentrations in the 2018 and 2019 GS at our research site (Section~\ref{subsec:stats}).
 In Section~\ref{subsec:do3se_results}, we assess the difference in predicted risk for ozone-induced damage for the local climate derived parameterizations compared to MM parameters.
-In Section~\ref{sec:conc}, we discuss our results concerning prevailing conditions and give an outlook on how future assessments of pollution risk in the subarctic could be improved.
+We discuss our results concerning prevailing conditions and give an outlook on how future assessments of pollution risk in the subarctic could be improved in Section~\ref{sec:conc}.
 
@@
 In the following, we briefly describe the $\mathrm{DO_3SE}$ model used for assessing the ozone risk, its key components, input data, and gap filling procedure (Section~\ref{sec:do3se}).
-To study the impact of model parameters, we propose a method to define a parameterization for a subarctic climate based on existing Boreal parameterizations and local climate conditions, study the impact of stomata aperture at low light intensities, and estimate the start and end of the growing season from satellite derived GPP (Section~\ref{sec:localization}).
+To study the impact of model parameters, we propose a method to define a parameterization for a subarctic climate based on existing Boreal parameterizations and local climate conditions, study the impact of stomata aperture parameterization at low light intensities, and estimate the start and end of the growing season from satellite derived GPP (Section~\ref{sec:localization}).
 
@@
 Where $g_\mathrm{max}$ is a species-specific maximum stomatal conductance value that is modified by factors that vary within a range of $0-1$.
-These factors are empirically determined and account for leaf phenology ($f_\text{phen}$), light ($f_\text{light}$), temperature ($f_\mathrm{T}$), water vapor pressure deficit ($f_\mathrm{VPD}$), and soil water potential ($f_\text{SWP}$).
+These factors are empirically determined and account for leaf phenology $f_\text{phen}$, light $f_\text{light}$, temperature $f_\mathrm{T}$, water vapor pressure deficit $f_\mathrm{VPD}$, and soil water potential $f_\text{SWP}$.
 All factors differ with plant functional type (PFT).
 
-The leaf phenology factor ($f_\mathrm{phen}$) is a stepwise defined linear function that allows for an increase of stomatal conductance in spring (start of growing season) and a decrease in fall (end of growing season).
+The leaf phenology factor $f_\mathrm{phen}$ is a stepwise defined linear function that allows for an increase of stomatal conductance in spring (start of growing season) and a decrease in fall (end of growing season).
 
@@
 
-$\Phi_\mathrm{sto}$ for a sunlit upper canopy leaf (vegetation height $h$) is then calculated under the assumption that the concentration of \chem{O_3} at the top of the canopy $h$ represents a reasonable estimate of the \chem{O_3} concentration at the upper surface of the laminar layer:
+$\Phi_\mathrm{sto}$ for a sunlit upper canopy leaf (vegetation height $h$) is then calculated under the assumption that the $\chem{[O_3]}_h$ at the top of the canopy $h$ represents a reasonable estimate of the \chem{[O_3]} at the upper surface of the laminar layer:
 %
@@
 \end{equation}
-
-$\mathrm{POD_y}$ is then computed according to:
+$\chem{[O_3]}_h$ is time dependent and typically 1-hourly averaged. Using this $\mathrm{POD_y}$ is then computed according to:
 %
@@
 %
-with an hourly averaged stomatal ozone flux $\Phi_\mathrm{sto}(t)$ (Eq.~(\ref{eq:flux_stomata})) and a stomatal ozone flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}^{-1}\,s^{-1}}.
+with a stomatal ozone flux threshold $\mathrm{y}$ and $\Phi_\mathrm{sto}(t)$ given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}^{-1}\,s^{-1}}.
 Herein PLA stands for Projected~Leaf~Area.
-The timestep $\mathrm{d}t$ is usually 1~hour.
+The integration timestep $\mathrm{d}t$ is usually 1~hour.
 The flux threshold~$\mathrm{y}$ represents the detoxification potential of the plant and is typically only exceeded during daylight hours (i.e.
@@
 with an ozone climatology (multiannual norm) \chem{\left<[O_3]\right>} and ozone anomalies (deviation from the norm) \chem{\Delta[O_3]}.
-\citet{ACP:Falk2021} computed a generalized climatology for northern Fennoscandia from long-term records of different close by sites and determined the station with the highest correlation coefficient compared to the site with the gap.
+\citet{ACP:Falk2021} computed a generalized climatology for northern Fennoscandia from long-term records of different close by atmospheric monitoring sites and determined the station with the highest correlation coefficient compared to the site with the gap.
 For this site, the anomalies with respect to the generalized climatology were calculated for the gap period.
@@
 $2\,\unit{m}$ temperature, precipitation, global irradiance, and soil temperature at different depths, and (3) a national surveillance of air pollution monitoring (\chem{NO_x}, \chem{SO_x}, particular matter (PM), radiological species) is ongoing.
-Most of these measurements are performed by the Norwegian Institute for Air Research (NILU), who also conducted ozone measurements on our behalf during the 2018/19 growing season.
-
-All essential climate variables for the 2018/19 growing season were provided by NILU through the web portal \href{luftkvalitet.no}{luftkvalitet.no} (last accessed May 2020).
+Most of these measurements are performed by the Norwegian Institute for Air Research (NILU), who also conducted ozone measurements on our behalf during the 2018 and 2019 growing seasons.
+
+All essential climate variables for the 2018 and 2019 growing seasons were provided by NILU through the web portal \href{luftkvalitet.no}{luftkvalitet.no} (last accessed May 2020).
 The \chem{O_3} monitoring data have been added to a database for air-quality monitoring called EBAS operated by NILU and are thus openly accessible.
@@
 
-To relate local ozone concentration \chem{[O_3]} with visible injuries on ozone sensitive plants, following \citet{GCB:Mills2011}, an ozone bio-monitor (referred to as ozone garden) has been installed at Svanhovd in the 2018/19 GS.
+To relate local ozone concentration \chem{[O_3]} with visible injuries on ozone sensitive plants, following \citet{GCB:Mills2011}, an ozone bio-monitor (referred to as ozone garden) has been installed at Svanhovd during the 2018 and 2019 GSs.
 The locations of the atmospheric monitoring site and the ozone garden are marked in the aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}a.
@@
 
-We determined the dominant species on-site at Svanhovd from Fig.~\ref{fig:svanhovd_research_station}a and found perennial grassland, birch (generalized as deciduous trees), and Scots pine (generalized as coniferous trees).
-Basic parameters for these PFTs are derived from Boreal paramterizations \citep{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as mapping manual (MM) parameterizations.
-The subarctic local climate derived parameterization for coniferous trees is based on MM's Boreal Norway spruce, deciduous trees on the MM's silver birch, and perennial grassland on MM's perennial grassland for Central Europe.
+We determined the dominant species on-site at Svanhovd and from Fig.~\ref{fig:svanhovd_research_station}a and found perennial grassland, birch (generalized as deciduous trees), and Scots pine (generalized as coniferous trees).
+Basic parameters for these PFTs are derived from Boreal paramterizations when available \citep{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as mapping manual (MM) parameterizations.
+In particular, the subarctic local climate derived parameterization for coniferous trees is based on MM's Boreal Norway spruce, deciduous trees on the MM's Boreal silver birch, and perennial grassland on MM's perennial grassland for Central Europe as a Boreal parameterization is not available.
 For a comprehensive list of MM parameters, consult Supplement~Table~S1.
@@
 
-Initial simulations with the Central European (Boreal is not available) MM parameterizations for perennial grassland revealed an unexpectedly low leaf level stomatal conductance ($G_\mathrm{sto}^\mathrm{leaf}$) in 2019 (Appendix~Fig.~\ref{fig:pody_mm_composit}).
+Initial simulations with the $\mathrm{DO_3SE}$ model using the Central European MM parameterization for perennial grassland revealed an unexpectedly low leaf level stomatal conductance ($G_\mathrm{sto}^\mathrm{leaf}$) in 2019 (Appendix~Fig.~\ref{fig:pody_mm_composit}).
 Substantial stomatal conductance occurred only during an extended warm period in late July.
@@
 
-To derive a subarctic parameterization for the temperature response function $f_\mathrm{T}$, we divided the temperature dataset into two subsets: 1992--2000 (near past) and 2011--2019 (present).
-From these two subsets, we selected only dates within the GS (May--September) and plotted the probability density functions (PDFs) as histograms with $1\,\unit{^\circ C}$ binning (Fig.~\ref{fig:f_temp_grassland}a).
+To derive a subarctic parameterization for the temperature response function $f_\mathrm{T}$, we divide the temperature dataset into two subsets: 1992--2000 (near past) and 2011--2019 (present).
+From these two subsets, we select only dates within the GS (May--September) and plott the probability density functions (PDFs) as histograms with $1\,\unit{^\circ C}$ binning (Fig.~\ref{fig:f_temp_grassland}a).
 A comparison between the two PDFs indicates that during the past decade the temperature distribution has shifted towards higher temperatures reducing the number of days below $5\,\unit{^\circ C}$.
@@
 Thus, the predominant acclimation at present is not known.
-Further, we overlaid these histograms with the MM $f_\mathrm{T}$.
+Further, we overlay these histograms with the MM $f_\mathrm{T}$.
 A low overlapping area suggests a low agreement with local climate conditions.
 
-We defined local climate derived parameterizations for two hypothetical species (\emph{subarctic} and \emph{cold}) by increasing the overlapping area between $f_\mathrm{T}$ and the temperature PDF during the GS.
-We constructed \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures than the MM species.
-This was accomplished by moving $T_\mathrm{min}$ and $T_\mathrm{opt}$ towards cooler temperatures but $T_\mathrm{max}$ was kept at its MM value.
-In the same way, \emph{subarctic} was constructed to represent a species that is very tolerant to cold but sensitive to high temperatures, and most efficient at cool temperatures.
-Therefore, we shifted both $T_\mathrm{min}$ and $T_\mathrm{max}$ to colder temperatures and chose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000.
-The most extreme of the two local climate derived parameterizations compared to the MM is the \emph{subarctic}.
+We define local climate derived parameterizations for two hypothetical species (\emph{subarctic} and \emph{cold}) by increasing the overlapping area between $f_\mathrm{T}$ and the temperature PDF during the GS.
+Then we construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures than the MM species.
+This is accomplished by moving $T_\mathrm{min}$ and $T_\mathrm{opt}$ towards cooler temperatures but $T_\mathrm{max}$ is kept at its MM value.
+In the same way, \emph{subarctic} is constructed to represent a species that is very tolerant to cold but sensitive to high temperatures, and most efficient at cool temperatures.
+Therefore, we shift both $T_\mathrm{min}$ and $T_\mathrm{max}$ to colder temperatures and choose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000.
+The most extreme of the two local climate derived parameterizations compared to the MM parameterization is the \emph{subarctic}.
 This methodology is analogous to the optimal temperature acclimation experimentally found for photosynthesis \citep{PR:Way2013}.
 
-For localizing $f_\mathrm{light}$, we presumed that the aperture of stomata at low light conditions differs in subarctic species compared to species in less extreme climates.
-We, therefore, adjusted the modelled light sensitivity of stomata such that the PPFD value needed to reach $50\,\unit{\%}$ opening is varied by $\pm 20\,\unit{\%}$.
-For this, we derive the inverse function $f_{\mathrm{light},k}^{-1}$ of Eq.~(\ref{eq:flight}) for each PFT $k$ analytically
+For localizing $f_\mathrm{light}$, we presume that the aperture of stomata at low light conditions differs in subarctic species compared to species in less extreme climates.
+We, therefore, adjust the modelled light sensitivity of stomata such that the PPFD value needed to reach $50\,\unit{\%}$ opening is varied by $\pm 20\,\unit{\%}$.
+We derive the inverse function $f_{\mathrm{light},k}^{-1}$ of Eq.~(\ref{eq:flight}) for each PFT $k$ analytically
 %
@@
 %
-First, we solved Eq.~(\ref{eq:inverse_function}) for the MM default value of $\alpha_\mathrm{MM}^k$ at $0.5$ ($50\,\%$ opening)
+First, we solve Eq.~(\ref{eq:inverse_function}) for the MM default value of $\alpha_\mathrm{MM}^k$ at $0.5$ ($50\,\%$ opening)
 \begin{equation}
@@
 \end{equation}
-and defined a variation $\gamma' = \gamma \cdot \eta$ with $\eta \in \{0.8, 1.2\}$. By solving Eq.~(\ref{eq:inverse_function}) for $\alpha$, we find $\alpha'(\gamma')$.
+and define a variation $\gamma' = \gamma \cdot \eta$ with $\eta \in \{0.8, 1.2\}$. By solving Eq.~(\ref{eq:inverse_function}) for $\alpha$, we find $\alpha'(\gamma')$.
 
 In Fig.~\ref{fig:f_temp_grassland}b, we show the PPFD PDF for near past and present together with the $f_\mathrm{light}$ for perennial grassland.
-As no Boreal parameterization is available for perennial grassland, the Central European MM parameterization is displayed as solid line.
+The Central European MM parameterization is displayed as solid line for comparison.
 
@@
 
-Similarly, we defined local climate derived $f_\mathrm{T}$ and localized $f_\mathrm{light}$ for coniferous and deciduous trees.
+Similarly, we define local climate derived $f_\mathrm{T}$ and localized $f_\mathrm{light}$ for coniferous and deciduous trees.
 The resulting response functions are shown in Appendix~Figs.~\ref{fig:f_temp_spruce} and \ref{fig:f_temp_birch}, respectively.
@@
 Thus, coniferous trees are already active at low air temperatures and can reach $60\,\unit{\%}$ photosynthetic activity as early as day of the year~(\unit{doy})~100 but cease activity in less favorable conditions \citep{TB:Kolari2007, TP:Wallin2013}.
-To determine the local climate derived temperature parameterization for coniferous trees, we used the time series of \chem{CO_2} uptake and temperature at observation sites in southern ($\rightarrow$ \emph{cold}) and northern ($\rightarrow$ \emph{subarctic}) Finland by \citet{TB:Kolari2007}.
-From these, we estimated an optimal temperature interval ($10\,\unit{^\circ C}\le T_\mathrm{opt} \le 15\,\unit{^\circ C}$).
-For simplicity, the same interval was assumed for deciduous trees.
+To determine the local climate derived temperature parameterization for coniferous trees, we use the time series of \chem{CO_2} uptake and temperature at observation sites in southern ($\rightarrow$ \emph{cold}) and northern ($\rightarrow$ \emph{subarctic}) Finland by \citet{TB:Kolari2007}.
+From these, we estimate an optimal temperature interval ($10\,\unit{^\circ C}\le T_\mathrm{opt} \le 15\,\unit{^\circ C}$).
+For simplicity, the same interval is assumed for deciduous trees.
 
@@
 \begin{table}[t]
-  \caption{Local climate derived light parameterizations. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
+  \caption{Localized light response parameterizations. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
   \label{tab:sensitivity_tests_light}
@@
 The accumulation of ozone dose is strongly depended on the timing and length of the GS.
-\citet{ICP:MappingManual2017} suggest a simple latitude model for beech, birch, and Norway spruce to localize start and end of the GS:
+\citet{ICP:MappingManual2017} suggested a simple latitude model for beech, birch, and Norway spruce to localize start and end of the GS:
 %
@@
 %
-At Svanhovd, this yields $A_\mathrm{start}=135~\unit{doy}$ and $A_\mathrm{end}=257~\unit{doy}$. We verified these as follows.
-
-For coniferous trees, we used a MODIS (Aqua/Terra) retrieved GPP product \citep{MODIS_PSN} on a $1\times 1\,\unit{km}$ patch centered at Svanhovd.
+At Svanhovd, this yields $A_\mathrm{start}=135~\unit{doy}$ and $A_\mathrm{end}=257~\unit{doy}$. We verify these as follows.
+
+For coniferous trees, we use a MODIS (Aqua/Terra) retrieved GPP product \citep{MODIS_PSN} on a $1\times 1\,\unit{km}$ patch centered at Svanhovd.
 GPP over time follows approximately a second-order polynomial function (Fig.~\ref{fig:modis_gpp}).
-By calculating the roots of the fitted polynomial, we found $A_\text{start}=122~\unit{doy}$ in 2018 and $A_\text{start}=106~\unit{doy}$ in 2019.
+By calculating the roots of the fitted polynomial, we find $A_\text{start}=122~\unit{doy}$ in 2018 and $A_\text{start}=106~\unit{doy}$ in 2019.
 $A_\text{end}$ amounts to \unit{doy}~261 and 274, respectively.
-These $A_\mathrm{end}$ values will be used for all PFTs alike.
-The resulting growing season for coniferous trees in 2019 was one month longer than in 2018.
-For $A_\text{start}$ of deciduous trees, we used temperature-degree-days (above $5\,\unit{^\circ C}$) on gridded temperature data from SeNorge.no (Appendix~Fig.~\ref{fig:greening_season_change_Svanvik}) which corresponds to the phenological timing of budburst for birch trees.
-We found \unit{doy}~129 and 130, respectively.
+These $A_\mathrm{end}$ values are used for all PFTs alike.
+The resulting growing season for coniferous trees in 2019 is one month longer than in 2018.
+For $A_\text{start}$ of deciduous trees, we use temperature-degree-days (above $5\,\unit{^\circ C}$) on gridded temperature data from SeNorge.no (Appendix~Fig.~\ref{fig:greening_season_change_Svanvik}) which corresponds to the phenological timing of budburst of birch trees.
+We find \unit{doy}~129 and 130, respectively.
 For both years, these dates coincide with the first snow-free day at the closest inland meteorological weather station at Øvre~Neiden (Sør-Varanger, NOR) and are very similar to the latitude model predictions.
@@
 This assumption is supported by observations in Rovaniemi, Finland \citep[][Supplement~Fig.~S1]{FCR:Korhonen2018}.
-Assuming the linear trend for temperture-degree-days defined GS over the past 6~decades prevails, GS at Svanhovd could converge with GS in Central Europe.
-All results are tabulated in Table~\ref{tab:sensitivity_tests_gs}.
+Supposed, the linear trend for temperture-degree-days defined GS over the past 6~decades prevails, GS at Svanhovd could converge with GS in Central Europe by the end of the century.
+All GS results are tabulated in Table~\ref{tab:sensitivity_tests_gs}.
 
@@
 ImageJ (\href{https://imagej.nih.gov/ij/}{see webpage}) was used for thresholding the images to give silhouettes of the leaves.
-After scaling the images, the program was used to find the minimum Feret diameter.
-This measure finds the largest width of a leaf at a $90^\circ$ angle to the length of the leaf (the Feret diameter).
-With this method we found that the 20 leaves had a mean width of $(3.0\pm 0.5)\,\unit{cm}$ that is smaller than in the MM parameterization ($5\,\unit{cm}$, Supplement~Table~S1).
+After scaling the images, the program was used to find the minimum Feret diameter which is defined as the largest width of a leaf at a $90^\circ$ angle to the length of the leaf.
+With this method we find that the 20~leaves had a mean width of $(3.0\pm 0.5)\,\unit{cm}$ that is smaller than in the MM parameterization ($5\,\unit{cm}$, Supplement~Table~S1).
 
 A report by \citet[][p.~52]{NINA2004} indicated an average tree height of $13.5\,\unit{m}$ in the Svanhovd area.
-One Scots pine forest at Svanhovd has been studied as a part of the ICP Forest mapping.
-There, heights were measured in 2004 when the stand was 90~years old.
-The mean tree height was $10.1\,\unit{m}$ and maximum height was $16.2\,\unit{m}$ (V.
+For one Scots pine forest at Svanhovd that has been studied as a part of the ICP Forest mapping, heights were measured in 2004 when the stand was 90~years old.
+The mean tree height then was $10.1\,\unit{m}$ and maximum height was $16.2\,\unit{m}$ (V.
 Timmerman, personal communication, June 2020).
-We used $13.5\,\unit{m}$ for both deciduous and coniferous trees as this height is found from a larger and more diverse population of species.
+For our $\mathrm{DO_3SE}$ model experiments, we use $13.5\,\unit{m}$ for both deciduous and coniferous trees as this height has been obtained from a larger and more diverse population of species compared to the IPC Forest mapping pine forest.
 
@@
 In this section, we first estimate the multiannual average stomatal conductance $\left<g_\mathrm{sto}\right>$ relative to $g_\mathrm{max}$ for each PFT from the long term meteorological data according to Eq.~(\ref{eq:stomatal}) and compare the different local climate derived parameterizations to MM (Section~\ref{subsec:res_local}).
-We also use the $\mathrm{DO_3SE}$ model to compute the ozone dose and predicted biomass loss due to ozone for the two recent GSs 2018 and 2019 for each species and set of local parameters (Section~\ref{subsec:do3se_results}).
-Finally, we put the results on ozone damage risk in 2018 and 2019 into a climatological perspective meaning how common these conditions have been in the past.
-To this end, we analyze both the past seasonality and the deviation from the norm of temperature, precipitation radiant energy, and ozone for these two years (Section~\ref{subsec:climatologies}).
+To put the $\mathrm{DO_3SE}$ modeling results on ozone dose and predicted biomass loss in 2018 and 2019 (Section~\ref{subsec:do3se_results}) into a climatological context, we analyze both the past seasonality and the deviation from the norm of temperature, precipitation radiant energy, and ozone for these two years (Section~\ref{subsec:stats}).
 
@@
 To compare the different parameterizations, we propose the following metric: From the selected set of meteorological data, we define two distinct periods in the diurnal cycle: noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and morning ($5-9\,\unit{am}$ local time) and calculate the $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ ratio (Eq.~(\ref{eq:stomatal}) assuming $f_\mathrm{SWP}=1$).
-That this choice is justified is shown in Section~\ref{subsec:do3se_results}.
+That this choice for $f_\mathrm{SWP}$ is justified is shown in Section~\ref{subsec:do3se_results}.
 Presumably, $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ around noon (the highest light intensity) is a good proxy for \chem{CO_2} uptake efficiency (growth).
@@
 
-\subsection{Climate and weather conditions in the growing season 2018/19}
-\label{sec:stats}
+\subsection{Climate and weather conditions in the growing seasons 2018 and 2019}
+\label{subsec:stats}
 In the following section, we consider the climate local to the Svanhovd region to understand the characteristics that might influence the extent of stomatal ozone dose and hence risk.
-We conduct this analysis for the two years 2018 and 2019 and relate these years to seasonality to understand how risk is influence by climate.
+We conduct this analysis for the two years 2018 and 2019 and relate these years to seasonality to understand how risk for ozone-induced damage is influence by climate.
 
@@
 
-As can be seen in Fig.~\ref{fig:data_svanvik_2018_2019}a, \chem{\chi_{O_3}} peaks in spring (April/May) and reaches its minimum in late summer (July/August).
+As can be seen in Fig.~\ref{fig:data_svanvik_2018_2019}a, ozone volume mixing ratio (VMR) \chem{\chi_{O_3}} peaks in spring (April/May) and reaches its minimum in late summer (July/August).
 The spring peak has not been captured completely in 2019, since data acquisition only started in late April.
-In summer 2018 (June--August), high ozone concentrations ($\chem{\chi_{O_3}} > 40\,\unit{ppb}$) were recorded 50 times.
+In summer 2018 (June--August), high ozone VMR ($\chem{\chi_{O_3}} > 40\,\unit{ppb}$) were recorded 50 times.
 The highest summer ozone VMR ($\chem{\chi_{O_3}} = 50.2\,\unit{ppb}$) was measured on July~25.
@@
 Hourly averaged $2\,\unit{m}$ temperatures above $20\,\unit{^\circ C}$ occurred more regularly in July 2018 than in 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}b).
-In 2018, spring temperature regularly rose above freezing only in May, while in 2019 this occurred already early in March/April indicating a later start of the GS in 2018.
+In 2018, spring temperature regularly rose above freezing only in May, while in 2019 this occurred already early in March/April.
+This indicates an earlier start of the GS in 2019 compared to 2018.
 More rain events with accumulated daily precipitation ($\sum_d \mathrm{Precip}$) above $10\,\unit{mm}$ occurred in the summer of 2018 compared to 2019 (Fig.~\ref{fig:data_svanvik_2018_2019}c).
@@
   %\includegraphics[width=12cm]{fig02}.
-  \caption{Observational data from atmospheric monitoring at Svanhovd in 2018/19.
+  \caption{Observational data from atmospheric monitoring at Svanhovd in 2018 and 2019.
 The hatched areas indicate periods without ozone monitoring data.
@@
 
-Climatologies (typically multi-year monthly averages) of temperature, precipitation, and radiant energy $Q_\mathrm{e}$ for Svanhovd were derived for the period 1992--2012 and are shown as box plots in Fig.~\ref{fig:climatediagram}.
-Therein, the colored horizontal lines indicate the median and the triangles the mean value for the respective month.
+Climatologies (typically multi-annual monthly averages) of temperature, precipitation, and radiant energy $Q_\mathrm{e}$ for Svanhovd were derived for the period 1992--2012 and are shown as box plots in Fig.~\ref{fig:climatediagram}.
+Radiant energy is the diurnal integral of $Q_0$.
+The colored horizontal lines indicate the median and the triangles the mean value for the respective month.
 A box denotes the interquartile range (IQR).
@@
 The coldest ever recorded temperature was $-45.2\,\unit{^\circ C}$ on January~27, 1999, while the highest temperature ($29.4\,\unit{^\circ C}$) occurred on July~16 the same year.
-As can be seen mean and median are well aligned in the center of the IQR for most months meaning that temperature anomalies are symmetrically distributed.
+As can be seen, mean and median are well aligned in the center of the IQR for most months meaning that temperature anomalies are symmetrically distributed.
 
@@
 
-The climatology for global radiant energy $Q_e$ in \unit{J\,m^{-2}} (the integral of $Q_0$) is zero in November--January (polar night) and reaches its maximum in May--July (midnight sun conditions).
+The climatology for global radiant energy $Q_e$ in \unit{J\,m^{-2}} is zero in November--January (polar night) and reaches its maximum in May--July (midnight sun conditions).
 Mean and median are well aligned.
@@
 Mean and median are well aligned and centered indicating a symmetric distribution for each month.
-The highest recorded \chem{\chi_{O_3}} occur in June most likely due to forest fires as mentioned above.
+The highest recorded \chem{\chi_{O_3}} occur in June most likely due to forest fires as mentioned previously.
 
@@
 
-\subsubsection{2018/19 growing season anomalies}
+\subsubsection{2018 and 2019 growing season anomalies}
 \label{subsec:anomalies}
@@
 %Ozone
-We used the bias-corrected and cross-calibrated ozone climatology \citep{ACP:Falk2021} to assess the ozone anomalies in 2018/19.
+We used the bias-corrected and cross-calibrated ozone climatology \citep{ACP:Falk2021} to assess the ozone anomalies in 2018 and 2019.
 Note that we have not recorded ozone during winter (refer to Section~\ref{subsec:weather}).
@@
   \includegraphics[width=12cm]{anomalies}
-  \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as difference from median for each month.
+  \caption{2018 and 2019 anomalies of key environmental variables at Svanhovd displayed as difference from median for each month.
 The hatched area between indicates the IQR.
@@
 By construction, $\mathrm{POD_1}$ from local climate derived parmeterizations is always larger than estimates with the MM parameterization.
-This means that the risk of ozone damage could be underestimated especially in more normal years like 2019 (see Section~\ref{sec:stats} for details).
+This means that the risk of ozone damage could be underestimated especially in more normal years like 2019 (see Section~\ref{subsec:stats} for details).
 Establishing dedicated subarctic parameterizations are most likely required to properly assess risk.
@@
 Our method differs from NDVI based determination of GS \citep{IGARSS:Hogda2001,IJB:Karlsen2007, RS:Hogda2013}.
-The \citet{IJB:Shutova2006} data indicate an earlier end of GS for birch compared to 2018/19.
+The \citet{IJB:Shutova2006} data indicate an earlier end of GS for birch compared to 2018 and 2019.
 Therefore, $\mathrm{POD_1}$ for deciduous trees could be underestimated by $1.8\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $3.0\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019, respectively if a climatological GS is used.
@@
 As shown by \citet{ACP:Bueker2012}, soil water potential is a very important factor for modeling stomatal conductance in the $\mathrm{DO_3SE}$ model.
-Using our best estimate of the soil type at Svanhovd, we found that SWP under 2018/19 meteorological conditions was mainly negligible when considering the localized growing season.
+Using our best estimate of the soil type at Svanhovd, we found that SWP under 2018 and 2019 meteorological conditions was mainly negligible when considering the localized growing season.
 We have not assessed the effect of the plants' response to VPD in detail, but the climatological data indicated that conditions at Svanhovd are shifting towards a more VPD limited regime that could limit $\mathrm{POD_y}$ under drought stress.
@@
 
-To assess the $G_\mathrm{start}$  and $G_\mathrm{end}$ of the growing season for coniferous trees at Svanhovd in 2018/19, we used the gross photosynthesis product (GPP) of MODIS AQUA/TERRA over a $1\times 1\,\unit{km}$ area centered at Svanhovd.
+To assess the $G_\mathrm{start}$  and $G_\mathrm{end}$ of the growing season for coniferous trees at Svanhovd in 2018 and 2019, we used the gross photosynthesis product (GPP) of MODIS AQUA/TERRA over a $1\times 1\,\unit{km}$ area centered at Svanhovd.
 MODIS data indicate a higher photosynthetic activity in 2018 than in 2019.

### 2022-06-02 VegOzone.tex
Last fine tuning

@@
 \begin{align}
-  A_\mathrm{start} &= 105 + (LAT-50\,\unit{^\circ N})\cdot 1.5\\
-  A_\mathrm{end} &= 297 - (LAT-50\,\unit{^\circ N})\cdot 2
+  GS_\mathrm{start} &= 105 + (LAT-50\,\unit{^\circ N})\cdot 1.5\\
+  GS_\mathrm{end} &= 297 - (LAT-50\,\unit{^\circ N})\cdot 2
 \end{align}
 %
-At Svanhovd, this yields $A_\mathrm{start}=135~\unit{doy}$ and $A_\mathrm{end}=257~\unit{doy}$. We verify these as follows.
+At Svanhovd, this yields $GS_\mathrm{start}=135~\unit{doy}$ and $GS_\mathrm{end}=257~\unit{doy}$. We verify these as follows.
 
@@
 GPP over time follows approximately a second-order polynomial function (Fig.~\ref{fig:modis_gpp}).
-By calculating the roots of the fitted polynomial, we find $A_\text{start}=122~\unit{doy}$ in 2018 and $A_\text{start}=106~\unit{doy}$ in 2019.
-$A_\text{end}$ amounts to \unit{doy}~261 and 274, respectively.
-These $A_\mathrm{end}$ values are used for all PFTs alike.
+By calculating the roots of the fitted polynomial, we find $GS_\text{start}=122~\unit{doy}$ in 2018 and $GS_\text{start}=106~\unit{doy}$ in 2019.
+$GS_\text{end}$ amounts to \unit{doy}~261 and 274, respectively.
+These $GS_\mathrm{end}$ values are used for all PFTs alike.
 The resulting growing season for coniferous trees in 2019 is one month longer than in 2018.
-For $A_\text{start}$ of deciduous trees, we use temperature-degree-days (above $5\,\unit{^\circ C}$) on gridded temperature data from SeNorge.no (Appendix~Fig.~\ref{fig:greening_season_change_Svanvik}) which corresponds to the phenological timing of budburst of birch trees.
+For $GS_\text{start}$ of deciduous trees, we use temperature-degree-days (above $5\,\unit{^\circ C}$) on gridded temperature data from SeNorge.no (Appendix~Fig.~\ref{fig:greening_season_change_Svanvik}) which corresponds to the phenological timing of budburst of birch trees.
 We find \unit{doy}~129 and 130, respectively.
@@
     \tophline
-    Species & Year & $A_\mathrm{start}$ & $A_\mathrm{end}$\\
+    Species & Year & $GS_\mathrm{start}$ & $GS_\mathrm{end}$\\
     \middlehline
@@
 In Figure~\ref{fig:anomalies_svanvik}, we show anomalies for both temperature, precipitation, radiant energy, and ozone VMR.
-We defined anomalies as median deviation from the climatological median.
+Here, we define anomalies as median deviation from the climatological median which we have shown to align well with the climatological mean.
 The hatched area indicates the respective IQR.
 
-Temperature anomalies (Fig.~\ref{fig:anomalies_svanvik}a) show enhanced temperatures throughout summer (May--September) 2018 breaching the IQR two times while March was unusually cold.
-No notably enhanced temperatures occurred in 2019 indicating that 2019 was an average year.
-
-Similarly, the precipitation anomalies are displayed in Fig.~\ref{fig:anomalies_svanvik}b.
+Temperature anomalies $\Delta T$ (Fig.~\ref{fig:anomalies_svanvik}a) show enhancements throughout summer (May--September) 2018 breaching the IQR two times while March was unusually cold.
+No notable enhancement occurred in 2019 indicating that 2019 was an average year with respect to temperature.
+Similarly, the precipitation anomalies $\Delta\Sigma\mathrm{Precip}$ are displayed in Fig.~\ref{fig:anomalies_svanvik}b.
 March 2018 saw more precipitation than usual but mainly as snow.
@@
 2019 was rather average throughout the year, but unusually wet in June and dry in July.
-
-Radiant energy anomalies are presented in Fig.~\ref{fig:anomalies_svanvik}c.
+Radiant energy anomalies $\Delta Q_e$ are presented in Fig.~\ref{fig:anomalies_svanvik}c.
 In summer 2018 (May/July), $Q_e$ was extraordinarily higher than usual, while 2019 was generally an average year with a rather dark spring (April/May) and late summer (August).
-
 %Ozone
 We used the bias-corrected and cross-calibrated ozone climatology \citep{ACP:Falk2021} to assess the ozone anomalies in 2018 and 2019.
-Note that we have not recorded ozone during winter (refer to Section~\ref{subsec:weather}).
+As no \chem{[O_3]} was recorded during winter (refer to Section~\ref{subsec:weather}), no anomalies are available for these months.
 In April/May 2018, ozone was unusually elevated and stayed higher than usual through the whole GS.
-The highest anomaly in 2019 occurred in September at the end of the GS.
-The forest fires in Sweden in July enhanced ozone notably at Svanhovd.
+The highest anomaly in 2019 occurred in September at the end of the GS, as the spring peak had not been fully captured.
+The forest fires in Sweden in July 2018 enhanced ozone notably at Svanhovd.
 
@@
 
-We modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per PLA for the three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using the MM parameterization and our climate derived local temperature (\emph{cold}, \emph{subarctic}) and light (PPFD0.8, PPFD1.2) parameterizations.
+Using the $\mathrm{DO_3SE}$ model, we modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per PLA for the three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using the MM parameterization and our climate derived local temperature (\emph{cold}, \emph{subarctic}) and localized light (PPFD0.8, PPFD1.2) parameterizations.
+All model experiments have been performed with localized GS, canopy height, and leaf dimensions.
+(Section~\ref{subsec:gs_est}--\ref{subsec:soil})
 
@@
 Critical levels for reduction in biomass \citep[deciduous forest $4\,\unit{\%}$, coniferous forest $2\,\unit{\%}$, and grasslands $10\,\unit{\%}$ adopted from][]{ICP:MappingManual2017,ESPR:Hayes2021} are shown as horizontal dashed lines.
-The different background colors indicate the local climate derived temperature response function.
+The different background colors highlight the temperature response functions.
 The light response functions are denoted by $+$ (PPFD1.2) and $-$ (PPFD0.8).
 Open symbols indicate, the effect on $\mathrm{POD_y}$ of any soil water stress is accounted for.
-All model experiments have been performed with localized GS, canopy height, and leaf dimensions.
-
-We identified that, in general, an increasing cold tolerance represented by a shift in the temperature response function (MM~$<$~cold~$<$~subarctic) leads to an increase in $\mathrm{POD_1}$ in both investigated years (2018 and 2019) compared to the MM parameterization.
+
+We identify that, in general, an increasing cold tolerance represented by a shift in the temperature response function (MM~$<$~cold~$<$~subarctic) leads to an increase in $\mathrm{POD_1}$ in both investigated years (2018 and 2019) compared to the MM parameterization.
 The more extreme the local climate derived temperature response function the higher $\mathrm{POD_1}$.
@@
 An earlier start and overall longer GS in 2019 leads to higher $\mathrm{POD_1}$ compared to 2018 despite the occurrence of more episodes of elevated \chem{O_3} in 2018 than in 2019 (see Section~\ref{subsec:anomalies} for details on the particular features of these two years).
-Taking into account SWP, $\mathrm{POD_1}$ was only notably reduced in 2018 (for deciduous trees).
-In all other model experiments SWP has little relevance justifying the choice in Section~\ref{subsec:res_local}.
+Taking into account SWP, $\mathrm{POD_1}$ is only notably reduced in 2018 (for deciduous trees).
+In all other model experiments, SWP has little impact on the results justifying the choice in Section~\ref{subsec:res_local}.
 A larger aperture of stomata at low light intensities (PPFD0.8) causes an increase in $\mathrm{POD_1}$ compared to MM while a smaller aperture (PPFD1.2) is leading to reduced $\mathrm{POD_1}$.
-Due to the shape of $f_\mathrm{light}$, a symmetric variation can lead to an asymmetric response in $\mathrm{POD_1}$.
-
-For deciduous trees (Fig.~\ref{fig:pody_rel}a), all model experiments exceed the CL by ($5-16)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
+Due to the shape of $f_\mathrm{light}$, the symmetric variation can lead to an asymmetric response in $\mathrm{POD_1}$.
+
+For deciduous trees (Fig.~\ref{fig:pody_rel}a), all model experiments exceed the CL by $5-16\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
 In 2019, the lowest $\mathrm{POD_1}$ estimate (MM, PPFD1.2) amounts to $10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and the highest (subarctic, PPFD0.8) to $21\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
-In 2018, this range is smaller $(12-18)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
+In 2018, this range is smaller ($12-18\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
 The sensitivity to $f_\mathrm{light}$ is largest in the subarctic parameterization and amounts to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
 Herein, the difference of PPFD0.8 compared to MM $f_\mathrm{light}$ is denoted as superscript and for PPFD1.2 as subscript.
-A low soil water potential caused a reduction in $\mathrm{POD_1}$ in 2018 in particular when using PPFD0.8.
+A low soil water potential caused a reduction in $\mathrm{POD_1}$ in 2018, in particular when using PPFD0.8.
 This indicates that water limiting conditions can reduce the ozone damage risk.
 
-For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL by $(0-5)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$.
+For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL by ($0-5\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
 The different climate derived local parameterizations do not show much impact compared to MM indicating that the MM's Boreal Norway spruce parameterization is well-suited for a subarctic climate.
 We find higher $\mathrm{POD_1}$ in 2019 than in 2018.
-The maximum difference in $\mathrm{POD_1}$ determined between subarctic and cold parameterizations with respect to MM amounts to $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+The maximum difference in $\mathrm{POD_1}$ determined between subarctic and cold parameterizations with respect to MM is $0.5-1.0\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $2.0-3.3\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
 The maximum difference due to $f_\mathrm{light}$ amounts to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
@@
 Perennial grassland has the smallest $\mathrm{POD_1}$, but a similarly large response to the local climate derived temperature parameterization as deciduous trees, and the lowest sensitivity to the light threshold.
-Soil water limitation has little relevance.
-The difference in $\mathrm{POD_1}$ for subarctic and cold types with respect to MM amounts to $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+Soil water limitation has little impact on the results.
+The difference in $\mathrm{POD_1}$ for subarctic and cold types with respect to MM amounts to $2.2-3.4\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $4.4-6.4\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
 The maximum difference in $\mathrm{POD_1}$ determined from the PPFD0.8 (superscript) and PPFD1.2 (subscript) variation of $f_\mathrm{light}$ amount to $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ (2018) and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ (2019), respectively.
@@
 
-We estimated the biomass reductions in accordance to the CLs \citep{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme climate derived local temperature parameterization (subarctic) for each PFT.
+In the following, we estimate the biomass reductions in accordance to the CLs \citep{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme climate derived local temperature parameterization (subarctic) for each PFT.
 The estimates, assuming no soil water stress, are given in Table~\ref{tab:biomass_reduction}.
-We defined an uncertainty range from the difference in biomass reduction caused by the variation of $f_\mathrm{light}$.
-Superscript refers to PPFD0.8 and subscript to PPFD1.2.
-Herein, the reported range indicated by dots corresponds to the limiting effect by soil water stress.
-Because soil water stress reduces $\mathrm{POD_1}$, biomass reduction is also reduced.
+We define an uncertainty range from the difference in biomass reduction caused by the variation of $f_\mathrm{light}$.
+Again, superscript refers to PPFD0.8 and subscript to PPFD1.2.
+The reported range indicated by dots corresponds to the limiting effect by soil water stress.
 The biomass reduction ranges between $2.5\,\unit{\%}$ (coniferous, 2018) and $17.4\,\unit{\%}$ (deciduous, 2019).
-For deciduous trees this amounts to a reduction by up to $55\,\unit{\%}$ compared to MM.
-Coniferous trees are the least affected PFT suggesting the Boreal MM parameterization is appropriate for this PFT.
+For deciduous trees, this amounts to a reduction by up to $55\,\unit{\%}$ compared to MM.
+Coniferous trees are least affected by ozone-induced damage.
 
@@
 In future scenarios in which persistent heatwaves like in 2018 may become more common, the risk for ozone-induced damage is probably reduced for vegetation adapted to cold growing conditions.
-A longer growing season promoted by climate change without weather extremes like in 2019 is likely to increase this risk.
+A longer growing season promoted by climate change without weather extremes like in 2019 is, hence, likely to increase the risk.
 
@@
 In Sweden, about 500 fires had been reported (five times more than in a usual summer), and an estimated total of $25000\,\unit{hectare}$ burned down in central Sweden \citep{SOU2019}.
-We found coincident peak \chem{\chi_{O_3}} in ozone monitoring data at Svanhovd in July.
-Overall, ozone was more enhanced in 2018 than 2019.
-It is very likely that elevated ozone, like during the 2003 drought period \citep{JGR:Solberg2018}, was promoted by a combination of various factors such as wildfires, reduced cloud cover (increased solar radiation), reduced dry deposition and turbulent mixing due to the stagnant weather conditions, and increased BVOC emissions.
-As such events may become more frequent in the future, conditions at Svanhovd for the growing seasons 2018 could serve as a reference for probable future conditions in northern Fennoscandia and arising ozone-induced damage risk.
-Conditions in 2019 represent the present climate.
+Overall, ozone was more enhanced in 2018 than 2019 and it is very likely that elevated ozone, like during the 2003 drought period \citep{JGR:Solberg2018}, was promoted by a combination of various factors such as wildfires, reduced cloud cover (increased solar radiation), reduced dry deposition and turbulent mixing due to the stagnant weather conditions, and increased BVOC emissions.
+As such events may become more frequent in the future, conditions at Svanhovd for the growing seasons 2018 could serve as a reference for probable future conditions in northern Fennoscandia.
 
@@
 It is essential for ozone risk assessments in the subarctic to define plant physiology and traits such as phenological states properly.
-As no independent ozone flux measurements are available at Svanhovd, we could not fully validate our methodology at this point.
-Dry deposition velocities $V_\mathrm{d}$ for coniferous trees from comprehensive records over Pine forests from Hyytilälä, Finland \citep{BER:Keronen2003} yield summer maximum $V_\mathrm{d}\sim 11\,\unit{mm\,s^{-1}}$.
-The $\mathrm{DO_3SE}$ model derived summer maximum $V_\mathrm{d}\sim 9\,\unit{mm\,s^{-1}}$ with minimal differences between the parameterizations.
+As no independent ozone flux measurements are available at Svanhovd, we cannot fully validate our methodology at this point.
+Dry deposition velocities $V_\mathrm{d}$ for coniferous trees from comprehensive records over Pine forests from Hyytilälä, Finland \citep{BER:Keronen2003} suggest a summer maximum $V_\mathrm{d}\sim 11\,\unit{mm\,s^{-1}}$.
+The $\mathrm{DO_3SE}$ model derived summer maximum $V_\mathrm{d}$ is $\sim 9\,\unit{mm\,s^{-1}}$ with minimal differences between the parameterizations (see supplementary material Figs.~S1--S3).
 The highest $V_\mathrm{d}$ is found in summer 2019 using the cold paramterization.
@@
 No comparable data is available for northern Fennoscandia or deciduous tress and perennial grassland.
-Perennial grassland and deciduous trees displayed strong susceptibility to our methodology of local climate derived parameterization (increase in frequency of maximum $V_\mathrm{d}$ throughout the GS, see supplementary material Figs.~S1--S3) but are within the range ($v\mathrm{d} = (2-9)\,\unit{mm\,s^{-1}}$ for grassland) given in the literature \citep[e.g.][]{JGR:Jacob1992,AE:Padro1996,RG:Clifton2020}.
+Perennial grassland and deciduous trees displayed strong susceptibility to our methodology of local climate derived parameterization but are within the range ($V_\mathrm{d} = 2-9\,\unit{mm\,s^{-1}}$ for grassland) given in the literature \citep[e.g.][]{JGR:Jacob1992,AE:Padro1996,RG:Clifton2020}.
 Independent flux measurements over these vegetation types in a subarctic climate are thus crucial for evaluation and validation.
@@
 In contrast, no such damage was found on the clovers in 2019.
-This might indicate a higher risk of ozone damage under warmer climate conditions for these specific plant species in a subarctic climate.
+This might indicate a higher risk of visible ozone injuries under warmer climate conditions for these specific plant species in a subarctic climate.
 However, visible injuries are not always correlated with a reduction in biomass \citep[and citations therein]{CRG:Felzer2007} and the $\mathrm{DO_3SE}$ model does not estimate visible injury which is often related to short periods of very high ozone uptake.
 In this regard, open-top chamber (OTC) experiments performed in northern Finland where Scots pine and downy birch seedlings were exposed to elevated ozone concentrations attributed a reduction in biomass and reproduction with visible injuries explicitly to peak \chem{O_3} concentrations and fast phenological development at high growth rate \citep{Amb:Manninen2009}.
-The \citet{Amb:Manninen2009} results from open-top chamber (OTC) experiments on downy birch (\emph{Betula pubescens}) and mountain birch (\emph{Betula pubescens toruosa}), native to subarctic regions, indicated reductions in both biomass, root:shoot ratio, and visible leaf damage under elevated ozone treatments ($\left<\chem{\chi_{O_3}}\right> = 36-54\,\unit{ppb}$).
+The \citet{Amb:Manninen2009} results on downy birch (\emph{Betula pubescens}) and mountain birch (\emph{Betula pubescens toruosa}), native to subarctic regions, indicated reductions in both biomass, root:shoot ratio, and visible leaf damage under elevated ozone treatments ($\left<\chem{\chi_{O_3}}\right> = 36-54\,\unit{ppb}$).
 Although, Scots pine (\emph{Pinus sylvestris}) is considered to be more ozone tolerant due to an absence of visible injuries \citep{Amb:Girgzdiene2009}, \citet{Amb:Manninen2009} found chlorophyll:carotenoid ratio and polyamine reductions under elevated ozone concentrations indicating susceptibility to ozone also in these species.
@@
 That might explain why we observed visible injuries in the ozone garden to a larger extent in 2018 than in 2019.
-An associated damage function relating ozone susceptibility to leaf age as proposed by \citet{AE:Musselman2006} or phenological state might improve the predictions.
-Such a damage function could be different from $f_\mathrm{phen}$.
+An associated damage function for all PFTs relating visible ozone-injuries or phenological state might improve such predictions.
 
@@
 The results coincide with the first snow free days in our target region.
-For the period $1994-2003$ \citet{IJB:Shutova2006} reported an observed average start of the GS for mountain birch at Svanhovd at $\unit{doy}~154$ and end of GS at $\unit{doy}~ 229$.
+For the period $1994-2003$, \citet{IJB:Shutova2006} reported an observed average start of the GS for mountain birch at Svanhovd at $\unit{doy}~154$ and end of GS at $\unit{doy}~229$.
 These dates are well in line with our estimated average start and end of GS for the period $1960-2020$ (Fig.~\ref{fig:greening_season_change_Svanvik}).
 Our estimates for perennial grassland, are supported by observational data from Rovaniemi (Finland) \citep{FCR:Korhonen2018}.
-The end of GS inferred from the zero-crossing of MODIS observed GPP is in line with independent estimates from the gridded observational temperature.
-Our method differs from NDVI based determination of GS \citep{IGARSS:Hogda2001,IJB:Karlsen2007, RS:Hogda2013}.
-The \citet{IJB:Shutova2006} data indicate an earlier end of GS for birch compared to 2018 and 2019.
-Therefore, $\mathrm{POD_1}$ for deciduous trees could be underestimated by $1.8\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $3.0\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019, respectively if a climatological GS is used.
-For perennial grassland the resulting underestimation would amount to $2.0\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $2.3\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
+The end of GS inferred from the zero-crossing of MODIS observed GPP method differs from NDVI based determination of GS \citep{IGARSS:Hogda2001,IJB:Karlsen2007, RS:Hogda2013}.
+The \citet{IJB:Shutova2006} data indicate an earlier end of GS for birch compared to our estimates for 2018 and 2019.
+Therefore, our resulting $\mathrm{POD_1}$ for deciduous trees could be overestimated by $1.8\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $3.0\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019, respectively compared to this climatological GS.
+For perennial grassland the resulting overestimation would amount to $2.0\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $2.3\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.
 This demonstrates the importance of an accurate determination of GS for $\mathrm{POD_1}$ based risk assessments.
@@
 As shown by \citet{ACP:Bueker2012}, soil water potential is a very important factor for modeling stomatal conductance in the $\mathrm{DO_3SE}$ model.
-Using our best estimate of the soil type at Svanhovd, we found that SWP under 2018 and 2019 meteorological conditions was mainly negligible when considering the localized growing season.
+Using our best estimate of the soil type at Svanhovd, we found that SWP under 2018 and 2019 meteorological conditions was mainly negligible.
 We have not assessed the effect of the plants' response to VPD in detail, but the climatological data indicated that conditions at Svanhovd are shifting towards a more VPD limited regime that could limit $\mathrm{POD_y}$ under drought stress.
 
-Due to a climate change-induced convergence of start of GS with the ozone spring peak \citep{EP:Karlsson2007}, vegetation might become more exposed to higher ozone concentrations in early phenological stages.
+Due to a climate change-induced convergence of the start of GS with the ozone spring peak \citep{EP:Karlsson2007}, vegetation might become more exposed to higher ozone concentrations in early phenological stages.
 However, the decline of this ozone spring peak is partly caused by dry deposition of ozone to vegetation itself.
@@
 We found that the difference in biomass reduction compared to Boreal parmeterizations \citep{ICP:MappingManual2017} could amount to $55\,\unit{\%}$ in deciduous trees and $29-43\,\unit{\%}$ in perennial grassland.
-But the biomass reduction functions were established based on studies from less extreme climates, hence, these results ought to be treated with care.
+But the used biomass reduction functions were established based on studies from less extreme climates, hence, these results ought to be treated with care.
 In particular, CLs for deciduous and coniferous trees were breached in both years and for all parameterizations.
@@
 
-Beyond the risk assessment of ozone-induced damage, our climate derived local parameterization Jarvis-type temperature response functions, once verified by observations, can have important implications on land-surface modeling still using only Boreal parameterizations in global models.
+Beyond the risk assessment of ozone-induced damage, our climate derived local Jarvis-type temperature response functions, once verified by observations, can have important implications on land-surface modeling still using only Boreal parameterizations in global models.
 Problems in estimating the productivity of species especially in the Arctic regions occur due to PFTs which are not suitable for the climate.
@@
 
-To assess the $G_\mathrm{start}$  and $G_\mathrm{end}$ of the growing season for coniferous trees at Svanhovd in 2018 and 2019, we used the gross photosynthesis product (GPP) of MODIS AQUA/TERRA over a $1\times 1\,\unit{km}$ area centered at Svanhovd.
+To assess the $GS_\mathrm{start}$  and $GS_\mathrm{end}$ of the growing season for coniferous trees at Svanhovd in 2018 and 2019, we used the gross photosynthesis product (GPP) of MODIS AQUA/TERRA over a $1\times 1\,\unit{km}$ area centered at Svanhovd.
 MODIS data indicate a higher photosynthetic activity in 2018 than in 2019.
@@
 %
-with the form parameters $m_i$ to the data. Numerically, we retrieved the roots as $G_\mathrm{start}$ 122/106 and $G_\mathrm{end}$ 261/274 for 2018 and 2019, respectively.
+with the form parameters $m_i$ to the data. Numerically, we retrieved the roots as $GS_\mathrm{start}$ 122/106 and $GS_\mathrm{end}$ 261/274 for 2018 and 2019, respectively.
 
@@
   \includegraphics[width=8.3cm]{modis_gpp}
-  \caption{Estimated $G_\mathrm{start}$ and $G_\mathrm{end}$ of growing season for coniferous trees from MODIS Aqua/Terra gross primary productivity (GPP) product.
+  \caption{Estimated $GS_\mathrm{start}$ and $GS_\mathrm{end}$ of growing season for coniferous trees from MODIS Aqua/Terra gross primary productivity (GPP) product.
 A $1\times 1\,\unit{km}$ area around Svanhovd was selected.

### 2022-06-02 VegOzone.tex
Remove newpage comand
Fix typo

@@
 To date, only a few studies have assessed the response of subarctic vegetation to variations in climate and air pollution.
-This study looks to fill this knowledge gap by examining the uncertainty in risk assessments that result from using Boreal rather than subrctic paramterizations.
-With no parameterization of stomatal response of subarctic vegetation available, we propose an adjustment of existing Boreal parameterizations based on local climate condition.
+This study looks to fill this knowledge gap by examining the uncertainty in risk assessments that result from using Boreal rather than subarctic paramterizations.
+With no parameterization of stomatal response of subarctic vegetation available, we propose an adjustment of existing Boreal parameterizations based on local climate conditions.
 To this end, we vary the temperature response function according to the climatological temperature distribution during the growing season at our site.
@@
 
-\newpage
-
 %\copyrightstatement{TEXT}  %% This section is optional and can be used for copyright transfers.
@@
 Ozone causes reduced photosynthesis, visible foliar injury, early senescence, and programmed cell death of plants \citep{PCE:Kangasjarvi2005}.
-Such ozone induced injury and damage can cause losses in crop \citep{PJ:Ainsworth2017}, forest \citep{GCB:Wittig2009,EP:Matyssek2012}, and productive grassland productivity and interfere with the global carbon cycle \citep{BGS:Lombardozzi2013}.
+Such ozone-induced injury and damage can cause losses in crop \citep{PJ:Ainsworth2017}, forest \citep{GCB:Wittig2009,EP:Matyssek2012}, and productive grassland productivity and interfere with the global carbon cycle \citep{BGS:Lombardozzi2013}.
 Even though long-term (pre-1950) observations of ground-level ozone concentrations are scarce; evidence from trend analysis of those few sites for which such data do exist suggests that since the industrial revolution, background concentrations in the northern hemisphere have at least doubled and continued to increase \citep{IPCC2013} until recently.

### 2022-06-02 VegOzone.tex
Check with gramarly

@@
 To this end, we vary the temperature response function according to the climatological temperature distribution during the growing season at our site.
-We refer to this approach as a 'local climate derived parameterization' of stomatal response functions.
+We refer to this approach as a 'local climate-derived parameterization' of stomatal response functions.
 We also study the impact of the stomata aperture parameterization at low light intensities.
@@
 Our modeling study suggests that using standard Boreal parameterizations in assessments likely leads to underestimating the risk of ozone damage in this region for some species.
-We find that local climate derived parameterizations for subarctic vegetation result in an ozone-induced biomass loss of $2.5-17.4\,\unit{\%}$.
+We find that local climate-derived parameterizations for subarctic vegetation result in an ozone-induced biomass loss of $2.5-17.4\,\unit{\%}$.
 For some species, this loss is up to $55\,\unit{\%}$ larger than that projected from Boreal parameterizations used in general assessments.
@@
 Ozone causes reduced photosynthesis, visible foliar injury, early senescence, and programmed cell death of plants \citep{PCE:Kangasjarvi2005}.
-Such ozone-induced injury and damage can cause losses in crop \citep{PJ:Ainsworth2017}, forest \citep{GCB:Wittig2009,EP:Matyssek2012}, and productive grassland productivity and interfere with the global carbon cycle \citep{BGS:Lombardozzi2013}.
+Such ozone-induced injury and damage can cause losses in crops \citep{PJ:Ainsworth2017}, forest \citep{GCB:Wittig2009,EP:Matyssek2012}, and productive grassland productivity and interfere with the global carbon cycle \citep{BGS:Lombardozzi2013}.
 Even though long-term (pre-1950) observations of ground-level ozone concentrations are scarce; evidence from trend analysis of those few sites for which such data do exist suggests that since the industrial revolution, background concentrations in the northern hemisphere have at least doubled and continued to increase \citep{IPCC2013} until recently.
 Recent trend analyses for Europe indicate that a maximum was reached in 2007 \citep{AE:Derwent2018}, after which tropospheric background ozone concentrations started to level off or even decline \citep{ESA:Cooper2014, ACP:Wespes2018,ESA:Gaudel2018}.
-This can be attributed to a successful implementation of air quality regulations coupled with economic restructuring, reducing the number of episodes of peak concentrations especially in summertime \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
+This can be attributed to a successful implementation of air quality regulations coupled with economic restructuring, reducing the number of episodes of peak concentrations, especially in summertime \citep[e.g.,][]{ESA:Fleming2018, ESA:Mills2018}.
 However, changes in environmental conditions associated with climate change, such as an increase in the frequency of heatwaves, could negate the effectiveness of emission reductions on ozone concentrations and associated impacts \citep{NCC:Lin2020}.
@@
 A longer growing season is prolonging the time in which vegetation can accumulate an ozone dose.
-For natural and semi-natural vegetation types that are already subject to rapid climatic changes known as Arctic amplification \citep{AMAP2012,IPCC2013}, these factors could promote a higher potential risk to ozone-induced damage in the not so distant future.
+For natural and semi-natural vegetation types that are already subject to rapid climatic changes known as Arctic amplification \citep{AMAP2012,IPCC2013}, these factors could promote a higher potential risk of ozone-induced damage in the not so distant future.
 Recently, \citet{ESPR:Hayes2021} showed that the highest sensitivity to future \chem{[O_3]} variability is expected in summer when vegetation is most productive.
@@
 Currently, parameterizations are available for some species for particular climatic regions e.g.
-Boreal, temperate and Mediterranean; there is no parameterization for subarctic vegetation.
+Boreal, temperate, and Mediterranean; there is no parameterization for subarctic vegetation.
 Given the unique conditions of the subarctic (i.e.
-long duration of daylight in the summer and cold temperatures) it would be expected that this vegetation type has a particular physiological response to the prevailing climatic conditions different to Boreal species.
+long duration of daylight in the summer and cold temperatures), it would be expected that this vegetation type has a particular physiological response to the prevailing climatic conditions different to Boreal species.
 
@@
 
-In Section~\ref{sec:dam}, we give a brief summary of the key components and meteorological input data of the $\mathrm{DO_3SE}$ model that we use to assess the ozone risk.
-Subsequently, we propose a local climate derived temperature response function and explore the impact of the stomata aperture parameterization at low light intensities.
+In Section~\ref{sec:dam}, we give a summary of the key components and meteorological input data of the $\mathrm{DO_3SE}$ model that we use to assess the ozone risk.
+Subsequently, we propose a local climate-derived temperature response function and explore the impact of the stomata aperture parameterization at low light intensities.
 Start and end of the growing season are inferred from satellite observed gross primary productivity (GPP).
 We explore how risk might be over- or underestimated using Boreal parameterization (Section~\ref{sec:results}).
-In Section~\ref{subsec:res_local}, we evaluate the local climate derived parameterizations by means of average relative stomatal conductance.
+In Section~\ref{subsec:res_local}, we evaluate the local climate-derived parameterizations by means of average relative stomatal conductance.
 For evaluating the $\mathrm{DO_3SE}$ model results in a climate perspective, we assess the climate anomalies of weather conditions and ozone concentrations in the 2018 and 2019 GS at our research site (Section~\ref{subsec:stats}).
-In Section~\ref{subsec:do3se_results}, we assess the difference in predicted risk for ozone-induced damage for the local climate derived parameterizations compared to MM parameters.
+In Section~\ref{subsec:do3se_results}, we assess the difference in predicted risk for ozone-induced damage for the local climate-derived parameterizations compared to MM parameters.
 We discuss our results concerning prevailing conditions and give an outlook on how future assessments of pollution risk in the subarctic could be improved in Section~\ref{sec:conc}.
@@
 Basic parameters for these PFTs are derived from Boreal paramterizations when available \citep{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as mapping manual (MM) parameterizations.
-In particular, the subarctic local climate derived parameterization for coniferous trees is based on MM's Boreal Norway spruce, deciduous trees on the MM's Boreal silver birch, and perennial grassland on MM's perennial grassland for Central Europe as a Boreal parameterization is not available.
+In particular, the subarctic local climate-derived parameterization for coniferous trees is based on MM's Boreal Norway spruce, deciduous trees on the MM's Boreal silver birch, and perennial grassland on MM's perennial grassland for Central Europe as a Boreal parameterization is not available.
 For a comprehensive list of MM parameters, consult Supplement~Table~S1.
@@
 
-We define local climate derived parameterizations for two hypothetical species (\emph{subarctic} and \emph{cold}) by increasing the overlapping area between $f_\mathrm{T}$ and the temperature PDF during the GS.
+We define local climate-derived parameterizations for two hypothetical species (\emph{subarctic} and \emph{cold}) by increasing the overlapping area between $f_\mathrm{T}$ and the temperature PDF during the GS.
 Then we construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures than the MM species.
@@
 Therefore, we shift both $T_\mathrm{min}$ and $T_\mathrm{max}$ to colder temperatures and choose $T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000.
-The most extreme of the two local climate derived parameterizations compared to the MM parameterization is the \emph{subarctic}.
+The most extreme of the two local climate-derived parameterizations compared to the MM parameterization is the \emph{subarctic}.
 This methodology is analogous to the optimal temperature acclimation experimentally found for photosynthesis \citep{PR:Way2013}.
@@
   %\includegraphics[width=8.3cm]{fig07}
-\caption{Construction of local climate derived response functions for grassland.
+\caption{Construction of local climate-derived response functions for grassland.
 (a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively.
@@
 
-Similarly, we define local climate derived $f_\mathrm{T}$ and localized $f_\mathrm{light}$ for coniferous and deciduous trees.
+Similarly, we define local climate-derived $f_\mathrm{T}$ and localized $f_\mathrm{light}$ for coniferous and deciduous trees.
 The resulting response functions are shown in Appendix~Figs.~\ref{fig:f_temp_spruce} and \ref{fig:f_temp_birch}, respectively.
@@
 Thus, coniferous trees are already active at low air temperatures and can reach $60\,\unit{\%}$ photosynthetic activity as early as day of the year~(\unit{doy})~100 but cease activity in less favorable conditions \citep{TB:Kolari2007, TP:Wallin2013}.
-To determine the local climate derived temperature parameterization for coniferous trees, we use the time series of \chem{CO_2} uptake and temperature at observation sites in southern ($\rightarrow$ \emph{cold}) and northern ($\rightarrow$ \emph{subarctic}) Finland by \citet{TB:Kolari2007}.
+To determine the local climate-derived temperature parameterization for coniferous trees, we use the time series of \chem{CO_2} uptake and temperature at observation sites in southern ($\rightarrow$ \emph{cold}) and northern ($\rightarrow$ \emph{subarctic}) Finland by \citet{TB:Kolari2007}.
 From these, we estimate an optimal temperature interval ($10\,\unit{^\circ C}\le T_\mathrm{opt} \le 15\,\unit{^\circ C}$).
@@
 All resulting temperature response functions are tabulated in Table~\ref{tab:sensitivity_tests_temp} and light response function in Table~\ref{tab:sensitivity_tests_light}.
-In Section~\ref{subsec:res_local}, we compare relative $g_\mathrm{sto}$ based on Eq.~(\ref{eq:stomatal}) for the different local climate derived parameterizations using the long term meteorological data.
+In Section~\ref{subsec:res_local}, we compare relative $g_\mathrm{sto}$ based on Eq.~(\ref{eq:stomatal}) for the different local climate-derived parameterizations using the long term meteorological data.
 
 \begin{table}[t]
-  \caption{Local climate derived temperature parameterizations. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
+  \caption{Local climate-derived temperature parameterizations. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
   \label{tab:sensitivity_tests_temp}
@@
 \label{sec:results}
-In this section, we first estimate the multiannual average stomatal conductance $\left<g_\mathrm{sto}\right>$ relative to $g_\mathrm{max}$ for each PFT from the long term meteorological data according to Eq.~(\ref{eq:stomatal}) and compare the different local climate derived parameterizations to MM (Section~\ref{subsec:res_local}).
+In this section, we first estimate the multiannual average stomatal conductance $\left<g_\mathrm{sto}\right>$ relative to $g_\mathrm{max}$ for each PFT from the long term meteorological data according to Eq.~(\ref{eq:stomatal}) and compare the different local climate-derived parameterizations to MM (Section~\ref{subsec:res_local}).
 To put the $\mathrm{DO_3SE}$ modeling results on ozone dose and predicted biomass loss in 2018 and 2019 (Section~\ref{subsec:do3se_results}) into a climatological context, we analyze both the past seasonality and the deviation from the norm of temperature, precipitation radiant energy, and ozone for these two years (Section~\ref{subsec:stats}).
 
-\subsection{Local climate derived parameterization}
+\subsection{Local climate-derived parameterization}
 \label{subsec:res_local}
@@
 For perennial grassland, the differences compared to MM are largest suggesting that the MM Central European parameterization is not appropriate in a subarctic climate.
-Deciduous trees display also a notable increase in mean stomatal conductance using our local climate derived parameterizations.
+Deciduous trees display also a notable increase in mean stomatal conductance using our local climate-derived parameterizations.
 
@@
   %\includegraphics[width=8.3cm]{fig08}
-  \caption{Proposed metric to test local climate derived parameterizations for response functions.
+  \caption{Proposed metric to test local climate-derived parameterizations for response functions.
 GS (May--August) climatological averages and standard deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon (averaged over $11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning (averaged over $5-9\,\unit{am}$ local time) are shown for (a) deciduous trees; (b) coniferous trees; (c) perennial grassland.}
@@
 
-Using the $\mathrm{DO_3SE}$ model, we modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per PLA for the three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using the MM parameterization and our climate derived local temperature (\emph{cold}, \emph{subarctic}) and localized light (PPFD0.8, PPFD1.2) parameterizations.
+Using the $\mathrm{DO_3SE}$ model, we modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per PLA for the three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using the MM parameterization and our climate-derived local temperature (\emph{cold}, \emph{subarctic}) and localized light (PPFD0.8, PPFD1.2) parameterizations.
 All model experiments have been performed with localized GS, canopy height, and leaf dimensions.
@@
 We identify that, in general, an increasing cold tolerance represented by a shift in the temperature response function (MM~$<$~cold~$<$~subarctic) leads to an increase in $\mathrm{POD_1}$ in both investigated years (2018 and 2019) compared to the MM parameterization.
-The more extreme the local climate derived temperature response function the higher $\mathrm{POD_1}$.
+The more extreme the local climate-derived temperature response function the higher $\mathrm{POD_1}$.
 In 2018, coniferous trees and perennial grassland display a reduction in $\mathrm{POD_1}$ for the subarctic parameterization compared to the cold paramerization in responds to the prescribed lower $g_\mathrm{sto}$ at higher temperatures.
@@
 For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL by ($0-5\,\unit{mmol\,\chem{O_3}\,m^{-2}}$).
-The different climate derived local parameterizations do not show much impact compared to MM indicating that the MM's Boreal Norway spruce parameterization is well-suited for a subarctic climate.
+The different climate-derived local parameterizations do not show much impact compared to MM indicating that the MM's Boreal Norway spruce parameterization is well-suited for a subarctic climate.
 We find higher $\mathrm{POD_1}$ in 2019 than in 2018.
@@
 For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations stay below the CL.
-Perennial grassland has the smallest $\mathrm{POD_1}$, but a similarly large response to the local climate derived temperature parameterization as deciduous trees, and the lowest sensitivity to the light threshold.
+Perennial grassland has the smallest $\mathrm{POD_1}$, but a similarly large response to the local climate-derived temperature parameterization as deciduous trees, and the lowest sensitivity to the light threshold.
 Soil water limitation has little impact on the results.
@@
 The magnitude of all described effects on $\mathrm{POD_1}$ varies between PFTs and years.
-By construction, $\mathrm{POD_1}$ from local climate derived parmeterizations is always larger than estimates with the MM parameterization.
+By construction, $\mathrm{POD_1}$ from local climate-derived parmeterizations is always larger than estimates with the MM parameterization.
 This means that the risk of ozone damage could be underestimated especially in more normal years like 2019 (see Section~\ref{subsec:stats} for details).
@@
 
-In the following, we estimate the biomass reductions in accordance to the CLs \citep{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme climate derived local temperature parameterization (subarctic) for each PFT.
+In the following, we estimate the biomass reductions in accordance to the CLs \citep{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme climate-derived local temperature parameterization (subarctic) for each PFT.
 The estimates, assuming no soil water stress, are given in Table~\ref{tab:biomass_reduction}.
@@
 No comparable data is available for northern Fennoscandia or deciduous tress and perennial grassland.
-Perennial grassland and deciduous trees displayed strong susceptibility to our methodology of local climate derived parameterization but are within the range ($V_\mathrm{d} = 2-9\,\unit{mm\,s^{-1}}$ for grassland) given in the literature \citep[e.g.][]{JGR:Jacob1992,AE:Padro1996,RG:Clifton2020}.
+Perennial grassland and deciduous trees displayed strong susceptibility to our methodology of local climate-derived parameterization but are within the range ($V_\mathrm{d} = 2-9\,\unit{mm\,s^{-1}}$ for grassland) given in the literature \citep[e.g.][]{JGR:Jacob1992,AE:Padro1996,RG:Clifton2020}.
 Independent flux measurements over these vegetation types in a subarctic climate are thus crucial for evaluation and validation.
@@
 
-Beyond the risk assessment of ozone-induced damage, our climate derived local Jarvis-type temperature response functions, once verified by observations, can have important implications on land-surface modeling still using only Boreal parameterizations in global models.
+Beyond the risk assessment of ozone-induced damage, our climate-derived local Jarvis-type temperature response functions, once verified by observations, can have important implications on land-surface modeling still using only Boreal parameterizations in global models.
 Problems in estimating the productivity of species especially in the Arctic regions occur due to PFTs which are not suitable for the climate.
@@
   %\includegraphics[width=8.3cm]{figB3}
-\caption{Construction of local climate derived response functions for coniferous trees.
+\caption{Construction of local climate-derived response functions for coniferous trees.
 (a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively.
@@
   %\includegraphics[width=8.3cm]{figB4}
-\caption{Construction of local climate derived response functions for deciduous trees.
+\caption{Construction of local climate-derived response functions for deciduous trees.
 (a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively.

### 2022-06-02 VegOzone.tex
Checked with gramarly

@@
 \label{sec:dam}
-In the following, we briefly describe the $\mathrm{DO_3SE}$ model used for assessing the ozone risk, its key components, input data, and gap filling procedure (Section~\ref{sec:do3se}).
+In the following, we briefly describe the $\mathrm{DO_3SE}$ model used for assessing the ozone risk, its key components, input data, and gap-filling procedure (Section~\ref{sec:do3se}).
 To study the impact of model parameters, we propose a method to define a parameterization for a subarctic climate based on existing Boreal parameterizations and local climate conditions, study the impact of stomata aperture parameterization at low light intensities, and estimate the start and end of the growing season from satellite derived GPP (Section~\ref{sec:localization}).
@@
 
-\subsubsection{Gap filling methodology}
+\subsubsection{Gap-filling methodology}
 \label{subsec:gap_filling}
-The $\mathrm{DO_3SE}$ model requires hourly, continuous input data. Therefore, the following gap filling methodology was devised for the meteorological input data:
+The $\mathrm{DO_3SE}$ model requires hourly, continuous input data. Therefore, the following gap-filling methodology was devised for the meteorological input data:
 \begin{itemize}
@@
 For this site, the anomalies with respect to the generalized climatology were calculated for the gap period.
-These anomalies were than added to the climatology of the site with the gap and corrected for an increase in present-day tropospheric background ozone compared to the early 1990s.
+These anomalies were then added to the climatology of the site with the gap and corrected for an increase in present-day tropospheric background ozone compared to the early 1990s.
 For more details and a comparison with regional air quality models refer to \citet{ACP:Falk2021}.
@@
 In contrast, no such visible injuries were found on the clovers in 2019, although it was observed on the sensitive tobacco cultivar (Bel-W3).
-Note that the extent of visible injuries is, in general, not equivalent to a quantitative reduction of above/below ground biomass of a species.
+Note that the extent of visible injuries is, in general, not equivalent to a quantitative reduction of above/below-ground biomass of a species.
 
@@
 
-Initial simulations with the $\mathrm{DO_3SE}$ model using the Central European MM parameterization for perennial grassland revealed an unexpectedly low leaf level stomatal conductance ($G_\mathrm{sto}^\mathrm{leaf}$) in 2019 (Appendix~Fig.~\ref{fig:pody_mm_composit}).
+Initial simulations with the $\mathrm{DO_3SE}$ model using the Central European MM parameterization for perennial grassland revealed an unexpectedly low leaf-level stomatal conductance ($G_\mathrm{sto}^\mathrm{leaf}$) in 2019 (Appendix~Fig.~\ref{fig:pody_mm_composit}).
 Substantial stomatal conductance occurred only during an extended warm period in late July.
@@
 To derive a subarctic parameterization for the temperature response function $f_\mathrm{T}$, we divide the temperature dataset into two subsets: 1992--2000 (near past) and 2011--2019 (present).
-From these two subsets, we select only dates within the GS (May--September) and plott the probability density functions (PDFs) as histograms with $1\,\unit{^\circ C}$ binning (Fig.~\ref{fig:f_temp_grassland}a).
+From these two subsets, we select only dates within the GS (May--September) and plot the probability density functions (PDFs) as histograms with $1\,\unit{^\circ C}$ binning (Fig.~\ref{fig:f_temp_grassland}a).
 A comparison between the two PDFs indicates that during the past decade the temperature distribution has shifted towards higher temperatures reducing the number of days below $5\,\unit{^\circ C}$.
@@
 We define local climate-derived parameterizations for two hypothetical species (\emph{subarctic} and \emph{cold}) by increasing the overlapping area between $f_\mathrm{T}$ and the temperature PDF during the GS.
-Then we construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures than the MM species.
+Then we construct \emph{cold} as representative of a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures than the MM species.
 This is accomplished by moving $T_\mathrm{min}$ and $T_\mathrm{opt}$ towards cooler temperatures but $T_\mathrm{max}$ is kept at its MM value.
@@
 
-The accumulation of ozone dose is strongly depended on the timing and length of the GS.
-\citet{ICP:MappingManual2017} suggested a simple latitude model for beech, birch, and Norway spruce to localize start and end of the GS:
+The accumulation of ozone dose is strongly dependent on the timing and length of the GS.
+\citet{ICP:MappingManual2017} suggested a simple latitude model for beech, birch, and Norway spruce to localize the start and end of the GS:
 %
@@
 This assumption is supported by observations in Rovaniemi, Finland \citep[][Supplement~Fig.~S1]{FCR:Korhonen2018}.
-Supposed, the linear trend for temperture-degree-days defined GS over the past 6~decades prevails, GS at Svanhovd could converge with GS in Central Europe by the end of the century.
+Supposedly, the linear trend for temperature-degree-days defined GS over the past 6~decades prevails, GS at Svanhovd could converge with GS in Central Europe by the end of the century.
 All GS results are tabulated in Table~\ref{tab:sensitivity_tests_gs}.
@@
 
-We did not  derive a local climate parameterization for the water vapour pressure deficit function $f_\mathrm{VPD}$ and soil water potential $f_\mathrm{SWP}$ response functions.
+We did not derive a local climate parameterization for the water vapor pressure deficit function $f_\mathrm{VPD}$ and soil water potential $f_\mathrm{SWP}$ response functions.
 But acknowledge that these are very important factors in the modeling of ozone dose as pointed out by~\citet{ACP:Bueker2012}.
 
-The quasi-laminar boundary layer resistance $R_b$ is a function of wind speed $u$ and influenced by the tree height and leaf dimensions.
+The quasi-laminar boundary layer resistance $R_b$ is a function of wind speed $u$ and is influenced by the tree height and leaf dimensions.
 For comparison with MM, we collected a sample of birch (B.
@@
 For one Scots pine forest at Svanhovd that has been studied as a part of the ICP Forest mapping, heights were measured in 2004 when the stand was 90~years old.
-The mean tree height then was $10.1\,\unit{m}$ and maximum height was $16.2\,\unit{m}$ (V.
+The mean tree height then was $10.1\,\unit{m}$ and the maximum height was $16.2\,\unit{m}$ (V.
 Timmerman, personal communication, June 2020).
@@
 We presume that perennial vegetation will likely maximize carbon acquisition during the short subarctic growing season.
-Because the MM version of the $\mathrm{DO_3SE}$ model currently does not simulate net photosynthesis $A_\mathrm{n}$, we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011}.
+Because the MM version of the $\mathrm{DO_3SE}$ model currently does not simulate net photosynthesis $A_\mathrm{n}$, we assume a first-order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011}.
 A higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ ratio, thus, indicates a higher growth potential in the local climate.
@@
 The coldest ever recorded temperature was $-45.2\,\unit{^\circ C}$ on January~27, 1999, while the highest temperature ($29.4\,\unit{^\circ C}$) occurred on July~16 the same year.
-As can be seen, mean and median are well aligned in the center of the IQR for most months meaning that temperature anomalies are symmetrically distributed.
+As can be seen, the mean and median are well aligned in the center of the IQR for most months meaning that temperature anomalies are symmetrically distributed.
 
@@
 The average annual accumulated precipitation ($\sum_m\mathrm{Precip}$) given with standard error of mean is $(383\pm 86)\,\unit{mm}$.
-Mean and median precipitation are less well aligned compared to temperature and point to a considerable interannual variability.
+The mean and median precipitation are less well aligned compared to temperature and point to a considerable interannual variability.
 
 The climatology for global radiant energy $Q_e$ in \unit{J\,m^{-2}} is zero in November--January (polar night) and reaches its maximum in May--July (midnight sun conditions).
-Mean and median are well aligned.
+The mean and median are well aligned.
 Only June displays a slightly skewed distribution towards higher $Q_e$.
@@
 As no \chem{[O_3]} was recorded during winter (refer to Section~\ref{subsec:weather}), no anomalies are available for these months.
-In April/May 2018, ozone was unusually elevated and stayed higher than usual through the whole GS.
+In April/May 2018, ozone was unusually elevated and stayed higher than usual throughout the whole GS.
 The highest anomaly in 2019 occurred in September at the end of the GS, as the spring peak had not been fully captured.
@@
 The more extreme the local climate-derived temperature response function the higher $\mathrm{POD_1}$.
-In 2018, coniferous trees and perennial grassland display a reduction in $\mathrm{POD_1}$ for the subarctic parameterization compared to the cold paramerization in responds to the prescribed lower $g_\mathrm{sto}$ at higher temperatures.
+In 2018, coniferous trees and perennial grassland display a reduction in $\mathrm{POD_1}$ for the subarctic parameterization compared to the cold parameterization in response to the prescribed lower $g_\mathrm{sto}$ at higher temperatures.
 This is not seen in 2019.
@@
 The magnitude of all described effects on $\mathrm{POD_1}$ varies between PFTs and years.
-By construction, $\mathrm{POD_1}$ from local climate-derived parmeterizations is always larger than estimates with the MM parameterization.
+By construction, $\mathrm{POD_1}$ from local climate-derived parameterizations is always larger than estimates with the MM parameterization.
 This means that the risk of ozone damage could be underestimated especially in more normal years like 2019 (see Section~\ref{subsec:stats} for details).
-Establishing dedicated subarctic parameterizations are most likely required to properly assess risk.
+Establishing dedicated subarctic parameterizations is most likely required to properly assess risk.
 
@@
 
-In the following, we estimate the biomass reductions in accordance to the CLs \citep{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme climate-derived local temperature parameterization (subarctic) for each PFT.
+In the following, we estimate the biomass reductions in accordance with the CLs \citep{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme climate-derived local temperature parameterization (subarctic) for each PFT.
 The estimates, assuming no soil water stress, are given in Table~\ref{tab:biomass_reduction}.
@@
 As we have shown, 2019 was a rather normal year compared to 2018 with respect to present-day climate and ozone concentrations.
-This indicates that risk of ozone-induced damage in subarctic vegetation has been previously underestimated in assessments using the Boreal MM parameterizations.
+This indicates that the risk of ozone-induced damage in subarctic vegetation has been previously underestimated in assessments using the Boreal MM parameterizations.
 In future scenarios in which persistent heatwaves like in 2018 may become more common, the risk for ozone-induced damage is probably reduced for vegetation adapted to cold growing conditions.
@@
 Overall, ozone was more enhanced in 2018 than 2019 and it is very likely that elevated ozone, like during the 2003 drought period \citep{JGR:Solberg2018}, was promoted by a combination of various factors such as wildfires, reduced cloud cover (increased solar radiation), reduced dry deposition and turbulent mixing due to the stagnant weather conditions, and increased BVOC emissions.
-As such events may become more frequent in the future, conditions at Svanhovd for the growing seasons 2018 could serve as a reference for probable future conditions in northern Fennoscandia.
+As such events may become more frequent in the future, conditions at Svanhovd for the growing season 2018 could serve as a reference for probable future conditions in northern Fennoscandia.
 
@@
 The $\mathrm{DO_3SE}$ model derived summer maximum $V_\mathrm{d}$ is $\sim 9\,\unit{mm\,s^{-1}}$ with minimal differences between the parameterizations (see supplementary material Figs.~S1--S3).
-The highest $V_\mathrm{d}$ is found in summer 2019 using the cold paramterization.
+The highest $V_\mathrm{d}$ is found in summer 2019 using the cold parameterization.
 The maximum observed spring $V_\mathrm{d}\sim 8\,\unit{mm\,s^{-1}}$ is slightly smaller compared to $V_\mathrm{d}\sim 10\,\unit{mm\,s^{-1}}$ from the $\mathrm{DO_3SE}$ model.
@@
 
-Along the conventional ozone monitoring at Svanhovd, we had installed a bio-monitoring with ozone sensitive species.
+Along with the conventional ozone monitoring at Svanhovd, we installed a bio-monitoring with ozone-sensitive species.
 Qualitatively, this showed visible ozone injuries on clover and tobacco in 2018.
@@
 That might explain why we observed visible injuries in the ozone garden to a larger extent in 2018 than in 2019.
-An associated damage function for all PFTs relating visible ozone-injuries or phenological state might improve such predictions.
+An associated damage function for all PFTs relating to visible ozone injuries or phenological states might improve such predictions.
 
@@
 We determined the start of the growing season for deciduous trees using growing-degree-days above $5\,\unit{^\circ C}$ on gridded observational temperature.
-The results coincide with the first snow free days in our target region.
+The results coincide with the first snow-free days in our target region.
 For the period $1994-2003$, \citet{IJB:Shutova2006} reported an observed average start of the GS for mountain birch at Svanhovd at $\unit{doy}~154$ and end of GS at $\unit{doy}~229$.
@@
 Due to a climate change-induced convergence of the start of GS with the ozone spring peak \citep{EP:Karlsson2007}, vegetation might become more exposed to higher ozone concentrations in early phenological stages.
-However, the decline of this ozone spring peak is partly caused by dry deposition of ozone to vegetation itself.
+However, the decline of this ozone spring peak is partly caused by the dry deposition of ozone to vegetation itself.
 It, therefore, remains unclear whether an earlier start of the growing season will increase the exposure of vegetation to ozone or lead to compensation due to an earlier decline of the peak ozone in the future.
@@
 The actual resilience of northern Fennoscandia plant communities to climate change strongly depends on both the range of acclimation and the type of acclimation.
-This means, vegetation that is acclimated to cold temperatures and less to heat- and drought stress might suffer more strongly from a more frequent occurrence of heatwaves than plants that have a higher temperature tolerance.
+This means that vegetation that is acclimated to cold temperatures and less to heat- and drought stress might suffer more strongly from a more frequent occurrence of heatwaves than plants that have a higher temperature tolerance.
 Concerning the ozone stress, our results point to the opposite, namely that the closure of stomata at higher temperatures coinciding with high surface air ozone concentrations will prevent a high ozone dose in species acclimated to cold climates.
 
-We found that the difference in biomass reduction compared to Boreal parmeterizations \citep{ICP:MappingManual2017} could amount to $55\,\unit{\%}$ in deciduous trees and $29-43\,\unit{\%}$ in perennial grassland.
+We found that the difference in biomass reduction compared to Boreal parameterizations \citep{ICP:MappingManual2017} could amount to $55\,\unit{\%}$ in deciduous trees and $29-43\,\unit{\%}$ in perennial grassland.
 But the used biomass reduction functions were established based on studies from less extreme climates, hence, these results ought to be treated with care.
@@
 Beyond the risk assessment of ozone-induced damage, our climate-derived local Jarvis-type temperature response functions, once verified by observations, can have important implications on land-surface modeling still using only Boreal parameterizations in global models.
-Problems in estimating the productivity of species especially in the Arctic regions occur due to PFTs which are not suitable for the climate.
-A verification of our proposed methodology by independent flux and dry deposition measurements over birch forest and perennial grassland would be necessary to establish a climate data based local parameterization of stomatal conductance.
+Problems in estimating the productivity of species, especially in the Arctic regions occur due to PFTs which are not suitable for the climate.
+Verification of our proposed methodology by independent flux and dry deposition measurements over birch forest and perennial grassland would be necessary to establish a climate data-based local parameterization of stomatal conductance.
 
@@
 (a) $f_\mathrm{T}$ and (b) $f_\mathrm{flight}$ are shown together with underlying $T_\mathrm{air}$ and $Q_0$ climatologies (probability density function - PDF), respectively.
-Original mapping manual parameterization is shown in comparison as solid line.
+Original mapping manual parameterization is shown in comparison as a solid line.
 Note that the y-axis has been truncated to $0.006$.

### 2022-06-02 review/final_response.tex
Add final response

% Authors' response
\documentclass{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{multirow}
\thispagestyle{empty}

\begin{document}
\section{Final response}
In response to the major concerns of the referees, we had to rephrase our research question. We now focus on the uncertainty in the assessment of ozone-induced damage risk that follow the UNECE LRTAP Convention concerning the subarctic region. In these assessments, it is assumed that Boreal parameterizations of stomatal conductance are sufficient to describe also the response of subarctic vegetation.

To this end, we propose a local climate-derived temperature parameterization of stomatal conductance by shifting the existing Boreal parameterizations to lower temperatures. This is analogous to what has been experimentally found for temperature acclimation of $T_\mathrm{opt}$ in plant photosynthesis. We adjust stomatal conductance because the used model doesn't compute photosynthesis.

Based on this, we restructured the manuscript taking all relevant referee comments into consideration. 
\begin{itemize}
    \item The introduction has been condensed.
    \item Speculative phrases and formulations removed or rephrased.
    \item We separated methods, results, and discussion into respective sections.
    \item The discussion about local climate was also condensed and put into the context of the research question.
    \item Considering the attribution of "significance", we refrain now from using this term inappropriately and
    \item reworked the statistical analysis and presentation of the environmental data.
    \item We use box plots to condense statistical information about the data distribution.
\end{itemize}

Considering the comparison with flux measurements, we were not able to find appropriate data for northern Fennoscandia. Therefore, we discuss the difference in modeled ozone dry deposition velocities for coniferous trees with what has been observed at the forest research station Hyytilälä in Finland that is further south and not applicable to deciduous forests or perennial grassland. The comparison showed good agreement in summer but probably too high modeled deposition in spring.

\end{document}
